### 2012-11-21 Publications/2013-IPMI/Latex/00-Main.tex
Added base for IPMI2013 #25 I upload here the introduction. I take the last paragraph almost literally from the original Dominique's document.

\documentclass{llncs}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{acronym}
\usepackage[numbers]{natbib}
\bibliographystyle{splncsnat}
%\bibliographystyle{IEEEtran}
\usepackage[hyphens]{url}
\usepackage{nameref}
%\rewewcommand\bibname{References}
%

\begin{document}
%
%\tableofcontents
%
\mainmatter              % start of the contributions
%
\title{Simultaneous segmentation and distortion correction on diffusion weighted MR using structural priors}
%
\titlerunning{Segmentation & EPI correction on DWI}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Oscar Esteban\inst{1,2} \and Alessandro Daducci \inst{2} \and
        Virginia Estellers \inst{2} \and Luminita Vese \inst{3} \and 
        Jean-Philippe Thiran \inst{2} \and Andr\'es Santos \inst{1} \and
        Dominique Zosso \inst{2,3}}
%
\authorrunning{Oscar Esteban et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
%Craig Chambers, Kim B. Bruce, and Elisa Bertino}
%

\institute{Biomedical Image Technologies (BIT),
ETSI Telecomunicaci\'on - Universidad Polit\'ecnica de Madrid
and CIBER-BBN, \\
Av. Complutense 30, E-28040 Madrid, Spain\\
\email{oscar.esteban@upm.es},\\
% WWW home page:
%\texttt{http://users/\homedir iekeland/web/welcome.html}
\and
Signal Processing Laboratory 5 (LTS5),
\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL)\\
EPFL-STI-IEL-LTS5, Station 11, CH-1015 Lausanne, Switzerland
\and
Department of Mathematics, ... 
University of California Los Angeles (UCLA),
Los Angeles,..
}

\maketitle              % typeset the title of the contribution

\input{01-Acronyms}

\input{02-Abstract}

\input{03-Introduction}

\input{04-Methods}

\input{05-Results}

\input{06-Conclusion}

\bibliography{99-References}

\end{document}

### 2012-11-21 Publications/2013-IPMI/Latex/02-Abstract.tex
Added base for IPMI2013 #25 I upload here the introduction. I take the last paragraph almost literally from the original Dominique's document.

\begin{abstract}
The abstract should summarize the contents of the paper
using at least 70 and at most 150 words. It will be set in 9-point
font size and be inset 1.0 cm from the right and left margins.
There will be two blank lines before and after the Abstract.
%\dots
%\keywords{diffusion weighted imaging, connectomics, echo planar imaging, magnetic resonance, segmentation, registration, distortion correction}
\end{abstract}

### 2012-11-21 Publications/2013-IPMI/Latex/03-Introduction.tex
Added base for IPMI2013 #25 I upload here the introduction. I take the last paragraph almost literally from the original Dominique's document.

\section{Introduction}
\label{sec:introduction}
%
\ac{dwi} is a widely used family of \ac{mr} techniques
\citep{sundgren_diffusion_2004} which recently has accounted for a growing
interest in its application to whole-brain structural connectivity analysis.
This emerging field, coined in 2005 as \emph{\ac{mr} Connectomics}
\citep{hagmann_diffusion_2005,sporns_human_2005}, currently includes a
large amount of imaging techniques for acquisition, processing and analysis
specifically tuned for the \ac{dwi} data \citep{hagmann_mr_2012}.\\

The whole-brain connectivity analysis has arisen some challenges
that should be overcome in order to get reliable structural information
about the neuronal tracts from \ac{dwi}. The earlier stages of this
processing tools generally include two necessary steps, brain tissue
segmentation on the diffusion space and the correction of geometrical
distortions produced by the imaging techniques.\\

In this work, we will refer as brain tissue segmentation to the precise
delineation of the \ac{csf}-\ac{gm} and \ac{gm}-\ac{wm} interface surfaces.
This segmentation is an important step on which strongly rely further
tasks. In tractography, a high-standard \ac{wm} mask is required. Otherwise,
there is an important risk for the algorithm to lose fiber bundles. This
requirement is usually satisfied by plainly thresholding the \ac{fa}, a
well-know scalar map derived from \ac{dwi} which depicts the isotropy of
water diffusion inside the brain. Additionally, a precise location of the
\ac{gm}-\ac{wm} surface is required in the final steps to
achieve a consistent parcellisation of the cortex to represent the nodes 
of the output network. This parcellisation is generally defined in a 
high-resolution and better understood structural \ac{mri} of the same 
subject (eg. T1 and/or T2 weighted acquisitions). Conversely, this 
problem is resolved with non-linear registration of a structural \ac{mri}
of the subject to the \ac{dwi} data.\\

On the other hand, the \ac{dwi} data is usually obtained with \ac{epi}
acquisition techniques, that often suffer from severe distortions due to 
local field inhomogeneities. Generally, it is easily appreciated in the anterior
part of the brain, along the phase-encoded direction. Some methodologies have
been developed [CITATIONS] and named as \emph{\ac{epi}-unwarp} techniques,
and they require the extra acquisition of the magnitude and phase of
the field (field-mapping), condition which is not always met. Some other 
methodologies do not make use the field-mapping, compensating the distortion
with non-linear registration from structural \ac{mri}.\\

In this paper we propose a novel registration framework to simultaneously
solve the segmentation and distortion challenges, by exploiting as strong 
shape-prior the detailed anatomy extracted from anatomical \ac{mri}. 
We reformulate the segmentation problem as an inverse problem, where
we seek for an underlying deformation field (the distortion) mapping 
from the structural space into the diffusion space.

### 2012-11-21 Publications/2013-IPMI/Latex/05-Results.tex
Added base for IPMI2013 #25 I upload here the introduction. I take the last paragraph almost literally from the original Dominique's document.

\section{Results and discussion}
\label{sec:results}
%

### 2012-11-23 Publications/2013-IPMI/Latex/04-Methods.tex
Another little step on writing (fast-writing, btw)

\section{Methods}
\label{sec:methods}
%
\subsection{Simulated datasets}
%
As suggested in \autoref{sec:introduction}, the general situation 
consists of having reliable segmentations on the \ac{t1} reference
space, obtained with 
\emph{FreeSurfer\footnote{\url{http://surfer.nmr.mgh.harvard.edu/}}}.
Therefore, regarding the proposed solution, we will have a precise
location of the tissue interfaces of interest in a reference space.
On the other hand, we have a \ac{dwi} volume, characterized by its
low resolution (typically around $2.2x2.2x3mm^3$). Depending on the
posterior reconstruction methodology and the angular resolution
intended, the \ac{dwi} raw data has to be processed in order to
extract the information in a manageable manner. Particularly, we
will use the \ac{fa} and \ac{md} maps. Whereas \ac{fa} describes
the \emph{shape} of diffusion, the \ac{md} depicts the
\emph{intensity} of the process. There exist to main reasons to 
justify their choice. First, they are well-understood and
standardized in clinical routine. Second, they are statistically 
orthogonal and together contain most of the information that is
usually extracted from the \ac{dwi}-derived scalar maps. \\
In order that demonstrating the functionality of the proposed
methodology and characterize its possibilities, we developed two
synthetic model, generating the \ac{dwi} data as described in
 \citep{tuch_q-ball_2004}. We selected 30 directions, for being
a very common protocol for \ac{dti} reconstruction. The first model
is a set of spherical shapes representing the different brain
tissues. The second model is based on the BrainWeb dataset. We
reconstructed the \ac{dwi} data with standard to approximate the
environment to the real one at maximum. There is no interest on the
anatomical reference, given that with the models we hold \emph{a priori}
precisely located surface of the interfaces of interest. \\

FIGURE OF THE MODELS AND EXTRACTED FA, MD

\subsection{\acl{acwe}-like variational segmentation model}

I will summarize here your mathematical formulation, but with simplifications:
- only gradient descent
- no mention to anisotropic and inhomogeneous regularization.

\subsection{Experiment}
%
For both models, we created manually a sound distortion visually similar
to real \ac{epi} distortions. We interpolated the distortion to a 
dense deformation field, necessary for warping the raw \ac{dwi} simulated
data. Once the signal was deformed, we proceeded to reconstruct the
\ac{dti} and subsequently obtained the scalars of interest (\ac{fa}, \ac{md}).\\

We evaluate the performance of our methodology to estimate the deformation
field, obtaining a precise segmentation on the diffusion space.

### 2012-12-05 Publications/2013-IPMI/Latex/04-Methods.tex
Added Dominique's formulations I just added the interesting parts, but lots of work should be done:   * Clear out all references to T1. We should only keep the term ``shape prior''   * Re-write the introduction to the ACWE model. Remove the part of w_j (don't they belong to the DWI only?)   * Some explanations are lacking (I didn't copy the Augmented Lagrangians part)   * Use the discrete version of the update since the begining.
Additionally, I don't like the subsection Simulated Datasets. I have to improve it significantly (and summarize).

@@
 the \emph{shape} of diffusion, the \ac{md} depicts the
-\emph{intensity} of the process. There exist to main reasons to 
+\emph{manitude} of the process. There exist to main reasons to 
 justify their choice. First, they are well-understood and
@@
 \subsection{\acl{acwe}-like variational segmentation model}
-
-I will summarize here your mathematical formulation, but with simplifications:
-- only gradient descent
-- no mention to anisotropic and inhomogeneous regularization.
+%
+Let us denote $\{c_i\}_{i=1..N_c}$ the nodes of the \ac{wm}-\ac{gm}
+interface, and $\{d_i\}_{i=1..N_d}$ the nodes of the pial surface. 
+Within the respective volumes, we have sets of sample points, denoted 
+$\{w_j\}_{j=1..N_w}$, $\{g_j\}_{j=1..N_g}$, and $\{o_j\}_{j=1..N_o}$ 
+for the white matter, grey matter and CSF, respectively. All those 
+are given in high-resolution T1 reference coordinates.\\
+%
+On the other hand, we have low resolution \ac{dwi}. At each voxel 
+we have a sampled ODF (or some extracted features).
+Let us denote by $x$ the voxel and $f(x) = [ f_1, f_2, \ldots, f_N]^T(x)$ 
+its associated feature vector.
+%
+The transformation from T1 into DW reference coordinate space is 
+achieved through a dense deformation field $u(x)$, such that for example
+the nodes of the w/g-interface are located in DW-space as follows:
+%
+\begin{equation}
+c_i' = T\{c_i\} = c_i + u(c_i)
+\end{equation}
+% 
+Since the nodes of the anatomical surfaces might lay off-grid, it is 
+required to derive $u(x)$ from a discrete set of parameters $\{u_k\}_{k=1..K}$.
+Densification is achieved through a set of associated basis functions 
+$\psi_k$ (e.g. rbf, interpolation splines):
+%
+\begin{equation}
+u(x) = \sum_k \psi_k(x) u_k
+\end{equation}
+%
+Consequently, the transformation writes
+%
+\begin{equation}
+c_i' = T\{c_i\} = c_i + u(c_i) = c_i + \sum_k \psi_k(c_i)u_k
+\end{equation} 
+%
+Note that, since $c_i$ remains constant in the DW segmentation process,
+the values of $\psi_k(c_i)$ can be precomputed. Also, provided compact 
+support of the basis functions, the system remains relatively sparse.
+%
+Based on the region-wise samples $\{w_j\}$, $\{g_j\}$, and $\{o_j\}$, 
+and the current estimate of the distortion $u$, we can compute 
+``expected samples'' of the respective regions in DW space, written
+$\{w_j'\}$, $\{g_j'\}$, and $\{o_j'\}$. Based on those samples, we may 
+now estimate region descriptors of the DW features $f(x)$ of the three 
+respective regions in DW space. In the simplest case, estimate the 
+parameters $\mu_R$ and $\Sigma_R$, i.e. the regions' mean feature 
+vector and covariance matrix, for each region $R\in \{w,g,o\}$.
+%
+Based on these Gaussian region descriptors, we propose an \ac{acwe}-like, 
+piece-wise constant, variational image segmentation model (where the 
+unknown is the deformation field)\cite{chan_active_2001}:
+\begin{align}
+E(u) &= \int_{w'(u)} (f-\mu_w)^T\Sigma_w^{-1}(f-\mu_w) dx\nonumber\\
+&\quad +\int_{g'(u)} (f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g) dx\\
+&\quad +\int_{o'(u)} (f-\mu_o)^T\Sigma_o^{-1}(f-\mu_o) dx\nonumber
+\end{align}
+where the integral domains depend on the deformation field u. Note that 
+minimizing this energy, $\argmin_u\{E\}$, yields the MAP estimate of a 
+piece-wise smooth image model affected by Gaussian additive noise. This 
+inverse problem is ill-posed \cite{hadamard_sur_1902,bertero_ill-posed_1988}.
+In order to account for deformation field regularity and to render the 
+problem well-posed, we include limiting and regularization terms into 
+the energy functional 
+\cite{tichonov_solution_1963,morozov_linear_1975}:
+\begin{align}
+E(u) &= \int_{w'(u)} (f-\mu_w)^T\Sigma_w^{-1}(f-\mu_w) dx\nonumber\\
+&\quad +\int_{g'(u)} (f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g) dx\\
+&\quad +\int_{o'(u)} (f-\mu_o)^T\Sigma_o^{-1}(f-\mu_o) dx\nonumber\\
+&\quad +\int u^T A u dx + \int \tr\{(\nabla u^T)^T B (\nabla u^T)\} dx\nonumber
+\end{align}
+These regularity terms ensure, that the segmenting contours in DW space 
+are still close to their native shape in T1. In the simplest case, both 
+$A=\alpha$ and $B=\beta$ are constant scalars, and the terms rewrite 
+$\alpha \int  \|u\|^2 dx + \beta \int \left( \|\nabla u_x\|^2 +
+\|\nabla u_y\|^2 + \|\nabla u_z\|^2\right) dx$, which are familiar. 
+Later, $A$ and $B$ might be spatially varying and/or $3\times 3$ 
+matrices, therefore allowing to incorporate inhomogeneous and 
+anisotropic regularization \cite{nagel_investigation_1986}.
+%
+At each iteration, we update the distortion along the steepest energy descent:
+%
+\begin{equation}
+\frac{\partial u_k^t}{\partial t} = -\frac{\partial E(u)}{\partial u_k^t}
+\end{equation}
+%
+At this point, we interpret the parameter field $u_k$ to be a 
+continuous function $u(x)$, sampled at the locations $x_k$, and 
+determine the gradient-descent equation\footnote{The same assumption 
+was being made above in the minimization of the regularity terms}:
+%
+\begin{align}
+\frac{\partial u^t}{\partial t} &= - \sum_{i=1}^{N_c} \left[(f(c_i')-\mu_g)^T\Sigma_g^{-1}(f(c_i')-\mu_g) - (f(c_i')-\mu_w)^T\Sigma_w^{-1}(f(c_i')-\mu_w)\right]\psi_{c_i'}(x)N_{c_i'}\nonumber\\
+&\quad -\sum_{i=1}^{N_d} \left[(f(d_i')-\mu_o)^T\Sigma_o^{-1}(f(d_i')-\mu_o) - (f(d_i')-\mu_g)^T\Sigma_g^{-1}(f(d_i')-\mu_g)\right]\psi_{d_i'}(x)N_{d_i'}\\
+&\quad -\alpha u + \beta\Delta u\nonumber
+\end{align}
+%
+where we have swapped $\psi_k(c_i')$ into $\psi_{c_i'}(x)$.
+%
+This gradient descent step can be efficiently tackled by discretizing 
+the time in a forward Euler scheme, and making the right hand side 
+semi-implicit in the regularization terms:
+%
+\begin{align}
+\frac{u^{t+1}-u^t}{\tau} &= - \sum_{i=1}^{N_c} \left[(f(c_i')-\mu_g)^T\Sigma_g^{-1}(f(c_i')-\mu_g) - (f(c_i')-\mu_w)^T\Sigma_w^{-1}(f(c_i')-\mu_w)\right]\psi_{c_i'}(x)N_{c_i'}\nonumber\\
+&\quad -\sum_{i=1}^{N_d} \left[(f(d_i')-\mu_o)^T\Sigma_o^{-1}(f(d_i')-\mu_o) - (f(d_i')-\mu_g)^T\Sigma_g^{-1}(f(d_i')-\mu_g)\right]\psi_{d_i'}(x)N_{d_i'}\nonumber\\
+&\quad -\alpha u^{t+1} + \beta\Delta u^{t+1}
+\end{align}
+%
+where the data terms remain functions of the current estimate $u^t$, i.e. 
+all $c_i' = c_i'(u^t)$ and $d_i' = d_i'(u^t)$. Again, we propose a spectral 
+approach to solve this implicit scheme:
+%
+\begin{equation}
+u^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{u^t/\delta - \sum_{i=1}^{N_c}(\ldots) - \sum_{i=1}^{N_d}(\ldots)  \}}{\mathcal{F}\{(1/\delta+\alpha)\mathcal{I}-\beta\Delta\}} \right\}
+\end{equation}
+%
+It is easily verified, that the same update can be obtained by plugging the 
+AL-update w.r.t. $u$ into the AL-update w.r.t. $v$, and by identifying 
+$r = 1/\delta$ (the only exception is the distortion on which the data-term 
+is being evaluated).
 

### 2012-12-08 Publications/2013-IPMI/Latex/04-Methods.tex
Reviewed equations on methods section. Close #32

--- 2012-11-23 Publications/2013-IPMI/Latex/04-Methods.tex
@@
 \subsection{\acl{acwe}-like variational segmentation model}
+%
+Let us denote $\{c_i\}_{i=1..N_c}$ the nodes of a shape prior surface. In
+our application, a precise \ac{wm}-\ac{gm} interface extracted from a
+high-resolution reference volume. All the formulations can be naturally
+extended to include more shape priors. On the other hand, we have a 
+number of \ac{dwi}-derived features at each
+voxel of the volume. Let us denote by $x$ the voxel and 
+$f(x) = [ f_1, f_2, \ldots, f_N]^T(x)$ its associated feature vector.\\
+%
+The transformation from reference into \ac{dwi} coordinate space is 
+achieved through a dense deformation field $u(x)$, such that:
+%
+\begin{equation}
+c_i' = T\{c_i\} = c_i + u(c_i)
+\end{equation}
+% 
+Since the nodes of the anatomical surfaces might lay off-grid, it is 
+required to derive $u(x)$ from a discrete set of parameters $\{u_k\}_{k=1..K}$.
+Densification is achieved through a set of associated basis functions 
+$\Psi_k$ (e.g. rbf, interpolation splines):
+%
+\begin{equation}
+u(x) = \sum_k \Psi_k(x) u_k
+\end{equation}
+%
+Consequently, the transformation writes
+%
+\begin{equation}
+\label{eq:transformation}
+c_i' = T\{c_i\} = c_i + u(c_i) = c_i + \sum_k \Psi_k(c_i)u_k
+\end{equation} 
+%
+% Comment: maybe this is not for IPMI 2013.
+%Note that, since $c_i$ remains constant in the DW segmentation process,
+%the values of $\Psi_k(c_i)$ can be precomputed. Also, provided compact 
+%support of the basis functions, the system remains relatively sparse.\\
+%
+Based on the current estimate of the distortion $u$, we can compute 
+``expected samples'' within the shape prior projected into the \ac{dwi}.
+Thus, we now estimate region descriptors of the \ac{dwi} features 
+$f(x)$ of the regions defined by the priors in \ac{dwi} space.
+%
+Using Gaussian distributions as region descriptors, we propose an
+\ac{acwe}-like, piece-wise constant, variational image segmentation
+model (where the unknown is the deformation field)
+\cite{chan_active_2001}:
+\begin{equation}
+\label{eq:gaussian_energy}
+E(u)= \sum_{\forall{R}} \int_{\Omega_R} (f-\mu_R)^T\Sigma_R^{-1}(f-\mu_R) dx
+\end{equation}
+where $R$ indexes the existing regions and the integral domains
+depend on the deformation field $u$. Note
+that minimizing this energy, $\argmin_u\{E\}$, yields the \ac{map} 
+estimate of a piece-wise smooth image model affected by Gaussian 
+additive noise. This inverse problem is ill-posed
+\cite{bertero_ill-posed_1988,hadamard_sur_1902}.
+In order to account for deformation field regularity and to render the 
+problem well-posed, we include limiting and regularization terms into 
+the energy functional \cite{morozov_linear_1975,tichonov_solution_1963}:
+%
+\begin{align}
+\label{eq:complete_energy}
+E(u) &= \sum_{\forall{R}} \lbrace \int_{\Omega_R} (f-\mu_R)^T\Sigma_R^{-1}(f-\mu_R) dx \rbrace \nonumber \\
+&\quad + \alpha \int  \|u\|^2 dx + \beta \int \left( \|\nabla u_x\|^2 + \|\nabla u_y\|^2 + \|\nabla u_z\|^2\right) dx
+\end{align}
+%
+These regularity terms ensure that the segmenting contours in 
+\ac{dwi} space are still close to their native shape. The model
+easily allows to incorporate inhomogeneous and anisotropic 
+regularization \cite{nagel_investigation_1986} to better regularize
+the \ac{epi} distortion. \\
+%
 
-I will summarize here your mathematical formulation, but with simplifications:
-- only gradient descent
-- no mention to anisotropic and inhomogeneous regularization.
+At each iteration, we update the distortion along the steepest 
+energy descent. This gradient descent step can be efficiently 
+tackled by discretizing the time in a forward Euler scheme, 
+and making the right hand side semi-implicit in the 
+regularization terms:
+%
+\begin{align}
+\frac{u^{t+1}-u^t}{\tau} &= - \sum_{i=1}^{N_c} \left[ \Delta E(f(c_i'))  \hat{n}_{c_i'} \Psi_{c_i}(x) \right] -\alpha u^{t+1} + \beta\Delta u^{t+1}
+\end{align}
+%
+where the data terms remain functions of the current estimate 
+$u^t$, thus $c_i' = c_i'(u^t)$. For simplicity on notation, we 
+restricted the number of priors to only 1. We also defined 
+$\Delta E(f(c_i')) = E_{out}(f(c_i')) - E_{in}(f(c_i'))$, 
+and $E_R(f(c_i')) = (f(c_i')-\mu_R)^T\Sigma_R^{-1}(f(c_i')-\mu_R)$.
+We applied a spectral approach to solve this implicit scheme:
+%
+\begin{equation}
+u^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{u^t/\tau
+- \sum_{i=1}^{N_c} \left[ \Delta E(f(c_i')) \hat{n}_{c_i'} \Psi_{c_i}(x) \right]  \}}{\mathcal{F}\{(1/\tau+\alpha)\mathcal{I}-\beta\Delta\}} \right\}
+\end{equation}
+%
 

### 2012-12-12 Publications/2013-IPMI/Latex/00-Main.tex
Another little step on writing (fast-writing, btw)
Draft update (#25) Latex was not compiling due to incompatibility between the hyperref package and the Springer template. The error is as follows
::
Package hyperref Warning: bookmark level for unknown title defaults to 0.
Package hyperref Warning: bookmark level for unknown author defaults to 0.
Latex Error: ./D1.3.tex:30 Missing \endcsname inserted.

To fix it, just install llncs2e >=2.15
Reviewed equations on methods section. Close #32
Initial draft for the conference. Close #25

@@
 \usepackage[numbers]{natbib}
-\bibliographystyle{splncsnat}
+\bibliographystyle{splncs03} 
+%\bibliographystyle{splncsnat} 
 %\bibliographystyle{IEEEtran}
+
 \usepackage[hyphens]{url}
 \usepackage{nameref}
-%\rewewcommand\bibname{References}
+\usepackage[breaklinks]{hyperref}
+\usepackage[USenglish]{babel}
+\usepackage[T1]{fontenc}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{pslatex}
 %
 
-\begin{document}
+\DeclareMathOperator{\tr}{tr}
+\DeclareMathOperator{\argmax}{argmax}
+\DeclareMathOperator{\argmin}{argmin}
+\DeclareMathOperator{\diag}{diag}
+\DeclareMathOperator{\dist}{dist}
+\DeclareMathOperator{\const}{Const.}
+
+\usepackage{graphicx}
+
 %
@@
 %
-\title{Simultaneous segmentation and distortion correction on diffusion weighted MR using structural priors}
-%
+\graphicspath{{../Figures/}}
+
+\begin{document}
+
+\title{Simultaneous segmentation and distortion correction on diffusion weighted MR using shape priors}
+%%
 \titlerunning{Segmentation & EPI correction on DWI}  % abbreviated title (for running head)
@@
 \author{Oscar Esteban\inst{1,2} \and Alessandro Daducci \inst{2} \and
-        Virginia Estellers \inst{2} \and Luminita Vese \inst{3} \and 
+		Meritxell Bach-Cuadra \inst{2} \and Virginia Estellers \inst{2}
+		\and Mar\'ia-Jes\'us Ledesma-Carbayo \and Luminita Vese \inst{3} \and 
         Jean-Philippe Thiran \inst{2} \and Andr\'es Santos \inst{1} \and
@@
 \authorrunning{Oscar Esteban et al.} % abbreviated author list (for running head)
-%
-%%%% list of authors for the TOC (use if author list has to be modified)
-%\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
-%Craig Chambers, Kim B. Bruce, and Elisa Bertino}
-%
 

### 2012-12-12 Publications/2013-IPMI/Latex/02-Abstract.tex
Initial draft for the conference. Close #25

@@
 \begin{abstract}
-The abstract should summarize the contents of the paper
-using at least 70 and at most 150 words. It will be set in 9-point
-font size and be inset 1.0 cm from the right and left margins.
-There will be two blank lines before and after the Abstract.
+Connectomics is the study of fiber connections in the brain, and in particular variations thereof.
+High-quality reconstructions of the connectome from diffusion weighted MR acquisitions require a precise delineation of the pial surface and the grey-matter-white-matter interface.
+Also, to enable significant group studies, these surfaces must be located and parcellated consistently across subjects.
+Precise segmentation and parcellation are readily available on anatomical images obtained e.g., through T1-weighted MR imaging.
+Diffusion weighted images, however, typically suffer from drastically lower resolution and more important image distortions.
+Here we propose a joint segmentation-registration framework by exploiting the detailed and consistent anatomy extracted from anatomical MRI as strong shape-prior for DW segmentation.
+We formulate an instance of Mumford-Shah-like active contours without edges, where we look for a regular deformation field that optimally maps the shape prior on the multivariate features in diffusion space.
 %\dots
-%\keywords{diffusion weighted imaging, connectomics, echo planar imaging, magnetic resonance, segmentation, registration, distortion correction}
+\keywords{diffusion weighted imaging, connectomics, echo planar imaging, magnetic resonance, segmentation, registration, distortion correction}
 \end{abstract}

### 2012-12-12 Publications/2013-IPMI/Latex/03-Introduction.tex
Initial draft for the conference. Close #25

@@
 \citep{hagmann_diffusion_2005,sporns_human_2005}, currently includes a
-large amount of imaging techniques for acquisition, processing and analysis
-specifically tuned for the \ac{dwi} data \citep{hagmann_mr_2012}.\\
+large amount of imaging techniques for acquisition, processing, and analysis
+specifically tuned for \ac{dwi} data.\\
 
@@
 that should be overcome in order to get reliable structural information
-about the neuronal tracts from \ac{dwi}. The earlier stages of this
-processing tools generally include two necessary steps, brain tissue
+about the neuronal tracts from \ac{dwi} \cite{johansen-berg_using_2009,
+jones_white_2012}. The earlier stages of these
+processing pipelines generally include two necessary steps, brain tissue
 segmentation on the diffusion space and the correction of geometrical
-distortions produced by the imaging techniques.\\
+distortions inherent to the imaging techniques \citep{hagmann_mr_2012}.\\
 
@@
 there is an important risk for the algorithm to lose fiber bundles. This
-requirement is usually satisfied by plainly thresholding the \ac{fa}, a
-well-know scalar map derived from \ac{dwi} which depicts the isotropy of
-water diffusion inside the brain. Additionally, a precise location of the
+requirement is usually solved in practice by plainly thresholding the 
+\ac{fa} (a well-know scalar map derived from \ac{dwi} which depicts 
+the isotropy of water diffusion inside the brain). 
+Additionally, a precise location of the
 \ac{gm}-\ac{wm} surface is required in the final steps to
@@
 high-resolution and better understood structural \ac{mri} of the same 
-subject (eg. T1 and/or T2 weighted acquisitions). Conversely, this 
+subject (e.g. T1 and/or T2 weighted acquisitions). Conversely, this 
 problem is resolved with non-linear registration of a structural \ac{mri}
-of the subject to the \ac{dwi} data.\\
+of the subject to the \ac{dwi} data. Even though some efforts have addressed
+the study of the robustness of tractography versus the intra-subject variability
+\cite{wakana_reproducibility_2007,heiervang_between_2006}, the results produced
+are restricted to relevant regions of the brain. Therefore, extremely robust
+and precise segmentation methods are required in the whole-brain application. \\
 
@@
 part of the brain, along the phase-encoded direction. Some methodologies have
-been developed [CITATIONS] and named as \emph{\ac{epi}-unwarp} techniques,
-and they require the extra acquisition of the magnitude and phase of
+been developed and generically named as \emph{\ac{epi}-unwarp} techniques
+\cite{holland_efficient_2010,hsu_correction_2009,jezzard_characterization_2005,
+reber_correction_2005}. They usually 
+require the extra acquisition of the magnitude and phase of
 the field (field-mapping), condition which is not always met. Some other 
 methodologies do not make use the field-mapping, compensating the distortion
-with non-linear registration from structural \ac{mri}.\\
+with non-linear registration from structural \ac{mri} or other means
+\citep{andersson_modeling_2001}. To our knowledge, there exists no study
+of the impact of the \ac{epi} distortion on the variability of tractography
+results. \\
 
@@
 solve the segmentation and distortion challenges, by exploiting as strong 
-shape-prior the detailed anatomy extracted from anatomical \ac{mri}. 
+shape-prior the detailed anatomy extracted from anatomical \ac{mri}. This
+is justified by the strong relationship between both problems, and the
+advantage of a significant increase of coherence among these steps in
+the complete processing pipeline.
 We reformulate the segmentation problem as an inverse problem, where

### 2012-12-12 Publications/2013-IPMI/Latex/04-Methods.tex
Initial draft for the conference. Close #25

@@
 %
-As suggested in \autoref{sec:introduction}, the general situation 
-consists of having reliable segmentations on the \ac{t1} reference
-space, obtained with 
-\emph{FreeSurfer\footnote{\url{http://surfer.nmr.mgh.harvard.edu/}}}.
-Therefore, regarding the proposed solution, we will have a precise
-location of the tissue interfaces of interest in a reference space.
-On the other hand, we have a \ac{dwi} volume, characterized by its
+As described in \autoref{sec:introduction}, the general situation in
+the connectivity pipelines consists on having 
+a reliable segmentation obtained from the high resolution \ac{t1} 
+reference image. Therefore, a precise location of the tissue interfaces
+of interest is available in a reference space. Given that there is no 
+interest on the anatomical reference segmentation,
+we directly obtained the shape priors from the models. \\
+
+On the other hand, the target \ac{dwi} data is characterized by its
 low resolution (typically around $2.2x2.2x3mm^3$). Depending on the
@@
 extract the information in a manageable manner. Particularly, we
-will use the \ac{fa} and \ac{md} maps. Whereas \ac{fa} describes
-the \emph{shape} of diffusion, the \ac{md} depicts the
-\emph{manitude} of the process. There exist to main reasons to 
-justify their choice. First, they are well-understood and
-standardized in clinical routine. Second, they are statistically 
-orthogonal and together contain most of the information that is
+will use the \ac{fa} and \ac{md} maps for convenience.
+Whereas \ac{fa} describes the \emph{shape} of diffusion, 
+the \ac{md} depicts the \emph{manitude} of the process. 
+There exist two main reasons to justify their choice. 
+First, they are well-understood and standardized in clinical routine.
+Second, together they contain most of the information that is
 usually extracted from the \ac{dwi}-derived scalar maps. \\
+
 In order that demonstrating the functionality of the proposed
 methodology and characterize its possibilities, we developed two
-synthetic model, generating the \ac{dwi} data as described in
- \citep{tuch_q-ball_2004}. We selected 30 directions, for being
-a very common protocol for \ac{dti} reconstruction. The first model
-is a set of spherical shapes representing the different brain
-tissues. The second model is based on the BrainWeb dataset. We
-reconstructed the \ac{dwi} data with standard to approximate the
-environment to the real one at maximum. There is no interest on the
-anatomical reference, given that with the models we hold \emph{a priori}
-precisely located surface of the interfaces of interest. \\
-
-FIGURE OF THE MODELS AND EXTRACTED FA, MD
+synthetic models and simulated their corresponding \ac{dwi}
+raw signal as described in \citep{tuch_q-ball_2004}. 
+(HERE WE NEED A GOOD DESCRIPTION OF THE DATA, directions, resolution, etc)
+. The first model consists of several spherical shapes emulating
+the different brain tissues (see \autoref{fig:fa}, first row). 
+The second model is based on the BrainWeb dataset.
+We reconstructed the \ac{dwi} signal with standard procedures to approximate the
+environment to the real one at maximum. \\
 
@@
 $\Delta E(f(c_i')) = E_{out}(f(c_i')) - E_{in}(f(c_i'))$, 
-and $E_R(f(c_i')) = (f(c_i')-\mu_R)^T\Sigma_R^{-1}(f(c_i')-\mu_R)$.
+and $E_R(f) = \sqrt{(f-\mu_R)^T\Sigma_R^{-1}(f-\mu_R)}$.
 We applied a spectral approach to solve this implicit scheme:
@@
 data. Once the signal was deformed, we proceeded to reconstruct the
-\ac{dti} and subsequently obtained the scalars of interest (\ac{fa}, \ac{md}).\\
+\ac{dti} and subsequently obtained the scalars of interest (\ac{fa}, \ac{md}) and estimated their parameters on the model.\\
 
 We evaluate the performance of our methodology to estimate the deformation
-field, obtaining a precise segmentation on the diffusion space.
+field and compare it to the original synthetic deformation field applied to the data.

### 2012-12-12 Publications/2013-IPMI/Latex/05-Results.tex
Initial draft for the conference. Close #25

@@
 %
+The proposed method successfully reverted the synthetic distortion
+field we applied to the data. With 16x16x16 control points, the
+displacements field is dense enough to correctly represent the
+synthetic field. (INCLUDE FIGURE). Figure XX shows the restored
+image and a difference map with the original model (we can also
+compute Dice indexes and that stuff).\\
+%
+
+\begin{figure}
+\begin{tabular}{ccccc}
+\includegraphics[width=0.19\textwidth]{fig_modelFA_00} &
+\includegraphics[width=0.19\textwidth]{fig_modelFA_01} &
+\includegraphics[width=0.19\textwidth]{fig_modelFA_02} &
+\includegraphics[width=0.19\textwidth]{fig_modelFA_03} &
+\includegraphics[width=0.19\textwidth]{fig_modelFA_04} \\
+\includegraphics[width=0.19\textwidth]{fig_contours_00} &
+\includegraphics[width=0.19\textwidth]{fig_contours_01} &
+\includegraphics[width=0.19\textwidth]{fig_contours_02} &
+\includegraphics[width=0.19\textwidth]{fig_contours_03} &
+\includegraphics[width=0.19\textwidth]{fig_contours_04}
+\end{tabular}
+\caption{First row presents several slices along Z axis of the \ac{fa} map obtained 
+after \ac{dti} fitting (using the original signal without simulated distortion) and
+the corresponding \ac{wm}-\ac{gm} contour. Second row contains the warped \ac{fa}
+map (using the distorted signal). The red contour corresponds to the interface
+detected after registration. In blue color, the initial \ac{wm}-\ac{gm} contour
+(the same as in previous row).}
+\label{fig:fa}
+\end{figure}

### 2012-12-12 Publications/2013-IPMI/Latex/06-Conclusion.tex
Initial draft for the conference. Close #25

\section{Conclusion}
\label{sec:conclusion}
%
A novel application for the \ac{acwe} framework is proposed,
with the aim at recovering the displacement field underlying 
the \ac{epi} geometrical distortions. Exploiting the segmentation
properties of the \ac{acwe} and optimizing the displacement
field, we describe a registration-segmentation methodology that
simultaneously segmented and restored the distortion on 
\ac{dwi}-like synthetic data. Visual results and quantitative
results are provided. \\

Once proven the aptness of the methodology to the application
with simplistic synthetic data, in further studies we will 
cover the actual performance on real images and the benefits 
of overcoming the described challenges (segmentation and 
\ac{epi} distortion correction) in one single step.\\

We conclude stressing on the importance of tackling with
the numerous challenges that exist on the \ac{dwi} data 
processing in order to achieve reliable results on the
whole-brain connectivity analysis.

### 2012-12-12 Code/Modules/LevelSets/include/MahalanobisLevelSets.h
Created the ParametersType to help implement n-contours (#14)

@@ -99,10 +99,15 @@ public:
 	typedef itk::Matrix
 			< PixelValueType, Components, Components >     CovarianceType;
 
+	struct ParametersType {
+		MeanType mean[2];
+		CovarianceType iCovariance[2];
+	};
+
 	MeasureType GetValue() const;
 	DeformationFieldPointer GetLevelSetsMap( DeformationFieldType* levelSetMap);
 
-	void SetParameters( MeanType& mean, CovarianceType& cov, bool inside);
+	void SetParameters( ParametersType& params );
 
 	itkSetObjectMacro(ReferenceImage, ReferenceImageType);
 	itkGetConstObjectMacro(ReferenceImage, ReferenceImageType);
@@ -113,9 +118,8 @@ protected:
 
 	void PrintSelf( std::ostream& os, itk::Indent indent) const;
 
+	ParametersType m_Parameters;
 	ReferenceImageConstPointer m_ReferenceImage;
-	MeanType m_Mean[2];
-	CovarianceType m_InverseCovariance[2];
 
 
 private:

### 2012-12-12 Publications/2013-IPMI/Latex/00-Main.tex
fixed affiliation

@@
 \usepackage{nameref}
-\usepackage[breaklinks]{hyperref}
+\usepackage{hyperref}
 \usepackage[USenglish]{babel}
@@
 
-\usepackage{graphicx}
-
 %
@@
 %
-\graphicspath{{../Figures/}}
 
@@
 \author{Oscar Esteban\inst{1,2} \and Alessandro Daducci \inst{2} \and
-		Meritxell Bach-Cuadra \inst{2} \and Virginia Estellers \inst{2}
-		\and Mar\'ia-Jes\'us Ledesma-Carbayo \and Luminita Vese \inst{3} \and 
         Jean-Philippe Thiran \inst{2} \and Andr\'es Santos \inst{1} \and
@@
 \and
-Signal Processing Laboratory 5 (LTS5),
+Signal Processing Laboratory (LTS5),
 \'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL)\\
-EPFL-STI-IEL-LTS5, Station 11, CH-1015 Lausanne, Switzerland
+EPFL STI IEL LTS5, Station 11, CH-1015 Lausanne, Switzerland
 \and
-Department of Mathematics, ... 
-University of California Los Angeles (UCLA),
-Los Angeles,..
+Department of Mathematics, University of California, Los Angeles (UCLA)\\
+520 Portola Plaza, Box 951555, Los Angeles, CA 90095-1555, U.S.A.
 }

### 2012-12-12 Publications/2013-IPMI/Latex/00-Main.tex
added acknowledgments
but masked in main.tex for review. Please complete your own grants, and
maybe add Meri if apropriate (i.e. if removed from author list).
Fixed affiliations. Close #37, Close #38

--- 2012-12-12 Publications/2013-IPMI/Latex/00-Main.tex
@@
 %
-\author{Oscar Esteban\inst{1,2} \and Alessandro Daducci \inst{2} \and
-		Meritxell Bach-Cuadra \inst{2} \and Virginia Estellers \inst{2}
-		\and Mar\'ia-Jes\'us Ledesma-Carbayo \and Luminita Vese \inst{3} \and 
-        Jean-Philippe Thiran \inst{2} \and Andr\'es Santos \inst{1} \and
-        Dominique Zosso \inst{2,3}}
+\author{Oscar Esteban\inst{1,2} \and
+        Alessandro Daducci \inst{2} \and
+		Meritxell Bach-Cuadra \inst{2,4} \and
+		Mar\'ia-Jes\'us Ledesma-Carbayo \inst{1} \and 
+        Jean-Philippe Thiran \inst{2,4} \and
+        Andr\'es Santos \inst{1} \and
+        Dominique Zosso \inst{3,2}}
 %
@@
 Av. Complutense 30, E-28040 Madrid, Spain\\
-\email{oscar.esteban@upm.es},\\
+\email{phd@oscaresteban.es},\\
 % WWW home page:
@@
 \and
-Signal Processing Laboratory 5 (LTS5),
+Signal Processing Laboratory (LTS5),
 \'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL)\\
@@
 \and
-Department of Mathematics, ... 
-University of California Los Angeles (UCLA),
-Los Angeles,..
+Department of Mathematics, 
+University of California, Los Angeles (UCLA) \\
+520 Portola Plaza, Box 951555, Los Angeles, CA 90095-1555, USA
+\and
+Department of Radiology, University Hospital Center (CHUV) and University of Lausanne (UNIL) \\
+Rue du Bugnon 46, CH-1011 Lausanne, Switzerland
 }

### 2012-12-12 Code/Modules/Core/include/SparseToDenseFieldResampleFilter.h
Proof of concept of N-Contours registration Close #14

@@ -71,6 +71,7 @@ public:
 	typedef typename Superclass::InputMeshType             InputMeshType;
 	typedef typename Superclass::InputMeshPointer          InputMeshPointer;
 	typedef typename Superclass::InputMeshConstPointer     InputMeshConstPointer;
+	typedef typename Superclass::MeshPointType             MeshPointType;
 	typedef typename Superclass::OutputImageType           OutputImageType;
 	typedef typename Superclass::OutputImagePointer        OutputImagePointer;
 	typedef typename Superclass::OutputImageConstPointer   OutputImageConstPointer;
@@ -93,23 +94,17 @@ public:
 			< TInputMesh, TOutputImage >                   InterpolatorType;
 	typedef typename InterpolatorType::Pointer             InterpolatorPointerType;
 
+	typedef typename std::vector<MeshPointType>      ControlPointList;
+
 
 	itkStaticConstMacro( OutputImageDimension, unsigned int, TOutputImage::ImageDimension );
 
-
-	virtual void SetInput( const InputMeshType *input ) {
-		this->SetInput( 0, input );
-	}
-	virtual void SetInput( unsigned int, const InputMeshType *input) {
-		this->m_Mesh = input;
-	}
-
-	const InputMeshType*   GetInput(void) { return this->GetInput(0); }
-	const InputMeshType*   GetInput(unsigned int idx) { return this->m_Mesh; }
 	//const OutputImageType* GetOutput( void ) { return this->GetOutput(0); }
 
-	itkSetConstObjectMacro( ShapePrior, InputMeshType );
-	itkGetConstObjectMacro( ShapePrior, InputMeshType );
+	void AddControlPoints( const InputMeshType* prior );
+	itkGetMacro( ControlPoints, ControlPointList );
+	//itkSetConstObjectMacro( ShapePrior, InputMeshType );
+	//itkGetConstObjectMacro( ShapePrior, InputMeshType );
 
 	/** Set the interpolator function.  The default is
 	 * itk::VectorLinearInterpolateImageFunction<InputImageType, TInterpolatorPrecisionType>.  */
@@ -188,7 +183,8 @@ private:
 	OutputIndexType                 m_OutputStartIndex;
 	OutputImageSizeType             m_OutputSize;
 
-	InputMeshConstPointer           m_ShapePrior;
+	//InputMeshConstPointer           m_ShapePrior;
+	ControlPointList                m_ControlPoints;
 
[…]

### 2012-12-12 Code/Modules/LevelSets/include/LevelSetsBase.h

@@ -105,6 +105,10 @@ public:
 	typedef typename ContourDeformationType::ConstPointer    ContourDeformationConstPointer;
 	typedef typename ContourDeformationType
 			                     ::PointDataContainerPointer PointDataContainerPointer;
+
+
+	typedef typename std::vector<ContourDeformationPointer>  ContourDeformationList;
+
 	typedef typename itk::QuadEdgeMeshToQuadEdgeMeshFilter
 			<ContourDeformationType,ContourDeformationType>  ContourCopyType;
 	typedef typename ContourCopyType::Pointer                ContourCopyPointer;
@@ -133,10 +137,10 @@ public:
 
 	void UpdateDeformationField( const DeformationFieldType* newField );
 
-	void SetShapePrior( ContourDeformationType* prior );
-	itkGetConstObjectMacro(ShapePrior, ContourDeformationType);
+	void AddShapePrior( ContourDeformationType* prior );
+	itkGetMacro(ShapePrior, ContourDeformationList);
 
-	itkGetConstObjectMacro(CurrentContourPosition, ContourDeformationType);
+	itkGetMacro(CurrentContourPosition, ContourDeformationList);
 
 	itkSetObjectMacro(DeformationField, DeformationFieldType);
 	itkGetConstObjectMacro(DeformationField, DeformationFieldType);
@@ -153,8 +157,8 @@ protected:
 
 	mutable MeasureType m_Value;
 	DeformationFieldPointer m_DeformationField;
-	ContourDeformationPointer m_ShapePrior;
-	ContourDeformationPointer m_CurrentContourPosition;
+	ContourDeformationList m_ShapePrior;
+	ContourDeformationList m_CurrentContourPosition;
 	ContourCopyPointer m_ContourCopier;
 	WarpContourPointer m_ContourUpdater;
 	SparseToDenseFieldResamplePointer m_SparseToDenseResampler;

### 2012-12-12 Code/Modules/LevelSets/include/MahalanobisLevelSets.h

@@ -104,10 +104,14 @@ public:
 		CovarianceType iCovariance[2];
 	};
 
+	typedef typename std::vector< ParametersType >         ParametersList;
+
 	MeasureType GetValue() const;
 	DeformationFieldPointer GetLevelSetsMap( DeformationFieldType* levelSetMap);
 
-	void SetParameters( ParametersType& params );
+	void SetParameters( size_t contour_id, ParametersType& params );
+
+	void AddShapePrior( ContourDeformationType* prior, ParametersType&params );
 
 	itkSetObjectMacro(ReferenceImage, ReferenceImageType);
 	itkGetConstObjectMacro(ReferenceImage, ReferenceImageType);
@@ -118,9 +122,8 @@ protected:
 
 	void PrintSelf( std::ostream& os, itk::Indent indent) const;
 
-	ParametersType m_Parameters;
+	ParametersList m_Parameters;
 	ReferenceImageConstPointer m_ReferenceImage;
-
 
 private:
 	MahalanobisLevelSets( const Self &); // purposely not implemented

### 2012-12-12 Code/Modules/Core/include/MeshToImageFilter.h
Proof of concept of N-Contours registration Close #14

@@ -71,17 +71,31 @@ public:
 	typedef          TInputMesh                            InputMeshType;
 	typedef typename InputMeshType::Pointer                InputMeshPointer;
 	typedef typename InputMeshType::ConstPointer           InputMeshConstPointer;
+	typedef typename InputMeshType::PointType              MeshPointType;
 	typedef          TOutputImage                          OutputImageType;
 	typedef typename OutputImageType::Pointer              OutputImagePointer;
 	typedef typename OutputImageType::ConstPointer         OutputImageConstPointer;
 	typedef typename OutputImageType::SizeType             OutputImageSizeType;
 	typedef typename OutputImageType::PixelType            ValueType;
 
+	typedef typename std::vector< InputMeshConstPointer >  InputMeshList;
+
 	itkStaticConstMacro( OutputImageDimension, unsigned int, TOutputImage::ImageDimension );
 
 
-	void SetInput( const InputMeshType *input ) { m_Mesh = input; }
-	const InputMeshType* GetInput(void) { return m_Mesh; }
+	virtual void SetInput( const InputMeshType *input ) {
+		this->SetInput( 0, input );
+	}
+	virtual void SetInput( size_t idx, const InputMeshType *input) {
+		if ( idx >= m_Mesh.size() ) m_Mesh.resize( idx+1 );
+
+		this->m_Mesh[idx] = input;
+	}
+
+	const InputMeshType*   GetInput(void) { return this->GetInput(0); }
+	const InputMeshType*   GetInput(unsigned int idx) { return this->m_Mesh[idx]; }
+
+	size_t GetNumberOfIndexedInputs() { return m_Mesh.size(); }
 
 protected:
 	MeshToImageFilter(){};
@@ -90,7 +104,7 @@ protected:
 	//virtual void GenerateData();
 	//virtual void PrintSelf( std::ostream &os, Indent indent) const;
 
-	InputMeshConstPointer m_Mesh;
+	InputMeshList m_Mesh;
 
 private:
 	MeshToImageFilter( const Self& ); // purposely not implemented

### 2012-12-13 Publications/2013-IPMI/Latex/02-Abstract.tex
Abstract v. 2.0

--- 2012-11-21 Publications/2013-IPMI/Latex/02-Abstract.tex
@@
 \begin{abstract}
-The abstract should summarize the contents of the paper
-using at least 70 and at most 150 words. It will be set in 9-point
-font size and be inset 1.0 cm from the right and left margins.
-There will be two blank lines before and after the Abstract.
+Connectomics is the study of fiber connections in the brain.
+High-quality reconstructions of the connectome from diffusion weighted MR acquisitions require a precise delineation of pial and grey-matter-white-matter interfaces.
+Furthermore, to enable significant group studies, these surfaces must be located and parcellated consistently across subjects.
+Precise segmentation and parcellation are readily available on anatomical images, such as T1-weighted MRI.
+Diffusion weighted images (DWI), however, typically have drastically lower resolution and exhibit more important distortions.
+Here, we propose a joint segmentation-registration framework that exploits the detailed and consistent anatomy extracted from anatomical MRI as strong shape-prior for DW segmentation.
+We use the Mumford-Shah (or active contours without edges) model to look for a regular deformation field that optimally maps the shape prior on the multivariate features in diffusion space.
+Preliminary results on synthetic images and simluated DWI confirm the effectiveness of our approach.
 %\dots
-%\keywords{diffusion weighted imaging, connectomics, echo planar imaging, magnetic resonance, segmentation, registration, distortion correction}
+\keywords{diffusion weighted imaging, connectomics, echo planar imaging, magnetic resonance, segmentation, registration, distortion correction}
 \end{abstract}

### 2012-12-13 Publications/2013-IPMI/Latex/02-Abstract.tex
New Abstract (v.3.0)

--- 2012-11-21 Publications/2013-IPMI/Latex/02-Abstract.tex
@@
-\begin{abstract}
-The abstract should summarize the contents of the paper
-using at least 70 and at most 150 words. It will be set in 9-point
-font size and be inset 1.0 cm from the right and left margins.
-There will be two blank lines before and after the Abstract.
-%\dots
-%\keywords{diffusion weighted imaging, connectomics, echo planar imaging, magnetic resonance, segmentation, registration, distortion correction}
-\end{abstract}
+Whole-brain connectivity analysis on diffusion weighted MR images (DWI) involves
+a large set of processing techniques. Particularly, an accurate white matter (WM)
+segmentation is required, with precise location of the interface with grey matter.
+High-standard segmentation and surface parcellations are readily available on 
+anatomical images, such as T1-weighted MRI. However DWI typically have drastically 
+lower resolution, and generally exhibit important distortions. Here, we propose a 
+segmentation-registration framework that exploits the detailed anatomy extracted 
+from anatomical MRI as reference shape-prior for DWI segmentation. We use the 
+active contours without edges model to look for a deformation field that optimally 
+maps the shape prior on the multivariate features in diffusion space. This joint
+approach attends the intrinsic coupling of segmentation, parcellation, and distortion 
+correction problems to finally obtain reliable connectivity networks. Preliminary
+results on synthetic datasets and simulated DWI confirm the effectiveness of our 
+approach.

### 2012-12-13 Code/Modules/Optimizers/include/GradientDescentLevelSetsOptimizer.h
Fixed #41. Hopefully this stupid bug is the reason for not having good results. Thank you so much :)

@@ -148,6 +148,15 @@ public:
 //	itkGetConstReferenceMacro(ReturnBestParametersAndValue, bool);
 //	itkBooleanMacro(ReturnBestParametersAndValue);
 
+	itkSetMacro( Alpha, InternalComputationValueType );
+	itkGetConstMacro( Alpha, InternalComputationValueType );
+
+	itkSetMacro( Beta, InternalComputationValueType );
+	itkGetConstMacro( Beta, InternalComputationValueType );
+
+	itkSetMacro( StepSize, InternalComputationValueType );
+	itkGetConstMacro( StepSize, InternalComputationValueType );
+
 	/** Start and run the optimization */
 	virtual void Start();
 

### 2012-12-15 Data/ModelGeneration/Model2/strel_sphere.m
Added Dominique's "cortex" model

@@ -1,0 +1,6 @@
+function se = strel_sphere( r )
+% create 3D structuring element : sphere of radius r
+r = r-1;
+[x y z] = meshgrid( -r:r, -r:r, -r:r );
+se = x.^2 + y.^2 + z.^2 <= r.^2;
+end

### 2012-12-15 Publications/2013-IPMI/Latex/00-Main.tex
Merged Dominique's and mine commits.
New Abstract (v.3.0)
Add llncs template. Close #43

@@
 \bibliographystyle{splncs03}
-%\bibliographystyle{splncsnat} 
-%\bibliographystyle{IEEEtran}
 
+\usepackage{nameref}
+\usepackage{hyperref}
 \usepackage[hyphens]{url}
-\usepackage{nameref}
-\usepackage[breaklinks]{hyperref}
+\usepackage[hyphenbreaks]{breakurl}
 \usepackage[USenglish]{babel}
@@
 %
-
 \DeclareMathOperator{\tr}{tr}
@@
 \usepackage{graphicx}
-
 %
@@
 		Meritxell Bach-Cuadra \inst{2,4} \and
-		Mar\'ia-Jes\'us Ledesma-Carbayo \inst{1} \and 
         Jean-Philippe Thiran \inst{2,4} \and
@@
 
+\begin{abstract}
 \input{02-Abstract}
+\keywords{ diffusion weighted imaging, connectomics,
+           echo planar imaging, magnetic resonance, 
+           segmentation, registration, distortion correction}
+\end{abstract}
 
@@
 
+%\input{07-Acknowledgments}
+
 \bibliography{99-References}

### 2012-12-15 Publications/2013-IPMI/Latex/02-Abstract.tex
Add llncs template. Close #43

@@
-Whole-brain connectivity analysis on diffusion weighted MR images (DWI) involves
-a large set of processing techniques. Particularly, an accurate white matter (WM)
-segmentation is required, with precise location of the interface with grey matter.
-High-standard segmentation and surface parcellations are readily available on 
-anatomical images, such as T1-weighted MRI. However DWI typically have drastically 
-lower resolution, and generally exhibit important distortions. Here, we propose a 
-segmentation-registration framework that exploits the detailed anatomy extracted 
-from anatomical MRI as reference shape-prior for DWI segmentation. We use the 
-active contours without edges model to look for a deformation field that optimally 
-maps the shape prior on the multivariate features in diffusion space. This joint
-approach attends the intrinsic coupling of segmentation, parcellation, and distortion 
-correction problems to finally obtain reliable connectivity networks. Preliminary
-results on synthetic datasets and simulated DWI confirm the effectiveness of our 
-approach.
+In whole-brain connectivity analysis of diffusion weighted MR images (DWI),
+an accurate delineation of the white-matter and grey-matter surfaces is required.
+While high-standard segmentation is readily available for anatomical MRI, such 
+as T1-weighted, DWI typically have drastically lower resolution and strong 
+geometrical distortions. We propose a DWI segmentation-registration framework 
+that exploits the detailed anatomy extracted from anatomical MRI as shape-prior.
+We use an ``active contours without edges''-like model to look for a deformation
+field that optimally maps the shape prior on the multivariate features 
+in diffusion space. This joint approach reflects the intrinsic coupling of 
+segmentation and distortion correction. Complementary, a precise and consistent
+cortical parcellation on DWI is straightforward by projection from T1 space. Thus,
+we expect to improve the reliability and robustness of the resulting connectivity 
+networks and their comparability within and across subjects. Preliminary results 
+on synthetic datasets and simulated DWI confirm the effectiveness of our approach.

### 2012-12-15 Publications/2013-IPMI/Latex/045-Experiments.tex
rearranging paper

--- 2012-12-12 Publications/2013-IPMI/Latex/04-Methods.tex
@@
-\section{Methods}
-\label{sec:methods}
+\section{Data and experiments}
+\label{sec:experiments}
 %
@@
 
-\subsection{\acl{acwe}-like variational segmentation model}
 %
-Let us denote $\{c_i\}_{i=1..N_c}$ the nodes of a shape prior surface. In
-our application, a precise \ac{wm}-\ac{gm} interface extracted from a
-high-resolution reference volume. All the formulations can be naturally
-extended to include more shape priors. On the other hand, we have a 
-number of \ac{dwi}-derived features at each
-voxel of the volume. Let us denote by $x$ the voxel and 
-$f(x) = [ f_1, f_2, \ldots, f_N]^T(x)$ its associated feature vector.\\
-%
-The transformation from reference into \ac{dwi} coordinate space is 
-achieved through a dense deformation field $u(x)$, such that:
-%
-\begin{equation}
-c_i' = T\{c_i\} = c_i + u(c_i)
-\end{equation}
-% 
-Since the nodes of the anatomical surfaces might lay off-grid, it is 
-required to derive $u(x)$ from a discrete set of parameters $\{u_k\}_{k=1..K}$.
-Densification is achieved through a set of associated basis functions 
-$\Psi_k$ (e.g. rbf, interpolation splines):
-%
-\begin{equation}
-u(x) = \sum_k \Psi_k(x) u_k
-\end{equation}
-%
-Consequently, the transformation writes
-%
-\begin{equation}
-\label{eq:transformation}
-c_i' = T\{c_i\} = c_i + u(c_i) = c_i + \sum_k \Psi_k(c_i)u_k
-\end{equation} 
-%
-% Comment: maybe this is not for IPMI 2013.
-%Note that, since $c_i$ remains constant in the DW segmentation process,
-%the values of $\Psi_k(c_i)$ can be precomputed. Also, provided compact 
-%support of the basis functions, the system remains relatively sparse.\\
-%
-Based on the current estimate of the distortion $u$, we can compute 
-``expected samples'' within the shape prior projected into the \ac{dwi}.
-Thus, we now estimate region descriptors of the \ac{dwi} features 
-$f(x)$ of the regions defined by the priors in \ac{dwi} space.
-%
-Using Gaussian distributions as region descriptors, we propose an
-\ac{acwe}-like, piece-wise constant, variational image segmentation
-model (where the unknown is the deformation field)
-\cite{chan_active_2001}:
-\begin{equation}
-\label{eq:gaussian_energy}
-E(u)= \sum_{\forall{R}} \int_{\Omega_R} (f-\mu_R)^T\Sigma_R^{-1}(f-\mu_R) dx
-\end{equation}
-where $R$ indexes the existing regions and the integral domains
-depend on the deformation field $u$. Note
-that minimizing this energy, $\argmin_u\{E\}$, yields the \ac{map} 
-estimate of a piece-wise smooth image model affected by Gaussian 
-additive noise. This inverse problem is ill-posed
-\cite{bertero_ill-posed_1988,hadamard_sur_1902}.
-In order to account for deformation field regularity and to render the 
-problem well-posed, we include limiting and regularization terms into 
-the energy functional \cite{morozov_linear_1975,tichonov_solution_1963}:
-%
-\begin{align}
-\label{eq:complete_energy}
-E(u) &= \sum_{\forall{R}} \lbrace \int_{\Omega_R} (f-\mu_R)^T\Sigma_R^{-1}(f-\mu_R) dx \rbrace \nonumber \\
-&\quad + \alpha \int  \|u\|^2 dx + \beta \int \left( \|\nabla u_x\|^2 + \|\nabla u_y\|^2 + \|\nabla u_z\|^2\right) dx
-\end{align}
-%
-These regularity terms ensure that the segmenting contours in 
-\ac{dwi} space are still close to their native shape. The model
-easily allows to incorporate inhomogeneous and anisotropic 
-regularization \cite{nagel_investigation_1986} to better regularize
-the \ac{epi} distortion. \\
-%
-
-At each iteration, we update the distortion along the steepest 
-energy descent. This gradient descent step can be efficiently 
-tackled by discretizing the time in a forward Euler scheme, 
-and making the right hand side semi-implicit in the 
-regularization terms:
-%
-\begin{align}
-\frac{u^{t+1}-u^t}{\tau} &= - \sum_{i=1}^{N_c} \left[ \Delta E(f(c_i'))  \hat{n}_{c_i'} \Psi_{c_i}(x) \right] -\alpha u^{t+1} + \beta\Delta u^{t+1}
-\end{align}
-%
-where the data terms remain functions of the current estimate 
-$u^t$, thus $c_i' = c_i'(u^t)$. For simplicity on notation, we 
-restricted the number of priors to only 1. We also defined 
-$\Delta E(f(c_i')) = E_{out}(f(c_i')) - E_{in}(f(c_i'))$, 
-and $E_R(f) = \sqrt{(f-\mu_R)^T\Sigma_R^{-1}(f-\mu_R)}$.
-We applied a spectral approach to solve this implicit scheme:
-%
-\begin{equation}
-u^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{u^t/\tau
-- \sum_{i=1}^{N_c} \left[ \Delta E(f(c_i')) \hat{n}_{c_i'} \Psi_{c_i}(x) \right]  \}}{\mathcal{F}\{(1/\tau+\alpha)\mathcal{I}-\beta\Delta\}} \right\}
-\end{equation}
-%
-
 \subsection{Experiment}

### 2012-12-15 Publications/2013-IPMI/Latex/00-Main.tex
rearranging and fighting with commits

--- 2012-12-12 Publications/2013-IPMI/Latex/00-Main.tex
@@
 \usepackage{nameref}
-\usepackage{hyperref}
+\usepackage[breaklinks]{hyperref}
 \usepackage[USenglish]{babel}
@@
 
+\usepackage{graphicx}
+
 %
@@
 %
+\graphicspath{{../Figures/}}
 
@@
 %%
-\titlerunning{Segmentation & EPI correction on DWI}  % abbreviated title (for running head)
+\titlerunning{Simultaneous segmentation and distortion correction on DWI}  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 %
-\author{Oscar Esteban\inst{1,2} \and Alessandro Daducci \inst{2} \and
-        Jean-Philippe Thiran \inst{2} \and Andr\'es Santos \inst{1} \and
-        Dominique Zosso \inst{2,3}}
+\author{Oscar Esteban\inst{1,2} \and
+        Alessandro Daducci \inst{2} \and
+		Meritxell Bach-Cuadra \inst{2,3} \and
+		Mar\'ia-Jes\'us Ledesma-Carbayo \inst{1} \and 
+        Jean-Philippe Thiran \inst{2,3} \and
+        Andr\'es Santos \inst{1} \and
+        Dominique Zosso \inst{2,4}}
 %
@@
 \and
-Department of Mathematics, University of California, Los Angeles (UCLA)\\
-520 Portola Plaza, Box 951555, Los Angeles, CA 90095-1555, U.S.A.
+Department of Radiology, University Hospital Center (CHUV) and University of Lausanne (UNIL) \\
+Rue du Bugnon 46, CH-1011 Lausanne, Switzerland
+\and
+Department of Mathematics, 
+University of California, Los Angeles (UCLA) \\
+520 Portola Plaza, Box 951555, Los Angeles, CA 90095-1555, USA
 }
@@
 
+\begin{abstract}
 \input{02-Abstract}
+\keywords{ magnetic resonance, diffusion weighted imaging, distortion correction, segmentation, registration, shape priors, connectomics, echo planar imaging
+           }
+\end{abstract}
 
@@
 
+\input{045-Experiments}
+
 \input{05-Results}

### 2012-12-15 Publications/2013-IPMI/Latex/04-Methods.tex
rearranging and fighting with commits

@@
 %
-\subsection{Simulated datasets}
-%
-As described in \autoref{sec:introduction}, the general situation in
-the connectivity pipelines consists on having 
-a reliable segmentation obtained from the high resolution \ac{t1} 
-reference image. Therefore, a precise location of the tissue interfaces
-of interest is available in a reference space. Given that there is no 
-interest on the anatomical reference segmentation,
-we directly obtained the shape priors from the models. \\
-
-On the other hand, the target \ac{dwi} data is characterized by its
-low resolution (typically around $2.2x2.2x3mm^3$). Depending on the
-posterior reconstruction methodology and the angular resolution
-intended, the \ac{dwi} raw data has to be processed in order to
-extract the information in a manageable manner. Particularly, we
-will use the \ac{fa} and \ac{md} maps for convenience.
-Whereas \ac{fa} describes the \emph{shape} of diffusion, 
-the \ac{md} depicts the \emph{manitude} of the process. 
-There exist two main reasons to justify their choice. 
-First, they are well-understood and standardized in clinical routine.
-Second, together they contain most of the information that is
-usually extracted from the \ac{dwi}-derived scalar maps. \\
-
-In order that demonstrating the functionality of the proposed
-methodology and characterize its possibilities, we developed two
-synthetic models and simulated their corresponding \ac{dwi}
-raw signal as described in \citep{tuch_q-ball_2004}. 
-(HERE WE NEED A GOOD DESCRIPTION OF THE DATA, directions, resolution, etc)
-. The first model consists of several spherical shapes emulating
-the different brain tissues (see \autoref{fig:fa}, first row). 
-The second model is based on the BrainWeb dataset.
-We reconstructed the \ac{dwi} signal with standard procedures to approximate the
-environment to the real one at maximum. \\
-
 \subsection{\acl{acwe}-like variational segmentation model}
@@
 %
-
-\subsection{Experiment}
-%
-For both models, we created manually a sound distortion visually similar
-to real \ac{epi} distortions. We interpolated the distortion to a 
-dense deformation field, necessary for warping the raw \ac{dwi} simulated
-data. Once the signal was deformed, we proceeded to reconstruct the
-\ac{dti} and subsequently obtained the scalars of interest (\ac{fa}, \ac{md}) and estimated their parameters on the model.\\
-
-We evaluate the performance of our methodology to estimate the deformation
-field and compare it to the original synthetic deformation field applied to the data.

### 2012-12-15 Publications/2013-IPMI/Latex/00-Main.tex
fixing affiliations and merge

@@
 		Meritxell Bach-Cuadra \inst{2,3} \and
-		Mar\'ia-Jes\'us Ledesma-Carbayo \inst{1} \and 
         Jean-Philippe Thiran \inst{2,3} \and

### 2012-12-17 Publications/2013-IPMI/Latex/045-Experiments.tex
reordering / adding experiments
work in progress

@@
 %
-\subsection{Simulated datasets}
+\subsection{Shape prior}
 %
@@
 reference image. Therefore, a precise location of the tissue interfaces
-of interest is available in a reference space. Given that there is no 
-interest on the anatomical reference segmentation,
-we directly obtained the shape priors from the models. \\
+of interest is available in a reference space. Given that the anatomical reference segmentation is beyond the scope of this manuscript, we simply rely on the shape priors obtained from the models. \\
 
-On the other hand, the target \ac{dwi} data is characterized by its
+%
+\subsection{Synthetic gray-scale data}
+%
+The first, toy example is inspired by a problem shown for coupled CSF/GM and GM/WM segmentation in \cite{MacDonald2000}.
+\textbf{FIXME description of the model, expected difficulties (PVE -- lack of contour in sulcus), and how proposed model is expected to cope with this}.
+
+%
+\subsection{Simulated diffusion data}
+%
+In order to demonstrate the specific functionality of the proposed
+methodology and characterize its possibilities with diffusion data, we developed a
+synthetic model and simulated its corresponding \ac{dwi}
+raw signal as described in \citep{tuch_q-ball_2004}.
+
+This second model consists of several spherical shapes emulating
+the different brain tissues (see \autoref{fig:fa}, first row). 
+We reconstructed the \ac{dwi} signal with standard procedures to approximate the
+environment to the real one at maximum. \\
+
+The target \ac{dwi} data is characterized by its
 low resolution (typically around $2.2x2.2x3mm^3$). Depending on the
@@
 intended, the \ac{dwi} raw data has to be processed in order to
-extract the information in a manageable manner. Particularly, we
+extract the information in a manageable manner. Here, we
 will use the \ac{fa} and \ac{md} maps for convenience.
 Whereas \ac{fa} describes the \emph{shape} of diffusion, 
-the \ac{md} depicts the \emph{manitude} of the process. 
+the \ac{md} depicts the \emph{magnitude} of the process. 
 There exist two main reasons to justify their choice. 
@@
 
-In order that demonstrating the functionality of the proposed
-methodology and characterize its possibilities, we developed two
-synthetic models and simulated their corresponding \ac{dwi}
-raw signal as described in \citep{tuch_q-ball_2004}. 
-(HERE WE NEED A GOOD DESCRIPTION OF THE DATA, directions, resolution, etc)
-. The first model consists of several spherical shapes emulating
-the different brain tissues (see \autoref{fig:fa}, first row). 
-The second model is based on the BrainWeb dataset.
-We reconstructed the \ac{dwi} signal with standard procedures to approximate the
-environment to the real one at maximum. \\
-
-%
-\subsection{Experiment}
-%
-For both models, we created manually a sound distortion visually similar
+For this model, we created manually a sound distortion visually similar
 to real \ac{epi} distortions. We interpolated the distortion to a 
@@
 data. Once the signal was deformed, we proceeded to reconstruct the
-\ac{dti} and subsequently obtained the scalars of interest (\ac{fa}, \ac{md}) and estimated their parameters on the model.\\
+\ac{dti} \textbf{FIXME is it really DTI, here, not rather DSI?} and subsequently obtained the scalars of interest (\ac{fa}, \ac{md}) and estimated their parameters on the model.
 
-We evaluate the performance of our methodology to estimate the deformation
-field and compare it to the original synthetic deformation field applied to the data.
+\textbf{FIXME Give parameters here?}\\

### 2012-12-18 Publications/2013-IPMI/Latex/02-Abstract.tex
Fixed some parts of latex

@@
 While high-standard segmentation is readily available for anatomical MRI, such 
-as T1-weighted, DWI typically have drastically lower resolution and strong 
+as T1-weighted, DWI typically have drastically lower resolution and severe 
 geometrical distortions. We propose a DWI segmentation-registration framework 

### 2012-12-18 Publications/2013-IPMI/Latex/03-Introduction.tex
Fixed some parts of latex

@@
 requirement is usually solved in practice by plainly thresholding the 
-\ac{fa} (a well-know scalar map derived from \ac{dwi} which depicts 
-the isotropy of water diffusion inside the brain). 
+\ac{fa}, a well-know scalar map derived from \ac{dwi} which depicts 
+the isotropy of water diffusion inside the brain. 
 Additionally, a precise location of the
@@
 acquisition techniques, that often suffer from severe distortions due to 
-local field inhomogeneities. Generally, it is easily appreciated in the anterior
+local field inhomogeneities. Generally, it is appreciated in the anterior
 part of the brain, along the phase-encoded direction. Some methodologies have
@@
 solve the segmentation and distortion challenges, by exploiting as strong 
-shape-prior the detailed anatomy extracted from anatomical \ac{mri}. This
-is justified by the strong relationship between both problems, and the
-advantage of a significant increase of coherence among these steps in
-the complete processing pipeline.
+shape-prior the detailed anatomy extracted from anatomical \ac{mri}. On
+one hand, the significant benefits of exploiting the anatomical \ac{mri}
+have been demonstrated \cite{zollei_improved_2010}, justifying the use
+of the shape prior information. On the other, the
+significant coupling between both problems demands for consistent
+segmentation, parcellation, and distortion. This need is satisfied by
+the use of a unique underlying model of our joint approach.
 We reformulate the segmentation problem as an inverse problem, where

### 2012-12-18 Publications/2013-IPMI/Latex/045-Experiments.tex
Fixed some parts of latex

@@
 As described in \autoref{sec:introduction}, the general situation in
-the connectivity pipelines consists on having 
+the connectivity pipelines consists of having 
 a reliable segmentation obtained from the high resolution \ac{t1} 
 reference image. Therefore, a precise location of the tissue interfaces
-of interest is available in a reference space. Given that the anatomical reference segmentation is beyond the scope of this manuscript, we simply rely on the shape priors obtained from the models. \\
+of interest is available in a reference space. Given that the anatomical 
+reference segmentation is beyond the scope of this manuscript, we simply 
+rely on the shape priors obtained from the models.
 
@@
 %
-The first, toy example is inspired by a problem shown for coupled CSF/GM and GM/WM segmentation in \cite{MacDonald2000}.
-\textbf{FIXME description of the model, expected difficulties (PVE -- lack of contour in sulcus), and how proposed model is expected to cope with this}.
+The first, toy example is inspired by a problem shown for coupled CSF/GM 
+and GM/WM segmentation in \cite{MacDonald2000}. \textbf{FIXME description of the model, expected difficulties (PVE -- lack of contour in sulcus), and how proposed model is expected to cope with this}.
 
@@
 the different brain tissues (see \autoref{fig:fa}, first row). 
-We reconstructed the \ac{dwi} signal with standard procedures to approximate the
-environment to the real one at maximum. \\
+We reconstructed the \ac{dwi} signal with standard procedures to 
+approximate the environment to the real one at maximum. \\
 
-The target \ac{dwi} data is characterized by its
+The target \ac{dwi} data is characterized by its distortions and its
 low resolution (typically around $2.2x2.2x3mm^3$). Depending on the
@@
 intended, the \ac{dwi} raw data has to be processed in order to
-extract the information in a manageable manner. Here, we
-will use the \ac{fa} and \ac{md} maps for convenience.
+extract the information in a manageable manner. The properties of
+the reconstructed tensors and derived scalar maps have been
+studied and presented on \cite{ennis_orthogonal_2006}. Based on their
+findings, the proposed energy model adapts to the \ac{fa} and \ac{md}
+for their properties.
 Whereas \ac{fa} describes the \emph{shape} of diffusion, 
@@
 data. Once the signal was deformed, we proceeded to reconstruct the
-\ac{dti} \textbf{FIXME is it really DTI, here, not rather DSI?} and subsequently obtained the scalars of interest (\ac{fa}, \ac{md}) and estimated their parameters on the model.
+\ac{dti} and subsequently obtained the scalars of interest (\ac{fa}, \ac{md}).
+Finally, we estimated their parameters using the tissue probability
+distribution maps from the original model (\autoref{table:parameters}).
 
-\textbf{FIXME Give parameters here?}\\
+\begin{table}
+\begin{tabular}{cccc}
+         & $\mathbf{\mu}_{FA}$ & $\mathbf{\mu}_{MD}$ & $\mathbf{\Sigma}$ \\
+\ac{wm}  & & & \\
+\ac{gm}  & & & \\
+\ac{csf} & & & \\
+\end{tabular}
+\caption{Model parameters}
+\label{table:parameters}
+\end{table}

### 2012-12-18 Publications/2013-IPMI/Latex/045-Experiments.tex
working on model description

--- 2012-12-17 Publications/2013-IPMI/Latex/045-Experiments.tex
@@
 The first, toy example is inspired by a problem shown for coupled CSF/GM and GM/WM segmentation in \cite{MacDonald2000}.
-\textbf{FIXME description of the model, expected difficulties (PVE -- lack of contour in sulcus), and how proposed model is expected to cope with this}.
+The authors note that ``partial volume effects blur the distinction between closely adjacent surfaces in deep sulci, leading to a well-known segmentation error in which the deeper reaches of sulci are not penetrated by the putative surface model.'' This problem is aggravated in DWI, since the resolution tends to be worse compared to the anatomical images considered in \cite{MacDonald2000}.
+They test their coupled segmentation algorithm on an image, ``representing a sulcus in which the distinction between opposing banks of the sulcus has been obscured by partial volume.''  As illustrated, conventional single surface segmentation of the CSF/GM boundary misses to capture the sulcus in its full depth.
+With our proposed model, we expect the joint segmentation-registration to be driven largely by the inner, GM/WM contour that exhibits sufficient contrast and lesser partial volume effects.
 

### 2012-12-18 Publications/2013-IPMI/Latex/045-Experiments.tex

--- 2012-12-18 Publications/2013-IPMI/Latex/045-Experiments.tex
@@
 rely on the shape priors obtained from the models.
-
 %
@@
 %
-The first, toy example is inspired by a problem shown for coupled CSF/GM 
-and GM/WM segmentation in \cite{MacDonald2000}. \textbf{FIXME description of the model, expected difficulties (PVE -- lack of contour in sulcus), and how proposed model is expected to cope with this}.
-
+The first, toy example is inspired by a problem shown for coupled CSF/GM and GM/WM segmentation in \cite{MacDonald2000}.
+The authors note that ``partial volume effects blur the distinction between closely adjacent surfaces in deep sulci, leading to a well-known segmentation error in which the deeper reaches of sulci are not penetrated by the putative surface model.'' This problem is aggravated in DWI, since the resolution tends to be worse compared to the anatomical images considered in \cite{MacDonald2000}.
+They test their coupled segmentation algorithm on an image, ``representing a sulcus in which the distinction between opposing banks of the sulcus has been obscured by partial volume.''  As illustrated, conventional single surface segmentation of the CSF/GM boundary misses to capture the sulcus in its full depth.
+With our proposed model, we expect the joint segmentation-registration to be driven largely by the inner, GM/WM contour that exhibits sufficient contrast and lesser partial volume effects.
 %

### 2012-12-18 Publications/2013-IPMI/Latex/045-Experiments.tex
Updated latex and model parameteres

@@
 %
-The first, toy example is inspired by a problem shown for coupled CSF/GM and GM/WM segmentation in \cite{MacDonald2000}.
-The authors note that ``partial volume effects blur the distinction between closely adjacent surfaces in deep sulci, leading to a well-known segmentation error in which the deeper reaches of sulci are not penetrated by the putative surface model.'' This problem is aggravated in DWI, since the resolution tends to be worse compared to the anatomical images considered in \cite{MacDonald2000}.
-They test their coupled segmentation algorithm on an image, ``representing a sulcus in which the distinction between opposing banks of the sulcus has been obscured by partial volume.''  As illustrated, conventional single surface segmentation of the CSF/GM boundary misses to capture the sulcus in its full depth.
-With our proposed model, we expect the joint segmentation-registration to be driven largely by the inner, GM/WM contour that exhibits sufficient contrast and lesser partial volume effects.
+The first, toy example is inspired by a problem shown for coupled CSF/GM and GM/WM 
+segmentation in \cite{macdonald_automated_2000}. The authors note that ``partial 
+volume effects blur the distinction between closely adjacent surfaces in deep sulci,
+leading to a well-known segmentation error in which the deeper reaches of sulci are 
+not penetrated by the putative surface model.'' This problem is aggravated in DWI, 
+since the resolution tends to be worse compared to the anatomical images considered 
+in \cite{macdonald_automated_2000}. They test their coupled segmentation algorithm 
+on an image, ``representing a sulcus in which the distinction between opposing banks 
+of the sulcus has been obscured by partial volume.''  As illustrated, conventional 
+single surface segmentation of the CSF/GM boundary misses to capture the sulcus in 
+its full depth. With our proposed model, we expect the joint segmentation-registration 
+to be driven largely by the inner, GM/WM contour that exhibits sufficient contrast 
+and lesser partial volume effects.
 %
@@
          & $\mathbf{\mu}_{FA}$ & $\mathbf{\mu}_{MD}$ & $\mathbf{\Sigma}$ \\
-\ac{wm}  & & & \\
-\ac{gm}  & & & \\
-\ac{csf} & & & \\
+\ac{wm}  & 0.778 & 6.94\e{-4} & 
+   $\begin{pmatrix}
+   	4.85\e{-3} & -6.90\e{-6} \\ -6.90\e{-6} & 1.03\e{-8}
+   \end{pmatrix}$
+\\
+\ac{gm}  & 0.119 & 8.95\e{-4} &  \\
+\ac{csf} & 0.103 & 2.99\e{-3} & \\
 \end{tabular}

### 2012-12-18 Publications/2013-IPMI/Latex/045-Experiments.tex
adding sulcus model pics etc.

--- 2012-12-18 Publications/2013-IPMI/Latex/045-Experiments.tex
@@
 %
-The first, toy example is inspired by a problem shown for coupled CSF/GM and GM/WM segmentation in \cite{MacDonald2000}.
-The authors note that ``partial volume effects blur the distinction between closely adjacent surfaces in deep sulci, leading to a well-known segmentation error in which the deeper reaches of sulci are not penetrated by the putative surface model.'' This problem is aggravated in DWI, since the resolution tends to be worse compared to the anatomical images considered in \cite{MacDonald2000}.
-They test their coupled segmentation algorithm on an image, ``representing a sulcus in which the distinction between opposing banks of the sulcus has been obscured by partial volume.''  As illustrated, conventional single surface segmentation of the CSF/GM boundary misses to capture the sulcus in its full depth.
+The first, toy example is inspired by a problem shown for coupled CSF/GM and GM/WM segmentation in \citep{MacDonald2000}.
+These authors note that ``partial volume effects blur the distinction between closely adjacent surfaces in deep sulci, leading to a well-known segmentation error in which the deeper reaches of sulci are not penetrated by the putative surface model.'' This problem is aggravated in DWI, since the resolution tends to be worse compared to the anatomical images considered in \citep{MacDonald2000}.
+They test their coupled segmentation algorithm on an image, ``representing a sulcus in which the distinction between opposing banks of the sulcus has been obscured by partial volume.''  Here, we have reproduced a very similar model.
+As illustrated in \autoref{fig:sulcusmodel}, conventional single surface segmentation of the CSF/GM boundary misses to capture the sulcus in its full depth.
 With our proposed model, we expect the joint segmentation-registration to be driven largely by the inner, GM/WM contour that exhibits sufficient contrast and lesser partial volume effects.
+The shape prior of the outer, difficult contour will then be co-aligned through the regularity of the estimated deformation field.
+
+\begin{figure}
+\begin{tabular}{ccccc}
+\includegraphics[width=0.19\textwidth]{contours_inner} &
+\includegraphics[width=0.19\textwidth]{contours_wrong} &
+\includegraphics[width=0.19\textwidth]{contours_both} &
+\includegraphics[width=0.19\textwidth]{pialsurf} &
+\includegraphics[width=0.19\textwidth]{gmwmsurf}
+\end{tabular}
+\caption{balbal}
+\label{fig:sulcusmodel}
+\end{figure}
 %

### 2012-12-18 Publications/2013-IPMI/Latex/045-Experiments.tex

@@
 %
-The first, toy example is inspired by a problem shown for coupled CSF/GM and GM/WM segmentation in \citep{MacDonald2000}.
-These authors note that ``partial volume effects blur the distinction between closely adjacent surfaces in deep sulci, leading to a well-known segmentation error in which the deeper reaches of sulci are not penetrated by the putative surface model.'' This problem is aggravated in DWI, since the resolution tends to be worse compared to the anatomical images considered in \citep{MacDonald2000}.
+The first, toy example is inspired by a problem shown for coupled CSF/GM and GM/WM segmentation in \citep{macdonald_automated_2000}.
+These authors note that ``partial volume effects blur the distinction between closely adjacent surfaces in deep sulci, leading to a well-known segmentation error in which the deeper reaches of sulci are not penetrated by the putative surface model.'' This problem is aggravated in DWI, since the resolution tends to be worse compared to the anatomical images considered in \citep{macdonald_automated_2000}.
 They test their coupled segmentation algorithm on an image, ``representing a sulcus in which the distinction between opposing banks of the sulcus has been obscured by partial volume.''  Here, we have reproduced a very similar model.
@@
 \end{figure}
+
 %

### 2012-12-18 Publications/2013-IPMI/Latex/045-Experiments.tex
added Alessandros lines

@@
 \begin{tabular}{ccccc}
-\includegraphics[width=0.19\textwidth]{contours_inner} &
-\includegraphics[width=0.19\textwidth]{contours_wrong} &
+\includegraphics[width=0.19\textwidth]{contours_wrong} & \includegraphics[width=0.19\textwidth]{contours_inner} &
 \includegraphics[width=0.19\textwidth]{contours_both} &
 \includegraphics[width=0.19\textwidth]{pialsurf} &
-\includegraphics[width=0.19\textwidth]{gmwmsurf}
+\includegraphics[width=0.19\textwidth]{gmwmsurf}\\
+a)&b)&c)&d)&e)
 \end{tabular}
-\caption{balbal}
+\caption{The gray-scale sulcus model.
+a) The apparent CSF/GM boundary is affected by partial volume in the sulcal cavity, and conventional segmentation is likely to miss it.
+b) The GM/WM interface here has consistently good contrast.
+c) Registering the two shape priors coupled through deformation field regularity is expected to guide the CSF/GM contour.
+d\&e) 3D view of the two shape priors.}
 \label{fig:sulcusmodel}
@@
 
+\subsubsection{Signal simulation}
+
+To numerically simulate the MRI signal attenuation when applying a diffusion gradient in a voxel with $N$ fiber populations we made use of the standard \emph{Multi-Tensor Model}~\cite{Tuch:2002aa}:
+%
+\begin{equation} \label{eqn:MultiTensor}
+S(\vec{q}) / S_{0} = \sum_{i=1}^{N} f_{i} \, \exp{ \left( -b \, \vec{q}^{T} \, \mathbf{D}_i \, \vec{q}\right) } + f_{iso} \, \exp{ \left( -b \, \mathbf{D}_{iso} \right) } ,
+\end{equation}
+%
+where $\vec{q} \in \mathbb{S}^2$ is the direction of the diffusion gradient applied, $b$ is the b-value accounting for its strength, $S_0 \equiv S(\vec{0})$ is the signal with no diffusion weighting, $f_i$ and $\mathbf{D}_i$ are respectively the volume fraction and the diffusion tensor characterizing the $i$-th fiber population.
+The diffusion tensors $\mathbf{D}_i$ and $\mathbf{D}_{iso}$ describe the diffusion processes of each fiber compartment and of contaminations from the CSF.
+In this work, these quantities have been taken from standard ranges typically observed in the human brain~\cite{Canales-Rodriguez:2009aa}.
+
+\subsubsection{Noise simulation}
+
+The diffusion MRI signal $S$ has been corrupted with \textit{Rician noise}~\cite{Gudbjartsson:1995aa} as follows:
+%
+\begin{equation}
+	\tilde{S} = \sqrt{ (S + \varepsilon_1)^2 + \varepsilon_2^2 }
+\end{equation}
+%
+where $\varepsilon_{1,2}$ are Gaussian distributed with zero mean and standard deviation $\sigma = S_0 / \mathit{SNR}$ and SNR is the signal-to-noise ratio (SNR) on the $S_0$ image.
+
+\subsubsection{Derived features}
+
 The target \ac{dwi} data is characterized by its distortions and its
@@
 the \ac{md} depicts the \emph{magnitude} of the process.
+
+\textbf{FIXME: can you give the respective formula, how you get from the above signal to FA/MD?}
+
 There exist two main reasons to justify their choice. 
@@
 usually extracted from the \ac{dwi}-derived scalar maps. \\
+
+\subsubsection{Simulated distortions}
 

### 2012-12-18 Publications/2013-IPMI/Latex/045-Experiments.tex
Inserted Alessandro\'s text on DWI model

--- 2012-12-18 Publications/2013-IPMI/Latex/045-Experiments.tex
@@
 %
-In order to demonstrate the specific functionality of the proposed
-methodology and characterize its possibilities with diffusion data, we developed a
-synthetic model and simulated its corresponding \ac{dwi}
-raw signal as described in \citep{tuch_q-ball_2004}.
-
-This second model consists of several spherical shapes emulating
+In order to demonstrate the functionality of the methodology, 
+and characterize its possibilities with diffusion data,
+we simulated the \acs{dw} signal of a synthetic phantom from a model
+consisting of several spherical shapes emulating
 the different brain tissues (see \autoref{fig:fa}, first row). 
-We reconstructed the \ac{dwi} signal with standard procedures to 
+We reconstructed the signal with standard procedures to 
 approximate the environment to the real one at maximum. \\
 
+\paragraph{Signal simulation}
+To numerically simulate the MRI signal attenuation when applying a diffusion 
+gradient in a voxel with $N$ fiber populations we made use of the standard 
+\emph{Multi-Tensor Model}~\cite{Tuch:2002aa}:
+%
+\begin{equation} 
+\label{eqn:MultiTensor}
+S(\vec{q}) / S_{0} = \sum_{i=1}^{N} f_{i} \, \exp{ \left( -b \, \vec{q}^{T} \, \mathbf{D}_i \, \vec{q}\right) } + f_{iso} \, \exp{ \left( -b \, \mathbf{D}_{iso} \right) } ,
+\end{equation}
+%
+where $\vec{q} \in \mathbb{S}^2$ is the direction of the diffusion gradient 
+applied, $b$ is the b-value accounting for its strength, $S_0 \equiv S(\vec{0})$ 
+is the signal with no diffusion weighting, $f_i$ and $\mathbf{D}_i$ are 
+respectively the volume fraction and the diffusion tensor characterizing the 
+$i$-th fiber population.
+
+The diffusion tensors $\mathbf{D}_i$ and $\mathbf{D}_{iso}$ describe the diffusion 
+processes of each fiber compartment and of contaminations from the CSF. In this work, 
+these quantities have been taken from standard ranges typically observed in the 
+human brain~\cite{Canales-Rodriguez:2009aa}.
+
+\paragraph{Noise simulation}
+The diffusion MRI signal $S$ has been corrupted with 
+\textit{Rician noise}~\cite{Gudbjartsson:1995aa} as follows:
+%
+\begin{equation}
+	\tilde{S} = \sqrt{ (S + \varepsilon_1)^2 + \varepsilon_2^2 }
+\end{equation}
+%
+where $\varepsilon_{1,2}$ are Gaussian distributed with zero mean
+and standard deviation $\sigma = S_0 / \mathit{SNR}$ and \acs{snr}
+is the \ac{snr} on the $S_0$ image.
+
+\paragraph{Extracted scalar features}
 The target \ac{dwi} data is characterized by its distortions and its
@@
 the reconstructed tensors and derived scalar maps have been
-studied and presented on \cite{ennis_orthogonal_2006}. Based on their
+studied by \cite{ennis_orthogonal_2006}. Based on their
 findings, the proposed energy model adapts to the \ac{fa} and \ac{md}
-for their properties.
-Whereas \ac{fa} describes the \emph{shape} of diffusion, 
+for their properties. Whereas \ac{fa} describes the \emph{shape} of diffusion, 
 the \ac{md} depicts the \emph{magnitude} of the process. 
@@
 \begin{table}
-\begin{tabular}{cccc}
-         & $\mathbf{\mu}_{FA}$ & $\mathbf{\mu}_{MD}$ & $\mathbf{\Sigma}$ \\
+\begin{tabularx}{1.0\textwidth}{c|XXX}
+\hline
+         & $\mathbf{\mu}_{FA}$ & $\mathbf{\mu}_{MD}$ & $\Sigma$ \\
+\hline
 \ac{wm}  & 0.778 & 6.94\e{-4} & 
@@
 \\
-\ac{gm}  & 0.119 & 8.95\e{-4} &  \\
-\ac{csf} & 0.103 & 2.99\e{-3} & \\
-\end{tabular}
+\hline
+\ac{gm}  & 0.119 & 8.95\e{-4} &
+   $\begin{pmatrix}
+   	5.90\e{-4} & -1.43\e{-6} \\ -1.43\e{-6} & 1.04\e{-8}
+   \end{pmatrix}$
+\\
+\hline
+\ac{csf} & 0.103 & 2.99\e{-3} &
+   $\begin{pmatrix}
+   	1.19\e{-3} & 2.22\e{-7} \\ 2.22\e{-7} & 1.56\e{-8}
+   \end{pmatrix}$
+\\
+\hline
+\end{tabularx}
 \caption{Model parameters}

### 2012-12-18 Publications/2013-IPMI/Latex/045-Experiments.tex

@@
 
-\paragraph{Extracted scalar features}
+\paragraph{Derived scalar features}
 The target \ac{dwi} data is characterized by its distortions and its
@@
 the \ac{md} depicts the \emph{magnitude} of the process.
+
+\textbf{FIXME: can you give the respective formula, how you get from the above signal to FA/MD?}
+\begin{equation}
+FA = \sqrt{ \frac{3}{2}},\frac{\sqrt{ (\lambda_1 - \hat{\lamda})^2 + (\lambda_2 - \hat{\lamda})^2 + (\lambda_3 - \hat{\lamda})^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}}
+\end{equation}
+
 There exist two main reasons to justify their choice. 

### 2012-12-19 Publications/2013-IPMI/Latex/045-Experiments.tex
Added FA, MD formulae

@@
 studied by \cite{ennis_orthogonal_2006}. Based on their
-findings, the proposed energy model adapts to the \ac{fa} and \ac{md}
-for their properties. Whereas \ac{fa} describes the \emph{shape} of diffusion, 
+findings, the proposed energy model adapts to the \ac{fa}~\eqref{eq:fa} 
+and \ac{md}~\eqref{eq:md} for their properties. 
+Whereas \ac{fa} \ describes the \emph{shape} of diffusion, 
 the \ac{md} depicts the \emph{magnitude} of the process.
 
-\textbf{FIXME: can you give the respective formula, how you get from the above signal to FA/MD?}
-\begin{equation}
-FA = \sqrt{ \frac{3}{2}},\frac{\sqrt{ (\lambda_1 - \hat{\lamda})^2 + (\lambda_2 - \hat{\lamda})^2 + (\lambda_3 - \hat{\lamda})^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}}
-\end{equation}
+\begin{align}
+\mathrm{FA} &= \sqrt{ \frac{1}{2}}\,\frac{\sqrt{ (\lambda_1 - \lambda_2)^2 + (\lambda_2 - \lambda_3)^2 + (\lambda_3 - \lambda_1)^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}} \label{eq:fa} \\
+\mathrm{MD} &= ( \lambda_1 + \lambda_2 + \lambda_3 ) / 3 \label{eq:md}
+\end{align}
 

### 2012-12-19 Publications/2013-IPMI/Latex/045-Experiments.tex
adding new model1 results -- can we increase initial misfit?

--- 2012-12-18 Publications/2013-IPMI/Latex/045-Experiments.tex
@@
 reference segmentation is beyond the scope of this manuscript, we simply 
-rely on the shape priors obtained from the models.
+rely on the true contours known from the underlying models, and do not seek to establish them in a separate segmentation step on ``anatomical'' images.
 %
@@
 These authors note that ``partial volume effects blur the distinction between closely adjacent surfaces in deep sulci, leading to a well-known segmentation error in which the deeper reaches of sulci are not penetrated by the putative surface model.'' This problem is aggravated in DWI, since the resolution tends to be worse compared to the anatomical images considered in \citep{macdonald_automated_2000}.
-They test their coupled segmentation algorithm on an image, ``representing a sulcus in which the distinction between opposing banks of the sulcus has been obscured by partial volume.''  Here, we have reproduced a very similar model.
+They test their coupled segmentation algorithm on an image, ``representing a sulcus in which the distinction between opposing banks of the sulcus has been obscured by partial volume.''
+
+Here, we have reproduced a very similar model.
+A volume consists of three piecewise-constant parts: a notched ball representing the white matter with a single sulcus ($\mu_{WM} = 0.8$), a cortical sheet of grey-matter obtained through dilation of the WM ($\mu_{GM} = 0.5$), and the surrounding background representing CSF ($\mu_{CSF} = 0.2$).
+The volume is then affected by additive Gaussian noise, effectively creating uniform standard deviation of $\sigma = 0.045$ per region.
+
 As illustrated in \autoref{fig:sulcusmodel}, conventional single surface segmentation of the CSF/GM boundary misses to capture the sulcus in its full depth.
@@
 %
-In order to demonstrate the functionality of the methodology, 
-and characterize its possibilities with diffusion data,
-we simulated the \acs{dw} signal of a synthetic phantom from a model
-consisting of several spherical shapes emulating
+In order to demonstrate the specific functionality of the proposed
+methodology and characterize its possibilities with diffusion data, we developed a
+synthetic model and simulated its corresponding \ac{dwi}
+raw signal as described in \citep{tuch_q-ball_2004}.
+
+This second model consists of several spherical shapes emulating
 the different brain tissues (see \autoref{fig:fa}, first row). 
-We reconstructed the signal with standard procedures to 
+We reconstructed the \ac{dwi} signal with standard procedures to 
 approximate the environment to the real one at maximum. \\
 
-\paragraph{Signal simulation}
-To numerically simulate the MRI signal attenuation when applying a diffusion 
-gradient in a voxel with $N$ fiber populations we made use of the standard 
-\emph{Multi-Tensor Model}~\cite{Tuch:2002aa}:
+\subsubsection{Signal simulation}
+
+To numerically simulate the MRI signal attenuation when applying a diffusion gradient in a voxel with $N$ fiber populations we made use of the standard \emph{Multi-Tensor Model}~\cite{Tuch:2002aa}:
 %
-\begin{equation} 
-\label{eqn:MultiTensor}
+\begin{equation} \label{eqn:MultiTensor}
 S(\vec{q}) / S_{0} = \sum_{i=1}^{N} f_{i} \, \exp{ \left( -b \, \vec{q}^{T} \, \mathbf{D}_i \, \vec{q}\right) } + f_{iso} \, \exp{ \left( -b \, \mathbf{D}_{iso} \right) } ,
@@
 %
-where $\vec{q} \in \mathbb{S}^2$ is the direction of the diffusion gradient 
-applied, $b$ is the b-value accounting for its strength, $S_0 \equiv S(\vec{0})$ 
-is the signal with no diffusion weighting, $f_i$ and $\mathbf{D}_i$ are 
-respectively the volume fraction and the diffusion tensor characterizing the 
-$i$-th fiber population.
+where $\vec{q} \in \mathbb{S}^2$ is the direction of the diffusion gradient applied, $b$ is the b-value accounting for its strength, $S_0 \equiv S(\vec{0})$ is the signal with no diffusion weighting, $f_i$ and $\mathbf{D}_i$ are respectively the volume fraction and the diffusion tensor characterizing the $i$-th fiber population.
+The diffusion tensors $\mathbf{D}_i$ and $\mathbf{D}_{iso}$ describe the diffusion processes of each fiber compartment and of contaminations from the CSF.
+In this work, these quantities have been taken from standard ranges typically observed in the human brain~\cite{Canales-Rodriguez:2009aa}.
 
-The diffusion tensors $\mathbf{D}_i$ and $\mathbf{D}_{iso}$ describe the diffusion 
-processes of each fiber compartment and of contaminations from the CSF. In this work, 
-these quantities have been taken from standard ranges typically observed in the 
-human brain~\cite{Canales-Rodriguez:2009aa}.
+\subsubsection{Noise simulation}
 
-\paragraph{Noise simulation}
-The diffusion MRI signal $S$ has been corrupted with 
-\textit{Rician noise}~\cite{Gudbjartsson:1995aa} as follows:
+The diffusion MRI signal $S$ has been corrupted with \textit{Rician noise}~\cite{Gudbjartsson:1995aa} as follows:
 %
@@
 %
-where $\varepsilon_{1,2}$ are Gaussian distributed with zero mean
-and standard deviation $\sigma = S_0 / \mathit{SNR}$ and \acs{snr}
-is the \ac{snr} on the $S_0$ image.
+where $\varepsilon_{1,2}$ are Gaussian distributed with zero mean and standard deviation $\sigma = S_0 / \mathit{SNR}$ and SNR is the signal-to-noise ratio (SNR) on the $S_0$ image.
 
-\paragraph{Derived scalar features}
+\subsubsection{Derived features}
+
 The target \ac{dwi} data is characterized by its distortions and its
@@
 the reconstructed tensors and derived scalar maps have been
-studied by \cite{ennis_orthogonal_2006}. Based on their
+studied and presented on \cite{ennis_orthogonal_2006}. Based on their
 findings, the proposed energy model adapts to the \ac{fa} and \ac{md}
-for their properties. Whereas \ac{fa} describes the \emph{shape} of diffusion, 
+for their properties.
+Whereas \ac{fa} describes the \emph{shape} of diffusion, 
 the \ac{md} depicts the \emph{magnitude} of the process.
@@
 \textbf{FIXME: can you give the respective formula, how you get from the above signal to FA/MD?}
-\begin{equation}
-FA = \sqrt{ \frac{3}{2}},\frac{\sqrt{ (\lambda_1 - \hat{\lamda})^2 + (\lambda_2 - \hat{\lamda})^2 + (\lambda_3 - \hat{\lamda})^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}}
-\end{equation}
 
@@
 \begin{table}
-\begin{tabularx}{1.0\textwidth}{c|XXX}
-\hline
-         & $\mathbf{\mu}_{FA}$ & $\mathbf{\mu}_{MD}$ & $\Sigma$ \\
-\hline
+\begin{tabular}{cccc}
+         & $\mathbf{\mu}_{FA}$ & $\mathbf{\mu}_{MD}$ & $\mathbf{\Sigma}$ \\
 \ac{wm}  & 0.778 & 6.94\e{-4} & 
@@
 \\
-\hline
-\ac{gm}  & 0.119 & 8.95\e{-4} &
-   $\begin{pmatrix}
-   	5.90\e{-4} & -1.43\e{-6} \\ -1.43\e{-6} & 1.04\e{-8}
-   \end{pmatrix}$
-\\
-\hline
-\ac{csf} & 0.103 & 2.99\e{-3} &
-   $\begin{pmatrix}
-   	1.19\e{-3} & 2.22\e{-7} \\ 2.22\e{-7} & 1.56\e{-8}
-   \end{pmatrix}$
-\\
-\hline
-\end{tabularx}
+\ac{gm}  & 0.119 & 8.95\e{-4} &  \\
+\ac{csf} & 0.103 & 2.99\e{-3} & \\
+\end{tabular}
 \caption{Model parameters}

### 2012-12-19 Publications/2013-IPMI/Latex/05-Results.tex
Inserted Alessandro\'s text on DWI model
adding new model1 results -- can we increase initial misfit?

--- 2012-11-21 Publications/2013-IPMI/Latex/05-Results.tex
@@
 \label{sec:results}
+
+\subsection{Synthetic gray-scale data}
+
+\subsection{Simulated diffusion data}
 %
+The proposed method successfully reverted the synthetic distortion
+field we applied to the data. With 16x16x16 control points, the
+displacements field is dense enough to correctly represent the
+synthetic field. (INCLUDE FIGURE). Figure XX shows the restored
+image and a difference map with the original model (we can also
+compute Dice indexes and that stuff).\\
+%
+
+\begin{figure}
+\begin{tabular}{ccccc}
+%\includegraphics[width=0.19\textwidth]{fig_modelFA_00} &
+%\includegraphics[width=0.19\textwidth]{fig_modelFA_01} &
+%\includegraphics[width=0.19\textwidth]{fig_modelFA_02} &
+%\includegraphics[width=0.19\textwidth]{fig_modelFA_03} &
+%\includegraphics[width=0.19\textwidth]{fig_modelFA_04} \\
+%\includegraphics[width=0.19\textwidth]{fig_contours_00} &
+%\includegraphics[width=0.19\textwidth]{fig_contours_01} &
+%\includegraphics[width=0.19\textwidth]{fig_contours_02} &
+%\includegraphics[width=0.19\textwidth]{fig_contours_03} &
+%\includegraphics[width=0.19\textwidth]{fig_contours_04}
+\includegraphics[width=0.19\textwidth]{model1result_b_1} &
+\includegraphics[width=0.19\textwidth]{model1result_b_2} &
+\includegraphics[width=0.19\textwidth]{model1result_b_3} &
+\includegraphics[width=0.19\textwidth]{model1result_b_4} &
+\includegraphics[width=0.19\textwidth]{model1result_b_5} \\
+\includegraphics[width=0.19\textwidth]{model1result_a_1} &
+\includegraphics[width=0.19\textwidth]{model1result_a_2} &
+\includegraphics[width=0.19\textwidth]{model1result_a_3} &
+\includegraphics[width=0.19\textwidth]{model1result_a_4} &
+\includegraphics[width=0.19\textwidth]{model1result_a_5} \\
+\end{tabular}
+\caption{First row presents several slices along Z axis of the distorted \ac{fa} map and
+the undistorted \ac{wm}-\ac{gm} and \ac{wm}-\ac{csf} contours given as shape priors. Second row contains the same map, now with the contours after joint segmentation-registration.}
+\label{fig:fa}
+\end{figure}

### 2012-12-19 Publications/2013-IPMI/Latex/03-Introduction.tex
adding a few references for the technical state of the art

@@
 
+\textbf{FIXME: talk about inherent coupling}
+The significant benefits of exploiting the anatomical \ac{mri}
+have been demonstrated \cite{zollei_improved_2010}, justifying the use
+of the shape prior information. This need is satisfied by
+the use of a unique underlying model of our joint approach.
+
+\textbf{FIXME: talk about SOA in joint reg-seg, shape-prior registration, atlas-based registration in more general}
+\citep{Chan2005,Bresson2006a,Cremers2006,Rousson2002,Schmid2008,Vemuri2003a,Gastaud2004,McInerney1996,Chan,Paragios2003,Ayvaci2007,Pohl2005,Wyatt2003,Yezzi2003,Wang2006,Soatto2002,Yezzi2001,Pohl2006,Chen2002,le_guyader_combined_2011}
+
+\textbf{FIXME: now introduce our own approach}
 In this paper we propose a novel registration framework to simultaneously
 solve the segmentation and distortion challenges, by exploiting as strong 
-shape-prior the detailed anatomy extracted from anatomical \ac{mri}. On
-one hand, the significant benefits of exploiting the anatomical \ac{mri}
-have been demonstrated \cite{zollei_improved_2010}, justifying the use
-of the shape prior information. On the other, the
-significant coupling between both problems demands for consistent
-segmentation, parcellation, and distortion. This need is satisfied by
-the use of a unique underlying model of our joint approach.
+shape-prior the detailed anatomy extracted from anatomical \ac{mri}. 
 We reformulate the segmentation problem as an inverse problem, where
@@
 from the structural space into the diffusion space.
+
+\textbf{Now quickly summarize structure of the paper}

### 2012-12-20 Publications/2013-IPMI/Latex/03-Introduction.tex
time for an intermediate coffee break; references all nicely included (save for LeGuyader/Vese, to be done), rest of introduction to be fixed soon

--- 2012-12-12 Publications/2013-IPMI/Latex/03-Introduction.tex
@@
 large amount of imaging techniques for acquisition, processing, and analysis
-specifically tuned for \ac{dwi} data.\\
+specifically tuned for \ac{dwi} data.
 
-The whole-brain connectivity analysis has arisen some challenges
-that should be overcome in order to get reliable structural information
+The whole-brain connectivity analysis has given rise to some challenges
+towards reliable structural information
 about the neuronal tracts from \ac{dwi} \cite{johansen-berg_using_2009,
@@
 segmentation on the diffusion space and the correction of geometrical
-distortions inherent to the imaging techniques \citep{hagmann_mr_2012}.\\
+distortions inherent to the imaging techniques \citep{hagmann_mr_2012}.
 
@@
 requirement is usually solved in practice by plainly thresholding the 
-\ac{fa} (a well-know scalar map derived from \ac{dwi} which depicts 
-the isotropy of water diffusion inside the brain). 
-Additionally, a precise location of the
-\ac{gm}-\ac{wm} surface is required in the final steps to
+\ac{fa}, a well-known scalar map derived from \ac{dwi} which depicts 
+the isotropy of water diffusion inside the brain. 
+A precise location of the
+\ac{gm}-\ac{wm} surface is also essential in the final steps to
 achieve a consistent parcellisation of the cortex to represent the nodes 
@@
 high-resolution and better understood structural \ac{mri} of the same 
-subject (e.g. T1 and/or T2 weighted acquisitions). Conversely, this 
-problem is resolved with non-linear registration of a structural \ac{mri}
-of the subject to the \ac{dwi} data. Even though some efforts have addressed
-the study of the robustness of tractography versus the intra-subject variability
-\cite{wakana_reproducibility_2007,heiervang_between_2006}, the results produced
-are restricted to relevant regions of the brain. Therefore, extremely robust
-and precise segmentation methods are required in the whole-brain application. \\
+subject (e.g. T1 and/or T2 weighted acquisitions). %Conversely, this problem is resolved with non-linear registration of a structural \ac{mri} of the subject to the \ac{dwi} data. 
+Even though some efforts have addressed
+the study of the robustness of tractography with respect to intra-subject variability
+\cite{wakana_reproducibility_2007,heiervang_between_2006}, these results are restricted to some regions of the brain, only. Therefore, extremely robust
+and precise segmentation methods are required in the whole-brain application.
 
-On the other hand, the \ac{dwi} data is usually obtained with \ac{epi}
+The \ac{dwi} data is usually obtained with \ac{epi}
 acquisition techniques, that often suffer from severe distortions due to 
-local field inhomogeneities. Generally, it is easily appreciated in the anterior
+local field inhomogeneities. Generally, it is appreciated in the anterior
 part of the brain, along the phase-encoded direction. Some methodologies have
@@
 \cite{holland_efficient_2010,hsu_correction_2009,jezzard_characterization_2005,
-reber_correction_2005}. They usually 
+reber_correction_2005}. These methods usually 
 require the extra acquisition of the magnitude and phase of
-the field (field-mapping), condition which is not always met. Some other 
-methodologies do not make use the field-mapping, compensating the distortion
+the field (field-mapping), a condition which is not always met. Some other 
+methodologies do not make use of the field-mapping, compensating the distortion
 with non-linear registration from structural \ac{mri} or other means
@@
 of the impact of the \ac{epi} distortion on the variability of tractography
-results. \\
+results.
+
+It is easy to see, that the problems of precise segmentation in \ac{dwi}-space and the spatial mapping between these contours and the corresponding surfaces in anatomical images bear important redundancy.
+Once the spatial relationship between T1 and \ac{dwi} space is established, the contours which are readily available in T1 space, can simply be projected on to the \ac{dwi}-data.
+Conversely, if a precise delineation in \ac{dwi}-space was achieved, the spatial mapping with T1-space could be derived from one-to-one correspondences on the contours.
+However, neither segmentation nor registration can be performed flawlessly, if considered independently.
+The significant benefits of exploiting the anatomical \ac{mri} when segmenting the \ac{dwi} data have been demonstrated \cite{zollei_improved_2010}, justifying the use
+of the shape prior information.
+
+The use of anatomical templates in medical image segmentation is not exactly new---for an early overview see \citep{McInerney1996} and references therein.
+More recently, we suggest clustering the diverse methods of template-based segmentation methods into three groups.
+The first group typically adds a shape prior term to the energy functional of an evolving active contour \citep{Rousson2002,Chen2002,Paragios2003,Vemuri2003a,Yezzi2003a,Gastaud2004,Chan2005,Cremers2006,Bresson2006a,Ayvaci2007,Schmid2008}.
+These methods have a more or less explicit description of the expected relative boundary locations of the object to be delineated, and some even model the statistical deviations from this average shape.
+Closely related to this group are atlas-based segmentation methods \citep{Pohl2005,Pohl2006,Wang2006,Gorthi2009,Gorthi2011}, where the prior imposes consistent voxel-based classification of contiguous regions.
+Here, the presence of more structures than just the actual ROI helps aligning the target image with the atlas in a hierarchical fashion.
+Finally, the third group generalizes the atlas to actual images, and the contour is to segment simultaneously two different target images, related by a spatial transform to be co-estimated \citep{Wyatt2003,Yezzi2003}.
 
 In this paper we propose a novel registration framework to simultaneously
-solve the segmentation and distortion challenges, by exploiting as strong 
-shape-prior the detailed anatomy extracted from anatomical \ac{mri}. This
-is justified by the strong relationship between both problems, and the
-advantage of a significant increase of coherence among these steps in
-the complete processing pipeline.
-We reformulate the segmentation problem as an inverse problem, where
+solving the segmentation and distortion challenges, by exploiting as strong 
+shape-prior the detailed anatomy extracted from anatomical \ac{mri}.
+Indeed, in our case we can consider the segmentation in anatomical images as a solved problem.
+Moreover, the shape prior is of very strong nature, since it is very specific to the particular subject.
+Also, after global alignment using existing approaches, the remaining spatial deformation between anatomical and diffusion space is due to MR distortion.
+Therefore, we can reduce the problem to finding the differences of spatial distortion in between anatomical and \ac{dwi} space.
+We thus reformulate the segmentation problem as an inverse problem, where
 we seek for an underlying deformation field (the distortion) mapping 
 from the structural space into the diffusion space.
+
+\textbf{Anticipate results}
+
+\textbf{Now quickly summarize structure of the paper}

### 2012-12-20 Publications/2013-IPMI/Latex/045-Experiments.tex
time for an intermediate coffee break; references all nicely included (save for LeGuyader/Vese, to be done), rest of introduction to be fixed soon

@@
 \end{align}
-
-There exist two main reasons to justify their choice. 
+where $\lambda_i$ are the eigenvalues of the diffusion tensor associated with the diffusion signal $S(\vec{q})$. There exist two main reasons to justify their choice. 
 First, they are well-understood and standardized in clinical routine.
@@
 \begin{table}
-\begin{tabularx}{1.0\textwidth}{c|XXX}
-\hline
-         & $\mathbf{\mu}_{FA}$ & $\mathbf{\mu}_{MD}$ & $\Sigma$ \\
-\hline
+\caption{Model means and covariances of \acl{fa} and \acl{md} estimated from the reconstructed simulated \ac{dwi} images for each modeled tissue, \acl{wm},\acl{gm}, and \acl{csf}.
+As expected, the two scalar features are complementary and the three tissues can well be discriminated.
+}
+\label{table:parameters}
+%\begin{tabularx}{1.0\textwidth}{c|XXX}
+\begin{tabular}{*{3}{@{}p{0.2\textwidth}}@{}p{0.4\textwidth}@{}}
+\toprule
+% tissue        & $\mathbf{\mu}_{FA}$ & $\mathbf{\mu}_{MD}$ & $\Sigma$ \newline
+        & $\mathbf{\mu}$ \\
+\cmidrule(r){2-3}
+Tissue& FA & MD &$\Sigma$\\
+\cmidrule(r){1-1}\cmidrule(r){2-2}\cmidrule(r){3-3}\cmidrule{4-4}
 \ac{wm}  & 0.778 & 6.94\e{-4} & 
@@
    	4.85\e{-3} & -6.90\e{-6} \\ -6.90\e{-6} & 1.03\e{-8}
-   \end{pmatrix}$
-\\
-\hline
+   \end{pmatrix}$\vspace{2mm}\\
+%
 \ac{gm}  & 0.119 & 8.95\e{-4} &
@@
    	5.90\e{-4} & -1.43\e{-6} \\ -1.43\e{-6} & 1.04\e{-8}
-   \end{pmatrix}$
+   \end{pmatrix}$\vspace{2mm}
 \\
-\hline
+%
 \ac{csf} & 0.103 & 2.99\e{-3} &
@@
 \\
-\hline
-\end{tabularx}
-\caption{Model parameters}
-\label{table:parameters}
+\bottomrule
+\end{tabular}
+%\end{tabularx}
+
 \end{table}

### 2012-12-20 Publications/2013-IPMI/Latex/03-Introduction.tex
almost complete introduction, now waiting for further results to conclude.

@@
 high-resolution and better understood structural \ac{mri} of the same 
-subject (e.g. T1 and/or T2 weighted acquisitions). %Conversely, this problem is resolved with non-linear registration of a structural \ac{mri} of the subject to the \ac{dwi} data. 
+subject (e.g. \ac{t1} and/or \ac{t2} weighted acquisitions). %Conversely, this problem is resolved with non-linear registration of a structural \ac{mri} of the subject to the \ac{dwi} data. 
 Even though some efforts have addressed
@@
 It is easy to see, that the problems of precise segmentation in \ac{dwi}-space and the spatial mapping between these contours and the corresponding surfaces in anatomical images bear important redundancy.
-Once the spatial relationship between T1 and \ac{dwi} space is established, the contours which are readily available in T1 space, can simply be projected on to the \ac{dwi}-data.
-Conversely, if a precise delineation in \ac{dwi}-space was achieved, the spatial mapping with T1-space could be derived from one-to-one correspondences on the contours.
+Once the spatial relationship between \ac{t1} and \ac{dwi} space is established, the contours which are readily available in \ac{t1} space, can simply be projected on to the \ac{dwi}-data.
+Conversely, if a precise delineation in \ac{dwi}-space was achieved, the spatial mapping with \ac{t1}-space could be derived from one-to-one correspondences on the contours.
 However, neither segmentation nor registration can be performed flawlessly, if considered independently.
@@
 Closely related to this group are atlas-based segmentation methods \citep{Pohl2005,Pohl2006,Wang2006,Gorthi2009,Gorthi2011}, where the prior imposes consistent voxel-based classification of contiguous regions.
-Here, the presence of more structures than just the actual ROI helps aligning the target image with the atlas in a hierarchical fashion.
+Here, the presence of more structures than just the actual \ac{roi} helps aligning the target image with the atlas in a hierarchical fashion.
 Finally, the third group generalizes the atlas to actual images, and the contour is to segment simultaneously two different target images, related by a spatial transform to be co-estimated \citep{Wyatt2003,Yezzi2003}.
@@
 Moreover, the shape prior is of very strong nature, since it is very specific to the particular subject.
-Also, after global alignment using existing approaches, the remaining spatial deformation between anatomical and diffusion space is due to MR distortion.
+Also, after global alignment using existing approaches, the remaining spatial deformation between anatomical and diffusion space is due to \ac{mr} distortion.
+Finally, we need to establish precise spatial correspondence between the surfaces in both spaces, including the tangential direction for parcellation.
 Therefore, we can reduce the problem to finding the differences of spatial distortion in between anatomical and \ac{dwi} space.
-We thus reformulate the segmentation problem as an inverse problem, where
-we seek for an underlying deformation field (the distortion) mapping 
-from the structural space into the diffusion space.
+We thus reformulate the segmentation problem as an inverse problem, where we seek for an underlying deformation field---the distortion---mapping
+from the structural space into the diffusion space, such that the structural contours segment optimally the \ac{dwi} data.
+In the process, the one-to-one correspondence between the contours in both spaces is guaranteed, and projection of parcellisation is trivial.
 
-\textbf{Anticipate results}
+We test our proposed joint segmentation-registration model on two different synthetic examples.
+The first example is a scalar sulcus model, where the \ac{csf}-\ac{gm} boundary particularly suffers from \ac{pve} and can only be segmented correctly thanks to the shape prior and its coupling with the inner, \ac{gm}-\ac{wm} boundary through the imposed deformation field regularity.
+The second case deals with more realistic \ac{dwi} data stemming from simulations of a simplistic brain data.
+Again, we can show that the proposed model successfully segments the \ac{dwi} data based on two derived scalar features, namely \ac{fa} and \ac{md}, while establishing an estimate of the dense distortion field.
 
-\textbf{Now quickly summarize structure of the paper}
+The rest of this paper is organized as follows.
+First, in \autoref{sec:methods} we introduce our proposed model for joint multivariate segmentation-registration.
+Then we provide a more detailed description of the data and experimental setup in \autoref{sec:experiments}.
+We present results in \autoref{sec:results} and conclude in \autoref{sec:conclusion}.

### 2012-12-20 Publications/2013-IPMI/Latex/03-Introduction.tex
Added model 2 results

@@
 delineation of the \ac{csf}-\ac{gm} and \ac{gm}-\ac{wm} interface surfaces.
-This segmentation is an important step on which strongly rely further
+This segmentation is an processing step on which strongly rely further
 tasks. In tractography, a high-standard \ac{wm} mask is required. Otherwise,
@@
 the study of the robustness of tractography with respect to intra-subject variability
-\cite{wakana_reproducibility_2007,heiervang_between_2006}, these results are restricted to some regions of the brain, only. Therefore, extremely robust
+\cite{wakana_reproducibility_2007,heiervang_between_2006}, these results are restricted to some regions of the brain, only. Therefore, robust
 and precise segmentation methods are required in the whole-brain application.
@@
 require the extra acquisition of the magnitude and phase of
-the field (field-mapping), a condition which is not always met. Some other 
+the field (``field-mapping''), a condition which is not always met. Some other 
 methodologies do not make use of the field-mapping, compensating the distortion
@@
 \citep{andersson_modeling_2001}. To our knowledge, there exists no study
-of the impact of the \ac{epi} distortion on the variability of tractography
+on the impact of the \ac{epi} distortion on the variability of tractography
 results.
 
-It is easy to see, that the problems of precise segmentation in \ac{dwi}-space and the spatial mapping between these contours and the corresponding surfaces in anatomical images bear important redundancy.
-Once the spatial relationship between \ac{t1} and \ac{dwi} space is established, the contours which are readily available in \ac{t1} space, can simply be projected on to the \ac{dwi}-data.
-Conversely, if a precise delineation in \ac{dwi}-space was achieved, the spatial mapping with \ac{t1}-space could be derived from one-to-one correspondences on the contours.
-However, neither segmentation nor registration can be performed flawlessly, if considered independently.
-The significant benefits of exploiting the anatomical \ac{mri} when segmenting the \ac{dwi} data have been demonstrated \cite{zollei_improved_2010}, justifying the use
-of the shape prior information.
+Therefore, the problems of precise segmentation in \ac{dwi}-space and the 
+spatial mapping between these contours and the corresponding surfaces in 
+anatomical images bear significant redundancy. Once the spatial relationship 
+between \ac{t1} and \ac{dwi} space is established, the contours which are 
+readily available in \ac{t1} space, can simply be projected on to the 
+\ac{dwi}-data. Conversely, if a precise delineation in \ac{dwi}-space 
+was achieved, the spatial mapping with \ac{t1}-space could be derived 
+from one-to-one correspondences on the contours. However, neither segmentation 
+nor registration can be performed flawlessly, if considered independently. 
+The significant benefits of exploiting the anatomical \ac{mri} when 
+segmenting the \ac{dwi} data have been demonstrated \cite{zollei_improved_2010}, 
+justifying the use of the shape prior information.
 
-The use of anatomical templates in medical image segmentation is not exactly new---for an early overview see \citep{McInerney1996} and references therein.
-More recently, we suggest clustering the diverse methods of template-based segmentation methods into three groups.
-The first group typically adds a shape prior term to the energy functional of an evolving active contour \citep{Rousson2002,Chen2002,Paragios2003,Vemuri2003a,Yezzi2003a,Gastaud2004,Chan2005,Cremers2006,Bresson2006a,Ayvaci2007,Schmid2008}.
-These methods have a more or less explicit description of the expected relative boundary locations of the object to be delineated, and some even model the statistical deviations from this average shape.
-Closely related to this group are atlas-based segmentation methods \citep{Pohl2005,Pohl2006,Wang2006,Gorthi2009,Gorthi2011}, where the prior imposes consistent voxel-based classification of contiguous regions.
-Here, the presence of more structures than just the actual \ac{roi} helps aligning the target image with the atlas in a hierarchical fashion.
-Finally, the third group generalizes the atlas to actual images, and the contour is to segment simultaneously two different target images, related by a spatial transform to be co-estimated \citep{Wyatt2003,Yezzi2003}.
+We suggest clustering the current methodologies of template-based segmentation 
+methods into three groups. The first group typically adds a shape prior term to 
+the energy functional of an evolving active contour~\citep{Rousson2002,Chen2002,
+Paragios2003,Vemuri2003a,Yezzi2003a,Gastaud2004,Chan2005,Cremers2006,Bresson2006a,
+Ayvaci2007,Schmid2008} 
+{\color{red}{(FIXME:I would select from here the most important,
+removing directly those appearing in conferences and never published on journal. Better
+if we found a review paper/book chapter. I don't like so many references on a row)}}.
+These methods generally have a explicit description of the expected relative boundary 
+locations of the object to be delineated, and some even model the statistical deviations
+from this average shape. Closely related to this group are atlas-based segmentation
+methods \citep{Pohl2005,Pohl2006,Wang2006,Gorthi2009,Gorthi2011}, where the prior 
+imposes consistent voxel-based classification of contiguous regions. Here, the 
+presence of more structures than one unique \ac{roi} helps aligning the target image 
+with the atlas in a hierarchical fashion. Finally, the third group generalizes 
+the atlas to actual images, and the contour is to segment simultaneously two 
+different target images, related by a spatial transform to be co-estimated~\citep{Wyatt2003,
+Yezzi2003}.
 
@@
 solving the segmentation and distortion challenges, by exploiting as strong 
-shape-prior the detailed anatomy extracted from anatomical \ac{mri}.
-Indeed, in our case we can consider the segmentation in anatomical images as a solved problem.
-Moreover, the shape prior is of very strong nature, since it is very specific to the particular subject.
-Also, after global alignment using existing approaches, the remaining spatial deformation between anatomical and diffusion space is due to \ac{mr} distortion.
-Finally, we need to establish precise spatial correspondence between the surfaces in both spaces, including the tangential direction for parcellation.
-Therefore, we can reduce the problem to finding the differences of spatial distortion in between anatomical and \ac{dwi} space.
-We thus reformulate the segmentation problem as an inverse problem, where we seek for an underlying deformation field---the distortion---mapping
-from the structural space into the diffusion space, such that the structural contours segment optimally the \ac{dwi} data.
-In the process, the one-to-one correspondence between the contours in both spaces is guaranteed, and projection of parcellisation is trivial.
+shape-prior the detailed anatomy extracted from anatomical \ac{mri}. Indeed,
+hereafter we consider the segmentation in anatomical images as a solved problem.
+Moreover, the shape prior is of very ``strong'' nature, since it is specific to 
+the particular subject. Also, after global alignment using existing approaches, 
+the remaining spatial deformation between anatomical and diffusion space is 
+due to \ac{mr} distortion. Finally, we need to establish precise spatial 
+correspondence between the surfaces in both spaces, including the tangential 
+direction for parcellation. Therefore, we can reduce the problem to finding 
+the differences of spatial distortion in between anatomical and \ac{dw} space.
+We thus reformulate the segmentation problem as an inverse problem, where we 
+seek for an underlying deformation field (the distortion) mapping 
+from the structural space into the diffusion space, such that the structural 
+contours segment optimally the \ac{dwi} data. In the process, the one-to-one 
+correspondence between the contours in both spaces is guaranteed, and projection 
+of parcellisation to \ac{dw} space is implicit and consistent.
 
-We test our proposed joint segmentation-registration model on two different synthetic examples.
-The first example is a scalar sulcus model, where the \ac{csf}-\ac{gm} boundary particularly suffers from \ac{pve} and can only be segmented correctly thanks to the shape prior and its coupling with the inner, \ac{gm}-\ac{wm} boundary through the imposed deformation field regularity.
-The second case deals with more realistic \ac{dwi} data stemming from simulations of a simplistic brain data.
-Again, we can show that the proposed model successfully segments the \ac{dwi} data based on two derived scalar features, namely \ac{fa} and \ac{md}, while establishing an estimate of the dense distortion field.
+We test our proposed joint segmentation-registration model on two different 
+synthetic examples. The first example is a scalar sulcus model, where the 
+\ac{csf}-\ac{gm} boundary particularly suffers from \ac{pve} and can only be 
+segmented correctly thanks to the shape prior and its coupling with the inner, 
+\ac{gm}-\ac{wm} boundary through the imposed deformation field regularity. 
+The second case deals with more realistic \ac{dwi} data stemming from 
+phantom simulations of a simplistic brain data. Again, we show that the 
+proposed model successfully segments the \ac{dwi} data based on two derived 
+scalar features, namely \ac{fa} and \ac{md}, while establishing an estimate 
+of the dense distortion field.
 
-The rest of this paper is organized as follows.
-First, in \autoref{sec:methods} we introduce our proposed model for joint multivariate segmentation-registration.
-Then we provide a more detailed description of the data and experimental setup in \autoref{sec:experiments}.
-We present results in \autoref{sec:results} and conclude in \autoref{sec:conclusion}.
+The rest of this paper is organized as follows. First, in \autoref{sec:methods}
+we introduce our proposed model for joint multivariate segmentation-registration.
+Then we provide a more detailed description of the data and experimental setup in
+\autoref{sec:experiments}. We present results in \autoref{sec:results} and conclude 
+in \autoref{sec:conclusion}.

### 2012-12-20 Publications/2013-IPMI/Latex/045-Experiments.tex
Added model 2 results

@@
 reference segmentation is beyond the scope of this manuscript, we simply 
-rely on the true contours known from the underlying models, and do not seek to establish them in a separate segmentation step on ``anatomical'' images.
+rely on the true contours known from the underlying models, and do not 
+seek to establish them in a separate segmentation step on ``anatomical'' images.
 %
@@
 %
-The first, toy example is inspired by a problem shown for coupled CSF/GM and GM/WM segmentation in \citep{macdonald_automated_2000}.
-These authors note that ``partial volume effects blur the distinction between closely adjacent surfaces in deep sulci, leading to a well-known segmentation error in which the deeper reaches of sulci are not penetrated by the putative surface model.'' This problem is aggravated in DWI, since the resolution tends to be worse compared to the anatomical images considered in \citep{macdonald_automated_2000}.
-They test their coupled segmentation algorithm on an image, ``representing a sulcus in which the distinction between opposing banks of the sulcus has been obscured by partial volume.''
+The first simplified model to test the approach is inspired by a problem 
+shown for coupled \ac{csf}/\ac{gm} and \ac{gm}/\ac{wm} segmentation in \citep{macdonald_automated_2000}. Authors note that ``partial volume effects 
+blur the distinction between closely adjacent surfaces in deep sulci, leading 
+to a well-known segmentation error in which the deeper reaches of sulci are 
+not penetrated by the putative surface model.'' This problem is aggravated 
+in DWI, since the resolution tends to be worse compared to the anatomical 
+images considered in \citep{macdonald_automated_2000}. They test their 
+coupled segmentation algorithm on an image, ``representing a sulcus in 
+which the distinction between opposing banks of the sulcus has been obscured 
+by partial volume.''
 
-Here, we have reproduced a very similar model.
-A volume consists of three piecewise-constant parts: a notched ball representing the white matter with a single sulcus ($\mu_{WM} = 0.8$), a cortical sheet of grey-matter obtained through dilation of the WM ($\mu_{GM} = 0.5$), and the surrounding background representing CSF ($\mu_{CSF} = 0.2$).
-The volume is then affected by additive Gaussian noise, effectively creating uniform standard deviation of $\sigma = 0.045$ per region.
+Here, we reproduce their model on a volume consisting of three 
+piecewise-constant parts: a notched ball representing the \ac{wm} with a single 
+sulcus ($\mu_{WM} = 0.8$), a cortical sheet of \ac{gm} obtained through dilation 
+of the \ac{wm} ($\mu_{GM} = 0.5$), and the surrounding background representing 
+\ac{csf} ($\mu_{CSF} = 0.2$). The volume is then affected by additive Gaussian 
+noise, effectively creating uniform standard deviation of $\sigma = 0.045$ per 
+region.
 
-As illustrated in \autoref{fig:sulcusmodel}, conventional single surface segmentation of the CSF/GM boundary misses to capture the sulcus in its full depth.
-With our proposed model, we expect the joint segmentation-registration to be driven largely by the inner, GM/WM contour that exhibits sufficient contrast and lesser partial volume effects.
-The shape prior of the outer, difficult contour will then be co-aligned through the regularity of the estimated deformation field.
+As illustrated in \autoref{fig:sulcusmodel}, conventional single surface 
+segmentation of the \ac{csf}/\ac{gm} boundary misses to capture the sulcus in 
+its full depth. With our proposed model, we expect the joint 
+segmentation-registration to be driven largely by the inner, \ac{gm}/\ac{wm}
+contour that exhibits sufficient contrast and lesser partial volume effects. 
+The shape prior of the outer, difficult contour will then be co-aligned through 
+the regularity of the estimated deformation field.
 
@@
 \end{align}
-where $\lambda_i$ are the eigenvalues of the diffusion tensor associated with the diffusion signal $S(\vec{q})$. There exist two main reasons to justify their choice. 
+where $\lambda_i$ are the eigenvalues of the diffusion tensor 
+associated with the diffusion signal $S(\vec{q})$. There exist 
+two main reasons to justify their choice. 
 First, they are well-understood and standardized in clinical routine.

### 2012-12-20 Publications/2013-IPMI/Latex/05-Results.tex
Added model 2 results

@@
 \subsection{Synthetic gray-scale data}
+The proposed method solved precisely the specific challenge
+created by the model. A severe distortion of the model is
+artificially created adding complexity to the problem of
+partial voluming in the outer contour of the sulcus.
+\autoref{fig:sulcusmodel_result} provides visual assessment
+for this result. With 16x16x16 control points, 
+computation time for this model was around
+10 minutes in a {\color{red}{(specific machine details)}}.
+\begin{figure}
+\includegraphics[width=1.0\textwidth]{model2result}
+\caption{Result of the segmentation performance on the first
+model. In red and blue colors the obtained \ac{wm}/\ac{gm} 
+and \ac{gm}/\ac{csf} contours for the deformed image, respectively.
+In white color, the initial shape priors for the \ac{wm}/\ac{gm}
+interface.
+{\color{red}{This figure is to be replaced by the real result,
+this is a proof of concept of registration that does not solve
+the pv problem. The current figure also removes the outer prior}}}
+\label{fig:sulcusmodel_result}
+\end{figure}
 
@@
 displacements field is dense enough to correctly represent the
-synthetic field. (INCLUDE FIGURE). Figure XX shows the restored
-image and a difference map with the original model (we can also
-compute Dice indexes and that stuff).\\
+synthetic field. Second row in \autoref{fig:fa} shows the fitted
+contours obtained by using the original surfaces of the model
+as shape priors, with a constant translation of (5.0mm,10mm.,-5.0mm)
+to illustrate briefly the extent of the capture range of the algorithm.
+Computation time in this case was around 4 minutes in the previously
+described platform. \\
 %
@@
 \begin{tabular}{ccccc}
-%\includegraphics[width=0.19\textwidth]{fig_modelFA_00} &
-%\includegraphics[width=0.19\textwidth]{fig_modelFA_01} &
-%\includegraphics[width=0.19\textwidth]{fig_modelFA_02} &
-%\includegraphics[width=0.19\textwidth]{fig_modelFA_03} &
-%\includegraphics[width=0.19\textwidth]{fig_modelFA_04} \\
-%\includegraphics[width=0.19\textwidth]{fig_contours_00} &
-%\includegraphics[width=0.19\textwidth]{fig_contours_01} &
-%\includegraphics[width=0.19\textwidth]{fig_contours_02} &
-%\includegraphics[width=0.19\textwidth]{fig_contours_03} &
-%\includegraphics[width=0.19\textwidth]{fig_contours_04}
 \includegraphics[width=0.19\textwidth]{model1result_b_1} &

### 2012-12-20 Publications/2013-IPMI/Latex/06-Conclusion.tex
Added model 2 results

@@
-\section{Conclusion}
+\section{Conclusions and outlook}
 \label{sec:conclusion}
@@
 
+We implemented the methodology upon the widely used
+Insight Registration and Segmentation 
+Toolkit~\footnote{\url{http://www.itk.org}} (ITK)
+for its computational benefits, the standardized code, and 
+with the aim at making the procedure publicly available 
+when ready for sharing with the research community.
+
 Once proven the aptness of the methodology to the application

### 2012-12-21 Publications/2013-IPMI/Latex/03-Introduction.tex

@@
 methods into three groups. The first group typically adds a shape prior term to 
-the energy functional of an evolving active contour~\citep{Rousson2002,Chen2002,
-Paragios2003,Vemuri2003a,Yezzi2003a,Gastaud2004,Chan2005,Cremers2006,Bresson2006a,
-Ayvaci2007,Schmid2008} 
-{\color{red}{(FIXME:I would select from here the most important,
-removing directly those appearing in conferences and never published on journal. Better
-if we found a review paper/book chapter. I don't like so many references on a row)}}.
+the energy functional of an evolving active contour~\citep{Chen2002,
+Paragios2003,Vemuri2003a,Yezzi2003a,Gastaud2004,Chan2005,Cremers2006,Bresson2006a} %removed Rousson2002,Ayvaci2007,Schmid2008
+%{\color{red}{(FIXME:I would select from here the most important,
+%removing directly those appearing in conferences and never published on journal. Better
+%if we found a review paper/book chapter. I don't like so many references on a row)}}.
 These methods generally have a explicit description of the expected relative boundary 

### 2012-12-21 Publications/2013-IPMI/Latex/05-Results.tex

@@
 10 minutes in a {\color{red}{(specific machine details)}}.
+
 \begin{figure}
-\includegraphics[width=1.0\textwidth]{model2result}
-\caption{Result of the segmentation performance on the first
-model. In red and blue colors the obtained \ac{wm}/\ac{gm} 
-and \ac{gm}/\ac{csf} contours for the deformed image, respectively.
-In white color, the initial shape priors for the \ac{wm}/\ac{gm}
-interface.
+\begin{tabular}{ccccc}
+\includegraphics[width=0.19\textwidth]{model2result_b_1} &
+\includegraphics[width=0.19\textwidth]{model2result_b_2} &
+\includegraphics[width=0.19\textwidth]{model2result_a_1} &
+\includegraphics[width=0.19\textwidth]{model2result_a_2} &
+\includegraphics[width=0.19\textwidth]{model2surf} \\
+\multicolumn{2}{c}{a)} & \multicolumn{2}{c}{b)} & c)
+\end{tabular}
+\caption{Result of the segmentation performance on the sulcus model.
+a) Two slices and prior contours before segmentation-registration.
+b) The same slices and contours after distortion estimation.
+c) 3D rendering of the deformed priors, with the traces of the two selected slices highlighted.
+%In red and blue colors the obtained \ac{wm}/\ac{gm} 
+%and \ac{gm}/\ac{csf} contours for the deformed image, respectively.
+%In white color, the initial shape priors for the \ac{wm}/\ac{gm}
+%interface.
 {\color{red}{This figure is to be replaced by the real result,

### 2012-12-22 Publications/2013-IPMI/Latex/00-Main.tex
fixing sqrt in methods
Updated latex and model parameteres
Inserted Alessandro\'s text on DWI model
time for an intermediate coffee break; references all nicely included (save for LeGuyader/Vese, to be done), rest of introduction to be fixed soon
Added model 2 results
Anonimized paper & Glossaries package
Changed from acronym to glossaries to be able to handle
capitalization

@@
 \usepackage{makeidx}  % allows for indexgeneration
-\usepackage{acronym}
+%\usepackage{acronym}
+%\usepackage{acro}
+\usepackage{glossaries}
 \usepackage[numbers]{natbib}
@@
 \usepackage{hyperref}
-\usepackage[hyphens]{url}
-\usepackage[hyphenbreaks]{breakurl}
+%\usepackage[hyphens]{url}
+%\usepackage[hyphenbreaks]{breakurl}
 \usepackage[USenglish]{babel}
@@
 \usepackage{pslatex}
+%\usepackage{tabularx}
+\usepackage{booktabs}
+\usepackage{xcolor}
+
+%\renewcommand{\arraystretch}{0.8}
 %
@@
 \DeclareMathOperator{\const}{Const.}
+\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
 
 \usepackage{graphicx}
+
 %
@@
 \graphicspath{{../Figures/}}
+
+\input{01-Acronyms}
 
@@
 %
-\author{Oscar Esteban\inst{1,2} \and
-        Alessandro Daducci \inst{2} \and
-		Meritxell Bach-Cuadra \inst{2,3} \and
-        Jean-Philippe Thiran \inst{2,3} \and
-        Andr\'es Santos \inst{1} \and
-        Dominique Zosso \inst{2,4}}
+%\author{Oscar Esteban\inst{1,2} \and
+%        Alessandro Daducci \inst{2} \and
+%		Meritxell Bach-Cuadra \inst{2,3} \and
+%        Jean-Philippe Thiran \inst{2,3} \and
+%        Andr\'es Santos \inst{1} \and
+%        Dominique Zosso \inst{2,4}}
 %
-\authorrunning{Oscar Esteban et al.} % abbreviated author list (for running head)
+%\authorrunning{Oscar Esteban et al.} % abbreviated author list (for running head)
 
-\institute{Biomedical Image Technologies (BIT),
-ETSI Telecomunicaci\'on - Universidad Polit\'ecnica de Madrid
-and CIBER-BBN, \\
-Av. Complutense 30, E-28040 Madrid, Spain\\
-\email{phd@oscaresteban.es},\\
-% WWW home page:
-%\texttt{http://users/\homedir iekeland/web/welcome.html}
-\and
-Signal Processing Laboratory (LTS5),
-\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL)\\
-EPFL STI IEL LTS5, Station 11, CH-1015 Lausanne, Switzerland
-\and
-Department of Radiology, University Hospital Center (CHUV) and University of Lausanne (UNIL) \\
-Rue du Bugnon 46, CH-1011 Lausanne, Switzerland
-\and
-Department of Mathematics, 
-University of California, Los Angeles (UCLA) \\
-520 Portola Plaza, Box 951555, Los Angeles, CA 90095-1555, USA
-}
+%\institute{Biomedical Image Technologies (BIT),
+%ETSI Telecomunicaci\'on - Universidad Polit\'ecnica de Madrid
+%and CIBER-BBN, \\
+%Av. Complutense 30, E-28040 Madrid, Spain\\
+%\email{phd@oscaresteban.es},\\
+%\and
+%Signal Processing Laboratory (LTS5),
+%\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL)\\
+%EPFL STI IEL LTS5, Station 11, CH-1015 Lausanne, Switzerland
+%\and
+%Dept. of Radiology, University Hospital Center (CHUV) and University of Lausanne (UNIL) \\
+%Rue du Bugnon 46, CH-1011 Lausanne, Switzerland
+%\and
+%Department of Mathematics, 
+%University of California, Los Angeles (UCLA) \\
+%520 Portola Plaza, Box 951555, Los Angeles, CA 90095-1555, USA
+%}
 
 \maketitle              % typeset the title of the contribution
-
-\input{01-Acronyms}
 
@@
 
-\bibliography{99-References}
+%\bibliography{99-References}
+\bibliography{98-References-short}
 

### 2012-12-22 Publications/2013-IPMI/Latex/03-Introduction.tex
Anonimized paper & Glossaries package
Changed from acronym to glossaries to be able to handle
capitalization

@@
 %
-\ac{dwi} is a widely used family of \ac{mr} techniques
+\Gls{dwi} is a widely used family of \gls{mr} techniques
 \citep{sundgren_diffusion_2004} which recently has accounted for a growing
 interest in its application to whole-brain structural connectivity analysis.
-This emerging field, coined in 2005 as \emph{\ac{mr} Connectomics}
+This emerging field, coined in 2005 as \emph{\gls{mr} Connectomics}
 \citep{hagmann_diffusion_2005,sporns_human_2005}, currently includes a
 large amount of imaging techniques for acquisition, processing, and analysis
-specifically tuned for \ac{dwi} data.
+specifically tuned for \gls{dwi} data.
 
@@
 towards reliable structural information
-about the neuronal tracts from \ac{dwi} \cite{johansen-berg_using_2009,
+about the neuronal tracts from \gls{dwi} \cite{johansen-berg_using_2009,
 jones_white_2012}. The earlier stages of these
@@
 In this work, we will refer as brain tissue segmentation to the precise
-delineation of the \ac{csf}-\ac{gm} and \ac{gm}-\ac{wm} interface surfaces.
+delineation of the \gls{csf}-\gls{gm} and \gls{gm}-\gls{wm} interface surfaces.
 This segmentation is an processing step on which strongly rely further
-tasks. In tractography, a high-standard \ac{wm} mask is required. Otherwise,
+tasks. In tractography, a high-standard \gls{wm} mask is required. Otherwise,
 there is an important risk for the algorithm to lose fiber bundles. This
 requirement is usually solved in practice by plainly thresholding the 
-\ac{fa}, a well-known scalar map derived from \ac{dwi} which depicts 
+\gls{fa}, a well-known scalar map derived from \gls{dwi} which depicts 
 the isotropy of water diffusion inside the brain. 
 A precise location of the
-\ac{gm}-\ac{wm} surface is also essential in the final steps to
+\gls{gm}-\gls{wm} surface is also essential in the final steps to
 achieve a consistent parcellisation of the cortex to represent the nodes 
 of the output network. This parcellisation is generally defined in a 
-high-resolution and better understood structural \ac{mri} of the same 
-subject (e.g. \ac{t1} and/or \ac{t2} weighted acquisitions). %Conversely, this problem is resolved with non-linear registration of a structural \ac{mri} of the subject to the \ac{dwi} data. 
+high-resolution and better understood structural \gls{mri} of the same 
+subject (e.g. \gls{t1} and/or \gls{t2} weighted acquisitions). %Conversely, this problem is resolved with non-linear registration of a structural \gls{mri} of the subject to the \gls{dwi} data. 
 Even though some efforts have addressed
 the study of the robustness of tractography with respect to intra-subject variability
-\cite{wakana_reproducibility_2007,heiervang_between_2006}, these results are restricted to some regions of the brain, only. Therefore, robust
+\cite{heiervang_between_2006,wakana_reproducibility_2007}, these results are restricted to some regions of the brain, only. Therefore, robust
 and precise segmentation methods are required in the whole-brain application.
 
-The \ac{dwi} data is usually obtained with \ac{epi}
+The \gls{dwi} data is usually obtained with \gls{epi}
 acquisition techniques, that often suffer from severe distortions due to 
@@
 part of the brain, along the phase-encoded direction. Some methodologies have
-been developed and generically named as \emph{\ac{epi}-unwarp} techniques
+been developed and generically named as \emph{\gls{epi}-unwarp} techniques
 \cite{holland_efficient_2010,hsu_correction_2009,jezzard_characterization_2005,
@@
 methodologies do not make use of the field-mapping, compensating the distortion
-with non-linear registration from structural \ac{mri} or other means
+with non-linear registration from structural \gls{mri} or other means
 \citep{andersson_modeling_2001}. To our knowledge, there exists no study
-on the impact of the \ac{epi} distortion on the variability of tractography
+on the impact of the \gls{epi} distortion on the variability of tractography
 results.
 
-Therefore, the problems of precise segmentation in \ac{dwi}-space and the 
+Therefore, the problems of precise segmentation in \gls{dwi}-space and the 
 spatial mapping between these contours and the corresponding surfaces in 
 anatomical images bear significant redundancy. Once the spatial relationship 
-between \ac{t1} and \ac{dwi} space is established, the contours which are 
-readily available in \ac{t1} space, can simply be projected on to the 
-\ac{dwi}-data. Conversely, if a precise delineation in \ac{dwi}-space 
-was achieved, the spatial mapping with \ac{t1}-space could be derived 
+between \gls{t1} and \gls{dwi} space is established, the contours which are 
+readily available in \gls{t1} space, can simply be projected on to the 
+\gls{dwi}-data. Conversely, if a precise delineation in \gls{dwi}-space 
+was achieved, the spatial mapping with \gls{t1}-space could be derived 
 from one-to-one correspondences on the contours. However, neither segmentation 
 nor registration can be performed flawlessly, if considered independently. 
-The significant benefits of exploiting the anatomical \ac{mri} when 
-segmenting the \ac{dwi} data have been demonstrated \cite{zollei_improved_2010}, 
+The significant benefits of exploiting the anatomical \gls{mri} when 
+segmenting the \gls{dwi} data have been demonstrated \cite{zollei_improved_2010}, 
 justifying the use of the shape prior information.
@@
 methods into three groups. The first group typically adds a shape prior term to 
-the energy functional of an evolving active contour~\citep{Chen2002,
-Paragios2003,Vemuri2003a,Yezzi2003a,Gastaud2004,Chan2005,Cremers2006,Bresson2006a} %removed Rousson2002,Ayvaci2007,Schmid2008
+the energy functional of an evolving active contour~\citep{Bresson2006a,Chan2005,
+Chen2002,Cremers2006,Gastaud2004,Paragios2003,Vemuri2003a,Yezzi2003a} %removed Rousson2002,Ayvaci2007,Schmid2008
 %{\color{red}{(FIXME:I would select from here the most important,
@@
 from this average shape. Closely related to this group are atlas-based segmentation
-methods \citep{Pohl2005,Pohl2006,Wang2006,Gorthi2009,Gorthi2011}, where the prior 
+methods \citep{Gorthi2011,Gorthi2009,Pohl2005,Pohl2006,Wang2006}, where the prior 
 imposes consistent voxel-based classification of contiguous regions. Here, the 
-presence of more structures than one unique \ac{roi} helps aligning the target image 
+presence of more structures than one unique \gls{roi} helps aligning the target image 
 with the atlas in a hierarchical fashion. Finally, the third group generalizes 
@@
 solving the segmentation and distortion challenges, by exploiting as strong 
-shape-prior the detailed anatomy extracted from anatomical \ac{mri}. Indeed,
+shape-prior the detailed anatomy extracted from anatomical \gls{mri}. Indeed,
 hereafter we consider the segmentation in anatomical images as a solved problem.
@@
 the remaining spatial deformation between anatomical and diffusion space is 
-due to \ac{mr} distortion. Finally, we need to establish precise spatial 
+due to \gls{mr} distortion. Finally, we need to establish precise spatial 
 correspondence between the surfaces in both spaces, including the tangential 
 direction for parcellation. Therefore, we can reduce the problem to finding 
-the differences of spatial distortion in between anatomical and \ac{dw} space.
+the differences of spatial distortion in between anatomical and \gls{dw} space.
 We thus reformulate the segmentation problem as an inverse problem, where we 
@@
 from the structural space into the diffusion space, such that the structural 
-contours segment optimally the \ac{dwi} data. In the process, the one-to-one 
+contours segment optimally the \gls{dwi} data. In the process, the one-to-one 
 correspondence between the contours in both spaces is guaranteed, and projection 
-of parcellisation to \ac{dw} space is implicit and consistent.
+of parcellisation to \gls{dw} space is implicit and consistent.
 
@@
 synthetic examples. The first example is a scalar sulcus model, where the 
-\ac{csf}-\ac{gm} boundary particularly suffers from \ac{pve} and can only be 
+\gls{csf}-\gls{gm} boundary particularly suffers from \gls{pve} and can only be 
 segmented correctly thanks to the shape prior and its coupling with the inner, 
-\ac{gm}-\ac{wm} boundary through the imposed deformation field regularity. 
-The second case deals with more realistic \ac{dwi} data stemming from 
+\gls{gm}-\gls{wm} boundary through the imposed deformation field regularity. 
+The second case deals with more realistic \gls{dwi} data stemming from 
 phantom simulations of a simplistic brain data. Again, we show that the 
-proposed model successfully segments the \ac{dwi} data based on two derived 
-scalar features, namely \ac{fa} and \ac{md}, while establishing an estimate 
+proposed model successfully segments the \gls{dwi} data based on two derived 
+scalar features, namely \gls{fa} and \gls{md}, while establishing an estimate 
 of the dense distortion field.

### 2012-12-22 Publications/2013-IPMI/Latex/04-Methods.tex
Anonimized paper & Glossaries package
Changed from acronym to glossaries to be able to handle
capitalization

@@
 %
-\subsection{\acl{acwe}-like variational segmentation model}
+\subsection{\Acrlong{acwe} based segmentation model}
 %
 Let us denote $\{c_i\}_{i=1..N_c}$ the nodes of a shape prior surface. In
-our application, a precise \ac{wm}-\ac{gm} interface extracted from a
+our application, a precise \gls{wm}-\gls{gm} interface extracted from a
 high-resolution reference volume. All the formulations can be naturally
 extended to include more shape priors. On the other hand, we have a 
-number of \ac{dwi}-derived features at each
+number of \gls{dwi}-derived features at each
 voxel of the volume. Let us denote by $x$ the voxel and 
@@
 %
-The transformation from reference into \ac{dwi} coordinate space is 
+The transformation from reference into \gls{dwi} coordinate space is 
 achieved through a dense deformation field $u(x)$, such that:
@@
 Based on the current estimate of the distortion $u$, we can compute 
-``expected samples'' within the shape prior projected into the \ac{dwi}.
-Thus, we now estimate region descriptors of the \ac{dwi} features 
-$f(x)$ of the regions defined by the priors in \ac{dwi} space.
+``expected samples'' within the shape prior projected into the \gls{dwi}.
+Thus, we now estimate region descriptors of the \gls{dwi} features 
+$f(x)$ of the regions defined by the priors in \gls{dwi} space.
 %
 Using Gaussian distributions as region descriptors, we propose an
-\ac{acwe}-like, piece-wise constant, variational image segmentation
+\gls{acwe}-like, piece-wise constant, variational image segmentation
 model (where the unknown is the deformation field)
@@
 depend on the deformation field $u$. Note
-that minimizing this energy, $\argmin_u\{E\}$, yields the \ac{map} 
+that minimizing this energy, $\argmin_u\{E\}$, yields the \gls{map} 
 estimate of a piece-wise smooth image model affected by Gaussian 
@@
 These regularity terms ensure that the segmenting contours in 
-\ac{dwi} space are still close to their native shape. The model
+\gls{dwi} space are still close to their native shape. The model
 easily allows to incorporate inhomogeneous and anisotropic 
 regularization \cite{nagel_investigation_1986} to better regularize
-the \ac{epi} distortion. \\
+the \gls{epi} distortion. \\
 %
@@
 \begin{align}
-\frac{u^{t+1}-u^t}{\tau} &= - \sum_{i=1}^{N_c} \left[ \Delta E(f(c_i'))  \hat{n}_{c_i'} \Psi_{c_i}(x) \right] -\alpha u^{t+1} + \beta\Delta u^{t+1}
+\frac{u^{t+1}-u^t}{\tau} &= - \sum_{i=1}^{N_c} \left[ e(f(c_i'))  \hat{n}_{c_i'} \Psi_{c_i}(x) \right] -\alpha u^{t+1} + \beta\Delta u^{t+1}
 \end{align}
@@
 restricted the number of priors to only 1. We also defined 
-$\Delta E(f(c_i')) = E_{out}(f(c_i')) - E_{in}(f(c_i'))$, 
-and $E_R(f) = \sqrt{(f-\mu_R)^T\Sigma_R^{-1}(f-\mu_R)}$.
+$e(f(c_i')) = E_{out}(f(c_i')) - E_{in}(f(c_i'))$, 
+and $E_R(f) = {(f-\mu_R)^T\Sigma_R^{-1}(f-\mu_R)}$.
 We applied a spectral approach to solve this implicit scheme:
@@
 u^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{u^t/\tau
-- \sum_{i=1}^{N_c} \left[ \Delta E(f(c_i')) \hat{n}_{c_i'} \Psi_{c_i}(x) \right]  \}}{\mathcal{F}\{(1/\tau+\alpha)\mathcal{I}-\beta\Delta\}} \right\}
+- \sum_{i=1}^{N_c} \left[e(f(c_i')) \hat{n}_{c_i'} \Psi_{c_i}(x) \right]  \}}{\mathcal{F}\{(1/\tau+\alpha)\mathcal{I}-\beta\Delta\}} \right\}
 \end{equation}

### 2012-12-22 Publications/2013-IPMI/Latex/045-Experiments.tex
Anonimized paper & Glossaries package
Changed from acronym to glossaries to be able to handle
capitalization

@@
 the connectivity pipelines consists of having 
-a reliable segmentation obtained from the high resolution \ac{t1} 
+a reliable segmentation obtained from the high resolution \gls{t1} 
 reference image. Therefore, a precise location of the tissue interfaces
@@
 The first simplified model to test the approach is inspired by a problem 
-shown for coupled \ac{csf}/\ac{gm} and \ac{gm}/\ac{wm} segmentation in \citep{macdonald_automated_2000}. Authors note that ``partial volume effects 
+shown for coupled \gls{csf}/\gls{gm} and \gls{gm}/\gls{wm} segmentation in \citep{macdonald_automated_2000}. Authors note that ``partial volume effects 
 blur the distinction between closely adjacent surfaces in deep sulci, leading 
@@
 Here, we reproduce their model on a volume consisting of three 
-piecewise-constant parts: a notched ball representing the \ac{wm} with a single 
-sulcus ($\mu_{WM} = 0.8$), a cortical sheet of \ac{gm} obtained through dilation 
-of the \ac{wm} ($\mu_{GM} = 0.5$), and the surrounding background representing 
-\ac{csf} ($\mu_{CSF} = 0.2$). The volume is then affected by additive Gaussian 
+piecewise-constant parts: a notched ball representing the \gls{wm} with a single 
+sulcus ($\mu_{WM} = 0.8$), a cortical sheet of \gls{gm} obtained through dilation 
+of the \gls{wm} ($\mu_{GM} = 0.5$), and the surrounding background representing 
+\gls{csf} ($\mu_{CSF} = 0.2$). The volume is then affected by additive Gaussian 
 noise, effectively creating uniform standard deviation of $\sigma = 0.045$ per 
@@
 As illustrated in \autoref{fig:sulcusmodel}, conventional single surface 
-segmentation of the \ac{csf}/\ac{gm} boundary misses to capture the sulcus in 
+segmentation of the \gls{csf}/\gls{gm} boundary misses to capture the sulcus in 
 its full depth. With our proposed model, we expect the joint 
-segmentation-registration to be driven largely by the inner, \ac{gm}/\ac{wm}
+segmentation-registration to be driven largely by the inner, \gls{gm}/\gls{wm}
 contour that exhibits sufficient contrast and lesser partial volume effects. 
@@
 and characterize its possibilities with diffusion data,
-we simulated the \acs{dw} signal of a synthetic phantom from a model
+we simulated the \gls{dw} signal of a synthetic phantom from a model
 consisting of several spherical shapes emulating
@@
 where $\varepsilon_{1,2}$ are Gaussian distributed with zero mean
-and standard deviation $\sigma = S_0 / \mathit{SNR}$ and \acs{snr}
-is the \ac{snr} on the $S_0$ image.
+and standard deviation $\sigma = S_0 / \mathit{SNR}$ and \gls{snr}
+is the \gls{snr} on the $S_0$ image.
 
 \paragraph{Derived scalar features}
-The target \ac{dwi} data is characterized by its distortions and its
+The target \gls{dwi} data is characterized by its distortions and its
 low resolution (typically around $2.2x2.2x3mm^3$). Depending on the
 posterior reconstruction methodology and the angular resolution
-intended, the \ac{dwi} raw data has to be processed in order to
+intended, the \gls{dwi} raw data has to be processed in order to
 extract the information in a manageable manner. The properties of
@@
 studied by \cite{ennis_orthogonal_2006}. Based on their
-findings, the proposed energy model adapts to the \ac{fa}~\eqref{eq:fa} 
-and \ac{md}~\eqref{eq:md} for their properties. 
-Whereas \ac{fa} \ describes the \emph{shape} of diffusion, 
-the \ac{md} depicts the \emph{magnitude} of the process.
+findings, the proposed energy model adapts to the \gls{fa}~\eqref{eq:fa} 
+and \gls{md}~\eqref{eq:md} for their properties. 
+Whereas \gls{fa} \ describes the \emph{shape} of diffusion, 
+the \gls{md} depicts the \emph{magnitude} of the process.
 
@@
 Second, together they contain most of the information that is
-usually extracted from the \ac{dwi}-derived scalar maps. \\
+usually extracted from the \gls{dwi}-derived scalar maps. \\
 
@@
 For this model, we created manually a sound distortion visually similar
-to real \ac{epi} distortions. We interpolated the distortion to a 
-dense deformation field, necessary for warping the raw \ac{dwi} simulated
+to real \gls{epi} distortions. We interpolated the distortion to a 
+dense deformation field, necessary for warping the raw \gls{dwi} simulated
 data. Once the signal was deformed, we proceeded to reconstruct the
-\ac{dti} and subsequently obtained the scalars of interest (\ac{fa}, \ac{md}).
+\gls{dti} and subsequently obtained the scalars of interest (\gls{fa}, \gls{md}).
 Finally, we estimated their parameters using the tissue probability
@@
 \begin{table}
-\caption{Model means and covariances of \acl{fa} and \acl{md} estimated from the reconstructed simulated \ac{dwi} images for each modeled tissue, \acl{wm},\acl{gm}, and \acl{csf}.
+\caption{Model means and covariances of \acrfull{fa} and \acrfull{md} estimated from the reconstructed simulated \gls{dwi} images for each modeled tissue, \acrfull{wm},\acrfull{gm}, and \acrfull{csf}.
 As expected, the two scalar features are complementary and the three tissues can well be discriminated.
@@
 \cmidrule(r){1-1}\cmidrule(r){2-2}\cmidrule(r){3-3}\cmidrule{4-4}
-\ac{wm}  & 0.778 & 6.94\e{-4} & 
+\gls{wm}  & 0.778 & 6.94\e{-4} & 
    $\begin{pmatrix}
@@
 %
-\ac{gm}  & 0.119 & 8.95\e{-4} &
+\gls{gm}  & 0.119 & 8.95\e{-4} &
    $\begin{pmatrix}
@@
 %
-\ac{csf} & 0.103 & 2.99\e{-3} &
+\gls{csf} & 0.103 & 2.99\e{-3} &
    $\begin{pmatrix}

### 2012-12-22 Publications/2013-IPMI/Latex/05-Results.tex
Anonimized paper & Glossaries package
Changed from acronym to glossaries to be able to handle
capitalization

@@
 c) 3D rendering of the deformed priors, with the traces of the two selected slices highlighted.
-%In red and blue colors the obtained \ac{wm}/\ac{gm} 
-%and \ac{gm}/\ac{csf} contours for the deformed image, respectively.
-%In white color, the initial shape priors for the \ac{wm}/\ac{gm}
+%In red and blue colors the obtained \gls{wm}/\gls{gm} 
+%and \gls{gm}/\gls{csf} contours for the deformed image, respectively.
+%In white color, the initial shape priors for the \gls{wm}/\gls{gm}
 %interface.
@@
 \end{tabular}
-\caption{First row presents several slices along Z axis of the distorted \ac{fa} map and
-the undistorted \ac{wm}-\ac{gm} and \ac{wm}-\ac{csf} contours given as shape priors. Second row contains the same map, now with the contours after joint segmentation-registration.}
+\caption{First row presents several slices along Z axis of the distorted \gls{fa} map and
+the undistorted \gls{wm}-\gls{gm} and \gls{wm}-\gls{csf} contours given as shape priors. Second row contains the same map, now with the contours after joint segmentation-registration.}
 \label{fig:fa}

### 2012-12-22 Publications/2013-IPMI/Latex/06-Conclusion.tex
Anonimized paper & Glossaries package
Changed from acronym to glossaries to be able to handle
capitalization

@@
 %
-A novel application for the \ac{acwe} framework is proposed,
+A novel application for the \gls{acwe} framework is proposed,
 with the aim at recovering the displacement field underlying 
-the \ac{epi} geometrical distortions. Exploiting the segmentation
-properties of the \ac{acwe} and optimizing the displacement
+the \gls{epi} geometrical distortions. Exploiting the segmentation
+properties of the \gls{acwe} and optimizing the displacement
 field, we describe a registration-segmentation methodology that
 simultaneously segmented and restored the distortion on 
-\ac{dwi}-like synthetic data. Visual results and quantitative
+\gls{dwi}-like synthetic data. Visual results and quantitative
 results are provided. \\
@@
 of overcoming the described challenges (segmentation and 
-\ac{epi} distortion correction) in one single step.\\
+\gls{epi} distortion correction) in one single step.\\
 
 We conclude stressing on the importance of tackling with
-the numerous challenges that exist on the \ac{dwi} data 
+the numerous challenges that exist on the \gls{dwi} data 
 processing in order to achieve reliable results on the

### 2012-12-22 Publications/2013-IPMI/Latex/00-Main.tex
Included Alessandro's suggestions
And re-structured the text regarding the model 1

@@
 
-\title{Simultaneous segmentation and distortion correction on diffusion weighted MR using shape priors}
+\title{Simultaneous segmentation and distortion correction on diffusion MRI data using shape priors}
 %%
-\titlerunning{Simultaneous segmentation and distortion correction on DWI}  % abbreviated title (for running head)
+\titlerunning{Simultaneous segmentation and distortion correction on dMRI}  % abbreviated title (for running head)
 %                                     also used for the TOC unless

### 2012-12-22 Publications/2013-IPMI/Latex/03-Introduction.tex
Included Alessandro's suggestions
And re-structured the text regarding the model 1

@@
 The whole-brain connectivity analysis has given rise to some challenges
-towards reliable structural information
-about the neuronal tracts from \gls{dwi} \cite{johansen-berg_using_2009,
-jones_white_2012}. The earlier stages of these
-processing pipelines generally include two necessary steps, brain tissue
-segmentation on the diffusion space and the correction of geometrical
-distortions inherent to the imaging techniques \citep{hagmann_mr_2012}.
+towards reliable structural information about the neuronal tracts 
+from \gls{dwi} \cite{johansen-berg_using_2009,jones_white_2012}. 
+The earlier stages of these processing pipelines generally include 
+two necessary steps, brain tissue segmentation on the diffusion space 
+and the correction of geometrical distortions inherent to the acquisition 
+sequence \citep{hagmann_mr_2012}.
 
@@
 delineation of the \gls{csf}-\gls{gm} and \gls{gm}-\gls{wm} interface surfaces.
-This segmentation is an processing step on which strongly rely further
-tasks. In tractography, a high-standard \gls{wm} mask is required. Otherwise,
-there is an important risk for the algorithm to lose fiber bundles. This
-requirement is usually solved in practice by plainly thresholding the 
+This segmentation is a processing step on which strongly rely further
+tasks. An accurate brain tissue segmentation is required
+to filter the fiber bundles obtained with \gls{dwi} tractography. 
+This requirement is usually solved in practice by plainly thresholding the 
 \gls{fa}, a well-known scalar map derived from \gls{dwi} which depicts 
-the isotropy of water diffusion inside the brain. 
-A precise location of the
-\gls{gm}-\gls{wm} surface is also essential in the final steps to
-achieve a consistent parcellisation of the cortex to represent the nodes 
-of the output network. This parcellisation is generally defined in a 
-high-resolution and better understood structural \gls{mri} of the same 
-subject (e.g. \gls{t1} and/or \gls{t2} weighted acquisitions). %Conversely, this problem is resolved with non-linear registration of a structural \gls{mri} of the subject to the \gls{dwi} data. 
-Even though some efforts have addressed
-the study of the robustness of tractography with respect to intra-subject variability
-\cite{heiervang_between_2006,wakana_reproducibility_2007}, these results are restricted to some regions of the brain, only. Therefore, robust
-and precise segmentation methods are required in the whole-brain application.
+the isotropy of water diffusion inside the brain. Also,
+it is necessary to locate the intersections of fiber bundles and \gls{gm}.
+Moreover, a precise location of the \gls{gm}-\gls{wm} surface is also essential 
+to finally achieve a consistent parcellisation of the cortex 
+to represent the nodes of the output network. This parcellisation 
+is generally defined in a high-resolution and better understood structural 
+\gls{mri} of the same 
+subject (e.g. \gls{t1} and/or \gls{t2} weighted acquisitions). Even though
+some efforts have addressed the study of the robustness of tractography with
+respect to intra-subject variability \cite{heiervang_between_2006,
+wakana_reproducibility_2007}, these results are restricted to certain regions 
+of the brain, only. Therefore, robust and precise segmentation methods are 
+required in the whole-brain application. The problem is challenging due to
+the much lower resolution of \gls{dwi} (typically around 
+$2.5\times2.5\times2.5mm^3.$) compared to structural \gls{mri}, and the
+existence of geometrical distortions.
 
-The \gls{dwi} data is usually obtained with \gls{epi}
-acquisition techniques, that often suffer from severe distortions due to 
-local field inhomogeneities. Generally, it is appreciated in the anterior
-part of the brain, along the phase-encoded direction. Some methodologies have
+\gls{dwi} data are usually acquired with \gls{epi} sequences 
+as they allow for very fast acquisitions but they are known to 
+suffer from geometrical distortions due to local field inhomogeneities. 
+These artifacts happen along the phase-encoding direction, and are most 
+appreciable in the front part of the brain for the strong air/tissue 
+interface around the frontal sinuses. Some methodologies have
 been developed and generically named as \emph{\gls{epi}-unwarp} techniques
@@
 between \gls{t1} and \gls{dwi} space is established, the contours which are 
-readily available in \gls{t1} space, can simply be projected on to the 
+readily available in \gls{t1} space can simply be projected on to the 
 \gls{dwi}-data. Conversely, if a precise delineation in \gls{dwi}-space 
@@
 the energy functional of an evolving active contour~\citep{Bresson2006a,Chan2005,
-Chen2002,Cremers2006,Gastaud2004,Paragios2003,Vemuri2003a,Yezzi2003a} %removed Rousson2002,Ayvaci2007,Schmid2008
-%{\color{red}{(FIXME:I would select from here the most important,
-%removing directly those appearing in conferences and never published on journal. Better
-%if we found a review paper/book chapter. I don't like so many references on a row)}}.
+Chen2002,Cremers2006,Gastaud2004,Paragios2003,Vemuri2003a,Yezzi2003a}.
 These methods generally have a explicit description of the expected relative boundary 
@@
 solving the segmentation and distortion challenges, by exploiting as strong 
-shape-prior the detailed anatomy extracted from anatomical \gls{mri}. Indeed,
-hereafter we consider the segmentation in anatomical images as a solved problem.
+shape-prior the detailed morphology extracted from high-resolution anatomical 
+\gls{mri}. Indeed, hereafter we assume this segmentation problem in anatomical 
+images as a solved by widely used available procedures.
 Moreover, the shape prior is of very ``strong'' nature, since it is specific to 
@@
 the remaining spatial deformation between anatomical and diffusion space is 
-due to \gls{mr} distortion. Finally, we need to establish precise spatial 
+due to \gls{epi} distortions. Finally, we need to establish precise spatial 
 correspondence between the surfaces in both spaces, including the tangential 

### 2012-12-22 Publications/2013-IPMI/Latex/045-Experiments.tex
Included Alessandro's suggestions
And re-structured the text regarding the model 1

@@
 and characterize its possibilities with diffusion data,
-we simulated the \gls{dw} signal of a synthetic phantom from a model
-consisting of several spherical shapes emulating
-the different brain tissues (see \autoref{fig:fa}, first row). 
-We reconstructed the signal with standard procedures to 
-approximate the environment to the real one at maximum. \\
+we built a synthetic phantom from a model consisting of several 
+spherical shapes emulating the different brain tissues (see 
+\autoref{fig:fa}, first row). We simulated a single-shell
+acquisition at $b=1000$ \eqref{eqn:MultiTensor} with 30 samples
+equally distributed on one hemisphere, corresponding to a standard
+\gls{dti} acquisition commonly used in clinical practice.
+We generated a synthetic displacement field to produce
+an \gls{epi}-like distortion on the \gls{dwi} signal. Finally,
+we reconstructed the deformed signal, and generated \gls{dwi}-derived
+scalar maps (specifically the \gls{fa} and \gls{md} maps).
+Then, the scalar maps were placed as features in 
+\eqref{eq:complete_energy} to drive our segmentation approach, 
+finding the location of the \gls{wm}-\gls{gm}
+and the \gls{csf}-\gls{wm} interfaces in the distorted space. \\
 
 \paragraph{Signal simulation}
-To numerically simulate the MRI signal attenuation when applying a diffusion 
+To numerically simulate the \gls{mri} signal attenuation when applying a diffusion 
 gradient in a voxel with $N$ fiber populations we made use of the standard 
@@
 \label{eqn:MultiTensor}
-S(\vec{q}) / S_{0} = \sum_{i=1}^{N} f_{i} \, \exp{ \left( -b \, \vec{q}^{T} \, \mathbf{D}_i \, \vec{q}\right) } + f_{iso} \, \exp{ \left( -b \, \mathbf{D}_{iso} \right) } ,
+S(q) / S_{0} = \sum_{i=1}^{N} f_{i} \, \exp{ \left( -b \, q^{T} \, \mathbf{D}_i \, q\right) } + f_{iso} \, \exp{ \left( -b \, \mathbf{D}_{iso} \right) }  + f_{gm} \, \exp{ \left( -b \, \mathbf{D}_{gm} \right) } ,
 \end{equation}
 %
-where $\vec{q} \in \mathbb{S}^2$ is the direction of the diffusion gradient 
-applied, $b$ is the b-value accounting for its strength, $S_0 \equiv S(\vec{0})$ 
-is the signal with no diffusion weighting, $f_i$ and $\mathbf{D}_i$ are 
-respectively the volume fraction and the diffusion tensor characterizing the 
-$i$-th fiber population.
-
-The diffusion tensors $\mathbf{D}_i$ and $\mathbf{D}_{iso}$ describe the diffusion 
-processes of each fiber compartment and of contaminations from the CSF. In this work, 
-these quantities have been taken from standard ranges typically observed in the 
-human brain~\cite{Canales-Rodriguez:2009aa}.
+where $q \in \mathbb{S}^2$ is the direction of the diffusion gradient applied,
+$b$ is the b-value accounting for its strength and $S_0 \equiv S(0)$ is the 
+signal with no diffusion weighting. $f_i$ and $\mathbf{D}_i$ are the volume 
+fraction and the diffusion tensor characterizing the $i$-th fiber population, 
+whereas the tensors $\mathbf{D}_{iso}$ and $\mathbf{D}_{gm}$ describe the 
+diffusion processes of partial volumes with \gls{csf} and \gls{gm} within the 
+voxel, which volume fractions are, respectively, $f_{iso}$ and $f_{gm}$, and
+$\sum_i f_i + f_{iso} + f_{gm} = 1$. In this work, the diffusion properties have
+been taken from standard ranges typically observed in the human 
+brain~\cite{Canales-Rodriguez:2009aa}. Moreover we assumed $N = 1$ and $S_0 = 1$ 
+without loss of generality.
 
@@
 
+\paragraph{Simulated \gls{epi} distortion}
+For this model, we created manually a sound distortion to approximate
+the real \gls{epi} distortions. We interpolated the distortion to a 
+dense deformation field, and applied it to generate the deformed \gls{dwi}
+signal.
+
 \paragraph{Derived scalar features}
-The target \gls{dwi} data is characterized by its distortions and its
-low resolution (typically around $2.2x2.2x3mm^3$). Depending on the
-posterior reconstruction methodology and the angular resolution
-intended, the \gls{dwi} raw data has to be processed in order to
-extract the information in a manageable manner. The properties of
+We obtained the local fiber configuration in each voxel with a commonly 
+used \gls{dti} reconstruction tool~\footnote{DTIFIT, included
+in the FMRIB's Software Library (FSL), 
+\url{http://fsl.fmrib.ox.ac.uk/fsl/fsl-4.1.9/fdt/fdt_dtifit.html}}.
+
+The properties of
 the reconstructed tensors and derived scalar maps have been
 studied by \cite{ennis_orthogonal_2006}. Based on their
-findings, the proposed energy model adapts to the \gls{fa}~\eqref{eq:fa} 
-and \gls{md}~\eqref{eq:md} for their properties. 
+findings, we chose \gls{fa}~\eqref{eq:fa} 
+and \gls{md}~\eqref{eq:md} as features for the energy model in driving
+the registration-segmentation process. 
 Whereas \gls{fa} \ describes the \emph{shape} of diffusion, 
@@
 Second, together they contain most of the information that is
-usually extracted from the \gls{dwi}-derived scalar maps. \\
-
-\subsubsection{Simulated distortions}
-
-For this model, we created manually a sound distortion visually similar
-to real \gls{epi} distortions. We interpolated the distortion to a 
-dense deformation field, necessary for warping the raw \gls{dwi} simulated
-data. Once the signal was deformed, we proceeded to reconstruct the
-\gls{dti} and subsequently obtained the scalars of interest (\gls{fa}, \gls{md}).
-Finally, we estimated their parameters using the tissue probability
-distribution maps from the original model (\autoref{table:parameters}).
+usually extracted from the \gls{dwi}-derived scalar maps. The properties
+of brain tissues on the joint distribution of \gls{fa} and \gls{md} are 
+summarized in \autoref{table:parameters}. \\
 

### 2012-12-22 Publications/2013-IPMI/Latex/05-Results.tex
Included Alessandro's suggestions
And re-structured the text regarding the model 1

@@
 b) The same slices and contours after distortion estimation.
-c) 3D rendering of the deformed priors, with the traces of the two selected slices highlighted.
-%In red and blue colors the obtained \gls{wm}/\gls{gm} 
-%and \gls{gm}/\gls{csf} contours for the deformed image, respectively.
-%In white color, the initial shape priors for the \gls{wm}/\gls{gm}
-%interface.
-{\color{red}{This figure is to be replaced by the real result,
-this is a proof of concept of registration that does not solve
-the pv problem. The current figure also removes the outer prior}}}
+c) 3D rendering of the deformed priors, with the traces of the two selected slices highlighted.}
 \label{fig:sulcusmodel_result}

### 2012-12-22 Publications/2013-IPMI/Latex/06-Conclusion.tex
Included Alessandro's suggestions
And re-structured the text regarding the model 1

@@
 with the aim at making the procedure publicly available 
-when ready for sharing with the research community.
+when ready for sharing with the research community.\\
 

### 2012-12-27 Publications/2013-IPMI/Latex/06-Conclusion.tex
Version submitted to IPMI2013

--- 2012-12-22 Publications/2013-IPMI/Latex/06-Conclusion.tex
@@
 \gls{dwi}-like synthetic data. Visual results and quantitative
-results are provided. \\
+results are provided.
 
@@
 of overcoming the described challenges (segmentation and 
-\gls{epi} distortion correction) in one single step.\\
+\gls{epi} distortion correction) in one single step. Additional
+research lines regard with the use of more adequate 
+optimization schemes and the use of an energy model 
+better adapted to the specific nature of the \gls{dwi} data.
 
-We conclude stressing on the importance of tackling with
+We conclude recalling the importance of tackling with
 the numerous challenges that exist on the \gls{dwi} data 

### 2012-12-27 Code/Modules/LevelSets/include/LevelSetsBase.h
Updated code for IPMI 2013

@@ -47,6 +47,7 @@
 #include <itkQuadEdgeMesh.h>
 #include <itkQuadEdgeMeshToQuadEdgeMeshFilter.h>
 #include <itkWarpMeshFilter.h>
+#include <itkMeshSpatialObject.h>
 
 
 #include "SparseToDenseFieldResampleFilter.h"
@@ -92,6 +93,7 @@ public:
 
 	typedef TReferenceImageType                              ReferenceImageType;
 	typedef typename ReferenceImageType::PixelType           PixelType;
+	typedef typename ReferenceImageType::PointType           PixelPointType;
 	typedef typename PixelType::ValueType                    PixelValueType;
 
 	typedef itk::VectorLinearInterpolateImageFunction
@@ -108,6 +110,11 @@ public:
 
 
 	typedef typename std::vector<ContourDeformationPointer>  ContourDeformationList;
+
+	typedef typename itk::MeshSpatialObject
+			                   <ContourDeformationType>      ContourSpatialObject;
+	typedef typename ContourSpatialObject::Pointer           ContourSpatialPointer;
+	typedef typename std::vector<ContourSpatialPointer>      SpatialObjectsVector;
 
 	typedef typename itk::QuadEdgeMeshToQuadEdgeMeshFilter
 			<ContourDeformationType,ContourDeformationType>  ContourCopyType;

### 2013-05-16 Publications/2013-TMI/05-experiments.tex
Added TMI scaffold.
Added a basic structure for TMI manuscript.

--- 2012-12-22 Publications/2013-IPMI/Latex/045-Experiments.tex
@@
 \url{http://fsl.fmrib.ox.ac.uk/fsl/fsl-4.1.9/fdt/fdt_dtifit.html}}. 
-
-The properties of
-the reconstructed tensors and derived scalar maps have been
-studied by \cite{ennis_orthogonal_2006}. Based on their
-findings, we chose \gls{fa}~\eqref{eq:fa} 
-and \gls{md}~\eqref{eq:md} as features for the energy model in driving
-the registration-segmentation process. 
-Whereas \gls{fa} \ describes the \emph{shape} of diffusion, 
-the \gls{md} depicts the \emph{magnitude} of the process.
+The properties of the reconstructed tensors and derived scalar maps have
+been studied by \cite{ennis_orthogonal_2006}. Based on their
+findings, \gls{fa}~\eqref{eq:fa} and \gls{md}~\eqref{eq:md} are
+considered complementary features, and therefore we selected them for the 
+energy model \eqref{eq:complete_energy} in driving the 
+registration-segmentation process. 
+Whereas \gls{fa} informs mainly about the \emph{shape} of diffusion, 
+the \gls{md} is more related to the \emph{magnitude} of the process:
 
@@
 Second, together they contain most of the information that is
-usually extracted from the \gls{dwi}-derived scalar maps. The properties
-of brain tissues on the joint distribution of \gls{fa} and \gls{md} are 
-summarized in \autoref{table:parameters}. \\
+usually extracted from the \gls{dwi}-derived scalar maps
+\cite{ennis_orthogonal_2006}. The different region properties 
+$(\mathbf{\mu}_R,\Sigma_R)$ measured for the \gls{fa} and \gls{md}
+joint distribution are summarized in \autoref{table:parameters}. \\
 
 \begin{table}
-\caption{Model means and covariances of \acrfull{fa} and \acrfull{md} estimated from the reconstructed simulated \gls{dwi} images for each modeled tissue, \acrfull{wm},\acrfull{gm}, and \acrfull{csf}.
-As expected, the two scalar features are complementary and the three tissues can well be discriminated.
-}
+\caption{Model means and covariances of \gls{fa} and \gls{md} estimated 
+from the reconstructed simulated \gls{dwi} images for each modeled tissue, 
+\gls{wm}, \gls{gm}, and \gls{csf}. As expected, the two scalar features 
+are complementary and the three tissues can well be discriminated. }
 \label{table:parameters}

### 2013-05-16 Publications/2013-TMI/03-introduction.tex

--- 2012-12-22 Publications/2013-IPMI/Latex/03-Introduction.tex
@@
 from \gls{dwi} \cite{johansen-berg_using_2009,jones_white_2012}. 
-The earlier stages of these processing pipelines generally include 
-two necessary steps, brain tissue segmentation on the diffusion space 
-and the correction of geometrical distortions inherent to the acquisition 
+Here, we shall address brain tissue segmentation on diffusion space 
+and correction of geometrical distortions inherent to the acquisition 
 sequence \citep{hagmann_mr_2012}.
@@
 delineation of the \gls{csf}-\gls{gm} and \gls{gm}-\gls{wm} interface surfaces.
-This segmentation is a processing step on which strongly rely further
-tasks. An accurate brain tissue segmentation is required
+An accurate brain tissue segmentation is required
 to filter the fiber bundles obtained with \gls{dwi} tractography. 
@@
 the much lower resolution of \gls{dwi} (typically around 
-$2.5\times2.5\times2.5mm^3.$) compared to structural \gls{mri}, and the
+$2.5\times2.5\times2.5mm^3$) compared to structural \gls{mri}, and the
 existence of geometrical distortions.
@@
 \gls{dwi} data are usually acquired with \gls{epi} sequences 
-as they allow for very fast acquisitions but they are known to 
+as they allow for very fast acquisitions, but they are known to 
 suffer from geometrical distortions due to local field inhomogeneities. 
@@
 appreciable in the front part of the brain for the strong air/tissue 
-interface around the frontal sinuses. Some methodologies have
-been developed and generically named as \emph{\gls{epi}-unwarp} techniques
+interface around the frontal sinuses. A number of methodologies have
+been developed to correct for the distortion, and are generically named 
+as \emph{\gls{epi}-unwarp} techniques
 \cite{holland_efficient_2010,hsu_correction_2009,jezzard_characterization_2005,

### 2013-05-16 Publications/2013-TMI/06-results.tex

--- 2012-12-22 Publications/2013-IPMI/Latex/05-Results.tex
@@
 \autoref{fig:sulcusmodel_result} provides visual assessment
-for this result. With 16x16x16 control points, 
-computation time for this model was around
-10 minutes in a {\color{red}{(specific machine details)}}.
+for this result. With 16x16x16 control points and an 
+approximate total of 29,000 nodes contained by the two prior 
+surfaces, computation time for this model was around
+14 minutes in an \textit{Intel\textsuperscript{\textregistered} 
+Core\textsuperscript{\texttrademark} i5 CPU M 430 @ 2.27GHz} and
+4GB RAM.
 
@@
 \subsection{Simulated diffusion data}
+\label{sec:simulated_dwi}
 %
 The proposed method successfully reverted the synthetic distortion
-field we applied to the data. With 16x16x16 control points, the
-displacements field is dense enough to correctly represent the
-synthetic field. Second row in \autoref{fig:fa} shows the fitted
-contours obtained by using the original surfaces of the model
-as shape priors, with a constant translation of (5.0mm,10mm.,-5.0mm)
+field we applied to the data. Second row in \autoref{fig:fa} shows 
+the fitted contours obtained by using the original surfaces of the model
+as shape priors, with a constant translation of [5.0,10.0,-5.0] mm.
 to illustrate briefly the extent of the capture range of the algorithm.
-Computation time in this case was around 4 minutes in the previously
-described platform. \\
+Computation time in this case was around 10 minutes in the previously
+described platform, with 16x16x16 control points on
+the dense deformation field and approximately 26,000 nodes in 
+total for both prior surfaces. \\
 %
@@
 \includegraphics[width=0.19\textwidth]{model1result_b_5} \\
-\includegraphics[width=0.19\textwidth]{model1result_a_1} &
-\includegraphics[width=0.19\textwidth]{model1result_a_2} &
-\includegraphics[width=0.19\textwidth]{model1result_a_3} &
-\includegraphics[width=0.19\textwidth]{model1result_a_4} &
-\includegraphics[width=0.19\textwidth]{model1result_a_5} \\
+\includegraphics[width=0.19\textwidth]{IPMI2013-011} &
+\includegraphics[width=0.19\textwidth]{IPMI2013-012} &
+\includegraphics[width=0.19\textwidth]{IPMI2013-013} &
+\includegraphics[width=0.19\textwidth]{IPMI2013-014} &
+\includegraphics[width=0.19\textwidth]{IPMI2013-015} \\
+%\includegraphics[width=0.19\textwidth]{model1result_a_1} &
+%\includegraphics[width=0.19\textwidth]{model1result_a_2} &
+%\includegraphics[width=0.19\textwidth]{model1result_a_3} &
+%\includegraphics[width=0.19\textwidth]{model1result_a_4} &
+%\includegraphics[width=0.19\textwidth]{model1result_a_5} \\
 \end{tabular}
 \caption{First row presents several slices along Z axis of the distorted \gls{fa} map and
-the undistorted \gls{wm}-\gls{gm} and \gls{wm}-\gls{csf} contours given as shape priors. Second row contains the same map, now with the contours after joint segmentation-registration.}
+the undistorted \gls{wm}-\gls{gm} and \gls{wm}-\gls{csf} contours given as shape priors,
+subjected to the shift of [5.0,10.0,-5.0] mm. described in \autoref{sec:simulated_dwi}.
+Second row contains the same map, now with the contours after joint segmentation-registration.}
 \label{fig:fa}

### 2013-05-17 Publications/2013-TMI/03-introduction.tex
Re-added changes
Updated files. Copied everything from IPMI
Also updated references to add one to the new PlosOne
article.

@@
 \label{sec:introduction}
+% The very first letter is a 2 line initial drop letter followed
+% by the rest of the first word in caps.
 % 
-\Gls{dwi} is a widely used family of \gls{mr} techniques
+% form to use if the first word consists of a single letter:
+% \IEEEPARstart{A}{demo} file is ....
+% 
+% form to use if you need the single drop letter followed by
+% normal text (unknown if ever used by IEEE):
+% \IEEEPARstart{A}{}demo file is ....
+% 
+% Some journals put the first two words in caps:
+% \IEEEPARstart{T}{his demo} file is ....
+% 
+% Here we have the typical use of a "T" for an initial drop letter
+% and "HIS" in caps to complete the first word.
+%
+\IEEEPARstart{D}{iffusion} MRI (dMRI) is a widely used family of \gls{mr} techniques
 \citep{sundgren_diffusion_2004} which recently has accounted for a growing
 interest in its application to whole-brain structural connectivity analysis.
-This emerging field, coined in 2005 as \emph{\gls{mr} Connectomics}
-\citep{hagmann_diffusion_2005,sporns_human_2005}, currently includes a
-large amount of imaging techniques for acquisition, processing, and analysis
-specifically tuned for \gls{dwi} data.
+This emerging field \citep{hagmann_diffusion_2005,sporns_human_2005}, currently 
+includes a large amount of imaging techniques for acquisition, processing, 
+and analysis specifically tuned for \gls{dwi} data \citep{daducci_connectome_2012}.
 
@@
 the much lower resolution of \gls{dwi} (typically around 
-$2.5\times2.5\times2.5mm^3$) compared to structural \gls{mri}, and the
+$2.0\times2.0\times2.0mm^3$) compared to structural \gls{mri}, and the
 existence of geometrical distortions.

### 2013-05-17 Publications/2013-TMI/08-appendix-01.tex
Updated files. Copied everything from IPMI
Also updated references to add one to the new PlosOne
article.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices

\section{Proof of duality between the probabilistic and the presented frameworks}
Here the demonstration that we are preparing.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
%\section{}
%Appendix two text goes here.

### 2013-05-20 Publications/2013-TMI/06-results.tex
Included multivariate formulations in #67
Added some new adapted formulations.

@@
 
-\begin{figure}
-\begin{tabular}{ccccc}
-\includegraphics[width=0.19\textwidth]{model1result_b_1} &
-\includegraphics[width=0.19\textwidth]{model1result_b_2} &
-\includegraphics[width=0.19\textwidth]{model1result_b_3} &
-\includegraphics[width=0.19\textwidth]{model1result_b_4} &
-\includegraphics[width=0.19\textwidth]{model1result_b_5} \\
-\includegraphics[width=0.19\textwidth]{IPMI2013-011} &
-\includegraphics[width=0.19\textwidth]{IPMI2013-012} &
-\includegraphics[width=0.19\textwidth]{IPMI2013-013} &
-\includegraphics[width=0.19\textwidth]{IPMI2013-014} &
-\includegraphics[width=0.19\textwidth]{IPMI2013-015} \\
-%\includegraphics[width=0.19\textwidth]{model1result_a_1} &
-%\includegraphics[width=0.19\textwidth]{model1result_a_2} &
-%\includegraphics[width=0.19\textwidth]{model1result_a_3} &
-%\includegraphics[width=0.19\textwidth]{model1result_a_4} &
-%\includegraphics[width=0.19\textwidth]{model1result_a_5} \\
-\end{tabular}
-\caption{First row presents several slices along Z axis of the distorted \gls{fa} map and
-the undistorted \gls{wm}-\gls{gm} and \gls{wm}-\gls{csf} contours given as shape priors,
-subjected to the shift of [5.0,10.0,-5.0] mm. described in \autoref{sec:simulated_dwi}.
-Second row contains the same map, now with the contours after joint segmentation-registration.}
-\label{fig:fa}
-\end{figure}
+%\begin{figure}
+%\begin{tabular}{ccccc}
+%\includegraphics[width=0.19\textwidth]{model1result_b_1} &
+%\includegraphics[width=0.19\textwidth]{model1result_b_2} &
+%\includegraphics[width=0.19\textwidth]{model1result_b_3} &
+%\includegraphics[width=0.19\textwidth]{model1result_b_4} &
+%\includegraphics[width=0.19\textwidth]{model1result_b_5} \\
+%\includegraphics[width=0.19\textwidth]{IPMI2013-011} &
+%\includegraphics[width=0.19\textwidth]{IPMI2013-012} &
+%\includegraphics[width=0.19\textwidth]{IPMI2013-013} &
+%\includegraphics[width=0.19\textwidth]{IPMI2013-014} &
+%\includegraphics[width=0.19\textwidth]{IPMI2013-015} \\
+%%\includegraphics[width=0.19\textwidth]{model1result_a_1} &
+%%\includegraphics[width=0.19\textwidth]{model1result_a_2} &
+%%\includegraphics[width=0.19\textwidth]{model1result_a_3} &
+%%\includegraphics[width=0.19\textwidth]{model1result_a_4} &
+%%\includegraphics[width=0.19\textwidth]{model1result_a_5} \\
+%\end{tabular}
+%\caption{First row presents several slices along Z axis of the distorted \gls{fa} map and
+%the undistorted \gls{wm}-\gls{gm} and \gls{wm}-\gls{csf} contours given as shape priors,
+%subjected to the shift of [5.0,10.0,-5.0] mm. described in \autoref{sec:simulated_dwi}.
+%Second row contains the same map, now with the contours after joint segmentation-registration.}
+%\label{fig:fa}
+%\end{figure}

### 2013-05-20 Publications/2013-TMI/08-appendix-01.tex
Started with #67.
Just a couple of (unfinished) formulae
Included multivariate formulations in #67
Added some new adapted formulations.

@@
 %Appendix two text goes here.
+
+In the probabilistic approach to image segmentation, the smoothness constraints
+can be introduced into the \gls{map} criterium as a \gls{mrf}. Accordingly to the
+Hammersley-Clifford theorem (\textbf{FIXME}: citations needed), an \gls{mrf} can
+be equivalently characterized by a Gibb's distribution:
+\begin{equation}
+P(\mathbf{K})=\underset{i}{\prod} Z(U)^{-1}\,e^{-U(\mathbf{x}_i \mid \beta,K)} = \mathit{const.}\,e^{- \sum\limits_i \left( U(\mathbf{x}_i \mid \beta,K) \right) } \sim e^{-\nu_B \left| K \right| }.
+\end{equation}
+This way, we include the assumption that the total length of the edge set $K$ is small
+and we draw the equivalence with the \gls{mrf} modeling, for $U(\mathbf{x}_i \mid \beta,K)$
+being the Potts model.
+
+In the probabilistic approach, the underlying concept to segmenting an image is the
+Bayes' rule. For a discrete image of $N$ pixels indexed by $i$, this rule reads as
+follows:
+\begin{equation}
+P(K|I)=P(I \mid K)\, P(K)=\underset{i}{\prod} p (\mathbf{x}_i | K ) \, p_i(K).
+\end{equation}
+The normalizer probability $P(I)$ has been ommited for simplicity.
+
+In order to express the likelihood as an absolute energy in our variation framework,
+the \textit{log}-likelihood is computed:
+\begin{equation}
+E(K)= -\log( P(K|I) ) = \sum\limits_k \int_{\Omega_k} -\log p_k(I(\mathbf{x}),\mathbf{x}) \,d\mathbf{x}+\nu_B \left|K\right|,
+\end{equation}
+where $\mathbf{s} = I(\mathbf{x})$, $\mathbf{s} \in \mathbb{R}^C$ (a vector of $C$ scalar features).
+Additionally, the discrete grid of $N$ pixels has been converted to a continuous space by assuming
+$d\mathbf{x}$ as infinitesimal bin size.
+
+To explicitly define the likelihood, we firstly define the squared \textit{Mahalanobis distance} that
+is the exponential of a multivariate normal distribution:
+\begin{equation}
+\Delta^2_k (\mathbf{s}) = (\mathbf{s} - \boldsymbol{\mu}_k)^T \, \Sigma^{-1}_k \, (\mathbf{s} - \boldsymbol{\mu}_k).
+\end{equation}
+
+Therefore, the likelihood is defined as follows:
+\begin{equation}
+p_k(I(\mathbf{x}),\mathbf{x}) = p_k(\mathbf{s},\mathbf{x})= p_k(\mathbf{s}) = \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{s}) \right)}}.
+\end{equation}
+
+Introducing this definition on (14), we have:
+\begin{equation}
+E(K)= \sum\limits_k \int_{\Omega_k} -\log{\left[ \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{s}) \right)} \right] } \,d\mathbf{x}+\nu_B \left|K\right|.
+\end{equation}
+
+\begin{equation}
+E(K) = \sum\limits_k \int_{\Omega_k} \left( \frac{1}{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right| \right)} + \frac{1}{2}  \Delta^2_k (\mathbf{s}) \right) \,d\mathbf{x}+\nu_B \left|K\right|
+\end{equation}
+
+\begin{equation}
+E(K) = \sum\limits_k \left( \frac{ V_k }{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right| \right)}+ \frac{1}{2} \int_{\Omega_k} \Delta^2_k (\mathbf{s}) \,d\mathbf{x}+\nu_B \left|K\right| \right)
+\end{equation}

### 2013-05-20 Publications/2013-TMI/03-introduction.tex
Drafted validation
I wrote renewed Data and methods section for the TMI version.

As long as this summarizes the issue, close #59. Some references
have been fixed regarding #55.

Added some people to the acks.

@@
 methods into three groups. The first group typically adds a shape prior term to 
-the energy functional of an evolving active contour~\citep{Bresson2006a,Chan2005,
-Chen2002,Cremers2006,Gastaud2004,Paragios2003,Vemuri2003a,Yezzi2003a}.
+the energy functional of an evolving active contour
+\citep{bresson_variational_2006,chan_level_2005,chen_using_2002,cremers_kernel_2006,
+gastaud_combining_2004,paragios_level_2003,vemuri_joint_2003,yezzi_variational_2003}.
 These methods generally have a explicit description of the expected relative boundary 
@@
 from this average shape. Closely related to this group are atlas-based segmentation
-methods \citep{Gorthi2011,Gorthi2009,Pohl2005,Pohl2006,Wang2006}, where the prior 
+methods \citep{gorthi_segmentation_2009,gorthi_active_2011,pohl_unifying_2005,
+pohl_bayesian_2006,wang_joint_2006}, where the prior 
 imposes consistent voxel-based classification of contiguous regions. Here, the 
@@
 the atlas to actual images, and the contour is to segment simultaneously two 
-different target images, related by a spatial transform to be co-estimated~\citep{Wyatt2003,
-Yezzi2003}.
+different target images, related by a spatial transform to be co-estimated
+\citep{wyatt_map_2003,yezzi_variational_2003}.
 

### 2013-05-20 Publications/2013-TMI/05-experiments.tex
Drafted validation
I wrote renewed Data and methods section for the TMI version.

As long as this summarizes the issue, close #59. Some references
have been fixed regarding #55.

Added some people to the acks.

@@
 %
-\subsection{Shape prior}
+
+\subsection{Image Data}
+\label{sec:datasets}
+
+\subsubsection{Simulated digital phantom} %
 %
-As described in \autoref{sec:introduction}, the general situation in
-the connectivity pipelines consists of having 
-a reliable segmentation obtained from the high resolution \gls{t1} 
-reference image. Therefore, a precise location of the tissue interfaces
-of interest is available in a reference space. Given that the anatomical 
-reference segmentation is beyond the scope of this manuscript, we simply 
-rely on the true contours known from the underlying models, and do not 
-seek to establish them in a separate segmentation step on ``anatomical'' images.
+The lack of a widely accepted \emph{gold}-standard in the application
+field has been addressed by several authors \citep{cote_tractometer:_2013}
+In this work, we use one of the most complete, and publicly available,
+digital phantoms (\url{http://hardi.epfl.ch/static/events/2013_ISBI/testing_data.html}).
+The phantom is a spherical volume containing a set of fiber bundles, that connect 
+one area of a ``cortex'' to another. The model accounts for \gls{pve} using 
+a similar approach to \citep{close_software_2009} and for \gls{csf} contamination
+as well.
+
+\subsubsection{Real \gls{mri} datasets} %
 %
-\subsection{Synthetic gray-scale data}
+We used image data from XX healthy volunteers with no history of neurological
+conditions (ages XX$\pm$XX, X female) to illustrate the applicability of our 
+approach. All the subjects were scanned in a 3T MR Scanner (Siemens Magnetom
+TrioTim blah blah)
+with an X-channel head coil. Subjects were scanned twice with the same protocol,
+described hereafter. After being scanned the first time, each subject exited the
+scan room for a short break and then reentered for an identical scan session.
+To note, there was a full repositioning of the volunteer, coils, blankets and pads
+before each scan and re-scan session.
+
+The scan session protocol was as follows:
+\begin{enumerate}
+\item Triplanar survey (Localizer).
+\item Field Mapping: field mapping using GRE sequence was performed before \gls{dwi}
+acquisition for susceptibility correction purposes.
+\item DTI: \gls{dwi} were acquired with axial in-plane 
+isotropic resolution $2mm$, slice thickness $2mm$, $XXX \times XXX \times XXX$ 
+image matrix, TR= XXXX ms, TE = XX ms, NEX = X, BW = XXXX Hz/pixel, GRAPPA
+acceleration factor 2. The series included images acquired with diffusion
+weighting along 30 non-collinear directions ( $b$=700sm$^{-2}$ ) sampled
+twice for averaging, and 5 interleaved images acquired without diffusion 
+weighting ( $b=0$ ).
+\item Field Mapping (same as before DTI)
+\item Structural T1: An MPRAGE T1-weighted acquisition, sagittal
+GRE sequence, in-plane isotropic resolution 1.0 mm, slice thickness 1.2mm
+$XXX \times XXX \times XXX$ image matrix, TR=2300 ms, TE=2.98 ms, FA= 9, NEX=X, BW= 240 Hz/pixel.
+\item Structural T2: A T2-weighted acquisition, oblique axial
+TSE sequence, in-plane isotropic resolution 1.0 mm, slice thickness 1.2mm
+$XXX \times XXX \times XXX$ image matrix, TR=3200 ms, TE= 408 ms, NEX=X, BW= 751 Hz/pixel.
+\end{enumerate}
+
+\subsection{Image preprocessing and shape-prior generation}
+\label{sec:preprocessing}
 %
-The first simplified model to test the approach is inspired by a problem 
-shown for coupled \gls{csf}/\gls{gm} and \gls{gm}/\gls{wm} segmentation in \citep{macdonald_automated_2000}. Authors note that ``partial volume effects 
-blur the distinction between closely adjacent surfaces in deep sulci, leading 
-to a well-known segmentation error in which the deeper reaches of sulci are 
-not penetrated by the putative surface model.'' This problem is aggravated 
-in DWI, since the resolution tends to be worse compared to the anatomical 
-images considered in \citep{macdonald_automated_2000}. They test their 
-coupled segmentation algorithm on an image, ``representing a sulcus in 
-which the distinction between opposing banks of the sulcus has been obscured 
-by partial volume.''
-
-Here, we reproduce their model on a volume consisting of three 
-piecewise-constant parts: a notched ball representing the \gls{wm} with a single 
-sulcus ($\mu_{WM} = 0.8$), a cortical sheet of \gls{gm} obtained through dilation 
-of the \gls{wm} ($\mu_{GM} = 0.5$), and the surrounding background representing 
-\gls{csf} ($\mu_{CSF} = 0.2$). The volume is then affected by additive Gaussian 
-noise, effectively creating uniform standard deviation of $\sigma = 0.045$ per 
-region.
-
-As illustrated in \autoref{fig:sulcusmodel}, conventional single surface 
-segmentation of the \gls{csf}/\gls{gm} boundary misses to capture the sulcus in 
-its full depth. With our proposed model, we expect the joint 
-segmentation-registration to be driven largely by the inner, \gls{gm}/\gls{wm}
-contour that exhibits sufficient contrast and lesser partial volume effects. 
-The shape prior of the outer, difficult contour will then be co-aligned through 
-the regularity of the estimated deformation field.
-
-\begin{figure}
-\begin{tabular}{ccccc}
-\includegraphics[width=0.19\textwidth]{contours_wrong} & \includegraphics[width=0.19\textwidth]{contours_inner} &
-\includegraphics[width=0.19\textwidth]{contours_both} &
-\includegraphics[width=0.19\textwidth]{pialsurf} &
-\includegraphics[width=0.19\textwidth]{gmwmsurf}\\
-a)&b)&c)&d)&e)
-\end{tabular}
-\caption{The gray-scale sulcus model.
-a) The apparent CSF/GM boundary is affected by partial volume in the sulcal cavity, and conventional segmentation is likely to miss it.
-b) The GM/WM interface here has consistently good contrast.
-c) Registering the two shape priors coupled through deformation field regularity is expected to guide the CSF/GM contour.
-d\&e) 3D view of the two shape priors.}
-\label{fig:sulcusmodel}
-\end{figure}
-
-%
-\subsection{Simulated diffusion data}
-%
-In order to demonstrate the functionality of the methodology, 
-and characterize its possibilities with diffusion data,
-we built a synthetic phantom from a model consisting of several 
-spherical shapes emulating the different brain tissues (see 
-\autoref{fig:fa}, first row). We simulated a single-shell
-acquisition at $b=1000$ \eqref{eqn:MultiTensor} with 30 samples
-equally distributed on one hemisphere, corresponding to a standard
-\gls{dti} acquisition commonly used in clinical practice.
-We generated a synthetic displacement field to produce
-an \gls{epi}-like distortion on the \gls{dwi} signal. Finally,
-we reconstructed the deformed signal, and generated \gls{dwi}-derived
-scalar maps (specifically the \gls{fa} and \gls{md} maps).
-Then, the scalar maps were placed as features in 
-\eqref{eq:complete_energy} to drive our segmentation approach, 
-finding the location of the \gls{wm}-\gls{gm}
-and the \gls{csf}-\gls{wm} interfaces in the distorted space. \\
-
-\paragraph{Signal simulation}
-To numerically simulate the \gls{mri} signal attenuation when applying a diffusion 
-gradient in a voxel with $N$ fiber populations we made use of the standard 
-\emph{Multi-Tensor Model}~\cite{Tuch:2002aa}:
-%
-\begin{equation} 
-\label{eqn:MultiTensor}
-S(q) / S_{0} = \sum_{i=1}^{N} f_{i} \, \exp{ \left( -b \, q^{T} \, \mathbf{D}_i \, q\right) } + f_{iso} \, \exp{ \left( -b \, \mathbf{D}_{iso} \right) }  + f_{gm} \, \exp{ \left( -b \, \mathbf{D}_{gm} \right) } ,
-\end{equation}
-%
-where $q \in \mathbb{S}^2$ is the direction of the diffusion gradient applied,
-$b$ is the b-value accounting for its strength and $S_0 \equiv S(0)$ is the 
-signal with no diffusion weighting. $f_i$ and $\mathbf{D}_i$ are the volume 
-fraction and the diffusion tensor characterizing the $i$-th fiber population, 
-whereas the tensors $\mathbf{D}_{iso}$ and $\mathbf{D}_{gm}$ describe the 
-diffusion processes of partial volumes with \gls{csf} and \gls{gm} within the 
-voxel, which volume fractions are, respectively, $f_{iso}$ and $f_{gm}$, and
-$\sum_i f_i + f_{iso} + f_{gm} = 1$. In this work, the diffusion properties have
-been taken from standard ranges typically observed in the human 
-brain~\cite{Canales-Rodriguez:2009aa}. Moreover we assumed $N = 1$ and $S_0 = 1$ 
-without loss of generality.
-
-\paragraph{Noise simulation}
-The diffusion MRI signal $S$ has been corrupted with 
-\textit{Rician noise}~\cite{Gudbjartsson:1995aa} as follows:
-%
-\begin{equation}
-	\tilde{S} = \sqrt{ (S + \varepsilon_1)^2 + \varepsilon_2^2 }
-\end{equation}
-%
-where $\varepsilon_{1,2}$ are Gaussian distributed with zero mean
-and standard deviation $\sigma = S_0 / \mathit{SNR}$ and \gls{snr}
-is the \gls{snr} on the $S_0$ image.
-
-\paragraph{Simulated \gls{epi} distortion}
-For this model, we created manually a sound distortion to approximate
-the real \gls{epi} distortions. We interpolated the distortion to a 
-dense deformation field, and applied it to generate the deformed \gls{dwi}
-signal.
-
-\paragraph{Derived scalar features}
-We obtained the local fiber configuration in each voxel with a commonly 
-used \gls{dti} reconstruction tool~\footnote{DTIFIT, included
-in the FMRIB's Software Library (FSL), 
-\url{http://fsl.fmrib.ox.ac.uk/fsl/fsl-4.1.9/fdt/fdt_dtifit.html}}. 
+Regardless the dataset type (simulated or real), all \gls{dwi} datasets were processed
+using the standard \gls{dti} reconstruction methods provided by FSL \footnote{DTIFIT, 
+included in the FMRIB's Software Library (FSL), 
+\url{http://fsl.fmrib.ox.ac.uk/fsl/fsl-4.1.9/fdt/fdt_dtifit.html}} to fit
+the tensor model and produce scalar maps of the required features.
 The properties of the reconstructed tensors and derived scalar maps have
@@
 usually extracted from the \gls{dwi}-derived scalar maps
-\cite{ennis_orthogonal_2006}. The different region properties 
-$(\mathbf{\mu}_R,\Sigma_R)$ measured for the \gls{fa} and \gls{md}
-joint distribution are summarized in \autoref{table:parameters}. \\
+\cite{ennis_orthogonal_2006}.
 
-\begin{table}
-\caption{Model means and covariances of \gls{fa} and \gls{md} estimated 
-from the reconstructed simulated \gls{dwi} images for each modeled tissue, 
-\gls{wm}, \gls{gm}, and \gls{csf}. As expected, the two scalar features 
-are complementary and the three tissues can well be discriminated. }
-\label{table:parameters}
-%\begin{tabularx}{1.0\textwidth}{c|XXX}
-\begin{tabular}{*{3}{@{}p{0.2\textwidth}}@{}p{0.4\textwidth}@{}}
-\toprule
-% tissue        & $\mathbf{\mu}_{FA}$ & $\mathbf{\mu}_{MD}$ & $\Sigma$ \newline
-        & $\mathbf{\mu}$ \\
-\cmidrule(r){2-3}
-Tissue& FA & MD &$\Sigma$\\
-\cmidrule(r){1-1}\cmidrule(r){2-2}\cmidrule(r){3-3}\cmidrule{4-4}
-\gls{wm}  & 0.778 & 6.94\e{-4} & 
-   $\begin{pmatrix}
-   	4.85\e{-3} & -6.90\e{-6} \\ -6.90\e{-6} & 1.03\e{-8}
-   \end{pmatrix}$\vspace{2mm}\\
+Preprocessing differed between simulated and real datasets in the 
+shape-prior surfaces generation, as we describe hereafter.
+
+\subsubsection{Simulated digital phantom} %
+A description goes here: surfaces extraction, synthetic susceptibility artifact generation.
+
+\subsubsection {Real \gls{mri} data} %
+
+We used a standard automated method 
+available in \emph{FreeSurfer} \citep{fischl_freesurfer_2012} to obtain the
+cortical gray/white boundary from the T1 scan \citep{greve_accurate_2009}.
+Parcellations used to evaluate the repeatability of the method are also
+obtained from \emph{FreeSurfer}.
+
+\subsection{Experiments and evaluation}
+\label{sec:experiments_evaluation}
 %
-\gls{gm}  & 0.119 & 8.95\e{-4} &
-   $\begin{pmatrix}
-   	5.90\e{-4} & -1.43\e{-6} \\ -1.43\e{-6} & 1.04\e{-8}
-   \end{pmatrix}$\vspace{2mm}
-\\
-%
-\gls{csf} & 0.103 & 2.99\e{-3} &
-   $\begin{pmatrix}
-   	1.19\e{-3} & 2.22\e{-7} \\ 2.22\e{-7} & 1.56\e{-8}
-   \end{pmatrix}$
-\\
-\bottomrule
-\end{tabular}
-%\end{tabularx}
+\subsubsection{Validation on the simulated digital phantom} %
+\label{sec:validation_phantom}
+We firstly evaluated our approach on the simulated data, using
+the distorted data with a deformation field similar to the
+susceptibility artifact that affects real \gls{dwi} data. To 
+this end, we report the following indices:
+\begin{itemize}
+\item \gls{se}, a distance between the one-to-one corresponding
+vertices, weighted by their respective Voronoi area.
+\begin{equation}
+MSE = \sum\limits_k \sum\limits_j^M w_j\,\| \mathbf{x}_j - \hat{\mathbf{x}}_j \|
+\end{equation}
+where $\mathbf{x}_j$ are the locations of the $M$ vertices of the $k$ priors,
+$\hat{\mathbf{x}}_j$ are the corresponding locations recovered, and $w_j$ the
+weighting factor as the relative surface of the Voronoi area.
+\item \gls{wi}, L2-distance between the theoretical and the recovered
+deformation field.
+\begin{equation}
+WI = \frac{1}{N} \sum\limits_i^N \| \mathbf{d}_i - \hat{\mathbf{d}}_i \|
+\end{equation}
+where $\mathbf{d}_i$ is the theoretical displacement vector at position $i$
+and $\hat{\mathbf{d}}_i$ is the recovered one at the same index position.
+\item Parcellation agreement, the \gls{se} averaged by defined \glspl{roi}
+between the theoretical and the recovered parcellations.
+\item \gls{nof} agreement, between the fibers recovered on the original
+data and the processed data.
+\end{itemize}
 
-\end{table}
+Additionally, the entropy of data is studied in both
+original and recovered datasets to draw another possible basis for the 
+assessment of the real datasets. We also report the same indices for the
+outcome of a widely-used methodology that combines field-map susceptibility 
+correction and T1-T2-\gls{dwi} registration (\textbf{FIXME}: I suggest to give 
+it a compact name and define it in the pre-processing section or even before
+in the intro).
+
+\subsubsection{Evaluation on real datasets}
+For the real datasets there is no published \emph{gold}-standard to validate results.
+Thus, visual results are provided to let compare the performance with the NICEACRONYM
+standard methodology. Additionally, cross-comparison of repeatability results are 
+provided. In this second evaluation strategy, all the indices defined in 
+\autoref{sec:validation_phantom} are reported.

### 2013-05-20 Publications/2013-TMI/06-results.tex
Drafted validation
I wrote renewed Data and methods section for the TMI version.

As long as this summarizes the issue, close #59. Some references
have been fixed regarding #55.

Added some people to the acks.

--- 2012-12-22 Publications/2013-IPMI/Latex/05-Results.tex
@@
 
-\subsection{Synthetic gray-scale data}
-The proposed method solved precisely the specific challenge
-created by the model. A severe distortion of the model is
-artificially created adding complexity to the problem of
-partial voluming in the outer contour of the sulcus.
-\autoref{fig:sulcusmodel_result} provides visual assessment
-for this result. With 16x16x16 control points, 
-computation time for this model was around
-10 minutes in a {\color{red}{(specific machine details)}}.
-
-\begin{figure}
-\begin{tabular}{ccccc}
-\includegraphics[width=0.19\textwidth]{model2result_b_1} &
-\includegraphics[width=0.19\textwidth]{model2result_b_2} &
-\includegraphics[width=0.19\textwidth]{model2result_a_1} &
-\includegraphics[width=0.19\textwidth]{model2result_a_2} &
-\includegraphics[width=0.19\textwidth]{model2surf} \\
-\multicolumn{2}{c}{a)} & \multicolumn{2}{c}{b)} & c)
-\end{tabular}
-\caption{Result of the segmentation performance on the sulcus model.
-a) Two slices and prior contours before segmentation-registration.
-b) The same slices and contours after distortion estimation.
-c) 3D rendering of the deformed priors, with the traces of the two selected slices highlighted.}
-\label{fig:sulcusmodel_result}
-\end{figure}
-
-\subsection{Simulated diffusion data}
-%
-The proposed method successfully reverted the synthetic distortion
-field we applied to the data. With 16x16x16 control points, the
-displacements field is dense enough to correctly represent the
-synthetic field. Second row in \autoref{fig:fa} shows the fitted
-contours obtained by using the original surfaces of the model
-as shape priors, with a constant translation of (5.0mm,10mm.,-5.0mm)
-to illustrate briefly the extent of the capture range of the algorithm.
-Computation time in this case was around 4 minutes in the previously
-described platform. \\
-%
-
-\begin{figure}
-\begin{tabular}{ccccc}
-\includegraphics[width=0.19\textwidth]{model1result_b_1} &
-\includegraphics[width=0.19\textwidth]{model1result_b_2} &
-\includegraphics[width=0.19\textwidth]{model1result_b_3} &
-\includegraphics[width=0.19\textwidth]{model1result_b_4} &
-\includegraphics[width=0.19\textwidth]{model1result_b_5} \\
-\includegraphics[width=0.19\textwidth]{model1result_a_1} &
-\includegraphics[width=0.19\textwidth]{model1result_a_2} &
-\includegraphics[width=0.19\textwidth]{model1result_a_3} &
-\includegraphics[width=0.19\textwidth]{model1result_a_4} &
-\includegraphics[width=0.19\textwidth]{model1result_a_5} \\
-\end{tabular}
-\caption{First row presents several slices along Z axis of the distorted \gls{fa} map and
-the undistorted \gls{wm}-\gls{gm} and \gls{wm}-\gls{csf} contours given as shape priors. Second row contains the same map, now with the contours after joint segmentation-registration.}
-\label{fig:fa}
-\end{figure}
+Please! Write me ASAP.

### 2013-05-21 Publications/2013-TMI/04-methods.tex
Re-written section II with @zosso comments
I am currently applying the comments received by email from Dominique.

--- 2012-12-22 Publications/2013-IPMI/Latex/04-Methods.tex
@@
 %
+\subsection{Maximum A-Posteriori Model}
+\label{sec:methods_map}
+%
+A widely used approach to image segmentation is derived from the
+Bayes' rule (\autoref{eq:bayes_rule}), where one seeks for a partitioning
+of a certain image $X$ in piecewise smooth regions $\Omega_k$, 
+with $k \in \lbrace 1 .. K \rbrace$, that maximizes the a posteriori 
+probability given the multivariate image $X \in \mathbb{R}^C$, 
+with $C$ being the number of image channels.
+\begin{equation}
+p(Y \mid X) = \frac{p(X \mid Y)\, p(Y)}{p(X)}.
+\label{eq:bayes_rule}
+\end{equation}
+
+Therefore, $Y$ is a certain realization of the piecewise 
+disjoint region set $\Omega_k$. $p(X \mid Y)$ is the \emph{likelihood} of 
+the realization of $X$ (the image) given a certain distribution model for 
+each region $k$. The second term, $p(Y)$, is the a-priori probability of 
+the partitioning $Y$. Finally, $p(X)$ is the probability of a certain image 
+realization, and thus, it will remain constant when computing the \gls{map}.
+Consequently, $p(Y \mid X) \propto p(X \mid Y)\, p(Y)$, and:
+\begin{equation}
+\textrm{argmax} \left\{ p(Y \mid X) \right\} = 
+\textrm{argmax} \left\{ p(X \mid Y)\, p(Y) \right\}.
+\end{equation}
+
+An extended assumption is that the feature vector realization $X$ is
+\emph{i.i.d.}, and thus, it is possible to write the a-posteriori
+probability $p(X \mid Y)$ as a continuous product with $d\mathbf{x}$ the
+infinitesimal voxel size:
+\begin{equation}
+p(X \mid Y) \, p(Y) = \underset{k}{\prod} \underset{\mathbf{x}\in \Omega_k}{\prod}
+p_k( X(\mathbf{x}) \mid Y(\mathbf{x}) )^{d\mathbf{x}},
+\label{eq:bayes_aposteriori}
+\end{equation}
+where the prior probability $p(Y)$ is implicitly
+defined by the regions definition.
+
+A second widely-accepted assumption is the multivariate normal 
+distribution of the different tissues in \gls{mri} data. Therefore,
+the posterior probability of an infinitesimal voxel
+can be written as:
+\begin{equation}
+p_k( X(\mathbf{x}) \mid Y(\mathbf{x}) ) = \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{X(\mathbf{x})}) \right)}}.
+\label{eq:bayes_mpdf}
+\end{equation}
+where we can identify the factor in the exponential, with 
+$\mathbf{s} = X(\mathbf{x})$ as the squared \emph{Mahalanobis 
+distance}:
+\begin{equation}
+\Delta^2_k (\mathbf{s}) = (\mathbf{s} - \boldsymbol{\mu}_k)^T \, \Sigma^{-1}_k \, (\mathbf{s} - \boldsymbol{\mu}_k).
+\label{eq:bayes_mahalanobis}
+\end{equation}
+
+Finally, we can turn the \gls{map} problem into a variational problem
+applying the following log-transform:
+\begin{align}
+E(X \mid Y)= -\log( p(X \mid Y) \, p(Y) ) = \\
+= -\log(\underset{k}{\prod} \underset{\mathbf{x}\in \Omega_k}{\prod}
+p_k( X(\mathbf{x}) \mid Y(\mathbf{x}) )^{d\mathbf{x}}) = \\
+= \sum\limits_k \int_{\Omega_k} p_k(X(\mathbf{x}) \mid Y(\mathbf{x} ) ) \, d\mathbf{x},
+\label{eq:energy_1}
+\end{align}
+and introducing the posterior probability term (\autoref{eq:bayes_mpdf}), we get:
+\begin{align}
+E(X \mid Y) = \sum\limits_k \int_{\Omega_k} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{X(\mathbf{x})}) \right)}} \, d\mathbf{x} = \\
+= \sum\limits_k \int_{\Omega_k} \left[ \frac{1}{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right| \right)} + \frac{1}{2}  \Delta^2_k (\mathbf{s}) \right] \,d\mathbf{x}.
+\end{align}
+Finally, after removing scaling factors and independent constants,
+we obtain
+\begin{align}
+E(X \mid Y) = \sum\limits_k V_k \, \log{\left|\boldsymbol{\Sigma}_{k}\right|} + \sum\limits_k \int_{\Omega_k} \Delta^2_k (\mathbf{s}) \,d\mathbf{x},
+\end{align}
+that is a constant varying with the corresponding total volume and the 
+determinant of the covariance matrix of the region $k$, plus an energy 
+term based on the squared \emph{Mahalanobis distance}.
+
 \subsection{\Acrlong{acwe} based segmentation model}

### 2013-05-21 Publications/2013-TMI/04-methods.tex
Fixed some minor mathematical details
Section II.A

@@
 Bayes' rule (\autoref{eq:bayes_rule}), where one seeks for a partitioning
-of a certain image $X$ in piecewise smooth regions $\Omega_k$, 
-with $k \in \lbrace 1 .. K \rbrace$, that maximizes the a posteriori 
+of a certain image $X$ in piecewise smooth regions $\Omega = \lbrace \Omega_k , 
+k\in\left[ 1 .. K \right] \rbrace$,  that maximizes the a posteriori 
 probability given the multivariate image $X \in \mathbb{R}^C$, 
@@
 Therefore, $Y$ is a certain realization of the piecewise 
-disjoint region set $\Omega_k$. $p(X \mid Y)$ is the \emph{likelihood} of 
+disjoint region set $\Omega$. $p(X \mid Y)$ is the \emph{likelihood} of 
 the realization of $X$ (the image) given a certain distribution model for 
-each region $k$. The second term, $p(Y)$, is the a-priori probability of 
+each region $\Omega_k$. The second term, $p(Y)$, is the a-priori probability of 
 the partitioning $Y$. Finally, $p(X)$ is the probability of a certain image 

### 2013-05-22 Publications/2013-TMI/04-methods.tex

@@
 \begin{equation}
-\textrm{argmax} \left\{ p(Y \mid X) \right\} = 
-\textrm{argmax} \left\{ p(X \mid Y)\, p(Y) \right\}.
+\underset{Y}{\argmax} \left\{ p(Y \mid X) \right\} = 
+\underset{Y}{\argmax} \left\{ p(X \mid Y)\, p(Y) \right\}.
 \end{equation}
@@
 where we can identify the factor in the exponential, with 
-$\mathbf{s} = X(\mathbf{x})$ as the squared \emph{Mahalanobis 
+$\mathbf{f} = X(\mathbf{x})$ as the squared \emph{Mahalanobis 
 distance}:
 \begin{equation}
-\Delta^2_k (\mathbf{s}) = (\mathbf{s} - \boldsymbol{\mu}_k)^T \, \Sigma^{-1}_k \, (\mathbf{s} - \boldsymbol{\mu}_k).
+\Delta^2_k (\mathbf{f}) = (\mathbf{f} - \boldsymbol{\mu}_k)^T \, \Sigma^{-1}_k \, (\mathbf{f} - \boldsymbol{\mu}_k).
 \label{eq:bayes_mahalanobis}
@@
 
-Finally, we can turn the \gls{map} problem into a variational problem
+Finally, we can turn the \gls{map} problem into a variational one
 applying the following log-transform:
 \begin{align}
-E(X \mid Y)= -\log( p(X \mid Y) \, p(Y) ) = \\
-= -\log(\underset{k}{\prod} \underset{\mathbf{x}\in \Omega_k}{\prod}
-p_k( X(\mathbf{x}) \mid Y(\mathbf{x}) )^{d\mathbf{x}}) = \\
-= \sum\limits_k \int_{\Omega_k} p_k(X(\mathbf{x}) \mid Y(\mathbf{x} ) ) \, d\mathbf{x},
+E(X \mid Y)= -\log \left[ p(X \mid Y) \, p(Y) \right] = \\
+= -\log \left[ \underset{k}{\prod} \underset{\mathbf{x}\in \Omega_k}{\prod}
+p_k( X(\mathbf{x}) \mid Y(\mathbf{x}) )^{d\mathbf{x}} \right] = \\
+= \sum\limits_k \int_{\Omega_k} -\log \left[ p_k(X(\mathbf{x}) \mid Y(\mathbf{x} ) ) \right] \, d\mathbf{x},
 \label{eq:energy_1}
@@
 \begin{align}
-E(X \mid Y) = \sum\limits_k \int_{\Omega_k} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{X(\mathbf{x})}) \right)}} \, d\mathbf{x} = \\
-= \sum\limits_k \int_{\Omega_k} \left[ \frac{1}{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right| \right)} + \frac{1}{2}  \Delta^2_k (\mathbf{s}) \right] \,d\mathbf{x}.
+E(X \mid Y) = \sum\limits_k \int_{\Omega_k} -\log \left[ \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{f}) \right)}} \right] \, d\mathbf{x} = \\
+= \sum\limits_k \int_{\Omega_k} \left[ \frac{1}{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right| \right)} + \frac{1}{2}  \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}.
 \end{align}
@@
 \begin{align}
-E(X \mid Y) = \sum\limits_k V_k \, \log{\left|\boldsymbol{\Sigma}_{k}\right|} + \sum\limits_k \int_{\Omega_k} \Delta^2_k (\mathbf{s}) \,d\mathbf{x},
+E(X \mid Y) = \sum\limits_k \left[ \log{\left|\boldsymbol{\Sigma}_{k}\right|} \int_{\Omega_k} d\mathbf{x} + \int_{\Omega_k} \Delta^2_k (\mathbf{f}) \,d\mathbf{x} \right],
+\label{eq:map_energy}
 \end{align}
-that is a constant varying with the corresponding total volume and the 
-determinant of the covariance matrix of the region $k$, plus an energy 
-term based on the squared \emph{Mahalanobis distance}.
+(\textbf{FIXME:} bad explanation) where we have a constant term scaled by
+the total volume of the partition $\Omega_k$ and the determinant of the 
+covariance matrix of the partition $\left|\boldsymbol{\Sigma}_{k}\right|$,
+plus an energy term based on the squared \emph{Mahalanobis distance}.
+
+\subsection{Deformation model}
+\label{sec:deformation}
+%
+The segmentation problem is transformed into a registration one if the initial
+partition $Y$ is derived from the initial shape-priors in reference space. 
+Introducing a dense deformation field $u$ that maps the original partition to
+a better fit of the regions in the target coordinate space. Thus, the minimization
+problem stated in \autoref{eq:map_energy} can be expressed so that $u$ is the 
+unknown:
+\begin{equation}
+E(u(\mathbf{x})) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}
+\label{eq:map_energy_deformation}
+\end{equation}
 
@@
 \cite{chan_active_2001}:
-\begin{equation}
-\label{eq:gaussian_energy}
-E(u)= \sum_{\forall{R}} \int_{\Omega_R} (f-\mu_R)^T\Sigma_R^{-1}(f-\mu_R) dx
-\end{equation}
+
 where $R$ indexes the existing regions and the integral domains

### 2013-05-22 Code/Modules/LevelSets/include/LevelSetsBase.h
Starting with #70
I am implementing the code needed for parameters estimation.

I also open #77, given that we need a tissue probablity map of
each class to plug in the estimators.

@@ -149,6 +149,10 @@ public:
 			          <ContourDeformationType, ROIType>	     BinarizeMeshFilterType;
 	typedef typename BinarizeMeshFilterType::Pointer         BinarizeMeshFilterPointer;
 
+	typedef itk::Image< float, Dimension >                   ProbabilityMapType;
+	typedef typename ProbabilityMapType::Pointer             ProbabilityMapPointer;
+	typedef typename ProbabilityMapType::ConstPointer        ProbabilityMapConstPointer;
+
 	typedef typename itk::MeshSpatialObject
 			                   <ContourDeformationType>      ContourSpatialObject;
 	typedef typename ContourSpatialObject::Pointer           ContourSpatialPointer;
@@ -197,6 +201,8 @@ protected:
 	virtual void InitializeSamplingGrid( void ) = 0;
 
 	inline virtual MeasureType GetEnergyAtPoint( PixelPointType& point, size_t cont ) = 0;
+
+	ProbabilityMapConstPointer GetCurrentRegion( size_t idx );
 
 	mutable MeasureType m_Value;
 	DeformationFieldPointer m_DeformationField;

### 2013-05-24 Code/Modules/LevelSets/include/LevelSetsBase.h
Implemented parameters estimation routine
Now it is possible to re-estimate parameters whenever we want, so
close #70.

On the other hand, I realized that our approach is 2-class-only
segmentation. I opened issue #76 raising the concerns regarding
surfaces overlap and this stuff... but the problem is bigger than
this.

In the update formulas, the gradients are computed ONLY with the
*inside*/*outside* statistics. The problem comes when a region
is neighbouring with more than 1 other region. In this case,
our current implementation doesn't take it into account when computing
the *outside* energ

@@ -211,8 +211,6 @@ protected:
 		os << std::endl;
 	}
 
-
-	void InitializeROIs( void );
 	virtual void InitializeSamplingGrid( void ) = 0;
 
 	inline virtual MeasureType GetEnergyAtPoint( PixelPointType& point, size_t cont ) = 0;
@@ -230,6 +228,8 @@ protected:
 	ROIList m_CurrentROIs;
 	DisplacementTransformPointer m_Transform;
 	ReferenceImageConstPointer m_ReferenceImage;
+	bool m_Modified;
+
 private:
 	LevelSetsBase(const Self &);  //purposely not implemented
 	void operator=(const Self &); //purposely not implemented

### 2013-05-24 Code/Modules/LevelSets/include/MahalanobisLevelSets.h

@@ -158,7 +158,7 @@ protected:
 		return this->GetEnergyAtPoint( point, cont, 0 );
 	}
 
-	ParametersType& UpdateParametersOfRegion( const size_t idx );
+	ParametersType UpdateParametersOfRegion( const size_t idx );
 	void ComputeParameters( void );
 
 	ParametersList m_Parameters;

### 2013-05-24 Publications/2013-TMI/04-methods.tex
Added some primes in (12)
I have to include the Jacobian here.

@@
 %
-\subsection{Maximum A-Posteriori Model}
+\subsection{Mumford-Shah Functional derivation from the Maximum A-Posteriori Model}
 \label{sec:methods_map}
@@
 A widely used approach to image segmentation is derived from the
-Bayes' rule (\autoref{eq:bayes_rule}), where one seeks for a partitioning
+Bayes' rule \eqref{eq:bayes_rule}, where one seeks for a partitioning
 of a certain image $X$ in piecewise smooth regions $\Omega = \lbrace \Omega_k , 
@@
 $\mathbf{f} = X(\mathbf{x})$ as the squared \emph{Mahalanobis 
-distance}:
+distance} with the parameters set $\Theta_k = \lbrace \boldsymbol{\mu}_k, \boldsymbol{\Sigma}_k \rbrace$:
 \begin{equation}
-\Delta^2_k (\mathbf{f}) = (\mathbf{f} - \boldsymbol{\mu}_k)^T \, \Sigma^{-1}_k \, (\mathbf{f} - \boldsymbol{\mu}_k).
+\Delta^2_k (\mathbf{f} \mid \Theta_k ) = (\mathbf{f} - \boldsymbol{\mu}_k)^T \, \boldsymbol{\Sigma}^{-1}_k \, (\mathbf{f} - \boldsymbol{\mu}_k).
 \label{eq:bayes_mahalanobis}
@@
 applying the following log-transform:
-\begin{align}
+\begin{multline}
 E(X \mid Y)= -\log \left[ p(X \mid Y) \, p(Y) \right] = \\
@@
 \label{eq:energy_1}
-\end{align}
-and introducing the posterior probability term (\autoref{eq:bayes_mpdf}), we get:
-\begin{align}
+\end{multline}
+and introducing the posterior probability term \eqref{eq:bayes_mpdf}, we get:
+\begin{multline}
 E(X \mid Y) = \sum\limits_k \int_{\Omega_k} -\log \left[ \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{f}) \right)}} \right] \, d\mathbf{x} = \\
 = \sum\limits_k \int_{\Omega_k} \left[ \frac{1}{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right| \right)} + \frac{1}{2}  \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}.
-\end{align}
+\end{multline}
 Finally, after removing scaling factors and independent constants,
-we obtain
+we can express the functional in terms of $\lbrace\Theta_k,\Omega_k\rbrace$:
 \begin{align}
-E(X \mid Y) = \sum\limits_k \left[ \log{\left|\boldsymbol{\Sigma}_{k}\right|} \int_{\Omega_k} d\mathbf{x} + \int_{\Omega_k} \Delta^2_k (\mathbf{f}) \,d\mathbf{x} \right],
+E(\Theta_k,K) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x},
 \label{eq:map_energy}
@@
 plus an energy term based on the squared \emph{Mahalanobis distance}.
+
+Equation \eqref{eq:map_energy} resembles the Mumford-Shah functional including variance,
+that modifies the original functional in a way that it can deal with more general
+distributions. This is necessary to avoid the assumption that regions $\Omega_k$
+have a fixed covariance matrix on their complete domain. One immediate advantage
+of this functional from the original one is the possibility to distinguish regions
+with the same mean vector but different covariance matrix \citep{brox_local_2009}:
+\begin{multline}
+E(\Theta_k,\Omega_k) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x} \\
++ \lambda \int_{\Omega_k - K}  ( \left| \nabla \mathbf{\mu} \right| ^2 + \left| \nabla \mathbf{\Sigma}_k \right| ^2 ) \, d\mathbf{x} 
++ \nu |K|,
+\end{multline}
+that is easily identifiable with \eqref{eq:map_energy} when we apply 
+the so-called \emph{cartoon limit}, 
+for $\lambda \to \infty$:
+\begin{equation}
+E(\Theta_k,K) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}
++ \nu |K|.
+\end{equation}
 
@@
 a better fit of the regions in the target coordinate space. Thus, the minimization
-problem stated in \autoref{eq:map_energy} can be expressed so that $u$ is the 
+problem stated in \eqref{eq:map_energy} can be expressed so that $u$ is the 
 unknown:
 \begin{equation}
-E(u(\mathbf{x})) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}
+E(u(\mathbf{x})) = \sum\limits_k \int_{\Omega'_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f'}) \right] \,d\mathbf{x}
 \label{eq:map_energy_deformation}
 \end{equation}
+where $\Omega'_k = u(\Omega_k)$ and $\mathbf{f'} = X(u(\mathbf{x}))$.
 

### 2013-05-24 Publications/2013-TMI/04-methods.tex
Several minor fixes
  * Updated title (#82)
  * Very little changes on methods (#80)

@@
 \end{multline}
-and introducing the posterior probability term \eqref{eq:bayes_mpdf}, we get:
+and introducing the posterior probability term \eqref{eq:bayes_mpdf}, 
+we can express the functional in terms of $\lbrace\Theta_k,\Omega_k\rbrace$:
 \begin{multline}
-E(X \mid Y) = \sum\limits_k \int_{\Omega_k} -\log \left[ \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{f}) \right)}} \right] \, d\mathbf{x} = \\
+E(\Theta_k,\Omega_k) = \\
+= \sum\limits_k \int_{\Omega_k} -\log \left[ \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{f}) \right)}} \right] \, d\mathbf{x} = \\
 = \sum\limits_k \int_{\Omega_k} \left[ \frac{1}{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right| \right)} + \frac{1}{2}  \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}.
@@
 Finally, after removing scaling factors and independent constants,
-we can express the functional in terms of $\lbrace\Theta_k,\Omega_k\rbrace$:
+we obtain:
 \begin{align}
-E(\Theta_k,K) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x},
+E(\Theta_k,\Omega_k) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x},
 \label{eq:map_energy}
@@
 + \nu |K|.
+\end{equation}
+
+As long as we do not penalize the edge set $K$ length, $\nu = 0$ and
+the result is dual to \eqref{eq:map_energy}:
+\begin{equation}
+E(\Theta_k,K) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}.
 \end{equation}

### 2013-05-24 Publications/2013-TMI/05-experiments.tex
Several minor fixes
  * Updated title (#82)
  * Very little changes on methods (#80)

@@
 approach. All the subjects were scanned in a 3T MR Scanner (Siemens Magnetom
-TrioTim blah blah)
-with an X-channel head coil. Subjects were scanned twice with the same protocol,
+TrioTim, Siemens, Erlangen, Germany) with a standard 12-channel head coil. 
+(\textbf{FIXME:} check coil number of channels!!)
+Subjects were scanned twice with the same protocol,
 described hereafter. After being scanned the first time, each subject exited the
@@
 To note, there was a full repositioning of the volunteer, coils, blankets and pads
-before each scan and re-scan session.
-
-The scan session protocol was as follows:
+before each scan and re-scan session. The scan session protocol was as follows:
 \begin{enumerate}

### 2013-05-28 Publications/2013-TMI/03-introduction.tex
Advancing Introduction (#81)
I re-structured introduction, pushed forward mathematical background
to the II section (methods). Currently I still have to rewrite
the registration stuff, comments of existing literature about
distortion correction evaluation and the final subsection of I. Also some
todo tags have been placed where needed.

--- 2013-05-17 Publications/2013-TMI/03-introduction.tex
@@
 %
-\IEEEPARstart{D}{iffusion} MRI (dMRI) is a widely used family of \gls{mr} techniques
-\citep{sundgren_diffusion_2004} which recently has accounted for a growing
-interest in its application to whole-brain structural connectivity analysis.
-This emerging field \citep{hagmann_diffusion_2005,sporns_human_2005}, currently 
-includes a large amount of imaging techniques for acquisition, processing, 
-and analysis specifically tuned for \gls{dwi} data \citep{daducci_connectome_2012}.
+\IEEEPARstart{D}{iffusion} \gls{mri} is a widely used family
+of \gls{mri} techniques \citep{sundgren_diffusion_2004} which recently 
+has accounted for a growing interest in its application to structural 
+connectivity analysis of the brain. This emerging field exploits
+\gls{dwi} data to derive the local axonal structure at each imaging voxel 
+\citep{basser_microstructural_2011} and estimate a whole-brain mapping of fiber 
+tracts represented by trajectories reconstructed from the local information.
+This comprehensive map of neural connections of the brain is called the 
+\emph{connectome} \citep{hagmann_diffusion_2005,sporns_human_2005}. The connectome
+analysis is a promising tool for neuroscience and clinical applications 
+(\textbf{FIXME:} why? or references?).
 
-The whole-brain connectivity analysis has given rise to some challenges
-towards reliable structural information about the neuronal tracts 
-from \gls{dwi} \cite{johansen-berg_using_2009,jones_white_2012}. 
-Here, we shall address brain tissue segmentation on diffusion space 
-and correction of geometrical distortions inherent to the acquisition 
-sequence \citep{hagmann_mr_2012}.
+Early \gls{dwi} research focused mainly on the improvements of imaging 
+methodologies better understanding the diffusion effect and improving
+image reconstruction methodologies. Currently, the connectome extraction 
+and analysis relies on a large amount of sophisticated computational techniques
+\citep{daducci_connectome_2012,hagmann_mr_2012} including acquisition,
+reconstruction, modeling and model fitting, image processing, fibre tracking,
+connectivity mapping, visualization, group studies, and inference. This 
+growing complexity has given rise to challenging issues towards reliable 
+structural information about the neuronal tracts \cite{johansen-berg_using_2009,
+jones_white_2012,soares_hitchhikers_2013}, and statistical analysis 
+\citep{meskaldji_comparing_????}. Here,we shall address three tasks included within 
+the image processing stage in a unified approach: brain tissue segmentation in 
+diffusion space (\autoref{sec:dwi_segmentation}), correction of geometrical 
+distortions (\autoref{sec:distortion}), and structural image registration to 
+diffusion coordinate space (\autoref{sec:registration}). These tasks are generally
+solved independently, or combined in pairs. However, there exist fundamental 
+coupling relationships that can be exploited to obtain a simultaneous solution to 
+the three problems. This joint approach satisfactory impacts the downstream
+outcomes of the whole pipeline with the increase of the internal consistency.
 
-In this work, we will refer as brain tissue segmentation to the precise
-delineation of the \gls{csf}-\gls{gm} and \gls{gm}-\gls{wm} interface surfaces.
-An accurate brain tissue segmentation is required
-to filter the fiber bundles obtained with \gls{dwi} tractography. 
-This requirement is usually solved in practice by plainly thresholding the 
-\gls{fa}, a well-known scalar map derived from \gls{dwi} which depicts 
-the isotropy of water diffusion inside the brain. Also,
-it is necessary to locate the intersections of fiber bundles and \gls{gm}.
-Moreover, a precise location of the \gls{gm}-\gls{wm} surface is also essential 
-to finally achieve a consistent parcellisation of the cortex 
-to represent the nodes of the output network. This parcellisation 
-is generally defined in a high-resolution and better understood structural 
-\gls{mri} of the same 
-subject (e.g. \gls{t1} and/or \gls{t2} weighted acquisitions). Even though
-some efforts have addressed the study of the robustness of tractography with
-respect to intra-subject variability \cite{heiervang_between_2006,
-wakana_reproducibility_2007}, these results are restricted to certain regions 
-of the brain, only. Therefore, robust and precise segmentation methods are 
-required in the whole-brain application. The problem is challenging due to
-the much lower resolution of \gls{dwi} (typically around 
-$2.0\times2.0\times2.0mm^3$) compared to structural \gls{mri}, and the
-existence of geometrical distortions.
+\subsection{\gls{dwi} data overview}
+\label{sec:dwi_overview}
+
+\todo[inline]{VERY brief look into signal generation HERE and cite reconstruction methods.}
+
+\gls{dwi} data is strongly affected by \glspl{pve} \citep{alexander_analysis_2001},
+which appear when several different tissues, or signal emitters, are present
+in the same imaging unit, producing an averaged intensity. The effect is directly
+related to the low resolution achievable with \gls{dwi} (typically around 
+$2.0\times2.0\times2.0mm^3$). An additional complication specific to
+\gls{dwi} is the \gls{csf} \emph{contamination} \citep{metzler-baddeley_how_2012},
+that is a particular \gls{pve} in which the signal sensed inside the affected voxel is 
+linear with respect the \gls{gm} and \gls{wm} contributions, but non-linear with
+respect \gls{csf}.
+
+Generally speaking, alongside the difficulty posed by the low resolution, \gls{dwi} 
+processing is also challenging due to the \emph{direction dependency} of raw data.
+\todo[inline]{Introduce here what are b0, fa, md, and direction dependency problem.}
+These low-$b$ (or $b = 0$) volumes are acquired without direction gradient as reference, 
+and present a T2-like contrast.
+
+\subsection{\gls{dwi} segmentation}
+\label{sec:dwi_segmentation}
+
+A precise delineation of the \gls{csf}, \gls{gm} and \gls{wm} interfacing surfaces
+is required with sub-pixel resolution.
+The resulting segmentation is necessary to perform the majority of tractography 
+algorithms and it is required to filter the resulting tractogram. The \gls{gm}-\gls{wm}
+interface is necessary to locate the starting and ending points of the detected
+fiber bundles, and \gls{csf}-\gls{wm} surface is critical for pruning spurious 
+and discontinued fiber bundles.
+
+A number of methodologies have been proposed for \gls{dwi} segmentation, ranging 
+from intensity thresholding to atlas-based segmentation. The first approach is performed 
+on the \gls{fa} \citep{ennis_orthogonal_2006}, a well-known scalar map derived from
+\gls{dwi} data which depicts the isotropy of water diffusion inside the brain.
+Although this methodology was popular among the premier tractography studies,
+they were generally limited to certain regions or significant fiber tracts, and thus,
+it cannot be applied in whole-brain tractography. Early approaches to \gls{dwi} segmentation 
+include level set formulations using scalar maps of direction invariants derived
+from the tensor model \citep{zhukov_level_2003}, directly on the diffusion raw data
+\citep{rousson_level_2004}, or finding alternative diffusion representations 
+\citep{jonasson_representing_2007}. Even though this latter case was restricted to the extraction of
+the corpus callosum from a real dataset, the density of the components of the diffusion tensor
+are approximated by multivariate Gaussians for first. Iterative clustering performed on the 
+\emph{low}-$b$ volumes of \gls{dwi} data was proposed by \citep{hadjiprocopis_unbiased_2005}.
+Later studies investigated the application of probabilistic frameworks combining mixtures of 
+gaussians, \gls{mrf} and labeling fusion techniques \citep{liu_brain_2007} using as features 
+widely-used \gls{dwi}-derived scalar maps as \gls{fa} or \gls{md}. A similar framework 
+combining co-registered structural information (T1 weighted) with \emph{independent orthogonal 
+invariants} derived from the \gls{dwi} tensor model was proposed by \citep{awate_multivariate_2008}. 
+Some proposals suggest the use of the raw diffusion data (directionally dependent) to avoid fitting 
+a certain model \citep{lu_segmentation_2008}. In \citep{han_experimental_2009}, graph-cuts 
+voxel-based techniques are proposed using the most common diffusion tensor derived features.
+Further developments of the probabilistic approach have been proposed adding more scalar maps
+as features and a more detailed treatment of \gls{pve} \citep{kumazawa_partial_2010}.
+
+A number of methods have been proposed using features not directly derived from \gls{dwi} data.
+Segmentation obtained by co-registering structural T1-weighted images will be covered in 
+\autoref{sec:registration}. Some other works delay the segmentation task after the tractogram 
+is obtained, performing clustering on features derived on the tracts alignment 
+\citep{jonasson_white_2005}, combined tract registration \citep{mayer_supervised_2011} or using
+tractography atlases \citep{odonnell_automatic_2007}. However, methods based on tractography
+usually address the tractogram segmentation problem, to later combine the solution to 
+answer the whole-brain segmentation problem.
+
+None of the presented methods have claimed for definite results, mainly due to the lack of 
+a \emph{gold-standard} evaluation methodology. Most of them are tested only on certain
+regions \todo{citations}, or do not provide sub-pixel resolution results
+\citep{hadjiprocopis_unbiased_2005,liu_brain_2007,awate_multivariate_2008,lu_segmentation_2008,
+han_experimental_2009}. Generally, results obtained with high resolution atlas co-registration
+(\autoref{sec:registration}) are more compelling, minimizing the activity on this line which is 
+currently being considered to be included in reconstruction algorithms 
+\citep{kumazawa_improvement_2013}. \emph{Golden}-standard evaluation frameworks have been 
+proposed for the segmentation validation \citep{jha_task-based_2012} on the task of lesion 
+detection in visceral organs.
+
+\subsection{Correction for susceptibility distortions}
+\label{sec:distortion}
 
@@
 as they allow for very fast acquisitions, but they are known to 
-suffer from geometrical distortions due to local field inhomogeneities. 
-These artifacts happen along the phase-encoding direction, and are most 
-appreciable in the front part of the brain for the strong air/tissue 
-interface around the frontal sinuses. A number of methodologies have
-been developed to correct for the distortion, and are generically named 
-as \emph{\gls{epi}-unwarp} techniques
-\cite{holland_efficient_2010,hsu_correction_2009,jezzard_characterization_2005,
-reber_correction_2005}. These methods usually 
-require the extra acquisition of the magnitude and phase of
-the field (``field-mapping''), a condition which is not always met. Some other 
-methodologies do not make use of the field-mapping, compensating the distortion
-with non-linear registration from structural \gls{mri} or other means
-\citep{andersson_modeling_2001}. To our knowledge, there exists no study
-on the impact of the \gls{epi} distortion on the variability of tractography
+suffer from geometrical distortions and artifacts due to, mainly,
+three sources: the subject motion between direction sampling, the 
+induced \emph{Eddy currents} on the scanner coils and finally
+distortions caused by the magnetic susceptibility inhomogeneity
+present at tissue interface. In this paper, we restrict ourselves to
+the last one, as it accounts for the major impact in the connectome
+analysis. \emph{Susceptibility distortions} happen along the 
+phase-encoding direction, and are most appreciable in the front part of 
+the brain for the strong air/tissue interface surrounding the frontal sinuses.
+
+One approach to correcting the susceptibility distortions was proposed
+with the earliest \gls{epi} sequences used in functional \gls{mri}, and
+relies on the acquisition of extra \gls{mri} data. Generally, a \gls{gre} is used to 
+obtain a map of magnitude and phase of the actual magnetic field inside the
+scanner. Based on this \emph{fieldmap}, a number of methodologies have 
+been developed to correct for the distortion, and are generically named as 
+\emph{\gls{epi}-unwarp} techniques \cite{holland_efficient_2010,
+hsu_correction_2009,jezzard_characterization_2005, reber_correction_2005}. 
+Unfortunately, the availability of the corresponding fieldmap is not always
+met.\todo[inline]{Mention other problems: low precission, unreliability}
+
+Some other methodologies do not make use of the fieldmaps, compensating the 
+distortion with non-linear registration from structural \gls{mri} 
+\citep{kybic_unwarping_2000} (see \autoref{sec:registration}), or other means 
+\citep{andersson_modeling_2001}. To our knowledge, there exists no study on 
+the impact of the \gls{epi} distortion on the variability of tractography 
 results.
+
+\subsection{Structural information co-registration}
+\label{sec:registration}
 
@@
 
-We suggest clustering the current methodologies of template-based segmentation 
-methods into three groups. The first group typically adds a shape prior term to 
-the energy functional of an evolving active contour~\citep{Bresson2006a,Chan2005,
-Chen2002,Cremers2006,Gastaud2004,Paragios2003,Vemuri2003a,Yezzi2003a}.
-These methods generally have a explicit description of the expected relative boundary 
-locations of the object to be delineated, and some even model the statistical deviations
-from this average shape. Closely related to this group are atlas-based segmentation
-methods \citep{Gorthi2011,Gorthi2009,Pohl2005,Pohl2006,Wang2006}, where the prior 
-imposes consistent voxel-based classification of contiguous regions. Here, the 
-presence of more structures than one unique \gls{roi} helps aligning the target image 
-with the atlas in a hierarchical fashion. Finally, the third group generalizes 
-the atlas to actual images, and the contour is to segment simultaneously two 
-different target images, related by a spatial transform to be co-estimated~\citep{Wyatt2003,
-Yezzi2003}.
+\subsection{Summary}
+\label{sec:contributions}
 

### 2013-05-28 Publications/2013-TMI/04-methods.tex

@@
 %
+\subsection{Methodological background}
+\label{sec:methods_background}
+
+We suggest clustering the current methodologies of template-based segmentation 
+methods into three groups. The first group typically adds a shape prior term to 
+the energy functional of an evolving active contour
+\citep{bresson_variational_2006,chan_level_2005,chen_using_2002,cremers_kernel_2006,
+gastaud_combining_2004,paragios_level_2003,vemuri_joint_2003,yezzi_variational_2003}.
+These methods generally have a explicit description of the expected relative boundary 
+locations of the object to be delineated, and some even model the statistical deviations
+from this average shape. Closely related to this group are atlas-based segmentation
+methods \citep{gorthi_segmentation_2009,gorthi_active_2011,pohl_unifying_2005,
+pohl_bayesian_2006,wang_joint_2006}, where the prior 
+imposes consistent voxel-based classification of contiguous regions. Here, the 
+presence of more structures than one unique \gls{roi} helps aligning the target image 
+with the atlas in a hierarchical fashion. Finally, the third group generalizes 
+the atlas to actual images, and the contour is to segment simultaneously two 
+different target images, related by a spatial transform to be co-estimated
+\citep{wyatt_map_2003,yezzi_variational_2003}.
+
 \subsection{Mumford-Shah Functional derivation from the Maximum A-Posteriori Model}

### 2013-05-28 Publications/2013-TMI/04-methods.tex
Added todo marks

@@
 \end{align}
-(\textbf{FIXME:} bad explanation) where we have a constant term scaled by
+\todo[inline]{improve this explanation} where we have a constant term scaled by
 the total volume of the partition $\Omega_k$ and the determinant of the 

### 2013-05-28 Publications/2013-TMI/05-experiments.tex
Added todo marks

@@
 TrioTim, Siemens, Erlangen, Germany) with a standard 12-channel head coil. 
-(\textbf{FIXME:} check coil number of channels!!)
+\todo[inline]{check coil \# channels}
 Subjects were scanned twice with the same protocol,
@@
 outcome of a widely-used methodology that combines field-map susceptibility 
-correction and T1-T2-\gls{dwi} registration (\textbf{FIXME}: I suggest to give 
-it a compact name and define it in the pre-processing section or even before
-in the intro).
+correction and T1-T2-\gls{dwi} registration .
+\todo[inline]{I suggest to give it a compact name and define it in the 
+pre-processing section or even before in the intro}
 

### 2013-05-29 Publications/2013-TMI/03-introduction.tex
Added todo marks
Another big iteration in #81
* Rewritten significant part of introduction.
* Fixed some references.

@@
 \emph{connectome} \citep{hagmann_diffusion_2005,sporns_human_2005}. The connectome
-analysis is a promising tool for neuroscience and clinical applications 
-(\textbf{FIXME:} why? or references?).
+analysis is a promising tool for neuroscience and clinical applications.
+\todo[inline]{why? or references?}
 
@@
 
+\gls{dwi} data are usually acquired with \gls{epi} sequences 
+as they allow for very fast acquisitions, but they are known to 
+suffer from geometrical distortions and artifacts due to, mainly,
+three sources: the subject motion in between acquiring different
+sampling directions, the induced \emph{Eddy currents} on the scanner 
+coils and finally distortions caused by the magnetic susceptibility inhomogeneity
+present at tissue interface. In this paper, we restrict ourselves to
+the last one, as it accounts for the major impact in the connectome
+analysis. \emph{Susceptibility distortions} happen along the 
+phase-encoding direction, and are most appreciable in the front part of 
+the brain for the strong air/tissue interface surrounding the frontal sinuses.
+Two implications are associated to this artifact: the signal loss caused on
+highly distorted regions, and a significant added difficulty in 
+registering with structural images (e.g. T1-weighted).
+
 \gls{dwi} data is strongly affected by \glspl{pve} \citep{alexander_analysis_2001},
@@
 respect \gls{csf}.
+\todo[inline]{Define \textit{raw data}}
 
@@
 \todo[inline]{Introduce here what are b0, fa, md, and direction dependency problem.}
-These low-$b$ (or $b = 0$) volumes are acquired without direction gradient as reference, 
-and present a T2-like contrast.
+These $b = 0$ volumes (also called \emph{\gls{epi} baseline}, low-$b$, or just B0) 
+are acquired without direction gradient as reference, and they present a T2-like 
+contrast.
 
@@
 are approximated by multivariate Gaussians for first. Iterative clustering performed on the 
-\emph{low}-$b$ volumes of \gls{dwi} data was proposed by \citep{hadjiprocopis_unbiased_2005}.
+B0 volumes of \gls{dwi} data was proposed by \citep{hadjiprocopis_unbiased_2005}.
 Later studies investigated the application of probabilistic frameworks combining mixtures of 
@@
 
-\gls{dwi} data are usually acquired with \gls{epi} sequences 
-as they allow for very fast acquisitions, but they are known to 
-suffer from geometrical distortions and artifacts due to, mainly,
-three sources: the subject motion between direction sampling, the 
-induced \emph{Eddy currents} on the scanner coils and finally
-distortions caused by the magnetic susceptibility inhomogeneity
-present at tissue interface. In this paper, we restrict ourselves to
-the last one, as it accounts for the major impact in the connectome
-analysis. \emph{Susceptibility distortions} happen along the 
-phase-encoding direction, and are most appreciable in the front part of 
-the brain for the strong air/tissue interface surrounding the frontal sinuses.
-
 One approach to correcting the susceptibility distortions was proposed
@@
 obtain a map of magnitude and phase of the actual magnetic field inside the
-scanner. Based on this \emph{fieldmap}, a number of methodologies have 
-been developed to correct for the distortion, and are generically named as 
-\emph{\gls{epi}-unwarp} techniques \cite{holland_efficient_2010,
-hsu_correction_2009,jezzard_characterization_2005, reber_correction_2005}. 
-Unfortunately, the availability of the corresponding fieldmap is not always
-met.\todo[inline]{Mention other problems: low precission, unreliability}
+scanner. Based on this \emph{fieldmap} and the theory underlying the 
+distortion, \citep{jezzard_characterization_2005} proposed a correction 
+methodology. A number of forked and improved methodologies have been 
+developed to correct for the susceptibility distortion and received the
+generic name of ``fieldmap correction'' techniques \cite{hsu_correction_2009,
+reber_correction_2005}.
 
-Some other methodologies do not make use of the fieldmaps, compensating the 
-distortion with non-linear registration from structural \gls{mri} 
-\citep{kybic_unwarping_2000} (see \autoref{sec:registration}), or other means 
-\citep{andersson_modeling_2001}. To our knowledge, there exists no study on 
-the impact of the \gls{epi} distortion on the variability of tractography 
-results.
+Some other retrospective methodologies do not make use of the fieldmaps, 
+as explicitly modeling the distortion \citep{andersson_modeling_2001},
+registering with (anatomically correct) T2-weighted \gls{mri} 
+\citep{kybic_unwarping_2000,studholme_accurate_2000,li_accounting_2007,
+tao_variational_2009} (see \autoref{sec:registration}), or acquiring an 
+extra B0 image with reversed phase encoding direction 
+\citep{holland_efficient_2010}.
+
+To our knowledge, there exists no study on 
+the impact of the susceptibility distortion over the subsequent tractography 
+and connectome analyses. However, a comparison of the diverse correction 
+techniques is found in \citep{wu_comparison_2008}. This
+study claimed that fieldmap correction methodologies are not entirely accurate
+and reliable, even though the method is correct in principle. This conclusion
+was later confirmed by \citep{tao_variational_2009}. Additional concerns regarding
+fieldmap correction are the requirement of an extra acquisition (that is not always
+met or it is impractical), or the accuracy of the measured fieldmap that is sensitive
+to various effects (such us respiration, blood flow, etc.). All these factors 
+have turned susceptibility distortion in \gls{epi} sequences an active field of 
+research for the last 10 years.
 
@@
 
+The last addressed task in the image processing stage of the studied 
+connectivity analysis pipelines is the structural information co-registration.
+The need for this registration step is mainly raised when defining the 
+nodes of the connectivity matrix. This definition step is named
+\textit{cortical parcellation} as it imposes the regions that will be
+considered to cluster the fiber bundles reaching the \gls{wm}-\gls{gm}
+surface. The common procedure to accomplish this clustering is using
+a predefined parcellation in a high-resolution structural T1-weighted 
+atlas. The second goal that has justified the applicability of structural
+images co-registration is susceptibility distortion correction
+as discussed in \autoref{sec:distortion}. In any case, susceptibility 
+distortion hinders a rigid-registration solution to this end.
+Additionally, when applying non-linear intensity-based
+registration algorithms other difficulties have to be addressed as
+the significant \gls{pve} in the \gls{wm}-\gls{gm} layer, or the
+inherent inadequacy of the B0 contrast (the only \textit{directionally
+independent} existing in the raw \gls{dwi} data) due to the almost null
+difference in intensity between \gls{wm} and \gls{gm} pixels. Early 
+registration methods appeared targeting the distortion correction.
+They quickly standardized the choice of T2 as anatomically correct source
+to be registered against the B0 in \gls{dwi}. B0 images have a very similar
+contrast to T2 weighted due to the duality of acquisition procedure. However,
+parcellation is defined in T1 weighted space, adding an additional registration
+step missing on the literature due to the novelty of the task. Even though this
+registration step has a very low complexity compared to the remaining tasks of 
+the pipeline process, it is one additional source of inconsistency and unreliability.
+
+One of the first proposals is \citep{kybic_unwarping_2000}, where the deformation
+field is modeled with B-Splines, the cost function is least squares and optimized
+in a multi-resolution gradient descent strategy. Their method is evaluated in 
+both synthetic and real 2D images. Similarly,
+\citep{studholme_accurate_2000} proposed a spline-based deformation but including
+a weighting factor proportional to the Jacobian of the transform to correct the
+intensity of the undistorted data. They also use the $\log$ of the signal to
+enhance the low-signal regions and optimize with a gradient descent algorithm
+the mutual information of the mapping. The basis in the procedure is still the most 
+extended. For instance, \citep{wu_comparison_2008} also proposed a B-Spline registration
+providing quantitative comparisons with fieldmap correction methods. Recent approaches
+take into account the signal loss due to dephasing \citep{li_accounting_2007}, or 
+introduce more complex variational frameworks \citep{tao_variational_2009}.
+
+The significant benefits of exploiting the anatomical \gls{mri} when 
+segmenting the \gls{dwi} data have been demonstrated \cite{zollei_improved_2010}, 
+justifying the use of the shape prior information. To our knowledge, there is
+no study simultaneously taking advantage from segmentation or distortion correction
+tasks to be applied to the co-registration problem.
+
+\subsection{Summary}
+\label{sec:contributions}
 Therefore, the problems of precise segmentation in \gls{dwi}-space and the 
@@
 readily available in \gls{t1} space can simply be projected on to the 
-\gls{dwi}-data. Conversely, if a precise delineation in \gls{dwi}-space 
+\gls{dwi} data. Conversely, if a precise delineation in \gls{dwi} space 
 was achieved, the spatial mapping with \gls{t1}-space could be derived 
@@
 nor registration can be performed flawlessly, if considered independently.
-The significant benefits of exploiting the anatomical \gls{mri} when 
-segmenting the \gls{dwi} data have been demonstrated \cite{zollei_improved_2010}, 
-justifying the use of the shape prior information.
-
-\subsection{Summary}
-\label{sec:contributions}
 
 In this paper we propose a novel registration framework to simultaneously
-solving the segmentation and distortion challenges, by exploiting as strong 
-shape-prior the detailed morphology extracted from high-resolution anatomical 
-\gls{mri}. Indeed, hereafter we assume this segmentation problem in anatomical 
-images as a solved by widely used available procedures.
-Moreover, the shape prior is of very ``strong'' nature, since it is specific to 
-the particular subject. Also, after global alignment using existing approaches, 
-the remaining spatial deformation between anatomical and diffusion space is 
-due to \gls{epi} distortions. Finally, we need to establish precise spatial 
-correspondence between the surfaces in both spaces, including the tangential 
-direction for parcellation. Therefore, we can reduce the problem to finding 
-the differences of spatial distortion in between anatomical and \gls{dw} space.
+solving the segmentation, distortion and cortical parcellation challenges, 
+by exploiting as strong shape-prior the detailed morphology extracted 
+from high-resolution anatomically correct \gls{mri}. Indeed, hereafter 
+we assume this segmentation problem in anatomical images is reliably and
+accurately solved with readily available tools. After global alignment 
+using existing approaches, the remaining spatial mismatch between 
+anatomical and diffusion space is due to susceptibility distortions.
+Finally, we need to establish precise spatial correspondence between the 
+surfaces in both spaces, including the tangential direction for parcellation.
+Therefore, we can reduce the problem to finding the differences of spatial 
+distortion in between anatomical and \gls{dwi} space.
 We thus reformulate the segmentation problem as an inverse problem, where we 
@@
 correspondence between the contours in both spaces is guaranteed, and projection 
-of parcellisation to \gls{dw} space is implicit and consistent.
+of parcellisation to \gls{dwi} space is implicit and consistent.
 
+\todo[inline]{ (Rewrite) 
 We test our proposed joint segmentation-registration model on two different 
@@
 \autoref{sec:experiments}. We present results in \autoref{sec:results} and conclude 
-in \autoref{sec:conclusion}.
+in \autoref{sec:conclusion}.}

### 2013-06-04 Code/Modules/LevelSets/include/LevelSetsBase.h
Reviewed extents checking methods
Now the problem is *even* more clear. For now, I correct the method
so it is not stopped if the new position is inside the reference image
extents. Close #88.

Anyways, a consistent way to check image boundaries should be implemented

@@ -194,7 +194,7 @@ public:
 
 	MeasureType GetValue();
 
-	virtual DeformationFieldPointer GetLevelSetsMap( DeformationFieldType* levelSetMap);
+	DeformationFieldPointer GetShapeGradients( DeformationFieldType* gradientsMap );
 
 	void UpdateDeformationField( const DeformationFieldType* newField );
 
@@ -224,6 +224,9 @@ protected:
 	}
 
 	virtual void InitializeSamplingGrid( void ) = 0;
+
+	inline bool IsInside( const PixelPointType p, ContinuousIndex& idx ) const;
+	bool CheckExtents( const ContourDeformationType* prior ) const;
 
 	inline virtual MeasureType GetEnergyAtPoint( PixelPointType& point, size_t roi ) = 0;
 

### 2013-06-04 Code/Modules/LevelSets/include/MahalanobisLevelSets.h

@@ -165,7 +165,6 @@ private:
 	MahalanobisLevelSets( const Self &); // purposely not implemented
 	void operator=(const Self &); // purposely not implemented
 
-	bool CheckExtents( ContourDeformationType* prior ) const;
 	bool ParametersInitialized() const;
 
 	InterpolatorPointer m_Interp;

### 2013-06-04 Code/Modules/Optimizers/include/GradientDescentLevelSetsOptimizer.h
Reviewed extents checking methods
Now the problem is *even* more clear. For now, I correct the method
so it is not stopped if the new position is inside the reference image
extents. Close #88.

Anyways, a consistent way to check image boundaries should be implemented

@@ -223,7 +223,7 @@ protected:
 
 	DeformationFieldPointer m_DeformationField;
 	DeformationFieldPointer m_NextDeformationField;
-	DeformationFieldPointer m_SpeedsField;
+	DeformationFieldPointer m_ShapeGradients;
 	TensorFieldPointer m_Denominator;
 	MeasureType m_CurrentLevelSetsValue;
 

### 2013-06-04 Publications/2013-FINF-ITK/01-abstract.tex

In whole-brain connectivity analysis of diffusion MRI (dMRI) data,
an accurate delineation of the white-matter and grey-matter surfaces is required.
While high-standard segmentation is readily available for anatomical MRI, such 
as T1-weighted, dMRI typically have lower resolution and severe 
geometrical distortions. We propose a dMRI segmentation-registration framework 
that exploits the detailed anatomy extracted from anatomical MRI as shape-prior.
We use an ``active contours without edges''-like model to look for a deformation
field that optimally maps the shape prior on the multivariate features 
in diffusion space. This joint approach reflects the intrinsic coupling of 
segmentation and distortion correction. Complementary, a precise and consistent
cortical parcellation on dMRI is straightforward by projection from T1 space. Thus,
we expect to improve the reliability and robustness of the resulting connectivity 
networks and their comparability within and across subjects. First results 
on synthetic datasets and simulated dMRI confirm the effectiveness of our approach.

### 2013-06-04 Code/Modules/LevelSets/include/LevelSetsBase.h
Minor changes to track deformation norm evolution

@@ -196,7 +196,7 @@ public:
 
 	DeformationFieldPointer GetShapeGradients( DeformationFieldType* gradientsMap );
 
-	void UpdateDeformationField( const DeformationFieldType* newField );
+	MeasureType UpdateDeformationField( const DeformationFieldType* newField );
 
 	virtual void Initialize( void );
 

### 2013-06-04 Publications/2013-TMI/03-introduction.tex
Added references
* Added reference to Greve et al (FreeSurfer) regarding Parcellation (close #86)
* Added reference to latest Alessandra's paper regarding structural connectivity analysis.

@@
 \emph{connectome} \citep{hagmann_diffusion_2005,sporns_human_2005}. The connectome
-analysis is a promising tool for neuroscience and clinical applications.
-\todo[inline]{why? or references?}
+analysis is a promising tool for neuroscience and clinical applications
+\citep{griffa_structural_2013}.
 
@@
 jones_white_2012,soares_hitchhikers_2013}, and statistical analysis 
-\citep{meskaldji_comparing_????}. Here,we shall address three tasks included within 
+\citep{meskaldji_comparing_2013}. Here,we shall address three tasks included within 
 the image processing stage in a unified approach: brain tissue segmentation in 
@@
 a predefined parcellation in a high-resolution structural T1-weighted 
-atlas. The second goal that has justified the applicability of structural
+atlas \citep{greve_accurate_2009}. The second goal that has justified the applicability of structural
 images co-registration is susceptibility distortion correction

### 2013-06-13 Publications/2012-SeedManuscript/diffseg.tex
Added original manuscript for reference

\documentclass[a4paper,12pt]{article}
\usepackage{fullpage}

\usepackage[USenglish]{babel}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{pslatex}
\usepackage[pdftex]{graphicx}

\usepackage{url}

%\usepackage{natbib}
%\usepackage{setspace}

\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\argmax}{argmax}
\DeclareMathOperator{\argmin}{argmin}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\const}{Const.}

\title{Subvoxel Segmentation in DW-MR Based on T1-Weighted MR Shape Priors}
\author{Oscar Esteban\thanks{Contributed equally} \and Dominique Zosso\footnotemark[1] \and Alessandro Daducci \and Meritxell Bach Cuadra \and (...) \and Jean-Philippe Thiran} 
\date{version 0.2, dz, \today}

\begin{document}
\maketitle

\section{Introduction}
Precise segmentation of diffusion-weighted MR images is a difficult task due to the very limited resolution (currently around $2.2\times 2.2\times 3 mm^3$). In brain tractography and subsequent connectivity analysis, precise segmentation of the CSF-grey-matter and white-matter-grey-matter interfaces is, however, an important task in order to achieve consistent parcellisation of the cortical surface. The diffusion-weighted acquisitions often suffer from severe distortions due to local field inhomogeneities (in particular in the anterior part of the brain and along the phase-encoded direction).

On the other hand, anatomical MR images (e.g. T1 and T2 weighted acquisitions) can be produced with substantially better resolution, relatively low distortions and well understood tissue contrast.

Currently [...]?

Here we propose to solve the segmentation problem by exploiting the anatomy easily extracted in T1-weighted images as strong shape-priors. We reformulate the segmentation problem as an inverse problem, where we look for an underlying deformation field (the distortion) mapping from T1 space into DW space, such that the structures identified in the T1 image, with high confidence, will optimally align with structures in DW space.

\section{Materials and Methods}

\subsection{T1 image and segmentations}
We have a T1-weighted (or other anatomical high-resolution and low-distortion) image of the subject, as well as reliable segmentations of the pial surface (CSF-grey-matter) and the white-matter-grey-matter interface. The segmentations are available both as surface meshs (e.g. freesurfer meshes) and as corresponding volumes (or sampling points within).

Let us denote $\{c_i\}_{i=1..N_c}$ the nodes of the white-matter-grey-matter interface, and $\{d_i\}_{i=1..N_d}$ the nodes of the pial surface. Within the respective volumes, we have sets of sample points, denoted $\{w_j\}_{j=1..N_w}$, $\{g_j\}_{j=1..N_g}$, and $\{o_j\}_{j=1..N_o}$ for the white matter, grey matter and CSF, respectively. All those are given in high-resolution T1 reference coordinates.

\subsection{DW image}
Low resolution of diffusion-weighted image. At each voxel we have a sampled ODF (or some extracted features).
Let us denote by $x$ the voxel and $f(x) = [ f_1, f_2, \ldots, f_N]^T(x)$ its associated feature vector.

\subsection{Distortion field parametrisation}
The transformation from T1 into DW reference coordinate space is achieved through a dense deformation field $u(x)$, such that for example the nodes of the w/g-interface are located in DW-space as follows:
\begin{equation}
c_i' = T\{c_i\} = c_i + u(c_i)
\end{equation} 
Since the nodes of the anatomical surfaces might lay off-grid, it is required to derive $u(x)$ from a discrete set of parameters $\{u_k\}_{k=1..K}$. Densification is achieved through a set of associated basis functions $\psi_k$ (e.g. rbf, interpolation splines):
\begin{equation}
u(x) = \sum_k \psi_k(x) u_k
\end{equation}
Consequently, the transformation writes
\begin{equation}
c_i' = T\{c_i\} = c_i + u(c_i) = c_i + \sum_k \psi_k(c_i)u_k
\end{equation} 
Note that, since $c_i$ remains constant in the DW segmentation process, the values of $\psi_k(c_i)$ can be precomputed. Also, provided compact support of the basis functions, the system remains relatively sparse.

\subsection{DW region descriptors}
Based on the region-wise samples $\{w_j\}$, $\{g_j\}$, and $\{o_j\}$, and the current estimate of the distortion $u$, we can compute ``expected samples'' of the respective regions in DW space, written $\{w_j'\}$, $\{g_j'\}$, and $\{o_j'\}$. Based on those samples, we may now estimate region descriptors of the DW features $f(x)$ of the three respective regions in DW space. In the simplest case, estimate the parameters $\mu_R$ and $\Sigma_R$, i.e. the regions' mean feature vector and covariance matrix, for each region $R\in \{w,g,o\}$.

\subsection{Active contours without edges}
Based on these Gaussian region descriptors, we propose an ACWE-like, piece-wise constant, variational image segmentation model (where the unknown is the deformation field)\cite{Chan2001}:
\begin{align}
E(u) &= \int_{w'(u)} (f-\mu_w)^T\Sigma_w^{-1}(f-\mu_w) dx\nonumber\\
&\quad +\int_{g'(u)} (f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g) dx\\
&\quad +\int_{o'(u)} (f-\mu_o)^T\Sigma_o^{-1}(f-\mu_o) dx\nonumber
\end{align}
where the integral domains depend on the deformation field u. Note that minimizing this energy, $\argmin_u\{E\}$, yields the MAP estimate of a piece-wise smooth image model affected by Gaussian additive noise. This inverse problem is ill-posed \cite{Hadamard1902,Bertero1988}. In order to account for deformation field regularity and to render the problem well-posed, we include limiting and regularization terms into the energy functional \cite{Tichonov1963,Morozov1975}:
\begin{align}
E(u) &= \int_{w'(u)} (f-\mu_w)^T\Sigma_w^{-1}(f-\mu_w) dx\nonumber\\
&\quad +\int_{g'(u)} (f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g) dx\\
&\quad +\int_{o'(u)} (f-\mu_o)^T\Sigma_o^{-1}(f-\mu_o) dx\nonumber\\
&\quad +\int u^T A u dx + \int \tr\{(\nabla u^T)^T B (\nabla u^T)\} dx\nonumber
\end{align}
These regularity terms ensure, that the segmenting contours in DW space are still close to their native shape in T1. In the simplest case, both $A=\alpha$ and $B=\beta$ are constant scalars, and the terms rewrite $\alpha \int  \|u\|^2 dx + \beta \int \left( \|\nabla u_x\|^2 + \|\nabla u_y\|^2 + \|\nabla u_z\|^2\right) dx$, which are familiar. Later, $A$ and $B$ might be spatially varying and/or $3\times 3$ matrices, therefore allowing to incorporate inhomogeneous and anisotropic regularization \cite{Nagel1986}.

\subsection{Operator splitting}
In order to make the Energy minimisation computationally more tractable, we propose the following operator splitting: Let us optimize the data terms and the regularity terms on separate copies of the deformation field, now called $u$ and $v$, constrained to be equal:
\begin{align}
E(u,v) &= \int_{w'(u)} (f-\mu_w)^T\Sigma_w^{-1}(f-\mu_w) dx\nonumber\\
&\quad +\int_{g'(u)} (f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g) dx\\
&\quad +\int_{o'(u)} (f-\mu_o)^T\Sigma_o^{-1}(f-\mu_o) dx\nonumber\\
&\quad +\int v^T A v dx + \int \tr\{(\nabla v^T)^T B (\nabla v^T)\} dx\nonumber
\end{align}
and now
\begin{equation}
\min_{u,v} \{ E \} \quad s.t. \quad u = v
\end{equation}
In order to take the equality constraint into account, we may make use of augmented Lagrangians (a combination of Lagrangian multipliers and penalty terms on the constraint) \cite{Bertsekas1976,Glowinski1989,Nocedal2006}:
\begin{equation}
AL(u,v,\lambda,r) = E(u,v) + \langle \lambda, u-v \rangle + \frac{r}{2} \| u - v \|_2^2
\end{equation}
To solve the constraint minimization problem, we may now optimize the Augmented Lagrangian in an iterative way:
\begin{equation}
\left\lbrace \begin{array}{rcl}
u^{t+1} &=& \argmin_{u} AL(u,v^t,\lambda^t,r)\\
v^{t+1} &=& \argmin_{v} AL(u^{t+1},v,\lambda^t,r)\\
\lambda^{t+1} &=& \lambda^t + \rho(u^{t+1}-v^{t+1})
\end{array}\right.\end{equation}
where typically $0 < \rho < r$. The two subminimization problems will now be much easier to handle than the original complete problem (``divide et impera'').

\subsection{Shape gradients}
To compute the gradient-descent of the data-term domain integrals with respect to the underlying deformation field, we want to make use of shape gradients \cite{Jehan-Besson2003,Herbulot2006}. A little bit of theory is therefore in order.

Let $\Omega$ be an image domain and $\omega$ its boundary. Further, $r(x)$ is an ``arbitrary'' function over the image domain. We now derive the domain integral w.r.t. the contour evolution parameter $\tau$ (~ time):
\begin{equation}
\frac{\partial}{\partial \tau} \int_\Omega r(x) dx = \int_\Omega \frac{\partial r}{\partial \tau}(x) dx - \int_\omega r(x) \left\langle \frac{\partial\omega}{\partial\tau}, N_\omega\right\rangle dx
\end{equation}
where $\left\langle\frac{\partial\omega}{\partial\tau}, N_\omega\right\rangle$ is the projection of the boundary movement on the unit inward normal.

\subsection{Min w.r.t. $u$}
The first minimization problem optimizes the data-term. The problem is equivalent to minimizing the following energy:
\begin{align}
E(u) &= \int_{w'(u)} (f-\mu_w)^T\Sigma_w^{-1}(f-\mu_w) dx\nonumber\\
&\quad +\int_{g'(u)} (f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g) dx\\
&\quad +\int_{o'(u)} (f-\mu_o)^T\Sigma_o^{-1}(f-\mu_o) dx\nonumber\\
&\quad + \langle \lambda, u-v \rangle + \frac{r}{2} \| u - v \|_2^2\nonumber
\end{align}
where we identify three instances of domain integrals of the form $\int_\Omega r(x) dx$. Optimality requires the derivative of this energy with respect to the parameters $u$ to be null. At this point, we may decide to ignore the influence of the boundary shift on the statistics of the regions (i.e. moving the boundary does not significantly impact the $\mu$ and $\Sigma$ descriptors). This means that we can drop the derivative of $r(x)$ w.r.t. contour evolution. What remains, are surface integrals at the two respective domain interfaces, $c'$ (wm/gm) and $d'$ (gm/CSF), plus the Lagrangian and penalty terms:
\begin{align}
\frac{\partial E}{\partial u_k^a} &= \int_{c'} \left[(f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g) - (f-\mu_w)^T\Sigma_w^{-1}(f-\mu_w)\right]\left\langle\frac{\partial c'(s)}{\partial u_k^a}, N_{c'}(s)\right\rangle ds \nonumber\\
&\quad + \int_{d'} \left[(f-\mu_o)^T\Sigma_o^{-1}(f-\mu_o) - (f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g)\right]\left\langle\frac{\partial d'(s)}{\partial u_k^a}, N_{d'}(s)\right\rangle ds \\
&\quad + \lambda_k^a + r(u_k^a - v_k^a)\nonumber\\
& = 0\nonumber
\end{align}
where $u_k^a$ is the $a$-th component of the parameter $u_k$, $a\in \{x,y,z\}$, $s$ is the surface parameter, $c'(s)$ and $d'(s)$ the corresponding points on the surfaces $c'$ and $d'$, and $N_{c'}(s)$ and $N_{d'}(s)$ the associated brainwise-inward unit normals.
Given the deformation field interpolation stated above, the boundary moves according to
\begin{equation}
\frac{\partial c'(s)}{\partial u_k^a} = \psi_k(c'(s))e_a
\end{equation}
where $e_a$ is the unit vector along direction $a$, and thus
\begin{equation}
\left\langle\frac{\partial c'(s)}{\partial u_k^a}, N_{c'}(s)\right\rangle = \psi_k(c'(s))N_{c'}^a(s)
\end{equation}
Now, we will discretize the surface integrals over $c'$ and $d'$ by simply summing over the nodes $c_i'$ and $d_i'$:
\begin{align}
\frac{\partial E}{\partial u_k^a} &= \sum_{i=1}^{N_c} \left[(f(c_i')-\mu_g)^T\Sigma_g^{-1}(f(c_i')-\mu_g) - (f(c_i')-\mu_w)^T\Sigma_w^{-1}(f(c_i')-\mu_w)\right]\psi_k(c_i')N_{c_i'}^a \nonumber\\
&\quad + \sum_{i=1}^{N_d} \left[(f(d_i')-\mu_o)^T\Sigma_o^{-1}(f(d_i')-\mu_o) - (f(d_i')-\mu_g)^T\Sigma_g^{-1}(f(d_i')-\mu_g)\right]\psi_k(d_i')N_{d_i'}^a \\
&\quad + \lambda_k^a + r(u_k^a - v_k^a)\nonumber\\
& = 0\nonumber
\end{align}
It is straightforward to solve this equation for $u_k^a$. The optimal distortion $u_k$ is found at each iteration as:
\begin{align}
u_k^{t+1} &= v_k^t - \frac{1}{r}\lambda_k^{t}\nonumber\\
&\quad - \frac{1}{r}\sum_{i=1}^{N_c} \left[(f(c_i')-\mu_g)^T\Sigma_g^{-1}(f(c_i')-\mu_g) - (f(c_i')-\mu_w)^T\Sigma_w^{-1}(f(c_i')-\mu_w)\right]\psi_k(c_i')N_{c_i'}\\
&\quad - \frac{1}{r}\sum_{i=1}^{N_d} \left[(f(d_i')-\mu_o)^T\Sigma_o^{-1}(f(d_i')-\mu_o) - (f(d_i')-\mu_g)^T\Sigma_g^{-1}(f(d_i')-\mu_g)\right]\psi_k(d_i')N_{d_i'}\nonumber
\end{align}

\subsection{Min w.r.t. $v$}

\textbf{It is important to realize that here we do not regularize the actual deformation field (i.e. after interpolation), but the underlying raw parameter field. }

For the optimization w.r.t. $v$, the relevant energy writes
\begin{align}
E(v) &= \int v^T A v dx + \int \tr\{(\nabla v^T)^T B (\nabla v^T)\} dx\\
&\quad + \langle \lambda, u-v \rangle + \frac{r}{2} \| u - v \|_2^2\nonumber
\end{align}
Let's assume the simplest, homogeneous isotropic case, $A = \alpha/2$ and $B = \beta/2$. The associated Euler-Lagrange equation is found as:
\begin{equation}
\alpha v - \beta\Delta v + rv = ru + \lambda
\end{equation}
which easily translates into Fourier domain:
\begin{equation}
v^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{ru + \lambda\}}{\mathcal{F}\{(\alpha+r)\mathcal{I}-\beta\Delta\}} \right\}
\end{equation}
where $\mathcal{I}$ denotes the identity operator.

Here, we rewrite the Laplacian as a linear combination of the identity and shift operators:
\begin{equation}
\Delta = \mathcal{S}_x^- + \mathcal{S}_x^+ + \mathcal{S}_y^- + \mathcal{S}_y^+ - 4 \mathcal{I}
\end{equation}
where $\mathcal{S}_{x,y}^{\pm}$ stands for the forward ($+$) and backward ($-$) shift operator along $x$ or $y$, respectively, of which the Fourier transform is found easily as
\begin{equation}
\mathcal{F}\{\mathcal{S}_{x,y}^{\pm}\} = e^{\pm i\omega_{x,y}},
\end{equation}
where $\omega_{x,y}$ is the normalized pulsation along $x$- and $y$-direction. Accordingly, the Fourier transform of the discrete Laplacian is found as
\begin{align}
\mathcal{F}\{\Delta\} &= e^{-i\omega_x } + e^{i\omega_x } + e^{-i\omega_y } + e^{i\omega_y } - 4\nonumber\\
&= 2\left( \cos(\omega_x) + \cos(\omega_y) - 2 \right)
\end{align}

The remaining transforms are trivial or can be computed using FFT (as in \cite{Estellers2011b}).

\subsection{Lagrangian multiplier update}
At each iteration, the Lagrangian multipliers are updated as noted before:
\begin{equation}
\lambda^{t+1} = \lambda^t + \rho(u^{t+1}-v^{t+1})
\end{equation}

\subsection{Region descriptor reestimation}
In regular intervals, i.e. after $n$ iterations, the parameters $\mu$ and $\Sigma$ of the involved regions need to be reestimated based on the shifted volumetric samples $w_j'$, $g_j'$ and $o_j'$.

\subsection{Convergence}
In order to ``fixate'' the evolution when close to convergence, it is advised to slightly increase the penalty weight $r$ at each iteration. Note that as can be seen in the above equations, $r$ governs the step-size or  leash-length at each iteration, i.e. the amount by which the new estimate $u$ may move away from the preceding $v$ and vice-versa.

\subsection{Alternative route: Gradient descent}
If one absolutely wants to avoid the operator splitting and the augmented Lagrangians, then simple gradient descent may work as well (at least under the same simple model circumstances).

At each iteration, we update the distortion along the steepest energy descent:
\begin{equation}
\frac{\partial u_k^t}{\partial t} = -\frac{\partial E(u)}{\partial u_k^t}
\end{equation}
At this point, we interpret the parameter field $u_k$ to be a continuous function $u(x)$, sampled at the locations $x_k$, and determine the gradient-descent equation\footnote{The same assumption was being made above in the minimization of the regularity terms}:
\begin{align}
\frac{\partial u^t}{\partial t} &= - \sum_{i=1}^{N_c} \left[(f(c_i')-\mu_g)^T\Sigma_g^{-1}(f(c_i')-\mu_g) - (f(c_i')-\mu_w)^T\Sigma_w^{-1}(f(c_i')-\mu_w)\right]\psi_{c_i'}(x)N_{c_i'}\nonumber\\
&\quad -\sum_{i=1}^{N_d} \left[(f(d_i')-\mu_o)^T\Sigma_o^{-1}(f(d_i')-\mu_o) - (f(d_i')-\mu_g)^T\Sigma_g^{-1}(f(d_i')-\mu_g)\right]\psi_{d_i'}(x)N_{d_i'}\\
&\quad -\alpha u + \beta\Delta u\nonumber
\end{align}
where we have swapped $\psi_k(c_i')$ into $\psi_{c_i'}(x)$.

This gradient descent step can be efficiently tackled by discretizing the time in a forward Euler scheme, and making the right hand side semi-implicit in the regularization terms:
\begin{align}
\frac{u^{t+1}-u^t}{\tau} &= - \sum_{i=1}^{N_c} \left[(f(c_i')-\mu_g)^T\Sigma_g^{-1}(f(c_i')-\mu_g) - (f(c_i')-\mu_w)^T\Sigma_w^{-1}(f(c_i')-\mu_w)\right]\psi_{c_i'}(x)N_{c_i'}\nonumber\\
&\quad -\sum_{i=1}^{N_d} \left[(f(d_i')-\mu_o)^T\Sigma_o^{-1}(f(d_i')-\mu_o) - (f(d_i')-\mu_g)^T\Sigma_g^{-1}(f(d_i')-\mu_g)\right]\psi_{d_i'}(x)N_{d_i'}\nonumber\\
&\quad -\alpha u^{t+1} + \beta\Delta u^{t+1}
\end{align}
where the data terms remain functions of the current estimate $u^t$, i.e. all $c_i' = c_i'(u^t)$ and $d_i' = d_i'(u^t)$. Again, we propose a spectral approach to solve this implicit scheme:
\begin{equation}
u^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{u^t/\delta - \sum_{i=1}^{N_c}(\ldots) - \sum_{i=1}^{N_d}(\ldots)  \}}{\mathcal{F}\{(1/\delta+\alpha)\mathcal{I}-\beta\Delta\}} \right\}
\end{equation}
It is easily verified, that the same update can be obtained by plugging the AL-update w.r.t. $u$ into the AL-update w.r.t. $v$, and by identifying $r = 1/\delta$ (the only exception is the distortion on which the data-term is being evaluated).

\section{Experiments}
\section{Results and Discussion}
\section{Conclusions \& Outlook}
Possible complications:
\begin{itemize}
\item non-Gaussian region models (e.g. penalize entropy instead)
\item inhomogeneous, anisotropic regularization (that complicates alot!)
\item regularize dense distortion field rather than raw parameter field (in particular if number of params is limited)
\item take the region terms in the shape gradient into account
\end{itemize}

%%%% Bibliography  %%%%%%%%%%
\bibliographystyle{IEEEtran}
\bibliography{library}

\end{document}

### 2013-06-19 Publications/2013-TMI/04-methods.tex

@@
 %
-\subsection{Methodological background}
+\subsection{Related work}
 \label{sec:methods_background}
@@
 methods into three groups. The first group typically adds a shape prior term to 
-the energy functional of an evolving active contour
-\citep{bresson_variational_2006,chan_level_2005,chen_using_2002,cremers_kernel_2006,
-gastaud_combining_2004,paragios_level_2003,vemuri_joint_2003,yezzi_variational_2003}.
+the energy functional of an evolving active contour \citep{bresson_variational_2006,c
+han_level_2005,chen_using_2002,cremers_kernel_2006,gastaud_combining_2004}.
 These methods generally have a explicit description of the expected relative boundary 
 locations of the object to be delineated, and some even model the statistical deviations
-from this average shape. Closely related to this group are atlas-based segmentation
-methods \citep{gorthi_segmentation_2009,gorthi_active_2011,pohl_unifying_2005,
+from this average shape. By including a coordinates mapping, it is possible to perform
+active contours based registration between timesteps in a time-series or between different
+images \citep{wyatt_map_2003,paragios_level_2003,vemuri_joint_2003,yezzi_variational_2003}.
+This second group is closely related to atlas-based segmentation methods 
+\citep{gorthi_segmentation_2009,gorthi_active_2011,pohl_unifying_2005,
 pohl_bayesian_2006,wang_joint_2006}, where the prior 
@@
 different target images, related by a spatial transform to be co-estimated
-\citep{wyatt_map_2003,yezzi_variational_2003}.
 
-\subsection{Mumford-Shah Functional derivation from the Maximum A-Posteriori Model}
+\subsection{From the Maximum A-Posteriori Model to the Mumford-Shah functional}
 \label{sec:methods_map}
@@
 Densification is achieved through a set of associated basis functions 
-$\Psi_k$ (e.g. rbf, interpolation splines):
+$\psi_k$ (e.g. rbf, interpolation splines):
 %
 \begin{equation}
-u(x) = \sum_k \Psi_k(x) u_k
+u(x) = \sum_k \psi_k(x) u_k
 \end{equation}
@@
 \label{eq:transformation}
-c_i' = T\{c_i\} = c_i + u(c_i) = c_i + \sum_k \Psi_k(c_i)u_k
+c_i' = T\{c_i\} = c_i + u(c_i) = c_i + \sum_k \psi_k(c_i)u_k
 \end{equation} 
@@
 %Note that, since $c_i$ remains constant in the DW segmentation process,
-%the values of $\Psi_k(c_i)$ can be precomputed. Also, provided compact 
+%the values of $\psi_k(c_i)$ can be precomputed. Also, provided compact 
 %support of the basis functions, the system remains relatively sparse.\\
@@
 \begin{align}
-\frac{u^{t+1}-u^t}{\tau} &= - \sum_{i=1}^{N_c} \left[ e(f(c_i'))  \hat{n}_{c_i'} \Psi_{c_i}(x) \right] -\alpha u^{t+1} + \beta\Delta u^{t+1}
+\frac{u^{t+1}-u^t}{\tau} &= - \sum_{i=1}^{N_c} \left[ e(f(c_i'))  \hat{n}_{c_i'} \psi_{c_i}(x) \right] -\alpha u^{t+1} + \beta\Delta u^{t+1}
 \end{align}
@@
 u^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{u^t/\tau
-- \sum_{i=1}^{N_c} \left[e(f(c_i')) \hat{n}_{c_i'} \Psi_{c_i}(x) \right]  \}}{\mathcal{F}\{(1/\tau+\alpha)\mathcal{I}-\beta\Delta\}} \right\}
+- \sum_{i=1}^{N_c} \left[e(f(c_i')) \hat{n}_{c_i'} \psi_{c_i}(x) \right]  \}}{\mathcal{F}\{(1/\tau+\alpha)\mathcal{I}-\beta\Delta\}} \right\}
 \end{equation}
 %
+
+\subsection{Operator splitting}
+In order to make the Energy minimisation computationally more tractable, we propose the following operator splitting: Let us optimize the data terms and the regularity terms on separate copies of the deformation field, now called $u$ and $v$, constrained to be equal:
+\begin{align}
+E(u,v) &= \int_{w'(u)} (f-\mu_w)^T\Sigma_w^{-1}(f-\mu_w) dx\nonumber\\
+&\quad +\int_{g'(u)} (f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g) dx\\
+&\quad +\int_{o'(u)} (f-\mu_o)^T\Sigma_o^{-1}(f-\mu_o) dx\nonumber\\
+&\quad +\int v^T A v dx + \int \tr\{(\nabla v^T)^T B (\nabla v^T)\} dx\nonumber
+\end{align}
+and now
+\begin{equation}
+\min_{u,v} \{ E \} \quad s.t. \quad u = v
+\end{equation}
+In order to take the equality constraint into account, we may make use of augmented Lagrangians (a combination of Lagrangian multipliers and penalty terms on the constraint)
+\cite{bertsekas_multiplier_1976,glowinski_augmented_1989,nocedal_numerical_2006}:
+\begin{equation}
+AL(u,v,\lambda,r) = E(u,v) + \langle \lambda, u-v \rangle + \frac{r}{2} \| u - v \|_2^2
+\end{equation}
+To solve the constraint minimization problem, we may now optimize the Augmented Lagrangian in an iterative way:
+\begin{equation}
+\left\lbrace \begin{array}{rcl}
+u^{t+1} &=& \argmin_{u} AL(u,v^t,\lambda^t,r)\\
+v^{t+1} &=& \argmin_{v} AL(u^{t+1},v,\lambda^t,r)\\
+\lambda^{t+1} &=& \lambda^t + \rho(u^{t+1}-v^{t+1})
+\end{array}\right.\end{equation}
+where typically $0 < \rho < r$. The two subminimization problems will now be much easier to handle than the original complete problem (``divide et impera'').
+
+\subsection{Shape gradients}
+To compute the gradient-descent of the data-term domain integrals with respect to the underlying deformation field, we want to make use of shape gradients \cite{Jehan-Besson2003,herbulot_segmentation_2006}.
+A little bit of theory is therefore in order.
+
+Let $\Omega$ be an image domain and $\omega$ its boundary.
+Further, $r(x)$ is an ``arbitrary'' function over the image domain.
+We now derive the domain integral w.r.t.
+the contour evolution parameter $\tau$ (~ time):
+\begin{equation}
+\frac{\partial}{\partial \tau} \int_\Omega r(x) dx = \int_\Omega \frac{\partial r}{\partial \tau}(x) dx - \int_\omega r(x) \left\langle \frac{\partial\omega}{\partial\tau}, N_\omega\right\rangle dx
+\end{equation}
+where $\left\langle\frac{\partial\omega}{\partial\tau}, N_\omega\right\rangle$ is the projection of the boundary movement on the unit inward normal.
+
+\subsection{Min w.r.t. $u$}
+The first minimization problem optimizes the data-term. The problem is equivalent to minimizing the following energy:
+\begin{align}
+E(u) &= \int_{w'(u)} (f-\mu_w)^T\Sigma_w^{-1}(f-\mu_w) dx\nonumber\\
+&\quad +\int_{g'(u)} (f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g) dx\\
+&\quad +\int_{o'(u)} (f-\mu_o)^T\Sigma_o^{-1}(f-\mu_o) dx\nonumber\\
+&\quad + \langle \lambda, u-v \rangle + \frac{r}{2} \| u - v \|_2^2\nonumber
+\end{align}
+where we identify three instances of domain integrals of the form $\int_\Omega r(x) dx$.
+Optimality requires the derivative of this energy with respect to the parameters $u$ to be null.
+At this point, we may decide to ignore the influence of the boundary shift on the statistics of the regions (i.e.
+moving the boundary does not significantly impact the $\mu$ and $\Sigma$ descriptors).
+This means that we can drop the derivative of $r(x)$ w.r.t.
+contour evolution.
+What remains, are surface integrals at the two respective domain interfaces, $c'$ (wm/gm) and $d'$ (gm/CSF), plus the Lagrangian and penalty terms:
+\begin{align}
+\frac{\partial E}{\partial u_k^a} &= \int_{c'} \left[(f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g) - (f-\mu_w)^T\Sigma_w^{-1}(f-\mu_w)\right]\left\langle\frac{\partial c'(s)}{\partial u_k^a}, N_{c'}(s)\right\rangle ds \nonumber\\
+&\quad + \int_{d'} \left[(f-\mu_o)^T\Sigma_o^{-1}(f-\mu_o) - (f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g)\right]\left\langle\frac{\partial d'(s)}{\partial u_k^a}, N_{d'}(s)\right\rangle ds \\
+&\quad + \lambda_k^a + r(u_k^a - v_k^a)\nonumber\\
+& = 0\nonumber
+\end{align}
+where $u_k^a$ is the $a$-th component of the parameter $u_k$, $a\in \{x,y,z\}$, $s$ is the surface parameter, $c'(s)$ and $d'(s)$ the corresponding points on the surfaces $c'$ and $d'$, and $N_{c'}(s)$ and $N_{d'}(s)$ the associated brainwise-inward unit normals.
+Given the deformation field interpolation stated above, the boundary moves according to
+\begin{equation}
+\frac{\partial c'(s)}{\partial u_k^a} = \psi_k(c'(s))e_a
+\end{equation}
+where $e_a$ is the unit vector along direction $a$, and thus
+\begin{equation}
+\left\langle\frac{\partial c'(s)}{\partial u_k^a}, N_{c'}(s)\right\rangle = \psi_k(c'(s))N_{c'}^a(s)
+\end{equation}
+Now, we will discretize the surface integrals over $c'$ and $d'$ by simply summing over the nodes $c_i'$ and $d_i'$:
+\begin{align}
+\frac{\partial E}{\partial u_k^a} &= \sum_{i=1}^{N_c} \left[(f(c_i')-\mu_g)^T\Sigma_g^{-1}(f(c_i')-\mu_g) - (f(c_i')-\mu_w)^T\Sigma_w^{-1}(f(c_i')-\mu_w)\right]\psi_k(c_i')N_{c_i'}^a \nonumber\\
+&\quad + \sum_{i=1}^{N_d} \left[(f(d_i')-\mu_o)^T\Sigma_o^{-1}(f(d_i')-\mu_o) - (f(d_i')-\mu_g)^T\Sigma_g^{-1}(f(d_i')-\mu_g)\right]\psi_k(d_i')N_{d_i'}^a \\
+&\quad + \lambda_k^a + r(u_k^a - v_k^a)\nonumber\\
+& = 0\nonumber
+\end{align}
+It is straightforward to solve this equation for $u_k^a$. The optimal distortion $u_k$ is found at each iteration as:
+\begin{align}
+u_k^{t+1} &= v_k^t - \frac{1}{r}\lambda_k^{t}\nonumber\\
+&\quad - \frac{1}{r}\sum_{i=1}^{N_c} \left[(f(c_i')-\mu_g)^T\Sigma_g^{-1}(f(c_i')-\mu_g) - (f(c_i')-\mu_w)^T\Sigma_w^{-1}(f(c_i')-\mu_w)\right]\psi_k(c_i')N_{c_i'}\\
+&\quad - \frac{1}{r}\sum_{i=1}^{N_d} \left[(f(d_i')-\mu_o)^T\Sigma_o^{-1}(f(d_i')-\mu_o) - (f(d_i')-\mu_g)^T\Sigma_g^{-1}(f(d_i')-\mu_g)\right]\psi_k(d_i')N_{d_i'}\nonumber
+\end{align}
+
+\subsection{Min w.r.t. $v$}
+
+\textbf{It is important to realize that here we do not regularize the actual deformation field (i.e. after interpolation), but the underlying raw parameter field. }
+
+For the optimization w.r.t. $v$, the relevant energy writes
+\begin{align}
+E(v) &= \int v^T A v dx + \int \tr\{(\nabla v^T)^T B (\nabla v^T)\} dx\\
+&\quad + \langle \lambda, u-v \rangle + \frac{r}{2} \| u - v \|_2^2\nonumber
+\end{align}
+Let's assume the simplest, homogeneous isotropic case, $A = \alpha/2$ and $B = \beta/2$. The associated Euler-Lagrange equation is found as:
+\begin{equation}
+\alpha v - \beta\Delta v + rv = ru + \lambda
+\end{equation}
+which easily translates into Fourier domain:
+\begin{equation}
+v^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{ru + \lambda\}}{\mathcal{F}\{(\alpha+r)\mathcal{I}-\beta\Delta\}} \right\}
+\end{equation}
+where $\mathcal{I}$ denotes the identity operator.
+
+Here, we rewrite the Laplacian as a linear combination of the identity and shift operators:
+\begin{equation}
+\Delta = \mathcal{S}_x^- + \mathcal{S}_x^+ + \mathcal{S}_y^- + \mathcal{S}_y^+ - 4 \mathcal{I}
+\end{equation}
+where $\mathcal{S}_{x,y}^{\pm}$ stands for the forward ($+$) and backward ($-$) shift operator along $x$ or $y$, respectively, of which the Fourier transform is found easily as
+\begin{equation}
+\mathcal{F}\{\mathcal{S}_{x,y}^{\pm}\} = e^{\pm i\omega_{x,y}},
+\end{equation}
+where $\omega_{x,y}$ is the normalized pulsation along $x$- and $y$-direction. Accordingly, the Fourier transform of the discrete Laplacian is found as
+\begin{align}
+\mathcal{F}\{\Delta\} &= e^{-i\omega_x } + e^{i\omega_x } + e^{-i\omega_y } + e^{i\omega_y } - 4\nonumber\\
+&= 2\left( \cos(\omega_x) + \cos(\omega_y) - 2 \right)
+\end{align}
+
+The remaining transforms are trivial or can be computed using FFT (as in \citep{estellers_efficient_2011}).
+
+\subsection{Lagrangian multiplier update}
+At each iteration, the Lagrangian multipliers are updated as noted before:
+\begin{equation}
+\lambda^{t+1} = \lambda^t + \rho(u^{t+1}-v^{t+1})
+\end{equation}
+
+\subsection{Region descriptor reestimation}
+In regular intervals, i.e. after $n$ iterations, the parameters $\mu$ and $\Sigma$ of the involved regions need to be reestimated based on the shifted volumetric samples $w_j'$, $g_j'$ and $o_j'$.
+
+\subsection{Convergence}
+In order to ``fixate'' the evolution when close to convergence, it is advised to slightly increase the penalty weight $r$ at each iteration.
+Note that as can be seen in the above equations, $r$ governs the step-size or  leash-length at each iteration, i.e.
+the amount by which the new estimate $u$ may move away from the preceding $v$ and vice-versa.
+
+\subsection{Alternative route: Gradient descent}
+If one absolutely wants to avoid the operator splitting and the augmented Lagrangians, then simple gradient descent may work as well (at least under the same simple model circumstances).
+
+At each iteration, we update the distortion along the steepest energy descent:
+\begin{equation}
+\frac{\partial u_k^t}{\partial t} = -\frac{\partial E(u)}{\partial u_k^t}
+\end{equation}
+At this point, we interpret the parameter field $u_k$ to be a continuous function $u(x)$, sampled at the locations $x_k$, and determine the gradient-descent equation\footnote{The same assumption was being made above in the minimization of the regularity terms}:
+\begin{align}
+\frac{\partial u^t}{\partial t} &= - \sum_{i=1}^{N_c} \left[(f(c_i')-\mu_g)^T\Sigma_g^{-1}(f(c_i')-\mu_g) - (f(c_i')-\mu_w)^T\Sigma_w^{-1}(f(c_i')-\mu_w)\right]\psi_{c_i}(x)N_{c_i'}\nonumber\\
+&\quad -\sum_{i=1}^{N_d} \left[(f(d_i')-\mu_o)^T\Sigma_o^{-1}(f(d_i')-\mu_o) - (f(d_i')-\mu_g)^T\Sigma_g^{-1}(f(d_i')-\mu_g)\right]\psi_{d_i}(x)N_{d_i'}\\
+&\quad -\alpha u + \beta\Delta u\nonumber
+\end{align}
+where we have swapped $\psi_k(c_i')$ into $\psi_{c_i}(x)$.
+
+This gradient descent step can be efficiently tackled by discretizing the time in a forward Euler scheme, and making the right hand side semi-implicit in the regularization terms:
+\begin{align}
+\frac{u^{t+1}-u^t}{\tau} &= - \sum_{i=1}^{N_c} \left[(f(c_i')-\mu_g)^T\Sigma_g^{-1}(f(c_i')-\mu_g) - (f(c_i')-\mu_w)^T\Sigma_w^{-1}(f(c_i')-\mu_w)\right]\psi_{c_i}(x)N_{c_i'}\nonumber\\
+&\quad -\sum_{i=1}^{N_d} \left[(f(d_i')-\mu_o)^T\Sigma_o^{-1}(f(d_i')-\mu_o) - (f(d_i')-\mu_g)^T\Sigma_g^{-1}(f(d_i')-\mu_g)\right]\psi_{d_i}(x)N_{d_i'}\nonumber\\
+&\quad -\alpha u^{t+1} + \beta\Delta u^{t+1}
+\end{align}
+where the data terms remain functions of the current estimate $u^t$, i.e. all $c_i' = c_i'(u^t)$ and $d_i' = d_i'(u^t)$. Again, we propose a spectral approach to solve this implicit scheme:
+\begin{equation}
+u^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{u^t/\delta - \sum_{i=1}^{N_c}(\ldots) - \sum_{i=1}^{N_d}(\ldots)  \}}{\mathcal{F}\{(1/\delta+\alpha)\mathcal{I}-\beta\Delta\}} \right\}
+\end{equation}
+It is easily verified, that the same update can be obtained by plugging the AL-update w.r.t.
+$u$ into the AL-update w.r.t.
+$v$, and by identifying $r = 1/\delta$ (the only exception is the distortion on which the data-term is being evaluated).

### 2013-06-25 Publications/2013-TMI/04-methods.tex
Updated paper
New references, new methods part, etc.

@@
 
-We suggest clustering the current methodologies of template-based segmentation 
-methods into three groups. The first group typically adds a shape prior term to 
-the energy functional of an evolving active contour \citep{bresson_variational_2006,c
-han_level_2005,chen_using_2002,cremers_kernel_2006,gastaud_combining_2004}.
+\Gls{acwe} and level-set based formulations have been widely applied in image
+processing, as reviewed in \citep{suri_shape_2002}. We suggest clustering the 
+current methodologies of template-based segmentation methods into three groups. 
+The first group typically adds a shape prior term to the energy functional of 
+an evolving active contour \citep{bresson_variational_2006,
+chan_level_2005,chen_using_2002,cremers_kernel_2006,gastaud_combining_2004}.
 These methods generally have a explicit description of the expected relative boundary 
@@
 pohl_bayesian_2006,wang_joint_2006}, where the prior 
-imposes consistent voxel-based classification of contiguous regions. Here, the 
-presence of more structures than one unique \gls{roi} helps aligning the target image 
-with the atlas in a hierarchical fashion. Finally, the third group generalizes 
-the atlas to actual images, and the contour is to segment simultaneously two 
-different target images, related by a spatial transform to be co-estimated
+imposes consistent voxel-based classification of contiguous regions.
+A comprehensive summary of the \gls{acwe}-derived methodologies with special attention
+to joint registration-segmentation methods is found in \citep{gorthi_active_2011}.
+
+Our joint registration and segmentation method is mainly inspired in the active
+deformation fields presented by \citep{gorthi_active_2011}, that is a generalization
+of the methodologies presented by \citep{bertalmio_morphing_2000,vemuri_joint_2003,
+yezzi_variational_2003}.
+All these formulations are deduced from the general level set evolution equation
+introduced by \citep{osher_fronts_1988}:
+\begin{equation}
+\frac{\partial \Phi_D(x,t)}{\partial t} = \nu ( \Phi_D(x,t)) \, \left| \nabla \Phi_D(x,t) \right|
+\end{equation}
+where $\nu$ is the velocity of the flow or speed function that contains the local
+segmentation and contour regularization constraints, and $\Phi_D: \Omega \to \mathbb{R}$
+is the signed distance function often used to represent implicitly the active contour
+by its zero level. An early extension to image registration was proposed by 
+\citep{vemuri_joint_2003} but replacing $\Phi_D$ by the intensity function of the
+image to register, $\Phi_I: \Omega \to \mathbb{R}$ (the moving image). 
+\citep{bertalmio_morphing_2000,vemuri_joint_2003} then define the speed function as
+$\nu(\Phi_I(x,t)) = \Phi_I(x,t) - \Phi_T(x,t)$, where $\Phi_T(x)$ is the intensity
+function of the target image. Finally, \citep{gorthi_active_2011} propose a different
+multiphase label function as level set function ($\Phi_L$) that improves registration
+results.
+
+Generalizing the objective function $\Phi_G$, it is possible to introduce the 
+dense deformation field of vectors $u: \mathbb{R}^n \to \mathbb{R}^n$ (tipically
+$n = \{ 2, 3 \}$) assuming the conservation of the morphological description, that is,
+$\Phi_G(x,t) = \Phi_G( x + du, t + dt ) \implies d\Phi_G(x,t) = 0$, where $d\Phi_G$ is
+the total derivative of $\Phi_G$. Using the chain rule, this can be rewritten as the 
+evolution equation of a vector flow:
+\begin{equation}
+\frac{\partial u(x,t)}{\partial t}= - \frac{\Phi_G}{\left| \nabla \Phi_G \right|} N_{\Phi_G},
+\end{equation}
+where $N_{\Phi_G}$ is the normal of the level set. Finally, they introduce the general
+level set evolution equation into the vector flow to perform the registration task:
+\begin{equation}
+\frac{\partial u(x,t)}{\partial t} = - \nu( \Phi_G(x,t) ) N_{\Phi_G}.
+\end{equation}
+Therefore, the position of the level set function $\Phi_G$ at time $t$ is given by the
+deformation field $u(x,t)$ and the initial level set function $\Phi_G(x,0)$, yielding:
+\begin{equation}
+\frac{\partial u(x,t)}{\partial t} = - \nu( \Phi_G(x + u(x,t), 0) ) N_{\Phi_G}.
+\end{equation}
+
+Here, we propose an active deformation field framework with a slightly different
+formulation. The main differences with respect the approach proposed in
+\citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
+$\Phi_G$, replacing the level set gradient computation $N_{\Phi_G}$ with shape
+gradients \citep{jehan-besson_dream2s:_2003,herbulot_segmentation_2006}; 2) 
+\todo[inline]{regularization is also based on linear diffusion smoothing REF-thirion,
+but they use a Gaussian filtering)}; 3) optimization is applied in the spectral
+domain, observing anisotropic and inhomogeneous mappings along each direction.
 
@@
 \subsection{Shape gradients}
-To compute the gradient-descent of the data-term domain integrals with respect to the underlying deformation field, we want to make use of shape gradients \cite{Jehan-Besson2003,herbulot_segmentation_2006}.
+To compute the gradient-descent of the data-term domain integrals with respect to the underlying deformation field, we want to make use of shape gradients \cite{jehan-besson_dream2s:_2003,herbulot_segmentation_2006}.
 A little bit of theory is therefore in order.

### 2013-06-28 Publications/2013-TMI/03-introduction.tex
Refined introduction (#81)
With applying @meribach's suggestions, I'll will be able to close
the issue.

--- 2013-05-29 Publications/2013-TMI/03-introduction.tex
@@
 has accounted for a growing interest in its application to structural 
-connectivity analysis of the brain. This emerging field exploits
+connectivity analyses of the brain at the macro-scale
+\citep{craddock_imaging_2013}. This emerging field exploits
 \gls{dwi} data to derive the local axonal structure at each imaging voxel 
 \citep{basser_microstructural_2011} and estimate a whole-brain mapping of fiber 
-tracts represented by trajectories reconstructed from the local information.
+tracts \citep{hagmann_mr_2012,johansen-berg_using_2009} represented by 
+trajectories reconstructed from the local information.
 This comprehensive map of neural connections of the brain is called the 
 \emph{connectome} \citep{hagmann_diffusion_2005,sporns_human_2005}. The connectome
-analysis is a promising tool for neuroscience and clinical applications.
-\todo[inline]{why? or references?}
+analysis is a promising tool for neuroscience \citep{morgan_why_2013} and 
+for the clinical application \citep{griffa_structural_2013} as well.
 
-Early \gls{dwi} research focused mainly on the improvements of imaging 
-methodologies better understanding the diffusion effect and improving
+Early \gls{dwi} research mainly focused on the improvements of imaging 
+methodologies by a better understanding of the diffusion effect, and improving
 image reconstruction methodologies. Currently, the connectome extraction 
-and analysis relies on a large amount of sophisticated computational techniques
-\citep{daducci_connectome_2012,hagmann_mr_2012} including acquisition,
-reconstruction, modeling and model fitting, image processing, fibre tracking,
-connectivity mapping, visualization, group studies, and inference. This 
-growing complexity has given rise to challenging issues towards reliable 
-structural information about the neuronal tracts \cite{johansen-berg_using_2009,
-jones_white_2012,soares_hitchhikers_2013}, and statistical analysis 
-\citep{meskaldji_comparing_????}. Here,we shall address three tasks included within 
-the image processing stage in a unified approach: brain tissue segmentation in 
-diffusion space (\autoref{sec:dwi_segmentation}), correction of geometrical 
-distortions (\autoref{sec:distortion}), and structural image registration to 
-diffusion coordinate space (\autoref{sec:registration}). These tasks are generally
+and analysis relies on a large chain of sophisticated computational techniques
+including acquisition, reconstruction, modeling and model fitting, image processing, 
+fibre tracking, connectivity mapping, visualization, group studies, and inference
+\citep{daducci_connectome_2012}. This growing complexity has given rise to 
+challenging issues towards reliable structural information about the neuronal tracts 
+presented in, and also difficulties on the statistical analysis 
+\citep{bullmore_complex_2009,meskaldji_comparing_2013}. To our knowledge, studies 
+regarding the impact of the different sources of inaccuracy and variability along
+the toolchain for connectivity analysis \citep{jones_twenty-five_2010,
+cheng_characteristics_2012,jones_white_2013} are scarce.
+Here,we shall address three tasks included within  the image processing stage in 
+a unified approach: brain tissue segmentation in diffusion space 
+(\autoref{sec:dwi_segmentation}), 
+correction of geometrical distortions (\autoref{sec:distortion}), and structural 
+image registration to diffusion coordinate space (\autoref{sec:registration}). 
+These tasks are generally
 solved independently, or combined in pairs. However, there exist fundamental 
@@
 
-\subsection{\gls{dwi} data overview}
+\subsection{\Gls{dwi} data overview}
 \label{sec:dwi_overview}
+\gls{dwi} data are usually acquired with \gls{epi} sequences 
+as they allow for very fast acquisition of large sets of slices each
+acquired after a single excitation \citep{mukherjee_diffusion_2008,
+soares_hitchhikers_2013}. \Gls{epi} volumes are collected
+in sequence and each one represents one probed diffusion direction.
+This first acquisition is called \emph{raw} data or \emph{diffusion
+signal}. One or more volumes without directional encoding (so-called \emph{baseline},
+$b = 0$, \emph{low $b$}, or just \emph{B0} volumes) are interleaved as
+reference, at the least one is acquired at the beginning of the
+\gls{epi} dataset. These B0 are acquired without direction gradient,
+showing a T2-like contrast.
 
-\todo[inline]{VERY brief look into signal generation HERE and cite reconstruction methods.}
-
-\gls{dwi} data are usually acquired with \gls{epi} sequences 
-as they allow for very fast acquisitions, but they are known to 
-suffer from geometrical distortions and artifacts due to, mainly,
-three sources: the subject motion in between acquiring different
-sampling directions, the induced \emph{Eddy currents} on the scanner 
+\Gls{epi} are known to suffer from geometrical distortions and artifacts
+due to, mainly, three sources: the subject motion in between acquiring 
+different sampling directions, the induced \emph{Eddy currents} on the scanner 
 coils and finally distortions caused by the magnetic susceptibility inhomogeneity
-present at tissue interface. In this paper, we restrict ourselves to
-the last one, as it accounts for the major impact in the connectome
-analysis. \emph{Susceptibility distortions} happen along the 
-phase-encoding direction, and are most appreciable in the front part of 
-the brain for the strong air/tissue interface surrounding the frontal sinuses.
+present at tissue interface. \emph{Susceptibility distortions} happen along the 
+phase-encoding direction, and are most appreciable in the frontal part of 
+the brain due to the air/tissue interface located around the sinuses.
 Two implications are associated to this artifact: the signal loss caused on
@@
 
-\gls{dwi} data is strongly affected by \glspl{pve} \citep{alexander_analysis_2001},
-which appear when several different tissues, or signal emitters, are present
-in the same imaging unit, producing an averaged intensity. The effect is directly
-related to the low resolution achievable with \gls{dwi} (typically around 
-$2.0\times2.0\times2.0mm^3$). An additional complication specific to
+After correcting for artifacts, the difficulty of performing analyses directly
+on the \emph{directionally dependent} raw data is not affordable. Thus, 
+\gls{dwi} data is reconstructed using one of a numerous set of readily 
+available algorithms. The outcome is a voxel-wise model of diffusion shape 
+(\todo[inline]{cite here dti, dsi, hardi, etc.and refs}). On top of the
+chosen model, some scalar maps describing tissue features is
+computed. For instance, the most popular \gls{dwi}-derived scalar map 
+is \gls{fa} which depicts the isotropy of water diffusion inside the brain.
+Another widely studied map is the \gls{adc} (or \gls{md} that has a dual definition),
+which reflects the intensity of the diffusion effect. Many other
+directionally invariant measurements have been defined. For a comprehensive study, 
+we refer to \citep{ennis_orthogonal_2006}.
+
+Alongside distortions and direction dependency, \gls{dwi} data is strongly 
+affected by \glspl{pve} \citep{alexander_analysis_2001} that appears when
+several different tissues, or signal 
+emitters, are present in the same imaging unit, producing an averaged intensity. 
+The effect is directly related to the low resolution achievable with \gls{dwi} 
+(typically around $2.0\times2.0\times2.0mm^3$). An additional complication specific to
 \gls{dwi} is the \gls{csf} \emph{contamination} \citep{metzler-baddeley_how_2012},
@@
 respect \gls{csf}.
-\todo[inline]{Define \textit{raw data}}
 
-Generally speaking, alongside the difficulty posed by the low resolution, \gls{dwi} 
-processing is also challenging due to the \emph{direction dependency} of raw data.
-\todo[inline]{Introduce here what are b0, fa, md, and direction dependency problem.}
-These $b = 0$ volumes (also called \emph{\gls{epi} baseline}, low-$b$, or just B0) 
-are acquired without direction gradient as reference, and they present a T2-like 
-contrast.
-
-\subsection{\gls{dwi} segmentation}
+\subsection{\Gls{dwi} segmentation}
 \label{sec:dwi_segmentation}
@@
 A precise delineation of the \gls{csf}, \gls{gm} and \gls{wm} interfacing surfaces
-is required with sub-pixel resolution.
-The resulting segmentation is necessary to perform the majority of tractography 
-algorithms and it is required to filter the resulting tractogram. The \gls{gm}-\gls{wm}
+is required with sub-pixel resolution, because whereas tractography algorithms are
+very good at estimating the location of bundles in deep \gls{wm}, they are not
+god (yet) at identifying where the tracts project into the \gls{gm} 
+\citep{jbabdi_tractography:_2011,craddock_imaging_2013}.
+Thus, a \gls{wm} segmentation is required by most of the available tractography 
+algorithms to \emph{filter} the resulting tractogram. The \gls{gm}-\gls{wm}
 interface is necessary to locate the starting and ending points of the detected
@@
 A number of methodologies have been proposed for \gls{dwi} segmentation, ranging 
-from intensity thresholding to atlas-based segmentation. The first approach is performed 
-on the \gls{fa} \citep{ennis_orthogonal_2006}, a well-known scalar map derived from
-\gls{dwi} data which depicts the isotropy of water diffusion inside the brain.
+from intensity thresholding to atlas-based segmentation. 
+The first approach is performed simply thresholding the \gls{fa} map.
 Although this methodology was popular among the premier tractography studies,
@@
 Segmentation obtained by co-registering structural T1-weighted images will be covered in 
-\autoref{sec:registration}. Some other works delay the segmentation task after the tractogram 
-is obtained, performing clustering on features derived on the tracts alignment 
-\citep{jonasson_white_2005}, combined tract registration \citep{mayer_supervised_2011} or using
-tractography atlases \citep{odonnell_automatic_2007}. However, methods based on tractography
-usually address the tractogram segmentation problem, to later combine the solution to 
-answer the whole-brain segmentation problem.
+\autoref{sec:registration}.
 
@@
 a \emph{gold-standard} evaluation methodology. Most of them are tested only on certain
-regions \todo{citations}, or do not provide sub-pixel resolution results
+regions, or do not provide sub-pixel resolution results
 \citep{hadjiprocopis_unbiased_2005,liu_brain_2007,awate_multivariate_2008,lu_segmentation_2008,
@@
 proposed for the segmentation validation \citep{jha_task-based_2012} on the task of lesion 
-detection in visceral organs.
+detection in visceral organs.  \todo[inline]{reference to a summary table}
 
@@
 extra B0 image with reversed phase encoding direction 
-\citep{holland_efficient_2010}.
+\citep{andersson_how_2003,holland_efficient_2010}.
 
-To our knowledge, there exists no study on 
-the impact of the susceptibility distortion over the subsequent tractography 
-and connectome analyses. However, a comparison of the diverse correction 
-techniques is found in \citep{wu_comparison_2008}. This
-study claimed that fieldmap correction methodologies are not entirely accurate
+A comparison of the diverse correction techniques is found in \citep{wu_comparison_2008}.
+This study claims that fieldmap correction methodologies are not entirely accurate
 and reliable, even though the method is correct in principle. This conclusion
@@
 a predefined parcellation in a high-resolution structural T1-weighted 
-atlas. The second goal that has justified the applicability of structural
+atlas (e.g. \citep{greve_accurate_2009}). The second goal that has 
+justified the applicability of structural
 images co-registration is susceptibility distortion correction
 as discussed in \autoref{sec:distortion}. In any case, susceptibility 
-distortion hinders a rigid-registration solution to this end.
+distortion hinders a rigid-registration solution to this problem.
 Additionally, when applying non-linear intensity-based
@@
 to be registered against the B0 in \gls{dwi}. B0 images have a very similar
-contrast to T2 weighted due to the duality of acquisition procedure. However,
-parcellation is defined in T1 weighted space, adding an additional registration
-step missing on the literature due to the novelty of the task. Even though this
-registration step has a very low complexity compared to the remaining tasks of 
-the pipeline process, it is one additional source of inconsistency and unreliability.
+contrast to T2 weighted due to the parallelism of acquisition sequences.
+However, parcellation is defined in T1 weighted space, adding an additional 
+registration step is generally overlooked by existing literature. Even though this
+processing stage has a very low complexity compared to the remaining tasks of 
+the pipeline process, it is one additional source of inconsistency
+and unreliability.
 
@@
 intensity of the undistorted data. They also use the $\log$ of the signal to
-enhance the low-signal regions and optimize with a gradient descent algorithm
-the mutual information of the mapping. The basis in the procedure is still the most 
-extended. For instance, \citep{wu_comparison_2008} also proposed a B-Spline registration
+enhance the low-signal regions, and optimize with a gradient descent algorithm
+the mutual information of the mapping. This framework set the basis for the following
+works. For instance, \citep{wu_comparison_2008} also proposed a B-Spline registration
 providing quantitative comparisons with fieldmap correction methods. Recent approaches
@@
 \label{sec:contributions}
-Therefore, the problems of precise segmentation in \gls{dwi}-space and the 
+
+All the available \gls{dwi}-derived models and scalar maps 
+are very sensitive to distortions, \gls{pve}, noise and axonal dispersion 
+\citep{jbabdi_tractography:_2011}, what yields numerous challenges in the
+processing tasks (artifact correction, segmentation, registration and 
+parcellation). Finally, as a result a vast number of challenges and pitfalls 
+emerge regarding accuracy (correctness) and precision (reproducibility) of 
+tractography-based analyses \citep{jones_white_2013} and connectome
+analyses.
+
+Moreover, the problems of precise segmentation in \gls{dwi}-space and the 
 spatial mapping between these contours and the corresponding surfaces in 
@@
 by exploiting as strong shape-prior the detailed morphology extracted 
-from high-resolution anatomically correct \gls{mri}. Indeed, hereafter 
+from high-resolution and anatomically correct \gls{mri}. Indeed, hereafter 
 we assume this segmentation problem in anatomical images is reliably and
-accurately solved with readily available tools. After global alignment 
-using existing approaches, the remaining spatial mismatch between 
-anatomical and diffusion space is due to susceptibility distortions.
+accurately solved with readily available tools (e.g. \citep{fischl_freesurfer_2012}).
+After global alignment with \gls{t1} using existing approaches, the remaining 
+spatial mismatch between anatomical and diffusion space is due to susceptibility
+distortions.
 Finally, we need to establish precise spatial correspondence between the 
@@
 distortion in between anatomical and \gls{dwi} space.
-We thus reformulate the segmentation problem as an inverse problem, where we 
+{\color{red} {We thus reformulate the segmentation problem as an inverse problem, where we 
 seek for an underlying deformation field (the distortion) mapping 
@@
 
-\todo[inline]{ (Rewrite) 
 We test our proposed joint segmentation-registration model on two different 
@@
 \autoref{sec:experiments}. We present results in \autoref{sec:results} and conclude 
-in \autoref{sec:conclusion}.}
+in \autoref{sec:conclusion}.}}

### 2013-06-28 Publications/2013-TMI/03-introduction.tex
Revised intro
Close #81. I also add the pdf containing @meribach comments on the
intro with the fixes.

@@
 \IEEEPARstart{D}{iffusion} \gls{mri} is a widely used family
-of \gls{mri} techniques \citep{sundgren_diffusion_2004} which recently 
-has accounted for a growing interest in its application to structural 
+of \gls{mri} techniques \citep{sundgren_diffusion_2004} which 
+accounts for a growing interest in its application to structural 
 connectivity analyses of the brain at the macro-scale
@@
 and analysis relies on a large chain of sophisticated computational techniques
-including acquisition, reconstruction, modeling and model fitting, image processing, 
-fibre tracking, connectivity mapping, visualization, group studies, and inference
-\citep{daducci_connectome_2012}. This growing complexity has given rise to 
-challenging issues towards reliable structural information about the neuronal tracts 
-presented in, and also difficulties on the statistical analysis 
+including acquisition, reconstruction, modeling and model fitting, image segmentation
+and registration, fibre tracking, connectivity mapping, visualization, statistical 
+group studies, and inference \citep{daducci_connectome_2012}. This growing complexity 
+has given rise to challenging issues towards reliable structural information about 
+the neuronal tracts presented in, and also difficulties on the statistical analysis 
 \citep{bullmore_complex_2009,meskaldji_comparing_2013}. To our knowledge, studies 
@@
 cheng_characteristics_2012,jones_white_2013} are scarce.
-Here,we shall address three tasks included within  the image processing stage in 
+Here, we shall address three tasks included within  the image processing stage in 
 a unified approach: brain tissue segmentation in diffusion space 
@@
 different sampling directions, the induced \emph{Eddy currents} on the scanner 
-coils and finally distortions caused by the magnetic susceptibility inhomogeneity
+coils, and, finally, distortions caused by the magnetic susceptibility inhomogeneity
 present at tissue interface. \emph{Susceptibility distortions} happen along the 
@@
 han_experimental_2009}. Generally, results obtained with high resolution atlas co-registration
-(\autoref{sec:registration}) are more compelling, minimizing the activity on this line which is 
-currently being considered to be included in reconstruction algorithms 
-\citep{kumazawa_improvement_2013}. \emph{Golden}-standard evaluation frameworks have been 
-proposed for the segmentation validation \citep{jha_task-based_2012} on the task of lesion 
-detection in visceral organs.  \todo[inline]{reference to a summary table}
+(\autoref{sec:registration}) are more compelling, limiting the activity on \gls{dwi} 
+segmentation. However, recently \citep{kumazawa_improvement_2013} proposed to include this
+task within reconstruction algorithms. Finally, some \emph{golden}-standard evaluation
+frameworks for task-based validation have been proposed for the segmentation results
+(e.g. lesion detection in visceral organs \citep{jha_task-based_2012}).
+\todo[inline]{reference to a summary table}
 
@@
 images co-registration is susceptibility distortion correction
-as discussed in \autoref{sec:distortion}. In any case, susceptibility 
+as presented in \autoref{sec:distortion}. In any case, susceptibility 
 distortion hinders a rigid-registration solution to this problem.
 Additionally, when applying non-linear intensity-based
-registration algorithms other difficulties have to be addressed as
+registration algorithms, other difficulties have to be addressed as
 the significant \gls{pve} in the \gls{wm}-\gls{gm} layer, or the

### 2013-07-01 Publications/2013-TMI/03-introduction.tex
Fixed some references

@@
 due to, mainly, three sources: the subject motion in between acquiring 
-different sampling directions, the induced \emph{Eddy currents} on the scanner 
-coils, and, finally, distortions caused by the magnetic susceptibility inhomogeneity
+different sampling directions, the induced \emph{eddy currents} on the scanner 
+coils due to the gradient switching \citep{obrien_3d_2013}, and, finally, 
+distortions caused by the magnetic susceptibility inhomogeneity
 present at tissue interface. \emph{Susceptibility distortions} happen along the 
@@
 available algorithms. The outcome is a voxel-wise model of diffusion shape 
-(\todo[inline]{cite here dti, dsi, hardi, etc.and refs}). On top of the
-chosen model, some scalar maps describing tissue features is
+(in \citep{daducci_isbi2013_2013} \todo{FIX ref} there is available a comprehensive review of 
+state-of-art methodologies). On top of the
+chosen model, some scalar maps describing tissue features is generally
 computed. For instance, the most popular \gls{dwi}-derived scalar map 
@@
 god (yet) at identifying where the tracts project into the \gls{gm} 
-\citep{jbabdi_tractography:_2011,craddock_imaging_2013}.
+(limitations on the termination criteria, \citep{jbabdi_tractography:_2011,
+craddock_imaging_2013}).
 Thus, a \gls{wm} segmentation is required by most of the available tractography 
@@
 The first approach is performed simply thresholding the \gls{fa} map.
-Although this methodology was popular among the premier tractography studies,
+Although this methodology was popular among the pioneer tractography studies,
 they were generally limited to certain regions or significant fiber tracts, and thus,
@@
 \citep{kybic_unwarping_2000,studholme_accurate_2000,li_accounting_2007,
-tao_variational_2009} (see \autoref{sec:registration}), or acquiring an 
+tao_variational_2009} (we shall cover these methods in 
+\autoref{sec:registration}), or acquiring an 
 extra B0 image with reversed phase encoding direction 
@@
 This study claims that fieldmap correction methodologies are not entirely accurate
-and reliable, even though the method is correct in principle. This conclusion
+and reliable, even though the method is correct in principle. The conclusion
 was later confirmed by \citep{tao_variational_2009}. Additional concerns regarding

### 2013-07-01 Publications/2013-TMI/04-methods.tex
Fixed some references

@@
 chan_level_2005,chen_using_2002,cremers_kernel_2006,gastaud_combining_2004}.
-These methods generally have a explicit description of the expected relative boundary 
+These methods generally have an explicit description of the expected relative boundary 
 locations of the object to be delineated, and some even model the statistical deviations
@@
 active contours based registration between timesteps in a time-series or between different
-images \citep{wyatt_map_2003,paragios_level_2003,vemuri_joint_2003,yezzi_variational_2003}.
-This second group is closely related to atlas-based segmentation methods 
+images \citep{bertalmio_morphing_2000,wyatt_map_2003,paragios_level_2003,vemuri_joint_2003,
+yezzi_variational_2003}.
+A summary of these second set of techniques is performed in \citep{droske_mumfordshah_2009},
+proposing two different approaches to applying the Mumford-Shah \citep{mumford_optimal_1989}
+functional in joint registration and segmentation. Finally, a derivation of the 
+latter groups is composed by atlas-based segmentation methods 
 \citep{gorthi_segmentation_2009,gorthi_active_2011,pohl_unifying_2005,
@@
 deformation fields presented by \citep{gorthi_active_2011}, that is a generalization
-of the methodologies presented by \citep{bertalmio_morphing_2000,vemuri_joint_2003,
-yezzi_variational_2003}.
-All these formulations are deduced from the general level set evolution equation
+of the methodologies under second and third groups before. As it is described in
+their work, the first joint ``morphing''-segmentation approach (it was not a proper
+registration method) was proposed in \citep{bertalmio_morphing_2000}. The first
+registration framework is presented by \citep{yezzi_variational_2001} where the
+energy functional is defined simultaneously in the moving and target images with
+an affine transformation supporting the coordinates mapping (``2 PDEs approach'').
+\citep{vemuri_joint_2003} proposed the first atlas-based registration based on the 
+level set framework and using only one PDE. \citep{unal_coupled_2005} extended the 
+idea of \citep{bertalmio_morphing_2000,yezzi_variational_2001} for non-linear 
+registration with a dense deformation field as mapping function. \citep{droske_mumfordshah_2009}
+combining ideas from both branches and proposed the propagation of the 
+deformation field from the contours to the whole image definition. Finally,
+\citep{gorthi_active_2011} present a comprehensive generalization of the methodologies.
+
+All the surveyed frameworks are deduced from the general level set evolution equation
 introduced by \citep{osher_fronts_1988}:
@@
 
-Generalizing the objective function $\Phi_G$, it is possible to introduce the 
-dense deformation field of vectors $u: \mathbb{R}^n \to \mathbb{R}^n$ (tipically
-$n = \{ 2, 3 \}$) assuming the conservation of the morphological description, that is,
+Naming $\Phi_G$ a generic objective function, a dense deformation field of vectors 
+$u: \mathbb{R}^n \to \mathbb{R}^n$ (tipically $n = \{ 2, 3 \}$) is introduced. 
+Then, the conservation of the morphological description is assumed, such as
 $\Phi_G(x,t) = \Phi_G( x + du, t + dt ) \implies d\Phi_G(x,t) = 0$, where $d\Phi_G$ is
-the total derivative of $\Phi_G$. Using the chain rule, this can be rewritten as the 
+the total derivative of $\Phi_G$. Using the chain rule, it can be rewritten as the 
 evolution equation of a vector flow:
@@
 
-Here, we propose an active deformation field framework with a slightly different
-formulation. The main differences with respect the approach proposed in
+Here, we propose an active deformation field framework with some differences in
+the formulation of the problem. The main diverging points with respect to
 \citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
-$\Phi_G$, replacing the level set gradient computation $N_{\Phi_G}$ with shape
+$\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
 gradients \citep{jehan-besson_dream2s:_2003,herbulot_segmentation_2006}; 2) 

### 2013-07-01 Publications/2013-TMI/04-methods.tex
Updating numerical implementation section
* Implemented #80, but still missing the consecuences of p(U) in the general framework
* Re-written methodology in a fashion a bit more consistent with the undetermined number of regions

@@
 \begin{equation}
-\frac{\partial \Phi_D(x,t)}{\partial t} = \nu ( \Phi_D(x,t)) \, \left| \nabla \Phi_D(x,t) \right|
+\frac{\partial \Phi_D(\mathbf{x},t)}{\partial t} = \nu ( \Phi_D(\mathbf{x},t)) \, \left| \nabla \Phi_D(\mathbf{x},t) \right|
 \end{equation}
@@
 \citep{bertalmio_morphing_2000,vemuri_joint_2003} then define the speed function as
-$\nu(\Phi_I(x,t)) = \Phi_I(x,t) - \Phi_T(x,t)$, where $\Phi_T(x)$ is the intensity
+$\nu(\Phi_I(\mathbf{x},t)) = \Phi_I(\mathbf{x},t) - \Phi_T(\mathbf{x},t)$, where $\Phi_T(\mathbf{x})$ is the intensity
 function of the target image. Finally, \citep{gorthi_active_2011} propose a different
@@
 Then, the conservation of the morphological description is assumed, such as
-$\Phi_G(x,t) = \Phi_G( x + du, t + dt ) \implies d\Phi_G(x,t) = 0$, where $d\Phi_G$ is
+$\Phi_G(\mathbf{x},t) = \Phi_G( \mathbf{x} + du, t + dt ) \implies d\Phi_G(\mathbf{x},t) = 0$, where $d\Phi_G$ is
 the total derivative of $\Phi_G$. Using the chain rule, it can be rewritten as the 
@@
 \begin{equation}
-\frac{\partial u(x,t)}{\partial t}= - \frac{\Phi_G}{\left| \nabla \Phi_G \right|} N_{\Phi_G},
+\frac{\partial u(\mathbf{x},t)}{\partial t}= - \frac{\Phi_G}{\left| \nabla \Phi_G \right|} N_{\Phi_G},
 \end{equation}
@@
 \begin{equation}
-\frac{\partial u(x,t)}{\partial t} = - \nu( \Phi_G(x,t) ) N_{\Phi_G}.
+\frac{\partial u(\mathbf{x},t)}{\partial t} = - \nu( \Phi_G(\mathbf{x},t) ) N_{\Phi_G}.
 \end{equation}
 Therefore, the position of the level set function $\Phi_G$ at time $t$ is given by the
-deformation field $u(x,t)$ and the initial level set function $\Phi_G(x,0)$, yielding:
+deformation field $u(\mathbf{x},t)$ and the initial level set function $\Phi_G(\mathbf{x},0)$, yielding:
 \begin{equation}
-\frac{\partial u(x,t)}{\partial t} = - \nu( \Phi_G(x + u(x,t), 0) ) N_{\Phi_G}.
+\frac{\partial u(\mathbf{x},t)}{\partial t} = - \nu( \Phi_G(\mathbf{x} + u(\mathbf{x},t), 0) ) N_{\Phi_G}.
 \end{equation}
@@
 
-\subsection{From the Maximum A-Posteriori Model to the Mumford-Shah functional}
+\subsection{Bridging segmentation frameworks}
 \label{sec:methods_map}
@@
 A widely used approach to image segmentation is derived from the
-Bayes' rule \eqref{eq:bayes_rule}, where one seeks for a partitioning
-of a certain image $X$ in piecewise smooth regions $\Omega = \lbrace \Omega_k , 
-k\in\left[ 1 .. K \right] \rbrace$,  that maximizes the a posteriori 
-probability given the multivariate image $X \in \mathbb{R}^C$, 
-with $C$ being the number of image channels.
+Bayes' rule \eqref{eq:bayes_rule}. This framework seeks for a 
+partitioning of a certain observation of features $F: \mathbb{R}^n \to %
+\mathbb{R}^C$ (an $n$-dimensional \emph{image} comprising $C$ different 
+channels) in piecewise smooth regions $\Omega = \lbrace \Omega_k , 
+k\in\left[ 1 \ldots K \right] \rbrace$,  that maximize the \emph{a posteriori}
+probability $p(\Omega \mid F)$:
 \begin{equation}
-p(Y \mid X) = \frac{p(X \mid Y)\, p(Y)}{p(X)}.
+p(\Omega \mid F) = \frac{p(F \mid \Omega)\, p(\Omega)}{p(F)},
 \label{eq:bayes_rule}
 \end{equation}
-
-Therefore, $Y$ is a certain realization of the piecewise 
-disjoint region set $\Omega$. $p(X \mid Y)$ is the \emph{likelihood} of 
-the realization of $X$ (the image) given a certain distribution model for 
-each region $\Omega_k$. The second term, $p(Y)$, is the a-priori probability of 
-the partitioning $Y$. Finally, $p(X)$ is the probability of a certain image 
-realization, and thus, it will remain constant when computing the \gls{map}.
-Consequently, $p(Y \mid X) \propto p(X \mid Y)\, p(Y)$, and:
+where $p(F \mid \Omega)$ is the \emph{likelihood} of the realization 
+of $F$ given a certain region set $\Omega$. The second term, $p(\Omega)$,
+is the a-priori probability of the partitioning. Finally, $p(F)$ is the 
+probability of a certain image realization, and thus, it will remain 
+constant when computing the \gls{map}. Consequently, the Bayes' rule
+can be simplified for the optimization as follows:
 \begin{equation}
-\underset{Y}{\argmax} \left\{ p(Y \mid X) \right\} = 
-\underset{Y}{\argmax} \left\{ p(X \mid Y)\, p(Y) \right\}.
+p(\Omega \mid F) \propto p(F \mid \Omega)\, p(\Omega).
+\label{eq:bayes_rule_simplified}
 \end{equation}
 
-An extended assumption is that the feature vector realization $X$ is
+In the simplest Bayesian segmentation framework, the unknown is the
+parameter set $\Theta_k$ of region-wise descriptors that minimize 
+\eqref{eq:bayes_rule}. Conversely, we can assume these descriptors as
+fixed (fulfilling the concept of conservation of morphological description
+mentioned before) and introduce a realization of a dense deformation field $U: %
+\mathbb{R}^n \to \mathbb{R}^n$ that maps the initial partition $\Omega_0$
+to the estimated one:
+\begin{equation}
+\Omega' = \Omega_0 + \hat{U}.
+\label{eq:omega_tf}
+\end{equation}
+
+Thus, \eqref{eq:bayes_rule_simplified} is rewritten as 
+$p(U \mid F) \propto p(F \mid U)\, p(U)$ that yields the
+following \gls{map} solution:
+
+\begin{equation}
+\hat{U} = \underset{U}{\argmax} \left\{ p(U \mid F) \right\} = 
+\underset{U}{\argmax} \left\{ p(F \mid U)\, p(U) \right\}.
+\label{eq:map_u}
+\end{equation}
+
+An extended assumption is that the feature vector realization $F$ is
 \emph{i.i.d.}, and thus, it is possible to write the a-posteriori
-probability $p(X \mid Y)$ as a continuous product with $d\mathbf{x}$ the
+probability $p(F \mid U)$ as a continuous product with $d\mathbf{x}$ the
 infinitesimal voxel size:
 \begin{equation}
-p(X \mid Y) \, p(Y) = \underset{k}{\prod} \underset{\mathbf{x}\in \Omega_k}{\prod}
-p_k( X(\mathbf{x}) \mid Y(\mathbf{x}) )^{d\mathbf{x}},
+p(F \mid U) \, p(U) = \underset{k}{\prod} \underset{\mathbf{x}\in \Omega'_k}{\prod}
+p_k\left( F(\mathbf{x} + U(\mathbf{x})) \mid U(\mathbf{x}) \right)^{d\mathbf{x}}.
 \label{eq:bayes_aposteriori}
 \end{equation}
-where the prior probability $p(Y)$ is implicitly
-defined by the regions definition.
+\todo[inline]{I'm not convinced about turning $p(\Omega)$ into $p(U)$.}
 
@@
 distribution of the different tissues in \gls{mri} data. Therefore,
-the posterior probability of an infinitesimal voxel
-can be written as:
+the posterior probability of an infinitesimal voxel can be written as:
 \begin{equation}
-p_k( X(\mathbf{x}) \mid Y(\mathbf{x}) ) = \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{X(\mathbf{x})}) \right)}}.
+p_k( F(\mathbf{x}) \mid U(\mathbf{x}) ) = \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{f}') \right)}}.
 \label{eq:bayes_mpdf}
 \end{equation}
-where we can identify the factor in the exponential, with 
-$\mathbf{f} = X(\mathbf{x})$ as the squared \emph{Mahalanobis 
-distance} with the parameters set $\Theta_k = \lbrace \boldsymbol{\mu}_k, \boldsymbol{\Sigma}_k \rbrace$:
+where we can identify the factor in the exponential as the squared \emph{Mahalanobis 
+distance} of the feature observed in the displaced position, $\mathbf{f}' = 
+F(\mathbf{x} + U(\mathbf{x}))$ with region descriptors
+$\Theta_k = \lbrace \boldsymbol{\mu}_k, \boldsymbol{\Sigma}_k \rbrace$:
 \begin{equation}
@@
 \begin{multline}
-E(X \mid Y)= -\log \left[ p(X \mid Y) \, p(Y) \right] = \\
-= -\log \left[ \underset{k}{\prod} \underset{\mathbf{x}\in \Omega_k}{\prod}
-p_k( X(\mathbf{x}) \mid Y(\mathbf{x}) )^{d\mathbf{x}} \right] = \\
-= \sum\limits_k \int_{\Omega_k} -\log \left[ p_k(X(\mathbf{x}) \mid Y(\mathbf{x} ) ) \right] \, d\mathbf{x},
+E(F \mid U)= -\log \left[ p(F \mid U) \, p(U) \right] = \\
+= -\log \left[ \underset{k}{\prod} \underset{\mathbf{x}\in \Omega'_k}{\prod}
+p_k( F(\mathbf{x}) \mid U(\mathbf{x}) )^{d\mathbf{x}} \right] = \\
+= \sum\limits_k \int_{\Omega'_k} -\log \left[ p_k(F(\mathbf{x}) \mid U(\mathbf{x} ) ) \right] \, d\mathbf{x},
 \label{eq:energy_1}
@@
 and introducing the posterior probability term \eqref{eq:bayes_mpdf}, 
-we can express the functional in terms of $\lbrace\Theta_k,\Omega_k\rbrace$:
+we can express the functional in terms of the deformation field $U$ and
+the region descriptors $\Theta_k$:
 \begin{multline}
-E(\Theta_k,\Omega_k) = \\
-= \sum\limits_k \int_{\Omega_k} -\log \left[ \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{f}) \right)}} \right] \, d\mathbf{x} = \\
-= \sum\limits_k \int_{\Omega_k} \left[ \frac{1}{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right| \right)} + \frac{1}{2}  \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}.
+E(U) = \\
+= \sum\limits_k \int_{\Omega'_k} -\log \left[ \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{f}) \right)}} \right] \, d\mathbf{x} = \\
+= \sum\limits_k \int_{\Omega'_k} \left[ \frac{1}{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right| \right)} + \frac{1}{2}  \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}.
 \end{multline}
@@
 \begin{align}
-E(\Theta_k,\Omega_k) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x},
+E(U) = \sum\limits_k \left[ \left|\Omega'_k\right|\,\log \left|\mathbf{\Sigma}_k \right| + \int_{\Omega'_k} \Delta^2_k (\mathbf{f}) \,d\mathbf{x} \right],
 \label{eq:map_energy}
 \end{align}
-\todo[inline]{improve this explanation} where we have a constant term scaled by
-the total volume of the partition $\Omega_k$ and the determinant of the 
-covariance matrix of the partition $\left|\boldsymbol{\Sigma}_{k}\right|$,
+where we have a constant term scaled by the total volume of the partition $\Omega_k$ and 
+the determinant of the  covariance matrix of the partition $\left|\boldsymbol{\Sigma}_{k}\right|$,
 plus an energy term based on the squared \emph{Mahalanobis distance}.
+As long as this shape-based registration framework is designed to allow small changes
+between boundaries (and therefore, their volumes), and the shape descriptors should not
+change significantly due to the conservation principle, the first term can be optionally
+dismissed with respect the second one, what yields:
+\begin{align}
+E(U) = \sum\limits_k \int_{\Omega'_k} \Delta^2_k (\mathbf{f}) \,d\mathbf{x}.
+\label{eq:final_map_energy}
+\end{align}
 
-Equation \eqref{eq:map_energy} resembles the Mumford-Shah functional including variance,
-that modifies the original functional in a way that it can deal with more general
-distributions. This is necessary to avoid the assumption that regions $\Omega_k$
-have a fixed covariance matrix on their complete domain. One immediate advantage
-of this functional from the original one is the possibility to distinguish regions
-with the same mean vector but different covariance matrix \citep{brox_local_2009}:
+This latter expression resembles the Mumford-Shah functional 
+\citep{mumford_optimal_1989} which is at the background of all \gls{acwe}-based
+methods. In this case, \eqref{eq:final_map_energy} includes covariance as
+region descriptor, what modifies the original functional in a way that it 
+can deal with more general distributions. This is necessary to avoid the 
+assumption that regions $\Omega_k$ have a fixed covariance matrix on their 
+complete domain. One immediate advantage of this functional from the original 
+one is the possibility to distinguish regions with the same mean vector but 
+different covariance matrix \citep{brox_local_2009}. In this work, the 
+Mumford-Shah functional is derived for this extension as follows:
 \begin{multline}
@@
 
-\subsection{Deformation model}
-\label{sec:deformation}
+\subsection{Introducing priors}
+\label{sec:priors}
 %
-The segmentation problem is transformed into a registration one if the initial
-partition $Y$ is derived from the initial shape-priors in reference space. 
-Introducing a dense deformation field $u$ that maps the original partition to
-a better fit of the regions in the target coordinate space. Thus, the minimization
-problem stated in \eqref{eq:map_energy} can be expressed so that $u$ is the 
-unknown:
+By introducing $U$ in \eqref{eq:bayes_rule_simplified}, we transformed 
+the segmentation problem into a registration one provided with $\Omega_0$ 
+(the initial partition) is derived from the shape-priors given in reference 
+space. Therefore, the priors can be expressed in terms of the deformation
+field:
 \begin{equation}
-E(u(\mathbf{x})) = \sum\limits_k \int_{\Omega'_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f'}) \right] \,d\mathbf{x}
-\label{eq:map_energy_deformation}
+P(U) = \underset{\mathbf{x}}{\prod}\,p(U(\mathbf{x}))\,^{d\mathbf{x}}.
+\label{eq:prior_u}
 \end{equation}
-where $\Omega'_k = u(\Omega_k)$ and $\mathbf{f'} = X(u(\mathbf{x}))$.
+We define $P(U(\mathbf{x})) = p(\mathbf{u}) = \underset{d}\prod \, p_d(\mathbf{u})$ 
+with $d$ the order of the derivative. We also assume that an initial
+affine registration let us consider that the deformation field ($d=0$)
+and its derivatives up to some order ($d>0$) are all zero mean, and
+have some variance $\Sigma_d$. We restrict ourselves to $d=\lbrace 0,1 \rbrace$
+(the field and its gradient):
+\begin{subequations}
+\begin{equation}
+p_0(\mathbf{u}) = \mathcal{N}( \mathbf{u} \mid 0, \alpha^{-1}),
+\end{equation}
+\begin{equation}
+p_1(\mathbf{u}) = \mathcal{N}( \nabla \mathbf{u} \mid 0, {\Sigma_\beta}^{-1}), 
+\end{equation}
+\end{subequations}
+And therefore:
+\begin{equation}
+P(U) = \underset{\mathbf{x}}\prod \left[ p_0(\mathbf{u}) \, p_1(\mathbf{u}) \right]^{d\mathbf{x}}.
+\end{equation}
 
-\subsection{\Acrlong{acwe} based segmentation model}
+\subsection{Numerical Implementation}
 %
-Let us denote $\{c_i\}_{i=1..N_c}$ the nodes of a shape prior surface. In
-our application, a precise \gls{wm}-\gls{gm} interface extracted from a
-high-resolution reference volume. All the formulations can be naturally
-extended to include more shape priors. On the other hand, we have a 
-number of \gls{dwi}-derived features at each
-voxel of the volume. Let us denote by $x$ the voxel and 
-$f(x) = [ f_1, f_2, \ldots, f_N]^T(x)$ its associated feature vector.\\
-%
+Let us denote $\{c_i\}_{i=1 \ldots N_c}$ the nodes of one or several shape-prior
+surface(s). In our application, precise tissue interfaces of interest 
+extracted from a high-resolution, anatomically correct reference volume. 
+On the other hand, we have a number of \gls{dwi}-derived features at each
+voxel of the volume. Let us denote by $\mathbf{x}$ the voxel and $f(\mathbf{x}) = 
+[ f_1, f_2, \ldots, f_N]^T(\mathbf{x})$ its associated feature vector.
+
+\subsubsection{Deformation field}
+\label{sec:deformation_field}
 The transformation from reference into \gls{dwi} coordinate space is 
-achieved through a dense deformation field $u(x)$, such that:
+achieved through a dense deformation field $u(\mathbf{x})$, such that:
+\begin{equation}
+c_i' = T\{c_i\} = c_i + u(c_i),
+\end{equation}
+what is equivalent to \eqref{eq:omega_tf}. Since the nodes of the anatomical 
+surfaces might lay off-grid, it is required to derive $u(\mathbf{x})$ from a discrete 
+set of parameters $\{u_k\}_{k=1 \ldots K}$. Densification is achieved through 
+a set of associated basis functions $\psi_k$ (e.g. rbf, interpolation splines):
 %
 \begin{equation}
-c_i' = T\{c_i\} = c_i + u(c_i)
+u(\mathbf{x}) = \sum_k \psi_k(\mathbf{x}) u_k
 \end{equation}
 %
-Since the nodes of the anatomical surfaces might lay off-grid, it is 
-required to derive $u(x)$ from a discrete set of parameters $\{u_k\}_{k=1..K}$.
-Densification is achieved through a set of associated basis functions 
-$\psi_k$ (e.g. rbf, interpolation splines):
-%
-\begin{equation}
-u(x) = \sum_k \psi_k(x) u_k
-\end{equation}
-%
-Consequently, the transformation writes
+In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel.
+Then, the transformation writes
 %
@@
 \end{equation}
-%
-% Comment: maybe this is not for IPMI 2013.
-%Note that, since $c_i$ remains constant in the DW segmentation process,
-%the values of $\psi_k(c_i)$ can be precomputed. Also, provided compact 
-%support of the basis functions, the system remains relatively sparse.\\
-%
+
 Based on the current estimate of the distortion $u$, we can compute 
@@
 Thus, we now estimate region descriptors of the \gls{dwi} features 
-$f(x)$ of the regions defined by the priors in \gls{dwi} space.
+$f(\mathbf{x})$ of the regions defined by the priors in \gls{dwi} space.
 %
-Using Gaussian distributions as region descriptors, we propose an
-\gls{acwe}-like, piece-wise constant, variational image segmentation
+Using the region descriptors derived in \autoref{sec:methods_map}, we propose
+an \gls{acwe}-like, piece-wise constant, variational image segmentation
 model (where the unknown is the deformation field)
-\cite{chan_active_2001}:
-
-where $R$ indexes the existing regions and the integral domains
-depend on the deformation field $u$. Note
-that minimizing this energy, $\argmin_u\{E\}$, yields the \gls{map} 
-estimate of a piece-wise smooth image model affected by Gaussian 
-additive noise. This inverse problem is ill-posed
+\cite{chan_active_2001} with the energy functional obtained in 
+\eqref{eq:final_map_energy}. This inverse problem is ill-posed
 \cite{bertero_ill-posed_1988,hadamard_sur_1902}.
@@
 %
-\begin{align}
+\begin{multline}
+E(u) = \sum\limits_k \int_{\Omega'_k} \Delta^2_k (\mathbf{f}) \,d\mathbf{x} \\
++ \int u^T \, A \, u \, d\mathbf{x} + \int \tr\{(\nabla u^T)^T B (\nabla u^T)\} d\mathbf{x}
 \label{eq:complete_energy}
-E(u) &= \sum_{\forall{R}} \lbrace \int_{\Omega_R} (f-\mu_R)^T\Sigma_R^{-1}(f-\mu_R) dx \rbrace \nonumber \\
-&\quad + \alpha \int  \|u\|^2 dx + \beta \int \left( \|\nabla u_x\|^2 + \|\nabla u_y\|^2 + \|\nabla u_z\|^2\right) dx
-\end{align}
+\end{multline}
 %
@@
 
-At each iteration, we update the distortion along the steepest 
-energy descent. This gradient descent step can be efficiently 
-tackled by discretizing the time in a forward Euler scheme, 
-and making the right hand side semi-implicit in the 
-regularization terms:
-%
-\begin{align}
-\frac{u^{t+1}-u^t}{\tau} &= - \sum_{i=1}^{N_c} \left[ e(f(c_i'))  \hat{n}_{c_i'} \psi_{c_i}(x) \right] -\alpha u^{t+1} + \beta\Delta u^{t+1}
-\end{align}
-%
-where the data terms remain functions of the current estimate 
-$u^t$, thus $c_i' = c_i'(u^t)$. For simplicity on notation, we 
-restricted the number of priors to only 1. We also defined 
-$e(f(c_i')) = E_{out}(f(c_i')) - E_{in}(f(c_i'))$, 
-and $E_R(f) = {(f-\mu_R)^T\Sigma_R^{-1}(f-\mu_R)}$.
-We applied a spectral approach to solve this implicit scheme:
-%
-\begin{equation}
-u^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{u^t/\tau
-- \sum_{i=1}^{N_c} \left[e(f(c_i')) \hat{n}_{c_i'} \psi_{c_i}(x) \right]  \}}{\mathcal{F}\{(1/\tau+\alpha)\mathcal{I}-\beta\Delta\}} \right\}
-\end{equation}
-%
-
-\subsection{Operator splitting}
-In order to make the Energy minimisation computationally more tractable, we propose the following operator splitting: Let us optimize the data terms and the regularity terms on separate copies of the deformation field, now called $u$ and $v$, constrained to be equal:
-\begin{align}
-E(u,v) &= \int_{w'(u)} (f-\mu_w)^T\Sigma_w^{-1}(f-\mu_w) dx\nonumber\\
-&\quad +\int_{g'(u)} (f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g) dx\\
-&\quad +\int_{o'(u)} (f-\mu_o)^T\Sigma_o^{-1}(f-\mu_o) dx\nonumber\\
-&\quad +\int v^T A v dx + \int \tr\{(\nabla v^T)^T B (\nabla v^T)\} dx\nonumber
-\end{align}
+\subsubsection{Operator splitting}
+In order to make the energy minimization computationally more tractable, 
+we propose the following operator splitting: Let us optimize the data terms 
+and the regularity terms on separate copies of the deformation field, 
+now called $u$ and $v$, constrained to be equal:
+\begin{multline}
+E(u,v) = \sum\limits_k \int_{\Omega'_k(u)} \Delta^2_k (\mathbf{f}) \,d\mathbf{x} \\
++ \int v^T \, A \, v \, d\mathbf{x} + \int \tr\{(\nabla v^T)^T B (\nabla v^T)\} d\mathbf{x}
+\label{eq:operator_splitting}
+\end{multline}
 and now
 \begin{equation}
-\min_{u,v} \{ E \} \quad s.t. \quad u = v
+\min_{u,v} \{ E \} \quad s.t. \quad u = v.
 \end{equation}
-In order to take the equality constraint into account, we may make use of augmented Lagrangians (a combination of Lagrangian multipliers and penalty terms on the constraint)
-\cite{bertsekas_multiplier_1976,glowinski_augmented_1989,nocedal_numerical_2006}:
+
+In order to take the equality constraint into account, we may make use of 
+augmented Lagrangians (a combination of Lagrangian multipliers and penalty 
+terms on the constraint) \cite{bertsekas_multiplier_1976,
+glowinski_augmented_1989,nocedal_numerical_2006}:
 \begin{equation}
@@
 \end{equation}
-To solve the constraint minimization problem, we may now optimize the Augmented Lagrangian in an iterative way:
+
+To solve the constraint minimization problem, we may now optimize the 
+Augmented Lagrangian in an iterative way:
 \begin{equation}
-\left\lbrace \begin{array}{rcl}
+\left\lbrace 
+\begin{array}{rcl}
 u^{t+1} &=& \argmin_{u} AL(u,v^t,\lambda^t,r)\\
@@
 \lambda^{t+1} &=& \lambda^t + \rho(u^{t+1}-v^{t+1})
-\end{array}\right.\end{equation}
-where typically $0 < \rho < r$. The two subminimization problems will now be much easier to handle than the original complete problem (``divide et impera'').
+\end{array}\right.
+\end{equation}
+where typically $0 < \rho < r$. The two sub-minimization problems will 
+now be much easier to handle than the original complete problem.
 
-\subsection{Shape gradients}
-To compute the gradient-descent of the data-term domain integrals with respect to the underlying deformation field, we want to make use of shape gradients \cite{jehan-besson_dream2s:_2003,herbulot_segmentation_2006}.
-A little bit of theory is therefore in order.
+\subsubsection{Shape gradients}
+To compute the gradient-descent of the data-term domain integrals with 
+respect to the underlying deformation field, we use shape gradients 
+\cite{jehan-besson_dream2s:_2003,herbulot_segmentation_2006}. A little 
+bit of theory is therefore in order.
 
-Let $\Omega$ be an image domain and $\omega$ its boundary.
-Further, $r(x)$ is an ``arbitrary'' function over the image domain.
-We now derive the domain integral w.r.t.
-the contour evolution parameter $\tau$ (~ time):
+Let $\Omega$ be an image domain and $d\Omega$ its boundary. Further, $r(\mathbf{x})$ 
+is an ``arbitrary'' function over the image domain. We now derive the domain 
+integral w.r.t. the contour evolution parameter $t$ (time):
 \begin{equation}
-\frac{\partial}{\partial \tau} \int_\Omega r(x) dx = \int_\Omega \frac{\partial r}{\partial \tau}(x) dx - \int_\omega r(x) \left\langle \frac{\partial\omega}{\partial\tau}, N_\omega\right\rangle dx
+\frac{\partial}{\partial t} \int_\Omega r(\mathbf{x}) d\mathbf{x} = \int_\Omega \frac{\partial r}{\partial t}(\mathbf{x}) d\mathbf{x} - \int_{d\Omega} r(\mathbf{x}) \left\langle \frac{\partial{d\Omega}}{\partial t}, N_{d\Omega}\right\rangle d\mathbf{x}
+\label{eq:shape_gradients}
 \end{equation}
-where $\left\langle\frac{\partial\omega}{\partial\tau}, N_\omega\right\rangle$ is the projection of the boundary movement on the unit inward normal.
+where $\left\langle\frac{\partial{d\Omega}}{\partial t}, N_{d\Omega}\right\rangle$ is 
+the projection of the boundary movement on the unit inward normal. We recall
+here that equation \eqref{eq:shape_gradients} stands the bridge between the 
+explicit level set formulations surveyed in \autoref{sec:methods_background} 
+with the presented approach in setting up the velocity function.
 
-\subsection{Min w.r.t. $u$}
-The first minimization problem optimizes the data-term. The problem is equivalent to minimizing the following energy:
+\subsubsection{Minimization w.r.t. $u$}
+The first minimization problem optimizes the data-term. The problem is equivalent 
+to minimizing the following energy:
+\begin{multline}
+E(u,v) = \sum\limits_k \int_{\Omega'_k(u)} \Delta^2_k (\mathbf{f}) \,d\mathbf{x}
++ \langle \lambda, u-v \rangle + \frac{r}{2} \| u - v \|_2^2
+\end{multline}
+where we identify one instance of domain integrals of the form $\int_\Omega r(\mathbf{x}) 
+d\mathbf{x}$ for each region $k$. Optimality requires the derivative of this energy 
+with respect to the parameters $u$ to be null. At this point, we decide to ignore 
+the influence of the boundary shift on the statistics of the regions (i.e. moving the 
+boundary does not significantly impact the $\mu$ and $\Sigma$ descriptors). This means 
+that we can drop the derivative of $r(\mathbf{x})$ w.r.t. contour evolution. 
+What remains are the surface integrals at the subdomain interfaces, plus the Lagrangian 
+and penalty terms:
+
+\begin{multline}
+\frac{\partial E}{\partial u_k} = \sum\limits_k \int_{d\Omega'_k} \left[ \Delta^2_{O(k)} (\mathbf{f}) - \Delta^2_k (\mathbf{f}) \right]
+\left\langle\frac{\partial{d\Omega'_k}}{\partial t}, N_{d\Omega'_k}\right\rangle \,d\mathbf{x} \\
++ \lambda_k + r(u_k - v_k) = 0
+\label{eq:energy_gradient}
+\end{multline}
+where $O(k)$ is a function that returns the neighboring region at $\mathbf{x}$ (i.e.
+the region \emph{outside} the current location in the contour). For the shake of
+simple notation, we will refer as $c'$ to the positions of points belonging to
+any of the existing contours $d\Omega_k$.
+Given the deformation field interpolation stated in \autoref{sec:deformation_field}, 
+the boundary moves according to
+\begin{equation}
+\frac{\partial c'}{\partial u_k} = \psi_k(c')\,e_a
+\end{equation}
+where $e_a$ is the unit vector along direction $a \in \mathbb{R}^n$ and $n$ the dimension
+of the image. Thus
+\begin{equation}
+\left\langle\frac{\partial c'}{\partial u_k}, N_{c'}\right\rangle = \psi_k(c')\,\hat{\mathbf{n}}_{c'}
+\end{equation}
+The discrete implementation of \eqref{eq:energy_gradient} is straightforward as
+we have a triangularized mesh representation of the interfacing surfaces:
+\begin{multline}
+\frac{\partial E}{\partial u_k} = \sum\limits_k \sum\limits_{c'} \left[ \Delta^2_{O(k)} (\mathbf{f}) - \Delta^2_k (\mathbf{f}) \right]
+\, \psi_k(c')\,\hat{\mathbf{n}}_{c'} \\
++ \lambda_k + r(u_k - v_k) = 0
+\label{eq:discrete_energy_gradient}
+\end{multline}
+The optimal distortion $u_k$ is found at each iteration as:
+\begin{multline}
+u_k^{t+1} = v_k^t - \frac{1}{r}\lambda_k^{t} \\
+- \frac{1}{r} \sum\limits_k \sum\limits_{c'} \left[ \Delta^2_{O(k)} (\mathbf{f}) - \Delta^2_k (\mathbf{f}) \right]
+\, \psi_k(c')\,\hat{\mathbf{n}}_{c'}
+\label{eq:update_u}
+\end{multline}
+
+\subsubsection{Minimization w.r.t. $v$}
+For the optimization w.r.t. $v$, the relevant energy writes
 \begin{align}
-E(u) &= \int_{w'(u)} (f-\mu_w)^T\Sigma_w^{-1}(f-\mu_w) dx\nonumber\\
-&\quad +\int_{g'(u)} (f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g) dx\\
-&\quad +\int_{o'(u)} (f-\mu_o)^T\Sigma_o^{-1}(f-\mu_o) dx\nonumber\\
+E(v) &= \int v^T A v d\mathbf{x} + \int \tr\{(\nabla v^T)^T B (\nabla v^T)\} d\mathbf{x}\\
 &\quad + \langle \lambda, u-v \rangle + \frac{r}{2} \| u - v \|_2^2\nonumber
 \end{align}
-where we identify three instances of domain integrals of the form $\int_\Omega r(x) dx$.
-Optimality requires the derivative of this energy with respect to the parameters $u$ to be null.
-At this point, we may decide to ignore the influence of the boundary shift on the statistics of the regions (i.e.
-moving the boundary does not significantly impact the $\mu$ and $\Sigma$ descriptors).
-This means that we can drop the derivative of $r(x)$ w.r.t.
-contour evolution.
-What remains, are surface integrals at the two respective domain interfaces, $c'$ (wm/gm) and $d'$ (gm/CSF), plus the Lagrangian and penalty terms:
-\begin{align}
-\frac{\partial E}{\partial u_k^a} &= \int_{c'} \left[(f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g) - (f-\mu_w)^T\Sigma_w^{-1}(f-\mu_w)\right]\left\langle\frac{\partial c'(s)}{\partial u_k^a}, N_{c'}(s)\right\rangle ds \nonumber\\
-&\quad + \int_{d'} \left[(f-\mu_o)^T\Sigma_o^{-1}(f-\mu_o) - (f-\mu_g)^T\Sigma_g^{-1}(f-\mu_g)\right]\left\langle\frac{\partial d'(s)}{\partial u_k^a}, N_{d'}(s)\right\rangle ds \\
-&\quad + \lambda_k^a + r(u_k^a - v_k^a)\nonumber\\
-& = 0\nonumber
-\end{align}
-where $u_k^a$ is the $a$-th component of the parameter $u_k$, $a\in \{x,y,z\}$, $s$ is the surface parameter, $c'(s)$ and $d'(s)$ the corresponding points on the surfaces $c'$ and $d'$, and $N_{c'}(s)$ and $N_{d'}(s)$ the associated brainwise-inward unit normals.
-Given the deformation field interpolation stated above, the boundary moves according to
-\begin{equation}
-\frac{\partial c'(s)}{\partial u_k^a} = \psi_k(c'(s))e_a
-\end{equation}
-where $e_a$ is the unit vector along direction $a$, and thus
-\begin{equation}
-\left\langle\frac{\partial c'(s)}{\partial u_k^a}, N_{c'}(s)\right\rangle = \psi_k(c'(s))N_{c'}^a(s)
-\end{equation}
-Now, we will discretize the surface integrals over $c'$ and $d'$ by simply summing over the nodes $c_i'$ and $d_i'$:
-\begin{align}
-\frac{\partial E}{\partial u_k^a} &= \sum_{i=1}^{N_c} \left[(f(c_i')-\mu_g)^T\Sigma_g^{-1}(f(c_i')-\mu_g) - (f(c_i')-\mu_w)^T\Sigma_w^{-1}(f(c_i')-\mu_w)\right]\psi_k(c_i')N_{c_i'}^a \nonumber\\
-&\quad + \sum_{i=1}^{N_d} \left[(f(d_i')-\mu_o)^T\Sigma_o^{-1}(f(d_i')-\mu_o) - (f(d_i')-\mu_g)^T\Sigma_g^{-1}(f(d_i')-\mu_g)\right]\psi_k(d_i')N_{d_i'}^a \\
-&\quad + \lambda_k^a + r(u_k^a - v_k^a)\nonumber\\
-& = 0\nonumber
-\end{align}
-It is straightforward to solve this equation for $u_k^a$. The optimal distortion $u_k$ is found at each iteration as:
-\begin{align}
-u_k^{t+1} &= v_k^t - \frac{1}{r}\lambda_k^{t}\nonumber\\
-&\quad - \frac{1}{r}\sum_{i=1}^{N_c} \left[(f(c_i')-\mu_g)^T\Sigma_g^{-1}(f(c_i')-\mu_g) - (f(c_i')-\mu_w)^T\Sigma_w^{-1}(f(c_i')-\mu_w)\right]\psi_k(c_i')N_{c_i'}\\
-&\quad - \frac{1}{r}\sum_{i=1}^{N_d} \left[(f(d_i')-\mu_o)^T\Sigma_o^{-1}(f(d_i')-\mu_o) - (f(d_i')-\mu_g)^T\Sigma_g^{-1}(f(d_i')-\mu_g)\right]\psi_k(d_i')N_{d_i'}\nonumber
-\end{align}
-
-\subsection{Min w.r.t. $v$}
-
-\textbf{It is important to realize that here we do not regularize the actual deformation field (i.e. after interpolation), but the underlying raw parameter field. }
-
-For the optimization w.r.t. $v$, the relevant energy writes
-\begin{align}
-E(v) &= \int v^T A v dx + \int \tr\{(\nabla v^T)^T B (\nabla v^T)\} dx\\
-&\quad + \langle \lambda, u-v \rangle + \frac{r}{2} \| u - v \|_2^2\nonumber
-\end{align}
-Let's assume the simplest, homogeneous isotropic case, $A = \alpha/2$ and $B = \beta/2$. The associated Euler-Lagrange equation is found as:
+\todo[inline]{Do not assume this}
+Let's assume the simplest, homogeneous isotropic case, $A = \alpha/2$ 
+and $B = \beta/2$. The associated Euler-Lagrange equation is found as:
 \begin{equation}
@@
 \begin{equation}
-\Delta = \mathcal{S}_x^- + \mathcal{S}_x^+ + \mathcal{S}_y^- + \mathcal{S}_y^+ - 4 \mathcal{I}
+\Delta = \sum\limits_d \mathcal{S}_d^- + \mathcal{S}_d^+ - 2 \mathcal{I}
 \end{equation}
-where $\mathcal{S}_{x,y}^{\pm}$ stands for the forward ($+$) and backward ($-$) shift operator along $x$ or $y$, respectively, of which the Fourier transform is found easily as
+where $\mathcal{S}_{d}^{\pm}$ stands for the forward ($+$) and backward ($-$) shift 
+operator along coordinates axis $d$, of which the Fourier transform is found easily as
 \begin{equation}
-\mathcal{F}\{\mathcal{S}_{x,y}^{\pm}\} = e^{\pm i\omega_{x,y}},
+\mathcal{F}\{\mathcal{S}_{d}^{\pm}\} = e^{\pm i\omega_{d}},
 \end{equation}
-where $\omega_{x,y}$ is the normalized pulsation along $x$- and $y$-direction. Accordingly, the Fourier transform of the discrete Laplacian is found as
+where $\omega_{d}$ is the normalized pulsation along $d$-direction. \todo{stop!} Accordingly, the 
+Fourier transform of the discrete Laplacian is found as
 \begin{align}
@@
 
-\subsection{Lagrangian multiplier update}
+\subsubsection{Lagrangian multiplier update}
 At each iteration, the Lagrangian multipliers are updated as noted before:
@@
 
-\subsection{Region descriptor reestimation}
+\subsubsection{Region descriptor reestimation}
 In regular intervals, i.e. after $n$ iterations, the parameters $\mu$ and $\Sigma$ of the involved regions need to be reestimated based on the shifted volumetric samples $w_j'$, $g_j'$ and $o_j'$.
 
-\subsection{Convergence}
+\subsubsection{Convergence}
 In order to ``fixate'' the evolution when close to convergence, it is advised to slightly increase the penalty weight $r$ at each iteration.
@@
 the amount by which the new estimate $u$ may move away from the preceding $v$ and vice-versa.
-
-\subsection{Alternative route: Gradient descent}
-If one absolutely wants to avoid the operator splitting and the augmented Lagrangians, then simple gradient descent may work as well (at least under the same simple model circumstances).
-
-At each iteration, we update the distortion along the steepest energy descent:
-\begin{equation}
-\frac{\partial u_k^t}{\partial t} = -\frac{\partial E(u)}{\partial u_k^t}
-\end{equation}
-At this point, we interpret the parameter field $u_k$ to be a continuous function $u(x)$, sampled at the locations $x_k$, and determine the gradient-descent equation\footnote{The same assumption was being made above in the minimization of the regularity terms}:
-\begin{align}
-\frac{\partial u^t}{\partial t} &= - \sum_{i=1}^{N_c} \left[(f(c_i')-\mu_g)^T\Sigma_g^{-1}(f(c_i')-\mu_g) - (f(c_i')-\mu_w)^T\Sigma_w^{-1}(f(c_i')-\mu_w)\right]\psi_{c_i}(x)N_{c_i'}\nonumber\\
-&\quad -\sum_{i=1}^{N_d} \left[(f(d_i')-\mu_o)^T\Sigma_o^{-1}(f(d_i')-\mu_o) - (f(d_i')-\mu_g)^T\Sigma_g^{-1}(f(d_i')-\mu_g)\right]\psi_{d_i}(x)N_{d_i'}\\
-&\quad -\alpha u + \beta\Delta u\nonumber
-\end{align}
-where we have swapped $\psi_k(c_i')$ into $\psi_{c_i}(x)$.
-
-This gradient descent step can be efficiently tackled by discretizing the time in a forward Euler scheme, and making the right hand side semi-implicit in the regularization terms:
-\begin{align}
-\frac{u^{t+1}-u^t}{\tau} &= - \sum_{i=1}^{N_c} \left[(f(c_i')-\mu_g)^T\Sigma_g^{-1}(f(c_i')-\mu_g) - (f(c_i')-\mu_w)^T\Sigma_w^{-1}(f(c_i')-\mu_w)\right]\psi_{c_i}(x)N_{c_i'}\nonumber\\
-&\quad -\sum_{i=1}^{N_d} \left[(f(d_i')-\mu_o)^T\Sigma_o^{-1}(f(d_i')-\mu_o) - (f(d_i')-\mu_g)^T\Sigma_g^{-1}(f(d_i')-\mu_g)\right]\psi_{d_i}(x)N_{d_i'}\nonumber\\
-&\quad -\alpha u^{t+1} + \beta\Delta u^{t+1}
-\end{align}
-where the data terms remain functions of the current estimate $u^t$, i.e. all $c_i' = c_i'(u^t)$ and $d_i' = d_i'(u^t)$. Again, we propose a spectral approach to solve this implicit scheme:
-\begin{equation}
-u^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{u^t/\delta - \sum_{i=1}^{N_c}(\ldots) - \sum_{i=1}^{N_d}(\ldots)  \}}{\mathcal{F}\{(1/\delta+\alpha)\mathcal{I}-\beta\Delta\}} \right\}
-\end{equation}
-It is easily verified, that the same update can be obtained by plugging the AL-update w.r.t.
-$u$ into the AL-update w.r.t.
-$v$, and by identifying $r = 1/\delta$ (the only exception is the distortion on which the data-term is being evaluated).

### 2013-07-01 Publications/2013-TMI/04-methods.tex
Replaced $k$ subscripts by $l$ (#80)
Fixed the inconsistency of having $k$ for both u_k parameters and
region naming. The latter has been replaced by $l$ (stands for label,
from the probabilistic point of view).

@@
 \mathbb{R}^C$ (an $n$-dimensional \emph{image} comprising $C$ different 
-channels) in piecewise smooth regions $\Omega = \lbrace \Omega_k , 
-k\in\left[ 1 \ldots K \right] \rbrace$,  that maximize the \emph{a posteriori}
+channels) in piecewise smooth regions $\Omega = \lbrace \Omega_l , 
+l\in\left[ 1 \ldots L \right] \rbrace$,  that maximize the \emph{a posteriori}
 probability $p(\Omega \mid F)$:
@@
 In the simplest Bayesian segmentation framework, the unknown is the
-parameter set $\Theta_k$ of region-wise descriptors that minimize 
+parameter set $\Theta_l$ of region-wise descriptors that minimize 
 \eqref{eq:bayes_rule}. Conversely, we can assume these descriptors as
@@
 \begin{equation}
-p(F \mid U) \, p(U) = \underset{k}{\prod} \underset{\mathbf{x}\in \Omega'_k}{\prod}
-p_k\left( F(\mathbf{x} + U(\mathbf{x})) \mid U(\mathbf{x}) \right)^{d\mathbf{x}}.
+p(F \mid U) \, p(U) = \underset{l}{\prod} \underset{\mathbf{x}\in \Omega'_l}{\prod}
+p_l\left( F(\mathbf{x} + U(\mathbf{x})) \mid U(\mathbf{x}) \right)^{d\mathbf{x}}.
 \label{eq:bayes_aposteriori}
@@
 \begin{equation}
-p_k( F(\mathbf{x}) \mid U(\mathbf{x}) ) = \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{f}') \right)}}.
+p_l( F(\mathbf{x}) \mid U(\mathbf{x}) ) = \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_l (\mathbf{f}') \right)}}.
 \label{eq:bayes_mpdf}
@@
 F(\mathbf{x} + U(\mathbf{x}))$ with region descriptors
-$\Theta_k = \lbrace \boldsymbol{\mu}_k, \boldsymbol{\Sigma}_k \rbrace$:
+$\Theta_l = \lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \rbrace$:
 \begin{equation}
-\Delta^2_k (\mathbf{f} \mid \Theta_k ) = (\mathbf{f} - \boldsymbol{\mu}_k)^T \, \boldsymbol{\Sigma}^{-1}_k \, (\mathbf{f} - \boldsymbol{\mu}_k).
+\Delta^2_l (\mathbf{f} \mid \Theta_l ) = (\mathbf{f} - \boldsymbol{\mu}_l)^T \, \boldsymbol{\Sigma}^{-1}_l \, (\mathbf{f} - \boldsymbol{\mu}_l).
 \label{eq:bayes_mahalanobis}
@@
 E(F \mid U)= -\log \left[ p(F \mid U) \, p(U) \right] = \\
-= -\log \left[ \underset{k}{\prod} \underset{\mathbf{x}\in \Omega'_k}{\prod}
-p_k( F(\mathbf{x}) \mid U(\mathbf{x}) )^{d\mathbf{x}} \right] = \\
-= \sum\limits_k \int_{\Omega'_k} -\log \left[ p_k(F(\mathbf{x}) \mid U(\mathbf{x} ) ) \right] \, d\mathbf{x},
+= -\log \left[ \underset{l}{\prod} \underset{\mathbf{x}\in \Omega'_l}{\prod}
+p_l( F(\mathbf{x}) \mid U(\mathbf{x}) )^{d\mathbf{x}} \right] = \\
+= \sum\limits_l \int_{\Omega'_l} -\log \left[ p_l(F(\mathbf{x}) \mid U(\mathbf{x} ) ) \right] \, d\mathbf{x},
 \label{eq:energy_1}
@@
 we can express the functional in terms of the deformation field $U$ and
-the region descriptors $\Theta_k$:
+the region descriptors $\Theta_l$:
 \begin{multline}
 E(U) = \\
-= \sum\limits_k \int_{\Omega'_k} -\log \left[ \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{f}) \right)}} \right] \, d\mathbf{x} = \\
-= \sum\limits_k \int_{\Omega'_k} \left[ \frac{1}{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right| \right)} + \frac{1}{2}  \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}.
+= \sum\limits_l \int_{\Omega'_l} -\log \left[ \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_l (\mathbf{f}) \right)}} \right] \, d\mathbf{x} = \\
+= \sum\limits_l \int_{\Omega'_l} \left[ \frac{1}{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right| \right)} + \frac{1}{2}  \Delta^2_l (\mathbf{f}) \right] \,d\mathbf{x}.
 \end{multline}
@@
 \begin{align}
-E(U) = \sum\limits_k \left[ \left|\Omega'_k\right|\,\log \left|\mathbf{\Sigma}_k \right| + \int_{\Omega'_k} \Delta^2_k (\mathbf{f}) \,d\mathbf{x} \right],
+E(U) = \sum\limits_l \left[ \left|\Omega'_l\right|\,\log \left|\mathbf{\Sigma}_l \right| + \int_{\Omega'_l} \Delta^2_l (\mathbf{f}) \,d\mathbf{x} \right],
 \label{eq:map_energy}
 \end{align}
-where we have a constant term scaled by the total volume of the partition $\Omega_k$ and 
-the determinant of the  covariance matrix of the partition $\left|\boldsymbol{\Sigma}_{k}\right|$,
+where we have a constant term scaled by the total volume of the partition $\Omega_l$ and 
+the determinant of the  covariance matrix of the partition $\left|\boldsymbol{\Sigma}_{l}\right|$,
 plus an energy term based on the squared \emph{Mahalanobis distance}.
@@
 \begin{align}
-E(U) = \sum\limits_k \int_{\Omega'_k} \Delta^2_k (\mathbf{f}) \,d\mathbf{x}.
+E(U) = \sum\limits_l \int_{\Omega'_l} \Delta^2_l (\mathbf{f}) \,d\mathbf{x}.
 \label{eq:final_map_energy}
@@
 can deal with more general distributions. This is necessary to avoid the 
-assumption that regions $\Omega_k$ have a fixed covariance matrix on their 
+assumption that regions $\Omega_l$ have a fixed covariance matrix on their 
 complete domain. One immediate advantage of this functional from the original 
@@
 \begin{multline}
-E(\Theta_k,\Omega_k) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x} \\
-+ \lambda \int_{\Omega_k - K}  ( \left| \nabla \mathbf{\mu} \right| ^2 + \left| \nabla \mathbf{\Sigma}_k \right| ^2 ) \, d\mathbf{x} 
-+ \nu |K|,
+E(\Theta_l,\Omega_l) = \sum\limits_l \int_{\Omega_l} \left[ \log \left|\mathbf{\Sigma}_l\right| + \Delta^2_l (\mathbf{f}) \right] \,d\mathbf{x} \\
++ \lambda \int_{\Omega_l - d\Omega_l}  ( \left| \nabla \mathbf{\mu} \right| ^2 + \left| \nabla \mathbf{\Sigma}_l \right| ^2 ) \, d\mathbf{x} 
++ \nu |d\Omega_l|,
 \end{multline}
@@
 \begin{equation}
-E(\Theta_k,K) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}
-+ \nu |K|.
+E(\Theta_l,d\Omega_l) = \sum\limits_l \int_{\Omega_l} \left[ \log \left|\mathbf{\Sigma}_l\right| + \Delta^2_l (\mathbf{f}) \right] \,d\mathbf{x}
++ \nu |d\Omega_l|.
 \end{equation}
 
-As long as we do not penalize the edge set $K$ length, $\nu = 0$ and
+As long as we do not penalize the edge set $d\Omega_l$ length, $\nu = 0$ and
 the result is dual to \eqref{eq:map_energy}:
 \begin{equation}
-E(\Theta_k,K) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}.
+E(\Theta_l,d\Omega_l) = \sum\limits_l \int_{\Omega_l} \left[ \log \left|\mathbf{\Sigma}_l\right| + \Delta^2_l (\mathbf{f}) \right] \,d\mathbf{x}.
 \end{equation}
@@
 \label{eq:transformation}
-c_i' = T\{c_i\} = c_i + u(c_i) = c_i + \sum_k \psi_k(c_i)u_k
+c_i' = T\{c_i\} = c_i + u(c_i) = c_i + \sum_k \psi_k(c_i)\,u_k
 \end{equation}
@@
 \begin{multline}
-E(u) = \sum\limits_k \int_{\Omega'_k} \Delta^2_k (\mathbf{f}) \,d\mathbf{x} \\
+E(u) = \sum\limits_l \int_{\Omega'_l} \Delta^2_l (\mathbf{f}) \,d\mathbf{x} \\
 + \int u^T \, A \, u \, d\mathbf{x} + \int \tr\{(\nabla u^T)^T B (\nabla u^T)\} d\mathbf{x}
@@
 \begin{multline}
-E(u,v) = \sum\limits_k \int_{\Omega'_k(u)} \Delta^2_k (\mathbf{f}) \,d\mathbf{x} \\
+E(u,v) = \sum\limits_l \int_{\Omega'_l(u)} \Delta^2_l (\mathbf{f}) \,d\mathbf{x} \\
 + \int v^T \, A \, v \, d\mathbf{x} + \int \tr\{(\nabla v^T)^T B (\nabla v^T)\} d\mathbf{x}
@@
 \begin{multline}
-E(u,v) = \sum\limits_k \int_{\Omega'_k(u)} \Delta^2_k (\mathbf{f}) \,d\mathbf{x}
+E(u,v) = \sum\limits_l \int_{\Omega'_l(u)} \Delta^2_l (\mathbf{f}) \,d\mathbf{x}
 + \langle \lambda, u-v \rangle + \frac{r}{2} \| u - v \|_2^2
@@
 where we identify one instance of domain integrals of the form $\int_\Omega r(\mathbf{x}) 
-d\mathbf{x}$ for each region $k$. Optimality requires the derivative of this energy 
+d\mathbf{x}$ for each region $l$. Optimality requires the derivative of this energy 
 with respect to the parameters $u$ to be null. At this point, we decide to ignore 
@@
 \begin{multline}
-\frac{\partial E}{\partial u_k} = \sum\limits_k \int_{d\Omega'_k} \left[ \Delta^2_{O(k)} (\mathbf{f}) - \Delta^2_k (\mathbf{f}) \right]
-\left\langle\frac{\partial{d\Omega'_k}}{\partial t}, N_{d\Omega'_k}\right\rangle \,d\mathbf{x} \\
+\frac{\partial E}{\partial u_k} = \sum\limits_l \int_{d\Omega'_l} \left[ \Delta^2_{O(l)} (\mathbf{f}) - \Delta^2_l (\mathbf{f}) \right]
+\left\langle\frac{\partial{d\Omega'_l}}{\partial t}, N_{d\Omega'_l}\right\rangle \,d\mathbf{x} \\
 + \lambda_k + r(u_k - v_k) = 0
@@
 \end{multline}
-where $O(k)$ is a function that returns the neighboring region at $\mathbf{x}$ (i.e.
+where $O(l)$ is a function that returns the neighboring region at $\mathbf{x}$ (i.e.
 the region \emph{outside} the current location in the contour). For the shake of
 simple notation, we will refer as $c'$ to the positions of points belonging to
-any of the existing contours $d\Omega_k$.
+any of the existing contours $d\Omega_l$.
 Given the deformation field interpolation stated in \autoref{sec:deformation_field}, 
@@
 \begin{multline}
-\frac{\partial E}{\partial u_k} = \sum\limits_k \sum\limits_{c'} \left[ \Delta^2_{O(k)} (\mathbf{f}) - \Delta^2_k (\mathbf{f}) \right]
-\, \psi_k(c')\,\hat{\mathbf{n}}_{c'} \\
+\frac{\partial E}{\partial u_k} = \sum\limits_l \sum\limits_{c'} \left[ \Delta^2_{O(l)} (\mathbf{f}) - \Delta^2_l (\mathbf{f}) \right]
+\, \psi_l(c')\,\hat{\mathbf{n}}_{c'} \\
 + \lambda_k + r(u_k - v_k) = 0
@@
 u_k^{t+1} = v_k^t - \frac{1}{r}\lambda_k^{t} \\
-- \frac{1}{r} \sum\limits_k \sum\limits_{c'} \left[ \Delta^2_{O(k)} (\mathbf{f}) - \Delta^2_k (\mathbf{f}) \right]
+- \frac{1}{r} \sum\limits_l \sum\limits_{c'} \left[ \Delta^2_{O(l)} (\mathbf{f}) - \Delta^2_l (\mathbf{f}) \right]
 \, \psi_k(c')\,\hat{\mathbf{n}}_{c'}

### 2013-07-01 Publications/2013-TMI/04-methods.tex
First serious draft of intro for TMI
This version contains the bayesian/acwe demonstration as required in the correspoinding issue. (Close #80)

@@
 Naming $\Phi_G$ a generic objective function, a dense deformation field of vectors 
-$u: \mathbb{R}^n \to \mathbb{R}^n$ (tipically $n = \{ 2, 3 \}$) is introduced. 
+$u: \mathbb{R}^n \to \mathbb{R}^n$ (typically $n = \{ 2, 3 \}$) is introduced. 
 Then, the conservation of the morphological description is assumed, such as
@@
 \frac{\partial u(\mathbf{x},t)}{\partial t} = - \nu( \Phi_G(\mathbf{x} + u(\mathbf{x},t), 0) ) N_{\Phi_G}.
+\label{eq:active_deformation_fields}
 \end{equation}
 
-Here, we propose an active deformation field framework with some differences in
-the formulation of the problem. The main diverging points with respect to
+Here, we propose an \gls{adf} -like framework with some differences in
+the inspiration of the solution. Even though the historical evolution
+of joint segmentation and registration procedures is strongly linked
+to the level sets approximation, our work is based in derivations of
+the \gls{acwe} framework \citep{chan_active_2001}, more specifically 
+\citep{le_guyader_combined_2011}.
+The main diverging points with respect to
 \citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
@@
 \todo[inline]{regularization is also based on linear diffusion smoothing REF-thirion,
-but they use a Gaussian filtering)}; 3) optimization is applied in the spectral
+but they use a Gaussian filtering). Maybe \citep{le_guyader_combined_2011} contains
+more regarding this difference, given that they use the same regularization as us. But,
+currently, I do not have enough background to write this second difference}
+\textbf{}; 3) optimization is applied in the spectral
 domain, observing anisotropic and inhomogeneous mappings along each direction.
+With respect \citep{le_guyader_combined_2011}, the main differences are
+the distance function, and the spectral solution to the optimization updates,
+as we shall cover in \autoref{sec:numerical_implementation}.
+
+\todo[inline]{Disclaimer: I've been exploring the joint registration-segmentation
+applications starting from Gorthi's paper. Later, Dominique pointed me out that
+the main inspiration of the design comes from the ACWE world. If one takes a quick
+look on \citep{le_guyader_combined_2011} this is clear. But I did not have time
+yet to read this paper in depth. Fortunately, it does not seem to contain a lot
+of references to literature, so I hope it will be ease to re-formulate this
+section with its contents. Also, the highlighted differences will require an
+update once I fully understand that method}
 
@@
 \subsection{Numerical Implementation}
+\label{sec:numerical_implementation}
 %
@@
 Using the region descriptors derived in \autoref{sec:methods_map}, we propose
-an \gls{acwe}-like, piece-wise constant, variational image segmentation
+an \gls{adf}-like, \gls{acwe}-based, piece-wise constant, image segmentation
 model (where the unknown is the deformation field)
@@
 regularization \cite{nagel_investigation_1986} to better regularize
-the \gls{epi} distortion. \\
-%
+the \gls{epi} distortion.
+
+\subsubsection{Shape gradients}
+To compute the gradient-descent of the data-term domain integrals
+in \eqref{eq:complete_energy} with respect to the underlying 
+deformation field, we use shape gradients 
+\cite{jehan-besson_dream2s:_2003,herbulot_segmentation_2006}.
+
+Let $\Omega$ be an image domain and $d\Omega$ its boundary. Further, 
+$r(\mathbf{x})$ is an ``arbitrary'' function over the image domain. 
+We now derive the domain integral w.r.t. the contour evolution 
+parameter $t$ (time):
+\begin{equation}
+\frac{\partial}{\partial t} \int_\Omega r(\mathbf{x}) d\mathbf{x} = \int_\Omega \frac{\partial r}{\partial t}(\mathbf{x}) d\mathbf{x} - \int_{d\Omega} r(\mathbf{x}) \left\langle \frac{\partial{d\Omega}}{\partial t}, N_{d\Omega}\right\rangle d\mathbf{x}
+\label{eq:shape_gradients}
+\end{equation}
+where $\left\langle\frac{\partial{d\Omega}}{\partial t}, N_{d\Omega}\right\rangle$ is 
+the projection of the boundary movement on the unit inward normal. We recall
+here that equation \eqref{eq:shape_gradients} stands the bridge between the 
+explicit level set formulations surveyed in \autoref{sec:methods_background}
+by updating the distortion along the steepest energy descent:
+\begin{equation}
+\frac{\partial u(\mathbf{x},t)}{\partial t} = - \frac{\partial E(u)}{\partial u},
+\end{equation}
+what closes the loop with the general expression of the \glspl{adf} introduced in
+\eqref{eq:active_deformation_fields}.
 
@@
 
-\subsubsection{Shape gradients}
-To compute the gradient-descent of the data-term domain integrals with 
-respect to the underlying deformation field, we use shape gradients 
-\cite{jehan-besson_dream2s:_2003,herbulot_segmentation_2006}. A little 
-bit of theory is therefore in order.
-
-Let $\Omega$ be an image domain and $d\Omega$ its boundary. Further, $r(\mathbf{x})$ 
-is an ``arbitrary'' function over the image domain. We now derive the domain 
-integral w.r.t. the contour evolution parameter $t$ (time):
-\begin{equation}
-\frac{\partial}{\partial t} \int_\Omega r(\mathbf{x}) d\mathbf{x} = \int_\Omega \frac{\partial r}{\partial t}(\mathbf{x}) d\mathbf{x} - \int_{d\Omega} r(\mathbf{x}) \left\langle \frac{\partial{d\Omega}}{\partial t}, N_{d\Omega}\right\rangle d\mathbf{x}
-\label{eq:shape_gradients}
-\end{equation}
-where $\left\langle\frac{\partial{d\Omega}}{\partial t}, N_{d\Omega}\right\rangle$ is 
-the projection of the boundary movement on the unit inward normal. We recall
-here that equation \eqref{eq:shape_gradients} stands the bridge between the 
-explicit level set formulations surveyed in \autoref{sec:methods_background} 
-with the presented approach in setting up the velocity function.
-
 \subsubsection{Minimization w.r.t. $u$}
@@
 the region \emph{outside} the current location in the contour). For the shake of
-simple notation, we will refer as $c'$ to the positions of points belonging to
-any of the existing contours $d\Omega_l$.
+simple notation, we will refer as $c_i$ to the positions 
+of points belonging to any of the existing contours $d\Omega_l$ (with 
+$i \in [1 \ldots N_c]$ and $N_c$ the total number of control points).
 Given the deformation field interpolation stated in \autoref{sec:deformation_field}, 
@@
 \begin{equation}
-\frac{\partial c'}{\partial u_k} = \psi_k(c')\,e_a
+\frac{\partial c_i}{\partial u_k} = \psi_k(c_i)\,e_a
 \end{equation}
 where $e_a$ is the unit vector along direction $a \in \mathbb{R}^n$ and $n$ the dimension
-of the image. Thus
+of the image. Thus, if we name the inward unit normal of the corresponding transformed 
+contour at position $c'_i$ as $\hat{\mathbf{n}}_{c'_i}$,
 \begin{equation}
-\left\langle\frac{\partial c'}{\partial u_k}, N_{c'}\right\rangle = \psi_k(c')\,\hat{\mathbf{n}}_{c'}
+\left\langle\frac{\partial c_i}{\partial u_k}, N_{c_i}\right\rangle = \psi_k(c_i)\,\hat{\mathbf{n}}_{c'_i}.
 \end{equation}
@@
 \begin{multline}
-\frac{\partial E}{\partial u_k} = \sum\limits_l \sum\limits_{c'} \left[ \Delta^2_{O(l)} (\mathbf{f}) - \Delta^2_l (\mathbf{f}) \right]
-\, \psi_l(c')\,\hat{\mathbf{n}}_{c'} \\
-+ \lambda_k + r(u_k - v_k) = 0
+\frac{\partial E}{\partial u_k} = \sum\limits_{N_c} \left[ \Delta^2_{O(l)} (\mathbf{f}) - \Delta^2_l (\mathbf{f}) \right]
+\, \psi_{c_i}(\mathbf{x}_k)\,\hat{\mathbf{n}}_{c'_i} \\
++ \lambda_k + r(u_k - v_k) = 0,
 \label{eq:discrete_energy_gradient}
 \end{multline}
-The optimal distortion $u_k$ is found at each iteration as:
+where we have swapped $\psi_k(c_i)$ by $\psi_{c_i}(\mathbf{x}_k)$ for
+convenience in implementation, taking the advantage that they are
+equivalent. Subsequently, the optimal distortion $u_k$ 
+is found at each iteration as:
 \begin{multline}
 u_k^{t+1} = v_k^t - \frac{1}{r}\lambda_k^{t} \\
-- \frac{1}{r} \sum\limits_l \sum\limits_{c'} \left[ \Delta^2_{O(l)} (\mathbf{f}) - \Delta^2_l (\mathbf{f}) \right]
-\, \psi_k(c')\,\hat{\mathbf{n}}_{c'}
+- \frac{1}{r} \sum\limits_{N_c} \left[ \Delta^2_{O(l)} (\mathbf{f}) - \Delta^2_l (\mathbf{f}) \right]
+\, \psi_{c_i}(\mathbf{x}_k)\,\hat{\mathbf{n}}_{c'_i}
 \label{eq:update_u}
@@
 \end{align}
-\todo[inline]{Do not assume this}
-Let's assume the simplest, homogeneous isotropic case, $A = \alpha/2$ 
-and $B = \beta/2$. The associated Euler-Lagrange equation is found as:
+The associated Euler-Lagrange equation is found as:
 \begin{equation}
-\alpha v - \beta\Delta v + rv = ru + \lambda
+2 ( A v - B \Delta v ) + rv = ru + \lambda
 \end{equation}
@@
 \begin{equation}
-v^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{ru + \lambda\}}{\mathcal{F}\{(\alpha+r)\mathcal{I}-\beta\Delta\}} \right\}
+v^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{ru + \lambda\}}{\mathcal{F}\{( 2 A+r)\mathcal{I}-2 B \Delta\}} \right\}
 \end{equation}
@@
 \end{equation}
-where $\omega_{d}$ is the normalized pulsation along $d$-direction. \todo{stop!} Accordingly, the 
+where $\omega_{d}$ is the normalized pulsation along $d$-direction. Accordingly, the 
 Fourier transform of the discrete Laplacian is found as
-\begin{align}
-\mathcal{F}\{\Delta\} &= e^{-i\omega_x } + e^{i\omega_x } + e^{-i\omega_y } + e^{i\omega_y } - 4\nonumber\\
-&= 2\left( \cos(\omega_x) + \cos(\omega_y) - 2 \right)
-\end{align}
+\begin{equation}
+\mathcal{F}\{\Delta\} = \sum\limits_d e^{-i\omega_d } + e^{i\omega_d } - 2 = \sum\limits_d \left( 2\cos(\omega_d) - 2 \right)
+\end{equation}
 
-The remaining transforms are trivial or can be computed using FFT (as in \citep{estellers_efficient_2011}).
+The remaining transforms are trivial or can be computed using the \gls{fft} 
+as in \citep{estellers_efficient_2011}.
 
@@
 \subsubsection{Region descriptor reestimation}
-In regular intervals, i.e. after $n$ iterations, the parameters $\mu$ and $\Sigma$ of the involved regions need to be reestimated based on the shifted volumetric samples $w_j'$, $g_j'$ and $o_j'$.
+In regular intervals, after certain number of iterations,
+the parameters $\Theta_l$ of the regions can be reestimated 
+based on the shifted volumetric samples 
+$\mathbf{x}' = \mathbf{x_0} + u(\mathbf{x})$.
 
 \subsubsection{Convergence}
-In order to ``fixate'' the evolution when close to convergence, it is advised to slightly increase the penalty weight $r$ at each iteration.
-Note that as can be seen in the above equations, $r$ governs the step-size or  leash-length at each iteration, i.e.
-the amount by which the new estimate $u$ may move away from the preceding $v$ and vice-versa.
+In order to ``fixate'' the evolution when close to convergence,
+it is advised to slightly increase the penalty weight $r$ at each 
+iteration. Note that as can be seen in the above equations, $r$ 
+governs the step-size or leash-length at each iteration, i.e. the 
+amount by which the new estimate $u$ may move away from the preceding 
+$v$ and vice-versa.
+
+\subsubsection{Efficient field densification}
+As long as the dense deformation field is iteratively interpolated
+from the same set of control points that define the $L-1$ contours,
+it is possible to pre-cache all the $\psi_k(c_i)$ weights into a
+sparse matrix for fast densification. As well, we achieve a 
+diffeomorphic transform by ... 
+\todo[inline]{this is actually false (IMHO), so far :(. We will need
+to add some more conditions to the interopolating kernel so that
+the original values of the transform in $c_i$ can be recovered from 
+interpolated values of $u_k$}

### 2013-09-18 Scripts/workflows/dmri.py
Scaffold for t2 registration pipeline

@@ -13,6 +13,34 @@ import nipype.interfaces.diffusion_toolkit as dtk
 import nipype.interfaces.mrtrix as mrtrix    #<---- The important new part!
 import nipype.interfaces.camino as camino
 
+
+def t2_registation_correct( name="T2_Registration" ):
+    pipeline = pe.Workflow( name=name )
+    
+    inputnode = pe.Node( niu.IdentityInterface( fields=['in_file',
+                                                        'in_t2',
+                                                        'in_mask']
+                         ), name='inputnode' )
+
+    outputnode = pe.Node( niu.IdentityInterface( fields=['epi_corrected']),
+                          name='outputnode' )
+
+    get_b0 = pe.Node(fsl.utils.ExtractROI(
+        t_size=1, t_min=0), name="getB0")
+
+    registration = pe.Node( ANTS , name='B0-to-T2' )
+
+    dwi_split = pe.Node(niu.Function(input_names=['in_file'], output_names=['out_files'], function=_split_dwi), name='split_DWI')
+    applytfm = pe.MapNode( , iterfield=[''], name='AntsInverseTf')
+    dwi_merge = pe.Node(fsl.utils.Merge(dimension='t'), name='merge_DWI')
+
+    pipeline.connect( [
+                         (inputnode,      get_b0, [ ('in_file','in_file') ] )
+                        ,(get_b0,        mask_b0, [ ('roi_file','') ] )
+                        ,(inputnode,     mask_t2, [ ('in_t2','') ] )
+
+                      ]
+
 def fugue_all_workflow(name="Fugue_WarpDWIs"):
     pipeline = pe.Workflow(name=name)
     

### 2013-09-18 Publications/2013-ISBI14/01-abstract.tex
Base for ISBI14 paper

\begin{abstract}

Connectivity analysis on \gls*{dmri} data of the whole-brain
suffers from distortions originated by the standard 
\gls*{epi} sequences used in acquisition. These artifacts
are present at interfaces between tissues, due to the inhomogeneity
of the field derived from the discontinuity of magnetic
susceptibility. As a result, registered signal shows characteristic
geometrical deformations and signal destruction. Thus, the artifact
is a significant pitfall that limits the success of tractography 
algorithms.

Several retrospective correction techniques are readily
available, generally inherited from \gls*{fmri} literature,
as those data are usually acquired with \gls*{epi} as well.
These correction methodologies have been evaluated and 
compared in physical phantoms designed for \gls*{fmri} 
experiments, but little is known about their impact
on \gls*{dmri} and more precisely on the resulting 
tractography.

In this work, we use a renowned digital phantom designed
for assessing tractography algorithms and subject it to
a theoretically correct and plausible deformation. We corrected
data with several available methodologies and provide
a ranking based on the accuracy of the resulting tractographies.

\end{abstract}

### 2013-09-18 Publications/2013-ISBI14/02-introduction.tex
Base for ISBI14 paper

\section{INTRODUCTION}

This template provides authors with most of the formatting specifications needed for preparing electronic versions of their papers. All standard paper components have been specified for three reasons: (1) ease of use when formatting individual papers, (2) automatic compliance to electronic requirements that facilitate the concurrent or later production of electronic products, and (3) conformity of style throughout a conference proceedings. Margins, column widths, line spacing, and type styles are built-in; examples of the type styles are provided throughout this document and are identified in italic type, within parentheses, following the example. Some components, such as multi-leveled equations, graphics, and tables are not prescribed, although the various table text styles are provided. The formatter will need to create these components, incorporating the applicable criteria that follow.

### 2013-09-18 Publications/2013-ISBI14/03-methods.tex
Base for ISBI14 paper

\section{PROCEDURE FOR PAPER SUBMISSION}

\subsection{Selecting a Template (Heading 2)}

First, confirm that you have the correct template for your paper size. This template has been tailored for output on the US-letter paper size. 
It may be used for A4 paper size if the paper size setting is suitably modified.

\subsection{Maintaining the Integrity of the Specifications}

The template is used to format your paper and style the text. All margins, column widths, line spaces, and text fonts are prescribed; please do not alter them. You may note peculiarities. For example, the head margin in this template measures proportionately more than is customary. This measurement and others are deliberate, using specifications that anticipate your paper as one part of the entire proceedings, and not as an independent document. Please do not revise any of the current designations

### 2013-09-18 Publications/2013-ISBI14/04-results.tex
Base for ISBI14 paper

\section{MATH}

Before you begin to format your paper, first write and save the content as a separate text file. Keep your text and graphic files separate until after the text has been formatted and styled. Do not use hard tabs, and limit use of hard returns to only one return at the end of a paragraph. Do not add any kind of pagination anywhere in the paper. Do not number text heads-the template will do that for you.

Finally, complete content and organizational editing before formatting. Please take note of the following items when proofreading spelling and grammar:

\subsection{Abbreviations and Acronyms} Define abbreviations and acronyms the first time they are used in the text, even after they have been defined in the abstract. Abbreviations such as IEEE, SI, MKS, CGS, sc, dc, and rms do not have to be defined. Do not use abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}

\begin{itemize}

\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as Ò3.5-inch disk driveÓ.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ÒWb/m2Ó or Òwebers per square meterÓ, not Òwebers/m2Ó.  Spell out units when they appear in text: Ò. . . a few henriesÓ, not Ò. . . a few HÓ.
\item Use a zero before decimal points: Ò0.25Ó, not Ò.25Ó. Use Òcm3Ó, not ÒccÓ. (bullet list)

\end{itemize}

\subsection{Equations}

The equations are an exception to the prescribed specifications of this template. You will need to determine whether or not your equation should be typed using either the Times New Roman or the Symbol font (please no other font). To create multileveled equations, it may be necessary to treat the equation as a graphic and insert it into the text after your paper is styled. Number equations consecutively. Equation numbers, within parentheses, are to position flush right, as in (1), using a right tab stop. To make your equations more compact, you may use the solidus ( / ), the exp function, or appropriate exponents. Italicize Roman symbols for quantities and variables, but not Greek symbols. Use a long dash rather than a hyphen for a minus sign. Punctuate equations with commas or periods when they are part of a sentence, as in

$$
\alpha + \beta = \chi \eqno{(1)}
$$

Note that the equation is centered using a center tab stop. Be sure that the symbols in your equation have been defined before or immediately following the equation. Use Ò(1)Ó, not ÒEq. (1)Ó or Òequation (1)Ó, except at the beginning of a sentence: ÒEquation (1) is . . .Ó

\subsection{Some Common Mistakes}
\begin{itemize}

\item The word ÒdataÓ is plural, not singular.
\item The subscript for the permeability of vacuum ?0, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ÒoÓ.
\item In American English, commas, semi-/colons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ÒinsetÓ, not an ÒinsertÓ. The word alternatively is preferred to the word ÒalternatelyÓ (unless you really mean something that alternates).
\item Do not use the word ÒessentiallyÓ to mean ÒapproximatelyÓ or ÒeffectivelyÓ.
\item In your paper title, if the words Òthat usesÓ can accurately replace the word ÒusingÓ, capitalize the ÒuÓ; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ÒaffectÓ and ÒeffectÓ, ÒcomplementÓ and ÒcomplimentÓ, ÒdiscreetÓ and ÒdiscreteÓ, ÒprincipalÓ and ÒprincipleÓ.
\item Do not confuse ÒimplyÓ and ÒinferÓ.
\item The prefix ÒnonÓ is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ÒetÓ in the Latin abbreviation Òet al.Ó.
\item The abbreviation Òi.e.Ó means Òthat isÓ, and the abbreviation Òe.g.Ó means Òfor exampleÓ.

\end{itemize}

### 2013-09-18 Publications/2013-ISBI14/05-conclusions.tex
Base for ISBI14 paper

\section{USING THE TEMPLATE}

Use this sample document as your LaTeX source file to create your document. Save this file as {\bf root.tex}. You have to make sure to use the cls file that came with this distribution. If you use a different style file, you cannot expect to get required margins. Note also that when you are creating your out PDF file, the source file is only part of the equation. {\it Your \TeX\ $\rightarrow$ PDF filter determines the output file size. Even if you make all the specifications to output a letter file in the source - if your filter is set to produce A4, you will only get A4 output. }

It is impossible to account for all possible situation, one would encounter using \TeX. If you are using multiple \TeX\ files you must make sure that the ``MAIN`` source file is called root.tex - this is particularly important if your conference is using PaperPlaza's built in \TeX\ to PDF conversion tool.

\subsection{Headings, etc}

Text heads organize the topics on a relational, hierarchical basis. For example, the paper title is the primary text head because all subsequent material relates and elaborates on this one topic. If there are two or more sub-topics, the next level head (uppercase Roman numerals) should be used and, conversely, if there are not at least two sub-topics, then no subheads should be introduced. Styles named ÒHeading 1Ó, ÒHeading 2Ó, ÒHeading 3Ó, and ÒHeading 4Ó are prescribed.

\subsection{Figures and Tables}

Positioning Figures and Tables: Place figures and tables at the top and bottom of columns. Avoid placing them in the middle of columns. Large figures and tables may span across both columns. Figure captions should be below the figures; table heads should appear above the tables. Insert figures and tables after they are cited in the text. Use the abbreviation ÒFig. 1Ó, even at the beginning of a sentence.

\begin{table}[h]
\caption{An Example of a Table}
\label{table_example}
\begin{center}
\begin{tabular}{|c||c|}
\hline
One & Two\\
\hline
Three & Four\\
\hline
\end{tabular}
\end{center}
\end{table}

   \begin{figure}[thpb]
      \centering
      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
}}
      %\includegraphics[scale=1.0]{figurefile}
      \caption{Inductance of oscillation winding on amorphous
       magnetic core versus DC bias magnetic field}
      \label{figurelabel}
   \end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words rather than symbols or abbreviations when writing Figure axis labels to avoid confusing the reader. As an example, write the quantity ÒMagnetizationÓ, or ÒMagnetization, MÓ, not just ÒMÓ. If including units in the label, present them within parentheses. Do not label axes only with units. In the example, write ÒMagnetization (A/m)Ó or ÒMagnetization {A[m(1)]}Ó, not just ÒA/mÓ. Do not label axes with a ratio of quantities and units. For example, write ÒTemperature (K)Ó, not ÒTemperature/K.Ó

### 2013-09-18 Publications/2013-ISBI14/01-abstract.tex
Introduction of ISBI14
Updated bibliography

@@
 susceptibility. As a result, registered signal shows characteristic
-geometrical deformations and signal destruction. Thus, the artifact
-is a significant pitfall that limits the success of tractography 
+geometrical deformations and signal destruction. These artifacts
+suppose a significant pitfall that limits the success of tractography 
 algorithms.
@@
 In this work, we use a renowned digital phantom designed
-for assessing tractography algorithms and subject it to
-a theoretically correct and plausible deformation. We corrected
-data with several available methodologies and provide
-a ranking based on the accuracy of the resulting tractographies.
+for assessing tractography algorithms, and subject it to
+a ``theoretically correct'' and plausible deformation. We
+corrected data back, with several available techniques 
+(namely fieldmap-based, reversed encoding-based, 
+registration-based). Finally, we ranked the methods 
+based on the geometrical accuracy, the dropout compensation, 
+and the impact on the resulting tractography.
 

### 2013-09-18 Publications/2013-ISBI14/02-introduction.tex
Introduction of ISBI14
Updated bibliography

@@
 
-This template provides authors with most of the formatting specifications needed for preparing electronic versions of their papers.
-All standard paper components have been specified for three reasons: (1) ease of use when formatting individual papers, (2) automatic compliance to electronic requirements that facilitate the concurrent or later production of electronic products, and (3) conformity of style throughout a conference proceedings.
-Margins, column widths, line spacing, and type styles are built-in; examples of the type styles are provided throughout this document and are identified in italic type, within parentheses, following the example.
-Some components, such as multi-leveled equations, graphics, and tables are not prescribed, although the various table text styles are provided.
-The formatter will need to create these components, incorporating the applicable criteria that follow.
+In-vivo whole-brain connectivity analysis has been a
+research topic of high interest for the last
+5 years \cite{jones_studying_2008}. 
+\Gls*{dmri} can be used to probe the
+orientation of fiber bundles within the brain,
+generally applying \gls*{epi} sequences.
+After a signal reconstruction step, tractography
+algorithms draw a map of sampled structure.
+These maps can represent the actual trajectories
+of fiber bundles (deterministic tractography,
+\cite{basser_microstructural_2011}) or pixel-wise
+probability of connection to a certain origin
+(probabilistic tractography). Finally, the
+information about these connections is collected
+into a network matrix that can be subjected to
+the so-called ``connectome analysis''.
+
+However, such a complex workflow is prone to
+flaws that have been detailedly reported 
+\cite{jones_twenty-five_2010,jbabdi_tractography:_2011}.
+In order to enable reliable and robust connectome
+analyses of the brain, these challenging needs for
+solutions must be addressed. One prominent source
+of inaccuracy are the susceptibility-derived artifacts,
+for which \Gls*{epi} schemes are specially sensitive.
+As susceptibility changes at tissue interfaces,
+so does the magnetic field. The consequence of the 
+inhomogeneity of the field translates in a highly 
+distorted imaged anatomy and a significant signal 
+destruction of certain regions of the brain 
+(e.g. the orbitofrontal lobe, for the proximity of the
+air surrounding sinuses). This artifact has been
+profoundly described, and a number of methodologies
+have been proposed. First approaches to
+retrospective correction \cite{jezzard_correction_1995},
+relied on an extra \gls*{mri} acquisition (so-called
+field mapping), that probes the inhomogeneity of the field.
+A second theory-based breed of methodologies acquire a 
+map of the \gls*{psf} of the \gls*{epi} readouts to correct 
+for the artifact \cite{robson_measurement_1997}. 
+Next generation of methodologies \cite{cordes_geometric_2000,
+chiou_simple_2000}, make use of the acquisition of 
+extra \gls*{epi} volumes 
+with specific differences that enable correction, for instance
+the opposed direction of the phase encoding gradients 
+Finally, the last family of methodologies acquire an 
+extra \gls*{t2} image, and use its anatomical correctness 
+as reference to find the deformation map through nonlinear 
+registration \cite{kybic_unwarping_2000,studholme_accurate_2000}.
+Registration-based methods usually map
+the \gls*{t2} image to the so-called baseline (B0) volume 
+of \gls*{dmri}. The choice of \gls*{t2} is due to the strong
+similarity of intensity distribution with the B0 of \gls*{dmri}
+datasets. More recent works report extensions of the existing
+techniques \cite{andersson_how_2003,holland_efficient_2010,
+andersson_comprehensive_2012}, and combined approaches
+\cite{zaitsev_point_2004}.
+
+Even though aforementioned techniques for distortion correction
+have been studied \cite{zeng_image_2002,wu_comparison_2008},
+the lack of a gold-standard limit the possible benchmarking 
+strategies. \cite{irfanoglu_effects_2012} opened the question
+of the impact of the distortions on tractography. In this work,
+we propose an evaluation framework using a detailed digital 
+phantom designed for tractography. Using this framework,
+we are able to compare several correction techniques and we
+provide results related to their performance in the task
+itself and their impact on subsequent tractography.

### 2013-09-18 Publications/2013-ISBI14/03-methods.tex
Introduction of ISBI14
Updated bibliography

@@
-\section{PROCEDURE FOR PAPER SUBMISSION}
+\section{METHODS}
 
-\subsection{Selecting a Template (Heading 2)}
+\subsection{Digital Phantom}
+We modified the digital phantom provided with the
+\emph{\gls*{hardi} reconstruction challenge} held
+in ISBI 2013, San Francisco, US. This model is
+built upon high-resolution \glspl*{tpm}, from which
+\gls*{dmri} signal can be simulated with different
+\glspl*{snr}. The new version of the phantom includes
+a \gls*{csf} distribution, which was not simulated
+in the version available for the contest. For this paper,
+with these new \glspl*{tpm} we produced \gls*{dti} at
+typical low-resolution, and corresponding \gls*{t1} and 
+\gls*{t2} images at high-resolution. All the
+modalities where numerically simulated, using previously
+reported \gls*{mri} typical parameters.
 
-First, confirm that you have the correct template for your paper size. This template has been tailored for output on the US-letter paper size. 
-It may be used for A4 paper size if the paper size setting is suitably modified.
+\begin{table}[h]
+\caption{Simulated \gls*{mri} modalities}
+\label{table:phantom}
+\begin{center}
+\begin{tabular}{c|c}
+\hline
+Modality & Descrition \\
+\hline
+\gls*{dti} & 30 directions, \\
+\hline
+\gls*{t1} & \\
+\hline
+\gls*{t2} & \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 
-\subsection{Maintaining the Integrity of the Specifications}
+\subsection{Theory-based synthetic distortion}
 
-The template is used to format your paper and style the text.
-All margins, column widths, line spaces, and text fonts are prescribed; please do not alter them.
-You may note peculiarities.
-For example, the head margin in this template measures proportionately more than is customary.
-This measurement and others are deliberate, using specifications that anticipate your paper as one part of the entire proceedings, and not as an independent document.
-Please do not revise any of the current designations
+   \begin{figure}[thpb]
+      \centering
+      \framebox{\parbox{3in}{Figure of the phantom}}
+      %\includegraphics[scale=1.0]{figurefile}
+      \caption{Original and distorted digital phantom}
+      \label{fig:label}
+   \end{figure}
+
+\subsection{Evaluation workflows}

### 2013-09-18 Publications/2013-ISBI14/02-introduction.tex
Fixes (minor)

@@
 have been studied \cite{zeng_image_2002,wu_comparison_2008},
-the lack of a gold-standard limit the possible benchmarking 
+the lack of a gold-standard limits the possible benchmarking 
 strategies. \cite{irfanoglu_effects_2012} opened the question

### 2013-09-18 Scripts/workflows/dmri.py
Added t2_registration_workflow
Now, I'm to check that it works

@@ -4,42 +4,84 @@
 
 import os.path as op
 
-import nipype.interfaces.io as nio           # Data i/o
-import nipype.interfaces.utility as niu      # utility
-import nipype.pipeline.engine as pe          # pypeline engine
-import nipype.interfaces.fsl as fsl          # fsl
-import nipype.interfaces.freesurfer as fs    # freesurfer
+import nipype.interfaces.io as nio              # Data i/o
+import nipype.interfaces.utility as niu         # utility
+import nipype.pipeline.engine as pe             # pypeline engine
+import nipype.interfaces.fsl as fsl             # fsl
+import nipype.interfaces.freesurfer as fs       # freesurfer
+import nipype.interfaces.ants as ants           # ANTS
+
+
 import nipype.interfaces.diffusion_toolkit as dtk
 import nipype.interfaces.mrtrix as mrtrix    #<---- The important new part!
 import nipype.interfaces.camino as camino
 
 
-def t2_registation_correct( name="T2_Registration" ):
+def t2_registration_correct( name="T2_Registration" ):
     pipeline = pe.Workflow( name=name )
     
     inputnode = pe.Node( niu.IdentityInterface( fields=['in_file',
                                                         'in_t2',
-                                                        'in_mask']
+                                                        'in_mask_dwi',
+                                                        'in_mask_t2']
                          ), name='inputnode' )
 
     outputnode = pe.Node( niu.IdentityInterface( fields=['epi_corrected']),
                           name='outputnode' )
 
-    get_b0 = pe.Node(fsl.utils.ExtractROI(
-        t_size=1, t_min=0), name="getB0")
+    get_b0 = pe.Node(fsl.ExtractROI( t_size=1, t_min=0 ), name="getB0")
 
-    registration = pe.Node( ANTS , name='B0-to-T2' )
+    mask_t2 = pe.Node( fsl.ApplyMask(), name='maskT2' )
+    mask_b0 = pe.Node( fsl.ApplyMask(), name='maskB0' )
 
-    dwi_split = pe.Node(niu.Function(input_names=['in_file'], output_names=['out_files'], function=_split_dwi), name='split_DWI')
-    applytfm = pe.MapNode( , iterfield=[''], name='AntsInverseTf')
[…]

### 2013-09-18 Scripts/workflows/dmri.py
T2 registration workflow with jacobian modulation
Added also a hd mask

@@ -43,8 +43,8 @@ def t2_registration_correct( name="T2_Registration" ):
     reg.inputs.write_composite_transform = True
 
     reg.inputs.metric = ['Mattes']
-    reg.inputs.metric_weight = [[1]] # Default (value ignored currently by ANTs)
-    reg.inputs.radius_or_number_of_bins = [64]
+    reg.inputs.metric_weight = [1] # Default (value ignored currently by ANTs)
+#    reg.inputs.radius_or_number_of_bins = 64
     reg.inputs.sampling_strategy = ['Random']
     reg.inputs.sampling_percentage = [0.1]
     reg.inputs.convergence_threshold = [1.e-9]
@@ -57,11 +57,12 @@ def t2_registration_correct( name="T2_Registration" ):
 #    reg.inputs.output_warped_image = 'output_warped_image.nii.gz'
 #    reg.inputs.output_transform_prefix = "output_"
 #    reg.inputs.collapse_output_transforms = False
-    applytfm = pe.Node( ants.WarpTimeSeriesImageMultiTransform(), name='ApplyTfm' )
 
-#    dwi_split = pe.Node(niu.Function(input_names=['in_file'], output_names=['out_files'], function=_split_dwi), name='split_DWI')
-#    applytfm = pe.MapNode( , iterfield=[''], name='AntsInverseTf')
-#    dwi_merge = pe.Node(fsl.Merge(dimension='t'), name='merge_DWI')
+    applytfm = pe.Node( ants.WarpTimeSeriesImageMultiTransform(), name='ApplyTfm' )
+    jacobian = pe.Node( ants.JacobianDeterminant(dimension=3,use_log=0), name='Jacobian' )
+    dwi_split = pe.Node(niu.Function(input_names=['in_file'], output_names=['out_files'], function=_split_dwi), name='split_DWI')
+    applyjacobian = pe.MapNode( ants.MultiplyImages(dimension=3), iterfield=['first_input'], name='ApplyJacobian')
+    dwi_merge = pe.Node(fsl.Merge(dimension='t'), name='merge_DWI')
 
 
 
@@ -76,7 +77,12 @@ def t2_registration_correct( name="T2_Registration" ):
                         ,(mask_b0,           reg, [ ( ('out_file',_aslist),'moving_image') ])
                         ,(inputnode,    applytfm, [ ('in_file','input_image'),('in_t2','reference_image') ])
                         ,(reg,          applytfm, [ ('forward_transforms','transformation_series') ])
[…]

### 2013-09-19 Publications/2013-ISBI14/01-abstract.tex
Added figure 1. Methods written down
Advancing with the ISBI 2014 paper

@@
 Several retrospective correction techniques are readily
-available, generally inherited from \gls*{fmri} literature,
-as those data are usually acquired with \gls*{epi} as well.
-These correction methodologies have been evaluated and 
-compared in physical phantoms designed for \gls*{fmri} 
-experiments, but little is known about their impact
-on \gls*{dmri} and more precisely on the resulting 
-tractography.
-
-In this work, we use a renowned digital phantom designed
-for assessing tractography algorithms, and subject it to
-a ``theoretically correct'' and plausible deformation. We
-corrected data back, with several available techniques 
-(namely fieldmap-based, reversed encoding-based, 
-registration-based). Finally, we ranked the methods 
-based on the geometrical accuracy, the dropout compensation, 
-and the impact on the resulting tractography.
+available. In this work, we use a renowned digital phantom 
+designed for evaluating tractography. We subject the phantom 
+to a ``theoretically correct'' and plausible deformation that
+resembles the artifact under investigation. We correct data back, 
+with three standard methodologies (namely fieldmap-based, 
+reversed encoding-based, and registration-based). 
+Finally, we rank the methods based on their geometrical 
+accuracy, the dropout compensation, and their impact on 
+the resulting tractography.
 

### 2013-09-19 Publications/2013-ISBI14/02-introduction.tex
Added figure 1. Methods written down
Advancing with the ISBI 2014 paper

@@
 generally applying \gls*{epi} sequences.
-After a signal reconstruction step, tractography
-algorithms draw a map of sampled structure.
+After a signal reconstruction step 
+(i.e. \cite{basser_microstructural_2011}), 
+tractography algorithms draw a map of the sampled 
+structures.
 These maps can represent the actual trajectories
-of fiber bundles (deterministic tractography,
-\cite{basser_microstructural_2011}) or pixel-wise
-probability of connection to a certain origin
+of fiber bundles (deterministic tractography) or
+pixel-wise probability of connection to a certain origin
 (probabilistic tractography). Finally, the
@@
 In order to enable reliable and robust connectome
-analyses of the brain, these challenging needs for
-solutions must be addressed. One prominent source
+analyses of the brain, some weak processing points must
+be improved. One prominent source
 of inaccuracy are the susceptibility-derived artifacts,
@@
 air surrounding sinuses). This artifact has been
-profoundly described, and a number of methodologies
-have been proposed. First approaches to
+profoundly described, generally within \gls*{fmri}
+literature because they are usually acquired with 
+\gls*{epi} as \gls*{dmri}.
+
+First approaches to
 retrospective correction \cite{jezzard_correction_1995},
@@
 the lack of a gold-standard limits the possible benchmarking 
-strategies. \cite{irfanoglu_effects_2012} opened the question
-of the impact of the distortions on tractography. In this work,
-we propose an evaluation framework using a detailed digital 
-phantom designed for tractography. Using this framework,
+strategies. \cite{irfanoglu_effects_2012} raised the question
+of distortion-derived impacts in tractography.
+In this work, we propose an evaluation framework
+using a detailed digital phantom designed for tractography.
+Using this framework,
 we are able to compare several correction techniques and we

### 2013-09-19 Publications/2013-ISBI14/03-methods.tex
Added figure 1. Methods written down
Advancing with the ISBI 2014 paper

@@
 \hline
-Modality & Descrition \\
+Modality & Description \\
 \hline
@@
 
+Fieldmap-based correction methodologies
+\cite{jezzard_correction_1995} use a map
+of the field in the scanner. More precisely, the
+phase difference between two subsequent samplings
+of the fieldmap. With that information, it is possible
+to compute the theoretical displacements that each
+voxel is to suffer, the so-called \gls*{vsm}. The
+most prominent feature of this \gls*{vsm} is that all
+the shifts have the same orientation (parallel to the
+phase-encoding direction of the \gls*{epi}) and their
+magnitude and direction depend on the \gls*{epi} 
+gradient increments (or \emph{blips}), and the actual
+phase difference at the voxel.
+
+In order to create a realistic distortion, we started
+by generating a synthetic phase-difference map 
+consistent with the phantom.  We defined two regions 
+of smooth phase inhomogeneity and computed the 
+corresponding \gls*{vsm} using the tools distributed 
+with the \texttt{FSL} package \cite{jenkinson_fsl_2012}.
+
+Then, it is straightforward to use the \gls*{vsm} to 
+produce the deformation of the original phantom data,
+using the \texttt{fugue} tool included in
+\cite{jenkinson_fsl_2012}. With the same procedure,
+just setting the opposite encoding direction, we obtained
+the corresponding reversed encoding ``acquisition'' of
+the same phantom, in order to use it in evaluating the
+second family of correction methods.
+
+In this simple manner, we generated a full gold-standard
+protocol, containing realistic \gls*{t1} and \gls*{t2}
+at high resolution, \gls*{dti} data acquired in two
+different phase encoding directions, and the ground-truth
+\gls*{dti} data that is not available for real datasets.
+
 \begin{figure}[thpb]
    \centering
-      \framebox{\parbox{3in}{Figure of the phantom}}
-      %\includegraphics[scale=1.0]{figurefile}
-      \caption{Original and distorted digital phantom}
+   \includegraphics[width=0.8\columnwidth]{Fig01-Phantom}
+   \caption{Original and distorted digital phantom (sagittal
+   planes). A) T2-weighted image; B) undistorted B0 volume;
+   C and D) distorted B0 volumes with opposed phase encoding 
+   directions.}
    \label{fig:label}
@@
 
-\subsection{Evaluation workflows}
+\subsection{Correction methods}
+
+\paragraph{\Gls*{fmb}} As it was previously described,
+this family of correction technologies infer a \gls*{vsm}
+from the phase difference map. In this work, we use the
+tool \texttt{fugue} \cite{jenkinson_fsl_2012} aforementioned.
+
+\paragraph{\Gls*{reb}} We use the tool
+\texttt{topup} from \cite{jenkinson_fsl_2012} to correct
+using the B0 of a reverse-encoding simulation of the \gls*{dti}.
+In this second case, the \gls*{vsm} is inferred from the
+differences between the two corresponding B0. This technique
+also uses least-squares-fitting to interpolate the undistorted
+signal, and therefore provides with a robust dropout correction.
+
+\paragraph{\Gls*{t2b}} For this last methodology
+we fine-tuned the \texttt{ANTs} tool \cite{avants_ants:_2013},
+in order that registering the B0 to the T2. Additionally,
+undistorted images are corrected for dropout using the determinant
+of the Jacobian of the deformation field.
+
+\subsection{Evaluation}
+
+\paragraph*{Proposed workflows}
+We used \texttt{nipype} \cite{gorgolewski_nipype:_2011}
+to conveniently integrate the correction methods in
+comprehensive evaluation workflows:
+
+\begin{itemize}
+\item \textbf{Phantom deformation}: We release the pipeline
+that generates the gold-standard, along with
+distorted versions of the \glspl*{tpm} for
+evaluation purposes.
+\item \textbf{Correction workflow}: A high-level pipeline
+integrates the three proposed correction methods,
+ensuring that all the requirements for inputs are
+met. As outputs this workflow produces the different
+corrected versions of the phantom (\gls*{fmb}, \gls*{reb}
+and \gls*{t2b}) alongside corrected versions of the
+distorted \glspl*{tpm}.
+\item \textbf{Reconstruction workflow}: This pipeline uses
+the Diffusion Toolkit \cite{wang_diffusion_2007}, 
+to produce deterministic tractography
+from \gls*{dmri} data at the input. Therefore,
+tractographies of original, and recovered data
+are produced for comparison.
+\item \textbf{Benchmarking workflow}: The final processing
+element collects all the outputs from the previous
+workflows and extracts the evaluation scores.
+\end{itemize}
+General-purpose 
+workflows have been contributed to the upstream code
+of the tool, under the branch \texttt{nipype.workflows.dmri}.
+Testing-specific workflows are also publicly released.
+
+\paragraph*{Evaluation scores}
+In order to assess the geometrical correctness of the
+recovered phantoms, we report overlap indices, specifically
+the fuzzy Jaccard index as proposed in
+\cite{crum_generalized_2006} weighted by the volumes of
+labels. The recovered signal \emph{sanity} (dropout 
+correction) is evaluated with correlation maps, averaged
+along the time axis to take into account all the volumes
+existing on the \gls*{dmri} data. This checkpoint
+assumes that geometrical distortion is precisely
+corrected. Finally, we report the impact on tractography
+using the \gls*{frr}, that informs the number of recovered
+fibers versus the gold-standard number for a specific
+reconstruction method, and the error on the resulting
+connectivity matrices.

### 2013-09-19 Publications/2013-ISBI14/04-results.tex
Added figure 1. Methods written down
Advancing with the ISBI 2014 paper

@@
-\section{MATH}
+\section{RESULTS AND DISCUSSION}
 
-Before you begin to format your paper, first write and save the content as a separate text file.
-Keep your text and graphic files separate until after the text has been formatted and styled.
-Do not use hard tabs, and limit use of hard returns to only one return at the end of a paragraph.
-Do not add any kind of pagination anywhere in the paper.
-Do not number text heads-the template will do that for you.
-
-Finally, complete content and organizational editing before formatting. Please take note of the following items when proofreading spelling and grammar:
-
-\subsection{Abbreviations and Acronyms} Define abbreviations and acronyms the first time they are used in the text, even after they have been defined in the abstract.
-Abbreviations such as IEEE, SI, MKS, CGS, sc, dc, and rms do not have to be defined.
-Do not use abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
-
-\begin{itemize}
-
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as Ò3.5-inch disk driveÓ.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ÒWb/m2Ó or Òwebers per square meterÓ, not Òwebers/m2Ó.
-Spell out units when they appear in text: Ò.
-.
-.
-a few henriesÓ, not Ò.
-.
-.
-a few HÓ.
-\item Use a zero before decimal points: Ò0.25Ó, not Ò.25Ó. Use Òcm3Ó, not ÒccÓ. (bullet list)
-
-\end{itemize}
-
-\subsection{Equations}
-
-The equations are an exception to the prescribed specifications of this template.
-You will need to determine whether or not your equation should be typed using either the Times New Roman or the Symbol font (please no other font).
-To create multileveled equations, it may be necessary to treat the equation as a graphic and insert it into the text after your paper is styled.
-Number equations consecutively.
-Equation numbers, within parentheses, are to position flush right, as in (1), using a right tab stop.
-To make your equations more compact, you may use the solidus ( / ), the exp function, or appropriate exponents.
-Italicize Roman symbols for quantities and variables, but not Greek symbols.
-Use a long dash rather than a hyphen for a minus sign.
-Punctuate equations with commas or periods when they are part of a sentence, as in
-
-$$
-\alpha + \beta = \chi \eqno{(1)}
-$$
-
-Note that the equation is centered using a center tab stop.
-Be sure that the symbols in your equation have been defined before or immediately following the equation.
-Use Ò(1)Ó, not ÒEq.
-(1)Ó or Òequation (1)Ó, except at the beginning of a sentence: ÒEquation (1) is .
-.
-.Ó
-
-\subsection{Some Common Mistakes}
-\begin{itemize}
-
-\item The word ÒdataÓ is plural, not singular.
-\item The subscript for the permeability of vacuum ?0, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ÒoÓ.
-\item In American English, commas, semi-/colons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ÒinsetÓ, not an ÒinsertÓ. The word alternatively is preferred to the word ÒalternatelyÓ (unless you really mean something that alternates).
-\item Do not use the word ÒessentiallyÓ to mean ÒapproximatelyÓ or ÒeffectivelyÓ.
-\item In your paper title, if the words Òthat usesÓ can accurately replace the word ÒusingÓ, capitalize the ÒuÓ; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ÒaffectÓ and ÒeffectÓ, ÒcomplementÓ and ÒcomplimentÓ, ÒdiscreetÓ and ÒdiscreteÓ, ÒprincipalÓ and ÒprincipleÓ.
-\item Do not confuse ÒimplyÓ and ÒinferÓ.
-\item The prefix ÒnonÓ is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ÒetÓ in the Latin abbreviation Òet al.Ó.
-\item The abbreviation Òi.e.Ó means Òthat isÓ, and the abbreviation Òe.g.Ó means Òfor exampleÓ.
-
-\end{itemize}
+WIP: work in progess..

### 2013-09-19 Publications/2013-ISBI14/05-conclusions.tex
Added figure 1. Methods written down
Advancing with the ISBI 2014 paper

@@
-\section{USING THE TEMPLATE}
+\section{CONCLUSION}
 
-Use this sample document as your LaTeX source file to create your document.
-Save this file as {\bf root.tex}.
-You have to make sure to use the cls file that came with this distribution.
-If you use a different style file, you cannot expect to get required margins.
-Note also that when you are creating your out PDF file, the source file is only part of the equation.
-{\it Your \TeX\ $\rightarrow$ PDF filter determines the output file size.
-Even if you make all the specifications to output a letter file in the source - if your filter is set to produce A4, you will only get A4 output.
-}
-
-It is impossible to account for all possible situation, one would encounter using \TeX.
-If you are using multiple \TeX\ files you must make sure that the ``MAIN`` source file is called root.tex - this is particularly important if your conference is using PaperPlaza's built in \TeX\ to PDF conversion tool.
-
-\subsection{Headings, etc}
-
-Text heads organize the topics on a relational, hierarchical basis.
-For example, the paper title is the primary text head because all subsequent material relates and elaborates on this one topic.
-If there are two or more sub-topics, the next level head (uppercase Roman numerals) should be used and, conversely, if there are not at least two sub-topics, then no subheads should be introduced.
-Styles named ÒHeading 1Ó, ÒHeading 2Ó, ÒHeading 3Ó, and ÒHeading 4Ó are prescribed.
-
-\subsection{Figures and Tables}
-
-Positioning Figures and Tables: Place figures and tables at the top and bottom of columns.
-Avoid placing them in the middle of columns.
-Large figures and tables may span across both columns.
-Figure captions should be below the figures; table heads should appear above the tables.
-Insert figures and tables after they are cited in the text.
-Use the abbreviation ÒFig.
-1Ó, even at the beginning of a sentence.
-
-\begin{table}[h]
-\caption{An Example of a Table}
-\label{table_example}
-\begin{center}
-\begin{tabular}{|c||c|}
-\hline
-One & Two\\
-\hline
-Three & Four\\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-   \begin{figure}[thpb]
-      \centering
-      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
-}}
-      %\includegraphics[scale=1.0]{figurefile}
-      \caption{Inductance of oscillation winding on amorphous
-       magnetic core versus DC bias magnetic field}
-      \label{figurelabel}
-   \end{figure}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels.
-Use words rather than symbols or abbreviations when writing Figure axis labels to avoid confusing the reader.
-As an example, write the quantity ÒMagnetizationÓ, or ÒMagnetization, MÓ, not just ÒMÓ.
-If including units in the label, present them within parentheses.
-Do not label axes only with units.
-In the example, write ÒMagnetization (A/m)Ó or ÒMagnetization {A[m(1)]}Ó, not just ÒA/mÓ.
-Do not label axes with a ratio of quantities and units.
-For example, write ÒTemperature (K)Ó, not ÒTemperature/K.Ó
+This paper proposes an evaluation framework to
+comprehensively analyze the impact of susceptibility
+induced distortion on tractography from \gls*{dmri}
+data. Inaccuracy on tractography leads to significant
+errors in posterior connectivity analyses of the
+whole-brain, specially in distortion affected regions.
+We publicly release the framework and also contribute
+with the evaluation of three widely-used available
+correction methodologies. We conclude that the 
+\gls*{reb} method resulted the first method in the 
+proposed rankings (geometric correctness, signal loss
+recovery and tractography impact), with the drawback
+of demanding some extra \gls*{dmri} data.

### 2013-09-19 Scripts/workflows/dmri.py
Update registration workflow before merge
T2 registration parameters have changed

@@ -37,8 +37,8 @@ def t2_registration_correct( name="T2_Registration" ):
     reg = pe.Node( ants.Registration() , name='B0-to-T2' )
 
     reg.inputs.transforms = ['SyN']
-    reg.inputs.transform_parameters = [(0.25, 5.0, 10.0)]
-    reg.inputs.number_of_iterations = [[100, 50, 30]]
+    reg.inputs.transform_parameters = [(1.0, 1.5, 6.0) ]
+    reg.inputs.number_of_iterations = [[300,200, 20]]
     reg.inputs.dimension = 3
     reg.inputs.write_composite_transform = True
 
@@ -48,7 +48,7 @@ def t2_registration_correct( name="T2_Registration" ):
     reg.inputs.sampling_strategy = ['Random']
     reg.inputs.sampling_percentage = [0.1]
     reg.inputs.convergence_threshold = [1.e-9]
-    reg.inputs.convergence_window_size = [20]
+    reg.inputs.convergence_window_size = [40,30,20]
     reg.inputs.smoothing_sigmas = [[2,1,0]]
     reg.inputs.sigma_units = ['vox']
     reg.inputs.shrink_factors = [[3,2,1]]

### 2013-09-26 Scripts/workflows/dmri.py
Still working on it...

@@ -167,7 +167,7 @@ def dtk_tractography_workflow( name='DTK_tractography' ):
                           name='outputnode' )
 
     dtifit = pe.Node(dtk.DTIRecon(),name='dtifit')
-    dtk_tracker = pe.Node(dtk.DTITracker(mask1_threshold=0.25,invert_x=False,mask2_threshold=0.25,args='-rseed 10'), name="dtk_tracker")
+    dtk_tracker = pe.Node(dtk.DTITracker(mask1_threshold=0.01,invert_x=False,random_seed=10), name="dtk_tracker")
     smooth_trk = pe.Node(dtk.SplineFilter(step_length=0.5), name="smooth_trk")
     matrix = pe.Node( ConnectivityMatrix(), name='BuildMatrix' )
 
@@ -176,7 +176,7 @@ def dtk_tractography_workflow( name='DTK_tractography' ):
     
     pipeline.connect([
                          (inputnode,        dtifit, [('in_dwi','DWI'),('in_bvec','bvecs'),('in_bval','bvals')])
-                        ,(inputnode,   dtk_tracker, [('in_mask','mask1_file'),('in_mask','mask2_file') ])
+                        ,(inputnode,   dtk_tracker, [('in_mask','mask1_file'),('roi_file','mask_seed') ])
                         ,(dtifit,      dtk_tracker, [('tensor','tensor_file') ])
                         ,(dtk_tracker,  smooth_trk, [('track_file', 'track_file')])
                         ,(dtifit,       outputnode, [('tensor', 'tensor'),('ADC','ADC'),('FA','FA') ])

### 2013-09-26 Scripts/workflows/tracks.py

@@ -100,11 +100,23 @@ class ConnectivityMatrix(BaseInterface):
 
 
 def track_analyze( points, roidata, vox_size=(1.0,1.0,1.0) ):
-    nPoints = np.shape(points)[0]
     startval = 0
     startidx = -1
     vox_size=np.array( vox_size, dtype='f32' )
-    
+    im_idx = np.array( np.shape(roidata) )-1
+
+    invalid_idx = []
+    for i,point in enumerate(points):
+        vox = tuple( ( np.around( point / vox_size ) ).astype( 'uint16' ))
+        if np.any( vox > im_idx ):
+            invalid_idx+= [i]    
+
+    points = np.delete( points, invalid_idx, 0 )    
+    nPoints = np.shape(points)[0]
+
+    if nPoints == 0:
+        return (None,0)
+
     for i,point in enumerate(points):
         vox = tuple( ( np.around( point / vox_size ) ).astype( 'uint16' ))
         val = roidata[vox]
@@ -123,7 +135,6 @@ def track_analyze( points, roidata, vox_size=(1.0,1.0,1.0) ):
             endval=val
             endidx = nPoints-1-i
             break
-            
     
     length = 0.0
     
@@ -152,12 +163,15 @@ def filter_track( trkfile, roidata, trkhdr=None, min_length=3.0 ):
     
     for idx,trk in enumerate(streams):
         nt,tf = track_analyze( trk, roidata, vox_size )
-        tlen = tf[-1]
+
+        if nt is None:
+            continue
        
-        if tf[-1]>min_length and tf[0][0]!=tf[0][1]:
-            filtered_ft.append( tf )
-            new_trk = (np.array(nt), trkfile[idx][1],trkfile[idx][2])
-            out_trk.append(new_trk)
+        if tf[-1]>min_length:
+            if tf[0][0]!=tf[0][1]:
+                filtered_ft.append( tf )
+                new_trk = (np.array(nt), trkfile[idx][1],trkfile[idx][2])
+                out_trk.append(new_trk)
             
     return out_trk, filtered_ft
 

### 2013-09-27 Scripts/workflows/dmri.py
Working ?? ISBI 2014

@@ -166,8 +166,8 @@ def dtk_tractography_workflow( name='DTK_tractography' ):
                                   'ADC','FA','connectome','avglen_map','filter_tracks' ] ),
                           name='outputnode' )
 
-    dtifit = pe.Node(dtk.DTIRecon(),name='dtifit')
-    dtk_tracker = pe.Node(dtk.DTITracker(mask1_threshold=0.01,invert_x=False,random_seed=10), name="dtk_tracker")
+    dtifit = pe.Node(dtk.DTIRecon(output_type='nii.gz'),name='dtifit')
+    dtk_tracker = pe.Node(dtk.DTITracker(mask1_threshold=[0.002,3.0], mask_seed_th=[0.5,27.5],invert_x=False,random_seed=10), name="dtk_tracker")
     smooth_trk = pe.Node(dtk.SplineFilter(step_length=0.5), name="smooth_trk")
     matrix = pe.Node( ConnectivityMatrix(), name='BuildMatrix' )
 

### 2013-09-27 Scripts/workflows/evaluation.py

@@ -48,6 +48,9 @@ def evaluation_workflow( name="Evaluation" ):
 
     sel1 = pe.Node( niu.Select(index=3 ), name='SelectWM' )
     sel2 = pe.Node( niu.Select(index=3 ), name='SelectWM_dist' )
+    sel3 = pe.Node( niu.Select(index=3 ), name='SelectWM_fmap' )
+    sel4 = pe.Node( niu.Select(index=3 ), name='SelectWM_topup' )
+    sel5 = pe.Node( niu.Select(index=3 ), name='SelectWM_t2' )
 
     sim1 = compare_dwis( name="Sim_dist" )
     sim1.inputs.inputnode.name='dist' 
@@ -128,15 +131,18 @@ def evaluation_workflow( name="Evaluation" ):
         ,(normalize_tpms,     ove1, [ ('out_files','inputnode.in_ref') ])
         ,(fmap_wf,            ove1, [ ('outputnode.out_vsm','inputnode.in_vsm') ])
         ,(dist1_wf,           ove1, [ ('outputnode.out_tpms','inputnode.in_tpms' ) ])
+        ,(ove1,               sel3, [ ('outputnode.out_tpms','inlist')])
         ,(inputnode,          ove2, [ ('out_csv','inputnode.out_csv') ])
         ,(normalize_tpms,     ove2, [ ('out_files','inputnode.in_ref') ])
         ,(topup_wf,           ove2, [ ('outputnode.out_topup','inputnode.in_topup'),('outputnode.out_enc_file','inputnode.in_enc_file') ])
         ,(dist1_wf,           ove2, [ ('outputnode.out_tpms','inputnode.in_tpms' ) ])
         ,(dist2_wf,           ove2, [ ('outputnode.out_tpms','inputnode.in_tpms_rev') ])
+        ,(ove2,               sel4, [ ('outputnode.out_tpms','inlist')])
         ,(inputnode,          ove3, [ ('out_csv','inputnode.out_csv') ])
         ,(normalize_tpms,     ove3, [ ('out_files','inputnode.in_ref') ])
         ,(t2reg_wf,           ove3, [ ('outputnode.fwd_tfm','inputnode.in_tfm') ])
         ,(dist1_wf,           ove3, [ ('outputnode.out_tpms','inputnode.in_tpms' ) ])
+        ,(ove3,               sel5, [ ('outputnode.out_tpms','inlist')])
         #
         # Tractographies
         #
@@ -150,17 +156,20 @@ def evaluation_workflow( name="Evaluation" ):
[…]

### 2013-09-27 Publications/2013-ISBI14/03-methods.tex
Paper ISBI

@@
 \paragraph*{Evaluation scores}
-In order to assess the geometrical correctness of the
-recovered phantoms, we report overlap indices, specifically
-the fuzzy Jaccard index as proposed in
-\cite{crum_generalized_2006} weighted by the volumes of
-labels. The recovered signal \emph{sanity} (dropout 
-correction) is evaluated with correlation maps, averaged
-along the time axis to take into account all the volumes
-existing on the \gls*{dmri} data. This checkpoint
-assumes that geometrical distortion is precisely
-corrected. Finally, we report the impact on tractography
-using the \gls*{frr}, that informs the number of recovered
-fibers versus the gold-standard number for a specific
-reconstruction method, and the error on the resulting
-connectivity matrices.
+We evaluated three characteristics of the corrected
+phantom. First, we assessed the geometrical correctness
+and reported overlap indices of three \glspl*{tpm} 
+(namely \gls*{csf},\gls*{wm}, and \gls*{gm}).
+More precisely, we computed the \gls*{fji} as 
+proposed in \cite{crum_generalized_2006} weighted by 
+the volumes of labels. 
+Second, to evaluate the quality of the actual 
+signal \emph{sanity} after recovery (dropout 
+correction), we studied the signal similarity
+frame by frame computing the $\ell_1$-norm
+based correlation index. We report this score
+on the B0 and the average of the remaining 32 
+\glspl*{dwi}. Third, we report the 
+impact on tractography using the \gls*{frr}, that 
+informs the number of recovered fibers versus the 
+gold-standard number for a specific reconstruction
+method.

### 2013-09-27 Publications/2013-ISBI14/04-results.tex
Paper ISBI

@@
 
-WIP: work in progess..
+\subsection{Geometrical correctness}
+
+The three methods under study achieved good results in
+terms of geometrical accuracy. In this ranking, the best
+results were obtained with the \gls*{reb} method. Second
+best were obtained with \gls*{t2b}. We visually checked
+the numbers reported in \autoref{table:results}. 
+\autoref{fig:results} shows the B0 volumes after correction.
+
+\subsection{Signal loss recovery}
+
+A second workflow evaluated the different correction methodologies
+with analysis of the outcome similarity with respect to the
+the original signal. In this second study, \gls*{reb} performed
+significantly better than the other two methods, as reported
+in \autoref{table:results}.
+
+\subsection{Tractography impact}
+
+Tractography was shown to be deeply influenced by distortion and
+the correction method choice. As expected from the scores obtained
+by \gls*{reb} in the signal similarity check, this is also the
+best ranked method in this category.
+
+\subsection{Discussion}
+
+\input{041-results-table}
+
+\begin{figure}[thpb]
+   \centering
+   \includegraphics[width=0.8\columnwidth]{Fig02-Results}
+   \caption{Visual results}
+   \label{fig:results}
+\end{figure}
+
+Even though all the surveyed methods produced visually sound results,
+\gls*{reb} showed the best results for all the tests. Therefore,
+this study suggests that \gls*{reb} is the best available
+methodology for susceptibility-induced distortions in
+\gls*{epi}. An important limitation of this work is that 
+\gls*{psf}-based methods could not be compared, as a
+\gls*{psf} map consistent with the synthetic phase-difference
+map could not be generated. Nonetheless, to our knowledge there are
+no \gls*{psf}-based methods with readily available implementations
+for comparison. Moreover, whereas acquiring one extra dataset
+among \gls*{t2}, fieldmap, or reversed-encoding B0 is included in
+every typical \gls*{dmri} acquisition protocol, the \gls*{psf} 
+mapping is not usually performed. Therefore, the most common
+correction methodologies are included in the proposed evaluation
+framework, along with the benchmarking results.

### 2013-09-27 Publications/2013-ISBI14/02-introduction.tex
ISBI 2014. Almost finished!!

@@
 research topic of high interest for the last
-5 years \cite{jones_studying_2008}. 
+5 years. 
 \Gls*{dmri} can be used to probe the
@@
 generally applying \gls*{epi} sequences.
-After a signal reconstruction step 
-(i.e. \cite{basser_microstructural_2011}), 
+After a signal reconstruction step, 
 tractography algorithms draw a map of the sampled 
@@
 flaws that have been detailedly reported 
-\cite{jones_twenty-five_2010,jbabdi_tractography:_2011}.
+\cite{jones_twenty-five_2010}.
 In order to enable reliable and robust connectome

### 2013-09-28 Publications/2013-ISBI14/03-methods.tex
Adapted to exact ISBI layout

@@
 
-\paragraph{\Gls*{fmb}} As it was previously described,
+\paragraph*{\Gls*{fmb}} As it was previously described,
 this family of correction technologies infer a \gls*{vsm}
@@
 
-\paragraph{\Gls*{reb}} We use the tool
+\paragraph*{\Gls*{reb}} We use the tool
 \texttt{topup} from \cite{jenkinson_fsl_2012} to correct
@@
 
-\paragraph{\Gls*{t2b}} For this last methodology
+\paragraph*{\Gls*{t2b}} For this last methodology
 we fine-tuned the \texttt{ANTs} tool \cite{avants_ants:_2013},

### 2013-09-28 Publications/2013-ISBI14/01-abstract.tex
First review after Manu's comments

@@
 \begin{abstract}
-
-Connectivity analysis on \gls*{dmri} data of the whole-brain
+Connectivity analysis on diffusion MRI data of the whole-brain
 suffers from distortions originated by the standard 
-\gls*{epi} sequences used in acquisition. These artifacts
-are present at interfaces between tissues, due to the inhomogeneity
-of the field derived from the discontinuity of magnetic
-susceptibility. As a result, registered signal shows characteristic
-geometrical deformations and signal destruction. These artifacts
-suppose a significant pitfall that limits the success of tractography 
-algorithms.
+echo-planar imaging sequences used in acquisition. %
+%These artifacts are present at interfaces between tissues,
+%due to the inhomogeneity of the field derived from the 
+% discontinuity of magnetic susceptibility.
+These images show characteristic geometrical deformations
+and signal destruction that suppose a significant pitfall
+limiting the success of tractography algorithms.
 
 Several retrospective correction techniques are readily
-available. In this work, we use a renowned digital phantom 
-designed for evaluating tractography. We subject the phantom 
-to a ``theoretically correct'' and plausible deformation that
-resembles the artifact under investigation. We correct data back, 
-with three standard methodologies (namely fieldmap-based, 
-reversed encoding-based, and registration-based). 
-Finally, we rank the methods based on their geometrical 
-accuracy, the dropout compensation, and their impact on 
-the resulting tractography.
-
-\end{abstract}
+available. In this work, we use a digital phantom designed
+for the evaluation of connectivity pipelines. We subject
+the phantom to a ``theoretically correct'' and plausible
+deformation that resembles the artifact under investigation.
+We correct data back, with three standard methodologies
+(namely fieldmap-based, reversed encoding-based, and
+registration-based). Finally, we rank the methods based
+on their geometrical accuracy, the dropout compensation,
+and their impact on the resulting tractography.
+\end{abstract}%
+\begin{keywords}
+susceptibility artifacts, diffusion MRI, tractography, connectivity.
+\end{keywords}

### 2013-09-28 Publications/2013-ISBI14/02-introduction.tex
First review after Manu's comments

@@
 
-However, such a complex workflow is prone to
-flaws that have been detailedly reported 
-\cite{jones_twenty-five_2010}.
-In order to enable reliable and robust connectome
-analyses of the brain, some weak processing points must
-be improved. One prominent source
-of inaccuracy are the susceptibility-derived artifacts,
+Among all the difficulties that such a complex workflow
+raises \cite{jones_twenty-five_2010}, we will
+address here the susceptibility-derived artifacts,
 for which \Gls*{epi} schemes are specially sensitive.
 As susceptibility changes at tissue interfaces,
-so does the magnetic field. The consequence of the 
-inhomogeneity of the field translates in a highly 
-distorted imaged anatomy and a significant signal 
-destruction of certain regions of the brain 
+so does the magnetic field. This inhomogeneities 
+of the field translate in a highly distorted imaged 
+anatomy and a significant signal  destruction of 
+certain regions of the brain 
 (e.g. the orbitofrontal lobe, for the proximity of the
 air surrounding sinuses). This artifact has been
-profoundly described, generally within \gls*{fmri}
-literature because they are usually acquired with 
+profoundly described, generally within the context of
+functional MRI because it is usually acquired with 
 \gls*{epi} as \gls*{dmri}.
@@
 relied on an extra \gls*{mri} acquisition (so-called
-field mapping), that probes the inhomogeneity of the field.
+\emph{field mapping}), that probes the inhomogeneity of the field.
 A second theory-based breed of methodologies acquire a 
@@
 Finally, the last family of methodologies acquire an 
-extra \gls*{t2} image, and use its anatomical correctness 
+extra T2-weighted image, and use its anatomical correctness 
 as reference to find the deformation map through nonlinear 
@@
 Registration-based methods usually map
-the \gls*{t2} image to the so-called baseline (B0) volume 
-of \gls*{dmri}. The choice of \gls*{t2} is due to the strong
+the T2 image to the so-called baseline (B0) volume 
+of \gls*{dmri}. The choice of T2 is due to the strong
 similarity of intensity distribution with the B0 of \gls*{dmri}
-datasets. More recent works report extensions of the existing
-techniques \cite{andersson_how_2003,holland_efficient_2010,
-andersson_comprehensive_2012}, and combined approaches
-\cite{zaitsev_point_2004}.
+datasets. More recent works report extensions or combined
+approaches of existing techniques
+\cite{andersson_how_2003,zaitsev_point_2004,%
+holland_efficient_2010,andersson_comprehensive_2012}.
 
@@
 the lack of a gold-standard limits the possible benchmarking 
-strategies. \cite{irfanoglu_effects_2012} raised the question
+strategies. Recently, \cite{irfanoglu_effects_2012} raised the question
 of distortion-derived impacts in tractography.
 In this work, we propose an evaluation framework
-using a detailed digital phantom designed for tractography.
-Using this framework,
-we are able to compare several correction techniques and we
-provide results related to their performance in the task
-itself and their impact on subsequent tractography.
+using a digital phantom designed for connectivity assessment.
+This framework enabled us to compare several correction 
+techniques and characterize their geometrical accuracy 
+and the dropout compensation. Finally, we report their 
+impact on subsequent tractography and resulting connectivity
+matrices.

### 2013-09-28 Publications/2013-ISBI14/03-methods.tex
First review after Manu's comments

@@
 \subsection{Digital Phantom}
-We modified the digital phantom provided with the
-\emph{\gls*{hardi} reconstruction challenge} held
-in ISBI 2013, San Francisco, US. This model is
-built upon high-resolution \glspl*{tpm}, from which
-\gls*{dmri} signal can be simulated with different
-\glspl*{snr}. The new version of the phantom includes
-a \gls*{csf} distribution, which was not simulated
-in the version available for the contest. For this paper,
-with these new \glspl*{tpm} we produced \gls*{dti} at
-typical low-resolution, and corresponding \gls*{t1} and 
-\gls*{t2} images at high-resolution. All the
-modalities where numerically simulated, using previously
-reported \gls*{mri} typical parameters.
-
-\begin{table}[h]
-\caption{Simulated \gls*{mri} modalities}
-\label{table:phantom}
-\begin{center}
-\begin{tabular}{c|c}
-\hline
-Modality & Description \\
-\hline
-\gls*{dti} & 30 directions, \\
-\hline
-\gls*{t1} & \\
-\hline
-\gls*{t2} & \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
+Based on the fiber geometries of the digital phantom 
+created for the High angular resolution diffusion imaging 
+reconstruction Challenge held in ISBI 2013 
+(San Francisco, US), we simulated high resolution 
+(0.5mm isotropic) T1 (TE/TR=10/1500ms) and T2 
+(TE/TR=90/5000ms) images, as well as a \gls*{dmri}
+(1.0mm isotropic, $b$=1200, 32 evenly-distributed 
+directions, 1 B0 image) image.
+The phantom includes \gls*{wm} fiber bundles, 
+\gls*{csf} and \gls*{gm}. Diffusion is modeled by a 
+restricted and a hindered compartment, similar to
+\cite{assaf_composite_2005}.
 
@@
 of the fieldmap. With that information, it is possible
-to compute the theoretical displacements that each
+to compute the theoretical displacement that each
 voxel is to suffer, the so-called \gls*{vsm}. The
@@
 
-In order to create a realistic distortion, we started
-by generating a synthetic phase-difference map 
+In order to create a realistic distortion, we
+generated a synthetic phase-difference map 
 consistent with the phantom.  We defined two regions 
-of smooth phase inhomogeneity and computed the 
-corresponding \gls*{vsm} using the tools distributed 
-with the \texttt{FSL} package \cite{jenkinson_fsl_2012}.
+of smooth de-phasing (up to $0.05\pi$ rad. excursions) 
+and computed the corresponding \gls*{vsm}, using the 
+tools distributed with the \texttt{FSL} package 
+\cite{jenkinson_fsl_2012}  and
+standard parameters ($\Delta$TE=2.46ms for the
+fieldmap and \emph{effective dwell time} of 
+0.77ms for the \gls*{epi}).
 
-Then, it is straightforward to use the \gls*{vsm} to 
-produce the deformation of the original phantom data,
-using the \texttt{fugue} tool included in
-\cite{jenkinson_fsl_2012}. With the same procedure,
-just setting the opposite encoding direction, we obtained
-the corresponding reversed encoding ``acquisition'' of
-the same phantom, in order to use it in evaluating the
-second family of correction methods.
+Once a realistic \gls*{vsm} was available, we generated
+the corresponding distorted \glspl*{dwi}, in two opposed
+phase-gradient encoding directions. The second simulated
+``acquisition'' of the same phantom was necessary 
+for evaluating the second family of correction methods.
 
 In this simple manner, we generated a full gold-standard
-protocol, containing realistic \gls*{t1} and \gls*{t2}
-at high resolution, \gls*{dti} data acquired in two
-different phase encoding directions, and the ground-truth
-\gls*{dti} data that is not available for real datasets.
+protocol, containing realistic T1 and T2
+at high resolution, \glspl*{dwi} acquired in two
+different phase encoding directions, and a ground-truth
+\gls*{dwi} data, which is not available for real datasets.
 

### 2013-09-30 Publications/2013-ISBI14/02-introduction.tex
ISBI2014

@@
 
-First approaches to
-retrospective correction \cite{jezzard_correction_1995},
-relied on an extra \gls*{mri} acquisition (so-called
-\emph{field mapping}), that probes the inhomogeneity of the field.
+\Gls*{fmb} methods supposed the first approach to retrospective
+correction \cite{jezzard_correction_1995}. They rely on one
+extra acquisition (\emph{field mapping}), that probes the 
+inhomogeneity of the field.
 A second theory-based breed of methodologies acquire a 
-map of the \gls*{psf} of the \gls*{epi} readouts to correct 
+map of the \acrlong*{psf} of the \gls*{epi} readouts to correct 
 for the artifact \cite{robson_measurement_1997}. 
@@
 with specific differences that enable correction, for instance
-the opposed direction of the phase encoding gradients 
+the opposed direction of the phase encoding gradients. We will
+refer to these methods as \gls*{reb}.
 Finally, the last family of methodologies acquire an 
@@
 registration \cite{kybic_unwarping_2000,studholme_accurate_2000}.
-Registration-based methods usually map
-the T2 image to the so-called baseline (B0) volume 
+\gls*{t2b} methods usually map
+the T2 image to the \emph{baseline} volume or ``B0''
 of \gls*{dmri}. The choice of T2 is due to the strong
-similarity of intensity distribution with the B0 of \gls*{dmri}
-datasets. More recent works report extensions or combined
+similarity of intensity distribution with respect the B0.
+More recent works report extensions or combined
 approaches of existing techniques

### 2013-09-30 Publications/2013-ISBI14/03-methods.tex

@@
 
-Fieldmap-based correction methodologies
-\cite{jezzard_correction_1995} use a map
+\Gls*{fmb} methodologies use a map
 of the field in the scanner. More precisely, the
@@
 In this simple manner, we generated a full gold-standard
-protocol, containing realistic T1 and T2
+containing realistic T1 and T2
 at high resolution, \glspl*{dwi} acquired in two

### 2013-09-30 Publications/2013-ISBI14/04-results.tex

--- 2013-09-19 Publications/2013-ISBI14/04-results.tex
@@
 
-WIP: work in progess..
+\input{041-results-table}
+
+\begin{figure*}[tpb]
+   \centering
+   \includegraphics[width=0.9\textwidth]{Fig02-Results}
+   \caption{The three major stages of the evaluation framework,
+   along with the associated visual results. First row represents
+   a sagittal section of the B0 volume. In second row, the outcome
+   of tractography, showing only tracks that connect two different
+   network nodes. Third row shows the connectivity matrix associated
+   to the corresponding tractography in second row. The }
+   \label{fig:results}
+\end{figure*}
+
+The results of the proposed experiments are summarized
+in \autoref{fig:results} and \autoref{table:results}.
+The remaining of this section provides extended descriptions
+and interpretation of results.
+
+\subsection{Geometrical correctness}
+
+The three methods under study achieved good results in
+terms of geometrical accuracy, as it is shown in the first
+row of \autoref{fig:results}. The best numerical scores
+were obtained with the \gls*{reb} method (\autoref{table:results}),
+that achieved a weighted average overlap of $96.64\%$. Second
+best were obtained with \gls*{fmb}. Geometrical correctness
+is fundamental in connectivity analysis to spatially locate the
+\glspl*{roi} that will define the nodes of the final connectivity
+matrix. The clear difference of accuracy between \gls*{reb} and \gls*{fmb}
+with respect the \gls*{t2b} might point out that the latter is
+not an appropriate method for susceptibility correction.
+
+\subsection{Signal loss recovery}
+
+A second workflow evaluated the different correction 
+methodologies to show the similarity of the recovered
+signal with respect the original (undistorted) one. 
+In this second study, \gls*{reb} performed significantly
+better than the other two methods, as reported
+in \autoref{table:results}. \gls*{reb} scored a $91.0\%$
+similarity index for the B0 volume and an average $97.65\%$
+for the remaining 31 \glspl*{dwi}. Again, the second
+qualified was \gls*{fmb}, which achieved very close results
+for the \glspl*{dwi} ($96.26\%$) but not as good for the B0
+volume. Visual inspection of the recovered data confirm the 
+results presented in the Table (\autoref{fig:results}, first
+row). Finally, \gls*{t2b} confirmed that its correction
+can not be compared, even with using intensity correction with
+the determinant of the deformation field.
+
+\subsection{Impact on tractography and connectivity}
+
+The final objective of the present work was to characterize
+in a reproducible manner the effects of susceptibility distortions
+on the connectivity analysis. The gold-standard connectivity
+matrix contained a total of 1888 tracks, with an average
+length of $40.88\pm8.06$mm. After distortion, this numbers
+changed to a total of 1855 tracks and a characteristic length
+of $40.64\pm8.12$mm. Therefore, very specific tracks were affected
+by distortion, as it was intended on the experiment. These differences
+are highlighted in the second row of \autoref{fig:results}. Two
+main effects can be expected from susceptibility distortion,
+depending on the intensity of the deformation. Those areas exposed
+to a slight deformation showed an increase of variability when
+the signal is not orthogonal to the phase encoding axis 
+(\autoref{fig:results}, second row, A) or an
+increase of directional anisotropy of diffusion
+(\autoref{fig:results}, second row, B), when the
+diffusivity is orthogonal to the distortion.
+Finally, regions subjected to severe susceptibility-induced 
+distortion, some tracks were lost (\autoref{fig:results},
+second row, C).
+
+After \gls*{fmb} correction, a total of 1905 tracks 
+of $40.53\pm8.07$mm. were obtained (\gls*{frr}=$100.9\%$). 
+For \gls*{reb} these figures are 2283 tracks (\gls*{frr}=$120.93\%$) 
+and $41.15\pm7.60$mm. 
+length. Finally, \gls*{t2b} yielded 2495 (\gls*{frr}=$132.1\%$) tracks and 
+$40.60\pm7.81$mm. Although these results pointed at
+\gls*{fmb} as the best method, the visual inspection
+of tractography suggested that again, only \gls*{reb}
+was successful recovering lost connections and reducing
+variability of tracks over impacted regions.
+
+\subsection{Discussion}
+
+Even though all the surveyed methods produced visually sound results,
+our study suggested that \gls*{reb} is the most accurate
+methodology for susceptibility-induced distortions in
+\gls*{epi}. Nonetheless, \gls*{fmb} achieved very accurate
+scores, with better results on tractography as compared to
+\gls*{reb}. Clearly, the \gls*{t2b} method did not achieve
+the necessary high-standards to recommend its use.
+We recall that this last conclusion is potentially biased 
+by the gold-standard generation
+methodology, as it also used \gls*{fmb}.
+A second limitation of this work is that \gls*{psf}-based methods 
+could not be included on comparisons for two reasons.
+First, the necessary \gls*{psf} map consistent with the synthetic
+phase-difference map was not available. Second, to our knowledge,
+there are no \gls*{psf}-based methods with readily available 
+implementations we could have used for comparisons. 
+In practice, this second limitation is not relevant because
+typical acquisition protocols usually include one or more
+datasets among field mapping, reversed-encoding B0 or \gls*{t2},
+but they do not include the necessary \gls*{psf} mapping.
+
+Therefore, this work reviewed the most common
+correction methodologies using a new evaluation
+framework, and reporting the benchmark results.

### 2013-09-30 Publications/2013-ISBI14/05-conclusions.tex

@@
 induced distortion on tractography from \gls*{dmri}
-data. Inaccuracy on tractography leads to significant
-errors in posterior connectivity analyses of the
-whole-brain, specially in distortion affected regions.
+data. Inaccuracy on tractography produce errors
+and increased variability in subsequent connectivity 
+analyses of the whole-brain. Specially, in regions 
+influenced by  significant distortion.
 We publicly release the framework and also contribute
-with the evaluation of three widely-used available
-correction methodologies. We conclude that the 
-\gls*{reb} method resulted the first method in the 
-proposed rankings (geometric correctness, signal loss
-recovery and tractography impact), with the drawback
-of demanding some extra \gls*{dmri} data.
+with the evaluation of the most widely-used and
+readily-available correction methodologies. 
+This work shows that \gls*{dmri} should be corrected
+for susceptibility-induced artifacts prior to tractography
+and connectivity analyses. In that sense, the \gls*{reb} 
+method was the best classified
+in two of the three rankings (geometric 
+correctness and signal recovery), whereas \gls*{fmb}
+achieved the lowest impact on the connectivity matrix
+extracted from tractography.

### 2013-09-30 Publications/2013-ISBI14/01-abstract.tex
Kieran's comments and some other improvements

@@
 These images show characteristic geometrical deformations
-and signal destruction that suppose a significant pitfall
+and signal destruction that are a significant pitfall
 limiting the success of tractography algorithms.

### 2013-09-30 Publications/2013-ISBI14/02-introduction.tex
Kieran's comments and some other improvements

@@
 \section{INTRODUCTION}
-
+\label{sec:intro}
 In-vivo whole-brain connectivity analysis has been a
@@
 orientation of fiber bundles within the brain,
-generally applying \gls*{epi} sequences.
+and is generally acquired with an \gls*{epi} sequence.
 After a signal reconstruction step, 
@@
 address here the susceptibility-derived artifacts,
-for which \Gls*{epi} schemes are specially sensitive.
-As susceptibility changes at tissue interfaces,
-so does the magnetic field. This inhomogeneities 
-of the field translate in a highly distorted imaged 
+for which \Gls*{epi} schemes are especially sensitive.
+Magnetic susceptibility disturbs the magnetic field 
+close to tissue interfaces. This inhomogeneity 
+of the field translates in a highly distorted imaged 
 anatomy and a significant signal destruction of 
@@
 profoundly described, generally within the context of
-functional MRI because it is usually acquired with 
-\gls*{epi} as \gls*{dmri}.
+functional MRI because they are also acquired with 
+\gls*{epi}.
 
@@
 extra acquisition (\emph{field mapping}), that probes the 
-inhomogeneity of the field.
+inhomogeneity of the B0 field.
 A second theory-based breed of methodologies acquire a 
@@
 \gls*{t2b} methods usually map
-the T2 image to the \emph{baseline} volume or ``B0''
+the T2 image to the \emph{baseline} volume or \textit{b0}
 of \gls*{dmri}. The choice of T2 is due to the strong
-similarity of intensity distribution with respect the B0.
+similarity of intensity distribution with respect the \textit{b0}.
 More recent works report extensions or combined

### 2013-09-30 Publications/2013-ISBI14/03-methods.tex
Kieran's comments and some other improvements

@@
 (1.0mm isotropic, $b$=1200, 32 evenly-distributed 
-directions, 1 B0 image) image.
+directions, 1 \textit{b0} image) image.
 The phantom includes \gls*{wm} fiber bundles, 
@@
 \subsection{Theory-based synthetic distortion}
-
+\label{sec:distortion}
 \Gls*{fmb} methodologies use a map
@@
 to compute the theoretical displacement that each
-voxel is to suffer, the so-called \gls*{vsm}. The
+voxel undergoes, the so-called \gls*{vsm}. The
 most prominent feature of this \gls*{vsm} is that all
@@
 and computed the corresponding \gls*{vsm}, using the 
-tools distributed with the \texttt{FSL} package 
+tools distributed with the \emph{FSL} package 
 \cite{jenkinson_fsl_2012}  and
@@
    \centering
-   \includegraphics[width=0.8\columnwidth]{Fig01-Phantom}
-   \caption{Original and distorted digital phantom (sagittal
-   planes). A) T2-weighted image; B) undistorted B0 volume;
-   C and D) distorted B0 volumes with opposed phase encoding 
+   \includegraphics[width=0.95\columnwidth]{Fig01-Phantom}
+   \caption{Original and distorted digital phantom.
+   A) T2-weighted image; B) undistorted \textit{b0} volume;
+   C, D) distorted \textit{b0} volumes with opposed phase encoding 
    directions.}
@@
 
-\paragraph*{\Gls*{fmb}} As it was previously described,
-this family of correction technologies infer a \gls*{vsm}
-from the phase difference map. In this work, we use the
-tool \texttt{fugue} \cite{jenkinson_fsl_2012} aforementioned.
+Three out of four methods presented in \autoref{sec:intro}
+were tested on the evaluation framework.
 
-\paragraph*{\Gls*{reb}} We use the tool
-\texttt{topup} from \cite{jenkinson_fsl_2012} to correct
-using the B0 of a reverse-encoding simulation of the \gls*{dti}.
-In this second case, the \gls*{vsm} is inferred from the
-differences between the two corresponding B0. This technique
-also uses least-squares-fitting to interpolate the undistorted
-signal, and therefore provides with a robust dropout correction.
+\paragraph*{\Gls*{fmb}} In this case, the pipeline essentially
+the same as the one used for generating the distortion, but
+applying the \gls*{vsm} in the opposite direction.
 
-\paragraph*{\Gls*{t2b}} For this last methodology
-we fine-tuned the \texttt{ANTs} tool \cite{avants_ants:_2013},
-in order that registering the B0 to the T2. Additionally,
-undistorted images are corrected for dropout using the determinant
+\paragraph*{\Gls*{reb}} We made use of the implementation found
+in \emph{FSL} (called \texttt{topup}) to correct using the \textit{b0} of a
+reverse-encoding simulation of the \glspl*{dwi}. In this second
+case, a \gls*{vsm} is inferred from the differences between the 
+two corresponding \textit{b0}.
+
+\paragraph*{\Gls*{t2b}} For the last surveyed method,
+we fine-tuned \emph{ANTs} \cite{avants_ants:_2013},
+in order to register \textit{b0} to T2. To this end, we used a
+multi-resolution scheme with 3 levels of subsampling and smoothing,
+mutual information metric, and symmetric diffeomorphic transform 
+(\texttt{SyN}). Several configurations of kernel widths for the 
+regularization smoothers were tested, and finally selected 
+0.5/1.0 voxels (gradient/deformation fields, respectively) 
+for its best result. Additionally, undistorted images are 
+corrected for dropout using the determinant
 of the Jacobian of the deformation field.
@@
 
-\paragraph*{Proposed workflows}
-We used \texttt{nipype} \cite{gorgolewski_nipype:_2011}
-to conveniently integrate the correction methods in
-comprehensive evaluation workflows:
+The four modules described so far (namely, phantom
+deformation, \gls*{fmb} correction, \gls*{reb}
+correction and \gls*{t2b} correction) were conveniently
+integrated in second order workflows using
+\emph{nipype} \cite{gorgolewski_nipype:_2011}.
+The choice of this tool grants the reproducibility
+of the experiments, among other advantages.
 
-\begin{itemize}
-\item \textbf{Phantom deformation}: We release the pipeline
-that generates the gold-standard, along with
-distorted versions of the \glspl*{tpm} for
-evaluation purposes.
-\item \textbf{Correction workflow}: A high-level pipeline
-integrates the three proposed correction methods,
-ensuring that all the requirements for inputs are
-met. As outputs this workflow produces the different
-corrected versions of the phantom (\gls*{fmb}, \gls*{reb}
-and \gls*{t2b}) alongside corrected versions of the
-distorted \glspl*{tpm}.
-\item \textbf{Reconstruction workflow}: This pipeline uses
-the Diffusion Toolkit \cite{wang_diffusion_2007}, 
-to produce deterministic tractography
-from \gls*{dmri} data at the input. Therefore,
-tractographies of original, and recovered data
-are produced for comparison.
-\item \textbf{Benchmarking workflow}: The final processing
-element collects all the outputs from the previous
-workflows and extracts the evaluation scores.
-\end{itemize}
-General-purpose 
-workflows have been contributed to the upstream code
-of the tool, under the branch \texttt{nipype.workflows.dmri}.
-Testing-specific workflows are also publicly released.
+In order to enable the analysis of the influence of
+distortions on the final connectivity analyses,
+a intermediate step was also implemented to
+provide with deterministic tractography,
+using the \emph{Diffusion Toolkit} \cite{wang_diffusion_2007}.
+The original phantom, one distorted version, and
+the corrected instances are fed to this module.
+Additionally, the original tissue probability maps
+are also distorted and corrected, for computing overlap
+indices.
+Each tractography was restricted by its corresponding 
+\gls*{wm} mask (e.g. the distorted and corrected with
+fieldmap for the tractography that followed this method). 
+In all cases, we used 10 random seeds at each seeding
+voxel and the seeding voxels were taken from the 
+ground-truth, mimicking the procedure with real data
+(in which this regions are typically projected from the
+anatomically correct T1).
 
-\paragraph*{Evaluation scores}
-We evaluated three characteristics of the corrected
-phantom. First, we assessed the geometrical correctness
-and reported overlap indices of three \glspl*{tpm} 
-(namely \gls*{csf},\gls*{wm}, and \gls*{gm}).
-More precisely, we computed the \gls*{fji} as 
-proposed in \cite{crum_generalized_2006} weighted by 
-the volumes of labels. 
+Finally, the proposed workflows are connected to performance
+assessment modules. We evaluated three characteristics
+of the correction methods. 
+First, we assessed the geometrical correctness
+and reported overlap indices of three tissue
+probability maps (namely \gls*{csf}, \gls*{wm},
+and \gls*{gm}),
+weighting the average by the volumes of labels.
 Second, to evaluate the quality of the actual 
-signal \emph{sanity} after recovery (dropout 
-correction), we studied the signal similarity
-frame by frame computing the $\ell_1$-norm
-based correlation index. We report this score
-on the B0 and the average of the remaining 32 
-\glspl*{dwi}. Third, we report the 
-impact on tractography using the \gls*{frr}, that 
-informs the number of recovered fibers versus the 
-gold-standard number for a specific reconstruction
-method.
+signal dropout correction, we studied the 
+signal similarity frame by frame computing the $\ell_1$-norm
+correlation index. We report this score
+on the \textit{b0} and the average of the remaining 32 
+\glspl*{dwi}. Third, we studied the impact on the
+connectivity matrices reporting the number of
+false positives (inexistent connections in the
+gold-standard) and false negatives (or connections
+lost).

### 2013-09-30 Publications/2013-ISBI14/04-results.tex
Kieran's comments and some other improvements

--- 2013-09-19 Publications/2013-ISBI14/04-results.tex
@@
 \section{RESULTS AND DISCUSSION}
+\label{sec:results}
+The results of the proposed experiments are summarized
+in \autoref{fig:results} and \autoref{table:results}.
+The remaining of this section provides extended descriptions
+and interpretation of results.
 
-WIP: work in progess..
+\input{041-results-table}
+
+\begin{figure*}[tpb]
+   \centering
+   \includegraphics[width=0.9\textwidth]{Fig02-Results}
+   \caption{The three major stages of the evaluation framework,
+   along with the associated visual results. First row represents
+   a coronal section of the \textit{b0} volume. In second row, the outcome
+   of tractography, showing only tracks that connect two different
+   network nodes. Third row shows the associated connectivity matrix. }
+   \label{fig:results}
+\end{figure*}
+
+\subsection{Geometrical correctness}
+
+The three methods under study achieved good results in
+terms of geometrical accuracy, as it is shown in the first
+row of \autoref{fig:results}. The best numerical scores
+were obtained with the \gls*{reb} method (\autoref{table:results}),
+that achieved a weighted average overlap of $96.64\%$. Second
+best were obtained with \gls*{fmb}. Geometrical correctness
+is fundamental in connectivity analysis to spatially locate the
+\glspl*{roi} that will define the nodes of the final connectivity
+matrix. The clear difference of accuracy between \gls*{reb} and \gls*{fmb}
+with respect to \gls*{t2b} infers the latter may not be
+an appropriate method for susceptibility correction.
+
+\subsection{Signal loss recovery}
+
+A second workflow evaluated the different correction 
+methodologies to show the similarity of the recovered
+signal with respect to the original (undistorted).
+In this second study, \gls*{reb} performed significantly
+better than the other two methods, as reported
+in \autoref{table:results}. \gls*{reb} scored a $91.0\%$
+similarity index for the \textit{b0} volume and an average $97.65\%$
+for the remaining 31 \glspl*{dwi}. Again, the second
+qualified was \gls*{fmb}, which achieved very close results
+for the \glspl*{dwi} ($96.26\%$) but not as good for the \textit{b0}
+volume. Visual inspection of the recovered data confirm the 
+results presented in the Table (\autoref{fig:results}, first
+row). Finally, \gls*{t2b} confirmed that its correction
+can not be compared, even when using the intensity compensation
+based on the determinant of the deformation field.
+
+\subsection{Impact on tractography and connectivity}
+
+The gold-standard contained a total of 1800 valid tracks
+connecting regions, with an average length of $40.88\pm8.06$mm.
+The corresponding connectivity matrix presented 25 connections 
+between the total of 27 different seeding regions.
+
+The distortion of the phantom caused 2 false positives (inexistent 
+connections created by distortion) and 2 false negatives 
+(connections vanished). The correction methods had an adverse impact
+on connectivity. Only \gls*{reb} could recover one lost connection
+but increasing the false connections count to 3.
+\gls*{fmb} ranked second, with 5 false and 2 lost connections.
+Finally, \gls*{t2b} presented 9 false and 3 lost connections.
+
+For the shake of completeness, we also document the outcome
+of tractography after each correction method. After distortion,
+we obtained 1750 tracks and a characteristic length of 
+$40.64\pm8.12$mm.
+\gls*{fmb} produced 1900 tracks/$40.53\pm8.07$mm., \gls*{reb}
+obtained 2200 tracks/$41.15\pm7.60$mm., and 2500 
+tracks/$40.60\pm7.81$mm. for \gls*{t2b}. Although track counting has
+been used as a measurement of success and as a weight of connectivity 
+strength, the number of tracks produced by tractography strongly rely
+on the seeding performed and the smoothing caused by interpolations 
+within correction methodologies.
+
+\subsection{Discussion}
+
+Even though all the surveyed methods produced visually 
+sound results, our study suggested that \gls*{reb} is the 
+most accurate methodology for correction of susceptibility-induced
+distortions in \gls*{dwi}. Moreover, \gls*{reb} also achieved
+the best results in terms of connectivity, being able to recover
+one of two lost connections after distortion. \Gls*{fmb} achieved 
+very positive scores. Finally, the \gls*{t2b} method did not achieve
+the necessary high-standards to recommend its use. We understand
+that specific methods with anisotropic regularization that completely
+restrict deformations to the phase-encoding axis would perform
+significantly better as compares to the standard methods in this study.
+
+One relevant limitation of this work is that the \gls*{fmb} is
+used both in distortion creation and correction, what could
+bias results. Nonetheless, \gls*{reb} performed better that \gls*{fmb}.
+A second limitation is found on the impossibility of integrating the
+point spread function-based methods for two reasons.
+First, the unmet need of a point spread function map consistent
+with the synthetic phase-difference map. Second, to our knowledge,
+there are no implementations of the method readily available.
+In practice, this second limitation is narrowed down by the
+fact that typical acquisition protocols usually include one or more
+datasets among field mapping, reversed-encoding \textit{b0} or T2,
+but they do not include the point-spread function mapping.

### 2013-09-30 Publications/2013-ISBI14/04-results.tex
Added future lines as Kieran suggested

@@
 but they do not include the point-spread function mapping.
+
+This work presents an evaluation framework that will be extended
+in order to evaluate numerous different susceptibility distortion
+configurations, better characterizing the phenomenon. Also, the
+framework can be used for studying the impact of other artifacts
+as subject's motion or eddy currents-derived distortions.

### 2013-10-01 Publications/2013-ISBI14/02-introduction.tex
Chus' changes

@@
 
-\Gls*{fmb} methods supposed the first approach to retrospective
-correction \cite{jezzard_correction_1995}. They rely on one
-extra acquisition (\emph{field mapping}), that probes the 
-inhomogeneity of the B0 field.
+\Gls*{fmb} methods \cite{jezzard_correction_1995} are
+the initial approach to retrospective corrections. 
+They rely on one extra acquisition (\emph{field mapping}),
+that probes the inhomogeneity of the B0 field.
 A second theory-based breed of methodologies acquire a 
@@
 have been studied \cite{zeng_image_2002,wu_comparison_2008},
-the lack of a gold-standard limits the possible benchmarking 
-strategies. Recently, \cite{irfanoglu_effects_2012} raised the question
+the lack of a gold-standard limits benchmarking 
+strategies. Recently, Irfanoglu et al.
+\cite{irfanoglu_effects_2012} raised the question
 of distortion-derived impacts in tractography.

### 2013-10-01 Publications/2013-ISBI14/03-methods.tex
Chus' changes

@@
 Based on the fiber geometries of the digital phantom 
-created for the High angular resolution diffusion imaging 
-reconstruction Challenge held in ISBI 2013 
+created for the \emph{High angular resolution diffusion imaging 
+reconstruction Challenge} held in ISBI 2013 
 (San Francisco, US), we simulated high resolution 
@@
 (1.0mm isotropic, $b$=1200, 32 evenly-distributed 
-directions, 1 \textit{b0} image) image.
+directions, 1 \textit{b0} volume) image.
 The phantom includes \gls*{wm} fiber bundles, 
@@
 ``acquisition'' of the same phantom was necessary 
-for evaluating the second family of correction methods.
+for evaluating \gls*{reb} methods.
 
-In this simple manner, we generated a full gold-standard
-containing realistic T1 and T2
+Therefore, following the described strategy we generated 
+a full gold-standard containing realistic T1 and T2
 at high resolution, \glspl*{dwi} acquired in two
@@
 Three out of four methods presented in \autoref{sec:intro}
-were tested on the evaluation framework.
+were tested on the evaluation framework. \Gls*{fmb} correction
+is essentially the same as the one used for generating the 
+distortion, using the synthetic fieldmap to infer a new \gls*{vsm}.
+For practical reasons (distortion on the mask, signal dropout), the
+method did not achieve a perfect geometrical correction as might be
+expected if the process was exactly inverse. For \Gls*{reb} correction,
+we made use of the implementation found in \emph{FSL} (\texttt{topup})
+that demands for the \textit{b0} of the reverse-encoding simulation.
+In this second case, the \gls*{vsm} is inferred from the differences
+between the two corresponding \textit{b0}.
 
-\paragraph*{\Gls*{fmb}} In this case, the pipeline essentially
-the same as the one used for generating the distortion, but
-applying the \gls*{vsm} in the opposite direction.
-
-\paragraph*{\Gls*{reb}} We made use of the implementation found
-in \emph{FSL} (called \texttt{topup}) to correct using the \textit{b0} of a
-reverse-encoding simulation of the \glspl*{dwi}. In this second
-case, a \gls*{vsm} is inferred from the differences between the 
-two corresponding \textit{b0}.
-
-\paragraph*{\Gls*{t2b}} For the last surveyed method,
+Finally, for evaluating \Gls*{t2b},
 we fine-tuned \emph{ANTs} \cite{avants_ants:_2013},
@@
 In order to enable the analysis of the influence of
-distortions on the final connectivity analyses,
-a intermediate step was also implemented to
-provide with deterministic tractography,
+distortions on the final connectivity outcome,
+an intermediate step was also implemented to
+produce deterministic tractography,
 using the \emph{Diffusion Toolkit} \cite{wang_diffusion_2007}.
@@
 indices.
-Each tractography was restricted by its corresponding 
-\gls*{wm} mask (e.g. the distorted and corrected with
-fieldmap for the tractography that followed this method). 
-In all cases, we used 10 random seeds at each seeding
+Each tractography was restricted to its corresponding 
+\gls*{wm} mask, after distortion and correction when
+applied. For all the cases, we used 10 random seeds at each seeding
 voxel and the seeding voxels were taken from the 
-ground-truth, mimicking the procedure with real data
-(in which this regions are typically projected from the
+ground-truth, mimicking the usual procedure on real data
+(where regions are typically mapped from the
 anatomically correct T1).
 
-Finally, the proposed workflows are connected to performance
+The evaluation framework is completed by automated 
 assessment modules. We evaluated three characteristics
 of the correction methods. 
-First, we assessed the geometrical correctness
-and reported overlap indices of three tissue
+Firstly, we assessed the geometrical correctness
+reporting overlap indices of three tissue
 probability maps (namely \gls*{csf}, \gls*{wm},
-and \gls*{gm}),
-weighting the average by the volumes of labels.
-Second, to evaluate the quality of the actual 
+and \gls*{gm}), weighting the average by tissue
+volumes.
+Secondly, to evaluate the quality of the actual 
 signal dropout correction, we studied the 
@@
 on the \textit{b0} and the average of the remaining 32 
-\glspl*{dwi}. Third, we studied the impact on the
+\glspl*{dwi}. Thirdly, we studied the impact on the
 connectivity matrices reporting the number of

### 2013-10-01 Publications/2013-ISBI14/04-results.tex
Last changes
Chus' changes

--- 2013-09-30 Publications/2013-ISBI14/04-results.tex
@@
-\section{RESULTS AND DISCUSSION}
+\section{RESULTS}
 \label{sec:results}
@@
    \includegraphics[width=0.9\textwidth]{Fig02-Results}
-   \caption{The three major stages of the evaluation framework,
-   along with the associated visual results. First row represents
+   \caption{Visual comparison of correction methods results. 
+   First row represents
    a coronal section of the \textit{b0} volume. In second row, the outcome
@@
 
-A second workflow evaluated the different correction 
-methodologies to show the similarity of the recovered
+A second workflow investigated the similarity of the recovered
 signal with respect to the original (undistorted).
@@
 For the shake of completeness, we also document the outcome
-of tractography after each correction method. After distortion,
+of tractography at each step of the workflow. After distortion,
 we obtained 1750 tracks and a characteristic length of 
@@
 strength, the number of tracks produced by tractography strongly rely
-on the seeding performed and the smoothing caused by interpolations 
+on the seeding performed and the smoothing caused by interpolation used
 within correction methodologies.
@@
 that specific methods with anisotropic regularization that completely
-restrict deformations to the phase-encoding axis would perform
-significantly better as compares to the standard methods in this study.
+restrict deformations to the phase-encoding direction would perform
+significantly better compared to the standard method included
+in this study.
 
@@
 point spread function-based methods for two reasons.
-First, the unmet need of a point spread function map consistent
-with the synthetic phase-difference map. Second, to our knowledge,
+Firstly, the unmet need of a point spread function map consistent
+with the synthetic phase-difference map. Secondly, to our knowledge,
 there are no implementations of the method readily available.
@@
 but they do not include the point-spread function mapping.
+
+Future extensions of this work will evaluate
+different realizations of the synthetic phase-difference map,
+better characterizing the phenomenon. Moreover, the
+framework can be enhanced for studying the impact of 
+other artifacts as subject's motion or eddy 
+currents-derived distortions.

### 2013-10-01 Publications/2013-ISBI14/05-conclusions.tex
Chus' changes

@@
 and increased variability in subsequent connectivity 
-analyses of the whole-brain. Specially, in regions 
+analyses of the whole-brain. Especially, in regions 
 influenced by  significant distortion.
@@
 readily-available correction methodologies. 
-This work shows that \gls*{dmri} should be corrected
-for susceptibility-induced artifacts prior to tractography
-and connectivity analyses. In that sense, the \gls*{reb} 
-method was the best classified
-in two of the three rankings (geometric 
-correctness and signal recovery), whereas \gls*{fmb}
-achieved the lowest impact on the connectivity matrix
-extracted from tractography.
+The reverse-encoding method (\gls*{reb}) achieved the
+best results correcting for susceptibility-induced 
+artifacts. We reported geometrical accuracy, signal
+dropout recovery and connectivity matrix detection as
+ranking indices.

### 2013-10-01 Publications/2013-ISBI14/01-abstract.tex
Manus' changes 2

@@
 Connectivity analysis on diffusion MRI data of the whole-brain
-suffers from distortions originated by the standard 
+suffers from distortions caused by the standard 
 echo-planar imaging sequences used in acquisition. %

### 2013-10-01 Publications/2013-ISBI14/02-introduction.tex
Manus' changes 2

@@
 profoundly described, generally within the context of
-functional MRI because they are also acquired with 
+functional MRI which also uses
 \gls*{epi}.
@@
 \cite{irfanoglu_effects_2012} raised the question
-of distortion-derived impacts in tractography.
+of distortion-derived impact on tractography.
 In this work, we propose an evaluation framework

### 2013-10-01 Publications/2013-ISBI14/03-methods.tex
Manus' changes 2

@@
 
-Once a realistic \gls*{vsm} was available, we generated
+From this synthetic \gls*{vsm},
+we generated
 the corresponding distorted \glspl*{dwi}, in two opposed
@@
 Three out of four methods presented in \autoref{sec:intro}
-were tested on the evaluation framework. \Gls*{fmb} correction
+were tested on the evaluation framework.
+
+\Gls*{fmb} correction
 is essentially the same as the one used for generating the 
 distortion, using the synthetic fieldmap to infer a new \gls*{vsm}.
-For practical reasons (distortion on the mask, signal dropout), the
-method did not achieve a perfect geometrical correction as might be
-expected if the process was exactly inverse. For \Gls*{reb} correction,
-we made use of the implementation found in \emph{FSL} (\texttt{topup})
+
+For \Gls*{reb} correction, we made use of the implementation found 
+in \emph{FSL} (\texttt{topup})
 that demands for the \textit{b0} of the reverse-encoding simulation.
@@
 
-In order to enable the analysis of the influence of
+To evaluate the influence of
 distortions on the final connectivity outcome,
@@
 signal dropout correction, we studied the 
-signal similarity frame by frame computing the $\ell_1$-norm
+similarity volume by volume computing the $\ell_1$-norm
 correlation index. We report this score

### 2013-10-01 Publications/2013-ISBI14/04-results.tex
Manus' changes 2

@@
 for the \glspl*{dwi} ($96.26\%$) but not as good for the \textit{b0}
-volume. Visual inspection of the recovered data confirm the 
+volume. Visual inspection of the recovered data confirms the 
 results presented in the Table (\autoref{fig:results}, first
@@
 
-The distortion of the phantom caused 2 false positives (inexistent 
-connections created by distortion) and 2 false negatives 
-(connections vanished). The correction methods had an adverse impact
+The distortion of the phantom caused 2 false connections 
+(positive connections that are not present on the ground-truth) 
+and 2 false negatives (connections vanished).
+The correction methods had an adverse impact
 on connectivity. Only \gls*{reb} could recover one lost connection
@@
 
-For the shake of completeness, we also document the outcome
+For the sake of completeness, we also document the outcome
 of tractography at each step of the workflow. After distortion,
@@
 used both in distortion creation and correction, what could
-bias results. Nonetheless, \gls*{reb} performed better that \gls*{fmb}.
+bias results. For practical reasons (distortion on the mask, 
+signal dropout), the \gls*{fmb} can not achieve a perfect 
+geometrical correction as might be
+expected if the process was exactly inverse.
+Nonetheless, \gls*{reb} performed better that \gls*{fmb}.
 A second limitation is found on the impossibility of integrating the

### 2013-10-01 Publications/2013-ISBI14/01-abstract.tex
Kieran's second comments

@@
 suffers from distortions caused by the standard 
-echo-planar imaging sequences used in acquisition. %
+echo-planar imaging acquisition strategies. %
 %These artifacts are present at interfaces between tissues,

### 2013-10-01 Publications/2013-ISBI14/02-introduction.tex
Kieran's second comments

@@
 and is generally acquired with an \gls*{epi} sequence.
-After a signal reconstruction step, 
+After signal reconstruction,
 tractography algorithms draw a map of the sampled 
@@
 close to tissue interfaces. This inhomogeneity
-of the field translates in a highly distorted imaged 
-anatomy and a significant signal destruction of 
+of the field translates in a highly distorted
+anatomy and a significant signal destruction in
 certain regions of the brain 
@@
 air surrounding sinuses). This artifact has been
-profoundly described, generally within the context of
-functional MRI which also uses
-\gls*{epi}.
+well described, generally within the context of
+functional MRI which also uses \gls*{epi}.
 
-\Gls*{fmb} methods \cite{jezzard_correction_1995} are
-the initial approach to retrospective corrections. 
-They rely on one extra acquisition (\emph{field mapping}),
+Various approaches have been proposed to correct
+for the disortions. \Gls*{fmb} methods
+\cite{jezzard_correction_1995} rely on one extra
+acquisition (\emph{field mapping}),
 that probes the inhomogeneity of the B0 field.
@@
 for the artifact \cite{robson_measurement_1997}.
-Next generation of methodologies \cite{cordes_geometric_2000,
-chiou_simple_2000}, make use of the acquisition of 
-extra \gls*{epi} volumes 
-with specific differences that enable correction, for instance
-the opposed direction of the phase encoding gradients. We will
-refer to these methods as \gls*{reb}.
-Finally, the last family of methodologies acquire an 
-extra T2-weighted image, and use its anatomical correctness 
-as reference to find the deformation map through nonlinear 
+Another approach referred to as \gls*{reb}, acquires an
+extra \gls*{epi} volume in the orthogonal or reversed
+phase encoding that can be combined to remove the
+geometric distortions
+\cite{cordes_geometric_2000,chiou_simple_2000}.
+The last family of methodologies acquires an
+extra T2-weighted image, and uses it as an anatomical
+reference to find the deformation map through nonlinear 
 registration \cite{kybic_unwarping_2000,studholme_accurate_2000}.
-\gls*{t2b} methods usually map
+These \gls*{t2b} methods usually map
 the T2 image to the \emph{baseline} volume or \textit{b0}
@@
 
-Even though aforementioned techniques for distortion correction
+Even though the aforementioned techniques for distortion correction
 have been studied \cite{zeng_image_2002,wu_comparison_2008},

### 2013-10-01 Publications/2013-ISBI14/03-methods.tex
Kieran's second comments

@@
 
-Therefore, following the described strategy we generated 
+In summary, we generated 
 a full gold-standard containing realistic T1 and T2
@@
 different phase encoding directions, and a ground-truth
-\gls*{dwi} data, which is not available for real datasets.
+\gls*{dwi} data, which is not available for real datasets
+(\autoref{fig:phantom}).
 
@@
    directions.}
-   \label{fig:label}
+   \label{fig:phantom}
 \end{figure}

### 2013-10-01 Publications/2013-ISBI14/04-results.tex
Kieran's second comments

@@
 signal with respect to the original (undistorted).
-In this second study, \gls*{reb} performed significantly
+In this second study, \gls*{reb} performed
 better than the other two methods, as reported
@@
 One relevant limitation of this work is that the \gls*{fmb} is
-used both in distortion creation and correction, what could
-bias results. For practical reasons (distortion on the mask, 
+used both in distortion creation and correction, which could
+bias the results. For practical reasons (distortion on the mask, 
 signal dropout), the \gls*{fmb} can not achieve a perfect 
@@
 expected if the process was exactly inverse.
-Nonetheless, \gls*{reb} performed better that \gls*{fmb}.
+Nonetheless, \gls*{reb} performed better than \gls*{fmb}.
 A second limitation is found on the impossibility of integrating the

### 2013-10-01 Publications/2013-ISBI14/03-methods.tex
Andres' comments

@@
 applied. For all the cases, we used 10 random seeds at each seeding
-voxel and the seeding voxels were taken from the 
+voxel (belonging to one out of 27 different seeding regions).
+The seeding regions were taken from the 
 ground-truth, mimicking the usual procedure on real data
@@
 on the \textit{b0} and the average of the remaining 32 
-\glspl*{dwi}. Thirdly, we studied the impact on the
+volumes. Thirdly, we studied the impact on the
 connectivity matrices reporting the number of

### 2013-10-01 Publications/2013-ISBI14/04-results.tex
Andres' comments
New figure 02, with binary connectivity matrices

@@
    a coronal section of the \textit{b0} volume. In second row, the outcome
-   of tractography, showing only tracks that connect two different
-   network nodes. Third row shows the associated connectivity matrix. }
+   of tractography after filtering tracks that did not connect two regions.
+   Third row shows the associated connectivity matrices. First column shows
+   the ground-truth. For the remaining steps, false positives have been marked
+   in red color and false negatives in blue. Underlaying in gray color,
+   the true positive (correctly detected) connections. }
    \label{fig:results}
@@
 connecting regions, with an average length of $40.88\pm8.06$mm.
-The corresponding connectivity matrix presented 25 connections 
-between the total of 27 different seeding regions.
+The corresponding connectivity matrix presented 25 connections.
+Results are presented in last row of \autoref{fig:results}.
 

### 2013-10-01 Publications/2013-ISBI14/01-abstract.tex
Final version

@@
 These images show characteristic geometrical deformations
-and signal destruction that are a significant pitfall
+and signal destruction that are an important pitfall
 limiting the success of tractography algorithms.
@@
 on their geometrical accuracy, the dropout compensation,
-and their impact on the resulting tractography.
+and their impact on the resulting connectivity matrices.
 \end{abstract}%

### 2013-10-01 Publications/2013-ISBI14/02-introduction.tex
Final version

@@
 In-vivo whole-brain connectivity analysis has been a
-research topic of high interest for the last
-5 years. 
+research topic of high interest for the last 5 years.
 \Gls*{dmri} can be used to probe the
@@
 well described, generally within the context of
-functional MRI which also uses \gls*{epi}.
+functional MRI, which also uses \gls*{epi}.
 
 Various approaches have been proposed to correct
-for the disortions. \Gls*{fmb} methods
+for the distortions. \Gls*{fmb} methods
 \cite{jezzard_correction_1995} rely on one extra

### 2013-10-03 Scripts/workflows/evaluation.py
Last version for 1st deadline ISBI 2014

@@ -5,6 +5,7 @@
 
 import os.path as op
 import os
+import numpy as np
 
 import nipype.pipeline.engine as pe
 import dmri as dmri
@@ -17,7 +18,14 @@ import nipype.interfaces.ants as ants
 import nipype.workflows.dmri.fsl as wf
 
 
-def evaluation_workflow( name="Evaluation" ):
+def evaluation_workflow( name="Evaluation",
+                         te_incr=2.46e-3,
+                         dwell_time=0.77e-3,
+                         intensity=5.0e-2,
+                         sigma=6.0,
+                         direction='y' ):
+
+
     evaluation = pe.Workflow( name=name )
     
     inputnode = pe.Node( niu.IdentityInterface( fields=['in_dwi','in_dwi_mask',
@@ -29,28 +37,23 @@ def evaluation_workflow( name="Evaluation" ):
    
     outputnode = pe.Node( niu.IdentityInterface( fields=['out_tpms_resampled'] ),
                           name='outputnode' )
- 
-    te_incr = 2.46e-3 # secs
-    dwell_time = 0.77e-3 # secs
-    intensity = 0.75
-    sigma = 8.0
     
-    dist1_wf = dmri.distortion_workflow( name='Distortion_Y' )
+    dist1_wf = dmri.distortion_workflow( name='Distortion' )
     dist1_wf.inputs.inputnode.te_incr = te_incr
     dist1_wf.inputs.inputnode.echospacing = dwell_time
-    dist1_wf.inputs.inputnode.encoding_direction = 'y'
+    dist1_wf.inputs.inputnode.encoding_direction = direction
     dist1_wf.inputs.inputnode.intensity = intensity
     dist1_wf.inputs.inputnode.sigma = sigma
 
 
     resample= pe.MapNode( fs.MRIConvert(out_type='niigz'), iterfield=['in_file'], name='Resample')
-    normalize_tpms = pe.Node( niu.Function( input_names=['in_files'], output_names=['out_files'], function=dmri.normalize ), name='Normalize' )
+    norm_tpms = pe.Node( niu.Function( input_names=['in_files'], output_names=['out_files'], function=dmri.normalize ), name='Normalize' )
 
-    sel1 = pe.Node( niu.Select(index=3 ), name='SelectWM' )
[…]

### 2013-10-03 Publications/2013-ISBI14/03-methods.tex
Final version of ISBI2014 paper, 1st deadline

@@
 consistent with the phantom.  We defined two regions 
-of smooth de-phasing (up to $0.05\pi$ rad. excursions) 
+of smooth de-phasing (up to $2.7\times10^{-3}$~rad. excursions) 
 and computed the corresponding \gls*{vsm}, using the 
@@
 \cite{jenkinson_fsl_2012}  and
-standard parameters ($\Delta$TE=2.46ms for the
+standard parameters ($\Delta$TE=2.46~ms. for the
 fieldmap and \emph{effective dwell time} of 
-0.77ms for the \gls*{epi}).
+0.77~ms. for the \gls*{epi}).
 

### 2013-10-10 Publications/2013-ISBI14/03-methods.tex
Updating for second deadline

@@
 Based on the fiber geometries of the digital phantom 
-created for the \emph{High angular resolution diffusion imaging 
-reconstruction Challenge} held in ISBI 2013 
+created for the High angular resolution diffusion imaging 
+reconstruction Challenge held in ISBI 2013 
 (San Francisco, US), we simulated high resolution 
 (0.5mm isotropic) T1 (TE/TR=10/1500ms) and T2 
-(TE/TR=90/5000ms) images, as well as a \gls*{dmri}
-(1.0mm isotropic, $b$=1200, 32 evenly-distributed 
-directions, 1 \textit{b0} volume) image.
+(TE/TR=90/5000ms) images, as well as a two \gls*{dmri}
+images (1.0mm isotropic, $b$=1200, 1 B0 image) with
+32 and 64 evenly-distributed directions.
+Diffusion is modeled by a restricted and a hindered
+compartment, similar to \cite{assaf_composite_2005}.
 The phantom includes \gls*{wm} fiber bundles,
-\gls*{csf} and \gls*{gm}. Diffusion is modeled by a 
-restricted and a hindered compartment, similar to
-\cite{assaf_composite_2005}.
+\gls*{gm} and \gls*{csf}. Physical properties
+(T1/T2 times in ms) used in simulation are
+($0.832\pm0.010$ / $79\times10^{-3}\pm0.6\times10^{-3}$)
+for \gls*{wm},
+($1.331\pm0.013$ / $110\times10^{-3}\pm2.0\times10^{-3}$) for
+\gls*{gm} and ($3.5\pm0.1$ / $0.25\pm0.01$) for \gls*{csf}.
 
@@
 \subsection{Correction methods}
-
+\label{sec:correction}
 Three out of four methods presented in \autoref{sec:intro}
@@
 is essentially the same as the one used for generating the 
-distortion, using the synthetic fieldmap to infer a new \gls*{vsm}.
+distortion, on the inverse direction. Normally distributed 
+noise was added to the fieldmap (for a signal-to-noise 
+ratio of 20dB) before correction.
 
@@
 \subsection{Evaluation}
+\label{sec:evaluation}
+The original phantom, one distorted version, and
+the corrected instances are then connected to a
+\gls*{dwi} reconstruction and tractography pipeline.
+Additionally, the original tissue probability maps
+are also distorted and corrected to provide tractography
+with the required \gls*{wm} masks.
+These maps are also used in a final assessment module.
 
-The four modules described so far (namely, phantom
-deformation, \gls*{fmb} correction, \gls*{reb}
-correction and \gls*{t2b} correction) were conveniently
-integrated in second order workflows using
-\emph{nipype} \cite{gorgolewski_nipype:_2011}.
-The choice of this tool grants the reproducibility
-of the experiments, among other advantages.
-
-To evaluate the influence of
-distortions on the final connectivity outcome,
-an intermediate step was also implemented to
-produce deterministic tractography,
-using the \emph{Diffusion Toolkit} \cite{wang_diffusion_2007}.
-The original phantom, one distorted version, and
-the corrected instances are fed to this module.
-Additionally, the original tissue probability maps
-are also distorted and corrected, for computing overlap
-indices.
-Each tractography was restricted to its corresponding 
-\gls*{wm} mask, after distortion and correction when
-applied. For all the cases, we used 10 random seeds at each seeding
-voxel (belonging to one out of 27 different seeding regions).
-The seeding regions were taken from the 
-ground-truth, mimicking the usual procedure on real data
-(where regions are typically mapped from the
-anatomically correct T1).
+The framework provides \gls*{dwi} reconstruction and
+deterministic tractography,
+with \emph{Diffusion Toolkit} \cite{wang_diffusion_2007}
+for diffusion tensor imaging (configured with 10 random
+seeds per voxel by default), and \emph{MRTrix}
+\cite{tournier_mrtrix:_2012} for high angular resolution
+diffusion tensor imaging (default parameters set to use
+constrained spherical deconvolution, maximum harmonic
+order of 6, 150000 desired tracks).
+For both options, the seeding regions can be set to use
+the distorted-corrected \gls*{wm} mask or the regions used
+to generate the ground-truth, this second case mimicking
+the usual procedure on real data (where regions are 
+typically mapped from the anatomically correct T1).
 

### 2013-10-10 Publications/2013-ISBI14/04-results.tex
Updating for second deadline

@@
 \label{sec:results}
+
+All the modules described in the previous Section were
+conveniently integrated in workflows using \emph{nipype}
+\cite{gorgolewski_nipype:_2011}. The choice of this tool
+grants the reproducibility of the experiments,
+and the evaluation workflow is publicly released.
 The results of the proposed experiments are summarized
@@
 and interpretation of results.
-
-\input{041-results-table}
 
@@
 
-\subsection{Geometrical correctness}
+\subsection{Geometrical correctness and signal recovery}
 
@@
 
-\subsection{Signal loss recovery}
+\input{041-results-table}
 
@@
 volume. Visual inspection of the recovered data confirms the 
-results presented in the Table (\autoref{fig:results}, first
-row). Finally, \gls*{t2b} confirmed that its correction
-can not be compared, even when using the intensity compensation
-based on the determinant of the deformation field.
+presented numerical results (\autoref{fig:results}, first row).
 
 \subsection{Impact on tractography and connectivity}
+
+Results derived from diffusion tensor imaging 
+(using 32 directions \glspl*{dwi})
+were completely hindered by the high complexity of the
+fibers contained in the phantom (crossing, fanning, etc.).
+The immediate conclusion is that, with highly complex anatomies,
+connectivity outcome is more sensitive to 
+reconstruction and tractography than distortions. Therefore,
+all results presented are based on the 64-directions dataset
+reconstructed with constrained spherical deconvolution.
 
@@
 in this study.
+%Finally, \gls*{t2b} confirmed that its correction
+%can not be compared, even when using the intensity compensation
+%based on the determinant of the deformation field.
 

### 2013-10-10 Publications/2013-ISBI14/03-methods.tex
Added new results and figure

@@
 generated a synthetic phase-difference map 
-consistent with the phantom.  We defined two regions 
-of smooth de-phasing (up to $2.7\times10^{-3}$~rad. excursions) 
-and computed the corresponding \gls*{vsm}, using the 
+consistent with the phantom, using the 
 tools distributed with the \emph{FSL} package 
@@
 standard parameters ($\Delta$TE=2.46~ms. for the
-fieldmap and \emph{effective dwell time} of 
+field mapping and \emph{effective dwell time} of 
 0.77~ms. for the \gls*{epi}).
+We defined two regions of smooth dephasing and 
+computed the corresponding \gls*{vsm}. Amplitude
+of the dephasing maps can be modulated, enabling
+us to evaluate the impact of the distortion magnitude.
+We generated several \glspl*{vsm} with maximum
+displacements raging from 3.80 mm to 7.60 mm. The
+latter maximum was chosen at the upper limit of the
+typical distortion observed in real datasets.
 
-From this synthetic \gls*{vsm},
-we generated
+From these synthetic \glspl*{vsm}, we generated
 the corresponding distorted \glspl*{dwi}, in two opposed
@@
    \includegraphics[width=0.95\columnwidth]{Fig01-Phantom}
-   \caption{Original and distorted digital phantom.
+   \caption{Ground-truth digital phantom.
    A) T2-weighted image; B) undistorted \textit{b0} volume;
-   C, D) distorted \textit{b0} volumes with opposed phase encoding 
-   directions.}
+   C, D) distorted \textit{b0} volumes with opposed phase 
+   encoding directions, maximum displacement of 3.80 mm.}
    \label{fig:phantom}

### 2013-10-10 Publications/2013-ISBI14/04-results.tex
Added new results and figure

@@
 
+\input{042-results-tractography}
+
 \subsection{Impact on tractography and connectivity}
 
-Results derived from diffusion tensor imaging 
-(using 32 directions \glspl*{dwi})
-were completely hindered by the high complexity of the
-fibers contained in the phantom (crossing, fanning, etc.).
-The immediate conclusion is that, with highly complex anatomies,
-connectivity outcome is more sensitive to 
-reconstruction and tractography than distortions. Therefore,
-all results presented are based on the 64-directions dataset
-reconstructed with constrained spherical deconvolution.
+Connectivity matrices derived from diffusion tensor imaging 
+(using 32 directions \glspl*{dwi}) were completely
+hindered by the high complexity of the fibers contained
+in the phantom. Tractography discovered successfully only 4 out 
+of 27 existing connections from the original (undistorted) data,
+with more than 45 false connections.
+With the 64-directions dataset (reconstructed with constrained 
+spherical deconvolution) we found 23/27 connections,
+but we still observed 40 false positives. Therefore, the 
+experiments using the 32-directions phantom were discarded. The
+results presented in this subsection only refer to the 
+64-directions dataset.
 
-The gold-standard contained a total of 1800 valid tracks
-connecting regions, with an average length of $40.88\pm8.06$mm.
-The corresponding connectivity matrix presented 25 connections.
-Results are presented in last row of \autoref{fig:results}.
+Using the strongest distortion (maximum displacement of 7.60 mm),
+the number of detected connections
+remained the same (23/27), slightly increasing the number of
+false positives to 42. The immediate conclusion is that, with highly
+complex anatomies (crossing, fanning, etc.) and limited number of
+ground-truth connections, connectivity matrices are more sensitive 
+to reconstruction and tractography than to distortions.
 
-The distortion of the phantom caused 2 false connections 
-(positive connections that are not present on the ground-truth) 
-and 2 false negatives (connections vanished).
-The correction methods had an adverse impact
-on connectivity. Only \gls*{reb} could recover one lost connection
-but increasing the false connections count to 3.
-\gls*{fmb} ranked second, with 5 false and 2 lost connections.
-Finally, \gls*{t2b} presented 9 false and 3 lost connections.
-
-For the sake of completeness, we also document the outcome
-of tractography at each step of the workflow. After distortion,
-we obtained 1750 tracks and a characteristic length of 
-$40.64\pm8.12$mm.
-\gls*{fmb} produced 1900 tracks/$40.53\pm8.07$mm., \gls*{reb}
-obtained 2200 tracks/$41.15\pm7.60$mm., and 2500 
-tracks/$40.60\pm7.81$mm. for \gls*{t2b}. Although track counting has
-been used as a measurement of success and as a weight of connectivity 
-strength, the number of tracks produced by tractography strongly rely
-on the seeding performed and the smoothing caused by interpolation used
-within correction methodologies.
+For the sake of completeness, \autoref{table:results-tractography}
+reports the characteristic features of the tractography results
+and connectivity matrices, for the fieldmap that caused a maximum
+displacement of 3.80 mm. Very similar results were obtained for
+7.60 mm, a distortion which is at the limit of the real case. This
+results, along with visual inspection (\autoref{fig:results}, second
+and third rows) might point to \gls*{fmb} as the best correction method.
 
@@
 %based on the determinant of the deformation field.
+%Although track counting has been used as a measurement of success and as a weight of connectivity 
+%strength, the number of tracks produced by tractography strongly rely
+%on the seeding performed and the smoothing caused by interpolation used
+%within correction methodologies.
 

### 2013-10-11 Publications/2013-ISBI14/03-methods.tex
Draft sent to all for ISBI abstract

@@
 Based on the fiber geometries of the digital phantom 
-created for the High angular resolution diffusion imaging 
+created for the \gls*{hardi} 
 reconstruction Challenge held in ISBI 2013 
@@
 (0.5mm isotropic) T1 (TE/TR=10/1500ms) and T2 
-(TE/TR=90/5000ms) images, as well as a two \gls*{dmri}
+(TE/TR=90/5000ms) images, as well as two \gls*{dmri}
 images (1.0mm isotropic, $b$=1200, 1 B0 image) with
@@
 \gls*{gm} and \gls*{csf}. Physical properties
-(T1/T2 times in ms) used in simulation are
+(T1/T2 times in sec) used in simulation are
 ($0.832\pm0.010$ / $79\times10^{-3}\pm0.6\times10^{-3}$)
@@
 us to evaluate the impact of the distortion magnitude.
-We generated several \glspl*{vsm} with maximum
-displacements raging from 3.80 mm to 7.60 mm. The
-latter maximum was chosen at the upper limit of the
-typical distortion observed in real datasets.
+We generated several \glspl*{vsm} with increasing
+maximum shifts, ranging 3.80-7.60~mm, covering the
+typical range of distortion observed in real datasets.
 
@@
    C, D) distorted \textit{b0} volumes with opposed phase 
-   encoding directions, maximum displacement of 3.80 mm.}
+   encoding directions, maximum displacement of 3.80~mm.}
    \label{fig:phantom}
@@
 
-The framework provides \gls*{dwi} reconstruction and
-deterministic tractography,
-with \emph{Diffusion Toolkit} \cite{wang_diffusion_2007}
-for diffusion tensor imaging (configured with 10 random
-seeds per voxel by default), and \emph{MRTrix}
-\cite{tournier_mrtrix:_2012} for high angular resolution
-diffusion tensor imaging (default parameters set to use
-constrained spherical deconvolution, maximum harmonic
-order of 6, 150000 desired tracks).
+The framework provides two different options for
+\gls*{dwi} reconstruction and deterministic tractography.
+Firstly, \emph{Diffusion Toolkit} \cite{wang_diffusion_2007}
+for \gls*{dti}, is configured with 10 random
+seeds per voxel by default. Secondly, \emph{MRTrix}
+\cite{tournier_mrtrix:_2012} for \gls*{hardi}, with
+default parameters set to use constrained spherical
+deconvolution, maximum harmonic order of 6, and 150000
+desired tracks.
 For both options, the seeding regions can be set to use
-the distorted-corrected \gls*{wm} mask or the regions used
-to generate the ground-truth, this second case mimicking
-the usual procedure on real data (where regions are 
-typically mapped from the anatomically correct T1).
+either the distorted-corrected \gls*{wm} mask, or the 
+regions used to generate the ground-truth. This second
+seeding strategy mimics the usual procedure on real 
+data, where regions are typically mapped from the
+anatomically correct T1.
 
@@
 correlation index. We report this score
-on the \textit{b0} and the average of the remaining 32 
+on the \textit{b0} and the average of the remaining \gls*{dwi}
 volumes. Thirdly, we studied the impact on the

### 2013-10-11 Publications/2013-ISBI14/04-results.tex
Draft sent to all for ISBI abstract

@@
 
-All the modules described in the previous Section were
+All the modules described in the previous section were
 conveniently integrated in workflows using \emph{nipype}
@@
 The remaining of this section provides extended descriptions
-and interpretation of results.
+and interpretation of the results.
 
@@
    \includegraphics[width=0.9\textwidth]{Fig02-Results}
-   \caption{Visual comparison of correction methods results. 
-   First row represents
-   a coronal section of the \textit{b0} volume. In second row, the outcome
-   of tractography after filtering tracks that did not connect two regions.
-   Third row shows the associated connectivity matrices. First column shows
-   the ground-truth. For the remaining steps, false positives have been marked
-   in red color and false negatives in blue. Underlaying in gray color,
-   the true positive (correctly detected) connections. }
+   \caption{Visual evaluation of the correction methods.
+   First row represents a coronal section of the \textit{b0} volume. 
+   In second row, the outcome of tractography after filtering tracks
+   that did not connect two regions.
+   Third row shows the associated connectivity matrices. All the matrices
+   are compared to the ground-truth connections in the phantom. The existing
+   connections that were correctly detected at each step are depicted in white
+   color. False positives have been marked in red and false negatives in blue.
+   Gray color represents the true negatives (non-existent connections correctly detected).}
    \label{fig:results}
@@
 
-The three methods under study achieved good results in
-terms of geometrical accuracy, as it is shown in the first
-row of \autoref{fig:results}. The best numerical scores
-were obtained with the \gls*{reb} method (\autoref{table:results}),
-that achieved a weighted average overlap of $96.64\%$. Second
-best were obtained with \gls*{fmb}. Geometrical correctness
-is fundamental in connectivity analysis to spatially locate the
-\glspl*{roi} that will define the nodes of the final connectivity
-matrix. The clear difference of accuracy between \gls*{reb} and \gls*{fmb}
+A summary of quantitative results, computed for the \gls*{dti} 
+dataset with maximum shift of 3.80~mm,
+is provided in \autoref{table:results}.
+The best scores were obtained with the \gls*{reb} 
+method, followed by \gls*{fmb}. 
+The clear difference of accuracy between \gls*{reb} and \gls*{fmb}
 with respect to \gls*{t2b} infers the latter may not be
 an appropriate method for susceptibility correction. 
+The described performance was constant for \gls*{fmb} 
+and \gls*{reb} along the different magnitudes of 
+distortion evaluated (maximum phase-encoding direction 
+shifts in the range of 3.80-7.60~mm).
 
@@
 similarity index for the \textit{b0} volume and an average $97.65\%$
-for the remaining 31 \glspl*{dwi}. Again, the second
+for the remaining \glspl*{dwi}. Again, the second
 qualified was \gls*{fmb}, which achieved very close results
@@
 volume. Visual inspection of the recovered data confirms the 
-presented numerical results (\autoref{fig:results}, first row).
+presented quantitative results (\autoref{fig:results}, first row).
 
@@
 
-Connectivity matrices derived from diffusion tensor imaging 
-(using 32 directions \glspl*{dwi}) were completely
-hindered by the high complexity of the fibers contained
+Connectivity matrices derived from the \gls*{dti} dataset
+were completely hindered by the high complexity of the 
+fibers contained
 in the phantom. Tractography discovered successfully only 4 out 
@@
 with more than 45 false connections.
-With the 64-directions dataset (reconstructed with constrained 
+With the \gls*{hardi} dataset (reconstructed with constrained 
 spherical deconvolution) we found 23/27 connections,
 but we still observed 40 false positives. Therefore, the 
-experiments using the 32-directions phantom were discarded. The
+experiments using the \gls*{dti} phantom were discarded. The
 results presented in this subsection only refer to the 
-64-directions dataset.
+\gls*{hardi} dataset.
 
-Using the strongest distortion (maximum displacement of 7.60 mm),
+Using the strongest distortion (maximum shift of 7.60~mm),
 the number of detected connections
@@
 and connectivity matrices, for the fieldmap that caused a maximum
-displacement of 3.80 mm. Very similar results were obtained for
-7.60 mm, a distortion which is at the limit of the real case. This
-results, along with visual inspection (\autoref{fig:results}, second
-and third rows) might point to \gls*{fmb} as the best correction method.
+shift of 3.80~mm. Very similar results were obtained for
+7.60~mm. This results, along with visual inspection
+(\autoref{fig:results}, second and third rows), might point
+to \gls*{fmb} as the best correction method.
 
@@
 sound results, our study suggested that \gls*{reb} is the 
-most accurate methodology for correction of susceptibility-induced
-distortions in \gls*{dwi}. Moreover, \gls*{reb} also achieved
-the best results in terms of connectivity, being able to recover
-one of two lost connections after distortion. \Gls*{fmb} achieved 
-very positive scores. Finally, the \gls*{t2b} method did not achieve
-the necessary high-standards to recommend its use. We understand
-that specific methods with anisotropic regularization that completely
+most accurate method in terms of geometrical accuracy and
+signal dropout recovery. 
+The \gls*{t2b} method did not achieve the necessary high-standards
+to recommend its use. Nonetheless, we understand that specific
+methods with anisotropic regularization that completely
 restrict deformations to the phase-encoding direction would perform
-significantly better compared to the standard method included
-in this study.
-%Finally, \gls*{t2b} confirmed that its correction
-%can not be compared, even when using the intensity compensation
-%based on the determinant of the deformation field.
-%Although track counting has been used as a measurement of success and as a weight of connectivity 
-%strength, the number of tracks produced by tractography strongly rely
-%on the seeding performed and the smoothing caused by interpolation used
-%within correction methodologies.
+significantly better than the standard method presented.
+Geometrical correctness of \gls*{dwi} data is fundamental 
+in connectivity analysis to spatially locate the
+regions which will define the nodes of the final connectivity
+matrix.
 
-One relevant limitation of this work is that the \gls*{fmb} is
-used both in distortion creation and correction, which could
-bias the results. For practical reasons (distortion on the mask, 
-signal dropout), the \gls*{fmb} can not achieve a perfect 
-geometrical correction as might be
-expected if the process was exactly inverse.
-Nonetheless, \gls*{reb} performed better than \gls*{fmb}.
-A second limitation is found on the impossibility of integrating the
-point spread function-based methods for two reasons.
-Firstly, the unmet need of a point spread function map consistent
-with the synthetic phase-difference map. Secondly, to our knowledge,
-there are no implementations of the method readily available.
-In practice, this second limitation is narrowed down by the
-fact that typical acquisition protocols usually include one or more
-datasets among field mapping, reversed-encoding \textit{b0} or T2,
-but they do not include the point-spread function mapping.
+Regarding tractography, this study revealed that signal
+reconstruction and tractography algorithms masked the 
+impact of susceptibility distortion on the final connectivity
+matrices. Due to this effect, experiments performed on 
+the \gls*{dti} phantom were discarded for comparison.
+With the \gls*{hardi} dataset, the extracted 
+connectivity matrix slightly
+changed with distortion. Quantitative differences reported in
+\autoref{table:results-tractography} could be more related
+to the smoothing derived from interpolation implemented by
+each method. Visual results might suggest that \gls*{fmb}
+achieved better results.
 
-Future extensions of this work will evaluate
-different realizations of the synthetic phase-difference map,
-better characterizing the phenomenon. Moreover, the
-framework can be enhanced for studying the impact of 
-other artifacts as subject's motion or eddy 
-currents-derived distortions.
+Although we found connectivity matrices rather invariant with
+respect distortion, they are suspected to be significantly
+impacted by the susceptibility artifact on real data 
+\cite{irfanoglu_effects_2012}.
+This hypothesis points to the need of more appropriate 
+phantoms with denser connectivity matrices.
+State of the art phantoms for tractography usually present a
+discrete set of simulated fiber bundles that translate in 
+very sparse fiber-endings regions and connectivity matrix.
+In the real case, the surface limiting the tracks is densely 
+covered by the regions mapped from the anatomical 
+dataset, what leads to larger sensitivity with respect 
+deformations.
+
+A possible limitation of this work is that \gls*{fmb} is
+used for both synthesis and correction of the distortion.
+However, practical reasons (i.e. noise, signal dropout)
+impede perfect geometrical correction with \gls*{fmb}.
+Moreover, \gls*{reb} performed more accurately 
+than \gls*{fmb}.
+
+Future extensions of this work will include a 
+refined digital phantom. Additional lines for this
+work will evaluate different realizations of the 
+synthetic phase-difference map, to better characterize
+the phenomenon. Also, the framework can be enhanced 
+for studying the impact of other artifacts as 
+subject's motion or eddy currents-derived distortions.

### 2013-10-11 Publications/2013-ISBI14/05-conclusions.tex
Draft sent to all for ISBI abstract

@@
 and increased variability in subsequent connectivity 
-analyses of the whole-brain. Especially, in regions 
-influenced by  significant distortion.
+analyses of the whole-brain. This effect is suspected
+to be larger in regions influenced by large distortions.
+Even though, this study found out that the connectivity
+matrix is more largely affected by the tractography method.
 We publicly release the framework and also contribute
@@
 best results correcting for susceptibility-induced 
-artifacts. We reported geometrical accuracy, signal
-dropout recovery and connectivity matrix detection as
-ranking indices.
+artifacts. We also reported visual evaluation of 
+tractography and quantitative results for the 
+extracted connectivity matrices.

### 2013-10-11 Publications/2013-ISBI14/03-methods.tex
Final version submitted before new deadline (Oct 11)

@@
 Based on the fiber geometries of the digital phantom 
-created for the \gls*{hardi} 
-reconstruction Challenge held in ISBI 2013 
+created for the \emph{\Gls*{hardi} 
+Reconstruction Challenge} held in ISBI 2013 
 (San Francisco, US), we simulated high resolution 
@@
 images (1.0mm isotropic, $b$=1200, 1 B0 image) with
-32 and 64 evenly-distributed directions.
+32 and 64 evenly-distributed directions intended
+for \gls*{dti} and \gls*{hardi} reconstructions, respectively.
 Diffusion is modeled by a restricted and a hindered
@@
 We generated several \glspl*{vsm} with increasing
-maximum shifts, ranging 3.80-7.60~mm, covering the
-typical range of distortion observed in real datasets.
+maximum shifts (from 3.80 to 7.60~mm), covering 
+the typical magnitudes of distortion observed 
+in real datasets.
 
@@
 
-The framework provides two different options for
+The framework supports two different methods for
 \gls*{dwi} reconstruction and deterministic tractography.
@@
 false positives (inexistent connections in the
-gold-standard) and false negatives (or connections
-lost).
+gold-standard) and false negatives (or missed connections).

### 2013-10-11 Publications/2013-ISBI14/04-results.tex
Final version submitted before new deadline (Oct 11)

@@
 and the evaluation workflow is publicly released.
-The results of the proposed experiments are summarized
+The results of this study are summarized
 in \autoref{fig:results} and \autoref{table:results}.
 The remaining of this section provides extended descriptions
-and interpretation of the results.
+and interpretation of these findings.
 
@@
 and \gls*{reb} along the different magnitudes of 
-distortion evaluated (maximum phase-encoding direction 
-shifts in the range of 3.80-7.60~mm).
+distortion evaluated (maximum shifts of 3.80-7.60~mm).
 
 \input{041-results-table}
+
+\input{042-results-tractography}
 
@@
 
-\input{042-results-tractography}
-
 \subsection{Impact on tractography and connectivity}
@@
 with more than 45 false connections.
-With the \gls*{hardi} dataset (reconstructed with constrained 
-spherical deconvolution) we found 23/27 connections,
+With the \gls*{hardi} dataset we found 23/27 connections,
 but we still observed 40 false positives. Therefore, the 
@@
 and connectivity matrices, for the fieldmap that caused a maximum
-shift of 3.80~mm. Very similar results were obtained for
-7.60~mm. This results, along with visual inspection
+shift of 3.80~mm. A very similar outcome was obtained for
+7.60~mm. These results, along with visual inspection
 (\autoref{fig:results}, second and third rows), might point
@@
 sound results, our study suggested that \gls*{reb} is the 
-most accurate method in terms of geometrical accuracy and
+most accurate method in terms of geometrical agreement and
 signal dropout recovery. 

### 2013-10-11 Publications/2013-ISBI14/05-conclusions.tex
Final version submitted before new deadline (Oct 11)

@@
 comprehensively analyze the impact of susceptibility
-induced distortion on tractography from \gls*{dmri}
+induced distortion on connectivity analyses from \gls*{dmri}
 data. Inaccuracy on tractography produce errors
-and increased variability in subsequent connectivity 
-analyses of the whole-brain. This effect is suspected
-to be larger in regions influenced by large distortions.
-Even though, this study found out that the connectivity
+and increased variability in the extracted connectivity 
+matrices of the whole-brain. This effect is suspected
+to influence those regions with larger distortions.
+However, this study found out that the connectivity
 matrix is more largely affected by the tractography method.
@@
 readily-available correction methodologies. 
-The reverse-encoding method (\gls*{reb}) achieved the
+The reverse encoding-based method achieved the
 best results correcting for susceptibility-induced 

### 2013-10-14 Publications/2013-TMI/03-introduction.tex
Applying @zosso's changes to draft
Finished up to the numerical implementation section

@@
 \label{sec:introduction}
-% The very first letter is a 2 line initial drop letter followed
-% by the rest of the first word in caps.
-% 
-% form to use if the first word consists of a single letter:
-% \IEEEPARstart{A}{demo} file is ....
-% 
-% form to use if you need the single drop letter followed by
-% normal text (unknown if ever used by IEEE):
-% \IEEEPARstart{A}{}demo file is ....
-% 
-% Some journals put the first two words in caps:
-% \IEEEPARstart{T}{his demo} file is ....
-% 
-% Here we have the typical use of a "T" for an initial drop letter
-% and "HIS" in caps to complete the first word.
-%
+
 \IEEEPARstart{D}{iffusion} \gls{mri} is a widely used family
@@
 
-Early \gls{dwi} research mainly focused on the improvements of imaging 
-methodologies by a better understanding of the diffusion effect, and improving
-image reconstruction methodologies. Currently, the connectome extraction 
-and analysis relies on a large chain of sophisticated computational techniques
-including acquisition, reconstruction, modeling and model fitting, image segmentation
-and registration, fibre tracking, connectivity mapping, visualization, statistical 
-group studies, and inference \citep{daducci_connectome_2012}. This growing complexity 
-has given rise to challenging issues towards reliable structural information about 
-the neuronal tracts presented in, and also difficulties on the statistical analysis 
-\citep{bullmore_complex_2009,meskaldji_comparing_2013}. To our knowledge, studies 
-regarding the impact of the different sources of inaccuracy and variability along
-the toolchain for connectivity analysis \citep{jones_twenty-five_2010,
-cheng_characteristics_2012,jones_white_2013} are scarce.
-Here, we shall address three tasks included within  the image processing stage in 
-a unified approach: brain tissue segmentation in diffusion space 
-(\autoref{sec:dwi_segmentation}), 
-correction of geometrical distortions (\autoref{sec:distortion}), and structural 
-image registration to diffusion coordinate space (\autoref{sec:registration}). 
-These tasks are generally
-solved independently, or combined in pairs. However, there exist fundamental 
-coupling relationships that can be exploited to obtain a simultaneous solution to 
-the three problems. This joint approach satisfactory impacts the downstream
-outcomes of the whole pipeline with the increase of the internal consistency.
+% REMOVE paragraph (Dom, 1st comment)
+%Early \gls{dwi} research mainly focused on the improvements of imaging 
+%methodologies by a better understanding of the diffusion effect, and improving
+%image reconstruction methodologies. Currently, the connectome extraction 
+%and analysis relies on a large chain of sophisticated computational techniques
+%including acquisition, reconstruction, modeling and model fitting, image segmentation
+%and registration, fibre tracking, connectivity mapping, visualization, statistical 
+%group studies, and inference \citep{daducci_connectome_2012}. This growing complexity 
+%has given rise to challenging issues towards reliable structural information about 
+%the neuronal tracts presented in, and also difficulties on the statistical analysis 
+%\citep{bullmore_complex_2009,meskaldji_comparing_2013}. To our knowledge, studies 
+%regarding the impact of the different sources of inaccuracy and variability along
+%the toolchain for connectivity analysis \citep{jones_twenty-five_2010,
+%cheng_characteristics_2012,jones_white_2013} are scarce.
 
@@
 \label{sec:dwi_overview}
-\gls{dwi} data are usually acquired with \gls{epi} sequences 
+\gls{dwi} data are usually acquired with \gls{epi} schemes
 as they allow for very fast acquisition of large sets of slices each
@@
 in sequence and each one represents one probed diffusion direction.
-This first acquisition is called \emph{raw} data or \emph{diffusion
-signal}. One or more volumes without directional encoding (so-called \emph{baseline},
-$b = 0$, \emph{low $b$}, or just \emph{B0} volumes) are interleaved as
-reference, at the least one is acquired at the beginning of the
-\gls{epi} dataset. These B0 are acquired without direction gradient,
-showing a T2-like contrast.
+%This first acquisition is called \emph{raw} data or \emph{diffusion
+%signal}. One or more volumes without directional encoding (so-called \emph{baseline},
+%$b = 0$, \emph{low $b$}, or just \emph{b0} volumes) are interleaved as
+%reference, at the least one is acquired at the beginning of the
+%\gls{epi} dataset. These B0 are acquired without direction gradient,
+%showing a T2-like contrast.
 
@@
 state-of-art methodologies). On top of the
-chosen model, some scalar maps describing tissue features is generally
-computed. For instance, the most popular \gls{dwi}-derived scalar map 
-is \gls{fa} which depicts the isotropy of water diffusion inside the brain.
-Another widely studied map is the \gls{adc} (or \gls{md} that has a dual definition),
-which reflects the intensity of the diffusion effect. Many other
-directionally invariant measurements have been defined. For a comprehensive study, 
-we refer to \citep{ennis_orthogonal_2006}.
+chosen model, some scalar maps describing tissue features are generally
+computed (e.g. \gls*{fa}, and \gls*{adc}).
+%For instance, the most popular \gls{dwi}-derived scalar map 
+%is \gls{fa} which depicts the isotropy of water diffusion inside the brain.
+%Another widely studied map is the \gls{adc} (or \gls{md} that has a dual definition),
+%which reflects the intensity of the diffusion effect. Many other
+%directionally invariant measurements have been defined. For a comprehensive study, 
+%we refer to \citep{ennis_orthogonal_2006}.
 
@@
 affected by \glspl{pve} \citep{alexander_analysis_2001} that appears when
-several different tissues, or signal 
-emitters, are present in the same imaging unit, producing an averaged intensity. 
-The effect is directly related to the low resolution achievable with \gls{dwi} 
-(typically around $2.0\times2.0\times2.0mm^3$). An additional complication specific to
-\gls{dwi} is the \gls{csf} \emph{contamination} \citep{metzler-baddeley_how_2012},
-that is a particular \gls{pve} in which the signal sensed inside the affected voxel is 
-linear with respect the \gls{gm} and \gls{wm} contributions, but non-linear with
-respect \gls{csf}.
+several different tissues are present in the same imaging unit, producing
+an averaged intensity. The effect is directly related to the resolution,
+typically around $2.0\times2.0\times2.0mm^3$ for \gls{dwi}.
+
+% the low resolution achievable with \gls{dwi} 
+%(typically around ). An additional complication specific to
+%\gls{dwi} is the \gls{csf} \emph{contamination} \citep{metzler-baddeley_how_2012},
+%that is a particular \gls{pve} in which the signal sensed inside the affected voxel is 
+%linear with respect the \gls{gm} and \gls{wm} contributions, but non-linear with
+%respect \gls{csf}.
+
+\subsection{Correction for susceptibility distortions}
+\label{sec:distortion}
+
+One approach to correcting the susceptibility distortions was proposed
+with the earliest \gls{epi} sequences used in functional \gls{mri}, and
+relies on the acquisition of extra \gls{mri} data. Generally, a \gls{gre} is used to 
+obtain a map of magnitude and phase of the actual magnetic field inside the
+scanner. Based on this \emph{fieldmap} and the theory underlying the 
+distortion, \citep{jezzard_characterization_2005} proposed a correction 
+methodology. A number of forked and improved methodologies have been 
+developed to correct for the susceptibility distortion and received the
+generic name of ``fieldmap correction'' techniques \cite{hsu_correction_2009,
+reber_correction_2005}.
+
+Some other retrospective methodologies do not make use of the fieldmaps, 
+as explicitly modeling the distortion \citep{andersson_modeling_2001},
+registering with (anatomically correct) T2-weighted \gls{mri} 
+\citep{kybic_unwarping_2000,studholme_accurate_2000,li_accounting_2007,
+tao_variational_2009}, or acquiring an 
+extra B0 image with reversed phase encoding direction 
+\citep{andersson_how_2003,holland_efficient_2010}.
+
+A comparison of the diverse correction techniques is found in \citep{wu_comparison_2008}.
+This study claims that fieldmap correction methodologies are not entirely accurate
+and reliable, even though the method is correct in principle. The conclusion
+was later confirmed by \citep{tao_variational_2009}. Additional concerns regarding
+fieldmap correction are the requirement of an extra acquisition (that is not always
+met or it is impractical), or the accuracy of the measured fieldmap that is sensitive
+to various effects (such us respiration, blood flow, etc.). All these factors 
+have turned susceptibility distortion in \gls{epi} sequences an active field of 
+research for the last 10 years.
 
@@
 
-A number of methods have been proposed using features not directly derived from \gls{dwi} data.
-Segmentation obtained by co-registering structural T1-weighted images will be covered in 
-\autoref{sec:registration}.
+%A number of methods have been proposed using features not directly derived from \gls{dwi} data.
+%Segmentation obtained by co-registering structural T1-weighted images will be covered in 
+%\autoref{sec:registration}.
 
@@
 \todo[inline]{reference to a summary table}
-
-\subsection{Correction for susceptibility distortions}
-\label{sec:distortion}
-
-One approach to correcting the susceptibility distortions was proposed
-with the earliest \gls{epi} sequences used in functional \gls{mri}, and
-relies on the acquisition of extra \gls{mri} data. Generally, a \gls{gre} is used to 
-obtain a map of magnitude and phase of the actual magnetic field inside the
-scanner. Based on this \emph{fieldmap} and the theory underlying the 
-distortion, \citep{jezzard_characterization_2005} proposed a correction 
-methodology. A number of forked and improved methodologies have been 
-developed to correct for the susceptibility distortion and received the
-generic name of ``fieldmap correction'' techniques \cite{hsu_correction_2009,
-reber_correction_2005}.
-
-Some other retrospective methodologies do not make use of the fieldmaps, 
-as explicitly modeling the distortion \citep{andersson_modeling_2001},
-registering with (anatomically correct) T2-weighted \gls{mri} 
-\citep{kybic_unwarping_2000,studholme_accurate_2000,li_accounting_2007,
-tao_variational_2009} (we shall cover these methods in 
-\autoref{sec:registration}), or acquiring an 
-extra B0 image with reversed phase encoding direction 
-\citep{andersson_how_2003,holland_efficient_2010}.
-
-A comparison of the diverse correction techniques is found in \citep{wu_comparison_2008}.
-This study claims that fieldmap correction methodologies are not entirely accurate
-and reliable, even though the method is correct in principle. The conclusion
-was later confirmed by \citep{tao_variational_2009}. Additional concerns regarding
-fieldmap correction are the requirement of an extra acquisition (that is not always
-met or it is impractical), or the accuracy of the measured fieldmap that is sensitive
-to various effects (such us respiration, blood flow, etc.). All these factors 
-have turned susceptibility distortion in \gls{epi} sequences an active field of 
-research for the last 10 years.
 
@@
 
-\subsection{Summary}
+\subsection{Proposed model}
 \label{sec:contributions}
@@
 
+% Paragraph taken from first subsection to here
+Here, we shall address three tasks included within  the image processing stage in 
+a unified approach: brain tissue segmentation in diffusion space 
+(\autoref{sec:dwi_segmentation}), 
+correction of geometrical distortions (\autoref{sec:distortion}), and structural 
+image registration to diffusion coordinate space (\autoref{sec:registration}). 
+These tasks are generally
+solved independently, or combined in pairs. However, there exist fundamental 
+coupling relationships that can be exploited to obtain a simultaneous solution to 
+the three problems. This joint approach satisfactory impacts the downstream
+outcomes of the whole pipeline with the increase of the internal consistency.
+
 In this paper we propose a novel registration framework to simultaneously
@@
 distortion in between anatomical and \gls{dwi} space.
-{\color{red} {We thus reformulate the segmentation problem as an inverse problem, where we 
+We thus reformulate the segmentation problem as an inverse problem, where we 
 seek for an underlying deformation field (the distortion) mapping 
@@
 \autoref{sec:experiments}. We present results in \autoref{sec:results} and conclude 
-in \autoref{sec:conclusion}.}}
+in \autoref{sec:conclusion}.

### 2013-10-14 Publications/2013-TMI/04-methods.tex
Applying @zosso's changes to draft
Finished up to the numerical implementation section

@@
 
-\Gls{acwe} and level-set based formulations have been widely applied in image
-processing, as reviewed in \citep{suri_shape_2002}. We suggest clustering the 
+We suggest clustering the 
 current methodologies of template-based segmentation methods into three groups. 
@@
 to joint registration-segmentation methods is found in \citep{gorthi_active_2011}.
+\Gls{acwe} and level-set based formulations have been widely applied in image
+processing, as reviewed in \citep{suri_shape_2002}.
 
-Our joint registration and segmentation method is mainly inspired in the active
-deformation fields presented by \citep{gorthi_active_2011}, that is a generalization
-of the methodologies under second and third groups before. As it is described in
-their work, the first joint ``morphing''-segmentation approach (it was not a proper
+The first joint ``morphing''-segmentation approach (it was not a proper
 registration method) was proposed in \citep{bertalmio_morphing_2000}. The first
@@
 
-All the surveyed frameworks are deduced from the general level set evolution equation
-introduced by \citep{osher_fronts_1988}:
-\begin{equation}
-\frac{\partial \Phi_D(\mathbf{x},t)}{\partial t} = \nu ( \Phi_D(\mathbf{x},t)) \, \left| \nabla \Phi_D(\mathbf{x},t) \right|
-\end{equation}
-where $\nu$ is the velocity of the flow or speed function that contains the local
-segmentation and contour regularization constraints, and $\Phi_D: \Omega \to \mathbb{R}$
-is the signed distance function often used to represent implicitly the active contour
-by its zero level. An early extension to image registration was proposed by 
-\citep{vemuri_joint_2003} but replacing $\Phi_D$ by the intensity function of the
-image to register, $\Phi_I: \Omega \to \mathbb{R}$ (the moving image). 
-\citep{bertalmio_morphing_2000,vemuri_joint_2003} then define the speed function as
-$\nu(\Phi_I(\mathbf{x},t)) = \Phi_I(\mathbf{x},t) - \Phi_T(\mathbf{x},t)$, where $\Phi_T(\mathbf{x})$ is the intensity
-function of the target image. Finally, \citep{gorthi_active_2011} propose a different
-multiphase label function as level set function ($\Phi_L$) that improves registration
-results.
-
-Naming $\Phi_G$ a generic objective function, a dense deformation field of vectors 
-$u: \mathbb{R}^n \to \mathbb{R}^n$ (typically $n = \{ 2, 3 \}$) is introduced. 
-Then, the conservation of the morphological description is assumed, such as
-$\Phi_G(\mathbf{x},t) = \Phi_G( \mathbf{x} + du, t + dt ) \implies d\Phi_G(\mathbf{x},t) = 0$, where $d\Phi_G$ is
-the total derivative of $\Phi_G$. Using the chain rule, it can be rewritten as the 
-evolution equation of a vector flow:
-\begin{equation}
-\frac{\partial u(\mathbf{x},t)}{\partial t}= - \frac{\Phi_G}{\left| \nabla \Phi_G \right|} N_{\Phi_G},
-\end{equation}
-where $N_{\Phi_G}$ is the normal of the level set. Finally, they introduce the general
-level set evolution equation into the vector flow to perform the registration task:
-\begin{equation}
-\frac{\partial u(\mathbf{x},t)}{\partial t} = - \nu( \Phi_G(\mathbf{x},t) ) N_{\Phi_G}.
-\end{equation}
-Therefore, the position of the level set function $\Phi_G$ at time $t$ is given by the
-deformation field $u(\mathbf{x},t)$ and the initial level set function $\Phi_G(\mathbf{x},0)$, yielding:
-\begin{equation}
-\frac{\partial u(\mathbf{x},t)}{\partial t} = - \nu( \Phi_G(\mathbf{x} + u(\mathbf{x},t), 0) ) N_{\Phi_G}.
-\label{eq:active_deformation_fields}
-\end{equation}
-
-Here, we propose an \gls{adf} -like framework with some differences in
-the inspiration of the solution. Even though the historical evolution
+Even though the historical evolution
 of joint segmentation and registration procedures is strongly linked
 to the level sets approximation, our work is based in derivations of
-the \gls{acwe} framework \citep{chan_active_2001}, more specifically 
+the \gls{acwe} framework \citep{chan_active_2001}, and related to
 \citep{le_guyader_combined_2011}.
-The main diverging points with respect to
+{\color{red} {The main diverging points with respect to
 \citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
 $\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
-gradients \citep{jehan-besson_dream2s:_2003,herbulot_segmentation_2006}; 2) 
-\todo[inline]{regularization is also based on linear diffusion smoothing REF-thirion,
-but they use a Gaussian filtering). Maybe \citep{le_guyader_combined_2011} contains
-more regarding this difference, given that they use the same regularization as us. But,
-currently, I do not have enough background to write this second difference}
-\textbf{}; 3) optimization is applied in the spectral
+gradients \citep{jehan-besson_dream2s:_2003,herbulot_segmentation_2006};
+2) regularization is also based on linear diffusion smoothing \citep{thirion_image_1998},
+but we replace the Gaussian filtering by other constraints
+studied in \citep{nagel_investigation_1986} to better the problem;
+3) optimization is applied in the spectral
 domain, observing anisotropic and inhomogeneous mappings along each direction.
@@
 the distance function, and the spectral solution to the optimization updates,
-as we shall cover in \autoref{sec:numerical_implementation}.
+as we shall cover in \autoref{sec:numerical_implementation}.}}
 
-\todo[inline]{Disclaimer: I've been exploring the joint registration-segmentation
-applications starting from Gorthi's paper. Later, Dominique pointed me out that
-the main inspiration of the design comes from the ACWE world. If one takes a quick
-look on \citep{le_guyader_combined_2011} this is clear. But I did not have time
-yet to read this paper in depth. Fortunately, it does not seem to contain a lot
-of references to literature, so I hope it will be ease to re-formulate this
-section with its contents. Also, the highlighted differences will require an
-update once I fully understand that method}
+\todo[inline]{The previous paragraphs in red color are marked by Dominique
+as ``to rework'', and I totally agree}
 
-\subsection{Bridging segmentation frameworks}
+\subsection{Proposed model}
 \label{sec:methods_map}
-%
-A widely used approach to image segmentation is derived from the
-Bayes' rule \eqref{eq:bayes_rule}. This framework seeks for a 
-partitioning of a certain observation of features $F: \mathbb{R}^n \to %
-\mathbb{R}^C$ (an $n$-dimensional \emph{image} comprising $C$ different 
-channels) in piecewise smooth regions $\Omega = \lbrace \Omega_l , 
-l\in\left[ 1 \ldots L \right] \rbrace$,  that maximize the \emph{a posteriori}
-probability $p(\Omega \mid F)$:
+In this paper we formulate the joint registration-segmentation
+problem as follows. We look for a spatial mapping (with, typically,
+$n=3$ for the 3-dimensional case):
+\begin{align*}
+U\colon T \subset \mathbb{R}^n &\to D \subset \mathbb{R}^n \\
+\mathbf{x} &\mapsto \mathbf{x}' =\mathbf{x}+u(\mathbf{x}),
+\end{align*}
+such that the known contours in anatomical space $T$ optimally segment
+the diffusion space $D$.
+
+In a Bayesian framework, different such mappings $U$ between ``T1-''
+and ``\gls*{dwi}-'' spaces are evaluated based on their posterior
+probability given the observed data $F$. Using the Bayes' rule, this
+posterior \emph{likelihood} can be computed as:
 \begin{equation}
-p(\Omega \mid F) = \frac{p(F \mid \Omega)\, p(\Omega)}{p(F)},
+P(U \mid F,\Gamma_{ij}) = \frac{P(F \mid U,\Gamma_{ij})\, P(U)}{P(F)},
 \label{eq:bayes_rule}
 \end{equation}
-where $p(F \mid \Omega)$ is the \emph{likelihood} of the realization 
-of $F$ given a certain region set $\Omega$. The second term, $p(\Omega)$,
-is the a-priori probability of the partitioning. Finally, $p(F)$ is the 
-probability of a certain image realization, and thus, it will remain 
-constant when computing the \gls{map}. Consequently, the Bayes' rule
-can be simplified for the optimization as follows:
+where $P(F \mid U,\Gamma_{ij})$ is the data-likelihood, and
+$\Gamma_{ij}$ are a set of contours in T1-space that correspond to
+two interfacing regions $\Omega$, such that 
+$\Gamma_{ij} = \partial \Omega_i \cap \partial \Omega_j$.
+
+We are interested in a mapping $\hat{U}$ which maximizes this
+posterior probability (\gls*{map} estimate, \citep{bishop_pattern_2006}):
+\begin{align}
+\hat{U} &= \underset{U}{\argmax} \left\{ P(U \mid F,\Gamma_{ij}) \right\} \notag\\
+ &= \underset{U}{\argmax} \left\{ P(F \mid U,\Gamma_{ij})\, P(U) \right\}.
+\label{eq:map_u}
+\end{align}
+
+First, we assume independence between pixels, and thus break down the
+global data likelihood into a product of pixel-wise conditional probabilities:
+
 \begin{equation}
-p(\Omega \mid F) \propto p(F \mid \Omega)\, p(\Omega).
-\label{eq:bayes_rule_simplified}
+P(F \mid U,\Gamma_{ij}) = 
+\underset{\mathbf{x}\in \Omega}{\prod} P\left( F(\mathbf{x}') \mid U(\Gamma_{ij}) \right).
+\label{eq:bayes_aposteriori}
 \end{equation}
 
-In the simplest Bayesian segmentation framework, the unknown is the
-parameter set $\Theta_l$ of region-wise descriptors that minimize 
-\eqref{eq:bayes_rule}. Conversely, we can assume these descriptors as
-fixed (fulfilling the concept of conservation of morphological description
-mentioned before) and introduce a realization of a dense deformation field $U: %
-\mathbb{R}^n \to \mathbb{R}^n$ that maps the initial partition $\Omega_0$
-to the estimated one:
+Further, within each region $\Omega_k$ defined by the contours $\Gamma$
+trough duality, we assume the features to be i.i.d.:
 \begin{equation}
-\Omega' = \Omega_0 + \hat{U}.
-\label{eq:omega_tf}
+P\left( F(\mathbf{x}') \mid U,\Gamma_{ij} \right) = p_k( F(\mathbf{x}')) \quad
+\forall \mathbf{x} \in \Omega_k.
+\label{eq:likelihood}
 \end{equation}
 
-Thus, \eqref{eq:bayes_rule_simplified} is rewritten as 
-$p(U \mid F) \propto p(F \mid U)\, p(U)$ that yields the
-following \gls{map} solution:
+For convenience, and because this has been shown to be an appropriate
+approximation \citep{bishop_pattern_2006}, each such region is modeled
+by a multivariate normal distribution with region descriptors
+$\Theta_k = \lbrace \boldsymbol{\mu}_k, \boldsymbol{\Sigma}_{k} \rbrace$:
 
 \begin{equation}
-\hat{U} = \underset{U}{\argmax} \left\{ p(U \mid F) \right\} = 
-\underset{U}{\argmax} \left\{ p(F \mid U)\, p(U) \right\}.
-\label{eq:map_u}
+p_k( F(\mathbf{x}') ) = \mathcal{N} ( f \mid \Theta_k ).
+\label{eq:multivariate_normal}
 \end{equation}
 
-An extended assumption is that the feature vector realization $F$ is
-\emph{i.i.d.}, and thus, it is possible to write the a-posteriori
-probability $p(F \mid U)$ as a continuous product with $d\mathbf{x}$ the
-infinitesimal voxel size:
-\begin{equation}
-p(F \mid U) \, p(U) = \underset{l}{\prod} \underset{\mathbf{x}\in \Omega'_l}{\prod}
-p_l\left( F(\mathbf{x} + U(\mathbf{x})) \mid U(\mathbf{x}) \right)^{d\mathbf{x}}.
-\label{eq:bayes_aposteriori}
-\end{equation}
-\todo[inline]{I'm not convinced about turning $p(\Omega)$ into $p(U)$.}
-
-A second widely-accepted assumption is the multivariate normal 
-distribution of the different tissues in \gls{mri} data. Therefore,
-the posterior probability of an infinitesimal voxel can be written as:
-\begin{equation}
-p_l( F(\mathbf{x}) \mid U(\mathbf{x}) ) = \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_l (\mathbf{f}') \right)}}.
-\label{eq:bayes_mpdf}
-\end{equation}
-where we can identify the factor in the exponential as the squared \emph{Mahalanobis 
-distance} of the feature observed in the displaced position, $\mathbf{f}' = 
-F(\mathbf{x} + U(\mathbf{x}))$ with region descriptors
-$\Theta_l = \lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \rbrace$:
-\begin{equation}
-\Delta^2_l (\mathbf{f} \mid \Theta_l ) = (\mathbf{f} - \boldsymbol{\mu}_l)^T \, \boldsymbol{\Sigma}^{-1}_l \, (\mathbf{f} - \boldsymbol{\mu}_l).
-\label{eq:bayes_mahalanobis}
-\end{equation}
-
-Finally, we can turn the \gls{map} problem into a variational one
-applying the following log-transform:
-\begin{multline}
-E(F \mid U)= -\log \left[ p(F \mid U) \, p(U) \right] = \\
-= -\log \left[ \underset{l}{\prod} \underset{\mathbf{x}\in \Omega'_l}{\prod}
-p_l( F(\mathbf{x}) \mid U(\mathbf{x}) )^{d\mathbf{x}} \right] = \\
-= \sum\limits_l \int_{\Omega'_l} -\log \left[ p_l(F(\mathbf{x}) \mid U(\mathbf{x} ) ) \right] \, d\mathbf{x},
-\label{eq:energy_1}
-\end{multline}
-and introducing the posterior probability term \eqref{eq:bayes_mpdf}, 
-we can express the functional in terms of the deformation field $U$ and
-the region descriptors $\Theta_l$:
-\begin{multline}
-E(U) = \\
-= \sum\limits_l \int_{\Omega'_l} -\log \left[ \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_l (\mathbf{f}) \right)}} \right] \, d\mathbf{x} = \\
-= \sum\limits_l \int_{\Omega'_l} \left[ \frac{1}{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right| \right)} + \frac{1}{2}  \Delta^2_l (\mathbf{f}) \right] \,d\mathbf{x}.
-\end{multline}
-Finally, after removing scaling factors and independent constants,
-we obtain:
+\todo{I modified this slightly}
+Thus, defining the feature observed as $\mathbf{f}'=F(\mathbf{x}')$ and 
+the squared \emph{Mahalanobis distance} of $\mathbf{f}'$ with respect
+of the descriptors of region $k$ as
+$\Delta^2_k (\mathbf{f}' \mid \Theta_k ) = (\mathbf{f}' - \boldsymbol{\mu}_k)^T \, \boldsymbol{\Sigma}^{-1}_k \, (\mathbf{f}' - \boldsymbol{\mu}_k)$,
+we write:
 \begin{align}
-E(U) = \sum\limits_l \left[ \left|\Omega'_l\right|\,\log \left|\mathbf{\Sigma}_l \right| + \int_{\Omega'_l} \Delta^2_l (\mathbf{f}) \,d\mathbf{x} \right],
-\label{eq:map_energy}
-\end{align}
-where we have a constant term scaled by the total volume of the partition $\Omega_l$ and 
-the determinant of the  covariance matrix of the partition $\left|\boldsymbol{\Sigma}_{l}\right|$,
-plus an energy term based on the squared \emph{Mahalanobis distance}.
-As long as this shape-based registration framework is designed to allow small changes
-between boundaries (and therefore, their volumes), and the shape descriptors should not
-change significantly due to the conservation principle, the first term can be optionally
-dismissed with respect the second one, what yields:
-\begin{align}
-E(U) = \sum\limits_l \int_{\Omega'_l} \Delta^2_l (\mathbf{f}) \,d\mathbf{x}.
-\label{eq:final_map_energy}
+P( F \mid U, \Omega ) &= \underset{k}{\prod} \underset{\mathbf{x} \in \Omega_k}{\prod}
+\mathcal{N} ( f \mid \boldsymbol{\mu}_k, \boldsymbol{\Sigma}_{k} ) \notag\\
+&= \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_l (\mathbf{f}' \mid \Theta_k ) \right)}}.
+\label{eq:pdf}
 \end{align}
 
-This latter expression resembles the Mumford-Shah functional 
-\citep{mumford_optimal_1989} which is at the background of all \gls{acwe}-based
-methods. In this case, \eqref{eq:final_map_energy} includes covariance as
-region descriptor, what modifies the original functional in a way that it 
-can deal with more general distributions. This is necessary to avoid the 
-assumption that regions $\Omega_l$ have a fixed covariance matrix on their 
-complete domain. One immediate advantage of this functional from the original 
-one is the possibility to distinguish regions with the same mean vector but 
-different covariance matrix \citep{brox_local_2009}. In this work, the 
-Mumford-Shah functional is derived for this extension as follows:
-\begin{multline}
-E(\Theta_l,\Omega_l) = \sum\limits_l \int_{\Omega_l} \left[ \log \left|\mathbf{\Sigma}_l\right| + \Delta^2_l (\mathbf{f}) \right] \,d\mathbf{x} \\
-+ \lambda \int_{\Omega_l - d\Omega_l}  ( \left| \nabla \mathbf{\mu} \right| ^2 + \left| \nabla \mathbf{\Sigma}_l \right| ^2 ) \, d\mathbf{x} 
-+ \nu |d\Omega_l|,
-\end{multline}
-that is easily identifiable with \eqref{eq:map_energy} when we apply 
-the so-called \emph{cartoon limit}, 
-for $\lambda \to \infty$:
-\begin{equation}
-E(\Theta_l,d\Omega_l) = \sum\limits_l \int_{\Omega_l} \left[ \log \left|\mathbf{\Sigma}_l\right| + \Delta^2_l (\mathbf{f}) \right] \,d\mathbf{x}
-+ \nu |d\Omega_l|.
-\end{equation}
+We choose a Thikonov regularization prior as follows:
 
-As long as we do not penalize the edge set $d\Omega_l$ length, $\nu = 0$ and
-the result is dual to \eqref{eq:map_energy}:
-\begin{equation}
-E(\Theta_l,d\Omega_l) = \sum\limits_l \int_{\Omega_l} \left[ \log \left|\mathbf{\Sigma}_l\right| + \Delta^2_l (\mathbf{f}) \right] \,d\mathbf{x}.
-\end{equation}
+\begin{equation*}
+P(U) = \underset{\mathbf{x}}{\prod} p(u)
+\end{equation*}
+with further:
+\begin{align*}
+p(u) &= p_0(u) \, p_1(u) \notag\\
+p_0(u) &= \mathcal{N}( u \mid 0, A^{-1}) \notag\\
+p_1(u) &= \mathcal{N}( Du \mid 0, B^{-1}),
+\end{align*}
+that consider that the distortion and its gradient have zero
+mean and variance governed by $A$ and $B$. Since the anisotropy
+is technically aligned with the imaging axes, these can
+be simplified:
+\begin{align}
+p_0(u) &= \underset{\mathbf{x}}{\sum} \mathcal{N}( u(\mathbf{x}) \mid 0, \frac{1}{\sqrt{\alpha_{x,y,z}}}) \notag\\
+p_1(u) &= \underset{\mathbf{x}}{\sum} \mathcal{N}( \nabla_{x,y,z}u(\mathbf{x}) \mid 0, \frac{1}{\sqrt{\beta_{x,y,z}}})
+\label{eq:priors}
+\end{align}
 
-\subsection{Introducing priors}
-\label{sec:priors}
-%
-By introducing $U$ in \eqref{eq:bayes_rule_simplified}, we transformed 
-the segmentation problem into a registration one provided with $\Omega_0$ 
-(the initial partition) is derived from the shape-priors given in reference 
-space. Therefore, the priors can be expressed in terms of the deformation
-field:
-\begin{equation}
-P(U) = \underset{\mathbf{x}}{\prod}\,p(U(\mathbf{x}))\,^{d\mathbf{x}}.
-\label{eq:prior_u}
-\end{equation}
-We define $P(U(\mathbf{x})) = p(\mathbf{u}) = \underset{d}\prod \, p_d(\mathbf{u})$ 
-with $d$ the order of the derivative. We also assume that an initial
-affine registration let us consider that the deformation field ($d=0$)
-and its derivatives up to some order ($d>0$) are all zero mean, and
-have some variance $\Sigma_d$. We restrict ourselves to $d=\lbrace 0,1 \rbrace$
-(the field and its gradient):
-\begin{subequations}
-\begin{equation}
-p_0(\mathbf{u}) = \mathcal{N}( \mathbf{u} \mid 0, \alpha^{-1}),
-\end{equation}
-\begin{equation}
-p_1(\mathbf{u}) = \mathcal{N}( \nabla \mathbf{u} \mid 0, {\Sigma_\beta}^{-1}), 
-\end{equation}
-\end{subequations}
-And therefore:
-\begin{equation}
-P(U) = \underset{\mathbf{x}}\prod \left[ p_0(\mathbf{u}) \, p_1(\mathbf{u}) \right]^{d\mathbf{x}}.
-\end{equation}
+Finally, we can turn the \gls{map} problem into 
+a variational one
+applying the following log-transform:
+\begin{align}
+E(F \mid U) &= -\log \underset{k}{\prod}
+\underset{\mathbf{x} \in \Omega_k}{\prod} 
+p_k( \mathbf{f}')\,p_0( u(\mathbf{x}))\,p_1( u(\mathbf{x})) \notag\\
+&= C + \underset{k}{\sum}
+\underset{\mathbf{x} \in \Omega_k}{\sum}
+\Delta^2_k (\mathbf{f}' \mid \Theta_k ) \notag\\
+&+ \underset{d=\{x,y,z\}}{\sum} \alpha_d u^2_d
++ \underset{d=\{x,y,z\}}{\sum} \beta_d \|\nabla u_d\|^2,
+\label{eq:energy}
+\end{align}
+that is the dual expression to the energy functional corresponding
+to a discrete \gls*{acwe} framework \citep{chan_active_2001}
+with anisotropic regularization as studied in
+\citep{nagel_investigation_1986}.
 

### 2013-10-15 Publications/2013-TMI/04-methods.tex
Finished inclusion of @zosso comments
Some subscripts and mathematical symbols must be reviewed
because they are not consistent.

@@
 $n=3$ for the 3-dimensional case):
-\begin{align*}
-U\colon T \subset \mathbb{R}^n &\to D \subset \mathbb{R}^n \\
+\begin{align}
+U\colon T \subset \mathbb{R}^n &\to D \subset \mathbb{R}^n \notag\\
 \mathbf{x} &\mapsto \mathbf{x}' =\mathbf{x}+u(\mathbf{x}),
-\end{align*}
+\label{eq:deformation_field}
+\end{align}
 such that the known contours in anatomical space $T$ optimally segment
@@
 
-Further, within each region $\Omega_k$ defined by the contours $\Gamma$
+Further, within each region $\Omega_l$ defined by the contours $\Gamma$
 trough duality, we assume the features to be i.i.d.:
 \begin{equation}
-P\left( F(\mathbf{x}') \mid U,\Gamma_{ij} \right) = p_k( F(\mathbf{x}')) \quad
-\forall \mathbf{x} \in \Omega_k.
+P\left( F(\mathbf{x}') \mid U,\Gamma_{ij} \right) = p_l( F(\mathbf{x}')) \quad
+\forall \mathbf{x} \in \Omega_l.
 \label{eq:likelihood}
@@
 by a multivariate normal distribution with region descriptors
-$\Theta_k = \lbrace \boldsymbol{\mu}_k, \boldsymbol{\Sigma}_{k} \rbrace$:
+$\Theta_l = \lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$:
 
 \begin{equation}
-p_k( F(\mathbf{x}') ) = \mathcal{N} ( f \mid \Theta_k ).
+p_l( F(\mathbf{x}') ) = \mathcal{N} ( f \mid \Theta_l ).
 \label{eq:multivariate_normal}
@@
 the squared \emph{Mahalanobis distance} of $\mathbf{f}'$ with respect
-of the descriptors of region $k$ as
-$\Delta^2_k (\mathbf{f}' \mid \Theta_k ) = (\mathbf{f}' - \boldsymbol{\mu}_k)^T \, \boldsymbol{\Sigma}^{-1}_k \, (\mathbf{f}' - \boldsymbol{\mu}_k)$,
+of the descriptors of region $l$ as
+$\Delta^2_l (\mathbf{f}' \mid \Theta_l ) = (\mathbf{f}' - \boldsymbol{\mu}_l)^T \, \boldsymbol{\Sigma}^{-1}_l \, (\mathbf{f}' - \boldsymbol{\mu}_l)$,
 we write:
 \begin{align}
-P( F \mid U, \Omega ) &= \underset{k}{\prod} \underset{\mathbf{x} \in \Omega_k}{\prod}
-\mathcal{N} ( f \mid \boldsymbol{\mu}_k, \boldsymbol{\Sigma}_{k} ) \notag\\
-&= \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_l (\mathbf{f}' \mid \Theta_k ) \right)}}.
+P( F \mid U, \Omega ) &= \underset{l}{\prod} \underset{\mathbf{x} \in \Omega_l}{\prod}
+\mathcal{N} ( f \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
+&= \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_l (\mathbf{f}' \mid \Theta_l ) \right)}}.
 \label{eq:pdf}
@@
 \begin{align}
-p_0(u) &= \underset{\mathbf{x}}{\sum} \mathcal{N}( u(\mathbf{x}) \mid 0, \frac{1}{\sqrt{\alpha_{x,y,z}}}) \notag\\
-p_1(u) &= \underset{\mathbf{x}}{\sum} \mathcal{N}( \nabla_{x,y,z}u(\mathbf{x}) \mid 0, \frac{1}{\sqrt{\beta_{x,y,z}}})
+p_0(u) &= \underset{\mathbf{x}}{\sum} \mathcal{N}( u(\mathbf{x}) \mid 0, \frac{1}{\sqrt{a_{x,y,z}}}) \notag\\
+p_1(u) &= \underset{\mathbf{x}}{\sum} \mathcal{N}( \nabla_{x,y,z}u(\mathbf{x}) \mid 0, \frac{1}{\sqrt{b_{x,y,z}}})
 \label{eq:priors}
@@
 \begin{align}
-E(F \mid U) &= -\log \underset{k}{\prod}
-\underset{\mathbf{x} \in \Omega_k}{\prod} 
-p_k( \mathbf{f}')\,p_0( u(\mathbf{x}))\,p_1( u(\mathbf{x})) \notag\\
-&= C + \underset{k}{\sum}
-\underset{\mathbf{x} \in \Omega_k}{\sum}
-\Delta^2_k (\mathbf{f}' \mid \Theta_k ) \notag\\
-&+ \underset{d=\{x,y,z\}}{\sum} \alpha_d u^2_d
-+ \underset{d=\{x,y,z\}}{\sum} \beta_d \|\nabla u_d\|^2,
+E(F \mid U) &= -\log \underset{l}{\prod}
+\underset{\mathbf{x} \in \Omega_l}{\prod} 
+p_l( \mathbf{f}')\,p_0( u(\mathbf{x}))\,p_1( u(\mathbf{x})) \notag\\
+&= C + \underset{l}{\sum}
+\underset{\mathbf{x} \in \Omega_l}{\sum}
+\Delta^2_l (\mathbf{f}' \mid \Theta_l ) \notag\\
+&+ \underset{x,y,z}{\sum} a_d u^2_d
++ \underset{x,y,z}{\sum} b_d \|\nabla u_d\|^2,
 \label{eq:energy}
@@
 
-\subsection{Numerical Implementation}
-\label{sec:numerical_implementation}
-%
-Let us denote $\{c_i\}_{i=1 \ldots N_c}$ the nodes of one or several shape-prior
-surface(s). In our application, precise tissue interfaces of interest 
-extracted from a high-resolution, anatomically correct reference volume. 
-On the other hand, we have a number of \gls{dwi}-derived features at each
-voxel of the volume. Let us denote by $\mathbf{x}$ the voxel and $f(\mathbf{x}) = 
-[ f_1, f_2, \ldots, f_N]^T(\mathbf{x})$ its associated feature vector.
-
-\subsubsection{Deformation field}
-\label{sec:deformation_field}
-The transformation from reference into \gls{dwi} coordinate space is 
-achieved through a dense deformation field $u(\mathbf{x})$, such that:
-\begin{equation}
-c_i' = T\{c_i\} = c_i + u(c_i),
-\end{equation}
-what is equivalent to \eqref{eq:omega_tf}. Since the nodes of the anatomical 
-surfaces might lay off-grid, it is required to derive $u(\mathbf{x})$ from a discrete 
-set of parameters $\{u_k\}_{k=1 \ldots K}$. Densification is achieved through 
-a set of associated basis functions $\psi_k$ (e.g. rbf, interpolation splines):
-%
-\begin{equation}
-u(\mathbf{x}) = \sum_k \psi_k(\mathbf{x}) u_k
-\end{equation}
-%
-In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel.
-Then, the transformation writes
-%
-\begin{equation}
-\label{eq:transformation}
-c_i' = T\{c_i\} = c_i + u(c_i) = c_i + \sum_k \psi_k(c_i)\,u_k
-\end{equation}
-
-Based on the current estimate of the distortion $u$, we can compute 
-``expected samples'' within the shape prior projected into the \gls{dwi}.
-Thus, we now estimate region descriptors of the \gls{dwi} features 
-$f(\mathbf{x})$ of the regions defined by the priors in \gls{dwi} space.
-%
-Using the region descriptors derived in \autoref{sec:methods_map}, we propose
+{\color{red} {Using the region descriptors derived in \autoref{sec:methods_map}, we propose
 an \gls{adf}-like, \gls{acwe}-based, piece-wise constant, image segmentation
@@
 regularization \cite{nagel_investigation_1986} to better regularize
-the \gls{epi} distortion.
+the \gls{epi} distortion.}}
 
-\subsubsection{Shape gradients}
-To compute the gradient-descent of the data-term domain integrals
-in \eqref{eq:complete_energy} with respect to the underlying 
-deformation field, we use shape gradients 
+\todo[inline]{this last paragraphs in red color were located a bit
+further, but I think we can keep some of the refs and still reproduce
+the continuous version of the energy functional}
+
+\subsection{Numerical Implementation}
+\label{sec:numerical_implementation}
+%
+Let us denote $\{c_i\}_{i=1 \ldots N_c}$ the nodes of one or several shape-prior
+surface(s). In our application, precise tissue interfaces of interest 
+extracted from a high-resolution, anatomically correct reference volume. 
+On the other hand, we have a number of \gls{dwi}-derived features at each
+voxel of the volume. Let us denote by $\mathbf{x}$ the voxel and $F(\mathbf{x}) = 
+[ f_1, f_2, \ldots, f_N]^T(\mathbf{x})$ its associated feature vector.
+
+\subsubsection{Deformation field}
+\label{sec:deformation_field}
+The transformation from ``T1-'' into ``\gls{dwi}-'' coordinate space is 
+achieved through a dense deformation field $u(\mathbf{x})$, such that:
+\begin{equation}
+c_i' = U\{c_i\} = c_i + u(c_i),
+\label{eq:nodes_tfm}
+\end{equation}
+where $U$ is defined in \eqref{eq:deformation_field}. Since the nodes of the anatomical 
+surfaces might lay off-grid, it is required to derive $u(\mathbf{x})$ from a discrete 
+set of parameters $\{u_k\}_{k=1 \ldots K}$. Densification is achieved through 
+a set of associated basis functions $\psi_k$:
+%
+\begin{equation}
+u(\mathbf{x}) = \sum_k \psi_k(\mathbf{x}) u_k
+\label{eq:intp_kernel}
+\end{equation}
+%
+In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
+of degree 3 ($B_3$).
+Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
+$\psi$ by the actual kernel function, the transformation writes
+%
+\begin{equation}
+c_i' = c_i + \sum_k \left[ u_k \, \underset{x,y,x}{\prod} B_3(c_{i,d} - \mathbf{x}_{k,d}) \right].
+\label{eq:transformation}
+\end{equation}
+
+Based on the current estimate of the distortion $u$, we can compute 
+``expected samples'' within the shape prior projected into the \gls{dwi}.
+Thus, we now estimate region descriptors of the \gls{dwi} features 
+$\mathbf{f}'$ of the regions defined by the priors in \gls{dwi} space.
+
+\subsubsection{Gradient descent}
+To find the minimum of the energy functional \eqref{eq:energy},
+we propose a gradient-descent approach with respect to the underlying 
+deformation field through the following partial differential equation:
+\begin{equation}
+\frac{\partial u(\mathbf{x},t)}{\partial t} \propto - \frac{\partial E(u)}{\partial u},
+\label{eq:general_gradient_descent}
+\end{equation}
+with $t$ being an artificial time parameter of the contour
+evolution, included to express the iterative character of the approach.
+Now, we can introduce \eqref{eq:energy} in
+\eqref{eq:general_gradient_descent}:
+\begin{align}
+\frac{\partial E(u)}{\partial u^d_k} &=
+\frac{ \partial }{\partial u^d_k} \lbrace
+C + \underset{l}{\sum}
+\underset{\mathbf{x} \in \Omega_l}{\sum}
+\Delta^2_l (\mathbf{f}' \mid \Theta_l ) \notag\\
+&+ \underset{a}{\sum} a^d ({u^d})^2
++ \underset{a}{\sum} b^d \|\nabla u^d\|^2
+\rbrace
+\label{eq:gradient_descent}
+\end{align}
+
+Whereas related \glspl*{adf} introduced in \autoref{sec:methods_background}
+make use of explicit level-set formulations to solve \eqref{eq:gradient_descent},
+we alternatively use \emph{shape-gradients}
 \cite{jehan-besson_dream2s:_2003,herbulot_segmentation_2006}.
-
-Let $\Omega$ be an image domain and $d\Omega$ its boundary. Further, 
-$r(\mathbf{x})$ is an ``arbitrary'' function over the image domain. 
-We now derive the domain integral w.r.t. the contour evolution 
-parameter $t$ (time):
+Let $r(\mathbf{x})$ be an ``arbitrary'' function over the image domain,
+and $\Omega$ a certain image region with $\Gamma_{ij}$ its corresponding
+outer boundary as defined in \autoref{sec:methods_map}.
+We now derive the domain integral w.r.t. $t$:
+\begin{equation*}
+\frac{\partial}{\partial t} \int_\Omega r(\mathbf{x}') d\mathbf{x} = \int_\Omega \frac{\partial}{\partial t}r(\mathbf{x}') d\mathbf{x} - \int_{\Gamma_{ij}} r(\mathbf{x}') \left\langle \frac{\partial \Gamma_{ij} }{\partial t}, N_{\Gamma_{ij}}\right\rangle d\mathbf{x}
+\end{equation*}
+where $\left\langle\frac{\partial\Gamma_{ij}}{\partial t}, N_{\Gamma_{ij}}\right\rangle$ is 
+the projection of the boundary movement on the unit inward normal $N_{\Gamma_{ij}}$. Assuming
+that the region descriptors $\Theta_l$ vary slowly enough, we can consider
+that $\frac{\partial}{\partial t} r(\mathbf{x}') = 0$ and thus:
 \begin{equation}
-\frac{\partial}{\partial t} \int_\Omega r(\mathbf{x}) d\mathbf{x} = \int_\Omega \frac{\partial r}{\partial t}(\mathbf{x}) d\mathbf{x} - \int_{d\Omega} r(\mathbf{x}) \left\langle \frac{\partial{d\Omega}}{\partial t}, N_{d\Omega}\right\rangle d\mathbf{x}
+\frac{\partial}{\partial t} \int_\Omega r(\mathbf{x}') d\mathbf{x} = 
+- \int_{\Gamma_{ij}} r(\mathbf{x}') \left\langle \frac{\partial \Gamma_{ij} }{\partial t}, N_{\Gamma_{ij}}\right\rangle d\mathbf{x}
 \label{eq:shape_gradients}
 \end{equation}
-where $\left\langle\frac{\partial{d\Omega}}{\partial t}, N_{d\Omega}\right\rangle$ is 
-the projection of the boundary movement on the unit inward normal. We recall
-here that equation \eqref{eq:shape_gradients} stands the bridge between the 
-explicit level set formulations surveyed in \autoref{sec:methods_background}
-by updating the distortion along the steepest energy descent:
-\begin{equation}
-\frac{\partial u(\mathbf{x},t)}{\partial t} = - \frac{\partial E(u)}{\partial u},
-\end{equation}
-what closes the loop with the general expression of the \glspl{adf} introduced in
-\eqref{eq:active_deformation_fields}.
 
-\subsubsection{Operator splitting}
-In order to make the energy minimization computationally more tractable, 
-we propose the following operator splitting: Let us optimize the data terms 
-and the regularity terms on separate copies of the deformation field, 
-now called $u$ and $v$, constrained to be equal:
-\begin{multline}
-E(u,v) = \sum\limits_l \int_{\Omega'_l(u)} \Delta^2_l (\mathbf{f}) \,d\mathbf{x} \\
-+ \int v^T \, A \, v \, d\mathbf{x} + \int \tr\{(\nabla v^T)^T B (\nabla v^T)\} d\mathbf{x}
-\label{eq:operator_splitting}
-\end{multline}
-and now
-\begin{equation}
-\min_{u,v} \{ E \} \quad s.t. \quad u = v.
-\end{equation}
+Therefore, we can apply a discretized interpretation of \eqref{eq:shape_gradients}
+to compute the data term in \eqref{eq:gradient_descent} as follows:
+\begin{align}
+\frac{\partial E_{data}(u)}{\partial u^d_k} &=
+\underset{l}{\sum} \frac{ \partial }{\partial u^d_k} \lbrace
+ \underset{\mathbf{x} \in \Omega_l}{\sum}
+\Delta^2_l (\mathbf{f}') \rbrace \notag\\
+&= \underset{ij}{\sum} \underset{\mathbf{c}' \in \Gamma_{ij}}{\sum}
+\left[ \Delta^2_i (\mathbf{f}' ) - \Delta^2_j (\mathbf{f}')\right]
+\left\langle \frac{\partial \mathbf{c}'}{\partial {u^a_k}'}, \mathbf{n_i}'\right\rangle,
+\label{eq:gradient_wshape}
+\end{align}
+in this case, the formulation has been adapted to the non-binary case, $ij$
+being any pair of neighboring regions, and $\Gamma_{ij}$ the contour separating
+them such that $\mathbf{x}' = \mathbf{c}' \in\Gamma_{ij} \iff \mathbf{x}\in \partial\Omega_i \cap \partial\Omega_j$
+and $\mathbf{n_i}'$ is the unit inward normal to the contour at $c_i'$.
 
-In order to take the equality constraint into account, we may make use of 
-augmented Lagrangians (a combination of Lagrangian multipliers and penalty 
-terms on the constraint) \cite{bertsekas_multiplier_1976,
-glowinski_augmented_1989,nocedal_numerical_2006}:
-\begin{equation}
-AL(u,v,\lambda,r) = E(u,v) + \langle \lambda, u-v \rangle + \frac{r}{2} \| u - v \|_2^2
-\end{equation}
+Finally, we can compute:
+\begin{align}
+\frac{\partial}{\partial {u^a_k}'}{\mathbf{c}_i}' &= \frac{\partial}{\partial {u^a_k}'} 
+\lbrace c_i + \sum_k \psi_k(\mathbf{c_i}) u_k \rbrace
+= \psi_k(\mathbf{c_i})\, \hat{\mathbf{e}}^a
+\end{align}
+where $\hat{\mathbf{e}}^a$ is the unit vector along direction $a$. Projecting
+this gradient onto the surface normal,
+$\left\langle \frac{\partial \mathbf{c}'}{\partial {u^a_k}'}, \mathbf{n_i}'\right\rangle
+= \psi_k(\mathbf{c_i})\, \hat{\mathbf{n}}_i^a$, the
+full gradient evolution equation \eqref{eq:gradient_descent} yields:
+\begin{align}
+\frac{\partial E(u)}{\partial u^d_k} 
+&= \underset{ij}{\sum} \underset{\mathbf{c}' \in \Gamma_{ij}}{\sum}
+\left[ \Delta^2_i (\mathbf{f}' ) - \Delta^2_j (\mathbf{f}')\right]
+\psi_k(\mathbf{c_i})\, \hat{\mathbf{n}}_i^a \notag\\
+&2\, \underset{x,y,z}{\sum} a_d u^a_{k,d}
+-2\, \underset{x,y,z}{\sum} b_d \nabla u^a_{k,d},
+\label{eq:gradient_final}
+\end{align}
 
-To solve the constraint minimization problem, we may now optimize the 
-Augmented Lagrangian in an iterative way:
-\begin{equation}
-\left\lbrace 
-\begin{array}{rcl}
-u^{t+1} &=& \argmin_{u} AL(u,v^t,\lambda^t,r)\\
-v^{t+1} &=& \argmin_{v} AL(u^{t+1},v,\lambda^t,r)\\
-\lambda^{t+1} &=& \lambda^t + \rho(u^{t+1}-v^{t+1})
-\end{array}\right.
-\end{equation}
-where typically $0 < \rho < r$. The two sub-minimization problems will 
-now be much easier to handle than the original complete problem.
+\subsubsection{Semi-implicit Euler-forward optimization}
+It is necessary to discretize $t$ in order to obtain a numerical
+implementation of the equation \eqref{eq:gradient_final}:
+\begin{align}
+\frac{u^{t+1}-u^{t}}{\tau} 
+&= - \underset{ij}{\sum} \underset{\mathbf{c}' \in \Gamma_{ij}}{\sum}
+\left[ \Delta^2_i (\mathbf{f}' ) - \Delta^2_j (\mathbf{f}')\right]
+\psi_k(\mathbf{c_i})\, \hat{\mathbf{n}}_i^a \notag\\
+&-2\, \underset{x,y,z}{\sum} a_d (u^a_{k,d})^{t+1}
++2\, \underset{x,y,z}{\sum} b_d \nabla (u^a_{k,d})^{t+1}.
+\end{align}
 
-\subsubsection{Minimization w.r.t. $u$}
-The first minimization problem optimizes the data-term. The problem is equivalent 
-to minimizing the following energy:
-\begin{multline}
-E(u,v) = \sum\limits_l \int_{\Omega'_l(u)} \Delta^2_l (\mathbf{f}) \,d\mathbf{x}
-+ \langle \lambda, u-v \rangle + \frac{r}{2} \| u - v \|_2^2
-\end{multline}
-where we identify one instance of domain integrals of the form $\int_\Omega r(\mathbf{x}) 
-d\mathbf{x}$ for each region $l$. Optimality requires the derivative of this energy 
-with respect to the parameters $u$ to be null. At this point, we decide to ignore 
-the influence of the boundary shift on the statistics of the regions (i.e. moving the 
-boundary does not significantly impact the $\mu$ and $\Sigma$ descriptors). This means 
-that we can drop the derivative of $r(\mathbf{x})$ w.r.t. contour evolution. 
-What remains are the surface integrals at the subdomain interfaces, plus the Lagrangian 
-and penalty terms:
-
-\begin{multline}
-\frac{\partial E}{\partial u_k} = \sum\limits_l \int_{d\Omega'_l} \left[ \Delta^2_{O(l)} (\mathbf{f}) - \Delta^2_l (\mathbf{f}) \right]
-\left\langle\frac{\partial{d\Omega'_l}}{\partial t}, N_{d\Omega'_l}\right\rangle \,d\mathbf{x} \\
-+ \lambda_k + r(u_k - v_k) = 0
-\label{eq:energy_gradient}
-\end{multline}
-where $O(l)$ is a function that returns the neighboring region at $\mathbf{x}$ (i.e.
-the region \emph{outside} the current location in the contour). For the shake of
-simple notation, we will refer as $c_i$ to the positions 
-of points belonging to any of the existing contours $d\Omega_l$ (with 
-$i \in [1 \ldots N_c]$ and $N_c$ the total number of control points).
-Given the deformation field interpolation stated in \autoref{sec:deformation_field}, 
-the boundary moves according to
-\begin{equation}
-\frac{\partial c_i}{\partial u_k} = \psi_k(c_i)\,e_a
-\end{equation}
-where $e_a$ is the unit vector along direction $a \in \mathbb{R}^n$ and $n$ the dimension
-of the image. Thus, if we name the inward unit normal of the corresponding transformed 
-contour at position $c'_i$ as $\hat{\mathbf{n}}_{c'_i}$,
-\begin{equation}
-\left\langle\frac{\partial c_i}{\partial u_k}, N_{c_i}\right\rangle = \psi_k(c_i)\,\hat{\mathbf{n}}_{c'_i}.
-\end{equation}
-The discrete implementation of \eqref{eq:energy_gradient} is straightforward as
-we have a triangularized mesh representation of the interfacing surfaces:
-\begin{multline}
-\frac{\partial E}{\partial u_k} = \sum\limits_{N_c} \left[ \Delta^2_{O(l)} (\mathbf{f}) - \Delta^2_l (\mathbf{f}) \right]
-\, \psi_{c_i}(\mathbf{x}_k)\,\hat{\mathbf{n}}_{c'_i} \\
-+ \lambda_k + r(u_k - v_k) = 0,
-\label{eq:discrete_energy_gradient}
-\end{multline}
-where we have swapped $\psi_k(c_i)$ by $\psi_{c_i}(\mathbf{x}_k)$ for
-convenience in implementation, taking the advantage that they are
-equivalent. Subsequently, the optimal distortion $u_k$ 
-is found at each iteration as:
-\begin{multline}
-u_k^{t+1} = v_k^t - \frac{1}{r}\lambda_k^{t} \\
-- \frac{1}{r} \sum\limits_{N_c} \left[ \Delta^2_{O(l)} (\mathbf{f}) - \Delta^2_l (\mathbf{f}) \right]
-\, \psi_{c_i}(\mathbf{x}_k)\,\hat{\mathbf{n}}_{c'_i}
-\label{eq:update_u}
-\end{multline}
-
-\subsubsection{Minimization w.r.t. $v$}
-For the optimization w.r.t. $v$, the relevant energy writes
-\begin{align}
-E(v) &= \int v^T A v d\mathbf{x} + \int \tr\{(\nabla v^T)^T B (\nabla v^T)\} d\mathbf{x}\\
-&\quad + \langle \lambda, u-v \rangle + \frac{r}{2} \| u - v \|_2^2\nonumber
-\end{align}
 The associated Euler-Lagrange equation is found as:
 \begin{equation}
-2 ( A v - B \Delta v ) + rv = ru + \lambda
+(\tau^{-1} +2\, a_d - 2\, b_d \nabla )\,(u^a_{k,d})^{t+1} = 
+\tau^{-1} (u^a_{k,d})^{t} - \frac{\partial E_{data}(u)}{\partial u^d_k},
 \end{equation}
-which easily translates into Fourier domain:
-\begin{equation}
-v^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{ru + \lambda\}}{\mathcal{F}\{( 2 A+r)\mathcal{I}-2 B \Delta\}} \right\}
-\end{equation}
-where $\mathcal{I}$ denotes the identity operator.
+that is a linear system that we translate into Fourier domain,
+to obtain the next deformation field $(u^a_{k,d})^{t+1}$:
+\begin{multline}
+(u^a_{k,d})^{t+1} = \\
+ \mathcal{F}^{-1} \left\lbrace
+\frac{\mathcal{F}\lbrace \tau^{-1} (u^a_{k,d})^{t} - \underset{ij}{\sum} \underset{\mathbf{c}' \in \Gamma_{ij}}{\sum}
+\left[ \Delta^2_i (\mathbf{f}' ) - \Delta^2_j (\mathbf{f}')\right] \rbrace}
+     {\mathcal{F}\lbrace (\tau^{-1} +2\,a_d)\mathcal{I} - 2\,b_d \Delta ) \rbrace}
+     \right\rbrace
+\end{multline}
 
@@
 
-\subsubsection{Lagrangian multiplier update}
-At each iteration, the Lagrangian multipliers are updated as noted before:
-\begin{equation}
-\lambda^{t+1} = \lambda^t + \rho(u^{t+1}-v^{t+1})
-\end{equation}
-
 \subsubsection{Region descriptor reestimation}
@@
 \subsubsection{Convergence}
-In order to ``fixate'' the evolution when close to convergence,
-it is advised to slightly increase the penalty weight $r$ at each 
-iteration. Note that as can be seen in the above equations, $r$ 
-governs the step-size or leash-length at each iteration, i.e. the 
-amount by which the new estimate $u$ may move away from the preceding 
-$v$ and vice-versa.
+\todo[inline]{discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition, etc.}
 
@@
 sparse matrix for fast densification. As well, we achieve a 
-diffeomorphic transform by ... 
-\todo[inline]{this is actually false (IMHO), so far :(. We will need
-to add some more conditions to the interopolating kernel so that
-the original values of the transform in $c_i$ can be recovered from 
-interpolated values of $u_k$}
+diffeomorphic transform by

### 2013-10-15 Publications/2013-TMI/04-methods.tex
Making mathematical notation homogeneous
Still WIP

@@
 U\colon T \subset \mathbb{R}^n &\to D \subset \mathbb{R}^n \notag\\
-\mathbf{x} &\mapsto \mathbf{x}' =\mathbf{x}+u(\mathbf{x}),
+\vec{x} &\mapsto \vec{x}' =\vec{x}+u(\vec{x}),
 \label{eq:deformation_field}
@@
 \begin{equation}
-P(U \mid F,\Gamma_{ij}) = \frac{P(F \mid U,\Gamma_{ij})\, P(U)}{P(F)},
+P(U \mid F,\Gamma_{l,m}) = \frac{P(F \mid U,\Gamma_{l,m})\, P(U)}{P(F)},
 \label{eq:bayes_rule}
 \end{equation}
-where $P(F \mid U,\Gamma_{ij})$ is the data-likelihood, and
-$\Gamma_{ij}$ are a set of contours in T1-space that correspond to
+where $P(F \mid U,\Gamma_{l,m})$ is the data-likelihood, and
+$\Gamma_{l,m}$ are a set of contours in T1-space that correspond to
 two interfacing regions $\Omega$, such that 
-$\Gamma_{ij} = \partial \Omega_i \cap \partial \Omega_j$.
+$\Gamma_{l,m} = \partial \Omega_l \cap \partial \Omega_m$.
 
@@
 \begin{align}
-\hat{U} &= \underset{U}{\argmax} \left\{ P(U \mid F,\Gamma_{ij}) \right\} \notag\\
- &= \underset{U}{\argmax} \left\{ P(F \mid U,\Gamma_{ij})\, P(U) \right\}.
+\hat{U} &= \underset{U}{\argmax} \left\{ P(U \mid F,\Gamma_{l,m}) \right\} \notag\\
+ &= \underset{U}{\argmax} \left\{ P(F \mid U,\Gamma_{l,m})\, P(U) \right\}.
 \label{eq:map_u}
@@
 \begin{equation}
-P(F \mid U,\Gamma_{ij}) = 
-\underset{\mathbf{x}\in \Omega}{\prod} P\left( F(\mathbf{x}') \mid U(\Gamma_{ij}) \right).
+P(F \mid U,\Gamma_{l,m}) = 
+\underset{\vec{x}\in \Omega}{\prod} P\left( F(\vec{x}') \mid U(\Gamma_{l,m}) \right).
 \label{eq:bayes_aposteriori}
@@
 \begin{equation}
-P\left( F(\mathbf{x}') \mid U,\Gamma_{ij} \right) = p_l( F(\mathbf{x}')) \quad
-\forall \mathbf{x} \in \Omega_l.
+P\left( F(\vec{x}') \mid U,\Gamma_{l,m} \right) = p_l( F(\vec{x}')) \quad
+\forall \vec{x} \in \Omega_l.
 \label{eq:likelihood}
@@
 \begin{equation}
-p_l( F(\mathbf{x}') ) = \mathcal{N} ( f \mid \Theta_l ).
+p_l( F(\vec{x}') ) = \mathcal{N} \left( F(\vec{x}') \mid \Theta_l \right).
 \label{eq:multivariate_normal}
@@
 \todo{I modified this slightly}
-Thus, defining the feature observed as $\mathbf{f}'=F(\mathbf{x}')$ and 
-the squared \emph{Mahalanobis distance} of $\mathbf{f}'$ with respect
+Thus, defining the feature observed as $\vec{f}=F(\vec{x})$ and 
+the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
 of the descriptors of region $l$ as
-$\Delta^2_l (\mathbf{f}' \mid \Theta_l ) = (\mathbf{f}' - \boldsymbol{\mu}_l)^T \, \boldsymbol{\Sigma}^{-1}_l \, (\mathbf{f}' - \boldsymbol{\mu}_l)$,
+\begin{equation}
+\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l),
+\end{equation}
 we write:
 \begin{align}
-P( F \mid U, \Omega ) &= \underset{l}{\prod} \underset{\mathbf{x} \in \Omega_l}{\prod}
-\mathcal{N} ( f \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
-&= \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_l (\mathbf{f}' \mid \Theta_l ) \right)}}.
+P( F \mid U, \Omega ) &= \underset{l}{\prod} \underset{\vec{x} \in \Omega_l}{\prod}
+\mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
+&= \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
+\mdist{f'}{l} \right)}}.
 \label{eq:pdf}
@@
 We choose a Thikonov regularization prior as follows:
-
 \begin{equation*}
-P(U) = \underset{\mathbf{x}}{\prod} p(u)
+P(U) = \underset{\vec{x}}{\prod} p(u(\vec{x}))
 \end{equation*}
@@
 \begin{align*}
-p(u) &= p_0(u) \, p_1(u) \notag\\
-p_0(u) &= \mathcal{N}( u \mid 0, A^{-1}) \notag\\
-p_1(u) &= \mathcal{N}( Du \mid 0, B^{-1}),
+p(u(\vec{x})) &= p_0(u(\vec{x})) \, p_1(u(\vec{x})) \notag\\
+p_0(u(\vec{x})) &= \mathcal{N}( u(\vec{x}) \mid 0, A^{-1}) \notag\\
+p_1(u(\vec{x})) &= \mathcal{N}( Du(\vec{x}) \mid 0, B^{-1}),
 \end{align*}
@@
 \begin{align}
-p_0(u) &= \underset{\mathbf{x}}{\sum} \mathcal{N}( u(\mathbf{x}) \mid 0, \frac{1}{\sqrt{a_{x,y,z}}}) \notag\\
-p_1(u) &= \underset{\mathbf{x}}{\sum} \mathcal{N}( \nabla_{x,y,z}u(\mathbf{x}) \mid 0, \frac{1}{\sqrt{b_{x,y,z}}})
+p_0(u(\vec{x})) &= \underset{\vec{x}}{\sum} \mathcal{N}( u(\vec{x}) \mid 0, (\boldsymbol{\alpha}^{\circ\frac12}\,\vec{I}_n)^{-1}) \notag\\
+p_1(u(\vec{x})) &= \underset{\vec{x}}{\sum} \mathcal{N}( \nabla \cdot u(\vec{x}) \mid 0, (\boldsymbol{\beta}^{\circ\frac12}\,\vec{I}_n)^{-1})
 \label{eq:priors}
@@
 E(F \mid U) &= -\log \underset{l}{\prod}
-\underset{\mathbf{x} \in \Omega_l}{\prod} 
-p_l( \mathbf{f}')\,p_0( u(\mathbf{x}))\,p_1( u(\mathbf{x})) \notag\\
+\underset{\vec{x} \in \Omega_l}{\prod} 
+p_l( \vec{f}')\,p_0( u(\vec{x}))\,p_1( u(\vec{x})) \notag\\
 &= C + \underset{l}{\sum}
-\underset{\mathbf{x} \in \Omega_l}{\sum}
-\Delta^2_l (\mathbf{f}' \mid \Theta_l ) \notag\\
-&+ \underset{x,y,z}{\sum} a_d u^2_d
-+ \underset{x,y,z}{\sum} b_d \|\nabla u_d\|^2,
+\underset{\vec{x} \in \Omega_l}{\sum}
+\mdist{f'}{l} \notag\\
+&+ \underset{\vec{x} \in \Omega}{\sum} \boldsymbol{\alpha} \cdot u(\vec{x})^{\circ2}
++ \underset{\vec{x} \in \Omega}{\sum} \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{x}))^{\circ2},
 \label{eq:energy}
@@
 \begin{multline}
-E(u) = \sum\limits_l \int_{\Omega'_l} \Delta^2_l (\mathbf{f}) \,d\mathbf{x} \\
-+ \int u^T \, A \, u \, d\mathbf{x} + \int \tr\{(\nabla u^T)^T B (\nabla u^T)\} d\mathbf{x}
+E(u) = \sum\limits_l \int_{\Omega'_l} \mdist{f}{l} \,d\vec{x} \\
++ \int \vec{u}^T \, A \, \vec{u} \, d\vec{x} + \int \tr\{(\nabla \vec{u}^T)^T B (\nabla \vec{u}^T)\} d\vec{x}
 \label{eq:complete_energy}
@@
 %
-Let us denote $\{c_i\}_{i=1 \ldots N_c}$ the nodes of one or several shape-prior
+Let us denote $\{\vec{c}_i\}_{i=1 \ldots N_c}$ the nodes of one or several shape-prior
 surface(s). In our application, precise tissue interfaces of interest 
@@
 On the other hand, we have a number of \gls{dwi}-derived features at each
-voxel of the volume. Let us denote by $\mathbf{x}$ the voxel and $F(\mathbf{x}) = 
-[ f_1, f_2, \ldots, f_N]^T(\mathbf{x})$ its associated feature vector.
+voxel of the volume. Let us denote by $\vec{x}$ the voxel and $F(\vec{x}) = 
+[ f_1, f_2, \ldots, f_N]^T(\vec{x})$ its associated feature vector.
 
@@
 The transformation from ``T1-'' into ``\gls{dwi}-'' coordinate space is 
-achieved through a dense deformation field $u(\mathbf{x})$, such that:
+achieved through a dense deformation field $u(\vec{x})$, such that:
 \begin{equation}
-c_i' = U\{c_i\} = c_i + u(c_i),
+\vec{c}_i' = U\{\vec{c}_i\} = \vec{c}_i + u(\vec{c}_i),
 \label{eq:nodes_tfm}
@@
 where $U$ is defined in \eqref{eq:deformation_field}. Since the nodes of the anatomical 
-surfaces might lay off-grid, it is required to derive $u(\mathbf{x})$ from a discrete 
-set of parameters $\{u_k\}_{k=1 \ldots K}$. Densification is achieved through 
+surfaces might lay off-grid, it is required to derive $u(\vec{x})$ from a discrete 
+set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$. Densification is achieved through 
 a set of associated basis functions $\psi_k$:
@@
 \begin{equation}
-u(\mathbf{x}) = \sum_k \psi_k(\mathbf{x}) u_k
+u(\vec{x}) = \sum_k \psi_k(\vec{x}) \vec{u}_k
 \label{eq:intp_kernel}
@@
 \begin{equation}
-c_i' = c_i + \sum_k \left[ u_k \, \underset{x,y,x}{\prod} B_3(c_{i,d} - \mathbf{x}_{k,d}) \right].
+c_i' = c_i + \sum_k \left[ \vec{u}_k \, \underset{x,y,x}{\prod} B_3(c_{i,d} - \vec{x}_{k,d}) \right].
 \label{eq:transformation}
@@
 Thus, we now estimate region descriptors of the \gls{dwi} features 
-$\mathbf{f}'$ of the regions defined by the priors in \gls{dwi} space.
+$\vec{f}'$ of the regions defined by the priors in \gls{dwi} space.
 
@@
 \begin{equation}
-\frac{\partial u(\mathbf{x},t)}{\partial t} \propto - \frac{\partial E(u)}{\partial u},
+\frac{\partial u(\vec{x},t)}{\partial t} \propto - \frac{\partial E(u)}{\partial u},
 \label{eq:general_gradient_descent}
@@
 \begin{align}
-\frac{\partial E(u)}{\partial u^d_k} &=
-\frac{ \partial }{\partial u^d_k} \lbrace
+\frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
+\frac{ \partial }{\partial \vec{u}_k} \Big\{
 C + \underset{l}{\sum}
-\underset{\mathbf{x} \in \Omega_l}{\sum}
-\Delta^2_l (\mathbf{f}' \mid \Theta_l ) \notag\\
-&+ \underset{a}{\sum} a^d ({u^d})^2
-+ \underset{a}{\sum} b^d \|\nabla u^d\|^2
-\rbrace
+\underset{\vec{x} \in \Omega_l}{\sum} \mdist{f'}{l} \notag\\
+&+ \underset{\vec{x} \in \Omega}{\sum} \boldsymbol{\alpha} \cdot u(\vec{x})^{\circ2}
++ \underset{\vec{x} \in \Omega}{\sum} \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{x}))^{\circ2}
+\Big\}.
 \label{eq:gradient_descent}
@@
 \cite{jehan-besson_dream2s:_2003,herbulot_segmentation_2006}.
-Let $r(\mathbf{x})$ be an ``arbitrary'' function over the image domain,
-and $\Omega$ a certain image region with $\Gamma_{ij}$ its corresponding
+Let $r(\vec{x})$ be an ``arbitrary'' function over the image domain,
+and $\Omega$ a certain image region with $\Gamma_{l,m}$ its corresponding
 outer boundary as defined in \autoref{sec:methods_map}.
@@
 \begin{equation*}
-\frac{\partial}{\partial t} \int_\Omega r(\mathbf{x}') d\mathbf{x} = \int_\Omega \frac{\partial}{\partial t}r(\mathbf{x}') d\mathbf{x} - \int_{\Gamma_{ij}} r(\mathbf{x}') \left\langle \frac{\partial \Gamma_{ij} }{\partial t}, N_{\Gamma_{ij}}\right\rangle d\mathbf{x}
+\frac{\partial}{\partial t} \int_\Omega r(\vec{x}') d\vec{x} = \int_\Omega \frac{\partial}{\partial t}r(\vec{x}') d\vec{x} - \int_{\Gamma_{l,m}} r(\vec{x}') \left\langle \frac{\partial \Gamma_{l,m} }{\partial t}, N_{\Gamma_{l,m}}\right\rangle d\vec{x}
 \end{equation*}
-where $\left\langle\frac{\partial\Gamma_{ij}}{\partial t}, N_{\Gamma_{ij}}\right\rangle$ is 
-the projection of the boundary movement on the unit inward normal $N_{\Gamma_{ij}}$. Assuming
+where $\left\langle\frac{\partial\Gamma_{l,m}}{\partial t}, N_{\Gamma_{l,m}}\right\rangle$ is 
+the projection of the boundary movement on the unit inward normal $N_{\Gamma_{l,m}}$. Assuming
 that the region descriptors $\Theta_l$ vary slowly enough, we can consider
-that $\frac{\partial}{\partial t} r(\mathbf{x}') = 0$ and thus:
+that $\frac{\partial}{\partial t} r(\vec{x}') = 0$ and thus:
 \begin{equation}
-\frac{\partial}{\partial t} \int_\Omega r(\mathbf{x}') d\mathbf{x} = 
-- \int_{\Gamma_{ij}} r(\mathbf{x}') \left\langle \frac{\partial \Gamma_{ij} }{\partial t}, N_{\Gamma_{ij}}\right\rangle d\mathbf{x}
+\frac{\partial}{\partial t} \int_\Omega r(\vec{x}') d\vec{x} = 
+- \int_{\Gamma_{l,m}} r(\vec{x}') \left\langle \frac{\partial \Gamma_{l,m} }{\partial t}, N_{\Gamma_{l,m}}\right\rangle d\vec{x}
 \label{eq:shape_gradients}
@@
 \begin{align}
-\frac{\partial E_{data}(u)}{\partial u^d_k} &=
+\frac{\partial E_{data}(u)}{\partial \vec{u}_k} &=
 \underset{l}{\sum} \frac{ \partial }{\partial u^d_k} \lbrace
- \underset{\mathbf{x} \in \Omega_l}{\sum}
-\Delta^2_l (\mathbf{f}') \rbrace \notag\\
-&= \underset{ij}{\sum} \underset{\mathbf{c}' \in \Gamma_{ij}}{\sum}
-\left[ \Delta^2_i (\mathbf{f}' ) - \Delta^2_j (\mathbf{f}')\right]
-\left\langle \frac{\partial \mathbf{c}'}{\partial {u^a_k}'}, \mathbf{n_i}'\right\rangle,
+ \underset{\vec{x} \in \Omega_l}{\sum}
+{\mathcal{D}_l^2}^2 (\vec{f}') \rbrace \notag\\
+&= \underset{l,m}{\sum} \underset{\vec{c}' \in \Gamma_{l,m}}{\sum}
+\left[ \Delta^2_i (\vec{f}' ) - \Delta^2_j (\vec{f}')\right]
+\left\langle \frac{\partial \vec{c}'}{\partial {u^a_k}'}, \vec{n_i}'\right\rangle,
 \label{eq:gradient_wshape}
 \end{align}
-in this case, the formulation has been adapted to the non-binary case, $ij$
-being any pair of neighboring regions, and $\Gamma_{ij}$ the contour separating
-them such that $\mathbf{x}' = \mathbf{c}' \in\Gamma_{ij} \iff \mathbf{x}\in \partial\Omega_i \cap \partial\Omega_j$
-and $\mathbf{n_i}'$ is the unit inward normal to the contour at $c_i'$.
+in this case, the formulation has been adapted to the non-binary case, $l,m$
+being any pair of neighboring regions, and $\Gamma_{l,m}$ the contour separating
+them such that $\vec{x}' = \vec{c}' \in\Gamma_{l,m} \iff \vec{x}\in \partial\Omega_i \cap \partial\Omega_j$
+and $\vec{n_i}'$ is the unit inward normal to the contour at $c_i'$.
 
@@
 \begin{align}
-\frac{\partial}{\partial {u^a_k}'}{\mathbf{c}_i}' &= \frac{\partial}{\partial {u^a_k}'} 
-\lbrace c_i + \sum_k \psi_k(\mathbf{c_i}) u_k \rbrace
-= \psi_k(\mathbf{c_i})\, \hat{\mathbf{e}}^a
+\frac{\partial}{\partial {\vec{u}_k}'}{\vec{c}_i}' &= \frac{\partial}{\partial {u^a_k}'} 
+\lbrace c_i + \sum_k \psi_k(\vec{c}_i) \vec{u}_k \rbrace
+= \psi_k(\vec{c_i})\, \hat{\vec{e}}^a
 \end{align}
-where $\hat{\mathbf{e}}^a$ is the unit vector along direction $a$. Projecting
+where $\hat{\vec{e}}^a$ is the unit vector along direction $a$. Projecting
 this gradient onto the surface normal,
-$\left\langle \frac{\partial \mathbf{c}'}{\partial {u^a_k}'}, \mathbf{n_i}'\right\rangle
-= \psi_k(\mathbf{c_i})\, \hat{\mathbf{n}}_i^a$, the
+$\left\langle \frac{\partial \vec{c}'}{\partial {u^a_k}'}, \vec{n_i}'\right\rangle
+= \psi_k(\vec{c_i})\, \hat{\vec{n}}_i^a$, the
 full gradient evolution equation \eqref{eq:gradient_descent} yields:
 \begin{align}
-\frac{\partial E(u)}{\partial u^d_k} 
-&= \underset{ij}{\sum} \underset{\mathbf{c}' \in \Gamma_{ij}}{\sum}
-\left[ \Delta^2_i (\mathbf{f}' ) - \Delta^2_j (\mathbf{f}')\right]
-\psi_k(\mathbf{c_i})\, \hat{\mathbf{n}}_i^a \notag\\
+\frac{\partial E(u)}{\partial \vec{u}_k} 
+&= \underset{l,m}{\sum} \underset{\vec{c}' \in \Gamma_{l,m}}{\sum}
+\left[ \Delta^2_i (\vec{f}' ) - \Delta^2_j (\vec{f}')\right]
+\psi_k(\vec{c_i})\, \hat{\vec{n}}_i^a \notag\\
 &2\, \underset{x,y,z}{\sum} a_d u^a_{k,d}
@@
 \frac{u^{t+1}-u^{t}}{\tau} 
-&= - \underset{ij}{\sum} \underset{\mathbf{c}' \in \Gamma_{ij}}{\sum}
-\left[ \Delta^2_i (\mathbf{f}' ) - \Delta^2_j (\mathbf{f}')\right]
-\psi_k(\mathbf{c_i})\, \hat{\mathbf{n}}_i^a \notag\\
+&= - \underset{l,m}{\sum} \underset{\vec{c}' \in \Gamma_{l,m}}{\sum}
+\left[ \Delta^2_i (\vec{f}' ) - \Delta^2_j (\vec{f}')\right]
+\psi_k(\vec{c_i})\, \hat{\vec{n}}_i^a \notag\\
 &-2\, \underset{x,y,z}{\sum} a_d (u^a_{k,d})^{t+1}
@@
  \mathcal{F}^{-1} \left\lbrace
-\frac{\mathcal{F}\lbrace \tau^{-1} (u^a_{k,d})^{t} - \underset{ij}{\sum} \underset{\mathbf{c}' \in \Gamma_{ij}}{\sum}
-\left[ \Delta^2_i (\mathbf{f}' ) - \Delta^2_j (\mathbf{f}')\right] \rbrace}
+\frac{\mathcal{F}\lbrace \tau^{-1} (u^a_{k,d})^{t} - \underset{l,m}{\sum} \underset{\vec{c}' \in \Gamma_{l,m}}{\sum}
+\left[ \Delta^2_i (\vec{f}' ) - \Delta^2_j (\vec{f}')\right] \rbrace}
      {\mathcal{F}\lbrace (\tau^{-1} +2\,a_d)\mathcal{I} - 2\,b_d \Delta ) \rbrace}
@@
 based on the shifted volumetric samples 
-$\mathbf{x}' = \mathbf{x_0} + u(\mathbf{x})$.
+$\vec{x}' = \vec{x_0} + u(\vec{x})$.
 

### 2013-10-15 Publications/2013-TMI/04-methods.tex
Still working on nomenclature

@@
-\section{Methods}
+1\section{Methods}
 \label{sec:methods}
@@
 energy functional is defined simultaneously in the moving and target images with
-an affine transformation supporting the coordinates mapping (``2 PDEs approach'').
+an affine transformation supporting the coordinates mapping (``2 \glspl*{pde} approach'').
 \citep{vemuri_joint_2003} proposed the first atlas-based registration based on the 
-level set framework and using only one PDE. \citep{unal_coupled_2005} extended the 
+level set framework and using only one \gls*{pde}. \citep{unal_coupled_2005} extended the 
 idea of \citep{bertalmio_morphing_2000,yezzi_variational_2001} for non-linear 
@@
 \begin{equation}
-c_i' = c_i + \sum_k \left[ \vec{u}_k \, \underset{x,y,x}{\prod} B_3(c_{i,d} - \vec{x}_{k,d}) \right].
+\vec{c}_i' = \vec{c}_i + \sum_k \left[ \vec{u}_k \, \underset{n}{\prod} B_3( (\vec{c}_i - \vec{x}_k) \cdot \hat{\mathbf{e}}_d ) \right].
 \label{eq:transformation}
@@
 
-Based on the current estimate of the distortion $u$, we can compute 
+Based on the current estimate of the distortion $\vec{u}$, we can compute 
 ``expected samples'' within the shape prior projected into the \gls{dwi}.
@@
 we propose a gradient-descent approach with respect to the underlying 
-deformation field through the following partial differential equation:
+deformation field through the following \gls*{pde}:
 \begin{equation}
-\frac{\partial u(\vec{x},t)}{\partial t} \propto - \frac{\partial E(u)}{\partial u},
+\frac{\partial u(\vec{x},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}},
 \label{eq:general_gradient_descent}
@@
 \begin{align}
-\frac{\partial E_{data}(u)}{\partial \vec{u}_k} &=
-\underset{l}{\sum} \frac{ \partial }{\partial u^d_k} \lbrace
- \underset{\vec{x} \in \Omega_l}{\sum}
-{\mathcal{D}_l^2}^2 (\vec{f}') \rbrace \notag\\
-&= \underset{l,m}{\sum} \underset{\vec{c}' \in \Gamma_{l,m}}{\sum}
-\left[ \Delta^2_i (\vec{f}' ) - \Delta^2_j (\vec{f}')\right]
-\left\langle \frac{\partial \vec{c}'}{\partial {u^a_k}'}, \vec{n_i}'\right\rangle,
+\frac{\partial E_{data}(\vec{u})}{\partial \vec{u}_k} &=
+\underset{l}{\sum} \frac{ \partial }{\partial \vec{u}_k} \left\{
+ \underset{\vec{x} \in \Omega_l}{\sum} \mdist{f'}{l} \right\} \notag\\
+&= \underset{l,m}{\sum} \underset{i}{\sum}
+\left[ \mdist{f_i'}{l} - \mdist{f_i'}{m} \right]
+\left\langle \frac{\partial \vec{c}_i'}{\partial \vec{u}_k}, \vec{n_i}'\right\rangle,
 \label{eq:gradient_wshape}
@@
 \begin{align}
-\frac{\partial}{\partial {\vec{u}_k}'}{\vec{c}_i}' &= \frac{\partial}{\partial {u^a_k}'} 
-\lbrace c_i + \sum_k \psi_k(\vec{c}_i) \vec{u}_k \rbrace
-= \psi_k(\vec{c_i})\, \hat{\vec{e}}^a
+\frac{\partial \vec{c}_i'}{\partial \vec{u}_k} &= \frac{\partial}{\partial \vec{u}_k} 
+\left\{ \vec{c}_i + \sum_k \psi_k(\vec{c}_i) \vec{u}_k \right\} 
+= \psi_k(\vec{c_i})\, \hat{\vec{e}}
 \end{align}
-where $\hat{\vec{e}}^a$ is the unit vector along direction $a$. Projecting
-this gradient onto the surface normal,
-$\left\langle \frac{\partial \vec{c}'}{\partial {u^a_k}'}, \vec{n_i}'\right\rangle
-= \psi_k(\vec{c_i})\, \hat{\vec{n}}_i^a$, the
+where $\hat{\vec{e}}$ is the coordinates system's unit vector.
+Projecting this gradient onto the surface normal,
+$\left\langle \frac{\partial}{\partial \vec{u}_k}{\vec{c}_i}', \vec{n_i}'\right\rangle
+= \psi_k(\vec{c_i})\, \hat{\vec{n}}_i$, then the
 full gradient evolution equation \eqref{eq:gradient_descent} yields:
 \begin{align}
-\frac{\partial E(u)}{\partial \vec{u}_k} 
-&= \underset{l,m}{\sum} \underset{\vec{c}' \in \Gamma_{l,m}}{\sum}
-\left[ \Delta^2_i (\vec{f}' ) - \Delta^2_j (\vec{f}')\right]
-\psi_k(\vec{c_i})\, \hat{\vec{n}}_i^a \notag\\
-&2\, \underset{x,y,z}{\sum} a_d u^a_{k,d}
--2\, \underset{x,y,z}{\sum} b_d \nabla u^a_{k,d},
+\frac{\partial E(\vec{u}_k)}{\partial \vec{u}_k} =
+&- \underset{l,m}{\sum} \underset{i}{\sum}
+\left[ \mdist{f_i'}{l} - \mdist{f_i'}{m} \right]
+\psi_k(\vec{c}_i)\, \hat{\vec{n}}_i \notag\\
+&+2\, \boldsymbol{\alpha} \vec{u}_k
+-2\, \boldsymbol{\beta} \Delta \vec{u}_k,
 \label{eq:gradient_final}
@@
 \begin{align}
-\frac{u^{t+1}-u^{t}}{\tau} 
-&= - \underset{l,m}{\sum} \underset{\vec{c}' \in \Gamma_{l,m}}{\sum}
-\left[ \Delta^2_i (\vec{f}' ) - \Delta^2_j (\vec{f}')\right]
-\psi_k(\vec{c_i})\, \hat{\vec{n}}_i^a \notag\\
-&-2\, \underset{x,y,z}{\sum} a_d (u^a_{k,d})^{t+1}
-+2\, \underset{x,y,z}{\sum} b_d \nabla (u^a_{k,d})^{t+1}.
+\frac{\vec{u}_k^{t+1}-\vec{u}_k^t}{\tau} =
+&- \underset{l,m}{\sum} \underset{i}{\sum}
+\left[ \mdist{f_i'}{l} - \mdist{f_i'}{m}\right]
+\psi_k(\vec{c}_i)\, \hat{\vec{n}}_i \notag\\
+&+2\, \boldsymbol{\alpha} \vec{u}_k^{t+1}
+-2\, \boldsymbol{\beta} \Delta \vec{u}_k^{t+1}.
 \end{align}
@@
 \begin{equation}
-(\tau^{-1} +2\, a_d - 2\, b_d \nabla )\,(u^a_{k,d})^{t+1} = 
-\tau^{-1} (u^a_{k,d})^{t} - \frac{\partial E_{data}(u)}{\partial u^d_k},
+(\tau^{-1} +2\, \boldsymbol{\alpha} - 2\, \boldsymbol{\beta} \Delta )\, \vec{u}_k^{t+1} = 
+\tau^{-1} \vec{u}_k^t - \frac{\partial}{\partial \vec{u}_k} E_{data}(\vec{u}_k^t),
 \end{equation}
 that is a linear system that we translate into Fourier domain,
-to obtain the next deformation field $(u^a_{k,d})^{t+1}$:
+to obtain the next deformation field $\vec{u}_k^{t+1}$:
 \begin{multline}
-(u^a_{k,d})^{t+1} = \\
+\vec{u}_k^{t+1} = \\
  \mathcal{F}^{-1} \left\lbrace
-\frac{\mathcal{F}\lbrace \tau^{-1} (u^a_{k,d})^{t} - \underset{l,m}{\sum} \underset{\vec{c}' \in \Gamma_{l,m}}{\sum}
-\left[ \Delta^2_i (\vec{f}' ) - \Delta^2_j (\vec{f}')\right] \rbrace}
-     {\mathcal{F}\lbrace (\tau^{-1} +2\,a_d)\mathcal{I} - 2\,b_d \Delta ) \rbrace}
+\frac{\mathcal{F}\lbrace \tau^{-1}\vec{u}_{k}^{t} - \underset{l,m}{\sum} \underset{i}{\sum}
+\left[ \mdist{f_i'}{l} - \mdist{f_i'}{m}\right] \rbrace}
+     {\mathcal{F}\lbrace (\tau^{-1} +2\,\boldsymbol{\alpha})\mathcal{I} - 2\,\boldsymbol{\beta} \Delta ) \rbrace}
      \right\rbrace
@@
 \begin{equation}
-\mathcal{F}\{\mathcal{S}_{d}^{\pm}\} = e^{\pm i\omega_{d}},
+\mathcal{F}\{\mathcal{S}_{d}^{\pm}\} = e^{\pm \vec{i}\omega_{d}},
 \end{equation}
@@
 \begin{equation}
-\mathcal{F}\{\Delta\} = \sum\limits_d e^{-i\omega_d } + e^{i\omega_d } - 2 = \sum\limits_d \left( 2\cos(\omega_d) - 2 \right)
+\mathcal{F}\{\Delta\} = \sum\limits_d e^{-\vec{i}\omega_d } + e^{\vec{i}\omega_d } - 2 = \sum\limits_d \left( 2\cos(\omega_d) - 2 \right)
 \end{equation}

### 2013-12-03 Publications/2013-ISBI14/03-methods.tex
Fixing reviews

@@
 Finally, for evaluating \Gls*{t2b},
-we fine-tuned \emph{ANTs} \cite{avants_ants:_2013},
+we fine-tuned \emph{ANTs} (\url{http://stnava.github.io/ANTs}),
 in order to register \textit{b0} to T2. To this end, we used a
@@
 connectivity matrices reporting the number of
-false positives (inexistent connections in the
+false positives (nonexistent connections in the
 gold-standard) and false negatives (or missed connections).

### 2013-12-03 Publications/2013-ISBI14/04-results.tex
Fixing reviews

@@
    \centering
-   \includegraphics[width=0.9\textwidth]{Fig02-Results}
+   \includegraphics[width=0.80\textwidth]{Fig02-Results}
    \caption{Visual evaluation of the correction methods.
@@
 Although we found connectivity matrices rather invariant with
-respect distortion, they are suspected to be significantly
+respect to distortion, they are suspected to be significantly
 impacted by the susceptibility artifact on real data 
@@
 covered by the regions mapped from the anatomical 
-dataset, what leads to larger sensitivity with respect 
+dataset, what leads to larger sensitivity with respect to
 deformations.

### 2013-12-04 Publications/2013-ISBI14/04-results.tex
Last draft.

@@
    \centering
-   \includegraphics[width=0.80\textwidth]{Fig02-Results}
-   \caption{Visual evaluation of the correction methods.
-   First row represents a coronal section of the \textit{b0} volume. 
-   In second row, the outcome of tractography after filtering tracks
-   that did not connect two regions.
-   Third row shows the associated connectivity matrices. All the matrices
-   are compared to the ground-truth connections in the phantom. The existing
-   connections that were correctly detected at each step are depicted in white
-   color. False positives have been marked in red and false negatives in blue.
-   Gray color represents the true negatives (non-existent connections correctly detected).}
+   \includegraphics[width=0.8\linewidth]{Fig02-Results}
+   \caption{Visual assessment of the correction methods.}
    \label{fig:results}

### 2013-12-10 Publications/2013-ISBI14/02-introduction.tex
Fixing reviews
Final proposal and response letter draft

@@
 \label{sec:intro}
-In-vivo whole-brain connectivity analysis has been a
-research topic of high interest for the last 5 years.
-\Gls*{dmri} can be used to probe the
-orientation of fiber bundles within the brain,
-and is generally acquired with an \gls*{epi} sequence.
-After signal reconstruction,
-tractography algorithms draw a map of the sampled 
+In-vivo whole-brain connectivity analysis has been a research topic of high
+  interest for the last 5 yr.
+\Gls*{dmri} can be used to probe the orientation of fiber bundles within the
+  brain, and it is generally acquired with an \gls*{epi} sequence.
+After signal reconstruction, tractography algorithms draw a map of the sampled
   structures.
-These maps can represent the actual trajectories
-of fiber bundles (deterministic tractography) or
-pixel-wise probability of connection to a certain origin
-(probabilistic tractography). Finally, the
-information about these connections is collected
-into a network matrix that can be subjected to
-the so-called ``connectome analysis''.
+These maps can represent the actual trajectories of fiber bundles
+  (deterministic tractography) or pixel-wise probability of connection
+  to a certain origin (probabilistic tractography).
+Finally, the information about these connections is collected into a network
+  matrix that is subjected to the ``connectome analysis''.
 
-Among all the difficulties that such a complex workflow
-raises \cite{jones_twenty-five_2010}, we will
-address here the susceptibility-derived artifacts,
+Among all the difficulties that such a complex workflow raises, % \cite{jones_twenty-five_2010},
+  we will address here the susceptibility-derived artifacts,
   for which \Gls*{epi} schemes are especially sensitive.
-Magnetic susceptibility disturbs the magnetic field
-close to tissue interfaces. This inhomogeneity
-of the field translates in a highly distorted
-anatomy and a significant signal destruction in
-certain regions of the brain 
-(e.g. the orbitofrontal lobe, for the proximity of the
-air surrounding sinuses). This artifact has been
-well described, generally within the context of
-functional MRI, which also uses \gls*{epi}.
+Magnetic susceptibility disturbs the magnetic field close to tissue interfaces.
+This inhomogeneity of the field translates into a highly distorted anatomy
+  and a significant signal destruction in certain regions of the brain 
+  (e.g. the orbitofrontal lobe, for the proximity of the air surrounding sinuses).
+This artifact has been well described, generally within the context of functional MRI,
+  which also uses \gls*{epi}.
 
-Various approaches have been proposed to correct
-for the distortions. \Gls*{fmb} methods
-\cite{jezzard_correction_1995} rely on one extra
-acquisition (\emph{field mapping}),
-that probes the inhomogeneity of the B0 field.
-A second theory-based breed of methodologies acquire a 
-map of the \acrlong*{psf} of the \gls*{epi} readouts to correct 
-for the artifact \cite{robson_measurement_1997}.
-Another approach referred to as \gls*{reb}, acquires an
-extra \gls*{epi} volume in the orthogonal or reversed
-phase encoding that can be combined to remove the
-geometric distortions
-\cite{cordes_geometric_2000,chiou_simple_2000}.
-The last family of methodologies acquires an
-extra T2-weighted image, and uses it as an anatomical
-reference to find the deformation map through nonlinear 
-registration \cite{kybic_unwarping_2000,studholme_accurate_2000}.
-These \gls*{t2b} methods usually map
-the T2 image to the \emph{baseline} volume or \textit{b0}
-of \gls*{dmri}. The choice of T2 is due to the strong
-similarity of intensity distribution with respect the \textit{b0}.
-More recent works report extensions or combined
-approaches of existing techniques
+Various approaches have been proposed to correct for this distortion.
+\Gls*{fmb} methods \cite{jezzard_correction_1995} rely on one extra
+  acquisition (\emph{field mapping}), that probes the inhomogeneity of the
+  B0 field.
+A second theory-based breed of methodologies acquire a map of the \acrlong*{psf}
+  of the \gls*{epi} readouts to correct for the artifact \cite{robson_measurement_1997}.
+Another approach referred to as \gls*{reb}, acquires an extra \gls*{epi} volume in the
+  orthogonal or reversed phase encoding that can be combined to remove the geometric
+  distortions \cite{cordes_geometric_2000,chiou_simple_2000}.
+The last set of methodologies uses an extra T2-weighted image as anatomical reference
+  to seek for the deformation map through nonlinear registration 
+  \cite{kybic_unwarping_2000,studholme_accurate_2000}.
+These \gls*{t2b} methods usually map the T2w image to the \emph{baseline} volume or
+  \textit{b0} of \gls*{dmri}, as the latter is a T2 weighted image as well.
+More recent works report extensions or combinations of existing techniques 
   \cite{andersson_how_2003,zaitsev_point_2004,%
@@
 
-Even though the aforementioned techniques for distortion correction
-have been studied \cite{zeng_image_2002,wu_comparison_2008},
-the lack of a gold-standard limits benchmarking
-strategies. Recently, Irfanoglu et al.
-\cite{irfanoglu_effects_2012} raised the question
-of distortion-derived impact on tractography.
-In this work, we propose an evaluation framework
-using a digital phantom designed for connectivity assessment.
-This framework enabled us to compare several correction 
-techniques and characterize their geometrical accuracy 
-and the dropout compensation. Finally, we report their 
-impact on subsequent tractography and resulting connectivity
-matrices.
+Even though the aforementioned techniques for distortion correction have been studied
+  \cite{zeng_image_2002,wu_comparison_2008}, the lack of a gold-standard limits benchmarking
+  strategies.
+Recently, Irfanoglu et al. \cite{irfanoglu_effects_2012} raised the question of distortion-derived
+  impact on tractography.
+In this work, we propose an evaluation framework using a digital phantom designed for connectivity
+  assessment.
+This framework enabled us to compare several correction techniques and characterize their geometrical
+  accuracy and the dropout compensation.
+Finally, we report their impact on subsequent tractography and connectivity.

### 2013-12-10 Publications/2013-ISBI14/03-methods.tex
Final proposal and response letter draft

@@
-\section{METHODS}
-
-\subsection{Digital Phantom}
-Based on the fiber geometries of the digital phantom 
-created for the \emph{\Gls*{hardi} 
-Reconstruction Challenge} held in ISBI 2013 
-(San Francisco, US), we simulated high resolution 
-(0.5mm isotropic) T1 (TE/TR=10/1500ms) and T2 
-(TE/TR=90/5000ms) images, as well as two \gls*{dmri}
-images (1.0mm isotropic, $b$=1200, 1 B0 image) with
-32 and 64 evenly-distributed directions intended
-for \gls*{dti} and \gls*{hardi} reconstructions, respectively.
-Diffusion is modeled by a restricted and a hindered
-compartment, similar to \cite{assaf_composite_2005}.
-The phantom includes \gls*{wm} fiber bundles,
-\gls*{gm} and \gls*{csf}. Physical properties
-(T1/T2 times in sec) used in simulation are
-($0.832\pm0.010$ / $79\times10^{-3}\pm0.6\times10^{-3}$)
-for \gls*{wm},
-($1.331\pm0.013$ / $110\times10^{-3}\pm2.0\times10^{-3}$) for
-\gls*{gm} and ($3.5\pm0.1$ / $0.25\pm0.01$) for \gls*{csf}.
+\section{METHODS}\label{sec:methods}
+\subsection{Digital Phantom}\label{sec:phantom}
+Based on the fiber geometries of the digital phantom created
+  for the \emph{\acrshort*{hardi} Reconstruction Challenge} held
+  in ISBI 2013 (San Francisco, US), we simulated high resolution
+  (0.5mm isotropic) T1 weighted (TE/TR= 10/1500ms) and T2w
+  (TE/TR= 90/5000ms) images, as well as two \gls*{dmri} images
+  (1.0mm isotropic, $b$=1200, 1 B0 image) with 32 and 64
+  evenly-distributed directions intended for \gls*{dti} and 
+  \gls*{hardi} reconstructions, respectively.
+Diffusion is modeled by a restricted and a hindered compartment,
+  similar to \cite{assaf_composite_2005}.
+The phantom includes \gls*{wm} fiber bundles, \gls*{gm} and \gls*{csf}.
+Physical properties (T1/T2 times in ms) used in simulation are
+  ($832\pm10/79\pm0.6$) for \gls*{wm},
+  ($1331\pm13/110\pm2.0$) for \gls*{gm} and
+  ($3500\pm100$ / $250\pm10$) for \gls*{csf}.
+T1w, T2w and \glspl*{dwi} were added normally distributed noise
+  up to a \gls*{snr} of 30dB.
 
@@
 \label{sec:distortion}
-\Gls*{fmb} methodologies use a map
-of the field in the scanner. More precisely, the
-phase difference between two subsequent samplings
-of the fieldmap. With that information, it is possible
-to compute the theoretical displacement that each
-voxel undergoes, the so-called \gls*{vsm}. The
-most prominent feature of this \gls*{vsm} is that all
-the shifts have the same orientation (parallel to the
-phase-encoding direction of the \gls*{epi}) and their
-magnitude and direction depend on the \gls*{epi} 
-gradient increments (or \emph{blips}), and the actual
-phase difference at the voxel.
+\Gls*{fmb} methodologies use a map of the field in the scanner.
+More precisely, the phase difference between two subsequent samplings
+  of the field map.
+With that information, it is possible to compute the theoretical
+  displacement that each voxel undergoes, the so-called \gls*{vsm}.
+The most prominent feature of this \gls*{vsm} is that all the shifts
+  have the same orientation (parallel to the phase-encoding direction
+  of the \gls*{epi}) and their magnitude and direction depend on the 
+  \gls*{epi} gradient increments (or \emph{blips})
+  and the actual phase difference at the voxel.
 
-In order to create a realistic distortion, we
-generated a synthetic phase-difference map 
-consistent with the phantom, using the 
-tools distributed with the \emph{FSL} package 
-\cite{jenkinson_fsl_2012}  and
-standard parameters ($\Delta$TE=2.46~ms. for the
-field mapping and \emph{effective dwell time} of 
-0.77~ms. for the \gls*{epi}).
-We defined two regions of smooth dephasing and 
-computed the corresponding \gls*{vsm}. Amplitude
-of the dephasing maps can be modulated, enabling
-us to evaluate the impact of the distortion magnitude.
-We generated several \glspl*{vsm} with increasing
-maximum shifts (from 3.80 to 7.60~mm), covering 
-the typical magnitudes of distortion observed 
-in real datasets.
+In order to create a realistic distortion, we generated a synthetic and noise-free
+  phase-difference map consistent with the phantom, using the tools distributed with
+  the \emph{FSL} package %\cite{jenkinson_fsl_2012}
+  (\url{www.fmrib.ox.ac.uk/fsl}) and standard parameters ($\Delta$TE=2.46~ms. for the
+  field mapping and \emph{effective dwell time} of 0.77~ms for the \gls*{dwi}).
+We defined two regions of smooth dephasing and computed the corresponding \gls*{vsm}.
+Amplitude of the dephasing maps can be modulated, enabling us to evaluate the
+  magnitude of distortion.
+We generated several \glspl*{vsm} with increasing maximum shifts (from 3.80 to 7.60~mm),
+  covering the typical magnitudes of distortion observed in real datasets.
 
-From these synthetic \glspl*{vsm}, we generated
-the corresponding distorted \glspl*{dwi}, in two opposed
-phase-gradient encoding directions. The second simulated
-``acquisition'' of the same phantom was necessary 
-for evaluating \gls*{reb} methods.
+From these synthetic \glspl*{vsm}, we generated the corresponding distorted \glspl*{dwi},
+  in two opposed phase-gradient encoding directions.
+The second simulated ``acquisition'' of the same phantom was necessary for evaluating \gls*{reb}
+  methods.
 
-In summary, we generated 
-a full gold-standard containing realistic T1 and T2
-at high resolution, \glspl*{dwi} acquired in two
-different phase encoding directions, and a ground-truth
-\gls*{dwi} data, which is not available for real datasets
+In summary, we generated a full gold-standard containing realistic T1w and T2w
+  at high resolution, \glspl*{dwi} acquired in two different phase encoding directions,
+  and a ground-truth \gls*{dwi} data, which is not available for real datasets
   (\autoref{fig:phantom}).
+%The dataset is completed with two more images, the noise-free field map used for
+%  generating the \gls*{vsm} and a field map with normally distributed noise for
+%  a \gls*{snr} of 20dB.
 
@@
    \caption{Ground-truth digital phantom.
-   A) T2-weighted image; B) undistorted \textit{b0} volume;
+   A) T2w; B) undistorted \textit{b0} volume;
    C, D) distorted \textit{b0} volumes with opposed phase 
@@
 \label{sec:correction}
-Three out of four methods presented in \autoref{sec:intro}
-were tested on the evaluation framework.
-
-\Gls*{fmb} correction
-is essentially the same as the one used for generating the 
-distortion, on the inverse direction. Normally distributed 
-noise was added to the fieldmap (for a signal-to-noise 
-ratio of 20dB) before correction.
-
-For \Gls*{reb} correction, we made use of the implementation found 
-in \emph{FSL} (\texttt{topup})
-that demands for the \textit{b0} of the reverse-encoding simulation.
-In this second case, the \gls*{vsm} is inferred from the differences
-between the two corresponding \textit{b0}.
-
-Finally, for evaluating \Gls*{t2b},
-we fine-tuned \emph{ANTs} (\url{http://stnava.github.io/ANTs}),
-in order to register \textit{b0} to T2. To this end, we used a
-multi-resolution scheme with 3 levels of subsampling and smoothing,
-mutual information metric, and symmetric diffeomorphic transform 
-(\texttt{SyN}). Several configurations of kernel widths for the 
-regularization smoothers were tested, and finally selected 
-0.5/1.0 voxels (gradient/deformation fields, respectively) 
-for its best result. Additionally, undistorted images are 
-corrected for dropout using the determinant
-of the Jacobian of the deformation field.
+Three out of four methods presented in \autoref{sec:intro} were tested on the
+  evaluation framework.
+\Gls*{fmb} correction is the same as the one used for generating the distortion,
+  working on the opposite direction.
+Normally distributed noise was added to the synthetic field map (\gls*{snr}=20dB)
+  before correction, for a more realistic evaluation.
+\Gls*{reb} correction is implemented in \emph{FSL} (\texttt{topup}), and it demands for the \textit{b0}
+  of the reverse-encoding simulation.
+In this second case, the \gls*{vsm} is inferred from the differences between the two corresponding 
+  \textit{b0}.
+Finally, we included \Gls*{t2b} methods with \emph{ANTs} (\url{stnava.github.io/ANTs}),
+  fine-tuned for \textit{b0}-to-T2w registration.
+To this end, we used a multi-resolution scheme with 3 levels of subsampling and smoothing,
+  mutual information metric, and the symmetric diffeomorphic transform model.
+Several configurations of kernel widths for the regularization smoothers were tested, and finally
+  selected 0.5/1.0 voxels (gradient/deformation fields, respectively) for its best result.
+Additionally, undistorted images are corrected for dropout using the determinant of the Jacobian 
+  of the deformation field.
 
@@
 \label{sec:evaluation}
-The original phantom, one distorted version, and
-the corrected instances are then connected to a
+The original phantom, one distorted version, and the corrected instances are then connected to a
   \gls*{dwi} reconstruction and tractography pipeline.
-Additionally, the original tissue probability maps
-are also distorted and corrected to provide tractography
-with the required \gls*{wm} masks.
+Additionally, the original tissue probability maps are also distorted and corrected to provide
+  tractography with the required \gls*{wm} masks.
 These maps are also used in a final assessment module.
 
-The framework supports two different methods for
-\gls*{dwi} reconstruction and deterministic tractography.
-Firstly, \emph{Diffusion Toolkit} \cite{wang_diffusion_2007}
-for \gls*{dti}, is configured with 10 random
-seeds per voxel by default. Secondly, \emph{MRTrix}
-\cite{tournier_mrtrix:_2012} for \gls*{hardi}, with
-default parameters set to use constrained spherical
-deconvolution, maximum harmonic order of 6, and 150000
-desired tracks.
-For both options, the seeding regions can be set to use
-either the distorted-corrected \gls*{wm} mask, or the 
-regions used to generate the ground-truth. This second
-seeding strategy mimics the usual procedure on real 
-data, where regions are typically mapped from the
-anatomically correct T1.
+The framework supports two different methods for \gls*{dwi} reconstruction and deterministic
+  tractography:
+1) \emph{Diffusion Toolkit} (\url{trackvis.org/dtk}) %\cite{wang_diffusion_2007}
+  for \gls*{dti}, is configured with 10 random seeds per voxel by default; and
+2) \emph{MRTrix} %\cite{tournier_mrtrix:_2012}
+  (\url{www.brain.org.au/software/mrtrix}) for \gls*{hardi}, with default parameters
+  set to use constrained spherical deconvolution, maximum harmonic order of 6, and
+  150000 desired tracks.
+For both options, the seeding regions can be set to use either the distorted-corrected 
+  \gls*{wm} mask, or the regions used to generate the ground-truth.
+This second seeding strategy mimics the usual procedure on real data,
+  where regions are typically mapped from the anatomically correct T1w.
 
-The evaluation framework is completed by automated 
-assessment modules. We evaluated three characteristics
-of the correction methods. 
-Firstly, we assessed the geometrical correctness
-reporting overlap indices of three tissue
-probability maps (namely \gls*{csf}, \gls*{wm},
-and \gls*{gm}), weighting the average by tissue
-volumes.
-Secondly, to evaluate the quality of the actual 
-signal dropout correction, we studied the 
-similarity volume by volume computing the $\ell_1$-norm
-correlation index. We report this score
-on the \textit{b0} and the average of the remaining \gls*{dwi}
-volumes. Thirdly, we studied the impact on the
-connectivity matrices reporting the number of
-false positives (nonexistent connections in the
-gold-standard) and false negatives (or missed connections).
+The evaluation framework is completed by automated assessment modules.
+We evaluated three characteristics of the correction methods.
+Firstly, we assessed the geometrical correctness reporting overlap indices of three tissue
+  probability maps (namely \gls*{csf}, \gls*{wm}, and \gls*{gm}), weighting the average by
+  tissue volumes.
+Secondly, to evaluate the quality of the actual signal dropout correction, we studied the 
+  similarity volume by volume computing the $\ell_1$-norm correlation index.
+We report this score on the \textit{b0} and the average of the remaining \gls*{dwi} volumes.
+Thirdly, we studied the impact on the connectivity matrices reporting the number of 
+  false positives (nonexistent connections in the gold-standard) and false negatives 
+  (or missed connections).

### 2013-12-10 Publications/2013-ISBI14/04-results.tex
Final proposal and response letter draft

@@
 \label{sec:results}
-
 All the modules described in the previous section were
-conveniently integrated in workflows using \emph{nipype}
-\cite{gorgolewski_nipype:_2011}. The choice of this tool
-grants the reproducibility of the experiments,
+  conveniently integrated in workflows using \emph{nipype} % \cite{gorgolewski_nipype:_2011}
+  (\url{nipy.sourceforge.net/nipype}).
+The choice of this tool grants the reproducibility of the experiments,
   and the evaluation workflow is publicly released.
-The results of this study are summarized
-in \autoref{fig:results} and \autoref{table:results}.
-The remaining of this section provides extended descriptions
-and interpretation of these findings.
-
-\begin{figure*}[tpb]
+The results of this study are summarized in \autoref{fig:results} and
+  \autoref{table:results}.
+The remaining of this section provides extended descriptions and interpretation of these
+  findings.
+\begin{figure*}[tp]
    \centering
-   \includegraphics[width=0.8\linewidth]{Fig02-Results}
+   \includegraphics[width=0.95\linewidth]{Fig02-Results}
    \caption{Visual assessment of the correction methods.}
@@
 
-\subsection{Geometrical correctness and signal recovery}
-
-A summary of quantitative results, computed for the \gls*{dti} 
-dataset with maximum shift of 3.80~mm,
+\subsection{Geometrical correctness and signal recovery}\label{sec:results_geometry}
+A summary of quantitative results, computed for the \gls*{dti} dataset with max. shift 3.80~mm,
   is provided in \autoref{table:results}.
-The best scores were obtained with the \gls*{reb} 
-method, followed by \gls*{fmb}. 
-The clear difference of accuracy between \gls*{reb} and \gls*{fmb}
-with respect to \gls*{t2b} infers the latter may not be
-an appropriate method for susceptibility correction. 
-The described performance was constant for \gls*{fmb} 
-and \gls*{reb} along the different magnitudes of 
-distortion evaluated (maximum shifts of 3.80-7.60~mm).
-
+The best scores were obtained with the \gls*{reb} method, followed by \gls*{fmb}.
+The clear difference of accuracy between \gls*{reb} and \gls*{fmb} with respect to \gls*{t2b}
+  infers the latter may not be an appropriate method for susceptibility correction.
+The described performance was constant for \gls*{fmb} and \gls*{reb} along the different magnitudes 
+  of distortion evaluated (max. shift 3.80-7.60~mm).
 \input{041-results-table}
-
 \input{042-results-tractography}
-
-A second workflow investigated the similarity of the recovered
-signal with respect to the original (undistorted).
-In this second study, \gls*{reb} performed
-better than the other two methods, as reported
-in \autoref{table:results}. \gls*{reb} scored a $91.0\%$
-similarity index for the \textit{b0} volume and an average $97.65\%$
-for the remaining \glspl*{dwi}. Again, the second
-qualified was \gls*{fmb}, which achieved very close results
-for the \glspl*{dwi} ($96.26\%$) but not as good for the \textit{b0}
-volume. Visual inspection of the recovered data confirms the 
-presented quantitative results (\autoref{fig:results}, first row).
+A second workflow investigated the similarity of the recovered signal with respect to the original
+  (undistorted).
+In this second study, \gls*{reb} performed better than the other two methods, as reported in
+  \autoref{table:results}.
+\Gls*{reb} scored a 91.0\% similarity index for the \textit{b0} volume and an average 97.65\%
+  for the remaining \glspl*{dwi}.
+Again, the second qualified was \gls*{fmb}, which achieved very close results for the \glspl*{dwi}
+  (96.26\%) but not as good for the \textit{b0} volume.
+Visual inspection of the recovered data confirms the presented quantitative results 
+  (\autoref{fig:results}).
 
 \subsection{Impact on tractography and connectivity}
+Connectivity matrices derived from the \gls*{dti} dataset were completely hindered by the high complexity
+  of the fibers contained in the phantom.
+Tractography discovered successfully only 4 out of 27 existing connections from the original (undistorted) data,
+  with more than 45 false connections.
+With the \gls*{hardi} dataset we found 23/27 connections, but we still observed 40 false positives.
+Therefore, the experiments using the \gls*{dti} phantom were discarded.
+The results presented in this subsection only refer to the \gls*{hardi} dataset.
 
-Connectivity matrices derived from the \gls*{dti} dataset
-were completely hindered by the high complexity of the 
-fibers contained
-in the phantom. Tractography discovered successfully only 4 out 
-of 27 existing connections from the original (undistorted) data,
-with more than 45 false connections.
-With the \gls*{hardi} dataset we found 23/27 connections,
-but we still observed 40 false positives. Therefore, the 
-experiments using the \gls*{dti} phantom were discarded. The
-results presented in this subsection only refer to the 
-\gls*{hardi} dataset.
-
-Using the strongest distortion (maximum shift of 7.60~mm),
-the number of detected connections
-remained the same (23/27), slightly increasing the number of
-false positives to 42. The immediate conclusion is that, with highly
-complex anatomies (crossing, fanning, etc.) and limited number of
-ground-truth connections, connectivity matrices are more sensitive 
+Using the strongest distortion (max. shift 7.60~mm), the number of detected connections
+  remained the same (23/27), slightly increasing the number of false positives to 42.
+The immediate conclusion is that, with highly complex anatomies (crossing, fanning, etc.)
+  and limited number of ground-truth connections, connectivity matrices are more sensitive
   to reconstruction and tractography than to distortions.
 
-For the sake of completeness, \autoref{table:results-tractography}
-reports the characteristic features of the tractography results
-and connectivity matrices, for the fieldmap that caused a maximum
-shift of 3.80~mm. A very similar outcome was obtained for
-7.60~mm. These results, along with visual inspection
-(\autoref{fig:results}, second and third rows), might point
-to \gls*{fmb} as the best correction method.
+For the sake of completeness, \autoref{table:results-tractography} reports the characteristic
+  features of the tractography results and connectivity matrices, for the fieldmap that caused a max.
+  shift of 3.80~mm.
+A very similar outcome was obtained for 7.60~mm.
+These results, along with visual inspection (\autoref{fig:results}), might point to \gls*{fmb} as
+  the best correction method.
 
-\subsection{Discussion}
+\subsection{Discussion}\label{sec:discussion}
+Even though all the surveyed methods produced visually sound results, our study suggested that
+  \gls*{reb} is the most accurate method in terms of geometrical agreement and signal dropout recovery.
+The \gls*{t2b} method did not achieve the necessary high-standards to recommend its use.
+Nonetheless, we understand that specific methods with anisotropic regularization that completely
+  restrict deformations to the phase-encoding direction would perform significantly better than
+  the standard method presented.
+Geometrical correctness of \gls*{dwi} data is fundamental in connectivity analysis to spatially
+  locate the regions which will define the nodes of the final connectivity matrix.
 
-Even though all the surveyed methods produced visually 
-sound results, our study suggested that \gls*{reb} is the 
-most accurate method in terms of geometrical agreement and
-signal dropout recovery. 
-The \gls*{t2b} method did not achieve the necessary high-standards
-to recommend its use. Nonetheless, we understand that specific
-methods with anisotropic regularization that completely
-restrict deformations to the phase-encoding direction would perform
-significantly better than the standard method presented.
-Geometrical correctness of \gls*{dwi} data is fundamental 
-in connectivity analysis to spatially locate the
-regions which will define the nodes of the final connectivity
-matrix.
+Regarding tractography, this study revealed that signal reconstruction and tractography algorithms
+  masked the impact of susceptibility distortion on the final connectivity matrices.
+Due to this effect, experiments performed on the \gls*{dti} phantom were discarded for comparison.
+With the \gls*{hardi} dataset, the extracted connectivity matrix slightly changed with distortion.
+Quantitative differences reported in \autoref{table:results-tractography} could be more related to
+  the smoothing derived from interpolation implemented by each method.
+Visual results might suggest that \gls*{fmb} achieved better results.
 
-Regarding tractography, this study revealed that signal
-reconstruction and tractography algorithms masked the 
-impact of susceptibility distortion on the final connectivity
-matrices. Due to this effect, experiments performed on 
-the \gls*{dti} phantom were discarded for comparison.
-With the \gls*{hardi} dataset, the extracted 
-connectivity matrix slightly
-changed with distortion. Quantitative differences reported in
-\autoref{table:results-tractography} could be more related
-to the smoothing derived from interpolation implemented by
-each method. Visual results might suggest that \gls*{fmb}
-achieved better results.
+Although we found connectivity matrices rather invariant with respect to distortion, they are suspected
+  to be significantly impacted by the susceptibility artifact on real data \cite{irfanoglu_effects_2012}.
+This hypothesis points to the need of more appropriate phantoms with denser connectivity matrices.
+State of the art phantoms for tractography usually present a discrete set of simulated fiber bundles that
+  translate in very sparse fiber-endings regions and connectivity matrix.
+In the real case, the surface limiting the tracks is densely covered by the regions mapped from the anatomical
+  dataset, what leads to larger sensitivity with respect to deformations.
 
-Although we found connectivity matrices rather invariant with
-respect to distortion, they are suspected to be significantly
-impacted by the susceptibility artifact on real data 
-\cite{irfanoglu_effects_2012}.
-This hypothesis points to the need of more appropriate 
-phantoms with denser connectivity matrices.
-State of the art phantoms for tractography usually present a
-discrete set of simulated fiber bundles that translate in 
-very sparse fiber-endings regions and connectivity matrix.
-In the real case, the surface limiting the tracks is densely 
-covered by the regions mapped from the anatomical 
-dataset, what leads to larger sensitivity with respect to
-deformations.
+A possible limitation of this work is that \gls*{fmb} is used for both synthesis and correction of the
+  distortion.
+However, practical reasons (i.e. noise, signal dropout) impede perfect geometrical correction with \gls*{fmb}.
+Moreover, \gls*{reb} performed more accurately than \gls*{fmb}.
 
-A possible limitation of this work is that \gls*{fmb} is
-used for both synthesis and correction of the distortion.
-However, practical reasons (i.e. noise, signal dropout)
-impede perfect geometrical correction with \gls*{fmb}.
-Moreover, \gls*{reb} performed more accurately 
-than \gls*{fmb}.
-
-Future extensions of this work will include a 
-refined digital phantom. Additional lines for this
-work will evaluate different realizations of the 
-synthetic phase-difference map, to better characterize
-the phenomenon. Also, the framework can be enhanced 
-for studying the impact of other artifacts as 
-subject's motion or eddy currents-derived distortions.
+Future extensions of this work will include a refined digital phantom.
+Additional lines for this work will evaluate different realizations of the synthetic phase-difference map,
+  to better characterize the phenomenon.
+Also, the framework can be enhanced for studying the impact of other artifacts as subject's motion or eddy
+  currents-derived distortions.

### 2013-12-10 Code/Modules/Functional/include/FunctionalBase.h
Implementing statistical conformation of gradients

@@ -202,6 +202,35 @@ public:
 	typedef typename std::vector< PointType >                PointsVector;
 	typedef typename std::vector< PointsVector >             PointsList;
 
+	struct GradientSample {
+		PointValueType grad;
+		size_t id;
+
+		GradientSample() {}
+		GradientSample( PointValueType g, size_t i ): grad(g), id(i) {}
+		GradientSample( const GradientSample &s ): grad(s.grad), id(s.id) {}
+
+		GradientSample operator+(const GradientSample& g) const {
+			return GradientSample( grad+g.grad, id );
+		}
+
+		GradientSample operator<(const GradientSample& g) const {
+			return grad < g.grad;
+		}
+
+		GradientSample operator>(const GradientSample& g) const {
+			return grad > g.grad;
+		}
+	};
+
+	struct by_grad {
+		bool operator()( GradientSample const &a, GradientSample const &b ) {
+			return a.grad < b.grad;
+		}
+	};
+	typedef typename std::vector< GradientSample >           SampleType;
+
+
 	void CopyInformation( const FieldType* field);
 
 	MeasureType GetValue();

### 2013-12-11 Publications/2013-ISBI14/02-introduction.tex
Fixed some typos. Fixed vera font

@@
 
-Among all the difficulties that such a complex workflow raises, % \cite{jones_twenty-five_2010},
+Among all the difficulties that such a complex workflow raises \cite{jones_twenty-five_2010},
   we will address here the susceptibility-derived artifacts,

### 2013-12-11 Publications/2013-TMI/03-introduction.tex
Fixed some references
Added Alessandro's paper on TMI about ISBI13

@@
 \gls{dwi} data to derive the local axonal structure at each imaging voxel 
-\citep{basser_microstructural_2011} and estimate a whole-brain mapping of fiber 
+\citep{basser_microstructural_1996} and estimate a whole-brain mapping of fiber 
 tracts \citep{hagmann_mr_2012,johansen-berg_using_2009} represented by 
@@
 available algorithms. The outcome is a voxel-wise model of diffusion shape 
-(in \citep{daducci_isbi2013_2013} \todo{FIX ref} there is available a comprehensive review of 
+(\citep{daducci_quantitative_2013} contains a comprehensive review of 
 state-of-art methodologies). On top of the
@@
 scanner. Based on this \emph{fieldmap} and the theory underlying the 
-distortion, \citep{jezzard_characterization_2005} proposed a correction 
+distortion, \citep{jezzard_correction_1995} proposed a correction 
 methodology. A number of forked and improved methodologies have been 
@@
 generic name of ``fieldmap correction'' techniques \cite{hsu_correction_2009,
-reber_correction_2005}.
+reber_correction_1998}.
 

### 2014-04-11 Scripts/pyacwereg/workflows/evaluation.py
Modified evaluation pipelines
This change should allow to close #144. I was passing wrong inputs.

@@ -6,7 +6,7 @@
 # @Author: Oscar Esteban - code@oscaresteban.es
 # @Date:   2014-03-12 16:59:14
 # @Last Modified by:   oesteban
-# @Last Modified time: 2014-04-10 15:35:08
+# @Last Modified time: 2014-04-11 13:44:30
 
 import os
 import os.path as op
@@ -172,9 +172,9 @@ def bspline( name='BSplineEvaluation', methods=None, results=None ):
              ( inputnode, evwfs[i], [ ('subject_id', 'infonode.subject_id') ] )
             ,( prep,     reg, [('outputnode.out_surfs','inputnode.in_surf'),
                                ('outputnode.out_smri_brain', 'inputnode.in_orig' ),
-                               ('outputnode.out_tpms', 'inputnode.in_tpms'),
                                ('outputnode.out_mask', 'inputnode.in_mask' ) ])
-            ,( dist,     reg, [('outputnode.out_file', 'inputnode.in_dist' )])
+            ,( dist,     reg, [('outputnode.out_file', 'inputnode.in_dist'),
+                               ('outputnode.out_tpms', 'inputnode.in_tpms') ])
             ,( prep,evwfs[i], [('outputnode.out_smri_brain', 'refnode.in_imag'),
                                ('outputnode.out_tpms',       'refnode.in_tpms'),
                                ('outputnode.out_surfs',      'refnode.in_surf'),

### 2014-04-11 Publications/2014-TMI/02-abstract.tex
New paper branch

In whole-brain connectivity analysis of diffusion MRI (dMRI) data,
an accurate delineation of the white-matter and grey-matter 
interfaces is required. While high-standard segmentation is readily 
available for structural MRI, dMRI typically present significant cerebrospinal
fluid contamination effects, mainly derived from its typical low resolution,
and severe geometrical distortions. We propose a segmentation-registration 
variational framework that exploits the detailed anatomy extracted from 
structural MRI as shape-prior. We use an ``active contours without edges''-like 
model to search for a deformation field that optimally maps the shape prior 
on multivariate dMRI-derived data in diffusion space, registering structural
and diffusion coordinate spaces and implicitly segmenting dMRI data. The approach
proven the intrinsic coupling of segmentation and distortion correction and
we evaluate its results on a digital simulated phantom and real datasets. 
Therefore, precise and consistent cortical parcellation on dMRI is 
straightforward by projection from T1 space, avoiding additional registration,
segmentation and/or surface matching steps.

### 2014-04-11 Scripts/pyacwereg/workflows/evaluation.py
Updated evaluation

@@ -6,7 +6,7 @@
 # @Author: Oscar Esteban - code@oscaresteban.es
 # @Date:   2014-03-12 16:59:14
 # @Last Modified by:   oesteban
-# @Last Modified time: 2014-04-11 13:44:30
+# @Last Modified time: 2014-04-11 13:50:21
 
 import os
 import os.path as op
@@ -162,22 +162,22 @@ def bspline( name='BSplineEvaluation', methods=None, results=None ):
         norm_tpms.append( pe.Node( niu.Function( input_names=['in_files','in_mask'], output_names=['out_files'], function=normalize ), name='Normalize%02d' % i ) )
 
         wf.connect( [
-             ( inputnode, evwfs[i], [ ('subject_id', 'infonode.subject_id') ] )
-            ,( prep,     reg, [('outputnode.out_surfs','inputnode.in_surf'),
-                               ('outputnode.out_smri_brain', 'inputnode.in_orig' ),
-                               ('outputnode.out_mask', 'inputnode.in_mask' ) ])
-            ,( dist,     reg, [('outputnode.out_file', 'inputnode.in_dist'),
-                               ('outputnode.out_tpms', 'inputnode.in_tpms') ])
-            ,( prep,evwfs[i], [('outputnode.out_smri_brain', 'refnode.in_imag'),
-                               ('outputnode.out_tpms',       'refnode.in_tpms'),
-                               ('outputnode.out_surfs',      'refnode.in_surf'),
-                               ('outputnode.out_mask',       'refnode.in_mask'), ])
-            ,( dist,evwfs[i], [('outputnode.out_field',      'refnode.in_field' ) ])
-            ,( reg, norm_tpms[i], [ ('outputnode.out_tpms', 'in_files') ] )
-            ,( reg, evwfs[i], [('outputnode.out_corr', 'tstnode.in_imag'),
-                               ('outputnode.out_surf', 'tstnode.in_surf'),
-                               ('outputnode.out_field','tstnode.in_field' ) ])
-            ,( norm_tpms[i],evwfs[i], [('out_files', 'tstnode.in_tpms')])
+             (inputnode,    evwfs[i], [ ('subject_id', 'infonode.subject_id') ] )
+            ,(prep,              reg, [('outputnode.out_surfs','inputnode.in_surf'),
+                                       ('outputnode.out_smri_brain', 'inputnode.in_orig' ),
+                                       ('outputnode.out_mask', 'inputnode.in_mask' ) ])
[…]

### 2014-04-11 Scripts/pyacwereg/utils/run_evaluations.py
Several fixes
* Removed mesh interface for debugging
* Fixed bug in _stats function of registration_ev causing the
  workflow to crash. Close #145

@@ -6,7 +6,7 @@
 # @Author: oesteban - code@oscaresteban.es
 # @Date:   2014-04-04 19:39:38
 # @Last Modified by:   oesteban
-# @Last Modified time: 2014-04-04 20:25:28
+# @Last Modified time: 2014-04-11 13:52:38
 
 from argparse import ArgumentParser
 from argparse import RawTextHelpFormatter
@@ -76,7 +76,7 @@ if __name__== '__main__':
     ])
 
     if options.out_csv is None:
-        mm.inputs.inputnode.out_csv = op.join( options.work_dir, mm.name, 'results.csv' )
+        mm.inputs.inputnode.out_csv = op.join( options.work_dir, wf.name, 'results.csv' )
     else:
         mm.inputs.inputnode.out_csv = options.out_csv
 

### 2014-04-11 Scripts/pyacwereg/workflows/evaluation.py

@@ -6,7 +6,7 @@
 # @Author: Oscar Esteban - code@oscaresteban.es
 # @Date:   2014-03-12 16:59:14
 # @Last Modified by:   oesteban
-# @Last Modified time: 2014-04-11 13:50:21
+# @Last Modified time: 2014-04-11 13:57:10
 
 import os
 import os.path as op
@@ -33,13 +33,14 @@ def registration_ev( name='EvaluateMapping' ):
     similarity, displacement fields difference, mesh distances, and overlap indices.
     """
 
-    def _average( in_file ):
+    def _stats( in_file ):
         import numpy as np
         import nibabel as nb
 
         data = nb.load( in_file ).get_data()
         data = np.ma.masked_equal( data, 0 )
-        return [ data.mean(), data.std(), data.max(), data.min(), np.ma.extras.median(data) ]
+        result = np.array([ data.mean(), data.std(), data.max(), data.min(), np.ma.extras.median(data)])
+        return result.tolist()
 
 
     wf = pe.Workflow( name=name )
@@ -62,9 +63,9 @@ def registration_ev( name='EvaluateMapping' ):
     diff_im = pe.Node( Similarity(metric='cc'), name='ContrastDiff')
     inv_fld = pe.Node( iface.InverseField(), name='InvertField' )
     diff_fld = pe.Node( namev.ErrorMap(), name='FieldDiff')
-    mesh = pe.MapNode( namesh.P2PDistance(weighting='surface'),
-                      iterfield=[ 'surface1','surface2' ],
-                      name='SurfDistance')
+#    mesh = pe.MapNode( namesh.P2PDistance(weighting='surface'),
+#                      iterfield=[ 'surface1','surface2' ],
+#                      name='SurfDistance')
 
     csv = pe.Node( namisc.AddCSVRow(), name="AddRow" )
 
@@ -83,13 +84,13 @@ def registration_ev( name='EvaluateMapping' ):
                ,( input_ref,    diff_fld, [( 'in_mask','mask')])
                ,( inv_fld,      diff_fld, [( 'out_field', 'in_ref') ])
                ,( input_tst,    diff_fld, [( 'in_field', 'in_tst')])
-               ,( input_ref,        mesh, [( 'in_surf', 'surface1')])
-               ,( input_tst,        mesh, [( 'in_surf', 'surface2')])
+#               ,( input_ref,        mesh, [( 'in_surf', 'surface1')])
+#               ,( input_tst,        mesh, [( 'in_surf', 'surface2')])
[…]

### 2014-04-11 Scripts/pyacwereg/workflows/evaluation.py
Added case when ErrorMap is 0.0

@@ -6,7 +6,7 @@
 # @Author: Oscar Esteban - code@oscaresteban.es
 # @Date:   2014-03-12 16:59:14
 # @Last Modified by:   oesteban
-# @Last Modified time: 2014-04-11 13:57:10
+# @Last Modified time: 2014-04-11 15:19:09
 
 import os
 import os.path as op
@@ -38,6 +38,9 @@ def registration_ev( name='EvaluateMapping' ):
         import nibabel as nb
 
         data = nb.load( in_file ).get_data()
+
+        if ( np.all( data<1.0e-5 ) ):
+          return [ 0.0 ]*5
         data = np.ma.masked_equal( data, 0 )
         result = np.array([ data.mean(), data.std(), data.max(), data.min(), np.ma.extras.median(data)])
         return result.tolist()

### 2014-04-11 Code/Applications/bspline_field.h
Prune negative value
Close #143

@@ -57,6 +57,7 @@
 #include <itkImageFileReader.h>
 #include <itkImageFileWriter.h>
 #include <itkBinaryThresholdImageFilter.h>
+#include <itkThresholdImageFilter.h>
 #include <itkRandomImageSource.h>
 #include <itkGaussianImageSource.h>
 #include <itkImageRandomNonRepeatingIteratorWithIndex.h>
@@ -126,6 +127,9 @@ typedef typename MultiplyFilter::Pointer                       MultiplyPointer;
 
 typedef itk::SubtractImageFilter< CoefficientsType >           SubtractFilter;
 typedef typename SubtractFilter::Pointer                       SubtractPointer;
+
+typedef itk::ThresholdImageFilter< ChannelType >               ThresholdFilter;
+typedef typename ThresholdFilter::Pointer                      ThresholdPointer;
 
 typedef itk::ResampleImageFilter< ChannelType, ChannelType, ScalarType >    ResampleFilter;
 typedef typename ResampleFilter::Pointer                       ResamplePointer;

### 2014-04-11 Code/Applications/warp_image.h

@@ -28,6 +28,7 @@
 #include <itkMaskImageFilter.h>
 #include <itkNearestNeighborInterpolateImageFunction.h>
 #include <itkBinaryThresholdImageFilter.h>
+#include <itkThresholdImageFilter.h>
 
 #include <itkDisplacementFieldTransform.h>
 #include <itkMesh.h>
@@ -66,9 +67,13 @@ typedef itk::BSplineInterpolateImageFunction< ChannelType, ScalarType >    BSpli
 typedef itk::NearestNeighborInterpolateImageFunction< ChannelType, ScalarType >    NearestNeighborInterpolateImageFunction;
 typedef itk::MaskImageFilter< ChannelType, MaskType, ChannelType > MaskFilter;
 
+
+typedef itk::ThresholdImageFilter< ChannelType >               ThresholdFilter;
+typedef typename ThresholdFilter::Pointer                      ThresholdPointer;
+
 typedef itk::WarpImageFilter
-		         < ChannelType, ChannelType, DisplacementFieldType >     WarpFilter;
-typedef typename WarpFilter::Pointer                         WarpFilterPointer;
+	    < ChannelType, ChannelType, DisplacementFieldType >    WarpFilter;
+typedef typename WarpFilter::Pointer                           WarpFilterPointer;
 
 typedef typename itk::DisplacementFieldTransform
 		                           < ScalarType, DIMENSION>    TransformType;

### 2014-04-11 Code/Applications/regseg.h
Prune negative value
Close #143

@@ -52,6 +52,7 @@
 #include <itkVector.h>
 #include <itkVectorImage.h>
 #include <itkImage.h>
+#include <itkThresholdImageFilter.h>
 #include <itkImageFileReader.h>
 #include <itkImageFileWriter.h>
 #include <itkQuadEdgeMesh.h>
@@ -89,6 +90,10 @@ typedef BSplineSparseMatrixTransform<ScalarType,
 		                                     DIMENSION, 3u>  TransformType;
 typedef TransformType::Pointer                               TransformPointer;
 typedef typename TransformType::CoefficientsImageType        CoefficientsType;
+
+
+typedef itk::ThresholdImageFilter< ChannelType >             ThresholdFilter;
+typedef typename ThresholdFilter::Pointer                    ThresholdPointer;
 
 typedef itk::ResampleImageFilter
 		         < ChannelType, ChannelType >                ResampleFilter;

### 2014-04-11 Code/Applications/bspline_field.h
Fixed bug on bspline_field with mask

@@ -134,8 +134,12 @@ typedef typename ThresholdFilter::Pointer                      ThresholdPointer;
 typedef itk::ResampleImageFilter< ChannelType, ChannelType, ScalarType >    ResampleFilter;
 typedef typename ResampleFilter::Pointer                       ResamplePointer;
 
+typedef itk::ResampleImageFilter
+		                 < MaskType, MaskType, ScalarType >    MaskResampleFilter;
+typedef typename MaskResampleFilter::Pointer                   MaskResamplePointer;
+
 typedef itk::BSplineInterpolateImageFunction< ChannelType, ScalarType >    BSplineInterpolateImageFunction;
-typedef itk::NearestNeighborInterpolateImageFunction< ChannelType, ScalarType >    NearestNeighborInterpolateImageFunction;
+typedef itk::NearestNeighborInterpolateImageFunction< MaskType, ScalarType >    NearestNeighborInterpolateImageFunction;
 
 typedef itk::MaskImageFilter< ChannelType, MaskType, ChannelType > MaskFilter;
 

### 2014-04-11 Scripts/pyacwereg/workflows/evaluation.py
Fixed mask orientation in warp_image
Additionally fixed error in evaluation.py

@@ -6,7 +6,7 @@
 # @Author: Oscar Esteban - code@oscaresteban.es
 # @Date:   2014-03-12 16:59:14
 # @Last Modified by:   oesteban
-# @Last Modified time: 2014-04-11 15:19:09
+# @Last Modified time: 2014-04-11 16:59:05
 
 import os
 import os.path as op
@@ -170,10 +170,10 @@ def bspline( name='BSplineEvaluation', methods=None, results=None ):
         wf.connect( [
              (inputnode,    evwfs[i], [('subject_id', 'infonode.subject_id') ] )
             ,(prep,              reg, [('outputnode.out_surfs','inputnode.in_surf'),
-                                       ('outputnode.out_smri_brain', 'inputnode.in_orig' ),
-                                       ('outputnode.out_mask', 'inputnode.in_mask' ) ])
+                                       ('outputnode.out_smri_brain', 'inputnode.in_orig' ) ])
             ,(dist,              reg, [('outputnode.out_file', 'inputnode.in_dist'),
-                                       ('outputnode.out_tpms', 'inputnode.in_tpms') ])
+                                       ('outputnode.out_tpms', 'inputnode.in_tpms'),
+                                       ('outputnode.out_mask', 'inputnode.in_mask') ])
             ,(prep,         evwfs[i], [('outputnode.out_smri_brain', 'refnode.in_imag'),
                                        ('outputnode.out_tpms',       'refnode.in_tpms'),
                                        ('outputnode.out_surfs',      'refnode.in_surf'),

### 2014-04-11 Code/Applications/warp_image.h
Fixed mask orientation in warp_image
Additionally fixed error in evaluation.py

@@ -64,9 +64,12 @@ typedef itk::ImageFileReader<DisplacementFieldType>          DisplacementFieldRe
 typedef typename DisplacementFieldReaderType::Pointer        DisplacementFieldReaderPointer;
 
 typedef itk::BSplineInterpolateImageFunction< ChannelType, ScalarType >    BSplineInterpolateImageFunction;
-typedef itk::NearestNeighborInterpolateImageFunction< ChannelType, ScalarType >    NearestNeighborInterpolateImageFunction;
+typedef itk::NearestNeighborInterpolateImageFunction< MaskType, ScalarType >    NearestNeighborInterpolateImageFunction;
 typedef itk::MaskImageFilter< ChannelType, MaskType, ChannelType > MaskFilter;
 
+typedef itk::WarpImageFilter
+	    < MaskType, MaskType, DisplacementFieldType >          WarpMaskFilter;
+typedef typename WarpMaskFilter::Pointer                       WarpMaskFilterPointer;
 
 typedef itk::ThresholdImageFilter< ChannelType >               ThresholdFilter;
 typedef typename ThresholdFilter::Pointer                      ThresholdPointer;

### 2014-04-11 Scripts/pyacwereg/workflows/evaluation.py
Implemented masks

@@ -6,7 +6,7 @@
 # @Author: Oscar Esteban - code@oscaresteban.es
 # @Date:   2014-03-12 16:59:14
 # @Last Modified by:   oesteban
-# @Last Modified time: 2014-04-11 16:59:05
+# @Last Modified time: 2014-04-11 17:11:22
 
 import os
 import os.path as op
@@ -151,7 +151,8 @@ def bspline( name='BSplineEvaluation', methods=None, results=None ):
             ,( inputnode,  dist, [ ('grid_size', 'inputnode.grid_size')])
             ,( prep,       dist, [ ('outputnode.out_smri_brain','inputnode.in_file'),
                                    ('outputnode.out_surfs', 'inputnode.in_surfs'),
-                                   ('outputnode.out_tpms', 'inputnode.in_tpms')])
+                                   ('outputnode.out_tpms', 'inputnode.in_tpms'),
+                                   ('outputnode.out_mask', 'inputnode.in_mask')])
             ,( dist, outputnode, [ ('outputnode.out_file','out_file'),
                                    ('outputnode.out_field','out_field'),
                                    ('outputnode.out_coeff','out_coeff')])

### 2014-04-22 Publications/2013-ISBI14/abstract/responseLetter.tex

\documentclass[letter]{memoir}

\newcommand{\pointRaised}[2]{\medskip \hrule \noindent 
               \textsl{{\fontseries{b}\selectfont #1}: #2}} 
\newcommand{\reply}{\noindent \textbf{Reply}:\ }

\begin{document}

\section*{Response letter}

The authors would like to thank the reviewers for their time and their valuable comments.

\subsection*{Reviewer 1:}
\pointRaised{Genneral comment}%
{This is an excellent paper, addressing an important topic:
  validation and quantification of the correctness of 
  tractography in the presence of susceptibility artifacts.
Study design, execution and document structure are very good to excellent,
  and this is one of the rare attempts to try to model the correctness of
  connectivity analyses.
Very good for presentation at ISBI.}
\reply{No further action or correction is required.
We really appreciate the comments.}

\subsection*{Reviewer 3:}
\pointRaised{On page 2, line 41-43}{Was normally distributed noise added
in all cases or only for the FMB correction method?}
\reply{We agree on the importance this detail, and modified the
manuscript consequently.
All the images included in the phantom dataset have normally distributed
  noise.
As it was not stated on the submitted manuscript, we have added this point
  in the last sentence of sec. 2.1.:
\emph{T1w, T2w and diffusion weighted images (DWIs) were added normally distributed
  noise up to a signal-to-noise ratio (SNR) of 30dB}.
Therefore, all methods were subjected to variability due to noise, the FMB method
  with the noise added to the field map, the REB method with the noise added to
  each DWI, and finally, the T2B method with the noise added to the T2w and the
  \emph{b0} images.
Finally, it is important to mention that the noise-free realization of the field map
  was the one use to generate the deformation map.}

\pointRaised{On page 4, lines 28 + 37}{"with respect to"}
\reply{We corrected the manuscript accordingly in these two appearances
and reviewed comprehensively the paper for other typos.}

\pointRaised{General comment}
{The authors reason that the susceptibility correction method has hardly
  any effect on the obtained tractography but instead depends on the
  reconstruction and the tractography algorithm.
Does this conclusion also hold for the point-spread function based correction
  method, which is not considered in this work?
Might there exist reconstruction and/or (deterministic or probabilistic)
  tractography algorithms that are more sensitive to the susceptibility correction
  method than the reconstruction/tractography algorithm employed in this work?
}
\reply{The reviewer is right about the first point.
The conclusion would also hold for the point-spread function based methods, given
  that these methods perform in a very similar manner to the FMB and FEB
  methods on finding the deformation map.
In this regard, they differ on the input data required to seek for the deformation
  (for FMB a field map, for FEB two opposed gradient \emph{b0} images and for the
  point-spread function based methods a mapping, in this case, of the point-spread function
  itself).
If we assumed that the three methods performed equally well, finding a perfect deformation
  field, then they provide different implementations for interpolation and drop-out correction.
This latter step might be of great impact if the point-spread function method provided
  a correction significantly better than the proposed methods, but of course this is
  independent from the deformation map finding strategy.
Finally, some very recent works are starting to apply informed priors to reconstruction.
This could dramatically help minimize the side effects of EPI distortion, avoiding
  the need of a high-standard correction technique.}

\subsection*{Reviewer 4:}
\pointRaised{General comment}%
{The paper presents a method for evaluating diffusion MRI
  image distortion correction algorithms.
The paper is technically sound and clearly written, however the
  conference template is not followed to format the paper.
}
\reply{We have removed some elements that were overriding the
  latex template distributed for the conference.
  As the text spanned over the 4-pages limit, we had to
  reformulate some paragraphs with minor changes for briefness.
  We also modified Figure 2 to contain a legend, therefore
  shortening the associated caption.}

\pointRaised{1.}{Headings should be "1","2" instead of "I","II" and
  sub-headings should be "1.1", "2.1" instead of "I-A,
  "II-A".}
\reply{Headings are now compliant.}
\pointRaised{2.}{The reference formatting is also incorrect, please use
  the font size and line spacing as shown in the template.}
\reply{We fixed this formatting, and reduced the number of references.
  We removed references of software tools (ANTs, nipype, FSL, Diffusion Toolkit
  and MRTrix), replacing them by URLs to their websites.
  We also removed reference [1]

  D. K. Jones et al., “Twenty-ﬁve pitfalls in the analysis of diffusion MRI data,” NMR in Biomed., 23(7):803–820,
  2010,

  that was cited to illustrate the complexity and drawbacks of pipelines for connectivity analysis.}

\pointRaised{Section II-D, 2nd to last line}{consider replacing
  "inexistent connections" with "nonexistent connections"}
\reply{Now it reads "nonexitent connections"}
\pointRaised{Section III-C, paragragh 3, line 2}{"with respect
  distortion" should be "with respect to distortion".}
\reply{As for those errors pointed out by reviewer \#3, we corrected
this one, and reviewed again the whole paper for grammar and typos.}

\end{document}

### 2014-04-22 Publications/2013-ISBI14/abstract/01-abstract.tex

--- 2013-10-01 Publications/2013-ISBI14/01-abstract.tex
@@
 Connectivity analysis on diffusion MRI data of the whole-brain
-suffers from distortions caused by the standard 
-echo-planar imaging acquisition strategies. %
+  suffers from distortions caused by the standard echo-planar
+  imaging acquisition strategies. %
 %These artifacts are present at interfaces between tissues,
@@
 These images show characteristic geometrical deformations
-and signal destruction that are an important pitfall
+  and signal destruction that are an important drawback
   limiting the success of tractography algorithms.
@@
 Several retrospective correction techniques are readily
-available. In this work, we use a digital phantom designed
-for the evaluation of connectivity pipelines. We subject
-the phantom to a ``theoretically correct'' and plausible
-deformation that resembles the artifact under investigation.
+  available.
+In this work, we use a digital phantom designed for the
+  evaluation of connectivity pipelines.
+We subject the phantom to a ``theoretically correct'' and
+  plausible deformation that resembles the artifact under
+  investigation.
 We correct data back, with three standard methodologies
   (namely fieldmap-based, reversed encoding-based, and
-registration-based). Finally, we rank the methods based
-on their geometrical accuracy, the dropout compensation,
-and their impact on the resulting connectivity matrices.
+  registration-based).
+Finally, we rank the methods based on their geometrical
+  accuracy, the dropout compensation, and their impact on
+  the resulting connectivity matrices.
 \end{abstract}%

### 2014-04-22 Publications/2013-ISBI14/abstract/03-methods.tex

--- 2013-12-10 Publications/2013-ISBI14/03-methods.tex
@@
   (TE/TR= 90/5000ms) images, as well as two \gls*{dmri} images
-  (1.0mm isotropic, $b$=1200, 1 B0 image) with 32 and 64
+  (1.0mm isotropic, $b$=1200, 1 \emph{b0} image) with 32 and 64
   evenly-distributed directions intended for \gls*{dti} and 

### 2014-04-22 Publications/2013-ISBI14/poster/00-main.tex
New poster for ISBI

% @Author: oesteban
% @Date:   2014-04-22 08:55:40
% @Last Modified by:   oesteban
% @Last Modified time: 2014-04-22 09:11:27

\documentclass[final,hyperref={pdfpagelabels=false}]{beamer}
\setbeamertemplate{bibliography item}[text]

\let\Tiny=\tiny
\providecommand\thispdfpagelabel[1]{}
\mode<presentation>{\usetheme{Blue}}
\usepackage[english]{babel}

\usepackage[T1]{fontenc}
\usepackage{textcomp}

\usepackage[natbib=true, style=numeric-comp,backend=biber]{biblatex}

\usepackage{amsmath,amsthm, amssymb, latexsym}
\usepackage{lmodern}
\usefonttheme[onlymath]{serif}
%\boldmath
\usepackage[orientation=portrait,size=a0,scale=1.4]{beamerposter}

\usepackage{graphicx,epstopdf}
\DeclareGraphicsExtensions{.pdf,.eps}
\graphicspath{{../figures/} {logos/}}

\usepackage{array,booktabs,tabularx}
\newcolumntype{Z}{>{\centering\arraybackslash}X} % centered tabularx columns
\newcommand{\sups}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\subs}[1]{\ensuremath{_{\textrm{#1}}}}
\newcommand{\pphantom}{\textcolor{ta2skyblue}} % phantom introduces a vertical space in p formatted table columns??!!
\usepackage[scaled=.7]{beramono}

\bibliography{99-references}
\appto\bibfont{\hskip1cm\tiny}

\DeclareSymbolFont{cmlargesymbols}{OMX}{cmex}{m}{n}
\let\sumop\relax
\DeclareMathSymbol{\sumop}{\mathop}{cmlargesymbols}{"50}

\listfiles

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\huge Simulation-based evaluation of susceptibility distortion
  correction methods in diffusion MRI for connectivity analysis}
\author{Oscar Esteban$^{1,2,*}$, %
Alessandro Daducci$^{2}$, %
Emmanuel Caruyer$^{3}$, %
Kieran O'Brien$^{4}$,\\%
Mar\'ia J. Ledesma-Carbayo$^{1}$, %
Meritxell Bach-Cuadra$^{5,2}$, %
and Andr\'es Santos$^{1}$}%}
\institute[LTS5] {
$^{1}$Biomedical Image Technologies (BIT), ETSI Telecom.,
U. Polit\'ecnica de Madrid and CIBER-BBN, Spain. \\
$^{2}$Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland. \\
$^{3}$Section of Biomedical Image Analysis, Dpt. of Radiology,
University of Pennsylvania, Philadelphia, PA, USA. \\
$^{4}$Centre d'Imagerie Biom\'edicale (CIBM), and Dpt. of Radiology, University of Geneva, Geneva, Switzerland. \\
$^{5}$Radiology Department, CIBM, Lausanne University Hospital Center (CHUV)
and Lausanne University (UNIL), Switzerland.
$^{*}$Corresponding author (email: \texttt{oscar.esteban@upm.es})
}
\date[May 5th, 2012]{May 5th, 2012}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newlength{\columnheight}
\setlength{\columnheight}{105cm}

\begin{document}

\begin{frame}
  \begin{columns}
    % ---------------------------------------------------------%
    % Set up a column 
    \begin{column}{.49\textwidth}
      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
        \begin{minipage}[T]{.95\textwidth}  % tweaks the width, makes a new \textwidth
          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
            % ---------------------------------------------------------%
            % fill each column with content            
            %\input{motivation}
%            \vfill
            %\input{methods}
%            \vfill
            %\input{conclusions}
            %\vfill
          }
        \end{minipage}
      \end{beamercolorbox}
    \end{column}
    % ---------------------------------------------------------%
    % end the column

    % ---------------------------------------------------------%
    % Set up a column 
    \begin{column}{.49\textwidth}
      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
        \begin{minipage}[T]{.95\textwidth} % tweaks the width, makes a new \textwidth
          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
            % ---------------------------------------------------------%
            % fill each column with content
	    %\input{dataset}
        %\input{results}
        %\input{acknowledgements}
          }
          % ---------------------------------------------------------%
          % end the column
        \end{minipage}
      \end{beamercolorbox}
    \end{column}
    % ---------------------------------------------------------%
    % end the column
  \end{columns}
\end{frame}

\end{document}

### 2014-04-22 Code/Modules/Transform/include/SparseMatrixTransform.h
Code for bspline fields inversion (#147)

@@ -176,7 +176,26 @@ public:
     itkGetConstReferenceObjectMacro(KernelFunction, KernelFunctionType);
 
     itkGetConstMacro( NumberOfSamples, size_t );
-    itkSetMacro( NumberOfSamples, size_t );
+
+    void SetNumberOfSamples(size_t n) {
+    	if( this->m_UseImageOutput ) {
+    		itkExceptionMacro(<< "SetNumberOfSamples should not be used with OutputReference");
+    	}
+
+    	size_t oldsize = this->m_OffGridFieldValues[0].size();
+    	if ( oldsize!=0 && oldsize!=n ) {
+    		itkWarningMacro( << "this action will empty off-grid values set so-far" );
+    	}
+
+    	this->m_NumberOfSamples = n;
+
+    	for (size_t i=0; i<Dimension; i++) {
+    		this->m_OffGridFieldValues[i].set_size( n );
+    		this->m_OffGridFieldValues[i].fill( 0.0 );
+    	}
+
+    	this->Modified();
+    }
 
     itkGetConstMacro( NumberOfParameters, size_t );
 

### 2014-04-22 Publications/2013-ISBI14/poster/00-main.tex
Starting with poster...

@@
 % @Last Modified by:   oesteban
-% @Last Modified time: 2014-04-22 09:11:27
+% @Last Modified time: 2014-04-22 12:22:50
 
@@
 
-\title{\huge Simulation-based evaluation of susceptibility distortion
-  correction methods in diffusion MRI for connectivity analysis}
-\author{Oscar Esteban$^{1,2,*}$, %
-Alessandro Daducci$^{2}$, %
-Emmanuel Caruyer$^{3}$, %
-Kieran O'Brien$^{4}$,\\%
-Mar\'ia J. Ledesma-Carbayo$^{1}$, %
-Meritxell Bach-Cuadra$^{5,2}$, %
-and Andr\'es Santos$^{1}$}%}
-\institute[LTS5] {
-$^{1}$Biomedical Image Technologies (BIT), ETSI Telecom.,
-U. Polit\'ecnica de Madrid and CIBER-BBN, Spain. \\
-$^{2}$Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
-F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland. \\
-$^{3}$Section of Biomedical Image Analysis, Dpt. of Radiology,
-University of Pennsylvania, Philadelphia, PA, USA. \\
-$^{4}$Centre d'Imagerie Biom\'edicale (CIBM), and Dpt. of Radiology, University of Geneva, Geneva, Switzerland. \\
-$^{5}$Radiology Department, CIBM, Lausanne University Hospital Center (CHUV)
-and Lausanne University (UNIL), Switzerland.
-$^{*}$Corresponding author (email: \texttt{oscar.esteban@upm.es})
+\title{\huge Simulation-based evaluation of susceptibility distortion correction methods in dMRI}
+\author[Esteban et al.]{O. Esteban$^{1,2}$\and%
+A. Daducci$^{2}$\and%
+E. Caruyer$^{3}$\and%
+K. O'Brien$^{4}$\and%
+MJ. Ledesma-Carbayo$^{1}$\and%
+M. Bach-Cuadra$^{5,2}$\and%
+A. Santos$^{1}$}%}
+\institute{
+$^{1}$~\includegraphics[height=2cm]{logos/BIT.pdf}~\includegraphics[height=2cm]{logos/UPM.pdf}\hfill%
+$^{2}$~\includegraphics[height=2cm]{logos/LTS5.pdf}~\includegraphics[height=2cm]{logos/EPFL.pdf}\hfill%
+$^{3}$~\includegraphics[height=2cm]{logos/SBIA.pdf}~\includegraphics[height=2cm]{logos/Penn.pdf}\hfill%
+$^{4}$~\includegraphics[height=2cm]{logos/CIBM.pdf}~\includegraphics[height=2cm]{logos/UNIGE.pdf}\hfill%
+$^{5}$~\includegraphics[height=2cm]{logos/CHUV.pdf}~\includegraphics[height=2cm]{logos/UNIL.pdf}%
 }

### 2014-04-22 Publications/2013-ISBI14/poster/00-main.tex
Finishing poster

@@
 % @Last Modified by:   oesteban
-% @Last Modified time: 2014-04-22 12:22:50
+% @Last Modified time: 2014-04-22 20:49:46
 
@@
 \setbeamertemplate{bibliography item}[text]
-
 \let\Tiny=\tiny
@@
 \mode<presentation>{\usetheme{Blue}}
+
+\usepackage{glossaries}
 \usepackage[english]{babel}
@@
 \usepackage{array,booktabs,tabularx}
-\newcolumntype{Z}{>{\centering\arraybackslash}X} % centered tabularx columns
-\newcommand{\sups}[1]{\ensuremath{^{\textrm{#1}}}}
-\newcommand{\subs}[1]{\ensuremath{_{\textrm{#1}}}}
-\newcommand{\pphantom}{\textcolor{ta2skyblue}} % phantom introduces a vertical space in p formatted table columns??!!
+\newcolumntype{Z}{>{\center\arraybackslash}X}
+%\newcolumntype{Z}{>{\centering\arraybackslash}X} % centered tabularx columns
+%\newcommand{\sups}[1]{\ensuremath{^{\textrm{#1}}}}
+%\newcommand{\subs}[1]{\ensuremath{_{\textrm{#1}}}}
+%\newcommand{\pphantom}{\textcolor{ta2skyblue}} % phantom introduces a vertical space in p formatted table columns??!!
 \usepackage[scaled=.7]{beramono}
@@
 \newlength{\columnheight}
-\setlength{\columnheight}{105cm}
+\setlength{\columnheight}{100cm}
+
+\input{../abstract/98-acronyms}
 
@@
 
-\begin{frame}
-  \begin{columns}
-    % ---------------------------------------------------------%
-    % Set up a column 
-    \begin{column}{.49\textwidth}
+\begin{frame}[t]
+  \noindent\makebox[\textwidth][c]{%}
+  \begin{minipage}[t][\columnheight][c]{0.95\textwidth}
+  \parbox[t][\columnheight][c]{\textwidth}{%
+  \begin{columns}[T,totalwidth=\textwidth]
+    \begin{column}{.34\linewidth}
       \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
         \begin{minipage}[T]{.95\textwidth}  % tweaks the width, makes a new \textwidth
-          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
+          \parbox[t]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
             % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
@@
             % fill each column with content            
-            %\input{motivation}
-%            \vfill
-            %\input{methods}
-%            \vfill
-            %\input{conclusions}
-            %\vfill
+            \begin{alertblock}{}
+              \begin{itemize}
+                \item Connectivity analyses rely on complex workflows to extract the network from dMRI datasets.
+                \item One important pitfall that potentially bias the extracted connectome is susceptibility
+                      distortion, a typical artifact on dMRI [Irfanoglu et al. 2012].
+                \item In this work, we evaluate three widely used methodologies for bias correction, originally
+                      proposed for fMRI data:
+                      fieldmap-based method (FMB, [Jezzard et al 1995]),
+                      reverse-encoding method (REB, [Cordes et al. 2000, Chiou et al. 2000]),
+                      and T2-weighted intensity-based registration (T2B, [Kybic et al. 2000]).
+                \item Benchmarking includes geometrical accuracy scores, signal recovery scores, and
+                      a preliminary study of impact on the extracted tractography and connectivity matrices.
+              \end{itemize}
+            \end{alertblock}
           }
@@
     % Set up a column 
-    \begin{column}{.49\textwidth}
+    \begin{column}{.64\linewidth}
       \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
         \begin{minipage}[T]{.95\textwidth} % tweaks the width, makes a new \textwidth
-          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
+          \parbox[t]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
             % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
@@
             % fill each column with content
-	    %\input{dataset}
-        %\input{results}
-        %\input{acknowledgements}
+	        \begin{block}{Digital dMRI phantom \& theory-based warping}
+          We generated a test set using low-resolution dMRI phantoms (online available$^{1}$)
+          with corresponding T1-weighted and T2-weighted images at high-resolution.
+
+          We simulate the artifact (geometrical distortion and signal dropout) using a synthetic fieldmap.
+              \begin{figure}[thpb]
+                 \centering
+                 \includegraphics[width=0.95\columnwidth]{Fig01-Phantom}
+              \end{figure}
+              A) T2w; B) undistorted \textit{b0} volume;
+                 C, D) distorted \textit{b0} volumes with opposed phase 
+                 encoding directions, maximum displacement of 3.80~mm.
+          \end{block}
+
+          \begin{block}{Evaluation framework}
+          We use \emph{nipype}$^{2}$, a powerful tool for building processing pipelines in
+          neuroimaging.
+
+          The evaluation framework includes the phantom distortion module, the three correction
+          methodologies, DTI\&HARDI reconstruction methodologies, tractography,
+          and a final module to analyse downstream impact on geometry, tractography and connectivity.
+          \end{block}
           }
@@
   \end{columns}
+  \noindent
+  \vfill
+      \begin{beamercolorbox}[wd=\textwidth]{postercolumn}
+        \begin{minipage}[T]{\textwidth} % tweaks the width, makes a new \textwidth
+          \parbox[t]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
+            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
+            % ---------------------------------------------------------%
+            % fill each column with content
+            \begin{block}{Visual results}
+              \begin{figure}
+                 \centering
+                 \includegraphics[width=\textwidth]{Fig02-Results}
+              \end{figure}
+            \end{block}
+            }
+        \end{minipage}
+      \end{beamercolorbox}
+  \vfill
+  \begin{columns}[T,totalwidth=\textwidth]
+    \begin{column}{.34\textwidth}
+      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
+        \begin{minipage}[T]{\textwidth}  % tweaks the width, makes a new \textwidth
+          \noindent\parbox[t]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
+            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
+            % ---------------------------------------------------------%
+            % fill each column with content
+            \vfill            
+            \begin{block}{Quantitative results}
+            \noindent\makebox[\textwidth][c]{%}
+            \begin{minipage}[T]{.75\textwidth} 
+            \input{../abstract/041-results-table}
+            \input{../abstract/042-results-tractography}
+            \end{minipage}
+            }
+            \end{block}
+            \vfill
+          }
+        \end{minipage}
+      \end{beamercolorbox}
+    \end{column}
+    % ---------------------------------------------------------%
+    % end the column
+
+    % ---------------------------------------------------------%
+    % Set up a column 
+    \begin{column}{.64\textwidth}
+      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
+        \begin{minipage}[T]{.95\textwidth} % tweaks the width, makes a new \textwidth
+          \parbox[t]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
+            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
+            % ---------------------------------------------------------%
+            % fill each column with content
+          \begin{block}{Conclusions and references}
+            \begin{itemize}
+              \item DTI dataset was rejected in evaluation: extracted
+                    connectivity matrices were biased, as the phantom
+                    is designed for high angular resolution methods.
+              \item In terms of geometry, the results indicate
+                    that REB method ranks first.
+              \item In terms of tractography, visual assessment and
+                    quantitative results suggest that FMB could be
+                    better.
+              \item Connectivity matrices are evaluated, but a more appropriate
+                    phantom is required. A phantom with the connecting interface
+                    densely covered by the seeding regions may be the key
+                    to characterize the impact on connectivity.
+              \item Connectome analyses demand the standardization of
+                    processing techniques and pipelining sofware tools to
+                    ensure replicability of experiments and reliability of
+                    results.
+            \end{itemize}
+
+            \noindent{\vskip1cm\textbf{Links and references}}\par
+            \begin{enumerate}
+            \item {\small\url{emmanuelcaruyer.com/phantomas.php}}\par
+            \item {\small\url{nipy.sourceforge.net/nipype}}
+            \end{enumerate}
+          \end{block}
+          }
+          % ---------------------------------------------------------%
+          % end the column
+        \end{minipage}
+      \end{beamercolorbox}
+    \end{column}
+    % ---------------------------------------------------------%
+    % end the column
+  \end{columns}
+%  }
+  %\vfill
+  %\begin{alertblock}{}
+  %    1. {\small\url{emmanuelcaruyer.com/phantomas.php}} \par
+  %    2. {\small\url{nipy.sourceforge.net/nipype}}
+  %\end{alertblock}
+  }
+  \end{minipage}
+}
 \end{frame}

### 2014-04-23 Publications/2013-ISBI14/poster/00-main.tex
Last version sent to MJLC and AS
Final version!

@@
 % @Last Modified by:   oesteban
-% @Last Modified time: 2014-04-22 20:49:46
+% @Last Modified time: 2014-04-23 16:15:17
 
@@
 
+\usepackage{float}
 \usepackage{glossaries}
@@
 %\boldmath
+\usepackage[labelformat=empty]{caption}
 \usepackage[orientation=portrait,size=a0,scale=1.4]{beamerposter}
@@
 \newlength{\columnheight}
-\setlength{\columnheight}{100cm}
+\setlength{\columnheight}{102cm}
 
@@
   \parbox[t][\columnheight][c]{\textwidth}{%
-  \begin{columns}[T,totalwidth=\textwidth]
+  \begin{columns}[t,totalwidth=\textwidth]
     \begin{column}{.34\linewidth}
@@
   \end{columns}
-  \noindent
   \vfill
-      \begin{beamercolorbox}[wd=\textwidth]{postercolumn}
-        \begin{minipage}[T]{\textwidth} % tweaks the width, makes a new \textwidth
-          \parbox[t]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
-            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
-            % ---------------------------------------------------------%
-            % fill each column with content
+  \noindent\makebox[\textwidth][l]{
+  \begin{minipage}{0.978\textwidth}
+      \begin{beamercolorbox}[]{postercolumn}
             \begin{block}{Visual results}
@@
             \end{block}
+      \end{beamercolorbox}
+  \end{minipage}
   }
-        \end{minipage}
-      \end{beamercolorbox}
   \vfill
@@
             \noindent\makebox[\textwidth][c]{%}
-            \begin{minipage}[T]{.75\textwidth} 
+            \begin{minipage}[T][][c]{.75\textwidth} 
             \input{../abstract/041-results-table}
@@
             \vfill
+            \begin{exampleblock}{}
+            \begin{figure}
+                \centering
+                \begin{minipage}{0.45\textwidth}
+                \centering
+                \includegraphics[height=3.5cm]{logos/qrcode-github}
+                \caption{\tiny My GitHub}
+                \end{minipage}\hfill
+                \begin{minipage}{0.45\textwidth}
+                \centering
+                \includegraphics[height=3.5cm]{logos/qrcode-paper}
+                \caption{\tiny Paper}
+                \end{minipage}
+                \end{figure}
+            \end{exampleblock}
           }
@@
             \begin{itemize}
-              \item DTI dataset was rejected in evaluation: extracted
-                    connectivity matrices were biased, as the phantom
-                    is designed for high angular resolution methods.
-              \item In terms of geometry, the results indicate
-                    that REB method ranks first.
+              \item In terms of geometry, REB ranked first.
               \item In terms of tractography, visual assessment and
-                    quantitative results suggest that FMB could be
-                    better.
-              \item Connectivity matrices are evaluated, but a more appropriate
-                    phantom is required. A phantom with the connecting interface
-                    densely covered by the seeding regions may be the key
-                    to characterize the impact on connectivity.
+                    quantitative results suggested that FMB and REB perform better.
+              \item Only HARDI yielded useful results. DTI-based tractography erroneously
+                    estimated crossing and kissing fibers, therefore it was discarded
+                    from evaluation.
+              \item Connectivity matrices from HARDI were evaluated. Still, a more appropriate
+                    phantom is required, presenting its connecting interface
+                    densely covered by the seeding regions.
               \item Connectome analyses demand the standardization of
                     processing techniques and pipelining sofware tools to
-                    ensure replicability of experiments and reliability of
-                    results.
+                    ensure the reproducibility of experiments and the 
+                    reliability of results.
             \end{itemize}

### 2014-04-23 Publications/2013-ISBI14/poster/00-main.tex
Fixed typos, title in caps

@@
 % @Last Modified by:   oesteban
-% @Last Modified time: 2014-04-23 16:15:17
+% @Last Modified time: 2014-04-23 16:31:02
 
@@
 
-\title{\huge Simulation-based evaluation of susceptibility distortion correction methods in dMRI}
+\title{\huge SIMULATION-BASED EVALUATION OF SUSCEPTIBILITY DISTORTION CORRECTION METHODS IN DIFFUSION MRI}
 \author[Esteban et al.]{O. Esteban$^{1,2}$\and%
@@
           The evaluation framework includes the phantom distortion module, the three correction
-          methodologies, DTI\&HARDI reconstruction methodologies, tractography,
-          and a final module to analyse downstream impact on geometry, tractography and connectivity.
+          methodologies, DTI\&HARDI reconstruction methodologies, streamline tractography,
+          and a final module to analyze downstream impact on geometry, tractography and connectivity.
           \end{block}
@@
               \item Connectome analyses demand the standardization of
-                    processing techniques and pipelining sofware tools to
+                    processing techniques and pipelining software tools to
                     ensure the reproducibility of experiments and the 

### 2014-10-20 Scripts/pyacwereg/workflows/evaluation.py
implemented warp_image for #146

@@ -96,7 +96,8 @@ def bspline(name='BSplineEvaluation', n_tissues=3, methods=None, results=None):
                                       ('in_surfs',   'refnode.in_surf'),
                                       ('in_mask',    'refnode.in_mask'), ]),
             (inputnode,         reg, [('in_surfs', 'inputnode.in_surf'),
-                                      ('in_file', 'inputnode.in_orig')]),
+                                      ('in_file', 'inputnode.in_orig'),
+                                      ('grid_size', 'inputnode.grid_size')]),
             (dist,              reg, [
                 ('outputnode.out_file', 'inputnode.in_dist'),
                 ('outputnode.out_tpms', 'inputnode.in_tpms'),

### 2014-10-20 Scripts/pyacwereg/workflows/registration.py

@@ -24,7 +24,7 @@ def default_regseg(name='REGSEGDefault'):
 
     inputnode = pe.Node(niu.IdentityInterface(
                         fields=['in_orig', 'in_dist', 'in_tpms', 'in_surf',
-                                'in_mask']), name='inputnode')
+                                'in_mask', 'grid_size']), name='inputnode')
 
     outputnode = pe.Node(niu.IdentityInterface(
                          fields=['out_corr', 'out_tpms',
@@ -40,7 +40,7 @@ def default_regseg(name='REGSEGDefault'):
     regseg.inputs.step_size = [1.0]
     regseg.inputs.alpha = [0.0]
     regseg.inputs.beta = [0.0]
-    regseg.inputs.grid_size = [6]
+    # regseg.inputs.grid_size = [10]
     regseg.inputs.convergence_energy = [True]
     regseg.inputs.convergence_window = [50]
 
@@ -51,7 +51,8 @@ def default_regseg(name='REGSEGDefault'):
     # Connect
     wf.connect([
         (inputnode,   regseg, [('in_surf', 'in_prior'),
-                               ('in_dist', 'in_fixed')]),
+                               ('in_dist', 'in_fixed'),
+                               ('grid_size', 'grid_size')]),
         (inputnode, applytfm, [('in_tpms', 'in_file'),
                                ('in_mask', 'in_mask')]),
         (regseg,    applytfm, [('out_field', 'in_field')]),
@@ -74,7 +75,7 @@ def identity_wf(name='Identity', n_tissues=3):
 
     inputnode = pe.Node(niu.IdentityInterface(
                         fields=['in_orig', 'in_dist', 'in_tpms', 'in_surf',
-                                'in_mask', 'in_field']),
+                                'in_mask', 'in_field', 'grid_size']),
                         name='inputnode')
 
     outputnode = pe.Node(niu.IdentityInterface(
@@ -98,10 +99,11 @@ def identity_wf(name='Identity', n_tissues=3):
         (inputnode,     merge, [('in_dist', 'in1'),
                                 ('in_tpms', 'in2')]),
         (inputnode,  applytfm, [('in_mask', 'in_mask'),
-                                ('in_surf', 'in_surf')]),
+                                ('in_surf', 'in_surf'),
+                                ('grid_size', 'grid_size')]),
         (merge,      applytfm, [('out', 'in_file')]),
-        (inputnode,  applytfm, [('in_field', 'in_field')]),
[…]

### 2014-10-20 Publications/2013-TMI/03-introduction.tex
update TMI paper

@@
+% -*- root: 00-main.tex -*-
 \section{Introduction}
@@
   connectivity analyses of the brain at the macro-scale
-\citep{craddock_imaging_2013}. This emerging field exploits
-\gls{dwi} data to derive the local axonal structure at each imaging voxel 
-\citep{basser_microstructural_1996} and estimate a whole-brain mapping of fiber 
-tracts \citep{hagmann_mr_2012,johansen-berg_using_2009} represented by 
-trajectories reconstructed from the local information.
+  \citep{craddock_imaging_2013}.
+This emerging field exploits \gls{dwi} data to derive the local axonal structure at
+  each imaging voxel \citep{basser_microstructural_1996} and estimate a whole-brain
+  mapping of fiber tracts \citep{hagmann_mr_2010,berg_using_2009} trajectories
+  reconstructed from the local information.
 This comprehensive map of neural connections of the brain is called the
-\emph{connectome} \citep{hagmann_diffusion_2005,sporns_human_2005}. The connectome
-analysis is a promising tool for neuroscience \citep{morgan_why_2013} and 
-for the clinical application \citep{griffa_structural_2013} as well.
+  \emph{connectome} \citep{hagmann_from_2005,sporns_human_2005}.
+The connectome analysis is a promising tool for neuroscience \citep{morgan_why_2013}
+  and for the clinical application \citep{griffa_structural_2013} as well.
 
@@
 %methodologies by a better understanding of the diffusion effect, and improving
-%image reconstruction methodologies. Currently, the connectome extraction 
+%image reconstruction methodologies.
+%Currently, the connectome extraction
 %and analysis relies on a large chain of sophisticated computational techniques
@@
 %and registration, fibre tracking, connectivity mapping, visualization, statistical
-%group studies, and inference \citep{daducci_connectome_2012}. This growing complexity 
+%group studies, and inference \citep{daducci_connectome_2012}.
+%This growing complexity
 %has given rise to challenging issues towards reliable structural information about
 %the neuronal tracts presented in, and also difficulties on the statistical analysis
-%\citep{bullmore_complex_2009,meskaldji_comparing_2013}. To our knowledge, studies 
+%\citep{bullmore_complex_2009,meskaldji_comparing_2013}.
+%To our knowledge, studies
 %regarding the impact of the different sources of inaccuracy and variability along
@@
 \label{sec:dwi_overview}
-\gls{dwi} data are usually acquired with \gls{epi} schemes
-as they allow for very fast acquisition of large sets of slices each
-acquired after a single excitation \citep{mukherjee_diffusion_2008,
-soares_hitchhikers_2013}. \Gls{epi} volumes are collected
-in sequence and each one represents one probed diffusion direction.
+\gls{dwi} data are usually acquired with \gls{epi} schemes as they allow for very
+  fast acquisition of large sets of slices each acquired after a single excitation
+  \citep{mukherjee_diffusion_2008,soares_hitchhikers_2013}.
+\Gls{epi} volumes are collected in sequence and each one represents one probed
+  diffusion direction.
 %This first acquisition is called \emph{raw} data or \emph{diffusion
-%signal}. One or more volumes without directional encoding (so-called \emph{baseline},
+%signal}.
+%One or more volumes without directional encoding (so-called \emph{baseline},
 %$b = 0$, \emph{low $b$}, or just \emph{b0} volumes) are interleaved as
 %reference, at the least one is acquired at the beginning of the
-%\gls{epi} dataset. These B0 are acquired without direction gradient,
+%\gls{epi} dataset.
+%These B0 are acquired without direction gradient,
 %showing a T2-like contrast.
@@
   different sampling directions, the induced \emph{eddy currents} on the scanner
-coils due to the gradient switching \citep{obrien_3d_2013}, and, finally, 
+  coils due to the gradient switching \citep{brien_3d_2013}, and, finally,
   distortions caused by the magnetic susceptibility inhomogeneity
-present at tissue interface. \emph{Susceptibility distortions} happen along the 
-phase-encoding direction, and are most appreciable in the frontal part of 
-the brain due to the air/tissue interface located around the sinuses.
+  present at tissue interfaces.
+\emph{Susceptibility distortions} happen along the phase-encoding direction,
+  and are most appreciable in the frontal part of the brain due to the air/tissue
+  interface located around the sinuses.
 Two implications are associated to this artifact: the signal loss caused on
   highly distorted regions, and a significant added difficulty in
-registering with structural images (e.g. T1-weighted).
+  registering with structural images (e.g.
+T1-weighted).
 
 After correcting for artifacts, the difficulty of performing analyses directly
-on the \emph{directionally dependent} raw data is not affordable. Thus, 
-\gls{dwi} data is reconstructed using one of a numerous set of readily 
-available algorithms. The outcome is a voxel-wise model of diffusion shape 
-(\citep{daducci_quantitative_2013} contains a comprehensive review of 
-state-of-art methodologies). On top of the
-chosen model, some scalar maps describing tissue features are generally
-computed (e.g. \gls*{fa}, and \gls*{adc}).
+  on the \emph{directionally dependent} raw data is not affordable.
+Thus, \gls{dwi} data is reconstructed using one of a numerous set of readily
+  available algorithms.
+The outcome is a voxel-wise model of diffusion shape (\citep{daducci_quantitative_2014}
+  contains a comprehensive review of state-of-art methodologies).
+On top of the chosen model, some scalar maps describing tissue features are generally
+  computed (e.g.
+\gls*{fa}, and \gls*{adc}).
 %For instance, the most popular \gls{dwi}-derived scalar map
@@
 %Another widely studied map is the \gls{adc} (or \gls{md} that has a dual definition),
-%which reflects the intensity of the diffusion effect. Many other
-%directionally invariant measurements have been defined. For a comprehensive study, 
+%which reflects the intensity of the diffusion effect.
+%Many other
+%directionally invariant measurements have been defined.
+%For a comprehensive study,
 %we refer to \citep{ennis_orthogonal_2006}.
 
-Alongside distortions and direction dependency, \gls{dwi} data is strongly 
-affected by \glspl{pve} \citep{alexander_analysis_2001} that appears when
-several different tissues are present in the same imaging unit, producing
-an averaged intensity. The effect is directly related to the resolution,
-typically around $2.0\times2.0\times2.0mm^3$ for \gls{dwi}.
+Alongside distortions and direction dependency, \gls{dwi} data is strongly affected
+  by \glspl{pve} \citep{alexander_analysis_2001} that appears when several different
+  tissues are present in the same imaging unit, producing an averaged intensity.
+The effect is directly related to the resolution, typically around
+  $2.0\times2.0\times2.0mm^3$ for \gls{dwi}.
 
 % the low resolution achievable with \gls{dwi}
-%(typically around ). An additional complication specific to
+%(typically around ).
+%An additional complication specific to
 %\gls{dwi} is the \gls{csf} \emph{contamination} \citep{metzler-baddeley_how_2012},
@@
 
-One approach to correcting the susceptibility distortions was proposed
-with the earliest \gls{epi} sequences used in functional \gls{mri}, and
-relies on the acquisition of extra \gls{mri} data. Generally, a \gls{gre} is used to 
-obtain a map of magnitude and phase of the actual magnetic field inside the
-scanner. Based on this \emph{fieldmap} and the theory underlying the 
-distortion, \citep{jezzard_correction_1995} proposed a correction 
-methodology. A number of forked and improved methodologies have been 
-developed to correct for the susceptibility distortion and received the
-generic name of ``fieldmap correction'' techniques \cite{hsu_correction_2009,
-reber_correction_1998}.
+One approach to correcting the susceptibility distortions was proposed with the
+  earliest \gls{epi} sequences used in functional \gls{mri}, and relies on the
+  acquisition of extra \gls{mri} data.
+Generally, a \gls{gre} is used to obtain a map of magnitude and phase of the actual
+  magnetic field inside the scanner.
+Based on this \emph{fieldmap} and the theory underlying the distortion,
+  \citep{jezzard_correction_1995} proposed a correction methodology.
+A number of forked and improved methodologies have been developed to correct
+  for the susceptibility distortion and received the generic name of
+  ``fieldmap correction'' techniques \cite{hsu_correction_2009,reber_correction_1998}.
 
-Some other retrospective methodologies do not make use of the fieldmaps, 
-as explicitly modeling the distortion \citep{andersson_modeling_2001},
-registering with (anatomically correct) T2-weighted \gls{mri} 
-\citep{kybic_unwarping_2000,studholme_accurate_2000,li_accounting_2007,
-tao_variational_2009}, or acquiring an 
-extra B0 image with reversed phase encoding direction 
-\citep{andersson_how_2003,holland_efficient_2010}.
+Some other retrospective methodologies do not make use of the fieldmaps, as explicitly
+  modeling the distortion \citep{andersson_modeling_2001}, registering with (anatomically
+  correct) T2-weighted \gls{mri} \citep{kybic_unwarping_2000,studholme_accurate_2000,
+  li_accounting_2007,tao_variational_2009}, or acquiring an extra B0 image with reversed
+  phase encoding direction \citep{andersson_how_2003,holland_efficient_2010}.
 
@@
 This study claims that fieldmap correction methodologies are not entirely accurate
-and reliable, even though the method is correct in principle. The conclusion
-was later confirmed by \citep{tao_variational_2009}. Additional concerns regarding
-fieldmap correction are the requirement of an extra acquisition (that is not always
-met or it is impractical), or the accuracy of the measured fieldmap that is sensitive
-to various effects (such us respiration, blood flow, etc.). All these factors 
-have turned susceptibility distortion in \gls{epi} sequences an active field of 
+  and reliable, even though the method is correct in principle.
+The conclusion was later confirmed by \citep{tao_variational_2009}.
+Additional concerns regarding fieldmap correction are the requirement of an extra acquisition
+  (that is not always met or it is impractical), or the accuracy of the measured fieldmap that
+  is sensitive to various effects (such us respiration, blood flow, etc.).
+All these factors have turned susceptibility distortion in \gls{epi} sequences an active field of
   research for the last 10 years.
@@
   god (yet) at identifying where the tracts project into the \gls{gm}
-(limitations on the termination criteria, \citep{jbabdi_tractography:_2011,
+  (limitations on the termination criteria, \citep{jbabdi_tractography_2011,
   craddock_imaging_2013}).
 Thus, a \gls{wm} segmentation is required by most of the available tractography
-algorithms to \emph{filter} the resulting tractogram. The \gls{gm}-\gls{wm}
-interface is necessary to locate the starting and ending points of the detected
-fiber bundles, and \gls{csf}-\gls{wm} surface is critical for pruning spurious 
-and discontinued fiber bundles.
+algorithms to \emph{filter} the resulting tractogram.
+The \gls{gm}-\gls{wm} interface is necessary to locate the starting and ending points
+  of the detected fiber bundles, and \gls{csf}-\gls{wm} surface is critical for pruning
+  spurious and discontinued fiber bundles.
 
@@
   they were generally limited to certain regions or significant fiber tracts, and thus,
-it cannot be applied in whole-brain tractography. Early approaches to \gls{dwi} segmentation 
+  it cannot be applied in whole-brain tractography.
+Early approaches to \gls{dwi} segmentation
   include level set formulations using scalar maps of direction invariants derived
@@
   \citep{rousson_level_2004}, or finding alternative diffusion representations
-\citep{jonasson_representing_2007}. Even though this latter case was restricted to the extraction of
+  \citep{jonasson_representing_2007}.
+Even though this latter case was restricted to the extraction of
   the corpus callosum from a real dataset, the density of the components of the diffusion tensor
-are approximated by multivariate Gaussians for first. Iterative clustering performed on the 
-B0 volumes of \gls{dwi} data was proposed by \citep{hadjiprocopis_unbiased_2005}.
+  are approximated by multivariate Gaussians for first.
+Iterative clustering performed on the B0 volumes of \gls{dwi} data was proposed by
+  \citep{hadjiprocopis_unbiased_2005}.
 Later studies investigated the application of probabilistic frameworks combining mixtures of
   gaussians, \gls{mrf} and labeling fusion techniques \citep{liu_brain_2007} using as features
-widely-used \gls{dwi}-derived scalar maps as \gls{fa} or \gls{md}. A similar framework 
-combining co-registered structural information (T1 weighted) with \emph{independent orthogonal 
-invariants} derived from the \gls{dwi} tensor model was proposed by \citep{awate_multivariate_2008}. 
+  widely-used \gls{dwi}-derived scalar maps as \gls{fa} or \gls{md}.
+A similar framework combining co-registered structural information (T1 weighted) with
+  \emph{independent orthogonal invariants} derived from the \gls{dwi} tensor model was proposed
+  by \citep{awate_multivariate_2008}.
 Some proposals suggest the use of the raw diffusion data (directionally dependent) to avoid fitting
-a certain model \citep{lu_segmentation_2008}. In \citep{han_experimental_2009}, graph-cuts 
-voxel-based techniques are proposed using the most common diffusion tensor derived features.
+  a certain model \citep{lu_segmentation_2008}.
+In \citep{han_experimental_2009}, graph-cuts voxel-based techniques are proposed using the most
+  common diffusion tensor derived features.
 Further developments of the probabilistic approach have been proposed adding more scalar maps
@@
 None of the presented methods have claimed for definite results, mainly due to the lack of
-a \emph{gold-standard} evaluation methodology. Most of them are tested only on certain
+  a \emph{gold-standard} evaluation methodology.
+Most of them are tested only on certain
   regions, or do not provide sub-pixel resolution results
   \citep{hadjiprocopis_unbiased_2005,liu_brain_2007,awate_multivariate_2008,lu_segmentation_2008,
-han_experimental_2009}. Generally, results obtained with high resolution atlas co-registration
+  han_experimental_2009}.
+Generally, results obtained with high resolution atlas co-registration
   (\autoref{sec:registration}) are more compelling, limiting the activity on \gls{dwi}
-segmentation. However, recently \citep{kumazawa_improvement_2013} proposed to include this
-task within reconstruction algorithms. Finally, some \emph{golden}-standard evaluation
-frameworks for task-based validation have been proposed for the segmentation results
-(e.g. lesion detection in visceral organs \citep{jha_task-based_2012}).
+  segmentation.
+However, recently \citep{kumazawa_improvement_2013} proposed to include this
+  task within reconstruction algorithms.
+Finally, some \emph{golden}-standard evaluation frameworks for task-based validation have been
+  proposed for the segmentation results (e.g. lesion detection in visceral organs
+  \citep{jha_taskbased_2012}).
 \todo[inline]{reference to a summary table}
@@
 The need for this registration step is mainly raised when defining the
-nodes of the connectivity matrix. This definition step is named
+nodes of the connectivity matrix.
+This definition step is named
 \textit{cortical parcellation} as it imposes the regions that will be
 considered to cluster the fiber bundles reaching the \gls{wm}-\gls{gm}
-surface. The common procedure to accomplish this clustering is using
+surface.
+The common procedure to accomplish this clustering is using
 a predefined parcellation in a high-resolution structural T1-weighted
-atlas (e.g. \citep{greve_accurate_2009}). The second goal that has 
+atlas (e.g.
+\citep{greve_accurate_2009}).
+The second goal that has
 justified the applicability of structural
 images co-registration is susceptibility distortion correction
-as presented in \autoref{sec:distortion}. In any case, susceptibility 
+as presented in \autoref{sec:distortion}.
+In any case, susceptibility
 distortion hinders a rigid-registration solution to this problem.
@@
 independent} existing in the raw \gls{dwi} data) due to the almost null
-difference in intensity between \gls{wm} and \gls{gm} pixels. Early 
+difference in intensity between \gls{wm} and \gls{gm} pixels.
+Early
 registration methods appeared targeting the distortion correction.
 They quickly standardized the choice of T2 as anatomically correct source
-to be registered against the B0 in \gls{dwi}. B0 images have a very similar
+to be registered against the B0 in \gls{dwi}.
+B0 images have a very similar
 contrast to T2 weighted due to the parallelism of acquisition sequences.
 However, parcellation is defined in T1 weighted space, adding an additional
-registration step is generally overlooked by existing literature. Even though this
+registration step is generally overlooked by existing literature.
+Even though this
 processing stage has a very low complexity compared to the remaining tasks of
@@
 field is modeled with B-Splines, the cost function is least squares and optimized
-in a multi-resolution gradient descent strategy. Their method is evaluated in 
-both synthetic and real 2D images. Similarly,
+in a multi-resolution gradient descent strategy.
+Their method is evaluated in
+both synthetic and real 2D images.
+Similarly,
 \citep{studholme_accurate_2000} proposed a spline-based deformation but including
 a weighting factor proportional to the Jacobian of the transform to correct the
-intensity of the undistorted data. They also use the $\log$ of the signal to
+intensity of the undistorted data.
+They also use the $\log$ of the signal to
 enhance the low-signal regions, and optimize with a gradient descent algorithm
-the mutual information of the mapping. This framework set the basis for the following
-works. For instance, \citep{wu_comparison_2008} also proposed a B-Spline registration
-providing quantitative comparisons with fieldmap correction methods. Recent approaches
+the mutual information of the mapping.
+This framework set the basis for the following
+works.
+For instance, \citep{wu_comparison_2008} also proposed a B-Spline registration
+providing quantitative comparisons with fieldmap correction methods.
+Recent approaches
 take into account the signal loss due to dephasing \citep{li_accounting_2007}, or
@@
 The significant benefits of exploiting the anatomical \gls{mri} when
-segmenting the \gls{dwi} data have been demonstrated \cite{zollei_improved_2010}, 
-justifying the use of the shape prior information. To our knowledge, there is
+segmenting the \gls{dwi} data have been demonstrated \cite{zoellei_improved_2010},
+justifying the use of the shape prior information.
+To our knowledge, there is
 no study simultaneously taking advantage from segmentation or distortion correction
@@
 are very sensitive to distortions, \gls{pve}, noise and axonal dispersion
-\citep{jbabdi_tractography:_2011}, what yields numerous challenges in the
+\citep{jbabdi_tractography_2011}, what yields numerous challenges in the
 processing tasks (artifact correction, segmentation, registration and
-parcellation). Finally, as a result a vast number of challenges and pitfalls 
+parcellation).
+Finally, as a result a vast number of challenges and pitfalls
 emerge regarding accuracy (correctness) and precision (reproducibility) of
@@
 spatial mapping between these contours and the corresponding surfaces in
-anatomical images bear significant redundancy. Once the spatial relationship 
+anatomical images bear significant redundancy.
+Once the spatial relationship
 between \gls{t1} and \gls{dwi} space is established, the contours which are
 readily available in \gls{t1} space can simply be projected on to the
-\gls{dwi} data. Conversely, if a precise delineation in \gls{dwi} space 
+\gls{dwi} data.
+Conversely, if a precise delineation in \gls{dwi} space
 was achieved, the spatial mapping with \gls{t1}-space could be derived
-from one-to-one correspondences on the contours. However, neither segmentation 
+from one-to-one correspondences on the contours.
+However, neither segmentation
 nor registration can be performed flawlessly, if considered independently.
@@
 These tasks are generally
-solved independently, or combined in pairs. However, there exist fundamental 
+solved independently, or combined in pairs.
+However, there exist fundamental
 coupling relationships that can be exploited to obtain a simultaneous solution to
-the three problems. This joint approach satisfactory impacts the downstream
+the three problems.
+This joint approach satisfactory impacts the downstream
 outcomes of the whole pipeline with the increase of the internal consistency.
@@
 by exploiting as strong shape-prior the detailed morphology extracted
-from high-resolution and anatomically correct \gls{mri}. Indeed, hereafter 
+from high-resolution and anatomically correct \gls{mri}.
+Indeed, hereafter
 we assume this segmentation problem in anatomical images is reliably and
-accurately solved with readily available tools (e.g. \citep{fischl_freesurfer_2012}).
+accurately solved with readily available tools (e.g.
+\citep{fischl_freesurfer_2012}).
 After global alignment with \gls{t1} using existing approaches, the remaining
@@
 from the structural space into the diffusion space, such that the structural
-contours segment optimally the \gls{dwi} data. In the process, the one-to-one 
+contours segment optimally the \gls{dwi} data.
+In the process, the one-to-one
 correspondence between the contours in both spaces is guaranteed, and projection
@@
 We test our proposed joint segmentation-registration model on two different
-synthetic examples. The first example is a scalar sulcus model, where the 
+synthetic examples.
+The first example is a scalar sulcus model, where the
 \gls{csf}-\gls{gm} boundary particularly suffers from \gls{pve} and can only be
@@
 The second case deals with more realistic \gls{dwi} data stemming from
-phantom simulations of a simplistic brain data. Again, we show that the 
+phantom simulations of a simplistic brain data.
+Again, we show that the
 proposed model successfully segments the \gls{dwi} data based on two derived
@@
 
-The rest of this paper is organized as follows. First, in \autoref{sec:methods}
+The rest of this paper is organized as follows.
+First, in \autoref{sec:methods}
 we introduce our proposed model for joint multivariate segmentation-registration.
 Then we provide a more detailed description of the data and experimental setup in
-\autoref{sec:experiments}. We present results in \autoref{sec:results} and conclude 
+\autoref{sec:experiments}.
+We present results in \autoref{sec:results} and conclude
 in \autoref{sec:conclusion}.

### 2014-10-20 Publications/2013-TMI/04-methods.tex
update TMI paper

@@
-1\section{Methods}
+% -*- root: 00-main.tex -*-
+\section{Methods}
 \label{sec:methods}
@@
 The first group typically adds a shape prior term to the energy functional of
-an evolving active contour \citep{bresson_variational_2006,
-chan_level_2005,chen_using_2002,cremers_kernel_2006,gastaud_combining_2004}.
+\citep{chan_level_2005,chen_using_2002,cremers_kernel_2006,gastaud_combining_2004,wyatt_map_2003,paragios_level_2003,yezzi_variational_2003,gorthi_segmentation_2009,pohl_unifying_2005,pohl_bayesian_2006,wang_joint_2006,guyader_combined_2011,besson_dream2s_2003,bertero_illposed_1988,cote_tractometer_2013}
+an evolving active contour \citep{bresson_variational_2006, chan_level_2005,
+chen_using_2002,cremers_kernel_2006,gastaud_combining_2004}.
 These methods generally have an explicit description of the expected relative boundary
@@
 the \gls{acwe} framework \citep{chan_active_2001}, and related to
-\citep{le_guyader_combined_2011}.
+\citep{guyader_combined_2011}.
 {\color{red} {The main diverging points with respect to
@@
 $\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
-gradients \citep{jehan-besson_dream2s:_2003,herbulot_segmentation_2006};
+gradients \citep{besson_dream2s_2003,herbulot_segmentation_2006};
 2) regularization is also based on linear diffusion smoothing \citep{thirion_image_1998},
@@
 domain, observing anisotropic and inhomogeneous mappings along each direction.
-With respect \citep{le_guyader_combined_2011}, the main differences are
+With respect \citep{guyader_combined_2011}, the main differences are
 the distance function, and the spectral solution to the optimization updates,
@@
 \eqref{eq:final_map_energy}. This inverse problem is ill-posed
-\cite{bertero_ill-posed_1988,hadamard_sur_1902}.
+\cite{bertero_illposed_1988,hadamard_sur_1902}.
 In order to account for deformation field regularity and to render the
@@
 \begin{equation}
-\vec{c}_i' = \vec{c}_i + \sum_k \left[ \vec{u}_k \, \underset{n}{\prod} B_3( (\vec{c}_i - \vec{x}_k) \cdot \hat{\mathbf{e}}_d ) \right].
+\vec{c}_i' = \vec{c}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod} B_3( (\vec{c}_i - \vec{x}_k) \cdot \hat{\mathbf{e}}_d ) \right].
 \label{eq:transformation}
@@
 we alternatively use \emph{shape-gradients}
-\cite{jehan-besson_dream2s:_2003,herbulot_segmentation_2006}.
+\cite{besson_dream2s_2003,herbulot_segmentation_2006}.
 Let $r(\vec{x})$ be an ``arbitrary'' function over the image domain,

### 2014-10-20 Publications/2013-TMI/05-experiments.tex
update TMI paper

@@
+% -*- root: 00-main.tex -*-
 \section{Data and experiments}
@@
 The lack of a widely accepted \emph{gold}-standard in the application
-field has been addressed by several authors \citep{cote_tractometer:_2013}
+field has been addressed by several authors \citep{cote_tractometer_2013}
 In this work, we use one of the most complete, and publicly available,

### 2014-10-20 Publications/2013-TMI/02-abstract.tex
updating intro

% -*- root: 00-main.tex -*-
Integrating information from different images within- and between-subjects
  is an ubiquitous necessity required in current neuroimage analyses.
Many algorithms have been proposed to register image pairs from the same
  subject and between subjects to resolve the problem.
Both images are required to present comparable features on top of which
  an objective function is built to evaluate fitness of the solution.
However, oftentimes one of the images is not available or the objective
  function is not adequate or numerically tractable.

We propose a surface-to-volume variational solution to the registration problem,
  that exploits the detailed anatomy extracted manually or from structural images.
We use an ``active contours without edges''-like model to search for a
  deformation field that optimally maps the surface prior onto the data
  in the target space, implicitly segmenting with sub-pixel resolution the
  volume.
We demonstrate the usefulness of the approach on a digital phantom, and on
  10 real datasets randomly selected from the Human Connectome Project.

### 2014-10-20 Publications/2013-TMI/03-introduction.tex
updating intro

@@
 
-\IEEEPARstart{D}{iffusion} \gls{mri} is a widely used family
-  of \gls{mri} techniques \citep{sundgren_diffusion_2004} which
-  accounts for a growing interest in its application to structural
-  connectivity analyses of the brain at the macro-scale
-  \citep{craddock_imaging_2013}.
-This emerging field exploits \gls{dwi} data to derive the local axonal structure at
-  each imaging voxel \citep{basser_microstructural_1996} and estimate a whole-brain
-  mapping of fiber tracts \citep{hagmann_mr_2010,berg_using_2009} trajectories
-  reconstructed from the local information.
-This comprehensive map of neural connections of the brain is called the
-  \emph{connectome} \citep{hagmann_from_2005,sporns_human_2005}.
-The connectome analysis is a promising tool for neuroscience \citep{morgan_why_2013}
-  and for the clinical application \citep{griffa_structural_2013} as well.
-
-% REMOVE paragraph (Dom, 1st comment)
-%Early \gls{dwi} research mainly focused on the improvements of imaging
-%methodologies by a better understanding of the diffusion effect, and improving
-%image reconstruction methodologies.
-%Currently, the connectome extraction
-%and analysis relies on a large chain of sophisticated computational techniques
-%including acquisition, reconstruction, modeling and model fitting, image segmentation
-%and registration, fibre tracking, connectivity mapping, visualization, statistical
-%group studies, and inference \citep{daducci_connectome_2012}.
-%This growing complexity
-%has given rise to challenging issues towards reliable structural information about
-%the neuronal tracts presented in, and also difficulties on the statistical analysis
-%\citep{bullmore_complex_2009,meskaldji_comparing_2013}.
-%To our knowledge, studies
-%regarding the impact of the different sources of inaccuracy and variability along
-%the toolchain for connectivity analysis \citep{jones_twenty-five_2010,
-%cheng_characteristics_2012,jones_white_2013} are scarce.
-
-\subsection{\Gls{dwi} data overview}
-\label{sec:dwi_overview}
-\gls{dwi} data are usually acquired with \gls{epi} schemes as they allow for very
-  fast acquisition of large sets of slices each acquired after a single excitation
-  \citep{mukherjee_diffusion_2008,soares_hitchhikers_2013}.
-\Gls{epi} volumes are collected in sequence and each one represents one probed
-  diffusion direction.
-%This first acquisition is called \emph{raw} data or \emph{diffusion
-%signal}.
-%One or more volumes without directional encoding (so-called \emph{baseline},
-%$b = 0$, \emph{low $b$}, or just \emph{b0} volumes) are interleaved as
-%reference, at the least one is acquired at the beginning of the
-%\gls{epi} dataset.
-%These B0 are acquired without direction gradient,
-%showing a T2-like contrast.
-
-\Gls{epi} are known to suffer from geometrical distortions and artifacts
-  due to, mainly, three sources: the subject motion in between acquiring
-  different sampling directions, the induced \emph{eddy currents} on the scanner
-  coils due to the gradient switching \citep{brien_3d_2013}, and, finally,
-  distortions caused by the magnetic susceptibility inhomogeneity
-  present at tissue interfaces.
-\emph{Susceptibility distortions} happen along the phase-encoding direction,
-  and are most appreciable in the frontal part of the brain due to the air/tissue
-  interface located around the sinuses.
-Two implications are associated to this artifact: the signal loss caused on
-  highly distorted regions, and a significant added difficulty in
-  registering with structural images (e.g.
-T1-weighted).
-
-After correcting for artifacts, the difficulty of performing analyses directly
-  on the \emph{directionally dependent} raw data is not affordable.
-Thus, \gls{dwi} data is reconstructed using one of a numerous set of readily
-  available algorithms.
-The outcome is a voxel-wise model of diffusion shape (\citep{daducci_quantitative_2014}
-  contains a comprehensive review of state-of-art methodologies).
-On top of the chosen model, some scalar maps describing tissue features are generally
-  computed (e.g.
-\gls*{fa}, and \gls*{adc}).
-%For instance, the most popular \gls{dwi}-derived scalar map
-%is \gls{fa} which depicts the isotropy of water diffusion inside the brain.
-%Another widely studied map is the \gls{adc} (or \gls{md} that has a dual definition),
-%which reflects the intensity of the diffusion effect.
-%Many other
-%directionally invariant measurements have been defined.
-%For a comprehensive study,
-%we refer to \citep{ennis_orthogonal_2006}.
-
-Alongside distortions and direction dependency, \gls{dwi} data is strongly affected
-  by \glspl{pve} \citep{alexander_analysis_2001} that appears when several different
-  tissues are present in the same imaging unit, producing an averaged intensity.
-The effect is directly related to the resolution, typically around
-  $2.0\times2.0\times2.0mm^3$ for \gls{dwi}.
-
-% the low resolution achievable with \gls{dwi}
-%(typically around ).
-%An additional complication specific to
-%\gls{dwi} is the \gls{csf} \emph{contamination} \citep{metzler-baddeley_how_2012},
-%that is a particular \gls{pve} in which the signal sensed inside the affected voxel is
-%linear with respect the \gls{gm} and \gls{wm} contributions, but non-linear with
-%respect \gls{csf}.
-
-\subsection{Correction for susceptibility distortions}
-\label{sec:distortion}
-
-One approach to correcting the susceptibility distortions was proposed with the
-  earliest \gls{epi} sequences used in functional \gls{mri}, and relies on the
-  acquisition of extra \gls{mri} data.
-Generally, a \gls{gre} is used to obtain a map of magnitude and phase of the actual
-  magnetic field inside the scanner.
-Based on this \emph{fieldmap} and the theory underlying the distortion,
-  \citep{jezzard_correction_1995} proposed a correction methodology.
-A number of forked and improved methodologies have been developed to correct
-  for the susceptibility distortion and received the generic name of
-  ``fieldmap correction'' techniques \cite{hsu_correction_2009,reber_correction_1998}.
-
-Some other retrospective methodologies do not make use of the fieldmaps, as explicitly
-  modeling the distortion \citep{andersson_modeling_2001}, registering with (anatomically
-  correct) T2-weighted \gls{mri} \citep{kybic_unwarping_2000,studholme_accurate_2000,
-  li_accounting_2007,tao_variational_2009}, or acquiring an extra B0 image with reversed
-  phase encoding direction \citep{andersson_how_2003,holland_efficient_2010}.
-
-A comparison of the diverse correction techniques is found in \citep{wu_comparison_2008}.
-This study claims that fieldmap correction methodologies are not entirely accurate
-  and reliable, even though the method is correct in principle.
-The conclusion was later confirmed by \citep{tao_variational_2009}.
-Additional concerns regarding fieldmap correction are the requirement of an extra acquisition
-  (that is not always met or it is impractical), or the accuracy of the measured fieldmap that
-  is sensitive to various effects (such us respiration, blood flow, etc.).
-All these factors have turned susceptibility distortion in \gls{epi} sequences an active field of
-  research for the last 10 years.
-
-\subsection{\Gls{dwi} segmentation}
-\label{sec:dwi_segmentation}
-
-A precise delineation of the \gls{csf}, \gls{gm} and \gls{wm} interfacing surfaces
-  is required with sub-pixel resolution, because whereas tractography algorithms are
-  very good at estimating the location of bundles in deep \gls{wm}, they are not
-  god (yet) at identifying where the tracts project into the \gls{gm}
-  (limitations on the termination criteria, \citep{jbabdi_tractography_2011,
-  craddock_imaging_2013}).
-Thus, a \gls{wm} segmentation is required by most of the available tractography
-algorithms to \emph{filter} the resulting tractogram.
-The \gls{gm}-\gls{wm} interface is necessary to locate the starting and ending points
-  of the detected fiber bundles, and \gls{csf}-\gls{wm} surface is critical for pruning
-  spurious and discontinued fiber bundles.
-
-A number of methodologies have been proposed for \gls{dwi} segmentation, ranging
-  from intensity thresholding to atlas-based segmentation.
-The first approach is performed simply thresholding the \gls{fa} map.
-Although this methodology was popular among the pioneer tractography studies,
-  they were generally limited to certain regions or significant fiber tracts, and thus,
-  it cannot be applied in whole-brain tractography.
-Early approaches to \gls{dwi} segmentation
-  include level set formulations using scalar maps of direction invariants derived
-  from the tensor model \citep{zhukov_level_2003}, directly on the diffusion raw data
-  \citep{rousson_level_2004}, or finding alternative diffusion representations
-  \citep{jonasson_representing_2007}.
-Even though this latter case was restricted to the extraction of
-  the corpus callosum from a real dataset, the density of the components of the diffusion tensor
-  are approximated by multivariate Gaussians for first.
-Iterative clustering performed on the B0 volumes of \gls{dwi} data was proposed by
-  \citep{hadjiprocopis_unbiased_2005}.
-Later studies investigated the application of probabilistic frameworks combining mixtures of
-  gaussians, \gls{mrf} and labeling fusion techniques \citep{liu_brain_2007} using as features
-  widely-used \gls{dwi}-derived scalar maps as \gls{fa} or \gls{md}.
-A similar framework combining co-registered structural information (T1 weighted) with
-  \emph{independent orthogonal invariants} derived from the \gls{dwi} tensor model was proposed
-  by \citep{awate_multivariate_2008}.
-Some proposals suggest the use of the raw diffusion data (directionally dependent) to avoid fitting
-  a certain model \citep{lu_segmentation_2008}.
-In \citep{han_experimental_2009}, graph-cuts voxel-based techniques are proposed using the most
-  common diffusion tensor derived features.
-Further developments of the probabilistic approach have been proposed adding more scalar maps
-  as features and a more detailed treatment of \gls{pve} \citep{kumazawa_partial_2010}.
-
-%A number of methods have been proposed using features not directly derived from \gls{dwi} data.
-%Segmentation obtained by co-registering structural T1-weighted images will be covered in
-%\autoref{sec:registration}.
-
-None of the presented methods have claimed for definite results, mainly due to the lack of
-  a \emph{gold-standard} evaluation methodology.
-Most of them are tested only on certain
-  regions, or do not provide sub-pixel resolution results
-  \citep{hadjiprocopis_unbiased_2005,liu_brain_2007,awate_multivariate_2008,lu_segmentation_2008,
-  han_experimental_2009}.
-Generally, results obtained with high resolution atlas co-registration
-  (\autoref{sec:registration}) are more compelling, limiting the activity on \gls{dwi}
-  segmentation.
-However, recently \citep{kumazawa_improvement_2013} proposed to include this
-  task within reconstruction algorithms.
-Finally, some \emph{golden}-standard evaluation frameworks for task-based validation have been
-  proposed for the segmentation results (e.g. lesion detection in visceral organs
-  \citep{jha_taskbased_2012}).
-\todo[inline]{reference to a summary table}
-
-\subsection{Structural information co-registration}
-\label{sec:registration}
-
-The last addressed task in the image processing stage of the studied
-connectivity analysis pipelines is the structural information co-registration.
-The need for this registration step is mainly raised when defining the
-nodes of the connectivity matrix.
-This definition step is named
-\textit{cortical parcellation} as it imposes the regions that will be
-considered to cluster the fiber bundles reaching the \gls{wm}-\gls{gm}
-surface.
-The common procedure to accomplish this clustering is using
-a predefined parcellation in a high-resolution structural T1-weighted
-atlas (e.g.
-\citep{greve_accurate_2009}).
-The second goal that has
-justified the applicability of structural
-images co-registration is susceptibility distortion correction
-as presented in \autoref{sec:distortion}.
-In any case, susceptibility
-distortion hinders a rigid-registration solution to this problem.
-Additionally, when applying non-linear intensity-based
-registration algorithms, other difficulties have to be addressed as
-the significant \gls{pve} in the \gls{wm}-\gls{gm} layer, or the
-inherent inadequacy of the B0 contrast (the only \textit{directionally
-independent} existing in the raw \gls{dwi} data) due to the almost null
-difference in intensity between \gls{wm} and \gls{gm} pixels.
-Early
-registration methods appeared targeting the distortion correction.
-They quickly standardized the choice of T2 as anatomically correct source
-to be registered against the B0 in \gls{dwi}.
-B0 images have a very similar
-contrast to T2 weighted due to the parallelism of acquisition sequences.
-However, parcellation is defined in T1 weighted space, adding an additional
-registration step is generally overlooked by existing literature.
-Even though this
-processing stage has a very low complexity compared to the remaining tasks of
-the pipeline process, it is one additional source of inconsistency
-and unreliability.
-
-One of the first proposals is \citep{kybic_unwarping_2000}, where the deformation
-field is modeled with B-Splines, the cost function is least squares and optimized
-in a multi-resolution gradient descent strategy.
-Their method is evaluated in
-both synthetic and real 2D images.
-Similarly,
-\citep{studholme_accurate_2000} proposed a spline-based deformation but including
-a weighting factor proportional to the Jacobian of the transform to correct the
-intensity of the undistorted data.
-They also use the $\log$ of the signal to
-enhance the low-signal regions, and optimize with a gradient descent algorithm
-the mutual information of the mapping.
-This framework set the basis for the following
-works.
-For instance, \citep{wu_comparison_2008} also proposed a B-Spline registration
-providing quantitative comparisons with fieldmap correction methods.
-Recent approaches
-take into account the signal loss due to dephasing \citep{li_accounting_2007}, or
-introduce more complex variational frameworks \citep{tao_variational_2009}.
-
-The significant benefits of exploiting the anatomical \gls{mri} when
-segmenting the \gls{dwi} data have been demonstrated \cite{zoellei_improved_2010},
-justifying the use of the shape prior information.
-To our knowledge, there is
-no study simultaneously taking advantage from segmentation or distortion correction
-tasks to be applied to the co-registration problem.
-
-\subsection{Proposed model}
-\label{sec:contributions}
-
-All the available \gls{dwi}-derived models and scalar maps
-are very sensitive to distortions, \gls{pve}, noise and axonal dispersion
-\citep{jbabdi_tractography_2011}, what yields numerous challenges in the
-processing tasks (artifact correction, segmentation, registration and
-parcellation).
-Finally, as a result a vast number of challenges and pitfalls
-emerge regarding accuracy (correctness) and precision (reproducibility) of
-tractography-based analyses \citep{jones_white_2013} and connectome
-analyses.
-
-Moreover, the problems of precise segmentation in \gls{dwi}-space and the
-spatial mapping between these contours and the corresponding surfaces in
-anatomical images bear significant redundancy.
-Once the spatial relationship
-between \gls{t1} and \gls{dwi} space is established, the contours which are
-readily available in \gls{t1} space can simply be projected on to the
-\gls{dwi} data.
-Conversely, if a precise delineation in \gls{dwi} space
-was achieved, the spatial mapping with \gls{t1}-space could be derived
-from one-to-one correspondences on the contours.
-However, neither segmentation
-nor registration can be performed flawlessly, if considered independently.
-
-% Paragraph taken from first subsection to here
-Here, we shall address three tasks included within  the image processing stage in
-a unified approach: brain tissue segmentation in diffusion space
-(\autoref{sec:dwi_segmentation}),
-correction of geometrical distortions (\autoref{sec:distortion}), and structural
-image registration to diffusion coordinate space (\autoref{sec:registration}).
-These tasks are generally
-solved independently, or combined in pairs.
-However, there exist fundamental
-coupling relationships that can be exploited to obtain a simultaneous solution to
-the three problems.
-This joint approach satisfactory impacts the downstream
-outcomes of the whole pipeline with the increase of the internal consistency.
+\IEEEPARstart{F}{usion} of information from different images is a difficult
+  task in image processing, especially when the scale of the imaged structures
+  is in the range of the image resolution or below.
+In the case of the human brain, the cortex has a typical thickness of up to 5 mm
+  \citep{fischl_measuring_2000}.
+Moreover, as the cortex is densely folded, points located afar in the topological
+  manifold of the brain's surface can be very close in the corresponding 3D space.
+Consequently, accuracy and precision are very difficult to achieve, but necessary properties
+  of registration algorithms.
 

### 2014-10-20 Publications/2013-TMI/02-abstract.tex
update

@@
 We propose a surface-to-volume variational solution to the registration problem,
-  that exploits the detailed anatomy extracted manually or from structural images.
-We use an ``active contours without edges''-like model to search for a
-  deformation field that optimally maps the surface prior onto the data
-  in the target space, implicitly segmenting with sub-pixel resolution the
-  volume.
+  that exploits the detailed anatomy extracted manually or automatically
+  from structural images.
+We use an ``active contours without edges'' model to search for a
+  deformation field of B-Spline functions that optimally maps the
+  surface prior onto the data in the target space.
+The strategy implicitly segments with sub-pixel resolution the image.
 We demonstrate the usefulness of the approach on a digital phantom, and on

### 2014-10-20 Publications/2013-TMI/03-introduction.tex
update

@@
 
+-> Volume-based registration is a traditional technique
+
+-> We have anatomical surfaces at high resolution that we want to impose to a
+lower resolution (typically different sequence), a follow-up image to track the surface
+or similar.
+
+-> Similar problem: atlas-based registration.
+
+Atlas-based segmentation Nonlinear registration and  are fundamental techniques applied in
+  image processing.
 In this paper we propose a novel registration framework to simultaneously

### 2014-10-21 Publications/2013-TMI/03-introduction.tex
updated methods

@@
 
-\IEEEPARstart{F}{usion} of information from different images is a difficult
-  task in image processing, especially when the scale of the imaged structures
-  is in the range of the image resolution or below.
+\IEEEPARstart{N}{onlinear} registration of images is a difficult task required
+  in an endless number of image analysis applications, producing a vast amount of
+  methods \citep{sotiras_deformable_2013} over the past decades.
+
+   especially when the scale of the imaged structures is in the range of the
+  image resolution or below.
 In the case of the human brain, the cortex has a typical thickness of up to 5 mm
@@
   image processing.
+
+\subsection{Related work}
+\label{sec:methods_background}
+
+We suggest clustering the
+current methodologies of template-based segmentation methods into three groups.
+The first group typically adds a shape prior term to the energy functional of
+\citep{chan_level_2005,chen_using_2002,cremers_kernel_2006,gastaud_combining_2004,wyatt_map_2003,paragios_level_2003,yezzi_variational_2003,gorthi_segmentation_2009,pohl_unifying_2005,pohl_bayesian_2006,wang_joint_2006,guyader_combined_2011,besson_dream2s_2003,bertero_illposed_1988,cote_tractometer_2013}
+an evolving active contour \citep{bresson_variational_2006, chan_level_2005,
+chen_using_2002,cremers_kernel_2006,gastaud_combining_2004}.
+These methods generally have an explicit description of the expected relative boundary
+locations of the object to be delineated, and some even model the statistical deviations
+from this average shape. By including a coordinates mapping, it is possible to perform
+active contours based registration between timesteps in a time-series or between different
+images \citep{bertalmio_morphing_2000,wyatt_map_2003,paragios_level_2003,vemuri_joint_2003,
+yezzi_variational_2003}.
+A summary of these second set of techniques is performed in \citep{droske_mumfordshah_2009},
+proposing two different approaches to applying the Mumford-Shah \citep{mumford_optimal_1989}
+functional in joint registration and segmentation. Finally, a derivation of the
+latter groups is composed by atlas-based segmentation methods
+\citep{gorthi_segmentation_2009,gorthi_active_2011,pohl_unifying_2005,
+pohl_bayesian_2006,wang_joint_2006}, where the prior
+imposes consistent voxel-based classification of contiguous regions.
+A comprehensive summary of the \gls{acwe}-derived methodologies with special attention
+to joint registration-segmentation methods is found in \citep{gorthi_active_2011}.
+\Gls{acwe} and level-set based formulations have been widely applied in image
+processing, as reviewed in \citep{suri_shape_2002}.
+
+The first joint ``morphing''-segmentation approach (it was not a proper
+registration method) was proposed in \citep{bertalmio_morphing_2000}. The first
+registration framework is presented by \citep{yezzi_variational_2001} where the
+energy functional is defined simultaneously in the moving and target images with
+an affine transformation supporting the coordinates mapping (``2 \glspl*{pde} approach'').
+\citep{vemuri_joint_2003} proposed the first atlas-based registration based on the
+level set framework and using only one \gls*{pde}. \citep{unal_coupled_2005} extended the
+idea of \citep{bertalmio_morphing_2000,yezzi_variational_2001} for non-linear
+registration with a dense deformation field as mapping function. \citep{droske_mumfordshah_2009}
+combining ideas from both branches and proposed the propagation of the
+deformation field from the contours to the whole image definition. Finally,
+\citep{gorthi_active_2011} present a comprehensive generalization of the methodologies.
+
+Even though the historical evolution
+of joint segmentation and registration procedures is strongly linked
+to the level sets approximation, our work is based in derivations of
+the \gls{acwe} framework \citep{chan_active_2001}, and related to
+\citep{guyader_combined_2011}.
+{\color{red} {The main diverging points with respect to
+\citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
+$\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
+gradients \citep{besson_dream2s_2003,herbulot_segmentation_2006};
+2) regularization is also based on linear diffusion smoothing \citep{thirion_image_1998},
+but we replace the Gaussian filtering by other constraints
+studied in \citep{nagel_investigation_1986} to better the problem;
+3) optimization is applied in the spectral
+domain, observing anisotropic and inhomogeneous mappings along each direction.
+With respect \citep{guyader_combined_2011}, the main differences are
+the distance function, and the spectral solution to the optimization updates,
+as we shall cover in \autoref{sec:numerical_implementation}.}}
+
+\todo[inline]{The previous paragraphs in red color are marked by Dominique
+as ``to rework'', and I totally agree}
+
+\subsection{Our framework}
+\label{sec:our_framework}
+
+In this paper we formulate the joint registration-segmentation
+problem as follows. We look for a spatial mapping (with, typically,
+$n=3$ for the 3-dimensional case):
+
+\begin{align}
+U\colon T \subset \mathbb{R}^n &\to D \subset \mathbb{R}^n \notag\\
+\vec{x} &\mapsto \vec{x}' =\vec{x}+u(\vec{x}),
+\label{eq:transform}
+\end{align}
+such that the known contours in anatomical space $T$ optimally segment
+the diffusion space $D$.
+
+Whereas related \glspl*{adf} introduced in \autoref{sec:methods_background}
+make use of explicit level-set formulations to solve \eqref{eq:gradient_descent},
+we alternatively use \emph{shape-gradients}
+\cite{besson_dream2s_2003,herbulot_segmentation_2006}.
+Let $r(\vec{x})$ be an ``arbitrary'' function over the image domain,
+and $\Omega$ a certain image region with $\Gamma_{l,m}$ its corresponding
+outer boundary as defined in \autoref{sec:methods_map}.
+We now derive the domain integral w.r.t. $t$:
+\begin{equation*}
+\frac{\partial}{\partial t} \int_\Omega r(\vec{x}') d\vec{x} = \int_\Omega \frac{\partial}{\partial t}r(\vec{x}') d\vec{x} - \int_{\Gamma_{l,m}} r(\vec{x}') \left\langle \frac{\partial \Gamma_{l,m} }{\partial t}, N_{\Gamma_{l,m}}\right\rangle d\vec{x}
+\end{equation*}
+where $\left\langle\frac{\partial\Gamma_{l,m}}{\partial t}, N_{\Gamma_{l,m}}\right\rangle$ is
+the projection of the boundary movement on the unit inward normal $N_{\Gamma_{l,m}}$. Assuming
+that the region descriptors $\Theta_l$ vary slowly enough, we can consider
+that $\frac{\partial}{\partial t} r(\vec{x}') = 0$ and thus:
+\begin{equation}
+\frac{\partial}{\partial t} \int_\Omega r(\vec{x}') d\vec{x} =
+- \int_{\Gamma_{l,m}} r(\vec{x}') \left\langle \frac{\partial \Gamma_{l,m} }{\partial t}, N_{\Gamma_{l,m}}\right\rangle d\vec{x}
+\label{eq:shape_gradients}
+\end{equation}
+
+Let us denote by $\vec{x}$ the voxel and $F(\vec{x}) = [ f_1, f_2, \ldots, f_N]^T$
+  its associated feature vector in the following.
+
+In our application, these surfaces are precise tissue interfaces of interest extracted
+  from a high-resolution, anatomically correct reference volume using
+  the well-established
+
 In this paper we propose a novel registration framework to simultaneously
@@
 in \autoref{sec:conclusion}.
+
+The properties of the reconstructed tensors and derived scalar maps have
+been studied by \cite{ennis_orthogonal_2006}. Based on their
+findings, \gls{fa}~\eqref{eq:fa} and \gls{md}~\eqref{eq:md} are
+considered complementary features, and therefore we selected them for the
+energy model \eqref{eq:complete_energy} in driving the
+registration-segmentation process.
+Whereas \gls{fa} informs mainly about the \emph{shape} of diffusion,
+the \gls{md} is more related to the \emph{magnitude} of the process:
+
+\begin{align}
+\mathrm{FA} &= \sqrt{ \frac{1}{2}}\,\frac{\sqrt{ (\lambda_1 - \lambda_2)^2 + (\lambda_2 - \lambda_3)^2 + (\lambda_3 - \lambda_1)^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}} \label{eq:fa} \\
+\mathrm{MD} &= ( \lambda_1 + \lambda_2 + \lambda_3 ) / 3 \label{eq:md}
+\end{align}
+where $\lambda_i$ are the eigenvalues of the diffusion tensor
+associated with the diffusion signal $S(\vec{q})$. There exist
+two main reasons to justify their choice.
+First, they are well-understood and standardized in clinical routine.
+Second, together they contain most of the information that is
+usually extracted from the \gls{dwi}-derived scalar maps
+\cite{ennis_orthogonal_2006}.

### 2014-10-21 Publications/2013-TMI/04-methods.tex
updated methods

--- 2013-10-15 Publications/2013-TMI/04-methods.tex
@@
-1\section{Methods}
+% -*- root: 00-main.tex -*-
+\section{Methods}
 \label{sec:methods}
 %
-\subsection{Related work}
-\label{sec:methods_background}
 
-We suggest clustering the 
-current methodologies of template-based segmentation methods into three groups. 
-The first group typically adds a shape prior term to the energy functional of 
-an evolving active contour \citep{bresson_variational_2006,
-chan_level_2005,chen_using_2002,cremers_kernel_2006,gastaud_combining_2004}.
-These methods generally have an explicit description of the expected relative boundary 
-locations of the object to be delineated, and some even model the statistical deviations
-from this average shape. By including a coordinates mapping, it is possible to perform
-active contours based registration between timesteps in a time-series or between different
-images \citep{bertalmio_morphing_2000,wyatt_map_2003,paragios_level_2003,vemuri_joint_2003,
-yezzi_variational_2003}.
-A summary of these second set of techniques is performed in \citep{droske_mumfordshah_2009},
-proposing two different approaches to applying the Mumford-Shah \citep{mumford_optimal_1989}
-functional in joint registration and segmentation. Finally, a derivation of the 
-latter groups is composed by atlas-based segmentation methods 
-\citep{gorthi_segmentation_2009,gorthi_active_2011,pohl_unifying_2005,
-pohl_bayesian_2006,wang_joint_2006}, where the prior 
-imposes consistent voxel-based classification of contiguous regions.
-A comprehensive summary of the \gls{acwe}-derived methodologies with special attention
-to joint registration-segmentation methods is found in \citep{gorthi_active_2011}.
-\Gls{acwe} and level-set based formulations have been widely applied in image
-processing, as reviewed in \citep{suri_shape_2002}.
+\subsection{Cost-function derivation}
+\label{sec:methods_map}
 
-The first joint ``morphing''-segmentation approach (it was not a proper
-registration method) was proposed in \citep{bertalmio_morphing_2000}. The first
-registration framework is presented by \citep{yezzi_variational_2001} where the
-energy functional is defined simultaneously in the moving and target images with
-an affine transformation supporting the coordinates mapping (``2 \glspl*{pde} approach'').
-\citep{vemuri_joint_2003} proposed the first atlas-based registration based on the 
-level set framework and using only one \gls*{pde}. \citep{unal_coupled_2005} extended the 
-idea of \citep{bertalmio_morphing_2000,yezzi_variational_2001} for non-linear 
-registration with a dense deformation field as mapping function. \citep{droske_mumfordshah_2009}
-combining ideas from both branches and proposed the propagation of the 
-deformation field from the contours to the whole image definition. Finally,
-\citep{gorthi_active_2011} present a comprehensive generalization of the methodologies.
-
-Even though the historical evolution
-of joint segmentation and registration procedures is strongly linked
-to the level sets approximation, our work is based in derivations of
-the \gls{acwe} framework \citep{chan_active_2001}, and related to
-\citep{le_guyader_combined_2011}.
-{\color{red} {The main diverging points with respect to
-\citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
-$\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
-gradients \citep{jehan-besson_dream2s:_2003,herbulot_segmentation_2006};
-2) regularization is also based on linear diffusion smoothing \citep{thirion_image_1998},
-but we replace the Gaussian filtering by other constraints
-studied in \citep{nagel_investigation_1986} to better the problem;
-3) optimization is applied in the spectral
-domain, observing anisotropic and inhomogeneous mappings along each direction.
-With respect \citep{le_guyader_combined_2011}, the main differences are
-the distance function, and the spectral solution to the optimization updates,
-as we shall cover in \autoref{sec:numerical_implementation}.}}
-
-\todo[inline]{The previous paragraphs in red color are marked by Dominique
-as ``to rework'', and I totally agree}
-
-\subsection{Proposed model}
-\label{sec:methods_map}
-In this paper we formulate the joint registration-segmentation
-problem as follows. We look for a spatial mapping (with, typically,
-$n=3$ for the 3-dimensional case):
-\begin{align}
-U\colon T \subset \mathbb{R}^n &\to D \subset \mathbb{R}^n \notag\\
-\vec{x} &\mapsto \vec{x}' =\vec{x}+u(\vec{x}),
-\label{eq:deformation_field}
-\end{align}
-such that the known contours in anatomical space $T$ optimally segment
-the diffusion space $D$.
-
-In a Bayesian framework, different such mappings $U$ between ``T1-''
-and ``\gls*{dwi}-'' spaces are evaluated based on their posterior
-probability given the observed data $F$. Using the Bayes' rule, this
-posterior \emph{likelihood} can be computed as:
+In a Bayesian framework, the mappings $U$ in \autoref{eq:transform} are
+  evaluated based on their posterior probability given the observed data
+  $F$.
+Using the Bayes' rule, the posterior likelihood is computed as:
   \begin{equation}
@@
   where $P(F \mid U,\Gamma_{l,m})$ is the data-likelihood, and
-$\Gamma_{l,m}$ are a set of contours in T1-space that correspond to
-two interfacing regions $\Omega$, such that 
-$\Gamma_{l,m} = \partial \Omega_l \cap \partial \Omega_m$.
-
-We are interested in a mapping $\hat{U}$ which maximizes this
-posterior probability (\gls*{map} estimate, \citep{bishop_pattern_2006}):
+  $\Gamma_{l,m}$ are a set of surfaces corresponding to the interfaces
+  between piecewise smooth regions $\Omega_i$, such that
+  $\Gamma_{l,m} = \partial \Omega_l \cap \partial \Omega_m$ is the
+  contour between tissues $l$ and $m$.
+Then, the \gls*{map} criterium \citep{leemput_automated_1999} is used
+  to find the mapping estimate $\hat{U}$ that aligns $\Gamma$ into $F$:
   \begin{align}
@@
   global data likelihood into a product of pixel-wise conditional probabilities:
-
   \begin{equation}
-P(F \mid U,\Gamma_{l,m}) = 
-\underset{\vec{x}\in \Omega}{\prod} P\left( F(\vec{x}') \mid U(\Gamma_{l,m}) \right).
+  P(F \mid U,\Omega) = \underset{l}{\prod} \underset{\vec{x}\in \Omega_l}{\prod}
+    P\left( \vec{f}' \mid U \right),
   \label{eq:bayes_aposteriori}
   \end{equation}
-
-Further, within each region $\Omega_l$ defined by the contours $\Gamma$
-trough duality, we assume the features to be i.i.d.:
+  where $\vec{f}' = F(\vec{x}')$ is the feature vector at the displaced
+  position $\vec{x}'$ (\autoref{eq:transform}) of the multispectral target
+  volume.
+For convenience, and because it has been shown to be appropriate approximations
+  \citep{cuadra_comparison_2005}, we introduce two assumptions for each
+  region $\Omega_l$:
+  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
+  distributions:
   \begin{equation}
-P\left( F(\vec{x}') \mid U,\Gamma_{l,m} \right) = p_l( F(\vec{x}')) \quad
-\forall \vec{x} \in \Omega_l.
+  P\left( \vec{f}' \mid U,\Omega_l \right) = \mathcal{N} \left( \vec{f}' \mid \Theta_l \right),
   \label{eq:likelihood}
   \end{equation}
-
-For convenience, and because this has been shown to be an appropriate
-approximation \citep{bishop_pattern_2006}, each such region is modeled
-by a multivariate normal distribution with region descriptors
-$\Theta_l = \lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$:
-
-\begin{equation}
-p_l( F(\vec{x}') ) = \mathcal{N} \left( F(\vec{x}') \mid \Theta_l \right).
-\label{eq:multivariate_normal}
-\end{equation}
-
-\todo{I modified this slightly}
-Thus, defining the feature observed as $\vec{f}=F(\vec{x})$ and 
-the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
-of the descriptors of region $l$ as
-\begin{equation}
-\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l),
-\end{equation}
-we write:
+ 	where $\Theta_l = \lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$ are the
+ 	descriptors of region $\Omega_l$.
+We then insert \autoref{eq:likelihood} into \autoref{eq:bayes_rule} to obtain the full
+  formulation of the \emph{data term} of our registration model:
  	\begin{align}
@@
   \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
-&= \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
-\mdist{f'}{l} \right)}}.
+  &= \underset{l}{\prod} \underset{\vec{x} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
+  \mdist{f'}{l} \right)}},
   \label{eq:pdf}
   \end{align}
+  using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
+  to the descriptors of region $l$ as
+  $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
 
@@
   \begin{equation*}
-P(U) = \underset{\vec{x}}{\prod} p(u(\vec{x}))
+  P(U) = \underset{\vec{x}}{\prod} p(u(\vec{x})),
   \end{equation*}
@@
   p_0(u(\vec{x})) &= \mathcal{N}( u(\vec{x}) \mid 0, A^{-1}) \notag\\
-p_1(u(\vec{x})) &= \mathcal{N}( Du(\vec{x}) \mid 0, B^{-1}),
+  p_1(u(\vec{x})) &= \mathcal{N}(  \nabla \cdot u(\vec{x}) \mid 0, B^{-1}),
   \end{align*}
-that consider that the distortion and its gradient have zero
-mean and variance governed by $A$ and $B$. Since the anisotropy
-is technically aligned with the imaging axes, these can
-be simplified:
+  imposing that the distortion and its gradient have zero
+  mean and variance governed by $A$ and $B$.
+Since the anisotropy is generally aligned with the imaging axes, these will be simplified
+  in the following for the sake of clarity:
   \begin{align}
-p_0(u(\vec{x})) &= \underset{\vec{x}}{\sum} \mathcal{N}( u(\vec{x}) \mid 0, (\boldsymbol{\alpha}^{\circ\frac12}\,\vec{I}_n)^{-1}) \notag\\
-p_1(u(\vec{x})) &= \underset{\vec{x}}{\sum} \mathcal{N}( \nabla \cdot u(\vec{x}) \mid 0, (\boldsymbol{\beta}^{\circ\frac12}\,\vec{I}_n)^{-1})
+    p_0(u(\vec{x})) &= \underset{\vec{x}}{\sum} \mathcal{N}( u(\vec{x}) \mid 0,
+      (\boldsymbol{\alpha}^{\circ\frac12}\,\vec{I}_n)^{-1}), \notag\\
+    p_1(u(\vec{x})) &= \underset{\vec{x}}{\sum} \mathcal{N}( \nabla \cdot u(\vec{x}) \mid 0,
+      (\boldsymbol{\beta}^{\circ\frac12}\,\vec{I}_n)^{-1}).
   \label{eq:priors}
@@
 
-Finally, we can turn the \gls{map} problem into 
-a variational one
-applying the following log-transform:
+Finally, the \gls{map} problem is adapted into a variational one applying the
+  following log-transform:
   \begin{align}
@@
   \underset{\vec{x} \in \Omega_l}{\prod}
-p_l( \vec{f}')\,p_0( u(\vec{x}))\,p_1( u(\vec{x})) \notag\\
+  \mathcal{N} \left( \vec{f}' \mid \Theta_l \right)\,p_0( u(\vec{x}))\,p_1( u(\vec{x})), \notag\\
   &= C + \underset{l}{\sum}
   \underset{\vec{x} \in \Omega_l}{\sum}
-\mdist{f'}{l} \notag\\
+  \mdist{f'}{l}, \notag\\
   &+ \underset{\vec{x} \in \Omega}{\sum} \boldsymbol{\alpha} \cdot u(\vec{x})^{\circ2}
@@
   \citep{nagel_investigation_1986}.
-
-{\color{red} {Using the region descriptors derived in \autoref{sec:methods_map}, we propose
-an \gls{adf}-like, \gls{acwe}-based, piece-wise constant, image segmentation
-model (where the unknown is the deformation field)
-\cite{chan_active_2001} with the energy functional obtained in 
-\eqref{eq:final_map_energy}. This inverse problem is ill-posed
-\cite{bertero_ill-posed_1988,hadamard_sur_1902}.
-In order to account for deformation field regularity and to render the 
-problem well-posed, we include limiting and regularization terms into 
-the energy functional \cite{morozov_linear_1975,tichonov_solution_1963}:
-%
-\begin{multline}
-E(u) = \sum\limits_l \int_{\Omega'_l} \mdist{f}{l} \,d\vec{x} \\
-+ \int \vec{u}^T \, A \, \vec{u} \, d\vec{x} + \int \tr\{(\nabla \vec{u}^T)^T B (\nabla \vec{u}^T)\} d\vec{x}
-\label{eq:complete_energy}
-\end{multline}
-%
-These regularity terms ensure that the segmenting contours in 
-\gls{dwi} space are still close to their native shape. The model
-easily allows to incorporate inhomogeneous and anisotropic 
-regularization \cite{nagel_investigation_1986} to better regularize
-the \gls{epi} distortion.}}
-
-\todo[inline]{this last paragraphs in red color were located a bit
-further, but I think we can keep some of the refs and still reproduce
-the continuous version of the energy functional}
+We provide a proof of duality in {\color{red} [REF figshare?]}.
 
@@
 \label{sec:numerical_implementation}
-%
-Let us denote $\{\vec{c}_i\}_{i=1 \ldots N_c}$ the nodes of one or several shape-prior
-surface(s). In our application, precise tissue interfaces of interest 
-extracted from a high-resolution, anatomically correct reference volume. 
-On the other hand, we have a number of \gls{dwi}-derived features at each
-voxel of the volume. Let us denote by $\vec{x}$ the voxel and $F(\vec{x}) = 
-[ f_1, f_2, \ldots, f_N]^T(\vec{x})$ its associated feature vector.
 
-\subsubsection{Deformation field}
-\label{sec:deformation_field}
-The transformation from ``T1-'' into ``\gls{dwi}-'' coordinate space is 
-achieved through a dense deformation field $u(\vec{x})$, such that:
+\subsubsection{Deformation model}
+\label{sec:deformation_model}
+Let us denote $\{\vec{c}_i\}_{i=1 \ldots N_c}$ the nodes of one or several prior
+  surface(s).
+In our application, these surfaces are extracted using \emph{FreeSurfer}
+  \citep{fischl_freesurfer_2012} in the case of real datasets.
+The transformation from the structural space into the coordinate space of the
+  target image is achieved through a dense deformation field $u(\vec{x})$, such that:
   \begin{equation}
@@
   \end{equation}
-where $U$ is defined in \eqref{eq:deformation_field}. Since the nodes of the anatomical 
-surfaces might lay off-grid, it is required to derive $u(\vec{x})$ from a discrete 
-set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$. Densification is achieved through 
-a set of associated basis functions $\psi_k$:
-%
+  where $U$ is defined in \eqref{eq:transform}.
+Since the nodes of the anatomical surfaces likely lay off-grid, it is required to
+  derive $u(\vec{x})$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
+Densification is achieved through a set of associated basis functions $\psi_k$:
   \begin{equation}
-u(\vec{x}) = \sum_k \psi_k(\vec{x}) \vec{u}_k
+  u(\vec{x}) = \sum_k \psi_k(\vec{x}) \vec{u}_k.
   \label{eq:intp_kernel}
@@
 Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
-$\psi$ by the actual kernel function, the transformation writes
-%
+  $\psi$ by the actual kernel function, the transformation writes:
   \begin{equation}
-\vec{c}_i' = \vec{c}_i + \sum_k \left[ \vec{u}_k \, \underset{n}{\prod} B_3( (\vec{c}_i - \vec{x}_k) \cdot \hat{\mathbf{e}}_d ) \right].
+    \vec{c}_i' = \vec{c}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
+      B_3( (\vec{c}_i - \vec{x}_k) \cdot \hat{\mathbf{e}}_d ) \right],
   \label{eq:transformation}
   \end{equation}
+  with $\hat{\mathbf{e}}_d$ being the unitary vector along axis $d$.
 
-Based on the current estimate of the distortion $\vec{u}$, we can compute 
-``expected samples'' within the shape prior projected into the \gls{dwi}.
-Thus, we now estimate region descriptors of the \gls{dwi} features 
-$\vec{f}'$ of the regions defined by the priors in \gls{dwi} space.
-
-\subsubsection{Gradient descent}
+\subsubsection{Optimization}
+\label{sec:gradient_descent}
 To find the minimum of the energy functional \eqref{eq:energy},
@@
   \begin{equation}
-\frac{\partial u(\vec{x},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}},
+  \frac{\partial u(\vec{x},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}_k},
   \label{eq:general_gradient_descent}
@@
   with $t$ being an artificial time parameter of the contour
-evolution, included to express the iterative character of the approach.
-Now, we can introduce \eqref{eq:energy} in
-\eqref{eq:general_gradient_descent}:
+  evolution, and $\vec{u}_k$ the parameters supporting the estimate
+  $\hat{U}$ of the transformation at the current time point.
+Now, we introduce \eqref{eq:energy} in \eqref{eq:general_gradient_descent}:
   \begin{align}
@@
   \end{align}
-
-Whereas related \glspl*{adf} introduced in \autoref{sec:methods_background}
-make use of explicit level-set formulations to solve \eqref{eq:gradient_descent},
-we alternatively use \emph{shape-gradients}
-\cite{jehan-besson_dream2s:_2003,herbulot_segmentation_2006}.
-Let $r(\vec{x})$ be an ``arbitrary'' function over the image domain,
-and $\Omega$ a certain image region with $\Gamma_{l,m}$ its corresponding
-outer boundary as defined in \autoref{sec:methods_map}.
-We now derive the domain integral w.r.t. $t$:
-\begin{equation*}
-\frac{\partial}{\partial t} \int_\Omega r(\vec{x}') d\vec{x} = \int_\Omega \frac{\partial}{\partial t}r(\vec{x}') d\vec{x} - \int_{\Gamma_{l,m}} r(\vec{x}') \left\langle \frac{\partial \Gamma_{l,m} }{\partial t}, N_{\Gamma_{l,m}}\right\rangle d\vec{x}
-\end{equation*}
-where $\left\langle\frac{\partial\Gamma_{l,m}}{\partial t}, N_{\Gamma_{l,m}}\right\rangle$ is 
-the projection of the boundary movement on the unit inward normal $N_{\Gamma_{l,m}}$. Assuming
-that the region descriptors $\Theta_l$ vary slowly enough, we can consider
-that $\frac{\partial}{\partial t} r(\vec{x}') = 0$ and thus:
-\begin{equation}
-\frac{\partial}{\partial t} \int_\Omega r(\vec{x}') d\vec{x} = 
-- \int_{\Gamma_{l,m}} r(\vec{x}') \left\langle \frac{\partial \Gamma_{l,m} }{\partial t}, N_{\Gamma_{l,m}}\right\rangle d\vec{x}
-\label{eq:shape_gradients}
-\end{equation}
 
@@
 
-\subsubsection{Semi-implicit Euler-forward optimization}
-It is necessary to discretize $t$ in order to obtain a numerical
-implementation of the equation \eqref{eq:gradient_final}:
-\begin{align}
-\frac{\vec{u}_k^{t+1}-\vec{u}_k^t}{\tau} =
-&- \underset{l,m}{\sum} \underset{i}{\sum}
-\left[ \mdist{f_i'}{l} - \mdist{f_i'}{m}\right]
-\psi_k(\vec{c}_i)\, \hat{\vec{n}}_i \notag\\
-&+2\, \boldsymbol{\alpha} \vec{u}_k^{t+1}
--2\, \boldsymbol{\beta} \Delta \vec{u}_k^{t+1}.
-\end{align}
-
-The associated Euler-Lagrange equation is found as:
-\begin{equation}
-(\tau^{-1} +2\, \boldsymbol{\alpha} - 2\, \boldsymbol{\beta} \Delta )\, \vec{u}_k^{t+1} = 
-\tau^{-1} \vec{u}_k^t - \frac{\partial}{\partial \vec{u}_k} E_{data}(\vec{u}_k^t),
-\end{equation}
-that is a linear system that we translate into Fourier domain,
-to obtain the next deformation field $\vec{u}_k^{t+1}$:
-\begin{multline}
-\vec{u}_k^{t+1} = \\
- \mathcal{F}^{-1} \left\lbrace
-\frac{\mathcal{F}\lbrace \tau^{-1}\vec{u}_{k}^{t} - \underset{l,m}{\sum} \underset{i}{\sum}
-\left[ \mdist{f_i'}{l} - \mdist{f_i'}{m}\right] \rbrace}
-     {\mathcal{F}\lbrace (\tau^{-1} +2\,\boldsymbol{\alpha})\mathcal{I} - 2\,\boldsymbol{\beta} \Delta ) \rbrace}
-     \right\rbrace
-\end{multline}
-
-Here, we rewrite the Laplacian as a linear combination of the identity and shift operators:
-\begin{equation}
-\Delta = \sum\limits_d \mathcal{S}_d^- + \mathcal{S}_d^+ - 2 \mathcal{I}
-\end{equation}
-where $\mathcal{S}_{d}^{\pm}$ stands for the forward ($+$) and backward ($-$) shift 
-operator along coordinates axis $d$, of which the Fourier transform is found easily as
-\begin{equation}
-\mathcal{F}\{\mathcal{S}_{d}^{\pm}\} = e^{\pm \vec{i}\omega_{d}},
-\end{equation}
-where $\omega_{d}$ is the normalized pulsation along $d$-direction. Accordingly, the 
-Fourier transform of the discrete Laplacian is found as
-\begin{equation}
-\mathcal{F}\{\Delta\} = \sum\limits_d e^{-\vec{i}\omega_d } + e^{\vec{i}\omega_d } - 2 = \sum\limits_d \left( 2\cos(\omega_d) - 2 \right)
-\end{equation}
-
-The remaining transforms are trivial or can be computed using the \gls{fft} 
-as in \citep{estellers_efficient_2011}.
-
 \subsubsection{Region descriptor reestimation}
@@
 \subsubsection{Convergence}
-\todo[inline]{discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition, etc.}
+{\color{red} discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition / Wolfe conditions, multi-resolutions on the bspline grid, image samples and surface sampling,  etc.}
 
@@
 diffeomorphic transform by
+
+\subsection{Image Data and preprocessing}
+\label{sec:datasets}
+
+\subsubsection{Simulated digital phantom} %
+\label{sec:digital_phantoms}
+%
+We first present the proof of the concept on a simplistic digital phantom,
+  simulating a spherical crust of thin tissue with a folding resembling
+  a gyrus of the brain cortex folding (see \autoref{fig1}).
+We simulated 2.0mm isotropic \gls*{t1} and \gls*{t2} images at
+  several rician noise levels, with parameters TE/TR=10/1500ms and
+  TE/TR=90/5000ms respectively using \cite{caruyer_phantomas_2014}.
+The reference surfaces were extracted using the marching-cubes algorithm
+  on high-resolution (1.0mm isotropic) binary versions of the corresponding
+  maps of tissue volume fractions used during the simulation step.
+
+\subsubsection{Real \gls{mri} datasets} %
+\label{sec:human_connectome}
+%
+For the evaluation of the algorithm with real brains with \gls*{dmri} data,
+  we randomly selected 10 datasets from the minimally preprocessed
+  \citep{glasser_minimal_2013} cohort of the \gls*{hcp}.
+These datasets comprehend the preprocessed (corrected for artifacts,
+  brain-extracted and spatially normalized) \gls*{t1}, \gls*{t2} and
+  multi-shell \gls*{dmri} images, along with the most prominent results of
+  the standard processing with \emph{FreeSurfer}: \emph{aseg} and
+  \emph{aparc} segmentations, and the surfaces corresponding to the cortical
+  \gls*{gm} exterior (\emph{pial}) and the \gls*{wm}/\gls*{gm} interface
+  (\emph{white}).
+This cohort is especially convenient for their careful preprocessing,
+  making available high-standard \gls*{mri} data accurately corrected
+  for distortions and artifacts.
+For the exact details of the scanning parameters we refer the reader to
+  \citep{essen_human_2012}.
+Selecting the appropriate labels in the \emph{aseg} segmentation, we applied
+  the marching-cubes algorithm again to extract the surfaces containing the
+  \gls*{csf} of the ventricular system.
+
+As we propose the \gls*{fa} and the \gls*{md} as target features, we
+  implemented a simplified pipeline for processing \gls*{dmri} using
+  \emph{MRtrix} \citep{tournier_mrtrix_2012}.
+\todo[inline]{insert here the summary of parameters}
+
+\subsection{Experiments and evaluation}
+\label{sec:experiments_evaluation}
+%
+\subsubsection{Synthetic distortion}
+\label{sec:distortion}
+
+\subsubsection{Evaluation indices} %
+\label{sec:evaluation}
+We report the following indices:
+  \begin{itemize}
+  \item \gls{se}, a distance between the one-to-one corresponding
+    vertices, weighted by their respective Voronoi area.
+    \begin{equation}
+    MSE = \sum\limits_k \sum\limits_j^M w_j\,\| \mathbf{x}_j - \hat{\mathbf{x}}_j \|
+    \end{equation}
+    where $\mathbf{x}_j$ are the locations of the $M$ vertices of the $k$ priors,
+    $\hat{\mathbf{x}}_j$ are the corresponding locations recovered, and $w_j$ the
+    weighting factor as the relative surface of the Voronoi area.
+  \item \gls{wi}, L2-distance between the theoretical and the recovered
+    deformation field.
+    \begin{equation}
+    WI = \frac{1}{N} \sum\limits_i^N \| \mathbf{d}_i - \hat{\mathbf{d}}_i \|
+    \end{equation}
+    where $\mathbf{d}_i$ is the theoretical displacement vector at position $i$
+    and $\hat{\mathbf{d}}_i$ is the recovered one at the same index position.
+  \item Parcellation agreement, the Dice overlap index of the \glspl*{roi}
+    contained by the parcellation.
+  \end{itemize}

### 2014-10-21 Publications/2013-TMI/05-experiments.tex
updated methods

@@
 \label{sec:experiments}
-%
-
-\subsection{Image Data}
-\label{sec:datasets}
-
-\subsubsection{Simulated digital phantom} %
-%
-The lack of a widely accepted \emph{gold}-standard in the application
-field has been addressed by several authors \citep{cote_tractometer_2013}
-In this work, we use one of the most complete, and publicly available,
-digital phantoms (\url{http://hardi.epfl.ch/static/events/2013_ISBI/testing_data.html}).
-The phantom is a spherical volume containing a set of fiber bundles, that connect
-one area of a ``cortex'' to another. The model accounts for \gls{pve} using
-a similar approach to \citep{close_software_2009} and for \gls{csf} contamination
-as well.
-
-\subsubsection{Real \gls{mri} datasets} %
-%
-We used image data from XX healthy volunteers with no history of neurological
-conditions (ages XX$\pm$XX, X female) to illustrate the applicability of our
-approach. All the subjects were scanned in a 3T MR Scanner (Siemens Magnetom
-TrioTim, Siemens, Erlangen, Germany) with a standard 12-channel head coil.
-\todo[inline]{check coil \# channels}
-Subjects were scanned twice with the same protocol,
-described hereafter. After being scanned the first time, each subject exited the
-scan room for a short break and then reentered for an identical scan session.
-To note, there was a full repositioning of the volunteer, coils, blankets and pads
-before each scan and re-scan session. The scan session protocol was as follows:
-\begin{enumerate}
-\item Triplanar survey (Localizer).
-\item Field Mapping: field mapping using GRE sequence was performed before \gls{dwi}
-acquisition for susceptibility correction purposes.
-\item DTI: \gls{dwi} were acquired with axial in-plane
-isotropic resolution $2mm$, slice thickness $2mm$, $XXX \times XXX \times XXX$
-image matrix, TR= XXXX ms, TE = XX ms, NEX = X, BW = XXXX Hz/pixel, GRAPPA
-acceleration factor 2. The series included images acquired with diffusion
-weighting along 30 non-collinear directions ( $b$=700sm$^{-2}$ ) sampled
-twice for averaging, and 5 interleaved images acquired without diffusion
-weighting ( $b=0$ ).
-\item Field Mapping (same as before DTI)
-\item Structural T1: An MPRAGE T1-weighted acquisition, sagittal
-GRE sequence, in-plane isotropic resolution 1.0 mm, slice thickness 1.2mm
-$XXX \times XXX \times XXX$ image matrix, TR=2300 ms, TE=2.98 ms, FA= 9, NEX=X, BW= 240 Hz/pixel.
-\item Structural T2: A T2-weighted acquisition, oblique axial
-TSE sequence, in-plane isotropic resolution 1.0 mm, slice thickness 1.2mm
-$XXX \times XXX \times XXX$ image matrix, TR=3200 ms, TE= 408 ms, NEX=X, BW= 751 Hz/pixel.
-\end{enumerate}
-
-\subsection{Image preprocessing and shape-prior generation}
-\label{sec:preprocessing}
-%
-Regardless the dataset type (simulated or real), all \gls{dwi} datasets were processed
-using the standard \gls{dti} reconstruction methods provided by FSL \footnote{DTIFIT,
-included in the FMRIB's Software Library (FSL),
-\url{http://fsl.fmrib.ox.ac.uk/fsl/fsl-4.1.9/fdt/fdt_dtifit.html}} to fit
-the tensor model and produce scalar maps of the required features.
-The properties of the reconstructed tensors and derived scalar maps have
-been studied by \cite{ennis_orthogonal_2006}. Based on their
-findings, \gls{fa}~\eqref{eq:fa} and \gls{md}~\eqref{eq:md} are
-considered complementary features, and therefore we selected them for the
-energy model \eqref{eq:complete_energy} in driving the
-registration-segmentation process.
-Whereas \gls{fa} informs mainly about the \emph{shape} of diffusion,
-the \gls{md} is more related to the \emph{magnitude} of the process:
-
-\begin{align}
-\mathrm{FA} &= \sqrt{ \frac{1}{2}}\,\frac{\sqrt{ (\lambda_1 - \lambda_2)^2 + (\lambda_2 - \lambda_3)^2 + (\lambda_3 - \lambda_1)^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}} \label{eq:fa} \\
-\mathrm{MD} &= ( \lambda_1 + \lambda_2 + \lambda_3 ) / 3 \label{eq:md}
-\end{align}
-where $\lambda_i$ are the eigenvalues of the diffusion tensor
-associated with the diffusion signal $S(\vec{q})$. There exist
-two main reasons to justify their choice.
-First, they are well-understood and standardized in clinical routine.
-Second, together they contain most of the information that is
-usually extracted from the \gls{dwi}-derived scalar maps
-\cite{ennis_orthogonal_2006}.
-
-Preprocessing differed between simulated and real datasets in the
-shape-prior surfaces generation, as we describe hereafter.
 
@@
 obtained from \emph{FreeSurfer}.
-
-\subsection{Experiments and evaluation}
-\label{sec:experiments_evaluation}
-%
-\subsubsection{Validation on the simulated digital phantom} %
-\label{sec:validation_phantom}
-We firstly evaluated our approach on the simulated data, using
-the distorted data with a deformation field similar to the
-susceptibility artifact that affects real \gls{dwi} data. To
-this end, we report the following indices:
-\begin{itemize}
-\item \gls{se}, a distance between the one-to-one corresponding
-vertices, weighted by their respective Voronoi area.
-\begin{equation}
-MSE = \sum\limits_k \sum\limits_j^M w_j\,\| \mathbf{x}_j - \hat{\mathbf{x}}_j \|
-\end{equation}
-where $\mathbf{x}_j$ are the locations of the $M$ vertices of the $k$ priors,
-$\hat{\mathbf{x}}_j$ are the corresponding locations recovered, and $w_j$ the
-weighting factor as the relative surface of the Voronoi area.
-\item \gls{wi}, L2-distance between the theoretical and the recovered
-deformation field.
-\begin{equation}
-WI = \frac{1}{N} \sum\limits_i^N \| \mathbf{d}_i - \hat{\mathbf{d}}_i \|
-\end{equation}
-where $\mathbf{d}_i$ is the theoretical displacement vector at position $i$
-and $\hat{\mathbf{d}}_i$ is the recovered one at the same index position.
-\item Parcellation agreement, the \gls{se} averaged by defined \glspl{roi}
-between the theoretical and the recovered parcellations.
-\item \gls{nof} agreement, between the fibers recovered on the original
-data and the processed data.
-\end{itemize}
-
-Additionally, the entropy of data is studied in both
-original and recovered datasets to draw another possible basis for the
-assessment of the real datasets. We also report the same indices for the
-outcome of a widely-used methodology that combines field-map susceptibility
-correction and T1-T2-\gls{dwi} registration .
-\todo[inline]{I suggest to give it a compact name and define it in the
-pre-processing section or even before in the intro}
-
-\subsubsection{Evaluation on real datasets}
-For the real datasets there is no published \emph{gold}-standard to validate results.
-Thus, visual results are provided to let compare the performance with the NICEACRONYM
-standard methodology. Additionally, cross-comparison of repeatability results are
-provided. In this second evaluation strategy, all the indices defined in
-\autoref{sec:validation_phantom} are reported.

### 2014-10-21 Publications/2013-TMI/08-appendix-01.tex
updated methods

@@
 
-\appendices
+\appendix
 
@@
 \end{equation}
+
+{\color{red} {Using the region descriptors derived in \autoref{sec:methods_map}, we propose
+an \gls{adf}-like, \gls{acwe}-based, piece-wise constant, image segmentation
+model (where the unknown is the deformation field)
+\cite{chan_active_2001} with the energy functional obtained in
+\eqref{eq:final_map_energy}. This inverse problem is ill-posed
+\cite{bertero_illposed_1988,hadamard_sur_1902}.
+In order to account for deformation field regularity and to render the
+problem well-posed, we include limiting and regularization terms into
+the energy functional \cite{morozov_linear_1975,tichonov_solution_1963}:
+%
+\begin{multline}
+E(u) = \sum\limits_l \int_{\Omega'_l} \mdist{f}{l} \,d\vec{x} \\
++ \int \vec{u}^T \, A \, \vec{u} \, d\vec{x} + \int \tr\{(\nabla \vec{u}^T)^T B (\nabla \vec{u}^T)\} d\vec{x}
+\label{eq:complete_energy}
+\end{multline}
+%
+These regularity terms ensure that the segmenting contours in
+\gls{dwi} space are still close to their native shape. The model
+easily allows to incorporate inhomogeneous and anisotropic
+regularization \cite{nagel_investigation_1986} to better regularize
+the \gls{epi} distortion.}}
+
+\todo[inline]{this last paragraphs in red color were located a bit
+further, but I think we can keep some of the refs and still reproduce
+the continuous version of the energy functional}
+
+\section{Extensions}
+
+\subsubsection{Semi-implicit Euler-forward optimization}
+It is necessary to discretize $t$ in order to obtain a numerical
+implementation of the equation \eqref{eq:gradient_final}:
+\begin{align}
+\frac{\vec{u}_k^{t+1}-\vec{u}_k^t}{\tau} =
+&- \underset{l,m}{\sum} \underset{i}{\sum}
+\left[ \mdist{f_i'}{l} - \mdist{f_i'}{m}\right]
+\psi_k(\vec{c}_i)\, \hat{\vec{n}}_i \notag\\
+&+2\, \boldsymbol{\alpha} \vec{u}_k^{t+1}
+-2\, \boldsymbol{\beta} \Delta \vec{u}_k^{t+1}.
+\end{align}
+
+The associated Euler-Lagrange equation is found as:
+\begin{equation}
+(\tau^{-1} +2\, \boldsymbol{\alpha} - 2\, \boldsymbol{\beta} \Delta )\, \vec{u}_k^{t+1} =
+\tau^{-1} \vec{u}_k^t - \frac{\partial}{\partial \vec{u}_k} E_{data}(\vec{u}_k^t),
+\end{equation}
+that is a linear system that we translate into Fourier domain,
+to obtain the next deformation field $\vec{u}_k^{t+1}$:
+\begin{multline}
+\vec{u}_k^{t+1} = \\
+ \mathcal{F}^{-1} \left\lbrace
+\frac{\mathcal{F}\lbrace \tau^{-1}\vec{u}_{k}^{t} - \underset{l,m}{\sum} \underset{i}{\sum}
+\left[ \mdist{f_i'}{l} - \mdist{f_i'}{m}\right] \rbrace}
+     {\mathcal{F}\lbrace (\tau^{-1} +2\,\boldsymbol{\alpha})\mathcal{I} - 2\,\boldsymbol{\beta} \Delta ) \rbrace}
+     \right\rbrace
+\end{multline}
+
+Here, we rewrite the Laplacian as a linear combination of the identity and shift operators:
+\begin{equation}
+\Delta = \sum\limits_d \mathcal{S}_d^- + \mathcal{S}_d^+ - 2 \mathcal{I}
+\end{equation}
+where $\mathcal{S}_{d}^{\pm}$ stands for the forward ($+$) and backward ($-$) shift
+operator along coordinates axis $d$, of which the Fourier transform is found easily as
+\begin{equation}
+\mathcal{F}\{\mathcal{S}_{d}^{\pm}\} = e^{\pm \vec{i}\omega_{d}},
+\end{equation}
+where $\omega_{d}$ is the normalized pulsation along $d$-direction. Accordingly, the
+Fourier transform of the discrete Laplacian is found as
+\begin{equation}
+\mathcal{F}\{\Delta\} = \sum\limits_d e^{-\vec{i}\omega_d } + e^{\vec{i}\omega_d } - 2 = \sum\limits_d \left( 2\cos(\omega_d) - 2 \right)
+\end{equation}
+
+The remaining transforms are trivial or can be computed using the \gls{fft}
+as in \citep{estellers_efficient_2011}.

### 2014-10-21 Publications/2013-TMI/03-introduction.tex
updated all

--- 2014-10-20 Publications/2013-TMI/03-introduction.tex
@@
 
-\IEEEPARstart{F}{usion} of information from different images is a difficult
-  task in image processing, especially when the scale of the imaged structures
-  is in the range of the image resolution or below.
-In the case of the human brain, the cortex has a typical thickness of up to 5 mm
-  \citep{fischl_measuring_2000}.
-Moreover, as the cortex is densely folded, points located afar in the topological
-  manifold of the brain's surface can be very close in the corresponding 3D space.
-Consequently, accuracy and precision are very difficult to achieve, but necessary properties
-  of registration algorithms.
+\IEEEPARstart{N}{onlinear} registration of images is a challenging task ubiquitous
+  in an endless number of image analysis applications.
+The problem is solved finding a spatial mapping $U$ that aligns the
+  information given in two different coordinate systems (generally called
+  \emph{reference}, $R$, and \emph{moving}, $M$):
+  \begin{align}
+  U\colon R \subset \mathbb{R}^n &\to M \subset \mathbb{R}^n \notag\\
+  \vec{r} &\mapsto \vec{r}' =\vec{r}+u(\vec{r}),
+  \label{eq:transform}
+  \end{align}
+  where $\vec{r}$ denotes a position in the reference domain and $\vec{r}'$
+  its corresponding location in the target, and $n$ the dimension of domains
+  ($n=3$ for 3D images).
+Finally, $\vec{u} = u(\vec{r})$ is the function that models the displacement
+  of every point in the reference domain.
 
-In this paper we propose a novel registration framework to simultaneously
-solving the segmentation, distortion and cortical parcellation challenges,
-by exploiting as strong shape-prior the detailed morphology extracted
-from high-resolution and anatomically correct \gls{mri}.
-Indeed, hereafter
-we assume this segmentation problem in anatomical images is reliably and
-accurately solved with readily available tools (e.g.
-\citep{fischl_freesurfer_2012}).
-After global alignment with \gls{t1} using existing approaches, the remaining
-spatial mismatch between anatomical and diffusion space is due to susceptibility
-distortions.
-Finally, we need to establish precise spatial correspondence between the
-surfaces in both spaces, including the tangential direction for parcellation.
-Therefore, we can reduce the problem to finding the differences of spatial
-distortion in between anatomical and \gls{dwi} space.
-We thus reformulate the segmentation problem as an inverse problem, where we
-seek for an underlying deformation field (the distortion) mapping
-from the structural space into the diffusion space, such that the structural
-contours segment optimally the \gls{dwi} data.
-In the process, the one-to-one
-correspondence between the contours in both spaces is guaranteed, and projection
-of parcellisation to \gls{dwi} space is implicit and consistent.
+The wide range of applications has derived in an intensive generation of diverse
+  registration methods.
+In a recent and comprehensive survey by \citeauthor{sotiras_deformable_2013}
+  \citep{sotiras_deformable_2013}, registration methods are classified by
+  matching criteria, deformation model, and optimization method.
+The most widespread matching criteria are probably the so-called \emph{iconic
+  methods} in \citep{sotiras_deformable_2013}.
+However they are not appropriate for all the registration problems
+  \citep{greve_accurate_2009}.
+They propose a \emph{hybrid method} \citep{sotiras_deformable_2013} for the
+  intra-subject registration of \gls*{t1} \gls*{mri} and \gls*{dmri} images
+  of the brain using a linear transformation model.
+As in \citep{greve_accurate_2009}, we exploit the precise structures that
+  can be readily extracted with widely-used tools at high resolution
+  and map this prior information to low resolution data,
+  in which the scale of the imaged structures is in the range of the image
+  resolution or above.
+Our approach differs with \citep{greve_accurate_2009} in two fundamental
+  choices: 1) We use a nonlinear deformation model that enables the
+  correction of the susceptibility-derived distortions
+  \citep{jezzard_correction_1995} that \gls*{dmri} data typically present.
+And 2) our matching criteria does not necessarily need for clear intensity
+  steps at the tissue interfaces where surfaces must be fitted.
 
-We test our proposed joint segmentation-registration model on two different
-synthetic examples.
-The first example is a scalar sulcus model, where the
-\gls{csf}-\gls{gm} boundary particularly suffers from \gls{pve} and can only be
-segmented correctly thanks to the shape prior and its coupling with the inner,
-\gls{gm}-\gls{wm} boundary through the imposed deformation field regularity.
-The second case deals with more realistic \gls{dwi} data stemming from
-phantom simulations of a simplistic brain data.
-Again, we show that the
-proposed model successfully segments the \gls{dwi} data based on two derived
-scalar features, namely \gls{fa} and \gls{md}, while establishing an estimate
-of the dense distortion field.
+\subsection{Related work}
+\label{sec:methods_background}
+Registration-segmentation methods are typically built including a deformation
+  model to guide the evolution of a segmentation algorithm.
+Even though some Bayesian approaches have been proposed to solve the joint problem
+  \citep{wyatt_map_2003}, segmentation is usually performed using an evolving
+  active contour \citep{chan_active_2001} that normally includes a shape prior
+  term to the energy functional \citep{chen_using_2002,bresson_variational_2006,
+  chan_level_2005,cremers_kernel_2006,gastaud_combining_2004}.
+These methods generally have an explicit description of the expected relative boundary
+  locations of the object to be delineated, and some, even model the statistical deviations
+  from this average shape.
 
-The rest of this paper is organized as follows.
-First, in \autoref{sec:methods}
-we introduce our proposed model for joint multivariate segmentation-registration.
-Then we provide a more detailed description of the data and experimental setup in
-\autoref{sec:experiments}.
-We present results in \autoref{sec:results} and conclude
-in \autoref{sec:conclusion}.
+An early ``morphing''-segmentation approach was proposed in \citep{bertalmio_morphing_2000},
+  and the first proper registration method was presented by \citeauthor{yezzi_variational_2001}
+  \citep{yezzi_variational_2001}, where the energy functional is defined simultaneously in the target
+  and reference images with an affine transformation supporting the coordinates mapping
+  (the so-called ``2 \glspl*{pde} approach'').
+\citeauthor{vemuri_joint_2003} proposed the first atlas-based registration \citep{vemuri_joint_2003}
+  based on the level-sets framework and using only one \gls*{pde}.
+The principal application of these techniques was the image registration between timesteps in
+  a time-series or between different images \citep{paragios_level_2003}.
+\citeauthor{unal_coupled_2005} \citep{unal_coupled_2005} extended the idea of
+  \citep{bertalmio_morphing_2000,yezzi_variational_2001} to nonlinear registration
+  using a free deformation field model.
+A summary of these second set of techniques is performed in \citep{droske_mumfordshah_2009},
+  proposing two different approaches to applying the Mumford-Shah \citep{mumford_optimal_1989}
+  functional in joint registration and segmentation, thorugh the propagation of the deformation
+  ield from the contours to the whole image definition.
+Finally, \citeauthor{gorthi_active_2011} presented a comprehensive generalization of the
+  existing methodologies \citep{gorthi_active_2011} based on the level-sets approximation.
+Even though the historical evolution of joint segmentation and registration procedures is strongly
+  linked to the level sets approximation, our work is based in derivations of the \gls{acwe}
+  framework \citep{chan_active_2001}, and related to \citep{guyader_combined_2011}.
+
+Finally, a derivation of the latter group is composed by atlas-based segmentation methods
+  \citep{gorthi_segmentation_2009,gorthi_active_2011,pohl_unifying_2005,pohl_bayesian_2006,
+  wang_joint_2006}, where the prior imposes consistent voxel-based classification of contiguous
+  regions.
+A comprehensive summary of the methodologies derived from the \gls{acwe} model, with special attention
+  to joint registration-segmentation methods is found in \citep{gorthi_active_2011}.
+
+%{\color{red} {The main diverging points with respect to
+%\citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
+%$\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
+%gradients \citep{besson_dream2s_2003,herbulot_segmentation_2006};
+%2) regularization is also based on linear diffusion smoothing \citep{thirion_image_1998},
+%but we replace the Gaussian filtering by other constraints
+%studied in \citep{nagel_investigation_1986} to better the problem;
+%3) optimization is applied in the spectral
+%domain, observing anisotropic and inhomogeneous mappings along each direction.
+%With respect \citep{guyader_combined_2011}, the main differences are
+%the distance function, and the spectral solution to the optimization updates,
+%as we shall cover in \autoref{sec:numerical_implementation}.}}
+
+\subsection{Susceptibility-derived distortions}
+
+\subsection{Our framework}
+\label{sec:our_framework}
+
+%In this paper we formulate the joint registration-segmentation
+%problem as follows.
+%such that the known contours in anatomical space $T$ optimally segment
+%the diffusion space $D$.
+%
+%
+%Whereas related \glspl*{adf} introduced in \autoref{sec:methods_background}
+%make use of explicit level-set formulations to solve \eqref{eq:gradient_descent},
+%we alternatively use \emph{shape-gradients}
+%\cite{besson_dream2s_2003,herbulot_segmentation_2006}.
+
+%
+%Let us denote by $\vec{x}$ the voxel and $F(\vec{x}) = [ f_1, f_2, \ldots, f_N]^T$
+%  its associated feature vector in the following.
+%
+%In our application, these surfaces are precise tissue interfaces of interest extracted
+%  from a high-resolution, anatomically correct reference volume using
+%  the well-established
+%
+%In this paper we propose a novel registration framework to simultaneously
+%solving the segmentation, distortion and cortical parcellation challenges,
+%by exploiting as strong shape-prior the detailed morphology extracted
+%from high-resolution and anatomically correct \gls{mri}.
+%Indeed, hereafter
+%we assume this segmentation problem in anatomical images is reliably and
+%accurately solved with readily available tools (e.g.
+%\citep{fischl_freesurfer_2012}).
+%After global alignment with \gls{t1} using existing approaches, the remaining
+%spatial mismatch between anatomical and diffusion space is due to susceptibility
+%distortions.
+%Finally, we need to establish precise spatial correspondence between the
+%surfaces in both spaces, including the tangential direction for parcellation.
+%Therefore, we can reduce the problem to finding the differences of spatial
+%distortion in between anatomical and \gls{dwi} space.
+%We thus reformulate the segmentation problem as an inverse problem, where we
+%seek for an underlying deformation field (the distortion) mapping
+%from the structural space into the diffusion space, such that the structural
+%contours segment optimally the \gls{dwi} data.
+%In the process, the one-to-one
+%correspondence between the contours in both spaces is guaranteed, and projection
+%of parcellisation to \gls{dwi} space is implicit and consistent.
+%
+%We test our proposed joint segmentation-registration model on two different
+%synthetic examples.
+%The first example is a scalar sulcus model, where the
+%\gls{csf}-\gls{gm} boundary particularly suffers from \gls{pve} and can only be
+%segmented correctly thanks to the shape prior and its coupling with the inner,
+%\gls{gm}-\gls{wm} boundary through the imposed deformation field regularity.
+%The second case deals with more realistic \gls{dwi} data stemming from
+%phantom simulations of a simplistic brain data.
+%Again, we show that the
+%proposed model successfully segments the \gls{dwi} data based on two derived
+%scalar features, namely \gls{fa} and \gls{md}, while establishing an estimate
+%of the dense distortion field.
+%
+%The rest of this paper is organized as follows.
+%First, in \autoref{sec:methods}
+%we introduce our proposed model for joint multivariate segmentation-registration.
+%Then we provide a more detailed description of the data and experimental setup in
+%\autoref{sec:experiments}.
+%We present results in \autoref{sec:results} and conclude
+%in \autoref{sec:conclusion}.
+%
+%
+%The properties of the reconstructed tensors and derived scalar maps have
+%been studied by \cite{ennis_orthogonal_2006}. Based on their
+%findings, \gls{fa}~\eqref{eq:fa} and \gls{md}~\eqref{eq:md} are
+%considered complementary features, and therefore we selected them for the
+%energy model \eqref{eq:complete_energy} in driving the
+%registration-segmentation process.
+%Whereas \gls{fa} informs mainly about the \emph{shape} of diffusion,
+%the \gls{md} is more related to the \emph{magnitude} of the process:
+%
+%\begin{align}
+%\mathrm{FA} &= \sqrt{ \frac{1}{2}}\,\frac{\sqrt{ (\lambda_1 - \lambda_2)^2 + (\lambda_2 - \lambda_3)^2 + (\lambda_3 - \lambda_1)^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}} \label{eq:fa} \\
+%\mathrm{MD} &= ( \lambda_1 + \lambda_2 + \lambda_3 ) / 3 \label{eq:md}
+%\end{align}
+%where $\lambda_i$ are the eigenvalues of the diffusion tensor
+%associated with the diffusion signal $S(\vec{q})$. There exist
+%two main reasons to justify their choice.
+%First, they are well-understood and standardized in clinical routine.
+%Second, together they contain most of the information that is
+%usually extracted from the \gls{dwi}-derived scalar maps
+%\cite{ennis_orthogonal_2006}.
+%

### 2014-10-21 Publications/2013-TMI/04-methods.tex
updated all

@@
   \begin{equation}
-  P(F \mid U,\Omega) = \underset{l}{\prod} \underset{\vec{x}\in \Omega_l}{\prod}
+  P(F \mid U,\Omega) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
     P\left( \vec{f}' \mid U \right),
@@
   \end{equation}
-  where $\vec{f}' = F(\vec{x}')$ is the feature vector at the displaced
-  position $\vec{x}'$ (\autoref{eq:transform}) of the multispectral target
+  where $\vec{f}' = F(\vec{r}')$ is the feature vector at the displaced
+  position $\vec{r}'$ (\autoref{eq:transform}) of the multispectral target
   volume.
@@
  	\begin{align}
-  P( F \mid U, \Omega ) &= \underset{l}{\prod} \underset{\vec{x} \in \Omega_l}{\prod}
+  P( F \mid U, \Omega ) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
   \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
-  &= \underset{l}{\prod} \underset{\vec{x} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
+  &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
   \mdist{f'}{l} \right)}},
@@
   \begin{equation*}
-  P(U) = \underset{\vec{x}}{\prod} p(u(\vec{x})),
+  P(U) = \underset{\vec{r}}{\prod} p(u(\vec{r})) =
+  \underset{\vec{r}}{\prod} p_0(u(\vec{r})) \, p_1(u(\vec{r})),
   \end{equation*}
@@
   \begin{align*}
-  p(u(\vec{x})) &= p_0(u(\vec{x})) \, p_1(u(\vec{x})) \notag\\
-  p_0(u(\vec{x})) &= \mathcal{N}( u(\vec{x}) \mid 0, A^{-1}) \notag\\
-  p_1(u(\vec{x})) &= \mathcal{N}(  \nabla \cdot u(\vec{x}) \mid 0, B^{-1}),
+  p_0(u(\vec{r})) &= \mathcal{N}( u(\vec{r}) \mid 0, A^{-1}), \notag\\
+  p_1(u(\vec{r})) &= \mathcal{N}(  \nabla \cdot u(\vec{r}) \mid 0, B^{-1}),
   \end{align*}
@@
   \begin{align}
-    p_0(u(\vec{x})) &= \underset{\vec{x}}{\sum} \mathcal{N}( u(\vec{x}) \mid 0,
+    p_0(u(\vec{r})) &= \underset{\vec{r}}{\sum} \mathcal{N}( u(\vec{r}) \mid 0,
       (\boldsymbol{\alpha}^{\circ\frac12}\,\vec{I}_n)^{-1}), \notag\\
-    p_1(u(\vec{x})) &= \underset{\vec{x}}{\sum} \mathcal{N}( \nabla \cdot u(\vec{x}) \mid 0,
+    p_1(u(\vec{r})) &= \underset{\vec{r}}{\sum} \mathcal{N}( \nabla \cdot u(\vec{r}) \mid 0,
       (\boldsymbol{\beta}^{\circ\frac12}\,\vec{I}_n)^{-1}).
@@
   E(F \mid U) &= -\log \underset{l}{\prod}
-  \underset{\vec{x} \in \Omega_l}{\prod}
-  \mathcal{N} \left( \vec{f}' \mid \Theta_l \right)\,p_0( u(\vec{x}))\,p_1( u(\vec{x})), \notag\\
+  \underset{\vec{r} \in \Omega_l}{\prod}
+  \mathcal{N} \left( \vec{f}' \mid \Theta_l \right)\,p_0( u(\vec{r}))\,p_1( u(\vec{r})) \notag\\
   &= C + \underset{l}{\sum}
-  \underset{\vec{x} \in \Omega_l}{\sum}
-  \mdist{f'}{l}, \notag\\
-  &+ \underset{\vec{x} \in \Omega}{\sum} \boldsymbol{\alpha} \cdot u(\vec{x})^{\circ2}
-  + \underset{\vec{x} \in \Omega}{\sum} \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{x}))^{\circ2},
+  \underset{\vec{r} \in \Omega_l}{\sum}
+  \mdist{f'}{l} \notag\\
+  &+ \underset{\vec{r} \in \Omega}{\sum} \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
+  + \underset{\vec{r} \in \Omega}{\sum} \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2},
   \label{eq:energy}
@@
 The transformation from the structural space into the coordinate space of the
-  target image is achieved through a dense deformation field $u(\vec{x})$, such that:
+  target image is achieved through a dense deformation field $u(\vec{r})$, such that:
   \begin{equation}
@@
 Since the nodes of the anatomical surfaces likely lay off-grid, it is required to
-  derive $u(\vec{x})$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
+  derive $u(\vec{r})$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
 Densification is achieved through a set of associated basis functions $\psi_k$:
   \begin{equation}
-  u(\vec{x}) = \sum_k \psi_k(\vec{x}) \vec{u}_k.
+  u(\vec{r}) = \sum_k \psi_k(\vec{r}) \vec{u}_k.
   \label{eq:intp_kernel}
@@
     \vec{c}_i' = \vec{c}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
-      B_3( (\vec{c}_i - \vec{x}_k) \cdot \hat{\mathbf{e}}_d ) \right],
+      B_3( (\vec{c}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
   \label{eq:transformation}
@@
   \begin{equation}
-  \frac{\partial u(\vec{x},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}_k},
+  \frac{\partial u(\vec{r},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}_k},
   \label{eq:general_gradient_descent}
@@
   C + \underset{l}{\sum}
-  \underset{\vec{x} \in \Omega_l}{\sum} \mdist{f'}{l} \notag\\
-  &+ \underset{\vec{x} \in \Omega}{\sum} \boldsymbol{\alpha} \cdot u(\vec{x})^{\circ2}
-  + \underset{\vec{x} \in \Omega}{\sum} \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{x}))^{\circ2}
+  \underset{\vec{r} \in \Omega_l}{\sum} \mdist{f'}{l} \notag\\
+  &+ \underset{\vec{r} \in \Omega}{\sum} \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
+  + \underset{\vec{r} \in \Omega}{\sum} \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2}
   \Big\}.

### 2014-10-21 Publications/2013-TMI/06-results.tex
updated all

@@
+% -*- root: 00-main.tex -*-
 \section{Results and discussion}
@@
 
-Please! Write me ASAP.
+\subsection{Proof of concept}
+\label{sec:results_phantom}
+
+\subsection{Susceptibility distortion correction}
+\label{sec:results_phantom}
+
+\subsection{Discussion}
+\label{sec:discussion}
+Don't forget to mention other potential applications.
+
+We implemented the methodology upon the widely used
+Insight Registration and Segmentation
+Toolkit~\footnote{\url{http://www.itk.org}} (ITK)
+for its computational benefits, the standardized code, and
+with the aim at making the procedure publicly available
+when ready for sharing with the research community.

### 2014-10-21 Publications/2013-TMI/07-conclusion.tex
updated all

% -*- root: 00-main.tex -*-
\section{Conclusion}
\label{sec:conclusion}
%
We propose and make publicly available a novel methodology for
  simultaneous distortion correction and consistent parcellation
  of \gls*{dmri} images using a joint registration-segmentation
  framework.
From the Bayesian approach to image segmentation, we derive
  a cost function that models piecewise smooth regions with
  multivariate normal distributions.
This is, implicitly, an \acrlong*{acwe} model, in wich the
  contours evolve optimizing a B-Spline deformation
  model.
We demonstrate the aptness of the method with quantitative
  evaluations on digital phantoms and real \gls*{mri} data.

### 2014-10-21 Publications/2013-TMI/08-appendix-00-shapepriors.tex
updated all

% -*- root: 00-main.tex -*-

\section{Extended formulations}

\subsection{Shape priors}
\label{app:shape_priors}
Let $R(\vec{r})$ be an ``arbitrary'' function over the image domain,
and $\Omega$ a certain image region with $\Gamma_{l,m}$ its corresponding
outer boundary as defined in \autoref{sec:methods_map}.
We now derive the domain integral w.r.t. $t$:
  \begin{align*}
  \frac{\partial}{\partial t} \int_\Omega R(\vec{r}') d\vec{r} &=
  \int_\Omega \frac{\partial}{\partial t}R(\vec{r}') d\vec{r} \notag\\
  &- \int_{\Gamma_{l,m}} R(\vec{r}') \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
  N_{\Gamma_{l,m}}\right\rangle d\vec{r},
  \end{align*}
  where $\left\langle\frac{\partial\Gamma_{l,m}}{\partial t}, N_{\Gamma_{l,m}}\right\rangle$ is
  the projection of the boundary movement on the unit inward normal $N_{\Gamma_{l,m}}$.
Assuming that the region descriptors $\Theta_l$ vary slowly enough, we can consider
  that $\frac{\partial}{\partial t} R(\vec{r}') = 0$ and thus:
  \begin{equation}
  \frac{\partial}{\partial t} \int_\Omega R(\vec{r}') d\vec{r} =
  - \int_{\Gamma_{l,m}} R(\vec{r}') \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
  N_{\Gamma_{l,m}}\right\rangle d\vec{r}
  \label{eq:shape_gradients}
  \end{equation}

### 2014-11-05 Publications/2013-TMI/04-methods.tex
update paper

@@
   evaluated based on their posterior probability given the observed data
-  $F$.
+  $R$.
 Using the Bayes' rule, the posterior likelihood is computed as:
   \begin{equation}
-  P(U \mid F,\Gamma_{l,m}) = \frac{P(F \mid U,\Gamma_{l,m})\, P(U)}{P(F)},
+  P(U \mid R,\Gamma_{l,m}) = \frac{P(R \mid U,\Gamma_{l,m})\, P(U)}{P(R)},
   \label{eq:bayes_rule}
   \end{equation}
-  where $P(F \mid U,\Gamma_{l,m})$ is the data-likelihood, and
+  where $P(R \mid U,\Gamma_{l,m})$ is the data-likelihood, and
   $\Gamma_{l,m}$ are a set of surfaces corresponding to the interfaces
@@
 Then, the \gls*{map} criterium \citep{leemput_automated_1999} is used
-  to find the mapping estimate $\hat{U}$ that aligns $\Gamma$ into $F$:
+  to find the mapping estimate $\hat{U}$ that aligns $\Gamma$ into $R$:
   \begin{align}
-  \hat{U} &= \underset{U}{\argmax} \left\{ P(U \mid F,\Gamma_{l,m}) \right\} \notag\\
-   &= \underset{U}{\argmax} \left\{ P(F \mid U,\Gamma_{l,m})\, P(U) \right\}.
+  \hat{U} &= \underset{U}{\argmax} \left\{ P(U \mid R,\Gamma_{l,m}) \right\} \notag\\
+   &= \underset{U}{\argmax} \left\{ P(R \mid U,\Gamma_{l,m})\, P(U) \right\}.
   \label{eq:map_u}
@@
   \begin{equation}
-  P(F \mid U,\Omega) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
+  P(R \mid U,\Omega) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
     P\left( \vec{f}' \mid U \right),
@@
   \end{equation}
-  where $\vec{f}' = F(\vec{r}')$ is the feature vector at the displaced
+  where $\vec{f}' = R(\vec{r}')$ is the feature vector at the displaced
   position $\vec{r}'$ (\autoref{eq:transform}) of the multispectral target
@@
  	\begin{align}
-  P( F \mid U, \Omega ) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
+  P( R \mid U, \Omega ) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
   \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
@@
   to the descriptors of region $l$ as
-  $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
+  $\mathcal{E}_l(\vec{r}) = \mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
 
@@
   \begin{align}
-  E(F \mid U) &= -\log \underset{l}{\prod}
+  E(R \mid U) &= -\log \underset{l}{\prod}
   \underset{\vec{r} \in \Omega_l}{\prod}
@@
 
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/figure01.pdf}
+	\caption{The active contours are defined as the interfacing surfaces of the segmenting
+	\glspl{roi} $\Omega_i$, and represented in green and dark blue colors in this
+	close-up.
+	They iteratively through their normals $\hat{n}_i$ at each vertex $\vec{v}_i$ of the mesh.
+	The gradient speeds $\vec{s}_i$ are computed as the disparity between data energies of
+	the vertex $\vec{v}_i$ in each limiting region as described in \autoref{eq:energy}.
+	In this figure, the gradient speed corresponding to $\vec{v}_1$ is written in the lower
+	box, $\Omega_{wm}$ being the inner limiting region and $\Omega_{gm}$ the outer.
+	Finally, every $\bar{s}_i$ is projected into the grid of control points $\vec{u}_k$ that
+	support the deformation field through the corresponding weights $\psi_k(\vec{v}_i)$ obtained
+	as in \autoref{eq:basis_derivative}.
+	}\label{fig:method}
+\end{figure}
+
 \subsubsection{Deformation model}
 \label{sec:deformation_model}
-Let us denote $\{\vec{c}_i\}_{i=1 \ldots N_c}$ the nodes of one or several prior
+Let us denote $\{\vec{v}_i\}_{i=1 \ldots N_c}$ the vertices of one or several prior
   surface(s).
-In our application, these surfaces are extracted using \emph{FreeSurfer}
-  \citep{fischl_freesurfer_2012} in the case of real datasets.
+In our application, these surfaces are triangularized meshes extracted using \emph{FreeSurfer}
+  \citep{fischl_freesurfer_2012}.
 The transformation from the structural space into the coordinate space of the
@@
   \begin{equation}
-  \vec{c}_i' = U\{\vec{c}_i\} = \vec{c}_i + u(\vec{c}_i),
+  \vec{v}_i' = U\{\vec{v}_i\} = \vec{v}_i + u(\vec{v}_i),
   \label{eq:nodes_tfm}
@@
   \begin{equation}
-    \vec{c}_i' = \vec{c}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
-      B_3( (\vec{c}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
+    \vec{v}_i' = \vec{v}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
+      B_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
   \label{eq:transformation}
@@
   C + \underset{l}{\sum}
-  \underset{\vec{r} \in \Omega_l}{\sum} \mdist{f'}{l} \notag\\
-  &+ \underset{\vec{r} \in \Omega}{\sum} \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
-  + \underset{\vec{r} \in \Omega}{\sum} \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2}
+  \int_{\Omega_l} \mdist{f'}{l} \notag\\
+  &+ \int_{\Omega_l} \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
+  + \int_{\Omega_l} \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2}
   \Big\}.
@@
   \underset{l}{\sum} \frac{ \partial }{\partial \vec{u}_k} \left\{
-   \underset{\vec{x} \in \Omega_l}{\sum} \mdist{f'}{l} \right\} \notag\\
-  &= \underset{l,m}{\sum} \underset{i}{\sum}
-  \left[ \mdist{f_i'}{l} - \mdist{f_i'}{m} \right]
-  \left\langle \frac{\partial \vec{c}_i'}{\partial \vec{u}_k}, \vec{n_i}'\right\rangle,
+   \underset{\vec{x} \in \Omega_l}{\sum} \mdist{f'}{l} \right\}
+  = \underset{l}{\sum} \underset{i}{\sum}
+   \left\langle \frac{\partial \vec{v}_i'}{\partial \vec{u}_k}, \bar{s}_i'\right\rangle,
   \label{eq:gradient_wshape}
@@
   being any pair of neighboring regions, and $\Gamma_{l,m}$ the contour separating
-  them such that $\vec{x}' = \vec{c}' \in\Gamma_{l,m} \iff \vec{x}\in \partial\Omega_i \cap \partial\Omega_j$
+  them such that $\vec{x}' = \vec{v}' \in\Gamma_{l,m} \iff \vec{x}\in \partial\Omega_i \cap \partial\Omega_j$
   and $\vec{n_i}'$ is the unit inward normal to the contour at $c_i'$.
@@
   \begin{align}
-  \frac{\partial \vec{c}_i'}{\partial \vec{u}_k} &= \frac{\partial}{\partial \vec{u}_k}
-  \left\{ \vec{c}_i + \sum_k \psi_k(\vec{c}_i) \vec{u}_k \right\}
-  = \psi_k(\vec{c_i})\, \hat{\vec{e}}
+  \frac{\partial \vec{v}_i'}{\partial \vec{u}_k} &= \frac{\partial}{\partial \vec{u}_k}
+  \left\{ \vec{v}_i + \sum_k \psi_k(\vec{v}_i) \vec{u}_k \right\}
+  = \psi_k(\vec{v}_i)\, \hat{\vec{e}}
+  \label{eq:basis_derivative}
   \end{align}
   where $\hat{\vec{e}}$ is the coordinates system's unit vector.
-  Projecting this gradient onto the surface normal,
-  $\left\langle \frac{\partial}{\partial \vec{u}_k}{\vec{c}_i}', \vec{n_i}'\right\rangle
-  = \psi_k(\vec{c_i})\, \hat{\vec{n}}_i$, then the
-  full gradient evolution equation \eqref{eq:gradient_descent} yields:
+The vertex speeds $\bar{s}_i'$ obtained by computing the shape-gradients \autoref{eq:shape_gradients},
+	are then projected to the deformation field in order to obtain the derivatives $\vec{g}_k$
+	corresponding to $\vec{u}_k$:
+  \begin{equation}
+  \vec{g}_{i,k} = \left\langle \frac{\partial}{\partial \vec{u}_k}{\vec{v}_i}', \bar{s}_i'\right\rangle
+  = - \left[ \mdist{f_i'}{l} - \mdist{f_i'}{m} \right] \psi_k(\vec{v}_i)\, \hat{\vec{e}},
+  \end{equation}
+  then the full gradient evolution equation \eqref{eq:gradient_descent} yields:
   \begin{align}
   \frac{\partial E(\vec{u}_k)}{\partial \vec{u}_k} =
-  &- \underset{l,m}{\sum} \underset{i}{\sum}
-  \left[ \mdist{f_i'}{l} - \mdist{f_i'}{m} \right]
-  \psi_k(\vec{c}_i)\, \hat{\vec{n}}_i \notag\\
-  &+2\, \boldsymbol{\alpha} \vec{u}_k
+  &\underset{l,m}{\sum} \underset{i}{\sum}
+  \vec{g}_{i,k} +2\, \boldsymbol{\alpha} \vec{u}_k
   -2\, \boldsymbol{\beta} \Delta \vec{u}_k,
@@
 from the same set of control points that define the $L-1$ contours,
-it is possible to pre-cache all the $\psi_k(c_i)$ weights into a
+it is possible to pre-cache all the $\psi_k(\vec{v}_i)$ weights into a
 sparse matrix for fast densification. As well, we achieve a

### 2014-11-06 Publications/2013-TMI/02-abstract.tex
one more update to the paper

@@
 The strategy implicitly segments with sub-pixel resolution the image.
-We demonstrate the usefulness of the approach on a digital phantom, and on
-  10 real datasets randomly selected from the Human Connectome Project.
+We illustrate the usefulness of the approach on a digital phantom.
+Finally, we apply the registration to correct for suscetibility distortions
+  10 real diffusion MRI datasets.
+The joint segmentation-registration method successfully estimated the
+  distortions.
+We also compare its performance to a widely-used correction method that
+  demands for an extra MR acquisition.

### 2014-11-06 Publications/2013-TMI/03-introduction.tex
one more update to the paper

@@
   proposing two different approaches to applying the Mumford-Shah \citep{mumford_optimal_1989}
-  functional in joint registration and segmentation, thorugh the propagation of the deformation
-  ield from the contours to the whole image definition.
+  functional in joint registration and segmentation, through the propagation of the deformation
+  field from the contours to the whole image definition.
 Finally, \citeauthor{gorthi_active_2011} presented a comprehensive generalization of the
@@
 Even though the historical evolution of joint segmentation and registration procedures is strongly
-  linked to the level sets approximation, our work is based in derivations of the \gls{acwe}
+  linked to the level sets approximation, our work is based in derivations of the \acrlong{acwe}
   framework \citep{chan_active_2001}, and related to \citep{guyader_combined_2011}.

### 2014-11-06 Publications/2013-TMI/04-methods.tex
one more update to the paper

@@
   \mathcal{N} \left( \vec{f}' \mid \Theta_l \right)\,p_0( u(\vec{r}))\,p_1( u(\vec{r})) \notag\\
-  &= C + \underset{l}{\sum}
-  \underset{\vec{r} \in \Omega_l}{\sum}
-  \mdist{f'}{l} \notag\\
-  &+ \underset{\vec{r} \in \Omega}{\sum} \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
-  + \underset{\vec{r} \in \Omega}{\sum} \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2},
+  &= C + \underset{l}{\sum} \int_{\Omega_l}
+  \mdist{f'}{l} \,d\vec{r} \notag\\
+  &+ \int_{\Omega} \left[ \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
+  + \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2} \right] \,d\vec{r},
   \label{eq:energy}
@@
 \begin{figure}
-	\includegraphics[width=\linewidth]{figures/figure01.pdf}
+	\includegraphics[width=\linewidth]{figures/figure01}
 	\caption{The active contours are defined as the interfacing surfaces of the segmenting
@@
 	They iteratively through their normals $\hat{n}_i$ at each vertex $\vec{v}_i$ of the mesh.
-	The gradient speeds $\vec{s}_i$ are computed as the disparity between data energies of
+	The gradient speeds $\bar{s}_i$ are computed as the disparity between data energies of
 	the vertex $\vec{v}_i$ in each limiting region as described in \autoref{eq:energy}.
@@
   \frac{ \partial }{\partial \vec{u}_k} \Big\{
-  C + \underset{l}{\sum}
-  \int_{\Omega_l} \mdist{f'}{l} \notag\\
-  &+ \int_{\Omega_l} \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
-  + \int_{\Omega_l} \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2}
+  \int_{\Omega} \underset{l}{\sum} \mdist{f'}{l} \,d\vec{r} \notag\\
+  &+ \int_{\Omega} [ \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
+  + \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2} ] \,d\vec{r}
   \Big\}.
@@
   \frac{\partial E_{data}(\vec{u})}{\partial \vec{u}_k} &=
-  \underset{l}{\sum} \frac{ \partial }{\partial \vec{u}_k} \left\{
-   \underset{\vec{x} \in \Omega_l}{\sum} \mdist{f'}{l} \right\}
-  = \underset{l}{\sum} \underset{i}{\sum}
+  \frac{ \partial }{\partial \vec{u}_k} \left\{
+   \underset{\vec{x} \in \Omega_l}{\sum} \underset{l}{\sum} \mdist{f'}{l} \right\}
+  = \underset{i}{\sum}
    \left\langle \frac{\partial \vec{v}_i'}{\partial \vec{u}_k}, \bar{s}_i'\right\rangle,
@@
   \end{align}
-  in this case, the formulation has been adapted to the non-binary case, $l,m$
+  in this case, the formulation has been adapted to the non-binary case, $\{l,m\}$
   being any pair of neighboring regions, and $\Gamma_{l,m}$ the contour separating
@@
   \frac{\partial E(\vec{u}_k)}{\partial \vec{u}_k} =
-  &\underset{l,m}{\sum} \underset{i}{\sum}
-  \vec{g}_{i,k} +2\, \boldsymbol{\alpha} \vec{u}_k
+  &\underset{i}{\sum} \vec{g}_{i,k} +2\, \boldsymbol{\alpha} \vec{u}_k
   -2\, \boldsymbol{\beta} \Delta \vec{u}_k,
@@
 
-\subsubsection{Region descriptor reestimation}
-In regular intervals, after certain number of iterations,
-the parameters $\Theta_l$ of the regions can be reestimated
-based on the shifted volumetric samples
-$\vec{x}' = \vec{x_0} + u(\vec{x})$.
+\subsection{Experiments and evaluation}
+\label{sec:experiments_evaluation}
+%
+The experiments follow a five-steps design:
+\begin{enumerate}
+	\item The prior surfaces are extracted from an undistorted and high-resolution dataset
+	  (e.g. \autoref{fig:phantom}).
+	\item A plausible, synthetic distortion $U^{-1}_{true}$ is generated.
+	\item The low-resolution (target) images are warped using the synthetic distortion.
+	\item Our method is run, obtaining a $\hat{U}_{tst}$.
+	  A cross-comparison methodology is also applied, to obtain a competing $\hat{U}_{cc}$.
+	\item Results are visually and quantitatively evaluated.
+\end{enumerate}
 
-\subsubsection{Convergence}
-{\color{red} discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition / Wolfe conditions, multi-resolutions on the bspline grid, image samples and surface sampling,  etc.}
+To quantitatively evaluate the registration results we computed the \gls*{swindex} as the one-to-one
+  distance between corresponding vertices of surfaces, weighted by their respective
+  Voronoi area $w_i$:
+  \begin{equation}
+  sWI = \frac{1}{P} \sum\limits_p^P \frac{1}{A_p} \sum\limits_i^{N_p} w_i\,\|
+  \vec{v}_i - \hat{\vec{v}}_i \|
+  \end{equation}
+  where $\vec{v}_i$ are the locations of the $N_p$ vertices in each $p \in \{1, \dots P\}$
+    prior, $A_p$ the total area of surface $p$, and $\hat{\vec{v}}_i$ is the location
+    recovered corresponding to the vertex $\vec{v}_i$.
 
-\subsubsection{Efficient field densification}
-As long as the dense deformation field is iteratively interpolated
-from the same set of control points that define the $L-1$ contours,
-it is possible to pre-cache all the $\psi_k(\vec{v}_i)$ weights into a
-sparse matrix for fast densification. As well, we achieve a
-diffeomorphic transform by
+In the application on real data, three experts blindly rated correction results obtained
+  with the proposed method and the cross-comparison reference.
+The full report of visual results is provided within the Supplemental Materials.
+
+  % \item \gls{wi}, L2-distance between the theoretical and the recovered
+  %   deformation field.
+  %   \begin{equation}
+  %   WI = \frac{1}{N} \sum\limits_i^N \| \mathbf{d}_i - \hat{\mathbf{d}}_i \|
+  %   \end{equation}
+  %   where $\mathbf{d}_i$ is the theoretical displacement vector at position $i$
+  %   and $\hat{\mathbf{d}}_i$ is the recovered one at the same index position.
+  % \item Parcellation agreement, the Dice overlap index of the \glspl*{roi}
+  %   contained by the parcellation.
+  % \end{itemize}
 
@@
 %
-We first present the proof of the concept on a simplistic digital phantom,
-  simulating a spherical crust of thin tissue with a folding resembling
-  a gyrus of the brain cortex folding (see \autoref{fig1}).
-We simulated 2.0mm isotropic \gls*{t1} and \gls*{t2} images at
-  several rician noise levels, with parameters TE/TR=10/1500ms and
-  TE/TR=90/5000ms respectively using \cite{caruyer_phantomas_2014}.
+We first present the proof of the concept on a simplistic digital phantom
+  (\autoref{fig:phantom}).
+Using \cite{caruyer_phantomas_2014}, we simulated 2.0mm isotropic \gls*{t1} and
+  \gls*{t2} images at several rician noise levels, with parameters TE/TR=10/1500ms
+  and TE/TR=90/5000ms respectively.
 The reference surfaces were extracted using the marching-cubes algorithm
-  on high-resolution (1.0mm isotropic) binary versions of the corresponding
-  maps of tissue volume fractions used during the simulation step.
+  shipped with \emph{FreeSurfer} \citep{fischl_freesurfer_2012}
+  on high-resolution (1.0mm isotropic) and binary images corresponding to
+  the tissue fraction maps used in simulations.
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/figure02}
+	\caption{The ``cortex'' phantom is a spherical shape with two sulci and an
+	 outer crust resembling the cortical folding.
+	When the model is used to generate low resolution images, the
+	  partial volume effect turns segmentation of the sulci a
+	  challenging problem with voxel-wise clustering methods,
+	  but it is successfully segmented with the proposed method (see
+	  \autoref{fig:results_phantom}).
+	Top row shows the central-axial plane of the simulated images at
+	  high resolution and noiseless (left is \gls*{t1} and right is
+	  \gls*{t2}).
+	Bottom row represents the extracted contours that are used as shape
+	  priors to be registered to low-resolution, noise corrupted versions
+	  of the phantom.
+	}\label{fig:phantom}
+\end{figure}
 
@@
 %
-For the evaluation of the algorithm with real brains with \gls*{dmri} data,
-  we randomly selected 10 datasets from the minimally preprocessed
-  \citep{glasser_minimal_2013} cohort of the \gls*{hcp}.
-These datasets comprehend the preprocessed (corrected for artifacts,
-  brain-extracted and spatially normalized) \gls*{t1}, \gls*{t2} and
-  multi-shell \gls*{dmri} images, along with the most prominent results of
-  the standard processing with \emph{FreeSurfer}: \emph{aseg} and
-  \emph{aparc} segmentations, and the surfaces corresponding to the cortical
-  \gls*{gm} exterior (\emph{pial}) and the \gls*{wm}/\gls*{gm} interface
-  (\emph{white}).
-This cohort is especially convenient for their careful preprocessing,
-  making available high-standard \gls*{mri} data accurately corrected
-  for distortions and artifacts.
-For the exact details of the scanning parameters we refer the reader to
-  \citep{essen_human_2012}.
-Selecting the appropriate labels in the \emph{aseg} segmentation, we applied
-  the marching-cubes algorithm again to extract the surfaces containing the
-  \gls*{csf} of the ventricular system.
-
-As we propose the \gls*{fa} and the \gls*{md} as target features, we
-  implemented a simplified pipeline for processing \gls*{dmri} using
-  \emph{MRtrix} \citep{tournier_mrtrix_2012}.
-\todo[inline]{insert here the summary of parameters}
-
-\subsection{Experiments and evaluation}
-\label{sec:experiments_evaluation}
-%
-\subsubsection{Synthetic distortion}
-\label{sec:distortion}
-
-\subsubsection{Evaluation indices} %
-\label{sec:evaluation}
-We report the following indices:
-  \begin{itemize}
-  \item \gls{se}, a distance between the one-to-one corresponding
-    vertices, weighted by their respective Voronoi area.
-    \begin{equation}
-    MSE = \sum\limits_k \sum\limits_j^M w_j\,\| \mathbf{x}_j - \hat{\mathbf{x}}_j \|
-    \end{equation}
-    where $\mathbf{x}_j$ are the locations of the $M$ vertices of the $k$ priors,
-    $\hat{\mathbf{x}}_j$ are the corresponding locations recovered, and $w_j$ the
-    weighting factor as the relative surface of the Voronoi area.
-  \item \gls{wi}, L2-distance between the theoretical and the recovered
-    deformation field.
-    \begin{equation}
-    WI = \frac{1}{N} \sum\limits_i^N \| \mathbf{d}_i - \hat{\mathbf{d}}_i \|
-    \end{equation}
-    where $\mathbf{d}_i$ is the theoretical displacement vector at position $i$
-    and $\hat{\mathbf{d}}_i$ is the recovered one at the same index position.
-  \item Parcellation agreement, the Dice overlap index of the \glspl*{roi}
-    contained by the parcellation.
-  \end{itemize}
+For the evaluation of the algorithm to real \gls*{dmri} data of human brains,
+  we downloaded 10 datasets from the ``minimally preprocessed''
+	 database of the \gls*{hcp}.
+We refer the reader to \citep{essen_human_2012} for exact details about acquisition
+  parameters, and \citep{glasser_minimal_2013} for preprocessing issues.
+The datasets comprehend \gls*{t1}, \gls*{t2} and multi-shell \gls*{dmri} images
+  already corrected for artifacts, brain-extracted and spatially normalized,
+  along with the most prominent results of the standard processing with
+  \emph{FreeSurfer}: \emph{aparc} segmentations, and the surfaces corresponding to
+  the cortical \gls*{gm} exterior and the \gls*{wm} interface with
+  cortical \gls*{gm}.

### 2014-11-06 Publications/2013-TMI/06-results.tex
one more update to the paper

--- 2013-05-20 Publications/2013-TMI/06-results.tex
@@
 
-Please! Write me ASAP.
+\subsection{Proof of concept}
+\label{sec:results_phantom}
+
+Applying the evaluation protocol introduced in \autoref{sec:experiments_evaluation} on
+  the ``cortex'' phantom (\autoref{fig:phantom}), we successfully tested the method and
+  fine tuned the algorithm for its use on real data.
+In this characterization experiment, we generated random smooth displacement fields $U^{-1}_{true}$
+  (step 4 in \autoref{sec:experiments_evaluation}) based on a B-Spline field with
+  control points evenly located in a $6 \times 6 \times 6$ grid covering the region of the phantom.
+Invertibility is ensured by controlling the maximum	displacement
+  \citep{rueckert_diffeomorphic_2006}.
+The software tool to generate these fields is included with the released package.
+In this experiment we used $\hat{U}_{cc} = (U^{-1}_{true})^{-1}$, i.e. the ground-truth
+  displacement field.
+We performed the experiment in two different realizations of the phantom, one at full resolution
+  and a second with half resolution to study the performance under strong partial volume effect
+  conditions.
+The distorted phantoms and visual results are presented in \autoref{fig:results_phantom}.
+Quantitative evaluation of the \gls*{swindex} is reported in \autoref{tab:swi_phantom}.
+
+\begin{table}
+	\caption{Quantitative evaluation on phantoms: \acrlong*{swindex}}
+	\label{tab:swi_phantom}
+    \begin{tabular}{llll}
+    Voxel size  & Interior surface & Exterior surface & Average \\
+    $1.0mm$ & $0.718mm$            & $0.724mm$            & $0.721mm$   \\
+    $2.0mm$ & $0.723mm$            & $0.726mm$            & $0.725mm$   \\
+    \end{tabular}
+\end{table}
+
+{\color{red}
+\subsubsection{Convergence evindence and settings}
+ discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition / Wolfe conditions, multi-resolutions on the bspline grid, image samples and surface sampling,  etc.
+\subsubsection{Efficient field densification}
+As long as the dense deformation field is iteratively interpolated
+from the same set of control points that define the $L-1$ contours,
+it is possible to pre-cache all the $\psi_k(\vec{v}_i)$ weights into a
+sparse matrix for fast densification. As well, we achieve a
+diffeomorphic transform by}
+
+% In regular intervals, after certain number of iterations,
+% the parameters $\Theta_l$ of the regions can be reestimated
+% based on the shifted volumetric samples
+% $\vec{x}' = \vec{x_0} + u(\vec{x})$.
+
+\begin{figure*}
+\includegraphics[width=\textwidth]{figures/figure03}
+\caption{Proof of concept of the registration method. First row shows the registration experiment
+  on the high-resolution ($1.0mm$ isometric) realization of the phantom, second row contains the
+  low-resolution ($2.0mm$ isometric) version.
+  Both phantoms were simulated with \gls*{snr} 30.
+  In the first column, the distorted phantoms presented with the undistorted contours.
+  For the second column, the contours have been replaced by the distorted contours using
+    the ground-truth warping $U^{-1}_{true}$.
+  The third column contains again the contours warped with the ground-truth field
+  (thin yellow line), and the corresponding contours (green and blue thick lines)
+  obtained with the presented method.
+  Finally, last row represents the recovered images unwarped using the $\hat{U}_{tst}$
+    found through our registration.
+  }\label{fig:results_phantom}
+\end{figure*}
+
+\subsection{Susceptibility distortion correction}
+\label{sec:results_hcp}
+
+As we propose the \gls*{fa} and the \gls*{md} as target features, we
+  implemented a simplified pipeline for processing \gls*{dmri} using
+  \emph{MRtrix} \citep{tournier_mrtrix_2012}.
+\todo[inline]{insert here the summary of parameters}
+
+Selecting the appropriate labels in the \emph{aparc} segmentation, we applied
+  the marching-cubes algorithm again to extract priors of the following
+  homogeneous regions in terms of \gls*{fa} and \gls*{md} joint distribution:
+\begin{itemize}
+	\item \gls*{csf} of the ventricular system,
+	\item deep \gls*{gm} structures,
+	\item \gls*{wm} surface as the ensemble of brain stem, and
+	  cerebellar and cerebral \gls*{wm},
+	\item cortical \gls*{gm} surface, including cerebellar \gls*{gm}.
+\end{itemize}
+
+\subsection{Discussion}
+\label{sec:discussion}
+We propose a registration-segmentation algorithm designed for biomedical
+  image analyses that are affected by all or some of the following circumstances:
+  \begin{itemize}
+  	\item Reference and target images do not present appropriate contrasts to
+  	perform volume-based registration reliably, or there is not available an apt image
+ 		to be used as reference.
+  	\item The low resolution of the target volume produces strong partial volume effects
+  	that exclude the use of voxel-wise segmentation algorithms and hinder volume-based
+  	registration.
+  	\item The unknown warping is smooth enough to be represented with a sufficient number
+  	of BSpline kernels, and its explicit need is justified (i.e. the distortion
+  	field in \gls*{dmri} images, object tracking applications, etc.).
+  \end{itemize}
+The algorithm is designed for the proposed application on correcting distortions of
+  \gls*{dmri} data, a required preliminary step to extract the structural connectivity
+  networks.
+As a consequence, our approach relies on the availability of precise segmentations or
+  surfaces of the objects that are to be segmented on the target volume(s).
+Connectome extraction protocols generally include the acquisition of \gls*{t1} images
+  to provide with prior information about anatomy with high accuracy.
+Our method solves in a single step a joint problem usually addressed in two steps.
+The first solves the linear registration of the \gls*{t1} and \gls*{dmri} data (i.e.
+  \cite{greve_accurate_2009}), whereas the second corrects for the nonlinear distortions
+  derived from the inhomogenety of magnetic susceptibility across tissues
+  \citep{jezzard_correction_1995}.
+
+We implemented the registration method upon the widely used Insight Registration and Segmentation
+	Toolkit (\url{http://www.itk.org}) with the aim to release a useful and free software bundle.
+The instrumentation framework is also distributed with the registration method,
+  and has been implemented using \emph{nipype} \citep{gorgolewski_nipype_2011}.
+The pipelines include the automatic generation of the ``cortex'' phantom.
+The real datasets are publicly available under the \gls*{hcp} project.
+Thus, all the experiments and results presented in this paper are intended to be
+  fully reproducible.

### 2014-11-06 Scripts/tools/run_evaluations.py
fix mask orientation. close #166

@@ -5,8 +5,8 @@
 #
 # @Author: oesteban - code@oscaresteban.es
 # @Date:   2014-04-04 19:39:38
-# @Last Modified by:   oesteban
-# @Last Modified time: 2014-11-03 17:13:15
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2014-11-06 20:16:01
 
 __author__ = "Oscar Esteban"
 __copyright__ = "Copyright 2013, Biomedical Image Technologies (BIT), \
@@ -131,6 +131,7 @@ def hcp_workflow(name='HCP_TMI2015', settings={}):
     ])
 
     dti = mrtrix_dti()
+    rlmsk = pe.Node(fs.MRIConvert(), name='MaskReslice')
     mdti = pe.Node(niu.Merge(2), name='MergeDTI')
     msk = pe.Node(niu.Select(index=[2]), name='SelectHIresMask')
 
@@ -148,17 +149,19 @@ def hcp_workflow(name='HCP_TMI2015', settings={}):
     regseg.inputs.inputnode.grid_size = [(6, 4, 6), 6, 8]
 
     wf.connect([
-        (st1,  dti,    [('out_dis_set.dwi', 'inputnode.in_dwi'),
-                        ('out_dis_set.dwi_mask', 'inputnode.in_mask')]),
-        (ds,   dti,    [('bvec', 'inputnode.in_bvec'),
-                        ('bval', 'inputnode.in_bval')]),
-        (dti,  mdti,   [('outputnode.fa', 'in1'),
-                        ('outputnode.md', 'in2')]),
-        (mdti, regseg, [('out', 'inputnode.in_fixed')]),
-        (st1,  regseg, [('out_dis_set.tpms', 'inputnode.in_tpms'),
-                        ('out_ref_set.surf', 'inputnode.in_surf')]),
-        (st1,  msk,    [('out_dis_set.segs', 'inlist')]),
-        (msk,  regseg, [('out', 'inputnode.in_mask')])
+        (st1,   dti,    [('out_dis_set.dwi', 'inputnode.in_dwi'),
+                         ('out_dis_set.dwi_mask', 'inputnode.in_mask')]),
+        (ds,    dti,    [('bvec', 'inputnode.in_bvec'),
+                         ('bval', 'inputnode.in_bval')]),
+        (dti,   mdti,   [('outputnode.fa', 'in1'),
[…]

### 2014-11-06 Code/Modules/Functional/include/FunctionalBase.h
improving masked energy

@@ -205,9 +205,9 @@ public:
 	typedef typename ProbabilityMapType::ConstPointer        ProbabilityMapConstPointer;
 	typedef std::vector< ProbabilityMapPointer >             ProbabilityMapList;
 
-	typedef itk::LinearInterpolateImageFunction
-			< ProbabilityMapType >                           ProbmapInterpolatorType;
-	typedef typename ProbmapInterpolatorType::Pointer        ProbmapInterpolatorPointer;
+	typedef itk::NearestNeighborInterpolateImageFunction
+			< ProbabilityMapType >                           MaskInterpolatorType;
+	typedef typename MaskInterpolatorType::Pointer           MaskInterpolatorPointer;
 	typedef itk::ResampleImageFilter
 			< ProbabilityMapType, ProbabilityMapType >       ProbmapResampleType;
 	typedef typename ProbmapResampleType::Pointer            ProbmapResamplePointer;
@@ -403,7 +403,7 @@ protected:
 
 
 	InterpolatorPointer m_Interp;
-	ProbmapInterpolatorPointer m_MaskInterp;
+	MaskInterpolatorPointer m_MaskInterp;
 	PointDataContainerPointer m_CurrentDisplacements;
 	PointsVector m_NodesPosition;
 

### 2015-02-16 Publications/2015-NeuroImage/03-introduction.tex
Change to Neuroimage layout

--- 2014-10-21 Publications/2013-TMI/03-introduction.tex
@@
   in an endless number of image analysis applications.
-The problem is solved finding a spatial mapping $U$ that aligns the
-  information given in two different coordinate systems (generally called
+The process aims to find a spatial mapping $U$ that aligns the information available
+  in two different coordinate systems (generally called
   \emph{reference}, $R$, and \emph{moving}, $M$):
@@
   \end{align}
-  where $\vec{r}$ denotes a position in the reference domain and $\vec{r}'$
-  its corresponding location in the target, and $n$ the dimension of domains
-  ($n=3$ for 3D images).
-Finally, $\vec{u} = u(\vec{r})$ is the function that models the displacement
-  of every point in the reference domain.
+  where $\vec{r}$ denotes a position in the reference domain $R$, $\vec{r}'$ is
+  its corresponding location in $M$, and $n$ the dimensionality of images.
+Finally, $\vec{u} = u(\vec{r})$ is the displacement of every point with respect
+  to the reference domain.
+A comprehensive survey by \citeauthor{sotiras_deformable_2013} \citep{sotiras_deformable_2013}
+  illustrates the wide variety of existing registration methodologies.
+They classify the algorithms by the three principal building blocks that can be identified
+  among methods: the matching criteria or similarity function to evaluate the proximity of
+  the solution, the deformation model that may implement theoretical properties and constraints
+  of the mapping in the selected application, and the optimization method or search strategy.
+The most widespread matching criteria are probably the so-called \emph{iconic
+  methods} \citep{sotiras_deformable_2013}.
+These algorithms require the definition of an appropriate function on the voxel-wise information
+  content of both images.
+A second group, called \emph{geometric methods}, measure distances between features derived from
+  corresponding objects defined in both spaces, such as landmarks, shapes, surfaces, etc.
+Finally, \emph{hybrid methods} combine properties of the iconic and geometric methods.
+As reviewed in \citep{sotiras_deformable_2013}, an important research effort has been devoted to
+  study application-oriented matching criteria.
 
-The wide range of applications has derived in an intensive generation of diverse
-  registration methods.
-In a recent and comprehensive survey by \citeauthor{sotiras_deformable_2013}
-  \citep{sotiras_deformable_2013}, registration methods are classified by
-  matching criteria, deformation model, and optimization method.
-The most widespread matching criteria are probably the so-called \emph{iconic
-  methods} in \citep{sotiras_deformable_2013}.
+Currently uprising applications, such as the extraction of connectivity networks from
+  diffusion \gls*{mri} \citep{craddock_imaging_2013}, demand for improved approaches
+  that outperform existing techniques.
+We propose the image registration -based solution to correct for the
+  susceptibility distortions in \gls*{dmri} of the brain ,
+  that is involved in the extraction of the structural connectivity map of the brain.
 However they are not appropriate for all the registration problems
@@
 
-\subsection{Related work}
-\label{sec:methods_background}
-Registration-segmentation methods are typically built including a deformation
-  model to guide the evolution of a segmentation algorithm.
-Even though some Bayesian approaches have been proposed to solve the joint problem
-  \citep{wyatt_map_2003}, segmentation is usually performed using an evolving
-  active contour \citep{chan_active_2001} that normally includes a shape prior
-  term to the energy functional \citep{chen_using_2002,bresson_variational_2006,
-  chan_level_2005,cremers_kernel_2006,gastaud_combining_2004}.
-These methods generally have an explicit description of the expected relative boundary
-  locations of the object to be delineated, and some, even model the statistical deviations
-  from this average shape.
-
-An early ``morphing''-segmentation approach was proposed in \citep{bertalmio_morphing_2000},
-  and the first proper registration method was presented by \citeauthor{yezzi_variational_2001}
-  \citep{yezzi_variational_2001}, where the energy functional is defined simultaneously in the target
-  and reference images with an affine transformation supporting the coordinates mapping
-  (the so-called ``2 \glspl*{pde} approach'').
-\citeauthor{vemuri_joint_2003} proposed the first atlas-based registration \citep{vemuri_joint_2003}
-  based on the level-sets framework and using only one \gls*{pde}.
-The principal application of these techniques was the image registration between timesteps in
-  a time-series or between different images \citep{paragios_level_2003}.
-\citeauthor{unal_coupled_2005} \citep{unal_coupled_2005} extended the idea of
-  \citep{bertalmio_morphing_2000,yezzi_variational_2001} to nonlinear registration
-  using a free deformation field model.
-A summary of these second set of techniques is performed in \citep{droske_mumfordshah_2009},
-  proposing two different approaches to applying the Mumford-Shah \citep{mumford_optimal_1989}
-  functional in joint registration and segmentation, thorugh the propagation of the deformation
-  ield from the contours to the whole image definition.
-Finally, \citeauthor{gorthi_active_2011} presented a comprehensive generalization of the
-  existing methodologies \citep{gorthi_active_2011} based on the level-sets approximation.
-Even though the historical evolution of joint segmentation and registration procedures is strongly
-  linked to the level sets approximation, our work is based in derivations of the \gls{acwe}
-  framework \citep{chan_active_2001}, and related to \citep{guyader_combined_2011}.
-
-Finally, a derivation of the latter group is composed by atlas-based segmentation methods
-  \citep{gorthi_segmentation_2009,gorthi_active_2011,pohl_unifying_2005,pohl_bayesian_2006,
-  wang_joint_2006}, where the prior imposes consistent voxel-based classification of contiguous
-  regions.
-A comprehensive summary of the methodologies derived from the \gls{acwe} model, with special attention
-  to joint registration-segmentation methods is found in \citep{gorthi_active_2011}.
+The hypothesis underpinning this research is that, in those situations,
+  image registration can be reliably performed by searching for homogeneous
+  regions in one image corresponding to a prior knowledge of the shape of objects in
+  the moving coordinate system.
 
@@
 %as we shall cover in \autoref{sec:numerical_implementation}.}}
-
-\subsection{Susceptibility-derived distortions}
-
-\subsection{Our framework}
-\label{sec:our_framework}
 

### 2015-02-16 Publications/2015-NeuroImage/04-methods.tex
Change to Neuroimage layout

--- 2014-11-06 Publications/2013-TMI/04-methods.tex
@@
 %
+\subsection{Related work}
+\label{sec:methods_background}
+Registration-segmentation methods are typically built including a deformation
+  model to guide the evolution of a segmentation algorithm.
+Even though some Bayesian approaches have been proposed to solve the joint problem
+  \citep{wyatt_map_2003}, segmentation is usually performed using an evolving
+  active contour \citep{chan_active_2001} that normally includes a shape prior
+  term to the energy functional \citep{chen_using_2002,bresson_variational_2006,
+  chan_level_2005,cremers_kernel_2006,gastaud_combining_2004}.
+These methods generally have an explicit description of the expected relative boundary
+  locations of the object to be delineated, and some, even model the statistical deviations
+  from this average shape.
+
+An early ``morphing''-segmentation approach was proposed in \citep{bertalmio_morphing_2000},
+  and the first proper registration method was presented by \citeauthor{yezzi_variational_2001}
+  \citep{yezzi_variational_2001}, where the energy functional is defined simultaneously in the target
+  and reference images with an affine transformation supporting the coordinates mapping
+  (the so-called ``2 \glspl*{pde} approach'').
+\citeauthor{vemuri_joint_2003} proposed the first atlas-based registration \citep{vemuri_joint_2003}
+  based on the level-sets framework and using only one \gls*{pde}.
+The principal application of these techniques was the image registration between timesteps in
+  a time-series or between different images \citep{paragios_level_2003}.
+\citeauthor{unal_coupled_2005} \citep{unal_coupled_2005} extended the idea of
+  \citep{bertalmio_morphing_2000,yezzi_variational_2001} to nonlinear registration
+  using a free deformation field model.
+A summary of these second set of techniques is performed in \citep{droske_mumfordshah_2009},
+  proposing two different approaches to applying the Mumford-Shah \citep{mumford_optimal_1989}
+  functional in joint registration and segmentation, through the propagation of the deformation
+  field from the contours to the whole image definition.
+Finally, \citeauthor{gorthi_active_2011} presented a comprehensive generalization of the
+  existing methodologies \citep{gorthi_active_2011} based on the level-sets approximation.
+Even though the historical evolution of joint segmentation and registration procedures is strongly
+  linked to the level sets approximation, our work is based in derivations of the \acrlong{acwe}
+  framework \citep{chan_active_2001}, and related to \citep{guyader_combined_2011}.
+
+Finally, a derivation of the latter group is composed by atlas-based segmentation methods
+  \citep{gorthi_segmentation_2009,gorthi_active_2011,pohl_unifying_2005,pohl_bayesian_2006,
+  wang_joint_2006}, where the prior imposes consistent voxel-based classification of contiguous
+  regions.
+A comprehensive summary of the methodologies derived from the \acrlong{acwe} model, with special attention
+  to joint registration-segmentation methods is found in \citep{gorthi_active_2011}.
 

### 2015-02-16 Publications/2015-NeuroImage/07-conclusion.tex
Change to Neuroimage layout

% -*- root: 00-main.tex -*-
\section{Conclusion}
\label{sec:conclusion}
%
We propose and make publicly available a joint
  segmentation-registration method with a proof of concept
  on a digital phantom and an application on diffusion \gls*{mri}
  images of the human brain.
Starting with the Bayesian approach to image segmentation, we derive
  a cost function that models piecewise smooth regions with
  multivariate normal distributions.
This is, implicitly, an \acrlong*{acwe} model, in which the
  evolution of the surfaces drives the optimization of a B-Spline
  deformation model.

The phantom is used to demonstrate the success of the approach,
  on a quantitative evaluation that demonstrated its performance
  with sub-voxel precision (average vertex displacement close to
  $0.7mm$ in phantoms with isometric pixels of $1.0mm$ and $2.0mm$
  side).
We apply the method on the correction of susceptibility-derived
  distortions of diffusion \gls*{mri} data, and compare to
  a widely-used method based on the acquisition of extra images
  mapping the inhomogeneity of tissue susceptibility inside
  the scanner.

### 2015-02-16 Publications/2015-NeuroImage/08-appendix-00-shapepriors.tex
Change to Neuroimage layout

% -*- root: 00-main.tex -*-

\section{Extended formulations}

\subsection{Shape priors}
\label{app:shape_priors}
Let $F(\vec{r})$ be an ``arbitrary'' function over the image domain
  $\Omega = \Omega_l \cup \Omega_m$ splitted in two regions $l$ and
  $m$, and $\Gamma_{l,m}$ a closed boundary between them
  as defined in \autoref{sec:methods_map}.
We now derive the domain integral w.r.t. $t$:
  \begin{align*}
  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) d\vec{r} &=
  \int_\Omega \frac{\partial}{\partial t}F(\vec{r}) d\vec{r} \notag\\
  &- \int_{\Gamma_{l,m}} F(\vec{r}) \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
  N_{\Gamma_{l,m}}\right\rangle d\vec{r},
  \end{align*}
  where $\left\langle\frac{\partial\Gamma_{l,m}}{\partial t}, N_{\Gamma_{l,m}}\right\rangle$ is
  the projection of the boundary movement on the unit inward normal $N_{\Gamma_{l,m}}$.
Assuming that the region descriptors $\Theta_l$ vary slowly enough, we can consider
  that $\frac{\partial}{\partial t} F(\vec{r}) = 0$ and thus:
  \begin{equation*}
  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) d\vec{r} =
  - \int_{\Gamma_{l,m}} F(\vec{r}) \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
  N_{\Gamma_{l,m}}\right\rangle d\vec{r}
  \end{equation*}

  In our formulation, $F(\vec{r})$ is the difference on
    sample likelihood between the inner ($\Omega_l$) and the outer ($\Omega_m$) regions.
  Therefore, we substitute:
    \begin{equation}
    \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) d\vec{r} =
    - \int_{\Gamma_{l,m}} \left[ \mathcal{E}_{l}(\vec{r}) - \mathcal{E}_{m}(\vec{r}) \right]
    \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
  N_{\Gamma_{l,m}}\right\rangle d\vec{r},
  	\label{eq:shape_gradients}
    \end{equation}
    where $\mathcal{E}_l(\vec{r}) = \mathcal{D}^2_{l}(R(\vec{r}))$ for briefness.

### 2015-02-16 Publications/2015-NeuroImage/00-main.tex
update layout. close #224

%\documentclass[numbers,sort&compress]{elsarticle}
\documentclass[5p,numbers,sort&compress]{elsarticle}
\journal{NeuroImage}
% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
\usepackage{glossaries}
\usepackage{booktabs} % to use \toprule

\usepackage[colorinlistoftodos]{todonotes}

%\usepackage[cmex10,intlimits]{amsmath}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{algorithmic}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\argmax}{argmax}
\DeclareMathOperator{\argmin}{argmin}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\const}{Const.}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
\providecommand{\mdist}[2]{ \mathcal{D}_{#2}^2(\mathbf{#1}) }
\let\oldhat\hat
\renewcommand{\vec}[1]{\mathbf{#1}}
\providecommand{\nvec}[1]{\hat{\mathbf{#1}}}

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
\usepackage[hidelinks]{hyperref}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\input{01-acronyms}

\begin{document}

\begin{frontmatter}

%% Use \dochead if there is an article header, e.g. \dochead{Short communication}
% \dochead{Preprint for ArXiv archiving and pre-submission review}

\title{Simultaneous segmentation-registration using active contours with application on
diffusion MR images}

\author[bit]{Oscar~Esteban\corref{corrauthor}}
\author[lts5]{Alessandro~Daducci}
\author[chuv,lts5]{Meritxell~Bach-Cuadra}
\author[lts5]{Jean-Philippe~Thiran}
\author[bit]{Andr\'es~Santos}
\author[bit]{Mar\'ia-J.~Ledesma-Carbayo}
\author[ucla]{Dominique~Zosso}
\cortext[corrauthor]{Corresponding author}
\ead{phd@oscaresteban.es}

\address[bit]{Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on, %
Universidad Polit\'ecnica de Madrid and CIBER-BBN, Madrid, Spain}
\address[lts5]{Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland}
\address[chuv]{Dept. of Radiology, University
Hospital Center (CHUV) and University of Lausanne (UNIL), Lausanne, Switzerland}
\address[ucla]{Department of Mathematics, University of California,
Los Angeles (UCLA), Los Angeles, CA, US}

% \author{Oscar~Esteban$^{*}$,
%         Alessandro~Daducci,
%         Meritxell~Bach-Cuadra,
% %        Jean-Philippe~Thiran,
% %        Andr\'es~Santos,
%         and~Dominique~Zosso% <-this % stops a space
% \thanks{Manuscript received XXX XX, 2013; revised XXX XX, 2013.
% This study is supported by: the Spain's Ministry of Science and Innovation
% (projects TEC2010-21619- C04-03, TEC2011-28972-C02-02, CDTI-CENIT
% AMIT and INNPACTO PRECISION), Comunidad de Madrid (ARTEMIS) and
% European Regional Development Funds; the Center for Biomedical Imaging
% (CIBM) of the Geneva and Lausanne Universities and the EPFL, as well as the
% Leenaards and Louis Jeantet foundations.\emph{~Asterisk indicates corresponding
% author}.}%
% \thanks{$^{*}$O. Esteban is with the Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on,
% Universidad Polit\'ecnica de Madrid and CIBER-BBN, Madrid, Spain
% (e-mail: phd@oscaresteban.es).}% <-this % stops a space
% \thanks{A. Daducci is with the Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
% F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland.}%
% \thanks{M. Bach-Cuadra is with the Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
% F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland, and Dept. of Radiology, University
% Hospital Center (CHUV) and University of Lausanne (UNIL), Lausanne, Switzerland.}%
% %\thanks{A. Santos is with the Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on,
% %Universidad Polit\'ecnica de Madrid and CIBER-BBN, Madrid, Spain.}% <-this % stops a space
% \thanks{D. Zosso is with the Department of Mathematics, University of California,
% Los Angeles (UCLA), Los Angeles, CA, US and is supported by the Swiss National
% Science Foundation (SNF) under grant PBELP2-137727.}% <-this % stops a space
% }

\begin{abstract}
\input{02-abstract}
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{keyword}
diffusion MRI \sep susceptibility~distortion \sep segmentation %
\sep registration \sep parcellation \sep shape-prior.
\end{keyword}

\end{frontmatter}

\input{03-introduction}
\input{04-methods}
%\input{05-experiments}
\input{06-results}

\input{07-conclusion}

% Add appendices
\appendix
%\numberwithin{equation}{section}
\input{08-appendix-00-shapepriors}

% ACKs
\input{08-acknowledgements}

\bibliographystyle{myplainnat}
\bibliography{Remote}
\end{document}

### 2015-02-16 Publications/2015-NeuroImage/08-acknowledgements.tex
update acks #224

% -*- root: 00-main.tex -*-
% use section* for acknowledgement
\section*{Acknowledgment}
The authors thank V. Estellers for early discussions at the beginning of this project,
  and L. Vese for her support during OE's research visits in her laboratory.

DZ is supported by the Swiss National Science Foundation (SNF) under grant PBELP2-137727.

This study is supported by: the Spain's Ministry of Science and Innovation
(projects TEC2010-21619- C04-03, TEC2011-28972-C02-02, CDTI-CENIT
AMIT and INNPACTO PRECISION), Comunidad de Madrid (ARTEMIS) and
European Regional Development Funds; the Center for Biomedical Imaging
(CIBM) of the Geneva and Lausanne Universities and the EPFL, as well as the
Leenaards and Louis Jeantet foundations.

### 2015-02-16 Publications/2015-NeuroImage/03-introduction.tex
update layout. close #224
update intro

@@
 
-\IEEEPARstart{N}{onlinear} registration of images is a challenging task ubiquitous
+Nonlinear registration of images is a challenging task, ubiquitous
   in an endless number of image analysis applications.
@@
   of the mapping in the selected application, and the optimization method or search strategy.
-The most widespread matching criteria are probably the so-called \emph{iconic
-  methods} \citep{sotiras_deformable_2013}.
+The possibly most widespread matching criteria are those classified as \emph{iconic
+  methods} in \citep{sotiras_deformable_2013}.
 These algorithms require the definition of an appropriate function on the voxel-wise information
@@
 
-Currently uprising applications, such as the extraction of connectivity networks from
-  diffusion \gls*{mri} \citep{craddock_imaging_2013}, demand for improved approaches
-  that outperform existing techniques.
-We propose the image registration -based solution to correct for the
-  susceptibility distortions in \gls*{dmri} of the brain ,
-  that is involved in the extraction of the structural connectivity map of the brain.
-However they are not appropriate for all the registration problems
-  \citep{greve_accurate_2009}.
-They propose a \emph{hybrid method} \citep{sotiras_deformable_2013} for the
-  intra-subject registration of \gls*{t1} \gls*{mri} and \gls*{dmri} images
+In this work we propose a hybrid approach to nonlinear registration of anatomically correct
+  and geometrically precise surfaces to low-resolution and distorted \gls*{dmri} of the brain.
+Currently, we are witnessing an uprising demand for such algorithms in the challenging context
+  of extracting and characterizing the structural connectivity networks of the
+  human brain \citep{craddock_imaging_2013}.
+Our registration method is demonstrated here as a solution to correct for the susceptibility
+  distortions that \gls*{dmri} of the head typically present \cite{jezzard_correction_1995}.
+A fundamental antecedent of this approach is found in \citep{greve_accurate_2009}.
+They proposed a registration method to align \gls*{t1} and \gls*{dmri} images
   of the brain using a linear transformation model.
-As in \citep{greve_accurate_2009}, we exploit the precise structures that
-  can be readily extracted with widely-used tools at high resolution
-  and map this prior information to low resolution data,
-  in which the scale of the imaged structures is in the range of the image
-  resolution or above.
-Our approach differs with \citep{greve_accurate_2009} in two fundamental
-  choices: 1) We use a nonlinear deformation model that enables the
-  correction of the susceptibility-derived distortions
-  \citep{jezzard_correction_1995} that \gls*{dmri} data typically present.
-And 2) our matching criteria does not necessarily need for clear intensity
-  steps at the tissue interfaces where surfaces must be fitted.
-
-The hypothesis underpinning this research is that, in those situations,
+As in \citep{greve_accurate_2009}, we exploit the precise structures that can be extracted with
+  widely-used software (i.e. \emph{FreeSurfer}~\citep{fischl_freesurfer_2012}) at high resolution,
+  and then map this prior information to low resolution data, in which the scale of the imaged
+  structures is in the range of the image resolution or above.
+Our approach differs from \citep{greve_accurate_2009} in two fundamental choices.
+First, we use a nonlinear deformation model that enables the correction of the
+  susceptibility-derived distortions \citep{jezzard_correction_1995}.
+Second, our matching criteria does not necessarily need for clear intensity
+  steps at the tissue interfaces where surfaces must be fitted, thanks to the use of
+  active contours without edges \cite{chan_active_2001}.
+The hypothesis underpinning our research is that, in those situations,
   image registration can be reliably performed by searching for homogeneous
@@
   the moving coordinate system.
-
-%{\color{red} {The main diverging points with respect to
-%\citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
-%$\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
-%gradients \citep{besson_dream2s_2003,herbulot_segmentation_2006};
-%2) regularization is also based on linear diffusion smoothing \citep{thirion_image_1998},
-%but we replace the Gaussian filtering by other constraints
-%studied in \citep{nagel_investigation_1986} to better the problem;
-%3) optimization is applied in the spectral
-%domain, observing anisotropic and inhomogeneous mappings along each direction.
-%With respect \citep{guyader_combined_2011}, the main differences are
-%the distance function, and the spectral solution to the optimization updates,
-%as we shall cover in \autoref{sec:numerical_implementation}.}}
-
-%In this paper we formulate the joint registration-segmentation
-%problem as follows.
-%such that the known contours in anatomical space $T$ optimally segment
-%the diffusion space $D$.
-%
-%
-%Whereas related \glspl*{adf} introduced in \autoref{sec:methods_background}
-%make use of explicit level-set formulations to solve \eqref{eq:gradient_descent},
-%we alternatively use \emph{shape-gradients}
-%\cite{besson_dream2s_2003,herbulot_segmentation_2006}.
-
-%
-%Let us denote by $\vec{x}$ the voxel and $F(\vec{x}) = [ f_1, f_2, \ldots, f_N]^T$
-%  its associated feature vector in the following.
-%
-%In our application, these surfaces are precise tissue interfaces of interest extracted
-%  from a high-resolution, anatomically correct reference volume using
-%  the well-established
-%
-%In this paper we propose a novel registration framework to simultaneously
-%solving the segmentation, distortion and cortical parcellation challenges,
-%by exploiting as strong shape-prior the detailed morphology extracted
-%from high-resolution and anatomically correct \gls{mri}.
-%Indeed, hereafter
-%we assume this segmentation problem in anatomical images is reliably and
-%accurately solved with readily available tools (e.g.
-%\citep{fischl_freesurfer_2012}).
-%After global alignment with \gls{t1} using existing approaches, the remaining
-%spatial mismatch between anatomical and diffusion space is due to susceptibility
-%distortions.
-%Finally, we need to establish precise spatial correspondence between the
-%surfaces in both spaces, including the tangential direction for parcellation.
-%Therefore, we can reduce the problem to finding the differences of spatial
-%distortion in between anatomical and \gls{dwi} space.
-%We thus reformulate the segmentation problem as an inverse problem, where we
-%seek for an underlying deformation field (the distortion) mapping
-%from the structural space into the diffusion space, such that the structural
-%contours segment optimally the \gls{dwi} data.
-%In the process, the one-to-one
-%correspondence between the contours in both spaces is guaranteed, and projection
-%of parcellisation to \gls{dwi} space is implicit and consistent.
-%
-%We test our proposed joint segmentation-registration model on two different
-%synthetic examples.
-%The first example is a scalar sulcus model, where the
-%\gls{csf}-\gls{gm} boundary particularly suffers from \gls{pve} and can only be
-%segmented correctly thanks to the shape prior and its coupling with the inner,
-%\gls{gm}-\gls{wm} boundary through the imposed deformation field regularity.
-%The second case deals with more realistic \gls{dwi} data stemming from
-%phantom simulations of a simplistic brain data.
-%Again, we show that the
-%proposed model successfully segments the \gls{dwi} data based on two derived
-%scalar features, namely \gls{fa} and \gls{md}, while establishing an estimate
-%of the dense distortion field.
-%
-%The rest of this paper is organized as follows.
-%First, in \autoref{sec:methods}
-%we introduce our proposed model for joint multivariate segmentation-registration.
-%Then we provide a more detailed description of the data and experimental setup in
-%\autoref{sec:experiments}.
-%We present results in \autoref{sec:results} and conclude
-%in \autoref{sec:conclusion}.
-%
-%
-%The properties of the reconstructed tensors and derived scalar maps have
-%been studied by \cite{ennis_orthogonal_2006}. Based on their
-%findings, \gls{fa}~\eqref{eq:fa} and \gls{md}~\eqref{eq:md} are
-%considered complementary features, and therefore we selected them for the
-%energy model \eqref{eq:complete_energy} in driving the
-%registration-segmentation process.
-%Whereas \gls{fa} informs mainly about the \emph{shape} of diffusion,
-%the \gls{md} is more related to the \emph{magnitude} of the process:
-%
-%\begin{align}
-%\mathrm{FA} &= \sqrt{ \frac{1}{2}}\,\frac{\sqrt{ (\lambda_1 - \lambda_2)^2 + (\lambda_2 - \lambda_3)^2 + (\lambda_3 - \lambda_1)^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}} \label{eq:fa} \\
-%\mathrm{MD} &= ( \lambda_1 + \lambda_2 + \lambda_3 ) / 3 \label{eq:md}
-%\end{align}
-%where $\lambda_i$ are the eigenvalues of the diffusion tensor
-%associated with the diffusion signal $S(\vec{q})$. There exist
-%two main reasons to justify their choice.
-%First, they are well-understood and standardized in clinical routine.
-%Second, together they contain most of the information that is
-%usually extracted from the \gls{dwi}-derived scalar maps
-%\cite{ennis_orthogonal_2006}.
-%

### 2015-02-16 Publications/2015-NeuroImage/04-methods.tex
update intro

@@
 % -*- root: 00-main.tex -*-
+
+%{\color{red} {The main diverging points with respect to
+%\citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
+%$\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
+%gradients \citep{besson_dream2s_2003,herbulot_segmentation_2006};
+%2) regularization is also based on linear diffusion smoothing \citep{thirion_image_1998},
+%but we replace the Gaussian filtering by other constraints
+%studied in \citep{nagel_investigation_1986} to better the problem;
+%3) optimization is applied in the spectral
+%domain, observing anisotropic and inhomogeneous mappings along each direction.
+%With respect \citep{guyader_combined_2011}, the main differences are
+%the distance function, and the spectral solution to the optimization updates,
+%as we shall cover in \autoref{sec:numerical_implementation}.}}
+
+%In this paper we formulate the joint registration-segmentation
+%problem as follows.
+%such that the known contours in anatomical space $T$ optimally segment
+%the diffusion space $D$.
+%
+%
+%Whereas related \glspl*{adf} introduced in \autoref{sec:methods_background}
+%make use of explicit level-set formulations to solve \eqref{eq:gradient_descent},
+%we alternatively use \emph{shape-gradients}
+%\cite{besson_dream2s_2003,herbulot_segmentation_2006}.
+
+%
+%Let us denote by $\vec{x}$ the voxel and $F(\vec{x}) = [ f_1, f_2, \ldots, f_N]^T$
+%  its associated feature vector in the following.
+%
+%In our application, these surfaces are precise tissue interfaces of interest extracted
+%  from a high-resolution, anatomically correct reference volume using
+%  the well-established
+%
+%In this paper we propose a novel registration framework to simultaneously
+%solving the segmentation, distortion and cortical parcellation challenges,
+%by exploiting as strong shape-prior the detailed morphology extracted
+%from high-resolution and anatomically correct \gls{mri}.
+%Indeed, hereafter
+%we assume this segmentation problem in anatomical images is reliably and
+%accurately solved with readily available tools (e.g.
+%\citep{fischl_freesurfer_2012}).
+%After global alignment with \gls{t1} using existing approaches, the remaining
+%spatial mismatch between anatomical and diffusion space is due to susceptibility
+%distortions.
+%Finally, we need to establish precise spatial correspondence between the
+%surfaces in both spaces, including the tangential direction for parcellation.
+%Therefore, we can reduce the problem to finding the differences of spatial
+%distortion in between anatomical and \gls{dwi} space.
+%We thus reformulate the segmentation problem as an inverse problem, where we
+%seek for an underlying deformation field (the distortion) mapping
+%from the structural space into the diffusion space, such that the structural
+%contours segment optimally the \gls{dwi} data.
+%In the process, the one-to-one
+%correspondence between the contours in both spaces is guaranteed, and projection
+%of parcellisation to \gls{dwi} space is implicit and consistent.
+%
+%We test our proposed joint segmentation-registration model on two different
+%synthetic examples.
+%The first example is a scalar sulcus model, where the
+%\gls{csf}-\gls{gm} boundary particularly suffers from \gls{pve} and can only be
+%segmented correctly thanks to the shape prior and its coupling with the inner,
+%\gls{gm}-\gls{wm} boundary through the imposed deformation field regularity.
+%The second case deals with more realistic \gls{dwi} data stemming from
+%phantom simulations of a simplistic brain data.
+%Again, we show that the
+%proposed model successfully segments the \gls{dwi} data based on two derived
+%scalar features, namely \gls{fa} and \gls{md}, while establishing an estimate
+%of the dense distortion field.
+%
+%The rest of this paper is organized as follows.
+%First, in \autoref{sec:methods}
+%we introduce our proposed model for joint multivariate segmentation-registration.
+%Then we provide a more detailed description of the data and experimental setup in
+%\autoref{sec:experiments}.
+%We present results in \autoref{sec:results} and conclude
+%in \autoref{sec:conclusion}.
+%
+%
+%The properties of the reconstructed tensors and derived scalar maps have
+%been studied by \cite{ennis_orthogonal_2006}. Based on their
+%findings, \gls{fa}~\eqref{eq:fa} and \gls{md}~\eqref{eq:md} are
+%considered complementary features, and therefore we selected them for the
+%energy model \eqref{eq:complete_energy} in driving the
+%registration-segmentation process.
+%Whereas \gls{fa} informs mainly about the \emph{shape} of diffusion,
+%the \gls{md} is more related to the \emph{magnitude} of the process:
+%
+%\begin{align}
+%\mathrm{FA} &= \sqrt{ \frac{1}{2}}\,\frac{\sqrt{ (\lambda_1 - \lambda_2)^2 + (\lambda_2 - \lambda_3)^2 + (\lambda_3 - \lambda_1)^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}} \label{eq:fa} \\
+%\mathrm{MD} &= ( \lambda_1 + \lambda_2 + \lambda_3 ) / 3 \label{eq:md}
+%\end{align}
+%where $\lambda_i$ are the eigenvalues of the diffusion tensor
+%associated with the diffusion signal $S(\vec{q})$. There exist
+%two main reasons to justify their choice.
+%First, they are well-understood and standardized in clinical routine.
+%Second, together they contain most of the information that is
+%usually extracted from the \gls{dwi}-derived scalar maps
+%\cite{ennis_orthogonal_2006}.
+%
+
 \section{Methods}
@@
   model to guide the evolution of a segmentation algorithm.
-Even though some Bayesian approaches have been proposed to solve the joint problem
-  \citep{wyatt_map_2003}, segmentation is usually performed using an evolving
-  active contour \citep{chan_active_2001} that normally includes a shape prior
+Even though some Bayesian approaches have been proposed to jointly solve the problem
+  \citep{wyatt_map_2003,gass_simultaneous_2014}, segmentation is usually performed using
+  an evolving active contour \citep{chan_active_2001} that normally includes a shape prior
   term to the energy functional \citep{chen_using_2002,bresson_variational_2006,
   chan_level_2005,cremers_kernel_2006,gastaud_combining_2004}.
-These methods generally have an explicit description of the expected relative boundary
-  locations of the object to be delineated, and some, even model the statistical deviations
-  from this average shape.
+These methods generally have an inplicit description of the relative boundary locations of the objects
+  to be delineated using level sets, and some even model the statistical deviations from original
+  shape.
 
@@
 A summary of these second set of techniques is performed in \citep{droske_mumfordshah_2009},
-  proposing two different approaches to applying the Mumford-Shah \citep{mumford_optimal_1989}
-  functional in joint registration and segmentation, through the propagation of the deformation
-  field from the contours to the whole image definition.
+  proposing two different approaches to applying the Mumford-Shah functiona
+  \citep{mumford_optimal_1989} in simultaneous registration-segmentation, through the propagation
+  of the deformation field from the contours to the whole image definition.
 Finally, \citeauthor{gorthi_active_2011} presented a comprehensive generalization of the
@@
 Even though the historical evolution of joint segmentation and registration procedures is strongly
-  linked to the level sets approximation, our work is based in derivations of the \acrlong{acwe}
-  framework \citep{chan_active_2001}, and related to \citep{guyader_combined_2011}.
+  linked to the level sets approach, our method conversely used the explicit representation of
+  the evolving surfaces.
 
@@
   to joint registration-segmentation methods is found in \citep{gorthi_active_2011}.
+is based in derivations of the \acrlong{acwe}
+  framework \citep{chan_active_2001}, and related to \citep{guyader_combined_2011}.
 

### 2015-02-16 Scripts/pyacwereg/workflows/registration.py
base code for #221

@@ -5,8 +5,8 @@
 #
 # @Author: oesteban - code@oscaresteban.es
 # @Date:   2014-03-28 20:38:30
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-02-13 13:19:43
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-02-16 13:43:08
 
 import os
 import os.path as op
@@ -108,6 +108,191 @@ def default_regseg(name='REGSEGDefault'):
     wf.inputs.inputnode.f_smooth = [None, None]
     wf.inputs.inputnode.images_verbosity = 1
     wf.inputs.inputnode.convergence_value = [1.0e-6, 1.0e-8]
+    return wf
+
+
+def t2b_workflow(name='T2B', minimal=False, enc_dir=None,
+                 enhance_b0=True, icorr=True):
+    """
+    The T2w-registration based method (T2B) implements an SDC by nonlinear
+    registration of the anatomically correct *T2w* image to the *b0* image
+    of the *dMRI* dataset. The implementation here tries to reproduce the one
+    included in ExploreDTI `(Leemans et al., 2009)
+    <http://www.exploredti.com/ref/ExploreDTI_ISMRM_2009.pdf>`_, which is
+    also based on `(Irfanoglu et al., 2012)
+    <http://dx.doi.org/10.1016/j.neuroimage.2012.02.054>`_.
+
+    :param str name: a unique name for the workflow.
+
+    :inputs:
+
+        * in_t2w: the reference T2w image
+
+    :outputs:
+
+        * outputnode.corrected_image: the dMRI image after correction
+
+
+    Example::
+
+    >>> t2b = t2b_workflow()
+    >>> t2b.inputs.inputnode.in_dwi = 'dwi_brain.nii'
+    >>> t2b.inputs.inputnode.in_bval = 'dwi.bval'
+    >>> t2b.inputs.inputnode.in_mask = 'b0_mask.nii'
+    >>> t2b.inputs.inputnode.in_t2w = 't2w_brain.nii'
+    >>> t2b.run() # doctest: +SKIP
+
+    """
+
+    wf = SDCWorkflow(name=name)
+
+    def _default_params(enc_dir):
+        import pysdcev.data as data
+        if len(enc_dir) == 2:
+            enc_dir = enc_dir[0]
+        return data.get('t2b_params')[enc_dir]
+
+    def _get_last(inlist):
[…]

### 2015-02-16 Scripts/pyacwereg/workflows/registration.py
add inputnode

@@ -6,7 +6,7 @@
 # @Author: oesteban - code@oscaresteban.es
 # @Date:   2014-03-28 20:38:30
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-02-16 13:43:08
+# @Last Modified time: 2015-02-16 14:47:13
 
 import os
 import os.path as op
@@ -143,28 +143,17 @@ def t2b_workflow(name='T2B', minimal=False, enc_dir=None,
     >>> t2b.run() # doctest: +SKIP
 
     """
+    inputnode = pe.Node(niu.IdentityInterface(
+        fields=['in_dwi', 'in_bval', 'in_t2w', 'dwi_mask', 't2w_mask',
+                'in_param']), name='inputnode')
 
-    wf = SDCWorkflow(name=name)
-
-    def _default_params(enc_dir):
-        import pysdcev.data as data
-        if len(enc_dir) == 2:
-            enc_dir = enc_dir[0]
-        return data.get('t2b_params')[enc_dir]
-
-    def _get_last(inlist):
-        return inlist[-1]
-
-    selbmsk = pe.Node(niu.Select(index=2), name='SelectT2msk')
     avg_b0 = pe.Node(pmisc.ComputeAveragedB0(), name='avg_b0')
     cache_b0 = pe.Node(niu.IdentityInterface(fields=['b0', 'mask']),
                        name='B0Cache')
-    t2_msk = pe.Node(fs.Binarize(min=200), name='T2mask')
     enh_t2 = pe.Node(pmisc.SigmoidFilter(upper_perc=92.0, lower_perc=65.0),
                      name='enh_T2')
-    reg_param = pe.Node(niu.Function(input_names=['enc_dir'],
-                                     output_names=['out_file'],
-                                     function=_default_params), name='T2BSettings')
+    reg_param = pe.Node(nii.JSONFileGrabber(defaults={'enc_dir': 'y'}),
+                        name='T2BSettings')
     reg = pe.Node(elastix.Registration(num_threads=1), name='Elastix')
     tfx_b0 = pe.Node(elastix.EditTransform(), name='tfm_b0')
     split_dwi = pe.Node(fsl.utils.Split(dimension='t'), name='split_dwi')
@@ -177,20 +166,22 @@ def t2b_workflow(name='T2B', minimal=False, enc_dir=None,
     thres = pe.MapNode(Threshold(thresh=0.0), iterfield=['in_file'],
                        name='RemoveNegs')
[…]

### 2015-02-17 Publications/2015-NeuroImage/03-introduction.tex
Update all the sections

@@
 % -*- root: 00-main.tex -*-
-\section{Introduction}
+\section*{Introduction}
 \label{sec:introduction}
@@
   to the reference domain.
-A comprehensive survey by \citeauthor{sotiras_deformable_2013} \citep{sotiras_deformable_2013}
+A comprehensive survey by \cite{sotiras_deformable_2013}
   illustrates the wide variety of existing registration methodologies.
@@
   content of both images.
-A second group, called \emph{geometric methods}, measure distances between features derived from
+A second group, called \emph{geometric methods}, measures distances between features derived from
   corresponding objects defined in both spaces, such as landmarks, shapes, surfaces, etc.
@@
 In this work we propose a hybrid approach to nonlinear registration of anatomically correct
-  and geometrically precise surfaces to low-resolution and distorted \gls*{dmri} of the brain.
+  and precise surfaces to warped and low-resolution \gls*{dmri} of the brain.
 Currently, we are witnessing an uprising demand for such algorithms in the challenging context
@@
 Our registration method is demonstrated here as a solution to correct for the susceptibility
-  distortions that \gls*{dmri} of the head typically present \cite{jezzard_correction_1995}.
-A fundamental antecedent of this approach is found in \citep{greve_accurate_2009}.
+  distortions that \gls*{dmri} of the head typically present \citep{jezzard_correction_1995}.
+\todo[inline]{talk about the distortion and cite SDCs}
+
+A fundamental antecedent of the present approach is found in \citep{greve_accurate_2009}.
 They proposed a registration method to align \gls*{t1} and \gls*{dmri} images
   of the brain using a linear transformation model.
-As in \citep{greve_accurate_2009}, we exploit the precise structures that can be extracted with
-  widely-used software (i.e. \emph{FreeSurfer}~\citep{fischl_freesurfer_2012}) at high resolution,
-  and then map this prior information to low resolution data, in which the scale of the imaged
-  structures is in the range of the image resolution or above.
+Similarly, we exploit the precise structures that can be extracted with
+  widely-used software (i.e. \emph{FreeSurfer}~\citep{fischl_freesurfer_2012}),
+  and then map these surfaces into warped data, in which the scale of the
+  imaged structures is in the range of the image resolution or above.
 Our approach differs from \citep{greve_accurate_2009} in two fundamental choices.

### 2015-02-17 Publications/2015-NeuroImage/04-methods.tex
Update all the sections

@@
 
-\section{Methods}
+\section*{Methods}
 \label{sec:methods}
 %
-\subsection{Related work}
+\subsection*{Related work}
 \label{sec:methods_background}
-Registration-segmentation methods are typically built including a deformation
-  model to guide the evolution of a segmentation algorithm.
-Even though some Bayesian approaches have been proposed to jointly solve the problem
-  \citep{wyatt_map_2003,gass_simultaneous_2014}, segmentation is usually performed using
-  an evolving active contour \citep{chan_active_2001} that normally includes a shape prior
-  term to the energy functional \citep{chen_using_2002,bresson_variational_2006,
-  chan_level_2005,cremers_kernel_2006,gastaud_combining_2004}.
-These methods generally have an inplicit description of the relative boundary locations of the objects
-  to be delineated using level sets, and some even model the statistical deviations from original
-  shape.
+Registration-segmentation methods are generally built including a deformation model to
+  support the evolution of a segmentation algorithm that drives the mapping process.
+Thus, solutions usually select first a segmentation strategy to derive the cost function and
+  regularization of the problem.
+Generally, evolving active contours is the most extended segmentation approach to be
+  integrated within the registration process.
+These methods either have an implicit description of the relative boundary locations of the
+  objects to be delineated using level sets \citep{chen_using_2002,bresson_variational_2006,
+  chan_level_2005}, or model the statistical deviations from original shape
+  \citep{cremers_kernel_2006,gastaud_combining_2004}.
+Some Bayesian approaches have been proposed as well \citep{wyatt_map_2003,gass_simultaneous_2014}.
+In \autoref{sec:methods_map} we start from a probabilistic framework to finally propose
+	a cost function defined on the location of explicit surfaces, showing the duality of the
+	existing approaches.
 
-An early ``morphing''-segmentation approach was proposed in \citep{bertalmio_morphing_2000},
-  and the first proper registration method was presented by \citeauthor{yezzi_variational_2001}
-  \citep{yezzi_variational_2001}, where the energy functional is defined simultaneously in the target
+An early ``morphing''-segmentation approach was proposed in \citep{bertalmio_morphing_2000}.
+Soonly, \cite{yezzi_variational_2001} presented the first method including a full solution to
+  the registration, where the energy functional is defined simultaneously in the target
   and reference images with an affine transformation supporting the coordinates mapping
   (the so-called ``2 \glspl*{pde} approach'').
-\citeauthor{vemuri_joint_2003} proposed the first atlas-based registration \citep{vemuri_joint_2003}
-  based on the level-sets framework and using only one \gls*{pde}.
+\cite{vemuri_joint_2003} proposed the first atlas-based registration based on the level sets
+  framework and using only one \gls*{pde}.
 The principal application of these techniques was the image registration between timesteps in
   a time-series or between different images \citep{paragios_level_2003}.
-\citeauthor{unal_coupled_2005} \citep{unal_coupled_2005} extended the idea of
-  \citep{bertalmio_morphing_2000,yezzi_variational_2001} to nonlinear registration
-  using a free deformation field model.
-A summary of these second set of techniques is performed in \citep{droske_mumfordshah_2009},
-  proposing two different approaches to applying the Mumford-Shah functiona
-  \citep{mumford_optimal_1989} in simultaneous registration-segmentation, through the propagation
-  of the deformation field from the contours to the whole image definition.
-Finally, \citeauthor{gorthi_active_2011} presented a comprehensive generalization of the
-  existing methodologies \citep{gorthi_active_2011} based on the level-sets approximation.
-Even though the historical evolution of joint segmentation and registration procedures is strongly
-  linked to the level sets approach, our method conversely used the explicit representation of
-  the evolving surfaces.
+\cite{unal_coupled_2005} extended the idea of \citep{yezzi_variational_2001} to nonlinear
+  registration using a free deformation field model.
+\cite{droske_mumfordshah_2009} reviewed the latter set of techniques, and proposed two different
+  approaches to apply the Mumford-Shah functional \citep{mumford_optimal_1989} in simultaneous
+  registration-segmentation, through the propagation of the deformation field from
+  the contours to the whole image definition.
+Finally, \cite{gorthi_active_2011} presented a comprehensive generalization of the
+  existing methodologies based on the level sets approximation.
 
-Finally, a derivation of the latter group is composed by atlas-based segmentation methods
-  \citep{gorthi_segmentation_2009,gorthi_active_2011,pohl_unifying_2005,pohl_bayesian_2006,
-  wang_joint_2006}, where the prior imposes consistent voxel-based classification of contiguous
-  regions.
-A comprehensive summary of the methodologies derived from the \acrlong{acwe} model, with special attention
-  to joint registration-segmentation methods is found in \citep{gorthi_active_2011}.
-is based in derivations of the \acrlong{acwe}
-  framework \citep{chan_active_2001}, and related to \citep{guyader_combined_2011}.
+\todo[inline]{Revise these references and include if necessary:
+  \citep{pohl_unifying_2005,pohl_bayesian_2006,wang_joint_2006,guyader_combined_2011}}
 
-\subsection{Cost-function derivation}
+\subsection*{Cost-function derivation}
 \label{sec:methods_map}
@@
   contour between tissues $l$ and $m$.
-Then, the \gls*{map} criterium \citep{leemput_automated_1999} is used
-  to find the mapping estimate $\hat{U}$ that aligns $\Gamma$ into $R$:
+Then, the \gls*{map} criterium is applied to find the mapping estimate $\hat{U}$
+  that aligns $\Gamma$ into $R$:
   \begin{align}
@@
   volume.
-For convenience, and because it has been shown to be appropriate approximations
-  \citep{cuadra_comparison_2005}, we introduce two assumptions for each
+For convenience, and because it has been shown to be an appropriate approximation
+  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
   region $\Omega_l$:
@@
 
-\subsection{Numerical Implementation}
+\subsection*{Numerical Implementation}
 \label{sec:numerical_implementation}
@@
 
-\subsubsection{Deformation model}
+\paragraph*{Deformation model}
 \label{sec:deformation_model}
@@
 
-\subsubsection{Optimization}
+\paragraph*{Optimization}
 \label{sec:gradient_descent}
@@
 
-\subsection{Experiments and evaluation}
+\subsection*{Experiments and evaluation}
 \label{sec:experiments_evaluation}
 %
-The experiments follow a five-steps design:
-\begin{enumerate}
-	\item The prior surfaces are extracted from an undistorted and high-resolution dataset
-	  (e.g. \autoref{fig:phantom}).
-	\item A plausible, synthetic distortion $U^{-1}_{true}$ is generated.
-	\item The low-resolution (target) images are warped using the synthetic distortion.
-	\item Our method is run, obtaining a $\hat{U}_{tst}$.
-	  A cross-comparison methodology is also applied, to obtain a competing $\hat{U}_{cc}$.
-	\item Results are visually and quantitatively evaluated.
-\end{enumerate}
-
-To quantitatively evaluate the registration results we computed the \gls*{swindex} as the one-to-one
-  distance between corresponding vertices of surfaces, weighted by their respective
-  Voronoi area $w_i$:
+In order to demonstrate the applicability of the registration method, we first
+  subjected it to a battery of random tests over synthetic data and then applied
+  it on real data.
+\autoref{fig:evworkflows} presents the workflow implementing the evaluation instruments.
+Besides a visual assessment of the results, we report quantitative evaluations using
+  two metrics.
+In the case of the phantoms, as we warped data along the three available dimensions,
+  we computed the Haussdorf distance, integrating the ``point-to-cell'' method
+	implemented by \cite{commandeur_vtk_2011}.
+Conversely, the susceptibility-derived distortions only happen along the phase
+  encoding axis of the image.
+Therefore, a \gls*{swindex} can be computed as the one-to-one distance between corresponding
+  vertices of surfaces, weighted by their respective Voronoi area $w_i$:
   \begin{equation}
@@
 
-In the application on real data, three experts blindly rated correction results obtained
-  with the proposed method and the cross-comparison reference.
-The full report of visual results is provided within the Supplemental Materials.
+\begin{figure}
+\caption{1) The prior surfaces are extracted from an undistorted and high-resolution dataset
+	  (e.g. \autoref{fig:phantom}). 2) A plausible, synthetic distortion $U^{-1}_{true}$ is generated.
+	3) The low-resolution (target) images are warped using the synthetic distortion.
+	4) Our method is run, obtaining a $\hat{U}_{tst}$.
+	  A cross-comparison methodology is also applied, to obtain a competing $\hat{U}_{cc}$.
+	5) Results are visually and quantitatively evaluated.}\label{fig:evworkflows}
+\end{figure}
 
-  % \item \gls{wi}, L2-distance between the theoretical and the recovered
-  %   deformation field.
-  %   \begin{equation}
-  %   WI = \frac{1}{N} \sum\limits_i^N \| \mathbf{d}_i - \hat{\mathbf{d}}_i \|
-  %   \end{equation}
-  %   where $\mathbf{d}_i$ is the theoretical displacement vector at position $i$
-  %   and $\hat{\mathbf{d}}_i$ is the recovered one at the same index position.
-  % \item Parcellation agreement, the Dice overlap index of the \glspl*{roi}
-  %   contained by the parcellation.
-  % \end{itemize}
-
-\subsection{Image Data and preprocessing}
+\subsection*{Image Data and preprocessing}
 \label{sec:datasets}
 
-\subsubsection{Simulated digital phantom} %
+\paragraph*{Simulated phantoms}%
 \label{sec:digital_phantoms}
-%
-We first present the proof of the concept on a simplistic digital phantom
-  (\autoref{fig:phantom}).
-Using \cite{caruyer_phantomas_2014}, we simulated 2.0mm isotropic \gls*{t1} and
-  \gls*{t2} images at several rician noise levels, with parameters TE/TR=10/1500ms
-  and TE/TR=90/5000ms respectively.
+We proved the concept on simplistic digital phantoms which we will name after their
+  appearance as ``box'', ``ball'', ``L-shape'', and ``gyrus'' (\autoref{fig:phantom}).
+Using \emph{phantomas} \citep{caruyer_phantomas_2014}, we simulated \gls*{t1} and
+  \gls*{t2} images (TE/TR=10/1500ms and TE/TR=90/5000ms
+  respectively) corresponding to each shape.
+Simulations were conducted with rician noise for a \gls*{snr} of 300.0, and
+  two image resolutions, $1.0mm$ and $2.0mm$, both isotropic.
 The reference surfaces were extracted using the marching-cubes algorithm
   shipped with \emph{FreeSurfer} \citep{fischl_freesurfer_2012}
-  on high-resolution (1.0mm isotropic) and binary images corresponding to
-  the tissue fraction maps used in simulations.
+  at the higher resolution (1.0mm isotropic)..
 
@@
 
-\subsubsection{Real \gls{mri} datasets} %
+\paragraph*{Real datasets} %
 \label{sec:human_connectome}
 %
-For the evaluation of the algorithm to real \gls*{dmri} data of human brains,
-  we downloaded 10 datasets from the ``minimally preprocessed''
+For the evaluation of the algorithm on real \gls*{dmri} data of human brains,
+  we collected 15 datasets from the ``minimally preprocessed''
 	 database of the \gls*{hcp}.
 We refer the reader to \citep{essen_human_2012} for exact details about acquisition
-  parameters, and \citep{glasser_minimal_2013} for preprocessing issues.
+  parameters, and \citep{glasser_minimal_2013} for the preprocessing issues.
 The datasets comprehend \gls*{t1}, \gls*{t2} and multi-shell \gls*{dmri} images

### 2015-02-17 Publications/2015-NeuroImage/06-results.tex
Change to Neuroimage layout
Update all the sections

--- 2014-11-06 Publications/2013-TMI/06-results.tex
@@
 % -*- root: 00-main.tex -*-
-\section{Results and discussion}
+\section*{Results and discussion}
 \label{sec:results}
 
-\subsection{Proof of concept}
+\subsection*{Proof of concept}
 \label{sec:results_phantom}
@@
 {\color{red}
-\subsubsection{Convergence evindence and settings}
+\paragraph*{Convergence evindence and settings}
  discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition / Wolfe conditions, multi-resolutions on the bspline grid, image samples and surface sampling,  etc.
-\subsubsection{Efficient field densification}
+\paragraph*{Efficient field densification}
 As long as the dense deformation field is iteratively interpolated
@@
 
-\subsection{Susceptibility distortion correction}
+\subsection*{Susceptibility distortion correction}
 \label{sec:results_hcp}
@@
 
-\subsection{Discussion}
+\subsection*{Discussion}
 \label{sec:discussion}
@@
   to provide with prior information about anatomy with high accuracy.
+\todo[inline]{Add sentence about comment RW\#1 of IPMI2013: incorporate knowledge of
+the EPI distortions into regularization}
 Our method solves in a single step a joint problem usually addressed in two steps.
@@
   \citep{jezzard_correction_1995}.
+
+{\color{red} We also considered the reproducibility of results a design requirement.
+Therefore, real data are fetched from a publicly available repository
+  (the Human Connectome Project \citep{essen_human_2012}) and all the software
+  involved in this paper is also publicly released.
+\autoref{fig:evworkflows} describes the general structure of the workflow we implemented
+  as evaluation instrument, using \emph{nipype} \citep{gorgolewski_nipype_2011} to ensure
+  reproducibility.}
 

### 2015-02-17 Publications/2015-NeuroImage/08-acknowledgements.tex
Update all the sections

@@
 % use section* for acknowledgement
+\section*{Author contributions}
+OE implemented the method, designed and implemented the experiments, wrote the paper and
+  simulated and/or collected the data.
+DZ drafted the registration method, generated early phantom datasets and collaborated with the
+  implementation of the method.
+AD, MBC and MJLC interpreted the results.
+AD, MBC, MJLC, JPT and AS advised on all aspects of the work.
+All the authors contributed in editing the paper.
+
 \section*{Acknowledgment}

### 2015-02-17 Publications/2015-NeuroImage/04-methods.tex
write phantom experiments. close #222

@@
 
+{\color{red}
+\paragraph*{Convergence evidence and settings}
+ discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition / Wolfe conditions, multi-resolutions on the bspline grid, image samples and surface sampling,  etc.
+\paragraph*{Efficient field densification}
+As long as the dense deformation field is iteratively interpolated
+from the same set of control points that define the $L-1$ contours,
+it is possible to pre-cache all the $\psi_k(\vec{v}_i)$ weights into a
+sparse matrix for fast densification. As well, we achieve a
+diffeomorphic transform by}
+
 \subsection*{Experiments and evaluation}
@@
 We proved the concept on simplistic digital phantoms which we will name after their
-  appearance as ``box'', ``ball'', ``L-shape'', and ``gyrus'' (\autoref{fig:phantom}).
+  appearance as ``box'', ``ball'', ``L-shape'', and ``gyrus'' (\autoref{fig:phantom},
+  boxes A and B).
 Using \emph{phantomas} \citep{caruyer_phantomas_2014}, we simulated \gls*{t1} and
@@
   shipped with \emph{FreeSurfer} \citep{fischl_freesurfer_2012}
-  at the higher resolution (1.0mm isotropic)..
+  at the higher resolution (1.0mm isotropic).
 
-\begin{figure}
+\begin{figure*}
 	\includegraphics[width=\linewidth]{figures/figure02}
-	\caption{The ``cortex'' phantom is a spherical shape with two sulci and an
-	 outer crust resembling the cortical folding.
-	When the model is used to generate low resolution images, the
-	  partial volume effect turns segmentation of the sulci a
-	  challenging problem with voxel-wise clustering methods,
-	  but it is successfully segmented with the proposed method (see
-	  \autoref{fig:results_phantom}).
-	Top row shows the central-axial plane of the simulated images at
-	  high resolution and noiseless (left is \gls*{t1} and right is
-	  \gls*{t2}).
-	Bottom row represents the extracted contours that are used as shape
-	  priors to be registered to low-resolution, noise corrupted versions
-	  of the phantom.
+	\caption{A. The ``cortex'' phantom is a spherical shape with two sulci and an
+	  outer crust resembling the cortical folding (left).
+	The model is used to generate \gls*{t1} and \gls*{t2} images after warping the
+	  contours using a random and plausible transformation $U_{true}^{-1}$ (right).
+	B. Visual assessment of the results on the low resolution sets:
+	  ``gyrus'' (top-left), ``L-shape'' (top-right), ``ball'' (bottom-left),
+	  and ``box'' at (bottom-right).
+	In yellow color, the recovered contours after registration are represented.
+	Our method showed high accuracy, as they are overlapping the ground truth surfaces
+	  depicted in green.
+	Partial volume effect turns segmentation of the sulci a challenging problem with voxel-wise
+	  clustering methods, but it is successfully segmented with our method.
+	C. Quantitative evaluation of registration error in terms of average Hausdorff distance of
+	  surfaces at high (left) and low (right) resolutions, demonstrating that the error is
+	  consistently below the image resolution.
 	  }\label{fig:phantom}
-\end{figure}
+\end{figure*}
 

### 2015-02-17 Publications/2015-NeuroImage/06-results.tex
write phantom experiments. close #222

@@
 
-\subsection*{Proof of concept}
+\subsection*{Proof of concept on digital phantoms}
 \label{sec:results_phantom}
+Using the experimental instrumentation described in section \nameref{sec:experiments_evaluation},
+  we fine tuned the algorithm for its use on real data and collected quantitative results.
+To warp the simulated datasets, we generated 150 realizations of random and smooth displacement
+  fields $U^{-1}_{true}$.
+Smoothness is achieved using two levels of B-Spline functions, with control points evenly
+  located in isotropic grids covering the full extent of the model.
+The first level had $50.5mm$ of separation between control points and the second $25.25mm$.
+The rationale behind this choice is producing large deformations on the contours with
+  the first level, and matching the properties of susceptibility-derived distortions
+  as previously reported by \cite{irfanoglu_susceptibility_2011} with the finer grid.
+Invertibility of $U^{-1}_{true}$ is ensured by controlling the maximum displacement at each
+  level ($20.2mm$ and $10.1mm$ respectively) as studied in \citep{rueckert_diffeomorphic_2006}.
 
-Applying the evaluation protocol introduced in \autoref{sec:experiments_evaluation} on
-  the ``cortex'' phantom (\autoref{fig:phantom}), we successfully tested the method and
-  fine tuned the algorithm for its use on real data.
-In this characterization experiment, we generated random smooth displacement fields $U^{-1}_{true}$
-  (step 4 in \autoref{sec:experiments_evaluation}) based on a B-Spline field with
-  control points evenly located in a $6 \times 6 \times 6$ grid covering the region of the phantom.
-Invertibility is ensured by controlling the maximum	displacement
-  \citep{rueckert_diffeomorphic_2006}.
-The software tool to generate these fields is included with the released package.
-In this experiment we used $\hat{U}_{cc} = (U^{-1}_{true})^{-1}$, i.e. the ground-truth
-  displacement field.
-We performed the experiment in two different realizations of the phantom, one at full resolution
-  and a second with half resolution to study the performance under strong partial volume effect
-  conditions.
-The distorted phantoms and visual results are presented in \autoref{fig:results_phantom}.
-Quantitative evaluation of the \gls*{swindex} is reported in \autoref{tab:swi_phantom}.
+A total of 1200 registration experiments (4 phantom types $\times$ 150 random warpings
+  $\times$ 2 resolutions) were run.
+For each registration, the averaged Hausdorff distance (see \nameref{sec:experiments_evaluation})
+  of the inner and the outer surfaces was computed.
+The results showed a consistent and high accuracy, below voxel resolution.
+\autoref{fig:phantom} (block C) presents the box plots by model type corresponding
+  to the two sets of resolutions of generated phantoms.
+To support that the missregistration errors averaged per experiment was significantly
+  under the resolution of the target image, we proceeded as follows.
+First, we confirmed that the error distributions were skewed using a Shapiro-Wilk test of
+  normality.
+All the distributions of errors under test (4 phantom types $\times$ 2 resolutions) resulted
+  non-normal with $p<0.001$.
+Consequently, we used the non-parametric Wilcoxon signed-rank test along with Bonferroni
+  correction for multiple comparisons ($N=150$).
+Averaged errors resulted significantly lower than voxel resolution with $p < (0.001 / 150)$
+  for all the tests (4 phantom types $\times$ 2 resolutions).
 
-\begin{table}
-	\caption{Quantitative evaluation on phantoms: \acrlong*{swindex}}
-	\label{tab:swi_phantom}
-    \begin{tabular}{llll}
-    Voxel size  & Interior surface & Exterior surface & Average \\
-    $1.0mm$ & $0.718mm$            & $0.724mm$            & $0.721mm$   \\
-    $2.0mm$ & $0.723mm$            & $0.726mm$            & $0.725mm$   \\
-    \end{tabular}
-\end{table}
-
-{\color{red}
-\paragraph*{Convergence evindence and settings}
- discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition / Wolfe conditions, multi-resolutions on the bspline grid, image samples and surface sampling,  etc.
-\paragraph*{Efficient field densification}
-As long as the dense deformation field is iteratively interpolated
-from the same set of control points that define the $L-1$ contours,
-it is possible to pre-cache all the $\psi_k(\vec{v}_i)$ weights into a
-sparse matrix for fast densification. As well, we achieve a
-diffeomorphic transform by}
-
-% In regular intervals, after certain number of iterations,
-% the parameters $\Theta_l$ of the regions can be reestimated
-% based on the shifted volumetric samples
-% $\vec{x}' = \vec{x_0} + u(\vec{x})$.
-
-\begin{figure*}
-\includegraphics[width=\textwidth]{figures/figure03}
-\caption{Proof of concept of the registration method. First row shows the registration experiment
-  on the high-resolution ($1.0mm$ isometric) realization of the phantom, second row contains the
-  low-resolution ($2.0mm$ isometric) version.
-  Both phantoms were simulated with \gls*{snr} 30.
-  In the first column, the distorted phantoms presented with the undistorted contours.
-  For the second column, the contours have been replaced by the distorted contours using
-    the ground-truth warping $U^{-1}_{true}$.
-  The third column contains again the contours warped with the ground-truth field
-  (thin yellow line), and the corresponding contours (green and blue thick lines)
-  obtained with the presented method.
-  Finally, last row represents the recovered images unwarped using the $\hat{U}_{tst}$
-    found through our registration.
-  }\label{fig:results_phantom}
-\end{figure*}
-
-\subsection*{Susceptibility distortion correction}
+\subsection*{Distortion correction on real datasets}
 \label{sec:results_hcp}

### 2015-02-17 Publications/2015-NeuroImage/05-results.tex
re-structure discussion and results

--- 2015-02-17 Publications/2015-NeuroImage/06-results.tex
@@
 % -*- root: 00-main.tex -*-
-\section*{Results and discussion}
+\section*{Results}
 \label{sec:results}
@@
 \end{itemize}
-
-\subsection*{Discussion}
-\label{sec:discussion}
-We propose a registration-segmentation algorithm designed for biomedical
-  image analyses that are affected by all or some of the following circumstances:
-  \begin{itemize}
-  	\item Reference and target images do not present appropriate contrasts to
-  	perform volume-based registration reliably, or there is not available an apt image
- 		to be used as reference.
-  	\item The low resolution of the target volume produces strong partial volume effects
-  	that exclude the use of voxel-wise segmentation algorithms and hinder volume-based
-  	registration.
-  	\item The unknown warping is smooth enough to be represented with a sufficient number
-  	of BSpline kernels, and its explicit need is justified (i.e. the distortion
-  	field in \gls*{dmri} images, object tracking applications, etc.).
-  \end{itemize}
-The algorithm is designed for the proposed application on correcting distortions of
-  \gls*{dmri} data, a required preliminary step to extract the structural connectivity
-  networks.
-As a consequence, our approach relies on the availability of precise segmentations or
-  surfaces of the objects that are to be segmented on the target volume(s).
-Connectome extraction protocols generally include the acquisition of \gls*{t1} images
-  to provide with prior information about anatomy with high accuracy.
-\todo[inline]{Add sentence about comment RW\#1 of IPMI2013: incorporate knowledge of
-the EPI distortions into regularization}
-Our method solves in a single step a joint problem usually addressed in two steps.
-The first solves the linear registration of the \gls*{t1} and \gls*{dmri} data (i.e.
-  \cite{greve_accurate_2009}), whereas the second corrects for the nonlinear distortions
-  derived from the inhomogenety of magnetic susceptibility across tissues
-  \citep{jezzard_correction_1995}.
-
-{\color{red} We also considered the reproducibility of results a design requirement.
-Therefore, real data are fetched from a publicly available repository
-  (the Human Connectome Project \citep{essen_human_2012}) and all the software
-  involved in this paper is also publicly released.
-\autoref{fig:evworkflows} describes the general structure of the workflow we implemented
-  as evaluation instrument, using \emph{nipype} \citep{gorgolewski_nipype_2011} to ensure
-  reproducibility.}
-
-We implemented the registration method upon the widely used Insight Registration and Segmentation
-	Toolkit (\url{http://www.itk.org}) with the aim to release a useful and free software bundle.
-The instrumentation framework is also distributed with the registration method,
-  and has been implemented using \emph{nipype} \citep{gorgolewski_nipype_2011}.
-The pipelines include the automatic generation of the ``cortex'' phantom.
-The real datasets are publicly available under the \gls*{hcp} project.
-Thus, all the experiments and results presented in this paper are intended to be
-  fully reproducible.

### 2015-02-17 Publications/2015-NeuroImage/06-discussion.tex
re-structure discussion and results

--- 2015-02-17 Publications/2015-NeuroImage/06-results.tex
@@
 % -*- root: 00-main.tex -*-
-\section*{Results and discussion}
-\label{sec:results}
-
-\subsection*{Proof of concept on digital phantoms}
-\label{sec:results_phantom}
-Using the experimental instrumentation described in section \nameref{sec:experiments_evaluation},
-  we fine tuned the algorithm for its use on real data and collected quantitative results.
-To warp the simulated datasets, we generated 150 realizations of random and smooth displacement
-  fields $U^{-1}_{true}$.
-Smoothness is achieved using two levels of B-Spline functions, with control points evenly
-  located in isotropic grids covering the full extent of the model.
-The first level had $50.5mm$ of separation between control points and the second $25.25mm$.
-The rationale behind this choice is producing large deformations on the contours with
-  the first level, and matching the properties of susceptibility-derived distortions
-  as previously reported by \cite{irfanoglu_susceptibility_2011} with the finer grid.
-Invertibility of $U^{-1}_{true}$ is ensured by controlling the maximum displacement at each
-  level ($20.2mm$ and $10.1mm$ respectively) as studied in \citep{rueckert_diffeomorphic_2006}.
-
-A total of 1200 registration experiments (4 phantom types $\times$ 150 random warpings
-  $\times$ 2 resolutions) were run.
-For each registration, the averaged Hausdorff distance (see \nameref{sec:experiments_evaluation})
-  of the inner and the outer surfaces was computed.
-The results showed a consistent and high accuracy, below voxel resolution.
-\autoref{fig:phantom} (block C) presents the box plots by model type corresponding
-  to the two sets of resolutions of generated phantoms.
-To support that the missregistration errors averaged per experiment was significantly
-  under the resolution of the target image, we proceeded as follows.
-First, we confirmed that the error distributions were skewed using a Shapiro-Wilk test of
-  normality.
-All the distributions of errors under test (4 phantom types $\times$ 2 resolutions) resulted
-  non-normal with $p<0.001$.
-Consequently, we used the non-parametric Wilcoxon signed-rank test along with Bonferroni
-  correction for multiple comparisons ($N=150$).
-Averaged errors resulted significantly lower than voxel resolution with $p < (0.001 / 150)$
-  for all the tests (4 phantom types $\times$ 2 resolutions).
-
-\subsection*{Distortion correction on real datasets}
-\label{sec:results_hcp}
-
-As we propose the \gls*{fa} and the \gls*{md} as target features, we
-  implemented a simplified pipeline for processing \gls*{dmri} using
-  \emph{MRtrix} \citep{tournier_mrtrix_2012}.
-\todo[inline]{insert here the summary of parameters}
-
-Selecting the appropriate labels in the \emph{aparc} segmentation, we applied
-  the marching-cubes algorithm again to extract priors of the following
-  homogeneous regions in terms of \gls*{fa} and \gls*{md} joint distribution:
-\begin{itemize}
-	\item \gls*{csf} of the ventricular system,
-	\item deep \gls*{gm} structures,
-	\item \gls*{wm} surface as the ensemble of brain stem, and
-	  cerebellar and cerebral \gls*{wm},
-	\item cortical \gls*{gm} surface, including cerebellar \gls*{gm}.
-\end{itemize}
-
-\subsection*{Discussion}
+\section*{Discussion}
 \label{sec:discussion}

### 2015-02-17 Publications/2015-NeuroImage/08-acknowledgements.tex
re-structure discussion and results

--- 2015-02-16 Publications/2015-NeuroImage/08-acknowledgements.tex
@@
 % use section* for acknowledgement
+\section*{Author contributions}
+OE implemented the method, designed and conducted the experiments, wrote the paper,
+  simulated the phantoms and prepared the real data.
+DZ devised and drafted the registration method, generated early phantom datasets and
+  collaborated with the implementation of the method.
+AD, MBC and MJLC interpreted the results.
+AD, MBC, MJLC, JPT and AS advised on all aspects of the work.
+All the authors contributed in editing the paper.
+
 \section*{Acknowledgment}
@@
   and L. Vese for her support during OE's research visits in her laboratory.
-
-DZ is supported by the Swiss National Science Foundation (SNF) under grant PBELP2-137727.
 
@@
   Leenaards and Louis Jeantet foundations.
+DZ is supported by the Swiss National Science Foundation (SNF) under grant PBELP2-137727.

### 2015-02-17 Publications/2015-NeuroImage/03-introduction.tex
finishing details of introduction

@@
   content of both images.
-A second group, called \emph{geometric methods}, measures distances between features derived from
+A second group, called \emph{geometric methods}, align geometrical features derived from
   corresponding objects defined in both spaces, such as landmarks, shapes, surfaces, etc.
-Finally, \emph{hybrid methods} combine properties of the iconic and geometric methods.
-As reviewed in \citep{sotiras_deformable_2013}, an important research effort has been devoted to
-  study application-oriented matching criteria.
+Finally, \emph{hybrid methods} map geometrical features and voxel-wise information.
 
@@
   steps at the tissue interfaces where surfaces must be fitted, thanks to the use of
-  active contours without edges \cite{chan_active_2001}.
+  active contours without edges \citep{chan_active_2001}.
 The hypothesis underpinning our research is that, in those situations,

### 2015-02-17 Publications/2015-NeuroImage/04-methods.tex
finishing details of introduction

@@
   \citep{cremers_kernel_2006,gastaud_combining_2004}.
-Some Bayesian approaches have been proposed as well \citep{wyatt_map_2003,gass_simultaneous_2014}.
+Some Bayesian approaches have been proposed as well
+  \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014}.
 In \autoref{sec:methods_map} we start from a probabilistic framework to finally propose
@@
 An early ``morphing''-segmentation approach was proposed in \citep{bertalmio_morphing_2000}.
-Soonly, \cite{yezzi_variational_2001} presented the first method including a full solution to
+Shortly, \cite{yezzi_variational_2001} presented the first method including a full solution to
   the registration, where the energy functional is defined simultaneously in the target
-  and reference images with an affine transformation supporting the coordinates mapping
-  (the so-called ``2 \glspl*{pde} approach'').
-\cite{vemuri_joint_2003} proposed the first atlas-based registration based on the level sets
-  framework and using only one \gls*{pde}.
+  and reference images with an affine transformation supporting the coordinates mapping.
+\cite{vemuri_joint_2003} proposed an atlas-based registration using a level sets
+  framework and only one \gls*{pde} for first.
 The principal application of these techniques was the image registration between timesteps in
   a time-series or between different images \citep{paragios_level_2003}.
-\cite{unal_coupled_2005} extended the idea of \citep{yezzi_variational_2001} to nonlinear
-  registration using a free deformation field model.
+\cite{unal_coupled_2005} and later \cite{wang_joint_2006},
+  extended the ``two \glspl*{pde}'' method of \cite{yezzi_variational_2001}
+  to nonlinear registration with a free deformation field.
 \cite{droske_mumfordshah_2009} reviewed the latter set of techniques, and proposed two different
@@
 Finally, \cite{gorthi_active_2011} presented a comprehensive generalization of the
-  existing methodologies based on the level sets approximation.
-
-\todo[inline]{Revise these references and include if necessary:
-  \citep{pohl_unifying_2005,pohl_bayesian_2006,wang_joint_2006,guyader_combined_2011}}
+  existing methodologies based on the level sets approximation in the application of
+  atlas-based segmentation.
+A recent solution of the segmentation-registration problem close to the one proposed here was
+  presented by \cite{guyader_combined_2011}.
+Our method differs on two key features:
+  it allows multiple active contours for segmentation, and the regularization strategy.
+Whereas \cite{guyader_combined_2011} proposed a nonlinear elasticity smoother, we rely on
+  the smoothing effects of the deformation field and the anisotropic regularization
+  proposed by \cite{nagel_investigation_1986}.
 
@@
   parameters, and \citep{glasser_minimal_2013} for the preprocessing issues.
-The datasets comprehend \gls*{t1}, \gls*{t2} and multi-shell \gls*{dmri} images
-  already corrected for artifacts, brain-extracted and spatially normalized,
-  along with the most prominent results of the standard processing with
-  \emph{FreeSurfer}: \emph{aparc} segmentations, and the surfaces corresponding to
-  the cortical \gls*{gm} exterior and the \gls*{wm} interface with
-  cortical \gls*{gm}.
+The datasets comprehend a large of images, containing \gls*{t1}, \gls*{t2} and
+  multi-shell \gls*{dmri} images.
+The original acquisitions are released within ``unprocessed'' packages, and
+  the ``minimally preprocessed'' are corrected for artifacts, brain-extracted
+  and spatially normalized, along with some results of the standard processing with
+  \emph{FreeSurfer}.
+
+Selecting the appropriate labels in the \emph{aparc} segmentation, we applied
+  the marching-cubes algorithm again to extract the surface of the following
+  six homogeneous regions $\Omega_l$:
+  1) the thalamus;
+  2) \gls*{csf} of the ventricular system and deep \gls*{gm} structures;
+  3) cerebral \gls*{wm};
+  4) brain stem and cerebellar \gls*{wm};
+	5) cerebellar \gls*{gm}; and
+	6) cortical \gls*{gm} surface.
+The choice of this particular model of homogeneous regions responds to the joint distribution
+  of \gls*{fa} and \gls*{adc} of the human brain, as shown in the {\color{red} Supplemental
+  Materials (Figure XX)}.
+
+In this case, we derived the deformation $U_{true}^{-1}$ from the field maps released with
+  the corresponding packages of each dataset from the \gls*{hcp},
+  mimicking the real distortions, using a derivation of our previous work
+  \citep{esteban_simulationbased_2014}.
+After warping the original \gls*{dmri} with $U_{true}^{-1}$, we computed the \gls*{dti} and
+  the derived scalars (\gls*{fa} and \gls*{adc}) using \emph{MRtrix} \citep{tournier_mrtrix_2012}.

### 2015-02-17 Publications/2015-NeuroImage/05-results.tex
finishing details of introduction

@@
 
-\subsection*{Distortion correction on real datasets}
+\subsection*{Correction of distortions on real datasets}
 \label{sec:results_hcp}
-
-As we propose the \gls*{fa} and the \gls*{md} as target features, we
-  implemented a simplified pipeline for processing \gls*{dmri} using
-  \emph{MRtrix} \citep{tournier_mrtrix_2012}.
-\todo[inline]{insert here the summary of parameters}
-
-Selecting the appropriate labels in the \emph{aparc} segmentation, we applied
-  the marching-cubes algorithm again to extract priors of the following
-  homogeneous regions in terms of \gls*{fa} and \gls*{md} joint distribution:
-\begin{itemize}
-	\item \gls*{csf} of the ventricular system,
-	\item deep \gls*{gm} structures,
-	\item \gls*{wm} surface as the ensemble of brain stem, and
-	  cerebellar and cerebral \gls*{wm},
-	\item cortical \gls*{gm} surface, including cerebellar \gls*{gm}.
-\end{itemize}

### 2015-02-17 Publications/2015-NeuroImage/06-discussion.tex
finishing details of introduction

@@
 \label{sec:discussion}
-We propose a registration-segmentation algorithm designed for biomedical
+We propose a joint registration and segmentation algorithm designed for biomedical
   image analyses that are affected by all or some of the following circumstances:
-  \begin{itemize}
-  	\item Reference and target images do not present appropriate contrasts to
+  i) reference and target images do not present appropriate contrasts to
   	perform volume-based registration reliably, or there is not available an apt image
- 		to be used as reference.
-  	\item The low resolution of the target volume produces strong partial volume effects
+ 		to be used as reference;
+  ii) the low resolution of the target volume produces strong partial volume effects
   	that exclude the use of voxel-wise segmentation algorithms and hinder volume-based
-  	registration.
-  	\item The unknown warping is smooth enough to be represented with a sufficient number
-  	of BSpline kernels, and its explicit need is justified (i.e. the distortion
-  	field in \gls*{dmri} images, object tracking applications, etc.).
-  \end{itemize}
+  	registration; and,
+  iii) the estimation of the unknown warping is explicitly needed and the
+    deformation field is smooth enough to be represented with a sufficient number
+  	of B-Spline kernels (i.e. the distortion in \gls*{dmri} images,
+  	object tracking in time series, etc.).
+The registration method performs the segmentation of the target image in homogeneous
+  and competing regions defined by explicit active contours.
+We chose active contours without edges as we were interested on segmenting homogeneous
+  regions of diffusion properties within the brain.
+Finally, a B-Spline deformation field supports and keeps track of the evolution of the
+  active contours.
+Iteratively, the active contours are updated following the instant shape gradients
+  \citep{herbulot_segmentation_2006} with a gradient descent strategy.
+In section \nameref{sec:methods_background}, we framed our method within the current
+  state-of-art and defined the novelty of our approach.
+
+To support our claim that images can be simultaneously segmented and registered to a
+  reference coordinate system, we conducted evaluation experiments on digital
+  simulated phantoms and real datasets.
+
 The algorithm is designed for the proposed application on correcting distortions of
@@
 
+\paragraph*{Availability and reproducibility statement}
 {\color{red} We also considered the reproducibility of results a design requirement.
@@
   as evaluation instrument, using \emph{nipype} \citep{gorgolewski_nipype_2011} to ensure
-  reproducibility.}
+  reproducibility.
 
@@
 Thus, all the experiments and results presented in this paper are intended to be
-  fully reproducible.
+  fully reproducible.}

### 2015-02-17 Publications/2015-NeuroImage/08-acknowledgements.tex
finishing details of introduction

--- 2015-02-16 Publications/2015-NeuroImage/08-acknowledgements.tex
@@
 % use section* for acknowledgement
+\section*{Author contributions}
+All the authors contributed in editing the paper.
+OE implemented the method, designed and conducted the experiments, wrote the paper,
+  simulated the phantoms and prepared the real data.
+DZ devised and drafted the registration method, generated early phantom datasets, and
+  collaborated with the implementation of the method.
+AD, MBC and MJLC interpreted the results.
+AD, MBC, MJLC, JPT and AS advised on all aspects of the work.
+
 \section*{Acknowledgment}

### 2015-02-18 Scripts/pyacwereg/workflows/realdata.py
update t2b workflow

@@ -242,7 +242,8 @@ def hcp_workflow(name='Evaluation_HCP', settings={}, cfg={}):
         (st1,       cmethod1, [
             ('out_dis_set.dwi', 'inputnode.in_dwi'),
             ('out_dis_set.dwi_mask', 'inputnode.dwi_mask'),
-            ('out_ref_set.t2w_brain', 'inputnode.in_t2w')]),
+            ('out_ref_set.t2w_brain', 'inputnode.in_t2w'),
+            ('out_ref_set.surf', 'inputnode.in_surf')]),
         (ds,        cmethod1, [('bval', 'inputnode.in_bval'),
                                ('mr_param', 'inputnode.in_param')]),
         (cmethod1,   export2, [('outputnode.out_surf', 'surfaces0')]),

### 2015-02-18 Scripts/pyacwereg/workflows/registration.py
replace enhancements by N4. close #226

@@ -5,8 +5,8 @@
 #
 # @Author: oesteban - code@oscaresteban.es
 # @Date:   2014-03-28 20:38:30
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-02-17 12:02:09
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-02-18 11:45:39
 
 import os
 import os.path as op
@@ -115,7 +115,7 @@ def default_regseg(name='REGSEGDefault'):
     return wf
 
 
-def sdc_t2b(name='SDC_T2B', enhance_b0=True, icorr=True):
+def sdc_t2b(name='SDC_T2B', icorr=True):
     """
     The T2w-registration based method (T2B) implements an SDC by nonlinear
     registration of the anatomically correct *T2w* image to the *b0* image
@@ -154,10 +154,9 @@ def sdc_t2b(name='SDC_T2B', enhance_b0=True, icorr=True):
         fields=['dwi', 'dwi_mask', 'out_surf']), name='outputnode')
 
     avg_b0 = pe.Node(pmisc.ComputeAveragedB0(), name='avg_b0')
-    cache_b0 = pe.Node(niu.IdentityInterface(fields=['b0', 'mask']),
-                       name='B0Cache')
-    enh_t2 = pe.Node(pmisc.SigmoidFilter(upper_perc=92.0, lower_perc=65.0),
-                     name='enh_T2')
+    n4_b0 = pe.Node(ants.N4BiasFieldCorrection(dimension=3), name='BiasB0')
+    n4_t2 = pe.Node(ants.N4BiasFieldCorrection(dimension=3), name='BiasT2')
+
     getparam = pe.Node(nio.JSONFileGrabber(defaults={'enc_dir': 'y'}),
                        name='GetEncDir')
     reg = pe.Node(nex.Registration(num_threads=1), name='Elastix')
@@ -190,15 +189,17 @@ def sdc_t2b(name='SDC_T2B', enhance_b0=True, icorr=True):
         (inputnode,   getparam, [('in_param', 'in_file')]),
         (inputnode,  split_dwi, [('in_dwi', 'in_file')]),
         (inputnode,   corr_msk, [('dwi_mask', 'moving_image')]),
-        (inputnode,     enh_t2, [('in_t2w', 'in_file'),
-                                 ('t2w_mask', 'in_mask')]),
         (inputnode,      swarp, [('in_surf', 'points_file')]),
[…]

### 2015-02-18 Scripts/pyacwereg/workflows/registration.py
fix errors in #226

@@ -18,6 +18,7 @@ from nipype.interfaces import fsl
 from nipype.interfaces.fsl.maths import MultiImageMaths, Threshold
 from nipype.interfaces import freesurfer as fs
 from nipype.interfaces import elastix as nex
+from nipype.interfaces.ants import N4BiasFieldCorrection
 
 from pyacwereg.interfaces.acwereg import ACWEReg, ACWEReport
 from pyacwereg.interfaces.warps import FieldBasedWarp, InverseField
@@ -154,8 +155,8 @@ def sdc_t2b(name='SDC_T2B', icorr=True):
         fields=['dwi', 'dwi_mask', 'out_surf']), name='outputnode')
 
     avg_b0 = pe.Node(pmisc.ComputeAveragedB0(), name='avg_b0')
-    n4_b0 = pe.Node(ants.N4BiasFieldCorrection(dimension=3), name='BiasB0')
-    n4_t2 = pe.Node(ants.N4BiasFieldCorrection(dimension=3), name='BiasT2')
+    n4_b0 = pe.Node(N4BiasFieldCorrection(dimension=3), name='BiasB0')
+    n4_t2 = pe.Node(N4BiasFieldCorrection(dimension=3), name='BiasT2')
 
     getparam = pe.Node(nio.JSONFileGrabber(defaults={'enc_dir': 'y'}),
                        name='GetEncDir')
@@ -192,7 +193,7 @@ def sdc_t2b(name='SDC_T2B', icorr=True):
         (inputnode,      swarp, [('in_surf', 'points_file')]),
         (inputnode,        reg, [('t2w_mask', 'fixed_mask'),
                                  ('dwi_mask', 'moving_mask')]),
-        (inputnode,      n4_t2, [('in_t2w', 'mask_image'),
+        (inputnode,      n4_t2, [('in_t2w', 'input_image'),
                                  ('t2w_mask', 'mask_image')]),
         (inputnode,      n4_b0, [('dwi_mask', 'mask_image')]),
         (avg_b0,         n4_b0, [('out_file', 'input_image')]),

### 2015-02-18 Publications/2015-NeuroImage/00-main.tex
update acks #224
update intro
Update all the sections
re-structure discussion and results
advancing on #223

@@
 %\documentclass[numbers,sort&compress]{elsarticle}
-\documentclass[5p,numbers,sort&compress]{elsarticle}
+\documentclass[5p,authoryear]{elsarticle}
 \journal{NeuroImage}
@@
 \usepackage{url}
-\usepackage[hidelinks]{hyperref}
+\usepackage[colorlinks=true,linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
 
@@
 
-\title{Simultaneous segmentation-registration using active contours with application on
-diffusion MR images}
+\title{Simultaneous segmentation and registration of diffusion MR images of the brain driven
+by active-contours}
 
 \author[bit]{Oscar~Esteban\corref{corrauthor}}
+\cortext[corrauthor]{Corresponding author}
+\ead{phd@oscaresteban.es}
+\author[ucla]{Dominique~Zosso}
 \author[lts5]{Alessandro~Daducci}
 \author[chuv,lts5]{Meritxell~Bach-Cuadra}
+\author[bit]{Mar\'ia-J.~Ledesma-Carbayo}
 \author[lts5]{Jean-Philippe~Thiran}
 \author[bit]{Andr\'es~Santos}
-\author[bit]{Mar\'ia-J.~Ledesma-Carbayo}
-\author[ucla]{Dominique~Zosso}
-\cortext[corrauthor]{Corresponding author}
-\ead{phd@oscaresteban.es}
 
@@
 Universidad Polit\'ecnica de Madrid and CIBER-BBN, Madrid, Spain}
+\address[ucla]{Department of Mathematics, University of California,
+Los Angeles (UCLA), Los Angeles, CA, US}
 \address[lts5]{Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
@@
 Hospital Center (CHUV) and University of Lausanne (UNIL), Lausanne, Switzerland}
-\address[ucla]{Department of Mathematics, University of California,
-Los Angeles (UCLA), Los Angeles, CA, US}
-
-% \author{Oscar~Esteban$^{*}$,
-%         Alessandro~Daducci,
-%         Meritxell~Bach-Cuadra,
-% %        Jean-Philippe~Thiran,
-% %        Andr\'es~Santos,
-%         and~Dominique~Zosso% <-this % stops a space
-% \thanks{Manuscript received XXX XX, 2013; revised XXX XX, 2013.
-% This study is supported by: the Spain's Ministry of Science and Innovation
-% (projects TEC2010-21619- C04-03, TEC2011-28972-C02-02, CDTI-CENIT
-% AMIT and INNPACTO PRECISION), Comunidad de Madrid (ARTEMIS) and
-% European Regional Development Funds; the Center for Biomedical Imaging
-% (CIBM) of the Geneva and Lausanne Universities and the EPFL, as well as the
-% Leenaards and Louis Jeantet foundations.\emph{~Asterisk indicates corresponding
-% author}.}%
-% \thanks{$^{*}$O. Esteban is with the Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on,
-% Universidad Polit\'ecnica de Madrid and CIBER-BBN, Madrid, Spain
-% (e-mail: phd@oscaresteban.es).}% <-this % stops a space
-% \thanks{A. Daducci is with the Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
-% F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland.}%
-% \thanks{M. Bach-Cuadra is with the Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
-% F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland, and Dept. of Radiology, University
-% Hospital Center (CHUV) and University of Lausanne (UNIL), Lausanne, Switzerland.}%
-% %\thanks{A. Santos is with the Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on,
-% %Universidad Polit\'ecnica de Madrid and CIBER-BBN, Madrid, Spain.}% <-this % stops a space
-% \thanks{D. Zosso is with the Department of Mathematics, University of California,
-% Los Angeles (UCLA), Los Angeles, CA, US and is supported by the Swiss National
-% Science Foundation (SNF) under grant PBELP2-137727.}% <-this % stops a space
-% }
 
@@
 \input{04-methods}
-%\input{05-experiments}
-\input{06-results}
+\input{05-results}
+\input{06-discussion}
 

### 2015-02-18 Publications/2015-NeuroImage/04-methods.tex
advancing on #223

@@
 
-In a Bayesian framework, the mappings $U$ in \autoref{eq:transform} are
+In a Bayesian framework, the mappings $U$ in \eqref{eq:transform} are
   evaluated based on their posterior probability given the observed data
@@
   where $\vec{f}' = R(\vec{r}')$ is the feature vector at the displaced
-  position $\vec{r}'$ (\autoref{eq:transform}) of the multispectral target
+  position $\vec{r}'$ \eqref{eq:transform} of the multispectral target
   volume.
@@
  	descriptors of region $\Omega_l$.
-We then insert \autoref{eq:likelihood} into \autoref{eq:bayes_rule} to obtain the full
+We then insert \eqref{eq:likelihood} into \eqref{eq:bayes_rule} to obtain the full
   formulation of the \emph{data term} of our registration model:
@@
 	The gradient speeds $\bar{s}_i$ are computed as the disparity between data energies of
-	the vertex $\vec{v}_i$ in each limiting region as described in \autoref{eq:energy}.
+	the vertex $\vec{v}_i$ in each limiting region as described in \eqref{eq:energy}.
 	In this figure, the gradient speed corresponding to $\vec{v}_1$ is written in the lower
@@
 	support the deformation field through the corresponding weights $\psi_k(\vec{v}_i)$ obtained
-	as in \autoref{eq:basis_derivative}.
+	as in \eqref{eq:basis_derivative}.
 	}\label{fig:method}
@@
   where $\hat{\vec{e}}$ is the coordinates system's unit vector.
-The vertex speeds $\bar{s}_i'$ obtained by computing the shape-gradients \autoref{eq:shape_gradients},
+The vertex speeds $\bar{s}_i'$ obtained by computing the shape-gradients \eqref{eq:shape_gradients},
 	are then projected to the deformation field in order to obtain the derivatives $\vec{g}_k$

### 2015-02-18 Publications/2015-NeuroImage/05-results.tex
advancing on #223

@@
 Using the experimental instrumentation described in section \nameref{sec:experiments_evaluation},
-  we fine tuned the algorithm for its use on real data and collected quantitative results.
+  we collected quantitative results on the digital phantoms.
 To warp the simulated datasets, we generated 150 realizations of random and smooth displacement
@@
 Smoothness is achieved using two levels of B-Spline functions, with control points evenly
-  located in isotropic grids covering the full extent of the model.
+  located in isotropic grids covering the full extent of the phantom.
 The first level had $50.5mm$ of separation between control points and the second $25.25mm$.
@@
 
-\subsection*{Correction of distortions on real datasets}
-\label{sec:results_hcp}
+\subsection*{Correction of distortions on real datasets}\label{sec:results_hcp}
+
+\paragraph*{Assessing the segmentation model}\label{sec:res_model_and_metric} %
+%
+Preliminarly, we investigated the aptness of our segmentation model to play as cost function
+  driving the registration process.
+First, we selected the 6-regions model described in \nameref{sec:human_connectome} after
+  visual evaluation of the joint distribution {\color{red} REF Supplemental} of several
+  alternative models for segmenting the target bivariate images (\gls*{fa} and \gls*{adc}).
+Since susceptibility-derived distortions are always aligned with the phase encoding axis
+  and feature the property of being linear with respect to the phase inhomogeneity map,
+  we reduced the dimensionality of the error space to one dimension.
+Thus, the energy functional \eqref{eq:energy} was evaluated for the contour positions
+  obtained along a lineralized space of distortions
+  $\hat{U}\{\vec{r}\} = \epsilon \cdot U_{true} = \vec{r} + \epsilon \cdot u(\vec{r})$,
+  with $\epsilon \in [-1.1, 1.1]$.
+The result of this experiment is shown in \autoref{fig:realdata}.
+The metric consistently displayed its minimum at the zero-error point.
+
+\paragraph*{Cross-comparison evaluation}\label{sec:res_cc_evaluation}
+%
+Finally, we evaluated the performance of our registration method, and compared it to correcting
+  for distortions using the \emph{b0}-to-\gls*{t2} registration approach.

### 2015-02-18 Scripts/pyacwereg/interfaces/acwereg.py
improvements on executing workflows

@@ -5,7 +5,7 @@
 # @Author: Oscar Esteban - code@oscaresteban.es
 # @Date:   2014-03-12 13:20:04
 # @Last Modified by:   oesteban
-# @Last Modified time: 2015-01-16 11:06:07
+# @Last Modified time: 2015-02-18 18:17:30
 
 import os
 import os.path as op
@@ -150,12 +150,12 @@ pial_lh.vtk pial_rh.vtk -o tests [ -i 30 -u 10 -f 1.0 -s 0.5 -a 0.0 -b 0.0 \
     _grouped_traits = []
     _cmd = 'regseg'
     _num_levels = 0
+    _singleworker = False
 
     def __init__(self, command=None, **inputs):
         """ Combine general and grouped inputs """
         super(ACWEReg, self).__init__(command=command, **inputs)
         self.groups = self.input_group_spec()
-
         general_names = ANTSCommandInputSpec().trait_names()
 
         for name in self.groups.trait_names():

### 2015-02-18 Scripts/pyacwereg/workflows/realdata.py

@@ -2,15 +2,14 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2015-01-15 10:47:12
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-02-17 12:13:07
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-02-18 16:18:54
 
 
-def hcp_workflow(name='Evaluation_HCP', settings={}, cfg={}):
+def hcp_workflow(name='Evaluation_HCP', settings={}):
     """
     The pyacwereg evaluation workflow for the human connectome project (HCP)
     """
-    from nipype import config, logging
     from nipype.pipeline import engine as pe
     from nipype.interfaces import utility as niu
     from nipype.interfaces import io as nio
@@ -29,9 +28,6 @@ def hcp_workflow(name='Evaluation_HCP', settings={}, cfg={}):
     from pysdcev.workflows.smri import preprocess_t2, preprocess_dwi
     from pysdcev.workflows.tractography import mrtrix_dti
     from pysdcev.stages.stage1 import stage1
-
-    config.update_config(cfg)
-    logging.update_logging(config)
 
     fnames = dict(t1w='T1w_acpc_dc_restore.nii.gz',
                   t1w_brain='T1w_acpc_dc_restore_brain.nii.gz',

### 2015-02-18 Scripts/tools/run_evaluations.py

@@ -5,8 +5,8 @@
 #
 # @Author: oesteban - code@oscaresteban.es
 # @Date:   2014-04-04 19:39:38
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-02-02 10:35:00
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-02-18 16:13:28
 
 
 import os
@@ -17,6 +17,7 @@ if __name__ == '__main__':
     import os.path as op
     from glob import glob
     import numpy as np
+    from nipype import config, logging
 
     try:
         from enthought.etsconfig.api import ETSConfig
@@ -98,25 +99,28 @@ if __name__ == '__main__':
         from multiprocessing import cpu_count
         nthreads = cpu_count()
 
-    cfg = {}
-    cfg['plugin'] = 'Linear'
+    plugin = 'Linear'
     if nthreads > 1:
-        cfg['plugin'] = 'MultiProc'
-        cfg['plugin_args'] = {'n_proc': nthreads}
+        plugin = 'MultiProc'
+        plugin_args = {'n_proc': nthreads}
 
     # Setup logging dir
     log_dir = op.abspath('logs')
-    cfg['logging'] = {'log_directory': log_dir, 'log_to_file': True,
-                      'workflow_level': 'INFO', 'interface_level': 'INFO'}
     if not op.exists(log_dir):
         os.makedirs(log_dir)
 
+    cfg = {}
+    cfg['logging'] = {'log_directory': log_dir, 'log_to_file': True,
+                      'workflow_level': 'INFO', 'interface_level': 'INFO'}
     # Setup debug mode
     if opts.debug:
         cfg['logging']['workflow_level'] = 'DEBUG'
         cfg['logging']['interface_level'] = 'DEBUG'
+        config.enable_debug_mode()
 
-    wf = hcp_workflow(name=opts.name, settings=settings, cfg=cfg)
+    logging.update_logging(config)
+
+    wf = hcp_workflow(name=opts.name, settings=settings)
     wf.base_dir = settings['work_dir']
     wf.write_graph(format='pdf')
-    wf.run()
+    wf.run(plugin=plugin, plugin_args=plugin_args)

### 2015-02-18 Publications/2015-NeuroImage/03-introduction.tex
modify font, add lineno

@@
 % -*- root: 00-main.tex -*-
-\section*{Introduction}
-\label{sec:introduction}
-
+\section*{Introduction}\label{sec:introduction}
 Nonlinear registration of images is a challenging task, ubiquitous
@@
   in two different coordinate systems (generally called
-  \emph{reference}, $R$, and \emph{moving}, $M$):
+  \emph{reference}, $R$, and \emph{moving}, $M$):%
+
   \begin{align}
@@
   \end{align}
+
   where $\vec{r}$ denotes a position in the reference domain $R$, $\vec{r}'$ is

### 2015-02-18 Publications/2015-NeuroImage/04-methods.tex
modify font, add lineno

@@
 
-\subsection*{Cost-function derivation}
-\label{sec:methods_map}
+\subsection*{Cost-function derivation}\label{sec:methods_map}
 
@@
 Using the Bayes' rule, the posterior likelihood is computed as:
+
   \begin{equation}
@@
   \end{equation}
+
   where $P(R \mid U,\Gamma_{l,m})$ is the data-likelihood, and
@@
   that aligns $\Gamma$ into $R$:
+
   \begin{align}
@@
   global data likelihood into a product of pixel-wise conditional probabilities:
+
   \begin{equation}
@@
   \end{equation}
+
   where $\vec{f}' = R(\vec{r}')$ is the feature vector at the displaced
@@
   distributions:
+
   \begin{equation}
@@
   \end{equation}
+
  	where $\Theta_l = \lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$ are the
@@
   formulation of the \emph{data term} of our registration model:
+
  	\begin{align}
@@
   \end{align}
+
   using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
@@
 We choose a Thikonov regularization prior as follows:
+
   \begin{equation*}
@@
   \end{equation*}
+
   with further:
+
   \begin{align*}
@@
   \end{align*}
+
   imposing that the distortion and its gradient have zero
@@
   in the following for the sake of clarity:
+
   \begin{align}
@@
   following log-transform:
+
   \begin{align}
@@
   \end{align}
+
   that is the dual expression to the energy functional corresponding
@@
 
-\paragraph*{Deformation model}
-\label{sec:deformation_model}
+\paragraph*{Deformation model}\label{sec:deformation_model}
 Let us denote $\{\vec{v}_i\}_{i=1 \ldots N_c}$ the vertices of one or several prior
@@
   target image is achieved through a dense deformation field $u(\vec{r})$, such that:
+
   \begin{equation}
@@
   \end{equation}
+
   where $U$ is defined in \eqref{eq:transform}.
@@
 Densification is achieved through a set of associated basis functions $\psi_k$:
+
   \begin{equation}
@@
   \end{equation}
-%
+
 In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
@@
   $\psi$ by the actual kernel function, the transformation writes:
+
   \begin{equation}
@@
   \end{equation}
+
   with $\hat{\mathbf{e}}_d$ being the unitary vector along axis $d$.
@@
   deformation field through the following \gls*{pde}:
+
   \begin{equation}
@@
   \end{equation}
+
   with $t$ being an artificial time parameter of the contour
@@
 Now, we introduce \eqref{eq:energy} in \eqref{eq:general_gradient_descent}:
+
   \begin{align}
@@
   to compute the data term in \eqref{eq:gradient_descent} as follows:
+
   \begin{align}
@@
   \end{align}
+
   in this case, the formulation has been adapted to the non-binary case, $\{l,m\}$
@@
 Finally, we can compute:
+
   \begin{align}
@@
   \end{align}
+
   where $\hat{\vec{e}}$ is the coordinates system's unit vector.
@@
 	corresponding to $\vec{u}_k$:
+
   \begin{equation}
@@
   \end{equation}
+
   then the full gradient evolution equation \eqref{eq:gradient_descent} yields:
+
   \begin{align}
@@
   vertices of surfaces, weighted by their respective Voronoi area $w_i$:
+
   \begin{equation}
@@
   \end{equation}
+
   where $\vec{v}_i$ are the locations of the $N_p$ vertices in each $p \in \{1, \dots P\}$

### 2015-02-18 Publications/2015-NeuroImage/06-discussion.tex
modify font, add lineno

@@
   i) reference and target images do not present appropriate contrasts to
-  	perform volume-based registration reliably, or there is not available an apt image
- 		to be used as reference;
+  	perform volume-based registration reliably;
   ii) the low resolution of the target volume produces strong partial volume effects

### 2015-02-18 Publications/2015-NeuroImage/08-appendix-00-shapepriors.tex
modify font, add lineno

@@
 % -*- root: 00-main.tex -*-
+\renewcommand{\theequation}{A.\arabic{equation}}
 
-\section{Extended formulations}
+\section*{Appendix}
 
-\subsection{Shape priors}
+\subsection*{Instant shape priors}
 \label{app:shape_priors}
@@
   $m$, and $\Gamma_{l,m}$ a closed boundary between them
-  as defined in \autoref{sec:methods_map}.
+  as defined in \nameref{sec:methods_map}.
 We now derive the domain integral w.r.t. $t$:
+
   \begin{align*}
@@
   \end{align*}
+
   where $\left\langle\frac{\partial\Gamma_{l,m}}{\partial t}, N_{\Gamma_{l,m}}\right\rangle$ is
@@
   that $\frac{\partial}{\partial t} F(\vec{r}) = 0$ and thus:
+
   \begin{equation*}
@@
   Therefore, we substitute:
+
     \begin{equation}
@@
     \end{equation}
+
     where $\mathcal{E}_l(\vec{r}) = \mathcal{D}^2_{l}(R(\vec{r}))$ for briefness.

### 2015-02-19 Scripts/pyacwereg/workflows/evaluation.py
fix minor error writing csv files

@@ -5,8 +5,8 @@
 #
 # @Author: Oscar Esteban - code@oscaresteban.es
 # @Date:   2014-03-12 16:59:14
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-02-14 11:36:05
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-02-19 10:31:00
 
 import os
 import os.path as op
@@ -301,7 +301,7 @@ def map_energy(name='EnergyMapping', out_csv='energiesmapping.csv'):
     getval = pe.Node(nio.JSONFileGrabber(), name='GetZeroEnergy')
     csv = pe.Node(namisc.AddCSVRow(in_file=out_csv),
                   name="AddReferenceRow")
-    csv.inputs.errfactor = 0.0
+    csv.inputs.error = 0.0
 
     mapper = warp_n_map(out_csv=out_csv)
     wf = pe.Workflow(name=name)

### 2015-02-19 Scripts/pyacwereg/viz.py
save figure to pdf

@@ -2,8 +2,8 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2014-12-11 15:08:23
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-02-14 11:43:52
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-02-19 10:46:19
 
 
 def add_annotations(values, ax, level, nlevels, color, lastidx, units=''):
@@ -497,7 +497,7 @@ def phantom_errors(in_csv, resolution='lo',
     return g
 
 
-def metric_map_plot(in_file):
+def metric_map_plot(in_file, out_file=None):
     import pandas as pd
     import numpy as np
     import seaborn as sn
@@ -505,7 +505,6 @@ def metric_map_plot(in_file):
     plt.style.use('ggplot')
     sn.set_context("talk", font_scale=1.25)
     plt.rcParams['font.size'] = 24
-
 
     df = pd.read_csv(in_file)
     df.error.fillna(0.0, inplace=True)
@@ -520,6 +519,10 @@ def metric_map_plot(in_file):
         series.append(ts)
 
     final = pd.DataFrame(np.array(series).T, index=ndf.error, columns=subjects)
-    final.plot(legend=False)
-    return final
+    ax = final.plot(legend=False)
+    ax.set_xlabel(r'Registration error ($\epsilon$)')
+    ax.set_ylabel('Relative metric value')
+    if out_file is not None:
+        plt.savefig(out_file, dpi=300, bbox_inches='tight')
 
+    return final

### 2015-02-19 Publications/2015-NeuroImage/04-methods.tex
minor updates

@@
   boxes A and B).
-Using \emph{phantomas} \citep{caruyer_phantomas_2014}, we simulated \gls*{t1} and
-  \gls*{t2} images (TE/TR=10/1500ms and TE/TR=90/5000ms
-  respectively) corresponding to each shape.
-Simulations were conducted with rician noise for a \gls*{snr} of 300.0, and
-  two image resolutions, $1.0mm$ and $2.0mm$, both isotropic.
+Using \emph{phantomas} \citep{caruyer_phantomas_2014}, we simulated
+  \gls*{t1} (TE/TR=10/1500ms) and \gls*{t2} images (TE/TR=90/5000ms)
+  corresponding to each appearance shape, with two resolutions each
+  ($1.0mm$ and $2.0mm$ isotropic).
+Simulations were corrupted with rician noise for a \gls*{snr} of 300.0.
 The reference surfaces were extracted using the marching-cubes algorithm

### 2015-02-19 Publications/2015-NeuroImage/05-results.tex
minor updates

@@
   with $\epsilon \in [-1.1, 1.1]$.
-The result of this experiment is shown in \autoref{fig:realdata}.
+The result of this experiment is shown in \autoref{fig:energymap}.
 The metric consistently displayed its minimum at the zero-error point.
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/figure04.pdf}
+	\caption{Assessment of the segmentation model, sampling the value of the energy functional
+	\eqref{eq:energy} versus several registration errors generated as explained in
+	\nameref{sec:res_model_and_metric}.
+	The metric of our registration method displayed a smooth gradient towards the minimum
+	located at the $\epsilon = 0.0$ point.}\label{fig:energymap}
+\end{figure}
 

### 2015-02-19 Publications/2015-NeuroImage/03-introduction.tex
finishing changes

@@
   and precise surfaces to warped and low-resolution \gls*{dmri} of the brain.
-Currently, we are witnessing an uprising demand for such algorithms in the challenging context
-  of extracting and characterizing the structural connectivity networks of the
+Currently, we are witnessing an uprising demand for such algorithms in the quest
+  of extracting and characterizing the networks of structural connectivity in the
   human brain \citep{craddock_imaging_2013}.
@@
   distortions that \gls*{dmri} of the head typically present \citep{jezzard_correction_1995}.
-\todo[inline]{talk about the distortion and cite SDCs}
+The precise segmentation of \gls*{dmri} data is also crucial in such applications, since the
+  definition of the nodes in the network is generally derived from a parcellation performed in
+  an anatomical \gls*{mri} of the same subject \citep{daducci_connectome_2012}.
 
-A fundamental antecedent of the present approach is found in \citep{greve_accurate_2009}.
+\Gls*{dmri} data are usually acquired with \gls*{epi} schemes.
+In order to quickly probe  in many different orientations the diffusion process within the living
+  brain, these sequences minimize the bandwidth of the \gls*{pe} direction.
+This low bandwith originates the artifactual displacement of locations presenting small deviations
+  from the main magnetic field $B_0$.
+The principal source of these deviations, $\Delta B_0$, is the discontinuity of magnetic
+  susceptibility at tissue interfaces.
+Tissue/air boundaries show large steps of susceptibility and this artifact is
+  particularly evident in the surroundings of the orbito-frontal lobe or the temporal
+  bones in both hemispheres.
+The displacements field $U_s$ along the \gls*{pe} axis can be estimated theoretically
+  \citep{jezzard_correction_1995}:
+
+  \begin{equation}
+  \vec{r}' = \vec{r} + \frac{\gamma \, T_{acq}\, s_{PE}}{2\pi}\Delta B_0(\vec{r}) \cdot \hat{\vec{e}}_{PE}.
+  \label{eq:fieldmap}
+  \end{equation}
+
+where $\vec{r}$ is the location in the reference domain as in \eqref{eq:transform},
+  $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time,
+  $s_{PE}$ is the pixel spacing along \gls*{pe}, and $\hat{\vec{e}}_{PE}$ the unitary
+  vector along \gls*{pe}.
+Several methodologies have been proposed to estimate $U_s$ and fix the artifact.
+The first solution was acquiring extra images to estimate $\Delta B_0$ using field mapping
+  techniques \citep{jezzard_correction_1995,andersson_modeling_2001}.
+A second set of methods aquired extra images reproducing the same \gls*{epi} scheme but switching
+  the \gls*{pe} axis \citep{chiou_simple_2000} or reversing the direction of gradient
+  increments in the same axis \citep{cordes_geometric_2000,holland_efficient_2010}.
+\cite{kybic_unwarping_2000} proposed a correction using nonlinear registration of the \gls*{dmri}
+  data versus a \gls*{t2} image.
+Further developments \citep{irfanoglu_susceptibility_2011} extended this method to the
+  three-dimensional problem.
+All methods require extra \gls*{mr} aquisitions. Thus, a tipical protocol for the extraction
+  of structural connectivity comprehends \gls*{dmri}, \gls*{t1}, and the acquisition demanded
+  by the correction method of choice.
+
+Our method is related to the \gls*{t2b} methods as we also render the correction as a
+  registration problem.
+Conversely, the method does not require any of the previously mentioned extra acquisitions.
+A fundamental antecedent to the present approach is found in \citep{greve_accurate_2009}.
 They proposed a registration method to align \gls*{t1} and \gls*{dmri} images

### 2015-02-19 Publications/2015-NeuroImage/04-methods.tex
finishing changes

@@
   atlas-based segmentation.
-A recent solution of the segmentation-registration problem close to the one proposed here was
-  presented by \cite{guyader_combined_2011}.
+Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
+  registration method close to the one proposed here.
 Our method differs on two key features:
@@
   $\Gamma_{l,m} = \partial \Omega_l \cap \partial \Omega_m$ is the
-  contour between tissues $l$ and $m$.
-Then, the \gls*{map} criterium is applied to find the mapping estimate $\hat{U}$
-  that aligns $\Gamma$ into $R$:
-
-  \begin{align}
-  \hat{U} &= \underset{U}{\argmax} \left\{ P(U \mid R,\Gamma_{l,m}) \right\} \notag\\
-   &= \underset{U}{\argmax} \left\{ P(R \mid U,\Gamma_{l,m})\, P(U) \right\}.
-  \label{eq:map_u}
-  \end{align}
+  contour between competing regions $\Omega_l$ and $\Omega_m$.
+The best estimate $\hat{U}$ then fulfills the maximum a posteriori criterium
+  \citep{bishop_pattern_2006}, and aligns $\Gamma$ into $R$.
 
@@
   where $\vec{f}' = R(\vec{r}')$ is the feature vector at the displaced
-  position $\vec{r}'$ \eqref{eq:transform} of the multispectral target
+  position $\vec{r}'$ \eqref{eq:transform} in the multispectral target
   volume.
@@
   1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
-  distributions:
-
-  \begin{equation}
-  P\left( \vec{f}' \mid U,\Omega_l \right) = \mathcal{N} \left( \vec{f}' \mid \Theta_l \right),
-  \label{eq:likelihood}
-  \end{equation}
-
- 	where $\Theta_l = \lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$ are the
- 	descriptors of region $\Omega_l$.
-We then insert \eqref{eq:likelihood} into \eqref{eq:bayes_rule} to obtain the full
+  distributions with parameters $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
+  for each region $\Omega_l$.
+We then insert the normal distribution into \eqref{eq:bayes_rule} to obtain the full
   formulation of the \emph{data term} of our registration model:
@@
 
-We choose a Thikonov regularization prior as follows:
-
-  \begin{equation*}
-  P(U) = \underset{\vec{r}}{\prod} p(u(\vec{r})) =
-  \underset{\vec{r}}{\prod} p_0(u(\vec{r})) \, p_1(u(\vec{r})),
-  \end{equation*}
-
-  with further:
+The smoothness of the resulting displacements field is induced by a Thikonov regularization
+  prior:
 
   \begin{align*}
+  P(U) = \underset{\vec{r}}{\prod}\, p(u(\vec{r})) &=
+  \underset{\vec{r}}{\prod}\, p_0(u(\vec{r})) \, p_1(u(\vec{r})), \\
   p_0(u(\vec{r})) &= \mathcal{N}( u(\vec{r}) \mid 0, A^{-1}), \notag\\
@@
   \citep{fischl_freesurfer_2012}.
-The transformation from the structural space into the coordinate space of the
-  target image is achieved through a dense deformation field $u(\vec{r})$, such that:
+The transform $\hat{U}$ \eqref{eq:transform} is supported by a dense deformation field
+  $u(\vec{r})$, such that:
 
   \begin{equation}
-  \vec{v}_i' = U\{\vec{v}_i\} = \vec{v}_i + u(\vec{v}_i),
+  \vec{v}_i' = \hat{U}\{\vec{v}_i\} = \vec{v}_i + u(\vec{v}_i).
   \label{eq:nodes_tfm}
@@
 
-  where $U$ is defined in \eqref{eq:transform}.
 Since the nodes of the anatomical surfaces likely lay off-grid, it is required to
@@
 In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
-  of degree 3 ($B_3$).
+  of degree 3 ($\beta_3$).
 Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
@@
     \vec{v}_i' = \vec{v}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
-      B_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
+      \beta_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
   \label{eq:transformation}
@@
 
-{\color{red}
-\paragraph*{Convergence evidence and settings}
- discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition / Wolfe conditions, multi-resolutions on the bspline grid, image samples and surface sampling,  etc.
-\paragraph*{Efficient field densification}
-As long as the dense deformation field is iteratively interpolated
-from the same set of control points that define the $L-1$ contours,
-it is possible to pre-cache all the $\psi_k(\vec{v}_i)$ weights into a
-sparse matrix for fast densification. As well, we achieve a
-diffeomorphic transform by}
+{\color{red} \paragraph*{Evidencing convergence and settings}
+discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition / Wolfe conditions,
+multi-resolutions on the bspline grid, image samples and surface sampling,  etc.
+}
 
@@
 
-\begin{figure}
-\caption{1) The prior surfaces are extracted from an undistorted and high-resolution dataset
-	  (e.g. \autoref{fig:phantom}). 2) A plausible, synthetic distortion $U^{-1}_{true}$ is generated.
-	3) The low-resolution (target) images are warped using the synthetic distortion.
-	4) Our method is run, obtaining a $\hat{U}_{tst}$.
-	  A cross-comparison methodology is also applied, to obtain a competing $\hat{U}_{cc}$.
+\begin{figure*}
+\includegraphics[width=\linewidth]{figures/figure02}
+\caption{Experimental workflow applied on real data from the \gls*{hcp}.
+  1) The prior surfaces are extracted from the anatomical reference (\gls*{t1} image).
+	2) To operate as ground truth, we generate plausible-synthetic distortion $U^{-1}_{true}$
+	  from the fieldmap using \eqref{eq:fieldmap}.
+	3) The \gls*{dmri} data are warped using $U^{-1}_{true}$ to reproduce the effects of real
+	  susceptibility-derived distortions.
+	Target diffusion scalars (\gls*{fa} and \gls*{adc}) are computed on the distorted data and
+		stacked to feed the multivariate input required by our algorithm.
+	4) Registration is run, obtaining a $U_{test} = \hat{U}_{true}$, an estimation of
+	  the ground-truth deformation $U_{true}$.
+	  A cross-comparison methodology is also applied, to obtain a competing $U_{cc}$.
 	5) Results are visually and quantitatively evaluated.}\label{fig:evworkflows}
-\end{figure}
+\end{figure*}
 
@@
 \begin{figure*}
-	\includegraphics[width=\linewidth]{figures/figure02}
+	\includegraphics[width=\linewidth]{figures/figure03}
 	\caption{A. The ``cortex'' phantom is a spherical shape with two sulci and an
@@
 In this case, we derived the deformation $U_{true}^{-1}$ from the field maps released with
-  the corresponding packages of each dataset from the \gls*{hcp},
+  the corresponding packages of each dataset from the \gls*{hcp} applying \eqref{eq:fieldmap},
   mimicking the real distortions, using a derivation of our previous work

### 2015-02-19 Publications/2015-NeuroImage/05-results.tex
finishing changes

@@
   we collected quantitative results on the digital phantoms.
-To warp the simulated datasets, we generated 150 realizations of random and smooth displacement
-  fields $U^{-1}_{true}$.
-Smoothness is achieved using two levels of B-Spline functions, with control points evenly
+To warp the simulated datasets, we randomly generated 150 different realizations of
+  the distortion $U^{-1}_{true}$.
+Smoothness is introduced using two levels of B-Spline functions, with control points evenly
   located in isotropic grids covering the full extent of the phantom.
@@
   for distortions using the \emph{b0}-to-\gls*{t2} registration approach.
+
+\begin{figure*}
+	\includegraphics[width=\linewidth]{figures/figure05.pdf}
+	\caption{Results on real data}\label{fig:results_real}
+\end{figure*}

### 2015-02-19 Publications/2015-NeuroImage/06-discussion.tex
finishing changes

@@
 \label{sec:discussion}
-We propose a joint registration and segmentation algorithm designed for biomedical
-  image analyses that are affected by all or some of the following circumstances:
+We present a variational framework using to simultaneously segment and
+  register 3D \gls*{dmri} of the human brain using the corresponding-anatomical image
+  as reference.
+The aproach is valid for image processing applications in which a precise segmentation is
+  available (from an atlas or corresponding anatomical data) and match any or all
+  the following circumstances:
   i) reference and target images do not present appropriate contrasts to
@@
   	object tracking in time series, etc.).
-The registration method performs the segmentation of the target image in homogeneous
-  and competing regions defined by explicit active contours.
-We chose active contours without edges as we were interested on segmenting homogeneous
-  regions of diffusion properties within the brain.
-Finally, a B-Spline deformation field supports and keeps track of the evolution of the
-  active contours.
-Iteratively, the active contours are updated following the instant shape gradients
-  \citep{herbulot_segmentation_2006} with a gradient descent strategy.
-In section \nameref{sec:methods_background}, we framed our method within the current
-  state-of-art and defined the novelty of our approach.
-
-To support our claim that images can be simultaneously segmented and registered to a
-  reference coordinate system, we conducted evaluation experiments on digital
-  simulated phantoms and real datasets.
 
@@
   to provide with prior information about anatomy with high accuracy.
-\todo[inline]{Add sentence about comment RW\#1 of IPMI2013: incorporate knowledge of
-the EPI distortions into regularization}
 Our method solves in a single step a joint problem usually addressed in two steps.
@@
   \citep{jezzard_correction_1995}.
+
+In section \nameref{sec:methods_background}, we framed our method within the current
+  state-of-art and defined the novelty of our approach.
+The registration method performs the segmentation of the target image in several
+  competing regions defined explicitly by their limiting surfaces.
+We chose active contours without edges as we were interested in the homogeneous
+  regions of diffusion properties within the brain.
+Finally, a B-Spline deformation field supports and keeps track of the evolution of the
+  active contours.
+Iteratively, the active contours are updated following the instant shape gradients
+  (see \nameref{app:shape_priors} in Appendix) with a gradient descent strategy.
+To support our claim that images can be simultaneously segmented and registered to a
+  reference coordinate system, we conducted evaluation experiments on digital
+  simulated phantoms and real datasets.
+
+\todo[inline]{Add sentence about comment RW\#1 of IPMI2013: incorporate knowledge of
+the EPI distortions into regularization}
 

### 2015-02-19 Publications/2015-NeuroImage/08-appendix-00-shapepriors.tex
finishing changes

@@
 
-\subsection*{Instant shape priors}
+\subsection*{Instant velocity of contours}
 \label{app:shape_priors}
+The computation of gradients at the locations of the active contours in the
+  instant $t$ is based on the work of \cite{herbulot_segmentation_2006}.
 Let $F(\vec{r})$ be an ``arbitrary'' function over the image domain
@@
   the projection of the boundary movement on the unit inward normal $N_{\Gamma_{l,m}}$.
-Assuming that the region descriptors $\Theta_l$ vary slowly enough, we can consider
+Assuming that the region descriptors $\{\mu_l, \Sigma_l\}$ vary slowly enough, we can consider
   that $\frac{\partial}{\partial t} F(\vec{r}) = 0$ and thus:
@@
 
-  In our formulation, $F(\vec{r})$ is the difference on
-    sample likelihood between the inner ($\Omega_l$) and the outer ($\Omega_m$) regions.
-  Therefore, we substitute:
+  As inner ($\Omega_l$) and outer ($\Omega_m$) regions are competing and introducing
+    the energy functional \eqref{eq:energy}:
 
@@
     where $\mathcal{E}_l(\vec{r}) = \mathcal{D}^2_{l}(R(\vec{r}))$ for briefness.
+
+\subsection*{Efficient field densification}
+As long as the dense deformation field is iteratively interpolated
+from the same set of control points that define the $L-1$ contours,
+it is possible to pre-cache all the $\psi_k(\vec{v}_i)$ weights into a
+sparse matrix for fast densification. As well, we achieve a
+diffeomorphic transform by

### 2015-02-19 Scripts/pyacwereg/workflows/realdata.py
add usemask to realdata registrations

@@ -3,7 +3,7 @@
 # @Author: oesteban
 # @Date:   2015-01-15 10:47:12
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-02-18 16:18:54
+# @Last Modified time: 2015-02-19 19:17:37
 
 
 def hcp_workflow(name='Evaluation_HCP', settings={}):
@@ -120,7 +120,7 @@ def hcp_workflow(name='Evaluation_HCP', settings={}):
     dti = mrtrix_dti()
     mdti = pe.Node(niu.Merge(2), name='MergeDTI')
 
-    regseg = regseg_wf()
+    regseg = regseg_wf(usemask=True)
     regseg.inputs.inputnode.images_verbosity = 3
     regseg.inputs.inputnode.alpha = [0.0, 0.0, 0.0]
     regseg.inputs.inputnode.beta = [0.0, 0.0, 0.0]

### 2015-02-20 Publications/2015-NeuroImage/04-methods.tex
discussion refactoring

@@
   \citep{nagel_investigation_1986}.
-We provide a proof of duality in {\color{red} [REF figshare?]}.
 
@@
 
-{\color{red} \paragraph*{Evidencing convergence and settings}
-discuss choice of $\tau$, Courant-Friedrichs-Lewy (CFL) condition / Wolfe conditions,
-multi-resolutions on the bspline grid, image samples and surface sampling,  etc.
-}
+\paragraph*{Settings, implementation details, and evidencing convergence}
+\label{sec:conv_report}
+The registration parameters (such as $\tau$, the multi-resolutions on the B-Spline grid,
+  target image smoothing, etc.) and implementation details (for instance the sparse matrix approach
+  to fast interpolation) are dicussed in the {\color{red} Supplemental Materials, section XX.XX}.
+The actual choice of parameter settings are publicly distributed with the source code of the experiments.
+Additionally, we release our software along with a tool for generating convergence reports to
+  demonstrate the behavior of the tool and help scientists configure their experiments.
+A sample of report is found in the {\color{red} Supplemental Materials, section XX.XX}.
 
@@
 For the evaluation of the algorithm on real \gls*{dmri} data of human brains,
-  we collected 15 datasets from the ``minimally preprocessed''
+  we collected 16 datasets from the ``minimally preprocessed''
 	 database of the \gls*{hcp}.
@@
 	6) cortical \gls*{gm} surface.
-The choice of this particular model of homogeneous regions responds to the joint distribution
-  of \gls*{fa} and \gls*{adc} of the human brain, as shown in the {\color{red} Supplemental
-  Materials (Figure XX)}.
+The choice of this particular model is further addressed in \nameref{sec:res_model_and_metric}.
 

### 2015-02-20 Publications/2015-NeuroImage/05-results.tex
discussion refactoring

@@
 \label{sec:results_phantom}
-Using the experimental instrumentation described in section \nameref{sec:experiments_evaluation},
-  we collected quantitative results on the digital phantoms.
+Using a simplified subset of the experimental instrumentation described in section
+  \nameref{sec:experiments_evaluation}, we collected quantitative results on the digital
+  phantoms.
 To warp the simulated datasets, we randomly generated 150 different realizations of
@@
   $\times$ 2 resolutions) were run.
+Parameter settings of every processing step in the experiments are made available along with
+  the source code.
 For each registration, the averaged Hausdorff distance (see \nameref{sec:experiments_evaluation})
@@
   driving the registration process.
-First, we selected the 6-regions model described in \nameref{sec:human_connectome} after
-  visual evaluation of the joint distribution {\color{red} REF Supplemental} of several
-  alternative models for segmenting the target bivariate images (\gls*{fa} and \gls*{adc}).
-Since susceptibility-derived distortions are always aligned with the phase encoding axis
-  and feature the property of being linear with respect to the phase inhomogeneity map,
-  we reduced the dimensionality of the error space to one dimension.
-Thus, the energy functional \eqref{eq:energy} was evaluated for the contour positions
-  obtained along a lineralized space of distortions
+First, we evaluated several alternative models to segment the target bivariate images
+  (\gls*{fa} and \gls*{adc}).
+Based on the observed joint distribution of each (see {\color{red} Supplemental Materials, ...}),
+  we selected the 6-regions scheme described in \nameref{sec:human_connectome}.
+
+The distortions follow expression \eqref{eq:fieldmap} and hence, we easily reduced the
+  error space to one dimension as follows.
+The energy functional \eqref{eq:energy} was evaluated for the warpings given by
+  a lineralized space of distortions
   $\hat{U}\{\vec{r}\} = \epsilon \cdot U_{true} = \vec{r} + \epsilon \cdot u(\vec{r})$,
@@
 The result of this experiment is shown in \autoref{fig:energymap}.
-The metric consistently displayed its minimum at the zero-error point.
+The metric consistently displayed its minimum at $\epsilon=0.0$ (ground-truth location).
 
@@
 
-\paragraph*{Cross-comparison evaluation}\label{sec:res_cc_evaluation}
+\paragraph*{Evaluation \& cross-comparison}\label{sec:res_cc_evaluation}
 %
-Finally, we evaluated the performance of our registration method, and compared it to correcting
-  for distortions using the \emph{b0}-to-\gls*{t2} registration approach.
+Finally, we tested the performance of our registration method, and compared it to a competing
+  method: the \emph{b0}-to-\gls*{t2} registration approach.
+To this end, we integrated the registration software \emph{elastix} \citep{klein_elastix_2010},
+  and reproduced the settings that internally uses \emph{ExploreDTI}
+  \citep{leemans_exploredti_2009}, a widely used toolkit for tractography analysis of
+  \gls*{dti}.
+A dual workflow to the general evaluation framework (\autoref{fig:evworkflows})
+  was also implemented to run the correction experiment with the competing method.
+First, the average \emph{b0} volume is computed using all the \emph{low-b} volumes within
+  the distorted \gls*{dmri} dataset.
+Using the ``preprocessed'' instance of the \gls*{t2} image of the corresponding subject as
+  reference image, and the \emph{b0} image as moving image, we performed the same correction
+  experiment.
 
@@
 	\includegraphics[width=\linewidth]{figures/figure05.pdf}
-	\caption{Results on real data}\label{fig:results_real}
+	\caption{The evaluation instrument automatically renders the overlay of the
+	  target \gls*{fa} image and the resulting contours after distortion correction (yellow color) and
+	  their theoretical location using the ground-truth (green color).
+	The first two rows show several axial slices for \emph{regseg} (``proposed'' row) and the
+	  competing method.
+	The last two rows represent the sagittal view.
+	We intentionally ommited the coronal slicing due to it is the less informative given the properties
+	  of the distortion.
+	Red arrows point to regions where the accuracy of the proposed method more clearly overperformed
+	  the competing method.
+	}\label{fig:results_real}
 \end{figure*}

### 2015-02-20 Publications/2015-NeuroImage/06-discussion.tex
discussion refactoring

@@
 \label{sec:discussion}
-We present a variational framework using to simultaneously segment and
-  register 3D \gls*{dmri} of the human brain using the corresponding-anatomical image
-  as reference.
+In section \nameref{sec:methods_background}, we framed our method within the current
+  state of art and defined the contributions in our approach.
+We chose active contours without edges as we were interested in the homogeneous
+  regions of diffusion properties within the brain.
+Iteratively, the active contours are updated following the instant shape gradients
+  (see \nameref{app:shape_priors} in Appendix) with a gradient descent strategy.
 The aproach is valid for image processing applications in which a precise segmentation is
-  available (from an atlas or corresponding anatomical data) and match any or all
+  available (from an atlas or within subject anatomical data), and match any or all
   the following circumstances:
@@
   	object tracking in time series, etc.).
+To support our claim that images can be simultaneously segmented and registered to a
+  reference coordinate system, we conducted evaluation experiments on digital
+  simulated phantoms and real datasets.
 
@@
   networks.
-As a consequence, our approach relies on the availability of precise segmentations or
-  surfaces of the objects that are to be segmented on the target volume(s).
 Connectome extraction protocols generally include the acquisition of \gls*{t1} images
@@
 Our method solves in a single step a joint problem usually addressed in two steps.
-The first solves the linear registration of the \gls*{t1} and \gls*{dmri} data (i.e.
-  \cite{greve_accurate_2009}), whereas the second corrects for the nonlinear distortions
+The first solves the linear registration of the \gls*{t1} and \gls*{dmri} data
+  \citep{greve_accurate_2009}, whereas the second corrects for the nonlinear distortions
   derived from the inhomogenety of magnetic susceptibility across tissues
@@
 
-In section \nameref{sec:methods_background}, we framed our method within the current
-  state-of-art and defined the novelty of our approach.
-The registration method performs the segmentation of the target image in several
-  competing regions defined explicitly by their limiting surfaces.
-We chose active contours without edges as we were interested in the homogeneous
-  regions of diffusion properties within the brain.
-Finally, a B-Spline deformation field supports and keeps track of the evolution of the
-  active contours.
-Iteratively, the active contours are updated following the instant shape gradients
-  (see \nameref{app:shape_priors} in Appendix) with a gradient descent strategy.
-To support our claim that images can be simultaneously segmented and registered to a
-  reference coordinate system, we conducted evaluation experiments on digital
-  simulated phantoms and real datasets.
+\paragraph*{Prospects}
+Future extensions or improvements of \emph{regseg} similar to the evolution of existing
+  alternative methodologies.
+Incorporating knowledge about the distortion initializing our method with the theory-based
+  correction using fieldmaps would be the most profitable.
+The main pitfall using fieldmap correction is the challenging problem of unwrapping the
+  range of the phase map, as it is clipped in the range $(-\pi, \pi]$.
+The most extended solution to this is \emph{prelude} \citep{jenkinson_fast_2003}, and some
+  alternatives have been proposed \citep{daga_susceptibility_2014} to circumvent the
+  inherent unreliability of the method.
+The \emph{regseg} algorithm would naturally correct missregistered areas due to errors in
+  phase unwrapping, and errors derived to the important smoothing processes that are
+  typically to clean up field maps.
+This integration has been recently applied in the \emph{b0}-to-\gls*{t2} approach
+  by \cite{irfanoglu_susceptibility_2011}, studied along with the virtues of using
+  nonregular B-Spline grids.
+This is also a potential extension to our deformation model.
 
-\todo[inline]{Add sentence about comment RW\#1 of IPMI2013: incorporate knowledge of
-the EPI distortions into regularization}
-
-\paragraph*{Availability and reproducibility statement}
-{\color{red} We also considered the reproducibility of results a design requirement.
-Therefore, real data are fetched from a publicly available repository
-  (the Human Connectome Project \citep{essen_human_2012}) and all the software
-  involved in this paper is also publicly released.
-\autoref{fig:evworkflows} describes the general structure of the workflow we implemented
-  as evaluation instrument, using \emph{nipype} \citep{gorgolewski_nipype_2011} to ensure
-  reproducibility.
-
-We implemented the registration method upon the widely used Insight Registration and Segmentation
-	Toolkit (\url{http://www.itk.org}) with the aim to release a useful and free software bundle.
-The instrumentation framework is also distributed with the registration method,
-  and has been implemented using \emph{nipype} \citep{gorgolewski_nipype_2011}.
-The pipelines include the automatic generation of the ``cortex'' phantom.
-The real datasets are publicly available under the \gls*{hcp} project.
-Thus, all the experiments and results presented in this paper are intended to be
-  fully reproducible.}
+Finally, we are currently studying substitute features to build the energy functional on,
+  enabling to perform \emph{regseg} directly on the raw \gls*{dmri} data.

### 2015-02-20 Publications/2015-NeuroImage/07-conclusion.tex
Update all the sections
discussion refactoring

@@
 % -*- root: 00-main.tex -*-
-\section{Conclusion}
+\section*{Conclusion}
 \label{sec:conclusion}
-%
-We propose and make publicly available a joint
-  segmentation-registration method with a proof of concept
-  on a digital phantom and an application on diffusion \gls*{mri}
-  images of the human brain.
-Starting with the Bayesian approach to image segmentation, we derive
-  a cost function that models piecewise smooth regions with
-  multivariate normal distributions.
-This is, implicitly, an \acrlong*{acwe} model, in which the
-  evolution of the surfaces drives the optimization of a B-Spline
-  deformation model.
 
-The phantom is used to demonstrate the success of the approach,
-  on a quantitative evaluation that demonstrated its performance
-  with sub-voxel precision (average vertex displacement close to
-  $0.7mm$ in phantoms with isometric pixels of $1.0mm$ and $2.0mm$
-  side).
-We apply the method on the correction of susceptibility-derived
-  distortions of diffusion \gls*{mri} data, and compare to
-  a widely-used method based on the acquisition of extra images
-  mapping the inhomogeneity of tissue susceptibility inside
-  the scanner.
+\emph{Regseg} is a variational framework to simultaneously segment and
+  register 3D \gls*{dmri} data of the human brain using within subject
+  anatomical information as reference.
+The registration method segments the target multivariate image in several competing regions
+  defined explicitly by their limiting surfaces.
+The surfaces are active, and evolve on a B-Spline displacements field, guided by shape
+  gradients computed on the current partition of the target image.
+\emph{Regseg} uses active contours without edges, and hence looks for
+  homogeneous regions within the image.
+We tested \emph{regseg} on digital phantoms simulating \gls*{t1} and \gls*{t2} \gls*{mri},
+	warped with smooth random deformations.
+The resulting average misplacement of the contours was significantly lower than the
+  image resolution of the phantoms.
+
+We propose \emph{regseg} to correct \gls*{dmri} datasets for the typically present distortions
+  caused by inhomogeneitis of the main magnetic field.
+Visual assessment of \emph{regseg} and cross-comparison with a widely-used competing
+  technique demonstrated its accuracy, and overperformed the competing method.
+Moreover, \emph{regseg} does not require additional images to the minimal acquisition protocol
+  that generally includes only \gls*{t1} and \gls*{dmri}.
+
+\section*{Availability and reproducibility statement}
+\label{sec:availability}
+We considered the reproducibility of our results a design requirement.
+Therefore, we used real data fetched from a publicly available repository
+  (the Human Connectome Project \citep{essen_human_2012}) and all the software
+  involved in this paper is also publicly released.
+The \emph{regseg} algorithm is implemented in C++, on top of ITK-4.6
+  (Insight Registration and Segmentation Toolkit, \url{http://www.itk.org}).
+The evaluation instruments (\autoref{fig:evworkflows}), are implemented in Python using
+  \emph{nipype} \citep{gorgolewski_nipype_2011} to assess their reproducibility.
+Phantom data are automatically generated by the evaluation framework.
+All research elements (data, source code, figures, manuscript sources, etc.) in this paper
+  are made publicly available under a unique digital object identifier in a
+  package \citep{esteban_acweregistration_2015}.

### 2015-02-20 Publications/2015-NeuroImage/08-appendix-00-shapepriors.tex
discussion refactoring

@@
     where $\mathcal{E}_l(\vec{r}) = \mathcal{D}^2_{l}(R(\vec{r}))$ for briefness.
-
-\subsection*{Efficient field densification}
-As long as the dense deformation field is iteratively interpolated
-from the same set of control points that define the $L-1$ contours,
-it is possible to pre-cache all the $\psi_k(\vec{v}_i)$ weights into a
-sparse matrix for fast densification. As well, we achieve a
-diffeomorphic transform by

### 2015-02-20 Publications/2015-NeuroImage/04-methods.tex
add supplementals

@@
 
-\paragraph*{Settings, implementation details, and evidencing convergence}
+\paragraph*{Settings, implementation details, and convergence}
 \label{sec:conv_report}
-The registration parameters (such as $\tau$, the multi-resolutions on the B-Spline grid,
+The registration parameters (such as $\tau$, the B-Spline grid resolutions,
   target image smoothing, etc.) and implementation details (for instance the sparse matrix approach
-  to fast interpolation) are dicussed in the {\color{red} Supplemental Materials, section XX.XX}.
+  to fast interpolation) are dicussed in the \suppl{}, {\color{red} section S1}.
 The actual choice of parameter settings are publicly distributed with the source code of the experiments.
@@
   demonstrate the behavior of the tool and help scientists configure their experiments.
-A sample of report is found in the {\color{red} Supplemental Materials, section XX.XX}.
+A sample of report is found in the \suppl{} {\color{red}, section S1.3}.
 

### 2015-02-20 Publications/2015-NeuroImage/05-results.tex
add supplementals

@@
   (\gls*{fa} and \gls*{adc}).
-Based on the observed joint distribution of each (see {\color{red} Supplemental Materials, ...}),
+Based on the observed joint distribution of each (\suppl{}, {\color{red} section S2}),
   we selected the 6-regions scheme described in \nameref{sec:human_connectome}.

### 2015-02-20 Publications/2015-NeuroImage/supplemental.tex
add supplementals

% @Author: Oscar Esteban
% @Date:   2015-02-20 15:40:09
% @Last Modified by:   Oscar Esteban
% @Last Modified time: 2015-02-20 16:27:53

\documentclass[a4paper,notitlepage]{report}

\usepackage{natbib}
%\usepackage[style=numeric-comp,doi=false]{biblatex}

% *** GRAPHICS RELATED PACKAGES ***
\usepackage[table]{xcolor}
\usepackage{graphicx}
\usepackage{ifpdf}
\graphicspath{{figures/},{../Figures/}}

% *** MATH PACKAGES ***
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}

\usepackage[T1]{fontenc}
\usepackage{charter}
%\usepackage[expert]{mathdesign}
%\usepackage[libertine,cmintegrals,cmbraces,vvarbb]{newtxmath}

% *** SPECIALIZED LIST PACKAGES ***
%\usepackage{algorithmic}
%\usepackage{algorithm}

% *** ALIGNMENT PACKAGES ***
%\usepackage{array}

% *** SUBFIGURE PACKAGES ***
\usepackage[font={small}]{caption}
%\usepackage{sidecap}

% *** FLOAT PACKAGES ***
\usepackage[framemethod=tikz]{mdframed}
% \usepackage{fixltx2e}
% \usepackage{stfloats}
% \usepackage{dblfloatfix}

% *** MISC UTILITY PACKAGES ***
%\usepackage{fancyhdr}
\usepackage{multicol}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{booktabs} % to use \toprule
\usepackage{makeidx}
%\usepackage[scale=2.0]{ccicons}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[marginparwidth=1.2in]{geometry}
\usepackage{changepage}
\usepackage{epigraph}
\usepackage[toc,nomain,acronym,shortcuts,translate=false]{glossaries}
\usepackage[hyphens]{url}
\usepackage[breaklinks,hidelinks]{hyperref}
\usepackage[hyphenbreaks]{breakurl} %fixes boxes spanning through pages

\renewcommand\thesection{S\arabic{section}}
\renewcommand\thesubsection{S\arabic{section}.\arabic{subsection}}

\begin{document}
\title{Supplemental Materials: \emph{Simultaneous segmentation and registration of
diffusion MR images of the brain driven by active-contours}}
\author{Oscar Esteban}
\date{February 2015}

\maketitle

\section{Parameter settings and implementation details of \emph{regseg}}

\subsection{Settings}

\subsection{Implementation}

\subsection{Convergence evidencing}

\section{Instruments for evaluation}

\section{Model considerations}

\begin{figure*}[h!]
	\includegraphics[width=\textwidth]{figures/Suppl-figure01.pdf}
	\caption{Evaluating the joint distribution}\label{fig:jointplot}
\end{figure*}

\section{Extended results on phantom data}

\section{Extended results on real data}

\end{document}

### 2015-02-20 Scripts/pyacwereg/viz.py
add jointplot_real

@@ -3,7 +3,7 @@
 # @Author: oesteban
 # @Date:   2014-12-11 15:08:23
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-02-19 10:46:19
+# @Last Modified time: 2015-02-20 15:29:00
 
 
 def add_annotations(values, ax, level, nlevels, color, lastidx, units=''):
@@ -324,6 +324,62 @@ def jointplot_data(im1data, im2data, in_seg, labels=None, out_file=None,
 
     plt.savefig(out_file, dpi=300, bbox_inches='tight')
     return g, out_file
+
+
+def jointplot_real(imageX, imageY, segmentation, mask,
+                   labels=None, xlabel='X', ylabel='Y',
+                   huelabel='tissue',
+                   xlims=None, ylims=None, out_file=None):
+    import os.path as op
+    import nibabel as nb
+    import numpy as np
+    import matplotlib.pyplot as plt
+    import seaborn as sn
+    import pandas as pd
+
+    fa = nb.load(imageX).get_data().reshape(-1)
+    md = nb.load(imageY).get_data().reshape(-1)
+    seg = nb.load(segmentation).get_data().astype(np.uint8).reshape(-1)
+    msk = nb.load(mask).get_data().astype(np.uint8).reshape(-1)
+
+    if xlims is None:
+        xlims = (fa.min(), fa.max())
+
+    if ylims is None:
+        ylims = (md.min(), md.max())
+
+    md[md < ylims[0]] = np.nan
+    md[md > ylims[1]] = np.nan
+
+    if labels is None:
+        labels = ['Label%02d' % i for i in range(seg.max() + 1)]
+
+    df_pieces = []
+    for i, l in zip(np.unique(seg), labels):
+        idxs = np.where((seg == i) & (msk > 0))
+        fa_f = fa[idxs]
+        md_f = md[idxs]
+        d = {xlabel: fa_f,
+             ylabel: md_f,
+             huelabel: [l] * len(fa_f)}
+        df_pieces.append(pd.DataFrame(d))
+
+    df = pd.concat(df_pieces)
+    df = df[df.tissue != 'do-not-show']
+    sn.set_context("talk", font_scale=1.8)
+
+    g = sn.JointGrid(xlabel, ylabel, df, hue=huelabel, xlim=xlims, ylim=ylims,
+                     size=20, inline_labels=True)
[…]

### 2015-02-23 Publications/2015-NeuroImage/supplemental.tex
update paper

@@
 % @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-02-20 16:27:53
+% @Last Modified time: 2015-02-23 13:43:56
 
-\documentclass[a4paper,notitlepage]{report}
+\documentclass[a4paper]{report}
+
+\usepackage{titlesec}
+\newcommand{\sectionbreak}{\clearpage}
+
+\usepackage{aliascnt}
+\usepackage{relsize}
+\usepackage[procnames]{listings}
 
@@
 \usepackage[framemethod=tikz]{mdframed}
+%\usepackage{float}
 % \usepackage{fixltx2e}
@@
 
+% Listings style -----------------------------------------------------------------------------
+\definecolor{pink}{RGB}{255,0,90}
+\definecolor{comments}{RGB}{50,120,110}
+\definecolor{string}{RGB}{160,0,0}
+\definecolor{keywords}{RGB}{0,150,0}
+\def\listingsfont{\ttfamily}
+\def\listingsfontinline{\ttfamily}
+
+\lstnewenvironment{bashcode}[1][]
+  {\lstset{language=bash}\lstset{%
+  showstringspaces=false,
+  formfeed=\newpage,
+  tabsize=4,
+  breaklines=true,
+  basicstyle=\ttfamily\smaller\relax,
+  keywordstyle=\color{keywords}\bfseries,
+  commentstyle=\color{comments}\itshape,
+  stringstyle=\color{string},
+  showstringspaces=false,
+  %identifierstyle=\color{green},
+  procnamekeys={def,class},
+  morekeywords={models, lambda, forms, as, from},
+  numbers=none,
+  frame=single,
+  xleftmargin= 10pt,
+  xrightmargin= 10pt,
+  framexleftmargin=10pt,
+  frameround=tttt,
+  fillcolor=\color{gray!10},
+  backgroundcolor=\color{gray!10}
+}}
+{}
+
+\lstnewenvironment{pythoncode}[1][]
+  {\lstset{language=bash}\lstset{%
+  showstringspaces=false,
+  formfeed=\newpage,
+  tabsize=4,
+  breaklines=true,
+  basicstyle=\ttfamily\smaller\relax,
+  keywordstyle=\color{keywords}\bfseries,
+  commentstyle=\color{comments}\itshape,
+  stringstyle=\color{string},
+  showstringspaces=false,
+  %identifierstyle=\color{green},
+  procnamekeys={def,class},
+  morekeywords={models, lambda, forms, as, from},
+  numbers=left,
+  numberstyle=\smaller\color{black!60},
+  stepnumber=1,
+  numbersep=5pt,
+  frame=single,
+  xleftmargin= 40pt,
+  xrightmargin= 20pt,
+  framexleftmargin=20pt,
+  frameround=tttt,
+  fillcolor=\color{gray!10},
+  backgroundcolor=\color{gray!10}
+}}
+{}
+
 \begin{document}
@@
 
-\subsection{Settings}
+\subsection{Implementation}
+\paragraph{General} The \emph{regseg} registration and segmentation tool is written in C++, using ITK-4.6 as
+  implementation core.
+We designed a modular implementation, enabling multithreading in several pieces of the software,
+  as the process is computationally expensive.
+The tool generates a log-file in JSON format to easily inter-operate with secondary tools (such
+  as the convergence report generation, \autoref{sec:convergence_evidence}).
 
-\subsection{Implementation}
+\paragraph{Efficient interpolation using sparse matrices}
+During the registration process, every iteration requires computing the product of all the speeds
+  $\mathbf{s}'_i$ computed at the current position $\mathbf{v}'_i$ by the corresponding weights
+  $\psi_k(\mathbf{v}_i)$ of interpolating functions (Eq. (15)).
+As these weights can be computed once in the beginning of the process and they do not change along
+  it, and they are different from zero only in the proximity of the corresponding control node $k$,
+  they can be pre-cached in a sparse matrix.
 
-\subsection{Convergence evidencing}
+\subsection{Interface and Settings}\label{sec:interface_settings}
+
+\paragraph{Command-line interface}
+The command line interface of \emph{regseg} supports general settings and level-wise settings.
+For each multi-resolution level, its corresponding settings are added between brackets.
+
+\begin{bashcode}
+regseg -F fa.gz adc.nii.gz -P white.vtk pial.vtk -o myprefix [ -a 0.00000 -b 0.00000 --convergence-energy -t 1.0e-06 -w 60 --adaptative-descriptors --grid-spacing 16.0 -i 500 -s 0.001] [ -a 0.
+-b 0.
+--convergence-energy -t 1.e-08 -w 5 --grid-spacing 8.0 -i 250 -s 0.01]
+\end{bashcode}
+
+It is possible to get the description of available options running
+  \lstinline!regseg -h!:
+
+\begin{bashcode}
+Usage:
+
+General options:
+  -h [ --help ]                         show help message
+  -F [ --fixed-images ] arg             fixed image file
+  -P [ --surface-priors ] arg           shape priors
+  -T [ --surface-target ] arg           final shapes to evaluate metric (only
+                                        testing purposes)
+  -M [ --fixed-mask ] arg               fixed image mask
+  -L [ --transform-levels ] arg         number of multi-resolution levels for
+                                        the transform
+  -o [ --output-prefix ] arg (=regseg)  prefix for output files
+  -l [ --logfile ] arg                  log filename
+  -v [ --monitoring-verbosity ] arg (=1)
+                                        verbosity level of intermediate results
+                                        monitoring ( 0 = no output; 5 = verbose
+                                        )
+
+Optimizer options (by levels):
+  -a [ --alpha ] arg              alpha value in regularization
+  -b [ --beta ] arg               beta value in regularization
+  -s [ --step-size ] arg          step-size value in optimization
+  -g [ --gradient-scales ] arg    alpha value in regularization
+  -r [ --learning-rate ] arg      learning rate to update step size
+  -i [ --iterations ] arg         number of iterations
+  -w [ --convergence-window ] arg number of iterations of convergence window
+  -t [ --convergence-thresh ] arg convergence value
+  --grid-size arg                 size of control points grid
+  --grid-spacing arg              spacing between control points
+  -u [ --update-descriptors ] arg frequency (iterations) to update descriptors
+                                  of regions (0=no update)
+  --adaptative-descriptors        recomputes descriptors more often at the
+                                  beginning of the process
+  --convergence-energy            disables lazy convergence tracking: instead
+                                  of fast computation of the mean norm of the
+                                  displacement field, it computes the full
+                                  energy functional
+
+Functional options (by levels):
+  --smoothing arg               apply isotropic smoothing filter on target
+                                image, with kernel sigma=S mm.
+  --smooth-auto                 apply isotropic smoothing filter on target
+                                image, with automatic computation of kernel
+                                sigma.
+  --uniform-bg-membership       consider last ROI as background and do not
+                                compute descriptors.
+  -d [ --decile-threshold ] arg set (decile) threshold to consider a computed
+                                gradient as outlier (ranges 0.0-0.5)
+
+\end{bashcode}
+
+\paragraph{\emph{Nipype} interface}
+Our registration algorithm is released with a \emph{nipype Interface} packaged in
+  \lstinline!pyacwereg.interfaces.acwereg!.
+This interface has been comprehensively used in the evaluation workflows.
+
+\begin{pythoncode}
+from pyacwereg.interfaces.acwereg import ACWEReg
+regseg = ACWEReg()
+regseg.inputs.in_fixed = ['T1w.nii.gz', 'T2w.nii.gz']
+regseg.inputs.in_pior = ['csf.vtk', 'white_lh.vtk', 'white_rh.vtk',
+                         'pial_lh.vtk', 'pial_rh.vtk']
+ifresult = regseg.run()
+\end{pythoncode}
+
+\subsection{Convergence evidencing}\label{sec:convergence_evidence}
+In order to track the evolution of the registration process, several internal variables
+  are saved in the JSON log-file.
+Using the JSON log-file as input for the \emph{nipype Interface}
+  \lstinline!pyacwereg.interfaces.ACWEReport!, it is straightforward to obtain
+  a visual assessment document presenting the convergence.
+Online checking is also possible as the algorithm writes to the standard output as well.
+
+\begin{pythoncode}
+from pyacwereg.interfaces.acwereg import ACWEReport
+report = ACWEReport()
+report.inputs.in_log = `myprefix.log'
+ifresult = report.run()
+\end{pythoncode}
+
+\begin{figure*}[!ht]
+	\includegraphics[width=\textwidth]{figures/Suppl-figure02.pdf}
+	\caption{The evolution of the registration and segmentation process can be
+	  checked using the \emph{Convergence report},
+	  easily generated using the appropriate \emph{nipype Interface}.
+	The report comprehends several plots tracking the evolution of the algorithm and several
+	  features to help researchers tune up the algorithm in their application.}%
+	  \label{fig:convreport}
+\end{figure*}
 
@@
 
-\begin{figure*}[h!]
+\begin{figure*}[!ht]
 	\includegraphics[width=\textwidth]{figures/Suppl-figure01.pdf}

### 2015-02-27 Publications/2015-NeuroImage/03-introduction.tex
paper version candidate

@@
   illustrates the wide variety of existing registration methodologies.
-They classify the algorithms by the three principal building blocks that can be identified
+They classified the algorithms by the three principal building blocks that can be identified
   among methods: the matching criteria or similarity function to evaluate the proximity of
@@
   of the mapping in the selected application, and the optimization method or search strategy.
-The possibly most widespread matching criteria are those classified as \emph{iconic
+The possibly most widespread matching criteria are referred to as \emph{iconic
   methods} in \citep{sotiras_deformable_2013}.
@@
   content of both images.
-A second group, called \emph{geometric methods}, align geometrical features derived from
+A second group, called \emph{geometric methods}, align spatial features derived from
   corresponding objects defined in both spaces, such as landmarks, shapes, surfaces, etc.
@@
 \Gls*{dmri} data are usually acquired with \gls*{epi} schemes.
-In order to quickly probe  in many different orientations the diffusion process within the living
-  brain, these sequences minimize the bandwidth of the \gls*{pe} direction.
+In order to quickly probe the diffusion process within the brain in many different orientations,
+  these sequences minimize the bandwidth of the \gls*{pe} direction.
 This low bandwith originates the artifactual displacement of locations presenting small deviations
   from the main magnetic field $B_0$.
-The principal source of these deviations, $\Delta B_0$, is the discontinuity of magnetic
+The principal source of these deviations ($\Delta B_0$) is the discontinuity of magnetic
   susceptibility at tissue interfaces.
@@
   particularly evident in the surroundings of the orbito-frontal lobe or the temporal
-  bones in both hemispheres.
+  bone of both hemispheres.
 The displacements field $U_s$ along the \gls*{pe} axis can be estimated theoretically
@@
 
-where $\vec{r}$ is the location in the reference domain as in \eqref{eq:transform},
+where $\vec{r}$ is the true location in the undistorted domain as in \eqref{eq:transform},
   $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time,
@@
 
-Our method is related to the \gls*{t2b} methods as we also render the correction as a
+\emph{Regseg} is related to the methods based on nonlinear registration of the \emph{b0}
+  image\footnote{The term \emph{b0} refers to one or more images within the
+  \gls*{dmri} dataset that are acquired with a null or very low gradient intensity or
+   \emph{b}-value used as signal reference.}
+  to the \gls*{t2} image, since we also render the correction as a
   registration problem.
@@
 A fundamental antecedent to the present approach is found in \citep{greve_accurate_2009}.
-They proposed a registration method to align \gls*{t1} and \gls*{dmri} images
-  of the brain using a linear transformation model.
-Similarly, we exploit the precise structures that can be extracted with
-  widely-used software (i.e. \emph{FreeSurfer}~\citep{fischl_freesurfer_2012}),
-  and then map these surfaces into warped data, in which the scale of the
+They proposed a registration method to align contours extracted from the \gls*{t1}
+  to \gls*{dmri} images using a linear transformation model.
+Similarly, we exploit the precise surfaces that can be extracted with
+  \emph{FreeSurfer}~\citep{fischl_freesurfer_2012},
+  and then map this prior into warped data, in which the scale of the
   imaged structures is in the range of the image resolution or above.
@@
   susceptibility-derived distortions \citep{jezzard_correction_1995}.
+\cite{greve_accurate_2009} circumvented the nonlinearity by identifying tipically
+  distorted regions and removing them from the metric computation.
 Second, our matching criteria does not necessarily need for clear intensity
@@
   regions in one image corresponding to a prior knowledge of the shape of objects in
-  the moving coordinate system.
+  the moving (distorted) coordinate system.

### 2015-02-27 Publications/2015-NeuroImage/04-methods.tex
paper version candidate

@@
 % -*- root: 00-main.tex -*-
-
-%{\color{red} {The main diverging points with respect to
-%\citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
-%$\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
-%gradients \citep{besson_dream2s_2003,herbulot_segmentation_2006};
-%2) regularization is also based on linear diffusion smoothing \citep{thirion_image_1998},
-%but we replace the Gaussian filtering by other constraints
-%studied in \citep{nagel_investigation_1986} to better the problem;
-%3) optimization is applied in the spectral
-%domain, observing anisotropic and inhomogeneous mappings along each direction.
-%With respect \citep{guyader_combined_2011}, the main differences are
-%the distance function, and the spectral solution to the optimization updates,
-%as we shall cover in \autoref{sec:numerical_implementation}.}}
-
-%In this paper we formulate the joint registration-segmentation
-%problem as follows.
-%such that the known contours in anatomical space $T$ optimally segment
-%the diffusion space $D$.
-%
-%
-%Whereas related \glspl*{adf} introduced in \autoref{sec:methods_background}
-%make use of explicit level-set formulations to solve \eqref{eq:gradient_descent},
-%we alternatively use \emph{shape-gradients}
-%\cite{besson_dream2s_2003,herbulot_segmentation_2006}.
-
-%
-%Let us denote by $\vec{x}$ the voxel and $F(\vec{x}) = [ f_1, f_2, \ldots, f_N]^T$
-%  its associated feature vector in the following.
-%
-%In our application, these surfaces are precise tissue interfaces of interest extracted
-%  from a high-resolution, anatomically correct reference volume using
-%  the well-established
-%
-%In this paper we propose a novel registration framework to simultaneously
-%solving the segmentation, distortion and cortical parcellation challenges,
-%by exploiting as strong shape-prior the detailed morphology extracted
-%from high-resolution and anatomically correct \gls{mri}.
-%Indeed, hereafter
-%we assume this segmentation problem in anatomical images is reliably and
-%accurately solved with readily available tools (e.g.
-%\citep{fischl_freesurfer_2012}).
-%After global alignment with \gls{t1} using existing approaches, the remaining
-%spatial mismatch between anatomical and diffusion space is due to susceptibility
-%distortions.
-%Finally, we need to establish precise spatial correspondence between the
-%surfaces in both spaces, including the tangential direction for parcellation.
-%Therefore, we can reduce the problem to finding the differences of spatial
-%distortion in between anatomical and \gls{dwi} space.
-%We thus reformulate the segmentation problem as an inverse problem, where we
-%seek for an underlying deformation field (the distortion) mapping
-%from the structural space into the diffusion space, such that the structural
-%contours segment optimally the \gls{dwi} data.
-%In the process, the one-to-one
-%correspondence between the contours in both spaces is guaranteed, and projection
-%of parcellisation to \gls{dwi} space is implicit and consistent.
-%
-%We test our proposed joint segmentation-registration model on two different
-%synthetic examples.
-%The first example is a scalar sulcus model, where the
-%\gls{csf}-\gls{gm} boundary particularly suffers from \gls{pve} and can only be
-%segmented correctly thanks to the shape prior and its coupling with the inner,
-%\gls{gm}-\gls{wm} boundary through the imposed deformation field regularity.
-%The second case deals with more realistic \gls{dwi} data stemming from
-%phantom simulations of a simplistic brain data.
-%Again, we show that the
-%proposed model successfully segments the \gls{dwi} data based on two derived
-%scalar features, namely \gls{fa} and \gls{md}, while establishing an estimate
-%of the dense distortion field.
-%
-%The rest of this paper is organized as follows.
-%First, in \autoref{sec:methods}
-%we introduce our proposed model for joint multivariate segmentation-registration.
-%Then we provide a more detailed description of the data and experimental setup in
-%\autoref{sec:experiments}.
-%We present results in \autoref{sec:results} and conclude
-%in \autoref{sec:conclusion}.
-%
-%
-%The properties of the reconstructed tensors and derived scalar maps have
-%been studied by \cite{ennis_orthogonal_2006}. Based on their
-%findings, \gls{fa}~\eqref{eq:fa} and \gls{md}~\eqref{eq:md} are
-%considered complementary features, and therefore we selected them for the
-%energy model \eqref{eq:complete_energy} in driving the
-%registration-segmentation process.
-%Whereas \gls{fa} informs mainly about the \emph{shape} of diffusion,
-%the \gls{md} is more related to the \emph{magnitude} of the process:
-%
-%\begin{align}
-%\mathrm{FA} &= \sqrt{ \frac{1}{2}}\,\frac{\sqrt{ (\lambda_1 - \lambda_2)^2 + (\lambda_2 - \lambda_3)^2 + (\lambda_3 - \lambda_1)^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}} \label{eq:fa} \\
-%\mathrm{MD} &= ( \lambda_1 + \lambda_2 + \lambda_3 ) / 3 \label{eq:md}
-%\end{align}
-%where $\lambda_i$ are the eigenvalues of the diffusion tensor
-%associated with the diffusion signal $S(\vec{q})$. There exist
-%two main reasons to justify their choice.
-%First, they are well-understood and standardized in clinical routine.
-%Second, together they contain most of the information that is
-%usually extracted from the \gls{dwi}-derived scalar maps
-%\cite{ennis_orthogonal_2006}.
-%
-
 \section*{Methods}
@@
 \label{sec:methods_background}
-Registration-segmentation methods are generally built including a deformation model to
+Registration-segmentation methods are generally built on a deformation model to
   support the evolution of a segmentation algorithm that drives the mapping process.
@@
   \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014}.
-In \autoref{sec:methods_map} we start from a probabilistic framework to finally propose
+In section \nameref{sec:methods_map}, we start from a probabilistic framework to finally propose
 	a cost function defined on the location of explicit surfaces, showing the duality of the
-	existing approaches.
+	two existing approaches.
 
@@
   and reference images with an affine transformation supporting the coordinates mapping.
-\cite{vemuri_joint_2003} proposed an atlas-based registration using a level sets
-  framework and only one \gls*{pde} for first.
-The principal application of these techniques was the image registration between timesteps in
-  a time-series or between different images \citep{paragios_level_2003}.
+\cite{vemuri_joint_2003} proposed an atlas-based registration using a level-sets
+  framework and only one \gls*{pde} for first time.
+The principal application of these techniques was to perform registration between timesteps in
+  time-series or between different images \citep{paragios_level_2003}.
 \cite{unal_coupled_2005} and later \cite{wang_joint_2006},
@@
 Finally, \cite{gorthi_active_2011} presented a comprehensive generalization of the
-  existing methodologies based on the level sets approximation in the application of
-  atlas-based segmentation.
+  existing methodologies based on level sets in the application of atlas-based segmentation.
 Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
@@
 Whereas \cite{guyader_combined_2011} proposed a nonlinear elasticity smoother, we rely on
-  the smoothing effects of the deformation field and the anisotropic regularization
+  the inherent smoothness of the deformation field, and on the anisotropic regularization
   proposed by \cite{nagel_investigation_1986}.
@@
   to the descriptors of region $l$ as
-  $\mathcal{E}_l(\vec{r}) = \mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
+  $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
 
@@
   \underset{\vec{r}}{\prod}\, p_0(u(\vec{r})) \, p_1(u(\vec{r})), \\
-  p_0(u(\vec{r})) &= \mathcal{N}( u(\vec{r}) \mid 0, A^{-1}), \notag\\
-  p_1(u(\vec{r})) &= \mathcal{N}(  \nabla \cdot u(\vec{r}) \mid 0, B^{-1}),
+  p_0(u(\vec{r})) &= \mathcal{N}( u(\vec{r}) \mid 0, \mathbf{A}^{-1}), \notag\\
+  p_1(u(\vec{r})) &= \mathcal{N}(  \nabla \cdot u(\vec{r}) \mid 0, \mathbf{B}^{-1}),
   \end{align*}
@@
   imposing that the distortion and its gradient have zero
-  mean and variance governed by $A$ and $B$.
+  mean and variance governed by the matrices $\mathbf{A}$ and $\mathbf{B}$.
 Since the anisotropy is generally aligned with the imaging axes, these will be simplified
@@
 
-Finally, the \gls{map} problem is adapted into a variational one applying the
-  following log-transform:
+Finally, the maximum a posteriori problem is adapted into a variational one where we look for
+  the minimum of an energy functional, by applying a log-transform:
 
@@
 
-  that is the dual expression to the energy functional corresponding
+  this expression is dual to the energy functional corresponding
   to a discrete \gls*{acwe} framework \citep{chan_active_2001}
-  with anisotropic regularization as studied in
-  \citep{nagel_investigation_1986}.
+  with the corresponding anisotropic regularization term \citep{nagel_investigation_1986}.
 
@@
 	\includegraphics[width=\linewidth]{figures/figure01}
-	\caption{The active contours are defined as the interfacing surfaces of the segmenting
-	\glspl{roi} $\Omega_i$, and represented in green and dark blue colors in this
-	close-up.
-	They iteratively through their normals $\hat{n}_i$ at each vertex $\vec{v}_i$ of the mesh.
+	\caption{The active contours are defined as the interfacing surfaces of the competing
+	  \glspl{roi} $\Omega_i$.
+	They are represented in green and dark blue colors in this close-up.
+	They iteratively evolve following their inner normals $\hat{n}_i$ at each vertex
+	  $\vec{v}_i$ of the mesh.
 	The gradient speeds $\bar{s}_i$ are computed as the disparity between data energies of
-	the vertex $\vec{v}_i$ in each limiting region as described in \eqref{eq:energy}.
+	  the vertex $\vec{v}_i$ in each limiting region (see \nameref{app:shape_priors}).
 	In this figure, the gradient speed corresponding to $\vec{v}_1$ is written in the lower
 	  box, $\Omega_{wm}$ being the inner limiting region and $\Omega_{gm}$ the outer.
-	Finally, every $\bar{s}_i$ is projected into the grid of control points $\vec{u}_k$ that
-	support the deformation field through the corresponding weights $\psi_k(\vec{v}_i)$ obtained
-	as in \eqref{eq:basis_derivative}.
+	Finally, every $\bar{s}_i$ is associated with the control points $\vec{u}_k$ through
+	  the corresponding weights $\psi_k(\vec{v}_i)$ \eqref{eq:basis_derivative}.
 	}\label{fig:method}
@@
 
-Therefore, we can apply a discretized interpretation of \eqref{eq:shape_gradients}
-  to compute the data term in \eqref{eq:gradient_descent} as follows:
+We can apply a discretized interpretation of \eqref{eq:shape_gradients} to compute
+  the data term in \eqref{eq:gradient_descent} using shape-gradients
+  \citep{herbulot_segmentation_2006}, and ultimately avoid level sets.
+We refer the reader to the appendix \nameref{app:shape_priors} in order to fill in the gap
+  between \eqref{eq:gradient_descent} and the following expression:
 
@@
     being any pair of neighboring regions, and $\Gamma_{l,m}$ the contour separating
-  them such that $\vec{x}' = \vec{v}' \in\Gamma_{l,m} \iff \vec{x}\in \partial\Omega_i \cap \partial\Omega_j$
-  and $\vec{n_i}'$ is the unit inward normal to the contour at $c_i'$.
+    them such that
+    $\vec{x}' = \vec{v}' \in\Gamma_{l,m} \iff \vec{x}\in \partial\Omega_i \cap \partial\Omega_j$
+    and $\bar{s}_i'$ is the speed vector projected on to the unit inward normal to the contour
+    at $c_i'$ (described in \autoref{fig:method}).
 
@@
   target image smoothing, etc.) and implementation details (for instance the sparse matrix approach
-  to fast interpolation) are dicussed in the \suppl{}, {\color{red} section S1}.
-The actual choice of parameter settings are publicly distributed with the source code of the experiments.
+  to fast interpolation) are dicussed in the \suppl{}, section S1.
+The actual choices of parameter settings are publicly distributed with the source code of the experiments.
 Additionally, we release our software along with a tool for generating convergence reports to
-  demonstrate the behavior of the tool and help scientists configure their experiments.
-A sample of report is found in the \suppl{} {\color{red}, section S1.3}.
+  demonstrate the behavior of \emph{regseg} and help scientists configure their own experiments.
+One sample report is found in the \suppl{}, section S1.3.
 
@@
 %
-In order to demonstrate the applicability of the registration method, we first
-  subjected it to a battery of random tests over synthetic data and then applied
-  it on real data.
+In order to demonstrate the performance of \emph{regseg}, we first conducted a battery of
+  accuarcy tests over synthetic data, and then applied it in the susceptibility distortion
+  correction of real data.
 \autoref{fig:evworkflows} presents the workflow implementing the evaluation instruments.
@@
   two metrics.
-In the case of the phantoms, as we warped data along the three available dimensions,
-  we computed the Haussdorf distance, integrating the ``point-to-cell'' method
-	implemented by \cite{commandeur_vtk_2011}.
-Conversely, the susceptibility-derived distortions only happen along the phase
-  encoding axis of the image.
+In the case of the phantoms, since we had produced distortions along the three
+  available dimensions, we computed the Haussdorf distance, by reusing the
+  ``point-to-cell'' method of \cite{commandeur_vtk_2011}.
+Conversely, the susceptibility-derived distortions only happen along the \gls*{pe}
+  axis of the image.
 Therefore, a \gls*{swindex} can be computed as the one-to-one distance between corresponding
@@
   \vec{v}_i - \hat{\vec{v}}_i \|
+  \label{eq:swindex}
   \end{equation}
@@
 \includegraphics[width=\linewidth]{figures/figure02}
-\caption{Experimental workflow applied on real data from the \gls*{hcp}.
+\caption{Experimental workflow applied on real data from the \acrfull*{hcp}.
   1) The prior surfaces are extracted from the anatomical reference (\gls*{t1} image).
-	2) To operate as ground truth, we generate plausible-synthetic distortion $U^{-1}_{true}$
+	2) To operate as ground truth, we generate a plausible-synthetic distortion $U_{true}$
 	  from the fieldmap using \eqref{eq:fieldmap}.
@@
 		stacked to feed the multivariate input required by our algorithm.
-	4) Registration is run, obtaining a $U_{test} = \hat{U}_{true}$, an estimation of
-	  the ground-truth deformation $U_{true}$.
-	  A cross-comparison methodology is also applied, to obtain a competing $U_{cc}$.
+	4) \emph{Regseg} is run, obtaining a $U_{test} = \hat{U}_{true}$, the estimation of
+	  the ground-truth deformation.
+	  A cross-comparison methodology is also applied, to obtain a competing $\hat{U}_{cc}$.
 	5) Results are visually and quantitatively evaluated.}\label{fig:evworkflows}
@@
   \gls*{t1} (TE/TR=10/1500ms) and \gls*{t2} images (TE/TR=90/5000ms)
-  corresponding to each appearance shape, with two resolutions each
+  corresponding to each phantom type, with two resolutions each
   ($1.0mm$ and $2.0mm$ isotropic).
@@
   parameters, and \citep{glasser_minimal_2013} for the preprocessing issues.
-The datasets comprehend a large of images, containing \gls*{t1}, \gls*{t2} and
+The datasets comprehend a large set of images, containing \gls*{t1}, \gls*{t2} and
   multi-shell \gls*{dmri} images.
@@
   the ``minimally preprocessed'' are corrected for artifacts, brain-extracted
-  and spatially normalized, along with some results of the standard processing with
-  \emph{FreeSurfer}.
+  and spatially normalized, along with some results of the standard processing
+  pipeline of \emph{FreeSurfer}.
 
@@
   six homogeneous regions $\Omega_l$:
-  1) the thalamus;
-  2) \gls*{csf} of the ventricular system and deep \gls*{gm} structures;
-  3) cerebral \gls*{wm};
-  4) brain stem and cerebellar \gls*{wm};
-	5) cerebellar \gls*{gm}; and
-	6) cortical \gls*{gm} surface.
+  1) the thalamus, $\Gamma_{Tha}$;
+  2) \gls*{csf} of the ventricular system and deep \gls*{gm} structures, $\Gamma_{VdGM}$;
+  3) cerebral \gls*{wm}, $\Gamma_{WM}$;
+  4) brain stem and cerebellar \gls*{wm}, $\Gamma_{Bst}$;
+	5) cerebellar \gls*{gm}, $\Gamma_{cbGM}$; and
+	6) cortical \gls*{gm} surface, $\Gamma_{pial}$.
 The choice of this particular model is further addressed in \nameref{sec:res_model_and_metric}.
 
-In this case, we derived the deformation $U_{true}^{-1}$ from the field maps released with
+In this case, we derived the deformation $U_{true}$ from the field maps released with
   the corresponding packages of each dataset from the \gls*{hcp} applying \eqref{eq:fieldmap},

### 2015-02-27 Publications/2015-NeuroImage/05-results.tex
paper version candidate

@@
 To warp the simulated datasets, we randomly generated 150 different realizations of
-  the distortion $U^{-1}_{true}$.
+  the distortion $U_{true}$.
 Smoothness is introduced using two levels of B-Spline functions, with control points evenly
@@
   as previously reported by \cite{irfanoglu_susceptibility_2011} with the finer grid.
-Invertibility of $U^{-1}_{true}$ is ensured by controlling the maximum displacement at each
+Invertibility of $U_{true}$ is ensured by controlling the maximum displacement at each
   level ($20.2mm$ and $10.1mm$ respectively) as studied in \citep{rueckert_diffeomorphic_2006}.
@@
   for all the tests (4 phantom types $\times$ 2 resolutions).
+Since statistical tests may not be conclusive enough, we also computed the confidence intervals
+  (\autoref{tab:ci_phantom}).
+
+\begin{table}
+		\centering
+		\footnotesize
+    \begin{tabular}{llcc}
+    Type      & Low-res.  & Hi-res. \\
+    \hline
+    ``gyrus'' & 0.58-0.60 & 0.18-0.36 \\
+    ``L''     & 0.67-0.73 & 0.35-0.39 \\
+    ``ball''  & 0.66-0.73 & 0.32-0.41 \\
+    ``box''   & 0.68-0.71 & 0.35-0.39 \\
+    \hline
+    \end{tabular}
+    \caption{All the 95\% confidence intervals (CI) were below the corresponding
+      half-voxel resolution in all the phantom types.
+    CIs reported in this table were computed using bootstrapping of $10^4$ samples,
+      for each phantom type and resolution.
+    Taking into account the non-normal distribution of errors, we used the median as location
+  		statistic.}\label{tab:ci_phantom}
+\end{table}
 
@@
   a lineralized space of distortions
-  $\hat{U}\{\vec{r}\} = \epsilon \cdot U_{true} = \vec{r} + \epsilon \cdot u(\vec{r})$,
+  $\hat{U} = \epsilon \cdot U_{true} = \vec{r} + \epsilon \cdot u(\vec{r})$,
   with $\epsilon \in [-1.1, 1.1]$.
-The result of this experiment is shown in \autoref{fig:energymap}.
+The results of this experiment are shown in \autoref{fig:energymap}.
 The metric consistently displayed its minimum at $\epsilon=0.0$ (ground-truth location).
@@
 Using the ``preprocessed'' instance of the \gls*{t2} image of the corresponding subject as
-  reference image, and the \emph{b0} image as moving image, we performed the same correction
-  experiment.
+  reference image, and the \emph{b0} image as moving image, we registered both and
+  projected the correct (undistorted) contours to the warped space using the resulting
+  transform.
+For both \emph{regseg} and the competing method, we computed the \gls*{swindex} \eqref{eq:swindex}
+  of every surface.
+Finally, to statistically compare the results from both methods we performed a one-way ANOVA test
+  on the warping indices for three surfaces.
+Following the nomenclature introduced in section \nameref{sec:human_connectome}, we selected
+  $\Gamma_{VdGM}$, $\Gamma_{WM}$, and $\Gamma_{pial}$ as surfaces of interest.
+These interfaces are critical in connectivity analyses, and small errors in their location may
+  lead to discarding fundamental paths between regions.
+We also reported the 95\% CIs of the warping indices for those surfaces.
+The quantitative results of the ANOVA tests and the CIs are presented in \autoref{tab:results_real}.
+
+\begin{table}
+		\centering
+		\footnotesize
+		\tabcolsep=0.05cm
+    \begin{tabular}{L{1.8cm}ccc}
+                            & Surf. 1  & Surf. 2 & Surf. 3 \\
+    \hline
+    CI \emph{regseg}        & 0.39 - 0.64 & 0.64 - 0.83 & 0.76 - 0.93 \\
+    CI competing            & 2.36 - 3.37 & 2.65 - 3.19 & 2.82 - 3.40 \\
+    \hline
+    ANOVA (f-stat/p-value)  & (51.81/1.5e-06) & (148.25/8.1e-10) & (143.01/1.1e-09) \\
+    \hline
+    \end{tabular}
+    \caption{All the 95\% CIs of the \gls*{swindex} distributions computed for the
+      surfaces of interest were lower for \emph{regseg} with respect to the competing
+      method.
+    CIs reported in this table were computed using bootstrapping with mean as location
+      statistic and $10^4$ samples.
+    One-way-ANOVA tests indicated that there is a significant difference between our method and
+      the competing method of choice.
+    }\label{tab:results_real}
+\end{table}
 

### 2015-02-27 Publications/2015-NeuroImage/06-discussion.tex
paper version candidate

@@
   ii) the low resolution of the target volume produces strong partial volume effects
-  	that exclude the use of voxel-wise segmentation algorithms and hinder volume-based
+  	that preclude voxel-wise segmentation algorithms and hinder volume-based
   	registration; and,
@@
 
-The algorithm is designed for the proposed application on correcting distortions of
-  \gls*{dmri} data, a required preliminary step to extract the structural connectivity
-  networks.
+We designed \emph{regseg} as a method to correct \gls*{dmri} data for susceptibility
+  distortions.
 Connectome extraction protocols generally include the acquisition of \gls*{t1} images
   to provide with prior information about anatomy with high accuracy.
-Our method solves in a single step a joint problem usually addressed in two steps.
-The first solves the linear registration of the \gls*{t1} and \gls*{dmri} data
-  \citep{greve_accurate_2009}, whereas the second corrects for the nonlinear distortions
-  derived from the inhomogenety of magnetic susceptibility across tissues
-  \citep{jezzard_correction_1995}.
+Our method solves in one single step a joint problem usually addressed in two steps:
+  linear registration of \gls*{t1} and \gls*{dmri} to fuse the anatomical information on
+  the diffusion data, and distortions correction using extra acquisitions.
+Moreover, our method does not require any additional acquistition over the minimal protocol
+  that only includes \gls*{t1} and \gls*{dmri} images, a very common scenario for
+  historical datasets.
+
+\paragraph*{Accuracy tests}
+We proved that active contours without edges can be used to successfully drive a
+  deformation supported by B-Spline functions in digital phantoms.
+We randomly deformed 4 different phantom models mimicking three homogeneous regions,
+  showing contrasts generated by a publicly available \gls*{mri} simulator.
+After registration with \emph{regseg} we measured the Hausdorff distance between the
+  projected contours using the ground-truth warping and our estimation.
+We concluded that errors are significantly lower than voxel resolution.
+We also assessed the 95\% intervals, yielding an aggregated interval of distances of
+  $[0.58 - 0.73]$ in the low resolution phantoms ($2.0mm$ isotropic voxel) and
+  $[0.18 - 0.41]$ in the high resolution phantoms ($1.0mm$ isotropic).
+These aggregated intervals by resolution take the lowest of the lower boundary
+  and the highest of the upper boundary of all CIs regardless the phantom type.
+Therefore, we also conclude that the error is bounded and below the half of the
+  voxel spacing to illustrate the accuracy of \emph{regseg}.
+
+\paragraph*{Application on real data}
+After a preliminary experiment to demonstrate the aptness of our segmentation model
+  in section \nameref{sec:res_model_and_metric}, we evaluated \emph{regseg} in
+  a real environment, using the experimental framework presented in \autoref{fig:evworkflows}.
+We processed 16 subjects downloaded from the \gls*{hcp} database with both \emph{regseg}
+  and an in-house replication of a competing method.
+This alternative approach solves distortion using \emph{b0}-to-\gls*{t2} registration.
+Our method presented significantly different results (ANOVA tests are reported in
+  \autoref{tab:results_real}), and the 95\% CIs revealed the high accuracy of
+  \emph{regseg}.
+The aggregated CI for \emph{regseg} was $[0.39 - 0.93] mm$, whereas the competing method
+  yielded an aggregated CI of $[2.36 - 3.40] mm$.
+Even though we carefully configured the competing method by using the same algorithm and
+  extending the same settings from a widely used software, cross-comparison
+  experiments are prone to the so-called instrumentation bias \citep{tustison_instrumentation_2013}.
+Therefore, these results do not reflect that the \emph{regseg} software \emph{is} better than the
+  competing method, but they are communicated to proof that \emph{regseg} is a solid option to
+  correct \gls*{dmri} data for distortions, and any other potential application.
 
@@
 The most extended solution to this is \emph{prelude} \citep{jenkinson_fast_2003}, and some
-  alternatives have been proposed \citep{daga_susceptibility_2014} to circumvent the
-  inherent unreliability of the method.
+  alternatives have been proposed \citep{zhou_reliable_2009,daga_susceptibility_2014} to
+  circumvent the inherent unreliability of the method.
 The \emph{regseg} algorithm would naturally correct missregistered areas due to errors in
-  phase unwrapping, and errors derived to the important smoothing processes that are
-  typically to clean up field maps.
+  phase unwrapping, and subestimations introduced by necessary smoothing processes generally
+  applied to clean up field maps.
 This integration has been recently applied in the \emph{b0}-to-\gls*{t2} approach

### 2015-02-27 Publications/2015-NeuroImage/supplemental.tex
paper version candidate

@@
 % @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-02-23 13:43:56
+% @Last Modified time: 2015-02-27 15:21:03
 
@@
 \usepackage{relsize}
+%\usepackage{inconsolata}
+\usepackage[scaled]{beramono}
 \usepackage[procnames]{listings}
@@
 \definecolor{keywords}{RGB}{0,150,0}
-\def\listingsfont{\ttfamily}
-\def\listingsfontinline{\ttfamily}
+\definecolor{listingbg}{gray}{0.95}
+
+\lstset{basicstyle=\ttfamily\smaller\relax}
+
+\newcommand*{\codeinline}[1]{\colorbox{listingbg}{\lstinline!#1!}}
+%\newcommand\codeinline{\lstinline}
 
 \lstnewenvironment{bashcode}[1][]
-  {\lstset{language=bash}\lstset{%
+  {\lstset{language={}}\lstset{%
   showstringspaces=false,
@@
   %identifierstyle=\color{green},
-  procnamekeys={def,class},
-  morekeywords={models, lambda, forms, as, from},
+  morekeywords={regseg},
   numbers=none,
@@
   frameround=tttt,
-  fillcolor=\color{gray!10},
-  backgroundcolor=\color{gray!10}
+  fillcolor=\color{listingbg},
+  backgroundcolor=\color{listingbg}
 }}
@@
   showstringspaces=false,
-  %identifierstyle=\color{green},
+  identifierstyle=\color{black!70},
   procnamekeys={def,class},
-  morekeywords={models, lambda, forms, as, from},
+  morekeywords={models, lambda, forms, as, from, import},
   numbers=left,
@@
 It is possible to get the description of available options running
-  \lstinline!regseg -h!:
+  \codeinline{regseg -h}:
 
@@
 Our registration algorithm is released with a \emph{nipype Interface} packaged in
-  \lstinline!pyacwereg.interfaces.acwereg!.
+  \codeinline{pyacwereg.interfaces.acwereg}.
 This interface has been comprehensively used in the evaluation workflows.
@@
 Using the JSON log-file as input for the \emph{nipype Interface}
-  \lstinline!pyacwereg.interfaces.ACWEReport!, it is straightforward to obtain
+  \codeinline{ACWEReport}, it is straightforward to obtain
   a visual assessment document presenting the convergence.
@@
 
+This work is supported by two \emph{nipype Workflows} in order to ensure the reproducibility
+  of the results.
+All the intermediate results and figures in this paper have been encapsulated into
+  the workflows and are available in \url{http://dx.doi.org/zenodolink}.
+
+An overview of the general workflow is presented in the Figure 2 of the paper.
+However, this figure has been extremely simplified for the best of visualization.
+In this section, we review the main elements of the evaluation pipelines.
+
+\paragraph{Preprocessing}
+
+\paragraph{Distortion}
+
 \section{Model considerations}
@@
 \end{document}
+
+%{\color{red} {The main diverging points with respect to
+%\citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
+%$\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
+%gradients \citep{besson_dream2s_2003,herbulot_segmentation_2006};
+%2) regularization is also based on linear diffusion smoothing \citep{thirion_image_1998},
+%but we replace the Gaussian filtering by other constraints
+%studied in \citep{nagel_investigation_1986} to better the problem;
+%3) optimization is applied in the spectral
+%domain, observing anisotropic and inhomogeneous mappings along each direction.
+%With respect \citep{guyader_combined_2011}, the main differences are
+%the distance function, and the spectral solution to the optimization updates,
+%as we shall cover in \autoref{sec:numerical_implementation}.}}
+
+%In this paper we formulate the joint registration-segmentation
+%problem as follows.
+%such that the known contours in anatomical space $T$ optimally segment
+%the diffusion space $D$.
+%
+%
+%Whereas related \glspl*{adf} introduced in \autoref{sec:methods_background}
+%make use of explicit level-set formulations to solve \eqref{eq:gradient_descent},
+%we alternatively use \emph{shape-gradients}
+%\cite{besson_dream2s_2003,herbulot_segmentation_2006}.
+
+%
+%Let us denote by $\vec{x}$ the voxel and $F(\vec{x}) = [ f_1, f_2, \ldots, f_N]^T$
+%  its associated feature vector in the following.
+%
+%In our application, these surfaces are precise tissue interfaces of interest extracted
+%  from a high-resolution, anatomically correct reference volume using
+%  the well-established
+%
+%In this paper we propose a novel registration framework to simultaneously
+%solving the segmentation, distortion and cortical parcellation challenges,
+%by exploiting as strong shape-prior the detailed morphology extracted
+%from high-resolution and anatomically correct \gls{mri}.
+%Indeed, hereafter
+%we assume this segmentation problem in anatomical images is reliably and
+%accurately solved with readily available tools (e.g.
+%\citep{fischl_freesurfer_2012}).
+%After global alignment with \gls{t1} using existing approaches, the remaining
+%spatial mismatch between anatomical and diffusion space is due to susceptibility
+%distortions.
+%Finally, we need to establish precise spatial correspondence between the
+%surfaces in both spaces, including the tangential direction for parcellation.
+%Therefore, we can reduce the problem to finding the differences of spatial
+%distortion in between anatomical and \gls{dwi} space.
+%We thus reformulate the segmentation problem as an inverse problem, where we
+%seek for an underlying deformation field (the distortion) mapping
+%from the structural space into the diffusion space, such that the structural
+%contours segment optimally the \gls{dwi} data.
+%In the process, the one-to-one
+%correspondence between the contours in both spaces is guaranteed, and projection
+%of parcellisation to \gls{dwi} space is implicit and consistent.
+%
+%We test our proposed joint segmentation-registration model on two different
+%synthetic examples.
+%The first example is a scalar sulcus model, where the
+%\gls{csf}-\gls{gm} boundary particularly suffers from \gls{pve} and can only be
+%segmented correctly thanks to the shape prior and its coupling with the inner,
+%\gls{gm}-\gls{wm} boundary through the imposed deformation field regularity.
+%The second case deals with more realistic \gls{dwi} data stemming from
+%phantom simulations of a simplistic brain data.
+%Again, we show that the
+%proposed model successfully segments the \gls{dwi} data based on two derived
+%scalar features, namely \gls{fa} and \gls{md}, while establishing an estimate
+%of the dense distortion field.
+%
+%The rest of this paper is organized as follows.
+%First, in \autoref{sec:methods}
+%we introduce our proposed model for joint multivariate segmentation-registration.
+%Then we provide a more detailed description of the data and experimental setup in
+%\autoref{sec:experiments}.
+%We present results in \autoref{sec:results} and conclude
+%in \autoref{sec:conclusion}.
+%
+%
+%The properties of the reconstructed tensors and derived scalar maps have
+%been studied by \cite{ennis_orthogonal_2006}. Based on their
+%findings, \gls{fa}~\eqref{eq:fa} and \gls{md}~\eqref{eq:md} are
+%considered complementary features, and therefore we selected them for the
+%energy model \eqref{eq:complete_energy} in driving the
+%registration-segmentation process.
+%Whereas \gls{fa} informs mainly about the \emph{shape} of diffusion,
+%the \gls{md} is more related to the \emph{magnitude} of the process:
+%
+%\begin{align}
+%\mathrm{FA} &= \sqrt{ \frac{1}{2}}\,\frac{\sqrt{ (\lambda_1 - \lambda_2)^2 + (\lambda_2 - \lambda_3)^2 + (\lambda_3 - \lambda_1)^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}} \label{eq:fa} \\
+%\mathrm{MD} &= ( \lambda_1 + \lambda_2 + \lambda_3 ) / 3 \label{eq:md}
+%\end{align}
+%where $\lambda_i$ are the eigenvalues of the diffusion tensor
+%associated with the diffusion signal $S(\vec{q})$. There exist
+%two main reasons to justify their choice.
+%First, they are well-understood and standardized in clinical routine.
+%Second, together they contain most of the information that is
+%usually extracted from the \gls{dwi}-derived scalar maps
+%\cite{ennis_orthogonal_2006}.
+%

### 2015-02-27 Publications/2015-NeuroImage/02-abstract.tex
paper version candidate. with abstract

% -*- root: 00-main.tex -*-
Integrating information from different images is an ubiquitous necessity
  in current neuroimage analyses.
However, oftentimes one of the images is not available or the objective
  function is not adequate or numerically tractable.
We propose a surfaces-to-volume registration method called \emph{regseg}
  that exploits the detailed anatomy extracted from structural images and
  simultaneously segments the target image with subvoxel precision.
We use active contours without edges to search for a deformation field
  of B-Spline functions that optimally maps the surfaces
  onto the data in the target space.
We test \emph{regseg} on four digital phantoms warped with random deformations,
  and conclude that the precision of the method is below the half-voxel spacing.
Finally, we apply \emph{regseg} in the correction for susceptibility-derived
  distortions of 16 real diffusion MRI datasets from the \acrlong*{hcp}
  database.
The results suggest that \emph{regseg} overperformed an extended competing method.
\emph{Regseg} showed a consistent error within the 95\% confidence interval of
  $0.39-0.93mm$ in images of $1.25mm$ isotropic resolution.

### 2015-02-27 Publications/2015-NeuroImage/03-introduction.tex
paper version candidate. with abstract

@@
   these sequences minimize the bandwidth of the \gls*{pe} direction.
-This low bandwith originates the artifactual displacement of locations presenting small deviations
+This low bandwidth originates the artifactual displacement of locations presenting small deviations
   from the main magnetic field $B_0$.
@@
   techniques \citep{jezzard_correction_1995,andersson_modeling_2001}.
-A second set of methods aquired extra images reproducing the same \gls*{epi} scheme but switching
+A second set of methods acquired extra images reproducing the same \gls*{epi} scheme but switching
   the \gls*{pe} axis \citep{chiou_simple_2000} or reversing the direction of gradient
@@
   three-dimensional problem.
-All methods require extra \gls*{mr} aquisitions. Thus, a tipical protocol for the extraction
+All methods require extra \gls*{mr} acquisitions. Thus, a typical protocol for the extraction
   of structural connectivity comprehends \gls*{dmri}, \gls*{t1}, and the acquisition demanded
@@
   susceptibility-derived distortions \citep{jezzard_correction_1995}.
-\cite{greve_accurate_2009} circumvented the nonlinearity by identifying tipically
+\cite{greve_accurate_2009} circumvented the nonlinearity by identifying typically
   distorted regions and removing them from the metric computation.

### 2015-02-27 Publications/2015-NeuroImage/04-methods.tex
paper version candidate. with abstract

@@
   framework and only one \gls*{pde} for first time.
-The principal application of these techniques was to perform registration between timesteps in
-  time-series or between different images \citep{paragios_level_2003}.
+The principal application of these techniques was to perform registration between time steps in
+  time series or between different images \citep{paragios_level_2003}.
 \cite{unal_coupled_2005} and later \cite{wang_joint_2006},
@@
 
+\todo[inline]{Include here the final update equation using the FT, introducing the
+step size $\tau$, and refer to the supplementals for intermediate steps}
+
 \paragraph*{Settings, implementation details, and convergence}
@@
   target image smoothing, etc.) and implementation details (for instance the sparse matrix approach
-  to fast interpolation) are dicussed in the \suppl{}, section S1.
+  to fast interpolation) are discussed in the \suppl{}, section S1.
 The actual choices of parameter settings are publicly distributed with the source code of the experiments.

### 2015-02-27 Publications/2015-NeuroImage/05-results.tex
paper version candidate. with abstract

@@
   to the two sets of resolutions of generated phantoms.
-To support that the missregistration errors averaged per experiment was significantly
+To support that the misregistration errors averaged per experiment was significantly
   under the resolution of the target image, we proceeded as follows.
@@
 %
-Preliminarly, we investigated the aptness of our segmentation model to play as cost function
+Preliminarily, we investigated the aptness of our segmentation model to play as cost function
   driving the registration process.
@@
 The energy functional \eqref{eq:energy} was evaluated for the warpings given by
-  a lineralized space of distortions
+  a linearized space of distortions
   $\hat{U} = \epsilon \cdot U_{true} = \vec{r} + \epsilon \cdot u(\vec{r})$,
@@
     \begin{tabular}{L{1.8cm}ccc}
-                            & Surf. 1  & Surf. 2 & Surf. 3 \\
+                            & $\Gamma_{VdGM}$  & $\Gamma_{WM}$ & $\Gamma_{pial}$ \\
     \hline
@@
 	\caption{The evaluation instrument automatically renders the overlay of the
-	  target \gls*{fa} image and the resulting contours after distortion correction (yellow color) and
+	  target \gls*{fa} image and the resulting contours after distortion estimation (yellow color) and
 	  their theoretical location using the ground-truth (green color).
@@
 	The last two rows represent the sagittal view.
-	We intentionally ommited the coronal slicing due to it is the less informative given the properties
+	We intentionally omitted the coronal slicing due to it is the less informative given the properties
 	  of the distortion.
@@
 	  the competing method.
+	\todo[inline]{This figure will also have the box-plots of the errors of surfaces for both regseg
+	and the t2b method}
 	}\label{fig:results_real}

### 2015-03-12 Publications/2015-NeuroImage/04-methods.tex

@@
 	 database of the \gls*{hcp}.
+\todo[inline]{This is partially true: I am currently processing the data and
+only 5 subjects with regseg and 10 with the competing method are processed.
+All the results presented afterwards are based on this processed cases, and will be
+updated for the final submission.}
 We refer the reader to \citep{essen_human_2012} for exact details about acquisition

### 2015-03-12 Publications/2015-NeuroImage/06-discussion.tex

@@
   (see \nameref{app:shape_priors} in Appendix) with a gradient descent strategy.
-The aproach is valid for image processing applications in which a precise segmentation is
+The approach is valid for image processing applications in which a precise segmentation is
   available (from an atlas or within subject anatomical data), and match any or all
@@
   the diffusion data, and distortions correction using extra acquisitions.
-Moreover, our method does not require any additional acquistition over the minimal protocol
+Moreover, our method does not require any additional acquisition over the minimal protocol
   that only includes \gls*{t1} and \gls*{dmri} images, a very common scenario for
@@
 \paragraph*{Prospects}
-Future extensions or improvements of \emph{regseg} similar to the evolution of existing
+Future extensions or improvements of \emph{regseg} are similar to the evolution of existing
   alternative methodologies.
-Incorporating knowledge about the distortion initializing our method with the theory-based
+Incorporating knowledge about the distortion by initializing our method with the theory-based
   correction using fieldmaps would be the most profitable.
@@
   circumvent the inherent unreliability of the method.
-The \emph{regseg} algorithm would naturally correct missregistered areas due to errors in
-  phase unwrapping, and subestimations introduced by necessary smoothing processes generally
+The \emph{regseg} algorithm would naturally correct misregistered areas due to errors in
+  phase unwrapping, and underestimations introduced by necessary smoothing
   applied to clean up field maps.

### 2015-03-12 Publications/2015-NeuroImage/07-conclusion.tex

@@
 \emph{Regseg} is a variational framework to simultaneously segment and
-  register 3D \gls*{dmri} data of the human brain using within subject
+  register 3D \gls*{dmri} data of the human brain, using within-subject
   anatomical information as reference.
@@
 We propose \emph{regseg} to correct \gls*{dmri} datasets for the typically present distortions
-  caused by inhomogeneitis of the main magnetic field.
+  caused by inhomogeneities of the main magnetic field.
 Visual assessment of \emph{regseg} and cross-comparison with a widely-used competing
-  technique demonstrated its accuracy, and overperformed the competing method.
+  technique demonstrated its accuracy.
 Moreover, \emph{regseg} does not require additional images to the minimal acquisition protocol
-  that generally includes only \gls*{t1} and \gls*{dmri}.
+  that includes only \gls*{t1} and \gls*{dmri}.
 

### 2015-03-12 Publications/2015-NeuroImage/03-introduction.tex
updating paper for second round of reviews

--- 2015-02-19 Publications/2015-NeuroImage/03-introduction.tex
@@
 % -*- root: 00-main.tex -*-
-\section*{Introduction}\label{sec:introduction}
-Nonlinear registration of images is a challenging task, ubiquitous
-  in an endless number of image analysis applications.
-The process aims to find a spatial mapping $U$ that aligns the information available
-  in two different coordinate systems (generally called
-  \emph{reference}, $R$, and \emph{moving}, $M$):%
+\section{Introduction}\label{sec:introduction}
+Image registration is the process to find a spatial mapping $U$ that aligns the information available
+  in two different coordinate systems (generally called \emph{reference},
+  $R$, and \emph{moving}, $M$):%
 
@@
   \end{align}
-
+%
   where $\vec{r}$ denotes a position in the reference domain $R$, $\vec{r}'$ is
@@
   to the reference domain.
-A comprehensive survey by \cite{sotiras_deformable_2013}
-  illustrates the wide variety of existing registration methodologies.
-They classify the algorithms by the three principal building blocks that can be identified
-  among methods: the matching criteria or similarity function to evaluate the proximity of
-  the solution, the deformation model that may implement theoretical properties and constraints
-  of the mapping in the selected application, and the optimization method or search strategy.
-The possibly most widespread matching criteria are those classified as \emph{iconic
-  methods} in \citep{sotiras_deformable_2013}.
-These algorithms require the definition of an appropriate function on the voxel-wise information
-  content of both images.
-A second group, called \emph{geometric methods}, align geometrical features derived from
-  corresponding objects defined in both spaces, such as landmarks, shapes, surfaces, etc.
-Finally, \emph{hybrid methods} map geometrical features and voxel-wise information.
+A comprehensive survey by \cite{sotiras_deformable_2013} illustrates the wide variety
+  of existing methodologies, and classifies them by their three principal building blocks:
+  the matching criteria to evaluate the proximity of the solution,
+  the deformation model that may implement theoretical properties of the mapping
+  (i.e. linearity, elasticity, viscosity, etc.), and the optimization method.
+The matching criteria can be defined as a function on the intensities of voxels of the
+  two images, the alignment of spatial features derived from the corresponding objects in
+  $R$ and $M$, or use both intensity and spatial informations.
+In \citep{sotiras_deformable_2013} these methods are referred to as \emph{iconic}, \emph{geometric},
+  and \emph{hybrid}, respectively.
+When segmentation is introduced to define the matching criteria, we observe a \emph{hybrid}
+  registration method that simultaneously defines a partition of the reference space to identify
+  relevant objects in it.
+An early integration of segmentation and registration by \cite{bertalmio_morphing_2000} proposed
+  a sequential deformation of active contours for object tracking in sequential 2D images.
+Shortly, \cite{yezzi_variational_2001} presented the first method including a full solution to
+  the registration problem with an affine transformation supporting the coordinates mapping.
+\cite{vemuri_joint_2003} proposed an atlas-based registration framework using level sets and only
+  one \gls*{pde} for first time.
+\cite{unal_coupled_2005} and later \cite{wang_joint_2006},
+  extended the ``two \glspl*{pde}'' method of \cite{yezzi_variational_2001}
+  to nonlinear registration implementing a free deformation field.
+\cite{droske_mumfordshah_2009} reviewed the latter set of techniques, and proposed two different
+  approaches to apply the Mumford-Shah functional \citep{mumford_optimal_1989} in simultaneous
+  registration and segmentation, through the propagation of the deformation field from
+  the contours to the whole image definition.
+\cite{greve_accurate_2009} presented a widely used registration method called \emph{bbregister},
+  included in \emph{FreeSurfer} \citep{fischl_freesurfer_2012}.
+Their framework performs robust registration of brain surfaces into the intensity information
+  of a target \gls*{mri}, using an affine transformation and active contours with edges.
+Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
+  registration method using level sets and a nonlinear elasticity smoother on the
+  displacement vector field, which preserves topology even with very large deformations.
+Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
+  level-set function for the registration of several active contours, in the application
+  of atlas-based segmentation.
+Alternatively to the historical use of active contours based methods, some Bayesian approaches
+  have been proposed as well \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014}.
 
-In this work we propose a hybrid approach to nonlinear registration of anatomically correct
-  and precise surfaces to warped and low-resolution \gls*{dmri} of the brain.
-Currently, we are witnessing an uprising demand for such algorithms in the quest
-  of extracting and characterizing the networks of structural connectivity in the
-  human brain \citep{craddock_imaging_2013}.
-Our registration method is demonstrated here as a solution to correct for the susceptibility
-  distortions that \gls*{dmri} of the head typically present \citep{jezzard_correction_1995}.
-The precise segmentation of \gls*{dmri} data is also crucial in such applications, since the
-  definition of the nodes in the network is generally derived from a parcellation performed in
-  an anatomical \gls*{mri} of the same subject \citep{daducci_connectome_2012}.
-
-\Gls*{dmri} data are usually acquired with \gls*{epi} schemes.
-In order to quickly probe  in many different orientations the diffusion process within the living
-  brain, these sequences minimize the bandwidth of the \gls*{pe} direction.
-This low bandwith originates the artifactual displacement of locations presenting small deviations
-  from the main magnetic field $B_0$.
-The principal source of these deviations, $\Delta B_0$, is the discontinuity of magnetic
-  susceptibility at tissue interfaces.
-Tissue/air boundaries show large steps of susceptibility and this artifact is
-  particularly evident in the surroundings of the orbito-frontal lobe or the temporal
-  bones in both hemispheres.
-The displacements field $U_s$ along the \gls*{pe} axis can be estimated theoretically
-  \citep{jezzard_correction_1995}:
-
-  \begin{equation}
-  \vec{r}' = \vec{r} + \frac{\gamma \, T_{acq}\, s_{PE}}{2\pi}\Delta B_0(\vec{r}) \cdot \hat{\vec{e}}_{PE}.
-  \label{eq:fieldmap}
-  \end{equation}
-
-where $\vec{r}$ is the location in the reference domain as in \eqref{eq:transform},
-  $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time,
-  $s_{PE}$ is the pixel spacing along \gls*{pe}, and $\hat{\vec{e}}_{PE}$ the unitary
-  vector along \gls*{pe}.
-Several methodologies have been proposed to estimate $U_s$ and fix the artifact.
-The first solution was acquiring extra images to estimate $\Delta B_0$ using field mapping
-  techniques \citep{jezzard_correction_1995,andersson_modeling_2001}.
-A second set of methods aquired extra images reproducing the same \gls*{epi} scheme but switching
-  the \gls*{pe} axis \citep{chiou_simple_2000} or reversing the direction of gradient
-  increments in the same axis \citep{cordes_geometric_2000,holland_efficient_2010}.
-\cite{kybic_unwarping_2000} proposed a correction using nonlinear registration of the \gls*{dmri}
-  data versus a \gls*{t2} image.
-Further developments \citep{irfanoglu_susceptibility_2011} extended this method to the
-  three-dimensional problem.
-All methods require extra \gls*{mr} aquisitions. Thus, a tipical protocol for the extraction
-  of structural connectivity comprehends \gls*{dmri}, \gls*{t1}, and the acquisition demanded
-  by the correction method of choice.
-
-Our method is related to the \gls*{t2b} methods as we also render the correction as a
-  registration problem.
-Conversely, the method does not require any of the previously mentioned extra acquisitions.
-A fundamental antecedent to the present approach is found in \citep{greve_accurate_2009}.
-They proposed a registration method to align \gls*{t1} and \gls*{dmri} images
-  of the brain using a linear transformation model.
-Similarly, we exploit the precise structures that can be extracted with
-  widely-used software (i.e. \emph{FreeSurfer}~\citep{fischl_freesurfer_2012}),
-  and then map these surfaces into warped data, in which the scale of the
-  imaged structures is in the range of the image resolution or above.
-Our approach differs from \citep{greve_accurate_2009} in two fundamental choices.
-First, we use a nonlinear deformation model that enables the correction of the
-  susceptibility-derived distortions \citep{jezzard_correction_1995}.
-Second, our matching criteria does not necessarily need for clear intensity
-  steps at the tissue interfaces where surfaces must be fitted, thanks to the use of
-  active contours without edges \citep{chan_active_2001}.
-The hypothesis underpinning our research is that, in those situations,
-  image registration can be reliably performed by searching for homogeneous
-  regions in one image corresponding to a prior knowledge of the shape of objects in
-  the moving coordinate system.
+We propose \emph{regseg}, a registration method based on active contours without edges
+  \citep{chan_active_2001} to drive registration and segment the target image.
+The hypothesis underpinning our research is that image registration can be reliably performed
+  by searching for homogeneous regions in the target image that correspond to precise contours
+  from an atlas, or extracted from other image (i.e. a different time step).
+We demonstrate the approach in correcting \gls*{dmri} for nonlinear distortions, an application
+  that is described in \autoref{sec:distortions}.
+In \autoref{sec:methods_map}, we start from a probabilistic framework to finally show the duality
+  between the methods described previously.
+Following sections introduce
+  the specific features of \emph{regseg} (\autoref{sec:numerical_implementation}),
+  state the evaluation experiments (\autoref{sec:experiments_evaluation}),
+  and describe the synthetic and real data involved in this study (\autoref{sec:datasets}).
+In \autoref{sec:results} we describe our findings to support the usefulness of \emph{regseg}
+  in the described application, compared to a well-established tool.
+The interpretation of the results, and prospects, are addressed in \autoref{sec:discussion}.

### 2015-03-12 Publications/2015-NeuroImage/04-methods.tex
updating paper for second round of reviews

@@
 % -*- root: 00-main.tex -*-
-\section*{Methods}
+\section{Methods}
 \label{sec:methods}
 %
-\subsection*{Related work}
-\label{sec:methods_background}
-Registration-segmentation methods are generally built on a deformation model to
-  support the evolution of a segmentation algorithm that drives the mapping process.
-Thus, solutions usually select first a segmentation strategy to derive the cost function and
-  regularization of the problem.
-Generally, evolving active contours is the most extended segmentation approach to be
-  integrated within the registration process.
-These methods either have an implicit description of the relative boundary locations of the
-  objects to be delineated using level sets \citep{chen_using_2002,bresson_variational_2006,
-  chan_level_2005}, or model the statistical deviations from original shape
-  \citep{cremers_kernel_2006,gastaud_combining_2004}.
-Some Bayesian approaches have been proposed as well
-  \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014}.
-In section \nameref{sec:methods_map}, we start from a probabilistic framework to finally propose
-	a cost function defined on the location of explicit surfaces, showing the duality of the
-	two existing approaches.
 
-An early ``morphing''-segmentation approach was proposed in \citep{bertalmio_morphing_2000}.
-Shortly, \cite{yezzi_variational_2001} presented the first method including a full solution to
-  the registration, where the energy functional is defined simultaneously in the target
-  and reference images with an affine transformation supporting the coordinates mapping.
-\cite{vemuri_joint_2003} proposed an atlas-based registration using a level-sets
-  framework and only one \gls*{pde} for first time.
-The principal application of these techniques was to perform registration between time steps in
-  time series or between different images \citep{paragios_level_2003}.
-\cite{unal_coupled_2005} and later \cite{wang_joint_2006},
-  extended the ``two \glspl*{pde}'' method of \cite{yezzi_variational_2001}
-  to nonlinear registration with a free deformation field.
-\cite{droske_mumfordshah_2009} reviewed the latter set of techniques, and proposed two different
-  approaches to apply the Mumford-Shah functional \citep{mumford_optimal_1989} in simultaneous
-  registration-segmentation, through the propagation of the deformation field from
-  the contours to the whole image definition.
-Finally, \cite{gorthi_active_2011} presented a comprehensive generalization of the
-  existing methodologies based on level sets in the application of atlas-based segmentation.
-Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
-  registration method close to the one proposed here.
-Our method differs on two key features:
-  it allows multiple active contours for segmentation, and the regularization strategy.
-Whereas \cite{guyader_combined_2011} proposed a nonlinear elasticity smoother, we rely on
-  the inherent smoothness of the deformation field, and on the anisotropic regularization
-  proposed by \cite{nagel_investigation_1986}.
+\subsection{Susceptibility-derived distortions}
+\label{sec:distortions}
+\Gls*{dmri} data are usually acquired with \gls*{epi} schemes.
+In order to quickly probe the diffusion process within the brain in many different orientations,
+  these sequences minimize the bandwidth of the \gls*{pe} direction.
+This low bandwidth originates the artifactual displacement of locations presenting small deviations
+  from the main magnetic field $B_0$.
+The principal source of these deviations ($\Delta B_0$) is the discontinuity of magnetic
+  susceptibility at tissue interfaces.
+Tissue/air boundaries show large steps of susceptibility and this artifact is
+  particularly evident in the surroundings of the orbito-frontal lobe or the temporal
+  bone of both hemispheres.
+The displacements field $U_s$ along the \gls*{pe} axis can be estimated theoretically
+  \citep{jezzard_correction_1995}:
 
-\subsection*{Cost-function derivation}\label{sec:methods_map}
+  \begin{equation}
+  \vec{r}' = \vec{r} + \frac{\gamma \, T_{acq}\, s_{PE}}{2\pi}\Delta B_0(\vec{r}) \cdot \hat{\vec{e}}_{PE},
+  \label{eq:fieldmap}
+  \end{equation}
+%
+where $\vec{r}$ is the true location in the undistorted domain as in \eqref{eq:transform},
+  $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time,
+  $s_{PE}$ is the pixel spacing along \gls*{pe}, and $\hat{\vec{e}}_{PE}$ the unitary
+  vector along \gls*{pe}.
+Several methodologies have been proposed to estimate $U_s$ and fix the artifact.
+The first solution consisted on estimating $\Delta B_0$ with field mapping
+  techniques \citep{andersson_modeling_2001}.
+A second family of methods require an extra \gls*{epi} acquisition, but switching
+  the \gls*{pe} axis \citep{chiou_simple_2000} or reversing the direction of gradient \gls*{pe}
+  increments \citep{cordes_geometric_2000,holland_efficient_2010}.
+\cite{kybic_unwarping_2000} proposed a correction using nonlinear registration of the \emph{b0}
+  image\footnote{The term \emph{b0} refers to one or more images within the
+  \gls*{dmri} dataset that are acquired with a null or very low gradient intensity or
+   \emph{b}-value used as signal reference.} of the \gls*{dmri} versus a \gls*{t2} image.
+Further developments \citep{irfanoglu_susceptibility_2011} extended this method to the
+  three-dimensional problem.
+Therefore, a typical protocol for the extraction of structural connectivity
+  comprehends \gls*{dmri}, \gls*{t1}, and the acquisition determined by the correction
+  method of choice \citep{daducci_connectome_2012}.
+
+\subsection{Our method and contributions}
+\label{sec:our_method}
+\emph{Regseg} is related to \emph{bbregister}, as both methods exploit the precise
+  surfaces that can be extracted with \emph{FreeSurfer}~\citep{fischl_freesurfer_2012},
+  to then map this prior information into a target image.
+However, \emph{bbregister} features an affine transform that is unable to compensate the
+  distortions.
+To avoid this problem, the active contours are dismissed in regions typically affected by
+  warpings.
+Additionally, \emph{bbregister} is robust in finding clear tissue boundaries.
+In the case of \gls*{dmri} data, oftentimes the registration process is restricted to finding
+  the pial surface, as it is the only clear boundary.
+
+to the methods based on nonlinear registration of the
+  to the \gls*{t2} image, since we also render the correction as a
+  registration problem.
+Conversely, the method does not require any of the previously mentioned extra acquisitions.
+A fundamental antecedent to \emph{regseg} is found in .
+They proposed a registration method to align surfaces extracted from the \gls*{t1}
+  to \gls*{dmri} images using a linear transformation model using active contours
+  with edges.
+Similarly, we  warped data, in which the scale of the
+  imaged structures is in the range of the image resolution or above.
+Our approach differs from \citep{greve_accurate_2009} in two fundamental choices.
+First, we use a nonlinear deformation model that enables the correction of the
+  susceptibility-derived distortions \citep{jezzard_correction_1995}.
+\cite{greve_accurate_2009} circumvented the nonlinearity by identifying typically
+  distorted regions and removing them from the metric computation.
+Second, our matching criteria does not necessarily need for clear intensity
+  steps at the tissue interfaces where surfaces must be fitted, thanks to the use of
+  active contours without edges \citep{chan_active_2001}.
+
+This choice is similar to \citep{guyader_combined_2011}, but \emph{regseg} differs in two
+  features.
+First, the regularization relies on the inherent smoothness of the deformation field,
+  and on the anisotropic approach proposed by \cite{nagel_investigation_1986}.
+Second, it accepts several contours to be mapped as in \citep{gorthi_active_2011}.
+However, \emph{regseg} makes use of the explicit representation of surfaces, and
+  does not implement multiphase level sets.
+Our registration method is demonstrated here as a solution to correct for the susceptibility
+  distortions that \gls*{dmri} of the head typically present \citep{jezzard_correction_1995}.
+The distortion correction and the precise segmentation of \gls*{dmri} data is crucial in
+	applications such as connectome extraction \citep{craddock_imaging_2013}, since the
+  definition of the nodes in the network is generally derived from a parcellation performed in
+  an anatomical \gls*{mri} of the same subject \citep{daducci_connectome_2012}.
+
+\subsection{Cost-function derivation}\label{sec:methods_map}
 
@@
 
-\subsection*{Numerical Implementation}
+\subsection{Numerical Implementation}
 \label{sec:numerical_implementation}
@@
 	The gradient speeds $\bar{s}_i$ are computed as the disparity between data energies of
-	  the vertex $\vec{v}_i$ in each limiting region (see \nameref{app:shape_priors}).
+	  the vertex $\vec{v}_i$ in each limiting region (see \autoref{app:shape_priors}).
 	In this figure, the gradient speed corresponding to $\vec{v}_1$ is written in the lower
@@
   \citep{herbulot_segmentation_2006}, and ultimately avoid level sets.
-We refer the reader to the appendix \nameref{app:shape_priors} in order to fill in the gap
+We refer the reader to the appendix \autoref{app:shape_priors} in order to fill in the gap
   between \eqref{eq:gradient_descent} and the following expression:
@@
 
-\subsection*{Experiments and evaluation}
+\subsection{Experiments and evaluation}
 \label{sec:experiments_evaluation}
@@
 
-\subsection*{Image Data and preprocessing}
+\subsection{Image Data and preprocessing}
 \label{sec:datasets}
@@
 	6) cortical \gls*{gm} surface, $\Gamma_{pial}$.
-The choice of this particular model is further addressed in \nameref{sec:res_model_and_metric}.
+The choice of this particular model is further addressed in \autoref{sec:res_model_and_metric}.
 

### 2015-03-13 Publications/2015-NeuroImage/03-introduction.tex
updating references

@@
   and \emph{hybrid}, respectively.
-When segmentation is introduced to define the matching criteria, we observe a \emph{hybrid}
+When segmentation is introduced to define the matching criteria, we face a \emph{hybrid}
   registration method that simultaneously defines a partition of the reference space to identify
@@
   a sequential deformation of active contours for object tracking in sequential 2D images.
-Shortly, \cite{yezzi_variational_2001} presented the first method including a full solution to
+Shortly, \cite{yezzi_variational_2003} presented the first method including a full solution to
   the registration problem with an affine transformation supporting the coordinates mapping.

### 2015-03-13 Publications/2015-NeuroImage/04-methods.tex
updating references

@@
 \emph{Regseg} is related to \emph{bbregister}, as both methods exploit the precise
-  surfaces that can be extracted with \emph{FreeSurfer}~\citep{fischl_freesurfer_2012},
+  surfaces that can be extracted with \emph{FreeSurfer} \citep{fischl_freesurfer_2012},
   to then map this prior information into a target image.

### 2015-03-13 Publications/2015-NeuroImage/00-main.tex
modify font, add lineno
finishing changes
discussion refactoring
add supplementals
paper version candidate
updating paper for second round of reviews
one more iteration to second round

@@
-%\documentclass[numbers,sort&compress]{elsarticle}
 \documentclass[5p,authoryear]{elsarticle}
-\journal{NeuroImage}
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
+\journal{ArXiv archiving for pre-submission review}
 
 % *** MISC UTILITY PACKAGES ***
-%
 \usepackage{glossaries}
 \usepackage{booktabs} % to use \toprule
-
+\usepackage{multirow}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 \providecommand{\nvec}[1]{\hat{\mathbf{#1}}}
+\providecommand{\suppl}{\href{http://dx.doi.org/mydoi}{Supplemental Materials}}
 
@@
 \usepackage[colorlinks=true,linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
+\usepackage{doi}
 
 % correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
+\hyphenation{op-tical net-works semi-conduc-tor an-isot-ropy regis-tra-tion iso-tropic Free-Surfer %
+align-ed pre-sent}
 
@@
 
+\usepackage[T1]{fontenc}
+\usepackage{charter}
+%\usepackage[expert]{mathdesign}
+%\usepackage[libertine,cmintegrals,cmbraces,vvarbb]{newtxmath}
+%\renewcommand*{\familydefault}{\sfdefault}
+\usepackage[switch]{lineno}
+\usepackage{array}
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+
 \begin{document}
-
 \begin{frontmatter}
@@
 
-\title{Simultaneous segmentation and registration of diffusion MR images of the brain driven
-by active-contours}
+\title{Active contours driven registration of diffusion MR images of the brain}
 
@@
 
+\linenumbers
 \input{03-introduction}
@@
 
+% ACKs
+\input{08-acknowledgements}
+
+%\bibliographystyle{plainnat}
+\bibliographystyle{mystyle}
+\bibliography{Remote}
+
 % Add appendices
@@
 
-% ACKs
-\input{08-acknowledgements}
-
-\bibliographystyle{myplainnat}
-\bibliography{Remote}
 \end{document}

### 2015-03-13 Publications/2015-NeuroImage/03-introduction.tex
one more iteration to second round

@@
 % -*- root: 00-main.tex -*-
-\section{Introduction}\label{sec:introduction}
-Image registration is the process to find a spatial mapping $U$ that aligns the information available
-  in two different coordinate systems (generally called \emph{reference},
-  $R$, and \emph{moving}, $M$):%
-
-  \begin{align}
-  U\colon R \subset \mathbb{R}^n &\to M \subset \mathbb{R}^n \notag\\
-  \vec{r} &\mapsto \vec{r}' =\vec{r}+u(\vec{r}),
-  \label{eq:transform}
-  \end{align}
-%
-  where $\vec{r}$ denotes a position in the reference domain $R$, $\vec{r}'$ is
-  its corresponding location in $M$, and $n$ the dimensionality of images.
-Finally, $\vec{u} = u(\vec{r})$ is the displacement of every point with respect
-  to the reference domain.
-A comprehensive survey by \cite{sotiras_deformable_2013} illustrates the wide variety
-  of existing methodologies, and classifies them by their three principal building blocks:
+\section{Introduction}%
+\label{sec:introduction}
+Image registration is the process to find a spatial mapping that aligns the information available
+  in two different coordinate systems.
+In their comprehensive survey, \cite{sotiras_deformable_2013} illustrated the wide variety
+  of existing methodologies, and classified them by their three principal building blocks:
   the matching criteria to evaluate the proximity of the solution,
@@
   two images, the alignment of spatial features derived from the corresponding objects in
-  $R$ and $M$, or use both intensity and spatial informations.
+  the spaces to be registered, or use both intensity and spatial informations.
 In \citep{sotiras_deformable_2013} these methods are referred to as \emph{iconic}, \emph{geometric},
@@
 \cite{unal_coupled_2005} and later \cite{wang_joint_2006},
-  extended the ``two \glspl*{pde}'' method of \cite{yezzi_variational_2001}
+  extended the ``two \glspl*{pde}'' method of \cite{yezzi_variational_2003}
   to nonlinear registration implementing a free deformation field.
@@
   of atlas-based segmentation.
-Alternatively to the historical use of active contours based methods, some Bayesian approaches
+Alternatively to the historical use of active contours, some Bayesian approaches
   have been proposed as well \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014}.

### 2015-03-13 Publications/2015-NeuroImage/04-methods.tex
one more iteration to second round

@@
   bone of both hemispheres.
+Distortion can be rendered as a registration problem between the undistorted $R$ (reference), 
+  and distorted $M$ (moving) spaces:
+
+  \begin{align}
+  U_s\colon R \subset \mathbb{R}^n &\to M \subset \mathbb{R}^n \notag\\
+  \vec{r} &\mapsto \vec{r}' =\vec{r}+u(\vec{r}),
+  \label{eq:transform}
+  \end{align}
+%
+  where $\vec{r}$ denotes a position in the reference domain $R$, $\vec{r}'$ is
+  its corresponding location in $M$, and $n$ the dimensionality of images.
+Finally, $\vec{u} = u(\vec{r})$ is the displacement of every point with respect
+  to the reference domain.
+
 The displacements field $U_s$ along the \gls*{pe} axis can be estimated theoretically
@@
 %
-where $\vec{r}$ is the true location in the undistorted domain as in \eqref{eq:transform},
-  $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time,
+where $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time,
   $s_{PE}$ is the pixel spacing along \gls*{pe}, and $\hat{\vec{e}}_{PE}$ the unitary
@@
 \label{sec:our_method}
-\emph{Regseg} is related to \emph{bbregister}, as both methods exploit the precise
-  surfaces that can be extracted with \emph{FreeSurfer} \citep{fischl_freesurfer_2012},
-  to then map this prior information into a target image.
-However, \emph{bbregister} features an affine transform that is unable to compensate the
-  distortions.
-To avoid this problem, the active contours are dismissed in regions typically affected by
-  warpings.
-Additionally, \emph{bbregister} is robust in finding clear tissue boundaries.
-In the case of \gls*{dmri} data, oftentimes the registration process is restricted to finding
-  the pial surface, as it is the only clear boundary.
-
-to the methods based on nonlinear registration of the
-  to the \gls*{t2} image, since we also render the correction as a
-  registration problem.
-Conversely, the method does not require any of the previously mentioned extra acquisitions.
-A fundamental antecedent to \emph{regseg} is found in .
-They proposed a registration method to align surfaces extracted from the \gls*{t1}
-  to \gls*{dmri} images using a linear transformation model using active contours
-  with edges.
-Similarly, we  warped data, in which the scale of the
-  imaged structures is in the range of the image resolution or above.
-Our approach differs from \citep{greve_accurate_2009} in two fundamental choices.
-First, we use a nonlinear deformation model that enables the correction of the
-  susceptibility-derived distortions \citep{jezzard_correction_1995}.
-\cite{greve_accurate_2009} circumvented the nonlinearity by identifying typically
-  distorted regions and removing them from the metric computation.
-Second, our matching criteria does not necessarily need for clear intensity
-  steps at the tissue interfaces where surfaces must be fitted, thanks to the use of
-  active contours without edges \citep{chan_active_2001}.
-
-This choice is similar to \citep{guyader_combined_2011}, but \emph{regseg} differs in two
-  features.
-First, the regularization relies on the inherent smoothness of the deformation field,
-  and on the anisotropic approach proposed by \cite{nagel_investigation_1986}.
-Second, it accepts several contours to be mapped as in \citep{gorthi_active_2011}.
-However, \emph{regseg} makes use of the explicit representation of surfaces, and
-  does not implement multiphase level sets.
-Our registration method is demonstrated here as a solution to correct for the susceptibility
-  distortions that \gls*{dmri} of the head typically present \citep{jezzard_correction_1995}.
-The distortion correction and the precise segmentation of \gls*{dmri} data is crucial in
-	applications such as connectome extraction \citep{craddock_imaging_2013}, since the
-  definition of the nodes in the network is generally derived from a parcellation performed in
-  an anatomical \gls*{mri} of the same subject \citep{daducci_connectome_2012}.
+% \emph{Regseg} is related to \emph{bbregister}, as both methods exploit the precise
+%   surfaces that can be extracted with \emph{FreeSurfer} \citep{fischl_freesurfer_2012},
+%   to then map this prior information into a target image.
+% However, \emph{bbregister} features an affine transform that is unable to compensate the
+%   distortions.
+% To avoid this problem, the active contours are dismissed in regions typically affected by
+%   warpings.
+% Additionally, \emph{bbregister} is robust in finding clear tissue boundaries.
+% In the case of \gls*{dmri} data, oftentimes the registration process is restricted to finding
+%   the pial surface, as it is the only clear boundary.
+% 
+% to the methods based on nonlinear registration of the
+%   to the \gls*{t2} image, since we also render the correction as a
+%   registration problem.
+% Conversely, the method does not require any of the previously mentioned extra acquisitions.
+% A fundamental antecedent to \emph{regseg} is found in .
+% They proposed a registration method to align surfaces extracted from the \gls*{t1}
+%   to \gls*{dmri} images using a linear transformation model using active contours
+%   with edges.
+% Similarly, we  warped data, in which the scale of the
+%   imaged structures is in the range of the image resolution or above.
+% Our approach differs from \citep{greve_accurate_2009} in two fundamental choices.
+% First, we use a nonlinear deformation model that enables the correction of the
+%   susceptibility-derived distortions \citep{jezzard_correction_1995}.
+% \cite{greve_accurate_2009} circumvented the nonlinearity by identifying typically
+%   distorted regions and removing them from the metric computation.
+% Second, our matching criteria does not necessarily need for clear intensity
+%   steps at the tissue interfaces where surfaces must be fitted, thanks to the use of
+%   active contours without edges \citep{chan_active_2001}.
+% This choice is similar to \citep{guyader_combined_2011}, but \emph{regseg} differs in two
+%   features.
+% First, the regularization relies on the inherent smoothness of the deformation field,
+%   and on the anisotropic approach proposed by \cite{nagel_investigation_1986}.
+% Second, it accepts several contours to be mapped as in \citep{gorthi_active_2011}.
+% However, \emph{regseg} makes use of the explicit representation of surfaces, and
+%   does not implement multiphase level sets.
+% Our registration method is demonstrated here as a solution to correct for the susceptibility
+%   distortions that \gls*{dmri} of the head typically present \citep{jezzard_correction_1995}.
+% The distortion correction and the precise segmentation of \gls*{dmri} data is crucial in
+% 	applications such as connectome extraction \citep{craddock_imaging_2013}, since the
+%   definition of the nodes in the network is generally derived from a parcellation performed in
+%   an anatomical \gls*{mri} of the same subject \citep{daducci_connectome_2012}.
 
@@
   \end{equation}
-
+%
   where $P(R \mid U,\Gamma_{l,m})$ is the data-likelihood, and
@@
   \end{equation}
-
+%
   where $\vec{f}' = R(\vec{r}')$ is the feature vector at the displaced
@@
   \end{align}
-
+%
   using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
@@
   \begin{align*}
-  P(U) = \underset{\vec{r}}{\prod}\, p(u(\vec{r})) &=
-  \underset{\vec{r}}{\prod}\, p_0(u(\vec{r})) \, p_1(u(\vec{r})), \\
-  p_0(u(\vec{r})) &= \mathcal{N}( u(\vec{r}) \mid 0, \mathbf{A}^{-1}), \notag\\
-  p_1(u(\vec{r})) &= \mathcal{N}(  \nabla \cdot u(\vec{r}) \mid 0, \mathbf{B}^{-1}),
+  P(U) = \underset{\vec{r}}{\prod}\, p(\vec{u}) &=
+  \underset{\vec{r}}{\prod}\, p_0(\vec{u}) \, p_1(\vec{u}), \\
+  p_0(\vec{u}) &= \mathcal{N}( \vec{u} \mid 0, \mathbf{A}^{-1}), \notag\\
+  p_1(\vec{u}) &= \mathcal{N}(  \nabla \vec{u} \mid 0, \mathbf{B}^{-1}),
   \end{align*}
-
+%
   imposing that the distortion and its gradient have zero
-  mean and variance governed by the matrices $\mathbf{A}$ and $\mathbf{B}$.
-Since the anisotropy is generally aligned with the imaging axes, these will be simplified
-  in the following for the sake of clarity:
-
-  \begin{align}
-    p_0(u(\vec{r})) &= \underset{\vec{r}}{\sum} \mathcal{N}( u(\vec{r}) \mid 0,
-      (\boldsymbol{\alpha}^{\circ\frac12}\,\vec{I}_n)^{-1}), \notag\\
-    p_1(u(\vec{r})) &= \underset{\vec{r}}{\sum} \mathcal{N}( \nabla \cdot u(\vec{r}) \mid 0,
-      (\boldsymbol{\beta}^{\circ\frac12}\,\vec{I}_n)^{-1}).
-  \label{eq:priors}
-  \end{align}
+  mean and variance governed by the matrices $\mathbf{A} = (a_{ij})$ and $\mathbf{B} = (b_{ij})$.
+Since the anisotropy is generally aligned with the imaging axes, $\mathbf{A}$ and $\mathbf{B}$
+  can be simplified to diagonal matrices.
+For convenience, we will set $\alpha_i = a_{ii}^2$ and $\beta_i = b_{ii}^2$, what can be reduced
+  using the Hadamard square root notation\footnote{The Hadamard power of a matrix or a vector
+  is the power of its elements $\mathbf{M}^{\circ p} = ({m_{ij}}^{p})$} as
+  $\mathbf{A}=\boldsymbol{\alpha}^{\circ\frac12}\,\vec{I}_n$ and
+  $\mathbf{B}=\boldsymbol{\beta}^{\circ\frac12}\,\vec{I}_n$.
 
@@
   \underset{\vec{r} \in \Omega_l}{\prod}
-  \mathcal{N} \left( \vec{f}' \mid \Theta_l \right)\,p_0( u(\vec{r}))\,p_1( u(\vec{r})) \notag\\
+  \mathcal{N} \left( \vec{f}' \mid \Theta_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) \notag\\
   &= C + \underset{l}{\sum} \int_{\Omega_l}
   \mdist{f'}{l} \,d\vec{r} \notag\\
-  &+ \int_{\Omega} \left[ \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
-  + \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2} \right] \,d\vec{r},
+  &+ \int_{\Omega} \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
+  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r},
   \label{eq:energy}
   \end{align}
-
+%
   this expression is dual to the energy functional corresponding
@@
 The transform $\hat{U}$ \eqref{eq:transform} is supported by a dense deformation field
-  $u(\vec{r})$, such that:
+  $\vec{u}$, such that:
 
@@
 Since the nodes of the anatomical surfaces likely lay off-grid, it is required to
-  derive $u(\vec{r})$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
+  derive $\vec{u}$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
 Densification is achieved through a set of associated basis functions $\psi_k$:
@@
   \begin{equation}
-  u(\vec{r}) = \sum_k \psi_k(\vec{r}) \vec{u}_k.
+  \vec{u} = \sum_k \psi_k(\vec{r}) \vec{u}_k.
   \label{eq:intp_kernel}
@@
   \end{equation}
-
+%
   with $\hat{\mathbf{e}}_d$ being the unitary vector along axis $d$.
@@
   \end{equation}
-
+%
   with $t$ being an artificial time parameter of the contour
@@
   \int_{\Omega} \underset{l}{\sum} \mdist{f'}{l} \,d\vec{r} \notag\\
-  &+ \int_{\Omega} [ \boldsymbol{\alpha} \cdot u(\vec{r})^{\circ2}
-  + \boldsymbol{\beta} \cdot (\nabla \cdot u(\vec{r}))^{\circ2} ] \,d\vec{r}
+  &+ \int_{\Omega} [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
+  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
   \Big\}.
@@
   \end{align}
-
+%
   in this case, the formulation has been adapted to the non-binary case, $\{l,m\}$
@@
   \end{equation}
-
+%
   then the full gradient evolution equation \eqref{eq:gradient_descent} yields:
@@
 
-\todo[inline]{Include here the final update equation using the FT, introducing the
-step size $\tau$, and refer to the supplementals for intermediate steps}
+Finally, we introduce a step size parameter $\delta$ to discretize the artificial time parameter
+  in the iterative optimization.
+Then, applying the Fourier transform as detailed in the \suppl{}, section S1, the
+  update equation is derived from \eqref{eq:gradient_final}:
+
+  \begin{align}
+  \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\vec{u}_k^t / \delta - \sum_i \vec{g}_{i,k}\}}%
+                  {\mathcal{F}\{(1/\delta+\boldsymbol{\alpha})I-\boldsymbol{\beta}\Delta\}} \right\}
+  \label{eq:update_equation}
+  \end{align}
 
@@
 \label{sec:conv_report}
-The registration parameters (such as $\tau$, the B-Spline grid resolutions,
-  target image smoothing, etc.) and implementation details (for instance the sparse matrix approach
+The registration parameters (such as $\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$,
+  the B-Spline grid resolutions, target image smoothing, etc.)
+  and other implementation details (for instance the sparse matrix approach
   to fast interpolation) are discussed in the \suppl{}, section S1.

### 2015-03-13 Publications/2015-NeuroImage/06-discussion.tex
one more iteration to second round

@@
 \label{sec:discussion}
-In section \nameref{sec:methods_background}, we framed our method within the current
-  state of art and defined the contributions in our approach.
-We chose active contours without edges as we were interested in the homogeneous
-  regions of diffusion properties within the brain.
-Iteratively, the active contours are updated following the instant shape gradients
-  (see \nameref{app:shape_priors} in Appendix) with a gradient descent strategy.
-The approach is valid for image processing applications in which a precise segmentation is
-  available (from an atlas or within subject anatomical data), and match any or all
-  the following circumstances:
-  i) reference and target images do not present appropriate contrasts to
-  	perform volume-based registration reliably;
-  ii) the low resolution of the target volume produces strong partial volume effects
-  	that preclude voxel-wise segmentation algorithms and hinder volume-based
-  	registration; and,
-  iii) the estimation of the unknown warping is explicitly needed and the
-    deformation field is smooth enough to be represented with a sufficient number
-  	of B-Spline kernels (i.e. the distortion in \gls*{dmri} images,
-  	object tracking in time series, etc.).
-To support our claim that images can be simultaneously segmented and registered to a
-  reference coordinate system, we conducted evaluation experiments on digital
-  simulated phantoms and real datasets.
+% In section \nameref{sec:methods_background}, we framed our method within the current
+%   state of art and defined the contributions in our approach.
+% We chose active contours without edges as we were interested in the homogeneous
+%   regions of diffusion properties within the brain.
+% Iteratively, the active contours are updated following the instant shape gradients
+%   (see \nameref{app:shape_priors} in Appendix) with a gradient descent strategy.
+% The approach is valid for image processing applications in which a precise segmentation is
+%   available (from an atlas or within subject anatomical data), and match any or all
+%   the following circumstances:
+%   i) reference and target images do not present appropriate contrasts to
+%   	perform volume-based registration reliably;
+%   ii) the low resolution of the target volume produces strong partial volume effects
+%   	that preclude voxel-wise segmentation algorithms and hinder volume-based
+%   	registration; and,
+%   iii) the estimation of the unknown warping is explicitly needed and the
+%     deformation field is smooth enough to be represented with a sufficient number
+%   	of B-Spline kernels (i.e. the distortion in \gls*{dmri} images,
+%   	object tracking in time series, etc.).
+% To support our claim that images can be simultaneously segmented and registered to a
+%   reference coordinate system, we conducted evaluation experiments on digital
+%   simulated phantoms and real datasets.
 

### 2015-03-13 Publications/2015-NeuroImage/08-acknowledgements.tex
modify font, add lineno
fix acks

@@
 
-\section*{Acknowledgment}
+\section*{Acknowledgments}
 The authors thank V. Estellers for early discussions at the beginning of this project,
-  and L. Vese for her support during OE's research visits in her laboratory.
+  and L. A. Vese for her support during OE's research visits in her laboratory.
+We also thank A. Leemans for kindly sharing a p-code version of \emph{ExploreDTI} from
+  which settings for \emph{elastix} could be extracted.
 
+DZ is supported by the Swiss National Science Foundation under grant PBELP2-137727.
 This study is supported by: the Spain's Ministry of Science and Innovation
-  (projects TEC2010-21619- C04-03, TEC2011-28972-C02-02, CDTI-CENIT
-  AMIT and INNPACTO PRECISION), Comunidad de Madrid (ARTEMIS) and
+  (projects TEC2013-48251-C2-2-R and INNPACTO XIORT), Comunidad de Madrid (TOPUS) and
   European Regional Development Funds; the Center for Biomedical Imaging
@@
   Leenaards and Louis Jeantet foundations.
-DZ is supported by the Swiss National Science Foundation (SNF) under grant PBELP2-137727.

### 2015-03-14 Publications/2015-NeuroImage/03-introduction.tex
updating intro

@@
 An early integration of segmentation and registration by \cite{bertalmio_morphing_2000} proposed
-  a sequential deformation of active contours for object tracking in sequential 2D images.
+  a sequential deformation of active contours for object tracking in series of 2D images.
 Shortly, \cite{yezzi_variational_2003} presented the first method including a full solution to
@@
 
-We propose \emph{regseg}, a registration method based on active contours without edges
-  \citep{chan_active_2001} to drive registration and segment the target image.
-The hypothesis underpinning our research is that image registration can be reliably performed
-  by searching for homogeneous regions in the target image that correspond to precise contours
-  from an atlas, or extracted from other image (i.e. a different time step).
-We demonstrate the approach in correcting \gls*{dmri} for nonlinear distortions, an application
-  that is described in \autoref{sec:distortions}.
+We propose \emph{regseg}, an active-contours driven registration method.
+Then, we demonstrate its aptness on the correction of the nonlinear distortions typically
+  presented by \gls*{dmri} data of the brain.
+This problem is briefly described in \autoref{sec:distortions}.
+The general overview of our method, along with the contributions with respect 
+  to the current state of art are summarized in \autoref{sec:contributions}.
 In \autoref{sec:methods_map}, we start from a probabilistic framework to finally show the duality
   between the methods described previously.
-Following sections introduce
-  the specific features of \emph{regseg} (\autoref{sec:numerical_implementation}),
+Following sections
+  introduce the specific features of \emph{regseg} (\autoref{sec:numerical_implementation}),
   state the evaluation experiments (\autoref{sec:experiments_evaluation}),

### 2015-03-14 Publications/2015-NeuroImage/04-methods.tex
updating intro

@@
   \gls*{dmri} dataset that are acquired with a null or very low gradient intensity or
-   \emph{b}-value used as signal reference.} of the \gls*{dmri} versus a \gls*{t2} image.
+   \emph{b}-value.} of the \gls*{dmri} versus a \gls*{t2} image.
 Further developments \citep{irfanoglu_susceptibility_2011} extended this method to the
@@
 \subsection{Our method and contributions}
-\label{sec:our_method}
+\label{sec:contributions}
+%
+\emph{Regseg} can be viewed as a derivation of \emph{bbregister} \citep{greve_accurate_2009}.
+Both methods exploit the anatomically correct and precise surfaces that are readily
+  extracted with \emph{FreeSurfer} \citep{fischl_freesurfer_2012}, mapping them
+  to a target \gls*{dmri} image using active contours.
+We implemented active contours without edges \citep{chan_active_2001}, that allow
+  \emph{regseg} to segment homogeneous regions within the brain, while \emph{bbregister}
+  use active contours with edges to look for intensity boundaries in the target image.
+Another identifying difference is the deformation model:
+  \emph{bbregister} includes an affine transformation and overcomes the distortion
+  problem by excluding typically affected regions from computation.
+Conversely, \emph{regseg} implements a free-deformation field supported with B-Spline
+  functions.
+Our method is close to \citep{guyader_combined_2011} and \citep{gorthi_active_2011}
+  in the use of active contours without edges for registration, however we avoided
+  the use of level sets with the application of shape gradients
+  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
+\cite{gorthi_active_2011} used a gaussian regularizer, and \cite{guyader_combined_2011}
+  proposed a nonlinear elasticity smoother.
+In contrast, \emph{regseg} implements the anisotropic regularization proposed by
+  \citep{nagel_investigation_1986}.
+We also present an ellaborated evaluation instrumentation to create realistic distortions
+  and test algorithms.
+Integrating \emph{regseg} and an in-house implementation of the \gls*{t2b} correction method
+  in the workflow, we compared the performance of both solutions.
+
 % \emph{Regseg} is related to \emph{bbregister}, as both methods exploit the precise
@@
 The transform $\hat{U}$ \eqref{eq:transform} is supported by a dense deformation field
-  $\vec{u}$, such that:
+  $\vec{u} = u(\vec{r})$, such that:
 
   \begin{equation}
-  \vec{v}_i' = \hat{U}\{\vec{v}_i\} = \vec{v}_i + u(\vec{v}_i).
+  \vec{v}_i' = \hat{U}\{\vec{v}_i\} = \vec{v}_i + u(\vec{v}_i) = \vec{v}_i + \vec{u}_i.
   \label{eq:nodes_tfm}
@@
 Since the nodes of the anatomical surfaces likely lay off-grid, it is required to
-  derive $\vec{u}$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
+  derive $u(\vec{r})$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
 Densification is achieved through a set of associated basis functions $\psi_k$:
@@
   \begin{equation}
-  \vec{u} = \sum_k \psi_k(\vec{r}) \vec{u}_k.
+  u(\vec{r}) = \sum_k \psi_k(\vec{r}) \vec{u}_k.
   \label{eq:intp_kernel}
@@
   \begin{equation}
-  \frac{\partial u(\vec{r},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}_k},
+  \frac{\partial u(\vec{r},t)}{\partial t} \propto - \frac{\partial E(R \mid U)}{\partial \vec{u}_k},
   \label{eq:general_gradient_descent}
@@
 	 database of the \gls*{hcp}.
-\todo[inline]{This is partially true: I am currently processing the data and
-only 5 subjects with regseg and 10 with the competing method are processed.
-All the results presented afterwards are based on this processed cases, and will be
-updated for the final submission.}
 We refer the reader to \citep{essen_human_2012} for exact details about acquisition

### 2015-03-16 Publications/2015-NeuroImage/00-main.tex
updating intro
almost ready to revision round 2

@@
 \documentclass[5p,authoryear]{elsarticle}
-\journal{ArXiv archiving for pre-submission review}
+\journal{ArXiv archiving for pre-publication review}
 
@@
 \providecommand{\nvec}[1]{\hat{\mathbf{#1}}}
-\providecommand{\suppl}{\href{http://dx.doi.org/mydoi}{Supplemental Materials}}
+\providecommand{\suppl}[1]{\href{http://dx.doi.org/mydoi}{Supplemental Materials}, #1}
 
@@
 
-\title{Active contours driven registration of diffusion MR images of the brain}
+\title{Active-contours driven registration of diffusion MR images of the brain}
 

### 2015-03-16 Publications/2015-NeuroImage/03-introduction.tex
almost ready to revision round 2

@@
   and \emph{hybrid}, respectively.
-When segmentation is introduced to define the matching criteria, we face a \emph{hybrid}
-  registration method that simultaneously defines a partition of the reference space to identify
-  relevant objects in it.
+When the matching criteria is defined over an optimizable partition of one or both reference an
+  moving spaces, we face a hybrid method which simultaneously segments and registers images.
 An early integration of segmentation and registration by \cite{bertalmio_morphing_2000} proposed
@@
 Shortly, \cite{yezzi_variational_2003} presented the first method including a full solution to
-  the registration problem with an affine transformation supporting the coordinates mapping.
+  the registration problem with an affine transformation supporting the mapping.
 \cite{vemuri_joint_2003} proposed an atlas-based registration framework using level sets and only
@@
 Their framework performs robust registration of brain surfaces into the intensity information
-  of a target \gls*{mri}, using an affine transformation and active contours with edges.
+  of a target \gls*{mri}, using an affine transformation and active contours.
 Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
@@
   to the current state of art are summarized in \autoref{sec:contributions}.
-In \autoref{sec:methods_map}, we start from a probabilistic framework to finally show the duality
-  between the methods described previously.
+In \autoref{sec:methods_map}, we start from a probabilistic framework to finally show its duality
+  with respect to the active-contours route.
 Following sections
@@
   in the described application, compared to a well-established tool.
-The interpretation of the results, and prospects, are addressed in \autoref{sec:discussion}.
+The interpretation of the results and prospects is addressed in \autoref{sec:discussion}.

### 2015-03-16 Publications/2015-NeuroImage/04-methods.tex
almost ready to revision round 2

@@
   from the main magnetic field $B_0$.
-The principal source of these deviations ($\Delta B_0$) is the discontinuity of magnetic
-  susceptibility at tissue interfaces.
+One important source of these deviations ($\Delta B_0$) is the discontinuity of magnetic
+  susceptibility across the imaged object.
 Tissue/air boundaries show large steps of susceptibility and this artifact is
@@
 %
-  where $\vec{r}$ denotes a position in the reference domain $R$, $\vec{r}'$ is
+  where $\vec{r}$ denotes a position in $R$, $\vec{r}'$ is
   its corresponding location in $M$, and $n$ the dimensionality of images.
@@
 where $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time,
-  $s_{PE}$ is the pixel spacing along \gls*{pe}, and $\hat{\vec{e}}_{PE}$ the unitary
+  $s_{PE}$ is the pixel size along \gls*{pe}, and $\hat{\vec{e}}_{PE}$ the unitary
   vector along \gls*{pe}.
@@
   increments \citep{cordes_geometric_2000,holland_efficient_2010}.
-\cite{kybic_unwarping_2000} proposed a correction using nonlinear registration of the \emph{b0}
+\cite{kybic_unwarping_2000} proposed nonlinear registration of the \emph{b0}
   image\footnote{The term \emph{b0} refers to one or more images within the
   \gls*{dmri} dataset that are acquired with a null or very low gradient intensity or
-   \emph{b}-value.} of the \gls*{dmri} versus a \gls*{t2} image.
-Further developments \citep{irfanoglu_susceptibility_2011} extended this method to the
-  three-dimensional problem.
+   \emph{b}-value.} of the \gls*{dmri} to an undistorted \gls*{t2} image of
+   the same subject.
+In the following, we will refer to this method as \gls*{t2b}.
+Further developments \citep{irfanoglu_susceptibility_2011} extended the \gls*{t2b} 
+  approach to the 3D problem.
 Therefore, a typical protocol for the extraction of structural connectivity
@@
   and test algorithms.
-Integrating \emph{regseg} and an in-house implementation of the \gls*{t2b} correction method
+Integrating \emph{regseg} and an in-house implementation of the \gls*{t2b} correction
   in the workflow, we compared the performance of both solutions.
 
-% \emph{Regseg} is related to \emph{bbregister}, as both methods exploit the precise
-%   surfaces that can be extracted with \emph{FreeSurfer} \citep{fischl_freesurfer_2012},
-%   to then map this prior information into a target image.
-% However, \emph{bbregister} features an affine transform that is unable to compensate the
-%   distortions.
-% To avoid this problem, the active contours are dismissed in regions typically affected by
-%   warpings.
-% Additionally, \emph{bbregister} is robust in finding clear tissue boundaries.
-% In the case of \gls*{dmri} data, oftentimes the registration process is restricted to finding
-%   the pial surface, as it is the only clear boundary.
+\subsection{Cost-function derivation}\label{sec:methods_map}
 %
-% to the methods based on nonlinear registration of the
-%   to the \gls*{t2} image, since we also render the correction as a
-%   registration problem.
-% Conversely, the method does not require any of the previously mentioned extra acquisitions.
-% A fundamental antecedent to \emph{regseg} is found in .
-% They proposed a registration method to align surfaces extracted from the \gls*{t1}
-%   to \gls*{dmri} images using a linear transformation model using active contours
-%   with edges.
-% Similarly, we  warped data, in which the scale of the
-%   imaged structures is in the range of the image resolution or above.
-% Our approach differs from \citep{greve_accurate_2009} in two fundamental choices.
-% First, we use a nonlinear deformation model that enables the correction of the
-%   susceptibility-derived distortions \citep{jezzard_correction_1995}.
-% \cite{greve_accurate_2009} circumvented the nonlinearity by identifying typically
-%   distorted regions and removing them from the metric computation.
-% Second, our matching criteria does not necessarily need for clear intensity
-%   steps at the tissue interfaces where surfaces must be fitted, thanks to the use of
-%   active contours without edges \citep{chan_active_2001}.
-% This choice is similar to \citep{guyader_combined_2011}, but \emph{regseg} differs in two
-%   features.
-% First, the regularization relies on the inherent smoothness of the deformation field,
-%   and on the anisotropic approach proposed by \cite{nagel_investigation_1986}.
-% Second, it accepts several contours to be mapped as in \citep{gorthi_active_2011}.
-% However, \emph{regseg} makes use of the explicit representation of surfaces, and
-%   does not implement multiphase level sets.
-% Our registration method is demonstrated here as a solution to correct for the susceptibility
-%   distortions that \gls*{dmri} of the head typically present \citep{jezzard_correction_1995}.
-% The distortion correction and the precise segmentation of \gls*{dmri} data is crucial in
-% 	applications such as connectome extraction \citep{craddock_imaging_2013}, since the
-%   definition of the nodes in the network is generally derived from a parcellation performed in
-%   an anatomical \gls*{mri} of the same subject \citep{daducci_connectome_2012}.
-
-\subsection{Cost-function derivation}\label{sec:methods_map}
-
 In a Bayesian framework, the mappings $U$ in \eqref{eq:transform} are
@@
 	The gradient speeds $\bar{s}_i$ are computed as the disparity between data energies of
-	  the vertex $\vec{v}_i$ in each limiting region (see \autoref{app:shape_priors}).
+	  the vertex $\vec{v}_i$ in each limiting region (see \ref{app:shape_priors}).
 	In this figure, the gradient speed corresponding to $\vec{v}_1$ is written in the lower
@@
 	Finally, every $\bar{s}_i$ is associated with the control points $\vec{u}_k$ through
-	  the corresponding weights $\psi_k(\vec{v}_i)$ \eqref{eq:basis_derivative}.
+	  the corresponding weights $\psi_k(\vec{v}_i)$ by Eq. \eqref{eq:gradient_wshape}.
 	}\label{fig:method}
@@
 
-We can apply a discretized interpretation of \eqref{eq:shape_gradients} to compute
+We apply a discretized interpretation of \eqref{eq:shape_gradients} to compute
   the data term in \eqref{eq:gradient_descent} using shape-gradients
@@
   \frac{ \partial }{\partial \vec{u}_k} \left\{
-   \underset{\vec{x} \in \Omega_l}{\sum} \underset{l}{\sum} \mdist{f'}{l} \right\}
-  = \underset{i}{\sum}
+   \underset{\vec{x} \in \Omega_l}{\sum} \underset{l}{\sum} \mdist{f'}{l} \right\} \notag\\
+  &= \underset{i}{\sum} w_i
    \left\langle \frac{\partial \vec{v}_i'}{\partial \vec{u}_k}, \bar{s}_i'\right\rangle,
@@
     them such that
-    $\vec{x}' = \vec{v}' \in\Gamma_{l,m} \iff \vec{x}\in \partial\Omega_i \cap \partial\Omega_j$
-    and $\bar{s}_i'$ is the speed vector projected on to the unit inward normal to the contour
-    at $c_i'$ (described in \autoref{fig:method}).
+    $\vec{x}' = \vec{v}' \in\Gamma_{l,m} \iff \vec{x}\in \partial\Omega_i \cap \partial\Omega_j$,
+    $\bar{s}_i'$ is the speed vector projected on to the unit inward normal to the contour
+    at $c_i'$ (described in \autoref{fig:method}), $w_i$ is the area contribution of vertex
+    $c_i'$ to the the total area of the surface it belongs.
 
@@
   \end{align}
-
+%
   where $\hat{\vec{e}}$ is the coordinates system's unit vector.
@@
   \begin{equation}
-  \vec{g}_{i,k} = \left\langle \frac{\partial}{\partial \vec{u}_k}{\vec{v}_i}', \bar{s}_i'\right\rangle
-  = - \left[ \mdist{f_i'}{l} - \mdist{f_i'}{m} \right] \psi_k(\vec{v}_i)\, \hat{\vec{e}},
+  \vec{g}_{i,k} = w_i \, \left\langle \frac{\partial}{\partial \vec{u}_k}{\vec{v}_i}', \bar{s}_i'\right\rangle
+  = - w_i \left[ \mdist{f_i'}{l} - \mdist{f_i'}{m} \right] \psi_k(\vec{v}_i)\, \hat{\vec{n}}_i,
   \end{equation}
@@
   in the iterative optimization.
-Then, applying the Fourier transform as detailed in the \suppl{}, section S1, the
+Then, applying the Fourier transform as detailed in the \suppl{section S1}, the
   update equation is derived from \eqref{eq:gradient_final}:
@@
   and other implementation details (for instance the sparse matrix approach
-  to fast interpolation) are discussed in the \suppl{}, section S1.
+  to fast interpolation) are discussed in the \suppl{section S1}.
 The actual choices of parameter settings are publicly distributed with the source code of the experiments.
@@
   demonstrate the behavior of \emph{regseg} and help scientists configure their own experiments.
-One sample report is found in the \suppl{}, section S1.3.
+One sample report is found in the \suppl{section S1.3}.
 
@@
 In order to demonstrate the performance of \emph{regseg}, we first conducted a battery of
-  accuarcy tests over synthetic data, and then applied it in the susceptibility distortion
+  accuracy tests over synthetic data, and then applied it in the susceptibility distortion
   correction of real data.
@@
 Therefore, a \gls*{swindex} can be computed as the one-to-one distance between corresponding
-  vertices of surfaces, weighted by their respective Voronoi area $w_i$:
+  vertices of surfaces, weighted by their respective Voronoi area $a_i$:
 
   \begin{equation}
-  sWI = \frac{1}{P} \sum\limits_p^P \frac{1}{A_p} \sum\limits_i^{N_p} w_i\,\|
+  sWI = \frac{1}{P} \sum\limits_p^P \frac{1}{A_p} \sum\limits_i^{N_p} a_i\,\|
   \vec{v}_i - \hat{\vec{v}}_i \|
@@
   \end{equation}
-
+%
   where $\vec{v}_i$ are the locations of the $N_p$ vertices in each $p \in \{1, \dots P\}$
@@
 	  the ground-truth deformation.
-	  A cross-comparison methodology is also applied, to obtain a competing $\hat{U}_{cc}$.
 	5) Results are visually and quantitatively evaluated.}\label{fig:evworkflows}
@@
 Simulations were corrupted with rician noise for a \gls*{snr} of 300.0.
-The reference surfaces were extracted using the marching-cubes algorithm
-  shipped with \emph{FreeSurfer} \citep{fischl_freesurfer_2012}
-  at the higher resolution (1.0mm isotropic).
+The reference surfaces were extracted at the higher resolution (1.0mm isotropic),
+  using \texttt{mri\_tessellate}\footnote{A marching-cubes algorithm shipped with 
+  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}}.
 
@@
 	In yellow color, the recovered contours after registration are represented.
-	Our method showed high accuracy, as they are overlapping the ground truth surfaces
-	  depicted in green.
+	Our method showed high accuracy, as demonstrates almost exact location of the registered
+    contours with respect to their ground truth position depicted in green.
 	Partial volume effect turns segmentation of the sulci a challenging problem with voxel-wise
@@
 	  surfaces at high (left) and low (right) resolutions, demonstrating that the error is
-	  consistently below the image resolution.
+	  consistently below the voxel size.
 	  }\label{fig:phantom}
 \end{figure*}
+
+To warp the simulated datasets, we generated 150 different realizations of
+  the distortion $U_{true}$.
+Smoothness is introduced using two levels of B-Spline functions, with control points evenly
+  located in isotropic grids covering the full extent of the phantom.
+The first level had $50.5mm$ of separation between control points and the second $25.25mm$.
+The rationale behind this choice is producing large deformations on the contours with
+  the first level, and matching the properties of susceptibility-derived distortions
+  as previously reported by \cite{irfanoglu_susceptibility_2011} with the finer grid.
+The coefficients of the B-Spline functions were randomly generated for each resolution level.
+Invertibility of $U_{true}$ is ensured by controlling the maximum displacement at each
+  level ($20.2mm$ and $10.1mm$ respectively) as studied in \citep{rueckert_diffeomorphic_2006}.
 
@@
 Selecting the appropriate labels in the \emph{aparc} segmentation, we applied
-  the marching-cubes algorithm again to extract the surface of the following
+  \texttt{mri\_tessellate} to extract the surface of the following
   six homogeneous regions $\Omega_l$:
@@
   3) cerebral \gls*{wm}, $\Gamma_{WM}$;
-  4) brain stem and cerebellar \gls*{wm}, $\Gamma_{Bst}$;
+  4) brain stem and cerebellar \gls*{wm}, $\Gamma_{bst}$;
 	5) cerebellar \gls*{gm}, $\Gamma_{cbGM}$; and
@@
   the derived scalars (\gls*{fa} and \gls*{adc}) using \emph{MRtrix} \citep{tournier_mrtrix_2012}.
+The stack of \gls*{fa} and \gls*{adc} conforms the multivariate input for \emph{regseg}.
+
+A dual workflow to the general evaluation framework (\autoref{fig:evworkflows})
+  was also implemented to integrate the \gls*{t2b} registration scheme.
+We reproduced the solution and settings provided with \emph{ExploreDTI}
+  \citep{leemans_exploredti_2009}, a widely used toolkit for tractography analysis of
+  \gls*{dti}.
+\emph{ExploreDTI}, internally uses \emph{elastix} \citep{klein_elastix_2010} to
+  perform registration.
+First, the average \emph{b0} volume is computed using all the \emph{low-b} volumes within
+  the distorted \gls*{dmri} dataset.
+Using the ``preprocessed'' instance of the \gls*{t2} image of the corresponding subject as
+  reference image, and the \emph{b0} image as moving image, we registered both and
+  projected the correct (undistorted) contours to the warped space using the resulting
+  transform.

### 2015-03-16 Publications/2015-NeuroImage/05-results.tex
updating paper for second round of reviews
almost ready to revision round 2

@@
 % -*- root: 00-main.tex -*-
-\section*{Results}
+\section{Results}
 \label{sec:results}
 
-\subsection*{Proof of concept on digital phantoms}
+\subsection{Proof of concept on digital phantoms}
 \label{sec:results_phantom}
-Using a simplified subset of the experimental instrumentation described in section
-  \nameref{sec:experiments_evaluation}, we collected quantitative results on the digital
+Using a simplified subset of the experimental instrumentation described in
+  \autoref{sec:experiments_evaluation}, we collected quantitative results on the digital
   phantoms.
-To warp the simulated datasets, we randomly generated 150 different realizations of
-  the distortion $U_{true}$.
-Smoothness is introduced using two levels of B-Spline functions, with control points evenly
-  located in isotropic grids covering the full extent of the phantom.
-The first level had $50.5mm$ of separation between control points and the second $25.25mm$.
-The rationale behind this choice is producing large deformations on the contours with
-  the first level, and matching the properties of susceptibility-derived distortions
-  as previously reported by \cite{irfanoglu_susceptibility_2011} with the finer grid.
-Invertibility of $U_{true}$ is ensured by controlling the maximum displacement at each
-  level ($20.2mm$ and $10.1mm$ respectively) as studied in \citep{rueckert_diffeomorphic_2006}.
-
 A total of 1200 registration experiments (4 phantom types $\times$ 150 random warpings
@@
   the source code.
-For each registration, the averaged Hausdorff distance (see \nameref{sec:experiments_evaluation})
+For each registration, the averaged Hausdorff distance (see \autoref{sec:experiments_evaluation})
   of the inner and the outer surfaces was computed.
-The results showed a consistent and high accuracy, below voxel resolution.
+The results showed a consistent and high accuracy, below the image resolution.
 \autoref{fig:phantom} (block C) presents the box plots by model type corresponding
@@
   correction for multiple comparisons ($N=150$).
-Averaged errors resulted significantly lower than voxel resolution with $p < (0.001 / 150)$
+Averaged errors resulted significantly lower than voxel size with $p < (0.001 / 150)$
   for all the tests (4 phantom types $\times$ 2 resolutions).
@@
     \end{tabular}
-    \caption{All the 95\% confidence intervals (CI) were below the corresponding
-      half-voxel resolution in all the phantom types.
-    CIs reported in this table were computed using bootstrapping of $10^4$ samples,
+    \caption{95\% Confidence Intervals (CIs) were computed using bootstrapping of $10^4$ samples,
       for each phantom type and resolution.
     Taking into account the non-normal distribution of errors, we used the median as location
-  		statistic.}\label{tab:ci_phantom}
+  		statistic.
+    All the CIs were below the corresponding half-voxel size in all the phantom types.}\label{tab:ci_phantom}
 \end{table}
 
-\subsection*{Correction of distortions on real datasets}\label{sec:results_hcp}
+\subsection{Correction of distortions on real datasets}\label{sec:results_hcp}
 
@@
   (\gls*{fa} and \gls*{adc}).
-Based on the observed joint distribution of each (\suppl{}, {\color{red} section S2}),
-  we selected the 6-regions scheme described in \nameref{sec:human_connectome}.
+Based on the observed joint distribution of each (\suppl{section S2}),
+  we selected the 6-regions scheme described in \autoref{sec:human_connectome}.
 
@@
 	\eqref{eq:energy} versus several registration errors generated as explained in
-	\nameref{sec:res_model_and_metric}.
+	\autoref{sec:res_model_and_metric}.
 	The metric of our registration method displayed a smooth gradient towards the minimum
@@
 %
-Finally, we tested the performance of our registration method, and compared it to a competing
-  method: the \emph{b0}-to-\gls*{t2} registration approach.
-To this end, we integrated the registration software \emph{elastix} \citep{klein_elastix_2010},
-  and reproduced the settings that internally uses \emph{ExploreDTI}
-  \citep{leemans_exploredti_2009}, a widely used toolkit for tractography analysis of
-  \gls*{dti}.
-A dual workflow to the general evaluation framework (\autoref{fig:evworkflows})
-  was also implemented to run the correction experiment with the competing method.
-First, the average \emph{b0} volume is computed using all the \emph{low-b} volumes within
-  the distorted \gls*{dmri} dataset.
-Using the ``preprocessed'' instance of the \gls*{t2} image of the corresponding subject as
-  reference image, and the \emph{b0} image as moving image, we registered both and
-  projected the correct (undistorted) contours to the warped space using the resulting
-  transform.
-For both \emph{regseg} and the competing method, we computed the \gls*{swindex} \eqref{eq:swindex}
+Finally, we tested the performance of our registration method, and compared it to the
+  a standard \gls*{t2b} correction.
+For both \emph{regseg} and the \gls*{t2b} methods, we computed the \gls*{swindex} \eqref{eq:swindex}
   of every surface.
-Finally, to statistically compare the results from both methods we performed a one-way ANOVA test
+Finally, to statistically compare the results, we performed a one-way ANOVA test
   on the warping indices for three surfaces.
-Following the nomenclature introduced in section \nameref{sec:human_connectome}, we selected
+Following the nomenclature introduced in section \autoref{sec:human_connectome}, we selected
   $\Gamma_{VdGM}$, $\Gamma_{WM}$, and $\Gamma_{pial}$ as surfaces of interest.
@@
 The quantitative results of the ANOVA tests and the CIs are presented in \autoref{tab:results_real}.
+Visual assessment of one case is shown in \autoref{fig:results_real}, along with the corresponding
+  box plots of error distributions by correction method and surface.
+Extended visual assessment of the 16 cases is aggregated in the \suppl{section S4}.
 
@@
 	\includegraphics[width=\linewidth]{figures/figure05.pdf}
-	\caption{The evaluation instrument automatically renders the overlay of the
+	\caption{The evaluation instrument automatically renders the overlay between the
 	  target \gls*{fa} image and the resulting contours after distortion estimation (yellow color) and
@@
 	The last two rows represent the sagittal view.
-	We intentionally omitted the coronal slicing due to it is the less informative given the properties
-	  of the distortion.
+	We intentionally omitted the coronal slicing as it is the least informative, given the directional property
+	  of distortions.
 	Red arrows point to regions where the accuracy of the proposed method more clearly overperformed

### 2015-03-16 Publications/2015-NeuroImage/06-discussion.tex
updating intro
almost ready to revision round 2

@@
 % -*- root: 00-main.tex -*-
-\section*{Discussion}
+\section{Discussion}
 \label{sec:discussion}
-% In section \nameref{sec:methods_background}, we framed our method within the current
-%   state of art and defined the contributions in our approach.
-% We chose active contours without edges as we were interested in the homogeneous
-%   regions of diffusion properties within the brain.
-% Iteratively, the active contours are updated following the instant shape gradients
-%   (see \nameref{app:shape_priors} in Appendix) with a gradient descent strategy.
-% The approach is valid for image processing applications in which a precise segmentation is
-%   available (from an atlas or within subject anatomical data), and match any or all
-%   the following circumstances:
-%   i) reference and target images do not present appropriate contrasts to
-%   	perform volume-based registration reliably;
-%   ii) the low resolution of the target volume produces strong partial volume effects
-%   	that preclude voxel-wise segmentation algorithms and hinder volume-based
-%   	registration; and,
-%   iii) the estimation of the unknown warping is explicitly needed and the
-%     deformation field is smooth enough to be represented with a sufficient number
-%   	of B-Spline kernels (i.e. the distortion in \gls*{dmri} images,
-%   	object tracking in time series, etc.).
-% To support our claim that images can be simultaneously segmented and registered to a
-%   reference coordinate system, we conducted evaluation experiments on digital
-%   simulated phantoms and real datasets.
+
+The hypothesis underpinning our research is that image registration can be reliably performed
+  by searching for homogeneous regions in the target image that correspond to precise contours
+  from an atlas, or extracted from other image (i.e. a different time step).
 
@@
   a real environment, using the experimental framework presented in \autoref{fig:evworkflows}.
-We processed 16 subjects downloaded from the \gls*{hcp} database with both \emph{regseg}
+We processed 16 subjects from the \gls*{hcp} database with both \emph{regseg}
   and an in-house replication of a competing method.
@@
   experiments are prone to the so-called instrumentation bias \citep{tustison_instrumentation_2013}.
-Therefore, these results do not reflect that the \emph{regseg} software \emph{is} better than the
+Therefore, these results do not show that the \emph{regseg} software \emph{is} better than the
   competing method, but they are communicated to proof that \emph{regseg} is a solid option to
@@
 This integration has been recently applied in the \emph{b0}-to-\gls*{t2} approach
-  by \cite{irfanoglu_susceptibility_2011}, studied along with the virtues of using
+  by \cite{irfanoglu_susceptibility_2011}, also investigating the usefulness of
   nonregular B-Spline grids.

### 2015-03-16 Publications/2015-NeuroImage/07-conclusion.tex
almost ready to revision round 2

@@
   defined explicitly by their limiting surfaces.
-The surfaces are active, and evolve on a B-Spline displacements field, guided by shape
-  gradients computed on the current partition of the target image.
+The surfaces are active, and evolve on a free-deformation field supported by B-Splines.
+A descent optimization strategy is guided by shape gradients computed on the current partition
+  of the target image.
 \emph{Regseg} uses active contours without edges, and hence looks for

### 2015-03-16 Publications/2015-NeuroImage/06-discussion.tex
add potential applications to discussion

@@
   from an atlas, or extracted from other image (i.e. a different time step).
-
 We designed \emph{regseg} as a method to correct \gls*{dmri} data for susceptibility
   distortions.
-Connectome extraction protocols generally include the acquisition of \gls*{t1} images
+Applications of whole-brain tractography generally include the acquisition of \gls*{t1} images
   to provide with prior information about anatomy with high accuracy.
@@
 Moreover, our method does not require any additional acquisition over the minimal protocol
-  that only includes \gls*{t1} and \gls*{dmri} images, a very common scenario for
+  that only includes \gls*{t1} and \gls*{dmri} images, a very common situation for
   historical datasets.
+Beyond the proposed correction, potential applications of \emph{regseg} are atlas-based
+  segmentation and tracking objects in time-series.
 

### 2015-03-16 Publications/2015-NeuroImage/02-abstract.tex
update figure, updates in manuscript

@@
 % -*- root: 00-main.tex -*-
-Integrating information from different images is an ubiquitous necessity
-  in current neuroimage analyses.
-However, oftentimes one of the images is not available or the objective
-  function is not adequate or numerically tractable.
-We propose a surfaces-to-volume registration method called \emph{regseg}
-  that exploits the detailed anatomy extracted from structural images and
-  simultaneously segments the target image with subvoxel precision.
-We use active contours without edges to search for a deformation field
-  of B-Spline functions that optimally maps the surfaces
-  onto the data in the target space.
-We test \emph{regseg} on four digital phantoms warped with random deformations,
-  and conclude that the precision of the method is below the half-voxel spacing.
-Finally, we apply \emph{regseg} in the correction for susceptibility-derived
-  distortions of 16 real diffusion MRI datasets from the \acrlong*{hcp}
-  database.
+We propose a simultaneous segmentation and registration method called \emph{regseg}
+  that uses active contours without edges extracted from structural images.
+The contours evolve through a free-deformation field of B-Spline functions that
+  optimally maps the surfaces onto the data in the target space.
+\emph{Regseg} is intended, but not limited, to correct susceptibility-derived distortions
+  in echo-planar imaging.
+We proof the concept behind \emph{regseg} on four digital phantoms warped with random
+  deformations, and conclude that the precision of the segmentation is subvoxel.
+Finally, we demonstrate \emph{regseg} in the correction task on 16 real diffusion MRI
+  datasets from the \acrlong*{hcp}.
 The results suggest that \emph{regseg} overperformed an extended competing method.

### 2015-03-16 Publications/2015-NeuroImage/04-methods.tex
update figure, updates in manuscript

@@
 \begin{figure*}
-	\includegraphics[width=\linewidth]{figures/figure03}
+    \centering
+    \resizebox{\textwidth}{!}{%
+      \begin{tikzpicture}
+        \node[inner sep=5pt, draw=black!40](fig03a) at (0,14)
+          {\includegraphics[width=\linewidth]{figures/figure03-a}};
+        \node[inner sep=5pt, draw=black!40](fig03b) at (0,7.6)
+          {\includegraphics[width=\linewidth]{figures/figure03-b}};
+        \node[inner sep=5pt, draw=black!40](fig03c) at (0,0)
+          {\includegraphics[width=\linewidth]{figures/figure03-c}};
+        \node[circle, text=black!75] at (-8.8,16.2) {\Huge \textbf{A}};
+        \node[circle, text=black!75] at (-8.8,10.05) {\Huge \textbf{B}};
+        \node[circle, text=black!75] at (-8.8,3.4) {\Huge \textbf{C}};
+      \end{tikzpicture}
+    }%
 	\caption{A. The ``cortex'' phantom is a spherical shape with two sulci and an

### 2015-03-16 Publications/2015-NeuroImage/06-discussion.tex
update figure, updates in manuscript

@@
 After a preliminary experiment to demonstrate the aptness of our segmentation model
-  in section \nameref{sec:res_model_and_metric}, we evaluated \emph{regseg} in
+  in section \autoref{sec:res_model_and_metric}, we evaluated \emph{regseg} in
   a real environment, using the experimental framework presented in \autoref{fig:evworkflows}.

### 2015-03-16 Publications/2015-NeuroImage/05-results.tex
new violinplot for realdata

@@
 \begin{figure*}
-	\includegraphics[width=\linewidth]{figures/figure05.pdf}
+    \centering
+    \resizebox{\textwidth}{!}{%
+      \begin{tikzpicture}
+        \node[inner sep=15pt, draw=black!40](fig05a) at (0,11)
+          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure05-a}};
+        \node[inner sep=15pt, draw=black!40](fig05b) at (0,0)
+          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure05-b}};
+        % \node[inner sep=5pt, draw=black!40](fig03c) at (0,0)
+        %   {\includegraphics[width=\linewidth]{figures/figure03-c}};
+        \node[circle, text=black!75] at (-9.5,17) {\Huge \textbf{A}};
+        \node[circle, text=black!75] at (-9.5,3) {\Huge \textbf{B}};
+        % \node[circle, text=black!75] at (-8.8,3.4) {\Huge \textbf{C}};
+      \end{tikzpicture}
+    }%
 	\caption{The evaluation instrument automatically renders the overlay between the
@@
 	  the competing method.
-	\todo[inline]{This figure will also have the box-plots of the errors of surfaces for both regseg
-	and the t2b method}
 	}\label{fig:results_real}

### 2015-03-16 Scripts/pyacwereg/viz.py
new violinplots!

@@ -2,8 +2,9 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2014-12-11 15:08:23
-# @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-02-25 11:36:26
+# @Last Modified by:   oesteban
+# @Last Modified time: 2015-03-16 21:20:04
+import os.path as op
 
 
 def add_annotations(values, ax, level, nlevels, color, lastidx, units=''):
@@ -527,8 +528,121 @@ def slices_gridplot(in_files, view=['axial'], size=(3, 3), discard=3,
     return out_file
 
 
-def phantom_errors(in_csv, resolution='lo',
-                   mtypes=['ball', 'box', 'L', 'gyrus']):
+def phantom_errors(in_csv, size=(80, 30), out_file=None):
+    import pandas as pd
+    import seaborn as sn
+    import matplotlib.pyplot as plt
+    sn.set_context("poster", font_scale=4.5)
+    df = pd.read_csv(in_csv).drop_duplicates(subset=['surf_id', 'repetition',
+                                             'surfdist_avg', 'model_type'])
+    df.surf_id[df.surf_id == 0] = 'internal'
+    df.surf_id[df.surf_id == 1] = 'external'
+    mtypes = df.model_type.unique()
+    cols = len(mtypes) * 2
+    fig, (ax1, ax2) = plt.subplots(1, 2, figsize=size)
+    
+    lodf = df[df.resolution == 'lo']
+    plot0 = sn.violinplot(x='model_type', y='surfdist_avg', hue='surf_id',
+                          hue_order=['internal', 'external'],
+                          data=lodf, scale_hue=.9, width=.9, ax=ax1)
+    plot0.set_xlabel('Model Type')
+    plot0.set_ylabel('Averaged error of surfaces (mm)')
+    plot0.set_ylim([0.0, 2.25])
+    plot0.set_title(
+        r'Registration error @ $%.1f \times %.1f \times %.1fmm^3$' %
+        tuple([2.0] * 3))
+
+    hidf = df[df.resolution == 'hi']
+    plot1 = sn.violinplot(x='model_type', y='surfdist_avg', hue='surf_id',
+                          hue_order=['internal', 'external'],
+                          data=hidf, scale_hue=.9, width=.9, ax=ax2)
+    plot1.set_xlabel('Model Type')
+    plot1.set_ylabel('')
+    plot1.set_ylim([0.0, 1.15])
+    plot1.set_title(
[…]

### 2015-03-16 Scripts/pyacwereg/viz.py
update violinplot in realdata

@@ -2,8 +2,8 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2014-12-11 15:08:23
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-03-16 21:20:04
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-03-16 21:45:50
 import os.path as op
 
 
@@ -609,10 +609,11 @@ def realdata_errors(in_csv, size=(80, 25), out_file=None,
     df2 = _extract_method(t2bdf, 'T2B', columns)
 
     plot0 = sn.violinplot(x='surf', y='error', hue='method', size=size,
+                          bw=0.2, inner='quartile', linewidth=8,
                           data=pd.concat([df1, df2]), scale_hue=.9, width=.4)
 
     ymax = plot0.get_ylim()[1]
-    plot0.set_ylim([0.0, ymax])
+    plot0.set_ylim([0.0, 4.0])
 
     plot0.set_xticklabels([r'$\Gamma_{VdGM}$', r'$\Gamma_{WM}$',
                            r'$\Gamma_{pial}$', 'Aggregated'])

### 2015-03-16 Publications/2015-NeuroImage/05-results.tex
introducing final results

@@
   a standard \gls*{t2b} correction.
-For both \emph{regseg} and the \gls*{t2b} methods, we computed the \gls*{swindex} \eqref{eq:swindex}
-  of every surface.
+We computed the \gls*{swindex} \eqref{eq:swindex} of every surface after correction,
+  for both \emph{regseg} and the \gls*{t2b} methods.
 Finally, to statistically compare the results, we performed a one-way ANOVA test
-  on the warping indices for three surfaces.
-Following the nomenclature introduced in section \autoref{sec:human_connectome}, we selected
-  $\Gamma_{VdGM}$, $\Gamma_{WM}$, and $\Gamma_{pial}$ as surfaces of interest.
-These interfaces are critical in connectivity analyses, and small errors in their location may
-  lead to discarding fundamental paths between regions.
-We also reported the 95\% CIs of the warping indices for those surfaces.
+  on the warping indices for three specific surfaces of interest
+  ($\Gamma_{VdGM}$, $\Gamma_{WM}$, and $\Gamma_{pial}$).
+These interfaces are critical in analyses on whole-brain tractography.
+The one-way ANOVA evidenced that error distributions were significantly different,
+  $p \approx 2 \times 10^{-9}$.
+We also reported the 95\% CIs of the \gls*{swindex} for those surfaces.
+The aggregated CI for \emph{regseg} was $[1.08 - 1.50] mm$, whereas the competing method
+  yielded an aggregated CI of $[2.06 - 2.43] mm$.
 The quantitative results of the ANOVA tests and the CIs are presented in \autoref{tab:results_real}.
@@
   box plots of error distributions by correction method and surface.
-Extended visual assessment of the 16 cases is aggregated in the \suppl{section S4}.
+Extended visual assessment of the 16 cases is included in the \suppl{section S4}.
 
@@
 		\tabcolsep=0.05cm
-    \begin{tabular}{L{1.8cm}ccc}
-                            & $\Gamma_{VdGM}$  & $\Gamma_{WM}$ & $\Gamma_{pial}$ \\
+    \begin{tabular}{cccccc}
+    & & $\Gamma_{VdGM}$  & $\Gamma_{WM}$ & $\Gamma_{pial}$ & Aggregated \\
     \hline
-    CI \emph{regseg}        & 0.39 - 0.64 & 0.64 - 0.83 & 0.76 - 0.93 \\
-    CI competing            & 2.36 - 3.37 & 2.65 - 3.19 & 2.82 - 3.40 \\
+    \multirow{2}{*}{CI}
+       & \emph{regseg}        & 0.77 - 1.18 & 0.95 - 1.85 & 1.33 - 2.13 & 1.08 - 1.50 \\
+       & T2B                  & 1.73 - 2.62 & 1.96 - 2.40 & 2.16 - 2.63 & 2.06 - 2.43 \\
     \hline
-    ANOVA (f-stat/p-value)  & (51.81/1.5e-06) & (148.25/8.1e-10) & (143.01/1.1e-09) \\
+    \multirow{2}{*}{ANOVA}
+       & p-value  & 1.4$\times$10$^{-5}$& 1.3$\times$10$^{-3}$& 2.5$\times$10$^{-3}$ & 2$\times$10$^{-9}$ \\
+       & f-stat   & 27.87               & 12.89               & 11.13                & 45.04 \\
     \hline

### 2015-03-16 Publications/2015-NeuroImage/06-discussion.tex
introducing final results

@@
 We processed 16 subjects from the \gls*{hcp} database with both \emph{regseg}
-  and an in-house replication of a competing method.
-This alternative approach solves distortion using \emph{b0}-to-\gls*{t2} registration.
+  and an in-house replication of the \gls*{t2b} correction method.
 Our method presented significantly different results (ANOVA tests are reported in
@@
   \emph{regseg}.
-The aggregated CI for \emph{regseg} was $[0.39 - 0.93] mm$, whereas the competing method
-  yielded an aggregated CI of $[2.36 - 3.40] mm$.
 Even though we carefully configured the competing method by using the same algorithm and

### 2015-03-17 Scripts/pyacwereg/viz.py
update voxel resolution in violinplot

@@ -2,8 +2,8 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2014-12-11 15:08:23
-# @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-03-16 21:45:50
+# @Last Modified by:   oesteban
+# @Last Modified time: 2015-03-17 00:40:27
 import os.path as op
 
 
@@ -619,10 +619,10 @@ def realdata_errors(in_csv, size=(80, 25), out_file=None,
                            r'$\Gamma_{pial}$', 'Aggregated'])
     plot0.set_xlabel('Surface')
     plot0.set_ylabel('Surface warping index ($sWI$, mm)')
-    l = plot0.axhline(y=2.5, lw=15, xmin=0.07, xmax=0.93,
+    l = plot0.axhline(y=1.25, lw=15, xmin=0.07, xmax=0.93,
                       color='gray', alpha=.4)
     plot0.annotate(
-        "voxel size", xy=(-0.15, 2.5), xytext=(-250, 105),
+        "voxel size", xy=(-0.15, 1.25), xytext=(-250, 105),
         xycoords='data', textcoords='offset points', va='center',
         color='w', fontsize=80,
         bbox=dict(boxstyle='round', fc='gray', ec='none', color='w'),

### 2015-03-18 Publications/2015-NeuroImage/visual_report.sh
automatically generate supplemental figures of real data

@@ -1,0 +1,13 @@
+#!/bin/bash
+
+rm -f suppl-realdata.tex
+
+for i in figures/HCP*.pdf; do
+    fname=$(basename $i)
+    fname=${fname::-4}
+    echo "\\begin{figure}[!ht]" >> suppl-realdata.tex
+    echo "  \\centering" >> suppl-realdata.tex
+    echo "  \\textbf{$fname}" >> suppl-realdata.tex
+    echo "  \\includegraphics[width=.9\textwidth]{$i}" >> suppl-realdata.tex
+    echo "\\end{figure}" >> suppl-realdata.tex
+done

### 2015-03-18 Publications/2015-NeuroImage/supplemental.tex
one more iteration to second round
automatically generate supplemental figures of real data
update supplementals

@@
 % @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-02-27 15:21:03
+% @Last Modified time: 2015-03-18 15:08:42
 
@@
 \usepackage[colorinlistoftodos]{todonotes}
-\usepackage[marginparwidth=1.2in]{geometry}
+\usepackage[top=2cm,bottom=1.8cm,left=2.1cm,right=1.7cm,marginparwidth=1.2in]{geometry}
 \usepackage{changepage}
@@
 \usepackage[toc,nomain,acronym,shortcuts,translate=false]{glossaries}
-\usepackage[hyphens]{url}
-\usepackage[breaklinks,hidelinks]{hyperref}
-\usepackage[hyphenbreaks]{breakurl} %fixes boxes spanning through pages
+
+\usepackage{mathtools}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{cancel}
+%\usepackage{algorithmic}
+\DeclareMathOperator{\tr}{tr}
+\DeclareMathOperator{\argmax}{argmax}
+\DeclareMathOperator{\argmin}{argmin}
+\DeclareMathOperator{\diag}{diag}
+\DeclareMathOperator{\dist}{dist}
+\DeclareMathOperator{\const}{Const.}
+\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
+\providecommand{\mdist}[2]{ \mathcal{D}_{#2}^2(\mathbf{#1}) }
+\let\oldhat\hat
+\renewcommand{\vec}[1]{\mathbf{#1}}
+\providecommand{\nvec}[1]{\hat{\mathbf{#1}}}
 
@@
 \renewcommand\thesubsection{S\arabic{section}.\arabic{subsection}}
+\renewcommand\thefigure{S\arabic{figure}}
+\renewcommand{\theequation}{SM\arabic{equation}}
 
@@
 
+\usepackage[hyphens]{url}
+\usepackage[hyphenbreaks]{breakurl} %fixes boxes spanning through pages
+\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}
+
+\input{01-acronyms}
+
 \begin{document}
@@
 \maketitle
+\section{Extensions to the mathematical formulation of the methods}
+
+\paragraph{Discretization of the shape gradients} The energy functional to be optimized
+  in \emph{regseg} is presented in Eq. 7:
+  \begin{equation}
+  E(R \mid U) = C + \underbracket{\underset{l}{\sum} \int_{\Omega_l}
+  \mdist{f'}{l} \,d\vec{r}}_{\text{Data term}}
+  + \underbracket{\underset{\Omega}{\int} \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
+  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}}_{\text{Regularization term}}.
+  \label{eq:energy}
+  \end{equation}
+
+To search for the minimum of $E(R \mid U)$ w.r.t. the coefficients $\vec{u}_k$, we use a gradient descent strategy.
+In Eq. 11 we introduced the derivative of \eqref{eq:energy}:
+  \begin{equation}
+  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} =
+  \frac{ \partial }{\partial \vec{u}_k} \Big\{
+  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r}
+  + \int_{\Omega} [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
+  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
+  \Big\}.
+  \label{eq:gradient_descent}
+  \end{equation}
+
+Let us compute the derivative of the data term first (Eq. 12):
+\begin{equation}
+  \frac{\partial E_{data}(\vec{u})}{\partial \vec{u}_k} =
+  \frac{ \partial }{\partial \vec{u}_k} \Big\{
+  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \Big\} =
+  \underset{l}{\sum} \, \frac{ \partial }{\partial \vec{u}_k} \Big\{
+  \int_{\Omega_l} \mdist{f'}{l} \Big\},
+  \label{eq:data_derivative}
+\end{equation}
+where we can introduce the shape-gradients theory from Appendix 1.
+The equation A.1 is discretized as follows.
+First, the surface between limiting regions $\{l, m\}$, $\Gamma_{l,m}$ is explicitly represented by
+  a discrete set of vertices $\vec{v}_i$, with $i \in \{0, \ldots, N_p -1 \}$.
+Complementarily, the inwards normal of the surface $N_{\Gamma_{l,m}}$ is represented by the discrete
+  set of normals $\hat{\vec{n}}_i$ at each vertex of the mesh.
+The resulting summation is, therefore, discrete and the integral operator is replaced by the sum.
+  \begin{align}
+  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) d\vec{r} &=
+  \underbracket{\cancel{\int_\Omega \frac{\partial}{\partial t}F(\vec{r}) d\vec{r} }}_{\text{Functional's evolution}}
+  - \underbracket{\int_{\Gamma_{l,m}} F(\vec{r}) \left\langle \frac{\partial \Gamma_{l,m}}{\partial t},
+  N_{\Gamma_{l,m}}\right\rangle d\vec{r}}_{\text{Shape's evolution}} \notag \\
+  & = - \underset{p}{\sum} \frac{1}{A_p} \underset{i}{\sum} \, a_i \, F(\vec{v}_i) \left\langle \underbracket{\frac{\partial \vec{v}_i}{\partial t}}_{\text{speed of }\vec{v}_i},
+  \hat{\vec{n}}_{i}\right\rangle.
+  \label{eq:shape_gradient_orig}
+  \end{align}
+where $a_i$ is the area corresponding to vertex $\vec{v}_i$, and $A_p = \sum_i a_i$ is the total area of surface $p$.
+In the following, we will refer as $w_{p,i} = a_i / A_p $ to the area contribution of $\vec{v}_i$ to the
+  total area of the surface it belongs to.
+For simplicity, the sum over $p$ can be also removed, as the vertices belong to only one of the total $P$ contours
+
+Then, the speed of $\vec{v}_i$ is discretized using the artificial timestep parameter $\delta$, as the displacement
+  $\vec{d}_i = \vec{v}_i(\delta = t+1) - \vec{v}_i(\delta = t)$:
+  \begin{equation}
+  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) \, d\vec{r} =
+  - \underset{i}{\sum} w_{p,i} F(\vec{v}_i) \vec{d}_i \cdot \hat{\vec{n}}_i.
+  \label{eq:shape_gradient_disc1}
+  \end{equation}
+
+Since the energy functional is defined over competing regions, the displacement of $\vec{v}_i$ will cause
+  an energy exchange between the limiting regions, and therefore $F(\vec{r})$ must be splitted in
+  two terms, $F_{in}(\vec{r})$ corresponding to the interior region and $F_{out}(\vec{r})$ to the exterior:
+  \begin{equation}
+  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) \, d\vec{r} =
+  - \underset{i}{\sum} w_{p,i} \, \vec{d}_i \cdot
+  \underbracket{\Big[ \big(F_{out}(\vec{v}_i) - F_{in}(\vec{v}_i) \big) \hat{\vec{n}}_i \Big]}_{\bar{s}_i \text{ in Figure 1}}.
+  \label{eq:shape_gradient_disc2}
+  \end{equation}
+
+Now we can identify $t$ with the set of time-dependent coefficients, $\{\vec{u}_k\}$, and the target function $F$ with $\mathcal{D}_l^2$.
+To introduce the speed of the vertices from equation 13, we replace $\vec{d}_i$:
+\begin{equation}
+  \frac{\partial E_{data}(\vec{u})}{\partial \vec{u}_k} =
+  - \underset{i}{\sum} w_{p,i} \, \psi_k(\vec{v}_i) \hat{\vec{e}} \cdot \bar{s}_i =
+  - \underset{i}{\sum} w_{p,i} \, \psi_k(\vec{v}_i) \bar{s}_i = - \underset{i}{\sum} \vec{g}_{i,k}.
+  \label{eq:shape_gradient_final}
+\end{equation}
+
+Inserting \eqref{eq:shape_gradient_final} in \eqref{eq:gradient_descent} we reach equation 15,
+  and renaming $\vec{g}_k = - \underset{i}{\sum} \vec{g}_{i,k}$:
+  \begin{equation}
+  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} =
+  \vec{g}_k + \boldsymbol{\alpha} \vec{u}_k
+  -\boldsymbol{\beta} \Delta \vec{u}_k.
+  \label{eq:gradient_final}
+  \end{equation}
+
+\paragraph{Obtaining the update equation} Once the shape gradients are obtained \eqref{eq:gradient_final},
+  we apply the Fourier Transform to solve the \gls*{pde}:
+\begin{align}
+  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} \equiv \frac{\vec{u}_k^{t+1} - \vec{u}_k^t}{\delta} \implies
+  \vec{u}_k^{t+1} &= \delta \Big( \vec{g}_k + \boldsymbol{\alpha} \vec{u}_k
+  - \boldsymbol{\beta} \Delta \vec{u}_k \Big) + \vec{u}_k^t, \notag \\
+  \vec{u}_k^{t+1} &= \delta \vec{g}_k +  \boldsymbol{\alpha} \delta \vec{u}_k
+  - \boldsymbol{\beta} \delta \Delta \vec{u}_k + \vec{u}_k^t
+  \label{eq:discrete_derivative}
+\end{align}
 
@@
 \paragraph{Efficient interpolation using sparse matrices}
-During the registration process, every iteration requires computing the product of all the speeds
-  $\mathbf{s}'_i$ computed at the current position $\mathbf{v}'_i$ by the corresponding weights
-  $\psi_k(\mathbf{v}_i)$ of interpolating functions (Eq. (15)).
+During the registration process, every iteration requires computing the product of all the gradients
+  $\vec{g}_k$ associated to the control point $k$, and computed at the current position
+  $\vec{v}_i$ by the corresponding weights $\psi_{ik} = \psi_k(\mathbf{v}_i)$ of interpolating functions
+  \eqref{eq:shape_gradient_final}.
+In order to optimize multiplications and summations, all the $\psi_{ik}$ are collected in a
+  matrix $\boldsymbol{\Psi} = (\psi_{ik})$.
+Given the limited support of the basis function $\psi$, $\boldsymbol{\Psi}$ will hold the property
+  of being sparse, as only few $\psi_{ik} > 0$ in the surroundings of $\vec{v}_i$.
+Then, the gradients $g_k$ are easily computed using the matrix product:
+\begin{equation}
+  \big(\vec{g}_k \big) = \boldsymbol{\Psi} \cdot \big( w_i \, \bar{s}_i \big)^T
+  \label{eq:sparse_matrix}
+\end{equation}
+
 As these weights can be computed once in the beginning of the process and they do not change along
-  it, and they are different from zero only in the proximity of the corresponding control node $k$,
-  they can be pre-cached in a sparse matrix.
+  it, $\boldsymbol{\Psi}$ can be pre-cached.
 
@@
 \subsection{Convergence evidencing}\label{sec:convergence_evidence}
+
 In order to track the evolution of the registration process, several internal variables
@@
   a visual assessment document presenting the convergence.
-Online checking is also possible as the algorithm writes to the standard output as well.
-
 \begin{pythoncode}
@@
 
-\begin{figure*}[!ht]
+Online checking is also possible as the algorithm writes to the standard output as well.
+A sample report is found in \autoref{fig:convreport}.
+
+\begin{figure*}[b]
   \includegraphics[width=\textwidth]{figures/Suppl-figure02.pdf}
@@
 
-\section{Extended results on phantom data}
+\newpage
+\section{Extended results on real data}
 
-\section{Extended results on real data}
+\immediate\write18{./visual_report.sh}
+\input{suppl-realdata.tex}
 

### 2015-03-18 Publications/2015-NeuroImage/supplemental.tex
fix typos

@@
 % @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-03-18 15:08:42
+% @Last Modified time: 2015-03-18 15:12:35
 
@@
   a discrete set of vertices $\vec{v}_i$, with $i \in \{0, \ldots, N_p -1 \}$.
-Complementarily, the inwards normal of the surface $N_{\Gamma_{l,m}}$ is represented by the discrete
+Consequently, the inwards normal of the surface $N_{\Gamma_{l,m}}$ is represented by the discrete
   set of normals $\hat{\vec{n}}_i$ at each vertex of the mesh.
@@
 
-Then, the speed of $\vec{v}_i$ is discretized using the artificial timestep parameter $\delta$, as the displacement
+Then, the speed of $\vec{v}_i$ is discretized using the artificial time-step parameter $\delta$, as the displacement
   $\vec{d}_i = \vec{v}_i(\delta = t+1) - \vec{v}_i(\delta = t)$:
@@
 Since the energy functional is defined over competing regions, the displacement of $\vec{v}_i$ will cause
-  an energy exchange between the limiting regions, and therefore $F(\vec{r})$ must be splitted in
+  an energy exchange between the limiting regions, and therefore $F(\vec{r})$ must be split in
   two terms, $F_{in}(\vec{r})$ corresponding to the interior region and $F_{out}(\vec{r})$ to the exterior:

### 2015-03-18 Scripts/pyacwereg/workflows/registration.py
update step_size in real data

@@ -5,8 +5,8 @@
 #
 # @Author: oesteban - code@oscaresteban.es
 # @Date:   2014-03-28 20:38:30
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-03-10 13:29:28
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-03-18 17:03:44
 
 import os
 import os.path as op
@@ -29,7 +29,7 @@ from pyacwereg.interfaces.warps import FieldBasedWarp, InverseField
 def regseg_wf(name='REGSEG', enhance_inputs=True, usemask=False):
     wf = pe.Workflow(name=name)
 
-    regseg_inputs = ['iterations', 'alpha', 'beta', 'step_auto',
+    regseg_inputs = ['iterations', 'alpha', 'beta', 'step_size',
                      'grid_spacing', 'convergence_energy',
                      'convergence_window', 'f_smooth',
                      'images_verbosity', 'scales', 'descript_update',

### 2015-03-18 Publications/2015-NeuroImage/04-methods.tex
final fixes
update supplementals
improving to second round

--- 2015-03-16 Publications/2015-NeuroImage/04-methods.tex
@@
   \frac{ \partial }{\partial \vec{u}_k} \Big\{
-  \int_{\Omega} \underset{l}{\sum} \mdist{f'}{l} \,d\vec{r} \notag\\
+  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \notag\\
   &+ \int_{\Omega} [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
@@
   \citep{herbulot_segmentation_2006}, and ultimately avoid level sets.
-We refer the reader to the appendix \autoref{app:shape_priors} in order to fill in the gap
+We refer the reader to \ref{app:shape_priors} in order to fill in the gap
   between \eqref{eq:gradient_descent} and the following expression:
@@
   \begin{equation}
-  sWI = \frac{1}{P} \sum\limits_p^P \frac{1}{A_p} \sum\limits_i^{N_p} a_i\,\|
+  sWI = \frac{1}{\sum_i a_i} \sum\limits_i^{N_p} a_i\,\|
   \vec{v}_i - \hat{\vec{v}}_i \|
@@
   where $\vec{v}_i$ are the locations of the $N_p$ vertices in each $p \in \{1, \dots P\}$
-  prior, $A_p$ the total area of surface $p$, and $\hat{\vec{v}}_i$ is the location
+  prior, and $\hat{\vec{v}}_i$ is the location
   recovered corresponding to the vertex $\vec{v}_i$.
@@
   using \texttt{mri\_tessellate}\footnote{A marching-cubes algorithm shipped with 
-  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}}.
+  \emph{FreeSurfer}.}.
 
 \begin{figure*}
-	\includegraphics[width=\linewidth]{figures/figure03}
+    \centering
+    \resizebox{\textwidth}{!}{%
+      \begin{tikzpicture}
+        \node[inner sep=10pt, draw=black!40](fig03a) at (0,14.8)
+          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure03-a}};
+        \node[inner sep=10pt, draw=black!40](fig03b) at (0,7.9)
+          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure03-b}};
+        \node[inner sep=10pt, draw=black!40](fig03c) at (0,0)
+          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure03-c}};
+        \node[circle, text=black!75] at (-9.4,17.3) {\Huge \textbf{A}};
+        \node[circle, text=black!75] at (-9.4,10.6) {\Huge \textbf{B}};
+        \node[circle, text=black!75] at (-9.4,3.5) {\Huge \textbf{C}};
+      \end{tikzpicture}
+    }%
 	\caption{A. The ``cortex'' phantom is a spherical shape with two sulci and an

### 2015-03-18 Publications/2015-NeuroImage/05-results.tex
final fixes
improving to second round

@@
 The results showed a consistent and high accuracy, below the image resolution.
-\autoref{fig:phantom} (block C) presents the box plots by model type corresponding
+\autoref{fig:phantom} (block C) presents the violin plots by model type corresponding
   to the two sets of resolutions of generated phantoms.
@@
 		\footnotesize
-    \begin{tabular}{llcc}
-    Type      & Low-res.  & Hi-res. \\
-    \hline
-    ``gyrus'' & 0.58-0.60 & 0.18-0.36 \\
-    ``L''     & 0.67-0.73 & 0.35-0.39 \\
-    ``ball''  & 0.66-0.73 & 0.32-0.41 \\
-    ``box''   & 0.68-0.71 & 0.35-0.39 \\
+    \begin{tabular}{lccccc}
+    Res. & ``gyrus'' & ``ball''  & ``box''   & ``L''     & Aggreg.    \\\hline
+    Low  & 0.59-0.60 & 0.65-0.76 & 0.68-0.71 & 0.67-0.77 & 0.64-0.66  \\
+    High & 0.18-0.38 & 0.31-0.45 & 0.34-0.42 & 0.34-0.40 & 0.34-0.38  \\
     \hline
@@
   (\gls*{fa} and \gls*{adc}).
-Based on the observed joint distribution of each (\suppl{section S2}),
+Based on the observed joint distribution of each (\suppl{section S4}),
   we selected the 6-regions scheme described in \autoref{sec:human_connectome}.
@@
 %
-Finally, we tested the performance of our registration method, and compared it to the
-  a standard \gls*{t2b} correction.
+Finally, we compared the performance of \emph{regseg} and the standard \gls*{t2b}
+  correction.
 We computed the \gls*{swindex} \eqref{eq:swindex} of every surface after correction,
@@
   ($\Gamma_{VdGM}$, $\Gamma_{WM}$, and $\Gamma_{pial}$).
-These interfaces are critical in analyses on whole-brain tractography.
+These interfaces are critical in analyses of whole-brain tractography.
 The one-way ANOVA evidenced that error distributions were significantly different,
@@
 Visual assessment of one case is shown in \autoref{fig:results_real}, along with the corresponding
-  box plots of error distributions by correction method and surface.
-Extended visual assessment of the 16 cases is included in the \suppl{section S4}.
+  violin plots of error distributions by correction method and surface.
+Extended visual assessment of the 16 cases is included in the \suppl{section S5}.
 

### 2015-03-18 Publications/2015-NeuroImage/06-discussion.tex
improving to second round

@@
 
-The hypothesis underpinning our research is that image registration can be reliably performed
+\paragraph*{Accuracy tests}
+The hypothesis underpinning our research was that image registration can be reliably performed
   by searching for homogeneous regions in the target image that correspond to precise contours
   from an atlas, or extracted from other image (i.e. a different time step).
+We proved that active contours without edges can be used to successfully drive a
+  deformation supported by B-Spline functions using digital phantoms.
+We randomly deformed 4 different phantom models mimicking three homogeneous regions,
+  showing contrasts generated by a publicly available \gls*{mri} simulator.
+After registration with \emph{regseg} we measured the Hausdorff distance between the
+  projected contours using the ground-truth warping and our estimation.
+We concluded that errors are significantly lower than voxel resolution.
+We also assessed the 95\% \gls*{ci}, yielding an aggregated interval of
+  $[0.64 - 0.66]mm$ in the low resolution phantoms ($2.0mm$ isotropic voxel) and
+  $[0.34 - 0.38]mm$ in the high resolution phantoms ($1.0mm$ isotropic).
+Therefore, we also conclude that the error is bounded and below the half of the
+  voxel spacing to illustrate the accuracy of \emph{regseg}.
+
+\paragraph*{Application on real data}
 We designed \emph{regseg} as a method to correct \gls*{dmri} data for susceptibility
@@
   linear registration of \gls*{t1} and \gls*{dmri} to fuse the anatomical information on
-  the diffusion data, and distortions correction using extra acquisitions.
+  the diffusion data, and correction of distortion using extra acquisitions.
 Moreover, our method does not require any additional acquisition over the minimal protocol
@@
 
-\paragraph*{Accuracy tests}
-We proved that active contours without edges can be used to successfully drive a
-  deformation supported by B-Spline functions in digital phantoms.
-We randomly deformed 4 different phantom models mimicking three homogeneous regions,
-  showing contrasts generated by a publicly available \gls*{mri} simulator.
-After registration with \emph{regseg} we measured the Hausdorff distance between the
-  projected contours using the ground-truth warping and our estimation.
-We concluded that errors are significantly lower than voxel resolution.
-We also assessed the 95\% intervals, yielding an aggregated interval of distances of
-  $[0.58 - 0.73]$ in the low resolution phantoms ($2.0mm$ isotropic voxel) and
-  $[0.18 - 0.41]$ in the high resolution phantoms ($1.0mm$ isotropic).
-These aggregated intervals by resolution take the lowest of the lower boundary
-  and the highest of the upper boundary of all CIs regardless the phantom type.
-Therefore, we also conclude that the error is bounded and below the half of the
-  voxel spacing to illustrate the accuracy of \emph{regseg}.
-
-\paragraph*{Application on real data}
 After a preliminary experiment to demonstrate the aptness of our segmentation model
@@
   and an in-house replication of the \gls*{t2b} correction method.
-Our method presented significantly different results (ANOVA tests are reported in
-  \autoref{tab:results_real}), and the 95\% CIs revealed the high accuracy of
-  \emph{regseg}.
-Even though we carefully configured the competing method by using the same algorithm and
-  extending the same settings from a widely used software, cross-comparison
-  experiments are prone to the so-called instrumentation bias \citep{tustison_instrumentation_2013}.
+\emph{Regseg} showed very high accuracy, with an aggregated 95\% \gls*{ci} of $[1.08-1.50]mm$,
+  almost below the pixel size of $1.25mm$.
+The misregistration error remaining after \emph{regseg} was significantly lower ($p < 0.01$) than the
+  error corresponding to the \gls*{t2b} correction, using the one-way ANOVA test
+  (\autoref{tab:results_real}).
+Visual inspection of all the results (\suppl{section S5}), and the violin plots included in
+  \autoref{fig:results_real}, confirmed this intuition.
+Even though we carefully configured the \gls*{t2b} method by using the same algorithm and
+  same settings from a widely used software, cross-comparison experiments are prone to
+  the so-called \emph{instrumentation bias} \citep{tustison_instrumentation_2013}.
 Therefore, these results do not show that the \emph{regseg} software \emph{is} better than the
-  competing method, but they are communicated to proof that \emph{regseg} is a solid option to
+  \gls*{t2b} method, but they are communicated to proof that \emph{regseg} is a solid option to
   correct \gls*{dmri} data for distortions, and any other potential application.
@@
 \paragraph*{Prospects}
-Future extensions or improvements of \emph{regseg} are similar to the evolution of existing
-  alternative methodologies.
+Future improvements of \emph{regseg} possibly will follow a similar line to the evolution of 
+  existing methodologies.
 Incorporating knowledge about the distortion by initializing our method with the theory-based
-  correction using fieldmaps would be the most profitable.
+  correction using fieldmaps may be the most cost-effective.
 The main pitfall using fieldmap correction is the challenging problem of unwrapping the
@@
   applied to clean up field maps.
-This integration has been recently applied in the \emph{b0}-to-\gls*{t2} approach
+This integration has been recently applied in the \gls*{t2b} approach
   by \cite{irfanoglu_susceptibility_2011}, also investigating the usefulness of

### 2015-03-18 Scripts/pyacwereg/viz.py
update violinplots of phantom data

@@ -2,8 +2,8 @@
 # -*- coding: utf-8 -*-
 # @Author: oesteban
 # @Date:   2014-12-11 15:08:23
-# @Last Modified by:   oesteban
-# @Last Modified time: 2015-03-17 11:59:18
+# @Last Modified by:   Oscar Esteban
+# @Last Modified time: 2015-03-18 19:01:24
 import os.path as op
 
 
@@ -551,15 +551,17 @@ def phantom_errors(in_csv, size=(80, 30), out_file=None):
     sn.set_context("poster", font_scale=4.5)
     df = pd.read_csv(in_csv).drop_duplicates(subset=['surf_id', 'repetition',
                                              'surfdist_avg', 'model_type'])
+    del df['Unnamed: 0']
+
     df.surf_id[df.surf_id == 0] = 'internal'
     df.surf_id[df.surf_id == 1] = 'external'
     mtypes = df.model_type.unique()
     cols = len(mtypes) * 2
     fig, (ax1, ax2) = plt.subplots(1, 2, figsize=size)
-    
+
     lodf = df[df.resolution == 'lo']
     plot0 = sn.violinplot(x='model_type', y='surfdist_avg', hue='surf_id',
-                          hue_order=['internal', 'external'],
+                          hue_order=['internal', 'external'], inner='quartile',
                           data=lodf, scale_hue=.9, width=.9, ax=ax1)
     plot0.set_xlabel('Model Type')
     plot0.set_ylabel('Averaged error of surfaces (mm)')
@@ -568,9 +570,24 @@ def phantom_errors(in_csv, size=(80, 30), out_file=None):
         r'Registration error @ $%.1f \times %.1f \times %.1fmm^3$' %
         tuple([2.0] * 3))
 
+    l = plot0.axhline(y=2.0, lw=15, xmin=0.07, xmax=0.93,
+                      color='gray', alpha=.4)
+    plot0.annotate(
+        "voxel size", xy=(2.5, 2.0), xytext=(100, -150),
+        xycoords='data', textcoords='offset points', va='center',
+        color='w', fontsize=80,
+        bbox=dict(boxstyle='round', fc='gray', ec='none', color='w'),
+        arrowprops=dict(arrowstyle='wedge,tail_width=.7',
+                        fc='gray', ec='none',
+                        ))
+    
+    frame = plot0.legend(loc=2, fancybox=True).get_frame()
+    frame.set_facecolor('white')
+    frame.set_edgecolor('white')
+
[…]

### 2015-03-18 Scripts/pyacwereg/viz.py
fix pep8 errors

@@ -3,7 +3,7 @@
 # @Author: oesteban
 # @Date:   2014-12-11 15:08:23
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-03-18 19:01:24
+# @Last Modified time: 2015-03-18 20:00:10
 import os.path as op
 
 
@@ -549,8 +549,8 @@ def phantom_errors(in_csv, size=(80, 30), out_file=None):
     import seaborn as sn
     import matplotlib.pyplot as plt
     sn.set_context("poster", font_scale=4.5)
-    df = pd.read_csv(in_csv).drop_duplicates(subset=['surf_id', 'repetition',
-                                             'surfdist_avg', 'model_type'])
+    df = pd.read_csv(in_csv).drop_duplicates(
+        subset=['surf_id', 'repetition', 'surfdist_avg', 'model_type'])
     del df['Unnamed: 0']
 
     df.surf_id[df.surf_id == 0] = 'internal'
@@ -580,7 +580,7 @@ def phantom_errors(in_csv, size=(80, 30), out_file=None):
         arrowprops=dict(arrowstyle='wedge,tail_width=.7',
                         fc='gray', ec='none',
                         ))
-    
+
     frame = plot0.legend(loc=2, fancybox=True).get_frame()
     frame.set_facecolor('white')
     frame.set_edgecolor('white')
@@ -697,7 +697,6 @@ def phantom_boxplot(in_csv, resolution='lo',
 
     sn.set(style="whitegrid")
     sn.set_context("poster", font_scale=1.5)
-
 
     g = sn.factorplot('model_type', 'surfdist_avg', 'surf_id', filtdf,
                       size=15, kind='box', x_order=mtypes, legend=False,

### 2015-03-20 Publications/2015-NeuroImage/00-main.tex
update figure, updates in manuscript
final fixes
update supplementals
editions after @daducci review

@@
 
-\title{Active-contours driven registration of diffusion MR images of the brain}
+\title{Active-contours driven registration applied in the correction of
+susceptibility-derived distortions of diffusion MR images}
 
@@
 \author[ucla]{Dominique~Zosso}
-\author[lts5]{Alessandro~Daducci}
+\author[scil,lts5]{Alessandro~Daducci}
 \author[chuv,lts5]{Meritxell~Bach-Cuadra}
@@
 Los Angeles (UCLA), Los Angeles, CA, US}
+\address[scil]{Computer Science Department, Faculty of Science, Universit\'e de Sherbrooke, 2500 Boulevard Universit\'e, Sherbrooke, QC J1K 2R1, Canada}
 \address[lts5]{Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
@@
 diffusion MRI \sep susceptibility~distortion \sep segmentation %
-\sep registration \sep parcellation \sep shape-prior.
+\sep nonlinear~registration \sep active~contours \sep echo-planar imaging.
 \end{keyword}
@@
 %\bibliographystyle{plainnat}
+% \section*{References}
 \bibliographystyle{mystyle}

### 2015-03-20 Publications/2015-NeuroImage/02-abstract.tex
introducing final results
editions after @daducci review

@@
 The results suggest that \emph{regseg} overperformed an extended competing method.
-\emph{Regseg} showed a consistent error within the 95\% confidence interval of
-  $0.39-0.93mm$ in images of $1.25mm$ isotropic resolution.
+We compute the geometrical disagreement of the surfaces registered with \emph{regseg} with
+  respect to a ground truth, obtaining a 95\% confidence interval of $[1.08 - 1.50]mm$ in images 
+  of $1.25mm$ isotropic resolution.

### 2015-03-20 Scripts/pyacwereg/viz.py
update graphs

@@ -3,7 +3,7 @@
 # @Author: oesteban
 # @Date:   2014-12-11 15:08:23
 # @Last Modified by:   Oscar Esteban
-# @Last Modified time: 2015-03-18 20:00:10
+# @Last Modified time: 2015-03-20 13:18:39
 import os.path as op
 
 
@@ -650,8 +650,8 @@ def realdata_errors(in_csv, size=(80, 25), out_file=None,
     df2 = _extract_method(t2bdf, 'T2B', columns)
 
     plot0 = sn.violinplot(x='surf', y='error', hue='method', size=size,
-                          bw=0.2, inner='quartile', linewidth=8,
-                          data=pd.concat([df1, df2]), scale_hue=.9, width=.4)
+                          inner='quartile', linewidth=8,  # bw=0.2,
+                          data=pd.concat([df1, df2]), scale_hue=.9, width=.7)
 
     ymax = plot0.get_ylim()[1]
     plot0.set_ylim([0.0, 4.0])

### 2015-03-20 Publications/2015-NeuroImage/02-abstract.tex
fix aggregate, update results

--- 2015-03-16 Publications/2015-NeuroImage/02-abstract.tex
@@
 The results suggest that \emph{regseg} overperformed an extended competing method.
-\emph{Regseg} showed a consistent error within the 95\% confidence interval of
-  $0.39-0.93mm$ in images of $1.25mm$ isotropic resolution.
+We compute the geometrical disagreement of the surfaces registered with \emph{regseg} with
+  respect to a ground truth, obtaining a 95\% confidence interval of 0.56 - 0.66 (mm)
+  for the error in images of 1.25mm isotropic resolution.

### 2015-03-20 Publications/2015-NeuroImage/06-discussion.tex
fix aggregate, update results

@@
 We concluded that errors are significantly lower than voxel resolution.
-We also assessed the 95\% \gls*{ci}, yielding an aggregated interval of
+We also assessed the 95\% \gls*{ci}, yielding an aggregate interval of
   $[0.64 - 0.66]mm$ in the low resolution phantoms ($2.0mm$ isotropic voxel) and
@@
   and an in-house replication of the \gls*{t2b} correction method.
-\emph{Regseg} showed very high accuracy, with an aggregated 95\% \gls*{ci} of $[1.08-1.50]mm$,
+\emph{Regseg} showed very high accuracy, with an aggregate 95\% \gls*{ci} of $[1.08-1.50]mm$,
   almost below the pixel size of $1.25mm$.

### 2015-03-24 Publications/2015-NeuroImage/02-abstract.tex
Updates after Yasser and Dominique

--- 2015-03-16 Publications/2015-NeuroImage/02-abstract.tex
@@
   deformations, and conclude that the precision of the segmentation is subvoxel.
-Finally, we demonstrate \emph{regseg} in the correction task on 16 real diffusion MRI
+We demonstrate \emph{regseg} in the correction task on 16 real diffusion MRI
   datasets from the \acrlong*{hcp}.
 The results suggest that \emph{regseg} overperformed an extended competing method.
-\emph{Regseg} showed a consistent error within the 95\% confidence interval of
-  $0.39-0.93mm$ in images of $1.25mm$ isotropic resolution.
+We compute the geometrical disagreement of the surfaces registered with \emph{regseg} with
+  respect to a ground truth, obtaining a 95\% confidence interval of 0.56 - 0.66 (mm)
+  for the Hausdorff distance of mesh vertices in images of 1.25mm isotropic resolution.

### 2015-03-24 Publications/2015-NeuroImage/03-introduction.tex
Updates after Yasser and Dominique

@@
 \label{sec:introduction}
-Image registration is the process to find a spatial mapping that aligns the information available
-  in two different coordinate systems.
-In their comprehensive survey, \cite{sotiras_deformable_2013} illustrated the wide variety
-  of existing methodologies, and classified them by their three principal building blocks:
-  the matching criteria to evaluate the proximity of the solution,
-  the deformation model that may implement theoretical properties of the mapping
-  (i.e. linearity, elasticity, viscosity, etc.), and the optimization method.
-The matching criteria can be defined as a function on the intensities of voxels of the
-  two images, the alignment of spatial features derived from the corresponding objects in
-  the spaces to be registered, or use both intensity and spatial informations.
-In \citep{sotiras_deformable_2013} these methods are referred to as \emph{iconic}, \emph{geometric},
-  and \emph{hybrid}, respectively.
-When the matching criteria is defined over an optimizable partition of one or both reference an
-  moving spaces, we face a hybrid method which simultaneously segments and registers images.
+Image registration and segmentation are crucial in current pipelines for
+  analysis of whole-brain connectivity using \gls*{dmri}, such as the extraction
+  of brain connectivity \citep{craddock_imaging_2013}.
+
 An early integration of segmentation and registration by \cite{bertalmio_morphing_2000} proposed
@@
 
-We propose \emph{regseg}, an active-contours driven registration method.
-Then, we demonstrate its aptness on the correction of the nonlinear distortions typically
-  presented by \gls*{dmri} data of the brain.
-This problem is briefly described in \autoref{sec:distortions}.
-The general overview of our method, along with the contributions with respect 
-  to the current state of art are summarized in \autoref{sec:contributions}.
-In \autoref{sec:methods_map}, we start from a probabilistic framework to finally show its duality
-  with respect to the active-contours route.
-Following sections
-  introduce the specific features of \emph{regseg} (\autoref{sec:numerical_implementation}),
-  state the evaluation experiments (\autoref{sec:experiments_evaluation}),
-  and describe the synthetic and real data involved in this study (\autoref{sec:datasets}).
-In \autoref{sec:results} we describe our findings to support the usefulness of \emph{regseg}
-  in the described application, compared to a well-established tool.
-The interpretation of the results and prospects is addressed in \autoref{sec:discussion}.
+We propose \emph{regseg} as a solution to joint segmentation and registration 
+  problems with a shape priors approach.
+An immediate application of \emph{regseg} is the correction of susceptibility-derived
+  distortions typically present in \gls*{dmri} data and its segmentation
+  after correction.
+Both correction and segmentation are required steps in current investigations of whole-brain
+  tractography, such as the connectome extraction \citep{daducci_connectome_2012}.
+Indeed, the joint segmentation and correction can exploit the anatomy extracted from
+  the \gls*{t1} as strong shape-priors.

### 2015-03-24 Publications/2015-NeuroImage/04-methods.tex
editions after @daducci review
Updates after Yasser and Dominique

@@
   these sequences minimize the bandwidth of the \gls*{pe} direction.
-This low bandwidth originates the artifactual displacement of locations presenting small deviations
-  from the main magnetic field $B_0$.
+This limitation originates a warping of the reconstructed image in regions affected by
+  small deviations from the main magnetic field $B_0$.
 One important source of these deviations ($\Delta B_0$) is the discontinuity of magnetic
@@
   bone of both hemispheres.
-Distortion can be rendered as a registration problem between the undistorted $R$ (reference), 
-  and distorted $M$ (moving) spaces:
+
+Distortion can be rendered as a registration problem where we look for an underlying deformation 
+  field $U_s$, mapping from an undistorted \gls*{t1} image (the reference $R$) to the distorted
+  \gls*{dmri} images (the moving space $M$), such that the structures identified in $R$ optimally
+  align with the corresponding features in $M$:
 
@@
   to the reference domain.
-
 The displacements field $U_s$ along the \gls*{pe} axis can be estimated theoretically
@@
   approach to the 3D problem.
-Therefore, a typical protocol for the extraction of structural connectivity
-  comprehends \gls*{dmri}, \gls*{t1}, and the acquisition determined by the correction
-  method of choice \citep{daducci_connectome_2012}.
 
-\subsection{Our method and contributions}
+\subsection{Contributions}
 \label{sec:contributions}
 %
-\emph{Regseg} can be viewed as a derivation of \emph{bbregister} \citep{greve_accurate_2009}.
-Both methods exploit the anatomically correct and precise surfaces that are readily
-  extracted with \emph{FreeSurfer} \citep{fischl_freesurfer_2012}, mapping them
-  to a target \gls*{dmri} image using active contours.
-We implemented active contours without edges \citep{chan_active_2001}, that allow
-  \emph{regseg} to segment homogeneous regions within the brain, while \emph{bbregister}
-  use active contours with edges to look for intensity boundaries in the target image.
-Another identifying difference is the deformation model:
-  \emph{bbregister} includes an affine transformation and overcomes the distortion
-  problem by excluding typically affected regions from computation.
-Conversely, \emph{regseg} implements a free-deformation field supported with B-Spline
-  functions.
-Our method is close to \citep{guyader_combined_2011} and \citep{gorthi_active_2011}
-  in the use of active contours without edges for registration, however we avoided
-  the use of level sets with the application of shape gradients
-  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
-\cite{gorthi_active_2011} used a gaussian regularizer, and \cite{guyader_combined_2011}
-  proposed a nonlinear elasticity smoother.
-In contrast, \emph{regseg} implements the anisotropic regularization proposed by
-  \citep{nagel_investigation_1986}.
-We also present an ellaborated evaluation instrumentation to create realistic distortions
-  and test algorithms.
-Integrating \emph{regseg} and an in-house implementation of the \gls*{t2b} correction
-  in the workflow, we compared the performance of both solutions.
+While all the correction methods require an extra acquisition, \emph{regseg} is especially
+  indicated in scanning protocols without specific data for correction,
+  such as historical datasets.
+
+Joint segmentation and distortion correction of \gls*{dmri} data is also interesting to
+  improve the overall consistence of complex processing workflows.
+The closest antecedent of \emph{regseg} is \emph{bbregister} \citep{greve_accurate_2009}.
+Both methods use active contours to align surfaces to same-subject \gls*{dmri} images.
+However, \emph{bbregister} is limited to affine transformations, and thus, it is not able to
+  solve the susceptibility-derived distortions, generally nonlinear.
+To enable registration of uncorrected data, \emph{bbregister} excludes from the cost-function
+  those regions frequently warped.
+Conversely, \emph{regseg} implements a free-deformation field, supported with B-Spline
+  functions to estimate deformations.
+A second diverging foundation is the active-contours framework.
+In \emph{regseg}, active contours are implemented without edges \citep{chan_active_2001},
+  enabling the tool to segment homogeneous regions within the brain.
+On the other hand, \emph{bbregister} requires that the target features present intensity
+  boundaries as it implements active contours with edges.
+The tissue interfaces imaged with \gls*{dmri} are affected by partial volume effects that
+  blur boundaries.
+In practice, registration of \gls*{t1} and \gls*{dmri} data with \emph{bbregister}
+  is limited to the linear match of the pial surface of the brain.
+
+Finally, we present an instrumentation framework to evaluate correction and
+  segmentation pipelines using realistic distortions as ground truth.
+With the integration of \emph{regseg} and an in-house implementation of the 
+  \gls*{t2b} correction in the framework, we compare the performance
+  of both solutions.
+
+% Our method is close to \citep{guyader_combined_2011} and \citep{gorthi_active_2011}
+%   in the use of active contours without edges for registration, however we avoided
+%   the use of level sets with the application of shape gradients
+%   \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
+% \cite{gorthi_active_2011} used a gaussian regularizer, and \cite{guyader_combined_2011}
+%   proposed a nonlinear elasticity smoother.
+% In contrast, \emph{regseg} implements the anisotropic regularization proposed by
+%   \citep{nagel_investigation_1986}.
+% We demonstrate its aptness on the correction of the nonlinear distortions typically
+%   presented by \gls*{dmri} data of the brain.
+% Therefore, a typical protocol for the extraction of structural connectivity
+%   comprehends \gls*{dmri}, \gls*{t1}, and the acquisition determined by the correction
+%   method of choice \citep{daducci_connectome_2012}.
 
@@
   \begin{equation}
-  P(U \mid R,\Gamma_{l,m}) = \frac{P(R \mid U,\Gamma_{l,m})\, P(U)}{P(R)},
+  P(U \mid R, \boldsymbol{\Gamma}) = \frac{P(R \mid U,\boldsymbol{\Gamma})\, P(U)}{P(R)},
   \label{eq:bayes_rule}
@@
 %
-  where $P(R \mid U,\Gamma_{l,m})$ is the data-likelihood, and
-  $\Gamma_{l,m}$ are a set of surfaces corresponding to the interfaces
-  between piecewise smooth regions $\Omega_i$, such that
-  $\Gamma_{l,m} = \partial \Omega_l \cap \partial \Omega_m$ is the
-  contour between competing regions $\Omega_l$ and $\Omega_m$.
+  where $P(R \mid U,\boldsymbol{\Gamma})$ is the data-likelihood, and
+  $\boldsymbol{\Gamma} = \{\Gamma_{\text{in},\text{out}}\}$ is a set of surfaces 
+  corresponding to the interfaces between pairs of competing regions $\Omega_\text{in}$
+  and $\Omega_\text{out}$ drawn from the full set of regions $\{\Omega_l \mid l \in \mathbb{N}, l \leq L\}$.
+As $\boldsymbol{\Gamma}$ are mapped by $U$, we simplify
+  $P(U, \boldsymbol{\Gamma}) = P(U) \implies P(R \mid U,\boldsymbol{\Gamma}) = P(R \mid U)$.
+
 The best estimate $\hat{U}$ then fulfills the maximum a posteriori criterium
-  \citep{bishop_pattern_2006}, and aligns $\Gamma$ into $R$.
-
+  \citep{bishop_pattern_2006}, and aligns $\boldsymbol{\Gamma}$ into $R$.
 First, we assume independence between pixels, and thus break down the
@@
   \begin{equation}
-  P(R \mid U,\Omega) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
+  P(R \mid U) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
     P\left( \vec{f}' \mid U \right),
@@
  	\begin{align}
-  P( R \mid U, \Omega ) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
+  P( R \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
   \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
@@
   $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
+$C$ is the number of channels comprised in the reference image $R$.
 
@@
   \underset{\vec{r} \in \Omega_l}{\prod}
-  \mathcal{N} \left( \vec{f}' \mid \Theta_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) \notag\\
-  &= C + \underset{l}{\sum} \int_{\Omega_l}
+  \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) \notag\\
+  &= \const + \underset{l}{\sum} \int_{\Omega_l}
   \mdist{f'}{l} \,d\vec{r} \notag\\
@@
   this expression is dual to the energy functional corresponding
-  to a discrete \gls*{acwe} framework \citep{chan_active_2001}
-  with the corresponding anisotropic regularization term \citep{nagel_investigation_1986}.
+  to a discrete \acrlong*{acwe} framework \citep{chan_active_2001}
+  with the corresponding anisotropic regularization term of \cite{nagel_investigation_1986}.
 
@@
   derive $u(\vec{r})$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
-Densification is achieved through a set of associated basis functions $\psi_k$:
+Densification is achieved through a set of associated basis functions $\psi_k$.
+In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
+  of degree 3: % ($\beta_3$).
 
@@
 
-In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
-  of degree 3 ($\beta_3$).
-Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
-  $\psi$ by the actual kernel function, the transformation writes:
-
-  \begin{equation}
-    \vec{v}_i' = \vec{v}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
-      \beta_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
-  \label{eq:transformation}
-  \end{equation}
+%Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
+%  $\psi$ by the actual kernel function, the transformation writes:
 %
-  with $\hat{\mathbf{e}}_d$ being the unitary vector along axis $d$.
+%  \begin{equation}
+%    \vec{v}_i' = \vec{v}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
+%      \beta_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
+%  \label{eq:transformation}
+%  \end{equation}
+%%
+%  with $\hat{\mathbf{e}}_d$ being the unitary vector along axis $d$.
 
@@
   corresponding to each phantom type, with two resolutions each
-  ($1.0mm$ and $2.0mm$ isotropic).
+  (1.0mm and 2.0mm isotropic).
+Field of view at both resolutions was \isores{100}mm.
 Simulations were corrupted with rician noise for a \gls*{snr} of 300.0.
-The reference surfaces were extracted at the higher resolution (1.0mm isotropic),
-  using \texttt{mri\_tessellate}\footnote{A marching-cubes algorithm shipped with 
+The reference surfaces were extracted at high resolution, using 
+  \texttt{mri\_tessellate}\footnote{A marching-cubes algorithm shipped with 
   \emph{FreeSurfer}.}.
@@
   located in isotropic grids covering the full extent of the phantom.
-The first level had $50.5mm$ of separation between control points and the second $25.25mm$.
+Separation between control points was isotropic, and set to 50.5mm for the first level
+  of distortion and 25.25mm for the second.
 The rationale behind this choice is producing large deformations on the contours with
@@
 Invertibility of $U_{true}$ is ensured by controlling the maximum displacement at each
-  level ($20.2mm$ and $10.1mm$ respectively) as studied in \citep{rueckert_diffeomorphic_2006}.
+  level (20.2mm and 10.1mm respectively) as studied in \citep{rueckert_diffeomorphic_2006}.
 

### 2015-03-24 Publications/2015-NeuroImage/08-acknowledgements.tex
Updates after Yasser and Dominique

@@
 
-DZ is supported by the Swiss National Science Foundation under grant PBELP2-137727.
+DZ was supported by the Swiss National Science Foundation under grants PBELP2\_137727 
+  and P300P2\_147778, and NSF-DMS 1418812
 This study is supported by: the Spain's Ministry of Science and Innovation

### 2015-03-24 Publications/2015-NeuroImage/supplemental.tex
editions after @daducci review
Updates after Yasser and Dominique

@@
 % @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-03-18 15:12:35
+% @Last Modified time: 2015-03-24 16:12:10
 
@@
 
-\paragraph{Obtaining the update equation} Once the shape gradients are obtained \eqref{eq:gradient_final},
-  we apply the Fourier Transform to solve the \gls*{pde}:
+\paragraph{Obtaining the update equation} To solve the differential equation in \eqref{eq:gradient_descent},
+  we use a semi-implicit Euler scheme, where the shape-gradients $\vec{g}_k$ are explicit:
 \begin{align}
-  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} \equiv \frac{\vec{u}_k^{t+1} - \vec{u}_k^t}{\delta} \implies
-  \vec{u}_k^{t+1} &= \delta \Big( \vec{g}_k + \boldsymbol{\alpha} \vec{u}_k
-  - \boldsymbol{\beta} \Delta \vec{u}_k \Big) + \vec{u}_k^t, \notag \\
-  \vec{u}_k^{t+1} &= \delta \vec{g}_k +  \boldsymbol{\alpha} \delta \vec{u}_k
-  - \boldsymbol{\beta} \delta \Delta \vec{u}_k + \vec{u}_k^t
+  \vec{g}_k^{t+1} &= \vec{g}_k^{t} - \vec{u}_k^{t}, \notag \\
+  \vec{u}_k^{t+1} &= \vec{u}_k^{t} + \delta \, \Big( \vec{g}_k^{t+1} + (\boldsymbol{\alpha} - \boldsymbol{\beta} \Delta) \, \vec{u}_k^{t+1} \Big),
+  \label{eq:semiimplicit_euler}
+\end{align}
+and introducing the $\vec{g}_k^{t+1}$ in the second equation:
+\begin{align}
+  \vec{u}_k^{t+1} &= \vec{u}_k^{t} + \delta \Big( \vec{g}_k^{t} + \delta \vec{u}_k^t + (\boldsymbol{\alpha} - \boldsymbol{\beta} \Delta) \, \vec{u}_k^{t+1} \Big) \notag \\
+  \vec{u}_k^{t+1} &= \frac{ (1+\delta^2) \, \vec{u}_k^{t} + \vec{g}_k^{t}} {1 - \boldsymbol{\alpha} + \boldsymbol{\beta} \Delta}
   \label{eq:discrete_derivative}
@@
   of being sparse, as only few $\psi_{ik} > 0$ in the surroundings of $\vec{v}_i$.
-Then, the gradients $g_k$ are easily computed using the matrix product:
+Then, the gradients $\vec{g}_k$ are easily computed using the matrix product:
 \begin{equation}

### 2015-03-26 Publications/2015-NeuroImage/03-introduction.tex
Removing formulas from paper

@@
 % -*- root: 00-main.tex -*-
-\section{Introduction}%
-\label{sec:introduction}
-Image registration and segmentation are crucial in current pipelines for
-  analysis of whole-brain connectivity using \gls*{dmri}, such as the extraction
-  of brain connectivity \citep{craddock_imaging_2013}.
+\section{Introduction}\label{sec:introduction}
 
-An early integration of segmentation and registration by \cite{bertalmio_morphing_2000} proposed
-  a sequential deformation of active contours for object tracking in series of 2D images.
+\Gls*{dmri} data of the brain present severe distortions, directly related to the
+  effort in minimizing acquisition time.
+In order to quickly probe the diffusion process in sufficiently different orientations,
+  \gls*{dmri} are acquired with \gls*{epi} sequences, which minimize the bandwidth of the \gls*{pe}
+  direction.
+This limitation originates a warping of the reconstructed image in regions affected by
+  small deviations from the main magnetic field $B_0$.
+One important source of these deviations is the discontinuity of magnetic susceptibility across
+  the imaged object.
+In fact, the distortion is the effect of the underlaying misregistration between correct and
+  warped data.
+Tissue/air boundaries show large steps of susceptibility and this artifact is
+  particularly evident in the surroundings of the orbito-frontal lobe and the temporal
+  bone of both hemispheres.
+Several approaches have been proposed to estimate the distortion and correct
+  for the artifact.
+The first solution consisted on estimating the deviations from $B_0$ with field mapping
+  techniques \citep{andersson_modeling_2001}.
+A second family of methods require an extra \gls*{epi} acquisition, but switching
+  the \gls*{pe} axis \citep{chiou_simple_2000} or reversing the direction of gradient \gls*{pe}
+  increments \citep{cordes_geometric_2000,holland_efficient_2010}.
+\cite{kybic_unwarping_2000} proposed nonlinear registration of the \emph{b0}
+  image of the \gls*{dmri} to an undistorted \gls*{t2} image of
+   the same subject.
+
+The artifact has been proven to cause an important impact on the anatomy extracted
+  from \gls*{dmri} \citep{irfanoglu_effects_2012}, particularly in certain fiber bundles
+  when used in whole-brain tractography.
+Currently, a broad range of applications make use of whole-brain tractography, such as
+  connectome extraction \citep{craddock_imaging_2013} or tract-based spatial statistics
+  \citep{smith_tractbased_2006}.
+These applications generally require the segmentation of the \gls*{dmri} data, also
+  hindered by distortions.
+One way to identify the regions of interest is extracting them from an appropriate 
+  \gls*{t1} of the same subject and map them into diffusion space,
+  in a sort of atlas-based segmentation.
+
+\paragraph{State of the art}\label{sec:state_of_art}
+  Active contours have been successfully integrated in image registration methods
+  for applications such as longitudinal object tracking \citep{paragios_level_2003} and
+  atlas-based segmentation \citep{gorthi_active_2011}.
+\cite{bertalmio_morphing_2000} proposed morphing active-contours for object tracking in 
+  series of 2D images.
 Shortly, \cite{yezzi_variational_2003} presented the first method including a full solution to
@@
   the contours to the whole image definition.
-\cite{greve_accurate_2009} presented a widely used registration method called \emph{bbregister},
-  included in \emph{FreeSurfer} \citep{fischl_freesurfer_2012}.
-Their framework performs robust registration of brain surfaces into the intensity information
-  of a target \gls*{mri}, using an affine transformation and active contours.
 Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
@@
   of atlas-based segmentation.
-Alternatively to the historical use of active contours, some Bayesian approaches
-  have been proposed as well \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014}.
 
-We propose \emph{regseg} as a solution to joint segmentation and registration 
-  problems with a shape priors approach.
-An immediate application of \emph{regseg} is the correction of susceptibility-derived
-  distortions typically present in \gls*{dmri} data and its segmentation
-  after correction.
-Both correction and segmentation are required steps in current investigations of whole-brain
-  tractography, such as the connectome extraction \citep{daducci_connectome_2012}.
-Indeed, the joint segmentation and correction can exploit the anatomy extracted from
-  the \gls*{t1} as strong shape-priors.
+\cite{greve_accurate_2009} presented a relevant registration method called \emph{bbregister},
+  now included in \emph{FreeSurfer} \citep{fischl_freesurfer_2012}.
+The method is a standard choice to map surfaces extracted in the \gls*{t1} image to the \emph{b0}
+  image\footnote{The term \emph{b0} refers to one or more images within the
+  \gls*{dmri} dataset that are acquired with a null or very low gradient intensity or
+   \emph{b}-value.} of the \gls*{dmri} dataset with two limitations.
+First, the transform is affine and therefore it does not cope with the nonlinear distortions.
+Second, \emph{bbregister} looks for intensity steps in the image, using active contours with edges.
+In practice, the target \emph{b0} does not present clear frontiers of intensity between
+  interfacing tissues, what limits \emph{bbregister} to mapping the cortical surface only.
+
+\paragraph{Contributions}\label{sec:contributions}
+The hypothesis under this research is that the registration and segmentation
+  problems can be solved jointly, increasing the geometrical accuracy of the process.
+We propose \emph{regseg} as a joint solution to distortion correction and segmentation,
+  using active contours without edges \citep{chan_active_2001} that evolve driving a
+  free-deformation field of B-Spline functions.
+Optimization is performed using a descent strategy of shape-gradients
+  \citep{herbulot_segmentation_2006,besson_dream2s_2003}, therefore \emph{regseg}
+  does not implement level sets as most of the presented methods.
+Given the properties of the distortion, \emph{regseg} includes the anisotropic regularization
+  of displacement fields of \cite{nagel_investigation_1986}.
+
+While all the correction methods require an extra acquisition, \emph{regseg} is especially
+  indicated in scanning protocols without specific data for correction,
+  such as historical datasets.
+Finally, we present an instrumentation framework to evaluate correction and
+  segmentation pipelines using realistic distortions as ground truth.
+With the integration of \emph{regseg} and an in-house implementation of the 
+  \gls*{t2b} correction in the framework, we compare the performance
+  of both solutions.

### 2015-03-26 Publications/2015-NeuroImage/04-methods.tex
Removing formulas from paper

@@
 %
-
-\subsection{Susceptibility-derived distortions}
-\label{sec:distortions}
-\Gls*{dmri} data are usually acquired with \gls*{epi} schemes.
-In order to quickly probe the diffusion process within the brain in many different orientations,
-  these sequences minimize the bandwidth of the \gls*{pe} direction.
-This limitation originates a warping of the reconstructed image in regions affected by
-  small deviations from the main magnetic field $B_0$.
-One important source of these deviations ($\Delta B_0$) is the discontinuity of magnetic
-  susceptibility across the imaged object.
-Tissue/air boundaries show large steps of susceptibility and this artifact is
-  particularly evident in the surroundings of the orbito-frontal lobe or the temporal
-  bone of both hemispheres.
-
+Let $\gammaset = \{\Gamma_m: m \in \mathbb{N}, m \leq N_c\}$ be the set of $N_c$ surfaces
+  extracted from the undistorted \gls*{t1} image (the reference space $R$).
 Distortion can be rendered as a registration problem where we look for an underlying deformation 
-  field $U_s$, mapping from an undistorted \gls*{t1} image (the reference $R$) to the distorted
-  \gls*{dmri} images (the moving space $M$), such that the structures identified in $R$ optimally
-  align with the corresponding features in $M$:
+  field $U_s$, mapping $R$ (with $\gammaset \subset R$) to the corresponding features in the distorted 
+  \gls*{dmri} images (the moving space $M$):
 
@@
   to the reference domain.
-The displacements field $U_s$ along the \gls*{pe} axis can be estimated theoretically
-  \citep{jezzard_correction_1995}:
+The displacements along the \gls*{pe} axis are related to the local field
+  deviation $\Delta B_0(\vec{r})$ \citep{jezzard_correction_1995}:
 
   \begin{equation}
-  \vec{r}' = \vec{r} + \frac{\gamma \, T_{acq}\, s_{PE}}{2\pi}\Delta B_0(\vec{r}) \cdot \hat{\vec{e}}_{PE},
+  u(\vec{r}) = \frac{\gamma \, T_{acq}\, s_{PE}}{2\pi}\Delta B_0(\vec{r}) \: \hat{\vec{e}}_{PE},
   \label{eq:fieldmap}
@@
   vector along \gls*{pe}.
-Several methodologies have been proposed to estimate $U_s$ and fix the artifact.
-The first solution consisted on estimating $\Delta B_0$ with field mapping
-  techniques \citep{andersson_modeling_2001}.
-A second family of methods require an extra \gls*{epi} acquisition, but switching
-  the \gls*{pe} axis \citep{chiou_simple_2000} or reversing the direction of gradient \gls*{pe}
-  increments \citep{cordes_geometric_2000,holland_efficient_2010}.
-\cite{kybic_unwarping_2000} proposed nonlinear registration of the \emph{b0}
-  image\footnote{The term \emph{b0} refers to one or more images within the
-  \gls*{dmri} dataset that are acquired with a null or very low gradient intensity or
-   \emph{b}-value.} of the \gls*{dmri} to an undistorted \gls*{t2} image of
-   the same subject.
-In the following, we will refer to this method as \gls*{t2b}.
-Further developments \citep{irfanoglu_susceptibility_2011} extended the \gls*{t2b} 
-  approach to the 3D problem.
 
-\subsection{Contributions}
-\label{sec:contributions}
-%
-While all the correction methods require an extra acquisition, \emph{regseg} is especially
-  indicated in scanning protocols without specific data for correction,
-  such as historical datasets.
-
-Joint segmentation and distortion correction of \gls*{dmri} data is also interesting to
-  improve the overall consistence of complex processing workflows.
-The closest antecedent of \emph{regseg} is \emph{bbregister} \citep{greve_accurate_2009}.
-Both methods use active contours to align surfaces to same-subject \gls*{dmri} images.
-However, \emph{bbregister} is limited to affine transformations, and thus, it is not able to
-  solve the susceptibility-derived distortions, generally nonlinear.
-To enable registration of uncorrected data, \emph{bbregister} excludes from the cost-function
-  those regions frequently warped.
-Conversely, \emph{regseg} implements a free-deformation field, supported with B-Spline
-  functions to estimate deformations.
-A second diverging foundation is the active-contours framework.
-In \emph{regseg}, active contours are implemented without edges \citep{chan_active_2001},
-  enabling the tool to segment homogeneous regions within the brain.
-On the other hand, \emph{bbregister} requires that the target features present intensity
-  boundaries as it implements active contours with edges.
-The tissue interfaces imaged with \gls*{dmri} are affected by partial volume effects that
-  blur boundaries.
-In practice, registration of \gls*{t1} and \gls*{dmri} data with \emph{bbregister}
-  is limited to the linear match of the pial surface of the brain.
-
-Finally, we present an instrumentation framework to evaluate correction and
-  segmentation pipelines using realistic distortions as ground truth.
-With the integration of \emph{regseg} and an in-house implementation of the 
-  \gls*{t2b} correction in the framework, we compare the performance
-  of both solutions.
+% Joint segmentation and distortion correction of \gls*{dmri} data is also interesting to
+%   improve the overall consistence of complex processing workflows.
+% The closest antecedent of \emph{regseg} is \emph{bbregister} \citep{greve_accurate_2009}.
+% Both methods use active contours to align surfaces to same-subject \gls*{dmri} images.
+% However, \emph{bbregister} is limited to affine transformations, and thus, it is not able to
+%   solve the susceptibility-derived distortions, generally nonlinear.
+% To enable registration of uncorrected data, \emph{bbregister} excludes from the cost-function
+%   those regions frequently warped.
+% Conversely, \emph{regseg} implements a free-deformation field, supported with B-Spline
+%   functions to estimate deformations.
+% A second diverging foundation is the active-contours framework.
+% In \emph{regseg}, active contours are implemented without edges \citep{chan_active_2001},
+%   enabling the tool to segment homogeneous regions within the brain.
+% On the other hand, \emph{bbregister} requires that the target features present intensity
+%   boundaries as it implements active contours with edges.
+% The tissue interfaces imaged with \gls*{dmri} are affected by partial volume effects that
+%   blur boundaries.
+% In practice, registration of \gls*{t1} and \gls*{dmri} data with \emph{bbregister}
+%   is limited to the linear match of the pial surface of the brain.
 
@@
 %
-In a Bayesian framework, the mappings $U$ in \eqref{eq:transform} are
+In a Bayesian framework for registration \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014},
+  the mappings $U$ in \eqref{eq:transform} are
   evaluated based on their posterior probability given the observed data
-  $R$.
+  $M$.
+Let $\omegaset = \{\Omega_l: l \in \mathbb{N}, l \leq L\}$ be the set of $L$ competing regions in
+  which $M$ is partitioned by the projection of \gammaset, 
+  the given surfaces serving as shape priors.
 Using the Bayes' rule, the posterior likelihood is computed as:
@@
   \begin{equation}
-  P(U \mid R, \boldsymbol{\Gamma}) = \frac{P(R \mid U,\boldsymbol{\Gamma})\, P(U)}{P(R)},
+  P(U \mid M, \omegaset ) = \frac{P(M \mid U,\omegaset )\, P(U)}{P(M)},
   \label{eq:bayes_rule}
@@
 %
-  where $P(R \mid U,\boldsymbol{\Gamma})$ is the data-likelihood, and
-  $\boldsymbol{\Gamma} = \{\Gamma_{\text{in},\text{out}}\}$ is a set of surfaces 
-  corresponding to the interfaces between pairs of competing regions $\Omega_\text{in}$
-  and $\Omega_\text{out}$ drawn from the full set of regions $\{\Omega_l \mid l \in \mathbb{N}, l \leq L\}$.
-As $\boldsymbol{\Gamma}$ are mapped by $U$, we simplify
-  $P(U, \boldsymbol{\Gamma}) = P(U) \implies P(R \mid U,\boldsymbol{\Gamma}) = P(R \mid U)$.
+  where $P(M \mid U,\omegaset)$ is the data-likelihood.
+As $\omegaset$ are mapped by $U$, we simplify
+  $P(U, \omegaset) = P(U) \implies P(M \mid U,\omegaset) = P(M \mid U)$.
 
 The best estimate $\hat{U}$ then fulfills the maximum a posteriori criterium
-  \citep{bishop_pattern_2006}, and aligns $\boldsymbol{\Gamma}$ into $R$.
+  \citep{bishop_pattern_2006}, and aligns \gammaset into $M$.
 First, we assume independence between pixels, and thus break down the
@@
   \begin{equation}
-  P(R \mid U) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
+  P(M \mid U) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
     P\left( \vec{f}' \mid U \right),
@@
 %
-  where $\vec{f}' = R(\vec{r}')$ is the feature vector at the displaced
+  where $\vec{f}' = M(\vec{r}')$ is the feature vector at the displaced
   position $\vec{r}'$ \eqref{eq:transform} in the multispectral target
@@
  	\begin{align}
-  P( R \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
+  P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
   \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
@@
   $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
-$C$ is the number of channels comprised in the reference image $R$.
+$C$ is the number of channels comprised in the reference image $M$.
 
@@
 
-  \begin{align*}
+  \begin{align}
   P(U) = \underset{\vec{r}}{\prod}\, p(\vec{u}) &=
@@
   p_1(\vec{u}) &= \mathcal{N}(  \nabla \vec{u} \mid 0, \mathbf{B}^{-1}),
-  \end{align*}
+  \label{eq:thikonov}
+  \end{align}
 %
   imposing that the distortion and its gradient have zero
-  mean and variance governed by the matrices $\mathbf{A} = (a_{ij})$ and $\mathbf{B} = (b_{ij})$.
-Since the anisotropy is generally aligned with the imaging axes, $\mathbf{A}$ and $\mathbf{B}$
-  can be simplified to diagonal matrices.
-For convenience, we will set $\alpha_i = a_{ii}^2$ and $\beta_i = b_{ii}^2$, what can be reduced
-  using the Hadamard square root notation\footnote{The Hadamard power of a matrix or a vector
-  is the power of its elements $\mathbf{M}^{\circ p} = ({m_{ij}}^{p})$} as
-  $\mathbf{A}=\boldsymbol{\alpha}^{\circ\frac12}\,\vec{I}_n$ and
-  $\mathbf{B}=\boldsymbol{\beta}^{\circ\frac12}\,\vec{I}_n$.
-
+  mean and variance governed by the matrices $\mathbf{A}$ and $\mathbf{B}$.
 Finally, the maximum a posteriori problem is adapted into a variational one where we look for
@@
   \begin{align}
-  E(R \mid U) &= -\log \underset{l}{\prod}
+  E(M \mid U) &= -\log \underset{l}{\prod}
   \underset{\vec{r} \in \Omega_l}{\prod}
@@
   &= \const + \underset{l}{\sum} \int_{\Omega_l}
-  \mdist{f'}{l} \,d\vec{r} \notag\\
-  &+ \int_{\Omega} \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
-  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r},
+  \mdist{f'}{l} \,d\vec{r} \, + \notag\\
+  &+ \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
   \label{eq:energy}
@@
 %
-  this expression is dual to the energy functional corresponding
+
+This expression is dual of the Mumford-Shah functional corresponding
   to a discrete \acrlong*{acwe} framework \citep{chan_active_2001}
@@
 \paragraph*{Deformation model}\label{sec:deformation_model}
-Let us denote $\{\vec{v}_i\}_{i=1 \ldots N_c}$ the vertices of one or several prior
-  surface(s).
-In our application, these surfaces are triangularized meshes extracted using \emph{FreeSurfer}
-  \citep{fischl_freesurfer_2012}.
-The transform $\hat{U}$ \eqref{eq:transform} is supported by a dense deformation field
-  $\vec{u} = u(\vec{r})$, such that:
-
-  \begin{equation}
-  \vec{v}_i' = \hat{U}\{\vec{v}_i\} = \vec{v}_i + u(\vec{v}_i) = \vec{v}_i + \vec{u}_i.
-  \label{eq:nodes_tfm}
-  \end{equation}
-
-Since the nodes of the anatomical surfaces likely lay off-grid, it is required to
-  derive $u(\vec{r})$ from a discrete set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$.
+Since the vertices of the surfaces $\{\vec{v}_i: \vec{v}_i \subset \gammaset \}_{i=1 \ldots V}$
+  likely lay off-grid, it is required to derive $u(\vec{v}_i)$ from a discrete set of parameters
+  $\{\vec{u}_k\}_{k=1 \ldots K}$.
 Densification is achieved through a set of associated basis functions $\psi_k$.
@@
   \begin{equation}
-  u(\vec{r}) = \sum_k \psi_k(\vec{r}) \vec{u}_k.
-  \label{eq:intp_kernel}
+  \vec{v}_i' = \vec{v}_i + u(\vec{v}_i) = \vec{v}_i + \sum_k \psi_k(\vec{r}) \: \vec{u}_k.
+  \label{eq:nodes_tfm}
   \end{equation}
-
-%Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
-%  $\psi$ by the actual kernel function, the transformation writes:
-%
-%  \begin{equation}
-%    \vec{v}_i' = \vec{v}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
-%      \beta_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
-%  \label{eq:transformation}
-%  \end{equation}
-%%
-%  with $\hat{\mathbf{e}}_d$ being the unitary vector along axis $d$.
 
@@
   \begin{equation}
-  \frac{\partial u(\vec{r},t)}{\partial t} \propto - \frac{\partial E(R \mid U)}{\partial \vec{u}_k},
+  \frac{\partial u(\vec{r},t)}{\partial t} \propto - \frac{\partial E(M \mid U)}{\partial \vec{u}_k},
   \label{eq:general_gradient_descent}
@@
   $\hat{U}$ of the transformation at the current time point.
-Now, we introduce \eqref{eq:energy} in \eqref{eq:general_gradient_descent}:
+Let us assume that anisotropy is aligned with the imaging axes to simplify
+  \eqref{eq:energy} into expression \eqref{eq:app_energy} of \ref{app:reg_term}
+  and compute the derivative \eqref{eq:general_gradient_descent}:
 
@@
   \frac{ \partial }{\partial \vec{u}_k} \Big\{
-  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \notag\\
-  &+ \int_{\Omega} [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
+  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, + \notag\\
+  &+ \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
   + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
@@
   \end{align}
-
-We apply a discretized interpretation of \eqref{eq:shape_gradients} to compute
-  the data term in \eqref{eq:gradient_descent} using shape-gradients
-  \citep{herbulot_segmentation_2006}, and ultimately avoid level sets.
-We refer the reader to \ref{app:shape_priors} in order to fill in the gap
-  between \eqref{eq:gradient_descent} and the following expression:
+%
+  where $\vec{u}^{\circ2} = \vec{u}^T \cdot \vec{u}$.
+Then, the data and regularization terms are split and discretized to compute their
+  derivatives.
+The intermediate steps to get to \eqref{eq:gradient_descent} from \eqref{eq:final_gradient} are
+  detailed in the \suppl{equations SM6-SM12}:
 
   \begin{align}
-  \frac{\partial E_{data}(\vec{u})}{\partial \vec{u}_k} &=
-  \frac{ \partial }{\partial \vec{u}_k} \left\{
-   \underset{\vec{x} \in \Omega_l}{\sum} \underset{l}{\sum} \mdist{f'}{l} \right\} \notag\\
-  &= \underset{i}{\sum} w_i
-   \left\langle \frac{\partial \vec{v}_i'}{\partial \vec{u}_k}, \bar{s}_i'\right\rangle,
-  \label{eq:gradient_wshape}
+  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
+  \vec{g}_k  + \boldsymbol{\alpha} \cdot \vec{u}_k + \boldsymbol{\beta} \cdot (\Delta \vec{u}_k).
+  \label{eq:final_gradient}
   \end{align}
-%
-  in this case, the formulation has been adapted to the non-binary case, $\{l,m\}$
-    being any pair of neighboring regions, and $\Gamma_{l,m}$ the contour separating
-    them such that
-    $\vec{x}' = \vec{v}' \in\Gamma_{l,m} \iff \vec{x}\in \partial\Omega_i \cap \partial\Omega_j$,
-    $\bar{s}_i'$ is the speed vector projected on to the unit inward normal to the contour
-    at $c_i'$ (described in \autoref{fig:method}), $w_i$ is the area contribution of vertex
-    $c_i'$ to the the total area of the surface it belongs.
 
-Finally, we can compute:
+Finally, to descend this gradient we establish a semi-implicit Euler scheme (see \suppl{section S1.3}),
+  with a step size parameter $\delta$, that we solve it in the spectral domain as follows:
 
   \begin{align}
-  \frac{\partial \vec{v}_i'}{\partial \vec{u}_k} &= \frac{\partial}{\partial \vec{u}_k}
-  \left\{ \vec{v}_i + \sum_k \psi_k(\vec{v}_i) \vec{u}_k \right\}
-  = \psi_k(\vec{v}_i)\, \hat{\vec{e}}
-  \label{eq:basis_derivative}
-  \end{align}
-%
-  where $\hat{\vec{e}}$ is the coordinates system's unit vector.
-The vertex speeds $\bar{s}_i'$ obtained by computing the shape-gradients \eqref{eq:shape_gradients},
-	are then projected to the deformation field in order to obtain the derivatives $\vec{g}_k$
-	corresponding to $\vec{u}_k$:
-
-  \begin{equation}
-  \vec{g}_{i,k} = w_i \, \left\langle \frac{\partial}{\partial \vec{u}_k}{\vec{v}_i}', \bar{s}_i'\right\rangle
-  = - w_i \left[ \mdist{f_i'}{l} - \mdist{f_i'}{m} \right] \psi_k(\vec{v}_i)\, \hat{\vec{n}}_i,
-  \end{equation}
-%
-  then the full gradient evolution equation \eqref{eq:gradient_descent} yields:
-
-  \begin{align}
-  \frac{\partial E(\vec{u}_k)}{\partial \vec{u}_k} =
-  &\underset{i}{\sum} \vec{g}_{i,k} +2\, \boldsymbol{\alpha} \vec{u}_k
-  -2\, \boldsymbol{\beta} \Delta \vec{u}_k,
-  \label{eq:gradient_final}
-  \end{align}
-
-Finally, we introduce a step size parameter $\delta$ to discretize the artificial time parameter
-  in the iterative optimization.
-Then, applying the Fourier transform as detailed in the \suppl{section S1}, the
-  update equation is derived from \eqref{eq:gradient_final}:
-
-  \begin{align}
-  \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\vec{u}_k^t / \delta - \sum_i \vec{g}_{i,k}\}}%
-                  {\mathcal{F}\{(1/\delta+\boldsymbol{\alpha})I-\boldsymbol{\beta}\Delta\}} \right\}
+  \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\delta^{-1} \, \vec{u}_k^t + \vec{g}_k\} }%
+                  {\mathcal{F}\{(\delta^{-1} - \boldsymbol{\alpha})\, I-\boldsymbol{\beta}\Delta\}} \right\}
   \label{eq:update_equation}

### 2015-03-26 Publications/2015-NeuroImage/08-appendix-00-shapepriors.tex
almost ready to revision round 2
update supplementals
improving to second round
Removing formulas from paper

--- 2015-02-18 Publications/2015-NeuroImage/08-appendix-00-shapepriors.tex
@@
 \renewcommand{\theequation}{A.\arabic{equation}}
+\renewcommand{\thesubsection}{Appendix \arabic{subsection}}
 
@@
 
-\subsection*{Instant shape priors}
-\label{app:shape_priors}
-Let $F(\vec{r})$ be an ``arbitrary'' function over the image domain
-  $\Omega = \Omega_l \cup \Omega_m$ splitted in two regions $l$ and
-  $m$, and $\Gamma_{l,m}$ a closed boundary between them
-  as defined in \nameref{sec:methods_map}.
-We now derive the domain integral w.r.t. $t$:
+\subsection{Simplifying the regularization term}\label{app:reg_term}
+The exponentials of the Thikonov regularization prior \eqref{eq:thikonov} have the general form
+  $\vec{v}^T \mathbf{M} \vec{v}$.
+If $\mathbf{M}$ is a $n \times n$ diagonal matrix such that $\mathbf{M} = \vec{m} \, \mathbf{I}_n$, 
+  then:
+\begin{equation*}
+\vec{v}^T \mathbf{M} \vec{v} = \vec{m} \cdot (\vec{v}^T \mathbf{I}_n \vec{v}) = \vec{m} \cdot \vec{v}^{\circ2},
+\end{equation*}
+%
+  where we have introduced the Hadamard power notation\footnote{The Hadamard power of a matrix or a vector
+  is the power of its elements $\mathbf{M}^{\circ p} = ({m_{ij}}^{p})$}.
 
-  \begin{align*}
-  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) d\vec{r} &=
-  \int_\Omega \frac{\partial}{\partial t}F(\vec{r}) d\vec{r} \notag\\
-  &- \int_{\Gamma_{l,m}} F(\vec{r}) \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
-  N_{\Gamma_{l,m}}\right\rangle d\vec{r},
-  \end{align*}
-
-  where $\left\langle\frac{\partial\Gamma_{l,m}}{\partial t}, N_{\Gamma_{l,m}}\right\rangle$ is
-  the projection of the boundary movement on the unit inward normal $N_{\Gamma_{l,m}}$.
-Assuming that the region descriptors $\Theta_l$ vary slowly enough, we can consider
-  that $\frac{\partial}{\partial t} F(\vec{r}) = 0$ and thus:
-
-  \begin{equation*}
-  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) d\vec{r} =
-  - \int_{\Gamma_{l,m}} F(\vec{r}) \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
-  N_{\Gamma_{l,m}}\right\rangle d\vec{r}
-  \end{equation*}
-
-  In our formulation, $F(\vec{r})$ is the difference on
-    sample likelihood between the inner ($\Omega_l$) and the outer ($\Omega_m$) regions.
-  Therefore, we substitute:
-
-    \begin{equation}
-    \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) d\vec{r} =
-    - \int_{\Gamma_{l,m}} \left[ \mathcal{E}_{l}(\vec{r}) - \mathcal{E}_{m}(\vec{r}) \right]
-    \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
-  N_{\Gamma_{l,m}}\right\rangle d\vec{r},
-  	\label{eq:shape_gradients}
-    \end{equation}
-
-    where $\mathcal{E}_l(\vec{r}) = \mathcal{D}^2_{l}(R(\vec{r}))$ for briefness.
+Since the anisotropy is generally aligned with the imaging axes, 
+  $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:energy} can be simplified to diagonal matrices, such that
+  $\mathbf{A}= \, \boldsymbol{\alpha}\,\vec{I}_n$ and
+  $\mathbf{B}= \, \boldsymbol{\beta}\,\vec{I}_n$.
+Substituting in the equation \eqref{eq:energy}, we get:
+  \begin{align}
+  E(M \mid U) &= \const + \underset{l}{\sum} \int_{\Omega_l}
+  \mdist{f'}{l} \,d\vec{r} \notag\\
+  &+ \int_{\Omega} \frac12 \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2} + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}.
+  \label{eq:app_energy}
+  \end{align}

### 2015-03-26 Publications/2015-NeuroImage/supplemental.tex
Removing formulas from paper

--- 2015-03-18 Publications/2015-NeuroImage/supplemental.tex
@@
 % @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-03-18 15:12:35
+% @Last Modified time: 2015-03-26 17:57:06
 
@@
 diffusion MR images of the brain driven by active-contours}}
-\author{Oscar Esteban}
+\author{Oscar Esteban and Dominique Zosso}
 \date{February 2015}
@@
 
-\paragraph{Discretization of the shape gradients} The energy functional to be optimized
-  in \emph{regseg} is presented in Eq. 7:
+\subsection{Computing the gradients of shape-priors}\label{sec:shape_priors}
+The computation of gradients at the locations of the active contours in the
+  instant $t$ is based on the work of \cite{herbulot_segmentation_2006}.
+Let $F(\vec{r})$ be an ``arbitrary'' function over the image domain
+  $\Omega = \Omega_l \cup \Omega_m$ splitted in two regions $l$ and
+  $m$, and $\Gamma_{l,m}$ a closed boundary between them.
+We now derive the domain integral w.r.t. $t$:
+
   \begin{equation}
-  E(R \mid U) = C + \underbracket{\underset{l}{\sum} \int_{\Omega_l}
-  \mdist{f'}{l} \,d\vec{r}}_{\text{Data term}}
-  + \underbracket{\underset{\Omega}{\int} \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
-  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}}_{\text{Regularization term}}.
-  \label{eq:energy}
+  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) d\vec{r} =
+  \int_\Omega \frac{\partial}{\partial t}F(\vec{r}) d\vec{r} 
+  - \int_{\Gamma_{l,m}} F(\vec{r}) \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
+  N_{\Gamma_{l,m}}\right\rangle d\vec{r},
+  \end{equation}
+%
+  where $\left\langle\frac{\partial\Gamma_{l,m}}{\partial t}, N_{\Gamma_{l,m}}\right\rangle$ is
+  the projection of the boundary movement on the unit inward normal $N_{\Gamma_{l,m}}$.
+Assuming that the region descriptors $\{\boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l\}$ vary slowly enough, we can consider
+  that $\frac{\partial}{\partial t} F(\vec{r}) = 0$ and thus:
+
+  \begin{equation}
+  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) d\vec{r} =
+  - \int_{\Gamma_{l,m}} F(\vec{r}) \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
+  N_{\Gamma_{l,m}}\right\rangle d\vec{r}.
+  \label{eq:shape_gradients}
   \end{equation}
 
-To search for the minimum of $E(R \mid U)$ w.r.t. the coefficients $\vec{u}_k$, we use a gradient descent strategy.
-In Eq. 11 we introduced the derivative of \eqref{eq:energy}:
-  \begin{equation}
-  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} =
-  \frac{ \partial }{\partial \vec{u}_k} \Big\{
-  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r}
-  + \int_{\Omega} [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
-  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
-  \Big\}.
-  \label{eq:gradient_descent}
-  \end{equation}
-
-Let us compute the derivative of the data term first (Eq. 12):
-\begin{equation}
-  \frac{\partial E_{data}(\vec{u})}{\partial \vec{u}_k} =
-  \frac{ \partial }{\partial \vec{u}_k} \Big\{
-  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \Big\} =
-  \underset{l}{\sum} \, \frac{ \partial }{\partial \vec{u}_k} \Big\{
-  \int_{\Omega_l} \mdist{f'}{l} \Big\},
-  \label{eq:data_derivative}
-\end{equation}
-where we can introduce the shape-gradients theory from Appendix 1.
-The equation A.1 is discretized as follows.
+The equation \eqref{eq:shape_gradients} is discretized as follows.
 First, the surface between limiting regions $\{l, m\}$, $\Gamma_{l,m}$ is explicitly represented by
@@
   set of normals $\hat{\vec{n}}_i$ at each vertex of the mesh.
-The resulting summation is, therefore, discrete and the integral operator is replaced by the sum.
+The resulting summation is, therefore, discrete and the integral operator is replaced by the sum:
   \begin{align}
@@
 Then, the speed of $\vec{v}_i$ is discretized using the artificial time-step parameter $\delta$, as the displacement
-  $\vec{d}_i = \vec{v}_i(\delta = t+1) - \vec{v}_i(\delta = t)$:
+  $\frac{\partial \vec{v}_i}{\partial t} = \vec{v}_i(\delta = t+1) - \vec{v}_i(\delta = t)$:
   \begin{equation}
   \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) \, d\vec{r} =
-  - \underset{i}{\sum} w_{p,i} F(\vec{v}_i) \vec{d}_i \cdot \hat{\vec{n}}_i.
+  - \underset{i}{\sum} w_{p,i} F(\vec{v}_i) \frac{\partial \vec{v}_i}{\partial t} \cdot \hat{\vec{n}}_i.
   \label{eq:shape_gradient_disc1}
@@
   \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) \, d\vec{r} =
-  - \underset{i}{\sum} w_{p,i} \, \vec{d}_i \cdot
-  \underbracket{\Big[ \big(F_{out}(\vec{v}_i) - F_{in}(\vec{v}_i) \big) \hat{\vec{n}}_i \Big]}_{\bar{s}_i \text{ in Figure 1}}.
+  - \underset{i}{\sum} \, \frac{\partial \vec{v}_i}{\partial t} \cdot
+  \underbracket{w_{p,i} \, \Big[ F_{out}(\vec{v}_i) - F_{in}(\vec{v}_i) \Big] \hat{\vec{n}}_i}_{\bar{s}_i \text{ in Figure 1}}.
   \label{eq:shape_gradient_disc2}
@@
 
-Now we can identify $t$ with the set of time-dependent coefficients, $\{\vec{u}_k\}$, and the target function $F$ with $\mathcal{D}_l^2$.
-To introduce the speed of the vertices from equation 13, we replace $\vec{d}_i$:
+\subsection{Gradient-descent optimization}\label{sec:gradient_descent}
+The energy functional to be optimized
+  in \emph{regseg} is presented in Eq. 7:
   \begin{equation}
-  \frac{\partial E_{data}(\vec{u})}{\partial \vec{u}_k} =
-  - \underset{i}{\sum} w_{p,i} \, \psi_k(\vec{v}_i) \hat{\vec{e}} \cdot \bar{s}_i =
-  - \underset{i}{\sum} w_{p,i} \, \psi_k(\vec{v}_i) \bar{s}_i = - \underset{i}{\sum} \vec{g}_{i,k}.
+  E(R \mid U) = C + \underbracket{\underset{l}{\sum} \int_{\Omega_l}
+  \mdist{f'}{l} \,d\vec{r}}_{\text{Data term } (E_{data})}
+  + \underbracket{\underset{\Omega}{\int} \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
+  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}}_{\text{Regularization term } (E_{reg})}.
+  \label{eq:energy}
+  \end{equation}
+
+To search for the minimum of $E(R \mid U)$ w.r.t. the coefficients $\vec{u}_k$, we use a gradient descent strategy.
+In Eq. 11 we introduced the derivative of \eqref{eq:energy}:
+  \begin{equation}
+  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} =
+  \frac{ \partial }{\partial \vec{u}_k} \Big\{
+  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r}
+  + \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
+  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
+  \Big\}.
+  \label{eq:gradient_descent}
+  \end{equation}
+
+We split \eqref{eq:gradient_descent} in the derivatives of its data and regularization terms.
+Let $\frac{\partial E_{data}}{\partial \vec{u}_k} = \vec{g}_k$ for simplicity, 
+  we compute the derivative of the data term and discretize the domain $\Omega$ as follows:
+\begin{equation}
+  \vec{g}_k =
+  \frac{ \partial }{\partial \vec{u}_k} \Big\{
+  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \Big\} =
+  \frac{ \partial }{\partial \vec{u}_k} \Big\{ \underset{l}{\sum} \underset{\vec{x} \in \Omega_l}{\sum} \, \mdist{f'}{l} \Big\},
+  \label{eq:data_derivative}
+\end{equation}
+  where we can apply the shape-gradients \eqref{eq:shape_gradient_disc2} introduced
+  in \autoref{sec:shape_priors}, and ultimately avoid implementing level sets:
+
+  \begin{align}
+  \vec{g}_k &= \underset{i}{\sum} \left\langle \frac{\partial \vec{v}_i'}{\partial \vec{u}_k}, \bar{s}_i'\right\rangle, \\
+  \text{with }
+  \bar{s}_i' &= - w_i \left[ \mdist{f_i'}{out} - \mdist{f_i'}{in} \right] \, \hat{\vec{n}}_i, \\
+  \text{and } 
+  \frac{\partial \vec{v}_i'}{\partial \vec{u}_k} &=
+  \frac{\partial}{\partial \vec{u}_k} \left\{ \vec{v}_i + \sum_k \psi_k(\vec{v}_i) \vec{u}_k \right\} = \psi_k(\vec{v}_i)\, \hat{\vec{e}}.
+  \label{eq:gradient_wshape}
+  \end{align}%
+  where $\hat{\vec{e}}$ is the coordinates system's unit vector.
+Therefore, the shape gradients projected to the grid of B-Spline control points is:
+\begin{equation}
+  \vec{g}_k =
+  - \underset{i}{\sum} \bar{s}_i \cdot \psi_k(\vec{v}_i) \, \hat{\vec{e}} = - \underset{i}{\sum} \vec{g}_{i,k}.
   \label{eq:shape_gradient_final}
@@
 
-Inserting \eqref{eq:shape_gradient_final} in \eqref{eq:gradient_descent} we reach equation 15,
-  and renaming $\vec{g}_k = - \underset{i}{\sum} \vec{g}_{i,k}$:
+It is also necessary to obtain and discretize the derivatives of the regularization term of \eqref{eq:gradient_descent}:
   \begin{equation}
-  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} =
-  \vec{g}_k + \boldsymbol{\alpha} \vec{u}_k
-  -\boldsymbol{\beta} \Delta \vec{u}_k.
+  \frac{\partial E_{reg}(\vec{u})}{\partial \vec{u}_k} = \frac{ \partial }{\partial \vec{u}_k} \Big\{
+  \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
+  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
+  \Big\} = \boldsymbol{\alpha} \cdot \vec{u}_k + \boldsymbol{\beta} \cdot \Delta \vec{u}_k.
+  \label{eq:gradient_regterm}
+  \end{equation}
+
+Inserting \eqref{eq:shape_gradient_final} and \eqref{eq:gradient_regterm} into \eqref{eq:gradient_descent} we get the
+  final evolution equation:
+  \begin{equation}
+  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} = \vec{g}_k + \boldsymbol{\alpha} \cdot \vec{u}_k + \boldsymbol{\beta} \cdot \Delta \vec{u}_k,
   \label{eq:gradient_final}
@@
 
-\paragraph{Obtaining the update equation} Once the shape gradients are obtained \eqref{eq:gradient_final},
-  we apply the Fourier Transform to solve the \gls*{pde}:
+\subsection{Obtaining the update equation} To solve the differential equation in \eqref{eq:gradient_descent},
+  we use a semi-implicit Euler scheme, referring to the discrete step size as $\delta$ and where the
+  shape-gradients $\vec{g}_k$ are explicit:
 \begin{align}
-  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} \equiv \frac{\vec{u}_k^{t+1} - \vec{u}_k^t}{\delta} \implies
-  \vec{u}_k^{t+1} &= \delta \Big( \vec{g}_k + \boldsymbol{\alpha} \vec{u}_k
-  - \boldsymbol{\beta} \Delta \vec{u}_k \Big) + \vec{u}_k^t, \notag \\
-  \vec{u}_k^{t+1} &= \delta \vec{g}_k +  \boldsymbol{\alpha} \delta \vec{u}_k
-  - \boldsymbol{\beta} \delta \Delta \vec{u}_k + \vec{u}_k^t
+  \vec{u}_k^{t+1} &= \vec{u}_k^{t} + \delta \, \Big( \vec{g}_k^{t+1} + (\boldsymbol{\alpha} - \boldsymbol{\beta} \Delta) \, \vec{u}_k^{t+1} \Big) \notag \\
+  (1 - \delta \boldsymbol{\alpha} - \delta \boldsymbol{\beta} \Delta)\, \vec{u}_k^{t+1} &= \vec{u}_k^{t} + \delta\, \vec{g}_k^{t}
   \label{eq:discrete_derivative}
 \end{align}
+
+This expression is easily translated into Fourier domain as follows:
+
+  \begin{align}
+  \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\delta^{-1} \, \vec{u}_k^t + \vec{g}_k\} }%
+                  {\mathcal{F}\{(\delta^{-1} - \boldsymbol{\alpha})\, I-\boldsymbol{\beta}\Delta\}} \right\},
+  \label{eq:update_equation}
+  \end{align}
+  where $I$ denotes the identity operator.
+Here, we rewrite the Laplacian as a linear combination of the identity and shift operators:
+  \begin{equation}
+  \Delta = \{\sum_{d=1}^n \mathcal{S}_d^- + \mathcal{S}_d^+\}  - 2n \mathcal{I}
+  \end{equation}
+  where $\mathcal{S}_d^{\pm}$ stands for the forward ($+$) and backward ($-$) shift operator along axis $d$, of which the Fourier transform is found easily as
+  \begin{equation}
+  \mathcal{F}\{\mathcal{S}_d^{\pm}\} = e^{\pm i\omega_d},
+  \end{equation}
+  where $\omega_d$ is the normalized pulsation along direction $d$.
+Accordingly, the Fourier transform of the discrete Laplacian is found as
+  \begin{align}
+  \mathcal{F}\{\Delta\} &= \sum_d e^{-i\omega_d} + e^{i\omega_d} - 2n = n \, \left( \sum_d \cos(\omega_d) - 2\, \right)
+  \end{align}
+
+The remaining transforms are trivial or can be computed using FFT as in \citep{estellers_efficient_2012}.
+
+%Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
+%  $\psi$ by the actual kernel function, the transformation writes:
+%
+%  \begin{equation}
+%    \vec{v}_i' = \vec{v}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod}
+%      \beta_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
+%  \label{eq:transformation}
+%  \end{equation}
+%%
+%  with $\hat{\mathbf{e}}_d$ being the unitary vector along axis $d$.
 
@@
 \subsection{Implementation}
-\paragraph{General} The \emph{regseg} registration and segmentation tool is written in C++, using ITK-4.6 as
+\subsection{General} The \emph{regseg} registration and segmentation tool is written in C++, using ITK-4.6 as
   implementation core.
@@
 
-\paragraph{Efficient interpolation using sparse matrices}
+\subsection{Efficient interpolation using sparse matrices}
 During the registration process, every iteration requires computing the product of all the gradients
@@
 
-\paragraph{Preprocessing}
+\subsection{Preprocessing}
 
-\paragraph{Distortion}
+\subsection{Distortion}
 
@@
 \input{suppl-realdata.tex}
+
+\bibliographystyle{mystyle}
+\bibliography{Remote}
 

### 2015-03-27 Publications/2015-NeuroImage/03-introduction.tex
reworking intro

@@
 \section{Introduction}\label{sec:introduction}
-
-\Gls*{dmri} data of the brain present severe distortions, directly related to the
-  effort in minimizing acquisition time.
-In order to quickly probe the diffusion process in sufficiently different orientations,
-  \gls*{dmri} are acquired with \gls*{epi} sequences, which minimize the bandwidth of the \gls*{pe}
-  direction.
-This limitation originates a warping of the reconstructed image in regions affected by
-  small deviations from the main magnetic field $B_0$.
-One important source of these deviations is the discontinuity of magnetic susceptibility across
-  the imaged object.
-In fact, the distortion is the effect of the underlaying misregistration between correct and
-  warped data.
-Tissue/air boundaries show large steps of susceptibility and this artifact is
-  particularly evident in the surroundings of the orbito-frontal lobe and the temporal
-  bone of both hemispheres.
-Several approaches have been proposed to estimate the distortion and correct
-  for the artifact.
-The first solution consisted on estimating the deviations from $B_0$ with field mapping
-  techniques \citep{andersson_modeling_2001}.
-A second family of methods require an extra \gls*{epi} acquisition, but switching
-  the \gls*{pe} axis \citep{chiou_simple_2000} or reversing the direction of gradient \gls*{pe}
-  increments \citep{cordes_geometric_2000,holland_efficient_2010}.
-\cite{kybic_unwarping_2000} proposed nonlinear registration of the \emph{b0}
-  image of the \gls*{dmri} to an undistorted \gls*{t2} image of
-   the same subject.
-
+The accurate delineation of \gls*{wm} in \gls*{dmri} and the fusion of prior
+  anatomical information extracted from a \gls*{t1} image of the same subject
+  are crucial in a range of applications based on tractography, such as
+  the extraction of structural connectivity \citep{craddock_imaging_2013} or
+  tract-based spatial statistics \citep{smith_tractbased_2006}.
+However, segmenting \gls*{dmri} data precisely and coregistration of anatomical
+  images are difficult for several reasons.
+First, \gls{dmri} images suffer from partial voluming of several tissues in 
+  interfacing voxels due to the low resolution of images (generally around
+  \isores{2.2}mm).
+Second, \gls*{dmri} schemes probe the diffusion process within the brain in 
+  many angles called \glspl*{dwi}, completed by one or more baseline (\emph{b0}) 
+  volumes without directional gradients.
+The extremely low \gls*{snr} and high dimensionality of \glspl*{dwi} disable their
+  use for direct segmentation.
+The contrast between \gls*{gm} and \gls*{wm} in the \emph{b0} volumes is not suitable for 
+  segmentation either.
+Finally, \gls*{dmri} images are acquired using \gls*{epi} to speed up acquisition
+  at the cost of presenting comprehensively reported geometrical distortion and 
+  signal losses \citep{jezzard_correction_1995} that hinder the segmentation by
+  registering an anatomical image as source of same-subject prior information.
 The artifact has been proven to cause an important impact on the anatomy extracted
@@
   when used in whole-brain tractography.
-Currently, a broad range of applications make use of whole-brain tractography, such as
-  connectome extraction \citep{craddock_imaging_2013} or tract-based spatial statistics
-  \citep{smith_tractbased_2006}.
-These applications generally require the segmentation of the \gls*{dmri} data, also
-  hindered by distortions.
-One way to identify the regions of interest is extracting them from an appropriate 
-  \gls*{t1} of the same subject and map them into diffusion space,
-  in a sort of atlas-based segmentation.
 
-\paragraph{State of the art}\label{sec:state_of_art}
-  Active contours have been successfully integrated in image registration methods
-  for applications such as longitudinal object tracking \citep{paragios_level_2003} and
-  atlas-based segmentation \citep{gorthi_active_2011}.
-\cite{bertalmio_morphing_2000} proposed morphing active-contours for object tracking in 
-  series of 2D images.
-Shortly, \cite{yezzi_variational_2003} presented the first method including a full solution to
-  the registration problem with an affine transformation supporting the mapping.
-\cite{vemuri_joint_2003} proposed an atlas-based registration framework using level sets and only
-  one \gls*{pde} for first time.
+Early attempts to delineate the \gls*{wm} proceeded by thresholding the 
+  \gls*{fa}\footnote{the \gls*{fa} is a scalar parameter of diffusion derived from
+  the \gls*{dmri} data.} map at values around 0.7.
+The approach is unreliable for the high variability of \gls*{fa} and its
+  decay with the increase of incoherence between fibers.
+\cite{zhukov_level_2003} proposed active contours with edges represented
+  by level sets, evolving on directionally invariant scalar maps.
+\cite{rousson_level_2004} successfully segmented the corpus callosum with
+  region-based level-sets on the eigenvalues of tensors fitted on the
+  \gls*{dmri} data.
+A large body of work targets the definition of appropriate features to segmentation,
+  such as the 5D representation proposed by \cite{jonasson_segmentation_2005}.
+Other efforts include mixture models on sets of directionally invariant maps
+  \citep{liu_brain_2007}, iterative \citep{hadjiprocopis_unbiased_2005} and
+  hierarchical \citep{lu_segmentation_2008} clusterings,
+  graph-cuts \citep{han_experimental_2009}, and volume fraction modeling
+  \citep{kumazawa_improvement_2013}.
+
+\cite{saad_new_2009} used Pearson correlation to perform linear registration
+  of the \gls*{t1} image to the \emph{b0} of the \gls*{dmri}.
+Another linear registration method is \emph{bbregister} \citep{greve_accurate_2009},
+  that uses active contours extracted from the \gls*{t1}\footnote{using \emph{FreeSurfer}
+  \citep{fischl_freesurfer_2012}} to look for intensity boundaries in the \emph{b0}
+  image.
+In practice, the target \emph{b0} does not present clear frontiers of intensity between
+  interfacing tissues, what limits \emph{bbregister} to mapping the cortical surface only.
+Since the distortions found in \gls*{dmri} are nonlinear, \emph{bbregister} excludes
+  from the boundary search those regions typically warped by the artifact.
+This tool has become the standard method for its proven robustness.
+Nonlinear registration has been successfully performed between \gls*{t2} and \emph{b0}
+  images for their similarity, but uniquely as a way to correct for distortions
+  \citep{kybic_unwarping_2000,studholme_accurate_2000,wu_comparison_2008,tao_variational_2009}.
+Further registration of \gls*{t1} and \gls*{t2} images was still required to map the anatomical
+  information (and the \gls*{wm} segmentation) into \gls*{dmri} space.
+
+To jointly solve registration and segmentation problems such as longitudinal object
+  tracking \citep{paragios_level_2003} and atlas-based segmentation 
+  \citep{gorthi_active_2011}, active contours have been successfully integrated 
+  in image registration methods.
 \cite{unal_coupled_2005} and later \cite{wang_joint_2006},
-  extended the ``two \glspl*{pde}'' method of \cite{yezzi_variational_2003}
-  to nonlinear registration implementing a free deformation field.
+  extended an existing linear registration \cite{yezzi_variational_2003}
+  to the nonlinear case implementing a free deformation field.
 \cite{droske_mumfordshah_2009} reviewed the latter set of techniques, and proposed two different
@@
   of atlas-based segmentation.
-
-\cite{greve_accurate_2009} presented a relevant registration method called \emph{bbregister},
-  now included in \emph{FreeSurfer} \citep{fischl_freesurfer_2012}.
-The method is a standard choice to map surfaces extracted in the \gls*{t1} image to the \emph{b0}
-  image\footnote{The term \emph{b0} refers to one or more images within the
-  \gls*{dmri} dataset that are acquired with a null or very low gradient intensity or
-   \emph{b}-value.} of the \gls*{dmri} dataset with two limitations.
-First, the transform is affine and therefore it does not cope with the nonlinear distortions.
-Second, \emph{bbregister} looks for intensity steps in the image, using active contours with edges.
-In practice, the target \emph{b0} does not present clear frontiers of intensity between
-  interfacing tissues, what limits \emph{bbregister} to mapping the cortical surface only.
 

### 2015-03-27 Publications/2015-NeuroImage/03-introduction.tex
reworking intro and methods

@@
 Finally, \gls*{dmri} images are acquired using \gls*{epi} to speed up acquisition
-  at the cost of presenting comprehensively reported geometrical distortion and 
+  at the cost of presenting geometrical distortion and 
   signal losses \citep{jezzard_correction_1995} that hinder the segmentation by
@@
   hierarchical \citep{lu_segmentation_2008} clusterings,
-  graph-cuts \citep{han_experimental_2009}, and volume fraction modeling
-  \citep{kumazawa_improvement_2013}.
+  and graph-cuts \citep{han_experimental_2009}.
+  % , and volume fraction modeling \citep{kumazawa_improvement_2013}.
 
-\cite{saad_new_2009} used Pearson correlation to perform linear registration
-  of the \gls*{t1} image to the \emph{b0} of the \gls*{dmri}.
+To solve registration, \cite{saad_new_2009} used Pearson correlation to perform linear
+  alignment of \gls*{t1} and \emph{b0}.
 Another linear registration method is \emph{bbregister} \citep{greve_accurate_2009},
@@
 \cite{unal_coupled_2005} and later \cite{wang_joint_2006},
-  extended an existing linear registration \cite{yezzi_variational_2003}
+  extended an existing method using linear registration \citep{yezzi_variational_2003}
   to the nonlinear case implementing a free deformation field.
@@
 
-\paragraph{Contributions}\label{sec:contributions}
 The hypothesis under this research is that the registration and segmentation
@@
 
-While all the correction methods require an extra acquisition, \emph{regseg} is especially
-  indicated in scanning protocols without specific data for correction,
-  such as historical datasets.
+% While all the correction methods require an extra acquisition, \emph{regseg} is especially
+%   indicated in scanning protocols without specific data for correction,
+%   such as historical datasets.
 Finally, we present an instrumentation framework to evaluate correction and

### 2015-03-27 Publications/2015-NeuroImage/03-introduction.tex
Update references

@@
   hierarchical \citep{lu_segmentation_2008} clusterings,
-  and graph-cuts \citep{han_experimental_2009}.
-  % , and volume fraction modeling \citep{kumazawa_improvement_2013}.
+  and graph-cuts \citep{han_experimental_2009},
+  and volume fraction modeling \citep{kumazawa_improvement_2013}.
 

### 2015-03-27 Publications/2015-NeuroImage/03-introduction.tex
update intro

@@
 Another linear registration method is \emph{bbregister} \citep{greve_accurate_2009},
-  that uses active contours extracted from the \gls*{t1}\footnote{using \emph{FreeSurfer}
-  \citep{fischl_freesurfer_2012}} to look for intensity boundaries in the \emph{b0}
+  uses active contours with edges to look for intensity boundaries in the \emph{b0}
   image.
-In practice, the target \emph{b0} does not present clear frontiers of intensity between
-  interfacing tissues, what limits \emph{bbregister} to mapping the cortical surface only.
+The active contours are initialized by surfaces extracted from the 
+  \gls*{t1}\footnote{using \emph{FreeSurfer}\citep{fischl_freesurfer_2012}}:
+  the pial surface (exterior of cortical \gls*{gm}) and the \gls*{wm}/\gls*{gm} interface.
+In practice, the \emph{b0} image only presents a detectable frontier for the pial surface
+  and \emph{bbregister} is limited to align the cortical surface.
 Since the distortions found in \gls*{dmri} are nonlinear, \emph{bbregister} excludes
@@
 The hypothesis under this research is that the registration and segmentation
-  problems can be solved jointly, increasing the geometrical accuracy of the process.
-We propose \emph{regseg} as a joint solution to distortion correction and segmentation,
-  using active contours without edges \citep{chan_active_2001} that evolve driving a
+  problems in \gls*{dmri} can be solved simultaneously, increasing the geometrical
+  accuracy of the process.
+We propose a tool called \emph{regseg} to exploit the prior information of shapes
+  extracted on \gls*{t1} images to register the anatomical reference
+  to \gls*{dmri} space, implicitly segmenting these data.
+The approach uses active contours without edges \citep{chan_active_2001} that evolve driving a
   free-deformation field of B-Spline functions.
 Optimization is performed using a descent strategy of shape-gradients
-  \citep{herbulot_segmentation_2006,besson_dream2s_2003}, therefore \emph{regseg}
-  does not implement level sets as most of the presented methods.
-Given the properties of the distortion, \emph{regseg} includes the anisotropic regularization
-  of displacement fields of \cite{nagel_investigation_1986}.
-
-% While all the correction methods require an extra acquisition, \emph{regseg} is especially
-%   indicated in scanning protocols without specific data for correction,
-%   such as historical datasets.
+  \citep{besson_dream2s_2003,herbulot_segmentation_2006}, therefore \emph{regseg}
+  does not implement levelsets as most of the presented methods.
+The nonlinear warping caused by distortion is aligned with one of the imaging axis due
+  to the theory underpinning \gls*{epi} sequences.
+Therefore, \emph{regseg} includes an anisotropic regularizer of the displacement field 
+  proposed by \cite{nagel_investigation_1986}.
 Finally, we present an instrumentation framework to evaluate correction and

### 2015-04-03 Publications/2015-NeuroImage/00-main.tex
Updates after Yasser and Dominique
Removing formulas from paper
revising paper with new focus

@@
 \providecommand{\mdist}[2]{ \mathcal{D}_{#2}^2(\mathbf{#1}) }
+\providecommand{\omegaset}{\ensuremath{\boldsymbol{\Omega}}}
+\providecommand{\gammaset}{\ensuremath{\boldsymbol{\Gamma}}}
+
 \let\oldhat\hat
 \renewcommand{\vec}[1]{\mathbf{#1}}
+
 \providecommand{\nvec}[1]{\hat{\mathbf{#1}}}
 \providecommand{\suppl}[1]{\href{http://dx.doi.org/mydoi}{Supplemental Materials}, #1}
+\providecommand{\isores}[1]{#1\ensuremath{\times}#1\ensuremath{\times}#1}
 
@@
 
-\author[bit]{Oscar~Esteban\corref{corrauthor}}
+\author[bit,ciber]{Oscar~Esteban\corref{corrauthor}}
 \cortext[corrauthor]{Corresponding author}
@@
 \author[chuv,lts5]{Meritxell~Bach-Cuadra}
-\author[bit]{Mar\'ia-J.~Ledesma-Carbayo}
+\author[bit,ciber]{Mar\'ia-J.~Ledesma-Carbayo}
 \author[lts5]{Jean-Philippe~Thiran}
-\author[bit]{Andr\'es~Santos}
+\author[bit,ciber]{Andres~Santos}
 
 \address[bit]{Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on, %
-Universidad Polit\'ecnica de Madrid and CIBER-BBN, Madrid, Spain}
+Universidad Polit\'ecnica de Madrid, Madrid, Spain}
+\address[ciber]{Centro de Investigaci\'on Biom\'edica en Red en Bioingenier\'ia, Biomateriales y Nanomedicina (CIBER-BBN), Zaragoza, Spain}
 \address[ucla]{Department of Mathematics, University of California,
@@
 \appendix
-%\numberwithin{equation}{section}
+\numberwithin{equation}{section}
 \input{08-appendix-00-shapepriors}

### 2015-04-03 Publications/2015-NeuroImage/03-introduction.tex
revising paper with new focus

@@
   interfacing voxels due to the low resolution of images (generally around
-  \isores{2.2}mm).
+  \isores{2.2} [mm]).
 Second, \gls*{dmri} schemes probe the diffusion process within the brain in 
@@
   \gls*{fa}\footnote{the \gls*{fa} is a scalar parameter of diffusion derived from
-  the \gls*{dmri} data.} map at values around 0.7.
+  the \gls*{dmri} data.} map.
+However, there is a large variability in the mask obtained depending on the value of
+  choice \citep{taoka_fractional_2009}.
 The approach is unreliable for the high variability of \gls*{fa} and its
@@
 The active contours are initialized by surfaces extracted from the 
-  \gls*{t1}\footnote{using \emph{FreeSurfer}\citep{fischl_freesurfer_2012}}:
+  \gls*{t1} using \emph{FreeSurfer} \citep{fischl_freesurfer_2012}:
   the pial surface (exterior of cortical \gls*{gm}) and the \gls*{wm}/\gls*{gm} interface.

### 2015-04-03 Publications/2015-NeuroImage/03-introduction.tex
refactoring, add figure of phantoms

@@
   interfacing voxels due to the low resolution of images (generally around
-  \isores{2.2} [mm]).
+  \isores[mm]{2.2}).
 Second, \gls*{dmri} schemes probe the diffusion process within the brain in 
@@
   the \gls*{dmri} data.} map.
-However, there is a large variability in the mask obtained depending on the value of
-  choice \citep{taoka_fractional_2009}.
-The approach is unreliable for the high variability of \gls*{fa} and its
+However, the mask and subsequent analyses strongly depend on the threshold set
+  \citep{taoka_fractional_2009} owing to the large variability of \gls*{fa} and its
   decay with the increase of incoherence between fibers.

### 2015-04-03 Publications/2015-NeuroImage/04-methods.tex
reworking intro and methods
refactoring, add figure of phantoms

@@
   U_s\colon R \subset \mathbb{R}^n &\to M \subset \mathbb{R}^n \notag\\
-  \vec{r} &\mapsto \vec{r}' =\vec{r}+u(\vec{r}),
+  \vec{r} &\mapsto \vec{r}' =\vec{r}+\vec{u}(\vec{r}),
   \label{eq:transform}
@@
   its corresponding location in $M$, and $n$ the dimensionality of images.
-Finally, $\vec{u} = u(\vec{r})$ is the displacement of every point with respect
+Finally, $\vec{u} = \vec{u}(\vec{r})$ is the displacement of every point with respect
   to the reference domain.
-The displacements along the \gls*{pe} axis are related to the local field
-  deviation $\Delta B_0(\vec{r})$ \citep{jezzard_correction_1995}:
-
-  \begin{equation}
-  u(\vec{r}) = \frac{\gamma \, T_{acq}\, s_{PE}}{2\pi}\Delta B_0(\vec{r}) \: \hat{\vec{e}}_{PE},
-  \label{eq:fieldmap}
-  \end{equation}
-%
-where $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time,
-  $s_{PE}$ is the pixel size along \gls*{pe}, and $\hat{\vec{e}}_{PE}$ the unitary
-  vector along \gls*{pe}.
-
-% Joint segmentation and distortion correction of \gls*{dmri} data is also interesting to
-%   improve the overall consistence of complex processing workflows.
-% The closest antecedent of \emph{regseg} is \emph{bbregister} \citep{greve_accurate_2009}.
-% Both methods use active contours to align surfaces to same-subject \gls*{dmri} images.
-% However, \emph{bbregister} is limited to affine transformations, and thus, it is not able to
-%   solve the susceptibility-derived distortions, generally nonlinear.
-% To enable registration of uncorrected data, \emph{bbregister} excludes from the cost-function
-%   those regions frequently warped.
-% Conversely, \emph{regseg} implements a free-deformation field, supported with B-Spline
-%   functions to estimate deformations.
-% A second diverging foundation is the active-contours framework.
-% In \emph{regseg}, active contours are implemented without edges \citep{chan_active_2001},
-%   enabling the tool to segment homogeneous regions within the brain.
-% On the other hand, \emph{bbregister} requires that the target features present intensity
-%   boundaries as it implements active contours with edges.
-% The tissue interfaces imaged with \gls*{dmri} are affected by partial volume effects that
-%   blur boundaries.
-% In practice, registration of \gls*{t1} and \gls*{dmri} data with \emph{bbregister}
-%   is limited to the linear match of the pial surface of the brain.
-
-% Our method is close to \citep{guyader_combined_2011} and \citep{gorthi_active_2011}
-%   in the use of active contours without edges for registration, however we avoided
-%   the use of level sets with the application of shape gradients
-%   \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
-% \cite{gorthi_active_2011} used a gaussian regularizer, and \cite{guyader_combined_2011}
-%   proposed a nonlinear elasticity smoother.
-% In contrast, \emph{regseg} implements the anisotropic regularization proposed by
-%   \citep{nagel_investigation_1986}.
-% We demonstrate its aptness on the correction of the nonlinear distortions typically
-%   presented by \gls*{dmri} data of the brain.
-% Therefore, a typical protocol for the extraction of structural connectivity
-%   comprehends \gls*{dmri}, \gls*{t1}, and the acquisition determined by the correction
-%   method of choice \citep{daducci_connectome_2012}.
 
@@
 
-\begin{figure*}
-\includegraphics[width=\linewidth]{figures/figure02}
-\caption{Experimental workflow applied on real data from the \acrfull*{hcp}.
-  1) The prior surfaces are extracted from the anatomical reference (\gls*{t1} image).
-	2) To operate as ground truth, we generate a plausible-synthetic distortion $U_{true}$
-	  from the fieldmap using \eqref{eq:fieldmap}.
-	3) The \gls*{dmri} data are warped using $U^{-1}_{true}$ to reproduce the effects of real
-	  susceptibility-derived distortions.
-	Target diffusion scalars (\gls*{fa} and \gls*{adc}) are computed on the distorted data and
-		stacked to feed the multivariate input required by our algorithm.
-	4) \emph{Regseg} is run, obtaining a $U_{test} = \hat{U}_{true}$, the estimation of
-	  the ground-truth deformation.
-	5) Results are visually and quantitatively evaluated.}\label{fig:evworkflows}
-\end{figure*}
-
 \subsection{Image Data and preprocessing}
@@
   (1.0mm and 2.0mm isotropic).
-Field of view at both resolutions was \isores{100}mm.
+The field of view at both resolutions was \isores[mm]{100}.
 Simulations were corrupted with rician noise for a \gls*{snr} of 300.0.
@@
 \begin{figure*}
-    \centering
-    \resizebox{\textwidth}{!}{%
-      \begin{tikzpicture}
-        \node[inner sep=10pt, draw=black!40](fig03a) at (0,14.8)
-          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure03-a}};
-        \node[inner sep=10pt, draw=black!40](fig03b) at (0,7.9)
-          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure03-b}};
-        \node[inner sep=10pt, draw=black!40](fig03c) at (0,0)
-          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure03-c}};
-        \node[circle, text=black!75] at (-9.4,17.3) {\Huge \textbf{A}};
-        \node[circle, text=black!75] at (-9.4,10.6) {\Huge \textbf{B}};
-        \node[circle, text=black!75] at (-9.4,3.5) {\Huge \textbf{C}};
-      \end{tikzpicture}
-    }%
-	\caption{A. The ``cortex'' phantom is a spherical shape with two sulci and an
-	  outer crust resembling the cortical folding (left).
-	The model is used to generate \gls*{t1} and \gls*{t2} images after warping the
-	  contours using a random and plausible transformation $U_{true}^{-1}$ (right).
-	B. Visual assessment of the results on the low resolution sets:
-	  ``gyrus'' (top-left), ``L-shape'' (top-right), ``ball'' (bottom-left),
-	  and ``box'' at (bottom-right).
-	In yellow color, the recovered contours after registration are represented.
-	Our method showed high accuracy, as demonstrates almost exact location of the registered
-    contours with respect to their ground truth position depicted in green.
-	Partial volume effect turns segmentation of the sulci a challenging problem with voxel-wise
-	  clustering methods, but it is successfully segmented with our method.
-	C. Quantitative evaluation of registration error in terms of average Hausdorff distance of
-	  surfaces at high (left) and low (right) resolutions, demonstrating that the error is
-	  consistently below the voxel size.
-	  }\label{fig:phantom}
+\includegraphics[width=\linewidth]{figures/figure03}
+\caption{Evaluation workflow of phantoms}\label{fig:evphantoms}
 \end{figure*}
@@
 
+\begin{figure*}
+\includegraphics[width=\linewidth]{figures/figure02}
+\caption{Experimental workflow applied on real data from the \acrfull*{hcp}.
+  1) The prior surfaces are extracted from the anatomical reference (\gls*{t1} image).
+  2) To operate as ground truth, we generate a plausible-synthetic distortion $U_{true}$
+    from the fieldmap using \eqref{eq:fieldmap}.
+  3) The \gls*{dmri} data are warped using $U^{-1}_{true}$ to reproduce the effects of real
+    susceptibility-derived distortions.
+  Target diffusion scalars (\gls*{fa} and \gls*{adc}) are computed on the distorted data and
+    stacked to feed the multivariate input required by our algorithm.
+  4) \emph{Regseg} is run, obtaining a $U_{test} = \hat{U}_{true}$, the estimation of
+    the ground-truth deformation.
+  5) Results are visually and quantitatively evaluated.}\label{fig:evworkflows}
+\end{figure*}
+
 Selecting the appropriate labels in the \emph{aparc} segmentation, we applied
@@
 The choice of this particular model is further addressed in \autoref{sec:res_model_and_metric}.
+
+The displacements along the \gls*{pe} axis are related to the local field
+  deviation $\Delta B_0(\vec{r})$ \citep{jezzard_correction_1995}:
+
+  \begin{equation}
+  u(\vec{r}) = \frac{\gamma \, T_{acq}\, s_{PE}}{2\pi}\Delta B_0(\vec{r}) \: \hat{\vec{e}}_{PE},
+  \label{eq:fieldmap}
+  \end{equation}
+%
+where $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time,
+  $s_{PE}$ is the pixel size along \gls*{pe}, and $\hat{\vec{e}}_{PE}$ the unitary
+  vector along \gls*{pe}.
 

### 2015-04-03 Publications/2015-NeuroImage/05-results.tex
fix aggregate, update results
refactoring, add figure of phantoms

@@
 \label{sec:results_phantom}
+
+\begin{figure*}
+    \centering
+    \resizebox{\textwidth}{!}{%
+      \begin{tikzpicture}
+        \node[inner sep=10pt, draw=black!40](fig03b) at (0,7.9)
+          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure03-b}};
+        \node[inner sep=10pt, draw=black!40](fig03c) at (0,0)
+          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure03-c}};
+        \node[circle, text=black!75] at (-9.4,10.6) {\Huge \textbf{A}};
+        \node[circle, text=black!75] at (-9.4,3.5) {\Huge \textbf{B}};
+      \end{tikzpicture}
+    }%
+  \caption{A. Visual assessment of the results on the low resolution sets:
+    ``gyrus'' (top-left), ``L-shape'' (top-right), ``ball'' (bottom-left),
+    and ``box'' at (bottom-right).
+  In yellow color, the recovered contours after registration are represented.
+  Our method showed high accuracy, as demonstrates almost exact location of the registered
+    contours with respect to their ground truth position depicted in green.
+  Partial volume effect turns segmentation of the sulci a challenging problem with voxel-wise
+    clustering methods, but it is successfully segmented with our method.
+  B. Quantitative evaluation of registration error in terms of average Hausdorff distance of
+    surfaces at high (left) and low (right) resolutions, demonstrating that the error is
+    consistently below the voxel size.
+    }\label{fig:phantom}
+\end{figure*}
+
 Using a simplified subset of the experimental instrumentation described in
@@
 
-\begin{figure}
-	\includegraphics[width=\linewidth]{figures/figure04.pdf}
-	\caption{Assessment of the segmentation model, sampling the value of the energy functional
-	\eqref{eq:energy} versus several registration errors generated as explained in
-	\autoref{sec:res_model_and_metric}.
-	The metric of our registration method displayed a smooth gradient towards the minimum
-	located at the $\epsilon = 0.0$ point.}\label{fig:energymap}
-\end{figure}
-
 \paragraph*{Evaluation \& cross-comparison}\label{sec:res_cc_evaluation}
@@
 We also reported the 95\% CIs of the \gls*{swindex} for those surfaces.
-The aggregated CI for \emph{regseg} was $[1.08 - 1.50] mm$, whereas the competing method
-  yielded an aggregated CI of $[2.06 - 2.43] mm$.
+The aggregate CI for \emph{regseg} was $[1.08 - 1.50] mm$, whereas the competing method
+  yielded an aggregate CI of $[2.06 - 2.43] mm$.
 The quantitative results of the ANOVA tests and the CIs are presented in \autoref{tab:results_real}.
@@
     \begin{tabular}{cccccc}
-    & & $\Gamma_{VdGM}$  & $\Gamma_{WM}$ & $\Gamma_{pial}$ & Aggregated \\
+    & & $\Gamma_{VdGM}$  & $\Gamma_{WM}$ & $\Gamma_{pial}$ & Aggregate \\
     \hline
     \multirow{2}{*}{CI}
-       & \emph{regseg}        & 0.77 - 1.18 & 0.95 - 1.85 & 1.33 - 2.13 & 1.08 - 1.50 \\
-       & T2B                  & 1.73 - 2.62 & 1.96 - 2.40 & 2.16 - 2.63 & 2.06 - 2.43 \\
+       & \emph{regseg}        & 0.50 - 0.78 & 0.50 - 0.55 & 0.66 - 0.73 & 0.56 - 0.66 \\
+       & T2B                  & 1.78 - 2.58 & 1.94 - 2.36 & 2.16 - 2.58 & 2.05 - 2.39 \\
     \hline
     \multirow{2}{*}{ANOVA}
-       & p-value  & 1.4$\times$10$^{-5}$& 1.3$\times$10$^{-3}$& 2.5$\times$10$^{-3}$ & 2$\times$10$^{-9}$ \\
-       & f-stat   & 27.87               & 12.89               & 11.13                & 45.04 \\
+       & p-value  & 1.5$\times$10$^{-8}$& 6.3$\times$10$^{-16}$& 4.2$\times$10$^{-16}$ & 4$\times$10$^{-34}$ \\
+       & f-stat   & 60.03               & 255.82               & 263.97                & 377.82              \\
     \hline

### 2015-04-03 Publications/2015-NeuroImage/supplemental.tex
refactoring, add figure of phantoms

@@
 
+\subsection{Assessment of the segmentation model}
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/figure04.pdf}
+  \caption{Assessment of the segmentation model, sampling the value of the energy functional
+  \eqref{eq:energy} versus several registration errors generated as explained in
+  \autoref{sec:res_model_and_metric}.
+  The metric of our registration method displayed a smooth gradient towards the minimum
+  located at the $\epsilon = 0.0$ point.}\label{fig:energymap}
+\end{figure}
+
 \subsection{Preprocessing}

### 2015-04-04 Publications/2015-NeuroImage/03-introduction.tex
my changes

@@
   use for direct segmentation.
-The contrast between \gls*{gm} and \gls*{wm} in the \emph{b0} volumes is not suitable for 
+The low contrast between \gls*{gm} and \gls*{wm} in the \emph{b0} volumes is not suitable for 
   segmentation either.
-Finally, \gls*{dmri} images are acquired using \gls*{epi} to speed up acquisition
+Lastly, \gls*{dmri} images are acquired using \gls*{epi} to speed up acquisition
   at the cost of presenting geometrical distortion and 
@@
 
-To solve registration, \cite{saad_new_2009} used Pearson correlation to perform linear
-  alignment of \gls*{t1} and \emph{b0}.
-Another linear registration method is \emph{bbregister} \citep{greve_accurate_2009},
-  uses active contours with edges to look for intensity boundaries in the \emph{b0}
+To solve the segmentation problem through registration, \cite{saad_new_2009} 
+  used Pearson correlation to perform linear alignment of the \gls*{t1} and 
+  the \emph{b0}.
+Similarly, an also linear registration method is \emph{bbregister} \citep{greve_accurate_2009}.
+It uses active contours with edges to look for intensity boundaries in the \emph{b0}
   image.
@@
   \gls*{t1} using \emph{FreeSurfer} \citep{fischl_freesurfer_2012}:
-  the pial surface (exterior of cortical \gls*{gm}) and the \gls*{wm}/\gls*{gm} interface.
-In practice, the \emph{b0} image only presents a detectable frontier for the pial surface
-  and \emph{bbregister} is limited to align the cortical surface.
-Since the distortions found in \gls*{dmri} are nonlinear, \emph{bbregister} excludes
-  from the boundary search those regions typically warped by the artifact.
-This tool has become the standard method for its proven robustness.
+  the \emph{pial} surface (exterior of cortical \gls*{gm}) and the \emph{white}
+  surface (the \gls*{wm}/\gls*{gm} interface). 
+Since the \emph{b0} image only presents a detectable frontier for the pial surface,
+  \emph{bbregister} is limited to align the cortical layer in this
+  application.
+This tool has become the standard method for its proven robustness, despite the 
+  distortions found in \gls*{dmri} are nonlinear.
+To overcome this issue, \emph{bbregister} excludes from the boundary search those
+  regions typically warped by the artifact.
+As the distortion is not accounted for, it should be addressed separately.
 Nonlinear registration has been successfully performed between \gls*{t2} and \emph{b0}
@@
 
-To jointly solve registration and segmentation problems such as longitudinal object
-  tracking \citep{paragios_level_2003} and atlas-based segmentation 
-  \citep{gorthi_active_2011}, active contours have been successfully integrated 
-  in image registration methods.
+Therefore, given the described circumstances, a method performing simultaneously
+  segmentation in \gls*{dmri} space and registration of the corresponding \gls*{t1} image
+  to \gls*{dmri} data could provide the optimal solution.
+Previously, joint segmentation and registration has been successfully applied to
+  longitudinal object tracking \citep{paragios_level_2003} and atlas-based
+  segmentation \citep{gorthi_active_2011}.
+The most extended approach is integrating active contours in image registration 
+  methods.
 \cite{unal_coupled_2005} and later \cite{wang_joint_2006},
-  extended an existing method using linear registration \citep{yezzi_variational_2003}
-  to the nonlinear case implementing a free deformation field.
+  improved an existing method \citep{yezzi_variational_2003} based on linear registration
+  to the nonlinear case implementing a free-form deformation field.
 \cite{droske_mumfordshah_2009} reviewed the latter set of techniques, and proposed two different
@@
   the contours to the whole image definition.
-Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
+Recently, \cite{guyader_combined_2011} proposed a 2D simultaneous segmentation and
   registration method using level sets and a nonlinear elasticity smoother on the
@@
   free-deformation field of B-Spline functions.
-Optimization is performed using a descent strategy of shape-gradients
+Optimization is performed using a descent strategy of explicit shape-gradients
   \citep{besson_dream2s_2003,herbulot_segmentation_2006}, therefore \emph{regseg}

### 2015-04-04 Publications/2015-NeuroImage/04-methods.tex
my changes

@@
   extracted from the undistorted \gls*{t1} image (the reference space $R$).
-Distortion can be rendered as a registration problem where we look for an underlying deformation 
-  field $U_s$, mapping $R$ (with $\gammaset \subset R$) to the corresponding features in the distorted 
-  \gls*{dmri} images (the moving space $M$):
+We reformulate the segmentation of the distorted \gls*{dmri} images (the moving space $M$)
+  as a registration problem where we look for an underlying deformation field $U$, such that 
+  the structures in $R$ defined by \gammaset{} optimally align with their corresponding
+  structures in $M$:
 
   \begin{align}
-  U_s\colon R \subset \mathbb{R}^n &\to M \subset \mathbb{R}^n \notag\\
+  U\colon R \subset \mathbb{R}^n &\to M \subset \mathbb{R}^n \notag\\
   \vec{r} &\mapsto \vec{r}' =\vec{r}+\vec{u}(\vec{r}),
@@
 The best estimate $\hat{U}$ then fulfills the maximum a posteriori criterium
-  \citep{bishop_pattern_2006}, and aligns \gammaset into $M$.
+  \citep{bishop_pattern_2006}, and aligns \gammaset{} into $M$.
 First, we assume independence between pixels, and thus break down the
@@
 \paragraph*{Deformation model}\label{sec:deformation_model}
-Since the vertices of the surfaces $\{\vec{v}_i: \vec{v}_i \subset \gammaset \}_{i=1 \ldots V}$
+Since the vertices of the surfaces $\{\vec{v}_i: \vec{v}_i \subset \gammaset \}_{i=1 \ldots N_v}$
   likely lay off-grid, it is required to derive $u(\vec{v}_i)$ from a discrete set of parameters
@@
   derivatives.
-The intermediate steps to get to \eqref{eq:gradient_descent} from \eqref{eq:final_gradient} are
+The computation of the explicit shape gradients at each $\vec{v}_i'$ is described in \autoref{fig:method}.
+The intermediate steps to get to \eqref{eq:final_gradient} from \eqref{eq:gradient_descent} are
   detailed in the \suppl{equations SM6-SM12}:
@@
   \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
-  \vec{g}_k  + \boldsymbol{\alpha} \cdot \vec{u}_k + \boldsymbol{\beta} \cdot (\Delta \vec{u}_k).
+  \vec{g}_k  + \boldsymbol{\alpha} \cdot \vec{u}_k - \boldsymbol{\beta} \cdot (\Delta \vec{u}_k).
   \label{eq:final_gradient}
@@
   \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\delta^{-1} \, \vec{u}_k^t + \vec{g}_k\} }%
-                  {\mathcal{F}\{(\delta^{-1} - \boldsymbol{\alpha})\, I-\boldsymbol{\beta}\Delta\}} \right\}
+                  {\mathcal{F}\{(\delta^{-1} + \boldsymbol{\alpha})\, I - \boldsymbol{\beta}\Delta\}} \right\}
   \label{eq:update_equation}
@@
 
-\paragraph*{Settings, implementation details, and convergence}
+\paragraph*{Implementation details, settings and convergence}
 \label{sec:conv_report}
-The registration parameters (such as $\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$,
-  the B-Spline grid resolutions, target image smoothing, etc.)
-  and other implementation details (for instance the sparse matrix approach
-  to fast interpolation) are discussed in the \suppl{section S1}.
+The \emph{regseg} tool includes a multiresolution strategy on the free-form deformation field.
+Registration pyramids are created by setting the spacing between control points of the B-Spline basis
+  functions for each level of the multiresolution.
+This implementation detail, along with other features (as the sparse matrix approach
+  to fast interpolation) and the main parameters 
+  (such as $\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, the B-Spline grid resolutions,
+  target image smoothing, etc.) are discussed in the \suppl{section S1}.
 The actual choices of parameter settings are publicly distributed with the source code of the experiments.
-Additionally, we release our software along with a tool for generating convergence reports to
-  demonstrate the behavior of \emph{regseg} and help scientists configure their own experiments.
-One sample report is found in the \suppl{section S1.3}.
+These final settings were manually obtained based on the feedback of post-registration convergence
+  reports (as the one found in the \suppl{section S1.3}).
+We release \emph{regseg} along with the tool for generating the convergence report.
 
@@
 In order to demonstrate the performance of \emph{regseg}, we first conducted a battery of
-  accuracy tests over synthetic data, and then applied it in the susceptibility distortion
-  correction of real data.
+  accuracy tests over synthetic data, and then applied to real \gls*{dmri} data.
 \autoref{fig:evworkflows} presents the workflow implementing the evaluation instruments.
@@
 	6) cortical \gls*{gm} surface, $\Gamma_{pial}$.
-The choice of this particular model is further addressed in \autoref{sec:res_model_and_metric}.
+The segmentation model corresponding to this selection of surfaces is presented in
+  \autoref{fig:model}, and further addressed in \autoref{sec:res_model_and_metric}.
 
-The displacements along the \gls*{pe} axis are related to the local field
-  deviation $\Delta B_0(\vec{r})$ \citep{jezzard_correction_1995}:
+The displacements along the \gls*{pe} axis are related to the local deviation of the
+  field $\Delta B_0(\vec{r})$ \citep{jezzard_correction_1995}:
 
@@
   vector along \gls*{pe}.
+With certain \gls*{mr} sequences it is possible to estimate $\Delta B_0$, obtaining
+  the so-called fieldmap.
 
-In this case, we derived the deformation $U_{true}$ from the field maps released with
+In this case, we derived the deformation $U_{true}$ from the fieldmaps released with
   the corresponding packages of each dataset from the \gls*{hcp} applying \eqref{eq:fieldmap},

### 2015-04-04 Publications/2015-NeuroImage/05-results.tex
my changes

@@
   In yellow color, the recovered contours after registration are represented.
-  Our method showed high accuracy, as demonstrates almost exact location of the registered
+  Our method showed high accuracy, as it demonstrates the almost exact location of the registered
     contours with respect to their ground truth position depicted in green.
@@
   B. Quantitative evaluation of registration error in terms of average Hausdorff distance of
-    surfaces at high (left) and low (right) resolutions, demonstrating that the error is
+    surfaces at low (left) and high (right) resolutions, demonstrating that the error is
     consistently below the voxel size.

### 2015-04-05 Publications/2015-NeuroImage/03-introduction.tex
my fixes

@@
   the \gls*{dmri} data.} map.
-However, the mask and subsequent analyses strongly depend on the threshold set
+However, the mask and subsequent analyses substantially depend on the threshold set
   \citep{taoka_fractional_2009} owing to the large variability of \gls*{fa} and its
@@
 \cite{zhukov_level_2003} proposed active contours with edges represented
-  by level sets, evolving on directionally invariant scalar maps.
+  by levelsets, evolving on directionally invariant scalar maps.
 \cite{rousson_level_2004} successfully segmented the corpus callosum with
-  region-based level-sets on the eigenvalues of tensors fitted on the
+  region-based levelsets on the eigenvalues of tensors fitted on the
   \gls*{dmri} data.
-A large body of work targets the definition of appropriate features to segmentation,
+A large body of work targeted the definition of appropriate features to cluster,
   such as the 5D representation proposed by \cite{jonasson_segmentation_2005}.
@@
   segmentation in \gls*{dmri} space and registration of the corresponding \gls*{t1} image
-  to \gls*{dmri} data could provide the optimal solution.
+  to the \gls*{dmri} data could provide the optimal solution.
 Previously, joint segmentation and registration has been successfully applied to
@@
   to the nonlinear case implementing a free-form deformation field.
-\cite{droske_mumfordshah_2009} reviewed the latter set of techniques, and proposed two different
+\cite{droske_mumfordshah_2009} reviewed the existing techniques, and proposed two different
   approaches to apply the Mumford-Shah functional \citep{mumford_optimal_1989} in simultaneous
@@
   the contours to the whole image definition.
-Recently, \cite{guyader_combined_2011} proposed a 2D simultaneous segmentation and
-  registration method using level sets and a nonlinear elasticity smoother on the
+Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
+  registration method in 2D using levelsets and a nonlinear elasticity smoother on the
   displacement vector field, which preserves topology even with very large deformations.
 Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
-  level-set function for the registration of several active contours, in the application
+  levelset function for the registration of several active contours, in the application
   of atlas-based segmentation.
@@
 The approach uses active contours without edges \citep{chan_active_2001} that evolve driving a
-  free-deformation field of B-Spline functions.
+  free-form deformation field of B-Spline basis functions.
 Optimization is performed using a descent strategy of explicit shape-gradients
@@
   does not implement levelsets as most of the presented methods.
-The nonlinear warping caused by distortion is aligned with one of the imaging axis due
-  to the theory underpinning \gls*{epi} sequences.
-Therefore, \emph{regseg} includes an anisotropic regularizer of the displacement field 
-  proposed by \cite{nagel_investigation_1986}.
-Finally, we present an instrumentation framework to evaluate correction and
-  segmentation pipelines using realistic distortions as ground truth.
-With the integration of \emph{regseg} and an in-house implementation of the 
-  \gls*{t2b} correction in the framework, we compare the performance
-  of both solutions.
+Since the nonlinear distortion is aligned with one of the imaging axis (see 
+  \autoref{sec:datasets}), \emph{regseg} includes an anisotropic regularizer of
+  the displacements field proposed by \cite{nagel_investigation_1986}.
+Finally, we evaluate \emph{regseg} using an extension of our instrumentation framework
+  \citep{esteban_simulationbased_2014} that simulates known and realistic distortions
+  on \gls*{dmri} data.
+The evaluation includes a comparison to the \gls*{t2b} correction by integrating in the framework
+  an in-house implementation of the method.

### 2015-04-05 Publications/2015-NeuroImage/04-methods.tex
my fixes

@@
   $M$.
-Let $\omegaset = \{\Omega_l: l \in \mathbb{N}, l \leq L\}$ be the set of $L$ competing regions in
-  which $M$ is partitioned by the projection of \gammaset, 
-  the given surfaces serving as shape priors.
+Let $\omegaset = \{\Omega_l: l \in \mathbb{N}, l \leq N_L\}$ be the set of $N_L$ competing regions in
+  which $M$ is partitioned by the projection of \gammaset.
 Using the Bayes' rule, the posterior likelihood is computed as:
@@
   distributions with parameters $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
-  for each region $\Omega_l$.
-We then insert the normal distribution into \eqref{eq:bayes_rule} to obtain the full
-  formulation of the \emph{data term} of our registration model:
+  for each region $\Omega_l$:
 
@@
   P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
-  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) \notag\\
+  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) = \notag\\
   &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
@@
 Finally, the maximum a posteriori problem is adapted into a variational one where we look for
-  the minimum of an energy functional, by applying a log-transform:
+  the minimum of an energy functional, by applying $E(\vec{u}) = -\log P(U)$:
 
   \begin{align}
-  E(M \mid U) &= -\log \underset{l}{\prod}
+  E(\vec{u}) &= -\log \underset{l}{\prod}
   \underset{\vec{r} \in \Omega_l}{\prod}
@@
 %
-
 This expression is dual of the Mumford-Shah functional corresponding
-  to a discrete \acrlong*{acwe} framework \citep{chan_active_2001}
+  to a framework of \acrlong*{acwe} \citep{chan_active_2001}
   with the corresponding anisotropic regularization term of \cite{nagel_investigation_1986}.
@@
 Since the vertices of the surfaces $\{\vec{v}_i: \vec{v}_i \subset \gammaset \}_{i=1 \ldots N_v}$
-  likely lay off-grid, it is required to derive $u(\vec{v}_i)$ from a discrete set of parameters
+  likely lay off-grid, it is required to derive $\vec{u}_i = \vec{u}(\vec{v}_i)$ from a discrete set of parameters
   $\{\vec{u}_k\}_{k=1 \ldots K}$.
-Densification is achieved through a set of associated basis functions $\psi_k$.
+Densification is achieved through a set of associated basis functions $\psi_k$ \eqref{eq:nodes_tfm}.
 In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
-  of degree 3: % ($\beta_3$).
+  of degree 3.
 
   \begin{equation}
-  \vec{v}_i' = \vec{v}_i + u(\vec{v}_i) = \vec{v}_i + \sum_k \psi_k(\vec{r}) \: \vec{u}_k.
+  \vec{v}_i' = \vec{v}_i + \vec{u}_i = \vec{v}_i + \sum_k \psi_k(\vec{r}) \: \vec{u}_k.
   \label{eq:nodes_tfm}
@@
   \begin{equation}
-  \frac{\partial u(\vec{r},t)}{\partial t} \propto - \frac{\partial E(M \mid U)}{\partial \vec{u}_k},
+  \frac{\partial \vec{u}(\vec{r},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}_k},
   \label{eq:general_gradient_descent}
@@
   \eqref{eq:energy} into expression \eqref{eq:app_energy} of \ref{app:reg_term}
-  and compute the derivative \eqref{eq:general_gradient_descent}:
+  and compute its derivative \eqref{eq:general_gradient_descent}:
 
@@
 
-\subsection{Experiments and evaluation}
-\label{sec:experiments_evaluation}
-%
-In order to demonstrate the performance of \emph{regseg}, we first conducted a battery of
-  accuracy tests over synthetic data, and then applied to real \gls*{dmri} data.
-\autoref{fig:evworkflows} presents the workflow implementing the evaluation instruments.
-Besides a visual assessment of the results, we report quantitative evaluations using
-  two metrics.
-In the case of the phantoms, since we had produced distortions along the three
-  available dimensions, we computed the Haussdorf distance, by reusing the
-  ``point-to-cell'' method of \cite{commandeur_vtk_2011}.
-Conversely, the susceptibility-derived distortions only happen along the \gls*{pe}
-  axis of the image.
-Therefore, a \gls*{swindex} can be computed as the one-to-one distance between corresponding
-  vertices of surfaces, weighted by their respective Voronoi area $a_i$:
+\subsection{Data and corresponding evaluations}
+\label{sec:datasets}
+In order to assess the performance of \emph{regseg}, we defined a general
+  evaluation protocol:
+1) Extraction of the set of undistorted surfaces \gammaset{};  
+2) computation of a ground-truth field of displacements that is applied to
+  generate warped images ($M$) to segment;
+3) execution of \emph{regseg} with \gammaset{} and the warped data as inputs; and
+4) visual assessment and computation of error metrics.
 
-  \begin{equation}
-  sWI = \frac{1}{\sum_i a_i} \sum\limits_i^{N_p} a_i\,\|
-  \vec{v}_i - \hat{\vec{v}}_i \|
-  \label{eq:swindex}
-  \end{equation}
-%
-  where $\vec{v}_i$ are the locations of the $N_p$ vertices in each $p \in \{1, \dots P\}$
-  prior, and $\hat{\vec{v}}_i$ is the location
-  recovered corresponding to the vertex $\vec{v}_i$.
-
-\subsection{Image Data and preprocessing}
-\label{sec:datasets}
-
-\paragraph*{Simulated phantoms}%
-\label{sec:digital_phantoms}
-We proved the concept on simplistic digital phantoms which we will name after their
-  appearance as ``box'', ``ball'', ``L-shape'', and ``gyrus'' (\autoref{fig:phantom},
-  boxes A and B).
-Using \emph{phantomas} \citep{caruyer_phantomas_2014}, we simulated
+\paragraph*{Simulated phantoms}\label{sec:digital_phantoms}
+The workflow to generate the digital phantoms and assess
+  \emph{regseg} on them is presented in \autoref{fig:evphantoms}.
+A set of four binary objects (called ``box'', ``ball'', ``L'',
+  and ``gyrus'') are generated combining binarization of
+  analytical shapes and mathematical morphology.
+Using \emph{FreeSurfer} tools, the reference surfaces \gammaset{}
+  are generated.
+The ground-truth distortion is generated with a chain of two displacement 
+  fields supported by grids of B-Spline basis functions.
+The coefficients of the basis functions are generated randomly for
+  both levels in their three dimensions.
+The three components of the displacements $\vec{u} = (u_d)$ 
+  are bounded above by 40\% of the separation between control points
+  at each level, to obtain diffeomorphic transforms
+  after concatenation \citep{rueckert_diffeomorphic_2006}.
+The first deformation field is intended to generate large warpings,
+  with control points separated by 50.50mm along the three dimensions
+  ($u_d\leq$ 20.20mm).
+With the second warping, we look for a resulting field with smoothness
+  close to the found in a typical distortion field of \gls*{dmri} data
+  \citep{irfanoglu_susceptibility_2011}.
+Therefore, the control points are separated 25.25mm ($u_d\leq$ 10.10mm).
+Once the ground-truth deformation is generated, the original surfaces
+  are warped interpolating the displacements field at each vertex.
+The warped surfaces are binarized to generate a distorted version of
+  the model in two target resolutions: \isores[mm]{2.0} and \isores[mm]{1.0}.
+Using a \gls*{mr} simulator \citep{caruyer_phantomas_2014}, we synthesized
   \gls*{t1} (TE/TR=10/1500ms) and \gls*{t2} images (TE/TR=90/5000ms)
@@
 Simulations were corrupted with rician noise for a \gls*{snr} of 300.0.
-The reference surfaces were extracted at high resolution, using 
-  \texttt{mri\_tessellate}\footnote{A marching-cubes algorithm shipped with 
-  \emph{FreeSurfer}.}.
+
+To evaluate the error, we computed the Haussdorf distance (as implemented
+ by \cite{commandeur_vtk_2011}) between the surfaces warped with the ground-truth
+ and their corresponding contours estimated with \emph{regseg}.
 
@@
 \includegraphics[width=\linewidth]{figures/figure03}
-\caption{Evaluation workflow of phantoms}\label{fig:evphantoms}
+\caption{Evaluation workflow of phantoms.
+
+}\label{fig:evphantoms}
 \end{figure*}
-
-To warp the simulated datasets, we generated 150 different realizations of
-  the distortion $U_{true}$.
-Smoothness is introduced using two levels of B-Spline functions, with control points evenly
-  located in isotropic grids covering the full extent of the phantom.
-Separation between control points was isotropic, and set to 50.5mm for the first level
-  of distortion and 25.25mm for the second.
-The rationale behind this choice is producing large deformations on the contours with
-  the first level, and matching the properties of susceptibility-derived distortions
-  as previously reported by \cite{irfanoglu_susceptibility_2011} with the finer grid.
-The coefficients of the B-Spline functions were randomly generated for each resolution level.
-Invertibility of $U_{true}$ is ensured by controlling the maximum displacement at each
-  level (20.2mm and 10.1mm respectively) as studied in \citep{rueckert_diffeomorphic_2006}.
 
@@
 % 
+\autoref{fig:evworkflows}
 For the evaluation of the algorithm on real \gls*{dmri} data of human brains,
@@
   transform.
+
+Conversely, the susceptibility-derived distortions only happen along the \gls*{pe}
+  axis of the image.
+Therefore, a \gls*{swindex} can be computed as the one-to-one distance between corresponding
+  vertices of surfaces, weighted by their respective Voronoi area $a_i$:
+
+  \begin{equation}
+  sWI = \frac{1}{\sum_i a_i} \sum\limits_i^{N_V} a_i\,\|
+  \vec{v}_i - \hat{\vec{v}}_i \|
+  \label{eq:swindex}
+  \end{equation}
+%
+  where $\vec{v}_i \subset \gammaset$ are the locations of the total $N_V$ vertices, and $\hat{\vec{v}}_i$ is the location
+  recovered corresponding to the vertex $\vec{v}_i$.

### 2015-04-06 Publications/2015-NeuroImage/04-methods.tex
added new figure

@@
 %
-Let $\gammaset = \{\Gamma_m: m \in \mathbb{N}, m \leq N_c\}$ be the set of $N_c$ surfaces
+Let $\gammaset_R = \{\Gamma_m: m \in \mathbb{N}, m \leq N_c\}$ be the set of $N_c$ surfaces
   extracted from the undistorted \gls*{t1} image (the reference space $R$).
@@
   as a registration problem where we look for an underlying deformation field $U$, such that 
-  the structures in $R$ defined by \gammaset{} optimally align with their corresponding
+  the structures in $R$ defined by $\gammaset_R$ optimally align with their corresponding
   structures in $M$:
@@
 Let $\omegaset = \{\Omega_l: l \in \mathbb{N}, l \leq N_L\}$ be the set of $N_L$ competing regions in
-  which $M$ is partitioned by the projection of \gammaset.
+  which $M$ is partitioned by the projection of $\gammaset_R$.
 Using the Bayes' rule, the posterior likelihood is computed as:
@@
 The best estimate $\hat{U}$ then fulfills the maximum a posteriori criterium
-  \citep{bishop_pattern_2006}, and aligns \gammaset{} into $M$.
+  \citep{bishop_pattern_2006}, and aligns $\gammaset_R$ into $M$.
 First, we assume independence between pixels, and thus break down the
@@
 $C$ is the number of channels comprised in the reference image $M$.
+In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
+  the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized.
 
@@
 \paragraph*{Simulated phantoms}\label{sec:digital_phantoms}
-The workflow to generate the digital phantoms and assess
+The workflow to simulate the digital phantoms and assess
   \emph{regseg} on them is presented in \autoref{fig:evphantoms}.
@@
   analytical shapes and mathematical morphology.
-Using \emph{FreeSurfer} tools, the reference surfaces \gammaset{}
-  are generated.
+The reference surfaces $\gammaset_R$ are extracted from the binary shapes
+  using \emph{FreeSurfer} tools.
 The ground-truth distortion is generated with a chain of two displacement 
@@
   are warped interpolating the displacements field at each vertex.
-The warped surfaces are binarized to generate a distorted version of
-  the model in two target resolutions: \isores[mm]{2.0} and \isores[mm]{1.0}.
-Using a \gls*{mr} simulator \citep{caruyer_phantomas_2014}, we synthesized
-  \gls*{t1} (TE/TR=10/1500ms) and \gls*{t2} images (TE/TR=90/5000ms)
+The warped surfaces $\gammaset_\text{true}$ are binarized to generate tissue fractions
+  in low (\isores[mm]{2.0}) and high (\isores[mm]{1.0}).
+Using a \gls*{mr} simulator \citep{caruyer_phantomas_2014}, we synthesize
+  \gls*{t1} (TE/TR= 10/1500ms) and \gls*{t2} images (TE/TR= 90/5000ms)
   corresponding to each phantom type, with two resolutions each
   (1.0mm and 2.0mm isotropic).
-The field of view at both resolutions was \isores[mm]{100}.
-Simulations were corrupted with rician noise for a \gls*{snr} of 300.0.
-
-To evaluate the error, we computed the Haussdorf distance (as implemented
- by \cite{commandeur_vtk_2011}) between the surfaces warped with the ground-truth
- and their corresponding contours estimated with \emph{regseg}.
+The field of view at both resolutions is \isores[mm]{100}.
+Then, \emph{regseg} is applied to map $\gammaset_R$ into the warped phantoms to
+  obtain the registered surfaces ($\hat{\gammaset}_{test}$).
+% Simulations were corrupted with rician noise for a \gls*{snr} of 300.0.
+To quantify the misregistration error, we compute the Hausdorff distance between
+ $\hat{\gammaset}_{test}$ and $\gammaset_\text{true}$, using \citep{commandeur_vtk_2011}.
 
 \begin{figure*}
-\includegraphics[width=\linewidth]{figures/figure03}
-\caption{Evaluation workflow of phantoms.
-
-}\label{fig:evphantoms}
+\includegraphics[width=\linewidth]{figures/figure02}
+\caption{The evaluation of \emph{regseg} on phantom data is performed with the presented instrumental workflow:
+  1) The reference surfaces $\gammaset_R$ are triangularized meshes extracted from the four binary shapes (``box'', ``ball'', ``L'', ``gyrus'');
+  2) A ground-truth displacement field is generated as described in \autoref{sec:digital_phantoms}, and applied to warp
+      $\gammaset_R$, obtaining $\gammaset_\text{true}$;
+  3) Once warped, $\gammaset_\text{true}$ are projected to the corresponding discrete 3D volume, and downsampled creating partial volume effects to two resolutions
+     of \isores[mm]{2.0} and \isores[mm]{1.0}, producing sets of tissue fractions maps;
+  4) The tissue fractions are fed to an \acrfull*{mr} simulator, generating \acrfull*{t1} and \acrfull*{t2} -like images at the
+     two available resolutions;
+  5) The \emph{regseg} is run, using the warped test images as moving multispectral-image and $\gammaset_R$ as shape-priors;
+  6) Assessment of the agreement between the surfaces fitted with \emph{regseg} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$
+     with automatically generated visual reports and computing the Hausdorff distance between corresponding surfaces.}\label{fig:evphantoms}
 \end{figure*}
@@
 % 
-\autoref{fig:evworkflows}
-For the evaluation of the algorithm on real \gls*{dmri} data of human brains,
-  we collected 16 datasets from the ``minimally preprocessed''
-	 database of the \gls*{hcp}.
+For the evaluation of \emph{regseg} on real \gls*{dmri} data of human brains,
+  we collected 16 datasets from the \gls*{hcp} database.
+The original acquisitions are released within ``unprocessed'' packages, and
+  the ``minimally preprocessed'' are corrected for artifacts, brain-extracted
+  and spatially normalized, along with some results of the standard processing
+  pipeline of \emph{FreeSurfer}.
 We refer the reader to \citep{essen_human_2012} for exact details about acquisition
@@
   multi-shell \gls*{dmri} images.
-The original acquisitions are released within ``unprocessed'' packages, and
-  the ``minimally preprocessed'' are corrected for artifacts, brain-extracted
-  and spatially normalized, along with some results of the standard processing
-  pipeline of \emph{FreeSurfer}.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{figures/figure04-fa-md-jointplot}
+\caption{Segmentation model. The prior surfaces $\gammaset_R$ projected into the
+  multispectral image to be segmented $M$ actually partition the image in the
+  \omegaset{} homogeneous regions.}\label{fig:model}
+\end{figure}
+
+The general experimental framework for the real datasets is presented in \autoref{fig:evworkflows}.
+We designed the workflow with based on two arbitrary decisions:
+First, we map $\gammaset_R$ into a bivariates image comprising the \gls*{fa} and the
+  \gls*{adc} maps derived from the distorted \gls*{dmri} dataset.
+Previous studies \citep{ennis_orthogonal_2006,}
+The underlying segmentation model \omegaset{}
+Selecting the appropriate labels in the \emph{aparc} segmentation, we applied
+  \texttt{mri\_tessellate} to extract the surface of the following
+  six homogeneous regions $\Omega_l$:
+  1) the thalamus, $\Omega_{Tha}$;
+  2) \gls*{csf} of the ventricular system and deep \gls*{gm} structures, $\Omega_{VdGM}$;
+  3) cerebral \gls*{wm}, $\Omega_{WM}$;
+  4) brain stem and cerebellar \gls*{wm}, $\Omega_{bst}$;
+	5) cerebellar \gls*{gm}, $\Omega_{cbGM}$; and
+	6) cortical \gls*{gm} surface, $\Omega_{pial}$.
+The segmentation model corresponding to this selection of surfaces is presented in
+  \autoref{fig:model}, and further addressed in \autoref{sec:res_model_and_metric}.
 
 \begin{figure*}
-\includegraphics[width=\linewidth]{figures/figure02}
-\caption{Experimental workflow applied on real data from the \acrfull*{hcp}.
-  1) The prior surfaces are extracted from the anatomical reference (\gls*{t1} image).
+\includegraphics[width=\linewidth]{figures/figure03}
+\caption{Experimental workflow real data from the \acrfull*{hcp}.
+  1) $\gammaset_R$ are extracted from the anatomical reference (\gls*{t1} image).
   2) To operate as ground truth, we generate a plausible-synthetic distortion $U_{true}$
@@
 \end{figure*}
-
-Selecting the appropriate labels in the \emph{aparc} segmentation, we applied
-  \texttt{mri\_tessellate} to extract the surface of the following
-  six homogeneous regions $\Omega_l$:
-  1) the thalamus, $\Gamma_{Tha}$;
-  2) \gls*{csf} of the ventricular system and deep \gls*{gm} structures, $\Gamma_{VdGM}$;
-  3) cerebral \gls*{wm}, $\Gamma_{WM}$;
-  4) brain stem and cerebellar \gls*{wm}, $\Gamma_{bst}$;
-	5) cerebellar \gls*{gm}, $\Gamma_{cbGM}$; and
-	6) cortical \gls*{gm} surface, $\Gamma_{pial}$.
-The segmentation model corresponding to this selection of surfaces is presented in
-  \autoref{fig:model}, and further addressed in \autoref{sec:res_model_and_metric}.
 

### 2015-04-06 Publications/2015-NeuroImage/04-methods.tex
add ISBI reference

@@
 The general experimental framework for the real datasets is presented in \autoref{fig:evworkflows}.
+\citep{esteban_brain_2012}
+
 We designed the workflow with based on two arbitrary decisions:
-First, we map $\gammaset_R$ into a bivariates image comprising the \gls*{fa} and the
+First, we map $\gammaset_R$ into a bivariate image comprising the \gls*{fa} and the
   \gls*{adc} maps derived from the distorted \gls*{dmri} dataset.
-Previous studies \citep{ennis_orthogonal_2006,}
+In a previous study
 The underlying segmentation model \omegaset{}

### 2015-04-06 Publications/2015-NeuroImage/00-main.tex
refactoring, add figure of phantoms
my fixes
added new figure
add ISBI reference
major update

@@
 \documentclass[5p,authoryear]{elsarticle}
+% \documentclass[authoryear,onecolumn]{elsarticle}
 \journal{ArXiv archiving for pre-publication review}
@@
 \usepackage[colorinlistoftodos]{todonotes}
+\usepackage{xifthen}
+\usepackage{xspace}
 
@@
 \providecommand{\suppl}[1]{\href{http://dx.doi.org/mydoi}{Supplemental Materials}, #1}
-\providecommand{\isores}[1]{#1\ensuremath{\times}#1\ensuremath{\times}#1}
+\providecommand{\isores}[2][]{#2\ensuremath{\times}#2\ensuremath{\times}#2%
+\ifthenelse{\equal{#1}{}}{}{ [#1]}\xspace}
 
@@
 \hyphenation{op-tical net-works semi-conduc-tor an-isot-ropy regis-tra-tion iso-tropic Free-Surfer %
-align-ed pre-sent}
+align-ed pre-sent work-flow data-set data-sets}
 
@@
 
+% http://tex.stackexchange.com/questions/56105/missing-references-heading
+\def\bibsection{\section*{References}}
 \begin{document}
@@
 
-\title{Active-contours driven registration applied in the correction of
-susceptibility-derived distortions of diffusion MR images}
+\title{Active-contours driven registration method for structure-informed segmentation of diffusion MR images}
 
@@
 \begin{keyword}
-diffusion MRI \sep susceptibility~distortion \sep segmentation %
+diffusion MRI \sep cortical~parcellation \sep segmentation %
 \sep nonlinear~registration \sep active~contours \sep echo-planar imaging.
@@
 
-%\bibliographystyle{plainnat}
-% \section*{References}
-\bibliographystyle{mystyle}
-\bibliography{Remote}
-
 % Add appendices
@@
 
+%\bibliographystyle{plainnat}
+\bibliographystyle{mystyle}
+\bibliography{Remote}
+
 \end{document}

### 2015-04-06 Publications/2015-NeuroImage/03-introduction.tex
major update

@@
 Since the nonlinear distortion is aligned with one of the imaging axis (see 
-  \autoref{sec:datasets}), \emph{regseg} includes an anisotropic regularizer of
+  \autoref{sec:human_connectome}), \emph{regseg} includes an anisotropic regularizer of
   the displacements field proposed by \cite{nagel_investigation_1986}.

### 2015-04-06 Publications/2015-NeuroImage/04-methods.tex
major update

@@
   where $\vec{f}' = M(\vec{r}')$ is the feature vector at the displaced
-  position $\vec{r}'$ \eqref{eq:transform} in the multispectral target
-  volume.
+  position $\vec{r}'$ \eqref{eq:transform} in the moving image.
 For convenience, and because it has been shown to be an appropriate approximation
@@
   $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
-$C$ is the number of channels comprised in the reference image $M$.
+$C$ is the number of channels comprised in the image $M$.
 In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
@@
 	The gradient speeds $\bar{s}_i$ are computed as the disparity between data energies of
-	  the vertex $\vec{v}_i$ in each limiting region (see \ref{app:shape_priors}).
+	  the vertex $\vec{v}_i$ in each limiting region (see \suppl{equation SM5})..
 	In this figure, the gradient speed corresponding to $\vec{v}_1$ is written in the lower
-	  box, $\Omega_{wm}$ being the inner limiting region and $\Omega_{gm}$ the outer.
-	Finally, every $\bar{s}_i$ is associated with the control points $\vec{u}_k$ through
-	  the corresponding weights $\psi_k(\vec{v}_i)$ by Eq. \eqref{eq:gradient_wshape}.
+	  box, $\Omega_{wm}$ being the inner limiting region and $\Omega_{gm}$ the outer, and
+    $w_{0,1}$ the relative area associated to vertex $\vec{v}_1$ with respect to
+    the total area of surface $\Gamma_0$.
+	% Finally, every $\bar{s}_i$ is associated with the control points $\vec{u}_k$ through
+	%   the corresponding weights $\psi_k(\vec{v}_i)$ by Eq. \eqref{eq:gradient_wshape}.
 	}\label{fig:method}
@@
   + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
-  \Big\}.
+  \Big\},
   \label{eq:gradient_descent}
@@
 Finally, to descend this gradient we establish a semi-implicit Euler scheme (see \suppl{section S1.3}),
-  with a step size parameter $\delta$, that we solve it in the spectral domain as follows:
+  with a step size parameter $\delta$, that we solve in the spectral domain as follows:
 
@@
   \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\delta^{-1} \, \vec{u}_k^t + \vec{g}_k\} }%
-                  {\mathcal{F}\{(\delta^{-1} + \boldsymbol{\alpha})\, I - \boldsymbol{\beta}\Delta\}} \right\}
+                  {\mathcal{F}\{(\delta^{-1} + \boldsymbol{\alpha})\, I - \boldsymbol{\beta}\Delta\}} \right\}.
   \label{eq:update_equation}
@@
 
-\subsection{Data and corresponding evaluations}
-\label{sec:datasets}
+\subsection{Evaluation protocol}\label{sec:evaluation_protocol}
 In order to assess the performance of \emph{regseg}, we defined a general
   evaluation protocol:
-1) Extraction of the set of undistorted surfaces \gammaset{};  
-2) computation of a ground-truth field of displacements that is applied to
+1) Extraction of the set of undistorted surfaces $\gammaset_R$;
+2) computation of a ground-truth field of displacements $U_{true}$ that is applied to
   generate warped images ($M$) to segment;
-3) execution of \emph{regseg} with \gammaset{} and the warped data as inputs; and
+3) execution of \emph{regseg} with $\gammaset_R$ and the warped data as inputs; and
 4) visual assessment and computation of error metrics.
 
-\paragraph*{Simulated phantoms}\label{sec:digital_phantoms}
+\subsection{Simulated phantoms}\label{sec:digital_phantoms}
 The workflow to simulate the digital phantoms and assess
@@
 
-\paragraph*{Real datasets} %
+\subsection{Real datasets} %
 \label{sec:human_connectome}
 % 
-For the evaluation of \emph{regseg} on real \gls*{dmri} data of human brains,
-  we collected 16 datasets from the \gls*{hcp} database.
-The original acquisitions are released within ``unprocessed'' packages, and
-  the ``minimally preprocessed'' are corrected for artifacts, brain-extracted
-  and spatially normalized, along with some results of the standard processing
-  pipeline of \emph{FreeSurfer}.
-We refer the reader to \citep{essen_human_2012} for exact details about acquisition
-  parameters, and \citep{glasser_minimal_2013} for the preprocessing issues.
-The datasets comprehend a large set of images, containing \gls*{t1}, \gls*{t2} and
-  multi-shell \gls*{dmri} images.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{figures/figure04-fa-md-jointplot}
-\caption{Segmentation model. The prior surfaces $\gammaset_R$ projected into the
-  multispectral image to be segmented $M$ actually partition the image in the
-  \omegaset{} homogeneous regions.}\label{fig:model}
-\end{figure}
-
 The general experimental framework for the real datasets is presented in \autoref{fig:evworkflows}.
-\citep{esteban_brain_2012}
-
-We designed the workflow with based on two arbitrary decisions:
-First, we map $\gammaset_R$ into a bivariate image comprising the \gls*{fa} and the
-  \gls*{adc} maps derived from the distorted \gls*{dmri} dataset.
-In a previous study
-The underlying segmentation model \omegaset{}
-Selecting the appropriate labels in the \emph{aparc} segmentation, we applied
-  \texttt{mri\_tessellate} to extract the surface of the following
-  six homogeneous regions $\Omega_l$:
-  1) the thalamus, $\Omega_{Tha}$;
-  2) \gls*{csf} of the ventricular system and deep \gls*{gm} structures, $\Omega_{VdGM}$;
-  3) cerebral \gls*{wm}, $\Omega_{WM}$;
-  4) brain stem and cerebellar \gls*{wm}, $\Omega_{bst}$;
-	5) cerebellar \gls*{gm}, $\Omega_{cbGM}$; and
-	6) cortical \gls*{gm} surface, $\Omega_{pial}$.
-The segmentation model corresponding to this selection of surfaces is presented in
-  \autoref{fig:model}, and further addressed in \autoref{sec:res_model_and_metric}.
+It extends our previous work \citep{esteban_simulationbased_2014} evaluating distortions
+  on \gls*{dmri} phantoms.
 
@@
 
-The displacements along the \gls*{pe} axis are related to the local deviation of the
-  field $\Delta B_0(\vec{r})$ \citep{jezzard_correction_1995}:
+\paragraph*{Data}
+For the evaluation of \emph{regseg} on real \gls*{dmri} data of human brains,
+  we collected 16 subjects from the \gls*{hcp} database.
+The original acquisitions are released within ``unprocessed'' packages, and
+  the ``minimally preprocessed'' are corrected for artifacts, brain-extracted
+  and spatially normalized, along with some results of the standard processing
+  pipeline of \emph{FreeSurfer}.
+We refer the reader to \citep{essen_human_2012} for exact details about acquisition
+  parameters, and \citep{glasser_minimal_2013} for the preprocessing issues.
+The datasets comprehend a large set of images, containing \gls*{t1}, \gls*{t2} and
+  multi-shell \gls*{dmri} images.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{figures/figure04-fa-md-jointplot}
+\caption{Segmentation model defined by the homogeneous regions $\Omega_i$ in which
+  the image to be segmented is partitioned by the theoretical projection 
+  ($\gammaset_{true}$) of $\gammaset_R$ into $M$.
+The plot represents a kernel density estimation of the location and spread of
+  each region $\Omega_i$ in the bivariate feature space of $M$, comprising 
+  the \gls*{fa} and \gls*{adc} maps.
+On top and right margins the marginal distributions of $\Omega_i$ are also plotted for
+  the \gls*{fa} and \gls*{adc}, respectively.
+}\label{fig:model}
+\end{figure}
+
+\paragraph*{Segmentation model}
+Based on our experience \citep{esteban_brain_2012} and previous studies \citep{ennis_orthogonal_2006},
+  we define the moving image as a stack of the \gls*{fa} and \gls*{adc} maps derived
+  from \gls*{dmri} data.
+After evaluating several alternative models, we empirically defined the partition \omegaset{} 
+  of the reference image $R$ as follows:
+  1) $\Omega_{Tha}$ including the thalamus;
+  2) $\Omega_{VdGM}$ including the ventricular system and deep \gls*{gm} structures;
+  3) $\Omega_{WM}$ including the cerebral \gls*{wm};
+  4) $\Omega_{bst}$ including the brain stem and cerebellar \gls*{wm};
+	5) $\Omega_{cbGM}$ including the cerebellar \gls*{gm}; and
+	6) $\Omega_{GM}$ including the cortical \gls*{gm}.
+Using tools of \emph{FreeSurfer}, and the appropriate selections of labels in the 
+  \emph{aparc} segmentation released with the \gls*{hcp} data, the $\gammaset_R$ set of
+  reference surfaces is extracted.
+The segmentation model corresponding to this partition is depicted in \autoref{fig:model},
+  and discussed with finer detail in \suppl{section S4}.
+
+\paragraph*{Ground-truth generation}
+Realistic deformation is achieved generating displacements fields that meet the theoretical
+  properties of distortion \citep{jezzard_correction_1995}.
+The displacements along the \gls*{pe} axis of the \gls*{dmri} image are related to the local
+  deviation of the field $\Delta B_0(\vec{r})$ from its nominal value $B_0$ as follows:
 
   \begin{equation}
-  u(\vec{r}) = \frac{\gamma \, T_{acq}\, s_{PE}}{2\pi}\Delta B_0(\vec{r}) \: \hat{\vec{e}}_{PE},
+  u_\text{PE} = \frac{\gamma \, T_{acq}\, s_\text{PE}}{2\pi}\Delta B_0(\vec{r})\text{ [mm]},
   \label{eq:fieldmap}
@@
 %
-where $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time,
-  $s_{PE}$ is the pixel size along \gls*{pe}, and $\hat{\vec{e}}_{PE}$ the unitary
-  vector along \gls*{pe}.
-With certain \gls*{mr} sequences it is possible to estimate $\Delta B_0$, obtaining
+where $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time, and
+  $s_\text{PE}$ is the pixel size along \gls*{pe}.
+Certain \gls*{mr} sequences are designed to estimate $\Delta B_0$, obtaining
   the so-called fieldmap.
+We derive the deformation $U_{true}$ from the fieldmap image released with
+  the corresponding packages of each dataset from the \gls*{hcp}.
+The fieldmap is unwrapped\footnote{fieldmaps are phase maps, clipped in the interval
+  of $[-\pi, \pi)$ [rads] or [rads/s].} and smoothed before applying \eqref{eq:fieldmap}.
+Then, the original \gls*{dmri} is warped using the resulting displacements field and fed into
+  a pipeline that process the corresponding \gls*{dti} and computes the derived scalars of
+  interest (\gls*{fa} and \gls*{adc}) using \emph{MRtrix} \citep{tournier_mrtrix_2012}.
 
-In this case, we derived the deformation $U_{true}$ from the fieldmaps released with
-  the corresponding packages of each dataset from the \gls*{hcp} applying \eqref{eq:fieldmap},
-  mimicking the real distortions, using a derivation of our previous work
-  \citep{esteban_simulationbased_2014}.
-After warping the original \gls*{dmri} with $U_{true}^{-1}$, we computed the \gls*{dti} and
-  the derived scalars (\gls*{fa} and \gls*{adc}) using \emph{MRtrix} \citep{tournier_mrtrix_2012}.
-The stack of \gls*{fa} and \gls*{adc} conforms the multivariate input for \emph{regseg}.
+\paragraph*{Metric assessment}
+Preliminarily, we investigated the aptness of the segmentation model.
+For five test datasets, we uniformly sampled the space of distortions
+  $\hat{U} = \epsilon \cdot U_{true} = \vec{r} + \epsilon \, u_\text{PE}$
+  (with $\epsilon \in [-1.1, 1.1]$ and $u_\text{PE}$ from \eqref{eq:fieldmap}),
+  and evaluated the data-term of the cost function \eqref{eq:energy}.
+The metric consistently displayed its minimum at the ground-truth location ($\epsilon=0.0$)
+  for all the cases (\suppl{Figure S2}).
 
-A dual workflow to the general evaluation framework (\autoref{fig:evworkflows})
-  was also implemented to integrate the \gls*{t2b} registration scheme.
+\paragraph*{Cross-comparison}
+A dual workflow to the general evaluation framework of \autoref{fig:evworkflows}
+  integrates the alternate \gls*{t2b} registration scheme.
 We reproduced the solution and settings provided with \emph{ExploreDTI}
@@
   perform registration.
-First, the average \emph{b0} volume is computed using all the \emph{low-b} volumes within
-  the distorted \gls*{dmri} dataset.
-Using the ``preprocessed'' instance of the \gls*{t2} image of the corresponding subject as
-  reference image, and the \emph{b0} image as moving image, we registered both and
-  projected the correct (undistorted) contours to the warped space using the resulting
-  transform.
 
-Conversely, the susceptibility-derived distortions only happen along the \gls*{pe}
-  axis of the image.
-Therefore, a \gls*{swindex} can be computed as the one-to-one distance between corresponding
-  vertices of surfaces, weighted by their respective Voronoi area $a_i$:
+\paragraph*{Error measurement}\label{sec:experiments_evaluation}
+Since the distortions only happen along the \gls*{pe} axis of the image, we compute the
+  \gls*{swindex} as the area-weighted distance between corresponding vertices of 
+  $\gammaset_{true}$ and their estimation by the method under test $\hat{\gammaset}_{test}$:
 
@@
 %
-  where $\vec{v}_i \subset \gammaset$ are the locations of the total $N_V$ vertices, and $\hat{\vec{v}}_i$ is the location
-  recovered corresponding to the vertex $\vec{v}_i$.
+  where $\vec{v}_i \subset \gammaset_\text{true}$ are the locations of the total $N_V$ vertices,
+  and $\hat{\vec{v}}_i \subset \hat{\gammaset}_{test}$ are the locations
+  recovered corresponding to $\vec{v}_i$.
+In practice, we report the \gls*{swindex} only for three surfaces of interest crucial in whole-brain
+  tractography.
+These three surfaces delineate the following regions in the model: $\Omega_{VdGM}$, $\Omega_{WM}$ and $\Omega_{GM}$.

### 2015-04-06 Publications/2015-NeuroImage/05-results.tex
major update

@@
 \end{figure*}
-
-Using a simplified subset of the experimental instrumentation described in
-  \autoref{sec:experiments_evaluation}, we collected quantitative results on the digital
-  phantoms.
-A total of 1200 registration experiments (4 phantom types $\times$ 150 random warpings
-  $\times$ 2 resolutions) were run.
-Parameter settings of every processing step in the experiments are made available along with
-  the source code.
-For each registration, the averaged Hausdorff distance (see \autoref{sec:experiments_evaluation})
-  of the inner and the outer surfaces was computed.
+A total of 1200 experiments (4 phantom types $\times$ 150 random warpings $\times$ 2 resolutions) were
+  run using the workflow given in \autoref{fig:evphantoms}.
+For each experiment, the misregistration error was measured using the Hausdorff distance
+  (see \autoref{sec:experiments_evaluation}) between the theoretical $\gammaset_\text{true}$ and
+  the estimation done by \emph{regseg} ($\hat{\gammaset}_{test}$).
 The results showed a consistent and high accuracy, below the image resolution.
@@
   for all the tests (4 phantom types $\times$ 2 resolutions).
-Since statistical tests may not be conclusive enough, we also computed the confidence intervals
-  (\autoref{tab:ci_phantom}).
+Since statistical tests may not be conclusive enough, we also computed the confidence intervals,
+  reported in \autoref{tab:ci_phantom}.
 
@@
 
-\subsection{Correction of distortions on real datasets}\label{sec:results_hcp}
-
-\paragraph*{Assessing the segmentation model}\label{sec:res_model_and_metric} %
-%
-Preliminarily, we investigated the aptness of our segmentation model to play as cost function
-  driving the registration process.
-First, we evaluated several alternative models to segment the target bivariate images
-  (\gls*{fa} and \gls*{adc}).
-Based on the observed joint distribution of each (\suppl{section S4}),
-  we selected the 6-regions scheme described in \autoref{sec:human_connectome}.
-
-The distortions follow expression \eqref{eq:fieldmap} and hence, we easily reduced the
-  error space to one dimension as follows.
-The energy functional \eqref{eq:energy} was evaluated for the warpings given by
-  a linearized space of distortions
-  $\hat{U} = \epsilon \cdot U_{true} = \vec{r} + \epsilon \cdot u(\vec{r})$,
-  with $\epsilon \in [-1.1, 1.1]$.
-The results of this experiment are shown in \autoref{fig:energymap}.
-The metric consistently displayed its minimum at $\epsilon=0.0$ (ground-truth location).
-
-\paragraph*{Evaluation \& cross-comparison}\label{sec:res_cc_evaluation}
+\subsection{Evaluation on real datasets and cross-comparison}\label{sec:results_hcp}
 %
 Finally, we compared the performance of \emph{regseg} and the standard \gls*{t2b}
-  correction.
-We computed the \gls*{swindex} \eqref{eq:swindex} of every surface after correction,
+  registration.
+We computed the \gls*{swindex} \eqref{eq:swindex} of every surface after registration,
   for both \emph{regseg} and the \gls*{t2b} methods.
@@
 We also reported the 95\% CIs of the \gls*{swindex} for those surfaces.
-The aggregate CI for \emph{regseg} was $[1.08 - 1.50] mm$, whereas the competing method
-  yielded an aggregate CI of $[2.06 - 2.43] mm$.
+The aggregate CI for \emph{regseg} was 1.08 - 1.50 [mm], whereas the \gls*{t2b} method
+  yielded an aggregate CI of 2.06 - 2.43 [mm].
 The quantitative results of the ANOVA tests and the CIs are presented in \autoref{tab:results_real}.
-Visual assessment of one case is shown in \autoref{fig:results_real}, along with the corresponding
-  violin plots of error distributions by correction method and surface.
-Extended visual assessment of the 16 cases is included in the \suppl{section S5}.
+Visual assessment of the 16 cases is included in the \suppl{section S5}.
+In \autoref{fig:results_real}, box A, the visual report for one subject is supplied.
+Box B presents the corresponding violin plots of error distributions by registration method
+  and surface.
 

### 2015-04-06 Publications/2015-NeuroImage/06-discussion.tex
major update

@@
 We proved that active contours without edges can be used to successfully drive a
-  deformation supported by B-Spline functions using digital phantoms.
+  deformation supported by B-Spline basis functions using digital phantoms.
 We randomly deformed 4 different phantom models mimicking three homogeneous regions,
-  showing contrasts generated by a publicly available \gls*{mri} simulator.
+  and use them to simulate \gls*{t1} and \gls*{t2} images at two resolution levels.
 After registration with \emph{regseg} we measured the Hausdorff distance between the
@@
 We also assessed the 95\% \gls*{ci}, yielding an aggregate interval of
-  $[0.64 - 0.66]mm$ in the low resolution phantoms ($2.0mm$ isotropic voxel) and
-  $[0.34 - 0.38]mm$ in the high resolution phantoms ($1.0mm$ isotropic).
-Therefore, we also conclude that the error is bounded and below the half of the
-  voxel spacing to illustrate the accuracy of \emph{regseg}.
+  0.64 - 0.66 [mm] in the low resolution phantoms (2.0mm isotropic voxel) and
+  0.34 - 0.38 [mm] in the high resolution phantoms (1.0mm isotropic).
+Therefore, we also conclude that the error is bounded above by the half of the
+  voxel spacing.
 
 \paragraph*{Application on real data}
-We designed \emph{regseg} as a method to correct \gls*{dmri} data for susceptibility
-  distortions.
-Applications of whole-brain tractography generally include the acquisition of \gls*{t1} images
-  to provide with prior information about anatomy with high accuracy.
-Our method solves in one single step a joint problem usually addressed in two steps:
-  linear registration of \gls*{t1} and \gls*{dmri} to fuse the anatomical information on
-  the diffusion data, and correction of distortion using extra acquisitions.
-Moreover, our method does not require any additional acquisition over the minimal protocol
-  that only includes \gls*{t1} and \gls*{dmri} images, a very common situation for
-  historical datasets.
-Beyond the proposed correction, potential applications of \emph{regseg} are atlas-based
-  segmentation and tracking objects in time-series.
+We designed \emph{regseg} as a method to segment \gls*{dmri} data exploiting the
+  anatomical information extracted in the corresponding \gls*{t1} image of the subject.
+Applications of whole-brain tractography \citep{smith_tractbased_2006,craddock_imaging_2013}
+  usually solve the problem in a two-step approach.
+First, images are corrected for nonlinear distortions using auxiliary acquisitions
+  such as fieldmaps \citep{jezzard_correction_1995}, \gls*{t2} images \citep{kybic_fast_2003},
+  etc.
+Second, the segmentation is projected from a reference \gls*{t1} image using linear
+  registration \citep{greve_accurate_2009}.
+\emph{Regseg} solves this joint problem in one single step, and it does not require any additional
+  acquisition over the minimal protocol including \gls*{t1} and \gls*{dmri} images only.
+This situation is commonly found in historical datasets.
+Beyond the proposed application in \gls*{dmri} data, other potential uses of \emph{regseg} are
+  atlas-based segmentation and tracking objects in time-series.
 
-After a preliminary experiment to demonstrate the aptness of our segmentation model
-  in section \autoref{sec:res_model_and_metric}, we evaluated \emph{regseg} in
-  a real environment, using the experimental framework presented in \autoref{fig:evworkflows}.
+We evaluated \emph{regseg} in a real environment, using the experimental framework presented
+  in \autoref{fig:evworkflows}.
 We processed 16 subjects from the \gls*{hcp} database with both \emph{regseg}
   and an in-house replication of the \gls*{t2b} correction method.
-\emph{Regseg} showed very high accuracy, with an aggregate 95\% \gls*{ci} of $[1.08-1.50]mm$,
-  almost below the pixel size of $1.25mm$.
+\emph{Regseg} showed very high accuracy, with an aggregate 95\% \gls*{ci} of 0.56-0.66 [mm],
+  bellow the pixel size of 1.25mm.
 The misregistration error remaining after \emph{regseg} was significantly lower ($p < 0.01$) than the
@@
 Even though we carefully configured the \gls*{t2b} method by using the same algorithm and
-  same settings from a widely used software, cross-comparison experiments are prone to
+  same settings of a widely used software, cross-comparison experiments are prone to
   the so-called \emph{instrumentation bias} \citep{tustison_instrumentation_2013}.
 Therefore, these results do not show that the \emph{regseg} software \emph{is} better than the
-  \gls*{t2b} method, but they are communicated to proof that \emph{regseg} is a solid option to
-  correct \gls*{dmri} data for distortions, and any other potential application.
+  \gls*{t2b} method, but they are communicated to proof that \emph{regseg} is a solid option
+  in the proposed application.
+Additionally, the \gls*{t2b} may introduce an additional (and small) error in the necessary
+  registration of \gls*{t2} to \gls*{t1} space.
 
-\paragraph*{Prospects}
-Future improvements of \emph{regseg} possibly will follow a similar line to the evolution of 
-  existing methodologies.
-Incorporating knowledge about the distortion by initializing our method with the theory-based
-  correction using fieldmaps may be the most cost-effective.
-The main pitfall using fieldmap correction is the challenging problem of unwrapping the
-  range of the phase map, as it is clipped in the range $(-\pi, \pi]$.
-The most extended solution to this is \emph{prelude} \citep{jenkinson_fast_2003}, and some
-  alternatives have been proposed \citep{zhou_reliable_2009,daga_susceptibility_2014} to
-  circumvent the inherent unreliability of the method.
-The \emph{regseg} algorithm would naturally correct misregistered areas due to errors in
-  phase unwrapping, and underestimations introduced by necessary smoothing
-  applied to clean up field maps.
-This integration has been recently applied in the \gls*{t2b} approach
-  by \cite{irfanoglu_susceptibility_2011}, also investigating the usefulness of
-  nonregular B-Spline grids.
-This is also a potential extension to our deformation model.
-
-Finally, we are currently studying substitute features to build the energy functional on,
-  enabling to perform \emph{regseg} directly on the raw \gls*{dmri} data.
+% \paragraph*{Prospects}
+% First extensions of this work will study more appropriate features to build the energy functional
+%   on, enabling to perform \emph{regseg} directly on the raw \gls*{dmri} data.
+% A second outlook covers incorporating knowledge about the distortion by initializing our method
+%   with the theory-based displacement field that can be estimated with fieldmaps.

### 2015-04-06 Publications/2015-NeuroImage/07-conclusion.tex
major update

@@
   defined explicitly by their limiting surfaces.
-The surfaces are active, and evolve on a free-deformation field supported by B-Splines.
+The surfaces are active, and evolve on a free-form deformation field supported by B-Splines.
 A descent optimization strategy is guided by shape gradients computed on the current partition
@@
 	warped with smooth random deformations.
-The resulting average misplacement of the contours was significantly lower than the
-  image resolution of the phantoms.
+The resulting misregistration of the contours was significantly lower than the image resolution
+  of the phantoms.
 
-We propose \emph{regseg} to correct \gls*{dmri} datasets for the typically present distortions
-  caused by inhomogeneities of the main magnetic field.
-Visual assessment of \emph{regseg} and cross-comparison with a widely-used competing
-  technique demonstrated its accuracy.
+We propose \emph{regseg} to simultaneously segment \gls*{dmri} data and register them to
+  their corresponding \gls*{t1} image of the same subject.
+Visual assessment of \emph{regseg} and cross-comparison with a widely-used technique demonstrated 
+  its accuracy.
 Moreover, \emph{regseg} does not require additional images to the minimal acquisition protocol
@@
   involved in this paper is also publicly released.
-The \emph{regseg} algorithm is implemented in C++, on top of ITK-4.6
-  (Insight Registration and Segmentation Toolkit, \url{http://www.itk.org}).
-The evaluation instruments (\autoref{fig:evworkflows}), are implemented in Python using
+\emph{Regseg} is implemented on top of ITK-4.6 (Insight Registration and 
+  Segmentation Toolkit, \url{http://www.itk.org}).
+The evaluation instruments (\autoref{fig:evworkflows}), are implemented using
   \emph{nipype} \citep{gorgolewski_nipype_2011} to assess their reproducibility.
-Phantom data are automatically generated by the evaluation framework.
 All research elements (data, source code, figures, manuscript sources, etc.) in this paper
-  are made publicly available under a unique digital object identifier in a
-  package \citep{esteban_acweregistration_2015}.
+  are made publicly available under a unique package \citep{esteban_acweregistration_2015}.

### 2015-04-07 Publications/2015-NeuroImage/06-discussion.tex
edits received from Andres

@@
   (\autoref{tab:results_real}).
-Visual inspection of all the results (\suppl{section S5}), and the violin plots included in
-  \autoref{fig:results_real}, confirmed this intuition.
+Visual inspection of all the results (\suppl{section S5}) and the violin plots included in
+  \autoref{fig:results_real} confirmed that \emph{regseg} overperformed the \gls*{t2b} method
+  in our settings.
 Even though we carefully configured the \gls*{t2b} method by using the same algorithm and
@@
   the so-called \emph{instrumentation bias} \citep{tustison_instrumentation_2013}.
-Therefore, these results do not show that the \emph{regseg} software \emph{is} better than the
-  \gls*{t2b} method, but they are communicated to proof that \emph{regseg} is a solid option
-  in the proposed application.
+Therefore, these results do not prove that the \emph{regseg} software \emph{is better than} the
+  \gls*{t2b} method.
+The results reported let us propose \emph{regseg} as a solid option in the application field.
 Additionally, the \gls*{t2b} may introduce an additional (and small) error in the necessary

### 2015-04-07 Publications/2015-NeuroImage/03-introduction.tex
edits received from Andres
edits from Chus

@@
 However, segmenting \gls*{dmri} data precisely and coregistration of anatomical
-  images are difficult for several reasons.
+  images are difficult due to several limitations.
 First, \gls{dmri} images suffer from partial voluming of several tissues in 
-  interfacing voxels due to the low resolution of images (generally around
-  \isores[mm]{2.2}).
+  voxels located at their interfaces due to the low resolution of images
+  (generally around \isores[mm]{2.2}).
 Second, \gls*{dmri} schemes probe the diffusion process within the brain in 
@@
 The extremely low \gls*{snr} and high dimensionality of \glspl*{dwi} disable their
-  use for direct segmentation.
+  direct use in segmentation.
 The low contrast between \gls*{gm} and \gls*{wm} in the \emph{b0} volumes is not suitable for 
@@
 Lastly, \gls*{dmri} images are acquired using \gls*{epi} to speed up acquisition
-  at the cost of presenting geometrical distortion and 
-  signal losses \citep{jezzard_correction_1995} that hinder the segmentation by
-  registering an anatomical image as source of same-subject prior information.
+  at the cost of presenting geometrical distortion and signal degradation and
+  destruction \citep{jezzard_correction_1995}.
 The artifact has been proven to cause an important impact on the anatomy extracted
@@
   when used in whole-brain tractography.
+These limitations preclude segmentation in native \gls*{dmri} space, including
+  registration-based approaches using anatomical images as source of structural
+  information.
 
@@
 Therefore, given the described circumstances, a method performing simultaneously
-  segmentation in \gls*{dmri} space and registration of the corresponding \gls*{t1} image
+  segmentation in native \gls*{dmri} space and registration of the corresponding \gls*{t1} image
   to the \gls*{dmri} data could provide the optimal solution.
@@
   \citep{besson_dream2s_2003,herbulot_segmentation_2006}, therefore \emph{regseg}
-  does not implement levelsets as most of the presented methods.
+  does not implement levelsets as most of the presented methods do.
 Since the nonlinear distortion is aligned with one of the imaging axis (see 

### 2015-04-07 Publications/2015-NeuroImage/05-results.tex
edits from Chus

@@
   to the two sets of resolutions of generated phantoms.
-To support that the misregistration errors averaged per experiment was significantly
-  under the resolution of the target image, we proceeded as follows.
-First, we confirmed that the error distributions were skewed using a Shapiro-Wilk test of
+In order to relate average misregistration error to the resolution of the moving image,
+  we proceeded as follows.
+First, we confirmed that the vertex-wise error distributions were skewed using a Shapiro-Wilk test of
   normality.
@@
 Consequently, we used the non-parametric Wilcoxon signed-rank test along with Bonferroni
-  correction for multiple comparisons ($N=150$).
+  correction for multiple comparisons ($N=150$, for each phantom type).
 Averaged errors resulted significantly lower than voxel size with $p < (0.001 / 150)$

### 2015-04-07 Publications/2015-NeuroImage/04-methods.tex
edits received from Andres
final improvements before sending to authors, removed ANOVA

@@
 	They are represented in green and dark blue colors in this close-up.
-	They iteratively evolve following their inner normals $\hat{n}_i$ at each vertex
+	They iteratively evolve following their inward normals $\hat{n}_i$ at each vertex
 	  $\vec{v}_i$ of the mesh.
-	The gradient speeds $\bar{s}_i$ are computed as the disparity between data energies of
-	  the vertex $\vec{v}_i$ in each limiting region (see \suppl{equation SM5})..
-	In this figure, the gradient speed corresponding to $\vec{v}_1$ is written in the lower
-	  box, $\Omega_{wm}$ being the inner limiting region and $\Omega_{gm}$ the outer, and
+	The gradient speeds $\bar{s}_i$ drive registration, and are computed as the disparity of data
+    energies with respect to the two limiting regions of $M(\vec{v}_i)$, the features of the image
+    $M$ in the location of vertex $\vec{v}_i$ (see \suppl{equation SM5}).
+	In this figure, $\bar{s}_1$ is written in the lower
+	  box, with $\Omega_{wm}$ being the inner limiting region, $\Omega_{gm}$ the outer region, and
     $w_{0,1}$ the relative area associated to vertex $\vec{v}_1$ with respect to
@@
   \begin{align}
-  \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\delta^{-1} \, \vec{u}_k^t + \vec{g}_k\} }%
-                  {\mathcal{F}\{(\delta^{-1} + \boldsymbol{\alpha})\, I - \boldsymbol{\beta}\Delta\}} \right\}.
+  \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\delta^{-1} \, \vec{u}_k^t - \vec{g}_k\} }%
+                  {\mathcal{F}\{(\delta^{-1} + \boldsymbol{\alpha})\, I - \boldsymbol{\beta}\Delta\}} \right\},
   \label{eq:update_equation}
   \end{align}
+  where $I$ denotes the identity operator.
 
@@
 4) visual assessment and computation of error metrics.
+How this protocol is adapted to the simulated phantoms and real data is covered in the
+  following sections.
 
@@
     from the fieldmap using \eqref{eq:fieldmap}.
-  3) The \gls*{dmri} data are warped using $U^{-1}_{true}$ to reproduce the effects of real
+  3) The \gls*{dmri} data are warped using $U_{true}$ to reproduce the effects of real
     susceptibility-derived distortions.
   Target diffusion scalars (\gls*{fa} and \gls*{adc}) are computed on the distorted data and
-    stacked to feed the multivariate input required by our algorithm.
-  4) \emph{Regseg} is run, obtaining a $U_{test} = \hat{U}_{true}$, the estimation of
+    stacked to feed the multivariate input required by \emph{regseg}.
+  4) the method is run, obtaining a $U_{test} = \hat{U}_{true}$, the estimation of
     the ground-truth deformation.

### 2015-04-07 Publications/2015-NeuroImage/05-results.tex
final improvements before sending to authors, removed ANOVA

@@
   \caption{A. Visual assessment of the results on the low resolution sets:
-    ``gyrus'' (top-left), ``L-shape'' (top-right), ``ball'' (bottom-left),
+    ``gyrus'' (top-left), ``L'' (top-right), ``ball'' (bottom-left),
     and ``box'' at (bottom-right).
@@
 		\footnotesize
+    \tabcolsep=0.1cm
     \begin{tabular}{lccccc}
     Res.   & ``gyrus'' & ``ball''  & ``box''   & ``L''     & Aggreg.    \\\hline
-    Low  & 0.59-0.60 & 0.65-0.76 & 0.68-0.71 & 0.67-0.77 & 0.64-0.66  \\
-    High & 0.18-0.38 & 0.31-0.45 & 0.34-0.42 & 0.34-0.40 & 0.34-0.38  \\
+    1.0mm  & 0.18-0.38 & 0.31-0.45 & 0.34-0.42 & 0.34-0.40 & 0.34-0.38  \\
+    2.0mm  & 0.59-0.60 & 0.65-0.76 & 0.68-0.71 & 0.67-0.77 & 0.64-0.66  \\
     \hline
     \end{tabular}
-    \caption{95\% Confidence Intervals (CIs) were computed using bootstrapping of $10^4$ samples,
-      for each phantom type and resolution.
-    Taking into account the non-normal distribution of errors, we used the median as location
-  		statistic.
-    All the CIs were below the corresponding half-voxel size in all the phantom types.}\label{tab:ci_phantom}
+    \caption{The vertex-wise Haussdorf distance between the ground-truth surfaces and their
+    corresponding estimation with \emph{regseg} presented distributions with 95\% CI below
+    the half-voxel size in all the phantom types.
+    CIs were computed using bootstrapping of 10$^4$ samples, and median as location statistic.}\label{tab:ci_phantom}
 \end{table}
@@
 Finally, we compared the performance of \emph{regseg} and the standard \gls*{t2b}
-  registration.
+  method.
+Visual assessment of the 16 cases is included in the \suppl{section S5}.
+In \autoref{fig:results_real}, box A, the visual report for one subject is supplied.
 We computed the \gls*{swindex} \eqref{eq:swindex} of every surface after registration,
   for both \emph{regseg} and the \gls*{t2b} methods.
-Finally, to statistically compare the results, we performed a one-way ANOVA test
-  on the warping indices for three specific surfaces of interest
-  ($\Gamma_{VdGM}$, $\Gamma_{WM}$, and $\Gamma_{pial}$).
-These interfaces are critical in analyses of whole-brain tractography.
-The one-way ANOVA evidenced that error distributions were significantly different,
-  $p \approx 2 \times 10^{-9}$.
+Finally, to statistically compare the results, we performed Kruskal-Wallis H-tests
+  (a non-parametric alternative to ANOVA) on the warping indices for three specific 
+  surfaces of interest selected in \autoref{sec:experiments_evaluation}.
+All the statistical tests evidenced that error distributions obtained with \emph{regseg} and
+  the \gls*{t2b} were significantly different, and using the violin plots in box B of
+  \autoref{fig:results_real} we observed that errors are always larger for \gls*{t2b}.
 We also reported the 95\% CIs of the \gls*{swindex} for those surfaces.
@@
   yielded an aggregate CI of 2.06 - 2.43 [mm].
-The quantitative results of the ANOVA tests and the CIs are presented in \autoref{tab:results_real}.
-Visual assessment of the 16 cases is included in the \suppl{section S5}.
-In \autoref{fig:results_real}, box A, the visual report for one subject is supplied.
-Box B presents the corresponding violin plots of error distributions by registration method
-  and surface.
+The results of the statistical tests and the CIs are summarized in \autoref{tab:results_real}.
 
@@
 		\footnotesize
-		\tabcolsep=0.05cm
+		\tabcolsep=0.08cm
     \begin{tabular}{cccccc}
@@
     \hline
-    \multirow{2}{*}{ANOVA}
-       & p-value  & 1.5$\times$10$^{-8}$& 6.3$\times$10$^{-16}$& 4.2$\times$10$^{-16}$ & 4$\times$10$^{-34}$ \\
-       & f-stat   & 60.03               & 255.82               & 263.97                & 377.82              \\
+    \multirow{2}{*}{H-tests}
+       & p-value  & 4.1$\times$10$^{-6}$& 2.3$\times$10$^{-6}$& 2.3$\times$10$^{-6}$ & 1.8$\times$10$^{-16}$ \\
+       & H-stat   & 21.20               & 22.31               & 22.31                & 67.85              \\
     \hline
     \end{tabular}
-    \caption{All the 95\% CIs of the \gls*{swindex} distributions computed for the
-      surfaces of interest were lower for \emph{regseg} with respect to the competing
-      method.
-    CIs reported in this table were computed using bootstrapping with mean as location
-      statistic and $10^4$ samples.
-    One-way-ANOVA tests indicated that there is a significant difference between our method and
-      the competing method of choice.
+    \caption{The statistical analysis of results on real data supported that \emph{regseg} overperformed
+    the alternative \acrfull{t2b} method.
+    The distribution of errors computed for the surfaces of interest ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$)
+      and the aggregate of all surfaces presented lower 95\% CIs for \emph{regseg}.
+    CIs reported in this table were computed using bootstrapping with mean as location statistic and 10$^4$ samples.
+    The Kruskal-Wallis H-tests indicated that there is a significant difference between \emph{regseg} and
+      the \gls*{t2b} method.
     }\label{tab:results_real}
@@
     }%
-	\caption{The evaluation instrument automatically renders the overlay between the
-	  target \gls*{fa} image and the resulting contours after distortion estimation (yellow color) and
-	  their theoretical location using the ground-truth (green color).
-	The first two rows show several axial slices for \emph{regseg} (``proposed'' row) and the
-	  competing method.
-	The last two rows represent the sagittal view.
-	We intentionally omitted the coronal slicing as it is the least informative, given the directional property
+	\caption{A. Example of one report for visual assessment, automatically generated by the evaluation instrument.
+    Each view shows one component of the input image (in this case, the \gls*{fa} map), the ground-truth location
+    of the surfaces (green contours), and the resulting surfaces with the method under test (yellow contours).
+  First two rows show axial slices for \emph{regseg} and the \acrfull*{t2b} method, and the last two rows
+    show corresponding sagittal views.
+  Coronal view is omitted since it is the least informative due to the the directional property
     of distortions.
-	Red arrows point to regions where the accuracy of the proposed method more clearly overperformed
-	  the competing method.
+	Red arrows point to regions where \emph{regseg} overperformed the \gls*{t2b} method.
+  B. Violin plots of error distributions of each surface, with indication of the voxel size of the \gls*{dmri} images
+    (1.25 mm), and supporting the improved results of \emph{regseg} in the proposed settings.
 	}\label{fig:results_real}

### 2015-04-07 Publications/2015-NeuroImage/06-discussion.tex
final improvements before sending to authors, removed ANOVA

@@
 The misregistration error remaining after \emph{regseg} was significantly lower ($p < 0.01$) than the
-  error corresponding to the \gls*{t2b} correction, using the one-way ANOVA test
+  error corresponding to the \gls*{t2b} correction, using Kruskal-Wallis H-tests
   (\autoref{tab:results_real}).

### 2015-04-08 Publications/2015-NeuroImage/05-results.tex
updating supplementals

@@
         \node[inner sep=10pt, draw=black!40](fig03b) at (0,7.9)
-          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure03-b}};
+          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure05A}};
         \node[inner sep=10pt, draw=black!40](fig03c) at (0,0)
-          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure03-c}};
+          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure05B}};
         \node[circle, text=black!75] at (-9.4,10.6) {\Huge \textbf{A}};
@@
         \node[inner sep=15pt, draw=black!40](fig05a) at (0,11)
-          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure05-a}};
+          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure06A}};
         \node[inner sep=15pt, draw=black!40](fig05b) at (0,0)
-          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure05-b}};
+          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure06B}};
         % \node[inner sep=5pt, draw=black!40](fig03c) at (0,0)

### 2015-04-08 Publications/2015-NeuroImage/supplemental.tex
updating supplementals

--- 2015-03-26 Publications/2015-NeuroImage/supplemental.tex
@@
 % @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-03-26 17:57:06
+% @Last Modified time: 2015-04-08 13:07:44
 
@@
 \usepackage{ifpdf}
-\graphicspath{{figures/},{../Figures/}}
+\graphicspath{{figures/},{supplemental/}}
 
@@
 % *** FLOAT PACKAGES ***
+\usepackage{floatrow}
 \usepackage[framemethod=tikz]{mdframed}
@@
 \usepackage[colorlinks=true,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}
+\usepackage{doi}
 
@@
 \begin{document}
-\title{Supplemental Materials: \emph{Simultaneous segmentation and registration of
-diffusion MR images of the brain driven by active-contours}}
+\title{Supplemental Materials: \\ \emph{Active-contours driven registration method for structure-informed segmentation of diffusion MR images}}
 \author{Oscar Esteban and Dominique Zosso}
@@
 \subsection{Gradient-descent optimization}\label{sec:gradient_descent}
-The energy functional to be optimized
-  in \emph{regseg} is presented in Eq. 7:
+The energy functional to be optimized in \emph{regseg} is presented in Eq. 7.
+After the simplifications described in equation (A.1) of Appendix 1, we obtain the
+  following energy functional:
   \begin{equation}
-  E(R \mid U) = C + \underbracket{\underset{l}{\sum} \int_{\Omega_l}
+  E(\vec{u}) = C + \underbracket{\underset{l}{\sum} \int_{\Omega_l}
   \mdist{f'}{l} \,d\vec{r}}_{\text{Data term } (E_{data})}
@@
 
-To search for the minimum of $E(R \mid U)$ w.r.t. the coefficients $\vec{u}_k$, we use a gradient descent strategy.
-In Eq. 11 we introduced the derivative of \eqref{eq:energy}:
+To search for the minimum of $E(\vec{u})$ w.r.t. the coefficients $\vec{u}_k$, we use a gradient descent strategy.
+In Eq. (9) we introduced the derivative of \eqref{eq:energy}:
   \begin{equation}
@@
 \begin{equation}
-  \vec{g}_k =
-  - \underset{i}{\sum} \bar{s}_i \cdot \psi_k(\vec{v}_i) \, \hat{\vec{e}} = - \underset{i}{\sum} \vec{g}_{i,k}.
+  \vec{g}_k = - \underset{i}{\sum} \bar{s}_i \cdot \psi_k(\vec{v}_i) \, \hat{\vec{e}} = - \underset{i}{\sum} \vec{g}_{i,k}.
   \label{eq:shape_gradient_final}
@@
   + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
-  \Big\} = \boldsymbol{\alpha} \cdot \vec{u}_k + \boldsymbol{\beta} \cdot \Delta \vec{u}_k.
+  \Big\} = \boldsymbol{\alpha} \cdot \vec{u}_k - \boldsymbol{\beta} \cdot \Delta \vec{u}_k.
   \label{eq:gradient_regterm}
@@
   final evolution equation:
-  \begin{equation}
-  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} = \vec{g}_k + \boldsymbol{\alpha} \cdot \vec{u}_k + \boldsymbol{\beta} \cdot \Delta \vec{u}_k,
+  \begin{align}
+  \frac{\partial \vec{u}}{\partial t} &\propto - \frac{\partial E(\vec{u})}{\partial \vec{u}_k}, \notag\\
+  \frac{\partial \vec{u}}{\partial t} &\propto - \Big( \vec{g}_k - \boldsymbol{\alpha} \cdot \vec{u}_k + \boldsymbol{\beta} \cdot \Delta \vec{u}_k \Big).
   \label{eq:gradient_final}
-  \end{equation}
+  \end{align}
 
-\subsection{Obtaining the update equation} To solve the differential equation in \eqref{eq:gradient_descent},
+\subsection{Obtaining the update equation} To solve the differential equation in \eqref{eq:gradient_final},
   we use a semi-implicit Euler scheme, referring to the discrete step size as $\delta$ and where the
@@
 \begin{align}
-  \vec{u}_k^{t+1} &= \vec{u}_k^{t} + \delta \, \Big( \vec{g}_k^{t+1} + (\boldsymbol{\alpha} - \boldsymbol{\beta} \Delta) \, \vec{u}_k^{t+1} \Big) \notag \\
-  (1 - \delta \boldsymbol{\alpha} - \delta \boldsymbol{\beta} \Delta)\, \vec{u}_k^{t+1} &= \vec{u}_k^{t} + \delta\, \vec{g}_k^{t}
+  \vec{u}_k^{t+1} &= \vec{u}_k^{t} + \delta \, \Big( - \vec{g}_k^{t} - ( \boldsymbol{\alpha} - \boldsymbol{\beta} \Delta) \, \vec{u}_k^{t+1} \Big) \notag \\
+  (1 + \delta \, \boldsymbol{\alpha} - \delta \, \boldsymbol{\beta} \Delta)\, \vec{u}_k^{t+1} &= \vec{u}_k^{t} - \delta\, \vec{g}_k^{t}
   \label{eq:discrete_derivative}
@@
   \begin{align}
-  \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\delta^{-1} \, \vec{u}_k^t + \vec{g}_k\} }%
-                  {\mathcal{F}\{(\delta^{-1} - \boldsymbol{\alpha})\, I-\boldsymbol{\beta}\Delta\}} \right\},
+  \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\delta^{-1} \, \vec{u}_k^t - \vec{g}_k^t\} }%
+                  {\mathcal{F}\{(\delta^{-1} + \boldsymbol{\alpha})\, I-\boldsymbol{\beta}\Delta\}} \right\},
   \label{eq:update_equation}
@@
   \begin{equation}
-  \Delta = \{\sum_{d=1}^n \mathcal{S}_d^- + \mathcal{S}_d^+\}  - 2n \mathcal{I}
+  \Delta = \Big\{\sum_{d=1}^n \mathcal{S}_d^- + \mathcal{S}_d^+\Big\}  - 2n \mathcal{I}
   \end{equation}
@@
 
+\newpage
 \section{Parameter settings and implementation details of \emph{regseg}}
@@
   it, $\boldsymbol{\Psi}$ can be pre-cached.
+
+\subsection{Assessment of the segmentation model}
+\begin{figure}[!ht]
+  \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,top},capbesidewidth=0.5\textwidth}}]{figure}[\FBwidth]
+  {\caption{\textbf{Assessment of the segmentation model}: 
+  Preliminarily, we investigated the aptness of the segmentation model.
+  For five test datasets, we uniformly sampled the space of distortions
+    $\hat{U} = \epsilon \cdot U_{true} = \vec{r} + \epsilon \, u_\text{PE}$
+    (with $\epsilon \in [-1.1, 1.1]$ and $u_\text{PE}$ from Equation 13 in the
+    paper), and evaluated the data-term of the cost function \eqref{eq:energy}.
+  The metric consistently displayed its minimum at the ground-truth location ($\epsilon=0.0$)
+    for all the cases, indicating that the segmentation model is appropriate for
+    registration.}}
+  {\includegraphics[width=\linewidth]{Suppl-figure04.pdf}}
+  \label{fig:energymap}
+\end{figure}
 
@@
 \begin{figure*}[b]
-  \includegraphics[width=\textwidth]{figures/Suppl-figure02.pdf}
+  \includegraphics[width=\linewidth]{Suppl-figure02.pdf}
   \caption{The evolution of the registration and segmentation process can be
@@
 
+\newpage
 \section{Instruments for evaluation}
@@
 All the intermediate results and figures in this paper have been encapsulated into
-  the workflows and are available in \url{http://dx.doi.org/zenodolink}.
+  the workflows and are available in \citep{esteban_acweregistration_2015}.
 
-An overview of the general workflow is presented in the Figure 2 of the paper.
-However, this figure has been extremely simplified for the best of visualization.
+An overview of the workflow applied in phantoms is presented in the Figure 2 of the paper.
+Subsequently, Figure 3 reproduces the evaluation on real data.
+The two figures have been extremely simplified for the best of visualization.
 In this section, we review the main elements of the evaluation pipelines.
 
-\subsection{Preprocessing}
+\subsection{Extraction of structural information}
+\begin{figure}[!ht]
+  \includegraphics[width=\linewidth]{surfaces.pdf}
+  \caption{\textbf{Generating Surfaces}.
+  Each surface extracted in the structural reference
+    (in both phantoms and real datasets) is obtained with a unitary pipeline,
+    for instance, the \emph{PialSurface} block.
+  In this figure, the composite workflow to extract the six surfaces that define the
+    segmentation model (\autoref{fig:jointplot}) is represented.
+  }\label{fig:wf_surfaces}
+\end{figure}
 
-\subsection{Distortion}
+\subsection{Processing diffusion MRI data}
+\begin{figure}[!ht]
+  \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,top},capbesidewidth=.5\linewidth}}]{figure}[\FBwidth]
+  {\includegraphics[width=.9\linewidth]{wf_dti.pdf}}
+  {\caption{\textbf{Workflow to generate \gls*{fa} and \gls*{adc} maps from \gls*{dmri} data}.
+    We use \emph{MRtrix} \citep{tournier_mrtrix_2012} to generate a \gls*{dti} from data and
+      computing the two feature maps of interest (\gls*{fa} and \gls*{adc}).
+    This workflow is performed in both undistorted and warped data.}\label{fig:wf_dti}}
+\end{figure}
 
+\subsection{T2-weighted registration to \emph{b0} -based (T2B) method}
+\begin{figure}[!ht]
+  \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,top},capbesidewidth=0.3\textwidth}}]{figure}[\FBwidth]
+  {\includegraphics[width=.85\linewidth]{wf_t2b.pdf}}
+  {\caption{\textbf{In-house implementation of the T2B method}.
+    Each surface extracted in the structural reference
+      (in both phantoms and real datasets) is obtained with a unitary pipeline,
+      for instance, the \emph{PialSurface} block.
+    In this figure, the composite workflow to extract the six surfaces that define the
+      segmentation model (\autoref{fig:jointplot}) is represented.}\label{fig:wf_t2b}}
+\end{figure}
+
+\subsection{Complete workflow for evaluation on real data}
+\begin{figure}[!ht]
+  \includegraphics[width=\linewidth]{hcp_wf.pdf}
+  \caption{\textbf{The \emph{nipype} design of the workflow for real data}.}\label{fig:wf_hcp}
+\end{figure}
+
+\newpage
 \section{Model considerations}
@@
 \begin{figure*}[!ht]
-	\includegraphics[width=\textwidth]{figures/Suppl-figure01.pdf}
+	\includegraphics[width=\linewidth]{Suppl-figure01.pdf}
 	\caption{Evaluating the joint distribution}\label{fig:jointplot}

### 2015-04-08 Publications/2015-NeuroImage/visual_report.sh

@@ -2,7 +2,7 @@
 
 rm -f suppl-realdata.tex
 
-for i in figures/HCP*.pdf; do
+for i in supplemental/HCP*.pdf; do
     fname=$(basename $i)
     fname=${fname::-4}
     echo "\\begin{figure}[!ht]" >> suppl-realdata.tex

### 2015-04-08 Publications/2015-NeuroImage/06-discussion.tex
add some missing dois

@@
 First, images are corrected for nonlinear distortions using auxiliary acquisitions
-  such as fieldmaps \citep{jezzard_correction_1995}, \gls*{t2} images \citep{kybic_fast_2003},
+  such as fieldmaps \citep{jezzard_correction_1995}, \gls*{t2} images \citep{kybic_unwarping_2000},
   etc.

### 2015-04-08 Code/Modules/Transform/include/SparseMatrixTransform.h
fixing segfault #210

@@ -338,7 +338,7 @@ protected:
 	FieldPointer          m_GradientField;
 
 	virtual void ComputeMatrix( WeightsMatrixType type, size_t dim = 0 );
-	virtual void AfterComputeMatrix( WeightsMatrixType type );
+	virtual void AfterComputeMatrix( WeightsMatrixType type, size_t dim = 0 );
 	virtual size_t ComputeRegionOfPoint(const PointType& point, VectorType& cvector, IndexType& start, IndexType& end, OffsetTableType offsetTable );
 
 	/** Support processing data in multiple threads. */

### 2015-04-08 Publications/2015-NeuroImage/05-results.tex
updates to export to text

@@
         \node[inner sep=10pt, draw=black!40](fig03b) at (0,7.9)
-          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure05A}};
+          {\hspace*{20pt}\includegraphics[width=\linewidth]{figure05A}};
         \node[inner sep=10pt, draw=black!40](fig03c) at (0,0)
-          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure05B}};
+          {\hspace*{20pt}\includegraphics[width=\linewidth]{figure05B}};
         \node[circle, text=black!75] at (-9.4,10.6) {\Huge \textbf{A}};
@@
         \node[inner sep=15pt, draw=black!40](fig05a) at (0,11)
-          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure06A}};
+          {\hspace*{20pt}\includegraphics[width=\linewidth]{figure06A}};
         \node[inner sep=15pt, draw=black!40](fig05b) at (0,0)
-          {\hspace*{20pt}\includegraphics[width=\linewidth]{figures/figure06B}};
+          {\hspace*{20pt}\includegraphics[width=\linewidth]{figure06B}};
         % \node[inner sep=5pt, draw=black!40](fig03c) at (0,0)
-        %   {\includegraphics[width=\linewidth]{figures/figure03-c}};
+        %   {\includegraphics[width=\linewidth]{figure03-c}};
         \node[circle, text=black!75] at (-9.5,17) {\Huge \textbf{A}};

### 2015-04-09 Publications/2015-NeuroImage/03-introduction.tex
added Gert's comments

@@
 Lastly, \gls*{dmri} images are acquired using \gls*{epi} to speed up acquisition
-  at the cost of presenting geometrical distortion and signal degradation and
+  at the cost of introducing geometrical distortion and signal degradation and
   destruction \citep{jezzard_correction_1995}.
@@
   hierarchical \citep{lu_segmentation_2008} clusterings,
-  and graph-cuts \citep{han_experimental_2009},
+  graph-cuts \citep{han_experimental_2009},
   and volume fraction modeling \citep{kumazawa_improvement_2013}.
@@
   the \emph{b0}.
-Similarly, an also linear registration method is \emph{bbregister} \citep{greve_accurate_2009}.
+Similarly, an also linear-registration method is \emph{bbregister} \citep{greve_accurate_2009}.
 It uses active contours with edges to look for intensity boundaries in the \emph{b0}
@@
   regions typically warped by the artifact.
-As the distortion is not accounted for, it should be addressed separately.
+As the distortion is not accounted for, it must be addressed separately.
 Nonlinear registration has been successfully performed between \gls*{t2} and \emph{b0}
@@
 Optimization is performed using a descent strategy of explicit shape-gradients
-  \citep{besson_dream2s_2003,herbulot_segmentation_2006}, therefore \emph{regseg}
-  does not implement levelsets as most of the presented methods do.
+  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
+Hence, unlike the most of presented methods, \emph{regseg} does not implement levelsets.
 Since the nonlinear distortion is aligned with one of the imaging axis (see 

### 2015-04-09 Publications/2015-NeuroImage/04-methods.tex
updates to export to text
added Gert's comments

@@
 \begin{figure}
-	\includegraphics[width=\linewidth]{figures/figure01}
+	\includegraphics[width=\linewidth]{figure01}
 	\caption{The active contours are defined as the interfacing surfaces of the competing
@@
 The computation of the explicit shape gradients at each $\vec{v}_i'$ is described in \autoref{fig:method}.
-The intermediate steps to get to \eqref{eq:final_gradient} from \eqref{eq:gradient_descent} are
-  detailed in the \suppl{equations SM6-SM12}:
+Then, \eqref{eq:gradient_descent} can be reformulated as (see \suppl{equations SM6-SM12}):
 
@@
 The reference surfaces $\gammaset_R$ are extracted from the binary shapes
-  using \emph{FreeSurfer} tools.
+  using \emph{FreeSurfer} tools \citep{fischl_freesurfer_2012}.
 The ground-truth distortion is generated with a chain of two displacement 
@@
 The warped surfaces $\gammaset_\text{true}$ are binarized to generate tissue fractions
-  in low (\isores[mm]{2.0}) and high (\isores[mm]{1.0}).
+  in low (\isores[mm]{2.0}) and high (\isores[mm]{1.0}) resolution.
 Using a \gls*{mr} simulator \citep{caruyer_phantomas_2014}, we synthesize
@@
 \begin{figure*}
-\includegraphics[width=\linewidth]{figures/figure02}
+\includegraphics[width=\linewidth]{figure02}
 \caption{The evaluation of \emph{regseg} on phantom data is performed with the presented instrumental workflow:
@@
 \begin{figure*}
-\includegraphics[width=\linewidth]{figures/figure03}
+\includegraphics[width=\linewidth]{figure03}
 \caption{Experimental workflow real data from the \acrfull*{hcp}.
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{figures/figure04-fa-md-jointplot}
+\includegraphics[width=\linewidth]{figure04}
 \caption{Segmentation model defined by the homogeneous regions $\Omega_i$ in which
@@
   from \gls*{dmri} data.
-After evaluating several alternative models, we empirically defined the partition \omegaset{} 
-  of the reference image $R$ as follows:
-  1) $\Omega_{Tha}$ including the thalamus;
-  2) $\Omega_{VdGM}$ including the ventricular system and deep \gls*{gm} structures;
-  3) $\Omega_{WM}$ including the cerebral \gls*{wm};
-  4) $\Omega_{bst}$ including the brain stem and cerebellar \gls*{wm};
-	5) $\Omega_{cbGM}$ including the cerebellar \gls*{gm}; and
-	6) $\Omega_{GM}$ including the cortical \gls*{gm}.
+After evaluating several alternative models, we empirically defined a partition \omegaset{}
+  into the following six regions:
+  1) the thalamus ($\Omega_{Tha}$);
+  2) the ventricular system and deep \gls*{gm} structures ($\Omega_{VdGM}$);
+  3) the cerebral \gls*{wm} ($\Omega_{WM}$);
+  4) the brain stem and cerebellar \gls*{wm} ($\Omega_{bst}$);
+	5) the cerebellar \gls*{gm} ($\Omega_{cbGM}$); and
+	6) the cortical \gls*{gm} ($\Omega_{GM}$).
 Using tools of \emph{FreeSurfer}, and the appropriate selections of labels in the 

### 2015-04-09 Publications/2015-NeuroImage/08-acknowledgements.tex
added Gert's comments

--- 2015-03-13 Publications/2015-NeuroImage/08-acknowledgements.tex
@@
 \section*{Acknowledgments}
-The authors thank V. Estellers for early discussions at the beginning of this project,
+The authors thank Y. Alem\'an and G. Wollny for their thorough reviews of the manuscript,
+  V. Estellers for early discussions at the beginning of this project,
   and L. A. Vese for her support during OE's research visits in her laboratory.
@@
 
-DZ is supported by the Swiss National Science Foundation under grant PBELP2-137727.
+DZ was supported by the Swiss National Science Foundation under grants PBELP2\_137727 
+  and P300P2\_147778, and NSF-DMS 1418812
 This study is supported by: the Spain's Ministry of Science and Innovation
-  (projects TEC2013-48251-C2-2-R and INNPACTO XIORT), Comunidad de Madrid (TOPUS) and
+  (projects TEC-2013-48251-C2-2-R and INNPACTO XIORT), Comunidad de Madrid (TOPUS) and
   European Regional Development Funds; the Center for Biomedical Imaging

### 2015-04-09 Publications/2015-NeuroImage/05-results.tex
updates and standalone figures

@@
   \centering
-    \resizebox{\textwidth}{!}{%
-      \begin{tikzpicture}
-        \node[inner sep=10pt, draw=black!40](fig03b) at (0,7.9)
-          {\hspace*{20pt}\includegraphics[width=\linewidth]{figure05A}};
-        \node[inner sep=10pt, draw=black!40](fig03c) at (0,0)
-          {\hspace*{20pt}\includegraphics[width=\linewidth]{figure05B}};
-        \node[circle, text=black!75] at (-9.4,10.6) {\Huge \textbf{A}};
-        \node[circle, text=black!75] at (-9.4,3.5) {\Huge \textbf{B}};
-      \end{tikzpicture}
-    }%
+  \includestandalone[width=\linewidth]{figure05}
   \caption{A. Visual assessment of the results on the low resolution sets:
@@
   \centering
-    \resizebox{\textwidth}{!}{%
-      \begin{tikzpicture}
-        \node[inner sep=15pt, draw=black!40](fig05a) at (0,11)
-          {\hspace*{20pt}\includegraphics[width=\linewidth]{figure06A}};
-        \node[inner sep=15pt, draw=black!40](fig05b) at (0,0)
-          {\hspace*{20pt}\includegraphics[width=\linewidth]{figure06B}};
-        % \node[inner sep=5pt, draw=black!40](fig03c) at (0,0)
-        %   {\includegraphics[width=\linewidth]{figure03-c}};
-        \node[circle, text=black!75] at (-9.5,17) {\Huge \textbf{A}};
-        \node[circle, text=black!75] at (-9.5,3) {\Huge \textbf{B}};
-        % \node[circle, text=black!75] at (-8.8,3.4) {\Huge \textbf{C}};
-      \end{tikzpicture}
-    }%
+  \includestandalone[width=\linewidth]{figure06}
   \caption{A. Example of one report for visual assessment, automatically generated by the evaluation instrument.

### 2015-04-09 Publications/2015-NeuroImage/08-acknowledgements.tex
updates and standalone figures

@@
 
-DZ was supported by the Swiss National Science Foundation under grants PBELP2\_137727 
-  and P300P2\_147778, and NSF-DMS 1418812
+DZ was supported by the Swiss National Science Foundation under grants PBELP2-137727 
+  and P300P2-147778, and NSF-DMS 1418812
 This study is supported by: the Spain's Ministry of Science and Innovation

### 2015-04-09 Publications/2015-NeuroImage/figure05.tex
updates and standalone figures

% -*- root: 00-main.tex -*-
% @Author: Oscar Esteban
% @Date:   2015-04-09 13:06:32
% @Last Modified by:   Oscar Esteban
% @Last Modified time: 2015-04-09 15:52:17

\documentclass[tikz]{standalone}
\usepackage{graphicx}
\graphicspath{{figures/}}

\usepackage[T1]{fontenc}
\usepackage{charter}

\begin{document}
\begin{tikzpicture}
  \node[inner sep=10pt, draw=black!40](fig03b) at (0,8.8)
    {\hspace*{20pt}\includegraphics[width=21cm]{figure05A}};
  \node[inner sep=10pt, draw=black!40](fig03c) at (0,0)
    {\hspace*{20pt}\includegraphics[width=21cm]{figure05B}};
  \node[circle, text=black!75, inner sep=0] at (-10.65,12.0) {\Huge \textbf{A}};
  \node[circle, text=black!75, inner sep=0] at (-10.65,4.05) {\Huge \textbf{B}};
\end{tikzpicture}
\end{document}

### 2015-04-09 Publications/2015-NeuroImage/figure06.tex
updates and standalone figures

% -*- root: 00-main.tex -*-
% @Author: Oscar Esteban
% @Date:   2015-04-09 14:39:30
% @Last Modified by:   Oscar Esteban
% @Last Modified time: 2015-04-09 15:54:48

\documentclass[tikz]{standalone}

\thispagestyle{empty}
\usepackage{graphicx}
\graphicspath{{figures/}}

\usepackage[T1]{fontenc}
\usepackage{charter}

\begin{document}
  \begin{tikzpicture}
    \node[inner sep=15pt, draw=black!40](fig05a) at (0,12.0)
      {\hspace*{20pt}\includegraphics[width=21cm]{figure06A}};
    \node[inner sep=15pt, draw=black!40](fig05b) at (0,0)
      {\hspace*{20pt}\includegraphics[width=21cm]{figure06B}};
    \node[circle, text=black!75, inner sep=0] at (-10.65,18.6) {\Huge \textbf{A}};
    \node[circle, text=black!75, inner sep=0] at (-10.65,3.5) {\Huge \textbf{B}};
  \end{tikzpicture}
\end{document}

### 2015-04-10 Publications/2015-NeuroImage/02-abstract.tex
regseg label

--- 2015-03-20 Publications/2015-NeuroImage/02-abstract.tex
@@
 % -*- root: 00-main.tex -*-
-We propose a simultaneous segmentation and registration method called \emph{regseg}
+Current applications of whole-brain tractography on \acrlong*{dmri} data require highly precise
+  delineations of anatomical structures, usually projected from anatomical \acrlong*{t1} images
+  through registration.
+We propose \regseg{}, a simultaneous segmentation and registration method
   that uses active contours without edges extracted from structural images.
-The contours evolve through a free-deformation field of B-Spline functions that
-  optimally maps the surfaces onto the data in the target space.
-\emph{Regseg} is intended, but not limited, to correct susceptibility-derived distortions
-  in echo-planar imaging.
-We proof the concept behind \emph{regseg} on four digital phantoms warped with random
-  deformations, and conclude that the precision of the segmentation is subvoxel.
-Finally, we demonstrate \emph{regseg} in the correction task on 16 real diffusion MRI
-  datasets from the \acrlong*{hcp}.
-The results suggest that \emph{regseg} overperformed an extended competing method.
-We compute the geometrical disagreement of the surfaces registered with \emph{regseg} with
-  respect to a ground truth, obtaining a 95\% confidence interval of $[1.08 - 1.50]mm$ in images 
-  of $1.25mm$ isotropic resolution.
+The contours evolve through a free-form deformation field supported by B-Spline basis
+  optimally mapping the contours onto the data in the target space.
+We tested the functionality of \regseg{} on four digital phantoms warped with known and
+  randomly generated deformations, to conclude that the accuracy achieved is subvoxel.
+We then apply \regseg{} to perform registration/segmentation task on 16 real diffusion MRI
+  datasets from the \acrlong*{hcp}, warped with realistic and nonlinear distortions typically
+  present in these data.
+We compute the misregistration error of the contours estimated by \regseg{} with respect
+  to their theoretical location using the ground-truth, obtaining a 95\% CI of 0.56 - 0.66 (mm)
+  distance between corresponding mesh vertices, below the 1.25mm isotropic resolution of the images.
+We also compared the performance to a widely-used registration tool, to find out
+  that \regseg{} overperformed the alternate method in our settings.

### 2015-04-10 Publications/2015-NeuroImage/03-introduction.tex
regseg label

@@
   accuracy of the process.
-We propose a tool called \emph{regseg} to exploit the prior information of shapes
+We propose a tool called \regseg{} to exploit the prior information of shapes
   extracted on \gls*{t1} images to register the anatomical reference
@@
   \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
-Hence, unlike the most of presented methods, \emph{regseg} does not implement levelsets.
+Hence, unlike the most of presented methods, \regseg{} does not implement levelsets.
 Since the nonlinear distortion is aligned with one of the imaging axis (see 
-  \autoref{sec:human_connectome}), \emph{regseg} includes an anisotropic regularizer of
+  \autoref{sec:human_connectome}), \regseg{} includes an anisotropic regularizer of
   the displacements field proposed by \cite{nagel_investigation_1986}.
-Finally, we evaluate \emph{regseg} using an extension of our instrumentation framework
+Finally, we evaluate \regseg{} using an extension of our instrumentation framework
   \citep{esteban_simulationbased_2014} that simulates known and realistic distortions

### 2015-04-10 Publications/2015-NeuroImage/04-methods.tex
updates and standalone figures
add missing comma after formula
regseg label

@@
   P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
-  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) = \notag\\
-  &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
+  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) = % \notag\\ &=
+  \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
   \mdist{f'}{l} \right)}},
@@
   P(U) = \underset{\vec{r}}{\prod}\, p(\vec{u}) &=
-  \underset{\vec{r}}{\prod}\, p_0(\vec{u}) \, p_1(\vec{u}), \\
+  \underset{\vec{r}}{\prod}\, p_0(\vec{u}) \, p_1(\vec{u}), \text{ with} \\
   p_0(\vec{u}) &= \mathcal{N}( \vec{u} \mid 0, \mathbf{A}^{-1}), \notag\\
@@
   \underset{\vec{r} \in \Omega_l}{\prod}
-  \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) \notag\\
-  &= \const + \underset{l}{\sum} \int_{\Omega_l}
-  \mdist{f'}{l} \,d\vec{r} \, + \notag\\
-  &+ \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
+  \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) = \notag\\ &=
+  \const + \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, + % \notag\\ &+
+  \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
   \label{eq:energy}
@@
   \frac{ \partial }{\partial \vec{u}_k} \Big\{
-  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, + \notag\\
-  &+ \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
+  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, + % \notag\\ &+
+  \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
   + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
@@
 \label{sec:conv_report}
-The \emph{regseg} tool includes a multiresolution strategy on the free-form deformation field.
+The \regseg{} tool includes a multiresolution strategy on the free-form deformation field.
 Registration pyramids are created by setting the spacing between control points of the B-Spline basis
@@
   reports (as the one found in the \suppl{section S1.3}).
-We release \emph{regseg} along with the tool for generating the convergence report.
+We release \regseg{} along with the tool for generating the convergence report.
 
 \subsection{Evaluation protocol}\label{sec:evaluation_protocol}
-In order to assess the performance of \emph{regseg}, we defined a general
+In order to assess the performance of \regseg{}, we defined a general
   evaluation protocol:
@@
   generate warped images ($M$) to segment;
-3) execution of \emph{regseg} with $\gammaset_R$ and the warped data as inputs; and
+3) execution of \regseg{} with $\gammaset_R$ and the warped data as inputs; and
 4) visual assessment and computation of error metrics.
@@
 The workflow to simulate the digital phantoms and assess
-  \emph{regseg} on them is presented in \autoref{fig:evphantoms}.
+  \regseg{} on them is presented in \autoref{fig:evphantoms}.
 A set of four binary objects (called ``box'', ``ball'', ``L'',
@@
 The field of view at both resolutions is \isores[mm]{100}.
-Then, \emph{regseg} is applied to map $\gammaset_R$ into the warped phantoms to
+Then, \regseg{} is applied to map $\gammaset_R$ into the warped phantoms to
   obtain the registered surfaces ($\hat{\gammaset}_{test}$).
@@
 \includegraphics[width=\linewidth]{figure02}
-\caption{The evaluation of \emph{regseg} on phantom data is performed with the presented instrumental workflow:
+\caption{The evaluation of \regseg{} on phantom data is performed with the presented instrumental workflow:
   1) The reference surfaces $\gammaset_R$ are triangularized meshes extracted from the four binary shapes (``box'', ``ball'', ``L'', ``gyrus'');
@@
      two available resolutions;
-  5) The \emph{regseg} is run, using the warped test images as moving multispectral-image and $\gammaset_R$ as shape-priors;
-  6) Assessment of the agreement between the surfaces fitted with \emph{regseg} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$
+  5) The \regseg{} is run, using the warped test images as moving multispectral-image and $\gammaset_R$ as shape-priors;
+  6) Assessment of the agreement between the surfaces fitted with \regseg{} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$
      with automatically generated visual reports and computing the Hausdorff distance between corresponding surfaces.}\label{fig:evphantoms}
@@
   Target diffusion scalars (\gls*{fa} and \gls*{adc}) are computed on the distorted data and
-    stacked to feed the multivariate input required by \emph{regseg}.
+    stacked to feed the multivariate input required by \regseg{}.
   4) the method is run, obtaining a $U_{test} = \hat{U}_{true}$, the estimation of
@@
 \paragraph*{Data}
-For the evaluation of \emph{regseg} on real \gls*{dmri} data of human brains,
+For the evaluation of \regseg{} on real \gls*{dmri} data of human brains,
   we collected 16 subjects from the \gls*{hcp} database.
@@
   sWI = \frac{1}{\sum_i a_i} \sum\limits_i^{N_V} a_i\,\|
-  \vec{v}_i - \hat{\vec{v}}_i \|
+  \vec{v}_i - \hat{\vec{v}}_i \|,
   \label{eq:swindex}

### 2015-04-10 Publications/2015-NeuroImage/05-results.tex
regseg label

@@
   (see \autoref{sec:experiments_evaluation}) between the theoretical $\gammaset_\text{true}$ and
-  the estimation done by \emph{regseg} ($\hat{\gammaset}_{test}$).
+  the estimation done by \regseg{} ($\hat{\gammaset}_{test}$).
 The results showed a consistent and high accuracy, below the image resolution.
@@
     \caption{The vertex-wise Haussdorf distance between the ground-truth surfaces and their
-    corresponding estimation with \emph{regseg} presented distributions with 95\% CI below
+    corresponding estimation with \regseg{} presented distributions with 95\% CI below
     the half-voxel size in all the phantom types.
@@
 %
-Finally, we compared the performance of \emph{regseg} and the standard \gls*{t2b}
+Finally, we compared the performance of \regseg{} and the standard \gls*{t2b}
   method.
@@
 We computed the \gls*{swindex} \eqref{eq:swindex} of every surface after registration,
-  for both \emph{regseg} and the \gls*{t2b} methods.
+  for both \regseg{} and the \gls*{t2b} methods.
 Finally, to statistically compare the results, we performed Kruskal-Wallis H-tests
@@
   surfaces of interest selected in \autoref{sec:experiments_evaluation}.
-All the statistical tests evidenced that error distributions obtained with \emph{regseg} and
+All the statistical tests evidenced that error distributions obtained with \regseg{} and
   the \gls*{t2b} were significantly different, and using the violin plots in box B of
@@
 We also reported the 95\% CIs of the \gls*{swindex} for those surfaces.
-The aggregate CI for \emph{regseg} was 1.08 - 1.50 [mm], whereas the \gls*{t2b} method
+The aggregate CI for \regseg{} was 1.08 - 1.50 [mm], whereas the \gls*{t2b} method
   yielded an aggregate CI of 2.06 - 2.43 [mm].
@@
     \multirow{2}{*}{CI}
-       & \emph{regseg}        & 0.50 - 0.78 & 0.50 - 0.55 & 0.66 - 0.73 & 0.56 - 0.66 \\
+       & \regseg{}        & 0.50 - 0.78 & 0.50 - 0.55 & 0.66 - 0.73 & 0.56 - 0.66 \\
        & T2B                  & 1.78 - 2.58 & 1.94 - 2.36 & 2.16 - 2.58 & 2.05 - 2.39 \\
@@
     \end{tabular}
-    \caption{The statistical analysis of results on real data supported that \emph{regseg} overperformed
+    \caption{The statistical analysis of results on real data supported that \regseg{} overperformed
     the alternative \acrfull{t2b} method.
     The distribution of errors computed for the surfaces of interest ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$)
-      and the aggregate of all surfaces presented lower 95\% CIs for \emph{regseg}.
+      and the aggregate of all surfaces presented lower 95\% CIs for \regseg{}.
     CIs reported in this table were computed using bootstrapping with mean as location statistic and 10$^4$ samples.
-    The Kruskal-Wallis H-tests indicated that there is a significant difference between \emph{regseg} and
+    The Kruskal-Wallis H-tests indicated that there is a significant difference between \regseg{} and
       the \gls*{t2b} method.
@@
     of the surfaces (green contours), and the resulting surfaces with the method under test (yellow contours).
-  First two rows show axial slices for \emph{regseg} and the \acrfull*{t2b} method, and the last two rows
+  First two rows show axial slices for \regseg{} and the \acrfull*{t2b} method, and the last two rows
     show corresponding sagittal views.
@@
     of distortions.
-	Red arrows point to regions where \emph{regseg} overperformed the \gls*{t2b} method.
+  Certain regions where \regseg{} overperformed the \gls*{t2b} have been zoomed in.
   B. Violin plots of error distributions of each surface, with indication of the voxel size of the \gls*{dmri} images
-    (1.25 mm), and supporting the improved results of \emph{regseg} in the proposed settings.
+    (1.25 mm), and supporting the improved results of \regseg{} in the proposed settings.
 	}\label{fig:results_real}

### 2015-04-10 Publications/2015-NeuroImage/06-discussion.tex
added Gert's comments
regseg label

@@
   and use them to simulate \gls*{t1} and \gls*{t2} images at two resolution levels.
-After registration with \emph{regseg} we measured the Hausdorff distance between the
+After registration with \regseg{} we measured the Hausdorff distance between the
   projected contours using the ground-truth warping and our estimation.
@@
 \paragraph*{Application on real data}
-We designed \emph{regseg} as a method to segment \gls*{dmri} data exploiting the
+We designed \regseg{} as a method to segment \gls*{dmri} data exploiting the
   anatomical information extracted in the corresponding \gls*{t1} image of the subject.
@@
 This situation is commonly found in historical datasets.
-Beyond the proposed application in \gls*{dmri} data, other potential uses of \emph{regseg} are
-  atlas-based segmentation and tracking objects in time-series.
 
-We evaluated \emph{regseg} in a real environment, using the experimental framework presented
+We evaluated \regseg{} in a real environment, using the experimental framework presented
   in \autoref{fig:evworkflows}.
-We processed 16 subjects from the \gls*{hcp} database with both \emph{regseg}
+We processed 16 subjects from the \gls*{hcp} database with both \regseg{}
   and an in-house replication of the \gls*{t2b} correction method.
@@
   bellow the pixel size of 1.25mm.
-The misregistration error remaining after \emph{regseg} was significantly lower ($p < 0.01$) than the
+The misregistration error remaining after \regseg{} was significantly lower ($p < 0.01$) than the
   error corresponding to the \gls*{t2b} correction, using Kruskal-Wallis H-tests
@@
 Visual inspection of all the results (\suppl{section S5}) and the violin plots included in
-  \autoref{fig:results_real} confirmed that \emph{regseg} overperformed the \gls*{t2b} method
+  \autoref{fig:results_real} confirmed that \regseg{} overperformed the \gls*{t2b} method
   in our settings.
@@
   the so-called \emph{instrumentation bias} \citep{tustison_instrumentation_2013}.
-Therefore, these results do not prove that the \emph{regseg} software \emph{is better than} the
+Therefore, these results do not prove that the \regseg{} software \emph{is better than} the
   \gls*{t2b} method.
-The results reported let us propose \emph{regseg} as a solid option in the application field.
+The results reported let us propose \regseg{} as a solid option in the application field.
 Additionally, the \gls*{t2b} may introduce an additional (and small) error in the necessary
@@
 % First extensions of this work will study more appropriate features to build the energy functional
-%   on, enabling to perform \emph{regseg} directly on the raw \gls*{dmri} data.
+%   on, enabling to perform \regseg{} directly on the raw \gls*{dmri} data.
 % A second outlook covers incorporating knowledge about the distortion by initializing our method

### 2015-04-10 Publications/2015-NeuroImage/07-conclusion.tex
added Gert's comments
regseg label

@@
   homogeneous regions within the image.
-We tested \emph{regseg} on digital phantoms simulating \gls*{t1} and \gls*{t2} \gls*{mri},
+We tested \regseg{} on digital phantoms simulating \gls*{t1} and \gls*{t2} \gls*{mri},
 	warped with smooth random deformations.
@@
 
-We propose \emph{regseg} to simultaneously segment \gls*{dmri} data and register them to
+We propose \regseg{} to simultaneously segment \gls*{dmri} data and register them to
   their corresponding \gls*{t1} image of the same subject.
-Visual assessment of \emph{regseg} and cross-comparison with a widely-used technique demonstrated 
+Visual assessment of \regseg{} and cross-comparison with a widely-used technique demonstrated 
   its accuracy.
-Moreover, \emph{regseg} does not require additional images to the minimal acquisition protocol
+Moreover, \regseg{} does not require additional images to the minimal acquisition protocol
   that includes only \gls*{t1} and \gls*{dmri}.
+Beyond the proposed application in \gls*{dmri} data, other potential uses of \regseg{} are
+  atlas-based segmentation and tracking objects in time-series.
 

### 2015-04-10 Publications/2015-NeuroImage/06-discussion.tex
add regseg label uppercase

@@
   registration \citep{greve_accurate_2009}.
-\emph{Regseg} solves this joint problem in one single step, and it does not require any additional
+\Regseg{} solves this joint problem in one single step, and it does not require any additional
   acquisition over the minimal protocol including \gls*{t1} and \gls*{dmri} images only.
@@
   and an in-house replication of the \gls*{t2b} correction method.
-\emph{Regseg} showed very high accuracy, with an aggregate 95\% \gls*{ci} of 0.56-0.66 [mm],
+\Regseg{} showed very high accuracy, with an aggregate 95\% \gls*{ci} of 0.56-0.66 [mm],
   bellow the pixel size of 1.25mm.

### 2015-04-10 Publications/2015-NeuroImage/07-conclusion.tex
add regseg label uppercase

@@
 
-\emph{Regseg} is a variational framework to simultaneously segment and
+\Regseg{} is a variational framework to simultaneously segment and
   register 3D \gls*{dmri} data of the human brain, using within-subject
@@
   of the target image.
-\emph{Regseg} uses active contours without edges, and hence looks for
+\Regseg{} uses active contours without edges, and hence looks for
   homogeneous regions within the image.
@@
   involved in this paper is also publicly released.
-\emph{Regseg} is implemented on top of ITK-4.6 (Insight Registration and 
+\Regseg{} is implemented on top of ITK-4.6 (Insight Registration and 
   Segmentation Toolkit, \url{http://www.itk.org}).

### 2015-04-13 Publications/2015-NeuroImage/03-introduction.tex
minor fixes

@@
   tract-based spatial statistics \citep{smith_tractbased_2006}.
-However, segmenting \gls*{dmri} data precisely and coregistration of anatomical
+However, segmenting \gls*{dmri} data precisely, and coregistration of anatomical
   images are difficult due to several limitations.
@@
 The artifact has been proven to cause an important impact on the anatomy extracted
-  from \gls*{dmri} \citep{irfanoglu_effects_2012}, particularly in certain fiber bundles
-  when used in whole-brain tractography.
+  from \gls*{dmri}, particularly in certain fiber bundles \citep{irfanoglu_effects_2012}.
 These limitations preclude segmentation in native \gls*{dmri} space, including

### 2015-04-15 Publications/2015-NeuroImage/00-main.tex
change kw
updates to export to text
added Gert's comments
updates and standalone figures
regseg label
add regseg label uppercase
add Meri's comments

@@
-\documentclass[5p,authoryear]{elsarticle}
-% \documentclass[authoryear,onecolumn]{elsarticle}
+% \documentclass[5p,authoryear,fleqn]{elsarticle}
+\documentclass[3p,authoryear,fleqn]{elsarticle}
 \journal{ArXiv archiving for pre-publication review}
@@
 % *** MISC UTILITY PACKAGES ***
+\usepackage[mode=buildnew]{standalone}
+\usepackage{gincltex}
 \usepackage{glossaries}
@@
 \usepackage{xspace}
+\usepackage{graphicx}
+\graphicspath{{figures/}}
 
@@
 \providecommand{\gammaset}{\ensuremath{\boldsymbol{\Gamma}}}
+\providecommand{\regseg}{\emph{regseg}}
+\providecommand{\Regseg}{\emph{Regseg}}
 
@@
 %\renewcommand*{\familydefault}{\sfdefault}
-\usepackage[switch]{lineno}
+
 \usepackage{array}
 \newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+
+\usepackage{ifpdf}
+
+\ifpdf
+  \usepackage[switch]{lineno}
+  % nothing to do
+\else
+  \makeatletter
+  \def\temp{dvips.def}
+  \ifx\Gin@driver\temp
+  \def\Ginclude@graphics#1{\def\temp{#1}---image \expandafter\strip@prefix\meaning\temp---}
+  \fi
+  \makeatother
+\fi
 
@@
 F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland}
-\address[chuv]{Dept. of Radiology, University
+\address[chuv]{Dept. of Radiology, CIBM, University
 Hospital Center (CHUV) and University of Lausanne (UNIL), Lausanne, Switzerland}
@@
 diffusion MRI \sep cortical~parcellation \sep segmentation %
-\sep nonlinear~registration \sep active~contours \sep echo-planar imaging.
+\sep nonlinear~registration \sep active~contours \sep susceptibility~distortion.
 \end{keyword}

### 2015-04-15 Publications/2015-NeuroImage/03-introduction.tex
add Meri's comments

@@
 \section{Introduction}\label{sec:introduction}
+% Current applications to investigate the fiber pathways of the brain
+%   using \gls*{dmri}, such as tract-based spatial statistics \citep{smith_tractbased_2006}
+%   and connectome extraction \citep{craddock_imaging_2013}, critically rely on the fusion of 
+%   prior anatomical information extracted from a \gls*{t1} image of the same subject.
 The accurate delineation of \gls*{wm} in \gls*{dmri} and the fusion of prior
@@
   images are difficult due to several limitations.
-First, \gls{dmri} images suffer from partial voluming of several tissues in 
-  voxels located at their interfaces due to the low resolution of images
-  (generally around \isores[mm]{2.2}).
+First, \gls{dmri} images present a resolution much higher than the imaged
+  microstructural features \citep{basser_microstructural_1996}.
+Therefore, voxels located at structural discontinuities are affected by partial
+  voluming of sources of signal.
 Second, \gls*{dmri} schemes probe the diffusion process within the brain in 
@@
 However, the mask and subsequent analyses substantially depend on the threshold set
-  \citep{taoka_fractional_2009} owing to the large variability of \gls*{fa} and its
-  decay with the increase of incoherence between fibers.
+  \citep{taoka_fractional_2009}.
+To overcome the unreliability of of \gls*{fa} thresholding,
   \cite{zhukov_level_2003} proposed active contours with edges represented
@@
   to the \gls*{dmri} data could provide the optimal solution.
-Previously, joint segmentation and registration has been successfully applied to
-  longitudinal object tracking \citep{paragios_level_2003} and atlas-based
+As far as we know, such strategy has never been proposed for the application at hand.
+Previously, joint segmentation and registration has been successfully applied to other problems.
+For instance, longitudinal object tracking \citep{paragios_level_2003} and atlas-based
   segmentation \citep{gorthi_active_2011}.

### 2015-04-16 Publications/2015-NeuroImage/05-results.tex
some final fixes

@@
 Finally, to statistically compare the results, we performed Kruskal-Wallis H-tests
-  (a non-parametric alternative to ANOVA) on the warping indices for three specific 
-  surfaces of interest selected in \autoref{sec:experiments_evaluation}.
+  (a non-parametric alternative to ANOVA) on the warping indices for the three surfaces of 
+  interest selected in \autoref{sec:experiments_evaluation}
+  ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$).
 All the statistical tests evidenced that error distributions obtained with \regseg{} and
-  the \gls*{t2b} were significantly different, and using the violin plots in box B of
+  \gls*{t2b} were significantly different, and using the violin plots in box B of
   \autoref{fig:results_real} we observed that errors are always larger for \gls*{t2b}.
@@
     \begin{tabular}{cccccc}
-    & & $\Gamma_{VdGM}$  & $\Gamma_{WM}$ & $\Gamma_{pial}$ & Aggregate \\
+    & & $\Gamma_{VdGM}$  & $\Gamma_{WM}$ & $\Gamma_{pial}$ & Aggreg. \\
     \hline
@@
     The distribution of errors computed for the surfaces of interest ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$)
-      and the aggregate of all surfaces presented lower 95\% CIs for \regseg{}.
+      and the aggregate of all surfaces (Aggreg. column) presented lower 95\% CIs for \regseg{}.
     CIs reported in this table were computed using bootstrapping with mean as location statistic and 10$^4$ samples.
-    The Kruskal-Wallis H-tests indicated that there is a significant difference between \regseg{} and
+    The Kruskal-Wallis H-tests indicated that there was a significant difference between \regseg{} and
       the \gls*{t2b} method.

### 2015-04-16 Publications/2015-NeuroImage/06-discussion.tex
some final fixes

@@
 The hypothesis underpinning our research was that image registration can be reliably performed
-  by searching for homogeneous regions in the target image that correspond to precise contours
+  by searching for homogeneous regions in a multispectral image that correspond to precise contours
   from an atlas, or extracted from other image (i.e. a different time step).
@@
   deformation supported by B-Spline basis functions using digital phantoms.
-We randomly deformed 4 different phantom models mimicking three homogeneous regions,
+We randomly deformed four different phantom models mimicking three homogeneous regions,
   and use them to simulate \gls*{t1} and \gls*{t2} images at two resolution levels.
@@
   0.34 - 0.38 [mm] in the high resolution phantoms (1.0mm isotropic).
-Therefore, we also conclude that the error is bounded above by the half of the
+Therefore, we also concluded that the error was bounded above by the half of the
   voxel spacing.
@@
 We processed 16 subjects from the \gls*{hcp} database with both \regseg{}
-  and an in-house replication of the \gls*{t2b} correction method.
+  and an in-house replication of the \acrfull*{t2b} method.
 \Regseg{} showed very high accuracy, with an aggregate 95\% \gls*{ci} of 0.56-0.66 [mm],
-  bellow the pixel size of 1.25mm.
+  below the pixel size of 1.25mm.
 The misregistration error remaining after \regseg{} was significantly lower ($p < 0.01$) than the
@@
   the so-called \emph{instrumentation bias} \citep{tustison_instrumentation_2013}.
-Therefore, these results do not prove that the \regseg{} software \emph{is better than} the
-  \gls*{t2b} method.
+Therefore, these results do not prove that \regseg{} \emph{is better than} \gls*{t2b}.
 The results reported let us propose \regseg{} as a solid option in the application field.

### 2015-04-16 Publications/2015-NeuroImage/04-methods.tex
fixed typos, fixed reference

@@
      two available resolutions;
-  5) The \regseg{} is run, using the warped test images as moving multispectral-image and $\gammaset_R$ as shape-priors;
+  5) The \regseg{} is run, using the warped test images as multispectral moving-image and $\gammaset_R$ as shape-priors;
   6) Assessment of the agreement between the surfaces fitted with \regseg{} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$

### 2015-04-16 Publications/2015-NeuroImage/05-results.tex
fixed typos, fixed reference

@@
     \end{tabular}
-    \caption{The vertex-wise Haussdorf distance between the ground-truth surfaces and their
+    \caption{The vertex-wise Hausdorff distance between the ground-truth surfaces and their
     corresponding estimation with \regseg{} presented distributions with 95\% CI below

### 2015-04-16 Publications/2015-NeuroImage/03-introduction.tex
version sent to proofreading

@@
   \citep{taoka_fractional_2009}.
-To overcome the unreliability of of \gls*{fa} thresholding,
+To overcome the unreliability of \gls*{fa} thresholding,
   \cite{zhukov_level_2003} proposed active contours with edges represented
@@
 To solve the segmentation problem through registration, \cite{saad_new_2009} 
-  used Pearson correlation to perform linear alignment of the \gls*{t1} and 
+  used Pearson correlation to perform a linear alignment of the \gls*{t1} and 
   the \emph{b0}.
@@
   distortions found in \gls*{dmri} are nonlinear.
-To overcome this issue, \emph{bbregister} excludes from the boundary search those
+To circumvent this issue, \emph{bbregister} excludes from the boundary search those
   regions typically warped by the artifact.
@@
   registration method in 2D using levelsets and a nonlinear elasticity smoother on the
-  displacement vector field, which preserves topology even with very large deformations.
+  displacement vector field, which preserves the topology even with very large deformations.
 Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase

### 2015-04-16 Publications/2015-NeuroImage/04-methods.tex
version sent to proofreading

@@
      two available resolutions;
-  5) The \regseg{} is run, using the warped test images as multispectral moving-image and $\gammaset_R$ as shape-priors;
+  5) The \regseg{} tool is run, using the warped test images as multispectral moving-image and $\gammaset_R$ as shape-priors;
   6) Assessment of the agreement between the surfaces fitted with \regseg{} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$
@@
 \includegraphics[width=\linewidth]{figure03}
-\caption{Experimental workflow real data from the \acrfull*{hcp}.
+\caption{Experimental workflow to process real data from the \acrfull*{hcp}.
   1) $\gammaset_R$ are extracted from the anatomical reference (\gls*{t1} image).

### 2015-04-16 Publications/2015-NeuroImage/05-results.tex
version sent to proofreading

@@
   In yellow color, the recovered contours after registration are represented.
-  Our method showed high accuracy, as it demonstrates the almost exact location of the registered
+  \Regseg showed high accuracy, as it demonstrates the almost exact location of the registered
     contours with respect to their ground truth position depicted in green.
   Partial volume effect turns segmentation of the sulci a challenging problem with voxel-wise
-    clustering methods, but it is successfully segmented with our method.
+    clustering methods, but it is successfully segmented with \regseg{}.
   B. Quantitative evaluation of registration error in terms of average Hausdorff distance of
@@
   \gls*{t2b} were significantly different, and using the violin plots in box B of
-  \autoref{fig:results_real} we observed that errors are always larger for \gls*{t2b}.
+  \autoref{fig:results_real} we observed that errors were always larger for \gls*{t2b}.
 We also reported the 95\% CIs of the \gls*{swindex} for those surfaces.

### 2015-04-16 Publications/2015-NeuroImage/06-discussion.tex
version sent to proofreading

@@
   deformation supported by B-Spline basis functions using digital phantoms.
-We randomly deformed four different phantom models mimicking three homogeneous regions,
-  and use them to simulate \gls*{t1} and \gls*{t2} images at two resolution levels.
-After registration with \regseg{} we measured the Hausdorff distance between the
+We randomly deformed four different phantom models mimicking three homogeneous regions
+  (\gls*{wm}, \gls*{gm}, and \acrlong*{csf}), and use them to simulate \gls*{t1} and \gls*{t2}
+  images at two resolution levels.
+After registration with \regseg{}, we measured the Hausdorff distance between the
   projected contours using the ground-truth warping and our estimation.

### 2015-04-28 Publications/2015-NeuroImage/00-main.tex
Elsevier's proofreader edits

@@
 
-\title{Active-contours driven registration method for structure-informed segmentation of diffusion MR images}
+\title{Active contours-driven registration method for the structure-informed segmentation of diffusion MR images}
 
@@
 
-\address[bit]{Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on, %
-Universidad Polit\'ecnica de Madrid, Madrid, Spain}
+\address[bit]{Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on, Universidad Polit\'ecnica de Madrid, Madrid, Spain}
 \address[ciber]{Centro de Investigaci\'on Biom\'edica en Red en Bioingenier\'ia, Biomateriales y Nanomedicina (CIBER-BBN), Zaragoza, Spain}
@@
 \begin{keyword}
-diffusion MRI \sep cortical~parcellation \sep segmentation %
-\sep nonlinear~registration \sep active~contours \sep susceptibility~distortion.
+active~contours \sep cortical~parcellation \sep diffusion MRI \sep nonlinear~registration \sep segmentation \sep susceptibility~distortion.
 \end{keyword}

### 2015-04-28 Publications/2015-NeuroImage/03-introduction.tex
Elsevier's proofreader edits

--- 2015-04-13 Publications/2015-NeuroImage/03-introduction.tex
@@
 The accurate delineation of \gls*{wm} in \gls*{dmri} and the fusion of prior
-  anatomical information extracted from a \gls*{t1} image of the same subject
+  anatomical information extracted from a \gls*{t1} image for the same subject
   are crucial in a range of applications based on tractography, such as
@@
   tract-based spatial statistics \citep{smith_tractbased_2006}.
-However, segmenting \gls*{dmri} data precisely, and coregistration of anatomical
+However, the precise segmentation of \gls*{dmri} data and the coregistration of anatomical
   images are difficult due to several limitations.
-First, \gls{dmri} images suffer from partial voluming of several tissues in 
-  voxels located at their interfaces due to the low resolution of images
-  (generally around \isores[mm]{2.2}).
-Second, \gls*{dmri} schemes probe the diffusion process within the brain in 
-  many angles called \glspl*{dwi}, completed by one or more baseline (\emph{b0}) 
+First, \gls{dmri} images have a resolution that is much higher than that of the imaged
+  microstructural features \citep{basser_microstructural_1996}.
+Therefore, voxels located in structural discontinuities are affected by partial
+  voluming of the signal sources.
+Second, \gls*{dmri} schemes probe the diffusion process within the brain at
+  many angles to obtain \glspl*{dwi}, which are completed by one or more baseline (\emph{b0})
   volumes without directional gradients.
-The extremely low \gls*{snr} and high dimensionality of \glspl*{dwi} disable their
+The extremely low \gls*{snr} and the high dimensionality of \glspl*{dwi} prevent their
   direct use in segmentation.
-The low contrast between \gls*{gm} and \gls*{wm} in the \emph{b0} volumes is not suitable for 
-  segmentation either.
-Lastly, \gls*{dmri} images are acquired using \gls*{epi} to speed up acquisition
-  at the cost of introducing geometrical distortion and signal degradation and
+The low contrast between \gls*{gm} and \gls*{wm} in the \emph{b0} volumes also makes them unsuitable for
+  segmentation.
+Finally, \gls*{dmri} images are acquired using \gls*{epi} to speed up the acquisition process but
+  at the cost of introducing geometrical distortion, as well as signal degradation and
   destruction \citep{jezzard_correction_1995}.
-The artifact has been proven to cause an important impact on the anatomy extracted
+The artifacts generated have major impacts on the anatomy extracted
   from \gls*{dmri}, particularly in certain fiber bundles \citep{irfanoglu_effects_2012}.
-These limitations preclude segmentation in native \gls*{dmri} space, including
-  registration-based approaches using anatomical images as source of structural
+These limitations prevent segmentation in the native \gls*{dmri} space, including
+  registration-based approaches using anatomical images as the sources of structural
   information.
 
-Early attempts to delineate the \gls*{wm} proceeded by thresholding the 
+Early attempts to delineate the \gls*{wm} involved thresholding the
   \gls*{fa}\footnote{the \gls*{fa} is a scalar parameter of diffusion derived from
   the \gls*{dmri} data.} map.
-However, the mask and subsequent analyses substantially depend on the threshold set
-  \citep{taoka_fractional_2009} owing to the large variability of \gls*{fa} and its
-  decay with the increase of incoherence between fibers.
+However, the mask and subsequent analyses are highly dependent on the threshold set
+  \citep{taoka_fractional_2009}.
+To overcome the unreliability of \gls*{fa} thresholding,
   \cite{zhukov_level_2003} proposed active contours with edges represented
-  by levelsets, evolving on directionally invariant scalar maps.
+  by level sets, which evolve on directionally invariant scalar maps.
 \cite{rousson_level_2004} successfully segmented the corpus callosum with
-  region-based levelsets on the eigenvalues of tensors fitted on the
+  region-based level sets based on the eigenvalues of tensors fitted using the
   \gls*{dmri} data.
-A large body of work targeted the definition of appropriate features to cluster,
+Many studies have addressed the definition of appropriate features for clustering,
   such as the 5D representation proposed by \cite{jonasson_segmentation_2005}.
-Other efforts include mixture models on sets of directionally invariant maps
+Other efforts include mixed models based on sets of directionally invariant maps
   \citep{liu_brain_2007}, iterative \citep{hadjiprocopis_unbiased_2005} and
-  hierarchical \citep{lu_segmentation_2008} clusterings,
+  hierarchical \citep{lu_segmentation_2008} clustering,
   graph-cuts \citep{han_experimental_2009},
@@
 
-To solve the segmentation problem through registration, \cite{saad_new_2009} 
-  used Pearson correlation to perform linear alignment of the \gls*{t1} and 
+To address the segmentation problem by registration, \cite{saad_new_2009}
+  used the Pearson’s correlation coefficient to obtain a linear alignment of the \gls*{t1} and
   the \emph{b0}.
-Similarly, an also linear-registration method is \emph{bbregister} \citep{greve_accurate_2009}.
-It uses active contours with edges to look for intensity boundaries in the \emph{b0}
+Similarly, a linear registration method was employed by \emph{bbregister} \citep{greve_accurate_2009},
+  which uses active contours with edges to search for intensity boundaries in the \emph{b0}
   image.
-The active contours are initialized by surfaces extracted from the 
+The active contours are initialized using surfaces extracted from the
   \gls*{t1} using \emph{FreeSurfer} \citep{fischl_freesurfer_2012}:
-  the \emph{pial} surface (exterior of cortical \gls*{gm}) and the \emph{white}
+  the \emph{pial} surface (exterior of the cortical \gls*{gm}) and the \emph{white}
   surface (the \gls*{wm}/\gls*{gm} interface).
-Since the \emph{b0} image only presents a detectable frontier for the pial surface,
-  \emph{bbregister} is limited to align the cortical layer in this
+The \emph{b0} image only includes a detectable frontier for the pial surface, so
+  \emph{bbregister} is limited to aligning the cortical layer in this
   application.
-This tool has become the standard method for its proven robustness, despite the 
+This tool has become the standard method because of its proven robustness, although the
   distortions found in \gls*{dmri} are nonlinear.
-To overcome this issue, \emph{bbregister} excludes from the boundary search those
-  regions typically warped by the artifact.
-As the distortion is not accounted for, it must be addressed separately.
-Nonlinear registration has been successfully performed between \gls*{t2} and \emph{b0}
-  images for their similarity, but uniquely as a way to correct for distortions
+To overcome this issue, \emph{bbregister} excludes the
+  regions that are typically warped by artifacts from the boundary search.
+However, because the distortion is not considered, it must be addressed separately.
+Nonlinear registration has been performed successfully between \gls*{t2} and \emph{b0}
+  images based on their similarity as a unique method for correcting distortion
   \citep{kybic_unwarping_2000,studholme_accurate_2000,wu_comparison_2008,tao_variational_2009}.
-Further registration of \gls*{t1} and \gls*{t2} images was still required to map the anatomical
-  information (and the \gls*{wm} segmentation) into \gls*{dmri} space.
+However, further registration of the \gls*{t1} and \gls*{t2} images is still required to map the anatomical
+  information (and for \gls*{wm} segmentation) into the \gls*{dmri} space.
 
-Therefore, given the described circumstances, a method performing simultaneously
-  segmentation in native \gls*{dmri} space and registration of the corresponding \gls*{t1} image
-  to the \gls*{dmri} data could provide the optimal solution.
-Previously, joint segmentation and registration has been successfully applied to
-  longitudinal object tracking \citep{paragios_level_2003} and atlas-based
+Therefore, given these issues, a method that simultaneously performs
+  segmentation in the native \gls*{dmri} space and registration of the corresponding \gls*{t1} image
+  onto the \gls*{dmri} data could provide the optimal solution.
+To the best of our knowledge, this strategy has never been proposed for the application described above.
+Previously, joint segmentation and registration have been applied successfully to other problems
+such as longitudinal object tracking \citep{paragios_level_2003} and atlas-based
   segmentation \citep{gorthi_active_2011}.
-The most extended approach is integrating active contours in image registration 
+The most thorough approach integrates active contours during image registration
   methods.
-\cite{unal_coupled_2005} and later \cite{wang_joint_2006},
+\cite{unal_coupled_2005}, and later \cite{wang_joint_2006},
   improved an existing method \citep{yezzi_variational_2003} based on linear registration
-  to the nonlinear case implementing a free-form deformation field.
-\cite{droske_mumfordshah_2009} reviewed the existing techniques, and proposed two different
-  approaches to apply the Mumford-Shah functional \citep{mumford_optimal_1989} in simultaneous
-  registration and segmentation, through the propagation of the deformation field from
-  the contours to the whole image definition.
+  to the nonlinear case by implementing a free-form deformation field.
+\cite{droske_mumfordshah_2009} reviewed the existing techniques and proposed two different
+  approaches for applying the Mumford-Shah functional \citep{mumford_optimal_1989} during simultaneous
+  registration and segmentation by propagating the deformation field from
+  the contours onto the whole image definition.
 Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
-  registration method in 2D using levelsets and a nonlinear elasticity smoother on the
-  displacement vector field, which preserves topology even with very large deformations.
+  registration method in 2D using level sets and a nonlinear elasticity smoother on the
+  displacement vector field, which preserves the topology even with very high deformation.
 Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
-  levelset function for the registration of several active contours, in the application
+  level set function to register several active contours during the application
   of atlas-based segmentation.
 
-The hypothesis under this research is that the registration and segmentation
-  problems in \gls*{dmri} can be solved simultaneously, increasing the geometrical
+The hypothesis tested in the present study is that registration and segmentation
+  problems in \gls*{dmri} can be solved simultaneously, thereby increasing the geometrical
   accuracy of the process.
-We propose a tool called \regseg{} to exploit the prior information of shapes
-  extracted on \gls*{t1} images to register the anatomical reference
-  to \gls*{dmri} space, implicitly segmenting these data.
-The approach uses active contours without edges \citep{chan_active_2001} that evolve driving a
-  free-form deformation field of B-Spline basis functions.
-Optimization is performed using a descent strategy of explicit shape-gradients
+Thus, we propose a tool called \regseg{} that exploits the prior information of shapes
+  extracted using \gls*{t1} images to register the anatomical reference
+  to \gls*{dmri} space, thereby implicitly segmenting these data.
+The proposed approach uses active contours without edges \citep{chan_active_2001}, which evolve to drive a
+  free-form deformation field of B-spline basis functions.
+Optimization is performed using a descent strategy with explicit shape gradients
   \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
-Hence, unlike the most of presented methods, \regseg{} does not implement levelsets.
-Since the nonlinear distortion is aligned with one of the imaging axis (see 
-  \autoref{sec:human_connectome}), \regseg{} includes an anisotropic regularizer of
+Therefore, unlike most previous methods, \regseg{} does not implement level sets.
+The nonlinear distortion is aligned with one of the imaging axes (see
+  \autoref{sec:human_connectome}), so \regseg{} includes an anisotropic regularizer for
   the displacements field proposed by \cite{nagel_investigation_1986}.
-Finally, we evaluate \regseg{} using an extension of our instrumentation framework
-  \citep{esteban_simulationbased_2014} that simulates known and realistic distortions
-  on \gls*{dmri} data.
-The evaluation includes a comparison to the \gls*{t2b} correction by integrating in the framework
-  an in-house implementation of the method.
+Finally, we evaluated \regseg{} using an extension of our instrumentation framework
+  \citep{esteban_simulationbased_2014}, which simulates known and realistic distortions
+  based on \gls*{dmri} data.
+The evaluation included a comparison with \gls*{t2b} correction by integrating it within the framework
+  of an in-house implementation of the method.

### 2015-04-28 Publications/2015-NeuroImage/06-discussion.tex
Elsevier's proofreader edits

--- 2015-04-16 Publications/2015-NeuroImage/06-discussion.tex
@@
 \paragraph*{Accuracy tests}
-The hypothesis underpinning our research was that image registration can be reliably performed
-  by searching for homogeneous regions in a multispectral image that correspond to precise contours
-  from an atlas, or extracted from other image (i.e. a different time step).
-We proved that active contours without edges can be used to successfully drive a
-  deformation supported by B-Spline basis functions using digital phantoms.
-We randomly deformed four different phantom models mimicking three homogeneous regions,
-  and use them to simulate \gls*{t1} and \gls*{t2} images at two resolution levels.
-After registration with \regseg{} we measured the Hausdorff distance between the
-  projected contours using the ground-truth warping and our estimation.
-We concluded that errors are significantly lower than voxel resolution.
-We also assessed the 95\% \gls*{ci}, yielding an aggregate interval of
-  0.64 - 0.66 [mm] in the low resolution phantoms (2.0mm isotropic voxel) and
-  0.34 - 0.38 [mm] in the high resolution phantoms (1.0mm isotropic).
-Therefore, we also concluded that the error was bounded above by the half of the
+The hypothesis tested in our study was that reliable image registration can be performed
+  by searching for homogeneous regions in a multispectral image, which correspond to precise contours
+  from an atlas, or extracted from another image (i.e., a different time step).
+We demonstrated that active contours without edges can be used successfully to drive a
+  deformation supported by B-spline basis functions with digital phantoms.
+We randomly deformed four different phantom models to mimic three homogeneous regions
+  (\gls*{wm}, \gls*{gm}, and \acrlong*{csf}) and we used them to simulate \gls*{t1} and \gls*{t2}
+  images at two resolution levels.
+After registration with \regseg{}, we measured the Hausdorff distance between the
+  projected contours obtained using the ground-truth warping and our estimates.
+We concluded that the errors were significantly lower than the voxel resolution.
+We also assessed the 95\% \gls*{ci}, which yielded an aggregate interval of
+  0.64--0.66 [mm] for the low resolution phantoms (2.0 mm isotropic voxel) and
+  0.34--0.38 [mm] for the high resolution phantoms (1.0 mm isotropic).
+Therefore, we also concluded that the error was bounded above by half of the
   voxel spacing.
 
-\paragraph*{Application on real data}
-We designed \regseg{} as a method to segment \gls*{dmri} data exploiting the
-  anatomical information extracted in the corresponding \gls*{t1} image of the subject.
-Applications of whole-brain tractography \citep{smith_tractbased_2006,craddock_imaging_2013}
-  usually solve the problem in a two-step approach.
-First, images are corrected for nonlinear distortions using auxiliary acquisitions
-  such as fieldmaps \citep{jezzard_correction_1995}, \gls*{t2} images \citep{kybic_unwarping_2000},
-  etc.
+\paragraph*{Application to real data}
+We designed \regseg{} as a method for segmenting \gls*{dmri} data by exploiting the
+  anatomical information extracted from the corresponding \gls*{t1} image of the subject.
+Applications in whole-brain tractography \citep{smith_tractbased_2006,craddock_imaging_2013}
+  usually solve this problem with a two-step approach.
+First, the images are corrected for nonlinear distortions using auxiliary acquisitions
+  such as field maps \citep{jezzard_correction_1995}, \gls*{t2} images \citep{kybic_unwarping_2000}.
 Second, the segmentation is projected from a reference \gls*{t1} image using linear
   registration \citep{greve_accurate_2009}.
-\Regseg{} solves this joint problem in one single step, and it does not require any additional
-  acquisition over the minimal protocol including \gls*{t1} and \gls*{dmri} images only.
-This situation is commonly found in historical datasets.
+\Regseg{} addresses this joint problem in a single step and it does not require any additional
+  acquisition other than the minimal protocol using only \gls*{t1} and \gls*{dmri} images.
+This situation is found commonly in historical datasets.
 
-We evaluated \regseg{} in a real environment, using the experimental framework presented
+We evaluated \regseg{} in a real environment using the experimental framework presented
   in \autoref{fig:evworkflows}.
-We processed 16 subjects from the \gls*{hcp} database with both \regseg{}
+We processed 16 subjects from the \gls*{hcp} database using both \regseg{}
   and an in-house replication of the \acrfull*{t2b} method.
-\Regseg{} showed very high accuracy, with an aggregate 95\% \gls*{ci} of 0.56-0.66 [mm],
-  below the pixel size of 1.25mm.
-The misregistration error remaining after \regseg{} was significantly lower ($p < 0.01$) than the
-  error corresponding to the \gls*{t2b} correction, using Kruskal-Wallis H-tests
+\Regseg{} obtained very high accuracy, with an aggregate 95\% \gls*{ci} of 0.56--0.66 [mm], which was
+  below the pixel size of 1.25 mm.
+The misregistration error that remained after \regseg{} was significantly lower ($p < 0.01$) than the
+  error corresponding to the \gls*{t2b} correction according to Kruskal-Wallis H-tests
   (\autoref{tab:results_real}).
-Visual inspection of all the results (\suppl{section S5}) and the violin plots included in
-  \autoref{fig:results_real} confirmed that \regseg{} overperformed the \gls*{t2b} method
+Visual inspections of all the results (\suppl{section S5}) and the violin plots in
+  \autoref{fig:results_real} confirmed that \regseg{} performed better than the \gls*{t2b} method
   in our settings.
-Even though we carefully configured the \gls*{t2b} method by using the same algorithm and
-  same settings of a widely used software, cross-comparison experiments are prone to
-  the so-called \emph{instrumentation bias} \citep{tustison_instrumentation_2013}.
+We carefully configured the \gls*{t2b} method using the same algorithm and the 
+  same settings employed widely in software, but cross-comparison experiments are prone to
+  so-called \emph{instrumentation bias} \citep{tustison_instrumentation_2013}.
 Therefore, these results do not prove that \regseg{} \emph{is better than} \gls*{t2b}.
-The results reported let us propose \regseg{} as a solid option in the application field.
-Additionally, the \gls*{t2b} may introduce an additional (and small) error in the necessary
-  registration of \gls*{t2} to \gls*{t1} space.
+However, our results suggest that \regseg{} is a reliable option in this application field.
+In addition, the \gls*{t2b} may introduce an additional (and small) error during the necessary
+  registration of \gls*{t2} in the \gls*{t1} space.
 

### 2015-04-28 Publications/2015-NeuroImage/08-appendix-00-shapepriors.tex
major update
updates and standalone figures
Elsevier's proofreader edits

@@
   then:
+
 \begin{equation*}
@@
 \end{equation*}
-%
   where we have introduced the Hadamard power notation\footnote{The Hadamard power of a matrix or a vector
@@
 
-Since the anisotropy is generally aligned with the imaging axes, 
+In general, the anisotropy is aligned with the imaging axes, so 
   $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:energy} can be simplified to diagonal matrices, such that
@@
   $\mathbf{B}= \, \boldsymbol{\beta}\,\vec{I}_n$.
-Substituting in the equation \eqref{eq:energy}, we get:
+By substituting into equation \eqref{eq:energy}, we obtain:
+
   \begin{align}
-  E(M \mid U) &= \const + \underset{l}{\sum} \int_{\Omega_l}
-  \mdist{f'}{l} \,d\vec{r} \notag\\
-  &+ \int_{\Omega} \frac12 \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2} + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}.
+  E(\vec{u}) &= \const + \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} +   \int_{\Omega} \frac12 \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2} + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}.
   \label{eq:app_energy}

### 2015-04-29 Publications/2015-NeuroImage/02-abstract.tex

@@
   through registration.
-We propose \regseg{}, a simultaneous segmentation and registration method
+We propose \emph{regseg}, a simultaneous segmentation and registration method
   that uses active contours without edges extracted from structural images.
@@
   optimally mapping the contours onto the data in the target space.
-We tested the functionality of \regseg{} on four digital phantoms warped with known and
+We tested the functionality of \emph{regseg} on four digital phantoms warped with known and
   randomly generated deformations, to conclude that the accuracy achieved is subvoxel.
-We then apply \regseg{} to perform registration/segmentation task on 16 real diffusion MRI
+We then apply \emph{regseg} to perform registration/segmentation task on 16 real diffusion MRI
   datasets from the \acrlong*{hcp}, warped with realistic and nonlinear distortions typically
   present in these data.
-We compute the misregistration error of the contours estimated by \regseg{} with respect
+We compute the misregistration error of the contours estimated by \emph{regseg} with respect
   to their theoretical location using the ground-truth, obtaining a 95\% CI of 0.56 - 0.66 (mm)
@@
 We also compared the performance to a widely-used registration tool, to find out
-  that \regseg{} overperformed the alternate method in our settings.
+  that \emph{regseg} overperformed the alternate method in our settings.

### 2015-04-29 Publications/2015-NeuroImage/03-introduction.tex

--- 2015-04-15 Publications/2015-NeuroImage/03-introduction.tex
@@
 \section{Introduction}\label{sec:introduction}
-% Current applications to investigate the fiber pathways of the brain
-%   using \gls*{dmri}, such as tract-based spatial statistics \citep{smith_tractbased_2006}
-%   and connectome extraction \citep{craddock_imaging_2013}, critically rely on the fusion of 
-%   prior anatomical information extracted from a \gls*{t1} image of the same subject.
 The accurate delineation of \gls*{wm} in \gls*{dmri} and the fusion of prior
@@
   tract-based spatial statistics \citep{smith_tractbased_2006}.
-However, segmenting \gls*{dmri} data precisely, and coregistration of anatomical
+However, segmenting \gls*{dmri} data precisely and coregistration of anatomical
   images are difficult due to several limitations.
-First, \gls{dmri} images present a resolution much higher than the imaged
-  microstructural features \citep{basser_microstructural_1996}.
-Therefore, voxels located at structural discontinuities are affected by partial
-  voluming of sources of signal.
+First, \gls{dmri} images suffer from partial voluming of several tissues in 
+  voxels located at their interfaces due to the low resolution of images
+  (generally around \isores[mm]{2.2}).
 Second, \gls*{dmri} schemes probe the diffusion process within the brain in 
@@
 Lastly, \gls*{dmri} images are acquired using \gls*{epi} to speed up acquisition
-  at the cost of introducing geometrical distortion and signal degradation and
+  at the cost of presenting geometrical distortion and signal degradation and
   destruction \citep{jezzard_correction_1995}.
 The artifact has been proven to cause an important impact on the anatomy extracted
-  from \gls*{dmri}, particularly in certain fiber bundles \citep{irfanoglu_effects_2012}.
+  from \gls*{dmri} \citep{irfanoglu_effects_2012}, particularly in certain fiber bundles
+  when used in whole-brain tractography.
 These limitations preclude segmentation in native \gls*{dmri} space, including
@@
 However, the mask and subsequent analyses substantially depend on the threshold set
-  \citep{taoka_fractional_2009}.
-To overcome the unreliability of of \gls*{fa} thresholding,
+  \citep{taoka_fractional_2009} owing to the large variability of \gls*{fa} and its
+  decay with the increase of incoherence between fibers.
 \cite{zhukov_level_2003} proposed active contours with edges represented
@@
   regions typically warped by the artifact.
-As the distortion is not accounted for, it must be addressed separately.
+As the distortion is not accounted for, it should be addressed separately.
 Nonlinear registration has been successfully performed between \gls*{t2} and \emph{b0}
@@
   to the \gls*{dmri} data could provide the optimal solution.
-As far as we know, such strategy has never been proposed for the application at hand.
-Previously, joint segmentation and registration has been successfully applied to other problems.
-For instance, longitudinal object tracking \citep{paragios_level_2003} and atlas-based
+Previously, joint segmentation and registration has been successfully applied to
+  longitudinal object tracking \citep{paragios_level_2003} and atlas-based
   segmentation \citep{gorthi_active_2011}.
@@
   accuracy of the process.
-We propose a tool called \regseg{} to exploit the prior information of shapes
+We propose a tool called \emph{regseg} to exploit the prior information of shapes
   extracted on \gls*{t1} images to register the anatomical reference
@@
 Optimization is performed using a descent strategy of explicit shape-gradients
-  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
-Hence, unlike the most of presented methods, \regseg{} does not implement levelsets.
+  \citep{besson_dream2s_2003,herbulot_segmentation_2006}, therefore \emph{regseg}
+  does not implement levelsets as most of the presented methods do.
 Since the nonlinear distortion is aligned with one of the imaging axis (see 
-  \autoref{sec:human_connectome}), \regseg{} includes an anisotropic regularizer of
+  \autoref{sec:human_connectome}), \emph{regseg} includes an anisotropic regularizer of
   the displacements field proposed by \cite{nagel_investigation_1986}.
-Finally, we evaluate \regseg{} using an extension of our instrumentation framework
+Finally, we evaluate \emph{regseg} using an extension of our instrumentation framework
   \citep{esteban_simulationbased_2014} that simulates known and realistic distortions

### 2015-04-29 Publications/2015-NeuroImage/04-methods.tex

--- 2015-04-16 Publications/2015-NeuroImage/04-methods.tex
@@
   P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
-  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) = % \notag\\ &=
-  \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
+  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) = \notag\\
+  &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
   \mdist{f'}{l} \right)}},
@@
   P(U) = \underset{\vec{r}}{\prod}\, p(\vec{u}) &=
-  \underset{\vec{r}}{\prod}\, p_0(\vec{u}) \, p_1(\vec{u}), \text{ with} \\
+  \underset{\vec{r}}{\prod}\, p_0(\vec{u}) \, p_1(\vec{u}), \\
   p_0(\vec{u}) &= \mathcal{N}( \vec{u} \mid 0, \mathbf{A}^{-1}), \notag\\
@@
   \underset{\vec{r} \in \Omega_l}{\prod}
-  \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) = \notag\\ &=
-  \const + \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, + % \notag\\ &+
-  \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
+  \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) \notag\\
+  &= \const + \underset{l}{\sum} \int_{\Omega_l}
+  \mdist{f'}{l} \,d\vec{r} \, + \notag\\
+  &+ \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
   \label{eq:energy}
@@
 \begin{figure}
-	\includegraphics[width=\linewidth]{figure01}
+	\includegraphics[width=\linewidth]{figures/figure01}
 	\caption{The active contours are defined as the interfacing surfaces of the competing
@@
   \frac{ \partial }{\partial \vec{u}_k} \Big\{
-  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, + % \notag\\ &+
-  \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
+  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, + \notag\\
+  &+ \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
   + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
@@
 The computation of the explicit shape gradients at each $\vec{v}_i'$ is described in \autoref{fig:method}.
-Then, \eqref{eq:gradient_descent} can be reformulated as (see \suppl{equations SM6-SM12}):
+The intermediate steps to get to \eqref{eq:final_gradient} from \eqref{eq:gradient_descent} are
+  detailed in the \suppl{equations SM6-SM12}:
 
@@
 \label{sec:conv_report}
-The \regseg{} tool includes a multiresolution strategy on the free-form deformation field.
+The \emph{regseg} tool includes a multiresolution strategy on the free-form deformation field.
 Registration pyramids are created by setting the spacing between control points of the B-Spline basis
@@
   reports (as the one found in the \suppl{section S1.3}).
-We release \regseg{} along with the tool for generating the convergence report.
+We release \emph{regseg} along with the tool for generating the convergence report.
 
 \subsection{Evaluation protocol}\label{sec:evaluation_protocol}
-In order to assess the performance of \regseg{}, we defined a general
+In order to assess the performance of \emph{regseg}, we defined a general
   evaluation protocol:
@@
   generate warped images ($M$) to segment;
-3) execution of \regseg{} with $\gammaset_R$ and the warped data as inputs; and
+3) execution of \emph{regseg} with $\gammaset_R$ and the warped data as inputs; and
 4) visual assessment and computation of error metrics.
@@
 The workflow to simulate the digital phantoms and assess
-  \regseg{} on them is presented in \autoref{fig:evphantoms}.
+  \emph{regseg} on them is presented in \autoref{fig:evphantoms}.
 A set of four binary objects (called ``box'', ``ball'', ``L'',
@@
 The reference surfaces $\gammaset_R$ are extracted from the binary shapes
-  using \emph{FreeSurfer} tools \citep{fischl_freesurfer_2012}.
+  using \emph{FreeSurfer} tools.
 The ground-truth distortion is generated with a chain of two displacement 
@@
 The warped surfaces $\gammaset_\text{true}$ are binarized to generate tissue fractions
-  in low (\isores[mm]{2.0}) and high (\isores[mm]{1.0}) resolution.
+  in low (\isores[mm]{2.0}) and high (\isores[mm]{1.0}).
 Using a \gls*{mr} simulator \citep{caruyer_phantomas_2014}, we synthesize
@@
 The field of view at both resolutions is \isores[mm]{100}.
-Then, \regseg{} is applied to map $\gammaset_R$ into the warped phantoms to
+Then, \emph{regseg} is applied to map $\gammaset_R$ into the warped phantoms to
   obtain the registered surfaces ($\hat{\gammaset}_{test}$).
@@
 \begin{figure*}
-\includegraphics[width=\linewidth]{figure02}
-\caption{The evaluation of \regseg{} on phantom data is performed with the presented instrumental workflow:
+\includegraphics[width=\linewidth]{figures/figure02}
+\caption{The evaluation of \emph{regseg} on phantom data is performed with the presented instrumental workflow:
   1) The reference surfaces $\gammaset_R$ are triangularized meshes extracted from the four binary shapes (``box'', ``ball'', ``L'', ``gyrus'');
@@
      two available resolutions;
-  5) The \regseg{} is run, using the warped test images as multispectral moving-image and $\gammaset_R$ as shape-priors;
-  6) Assessment of the agreement between the surfaces fitted with \regseg{} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$
+  5) The \emph{regseg} is run, using the warped test images as moving multispectral-image and $\gammaset_R$ as shape-priors;
+  6) Assessment of the agreement between the surfaces fitted with \emph{regseg} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$
      with automatically generated visual reports and computing the Hausdorff distance between corresponding surfaces.}\label{fig:evphantoms}
@@
 \begin{figure*}
-\includegraphics[width=\linewidth]{figure03}
+\includegraphics[width=\linewidth]{figures/figure03}
 \caption{Experimental workflow real data from the \acrfull*{hcp}.
@@
   Target diffusion scalars (\gls*{fa} and \gls*{adc}) are computed on the distorted data and
-    stacked to feed the multivariate input required by \regseg{}.
+    stacked to feed the multivariate input required by \emph{regseg}.
   4) the method is run, obtaining a $U_{test} = \hat{U}_{true}$, the estimation of
@@
 \paragraph*{Data}
-For the evaluation of \regseg{} on real \gls*{dmri} data of human brains,
+For the evaluation of \emph{regseg} on real \gls*{dmri} data of human brains,
   we collected 16 subjects from the \gls*{hcp} database.
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{figure04}
+\includegraphics[width=\linewidth]{figures/figure04}
 \caption{Segmentation model defined by the homogeneous regions $\Omega_i$ in which
@@
   from \gls*{dmri} data.
-After evaluating several alternative models, we empirically defined a partition \omegaset{}
-  into the following six regions:
-  1) the thalamus ($\Omega_{Tha}$);
-  2) the ventricular system and deep \gls*{gm} structures ($\Omega_{VdGM}$);
-  3) the cerebral \gls*{wm} ($\Omega_{WM}$);
-  4) the brain stem and cerebellar \gls*{wm} ($\Omega_{bst}$);
-	5) the cerebellar \gls*{gm} ($\Omega_{cbGM}$); and
-	6) the cortical \gls*{gm} ($\Omega_{GM}$).
+After evaluating several alternative models, we empirically defined the partition \omegaset{} 
+  of the reference image $R$ as follows:
+  1) $\Omega_{Tha}$ including the thalamus;
+  2) $\Omega_{VdGM}$ including the ventricular system and deep \gls*{gm} structures;
+  3) $\Omega_{WM}$ including the cerebral \gls*{wm};
+  4) $\Omega_{bst}$ including the brain stem and cerebellar \gls*{wm};
+	5) $\Omega_{cbGM}$ including the cerebellar \gls*{gm}; and
+	6) $\Omega_{GM}$ including the cortical \gls*{gm}.
 Using tools of \emph{FreeSurfer}, and the appropriate selections of labels in the 
@@
   sWI = \frac{1}{\sum_i a_i} \sum\limits_i^{N_V} a_i\,\|
-  \vec{v}_i - \hat{\vec{v}}_i \|,
+  \vec{v}_i - \hat{\vec{v}}_i \|
   \label{eq:swindex}

### 2015-04-29 Publications/2015-NeuroImage/supplemental.tex

@@
 % @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-04-08 13:07:44
+% @Last Modified time: 2015-04-08 13:28:07
 
@@
 \subsection{Implementation}
-\subsection{General} The \emph{regseg} registration and segmentation tool is written in C++, using ITK-4.6 as
-  implementation core.
+\subsection{General} The \emph{regseg} registration and segmentation tool is written in C++, using ITK-4.6
+  \citep{ibanez_itk_2005} as  implementation core.
 We designed a modular implementation, enabling multithreading in several pieces of the software,
@@
   as the convergence report generation, \autoref{sec:convergence_evidence}).
+The evaluation workflow is independent, and implemented using \emph{nipype} \citep{gorgolewski_nipype_2011}.
 

### 2015-04-29 Publications/2015-NeuroImage/00-main.tex

@@
 \providecommand{\nvec}[1]{\hat{\mathbf{#1}}}
-\providecommand{\suppl}[1]{\href{http://dx.doi.org/mydoi}{Supplemental Materials}, #1}
+\providecommand{\suppl}[1]{\href{http://figshare.com/s/459c26b4ee8211e493b306ec4bbcf141}{Supplemental Materials}, #1}
 \providecommand{\isores}[2][]{#2\ensuremath{\times}#2\ensuremath{\times}#2%
@@
 
-\title{Active contours-driven registration method for the structure-informed segmentation of diffusion MR images}
+\title{Active contours-driven registration method for the structure-informed segmentation of diffusion MR images\tnoteref{t1}}
+\tnotetext[t1]{Supplemental materials to this work will be publicly available upon publication with doi:\href{http://figshare.com/s/459c26b4ee8211e493b306ec4bbcf141}{10.6084/m9.figshare.1397502}.
+During the review process, the document is available with the private link.}
 

### 2015-04-29 Publications/2015-NeuroImage/02-abstract.tex

--- 2015-03-24 Publications/2015-NeuroImage/02-abstract.tex
@@
 % -*- root: 00-main.tex -*-
-We propose a simultaneous segmentation and registration method called \emph{regseg}
+Current applications of whole-brain tractography to \acrlong*{dmri} data require highly precise
+  delineations of anatomical structures, which are usually projected from anatomical \acrlong*{t1} images
+  by registration.
+In this study, we propose \regseg{}, which is a simultaneous segmentation and registration method
   that uses active contours without edges extracted from structural images.
-The contours evolve through a free-deformation field of B-Spline functions that
-  optimally maps the surfaces onto the data in the target space.
-\emph{Regseg} is intended, but not limited, to correct susceptibility-derived distortions
-  in echo-planar imaging.
-We proof the concept behind \emph{regseg} on four digital phantoms warped with random
-  deformations, and conclude that the precision of the segmentation is subvoxel.
-We demonstrate \emph{regseg} in the correction task on 16 real diffusion MRI
-  datasets from the \acrlong*{hcp}.
-The results suggest that \emph{regseg} overperformed an extended competing method.
-We compute the geometrical disagreement of the surfaces registered with \emph{regseg} with
-  respect to a ground truth, obtaining a 95\% confidence interval of 0.56 - 0.66 (mm)
-  for the Hausdorff distance of mesh vertices in images of 1.25mm isotropic resolution.
+The contours evolve through a free-form deformation field supported by the B-spline basis
+  to optimally map the contours onto the data in the target space.
+We tested the functionality of \regseg{} using four digital phantoms warped with known and
+  randomly generated deformations, where subvoxel accuracy was achieved.
+We then applied \regseg{} to a registration/segmentation task using 16 real diffusion MRI
+  datasets from the \acrlong*{hcp}, which were warped by realistic and nonlinear distortions that are typically
+  present in these data.
+We computed the misregistration error of the contours estimated by \regseg{} with respect
+  to their theoretical location using the ground truth, thereby obtaining a 95\% CI of 0.56--0.66 mm
+  distance between corresponding mesh vertices, which was below the 1.25 mm isotropic resolution of the images.
+We also compared the performance of our proposed method with a widely used registration tool, which showed 
+  that \regseg{} outperformed this method in our settings.

### 2015-04-29 Publications/2015-NeuroImage/04-methods.tex

--- 2015-04-16 Publications/2015-NeuroImage/04-methods.tex
@@
 \label{sec:methods}
-%
 Let $\gammaset_R = \{\Gamma_m: m \in \mathbb{N}, m \leq N_c\}$ be the set of $N_c$ surfaces
@@
 We reformulate the segmentation of the distorted \gls*{dmri} images (the moving space $M$)
-  as a registration problem where we look for an underlying deformation field $U$, such that 
-  the structures in $R$ defined by $\gammaset_R$ optimally align with their corresponding
+  as a registration problem where we search for an underlying deformation field $U$ such that 
+  the structures in $R$ defined by $\gammaset_R$ align optimally with their corresponding
   structures in $M$:
@@
   \end{align}
-%
   where $\vec{r}$ denotes a position in $R$, $\vec{r}'$ is
-  its corresponding location in $M$, and $n$ the dimensionality of images.
+  its corresponding location in $M$, and $n$ denotes the dimensionality of images.
 Finally, $\vec{u} = \vec{u}(\vec{r})$ is the displacement of every point with respect
@@
 \subsection{Cost-function derivation}\label{sec:methods_map}
-%
 In a Bayesian framework for registration \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014},
@@
   which $M$ is partitioned by the projection of $\gammaset_R$.
-Using the Bayes' rule, the posterior likelihood is computed as:
+Using Bayes' rule, the posterior likelihood is computed as:
 
@@
   \end{equation}
-%
-  where $P(M \mid U,\omegaset)$ is the data-likelihood.
-As $\omegaset$ are mapped by $U$, we simplify
+  where $P(M \mid U,\omegaset)$ is the data likelihood.
+Because $\omegaset$ are mapped by $U$, then we simplify
   $P(U, \omegaset) = P(U) \implies P(M \mid U,\omegaset) = P(M \mid U)$.
 
-The best estimate $\hat{U}$ then fulfills the maximum a posteriori criterium
-  \citep{bishop_pattern_2006}, and aligns $\gammaset_R$ into $M$.
-First, we assume independence between pixels, and thus break down the
+The best estimate $\hat{U}$ then satisfies the maximum a posteriori criterion
+  \citep{bishop_pattern_2006} and it aligns $\gammaset_R$ into $M$.
+First, we assume independence between pixels, and thus we break down the
   global data likelihood into a product of pixel-wise conditional probabilities:
@@
   \end{equation}
-%
   where $\vec{f}' = M(\vec{r}')$ is the feature vector at the displaced
   position $\vec{r}'$ \eqref{eq:transform} in the moving image.
-For convenience, and because it has been shown to be an appropriate approximation
+For convenience and because it has been shown to be an appropriate approximation
   \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
@@
   1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
-  distributions with parameters $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
+  distributions with the parameters $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
   for each region $\Omega_l$:
@@
   P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
-  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) = % \notag\\ &=
-  \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
+  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) =   \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
   \mdist{f'}{l} \right)}},
@@
   \end{align}
-%
   using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
@@
   $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
-$C$ is the number of channels comprised in the image $M$.
+$C$ is the number of channels in the image $M$.
 In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
@@
   \end{align}
-%
-  imposing that the distortion and its gradient have zero
+ which requires that the distortion and its gradient have zero
   mean and variance governed by the matrices $\mathbf{A}$ and $\mathbf{B}$.
-Finally, the maximum a posteriori problem is adapted into a variational one where we look for
-  the minimum of an energy functional, by applying $E(\vec{u}) = -\log P(U)$:
+Finally, the maximum a posteriori problem is adapted to a variational problem where we search for
+  the minimum of an energy functional by applying $E(\vec{u}) = -\log P(U)$:
 
@@
   \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) = \notag\\ &=
-  \const + \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, + % \notag\\ &+
-  \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
+  \const + \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, +   \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
   \label{eq:energy}
   \end{align}
-%
-This expression is dual of the Mumford-Shah functional corresponding
-  to a framework of \acrlong*{acwe} \citep{chan_active_2001}
+This expression is the dual of the Mumford-Shah functional that corresponds
+  to the framework of \acrlong*{acwe} \citep{chan_active_2001}
   with the corresponding anisotropic regularization term of \cite{nagel_investigation_1986}.
@@
   \caption{The active contours are defined as the interfacing surfaces of the competing
-	  \glspl{roi} $\Omega_i$.
-	They are represented in green and dark blue colors in this close-up.
-	They iteratively evolve following their inward normals $\hat{n}_i$ at each vertex
+    \glspl{roi} $\Omega_i$, which
+  are represented in green and dark blue in this close-up.
+  They evolve iteratively following their inward normals $\hat{n}_i$ at each vertex
     $\vec{v}_i$ of the mesh.
-	The gradient speeds $\bar{s}_i$ drive registration, and are computed as the disparity of data
+  The gradient speeds $\bar{s}_i$ drive registration, which are computed as the disparity of the data
     energies with respect to the two limiting regions of $M(\vec{v}_i)$, the features of the image
@@
   In this figure, $\bar{s}_1$ is written in the lower
-	  box, with $\Omega_{wm}$ being the inner limiting region, $\Omega_{gm}$ the outer region, and
-    $w_{0,1}$ the relative area associated to vertex $\vec{v}_1$ with respect to
+    box, with $\Omega_{wm}$ in the inner limiting region, $\Omega_{gm}$ the outer region, and
+    $w_{0,1}$ is the relative area associated with vertex $\vec{v}_1$ with respect to
     the total area of surface $\Gamma_0$.
-	% Finally, every $\bar{s}_i$ is associated with the control points $\vec{u}_k$ through
-	%   the corresponding weights $\psi_k(\vec{v}_i)$ by Eq. \eqref{eq:gradient_wshape}.
       }\label{fig:method}
@@
 Since the vertices of the surfaces $\{\vec{v}_i: \vec{v}_i \subset \gammaset \}_{i=1 \ldots N_v}$
-  likely lay off-grid, it is required to derive $\vec{u}_i = \vec{u}(\vec{v}_i)$ from a discrete set of parameters
+  are probably located off-grid, it is necessary to derive $\vec{u}_i = \vec{u}(\vec{v}_i)$ from a discrete set of parameters
   $\{\vec{u}_k\}_{k=1 \ldots K}$.
-Densification is achieved through a set of associated basis functions $\psi_k$ \eqref{eq:nodes_tfm}.
-In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
-  of degree 3.
+Densification is achieved using a set of associated basis functions $\psi_k$ \eqref{eq:nodes_tfm}.
+In our implementation, $\psi_k$ is selected as a tensor-product B-spline kernel
+  of degree three.
 
@@
   we propose a gradient-descent approach with respect to the underlying
-  deformation field through the following \gls*{pde}:
+  deformation field using the following \gls*{pde}:
 
@@
   \end{equation}
-%
-  with $t$ being an artificial time parameter of the contour
-  evolution, and $\vec{u}_k$ the parameters supporting the estimate
+  where $t$ is an artificial time parameter of the contour
+  evolution and $\vec{u}_k$ are the parameters that support the estimate
   $\hat{U}$ of the transformation at the current time point.
-Let us assume that anisotropy is aligned with the imaging axes to simplify
-  \eqref{eq:energy} into expression \eqref{eq:app_energy} of \ref{app:reg_term}
-  and compute its derivative \eqref{eq:general_gradient_descent}:
+Let us assume that the anisotropy is aligned with the imaging axes to simplify
+  \eqref{eq:energy} as expression \eqref{eq:app_energy} in \ref{app:reg_term},
+  and thus to compute its derivative \eqref{eq:general_gradient_descent}:
 
@@
   \frac{ \partial }{\partial \vec{u}_k} \Big\{
-  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, + % \notag\\ &+
-  \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
+  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, +   \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
   + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
@@
   \end{align}
-%
   where $\vec{u}^{\circ2} = \vec{u}^T \cdot \vec{u}$.
@@
   derivatives.
-The computation of the explicit shape gradients at each $\vec{v}_i'$ is described in \autoref{fig:method}.
+The computation of the explicit shape gradients at each $\vec{v}_i'$ is illustrated in \autoref{fig:method}.
 Then, \eqref{eq:gradient_descent} can be reformulated as (see \suppl{equations SM6-SM12}):
@@
 
-Finally, to descend this gradient we establish a semi-implicit Euler scheme (see \suppl{section S1.3}),
-  with a step size parameter $\delta$, that we solve in the spectral domain as follows:
+Finally, to descend this gradient, we establish a semi-implicit Euler scheme (see \suppl{section S1.3}),
+  with a step size parameter $\delta$, which we solve in the spectral domain as follows:
 
   \begin{align}
-  \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\delta^{-1} \, \vec{u}_k^t - \vec{g}_k\} }%
-                  {\mathcal{F}\{(\delta^{-1} + \boldsymbol{\alpha})\, I - \boldsymbol{\beta}\Delta\}} \right\},
+  \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\delta^{-1} \, \vec{u}_k^t - \vec{g}_k\} }                  {\mathcal{F}\{(\delta^{-1} + \boldsymbol{\alpha})\, I - \boldsymbol{\beta}\Delta\}} \right\},
   \label{eq:update_equation}
@@
 
-\paragraph*{Implementation details, settings and convergence}
+\paragraph*{Implementation details, settings, and convergence}
 \label{sec:conv_report}
 The \regseg{} tool includes a multiresolution strategy on the free-form deformation field.
-Registration pyramids are created by setting the spacing between control points of the B-Spline basis
-  functions for each level of the multiresolution.
-This implementation detail, along with other features (as the sparse matrix approach
+Registration pyramids are created by setting the spacing between the control points of the B-spline basis
+  functions for each level of the multiresolution strategy.
+The implementation details as well as other features (such as the sparse matrix approach
   to fast interpolation) and the main parameters 
-  (such as $\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, the B-Spline grid resolutions,
-  target image smoothing, etc.) are discussed in the \suppl{section S1}.
-The actual choices of parameter settings are publicly distributed with the source code of the experiments.
-These final settings were manually obtained based on the feedback of post-registration convergence
-  reports (as the one found in the \suppl{section S1.3}).
-We release \regseg{} along with the tool for generating the convergence report.
+  (such as $\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, the B-spline grid resolutions,
+ and target image smoothing) are discussed in \suppl{section S1}.
+The actual choices of the parameter settings are publicly distributed with the source code for the experiments.
+The final settings were obtained manually based on feedback from the post-registration convergence
+  reports (such as that found in \suppl{section S1.3}).
+We released \regseg{} along with the tool to generate the convergence report.
 
 \subsection{Evaluation protocol}\label{sec:evaluation_protocol}
-In order to assess the performance of \regseg{}, we defined a general
+In order to assess the performance of \regseg{}, we defined the following general
   evaluation protocol:
-1) Extraction of the set of undistorted surfaces $\gammaset_R$;
-2) computation of a ground-truth field of displacements $U_{true}$ that is applied to
-  generate warped images ($M$) to segment;
-3) execution of \regseg{} with $\gammaset_R$ and the warped data as inputs; and
-4) visual assessment and computation of error metrics.
-How this protocol is adapted to the simulated phantoms and real data is covered in the
+1) Extract the set of undistorted surfaces $\gammaset_R$;
+2) Compute a ground-truth field of displacements $U_{true}$, which is applied to
+  generate warped images ($M$) for segmentation;
+3) Execute \regseg{} with $\gammaset_R$ and use the warped data as inputs; and
+4) Perform a visual assessment and compute the error metrics.
+The adaptation of this protocol to the simulated phantoms and real data is explained in the
   following sections.
@@
 \subsection{Simulated phantoms}\label{sec:digital_phantoms}
-The workflow to simulate the digital phantoms and assess
-  \regseg{} on them is presented in \autoref{fig:evphantoms}.
-A set of four binary objects (called ``box'', ``ball'', ``L'',
-  and ``gyrus'') are generated combining binarization of
+The workflow required to simulate the digital phantoms and to assess the performance of
+  \regseg{} with them is presented in \autoref{fig:evphantoms}.
+A set of four binary objects (i.e., ``box'', ``ball'', ``L'',
+  and ``gyrus'') was generated by combining the binarization of
   analytical shapes and mathematical morphology.
-The reference surfaces $\gammaset_R$ are extracted from the binary shapes
+The reference surfaces $\gammaset_R$ were extracted from the binary shapes
   using \emph{FreeSurfer} tools \citep{fischl_freesurfer_2012}.
-The ground-truth distortion is generated with a chain of two displacement 
-  fields supported by grids of B-Spline basis functions.
-The coefficients of the basis functions are generated randomly for
+The ground-truth distortion was generated using a chain of two displacement 
+  fields supported by grids of B-spline basis functions.
+The coefficients of the basis functions were generated randomly for
   both levels in their three dimensions.
 The three components of the displacements $\vec{u} = (u_d)$ 
-  are bounded above by 40\% of the separation between control points
-  at each level, to obtain diffeomorphic transforms
+  were bounded above by 40\% of the separation between the control points
+  at each level to obtain diffeomorphic transforms
   after concatenation \citep{rueckert_diffeomorphic_2006}.
-The first deformation field is intended to generate large warpings,
-  with control points separated by 50.50mm along the three dimensions
-  ($u_d\leq$ 20.20mm).
-With the second warping, we look for a resulting field with smoothness
-  close to the found in a typical distortion field of \gls*{dmri} data
+The first deformation field was applied to generate large warpings
+  with control points separated by 50.50 mm in the three dimensions
+  ($u_d\leq$ 20.20 mm).
+With the second warping, we aimed to obtain a field with smoothness
+  close to that found in a typical distortion field of \gls*{dmri} data
   \citep{irfanoglu_susceptibility_2011}.
-Therefore, the control points are separated 25.25mm ($u_d\leq$ 10.10mm).
-Once the ground-truth deformation is generated, the original surfaces
-  are warped interpolating the displacements field at each vertex.
-The warped surfaces $\gammaset_\text{true}$ are binarized to generate tissue fractions
-  in low (\isores[mm]{2.0}) and high (\isores[mm]{1.0}) resolution.
-Using a \gls*{mr} simulator \citep{caruyer_phantomas_2014}, we synthesize
-  \gls*{t1} (TE/TR= 10/1500ms) and \gls*{t2} images (TE/TR= 90/5000ms)
-  corresponding to each phantom type, with two resolutions each
-  (1.0mm and 2.0mm isotropic).
-The field of view at both resolutions is \isores[mm]{100}.
-Then, \regseg{} is applied to map $\gammaset_R$ into the warped phantoms to
+Therefore, the control points were separated by 25.25 mm ($u_d\leq$ 10.10 mm).
+After generating the ground-truth deformation, the original surfaces
+  were warped by interpolating the displacements field at each vertex.
+The warped surfaces $\gammaset_\text{true}$ were binarized to generate tissue fractions
+  at low (\isores[mm]{2.0}) and high (\isores[mm]{1.0}) resolutions.
+Using a \gls*{mr} simulator \citep{caruyer_phantomas_2014}, we synthesized
+  \gls*{t1} (TE/TR= 10/1500 ms) and \gls*{t2} images (TE/TR= 90/5000 ms), which
+  corresponded to each phantom type, with each at two resolutions 
+  (1.0 mm and 2.0 mm isotropic).
+The field of view at both resolutions was \isores[mm]{100}.
+Next, \regseg{} was applied to map $\gammaset_R$ onto the warped phantoms to
   obtain the registered surfaces ($\hat{\gammaset}_{test}$).
-% Simulations were corrupted with rician noise for a \gls*{snr} of 300.0.
-To quantify the misregistration error, we compute the Hausdorff distance between
- $\hat{\gammaset}_{test}$ and $\gammaset_\text{true}$, using \citep{commandeur_vtk_2011}.
+To quantify the misregistration error, we computed the Hausdorff distance between
+ $\hat{\gammaset}_{test}$ and $\gammaset_\text{true}$ using \citep{commandeur_vtk_2011}.
 
@@
 \includegraphics[width=\linewidth]{figure02}
-\caption{The evaluation of \regseg{} on phantom data is performed with the presented instrumental workflow:
-  1) The reference surfaces $\gammaset_R$ are triangularized meshes extracted from the four binary shapes (``box'', ``ball'', ``L'', ``gyrus'');
-  2) A ground-truth displacement field is generated as described in \autoref{sec:digital_phantoms}, and applied to warp
-      $\gammaset_R$, obtaining $\gammaset_\text{true}$;
-  3) Once warped, $\gammaset_\text{true}$ are projected to the corresponding discrete 3D volume, and downsampled creating partial volume effects to two resolutions
-     of \isores[mm]{2.0} and \isores[mm]{1.0}, producing sets of tissue fractions maps;
-  4) The tissue fractions are fed to an \acrfull*{mr} simulator, generating \acrfull*{t1} and \acrfull*{t2} -like images at the
-     two available resolutions;
-  5) The \regseg{} is run, using the warped test images as multispectral moving-image and $\gammaset_R$ as shape-priors;
-  6) Assessment of the agreement between the surfaces fitted with \regseg{} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$
-     with automatically generated visual reports and computing the Hausdorff distance between corresponding surfaces.}\label{fig:evphantoms}
+\caption{Evaluation of \regseg{} using phantom data according to the following instrumental workflow.
+  1) The reference surfaces $\gammaset_R$ are triangularized meshes extracted from the four binary shapes (i.e., ``box'', ``ball'', ``L'', ``gyrus'').
+  2) A ground-truth displacement field was generated as described in \autoref{sec:digital_phantoms}, and applied to warp
+      $\gammaset_R$, thereby obtaining $\gammaset_\text{true}$.
+  3) After being warped, $\gammaset_\text{true}$ were projected onto the corresponding discrete 3D volume and downsampled to create partial volume effects at two resolutions,
+     i.e., \isores[mm]{2.0} and \isores[mm]{1.0}, thereby producing sets of tissue fractions maps.
+  4) The tissue fractions were fed into an \acrfull*{mr} simulator, which generated \acrfull*{t1} and \acrfull*{t2} -like images at the
+     two possible resolutions.
+  5) The \regseg{} tool was applied using the warped test images as multispectral moving images and $\gammaset_R$ as shape priors.
+  6) The agreement between the surfaces fitted by \regseg{} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$
+     using automatically generated visual reports was assessed by computing the Hausdorff distance between the corresponding surfaces.}\label{fig:evphantoms}
 \end{figure*}
 
-\subsection{Real datasets} %
-\label{sec:human_connectome}
-% 
-The general experimental framework for the real datasets is presented in \autoref{fig:evworkflows}.
-It extends our previous work \citep{esteban_simulationbased_2014} evaluating distortions
-  on \gls*{dmri} phantoms.
+\subsection{Real datasets} \label{sec:human_connectome}
+The general experimental framework for the real datasets is presented in \autoref{fig:evworkflows},
+which extends our previous evaluation \citep{esteban_simulationbased_2014} of distortions
+  using \gls*{dmri} phantoms.
 
@@
 \includegraphics[width=\linewidth]{figure03}
-\caption{Experimental workflow real data from the \acrfull*{hcp}.
-  1) $\gammaset_R$ are extracted from the anatomical reference (\gls*{t1} image).
-  2) To operate as ground truth, we generate a plausible-synthetic distortion $U_{true}$
-    from the fieldmap using \eqref{eq:fieldmap}.
-  3) The \gls*{dmri} data are warped using $U_{true}$ to reproduce the effects of real
+\caption{Experimental workflow employed to process real data from the \acrfull*{hcp}.
+  1) $\gammaset_R$ were extracted from the anatomical reference (\gls*{t1} image).
+  2) For use as the ground truth, we generated a plausible synthetic distortion $U_{true}$
+    from the field map with \eqref{eq:fieldmap}.
+  3) The \gls*{dmri} data were warped using $U_{true}$ to reproduce the effects of real
     susceptibility-derived distortions.
-  Target diffusion scalars (\gls*{fa} and \gls*{adc}) are computed on the distorted data and
+  Target diffusion scalars (\gls*{fa} and \gls*{adc}) were computed with the distorted data and
     stacked to feed the multivariate input required by \regseg{}.
-  4) the method is run, obtaining a $U_{test} = \hat{U}_{true}$, the estimation of
+  4) The method was run to obtain $U_{test} = \hat{U}_{true}$, i.e., the estimate of
     the ground-truth deformation.
-  5) Results are visually and quantitatively evaluated.}\label{fig:evworkflows}
+  5) The results were evaluated visually and quantitatively.}\label{fig:evworkflows}
 \end{figure*}
@@
 \paragraph*{Data}
-For the evaluation of \regseg{} on real \gls*{dmri} data of human brains,
+To evaluate \regseg{} using real \gls*{dmri} data obtained from human brains,
   we collected 16 subjects from the \gls*{hcp} database.
-The original acquisitions are released within ``unprocessed'' packages, and
-  the ``minimally preprocessed'' are corrected for artifacts, brain-extracted
-  and spatially normalized, along with some results of the standard processing
+The original acquisitions are released within ``unprocessed'' packages, whereas
+  the ``minimally preprocessed'' are corrected for artifacts, brain-extracted,
+  and spatially normalized, while some results have been subjected to the standard processing
   pipeline of \emph{FreeSurfer}.
-We refer the reader to \citep{essen_human_2012} for exact details about acquisition
-  parameters, and \citep{glasser_minimal_2013} for the preprocessing issues.
-The datasets comprehend a large set of images, containing \gls*{t1}, \gls*{t2} and
+We refer the reader to \citep{essen_human_2012} for exact details of the acquisition
+  parameters and \citep{glasser_minimal_2013} for the preprocessing issues.
+These datasets comprise a large set of images, including \gls*{t1}, \gls*{t2}, and
   multi-shell \gls*{dmri} images.
@@
 \includegraphics[width=\linewidth]{figure04}
-\caption{Segmentation model defined by the homogeneous regions $\Omega_i$ in which
-  the image to be segmented is partitioned by the theoretical projection 
-  ($\gammaset_{true}$) of $\gammaset_R$ into $M$.
-The plot represents a kernel density estimation of the location and spread of
-  each region $\Omega_i$ in the bivariate feature space of $M$, comprising 
+\caption{Segmentation model defined by the homogeneous regions $\Omega_i$, where
+  the image for segmentation is partitioned by the theoretical projection 
+  ($\gammaset_{true}$) of $\gammaset_R$ onto $M$.
+The plot represents a kernel density estimate of the location and spread of
+  each region $\Omega_i$ in the bivariate feature space of $M$, which comprises 
   the \gls*{fa} and \gls*{adc} maps.
-On top and right margins the marginal distributions of $\Omega_i$ are also plotted for
+At the top and the right margins, the marginal distributions of $\Omega_i$ are also plotted for
   the \gls*{fa} and \gls*{adc}, respectively.
@@
 Based on our experience \citep{esteban_brain_2012} and previous studies \citep{ennis_orthogonal_2006},
-  we define the moving image as a stack of the \gls*{fa} and \gls*{adc} maps derived
+  we defined the moving image as a stack of the \gls*{fa} and \gls*{adc} maps derived
   from \gls*{dmri} data.
 After evaluating several alternative models, we empirically defined a partition \omegaset{}
-  into the following six regions:
-  1) the thalamus ($\Omega_{Tha}$);
-  2) the ventricular system and deep \gls*{gm} structures ($\Omega_{VdGM}$);
-  3) the cerebral \gls*{wm} ($\Omega_{WM}$);
-  4) the brain stem and cerebellar \gls*{wm} ($\Omega_{bst}$);
-	5) the cerebellar \gls*{gm} ($\Omega_{cbGM}$); and
-	6) the cortical \gls*{gm} ($\Omega_{GM}$).
-Using tools of \emph{FreeSurfer}, and the appropriate selections of labels in the 
-  \emph{aparc} segmentation released with the \gls*{hcp} data, the $\gammaset_R$ set of
-  reference surfaces is extracted.
-The segmentation model corresponding to this partition is depicted in \autoref{fig:model},
-  and discussed with finer detail in \suppl{section S4}.
+  according to the following six regions:
+  1) thalamus ($\Omega_{Tha}$);
+  2) ventricular system and deep \gls*{gm} structures ($\Omega_{VdGM}$);
+  3) cerebral \gls*{wm} ($\Omega_{WM}$);
+  4) brain stem and cerebellar \gls*{wm} ($\Omega_{bst}$);
+  5) cerebellar \gls*{gm} ($\Omega_{cbGM}$); and
+  6) cortical \gls*{gm} ($\Omega_{GM}$).
+Using tools in \emph{FreeSurfer} and appropriate selections of labels in the 
+  \emph{aparc} segmentation released with the \gls*{hcp} data, we extracted the $\gammaset_R$ set for the
+  reference surfaces.
+The segmentation model corresponding to this partition is shown in \autoref{fig:model}
+  and discussed in greater detail in \suppl{section S4}.
 
 \paragraph*{Ground-truth generation}
-Realistic deformation is achieved generating displacements fields that meet the theoretical
+Realistic deformation is achieved by generating displacements fields that satisfy the theoretical
   properties of distortion \citep{jezzard_correction_1995}.
 The displacements along the \gls*{pe} axis of the \gls*{dmri} image are related to the local
-  deviation of the field $\Delta B_0(\vec{r})$ from its nominal value $B_0$ as follows:
+  deviation of the field $\Delta B_0(\vec{r})$ from its nominal value $B_0$, as follows:
 
@@
   \end{equation}
-%
 where $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time, and
   $s_\text{PE}$ is the pixel size along \gls*{pe}.
-Certain \gls*{mr} sequences are designed to estimate $\Delta B_0$, obtaining
-  the so-called fieldmap.
-We derive the deformation $U_{true}$ from the fieldmap image released with
-  the corresponding packages of each dataset from the \gls*{hcp}.
-The fieldmap is unwrapped\footnote{fieldmaps are phase maps, clipped in the interval
+Certain \gls*{mr} sequences are designed to estimate $\Delta B_0$, thereby obtaining
+  the so-called field map.
+We derived the deformation $U_{true}$ from the field map image released with
+  the corresponding packages of each dataset in the \gls*{hcp}.
+The field map was unwrapped\footnote{fieldmaps are phase maps, which have been clipped in the interval
   of $[-\pi, \pi)$ [rads] or [rads/s].} and smoothed before applying \eqref{eq:fieldmap}.
-Then, the original \gls*{dmri} is warped using the resulting displacements field and fed into
-  a pipeline that process the corresponding \gls*{dti} and computes the derived scalars of
+Next, the original \gls*{dmri} was warped using the resulting displacement field and fed into
+  a pipeline to process the corresponding \gls*{dti}, thereby computing the derived scalars of
   interest (\gls*{fa} and \gls*{adc}) using \emph{MRtrix} \citep{tournier_mrtrix_2012}.
@@
 \paragraph*{Metric assessment}
-Preliminarily, we investigated the aptness of the segmentation model.
+Initially, we investigated the appropriateness of the segmentation model.
 For five test datasets, we uniformly sampled the space of distortions
@@
   (with $\epsilon \in [-1.1, 1.1]$ and $u_\text{PE}$ from \eqref{eq:fieldmap}),
-  and evaluated the data-term of the cost function \eqref{eq:energy}.
-The metric consistently displayed its minimum at the ground-truth location ($\epsilon=0.0$)
-  for all the cases (\suppl{Figure S2}).
+  and we evaluated the data term of the cost function \eqref{eq:energy}.
+The minimum metric was consistently at the ground-truth location ($\epsilon=0.0$)
+  in all cases (\suppl{Figure S2}).
 
 \paragraph*{Cross-comparison}
-A dual workflow to the general evaluation framework of \autoref{fig:evworkflows}
-  integrates the alternate \gls*{t2b} registration scheme.
+A similar workflow to the general evaluation framework used for \autoref{fig:evworkflows}
+  was employed to integrate the alternate \gls*{t2b} registration scheme.
 We reproduced the solution and settings provided with \emph{ExploreDTI}
-  \citep{leemans_exploredti_2009}, a widely used toolkit for tractography analysis of
+  \citep{leemans_exploredti_2009}, which is a widely used toolkit for tractography analysis of
   \gls*{dti}.
-\emph{ExploreDTI}, internally uses \emph{elastix} \citep{klein_elastix_2010} to
+\emph{ExploreDTI} internally employs \emph{elastix} \citep{klein_elastix_2010} to
   perform registration.
@@
 \paragraph*{Error measurement}\label{sec:experiments_evaluation}
-Since the distortions only happen along the \gls*{pe} axis of the image, we compute the
-  \gls*{swindex} as the area-weighted distance between corresponding vertices of 
-  $\gammaset_{true}$ and their estimation by the method under test $\hat{\gammaset}_{test}$:
+Distortion only occurs along the \gls*{pe} axis of the image, so we computed the
+  \gls*{swindex} as the area-weighted distance between the corresponding vertices of 
+  $\gammaset_{true}$ and their estimate obtained by the method under the test $\hat{\gammaset}_{test}$:
 
@@
   \end{equation}
-%
-  where $\vec{v}_i \subset \gammaset_\text{true}$ are the locations of the total $N_V$ vertices,
-  and $\hat{\vec{v}}_i \subset \hat{\gammaset}_{test}$ are the locations
-  recovered corresponding to $\vec{v}_i$.
-In practice, we report the \gls*{swindex} only for three surfaces of interest crucial in whole-brain
+  where $\vec{v}_i \subset \gammaset_\text{true}$ are the locations of the total $N_V$ vertices
+  and $\hat{\vec{v}}_i \subset \hat{\gammaset}_{test}$ are the recovered locations
+  that correspond to $\vec{v}_i$.
+In practice, we only report the \gls*{swindex} for three surfaces of crucial interest in whole-brain
   tractography.
-These three surfaces delineate the following regions in the model: $\Omega_{VdGM}$, $\Omega_{WM}$ and $\Omega_{GM}$.
+These three surfaces delineated the following regions in the model: $\Omega_{VdGM}$, $\Omega_{WM}$, and $\Omega_{GM}$.

### 2015-04-29 Publications/2015-NeuroImage/05-results.tex

@@
 
-\subsection{Proof of concept on digital phantoms}
+\subsection{Proof of concept using digital phantoms}
 \label{sec:results_phantom}
@@
   \includestandalone[width=\linewidth]{figure05}
-  \caption{A. Visual assessment of the results on the low resolution sets:
-    ``gyrus'' (top-left), ``L'' (top-right), ``ball'' (bottom-left),
-    and ``box'' at (bottom-right).
-  In yellow color, the recovered contours after registration are represented.
-  \Regseg showed high accuracy, as it demonstrates the almost exact location of the registered
-    contours with respect to their ground truth position depicted in green.
-  Partial volume effect turns segmentation of the sulci a challenging problem with voxel-wise
-    clustering methods, but it is successfully segmented with \regseg{}.
-  B. Quantitative evaluation of registration error in terms of average Hausdorff distance of
-    surfaces at low (left) and high (right) resolutions, demonstrating that the error is
-    consistently below the voxel size.
+  \caption{A. Visual assessment of the results obtained with the low resolution sets:
+    ``gyrus'' (top left), ``L'' (top right), ``ball'' (bottom left),
+    and ``box'' at (bottom right).
+  The contours recovered after registration are represented in yellow.
+  \Regseg achieved high accuracy because it determined the almost exact locations of the registered
+    contours with respect to their ground truth positions (shown in green).
+  The partial volume effect makes segmentation of the sulci a challenging problem with voxel-wise
+    clustering methods, but they were successfully segmented with \regseg{}.
+  B. Quantitative evaluation of registration errors in terms of the average Hausdorff distances between
+    surfaces at low (left) and high (right) resolutions, which demonstrate that the errors were
+    consistently below the size of the voxels.
     }\label{fig:phantom}
 \end{figure*}
-A total of 1200 experiments (4 phantom types $\times$ 150 random warpings $\times$ 2 resolutions) were
-  run using the workflow given in \autoref{fig:evphantoms}.
+In total, 1200 experiments (four phantom types $\times$ 150 random warpings $\times$ two resolutions) were
+  performed according to the workflow illustrated in \autoref{fig:evphantoms}.
 For each experiment, the misregistration error was measured using the Hausdorff distance
   (see \autoref{sec:experiments_evaluation}) between the theoretical $\gammaset_\text{true}$ and
-  the estimation done by \regseg{} ($\hat{\gammaset}_{test}$).
-The results showed a consistent and high accuracy, below the image resolution.
-\autoref{fig:phantom} (block C) presents the violin plots by model type corresponding
-  to the two sets of resolutions of generated phantoms.
-In order to relate average misregistration error to the resolution of the moving image,
+  the estimate obtained by \regseg{} ($\hat{\gammaset}_{test}$).
+The results demonstrated that the accuracy was consistent and high, and below the image resolution.
+\autoref{fig:phantom} (block C) shows the violin plots for each model type corresponding
+  to the two sets of resolutions for the generated phantoms.
+In order to relate the average misregistration error to the resolution of the moving image,
   we proceeded as follows.
-First, we confirmed that the vertex-wise error distributions were skewed using a Shapiro-Wilk test of
+First, we confirmed that the vertex-wise error distributions were skewed by using Shapiro-Wilk’s test of
   normality.
-All the distributions of errors under test (4 phantom types $\times$ 2 resolutions) resulted
+All of the distributions of errors in the tests (four phantom types $\times$ two resolutions) were
   non-normal with $p<0.001$.
-Consequently, we used the non-parametric Wilcoxon signed-rank test along with Bonferroni
+Consequently, we used the non-parametric Wilcoxon signed-rank test with the Bonferroni
   correction for multiple comparisons ($N=150$, for each phantom type).
-Averaged errors resulted significantly lower than voxel size with $p < (0.001 / 150)$
-  for all the tests (4 phantom types $\times$ 2 resolutions).
-Since statistical tests may not be conclusive enough, we also computed the confidence intervals,
-  reported in \autoref{tab:ci_phantom}.
+The average errors were significantly lower than the voxel size with $p < (0.001 / 150)$
+  in all tests (four phantom types $\times$ two resolutions).
+Statistical tests might not be sufficiently conclusive, so we also computed the confidence intervals,
+  as shown in \autoref{tab:ci_phantom}.
 
@@
     \end{tabular}
-    \caption{The vertex-wise Hausdorff distance between the ground-truth surfaces and their
-    corresponding estimation with \regseg{} presented distributions with 95\% CI below
-    the half-voxel size in all the phantom types.
-    CIs were computed using bootstrapping of 10$^4$ samples, and median as location statistic.}\label{tab:ci_phantom}
+    \caption{Vertex-wise Hausdorff distances between the ground-truth surfaces and their
+    corresponding estimates obtained with \regseg{}, where the distributions are shown with the 95\% CI below
+    the half-voxel size for all of the phantom types.
+    The CIs were computed by bootstrapping using 10$^4$ samples, with the median as the location statistic.}\label{tab:ci_phantom}
 \end{table}
 
-\subsection{Evaluation on real datasets and cross-comparison}\label{sec:results_hcp}
-%
-Finally, we compared the performance of \regseg{} and the standard \gls*{t2b}
+\subsection{Evaluation using real datasets and cross-comparison}\label{sec:results_hcp}
+Finally, we compared the performance of \regseg{} with that of the standard \gls*{t2b}
   method.
-Visual assessment of the 16 cases is included in the \suppl{section S5}.
-In \autoref{fig:results_real}, box A, the visual report for one subject is supplied.
-We computed the \gls*{swindex} \eqref{eq:swindex} of every surface after registration,
-  for both \regseg{} and the \gls*{t2b} methods.
-Finally, to statistically compare the results, we performed Kruskal-Wallis H-tests
+Visual assessments of the 16 cases are described in \suppl{section S5}.
+In \autoref{fig:results_real}, box A, the visual report is shown for one subject.
+We computed the \gls*{swindex} \eqref{eq:swindex} of every surface after registration
+  using both the \regseg{} and \gls*{t2b} methods.
+Finally, to compare the results, we performed Kruskal-Wallis H-tests
   (a non-parametric alternative to ANOVA) on the warping indices for the three surfaces of 
@@
   ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$).
-All the statistical tests evidenced that error distributions obtained with \regseg{} and
-  \gls*{t2b} were significantly different, and using the violin plots in box B of
-  \autoref{fig:results_real} we observed that errors were always larger for \gls*{t2b}.
-We also reported the 95\% CIs of the \gls*{swindex} for those surfaces.
-The aggregate CI for \regseg{} was 1.08 - 1.50 [mm], whereas the \gls*{t2b} method
-  yielded an aggregate CI of 2.06 - 2.43 [mm].
+All of the statistical tests showed that the error distributions obtained with \regseg{} and
+  \gls*{t2b} were significantly different, and the violin plots in box B of
+  \autoref{fig:results_real} demonstrate that the errors were always larger with \gls*{t2b}.
+We also show the 95\% CIs of the \gls*{swindex} for these surfaces.
+The aggregate CI for \regseg{} was 1.08--1.50 [mm], whereas the \gls*{t2b} method
+  yielded an aggregate CI of 2.06--2.43 [mm].
 The results of the statistical tests and the CIs are summarized in \autoref{tab:results_real}.
@@
     \end{tabular}
-    \caption{The statistical analysis of results on real data supported that \regseg{} overperformed
+    \caption{Statistical analysis of results obtained using real data, which show that \regseg{} performed better than
     the alternative \acrfull{t2b} method.
-    The distribution of errors computed for the surfaces of interest ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$)
-      and the aggregate of all surfaces (Aggreg. column) presented lower 95\% CIs for \regseg{}.
-    CIs reported in this table were computed using bootstrapping with mean as location statistic and 10$^4$ samples.
-    The Kruskal-Wallis H-tests indicated that there was a significant difference between \regseg{} and
+    The distribution of the errors computed for the surfaces of interest ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$)
+      and the aggregate of all surfaces (Aggreg. column) had lower 95\% CIs with \regseg{}.
+   The CIs in this table were computed by bootstrapping using the mean as the location statistic and with 10$^4$ samples.
+    The Kruskal-Wallis H-tests indicated that there was a significant difference between the results obtained using \regseg{} and
       the \gls*{t2b} method.
@@
   \includestandalone[width=\linewidth]{figure06}
-  \caption{A. Example of one report for visual assessment, automatically generated by the evaluation instrument.
-    Each view shows one component of the input image (in this case, the \gls*{fa} map), the ground-truth location
-    of the surfaces (green contours), and the resulting surfaces with the method under test (yellow contours).
-  First two rows show axial slices for \regseg{} and the \acrfull*{t2b} method, and the last two rows
-    show corresponding sagittal views.
-  Coronal view is omitted since it is the least informative due to the the directional property
-    of distortions.
-  Certain regions where \regseg{} overperformed the \gls*{t2b} have been zoomed in.
-  B. Violin plots of error distributions of each surface, with indication of the voxel size of the \gls*{dmri} images
-    (1.25 mm), and supporting the improved results of \regseg{} in the proposed settings.
+  \caption{A. Example of a visual assessment report, which was generated automatically by the evaluation tool.
+    Each view shows one component of the input image (in this case, the \gls*{fa} map), the ground-truth locations
+    of the surfaces (green contours), and the resulting surfaces obtained with the test method (yellow contours).
+  The first two rows show axial slices for \regseg{} and the \acrfull*{t2b} method, while the last two rows
+    show the corresponding sagittal views.
+  The coronal view is omitted because it was the least informative due to the directional property
+    of the distortions.
+  Specific regions where \regseg{} outperformed \gls*{t2b} are enlarged.
+  B. Violin plots of the error distributions for each surface, which show the voxel size of the \gls*{dmri} images
+    (1.25 mm), thereby supporting the improved results obtained by \regseg{} with the proposed settings.
   }\label{fig:results_real}

### 2015-04-29 Publications/2015-NeuroImage/07-conclusion.tex
some final fixes

--- 2015-04-06 Publications/2015-NeuroImage/07-conclusion.tex
@@
 
-\emph{Regseg} is a variational framework to simultaneously segment and
-  register 3D \gls*{dmri} data of the human brain, using within-subject
-  anatomical information as reference.
-The registration method segments the target multivariate image in several competing regions
+\Regseg{} is a variational framework for the simultaneous segmentation and
+  registration of 3D \gls*{dmri} data obtained from the human brain, where within-subject
+  anatomical information is used as a reference.
+The registration method segments the target multivariate image into several competing regions, which are
   defined explicitly by their limiting surfaces.
-The surfaces are active, and evolve on a free-form deformation field supported by B-Splines.
+The surfaces are active and they evolve on a free-form deformation field supported by the B-spline basis.
 A descent optimization strategy is guided by shape gradients computed on the current partition
   of the target image.
-\emph{Regseg} uses active contours without edges, and hence looks for
+\Regseg{} uses active contours without edges and it searches for
   homogeneous regions within the image.
-We tested \emph{regseg} on digital phantoms simulating \gls*{t1} and \gls*{t2} \gls*{mri},
+We tested \regseg{} using digital phantoms by simulating \gls*{t1} and \gls*{t2} \gls*{mri}
   warped with smooth random deformations.
@@
 
-We propose \emph{regseg} to simultaneously segment \gls*{dmri} data and register them to
-  their corresponding \gls*{t1} image of the same subject.
-Visual assessment of \emph{regseg} and cross-comparison with a widely-used technique demonstrated 
-  its accuracy.
-Moreover, \emph{regseg} does not require additional images to the minimal acquisition protocol
-  that includes only \gls*{t1} and \gls*{dmri}.
+We proposed \regseg{} for simultaneously segmenting \gls*{dmri} data and registering them according to
+  their corresponding \gls*{t1} image from the same subject.
+We demonstrated 
+  the accuracy of the proposed method based on visual assessments of the results obtained by \regseg{} and cross-comparisons with a widely used technique.
+Moreover, \regseg{} does not require any images in addition to the minimal acquisition protocol,
+  which only utilizes \gls*{t1} and \gls*{dmri}.
+As well as the proposed application to \gls*{dmri} data, other potential uses of \regseg{} are
+  atlas-based segmentation and tracking objects in time-series.
 
@@
 \label{sec:availability}
-We considered the reproducibility of our results a design requirement.
-Therefore, we used real data fetched from a publicly available repository
-  (the Human Connectome Project \citep{essen_human_2012}) and all the software
-  involved in this paper is also publicly released.
-\emph{Regseg} is implemented on top of ITK-4.6 (Insight Registration and 
+We considered the reproducibility of our results as a design requirement.
+Therefore, we used real data obtained from a publicly available repository
+  (the Human Connectome Project \citep{essen_human_2012}) and all of the software
+  utilized in this study is also publicly available.
+\Regseg{} was implemented on top of ITK-4.6 (Insight Registration and 
   Segmentation Toolkit, \url{http://www.itk.org}).
-The evaluation instruments (\autoref{fig:evworkflows}), are implemented using
+The evaluation instruments (\autoref{fig:evworkflows}) were implemented using
   \emph{nipype} \citep{gorgolewski_nipype_2011} to assess their reproducibility.
-All research elements (data, source code, figures, manuscript sources, etc.) in this paper
-  are made publicly available under a unique package \citep{esteban_acweregistration_2015}.
+All of the research elements (data, source code, figures, manuscript sources, etc.) involved in this study
+  are publicly available under a unique package \citep{esteban_acweregistration_2015}.

### 2015-04-29 Publications/2015-NeuroImage/08-acknowledgements.tex

@@
 \section*{Author contributions}
-All the authors contributed in editing the paper.
+All the authors contributed to this study.
 OE implemented the method, designed and conducted the experiments, wrote the paper,
-  simulated the phantoms and prepared the real data.
+  simulated the phantoms, and prepared the real data.
 DZ devised and drafted the registration method, generated early phantom datasets, and
-  collaborated with the implementation of the method.
-AD, MBC and MJLC interpreted the results.
-AD, MBC, MJLC, JPT and AS advised on all aspects of the work.
+  collaborated in the implementation of the method.
+AD, MBC, and MJLC interpreted the results.
+AD, MBC, MJLC, JPT, and AS advised on all aspects of the study.
 
@@
   V. Estellers for early discussions at the beginning of this project,
-  and L. A. Vese for her support during OE's research visits in her laboratory.
+  and L. A. Vese for her support during OE's research visits to her laboratory.
 We also thank A. Leemans for kindly sharing a p-code version of \emph{ExploreDTI} from
-  which settings for \emph{elastix} could be extracted.
+  which the settings for \emph{elastix} could be extracted.
 
-DZ was supported by the Swiss National Science Foundation under grants PBELP2-137727 
-  and P300P2-147778, and NSF-DMS 1418812
-This study is supported by: the Spain's Ministry of Science and Innovation
+DZ was supported by the Swiss National Science Foundation under grants PBELP2-137727, 
+  P300P2-147778, and NSF-DMS 1418812
+This study was supported by the Spanish Ministry of Science and Innovation
   (projects TEC-2013-48251-C2-2-R and INNPACTO XIORT), Comunidad de Madrid (TOPUS) and
-  European Regional Development Funds; the Center for Biomedical Imaging
+  European Regional Development Funds, the Center for Biomedical Imaging
   (CIBM) of the Geneva and Lausanne Universities and the EPFL, as well as the
-  Leenaards and Louis Jeantet foundations.
+  Leenaards and Louis Jeantet Foundations.

### 2015-04-29 Publications/2015-NeuroImage/ELS285_article_edited.tex

--- 2015-04-29 Publications/2015-NeuroImage/04-methods.tex
@@
-% -*- root: 00-main.tex -*-
-\section{Methods}
+\documentclass[3p,authoryear,fleqn]{elsarticle}
+\journal{ArXiv archiving for pre-publication review}
+
+\usepackage[mode=buildnew]{standalone}
+\usepackage{gincltex}
+\usepackage{glossaries}
+\usepackage{booktabs} \usepackage{multirow}
+\usepackage[colorinlistoftodos]{todonotes}
+\usepackage{xifthen}
+\usepackage{xspace}
+\usepackage{graphicx}
+\graphicspath{{figures/}}
+
+\usepackage{amsmath}
+\usepackage{amssymb}
+\DeclareMathOperator{\tr}{tr}
+\DeclareMathOperator{\argmax}{argmax}
+\DeclareMathOperator{\argmin}{argmin}
+\DeclareMathOperator{\diag}{diag}
+\DeclareMathOperator{\dist}{dist}
+\DeclareMathOperator{\const}{Const.}
+\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
+\providecommand{\mdist}[2]{ \mathcal{D}_{#2}^2(\mathbf{#1}) }
+\providecommand{\omegaset}{\ensuremath{\boldsymbol{\Omega}}}
+\providecommand{\gammaset}{\ensuremath{\boldsymbol{\Gamma}}}
+\providecommand{\regseg}{\emph{regseg}}
+\providecommand{\Regseg}{\emph{Regseg}}
+
+\let\oldhat\hat
+\renewcommand{\vec}[1]{\mathbf{#1}}
+
+\providecommand{\nvec}[1]{\hat{\mathbf{#1}}}
+\providecommand{\suppl}[1]{\href{http://dx.doi.org/mydoi}{Supplemental Materials}, #1}
+\providecommand{\isores}[2][]{#2\ensuremath{\times}#2\ensuremath{\times}#2\ifthenelse{\equal{#1}{}}{}{ [#1]}\xspace}
+
+\usepackage{url}
+\usepackage[colorlinks=true,linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
+\usepackage{doi}
+
+\hyphenation{op-tical net-works semi-conduc-tor an-isot-ropy regis-tra-tion iso-tropic Free-Surfer align-ed pre-sent work-flow data-set data-sets}
+
+\makeatletter{}\newacronym{mr}{MR}{magnetic resonance}
+\newacronym{mri}{MRI}{magnetic resonance imaging}
+\newacronym{dwi}{DWI}{diffusion weighted image}
+\newacronym{dmri}{dMRI}{diffusion MRI}
+\newacronym{dw}{DW}{diffusion weighted}
+\newacronym{dti}{DTI}{diffusion tensor image}
+\newacronym{t1}{T1w}{T1 weighted}
+\newacronym{t2}{T2w}{T2 weighted}
+\newacronym{csf}{CSF}{cerebrospinal fluid}
+\newacronym{wm}{WM}{white matter}
+\newacronym{gm}{GM}{grey matter}
+\newacronym{epi}{EPI}{echo-planar imaging}
+\newacronym{gre}{GRE}{gradient echo sequence}
+\newacronym{fa}{FA}{fractional anisotropy}
+\newacronym{md}{MD}{mean diffusivity}
+\newacronym{adc}{ADC}{apparent diffusion coefficient}
+\newacronym{acwe}{ACWE}{active contours without edges}
+\newacronym{adf}{ADF}{active deformation field}
+\newacronym{map}{MAP}{maximum a posteriori}
+\newacronym{snr}{SNR}{signal-to-noise ratio}
+\newacronym{pve}{PVE}{partial volume effect}
+\newacronym{roi}{ROI}{region of interest}
+\newacronym{mrf}{MRF}{Markov Random Field}
+\newacronym{pde}{PDE}{partial differential equation}
+\newacronym{swindex}{sWI}{surface warping index}
+\newacronym{wi}{WI}{Warping index}
+\newacronym{nof}{NoF}{number of fibers}
+\newacronym{t2b}{T2B}{T2w-registration based}
+\newacronym{fft}{FFT}{fast fourier transform}
+\newacronym{hcp}{HCP}{Human Connectome Project}
+\newacronym{pe}{PE}{phase-encoding}
+\newacronym{ci}{CI}{confidence interval}
+
+\makeglossaries
+
+\usepackage[T1]{fontenc}
+\usepackage{charter}
+
+\usepackage{array}
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+
+\usepackage{ifpdf}
+
+\ifpdf
+  \usepackage[switch]{lineno}
+  \else
+  \makeatletter
+  \def\temp{dvips.def}
+  \ifx\Gin@driver\temp
+  \def\Ginclude@graphics#1{\def\temp{#1}---image \expandafter\strip@prefix\meaning\temp---}
+  \fi
+  \makeatother
+\fi
+
+\def\bibsection{\section*{References}}
+\begin{document}
+\begin{frontmatter}
+
+\title{Active contours-driven registration method for the structure-informed segmentation of diffusion MR images}
+
+\author[bit,ciber]{Oscar~Esteban\corref{corrauthor}}
+\cortext[corrauthor]{Corresponding author}
+\ead{phd@oscaresteban.es}
+\author[ucla]{Dominique~Zosso}
+\author[scil,lts5]{Alessandro~Daducci}
+\author[chuv,lts5]{Meritxell~Bach-Cuadra}
+\author[bit,ciber]{Mar\'ia-J.~Ledesma-Carbayo}
+\author[lts5]{Jean-Philippe~Thiran}
+\author[bit,ciber]{Andres~Santos}
+
+\address[bit]{Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on, Universidad Polit\'ecnica de Madrid, Madrid, Spain}
+\address[ciber]{Centro de Investigaci\'on Biom\'edica en Red en Bioingenier\'ia, Biomateriales y Nanomedicina (CIBER-BBN), Zaragoza, Spain}
+\address[ucla]{Department of Mathematics, University of California,
+Los Angeles (UCLA), Los Angeles, CA, US}
+\address[scil]{Computer Science Department, Faculty of Science, Universit\'e de Sherbrooke, 2500 Boulevard Universit\'e, Sherbrooke, QC J1K 2R1, Canada}
+\address[lts5]{Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
+F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland}
+\address[chuv]{Dept. of Radiology, CIBM, University
+Hospital Center (CHUV) and University of Lausanne (UNIL), Lausanne, Switzerland}
+
+\begin{abstract}
+\makeatletter{}Current applications of whole-brain tractography to \acrlong*{dmri} data require highly precise
+  delineations of anatomical structures, which are usually projected from anatomical \acrlong*{t1} images
+  by registration.
+In this study, we propose \regseg{}, which is a simultaneous segmentation and registration method
+  that uses active contours without edges extracted from structural images.
+The contours evolve through a free-form deformation field supported by the B-spline basis
+  to optimally map the contours onto the data in the target space.
+We tested the functionality of \regseg{} using four digital phantoms warped with known and
+  randomly generated deformations, where subvoxel accuracy was achieved.
+We then applied \regseg{} to a registration/segmentation task using 16 real diffusion MRI
+  datasets from the \acrlong*{hcp}, which were warped by realistic and nonlinear distortions that are typically
+  present in these data.
+We computed the misregistration error of the contours estimated by \regseg{} with respect
+  to their theoretical location using the ground truth, thereby obtaining a 95\% CI of 0.56--0.66 mm
+  distance between corresponding mesh vertices, which was below the 1.25 mm isotropic resolution of the images.
+We also compared the performance of our proposed method with a widely used registration tool, which showed 
+  that \regseg{} outperformed this method in our settings.
+
+\end{abstract}
+
+\begin{keyword}
+active~contours \sep cortical~parcellation \sep diffusion MRI \sep nonlinear~registration \sep segmentation \sep susceptibility~distortion.
+\end{keyword}
+
+\end{frontmatter}
+
+\linenumbers
+\makeatletter{}\section{Introduction}\label{sec:introduction}
+The accurate delineation of \gls*{wm} in \gls*{dmri} and the fusion of prior
+  anatomical information extracted from a \gls*{t1} image for the same subject
+  are crucial in a range of applications based on tractography, such as
+  the extraction of structural connectivity \citep{craddock_imaging_2013} or
+  tract-based spatial statistics \citep{smith_tractbased_2006}.
+However, the precise segmentation of \gls*{dmri} data and the coregistration of anatomical
+  images are difficult due to several limitations.
+First, \gls{dmri} images have a resolution that is much higher than that of the imaged
+  microstructural features \citep{basser_microstructural_1996}.
+Therefore, voxels located in structural discontinuities are affected by partial
+  voluming of the signal sources.
+Second, \gls*{dmri} schemes probe the diffusion process within the brain at 
+  many angles to obtain \glspl*{dwi}, which are completed by one or more baseline (\emph{b0}) 
+  volumes without directional gradients.
+The extremely low \gls*{snr} and the high dimensionality of \glspl*{dwi} prevent their
+  direct use in segmentation.
+The low contrast between \gls*{gm} and \gls*{wm} in the \emph{b0} volumes also makes them unsuitable for 
+  segmentation.
+Finally, \gls*{dmri} images are acquired using \gls*{epi} to speed up the acquisition process but
+  at the cost of introducing geometrical distortion, as well as signal degradation and
+  destruction \citep{jezzard_correction_1995}.
+The artifacts generated have major impacts on the anatomy extracted
+  from \gls*{dmri}, particularly in certain fiber bundles \citep{irfanoglu_effects_2012}.
+These limitations prevent segmentation in the native \gls*{dmri} space, including
+  registration-based approaches using anatomical images as the sources of structural
+  information.
+
+Early attempts to delineate the \gls*{wm} involved thresholding the 
+  \gls*{fa}\footnote{the \gls*{fa} is a scalar parameter of diffusion derived from
+  the \gls*{dmri} data.} map.
+However, the mask and subsequent analyses are highly dependent on the threshold set
+  \citep{taoka_fractional_2009}.
+To overcome the unreliability of \gls*{fa} thresholding,
+  \cite{zhukov_level_2003} proposed active contours with edges represented
+  by level sets, which evolve on directionally invariant scalar maps.
+\cite{rousson_level_2004} successfully segmented the corpus callosum with
+  region-based level sets based on the eigenvalues of tensors fitted using the
+  \gls*{dmri} data.
+Many studies have addressed the definition of appropriate features for clustering,
+  such as the 5D[A1] representation proposed by \cite{jonasson_segmentation_2005}.
+Other efforts include mixed models based on sets of directionally invariant maps
+  \citep{liu_brain_2007}, iterative \citep{hadjiprocopis_unbiased_2005} and
+  hierarchical \citep{lu_segmentation_2008} clustering,
+  graph-cuts \citep{han_experimental_2009},
+  and volume fraction modeling \citep{kumazawa_improvement_2013}.
+
+To address the segmentation problem by registration, \cite{saad_new_2009} 
+  used Pearson’s correlation coefficient to obtain a linear alignment of the \gls*{t1} and 
+  the \emph{b0}.
+Similarly, a linear registration method was employed by \emph{bbregister} \citep{greve_accurate_2009},
+which uses active contours with edges to search for intensity boundaries in the \emph{b0}
+  image.
+The active contours are initialized using surfaces extracted from the 
+  \gls*{t1} using \emph{FreeSurfer} \citep{fischl_freesurfer_2012}:
+  the \emph{pial} surface (exterior of the cortical \gls*{gm}) and the \emph{white}
+  surface (the \gls*{wm}/\gls*{gm} interface). 
+The \emph{b0} image only includes a detectable frontier for the pial surface, so 
+  \emph{bbregister} is limited to aligning the cortical layer in this
+  application.
+This tool has become the standard method because of its proven robustness, although the 
+  distortions found in \gls*{dmri} are nonlinear.
+To overcome this issue, \emph{bbregister} excludes the
+  regions that are typically warped by artifacts from the boundary search.
+However, because the distortion is not considered, it must be addressed separately.
+Nonlinear registration has been performed successfully between \gls*{t2} and \emph{b0}
+  images based on their similarity as a unique method for correcting distortion
+  \citep{kybic_unwarping_2000,studholme_accurate_2000,wu_comparison_2008,tao_variational_2009}.
+However, further registration of the \gls*{t1} and \gls*{t2} images is still required to map the anatomical
+  information (and for \gls*{wm} segmentation) into the \gls*{dmri} space.
+
+Therefore, given these issues, a method that simultaneously performs
+  segmentation in the native \gls*{dmri} space and registration of the corresponding \gls*{t1} image
+  onto the \gls*{dmri} data could provide the optimal solution.
+To the best of our knowledge, this strategy has never been proposed for the application described above.
+Previously, joint segmentation and registration have been applied successfully to other problems
+such as longitudinal object tracking \citep{paragios_level_2003} and atlas-based
+  segmentation \citep{gorthi_active_2011}.
+The most thorough approach integrates active contours during image registration 
+  methods.
+\cite{unal_coupled_2005}, and later \cite{wang_joint_2006},
+  improved an existing method \citep{yezzi_variational_2003} based on linear registration
+  to the nonlinear case by implementing a free-form deformation field.
+\cite{droske_mumfordshah_2009} reviewed the existing techniques and proposed two different
+  approaches for applying the Mumford-Shah functional \citep{mumford_optimal_1989} during simultaneous
+  registration and segmentation by propagating the deformation field from
+  the contours onto the whole image definition.
+Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
+  registration method in 2D using level sets and a nonlinear elasticity smoother on the
+  displacement vector field, which preserves the topology even with very high deformation.
+Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
+  level set function to register several active contours during the application
+  of atlas-based segmentation.
+
+The hypothesis tested in the present study is that registration and segmentation
+  problems in \gls*{dmri} can be solved simultaneously, thereby increasing the geometrical
+  accuracy of the process.
+Thus, we propose a tool called \regseg{} that exploits the prior information of shapes
+  extracted using \gls*{t1} images to register the anatomical reference
+  to \gls*{dmri} space, thereby implicitly segmenting these data.
+The proposed approach uses active contours without edges \citep{chan_active_2001}, which evolve to drive a
+  free-form deformation field of B-spline basis functions.
+Optimization is performed using a descent strategy with explicit shape gradients
+  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
+Therefore, unlike most previous methods, \regseg{} does not implement level sets.
+The nonlinear distortion is aligned with one of the imaging axes (see 
+  \autoref{sec:human_connectome}), so \regseg{} includes an anisotropic regularizer for
+  the displacements field proposed by \cite{nagel_investigation_1986}.
+Finally, we evaluated \regseg{} using an extension of our instrumentation framework
+  \citep{esteban_simulationbased_2014}, which simulates known and realistic distortions
+  based on \gls*{dmri} data.
+The evaluation included a comparison with \gls*{t2b} correction by integrating it within the framework
+  of an in-house implementation of the method.
+
+\makeatletter{}\section{Methods}
 \label{sec:methods}
@@
 These three surfaces delineated the following regions in the model: $\Omega_{VdGM}$, $\Omega_{WM}$, and $\Omega_{GM}$. 
+\makeatletter{}\section{Results}
+\label{sec:results}
+
+\subsection{Proof of concept using digital phantoms}
+\label{sec:results_phantom}
+
+\begin{figure*}
+  \centering
+  \includestandalone[width=\linewidth]{figure05}
+  \caption{A. Visual assessment of the results obtained with the low resolution sets:
+    ``gyrus'' (top left), ``L'' (top right), ``ball'' (bottom left),
+    and ``box'' at (bottom right).
+  The contours recovered after registration are represented in yellow.
+  \Regseg achieved high accuracy because it determined the almost exact locations of the registered
+    contours with respect to their ground truth positions (shown in green).
+  The partial volume effect makes segmentation of the sulci a challenging problem with voxel-wise
+    clustering methods, but they were successfully segmented with \regseg{}.
+  B. Quantitative evaluation of registration errors in terms of the average Hausdorff distances between
+    surfaces at low (left) and high (right) resolutions, which demonstrate that the errors were
+    consistently below the size of the voxels.
+    }\label{fig:phantom}
+\end{figure*}
+In total, 1200 experiments (four phantom types $\times$ 150 random warpings $\times$ two resolutions) were
+  performed according to the workflow illustrated in \autoref{fig:evphantoms}.
+For each experiment, the misregistration error was measured using the Hausdorff distance
+  (see \autoref{sec:experiments_evaluation}) between the theoretical $\gammaset_\text{true}$ and
+  the estimate obtained by \regseg{} ($\hat{\gammaset}_{test}$).
+The results demonstrated that the accuracy was consistent and high, and below the image resolution.
+\autoref{fig:phantom} (block C) shows the violin plots for each model type corresponding
+  to the two sets of resolutions for the generated phantoms.
+In order to relate the average misregistration error to the resolution of the moving image,
+  we proceeded as follows.
+First, we confirmed that the vertex-wise error distributions were skewed by using Shapiro-Wilk’s test of
+  normality.
+All of the distributions of errors in the tests (four phantom types $\times$ two resolutions) were
+  non-normal with $p<0.001$.
+Consequently, we used the non-parametric Wilcoxon signed-rank test with the Bonferroni
+  correction for multiple comparisons ($N=150$, for each phantom type).
+The average errors were significantly lower than the voxel size with $p < (0.001 / 150)$
+  in all tests (four phantom types $\times$ two resolutions).
+Statistical tests might not be sufficiently conclusive, so we also computed the confidence intervals,
+  as shown in \autoref{tab:ci_phantom}.
+
+\begin{table}
+		\centering
+		\footnotesize
+    \tabcolsep=0.1cm
+    \begin{tabular}{lccccc}
+    Res.   & ``gyrus'' & ``ball''  & ``box''   & ``L''     & Aggreg.    \\\hline
+    1.0mm  & 0.18-0.38 & 0.31-0.45 & 0.34-0.42 & 0.34-0.40 & 0.34-0.38  \\
+    2.0mm  & 0.59-0.60 & 0.65-0.76 & 0.68-0.71 & 0.67-0.77 & 0.64-0.66  \\
+    \hline
+    \end{tabular}
+    \caption{Vertex-wise Hausdorff distances between the ground-truth surfaces and their
+    corresponding estimates obtained with \regseg{}, where the distributions are shown with the 95\% CI below
+    the half-voxel size for all of the phantom types.
+    The CIs were computed by bootstrapping using 10$^4$ samples, with the median as the location statistic.}\label{tab:ci_phantom}
+\end{table}
+
+\subsection{Evaluation using real datasets and cross-comparison}\label{sec:results_hcp}
+Finally, we compared the performance of \regseg{} with that of the standard \gls*{t2b}
+  method.
+Visual assessments of the 16 cases are described in \suppl{section S5}.
+In \autoref{fig:results_real}, box A, the visual report is shown for one subject.
+We computed the \gls*{swindex} \eqref{eq:swindex} of every surface after registration
+  using both the \regseg{} and \gls*{t2b} methods.
+Finally, to compare the results, we performed Kruskal-Wallis H-tests
+  (a non-parametric alternative to ANOVA) on the warping indices for the three surfaces of 
+  interest selected in \autoref{sec:experiments_evaluation}
+  ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$).
+All of the statistical tests showed that the error distributions obtained with \regseg{} and
+  \gls*{t2b} were significantly different, and the violin plots in box B of
+  \autoref{fig:results_real} demonstrate that the errors were always larger with \gls*{t2b}.
+We also show the 95\% CIs of the \gls*{swindex} for these surfaces.
+The aggregate CI for \regseg{} was 1.08--1.50 [mm], whereas the \gls*{t2b} method
+  yielded an aggregate CI of 2.06--2.43 [mm].
+The results of the statistical tests and the CIs are summarized in \autoref{tab:results_real}.
+
+\begin{table}
+		\centering
+		\footnotesize
+		\tabcolsep=0.08cm
+    \begin{tabular}{cccccc}
+    & & $\Gamma_{VdGM}$  & $\Gamma_{WM}$ & $\Gamma_{pial}$ & Aggreg. \\
+    \hline
+    \multirow{2}{*}{CI}
+       & \regseg{}        & 0.50 - 0.78 & 0.50 - 0.55 & 0.66 - 0.73 & 0.56 - 0.66 \\
+       & T2B                  & 1.78 - 2.58 & 1.94 - 2.36 & 2.16 - 2.58 & 2.05 - 2.39 \\
+    \hline
+    \multirow{2}{*}{H-tests}
+       & p-value  & 4.1$\times$10$^{-6}$& 2.3$\times$10$^{-6}$& 2.3$\times$10$^{-6}$ & 1.8$\times$10$^{-16}$ \\
+       & H-stat   & 21.20               & 22.31               & 22.31                & 67.85              \\
+    \hline
+    \end{tabular}
+    \caption{Statistical analysis of results obtained using real data, which show that \regseg{} performed better than
+    the alternative \acrfull{t2b} method.
+    The distribution of the errors computed for the surfaces of interest ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$)
+      and the aggregate of all surfaces (Aggreg. column) had lower 95\% CIs with \regseg{}.
+   The CIs in this table were computed by bootstrapping using the mean as the location statistic and with 10$^4$ samples.
+    The Kruskal-Wallis H-tests indicated that there was a significant difference between the results obtained using \regseg{} and
+      the \gls*{t2b} method.
+    }\label{tab:results_real}
+\end{table}
+
+\begin{figure*}
+  \centering
+  \includestandalone[width=\linewidth]{figure06}
+  \caption{A. Example of a visual assessment report, which was generated automatically by the evaluation tool.
+    Each view shows one component of the input image (in this case, the \gls*{fa} map), the ground-truth locations
+    of the surfaces (green contours), and the resulting surfaces obtained with the test method (yellow contours).
+  The first two rows show axial slices for \regseg{} and the \acrfull*{t2b} method, while the last two rows
+    show the corresponding sagittal views.
+  The coronal view is omitted because it was the least informative due to the directional property
+    of the distortions.
+  Specific regions where \regseg{} outperformed \gls*{t2b} are enlarged.
+  B. Violin plots of the error distributions for each surface, which show the voxel size of the \gls*{dmri} images
+    (1.25 mm), thereby supporting the improved results obtained by \regseg{} with the proposed settings.
+	}\label{fig:results_real}
+\end{figure*} 
+\makeatletter{}\section{Discussion}
+\label{sec:discussion}
+
+\paragraph*{Accuracy tests}
+The hypothesis tested in our study was that reliable image registration can be performed
+  by searching for homogeneous regions in a multispectral image, which correspond to precise contours
+  from an atlas, or extracted from another image (i.e., a different time step).
+We demonstrated that active contours without edges can be used successfully to drive a
+  deformation supported by B-spline basis functions with digital phantoms.
+We randomly deformed four different phantom models to mimic three homogeneous regions
+  (\gls*{wm}, \gls*{gm}, and \acrlong*{csf}) and we used them to simulate \gls*{t1} and \gls*{t2}
+  images at two resolution levels.
+After registration with \regseg{}, we measured the Hausdorff distance between the
+  projected contours obtained using the ground-truth warping and our estimates.
+We concluded that the errors were significantly lower than the voxel resolution.
+We also assessed the 95\% \gls*{ci}, which yielded an aggregate interval of
+  0.64--0.66 [mm] for the low resolution phantoms (2.0 mm isotropic voxel) and
+  0.34--0.38 [mm] for the high resolution phantoms (1.0 mm isotropic).
+Therefore, we also concluded that the error was bounded above by half of the
+  voxel spacing.
+
+\paragraph*{Application to real data}
+We designed \regseg{} as a method for segmenting \gls*{dmri} data by exploiting the
+  anatomical information extracted from the corresponding \gls*{t1} image of the subject.
+Applications in whole-brain tractography \citep{smith_tractbased_2006,craddock_imaging_2013}
+  usually solve this problem with a two-step approach.
+First, the images are corrected for nonlinear distortions using auxiliary acquisitions
+  such as field maps \citep{jezzard_correction_1995}, \gls*{t2} images \citep{kybic_unwarping_2000}.
+Second, the segmentation is projected from a reference \gls*{t1} image using linear
+  registration \citep{greve_accurate_2009}.
+\Regseg{} addresses this joint problem in a single step and it does not require any additional
+  acquisition other than the minimal protocol using only \gls*{t1} and \gls*{dmri} images.
+This situation is found commonly in historical datasets.
+
+We evaluated \regseg{} in a real environment using the experimental framework presented
+  in \autoref{fig:evworkflows}.
+We processed 16 subjects from the \gls*{hcp} database using both \regseg{}
+  and an in-house replication of the \acrfull*{t2b} method.
+\Regseg{} obtained very high accuracy, with an aggregate 95\% \gls*{ci} of 0.56--0.66 [mm], which was
+  below the pixel size of 1.25 mm.
+The misregistration error that remained after \regseg{} was significantly lower ($p < 0.01$) than the
+  error corresponding to the \gls*{t2b} correction according to Kruskal-Wallis H-tests
+  (\autoref{tab:results_real}).
+Visual inspections of all the results (\suppl{section S5}) and the violin plots in
+  \autoref{fig:results_real} confirmed that \regseg{} performed better than the \gls*{t2b} method
+  in our settings.
+We carefully configured the \gls*{t2b} method using the same algorithm and the 
+  same settings employed widely in software, but cross-comparison experiments are prone to
+  so-called \emph{instrumentation bias} \citep{tustison_instrumentation_2013}.
+Therefore, these results do not prove that \regseg{} \emph{is better than} \gls*{t2b}.
+However, our results suggest that \regseg{} is a reliable option in this application field.
+In addition, the \gls*{t2b} may introduce an additional (and small) error during the necessary
+  registration of \gls*{t2} in the \gls*{t1} space.
+
+%   with the theory-based displacement field that can be estimated with fieldmaps.
+
+\makeatletter{}\section*{Conclusion}
+\label{sec:conclusion}
+
+\Regseg{} is a variational framework for the simultaneous segmentation and
+  registration of 3D \gls*{dmri} data obtained from the human brain, where within-subject
+  anatomical information is used as a reference.
+The registration method segments the target multivariate image into several competing regions, which are
+  defined explicitly by their limiting surfaces.
+The surfaces are active and they evolve on a free-form deformation field supported by the B-spline basis.
+A descent optimization strategy is guided by shape gradients computed on the current partition
+  of the target image.
+\Regseg{} uses active contours without edges and it searches for
+  homogeneous regions within the image.
+We tested \regseg{} using digital phantoms by simulating \gls*{t1} and \gls*{t2} \gls*{mri}
+	warped with smooth random deformations.
+The resulting misregistration of the contours was significantly lower than the image resolution
+  of the phantoms.
+
+We proposed \regseg{} for simultaneously segmenting \gls*{dmri} data and registering them according to
+  their corresponding \gls*{t1} image from the same subject.
+We demonstrated 
+  the accuracy of the proposed method based on visual assessments of the results obtained by \regseg{} and cross-comparisons with a widely used technique.
+Moreover, \regseg{} does not require any images in addition to the minimal acquisition protocol,
+  which only utilizes \gls*{t1} and \gls*{dmri}.
+As well as the proposed application to \gls*{dmri} data, other potential uses of \regseg{} are
+  atlas-based segmentation and tracking objects in time-series.
+
+\section*{Availability and reproducibility statement}
+\label{sec:availability}
+We considered the reproducibility of our results as a design requirement.
+Therefore, we used real data obtained from a publicly available repository
+  (the Human Connectome Project \citep{essen_human_2012}) and all of the software
+  utilized in this study is also publicly available.
+\Regseg{} was implemented on top of ITK-4.6 (Insight Registration and 
+  Segmentation Toolkit, \url{http://www.itk.org}).
+The evaluation instruments (\autoref{fig:evworkflows}) were implemented using
+  \emph{nipype} \citep{gorgolewski_nipype_2011} to assess their reproducibility.
+All of the research elements (data, source code, figures, manuscript sources, etc.) involved in this study
+  are publicly available under a unique package \citep{esteban_acweregistration_2015}.
+
+\makeatletter{}\section*{Author contributions}
+All the authors contributed to this study.
+OE implemented the method, designed and conducted the experiments, wrote the paper,
+  simulated the phantoms, and prepared the real data.
+DZ devised and drafted the registration method, generated early phantom datasets, and
+  collaborated in the implementation of the method.
+AD, MBC, and MJLC interpreted the results.
+AD, MBC, MJLC, JPT, and AS advised on all aspects of the study.
+
+\section*{Acknowledgments}
+The authors thank Y. Alem\'an and G. Wollny for their thorough reviews of the manuscript,
+  V. Estellers for early discussions at the beginning of this project,
+  and L. A. Vese for her support during OE's research visits to her laboratory.
+We also thank A. Leemans for kindly sharing a p-code version of \emph{ExploreDTI} from
+  which the settings for \emph{elastix} could be extracted.
+
+DZ was supported by the Swiss National Science Foundation under grants PBELP2-137727, 
+  P300P2-147778, and NSF-DMS 1418812
+This study was supported by the Spanish Ministry of Science and Innovation
+  (projects TEC-2013-48251-C2-2-R and INNPACTO XIORT), Comunidad de Madrid (TOPUS) and
+  European Regional Development Funds, the Center for Biomedical Imaging
+  (CIBM) of the Geneva and Lausanne Universities and the EPFL, as well as the
+  Leenaards and Louis Jeantet Foundations.
+
+\appendix
+\numberwithin{equation}{section}
+\makeatletter{}\renewcommand{\theequation}{A.\arabic{equation}}
+\renewcommand{\thesubsection}{Appendix \arabic{subsection}}
+
+\section*{Appendix}
+
+\subsection{Simplifying the regularization term}\label{app:reg_term}
+The exponentials of the Thikonov regularization prior \eqref{eq:thikonov} have the general form
+  $\vec{v}^T \mathbf{M} \vec{v}$.
+If $\mathbf{M}$ is a $n \times n$ diagonal matrix such that $\mathbf{M} = \vec{m} \, \mathbf{I}_n$, 
+  then:
+
+\begin{equation*}
+\vec{v}^T \mathbf{M} \vec{v} = \vec{m} \cdot (\vec{v}^T \mathbf{I}_n \vec{v}) = \vec{m} \cdot \vec{v}^{\circ2},
+\end{equation*}
+  where we have introduced the Hadamard power notation\footnote{The Hadamard power of a matrix or a vector
+  is the power of its elements $\mathbf{M}^{\circ p} = ({m_{ij}}^{p})$}.
+
+In general, the anisotropy is aligned with the imaging axes, so 
+  $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:energy} can be simplified to diagonal matrices, such that
+  $\mathbf{A}= \, \boldsymbol{\alpha}\,\vec{I}_n$ and
+  $\mathbf{B}= \, \boldsymbol{\beta}\,\vec{I}_n$.
+By substituting into equation \eqref{eq:energy}, we obtain:
+
+  \begin{align}
+  E(\vec{u}) &= \const + \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} +   \int_{\Omega} \frac12 \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2} + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}.
+  \label{eq:app_energy}
+  \end{align}
+
+\bibliographystyle{mystyle}
+\bibliography{Remote}
+
+\end{document}
+[A1]I understand that many of the abbreviations in this manuscript may be common abbreviations, however, depending on the journal to which you are submitting, it may be necessary to provide definitions in accordance with your target journal author abbreviation guidelines.

### 2015-04-29 Publications/2015-NeuroImage/supplemental.tex

@@
 % @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-04-08 13:28:07
+% @Last Modified time: 2015-04-21 13:22:12
 
 \documentclass[a4paper]{report}
-
-\usepackage{titlesec}
-\newcommand{\sectionbreak}{\clearpage}
-
 \usepackage{aliascnt}
@@
 \usepackage{cancel}
+
+\usepackage{tikz,caption}
+\usetikzlibrary{calc}
+\DeclareCaptionFormat{overlay}{\gdef\capoverlay{#1#2#3\par}}
+\DeclareCaptionStyle{overlay}{format=overlay}
+
 %\usepackage{algorithmic}
@@
 \usepackage{doi}
+\def\bibsection{\section*{References}}
 
@@
 \title{Supplemental Materials: \\ \emph{Active-contours driven registration method for structure-informed segmentation of diffusion MR images}}
-\author{Oscar Esteban and Dominique Zosso}
-\date{February 2015}
+\author{Esteban O., Zosso D., Daducci A., Bach-Cuadra M., \\ Ledesma-Carbayo MJ., Thiran JP, and Santos A.}
+\date{April 2015}
 
 \maketitle
-\section{Extensions to the mathematical formulation of the methods}
+\section{Extensions to the mathematical formulation of the methods}\label{sec:maths_extension}
 
@@
 
-\newpage
+\clearpage
 \section{Parameter settings and implementation details of \emph{regseg}}
@@
 
-\begin{figure*}[b]
+\begin{figure}[b]
   \includegraphics[width=\linewidth]{Suppl-figure02.pdf}
@@
     \label{fig:convreport}
-\end{figure*}
+\end{figure}
 
-\newpage
+\clearpage
 \section{Instruments for evaluation}
@@
 
-\newpage
+\clearpage
 \section{Model considerations}
 
-\begin{figure*}[!ht]
-	\includegraphics[width=\linewidth]{Suppl-figure01.pdf}
-	\caption{Evaluating the joint distribution}\label{fig:jointplot}
-\end{figure*}
+% \begin{figure*}
+%   \centering
+%   \includestandalone[width=\linewidth]{figure06}
+%   \caption{}label{fig:results_real}
+% \end{figure*}
 
-\newpage
+\begin{figure}[!ht]
+  \centering
+  \captionsetup{format=overlay}
+  \caption{\textbf{Evaluating the joint distribution}.
+  Using kernel density estimation, the joint distribution of the \gls*{fa} and the \gls{adc}
+    is presented for our selected model (bottom-left).
+  Along with the joint plot, the marginal distributions of each model component
+    are represented below (for the \gls*{fa}) and besides (for the \gls*{adc}) of their corresponding
+    scalar map.}
+  \begin{tikzpicture}
+  \node[anchor=south west, inner sep=0](image) at (0,0)
+	{\includegraphics[width=\linewidth]{Suppl-figure01.pdf}};
+    \begin{scope}[x={(image.north east)},y={(image.south west)}]
+      % \draw[white, fill=gray,opacity=0.5] (0.55,0.55) rectangle (1,1);
+      % \draw let \p1 = (0.55,0)
+      % in node[opacity=0.5,text width=\x1,align=center,color=white] at
+      % (0.5, 0) 
+      % {\capoverlay};
+      \draw let \p1 = (0.45,0)
+      in node[text width=\x1, align=justify] at (0.80, 1.0) 
+      {\capoverlay};
+    \end{scope}  
+  \end{tikzpicture}\label{fig:jointplot}
+\end{figure}
+
+\clearpage
 \section{Extended results on real data}
@@
 
+\clearpage
+\section*{Author Contributions}
+This report has been drafted and prepared by OE.
+Mathematical formulations in \autoref{sec:maths_extension} were originally devised and drafted
+  by DZ and rewritten for this document by OE.
+All the authors have read and critically contributed to this document.
+
 \bibliographystyle{mystyle}
@@
 \end{document}
-
-%{\color{red} {The main diverging points with respect to
-%\citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
-%$\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
-%gradients \citep{besson_dream2s_2003,herbulot_segmentation_2006};
-%2) regularization is also based on linear diffusion smoothing \citep{thirion_image_1998},
-%but we replace the Gaussian filtering by other constraints
-%studied in \citep{nagel_investigation_1986} to better the problem;
-%3) optimization is applied in the spectral
-%domain, observing anisotropic and inhomogeneous mappings along each direction.
-%With respect \citep{guyader_combined_2011}, the main differences are
-%the distance function, and the spectral solution to the optimization updates,
-%as we shall cover in \autoref{sec:numerical_implementation}.}}
-
-%In this paper we formulate the joint registration-segmentation
-%problem as follows.
-%such that the known contours in anatomical space $T$ optimally segment
-%the diffusion space $D$.
-%
-%
-%Whereas related \glspl*{adf} introduced in \autoref{sec:methods_background}
-%make use of explicit level-set formulations to solve \eqref{eq:gradient_descent},
-%we alternatively use \emph{shape-gradients}
-%\cite{besson_dream2s_2003,herbulot_segmentation_2006}.
-
-%
-%Let us denote by $\vec{x}$ the voxel and $F(\vec{x}) = [ f_1, f_2, \ldots, f_N]^T$
-%  its associated feature vector in the following.
-%
-%In our application, these surfaces are precise tissue interfaces of interest extracted
-%  from a high-resolution, anatomically correct reference volume using
-%  the well-established
-%
-%In this paper we propose a novel registration framework to simultaneously
-%solving the segmentation, distortion and cortical parcellation challenges,
-%by exploiting as strong shape-prior the detailed morphology extracted
-%from high-resolution and anatomically correct \gls{mri}.
-%Indeed, hereafter
-%we assume this segmentation problem in anatomical images is reliably and
-%accurately solved with readily available tools (e.g.
-%\citep{fischl_freesurfer_2012}).
-%After global alignment with \gls{t1} using existing approaches, the remaining
-%spatial mismatch between anatomical and diffusion space is due to susceptibility
-%distortions.
-%Finally, we need to establish precise spatial correspondence between the
-%surfaces in both spaces, including the tangential direction for parcellation.
-%Therefore, we can reduce the problem to finding the differences of spatial
-%distortion in between anatomical and \gls{dwi} space.
-%We thus reformulate the segmentation problem as an inverse problem, where we
-%seek for an underlying deformation field (the distortion) mapping
-%from the structural space into the diffusion space, such that the structural
-%contours segment optimally the \gls{dwi} data.
-%In the process, the one-to-one
-%correspondence between the contours in both spaces is guaranteed, and projection
-%of parcellisation to \gls{dwi} space is implicit and consistent.
-%
-%We test our proposed joint segmentation-registration model on two different
-%synthetic examples.
-%The first example is a scalar sulcus model, where the
-%\gls{csf}-\gls{gm} boundary particularly suffers from \gls{pve} and can only be
-%segmented correctly thanks to the shape prior and its coupling with the inner,
-%\gls{gm}-\gls{wm} boundary through the imposed deformation field regularity.
-%The second case deals with more realistic \gls{dwi} data stemming from
-%phantom simulations of a simplistic brain data.
-%Again, we show that the
-%proposed model successfully segments the \gls{dwi} data based on two derived
-%scalar features, namely \gls{fa} and \gls{md}, while establishing an estimate
-%of the dense distortion field.
-%
-%The rest of this paper is organized as follows.
-%First, in \autoref{sec:methods}
-%we introduce our proposed model for joint multivariate segmentation-registration.
-%Then we provide a more detailed description of the data and experimental setup in
-%\autoref{sec:experiments}.
-%We present results in \autoref{sec:results} and conclude
-%in \autoref{sec:conclusion}.
-%
-%
-%The properties of the reconstructed tensors and derived scalar maps have
-%been studied by \cite{ennis_orthogonal_2006}. Based on their
-%findings, \gls{fa}~\eqref{eq:fa} and \gls{md}~\eqref{eq:md} are
-%considered complementary features, and therefore we selected them for the
-%energy model \eqref{eq:complete_energy} in driving the
-%registration-segmentation process.
-%Whereas \gls{fa} informs mainly about the \emph{shape} of diffusion,
-%the \gls{md} is more related to the \emph{magnitude} of the process:
-%
-%\begin{align}
-%\mathrm{FA} &= \sqrt{ \frac{1}{2}}\,\frac{\sqrt{ (\lambda_1 - \lambda_2)^2 + (\lambda_2 - \lambda_3)^2 + (\lambda_3 - \lambda_1)^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}} \label{eq:fa} \\
-%\mathrm{MD} &= ( \lambda_1 + \lambda_2 + \lambda_3 ) / 3 \label{eq:md}
-%\end{align}
-%where $\lambda_i$ are the eigenvalues of the diffusion tensor
-%associated with the diffusion signal $S(\vec{q})$. There exist
-%two main reasons to justify their choice.
-%First, they are well-understood and standardized in clinical routine.
-%Second, together they contain most of the information that is
-%usually extracted from the \gls{dwi}-derived scalar maps
-%\cite{ennis_orthogonal_2006}.
-%

### 2015-05-04 Publications/2015-NeuroImage/05-results.tex
Fixes to the English editing

@@
   The contours recovered after registration are represented in yellow.
-  \Regseg achieved high accuracy because it determined the almost exact locations of the registered
+  \Regseg{} achieved high accuracy because it determined the almost exact locations of the registered
     contours with respect to their ground truth positions (shown in green).
@@
   we proceeded as follows.
-First, we confirmed that the vertex-wise error distributions were skewed by using Shapiro-Wilk’s test of
+First, we confirmed that the vertex-wise error distributions were skewed by using Shapiro-Wilk's test of
   normality.
@@
     \end{tabular}
-    \caption{Vertex-wise Hausdorff distances between the ground-truth surfaces and their
-    corresponding estimates obtained with \regseg{}, where the distributions are shown with the 95\% CI below
-    the half-voxel size for all of the phantom types.
+    \caption{The distributions of vertex-wise Hausdorff distances between the ground-truth surfaces and their
+    corresponding estimates obtained with \regseg{} presented a 95\% CI below the half-voxel size for all of 
+    the phantom types.
     The CIs were computed by bootstrapping using 10$^4$ samples, with the median as the location statistic.}\label{tab:ci_phantom}
@@
   method.
-Visual assessments of the 16 cases are described in \suppl{section S5}.
+Summary reports for visual assessment of the 16 cases are included in \suppl{section S5}.
 In \autoref{fig:results_real}, box A, the visual report is shown for one subject.
@@
 We also show the 95\% CIs of the \gls*{swindex} for these surfaces.
-The aggregate CI for \regseg{} was 1.08--1.50 [mm], whereas the \gls*{t2b} method
-  yielded an aggregate CI of 2.06--2.43 [mm].
+The aggregate CI for \regseg{} was 0.56--0.66 [mm], whereas the \gls*{t2b} method
+  yielded an aggregate CI of 2.05--2.39 [mm].
 The results of the statistical tests and the CIs are summarized in \autoref{tab:results_real}.

### 2015-05-04 Publications/2015-NeuroImage/06-discussion.tex
Fixes to the English editing

@@
 We carefully configured the \gls*{t2b} method using the same algorithm and the 
-  same settings employed widely in software, but cross-comparison experiments are prone to
-  so-called \emph{instrumentation bias} \citep{tustison_instrumentation_2013}.
+  same settings employed in a widely-used tool.
+However, cross-comparison experiments are prone to so-called \emph{instrumentation bias}
+  \citep{tustison_instrumentation_2013}.
 Therefore, these results do not prove that \regseg{} \emph{is better than} \gls*{t2b}.
-However, our results suggest that \regseg{} is a reliable option in this application field.
+Our results suggest that \regseg{} is a reliable option in this application field.
 In addition, the \gls*{t2b} may introduce an additional (and small) error during the necessary

### 2015-05-04 Publications/2015-NeuroImage/03-introduction.tex

--- 2015-04-28 Publications/2015-NeuroImage/03-introduction.tex
@@
 To address the segmentation problem by registration, \cite{saad_new_2009}
-  used the Pearson’s correlation coefficient to obtain a linear alignment of the \gls*{t1} and
+  used the Pearson's correlation coefficient to obtain a linear alignment of the \gls*{t1} and
   the \emph{b0}.

### 2015-05-04 Publications/2015-NeuroImage/07-conclusion.tex

@@
 
-We proposed \regseg{} for simultaneously segmenting \gls*{dmri} data and registering them according to
+We proposed \regseg{} for simultaneously segmenting and registering \gls*{dmri} data to
   their corresponding \gls*{t1} image from the same subject.
-We demonstrated 
-  the accuracy of the proposed method based on visual assessments of the results obtained by \regseg{} and cross-comparisons with a widely used technique.
+We demonstrated the accuracy of the proposed method based on visual assessments of the results
+  obtained by \regseg{} and cross-comparisons with a widely used technique.
 Moreover, \regseg{} does not require any images in addition to the minimal acquisition protocol,

### 2015-05-04 Publications/2015-NeuroImage/08-acknowledgements.tex

@@
 DZ was supported by the Swiss National Science Foundation under grants PBELP2-137727, 
-  P300P2-147778, and NSF-DMS 1418812
+  P300P2-147778, and NSF-DMS 1418812.
 This study was supported by the Spanish Ministry of Science and Innovation

### 2015-05-04 Code/Modules/Optimizers/include/SpectralOptimizer.h
Raise exception when using beta #210

@@ -187,6 +187,8 @@ public:
 	}
 
 	void SetBeta(const InternalComputationValueType v ) {
+		itkExceptionMacro( << "Error: use of beta was broken during refactoring involved in https://github.com/oesteban/RegSeg/issues/228." <<
+				" Bug https://github.com/oesteban/RegSeg/issues/210 has been already reported.	")
 		this->m_Beta.Fill(v);
 		this->Modified();
 	}

### 2015-05-04 Publications/2015-NeuroImage/00-main.tex
fix affiliation, fix reference

@@
 \author[ucla]{Dominique~Zosso}
-\author[scil,lts5]{Alessandro~Daducci}
+\author[lts5]{Alessandro~Daducci}
 \author[chuv,lts5]{Meritxell~Bach-Cuadra}
@@
 \address[bit]{Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on, Universidad Polit\'ecnica de Madrid, Madrid, Spain}
-\address[ciber]{Centro de Investigaci\'on Biom\'edica en Red en Bioingenier\'ia, Biomateriales y Nanomedicina (CIBER-BBN), Zaragoza, Spain}
+\address[ciber]{Centro de Investigaci\'on Biom\'edica en Red en Bioingenier\'ia, Biomateriales y Nanomedicina (CIBER-BBN), Spain}
 \address[ucla]{Department of Mathematics, University of California,
 Los Angeles (UCLA), Los Angeles, CA, US}
-\address[scil]{Computer Science Department, Faculty of Science, Universit\'e de Sherbrooke, 2500 Boulevard Universit\'e, Sherbrooke, QC J1K 2R1, Canada}
+% \address[scil]{Computer Science Department, Faculty of Science, Universit\'e de Sherbrooke, 2500 Boulevard Universit\'e, Sherbrooke, QC J1K 2R1, Canada}
 \address[lts5]{Signal Processing Laboratory (LTS5), \'Ecole Polytechnique

### 2015-05-04 Publications/2015-NeuroImage/04-methods.tex

@@
 \label{sec:methods}
-Let $\gammaset_R = \{\Gamma_m: m \in \mathbb{N}, m \leq N_c\}$ be the set of $N_c$ surfaces
+Let $\gammaset_R = \{\Gamma_m: m \in \mathbb{N}, m \leq N_S\}$ be the set of $N_S$ surfaces
   extracted from the undistorted \gls*{t1} image (the reference space $R$).
@@
   $P(U, \omegaset) = P(U) \implies P(M \mid U,\omegaset) = P(M \mid U)$.
-
 The best estimate $\hat{U}$ then satisfies the maximum a posteriori criterion
@@
   $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
-$C$ is the number of channels in the image $M$.
+$C$ is the number of channels comprised in the image $M$.
 In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
@@
 Finally, the maximum a posteriori problem is adapted to a variational problem where we search for
-  the minimum of an energy functional by applying $E(\vec{u}) = -\log P(U)$:
+  the minimum of an energy functional by applying $E(\vec{u}) = -\log \{P( M \mid U) \, P(U)\}$:
 
@@
   to the framework of \acrlong*{acwe} \citep{chan_active_2001}
-  with the corresponding anisotropic regularization term of \cite{nagel_investigation_1986}.
+  with the anisotropic regularization term of \cite{nagel_investigation_1986}.
 
@@
 \paragraph*{Deformation model}\label{sec:deformation_model}
-Since the vertices of the surfaces $\{\vec{v}_i: \vec{v}_i \subset \gammaset \}_{i=1 \ldots N_v}$
+Since the vertices of the surfaces $\{\vec{v}_i: \vec{v}_i \subset \gammaset \}_{i=1 \ldots N_V}$
   are probably located off-grid, it is necessary to derive $\vec{u}_i = \vec{u}(\vec{v}_i)$ from a discrete set of parameters
@@
      i.e., \isores[mm]{2.0} and \isores[mm]{1.0}, thereby producing sets of tissue fractions maps.
-  4) The tissue fractions were fed into an \acrfull*{mr} simulator, which generated \acrfull*{t1} and \acrfull*{t2} -like images at the
+  4) The tissue fractions were fed into a \acrfull*{mr} simulator, which generated \acrfull*{t1} and \acrfull*{t2} -like images at the
      two possible resolutions.
   5) The \regseg{} tool was applied using the warped test images as multispectral moving images and $\gammaset_R$ as shape priors.
-  6) The agreement between the surfaces fitted by \regseg{} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$
-     using automatically generated visual reports was assessed by computing the Hausdorff distance between the corresponding surfaces.}\label{fig:evphantoms}
+  6) The agreement between the surfaces fitted by \regseg{} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$ were assessed
+     visually using automatically generated visual reports and quantitativelly with the Hausdorff distance between the 
+     corresponding surfaces.}\label{fig:evphantoms}
 \end{figure*}
@@
 The original acquisitions are released within ``unprocessed'' packages, whereas
-  the ``minimally preprocessed'' are corrected for artifacts, brain-extracted,
-  and spatially normalized, while some results have been subjected to the standard processing
-  pipeline of \emph{FreeSurfer}.
+  the ``minimally preprocessed'' packages contain the corresponding images after
+  some processing (correction for several artifacts, brain-extraction, spatial 
+  normalization, etc.).
 We refer the reader to \citep{essen_human_2012} for exact details of the acquisition
@@
 \paragraph*{Segmentation model}
-Based on our experience \citep{esteban_brain_2012} and previous studies \citep{ennis_orthogonal_2006},
+Based on our experience % \citep{esteban_brain_2012}
+  and previous studies \citep{ennis_orthogonal_2006},
   we defined the moving image as a stack of the \gls*{fa} and \gls*{adc} maps derived
@@
   the corresponding packages of each dataset in the \gls*{hcp}.
-The field map was unwrapped\footnote{fieldmaps are phase maps, which have been clipped in the interval
+The field map was unwrapped\footnote{fieldmaps are phase maps, which are intrinsically clipped in the interval
   of $[-\pi, \pi)$ [rads] or [rads/s].} and smoothed before applying \eqref{eq:fieldmap}.
@@
   and we evaluated the data term of the cost function \eqref{eq:energy}.
-The minimum metric was consistently at the ground-truth location ($\epsilon=0.0$)
-  in all cases (\suppl{Figure S2}).
+The minimum of the cost function (\autoref{sec:methods_map}) was consistently located at 
+  $\epsilon=0.0$ (the ground-truth) for all of the cases (\suppl{Figure S2}).
 

### 2015-05-04 Publications/2015-NeuroImage/02-abstract.tex

@@
 Current applications of whole-brain tractography to \acrlong*{dmri} data require highly precise
-  delineations of anatomical structures, which are usually projected from anatomical \acrlong*{t1} images
+  delineations of anatomical structures, which are usually projected from \acrlong*{t1} images
   by registration.

### 2015-05-04 Code/Applications/binarize_surfaces.h
Update version in headers

@@ -2,7 +2,7 @@
 // File:          binarize_surfaces.h
 // Date:          Jan 12, 2015
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Applications/bspline_field.h

@@ -2,7 +2,7 @@
 // File:          bspline_field.h
 // Date:          Mar 5, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Applications/hdistance.h

@@ -2,7 +2,7 @@
 // File:          hdistance.h
 // Date:          Jan 19, 2015
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Applications/regridavg.h

@@ -2,7 +2,7 @@
 // File:          regridavg.h
 // Date:          Oct 23, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Core/include/ConfigurableObject.h

@@ -2,7 +2,7 @@
 // File:          ConfigurableObject.h
 // Date:          Feb 3, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Core/include/itkQuadEdgeMeshCopyGeometry.h

@@ -2,7 +2,7 @@
 // File:          itkQuadEdgeMeshCopyGeometry.h
 // Date:          Jan 9, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Core/include/rstkMacro.h

@@ -2,7 +2,7 @@
 // File:          rstkMacro.h
 // Date:          Jan 31, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Filtering/include/CopyCastMeshFilter.h

@@ -2,7 +2,7 @@
 // File:          CopyCastMeshFilter.h
 // Date:          Mar 6, 2015
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Filtering/include/CopyQEMeshStructureFilter.h

@@ -2,7 +2,7 @@
 // File:          CopyQEMeshStructureFilter.h
 // Date:          Jan 9, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Filtering/include/MultilabelBinarizeMeshFilter.h

@@ -2,7 +2,7 @@
 // File:          MultilabelBinarizeMeshFilter.h
 // Date:          Jan 9, 2015
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Functional/include/ModelBase.h

@@ -2,7 +2,7 @@
 // File:          ModelBase.h
 // Date:          Dec 23, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Functional/include/UniformMembershipFunction.h

@@ -2,7 +2,7 @@
 // File:          UniformMembershipFunction.h
 // Date:          Jan 10, 2015
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/IO/include/rstkVTKPolyDataWriter.h

@@ -2,7 +2,7 @@
 // File:          rstkVTKPolyDataWriter.h
 // Date:          Mar 17, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Observers/include/IterationStdOutUpdate.h

@@ -2,7 +2,7 @@
 // File:          IterationStdOutUpdate.h
 // Date:          Mar 18, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Optimizers/include/SegmentationOptimizer.h

@@ -2,7 +2,7 @@
 // File:          SegmentationOptimizer.h
 // Date:          Feb 10, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Optimizers/include/SpectralGradientDescentOptimizer.h

@@ -2,7 +2,7 @@
 // File:          SpectralGradientDescentOptimizer.h
 // Date:          Jul 31, 2013
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Optimizers/include/SpectralOptimizer.h

@@ -2,7 +2,7 @@
 // File:          SpectralOptimizer.h
 // Date:          Jul 31, 2013
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Transform/include/CachedMatrixTransform.h

@@ -2,7 +2,7 @@
 // File:          CachedMatrixTransform.h
 // Date:          Nov 25, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Transform/include/SparseMatrixTransform.h

@@ -2,7 +2,7 @@
 // File:          SparseMatrixTransform.h
 // Date:          Jun 6, 2013
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Applications/bspline_invert.h
Update version in headers

@@ -2,7 +2,7 @@
 // File:          bspline_invert.h
 // Date:          Apr 21, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Applications/slice_contours.h

@@ -2,7 +2,7 @@
 // File:          slice_contours.h
 // Date:          Nov 18, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Core/include/ACWERegistrationMethod.h

@@ -2,7 +2,7 @@
 // File:          ACWERegistrationMethod.h
 // Date:          Jan 29, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Core/include/BSplineSecondDerivativeKernelFunction.h

@@ -2,7 +2,7 @@
 // File:          BSplineSecondDerivativeKernelFunction.h
 // Date:          Feb 6, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Core/include/WarpQEMeshFilter.h

@@ -2,7 +2,7 @@
 // File:          WarpQEMeshFilter.h
 // Date:          Jan 14, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Functional/include/EnergyCalculatorFilter.h

@@ -2,7 +2,7 @@
 // File:          EnergyCalculatorFilter.h
 // Date:          Dec 22, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Functional/include/MahalanobisDistanceModel.h

@@ -2,7 +2,7 @@
 // File:          MahalanobisDistanceModel.h
 // Date:          Dec 23, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/IO/include/rstkCoefficientsWriter.h

@@ -2,7 +2,7 @@
 // File:          rstkCoefficientsWriter.h
 // Date:          Nov 27, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Observers/include/LevelObserver.h

@@ -2,7 +2,7 @@
 // File:          LevelObserver.h
 // Date:          Jan 29, 2015
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Optimizers/include/OptimizerBase.h

@@ -2,7 +2,7 @@
 // File:          OptimizerBase.h
 // Date:          Feb 10, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Optimizers/include/SpectralADMMOptimizer.h

@@ -2,7 +2,7 @@
 // File:          SpectralADMMOptimizer.h
 // Date:          Jul 31, 2013
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Transform/include/BSplineSparseMatrixTransform.h

@@ -2,7 +2,7 @@
 // File:          BSplineSparseMatrixTransform.h
 // Date:          Jan 15, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-05-04 Code/Modules/Transform/include/CompositeMatrixTransform.h

@@ -2,7 +2,7 @@
 // File:          CompositeMatrixTransform.h
 // Date:          Oct 30, 2014
 // Author:        code@oscaresteban.es (Oscar Esteban)
-// Version:       1.0 beta
+// Version:       1.5.5
 // License:       GPLv3 - 29 June 2007
 // Short Summary:
 // --------------------------------------------------------------------------------------

### 2015-09-03 Publications/2015-NeuroImage/responseLetter.tex
start with response letter

\documentclass[9pt]{memoir}

\usepackage[numbers,sort&compress]{natbib}
\usepackage{xcolor}
\usepackage[T1]{fontenc}
\usepackage{charter}

\usepackage{xcite}
\externalcitedocument{00-main}

\usepackage[hidelinks]{hyperref}
\usepackage{nameref}

\newcounter{reviewpoint}
\makeatletter
\newenvironment{reviewpoint}%
{\refstepcounter{reviewpoint}\par\medskip\vspace{3ex}\hrule\vspace{1.5ex}\par\noindent%
   {\fontseries{b}\selectfont Comment \arabic{reviewpoint}:}
   \begingroup%
   \color{black!60}
   \fontshape{it}\selectfont %

}
{\endgroup\label{com:\thereviewpoint}\par\medskip}
\def\reviewpointautorefname{Comment}
\makeatother

\newcommand{\reply}{\par\fontshape{n}\selectfont \noindent \textbf{Reply}:\ }

\setlrmarginsandblock{3cm}{2.5cm}{*}
\setulmarginsandblock{2.5cm}{2.5cm}{*}
\checkandfixthelayout

\begin{document}

\section*{Response letter to the manuscript no. NIMG-15-1151}
\medskip
Original title: \emph{Active contours-driven registration method for the structure-informed segmentation of
  diffusion MR images}.

\noindent New title: \emph{\color{red} Segmentation of homogeneous regions of diffusion MR features by mapping
  same-subject surfaces extracted in structural space}.

\bigskip
\noindent We thank the editors and the reviewers for revising our manuscript no. NIMG-15-1151 submitted to NeuroImage.
We appreciate the comments made by both reviewers, and considered them fully to improve our paper.
Please, find in the following our point-by-point answer to the concerns raised by the reviewers,
  along with the description of the corresponding changes in the manuscript.

\bigskip
\bigskip
\subsection*{Reviewer \#1:}
\begin{reviewpoint}
The paper describes an approach for the delineation of selected anatomical structures ([...]) in multi-shell diffusion MR images
  by registering a segmented T1 weighted reference image with fractional anisotropy and apparent diffusion coefficient maps derived
  from the diffusion weighted images.

The approach corresponds fully to the usual registration of an atlas to the image to be segmented, leaving me rather clueless why do
  they call the method ``simultaneous registration and segmentation''.

\end{reviewpoint}
\reply{
The first sentence of the comment is a rather accurate interpretation of our work.
However, it is important to clarify that our method maps anatomically-correct surfaces of the same subject into dMRI space,
  thereby segmenting this target space.
The method works regardless how the surfaces were extracted, or if they were extracted from T2w images or manually delineated.
In this work, the surfaces are extracted from T1w for two reasons: 1) it is the widely-accepted MR sequence for such extraction,
  and 2) the structural information used in reconstruction, tractography and network definition of connectivity extraction pipelines
  is usually defined in T1w-space as well.
We understand that the manuscript was not clear to this respect, and therefore included the following changes:
\begin{enumerate}
\item Changed the title
\item Changed abstract: {\color{red} say what here}
\item Changed introduction: {\color{red} say what here}
\end{enumerate}

Regarding the second sentence, we agree on that our method can be seen as a registration of an atlas to the
  image being segmented.
However, we need to point out some particular features of the method.
First is that surfaces in our method play the role of \emph{strong} priors, since the method does not allow
  for changes in their topology (affecting the segmentation side of the approach) and they need to be
  extremely accurate so that the mapping from structural space is correct (registration).
This might not be the case on atlas-based segmentation, which generally needs to deal with problems by the
  inappropriate topology of surfaces in the atlas, and the result of registration is relatively tolerant
  to the imperfect location of surfaces from the template.
Our method can be seen as a segmentation framework using active contours without edges in which the evolution
  of the contours is governed by an elastic registration method;
or as a nonlinear registration method which cost function is built on top of a segmentation model of active contours
  without edges.
In our opinion, the presented approach is mainly a segmentation method.
However, since the surfaces are supposed to be very accurate and defined in a different coordinate system,
  the method implicitly registers that space to the target image.
Some comments of Reviewer \#2 are very similar to this one (particularly \autoref{com:11}).

In order to clarify this issue, we have modified the manuscript as follows: {\color{red} say what here}.
}

\begin{reviewpoint}
The paper describes an approach [...] (quite some are listed as possible examples but results are only reported for the white matter,
  deep gray matter and the pial surface) in multi-shell [...].
\end{reviewpoint}
\reply{
We have noticed that the way we explained the surfaces necessary to segment real dMRI datasets
  was not appropriate.
On one hand, in this paper we propose a segmentation model to cluster the multivariate space of features
  provided by the FA and ADC map.
As correctly interpreted by reviewer \#2 mentioning it explicitly, and potentially reviewer \#1 noticed,
  the model is given by a set of nested surfaces.
However, it was probably unclear in the manuscript that those surfaces and their nesting is the cornerstone
  of the method, while the set of surfaces selected in validation are only those that will have an important
  impact in downstream processing, such as reconstruction or tractography.

In order to separate the concepts of validation and the definition of the model, we have modified the
  manuscript as follows: {\color{red} say what here}.
}

\begin{reviewpoint}
The proposed algorithm is rather standard, relying on a region-based elastic contour deformation approach.
The cost function is derived in the usual Bayesian framework, relying on a multivariate normal distribution prior for the image
  intensities and a Tikhonov regularizer enforcing a predefined normal distribution over the distortions along the imaging axes.
In this respect the novelty of the paper can be questioned, while the application context and the very promising results may still
  justify a publication in NeuroImage.
\end{reviewpoint}
\reply{
We agree with the reviewer on that the novelty of the methodological elements of our approach is very limited.
However, to our knowledge there exists no method with a paired combination of the following elements:
\begin{itemize}
\item Region based (i.e. without edges) active contours model
\item The use of surfaces as prior knowledge to initialize the segmentation model,
  with property of allowing precise registration if the surface is accurately representative
  of structures being aligned.
  Moreover, the multi-phase definition of levelsets is dual to our surface-nesting model.
\item The anisotropic regularization model of choice, since most of the levelset frameworks
  use Gaussian smoothing of the deformation field.
\item Even though the shape-gradients theory was proposed back in the early 2000's, atlas-based
  registration, and registration-segmentation methods have been generally implemented on levelsets.
\end{itemize}
To our knowledge, the application of a framework as such is new and an important contribution is done
  with the proposed segmentation model of the FA and ADC maps.
We would have appreciated if the reviewers would have provided specific references being close to our approach
  beyond those given in lines 43-57 of the original paper, so that we could point to the differences of our
  method.
In this regard, we have investigated some more potential work related to ours, modifying the manuscript as
  follows: {\color{red} say what here}.
}

\begin{reviewpoint}
However, in such a case the paper should be thoroughly revised to address the following major issues:
1. The assumption of independence between pixels is quite usual but obviously wrong.
The authors may want to comment on the consequences.
\end{reviewpoint}
\reply{
  The reviewer is right, assuming that pixels are i.i.d. is wrong but a widely accepted simplification
that performs correctly with the following details:
\begin{enumerate}
\item Pixels are not generally independent, and this issue is generally addressed imposing prior
  information over neighboring, such as Markov Random Fields.
In our framework, local dependency of pixels is implicitly addressed by the nested surfaces that impose
  the piecewise smoothness of the model.
\item Pixels are not identically distributed, and this issue posed important problems when defining
  our segmentation model.
However, we empirically found that the surfaces described in {\color{red} say line here} and their
  nesting made a good representation of the distribution of each region.
Actually, this reason may be at basis of the poor performance of most of the methods originally referenced
  in lines 19-28.
Our segmentation model ensures that pixels belonging to a certain region are ``almost'' identically distributed.
The bivariate approach using FA and ADC facilitates this, as demonstrated in Figure 3.
\end{enumerate}

Nonetheless, we have specifically introduced this discussion in our manuscript {\color{red} say where}
  and improved Figure 3 to better describe this point.
}

\begin{reviewpoint}
2. It is unclear, under which conditions enforcing the optimality condition guarantees that the region identified corresponds
  to the targeted one (it probably strongly depends on its intensity properties and those of neighbouring structures).
A corresponding analysis of the goal function and the presence of (undesired) local minima is missing.
Figure S1 is not particularly helpful in this respect.
Some information about the initialization of the distortion map should also be given.
\end{reviewpoint}
\reply{

}

\begin{reviewpoint}
3. I miss a reasonable description of the parameter settings necessary for generating the results presented.
It is not even clear if all experiments have been performed by the same parameter settings.
A reasonable sensitivity analysis is also missing.
\end{reviewpoint}
\reply{}

\begin{reviewpoint}
4. I could not understand how the ground truth can be generated for the real datasets.
As far as I can see, in these cases there is a (probably unknown) distortion between the T1 weighted and the
  diffusion weighted images.
How an emulated distortion can get around this fact?
\end{reviewpoint}
\reply{}

\begin{reviewpoint}
5. The vertex-wise error distributions are not only skewed, but quite clearly structured im many cases
  (especially for the phantom images).
What is the underlying reason?
\end{reviewpoint}
\reply{}

\bigskip
\bigskip
\subsection*{Reviewer \#2:}
\begin{reviewpoint}
This paper seems to present a method that can perform wM-GM segmentation and simultaneously estimate a deformation field to
  correct for EPI distortion.
Apparently, the idea is to use nested WM-GM surfaces, as well as other surfaces, to provide better cortical registration than
  a competing method, which is not clearly described.
The overall goal and motivation of the paper are not clearly explained, and almost all related references for EPI distortion
  correction are missing.
It is unclear if the deformation field solution is restricted to the phase encode direction.
It is possible that this is a very useful method, as the results seem to indicate, but the paper is unfortunately too unclear to
  make a full judgement about the method.
\end{reviewpoint}
\reply{}

\begin{reviewpoint}
abstract: Not immediately clear what this means: ``active contours without edges extracted from structural images''
  (Where do the edges come from? What modalities are actually input to regseg?)
\end{reviewpoint}
\reply{Active contours without edges \citep{chan_active_2001} is a framework for subpixel segmentation of images in
  which the contours evolve enclosing homogeneous regions inside.
Region-based active contours is an alternative terminology.
As previously described in \autoref{com:1} and \autoref{com:2}, we have made an effort to clearly state that the outputs
  to our method are a set of nested surfaces and a target multivariate image, that for the case of dMRI data comprehends
  FA and ADC.

However, and for the sake of clarity, we have modified the manuscript as follows:  {\color{red} say what here}.}

\begin{reviewpoint}
I note that the text in the abstract seems to have no correspondence with the graphical abstract.
The text abstract mentions nothing about diffusion MRI, FA, ADC, or phase maps, so it is unclear how these may be used in the method.
These images figure prominently in the graphical abstract, which may be a better overview than the text abstract.
\end{reviewpoint}
\reply{We agree with the reviewer, we have updated the text abstract, and proposed a completely new graphical abstract.}

\begin{reviewpoint}
Intro line 6-7. I am sure you mean to say that the resolution of dMRI is much LOWER than the images microstructure (not higher).
  larger voxels == lower resolution
\end{reviewpoint}
\reply{Absolutely, we have fixed the manuscript accordingly.}

\begin{reviewpoint}
Intro 9. Orientations is the usual term for diffusion weighting, rather than ``angles''.
\end{reviewpoint}
\reply{We agreed, and updated the manuscript accordingly.}

\begin{reviewpoint}
``These limitations prevent segmentation in the native dMRI space''  I disagree.
In fact, many methods have been used to segment fiber tracts, then to quantitatively measure them,
  in the dMRI space.
It is true the correspondence with T1 is imperfect, but segmentation of fiber tracts is possible.
\end{reviewpoint}
\reply{}

\begin{reviewpoint}
Please define what you mean here by segmentation.
It seems there is an assumption that it means specifically segmentation via registration, where the
  structures are defined using T1 images, but that is not stated anywhere.

Now on line 19 it seems the goal of the segmentation may simply be the segmentation of the WM.
So far, I have no idea what the overall goal of the approach is.
\end{reviewpoint}
\reply{}

\begin{reviewpoint}
Line 24. clustering of what?
\end{reviewpoint}
\reply{}

\begin{reviewpoint}
Now in the paragraph beginning on line 29, it seems the segmentation of WM/GM interface is of interest.
Still, I have no idea if this is the goal of the approach, or why.

Line 39. I note that there are many approaches, in addition to nonlinear T2->b0, that have been proposed
  for correction of EPI distortions in dMRI.
If that is the goal of this approach, the entire intro should be rewritten focusing on this related work.
However, I do not know yet what the goal is.

The entire Introduction should be rewritten to improve organization and clearly state the paper's goal much
  earlier (in abstract and introduction).
  Truly related work should be described in more detail, and all of the unrelated references should be removed.

Line 58. It seems the paper goal may be stated here. It is still completely unclear what type of segmentation should be achieved.
\end{reviewpoint}
\reply{}

\begin{reviewpoint}
It seems by line 69 that EPI distortion correction and segmentation by registration are to be performed jointly
  in some fashion.
There are many, many references for distortion correction that are missing from this introduction and that must
  be added.
See for example multiple papers by Carlo Pierpaoli's group, including this most recent:

Irfanoglu, M. Okan, et al. ``DR-BUDDI (Diffeomorphic Registration for Blip-Up blip-Down Diffusion Imaging) method for correcting echo planar imaging distortions.''
  NeuroImage 106 (2015): 284-299.
\end{reviewpoint}
\reply{}

\begin{reviewpoint}
The majority of the EPI distortion correction methods use T2 and B0 due to similar contrast. Why was T1 chosen in this framework?

Suddenly in Section 2 it becomes clear that there are surfaces being registered.
Perhaps this should have been obvious from the title, which mentioned active contours, but before line 72 it was not clearly
  stated that the method was posed as a surface registration problem.
\end{reviewpoint}
\reply{}

\begin{reviewpoint}
Was the warp restricted to the phase encode direction?
If so, the r in eq (1) should be restricted to this direction.
I do not see this mentioned anywhere near the equations in the Methods section.
If the solution found for the deformation field was not restricted to the phase encode direction,
  then this is not a reasonable way to perform EPI distortion correction. Please clarify.

[...]

I still see nothing in the results indicating that the estimated distortions were restricted to the phase encode direction.
If not, the added degrees of freedom might result in improved surface segmentation, but the distortion field would not correspond at
  all to the physical problem under study.
Furthermore, the dMRI gradients would have to be re-oriented independently at each voxel, and this is not currently possible in any
  software I know of.
\end{reviewpoint}
\reply{}

\begin{reviewpoint}
The experiments seem reasonable and clearly presented in general. But far too little detail is given on the B0-T2 comparison method.
\end{reviewpoint}
\reply{}

\begin{reviewpoint}
Please define the phrase ``active contours without edges''.
\end{reviewpoint}
\reply{}

\begin{reviewpoint}
Figure 3: What are the arrows in images 5? DEformations? Indications of important regions?
\end{reviewpoint}
\reply{}

\bibliographystyle{mystyle}
\bibliography{Remote}

\end{document}

### 2015-09-03 Publications/2015-NeuroImage/responseLetter.tex
add some answers

@@
 Our segmentation model ensures that pixels belonging to a certain region are ``almost'' identically distributed.
-The bivariate approach using FA and ADC facilitates this, as demonstrated in Figure 3.
+The bivariate approach using FA and ADC facilitates this, as demonstrated in Figure 4.
 \end{enumerate}
@@
 Nonetheless, we have specifically introduced this discussion in our manuscript {\color{red} say where}
-  and improved Figure 3 to better describe this point.
+  and improved Figure 4 to better describe this point.
 }
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+We have made a strong effort in providing reproducible experiments using openly available data.
+Therefore, parameter settings are publicly available in GitHub.
+Particularly, under folder \url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data}
+  there are several configuration files.
+Some parameters are also discussed in the Supplemental Materials, Section S2.
+However, we understand that a better description of the parameter settings should be included within
+  the manuscript and modified it accordingly as follows: {\color{red} say what here}.
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+As commented in line 160, we are using some ``minimally preprocessed'' images from the database of the
+  Human Connectome Project (HCP).
+Particularly, the dMRI datasets from these preprocessed images are given corrected for distortions
+  and registered in structural space.
+Therefore, there could be a residual misregistration between the T1w and the dMRI data, but it should be
+  very minimal.
+We appreciate the comment, and modified the manuscript to clarify this point {\color{red} say where here}.
+We also modified Figure 3 to reflect this aspect.
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+The displacements field supporting the deformation model is not restricted to the phase encoding direction by default.
+Therefore, equations in Methods section are presented under such conditions.
+The proof of concept presented in the first experiment with phantoms are conducted without preferential directions
+  for the deformation.
+However, the point raised about the distortions derived from susceptibility inhomogeneity in dMRI data is also true.
+Therefore, experiments with real data mimicking such distortions are performed with the deformation aligned with the phase encoding
+  axis.
+We understand that this point should be clearly stated in the manuscript and modified it as follows:
+{\color{red} say what here}.
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{We hope this point is answered in \autoref{com:10}.}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+The arrows in Figure 3, panel 5, are intended to show regions subjected to large deformations, where the contours
+  do not fit data before applying \emph{regseg}.
+They also happen to show areas of the forceps minor, one fiber bundle importantly affected
+  by the susceptibility distortions.
+We have modified the caption of Figure 3 to describe the meaning of the arrows correctly.
+}
 

### 2015-09-04 Publications/2015-NeuroImage/03-introduction.tex
rewriting intro

@@
 \section{Introduction}\label{sec:introduction}
-The accurate delineation of \gls*{wm} in \gls*{dmri} and the fusion of prior
-  anatomical information extracted from a \gls*{t1} image for the same subject
-  are crucial in a range of applications based on tractography, such as
-  the extraction of structural connectivity \citep{craddock_imaging_2013} or
-  tract-based spatial statistics \citep{smith_tractbased_2006}.
-However, the precise segmentation of \gls*{dmri} data and the coregistration of anatomical
-  images are difficult due to several limitations.
-First, \gls{dmri} images have a resolution that is much higher than that of the imaged
+\Gls*{dmri} enables the mapping of microstructure \citep{basser_microstructural_1996}
+  and connectivity \citep{craddock_imaging_2013} of the human brain \emph{in-vivo}.
+It is generally acquired using \gls*{epi} schemes, since they are very fast at
+  scanning a large sequence of images called \glspl*{dwi}, each sensitized with
+  a gradient to probe proton diffusion in a certain orientation.
+Subsequent processing involves describing local microstructure with one available
+  model, which range from the early \gls*{dti} proposed by \cite{basser_microstructural_1996}
+  to current models such as \citep{daducci_accelerated_2015}.
+The microstructural map is then used to draw the preferential orientations of diffusion
+  across regions using tractography \citep{mori_threedimensional_1999}.
+Finally, the streamlines found are filtered with a cortical parcellation
+  to build the graph representing the corresponding structural
+  network \citep{hagmann_mapping_2008}.
+Since the proposal of the earliest methodologies to solve the three problems,
+  there is a need for precisely identifying the anatomy on the \gls*{dmri} space.
+In fact, current trends in reconstruction \citep{jeurissen_multitissue_2014} and
+  tractography \citep{smith_anatomicallyconstrained_2012} are increasingly using
+  structural information to improve the microstructural mapping and fiber-tracking.
+
+The structural information required by pioneer tractography studies was a simple
+  \gls*{wm} mask to provide with a termination criteria for the streamlines.
+The first attempts to segment the \gls*{wm} in \gls*{dti} involved thresholding the
+  \gls*{fa}\footnote{the \gls*{fa} is a scalar parameter of diffusion derived from
+  the \gls*{dti}.} map.
+However, the mask and subsequent analyses are highly dependent on the threshold set
+  \citep{taoka_fractional_2009}.
+To overcome the unreliability of \gls*{fa} thresholding and broaden \gls*{wm} segmentation
+  to brain tissue segmentation, \cite{zhukov_level_2003} proposed active contours with
+  edges represented by level sets, which evolve on directionally invariant scalar maps
+  computed from \gls*{dti}.
+\cite{rousson_level_2004} successfully segmented the corpus callosum with
+  region-based level sets based on the maps of eigenvalues corresponding to the
+  \gls*{dti}.
+Many studies have addressed the definition of appropriate features for segmenting
+  regions and/or tissues in \gls*{dmri} datasets, such as the 5D representation
+  proposed by \cite{jonasson_segmentation_2005}.
+Other efforts include mixture models based on sets of directionally invariant maps
+  \citep{liu_brain_2007}, iterative clustering of the \lowb{}\footnote{The \lowb{}
+  image is one \gls*{epi} volume acquired as signal reference without sensitizing gradient.}
+  image \citep{hadjiprocopis_unbiased_2005},
+  hierarchical clustering of the \gls*{fa} and \gls*{adc} maps \citep{lu_segmentation_2008},
+  and graph-cuts on combinations of scalar maps \citep{han_experimental_2009}.
+However, the precise segmentation of \gls*{dmri} is difficult for several limitations.
+First, \gls{dmri} images have a resolution that is much lower than that of the imaged
   microstructural features \citep{basser_microstructural_1996}.
@@
   voluming of the signal sources.
-Second, \gls*{dmri} schemes probe the diffusion process within the brain at
-  many angles to obtain \glspl*{dwi}, which are completed by one or more baseline (\emph{b0})
-  volumes without directional gradients.
-The extremely low \gls*{snr} and the high dimensionality of \glspl*{dwi} prevent their
-  direct use in segmentation.
-The low contrast between \gls*{gm} and \gls*{wm} in the \emph{b0} volumes also makes them unsuitable for
-  segmentation.
-Finally, \gls*{dmri} images are acquired using \gls*{epi} to speed up the acquisition process but
-  at the cost of introducing geometrical distortion, as well as signal degradation and
-  destruction \citep{jezzard_correction_1995}.
-The artifacts generated have major impacts on the anatomy extracted
+Second, the extremely low \gls*{snr} and the high dimensionality of the \glspl*{dwi} prevent
+  their direct use in segmentation.
+Third, the low contrast between \gls*{gm} and \gls*{wm} in the \lowb{} volumes also makes
+  them unsuitable for brain tissue segmentation.
+
+An alternative route to segmentation in native \gls*{dmri} space the mapping of the
+  structural information extracted from anatomical MR images, such as \gls*{t1}, through
+  an image registration process.
+Generally, intra-subject registration of MR images of the brain involves only a linear
+  mapping to compensate for head motion between scans.
+However, \gls*{epi} images speed up the acquisition process at the cost of introducing geometrical
+  distortion, as well as signal degradation and destruction \citep{jezzard_correction_1995}.
+The so-called \emph{\gls*{epi} distortion} have major impacts on the anatomy extracted
   from \gls*{dmri}, particularly in certain fiber bundles \citep{irfanoglu_effects_2012}.
-These limitations prevent segmentation in the native \gls*{dmri} space, including
-  registration-based approaches using anatomical images as the sources of structural
-  information.
-
-Early attempts to delineate the \gls*{wm} involved thresholding the
-  \gls*{fa}\footnote{the \gls*{fa} is a scalar parameter of diffusion derived from
-  the \gls*{dmri} data.} map.
-However, the mask and subsequent analyses are highly dependent on the threshold set
-  \citep{taoka_fractional_2009}.
-To overcome the unreliability of \gls*{fa} thresholding,
-  \cite{zhukov_level_2003} proposed active contours with edges represented
-  by level sets, which evolve on directionally invariant scalar maps.
-\cite{rousson_level_2004} successfully segmented the corpus callosum with
-  region-based level sets based on the eigenvalues of tensors fitted using the
-  \gls*{dmri} data.
-Many studies have addressed the definition of appropriate features for clustering,
-  such as the 5D representation proposed by \cite{jonasson_segmentation_2005}.
-Other efforts include mixed models based on sets of directionally invariant maps
-  \citep{liu_brain_2007}, iterative \citep{hadjiprocopis_unbiased_2005} and
-  hierarchical \citep{lu_segmentation_2008} clustering,
-  graph-cuts \citep{han_experimental_2009},
-  and volume fraction modeling \citep{kumazawa_improvement_2013}.
-
-To address the segmentation problem by registration, \cite{saad_new_2009}
-  used the Pearson's correlation coefficient to obtain a linear alignment of the \gls*{t1} and
-  the \emph{b0}.
+To overcome this problem, two broad families of solutions are available.
+First, it is possible to use an existing ``\gls*{epi} distortion correction''
+  method to compensate for the nonlinear part of the misalignment with the structural space, combined
+  with a rigid registration method.
+Second, it is possible to apply a nonlinear registration method with specific constraints
+  related to the theoretical properties of the \gls*{epi} distortion.
+However, both solutions require extra acquisitions included in the \gls*{dmri} scanning protocol,
+  such as fieldmaps \citep{jezzard_correction_1995}, \glspl*{dwi} with a different \gls*{pe}
+  scheme \citep{cordes_geometric_2000,chiou_simple_2000}, or \gls*{t2} images.
+\cite{saad_new_2009} used the Pearson's correlation coefficient to obtain a linear alignment of
+  the \gls*{t1} and the \lowb{}.
 Similarly, a linear registration method was employed by \emph{bbregister} \citep{greve_accurate_2009},
-  which uses active contours with edges to search for intensity boundaries in the \emph{b0}
+  which uses active contours with edges to search for intensity boundaries in the \lowb{}
   image.
-The active contours are initialized using surfaces extracted from the
-  \gls*{t1} using \emph{FreeSurfer} \citep{fischl_freesurfer_2012}:
-  the \emph{pial} surface (exterior of the cortical \gls*{gm}) and the \emph{white}
-  surface (the \gls*{wm}/\gls*{gm} interface).
-The \emph{b0} image only includes a detectable frontier for the pial surface, so
+The active contours are initialized using surfaces extracted from the \gls*{t1} using
+  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}: the \emph{pial} surface (exterior of the
+  cortical \gls*{gm}) and the \emph{white} surface (the \gls*{wm}/\gls*{gm} interface).
+The \lowb{} image only includes a detectable frontier for the pial surface, so
   \emph{bbregister} is limited to aligning the cortical layer in this
@@
 However, because the distortion is not considered, it must be addressed separately.
-Nonlinear registration has been performed successfully between \gls*{t2} and \emph{b0}
+Nonlinear registration has been performed successfully between \gls*{t2} and \lowb{}
   images based on their similarity as a unique method for correcting distortion
@@
 However, further registration of the \gls*{t1} and \gls*{t2} images is still required to map the anatomical
-  information (and for \gls*{wm} segmentation) into the \gls*{dmri} space.
+  information into the \gls*{dmri} space.
 

### 2015-09-04 Publications/2015-NeuroImage/03-introduction.tex
rewriting intro (second part)

@@
   a gradient to probe proton diffusion in a certain orientation.
-Subsequent processing involves describing local microstructure with one available
-  model, which range from the early \gls*{dti} proposed by \cite{basser_microstructural_1996}
+Subsequent processing involves describing the local microstructure with one available
+  model, which ranges from the early \gls*{dti} proposed by \cite{basser_microstructural_1996}
   to current models such as \citep{daducci_accelerated_2015}.
@@
   \citep{taoka_fractional_2009}.
-To overcome the unreliability of \gls*{fa} thresholding and broaden \gls*{wm} segmentation
-  to brain tissue segmentation, \cite{zhukov_level_2003} proposed active contours with
-  edges represented by level sets, which evolve on directionally invariant scalar maps
-  computed from \gls*{dti}.
-\cite{rousson_level_2004} successfully segmented the corpus callosum with
-  region-based level sets based on the maps of eigenvalues corresponding to the
-  \gls*{dti}.
-Many studies have addressed the definition of appropriate features for segmenting
-  regions and/or tissues in \gls*{dmri} datasets, such as the 5D representation
-  proposed by \cite{jonasson_segmentation_2005}.
-Other efforts include mixture models based on sets of directionally invariant maps
-  \citep{liu_brain_2007}, iterative clustering of the \lowb{}\footnote{The \lowb{}
-  image is one \gls*{epi} volume acquired as signal reference without sensitizing gradient.}
-  image \citep{hadjiprocopis_unbiased_2005},
-  hierarchical clustering of the \gls*{fa} and \gls*{adc} maps \citep{lu_segmentation_2008},
-  and graph-cuts on combinations of scalar maps \citep{han_experimental_2009}.
+To overcome the unreliability of \gls*{fa} thresholding, and to broaden
+  \gls*{wm} segmentation to brain tissue segmentation, a large number of
+  methods have been proposed using \glspl*{dwi}, the \lowb{}, and \gls*{dti}-derived
+  scalar maps such as \gls*{fa}, \gls*{adc} and others \citep{zhukov_level_2003,
+  rousson_level_2004,jonasson_segmentation_2005,liu_brain_2007,hadjiprocopis_unbiased_2005,
+  lu_segmentation_2008,han_experimental_2009}.
 However, the precise segmentation of \gls*{dmri} is difficult for several limitations.
 First, \gls{dmri} images have a resolution that is much lower than that of the imaged
-  microstructural features \citep{basser_microstructural_1996}.
+  microstructural features.
 Therefore, voxels located in structural discontinuities are affected by partial
@@
 
-An alternative route to segmentation in native \gls*{dmri} space the mapping of the
+An alternative route to segmentation in native \gls*{dmri} space is the mapping of the
   structural information extracted from anatomical MR images, such as \gls*{t1}, through
@@
   distortion, as well as signal degradation and destruction \citep{jezzard_correction_1995}.
-The so-called \emph{\gls*{epi} distortion} have major impacts on the anatomy extracted
+The so-called \emph{\gls*{epi} distortions} have a major impact on the anatomy extracted
   from \gls*{dmri}, particularly in certain fiber bundles \citep{irfanoglu_effects_2012}.
@@
 First, it is possible to use an existing ``\gls*{epi} distortion correction''
-  method to compensate for the nonlinear part of the misalignment with the structural space, combined
-  with a rigid registration method.
-Second, it is possible to apply a nonlinear registration method with specific constraints
-  related to the theoretical properties of the \gls*{epi} distortion.
-However, both solutions require extra acquisitions included in the \gls*{dmri} scanning protocol,
-  such as fieldmaps \citep{jezzard_correction_1995}, \glspl*{dwi} with a different \gls*{pe}
-  scheme \citep{cordes_geometric_2000,chiou_simple_2000}, or \gls*{t2} images.
-\cite{saad_new_2009} used the Pearson's correlation coefficient to obtain a linear alignment of
-  the \gls*{t1} and the \lowb{}.
-Similarly, a linear registration method was employed by \emph{bbregister} \citep{greve_accurate_2009},
-  which uses active contours with edges to search for intensity boundaries in the \lowb{}
-  image.
+  method to compensate for the nonlinear part of the misalignment with the structural space,
+  combined with a rigid registration method.
+Second, to apply a nonlinear registration method with specific constraints derived from the
+  theoretical properties of the \gls*{epi} distortion.
+However, both solutions require extra acquisitions to be included within the \gls*{dmri} scanning
+  protocol, such as fieldmaps \citep{jezzard_correction_1995}, \glspl*{dwi} with a different
+  \gls*{pe} scheme \citep{cordes_geometric_2000,chiou_simple_2000}, \gls*{t2} images
+  \citep{kybic_unwarping_2000}.
+We will refer to this approach as \gls*{t2b} correction.
+\Gls*{epi} corrections have been an active field of research, with numerous refinements and
+  combinations of the original methods \citep{holland_efficient_2010,andersson_comprehensive_2012,
+  irfanoglu_drbuddi_2014}.
+
+The tool \emph{bbregister} \citep{greve_accurate_2009} has become a rather standard linear
+  registration method to be used with \gls*{epi} corrections.
+It uses 3D active contours with edges\footnote{A three-dimensional active contour is a surface
+  and it is ``with edges'' when the contours evolve in search of abrupt intensity steps (the edges) within
+  the target image.} to search for intensity boundaries in the \lowb{} image.
 The active contours are initialized using surfaces extracted from the \gls*{t1} using
@@
   cortical \gls*{gm}) and the \emph{white} surface (the \gls*{wm}/\gls*{gm} interface).
-The \lowb{} image only includes a detectable frontier for the pial surface, so
+The \lowb{} image only includes a detectable frontier for the pial surface, and thus
   \emph{bbregister} is limited to aligning the cortical layer in this
   application.
-This tool has become the standard method because of its proven robustness, although the
-  distortions found in \gls*{dmri} are nonlinear.
-To overcome this issue, \emph{bbregister} excludes the
+To overcome the problem of nonlinear distortions, \emph{bbregister} excludes the
   regions that are typically warped by artifacts from the boundary search.
 However, because the distortion is not considered, it must be addressed separately.
-Nonlinear registration has been performed successfully between \gls*{t2} and \lowb{}
-  images based on their similarity as a unique method for correcting distortion
-  \citep{kybic_unwarping_2000,studholme_accurate_2000,wu_comparison_2008,tao_variational_2009}.
+The nonlinear registration approach has been successfully solved with \gls*{t2} and \lowb{}
+  images, exploiting their similarity \citep{kybic_unwarping_2000,studholme_accurate_2000}.
 However, further registration of the \gls*{t1} and \gls*{t2} images is still required to map the anatomical
@@
 
-Therefore, given these issues, a method that simultaneously performs
-  segmentation in the native \gls*{dmri} space and registration of the corresponding \gls*{t1} image
-  onto the \gls*{dmri} data could provide the optimal solution.
-To the best of our knowledge, this strategy has never been proposed for the application described above.
-Previously, joint segmentation and registration have been applied successfully to other problems
-such as longitudinal object tracking \citep{paragios_level_2003} and atlas-based
-  segmentation \citep{gorthi_active_2011}.
-The most thorough approach integrates active contours during image registration
-  methods.
-\cite{unal_coupled_2005}, and later \cite{wang_joint_2006},
-  improved an existing method \citep{yezzi_variational_2003} based on linear registration
-  to the nonlinear case by implementing a free-form deformation field.
-\cite{droske_mumfordshah_2009} reviewed the existing techniques and proposed two different
-  approaches for applying the Mumford-Shah functional \citep{mumford_optimal_1989} during simultaneous
-  registration and segmentation by propagating the deformation field from
-  the contours onto the whole image definition.
-Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
-  registration method in 2D using level sets and a nonlinear elasticity smoother on the
-  displacement vector field, which preserves the topology even with very high deformation.
-Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
-  level set function to register several active contours during the application
-  of atlas-based segmentation.
-
-The hypothesis tested in the present study is that registration and segmentation
-  problems in \gls*{dmri} can be solved simultaneously, thereby increasing the geometrical
-  accuracy of the process.
-Thus, we propose a tool called \regseg{} that exploits the prior information of shapes
-  extracted using \gls*{t1} images to register the anatomical reference
-  to \gls*{dmri} space, thereby implicitly segmenting these data.
-The proposed approach uses active contours without edges \citep{chan_active_2001}, which evolve to drive a
-  free-form deformation field of B-spline basis functions.
+Here we show a method that simultaneously registers anatomically correct surfaces representing
+  the structure of the subject's brain into the bivariate image comprehending \gls*{fa} and
+  \gls*{adc}, thereby implicitly segmenting the \gls*{dmri} space.
+The hypothesis tested in the present study is that the registration and segmentation
+  problems in \gls*{dmri} can be solved simultaneously, effectively establishing the
+  necessary mapping to project structural information from anatomical images, increasing the
+  geometrical accuracy of the overall process, and avoiding data resampling of \gls*{epi}
+  correction methods.
+Thus, we propose a tool called \regseg{} that 3D active contours (surfaces) without edges
+  \citep{chan_active_2001}, which evolve to drive a free-form deformation field of B-spline
+  basis functions.
+The surfaces are nested and they enclose regions with homogeneous diffusion features
+  (the \gls*{fa} and \gls*{adc} maps for the case of \gls*{dmri} data).
 Optimization is performed using a descent strategy with explicit shape gradients
@@
 Therefore, unlike most previous methods, \regseg{} does not implement level sets.
-The nonlinear distortion is aligned with one of the imaging axes (see
-  \autoref{sec:human_connectome}), so \regseg{} includes an anisotropic regularizer for
+Since the \gls*{epi} distortion is aligned with the \gls*{pe} axis (see
+  \autoref{sec:human_connectome}), \regseg{} includes an anisotropic regularizer for
   the displacements field proposed by \cite{nagel_investigation_1986}.
-Finally, we evaluated \regseg{} using an extension of our instrumentation framework
+Finally, we evaluate \regseg{} using an extension of our instrumentation framework
   \citep{esteban_simulationbased_2014}, which simulates known and realistic distortions
   based on \gls*{dmri} data.
-The evaluation included a comparison with \gls*{t2b} correction by integrating it within the framework
-  of an in-house implementation of the method.
+The evaluation includes a comparison of \regseg{} to an in-house implementation of
+  the \gls*{t2b} correction.

### 2015-09-04 Publications/2015-NeuroImage/04-methods.tex
rewriting intro (second part)

@@
 \label{sec:methods}
+\subsection{Related work}\label{sec:related_work}
+Previously, joint segmentation and registration have been applied successfully to other problems
+such as longitudinal object tracking \citep{paragios_level_2003} and atlas-based
+  segmentation \citep{gorthi_active_2011}.
+The most thorough approach integrates active contours during image registration
+  methods.
+\cite{unal_coupled_2005}, and later \cite{wang_joint_2006},
+  improved an existing method \citep{yezzi_variational_2003} based on linear registration
+  to the nonlinear case by implementing a free-form deformation field.
+\cite{droske_mumfordshah_2009} reviewed the existing techniques and proposed two different
+  approaches for applying the Mumford-Shah functional \citep{mumford_optimal_1989} during simultaneous
+  registration and segmentation by propagating the deformation field from
+  the contours onto the whole image definition.
+Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
+  registration method in 2D using level sets and a nonlinear elasticity smoother on the
+  displacement vector field, which preserves the topology even with very high deformation.
+Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
+  level set function to register several active contours during the application
+  of atlas-based segmentation.
+
+\subsection{Cost-function derivation}\label{sec:methods_map}
 Let $\gammaset_R = \{\Gamma_m: m \in \mathbb{N}, m \leq N_S\}$ be the set of $N_S$ surfaces
@@
 
-\subsection{Cost-function derivation}\label{sec:methods_map}
 In a Bayesian framework for registration \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014},

### 2015-09-07 Publications/2015-NeuroImage/03-introduction.tex

@@
   \citep{kybic_unwarping_2000}.
-We will refer to this approach as \gls*{t2b} correction.
+We will refer to this latter approach as \gls*{t2b} correction in the following.
 \Gls*{epi} corrections have been an active field of research, with numerous refinements and
@@
 To overcome the problem of nonlinear distortions, \emph{bbregister} excludes the
-  regions that are typically warped by artifacts from the boundary search.
+  regions that are typically warped from the boundary search.
 However, because the distortion is not considered, it must be addressed separately.
-The nonlinear registration approach has been successfully solved with \gls*{t2} and \lowb{}
-  images, exploiting their similarity \citep{kybic_unwarping_2000,studholme_accurate_2000}.
-However, further registration of the \gls*{t1} and \gls*{t2} images is still required to map the anatomical
-  information into the \gls*{dmri} space.
+The nonlinear registration approach has been successfully solved with \gls*{t2b} methods,
+  which align the \gls*{t2} and \lowb{} images by exploiting their similarity
+  \citep{kybic_unwarping_2000,studholme_accurate_2000}.
+However, further rigid registration of the \gls*{t1} and \gls*{t2} images is still required to
+  map the anatomical information onto \gls*{dmri}-space.
 
-Here we show a method that simultaneously registers anatomically correct surfaces representing
-  the structure of the subject's brain into the bivariate image comprehending \gls*{fa} and
-  \gls*{adc}, thereby implicitly segmenting the \gls*{dmri} space.
+Here we show a method that registers anatomically-correct surfaces to multivariate images,
+  thereby implicitly segmenting the \gls*{dmri}-space into those structures enclosed by the contours
+  of the subject's brain.
+% It is important that surfaces are defined in the coordinate system of the image used as
+%   source of structural information (typically, but not restricted to, \gls*{t1} images), otherwise segmentation performs
+%   correctly but the mapping found in registration does not represent any valuable information.
 The hypothesis tested in the present study is that the registration and segmentation
@@
   correction methods.
-Thus, we propose a tool called \regseg{} that 3D active contours (surfaces) without edges
+Thus, we propose a tool called \regseg{} implementing 3D active contours (surfaces) without edges
   \citep{chan_active_2001}, which evolve to drive a free-form deformation field of B-spline

### 2015-09-07 Publications/2015-NeuroImage/03-introduction.tex
add expanded and more fixes

@@
   protocol, such as fieldmaps \citep{jezzard_correction_1995}, \glspl*{dwi} with a different
-  \gls*{pe} scheme \citep{cordes_geometric_2000,chiou_simple_2000}, \gls*{t2} images
+  \gls*{pe} scheme \citep{cordes_geometric_2000,chiou_simple_2000}, or \gls*{t2} images
   \citep{kybic_unwarping_2000}.
-We will refer to this latter approach as \gls*{t2b} correction in the following.
+In the following, we will refer to this latter approach as \gls*{t2b} correction.
 \Gls*{epi} corrections have been an active field of research, with numerous refinements and
@@
 The tool \emph{bbregister} \citep{greve_accurate_2009} has become a rather standard linear
-  registration method to be used with \gls*{epi} corrections.
+  registration method to be used along with \gls*{epi} corrections.
 It uses 3D active contours with edges\footnote{A three-dimensional active contour is a surface
@@
 Thus, we propose a tool called \regseg{} implementing 3D active contours (surfaces) without edges
-  \citep{chan_active_2001}, which evolve to drive a free-form deformation field of B-spline
+  \citep{chan_active_2001}, which evolve over a free-form deformation field of B-spline
   basis functions.

### 2015-09-07 Publications/2015-NeuroImage/04-methods.tex
add expanded and more fixes

@@
   segmentation \citep{gorthi_active_2011}.
-The most thorough approach integrates active contours during image registration
-  methods.
-\cite{unal_coupled_2005}, and later \cite{wang_joint_2006},
-  improved an existing method \citep{yezzi_variational_2003} based on linear registration
-  to the nonlinear case by implementing a free-form deformation field.
+The most common approach to solve the problem simultaneously is optimizing a deformation model
+  (registration) that supports the evolution the active contours (segmentation), like
+  \citep{paragios_level_2003,yezzi_variational_2003}.
+\cite{unal_coupled_2005}, and later \cite{wang_joint_2006}, replaced the linear registration
+  transform of \citep{yezzi_variational_2003} with a free-form deformation field.
 \cite{droske_mumfordshah_2009} reviewed the existing techniques and proposed two different
@@
   registration method in 2D using level sets and a nonlinear elasticity smoother on the
-  displacement vector field, which preserves the topology even with very high deformation.
+  displacement vector field, which preserves the topology even with very large deformations.
 Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
@@
 
-\subsection{Cost-function derivation}\label{sec:methods_map}
+\subsection{Registration framework and segmentation model derivation}\label{sec:methods_map}
 Let $\gammaset_R = \{\Gamma_m: m \in \mathbb{N}, m \leq N_S\}$ be the set of $N_S$ surfaces
@@
 
+\paragraph*{Cost-function derivation}
 In a Bayesian framework for registration \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014},
@@
   distributions with the parameters $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
-  for each region $\Omega_l$:
+  for each region $\Omega_l$.
+Even though the features being segmented are not generally i.i.d., the spatial interdependency of
+  voxels is implicitly supported by the piecewise smooth partition of the space \omegaset{}.
+In \autoref{fig:model} it is shown that marginal distributions of data can be approximated by
+  multivariate normal distributions. Therefore,
 
@@
 In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
-  the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized.
+  the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized. In \autoref{fig:model},
+  this minimization is illustrated for the segmentation of the \gls*{fa} and the \gls*{adc} maps
+  of one subject.
 
+\begin{figure}
+% \includegraphics[width=\linewidth]{figure04A}
+% \includegraphics[width=\linewidth]{figure04B}
+% \includegraphics[width=\linewidth]{figure04C}
+\caption{%
+Evolution of the segmentation model defined by the homogeneous regions $\Omega_i$
+  for a real dataset.
+Panel A shows how the \gls*{fa} and the \gls*{adc} are clustered by the surfaces mapped to
+  $M$ through the ground-truth.
+Panel B shows the same information at the initialization step of \regseg{}.
+Since the surfaces are misaligned in $M$, the spreads of the sampled distribution for
+  each region are large.
+Panel C, after \regseg{} registration, shows how the covariance matrices $\boldsymbol{\Sigma}_l$ 
+  are minimized with respect to panel B, and the resulting segmentation is close
+  to the theoretical situation presented in panel A.
+The left (large) plot in each panel represents a kernel density estimate of the location and spread of
+  each region $\Omega_i$ in the bivariate feature space of $M$, which comprises
+  the \gls*{fa} and \gls*{adc} maps.
+At the top and the right margins, the marginal distributions of $\Omega_i$ are also plotted for
+  the \gls*{fa} and \gls*{adc}, respectively.
+On the right, the sampled distribution of each region is shown in small scatter plot, in contrast to
+  the complete bivariate data.
+}\label{fig:model}
+\end{figure}
+
+\paragraph*{Regularization}
 The smoothness of the resulting displacements field is induced by a Thikonov regularization
@@
 1) Extract the set of undistorted surfaces $\gammaset_R$;
-2) Compute a ground-truth field of displacements $U_{true}$, which is applied to
+2) Compute a ground-truth field of displacements $U_\text{true}$, which is applied to
   generate warped images ($M$) for segmentation;
@@
 4) Perform a visual assessment and compute the error metrics.
+
 The adaptation of this protocol to the simulated phantoms and real data is explained in the
   following sections.
+A first proof of concept to demonstrate \regseg{} in phantoms with simple
+  geometries is introduced, using $U_\text{true}$ without directional restrictions.
+Then, a more realistic environment with $U_\text{true}$ restricted to the
+  \gls*{pe} axis of real \gls*{dmri} datasets is presented.
 
@@
 Next, \regseg{} was applied to map $\gammaset_R$ onto the warped phantoms to
-  obtain the registered surfaces ($\hat{\gammaset}_{test}$).
+  obtain the registered surfaces ($\hat{\gammaset}_\text{test}$).
 To quantify the misregistration error, we computed the Hausdorff distance between
- $\hat{\gammaset}_{test}$ and $\gammaset_\text{true}$ using \citep{commandeur_vtk_2011}.
+ $\hat{\gammaset}_\text{test}$ and $\gammaset_\text{true}$ using \citep{commandeur_vtk_2011}.
 
@@
   5) The \regseg{} tool was applied using the warped test images as multispectral moving images and $\gammaset_R$ as shape priors.
-  6) The agreement between the surfaces fitted by \regseg{} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$ were assessed
+  6) The agreement between the surfaces fitted by \regseg{} ($\hat{\gammaset}_\text{test}$) and $\gammaset_\text{true}$ were assessed
      visually using automatically generated visual reports and quantitativelly with the Hausdorff distance between the
@@
   1) $\gammaset_R$ were extracted from the anatomical reference (\gls*{t1} image).
-  2) For use as the ground truth, we generated a plausible synthetic distortion $U_{true}$
+  2) For use as the ground truth, we generated a plausible synthetic distortion $U_\text{true}$
     from the field map with \eqref{eq:fieldmap}.
-  3) The \gls*{dmri} data were warped using $U_{true}$ to reproduce the effects of real
+  3) The \gls*{dmri} data were warped using $U_\text{true}$ to reproduce the effects of real
     susceptibility-derived distortions.
@@
   multi-shell \gls*{dmri} images.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{figure04}
-\caption{Segmentation model defined by the homogeneous regions $\Omega_i$, where
-  the image for segmentation is partitioned by the theoretical projection
-  ($\gammaset_{true}$) of $\gammaset_R$ onto $M$.
-The plot represents a kernel density estimate of the location and spread of
-  each region $\Omega_i$ in the bivariate feature space of $M$, which comprises
-  the \gls*{fa} and \gls*{adc} maps.
-At the top and the right margins, the marginal distributions of $\Omega_i$ are also plotted for
-  the \gls*{fa} and \gls*{adc}, respectively.
-}\label{fig:model}
-\end{figure}
 
@@
   the so-called field map.
-We derived the deformation $U_{true}$ from the field map image released with
+We derived the deformation $U_\text{true}$ from the field map image released with
   the corresponding packages of each dataset in the \gls*{hcp}.
@@
 For five test datasets, we uniformly sampled the space of distortions
-  $\hat{U} = \epsilon \cdot U_{true} = \vec{r} + \epsilon \, u_\text{PE}$
+  $\hat{U} = \epsilon \cdot U_\text{true} = \vec{r} + \epsilon \, u_\text{PE}$
   (with $\epsilon \in [-1.1, 1.1]$ and $u_\text{PE}$ from \eqref{eq:fieldmap}),
@@
   \gls*{swindex} as the area-weighted distance between the corresponding vertices of
-  $\gammaset_{true}$ and their estimate obtained by the method under the test $\hat{\gammaset}_{test}$:
+  $\gammaset_{true}$ and their estimate obtained by the method under the test $\hat{\gammaset}_\text{test}$:
 
@@
   where $\vec{v}_i \subset \gammaset_\text{true}$ are the locations of the total $N_V$ vertices
-  and $\hat{\vec{v}}_i \subset \hat{\gammaset}_{test}$ are the recovered locations
+  and $\hat{\vec{v}}_i \subset \hat{\gammaset}_\text{test}$ are the recovered locations
   that correspond to $\vec{v}_i$.

### 2015-09-07 Publications/2015-NeuroImage/expanded.tex
add expanded and more fixes

--- 2015-04-29 Publications/2015-NeuroImage/ELS285_article_edited.tex
@@
 \documentclass[3p,authoryear,fleqn]{elsarticle}
-\journal{ArXiv archiving for pre-publication review}
+\journal{NeuroImage}
 
@@
 \providecommand{\Regseg}{\emph{Regseg}}
+\providecommand{\lowb}{\textit{b0}}
 
@@
 \providecommand{\nvec}[1]{\hat{\mathbf{#1}}}
-\providecommand{\suppl}[1]{\href{http://dx.doi.org/mydoi}{Supplemental Materials}, #1}
+\providecommand{\suppl}[1]{\href{http://figshare.com/s/459c26b4ee8211e493b306ec4bbcf141}{Supplemental Materials}, #1}
 \providecommand{\isores}[2][]{#2\ensuremath{\times}#2\ensuremath{\times}#2\ifthenelse{\equal{#1}{}}{}{ [#1]}\xspace}
@@
 \address[bit]{Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on, Universidad Polit\'ecnica de Madrid, Madrid, Spain}
-\address[ciber]{Centro de Investigaci\'on Biom\'edica en Red en Bioingenier\'ia, Biomateriales y Nanomedicina (CIBER-BBN), Zaragoza, Spain}
+\address[ciber]{Centro de Investigaci\'on Biom\'edica en Red en Bioingenier\'ia, Biomateriales y Nanomedicina (CIBER-BBN), Spain}
 \address[ucla]{Department of Mathematics, University of California,
 Los Angeles (UCLA), Los Angeles, CA, US}
-\address[scil]{Computer Science Department, Faculty of Science, Universit\'e de Sherbrooke, 2500 Boulevard Universit\'e, Sherbrooke, QC J1K 2R1, Canada}
 \address[lts5]{Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
@@
 \makeatletter{}Current applications of whole-brain tractography to \acrlong*{dmri} data require highly precise
-  delineations of anatomical structures, which are usually projected from anatomical \acrlong*{t1} images
+  delineations of anatomical structures, which are usually projected from \acrlong*{t1} images
   by registration.
@@
   that \regseg{} outperformed this method in our settings. 
-
 \end{abstract}
@@
 \makeatletter{}\section{Introduction}\label{sec:introduction}
-The accurate delineation of \gls*{wm} in \gls*{dmri} and the fusion of prior
-  anatomical information extracted from a \gls*{t1} image for the same subject
-  are crucial in a range of applications based on tractography, such as
-  the extraction of structural connectivity \citep{craddock_imaging_2013} or
-  tract-based spatial statistics \citep{smith_tractbased_2006}.
-However, the precise segmentation of \gls*{dmri} data and the coregistration of anatomical
-  images are difficult due to several limitations.
-First, \gls{dmri} images have a resolution that is much higher than that of the imaged
-  microstructural features \citep{basser_microstructural_1996}.
+\Gls*{dmri} enables the mapping of microstructure \citep{basser_microstructural_1996}
+  and connectivity \citep{craddock_imaging_2013} of the human brain \emph{in-vivo}.
+It is generally acquired using \gls*{epi} schemes, since they are very fast at
+  scanning a large sequence of images called \glspl*{dwi}, each sensitized with
+  a gradient to probe proton diffusion in a certain orientation.
+Subsequent processing involves describing the local microstructure with one available
+  model, which ranges from the early \gls*{dti} proposed by \cite{basser_microstructural_1996}
+  to current models such as \citep{daducci_accelerated_2015}.
+The microstructural map is then used to draw the preferential orientations of diffusion
+  across regions using tractography \citep{mori_threedimensional_1999}.
+Finally, the streamlines found are filtered with a cortical parcellation
+  to build the graph representing the corresponding structural
+  network \citep{hagmann_mapping_2008}.
+Since the proposal of the earliest methodologies to solve the three problems,
+  there is a need for precisely identifying the anatomy on the \gls*{dmri} space.
+In fact, current trends in reconstruction \citep{jeurissen_multitissue_2014} and
+  tractography \citep{smith_anatomicallyconstrained_2012} are increasingly using
+  structural information to improve the microstructural mapping and fiber-tracking.
+
+The structural information required by pioneer tractography studies was a simple
+  \gls*{wm} mask to provide with a termination criteria for the streamlines.
+The first attempts to segment the \gls*{wm} in \gls*{dti} involved thresholding the
+  \gls*{fa}\footnote{the \gls*{fa} is a scalar parameter of diffusion derived from
+  the \gls*{dti}.} map.
+However, the mask and subsequent analyses are highly dependent on the threshold set
+  \citep{taoka_fractional_2009}.
+To overcome the unreliability of \gls*{fa} thresholding, and to broaden
+  \gls*{wm} segmentation to brain tissue segmentation, a large number of
+  methods have been proposed using \glspl*{dwi}, the \lowb{}, and \gls*{dti}-derived
+  scalar maps such as \gls*{fa}, \gls*{adc} and others \citep{zhukov_level_2003,
+  rousson_level_2004,jonasson_segmentation_2005,liu_brain_2007,hadjiprocopis_unbiased_2005,
+  lu_segmentation_2008,han_experimental_2009}.
+However, the precise segmentation of \gls*{dmri} is difficult for several limitations.
+First, \gls{dmri} images have a resolution that is much lower than that of the imaged
+  microstructural features.
 Therefore, voxels located in structural discontinuities are affected by partial
   voluming of the signal sources.
-Second, \gls*{dmri} schemes probe the diffusion process within the brain at 
-  many angles to obtain \glspl*{dwi}, which are completed by one or more baseline (\emph{b0}) 
-  volumes without directional gradients.
-The extremely low \gls*{snr} and the high dimensionality of \glspl*{dwi} prevent their
-  direct use in segmentation.
-The low contrast between \gls*{gm} and \gls*{wm} in the \emph{b0} volumes also makes them unsuitable for 
-  segmentation.
-Finally, \gls*{dmri} images are acquired using \gls*{epi} to speed up the acquisition process but
-  at the cost of introducing geometrical distortion, as well as signal degradation and
-  destruction \citep{jezzard_correction_1995}.
-The artifacts generated have major impacts on the anatomy extracted
+Second, the extremely low \gls*{snr} and the high dimensionality of the \glspl*{dwi} prevent
+  their direct use in segmentation.
+Third, the low contrast between \gls*{gm} and \gls*{wm} in the \lowb{} volumes also makes
+  them unsuitable for brain tissue segmentation.
+
+An alternative route to segmentation in native \gls*{dmri} space is the mapping of the
+  structural information extracted from anatomical MR images, such as \gls*{t1}, through
+  an image registration process.
+Generally, intra-subject registration of MR images of the brain involves only a linear
+  mapping to compensate for head motion between scans.
+However, \gls*{epi} images speed up the acquisition process at the cost of introducing geometrical
+  distortion, as well as signal degradation and destruction \citep{jezzard_correction_1995}.
+The so-called \emph{\gls*{epi} distortions} have a major impact on the anatomy extracted
   from \gls*{dmri}, particularly in certain fiber bundles \citep{irfanoglu_effects_2012}.
-These limitations prevent segmentation in the native \gls*{dmri} space, including
-  registration-based approaches using anatomical images as the sources of structural
-  information.
+To overcome this problem, two broad families of solutions are available.
+First, it is possible to use an existing ``\gls*{epi} distortion correction''
+  method to compensate for the nonlinear part of the misalignment with the structural space,
+  combined with a rigid registration method.
+Second, to apply a nonlinear registration method with specific constraints derived from the
+  theoretical properties of the \gls*{epi} distortion.
+However, both solutions require extra acquisitions to be included within the \gls*{dmri} scanning
+  protocol, such as fieldmaps \citep{jezzard_correction_1995}, \glspl*{dwi} with a different
+  \gls*{pe} scheme \citep{cordes_geometric_2000,chiou_simple_2000}, or \gls*{t2} images
+  \citep{kybic_unwarping_2000}.
+In the following, we will refer to this latter approach as \gls*{t2b} correction.
+\Gls*{epi} corrections have been an active field of research, with numerous refinements and
+  combinations of the original methods \citep{holland_efficient_2010,andersson_comprehensive_2012,
+  irfanoglu_drbuddi_2014}.
 
-Early attempts to delineate the \gls*{wm} involved thresholding the 
-  \gls*{fa}\footnote{the \gls*{fa} is a scalar parameter of diffusion derived from
-  the \gls*{dmri} data.} map.
-However, the mask and subsequent analyses are highly dependent on the threshold set
-  \citep{taoka_fractional_2009}.
-To overcome the unreliability of \gls*{fa} thresholding,
-  \cite{zhukov_level_2003} proposed active contours with edges represented
-  by level sets, which evolve on directionally invariant scalar maps.
-\cite{rousson_level_2004} successfully segmented the corpus callosum with
-  region-based level sets based on the eigenvalues of tensors fitted using the
-  \gls*{dmri} data.
-Many studies have addressed the definition of appropriate features for clustering,
-  such as the 5D[A1] representation proposed by \cite{jonasson_segmentation_2005}.
-Other efforts include mixed models based on sets of directionally invariant maps
-  \citep{liu_brain_2007}, iterative \citep{hadjiprocopis_unbiased_2005} and
-  hierarchical \citep{lu_segmentation_2008} clustering,
-  graph-cuts \citep{han_experimental_2009},
-  and volume fraction modeling \citep{kumazawa_improvement_2013}.
-
-To address the segmentation problem by registration, \cite{saad_new_2009} 
-  used Pearson’s correlation coefficient to obtain a linear alignment of the \gls*{t1} and 
-  the \emph{b0}.
-Similarly, a linear registration method was employed by \emph{bbregister} \citep{greve_accurate_2009},
-which uses active contours with edges to search for intensity boundaries in the \emph{b0}
-  image.
-The active contours are initialized using surfaces extracted from the 
-  \gls*{t1} using \emph{FreeSurfer} \citep{fischl_freesurfer_2012}:
-  the \emph{pial} surface (exterior of the cortical \gls*{gm}) and the \emph{white}
-  surface (the \gls*{wm}/\gls*{gm} interface). 
-The \emph{b0} image only includes a detectable frontier for the pial surface, so 
+The tool \emph{bbregister} \citep{greve_accurate_2009} has become a rather standard linear
+  registration method to be used along with \gls*{epi} corrections.
+It uses 3D active contours with edges\footnote{A three-dimensional active contour is a surface
+  and it is ``with edges'' when the contours evolve in search of abrupt intensity steps (the edges) within
+  the target image.} to search for intensity boundaries in the \lowb{} image.
+The active contours are initialized using surfaces extracted from the \gls*{t1} using
+  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}: the \emph{pial} surface (exterior of the
+  cortical \gls*{gm}) and the \emph{white} surface (the \gls*{wm}/\gls*{gm} interface).
+The \lowb{} image only includes a detectable frontier for the pial surface, and thus
   \emph{bbregister} is limited to aligning the cortical layer in this
   application.
-This tool has become the standard method because of its proven robustness, although the 
-  distortions found in \gls*{dmri} are nonlinear.
-To overcome this issue, \emph{bbregister} excludes the
-  regions that are typically warped by artifacts from the boundary search.
+To overcome the problem of nonlinear distortions, \emph{bbregister} excludes the
+  regions that are typically warped from the boundary search.
 However, because the distortion is not considered, it must be addressed separately.
-Nonlinear registration has been performed successfully between \gls*{t2} and \emph{b0}
-  images based on their similarity as a unique method for correcting distortion
-  \citep{kybic_unwarping_2000,studholme_accurate_2000,wu_comparison_2008,tao_variational_2009}.
-However, further registration of the \gls*{t1} and \gls*{t2} images is still required to map the anatomical
-  information (and for \gls*{wm} segmentation) into the \gls*{dmri} space.
+The nonlinear registration approach has been successfully solved with \gls*{t2b} methods,
+  which align the \gls*{t2} and \lowb{} images by exploiting their similarity
+  \citep{kybic_unwarping_2000,studholme_accurate_2000}.
+However, further rigid registration of the \gls*{t1} and \gls*{t2} images is still required to
+  map the anatomical information onto \gls*{dmri}-space.
 
-Therefore, given these issues, a method that simultaneously performs
-  segmentation in the native \gls*{dmri} space and registration of the corresponding \gls*{t1} image
-  onto the \gls*{dmri} data could provide the optimal solution.
-To the best of our knowledge, this strategy has never been proposed for the application described above.
+Here we show a method that registers anatomically-correct surfaces to multivariate images,
+  thereby implicitly segmenting the \gls*{dmri}-space into those structures enclosed by the contours
+  of the subject's brain.
+The hypothesis tested in the present study is that the registration and segmentation
+  problems in \gls*{dmri} can be solved simultaneously, effectively establishing the
+  necessary mapping to project structural information from anatomical images, increasing the
+  geometrical accuracy of the overall process, and avoiding data resampling of \gls*{epi}
+  correction methods.
+Thus, we propose a tool called \regseg{} implementing 3D active contours (surfaces) without edges
+  \citep{chan_active_2001}, which evolve over a free-form deformation field of B-spline
+  basis functions.
+The surfaces are nested and they enclose regions with homogeneous diffusion features
+  (the \gls*{fa} and \gls*{adc} maps for the case of \gls*{dmri} data).
+Optimization is performed using a descent strategy with explicit shape gradients
+  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
+Therefore, unlike most previous methods, \regseg{} does not implement level sets.
+Since the \gls*{epi} distortion is aligned with the \gls*{pe} axis (see
+  \autoref{sec:human_connectome}), \regseg{} includes an anisotropic regularizer for
+  the displacements field proposed by \cite{nagel_investigation_1986}.
+Finally, we evaluate \regseg{} using an extension of our instrumentation framework
+  \citep{esteban_simulationbased_2014}, which simulates known and realistic distortions
+  based on \gls*{dmri} data.
+The evaluation includes a comparison of \regseg{} to an in-house implementation of
+  the \gls*{t2b} correction.
+
+\makeatletter{}\section{Methods}
+\label{sec:methods}
+\subsection{Related work}\label{sec:related_work}
 Previously, joint segmentation and registration have been applied successfully to other problems
@@
   segmentation \citep{gorthi_active_2011}.
-The most thorough approach integrates active contours during image registration 
-  methods.
-\cite{unal_coupled_2005}, and later \cite{wang_joint_2006},
-  improved an existing method \citep{yezzi_variational_2003} based on linear registration
-  to the nonlinear case by implementing a free-form deformation field.
+The most common approach to solve the problem simultaneously is optimizing a deformation model
+  (registration) that supports the evolution the active contours (segmentation), like
+  \citep{paragios_level_2003,yezzi_variational_2003}.
+\cite{unal_coupled_2005}, and later \cite{wang_joint_2006}, replaced the linear registration
+  transform of \citep{yezzi_variational_2003} with a free-form deformation field.
 \cite{droske_mumfordshah_2009} reviewed the existing techniques and proposed two different
@@
 
-The hypothesis tested in the present study is that registration and segmentation
-  problems in \gls*{dmri} can be solved simultaneously, thereby increasing the geometrical
-  accuracy of the process.
-Thus, we propose a tool called \regseg{} that exploits the prior information of shapes
-  extracted using \gls*{t1} images to register the anatomical reference
-  to \gls*{dmri} space, thereby implicitly segmenting these data.
-The proposed approach uses active contours without edges \citep{chan_active_2001}, which evolve to drive a
-  free-form deformation field of B-spline basis functions.
-Optimization is performed using a descent strategy with explicit shape gradients
-  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
-Therefore, unlike most previous methods, \regseg{} does not implement level sets.
-The nonlinear distortion is aligned with one of the imaging axes (see 
-  \autoref{sec:human_connectome}), so \regseg{} includes an anisotropic regularizer for
-  the displacements field proposed by \cite{nagel_investigation_1986}.
-Finally, we evaluated \regseg{} using an extension of our instrumentation framework
-  \citep{esteban_simulationbased_2014}, which simulates known and realistic distortions
-  based on \gls*{dmri} data.
-The evaluation included a comparison with \gls*{t2b} correction by integrating it within the framework
-  of an in-house implementation of the method.
-
-\makeatletter{}\section{Methods}
-\label{sec:methods}
-Let $\gammaset_R = \{\Gamma_m: m \in \mathbb{N}, m \leq N_c\}$ be the set of $N_c$ surfaces
+\subsection{Registration framework and segmentation model derivation}\label{sec:methods_map}
+Let $\gammaset_R = \{\Gamma_m: m \in \mathbb{N}, m \leq N_S\}$ be the set of $N_S$ surfaces
   extracted from the undistorted \gls*{t1} image (the reference space $R$).
@@
 
-\subsection{Cost-function derivation}\label{sec:methods_map}
+\paragraph*{Cost-function derivation}
 In a Bayesian framework for registration \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014},
@@
   $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
-$C$ is the number of channels in the image $M$.
+$C$ is the number of channels comprised in the image $M$.
 In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
-  the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized.
+  the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized. In \autoref{fig:model},
+  this minimization is illustrated for the segmentation of the \gls*{fa} and the \gls*{adc} maps
+  of one subject.
 
+\begin{figure}
+\caption{Evolution of the segmentation model defined by the homogeneous regions $\Omega_i$
+  for a real dataset.
+Panel A shows how the \gls*{fa} and the \gls*{adc} are clustered by the surfaces mapped to
+  $M$ through the ground-truth.
+Panel B shows the same information at the initialization step of \regseg{}.
+Since the surfaces are misaligned in $M$, the spreads of the sampled distribution for
+  each region are large.
+Panel C, after \regseg{} registration, shows how the covariance matrices $\boldsymbol{\Sigma}_l$ 
+  are minimized with respect to panel B, and the resulting segmentation is close
+  to the theoretical situation presented in panel A.
+The left (large) plot in each panel represents a kernel density estimate of the location and spread of
+  each region $\Omega_i$ in the bivariate feature space of $M$, which comprises
+  the \gls*{fa} and \gls*{adc} maps.
+At the top and the right margins, the marginal distributions of $\Omega_i$ are also plotted for
+  the \gls*{fa} and \gls*{adc}, respectively.
+On the right, the sampled distribution of each region is shown in small scatter plot, in contrast to
+  the complete bivariate data.
+}\label{fig:model}
+\end{figure}
+
+\paragraph*{Regularization}
 The smoothness of the resulting displacements field is induced by a Thikonov regularization
@@
      i.e., \isores[mm]{2.0} and \isores[mm]{1.0}, thereby producing sets of tissue fractions maps.
-  4) The tissue fractions were fed into an \acrfull*{mr} simulator, which generated \acrfull*{t1} and \acrfull*{t2} -like images at the
+  4) The tissue fractions were fed into a \acrfull*{mr} simulator, which generated \acrfull*{t1} and \acrfull*{t2} -like images at the
      two possible resolutions.
   5) The \regseg{} tool was applied using the warped test images as multispectral moving images and $\gammaset_R$ as shape priors.
-  6) The agreement between the surfaces fitted by \regseg{} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$
-     using automatically generated visual reports was assessed by computing the Hausdorff distance between the corresponding surfaces.}\label{fig:evphantoms}
+  6) The agreement between the surfaces fitted by \regseg{} ($\hat{\gammaset}_\text{test}$) and $\gammaset_\text{true}$ were assessed
+     visually using automatically generated visual reports and quantitativelly with the Hausdorff distance between the
+     corresponding surfaces.}\label{fig:evphantoms}
 \end{figure*}
@@
 The original acquisitions are released within ``unprocessed'' packages, whereas
-  the ``minimally preprocessed'' are corrected for artifacts, brain-extracted,
-  and spatially normalized, while some results have been subjected to the standard processing
-  pipeline of \emph{FreeSurfer}.
+  the ``minimally preprocessed'' packages contain the corresponding images after
+  some processing (correction for several artifacts, brain-extraction, spatial
+  normalization, etc.).
 We refer the reader to \citep{essen_human_2012} for exact details of the acquisition
@@
 
-\begin{figure}
-\includegraphics[width=\linewidth]{figure04}
-\caption{Segmentation model defined by the homogeneous regions $\Omega_i$, where
-  the image for segmentation is partitioned by the theoretical projection 
-  ($\gammaset_{true}$) of $\gammaset_R$ onto $M$.
-The plot represents a kernel density estimate of the location and spread of
-  each region $\Omega_i$ in the bivariate feature space of $M$, which comprises 
-  the \gls*{fa} and \gls*{adc} maps.
-At the top and the right margins, the marginal distributions of $\Omega_i$ are also plotted for
-  the \gls*{fa} and \gls*{adc}, respectively.
-}\label{fig:model}
-\end{figure}
-
 \paragraph*{Segmentation model}
-Based on our experience \citep{esteban_brain_2012} and previous studies \citep{ennis_orthogonal_2006},
+Based on our experience   and previous studies \citep{ennis_orthogonal_2006},
   we defined the moving image as a stack of the \gls*{fa} and \gls*{adc} maps derived
@@
   the so-called field map.
-We derived the deformation $U_{true}$ from the field map image released with
+We derived the deformation $U_\text{true}$ from the field map image released with
   the corresponding packages of each dataset in the \gls*{hcp}.
-The field map was unwrapped\footnote{fieldmaps are phase maps, which have been clipped in the interval
+The field map was unwrapped\footnote{fieldmaps are phase maps, which are intrinsically clipped in the interval
   of $[-\pi, \pi)$ [rads] or [rads/s].} and smoothed before applying \eqref{eq:fieldmap}.
@@
 For five test datasets, we uniformly sampled the space of distortions
-  $\hat{U} = \epsilon \cdot U_{true} = \vec{r} + \epsilon \, u_\text{PE}$
+  $\hat{U} = \epsilon \cdot U_\text{true} = \vec{r} + \epsilon \, u_\text{PE}$
   (with $\epsilon \in [-1.1, 1.1]$ and $u_\text{PE}$ from \eqref{eq:fieldmap}),
   and we evaluated the data term of the cost function \eqref{eq:energy}.
-The minimum metric was consistently at the ground-truth location ($\epsilon=0.0$)
-  in all cases (\suppl{Figure S2}).
+The minimum of the cost function (\autoref{sec:methods_map}) was consistently located at
+  $\epsilon=0.0$ (the ground-truth) for all of the cases (\suppl{Figure S2}).
 
@@
 \end{document}
-[A1]I understand that many of the abbreviations in this manuscript may be common abbreviations, however, depending on the journal to which you are submitting, it may be necessary to provide definitions in accordance with your target journal author abbreviation guidelines.

### 2015-09-07 Publications/2015-NeuroImage/expanded.tex
add expanded

--- 2015-04-29 Publications/2015-NeuroImage/ELS285_article_edited.tex
@@
 Many studies have addressed the definition of appropriate features for clustering,
-  such as the 5D[A1] representation proposed by \cite{jonasson_segmentation_2005}.
+  such as the 5D representation proposed by \cite{jonasson_segmentation_2005}.
 Other efforts include mixed models based on sets of directionally invariant maps
@@
 To address the segmentation problem by registration, \cite{saad_new_2009}
-  used Pearson’s correlation coefficient to obtain a linear alignment of the \gls*{t1} and 
+  used the Pearson's correlation coefficient to obtain a linear alignment of the \gls*{t1} and
   the \emph{b0}.
@@
 \paragraph*{Segmentation model}
-Based on our experience \citep{esteban_brain_2012} and previous studies \citep{ennis_orthogonal_2006},
+Based on our experience   and previous studies \citep{ennis_orthogonal_2006},
   we defined the moving image as a stack of the \gls*{fa} and \gls*{adc} maps derived

### 2015-09-07 Publications/2015-NeuroImage/03-introduction.tex
update changes

@@
   combinations of the original methods \citep{holland_efficient_2010,andersson_comprehensive_2012,
-  irfanoglu_drbuddi_2014}.
+  irfanoglu_drbuddi_2015}
 

### 2015-09-08 Publications/2015-NeuroImage/responseLetter2.tex
v1.0 of resubmission

--- 2015-09-03 Publications/2015-NeuroImage/responseLetter.tex
@@
 \usepackage{nameref}
+
+\usepackage{amsmath}
+\usepackage{amssymb}
 
@@
 
-\noindent New title: \emph{\color{red} Segmentation of homogeneous regions of diffusion MR features by mapping
-  same-subject surfaces extracted in structural space}.
+\noindent New title: \emph{{\color{blue} Surface}-driven registration method for the structure-informed segmentation of
+  diffusion MR images}.
 
@@
 \begin{enumerate}
-\item Changed the title
-\item Changed abstract: {\color{red} say what here}
-\item Changed introduction: {\color{red} say what here}
+\item Changed abstract, more oriented on the problem now.
+\item Re-structured introduction: lines 1-14 state the background and the problem more clearly, lines 15-27 review
+  the literature existing about direct segmentation of dMRI data and lines 28-55 describe the alternative approach
+  by registering structural images and the problem of EPI distortions.
+  The methodological background of active-contours and registration-segmentation methods has been relocated in
+    the beginning of section 2, lines 73-85.
 \end{enumerate}
 
-Regarding the second sentence, we agree on that our method can be seen as a registration of an atlas to the
-  image being segmented.
-However, we need to point out some particular features of the method.
-First is that surfaces in our method play the role of \emph{strong} priors, since the method does not allow
-  for changes in their topology (affecting the segmentation side of the approach) and they need to be
-  extremely accurate so that the mapping from structural space is correct (registration).
-This might not be the case on atlas-based segmentation, which generally needs to deal with problems by the
-  inappropriate topology of surfaces in the atlas, and the result of registration is relatively tolerant
-  to the imperfect location of surfaces from the template.
 Our method can be seen as a segmentation framework using active contours without edges in which the evolution
@@
 Some comments of Reviewer \#2 are very similar to this one (particularly \autoref{com:11}).
-
-In order to clarify this issue, we have modified the manuscript as follows: {\color{red} say what here}.
 }
@@
 \reply{
-We have noticed that the way we explained the surfaces necessary to segment real dMRI datasets
-  was not appropriate.
-On one hand, in this paper we propose a segmentation model to cluster the multivariate space of features
-  provided by the FA and ADC map.
-As correctly interpreted by reviewer \#2 mentioning it explicitly, and potentially reviewer \#1 noticed,
-  the model is given by a set of nested surfaces.
-However, it was probably unclear in the manuscript that those surfaces and their nesting is the cornerstone
-  of the method, while the set of surfaces selected in validation are only those that will have an important
-  impact in downstream processing, such as reconstruction or tractography.
-
-In order to separate the concepts of validation and the definition of the model, we have modified the
-  manuscript as follows: {\color{red} say what here}.
+We have defined clearly that the choice of surfaces to segment the target data is crucial, and define the
+  segmentation model that drives the registration process (lines 110-113 are new).
+We have also improved former Figure 4, and brought it earlier in the paper (now it is Figure 2).
+Now, the figure shows how the segmentation model works: A is the ideal segmentation, B shows the model in
+  the initialization settings, and C shows the model after using \emph{regseg}, demonstrating how the covariance
+  matrices for each region are reduced.
+Since the former visualization was a bit clumsy, we have extracted the distribution of diffusion features by region on
+  the right side of each row.
 }
@@
 \reply{
-We agree with the reviewer on that the novelty of the methodological elements of our approach is very limited.
-However, to our knowledge there exists no method with a paired combination of the following elements:
+We agree with the reviewer on that the novelty of the methodological elements of our approach is limited.
+However, to our knowledge there exists no alternative method with a paired combination of the following elements:
 \begin{itemize}
@@
   method.
-In this regard, we have investigated some more potential work related to ours, modifying the manuscript as
-  follows: {\color{red} say what here}.
+For the sake of clarity, we have removed the methodological background from introduction to section 2, lines 73-85.
 }
@@
 \reply{
-  The reviewer is right, assuming that pixels are i.i.d. is wrong but a widely accepted simplification
-that performs correctly with the following details:
-\begin{enumerate}
-\item Pixels are not generally independent, and this issue is generally addressed imposing prior
-  information over neighboring, such as Markov Random Fields.
-In our framework, local dependency of pixels is implicitly addressed by the nested surfaces that impose
-  the piecewise smoothness of the model.
-\item Pixels are not identically distributed, and this issue posed important problems when defining
-  our segmentation model.
-However, we empirically found that the surfaces described in {\color{red} say line here} and their
-  nesting made a good representation of the distribution of each region.
-Actually, this reason may be at basis of the poor performance of most of the methods originally referenced
-  in lines 19-28.
-Our segmentation model ensures that pixels belonging to a certain region are ``almost'' identically distributed.
-The bivariate approach using FA and ADC facilitates this, as demonstrated in Figure 4.
-\end{enumerate}
-
-Nonetheless, we have specifically introduced this discussion in our manuscript {\color{red} say where}
-  and improved Figure 4 to better describe this point.
+The reviewer is right, assuming that pixels are i.i.d. is wrong but a widely accepted simplification.
+We have supported the assumptions with new lines 105-107:
+\textit{Even though the features being segmented are not generally i.i.d., the spatial interdependency of voxels is
+implicitly supported by the piecewise smooth partition of the space $\boldsymbol\Omega$.
+In Figure 2 it is shown that marginal distributions of data can be approximated by multivariate normal distributions.}
 }
@@
 \reply{
-
+We have improved former Figure 4 (now, Figure 2) to clarify that the model strongly depends on the choice of surfaces.
+Adding more surfaces makes the model more robust to local minima at the cost of largely increased computation times.
+We agree with the reviewer on that Figure S1 is not illustrative of the metric for free-deformation fields, but
+  our intention was to show the performance of the method in realistic settings.
+This implies distortions with nonzero components on the phase-encoding axis (the $y$-axis in our experiments), and
+  therefore Figure S1 becomes representative since it samples the distortions space covering plausible deformations
+  for both encoding directions (Anterior-to-posterior and posterior-to-anterior).
+We have included information about the maximum distortion for which the metric was sampled.
 }
@@
   very minimal.
-We appreciate the comment, and modified the manuscript to clarify this point {\color{red} say where here}.
-We also modified Figure 3 to reflect this aspect.
+We have modified Figure 3 to reflect this aspect.
 }
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+We have added the following in the Discussion section (lines 262-267):
+
+\textit{The distributions of errors along surfaces vary importantly depending on the shape of the
+  phantom (see Figure 5B).
+The misregistration error of the ``gyrus'' phantom showed a much lower spread than that for the other shapes.
+We argue that the symmetry of those other shapes posed problem in driving the contours towards the appropriate
+  region and producing some ``sliding'' effect between the faces of the surface and their ground-truth position.
+This effect should not be present in real datasets, thanks to the very convoluted cortical layer, and the
+  directional restriction of distortion.}
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+We thank the reviewer for pointing out that the overall goal is not clearly explained.
+We have replaced the term ``active-contours'' by ``surfaces'' whenever possible.
+We have re-organized the introduction fully, as suggested (see reply to \autoref{com:1} for a brief explanation).
+We have included references of the main EPI distortion corrections (see lines 31-43).
+
+The \emph{regseg} tool is designed to allow any displacement, as demonstrated with the phantoms.
+However, as the reviewer points out, the deformation field is completelly restricted to the phase-encode direction
+  for the real datasets.
+We have added lines 204-206 to clarify this point.
+}
 
@@
   FA and ADC.
-
-However, and for the sake of clarity, we have modified the manuscript as follows:  {\color{red} say what here}.}
+The manuscript has been re-phrased accordingly in the title, abstract and introduction.}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+The reviewer is right, we have changed the paragraph about dMRI segmentation (lines 15-27) to
+  be more precise.
+We agree with the reviewer that there actually exist many methods to segment fiber tracts.
+However, our method is intended to support fiber tracking methods requiring detailed anatomical
+  information, such as \citep{smith_anatomicallyconstrained_2012}.
+We have supported the applications more clearly on lines 11-14.
+Clustering tractograms could be another potential application of \emph{regseg},
+  but we agree with the reviewer that current methods perform well without
+  additional structural information.
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+This comment is very related to \autoref{com:14}.
+We have modified the abstract and introduction accordingly.
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+This comment is a consequence of the two previous comments.
+The reviewer is right in the sense that we did not clearly state what we are segmenting, and therefore
+  the references in line 24 are a bit orphaned.
+All methods reviewed in lines 19-28 cluster/segment/classify \emph{voxels} in dMRI space based on
+  one or another feature derived from dMRI data.
+We have clarified the manuscript according to \autoref{com:14} and \autoref{com:15}, and specifically
+  rewritten lines 19-28 to explicitly state what those methods are clustering.
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+Again, the reviewer finds that the main goal of the paper is not clearly stated.
+We understand that modifications derived from the previous comments satisfy the answer to the concern,
+  and changes include abstract and introduction.
+We have stated more clearly the importance of the set of surfaces being registered to produce a reliable
+  segmentation model of the two diffusion features selected: FA and ADC
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+We have added the suggested reference (among other of interest) in lines 204-206.
+In lines 59-61 we have stated the differences of \emph{regseg} w.r.t. EPI distortion corrections.
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+We have updated the title to represent this aspect.
+We have also included all previous changes to fulfill this request.
+}
 
@@
   axis.
-We understand that this point should be clearly stated in the manuscript and modified it as follows:
-{\color{red} say what here}.
+New lines 203-205 now state this point clearly.
 }
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+We thank the reviewer for positively weighing our experiments and results.
+As stated in line 214, we use the settings from the \emph{ExploreDTI} software package \citep{leemans_exploredti_2009}, that
+  internally uses \emph{elastix} \citep{klein_elastix_2010} in registration.
+A footnote in line 215 directly links the settings for this method.
+}
 
@@
 \end{reviewpoint}
-\reply{We hope this point is answered in \autoref{com:10}.}
+\reply{We hope this point is answered in \autoref{com:10}, and tryied to avoid this concept
+  coming from the 2D processing field.
+Now, we refer to the contours as surfaces whenever possible.
+Regarding the ``without edges'' feature, we have clarified how the surfaces enclose homogeneous
+  regions (and therefore, they are not used to search for image edges).
+}
 

### 2015-09-08 Publications/2015-NeuroImage/02-abstract.tex
v1.0 of resubmission

@@
 % -*- root: 00-main.tex -*-
-Current applications of whole-brain tractography to \acrlong*{dmri} data require highly precise
-  delineations of anatomical structures, which are usually projected from \acrlong*{t1} images
-  by registration.
-In this study, we propose \regseg{}, which is a simultaneous segmentation and registration method
-  that uses active contours without edges extracted from structural images.
-The contours evolve through a free-form deformation field supported by the B-spline basis
+Current methods using diffusion MRI to map the microstructure and connectivity of the
+  human brain increasingly require precise delineations of anatomical structures.
+The problem is typically solved by projecting the structural information extracted
+  in T1-weighted images, by means of nonlinear registration since diffusion
+  MRI data present geometrical distortions.
+Here we propose \regseg{}, which segments homogeneous regions within multivariate
+  images by projecting a set of nested surfaces extracted from structural images.
+The surfaces evolve through a free-form deformation field supported by the B-spline basis
   to optimally map the contours onto the data in the target space.
@@
   randomly generated deformations, where subvoxel accuracy was achieved.
-We then applied \regseg{} to a registration/segmentation task using 16 real diffusion MRI
-  datasets from the \acrlong*{hcp}, which were warped by realistic and nonlinear distortions that are typically
-  present in these data.
-We computed the misregistration error of the contours estimated by \regseg{} with respect
+We proposed a set of surfaces to segment the fractional anisotropy and the apparent diffusion
+  coefficient maps derived from diffusion MRI datasets.
+These surfaces extracted from the same-subject T1-weighted image successfully segmented 16 real
+  diffusion MRI datasets from the Human Connectome Project that were warped by realistic and nonlinear
+  distortions.
+We computed the misregistration error of the surfaces estimated by \regseg{} with respect
   to their theoretical location using the ground truth, thereby obtaining a 95\% CI of 0.56--0.66 mm
   distance between corresponding mesh vertices, which was below the 1.25 mm isotropic resolution of the images.
-We also compared the performance of our proposed method with a widely used registration tool, which showed
+We also compared the performance of the proposed method with a widely used registration tool, which showed
   that \regseg{} outperformed this method in our settings.

### 2015-09-08 Publications/2015-NeuroImage/03-introduction.tex
v1.0 of resubmission

@@
 
-The tool \emph{bbregister} \citep{greve_accurate_2009} has become a rather standard linear
-  registration method to be used along with \gls*{epi} corrections.
-It uses 3D active contours with edges\footnote{A three-dimensional active contour is a surface
-  and it is ``with edges'' when the contours evolve in search of abrupt intensity steps (the edges) within
-  the target image.} to search for intensity boundaries in the \lowb{} image.
+The tool \emph{bbregister} \citep{greve_accurate_2009} has been widely adopted as the standard
+  registration method to be used along with the \gls*{epi} correction of choice.
+It implements a linear mapping, and uses 3D active contours with edges\footnote{A three-dimensional active contour
+  is a surface and it is ``with edges'' when the contours evolve in search of abrupt intensity steps
+  (the edges) within the target image.} to search for intensity boundaries in the \lowb{} image.
 The active contours are initialized using surfaces extracted from the \gls*{t1} using
@@
 %   correctly but the mapping found in registration does not represent any valuable information.
-The hypothesis tested in the present study is that the registration and segmentation
-  problems in \gls*{dmri} can be solved simultaneously, effectively establishing the
-  necessary mapping to project structural information from anatomical images, increasing the
-  geometrical accuracy of the overall process, and avoiding data resampling of \gls*{epi}
-  correction methods.
+The hypothesis underlying is that the registration and segmentation problems in \gls*{dmri}
+  can be solved simultaneously with three key benefits:
+  1) effectively establishing the necessary mapping to project structural information from anatomical images,
+  2) increasing the geometrical accuracy of the overall process,
+  and 3) avoiding data resampling of \gls*{epi} correction methods.
 Thus, we propose a tool called \regseg{} implementing 3D active contours (surfaces) without edges

### 2015-09-08 Publications/2015-NeuroImage/04-methods.tex
update changes
updating references
v1.0 of resubmission

@@
 
-\subsection{Registration framework and segmentation model derivation}\label{sec:methods_map}
+\subsection{Registration framework and segmentation model}\label{sec:methods_map}
 Let $\gammaset_R = \{\Gamma_m: m \in \mathbb{N}, m \leq N_S\}$ be the set of $N_S$ surfaces
@@
   to the reference domain.
+The general overview of how the surfaces interact with the registration framework
+  is presented in \autoref{fig:method}.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figure01}
+  \caption{The active contours are defined as the interfacing surfaces of the competing
+    \glspl{roi} $\Omega_i$, which
+  are represented in green and dark blue in this close-up.
+  They evolve iteratively following their inward normals $\hat{n}_i$ at each vertex
+    $\vec{v}_i$ of the mesh.
+  The gradient speeds $\bar{s}_i$ drive registration, which are computed as the disparity of the data
+    energies with respect to the two limiting regions of $M(\vec{v}_i)$, the features of the image
+    $M$ in the location of vertex $\vec{v}_i$ (see \suppl{equation SM5}).
+  In this figure, $\bar{s}_1$ is written in the lower
+    box, with $\Omega_{wm}$ in the inner limiting region, $\Omega_{gm}$ the outer region, and
+    $w_{0,1}$ is the relative area associated with vertex $\vec{v}_1$ with respect to
+    the total area of surface $\Gamma_0$.
+      }\label{fig:method}
+\end{figure}
 
@@
 \begin{figure}
+\includestandalone[width=\linewidth]{figure02-model}
 % \includegraphics[width=\linewidth]{figure04A}
@@
 Since the surfaces are misaligned in $M$, the spreads of the sampled distribution for
-  each region are large.
+  each region are larger than those in A.
 Panel C, after \regseg{} registration, shows how the covariance matrices $\boldsymbol{\Sigma}_l$
@@
 \label{sec:numerical_implementation}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figure01}
-  \caption{The active contours are defined as the interfacing surfaces of the competing
-    \glspl{roi} $\Omega_i$, which
-  are represented in green and dark blue in this close-up.
-  They evolve iteratively following their inward normals $\hat{n}_i$ at each vertex
-    $\vec{v}_i$ of the mesh.
-  The gradient speeds $\bar{s}_i$ drive registration, which are computed as the disparity of the data
-    energies with respect to the two limiting regions of $M(\vec{v}_i)$, the features of the image
-    $M$ in the location of vertex $\vec{v}_i$ (see \suppl{equation SM5}).
-  In this figure, $\bar{s}_1$ is written in the lower
-    box, with $\Omega_{wm}$ in the inner limiting region, $\Omega_{gm}$ the outer region, and
-    $w_{0,1}$ is the relative area associated with vertex $\vec{v}_1$ with respect to
-    the total area of surface $\Gamma_0$.
-      }\label{fig:method}
-\end{figure}
 
@@
   \begin{align}
-  \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\delta^{-1} \, \vec{u}_k^t - \vec{g}_k\} }                  {\mathcal{F}\{(\delta^{-1} + \boldsymbol{\alpha})\, I - \boldsymbol{\beta}\Delta\}} \right\},
+  \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\delta^{-1} \, \vec{u}_k^t - \vec{g}_k\} }
+                    {\mathcal{F}\{(\delta^{-1} + \boldsymbol{\alpha})\, I - \boldsymbol{\beta}\Delta\}} \right\},
   \label{eq:update_equation}
@@
  and target image smoothing) are discussed in \suppl{section S1}.
-The actual choices of the parameter settings are publicly distributed with the source code for the experiments.
+The actual choices of the parameter settings are publicly distributed with the source code for the
+  experiments\footnote{\url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data}}.
 The final settings were obtained manually based on feedback from the post-registration convergence
@@
     stacked to feed the multivariate input required by \regseg{}.
-  4) The method was run to obtain $U_{test} = \hat{U}_{true}$, i.e., the estimate of
+  4) The method was run to obtain $U_\text{test} = \hat{U}_\text{true}$, i.e., the estimate of
     the ground-truth deformation.
-  5) The results were evaluated visually and quantitatively.}\label{fig:evworkflows}
+  5) The results were evaluated visually and quantitatively.
+  The arrows point to edges in the target images (yellow arrows for \gls*{fa}, blue for \gls*{adc}) that
+    should be aligned with a surface, showing how distortion breaks the mapping with the structural space in which
+    the contours are defined.}\label{fig:evworkflows}
 \end{figure*}
@@
 \paragraph*{Ground-truth generation}
-Realistic deformation is achieved by generating displacements fields that satisfy the theoretical
-  properties of distortion \citep{jezzard_correction_1995}.
+Realistic deformation was achieved by generating displacements fields that satisfy the theoretical
+  properties of distortion.
 The displacements along the \gls*{pe} axis of the \gls*{dmri} image are related to the local
-  deviation of the field $\Delta B_0(\vec{r})$ from its nominal value $B_0$, as follows:
+  deviation of the field $\Delta B_0(\vec{r})$ from its nominal value $B_0$  \citep{jezzard_correction_1995},
+  as follows:
 
@@
   interest (\gls*{fa} and \gls*{adc}) using \emph{MRtrix} \citep{tournier_mrtrix_2012}.
+Since the distortion only happens along the \gls*{pe} direction ($y$-axis in our settings),
+  \regseg{} was configured to allow nonzero displacements only on the corresponding axis.
+We also configured the \gls*{t2b} method accordingly.
 
@@
   perform registration.
+Settings used with \emph{elastix} are conveniently available in its format\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/t2b_elastix_y.txt}}.
 
@@
   \gls*{swindex} as the area-weighted distance between the corresponding vertices of
-  $\gammaset_{true}$ and their estimate obtained by the method under the test $\hat{\gammaset}_\text{test}$:
+  $\gammaset_\text{true}$ and their estimate obtained by the method under the test $\hat{\gammaset}_\text{test}$:
 

### 2015-09-08 Publications/2015-NeuroImage/05-results.tex
v1.0 of resubmission

@@
   (see \autoref{sec:experiments_evaluation}) between the theoretical $\gammaset_\text{true}$ and
-  the estimate obtained by \regseg{} ($\hat{\gammaset}_{test}$).
+  the estimate obtained by \regseg{} ($\hat{\gammaset}_\text{test}$).
 The results demonstrated that the accuracy was consistent and high, and below the image resolution.

### 2015-09-08 Publications/2015-NeuroImage/06-discussion.tex
v1.0 of resubmission

@@
   voxel spacing.
+The distributions of errors along surfaces vary importantly depending on the shape of the
+  phantom (see \autoref{fig:phantom}B).
+The misregistration error of the ``gyrus'' phantom showed a much lower spread than that
+  for the other shapes.
+We argue that the symmetry of those other shapes posed problem in driving the contours
+  towards the appropriate region and producing some ``sliding'' effect between the
+  faces of the surface and their ground-truth position.
+This effect should not be present in real datasets, thanks to the very convoluted cortical
+  layer, and the directional restriction of distortion.
 
@@
 We designed \regseg{} as a method for segmenting \gls*{dmri} data by exploiting the
-  anatomical information extracted from the corresponding \gls*{t1} image of the subject.
-Applications in whole-brain tractography \citep{smith_tractbased_2006,craddock_imaging_2013}
-  usually solve this problem with a two-step approach.
+  anatomical information extracted from a structural image (typically a \gls*{t1} image)
+  of the subject.
+Current applications based on whole-brain tractography usually solve this problem with a two-step approach.
 First, the images are corrected for nonlinear distortions using auxiliary acquisitions

### 2015-09-08 Publications/2015-NeuroImage/figure02-model.tex
v1.0 of resubmission

% -*- root: 00-main.tex -*-
% @Author: Oscar Esteban
% @Date:   2015-04-09 13:06:32
% @Last Modified by:   Oscar Esteban
% @Last Modified time: 2015-09-08 13:12:00

\documentclass[tikz]{standalone}
\usepackage{graphicx}
\graphicspath{{figures/}}

\usepackage[T1]{fontenc}
\usepackage{charter}

\begin{document}
\begin{tikzpicture}
  \node[inner sep=10pt](fig01a) at (0,15.5)
    {\hspace*{20pt}\includegraphics[width=20cm]{figure04A}};
  \node[inner sep=10pt](fig01b) at (0,7.8)
    {\hspace*{20pt}\includegraphics[width=20cm]{figure04B}};
  \node[inner sep=10pt](fig01c) at (0,0)
    {\hspace*{20pt}\includegraphics[width=20cm]{figure04C}};
  \node[circle, text=black!75, inner sep=0] at (-10.65,18.5) {\Huge \textbf{A}};
  \node[circle, text=black!75, inner sep=0] at (-10.65,11.0) {\Huge \textbf{B}};
  \node[circle, text=black!75, inner sep=0] at (-10.65,3.15) {\Huge \textbf{C}};
\end{tikzpicture}
\end{document}

### 2015-09-15 Publications/2015-NeuroImage/00-main.tex
update changes
updating references
more changes, settings, contributions...

@@
 \documentclass[3p,authoryear,fleqn]{elsarticle}
-\journal{ArXiv archiving for pre-publication review}
+\journal{NeuroImage}
+\usepackage[utf8]{inputenc}
 
@@
 
-\title{Active contours-driven registration method for the structure-informed segmentation of diffusion MR images\tnoteref{t1}}
+\title{Surface-driven registration method for the structure-informed segmentation of diffusion MR images\tnoteref{t1}}
 \tnotetext[t1]{Supplemental materials to this work will be publicly available upon publication with doi:\href{http://figshare.com/s/459c26b4ee8211e493b306ec4bbcf141}{10.6084/m9.figshare.1397502}.
@@
 \begin{keyword}
-active~contours \sep cortical~parcellation \sep diffusion MRI \sep nonlinear~registration \sep segmentation \sep susceptibility~distortion.
+active~surfaces \sep cortical~parcellation \sep diffusion MRI \sep nonlinear~registration \sep segmentation \sep susceptibility~distortion.
 \end{keyword}

### 2015-09-15 Publications/2015-NeuroImage/03-introduction.tex
more changes, settings, contributions...

@@
 The microstructural map is then used to draw the preferential orientations of diffusion
-  across regions using tractography \citep{mori_threedimensional_1999}.
+  across the brain using tractography \citep{mori_threedimensional_1999}.
 Finally, the streamlines found are filtered with a cortical parcellation
@@
   theoretical properties of the \gls*{epi} distortion.
-However, both solutions require extra acquisitions to be included within the \gls*{dmri} scanning
+Both solutions require extra acquisitions to be included within the \gls*{dmri} scanning
   protocol, such as fieldmaps \citep{jezzard_correction_1995}, \glspl*{dwi} with a different
@@
   \citep{kybic_unwarping_2000}.
-In the following, we will refer to this latter approach as \gls*{t2b} correction.
-\Gls*{epi} corrections have been an active field of research, with numerous refinements and
-  combinations of the original methods \citep{holland_efficient_2010,andersson_comprehensive_2012,
-  irfanoglu_drbuddi_2015}
-
-The tool \emph{bbregister} \citep{greve_accurate_2009} has been widely adopted as the standard
-  registration method to be used along with the \gls*{epi} correction of choice.
-It implements a linear mapping, and uses 3D active contours with edges\footnote{A three-dimensional active contour
-  is a surface and it is ``with edges'' when the contours evolve in search of abrupt intensity steps
-  (the edges) within the target image.} to search for intensity boundaries in the \lowb{} image.
-The active contours are initialized using surfaces extracted from the \gls*{t1} using
-  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}: the \emph{pial} surface (exterior of the
-  cortical \gls*{gm}) and the \emph{white} surface (the \gls*{wm}/\gls*{gm} interface).
-The \lowb{} image only includes a detectable frontier for the pial surface, and thus
-  \emph{bbregister} is limited to aligning the cortical layer in this
-  application.
-To overcome the problem of nonlinear distortions, \emph{bbregister} excludes the
-  regions that are typically warped from the boundary search.
-However, because the distortion is not considered, it must be addressed separately.
 The nonlinear registration approach has been successfully solved with \gls*{t2b} methods,
@@
   map the anatomical information onto \gls*{dmri}-space.
+\Gls*{epi} corrections have been an active field of research, with numerous refinements and
+  combinations of the original methods \citep{holland_efficient_2010,andersson_comprehensive_2012,
+  irfanoglu_drbuddi_2015}
 
@@
   \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
-Therefore, unlike most previous methods, \regseg{} does not implement level sets.
-Since the \gls*{epi} distortion is aligned with the \gls*{pe} axis (see
-  \autoref{sec:human_connectome}), \regseg{} includes an anisotropic regularizer for
-  the displacements field proposed by \cite{nagel_investigation_1986}.
 Finally, we evaluate \regseg{} using an extension of our instrumentation framework

### 2015-09-15 Publications/2015-NeuroImage/04-methods.tex
more changes, settings, contributions...

@@
 \label{sec:methods}
-\subsection{Related work}\label{sec:related_work}
+\subsection{Related work and contributions}\label{sec:related_work}
 Previously, joint segmentation and registration have been applied successfully to other problems
@@
 \cite{unal_coupled_2005}, and later \cite{wang_joint_2006}, replaced the linear registration
-  transform of \citep{yezzi_variational_2003} with a free-form deformation field.
+  transform of \cite{yezzi_variational_2003} with a free-form deformation field.
 \cite{droske_mumfordshah_2009} reviewed the existing techniques and proposed two different
@@
   of atlas-based segmentation.
+
+The tool \emph{bbregister} \citep{greve_accurate_2009} has been widely adopted as the standard
+  registration method to be used along with the \gls*{epi} correction of choice.
+It implements a linear mapping, and uses 3D active contours with edges\footnote{A three-dimensional active contour
+  is a surface and it is ``with edges'' when the contours evolve in search of abrupt intensity steps
+  (the edges) within the target image.} to search for intensity boundaries in the \lowb{} image.
+The active contours are initialized using surfaces extracted from the \gls*{t1} using
+  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}: the \emph{pial} surface (exterior of the
+  cortical \gls*{gm}) and the \emph{white} surface (the \gls*{wm}/\gls*{gm} interface).
+The \lowb{} image only includes a detectable frontier for the pial surface, and thus
+  \emph{bbregister} is limited to aligning the cortical layer in this
+  application.
+To overcome the problem of nonlinear distortions, \emph{bbregister} excludes the
+  regions that are typically warped from the boundary search.
+Indeed, the distortion must be addressed separately because it is not supported by
+  the affine transformation model.
+
+\Regseg{} uses shape-gradients computed on a hierarchical set of explicit surfaces
+  (triangularized meshes), unlike the most of previous registration-and-segmentation
+  methods that implement multiphase level sets.
+\Regseg{} can be seen as a derivative of \emph{bbregister} for the use of active surfaces.
+However, the deformation model is nonlinear and the surfaces in \regseg{} look for the
+  homogeneity within the regions enclosed, instead of image edges.
+\Regseg{} includes an anisotropic regularizer for the displacements field proposed by
+  \cite{nagel_investigation_1986}.
+This regularization strategy conceptually falls in the midway between the Gaussian smoothing
+  generally included in most of the existing methodologies, and the complexity of
+  the elasticity smoother of \cite{guyader_combined_2011}.
+Finally, we also contribute proposing a piecewise smooth segmentation model defined by
+  a selection of nested surfaces to partition the multispectral space
+  comprehending the \gls*{fa} and the \gls*{adc} maps and ultimately identify these
+  structures in \gls*{dmri} space.
+With the definition of appropriate models, \regseg{} may be applied in other fields.
 
@@
   where $P(M \mid U,\omegaset)$ is the data likelihood.
-Because $\omegaset$ are mapped by $U$, then we simplify
+Because $\omegaset$ are mapped by $U$, we simplify
   $P(U, \omegaset) = P(U) \implies P(M \mid U,\omegaset) = P(M \mid U)$.
@@
 Densification is achieved using a set of associated basis functions $\psi_k$ \eqref{eq:nodes_tfm}.
-In our implementation, $\psi_k$ is selected as a tensor-product B-spline kernel
-  of degree three.
+In our implementation, $\psi_k$ is a tensor-product B-spline kernel of degree three.
 
@@
 
-\paragraph*{Implementation details, settings, and convergence}
+\paragraph*{Implementation details and convergence}
 \label{sec:conv_report}
@@
   and target image smoothing) are discussed in \suppl{section S1}.
-The actual choices of the parameter settings are publicly distributed with the source code for the
-  experiments\footnote{\url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data}}.
-The final settings were obtained manually based on feedback from the post-registration convergence
-  reports (such as that found in \suppl{section S1.3}).
-We released \regseg{} along with the tool to generate the convergence report.
 
@@
 
+\paragraph*{Segmentation model and settings}
+The segmentation model of the phantoms is implicitly defined in each shape.
+All phantoms comprehend an inner surface enclosing a uniform \gls*{wm}-like region,
+  and an outer surface wrapping a \gls*{gm}-like layer.
+The outside is filled with uniform background (see \autoref{fig:evphantoms}).
+All the experimental settings used for the phantoms are made available in
+  a unique configuration file\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/regseg_default.json}}.
+
 \subsection{Real datasets} \label{sec:human_connectome}
@@
   multi-shell \gls*{dmri} images.
+Since we obtained the \gls*{dmri} data from the minimally preprocessed package, these
+  images are corrected for \gls*{epi} distortions and spatially normalized in
+  \gls*{t1} space.
 
@@
   interest (\gls*{fa} and \gls*{adc}) using \emph{MRtrix} \citep{tournier_mrtrix_2012}.
-Since the distortion only happens along the \gls*{pe} direction ($y$-axis in our settings),
-  \regseg{} was configured to allow nonzero displacements only on the corresponding axis.
-We also configured the \gls*{t2b} method accordingly.
 
@@
 
-\paragraph*{Cross-comparison}
+\paragraph*{\Regseg{} settings}
+Since the distortion only happens along the \gls*{pe} direction ($y$-axis in our settings),
+  \regseg{} was configured to allow nonzero displacements only on the corresponding direction.
+For the experiments on real data, \regseg{} established a multi-resolution pyramid of
+  B-Spline basis, with control points distributed on grids of the following spacings:
+  40$\times$100$\times$40 [mm] for the first (coarser level), \isores[mm]{30} for the second,
+  and 20$\times$30$\times$10 [mm] in the third.
+Only the first and second levels included Gaussian smoothing of the target image ($\sigma$ = [2.0, 0.5] mm,
+  respectively).
+The actual choices of the parameter settings are publicly distributed with the source code for the
+  experiments\footnote{\url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data/regseg_hcp.json}}.
+These settings were obtained manually, driven by the feedback obtained from the post-registration
+  convergence reports (like that found in \suppl{section S1.3}).
+We released \regseg{} along with the tool to generate such convergence reports.
+
+\paragraph*{Cross-comparison and settings}
 A similar workflow to the general evaluation framework used for \autoref{fig:evworkflows}
@@
   perform registration.
+The deformation field is correspondingly restricted to the \gls*{pe} direction.
 Settings used with \emph{elastix} are conveniently available in its format\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/t2b_elastix_y.txt}}.

### 2015-09-15 Publications/2015-NeuroImage/06-discussion.tex
more changes, settings, contributions...

@@
 The hypothesis tested in our study was that reliable image registration can be performed
-  by searching for homogeneous regions in a multispectral image, which correspond to precise contours
-  from an atlas, or extracted from another image (i.e., a different time step).
+  by searching for the homogeneous regions defined by a set of nested surfaces in a
+  multispectral image
+The surfaces correspond to precise contours extracted from another image of the same subject
+  (i.e. a \gls*{t1} image, or a different time step).
 We demonstrated that active contours without edges can be used successfully to drive a
@@
   for the other shapes.
-We argue that the symmetry of those other shapes posed problem in driving the contours
+We argue that the symmetry of those other shapes posed difficulties in driving the contours
   towards the appropriate region and producing some ``sliding'' effect between the
-  faces of the surface and their ground-truth position.
+  faces of the surfaces and their ground-truth position.
 This effect should not be present in real datasets, thanks to the very convoluted cortical
@@
   of the subject.
-Current applications based on whole-brain tractography usually solve this problem with a two-step approach.
+Current pipelines using whole-brain tractography usually solve this problem with a two-step approach.
 First, the images are corrected for nonlinear distortions using auxiliary acquisitions
-  such as field maps \citep{jezzard_correction_1995}, \gls*{t2} images \citep{kybic_unwarping_2000}.
+  such as fieldmaps \citep{jezzard_correction_1995}, \glspl*{dwi} with reversed \gls*{pe}
+  blips \citep{chiou_simple_2000}, or \gls*{t2} images \citep{kybic_unwarping_2000}.
 Second, the segmentation is projected from a reference \gls*{t1} image using linear
@@
 This situation is found commonly in historical datasets.
+Moreover, since the structural information is projected into the native space of \gls*{dmri}
+  reconstruction and tractography can be performed without resampling data to an undistorted
+  space.
 

### 2015-09-30 Publications/2015-NeuroImage/responseLetter2.tex
fixes for version 1.2

@@
 
-\noindent New title: \emph{{\color{blue} Surface}-driven registration method for the structure-informed segmentation of
+\noindent New title: \emph{{\color{blue!70} Surface}-driven registration method for the structure-informed segmentation of
   diffusion MR images}.
@@
   along with the description of the corresponding changes in the manuscript.
+We have also changed the color of text for the modifications in the manuscript, along with side notes indicating which reviewer
+  and comment suggested the edition.
 
@@
 \end{reviewpoint}
-\reply{
-The first sentence of the comment is a rather accurate interpretation of our work.
-However, it is important to clarify that our method maps anatomically-correct surfaces of the same subject into dMRI space,
-  thereby segmenting this target space.
-The method works regardless how the surfaces were extracted, or if they were extracted from T2w images or manually delineated.
-In this work, the surfaces are extracted from T1w for two reasons: 1) it is the widely-accepted MR sequence for such extraction,
-  and 2) the structural information used in reconstruction, tractography and network definition of connectivity extraction pipelines
-  is usually defined in T1w-space as well.
-We understand that the manuscript was not clear to this respect, and therefore included the following changes:
-\begin{enumerate}
-\item Changed abstract, more oriented on the problem now.
-\item Re-structured introduction: lines 1-14 state the background and the problem more clearly, lines 15-27 review
-  the literature existing about direct segmentation of dMRI data and lines 28-55 describe the alternative approach
-  by registering structural images and the problem of EPI distortions.
-  The methodological background of active-contours and registration-segmentation methods has been relocated in
-    the beginning of section 2, lines 73-85.
-\end{enumerate}
-
-Our method can be seen as a segmentation framework using active contours without edges in which the evolution
-  of the contours is governed by an elastic registration method;
-or as a nonlinear registration method which cost function is built on top of a segmentation model of active contours
-  without edges.
-In our opinion, the presented approach is mainly a segmentation method.
-However, since the surfaces are supposed to be very accurate and defined in a different coordinate system,
-  the method implicitly registers that space to the target image.
-Some comments of Reviewer \#2 are very similar to this one (particularly \autoref{com:11}).
+\reply{%
+We decided to call the method ``simultaneous registration and segmentation'' because antecedent methods
+  using active contours that evolve driven by the optimization of a deformation field use that nomenclature.
+These methods are now summarized and discussed in lines 234--252 of the current version.
 }
@@
 We have defined clearly that the choice of surfaces to segment the target data is crucial, and define the
-  segmentation model that drives the registration process (lines 110-113 are new).
+  segmentation model that drives the registration process (lines 85--88 are new).
 We have also improved former Figure 4, and brought it earlier in the paper (now it is Figure 2).
-Now, the figure shows how the segmentation model works: A is the ideal segmentation, B shows the model in
-  the initialization settings, and C shows the model after using \emph{regseg}, demonstrating how the covariance
-  matrices for each region are reduced.
-Since the former visualization was a bit clumsy, we have extracted the distribution of diffusion features by region on
-  the right side of each row.
+Since the former visualization of the conditional (region-wise) joint distribution of FA and ADC was a bit clumsy,
+  we have extracted the conditional distributions to the right side of each panel.
 }
@@
 We agree with the reviewer on that the novelty of the methodological elements of our approach is limited.
-However, to our knowledge there exists no alternative method with a paired combination of the following elements:
-\begin{itemize}
-\item Region based (i.e. without edges) active contours model
-\item The use of surfaces as prior knowledge to initialize the segmentation model,
-  with property of allowing precise registration if the surface is accurately representative
-  of structures being aligned.
-  Moreover, the multi-phase definition of levelsets is dual to our surface-nesting model.
-\item The anisotropic regularization model of choice, since most of the levelset frameworks
-  use Gaussian smoothing of the deformation field.
-\item Even though the shape-gradients theory was proposed back in the early 2000's, atlas-based
-  registration, and registration-segmentation methods have been generally implemented on levelsets.
-\end{itemize}
-To our knowledge, the application of a framework as such is new and an important contribution is done
-  with the proposed segmentation model of the FA and ADC maps.
-We would have appreciated if the reviewers would have provided specific references being close to our approach
-  beyond those given in lines 43-57 of the original paper, so that we could point to the differences of our
-  method.
-For the sake of clarity, we have removed the methodological background from introduction to section 2, lines 73-85.
+However, to our knowledge there exists no alternative method combining the same elements to segment
+  multivariate data.
+We have moved the related methods and our methodological contributions from the Introduction to a more appropriate
+  place in the Discussion (ll. 234--252), additionally satisfying \autoref{com:9} suggesting to include in the introduction
+  only the references related to the focus of the paper.
 }
@@
 The reviewer is right, assuming that pixels are i.i.d. is wrong but a widely accepted simplification.
-We have supported the assumptions with new lines 105-107:
-\textit{Even though the features being segmented are not generally i.i.d., the spatial interdependency of voxels is
-implicitly supported by the piecewise smooth partition of the space $\boldsymbol\Omega$.
-In Figure 2 it is shown that marginal distributions of data can be approximated by multivariate normal distributions.}
-}
+We have supported the assumptions inserting the new lines 80--83.}
 
@@
 \reply{
-We have improved former Figure 4 (now, Figure 2) to clarify that the model strongly depends on the choice of surfaces.
-Adding more surfaces makes the model more robust to local minima at the cost of largely increased computation times.
-We agree with the reviewer on that Figure S1 is not illustrative of the metric for free-deformation fields, but
-  our intention was to show the performance of the method in realistic settings.
-This implies distortions with nonzero components on the phase-encoding axis (the $y$-axis in our experiments), and
-  therefore Figure S1 becomes representative since it samples the distortions space covering plausible deformations
-  for both encoding directions (Anterior-to-posterior and posterior-to-anterior).
-We have included information about the maximum distortion for which the metric was sampled.
+We have improved former Figure 4 (now, it is Figure 2) to clarify that the segmentation model is inherently defined by
+  the choice of surfaces and their nesting.
+This is an actual contribution of the work, as stated in ll. 264--266.
+
+The model behaves robustly when a) the target data are simple as in our phantom settings; and b) when the distortion
+  has only one nonzero component like in our real-data settings.
+The analysis of the goal function for the phantoms (without directional restrictions) was not necessary since
+  convergence was straigthforward.
+In the case of real data, since the distortion under investigation is aligned with one axis, the Figure S1 shows
+  that the goal function is well defined.
+
+However, the reviewer is right when raising convergence issues.
+We noticed that, when decimating the meshes in the phantom experiments, the zero of shape-gradients moved away from
+  the minimum of the goal function.
+Since decimation factors applied in this experiment were very large, and it did not affect our results with
+  the meshes directly obtained with \emph{Freesurfer} (in both phantoms and real data), we did include this discussion
+  within the paper.
+
+Finally, we commented on the initialization issue with the insertion of ll. 188--190.
 }
@@
 Some parameters are also discussed in the Supplemental Materials, Section S2.
+
 However, we understand that a better description of the parameter settings should be included within
-  the manuscript and modified it accordingly as follows: {\color{red} say what here}.
+  the manuscript and modified it accordingly, in ll. 192--196.
 }
@@
   very minimal.
-We have modified Figure 3 to reflect this aspect.
+We have modified former Figure 3 (now Fig. 4) to reflect this aspect.
 }
@@
 \reply{
-We have added the following in the Discussion section (lines 262-267):
-
-\textit{The distributions of errors along surfaces vary importantly depending on the shape of the
-  phantom (see Figure 5B).
-The misregistration error of the ``gyrus'' phantom showed a much lower spread than that for the other shapes.
-We argue that the symmetry of those other shapes posed problem in driving the contours towards the appropriate
-  region and producing some ``sliding'' effect between the faces of the surface and their ground-truth position.
-This effect should not be present in real datasets, thanks to the very convoluted cortical layer, and the
-  directional restriction of distortion.}
+We have added a discussion on this in the corresponding section, lines 286-291. Thanks for pointing out.
 }
@@
   for the real datasets.
-We have added lines 204-206 to clarify this point.
+We have added ll. 125--127 to indicate that the ground-truth in the experiments done with real data is directionally
+  restricted. Then, in lines 190-192 to clarify that \emph{regseg} is configured to fulfill this feature.
 }
@@
 \end{reviewpoint}
-\reply{Active contours without edges \citep{chan_active_2001} is a framework for subpixel segmentation of images in
-  which the contours evolve enclosing homogeneous regions inside.
-Region-based active contours is an alternative terminology.
-As previously described in \autoref{com:1} and \autoref{com:2}, we have made an effort to clearly state that the outputs
-  to our method are a set of nested surfaces and a target multivariate image, that for the case of dMRI data comprehends
-  FA and ADC.
-The manuscript has been re-phrased accordingly in the title, abstract and introduction.}
+\reply{The manuscript has been re-phrased accordingly in the title, abstract and introduction.}
 
@@
 \reply{
-The arrows in Figure 3, panel 5, are intended to show regions subjected to large deformations, where the contours
-  do not fit data before applying \emph{regseg}.
-They also happen to show areas of the forceps minor, one fiber bundle importantly affected
-  by the susceptibility distortions.
-We have modified the caption of Figure 3 to describe the meaning of the arrows correctly.
+We have modified the caption of former Figure 3 (now Fig. 4) to describe the meaning of the arrows correctly.
 }

### 2015-09-30 Publications/2015-NeuroImage/02-abstract.tex
fixes for version 1.2

@@
 % -*- root: 00-main.tex -*-
+\newcomment[RV\#1(C.1)]{%
 Current methods using diffusion MRI to map the microstructure and connectivity of the
@@
 Here we propose \regseg{}, which segments homogeneous regions within multivariate
-  images by projecting a set of nested surfaces extracted from structural images.
+  images by projecting a set of nested surfaces extracted from structural images.}
 The surfaces evolve through a free-form deformation field supported by the B-spline basis
@@
   randomly generated deformations, where subvoxel accuracy was achieved.
-We proposed a set of surfaces to segment the fractional anisotropy and the apparent diffusion
-  coefficient maps derived from diffusion MRI datasets.
+\newcomment[RV\#2(C.11)]{%
+We proposed a set of surfaces that define a segmentation model of the fractional anisotropy and
+  the apparent diffusion coefficient maps derived from diffusion MRI datasets.}
 These surfaces extracted from the same-subject T1-weighted image successfully segmented 16 real

### 2015-09-30 Publications/2015-NeuroImage/03-introduction.tex
fixes for version 1.2

@@
 It is generally acquired using \gls*{epi} schemes, since they are very fast at
-  scanning a large sequence of images called \glspl*{dwi}, each sensitized with
-  a gradient to probe proton diffusion in a certain orientation.
+  scanning a large sequence of images called \glspl*{dwi}.
+Each \gls*{dwi} is sensitized with a gradient to probe proton diffusion in a certain
+  \newcomment[RV\#2(C.13)]{orientation}.
 Subsequent processing involves describing the local microstructure with one available
@@
   to current models such as \citep{daducci_accelerated_2015}.
-The microstructural map is then used to draw the preferential orientations of diffusion
+The microstructural map is then used to draw the preferential \newcomment[RV\#2(C.13)]{orientations} of diffusion
   across the brain using tractography \citep{mori_threedimensional_1999}.
-Finally, the streamlines found are filtered with a cortical parcellation
-  to build the graph representing the corresponding structural
-  network \citep{hagmann_mapping_2008}.
-Since the proposal of the earliest methodologies to solve the three problems,
-  there is a need for precisely identifying the anatomy on the \gls*{dmri} space.
-In fact, current trends in reconstruction \citep{jeurissen_multitissue_2014} and
+Finally, a graph representing the corresponding structural network is built using
+  the regions of a cortical parcellation as nodes and the streamlines of
+  tractography as edges \citep{hagmann_mapping_2008}.
+\newcomment[RV\#2(C.17)]{Since the proposal of the earliest methodologies to solve the three problems,
+  there is a need for precisely identifying the anatomy in the \gls*{dmri} space.
+In fact, current trends on reconstruction \citep{jeurissen_multitissue_2014} and
   tractography \citep{smith_anatomicallyconstrained_2012} are increasingly using
-  structural information to improve the microstructural mapping and fiber-tracking.
+  structural information to improve the microstructural mapping and fiber-tracking.}
 
-The structural information required by pioneer tractography studies was a simple
+\newcomment[RV\#2(C.15)]{The structural information required by pioneer tractography studies was a simple
   \gls*{wm} mask to provide with a termination criteria for the streamlines.
@@
   rousson_level_2004,jonasson_segmentation_2005,liu_brain_2007,hadjiprocopis_unbiased_2005,
-  lu_segmentation_2008,han_experimental_2009}.
-However, the precise segmentation of \gls*{dmri} is difficult for several limitations.
-First, \gls{dmri} images have a resolution that is much lower than that of the imaged
+  lu_segmentation_2008,han_experimental_2009}.}
+\todo{RV\#2(C.14)}However, the precise segmentation of \gls*{dmri} is difficult for several limitations.
+First, \gls{dmri} images have a resolution that is \newcomment[RV\#2(C.12)]{much lower} than that of the imaged
   microstructural features.
@@
 
+\newcomment[RV\#2(C.9)]{%
 An alternative route to segmentation in native \gls*{dmri} space is the mapping of the
@@
   combinations of the original methods \citep{holland_efficient_2010,andersson_comprehensive_2012,
-  irfanoglu_drbuddi_2015}
+  irfanoglu_drbuddi_2015}.}\todo{RV\#2(C.18)}
 
-Here we show a method that registers anatomically-correct surfaces to multivariate images,
+Here we show a method that registers \newcomment[RV\#2(C.10)]{anatomically-correct surfaces to multivariate images},
   thereby implicitly segmenting the \gls*{dmri}-space into those structures enclosed by the contours

### 2015-09-30 Publications/2015-NeuroImage/04-methods.tex
fixes for version 1.2

--- 2015-09-08 Publications/2015-NeuroImage/04-methods.tex
@@
 % -*- root: 00-main.tex -*-
-\section{Methods}
-\label{sec:methods}
-\subsection{Related work}\label{sec:related_work}
-Previously, joint segmentation and registration have been applied successfully to other problems
-such as longitudinal object tracking \citep{paragios_level_2003} and atlas-based
-  segmentation \citep{gorthi_active_2011}.
-The most common approach to solve the problem simultaneously is optimizing a deformation model
-  (registration) that supports the evolution the active contours (segmentation), like
-  \citep{paragios_level_2003,yezzi_variational_2003}.
-\cite{unal_coupled_2005}, and later \cite{wang_joint_2006}, replaced the linear registration
-  transform of \citep{yezzi_variational_2003} with a free-form deformation field.
-\cite{droske_mumfordshah_2009} reviewed the existing techniques and proposed two different
-  approaches for applying the Mumford-Shah functional \citep{mumford_optimal_1989} during simultaneous
-  registration and segmentation by propagating the deformation field from
-  the contours onto the whole image definition.
-Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
-  registration method in 2D using level sets and a nonlinear elasticity smoother on the
-  displacement vector field, which preserves the topology even with very large deformations.
-Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
-  level set function to register several active contours during the application
-  of atlas-based segmentation.
+\section{Methods}\label{sec:methods}
 
@@
   \caption{The active contours are defined as the interfacing surfaces of the competing
-    \glspl{roi} $\Omega_i$, which
+    \glspl{roi} $\Omega_l$, which
   are represented in green and dark blue in this close-up.
@@
   for each region $\Omega_l$.
+\newcomment[RV\#1(C.4)]{%
 Even though the features being segmented are not generally i.i.d., the spatial interdependency of
@@
 In \autoref{fig:model} it is shown that marginal distributions of data can be approximated by
-  multivariate normal distributions. Therefore,
+  multivariate normal distributions.} Therefore,
 
@@
 $C$ is the number of channels comprised in the image $M$.
+\newcomment[RV\#1(C.2)]{%
 In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
@@
   this minimization is illustrated for the segmentation of the \gls*{fa} and the \gls*{adc} maps
-  of one subject.
+  of one subject.}
 
@@
 \caption{%
-Evolution of the segmentation model defined by the homogeneous regions $\Omega_i$
+\newcomment{%
+Evolution of the segmentation model defined by the homogeneous regions $\Omega_l$
   for a real dataset.
-Panel A shows how the \gls*{fa} and the \gls*{adc} are clustered by the surfaces mapped to
-  $M$ through the ground-truth.
-Panel B shows the same information at the initialization step of \regseg{}.
-Since the surfaces are misaligned in $M$, the spreads of the sampled distribution for
-  each region are larger than those in A.
+A) shows the initialization of the model, when the anatomically-correct surfaces are not
+  mapped into the \gls*{dmri} space.
+The panel shows the resulting conditional distributions of segmenting the (distorted) \gls*{fa} and
+  \gls*{adc} maps with the original (undistorted) surfaces.
+Panel B shows the equivalent information, using the surfaces with perfect matching with the distorted space
+  of \gls*{dmri}.
+Since the surfaces are now aligned in $M$, the spreads of the conditional distributions for each region are
+  reduced with respect to those in A.
 Panel C, after \regseg{} registration, shows how the covariance matrices $\boldsymbol{\Sigma}_l$
-  are minimized with respect to panel B, and the resulting segmentation is close
-  to the theoretical situation presented in panel A.
+  are minimized with respect to panel A, and the resulting segmentation is close
+  to the theoretical situation presented in panel B.
 The left (large) plot in each panel represents a kernel density estimate of the location and spread of
-  each region $\Omega_i$ in the bivariate feature space of $M$, which comprises
+  each region $\Omega_l$ in the bivariate feature space of $M$, which comprises
   the \gls*{fa} and \gls*{adc} maps.
-At the top and the right margins, the marginal distributions of $\Omega_i$ are also plotted for
+At the top and the right margins, the marginal distributions of $\Omega_l$ are also plotted for
   the \gls*{fa} and \gls*{adc}, respectively.
 On the right, the sampled distribution of each region is shown in small scatter plot, in contrast to
-  the complete bivariate data.
+  the complete bivariate data.}
 }\label{fig:model}
+\todo[inline]{RV\#1(C.5)}
 \end{figure}
@@
   following sections.
+\newcomment[RV\#2(C.20)]{%
 A first proof of concept to demonstrate \regseg{} in phantoms with simple
@@
 Then, a more realistic environment with $U_\text{true}$ restricted to the
-  \gls*{pe} axis of real \gls*{dmri} datasets is presented.
+  \gls*{pe} axis of real \gls*{dmri} datasets is presented.}
 
@@
  $\hat{\gammaset}_\text{test}$ and $\gammaset_\text{true}$ using \citep{commandeur_vtk_2011}.
+In total, 1200 experiments (four phantom types $\times$ 150 random warpings $\times$ two resolutions) were
+  performed according to the workflow illustrated in \autoref{fig:evphantoms}.
 
@@
 \end{figure*}
+
+\paragraph*{Segmentation model and settings}
+The segmentation model of the phantoms is implicitly defined in each shape.
+All phantoms comprehend an inner surface enclosing a uniform \gls*{wm}-like region,
+  and an outer surface wrapping a \gls*{gm}-like layer.
+The outside is filled with uniform background (see \autoref{fig:evphantoms}).
+All the experimental settings used for the phantoms are made available in
+  a unique configuration file\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/regseg_default.json}}.
 
@@
   5) The results were evaluated visually and quantitatively.
+  \newcomment{%
   The arrows point to edges in the target images (yellow arrows for \gls*{fa}, blue for \gls*{adc}) that
     should be aligned with a surface, showing how distortion breaks the mapping with the structural space in which
-    the contours are defined.}\label{fig:evworkflows}
+    the contours are defined.}}\label{fig:evworkflows}
+  \todo[inline]{RV\#2(C.23)}%
 \end{figure*}
@@
 
-\paragraph*{Cross-comparison}
-A similar workflow to the general evaluation framework used for \autoref{fig:evworkflows}
+\paragraph*{\Regseg{} settings} %
+\newcomment[RV\#1(C.5)]{%
+\Regseg{} accepts an affine mapping from surface-space to the \gls*{dmri} data as initialization.
+However, the images provided by the \gls*{hcp} are already spatially normalized.
+Therefore, the initial estimation of distortion is zero in this experiment.}
+\newcomment[RV\#1(C.20)]{%
+Since the distortion $U_\text{true}$ is aligned along the \gls*{pe} direction ($y$-axis in our settings),
+  \regseg{} was configured to allow nonzero displacements only on that corresponding direction.}
+\newcomment[RV\#1(C.6)]{%
+For the experiments on real data, \regseg{} established a multi-resolution pyramid of
+  B-Spline basis, with control points distributed on grids of the following spacings:
+  40$\times$100$\times$40 [mm] for the first (coarser level), \isores[mm]{30} for the second,
+  and 20$\times$30$\times$10 [mm] in the third.
+Only the first and second levels included Gaussian smoothing of the target image ($\sigma$ = [2.0, 0.5] mm,
+  respectively).
+The actual choices of the parameter settings are publicly distributed with the source code for the
+  experiments\footnote{\url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data/regseg_hcp.json}}.}
+These settings were obtained manually, driven by the feedback obtained from the post-registration
+  convergence reports (like that found in \suppl{section S1.3}).
+We released \regseg{} along with the tool to generate such convergence reports.
+
+\paragraph*{Cross-comparison and settings} %
+\newcomment[RV\#2(C.21)]{%
+A dual workflow to the general evaluation framework used for \regseg{} (\autoref{fig:evworkflows}),
   was employed to integrate the alternate \gls*{t2b} registration scheme.
@@
   perform registration.
-Settings used with \emph{elastix} are conveniently available in its format\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/t2b_elastix_y.txt}}.
+The deformation field is correspondingly restricted to the \gls*{pe} direction.
+Settings used with \emph{elastix} are conveniently available in its format\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/t2b_elastix_y.txt}}.}
 

### 2015-09-30 Publications/2015-NeuroImage/05-results.tex
fixes for version 1.2

--- 2015-05-04 Publications/2015-NeuroImage/05-results.tex
@@
 
-\begin{figure*}
-  \centering
-  \includestandalone[width=\linewidth]{figure05}
-  \caption{A. Visual assessment of the results obtained with the low resolution sets:
-    ``gyrus'' (top left), ``L'' (top right), ``ball'' (bottom left),
-    and ``box'' at (bottom right).
-  The contours recovered after registration are represented in yellow.
-  \Regseg{} achieved high accuracy because it determined the almost exact locations of the registered
-    contours with respect to their ground truth positions (shown in green).
-  The partial volume effect makes segmentation of the sulci a challenging problem with voxel-wise
-    clustering methods, but they were successfully segmented with \regseg{}.
-  B. Quantitative evaluation of registration errors in terms of the average Hausdorff distances between
-    surfaces at low (left) and high (right) resolutions, which demonstrate that the errors were
-    consistently below the size of the voxels.
-    }\label{fig:phantom}
-\end{figure*}
-In total, 1200 experiments (four phantom types $\times$ 150 random warpings $\times$ two resolutions) were
-  performed according to the workflow illustrated in \autoref{fig:evphantoms}.
-For each experiment, the misregistration error was measured using the Hausdorff distance
-  (see \autoref{sec:experiments_evaluation}) between the theoretical $\gammaset_\text{true}$ and
-  the estimate obtained by \regseg{} ($\hat{\gammaset}_{test}$).
 The results demonstrated that the accuracy was consistent and high, and below the image resolution.
@@
 
-\begin{table}
+\begin{figure*}[!h]
   \centering
-    \footnotesize
-    \tabcolsep=0.1cm
-    \begin{tabular}{lccccc}
-    Res.   & ``gyrus'' & ``ball''  & ``box''   & ``L''     & Aggreg.    \\\hline
-    1.0mm  & 0.18-0.38 & 0.31-0.45 & 0.34-0.42 & 0.34-0.40 & 0.34-0.38  \\
-    2.0mm  & 0.59-0.60 & 0.65-0.76 & 0.68-0.71 & 0.67-0.77 & 0.64-0.66  \\
-    \hline
-    \end{tabular}
-    \caption{The distributions of vertex-wise Hausdorff distances between the ground-truth surfaces and their
-    corresponding estimates obtained with \regseg{} presented a 95\% CI below the half-voxel size for all of 
-    the phantom types.
-    The CIs were computed by bootstrapping using 10$^4$ samples, with the median as the location statistic.}\label{tab:ci_phantom}
-\end{table}
+  \includestandalone[width=\linewidth]{figure05}
+  \caption{A. Visual assessment of the results obtained with the low resolution sets:
+    ``gyrus'' (top left), ``L'' (top right), ``ball'' (bottom left),
+    and ``box'' at (bottom right).
+  The contours recovered after registration are represented in yellow.
+  \Regseg{} achieved high accuracy because it determined the almost exact locations of the registered
+    contours with respect to their ground truth positions (shown in green).
+  The partial volume effect makes segmentation of the sulci a challenging problem with voxel-wise
+    clustering methods, but they were successfully segmented with \regseg{}.
+  B. Quantitative evaluation of registration errors in terms of the average Hausdorff distances between
+    surfaces at low (left) and high (right) resolutions, which demonstrate that the errors were
+    consistently below the size of the voxels.
+    }\label{fig:phantom}
+\end{figure*}
+\begin{figure*}[!h]
+  \centering
+  \includestandalone[width=\linewidth]{figure06}
+  \caption{A. Example of a visual assessment report, which was generated automatically by the evaluation tool.
+    Each view shows one component of the input image (in this case, the \gls*{fa} map), the ground-truth locations
+    of the surfaces (green contours), and the resulting surfaces obtained with the test method (yellow contours).
+  The first two rows show axial slices for \regseg{} and the \acrfull*{t2b} method, while the last two rows
+    show the corresponding sagittal views.
+  The coronal view is omitted because it was the least informative due to the directional property
+    of the distortions.
+  Specific regions where \regseg{} outperformed \gls*{t2b} are enlarged.
+  B. Violin plots of the error distributions for each surface, which show the voxel size of the \gls*{dmri} images
+    (1.25 mm), thereby supporting the improved results obtained by \regseg{} with the proposed settings.
+  }\label{fig:results_real}
+\end{figure*}
 
@@
 \footnotesize
-    \tabcolsep=0.08cm
+\tabcolsep=0.1cm
+\begin{minipage}{.48\textwidth}
+    \vskip-4.7ex
+    \begin{tabular}{lccccc}
+    Res.   & ``gyrus'' & ``ball''  & ``box''   & ``L''     & Aggreg.    \\\hline
+    1.0mm  & 0.18-0.38 & 0.31-0.45 & 0.34-0.42 & 0.34-0.40 & 0.34-0.38  \\
+    2.0mm  & 0.59-0.60 & 0.65-0.76 & 0.68-0.71 & 0.67-0.77 & 0.64-0.66  \\
+    \hline
+    \end{tabular}
+    \caption{The distributions of vertex-wise Hausdorff distances between the ground-truth surfaces and their
+    corresponding estimates obtained with \regseg{} presented a 95\% CI below the half-voxel size for all of
+    the phantom types.
+    The CIs were computed by bootstrapping using 10$^4$ samples, with the median as the location statistic.}\label{tab:ci_phantom}
+\end{minipage}
+\hfill
+\begin{minipage}{.48\textwidth}
     \begin{tabular}{cccccc}
@@
     \multirow{2}{*}{CI}
-       & \regseg{}        & 0.50 - 0.78 & 0.50 - 0.55 & 0.66 - 0.73 & 0.56 - 0.66 \\
-       & T2B                  & 1.78 - 2.58 & 1.94 - 2.36 & 2.16 - 2.58 & 2.05 - 2.39 \\
+       & \regseg{} & 0.50--0.78 & 0.50--0.55 & 0.66--0.73 & 0.56--0.66 \\
+       & T2B       & 1.78--2.58 & 1.94--2.36 & 2.16--2.58 & 2.05--2.39 \\
     \hline
     \multirow{2}{*}{H-tests}
-       & p-value  & 4.1$\times$10$^{-6}$& 2.3$\times$10$^{-6}$& 2.3$\times$10$^{-6}$ & 1.8$\times$10$^{-16}$ \\
+       & p-value  & 4.1\e{-6} & 2.3\e{-6} & 2.3\e{-6} & 1.8\e{-16} \\
        & H-stat   & 21.20 & 22.31 & 22.31 & 67.85 \\
@@
     }\label{tab:results_real}
+\end{minipage}
 \end{table}
-
-\begin{figure*}
-  \centering
-  \includestandalone[width=\linewidth]{figure06}
-  \caption{A. Example of a visual assessment report, which was generated automatically by the evaluation tool.
-    Each view shows one component of the input image (in this case, the \gls*{fa} map), the ground-truth locations
-    of the surfaces (green contours), and the resulting surfaces obtained with the test method (yellow contours).
-  The first two rows show axial slices for \regseg{} and the \acrfull*{t2b} method, while the last two rows
-    show the corresponding sagittal views.
-  The coronal view is omitted because it was the least informative due to the directional property
-    of the distortions.
-  Specific regions where \regseg{} outperformed \gls*{t2b} are enlarged.
-  B. Violin plots of the error distributions for each surface, which show the voxel size of the \gls*{dmri} images
-    (1.25 mm), thereby supporting the improved results obtained by \regseg{} with the proposed settings.
-  }\label{fig:results_real}
-\end{figure*}

### 2015-09-30 Publications/2015-NeuroImage/06-discussion.tex
fixes for version 1.2

@@
 \label{sec:discussion}
+
+\newcomment[RV\#1(C.3)]{%
+\paragraph*{Contributions}\label{sec:related_work} %
+Previously, joint segmentation and registration have been applied successfully to other problems
+such as longitudinal object tracking \citep{paragios_level_2003} and atlas-based
+  segmentation \citep{gorthi_active_2011}.
+The most common approach to solve the problem simultaneously is optimizing a deformation model
+  (registration) that supports the evolution the active contours (segmentation), like
+  \citep{paragios_level_2003,yezzi_variational_2003}.
+\cite{unal_coupled_2005}, and later \cite{wang_joint_2006}, replaced the linear registration
+  transform of \cite{yezzi_variational_2003} with a free-form deformation field.
+\cite{droske_mumfordshah_2009} reviewed the existing techniques and proposed two different
+  approaches for applying the Mumford-Shah functional \citep{mumford_optimal_1989} during simultaneous
+  registration and segmentation by propagating the deformation field from
+  the contours onto the whole image definition.
+Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
+  level set function to register several active contours during the application
+  of atlas-based segmentation.
+Unlike the aforementioned registration-and-segmentation methods, \regseg{} uses shape-gradients
+  \citep{herbulot_segmentation_2006} computed on a hierarchical set of explicit surfaces
+  (triangularized meshes) that substitute the multiphase level sets.
+Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
+  registration method in 2D using level sets and a nonlinear elasticity smoother on the
+  displacement vector field, which preserves the topology even with very large deformations.
+\Regseg{} includes an anisotropic regularizer for the displacements field proposed by
+  \cite{nagel_investigation_1986}.
+This regularization strategy conceptually falls in the midway between the Gaussian smoothing
+  generally included in most of the existing methodologies, and the complexity of
+  the elasticity smoother of \cite{guyader_combined_2011}.
+
+An important antecedent of \regseg{} is \emph{bbregister} \citep{greve_accurate_2009}.
+The tool has been widely adopted as the standard registration method to be used along with the \gls*{epi}
+  correction of choice.
+It implements a linear mapping, and uses 3D active contours with edges\footnote{A three-dimensional active contour
+  is a surface and it is ``with edges'' when the contours evolve in search of abrupt intensity steps
+  (the edges) within the target image.} to search for intensity boundaries in the \lowb{} image.
+The active contours are initialized using surfaces extracted from the \gls*{t1} using
+  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}: the \emph{pial} surface (exterior of the
+  cortical \gls*{gm}) and the \emph{white} surface (the \gls*{wm}/\gls*{gm} interface).
+The \lowb{} image only includes a detectable frontier for the pial surface, and thus
+  \emph{bbregister} is limited to aligning the cortical layer in this
+  application.
+To overcome the problem of nonlinear distortions, \emph{bbregister} excludes the
+  regions that are typically warped from the boundary search.
+Indeed, the distortion must be addressed separately because it is not supported by
+  the affine transformation model.
+\Regseg{} can be seen as a derivative of \emph{bbregister} for the use of active surfaces.
+However, the deformation model is nonlinear and the surfaces in \regseg{} look for the
+  homogeneity within the regions enclosed, instead of image edges.}
+
+\newcomment[RV\#1(C.5)]{%
+Finally, we also contribute proposing a piecewise-smooth segmentation model defined by
+  a selection of nested surfaces to partition the multispectral space
+  comprehending the \gls*{fa} and the \gls*{adc} maps and ultimately identify these
+  structures in \gls*{dmri} space.}
+With the definition of appropriate models, \regseg{} may be applied in other fields like
+  neonatal brain image segmentation in longitudinal \gls*{mri} studies like \citep{shi_neonatal_2010}
+  in which the surfaces obtained in a mature timepoint of the brain can be retrospectively
+  propagated to the initial timepoints, regardless the changes of the contrast and spatial
+  development between timepoints.
 
@@
   voxel spacing.
+\newcomment[RV\#1(C.8)]{%}
 The distributions of errors along surfaces vary importantly depending on the shape of the
@@
 This effect should not be present in real datasets, thanks to the very convoluted cortical
-  layer, and the directional restriction of distortion.
+  layer, and the directional restriction of distortion.}
 

### 2015-09-30 Publications/2015-NeuroImage/figure02-model.tex
fixes for version 1.2

@@
 % @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-09-08 13:12:00
+% @Last Modified time: 2015-09-29 15:31:02
 
@@
   \node[inner sep=10pt](fig01a) at (0,15.5)
+    {\hspace*{20pt}\includegraphics[width=20cm]{figure04B}};
+  \node[inner sep=10pt](fig01b) at (0,7.8)
     {\hspace*{20pt}\includegraphics[width=20cm]{figure04A}};
-  \node[inner sep=10pt](fig01b) at (0,7.8)
-    {\hspace*{20pt}\includegraphics[width=20cm]{figure04B}};
   \node[inner sep=10pt](fig01c) at (0,0)

### 2015-10-01 Publications/2015-NeuroImage/04-methods.tex
updates before sending @zosso and @daducci

@@
 For convenience and because it has been shown to be an appropriate approximation
-  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
+  \citep{leemput_automated_1999a,cuadra_comparison_2005}, we introduce two assumptions for each
   region $\Omega_l$:

### 2015-10-01 Publications/2015-NeuroImage/responseLetter2.tex
fixed typos, added comment labels in manuscript

@@
 The analysis of the goal function for the phantoms (without directional restrictions) was not necessary since
-  convergence was straigthforward.
+  convergence was straightforward.
 In the case of real data, since the distortion under investigation is aligned with one axis, the Figure S1 shows
@@
 The \emph{regseg} tool is designed to allow any displacement, as demonstrated with the phantoms.
-However, as the reviewer points out, the deformation field is completelly restricted to the phase-encode direction
+However, as the reviewer points out, the deformation field is completely restricted to the phase-encode direction
   for the real datasets.
@@
 \end{reviewpoint}
-\reply{The manuscript has been re-phrased accordingly in the title, abstract and introduction.}
+\reply{%
+In our adaptation of ``active contours without edges'' \citep{chan_active_2001}, the active contours are
+  ``extracted from structural images''.
+Then, they are ``without edges'' because the target features do not show clear intensity steps
+  where the contours are intended to be finally placed.
+In practice, active contours are ``without edges'' when they look for homogeneous regions enclosed
+  by the contours, as is the case of the FA and ADC.
+We agree with the reviewer on that this phrase is not clear, and therefore reformulated the abstract.
+We have also modified the introduction to state more clearly the actual modalities that input to \emph{regseg}.
+}
 
@@
 \end{reviewpoint}
-\reply{We hope this point is answered in \autoref{com:10}, and tryied to avoid this concept
+\reply{We hope this point is answered in \autoref{com:10}, and tried to avoid this concept
   coming from the 2D processing field.

### 2015-10-01 Publications/2015-NeuroImage/02-abstract.tex
fixed typos, added comment labels in manuscript

@@
 % -*- root: 00-main.tex -*-
-\newcomment[RV\#1(C.1)]{%
+\newcomment[RV\#1(C.1)\par RV\#2(C.10)]{%
 Current methods using diffusion MRI to map the microstructure and connectivity of the

### 2015-10-01 Publications/2015-NeuroImage/05-results.tex
fixed some numbers with text style

@@
 All of the distributions of errors in the tests (four phantom types $\times$ two resolutions) were
-  non-normal with $p<0.001$.
+  non-normal with $p<$ 0.001.
 Consequently, we used the non-parametric Wilcoxon signed-rank test with the Bonferroni
-  correction for multiple comparisons ($N=150$, for each phantom type).
-The average errors were significantly lower than the voxel size with $p < (0.001 / 150)$
+  correction for multiple comparisons ($N$=150, for each phantom type).
+The average errors were significantly lower than the voxel size with $p <$ (0.001 / 150)
   in all tests (four phantom types $\times$ two resolutions).
@@
     the phantom types.
-    The CIs were computed by bootstrapping using 10$^4$ samples, with the median as the location statistic.}\label{tab:ci_phantom}
+    The CIs were computed by bootstrapping using 10$^\text{4}$ samples, with the median as the location statistic.}\label{tab:ci_phantom}
 \end{minipage}
@@
       and the aggregate of all surfaces (Aggreg. column) had lower 95\% CIs with \regseg{}.
-   The CIs in this table were computed by bootstrapping using the mean as the location statistic and with 10$^4$ samples.
+   The CIs in this table were computed by bootstrapping using the mean as the location statistic and with 10$^\text{4}$ samples.
     The Kruskal-Wallis H-tests indicated that there was a significant difference between the results obtained using \regseg{} and

### 2015-10-01 Publications/2015-NeuroImage/06-discussion.tex
fixed some numbers with text style

@@
   below the pixel size of 1.25 mm.
-The misregistration error that remained after \regseg{} was significantly lower ($p < 0.01$) than the
+The misregistration error that remained after \regseg{} was significantly lower ($p <$ 0.01) than the
   error corresponding to the \gls*{t2b} correction according to Kruskal-Wallis H-tests

### 2015-10-01 Publications/2015-NeuroImage/04-methods.tex
fix labels with prefix

@@
 % -*- root: 00-main.tex -*-
-\section{Methods}\label{sec:methods}
+\section{Methods}\label{sec:regseg-methods}
 
-\subsection{Registration framework and segmentation model}\label{sec:methods_map}
+\subsection{Registration framework and segmentation model}\label{sec:regseg-methods_map}
 Let $\gammaset_R = \{\Gamma_m: m \in \mathbb{N}, m \leq N_S\}$ be the set of $N_S$ surfaces
@@
   \vec{r} &\mapsto \vec{r}' =\vec{r}+\vec{u}(\vec{r}),
-  \label{eq:transform}
+  \label{eq:regseg-transform}
   \end{align}
@@
 The general overview of how the surfaces interact with the registration framework
-  is presented in \autoref{fig:method}.
+  is presented in \autoref{fig:regseg-method}.
 
@@
     the total area of surface $\Gamma_0$.
-      }\label{fig:method}
+      }\label{fig:regseg-method}
 \end{figure}
@@
 In a Bayesian framework for registration \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014},
-  the mappings $U$ in \eqref{eq:transform} are
+  the mappings $U$ in \eqref{eq:regseg-transform} are
   evaluated based on their posterior probability given the observed data
@@
   P(U \mid M, \omegaset ) = \frac{P(M \mid U,\omegaset )\, P(U)}{P(M)},
-  \label{eq:bayes_rule}
+  \label{eq:regseg-bayes_rule}
   \end{equation}
@@
     P\left( \vec{f}' \mid U \right),
-  \label{eq:bayes_aposteriori}
+  \label{eq:regseg-bayes_aposteriori}
   \end{equation}
   where $\vec{f}' = M(\vec{r}')$ is the feature vector at the displaced
-  position $\vec{r}'$ \eqref{eq:transform} in the moving image.
+  position $\vec{r}'$ \eqref{eq:regseg-transform} in the moving image.
 For convenience and because it has been shown to be an appropriate approximation
@@
   voxels is implicitly supported by the piecewise smooth partition of the space \omegaset{}.
-In \autoref{fig:model} it is shown that marginal distributions of data can be approximated by
+In \autoref{fig:regseg-model} it is shown that marginal distributions of data can be approximated by
   multivariate normal distributions.} Therefore,
@@
   \mdist{f'}{l} \right)}},
-  \label{eq:pdf}
+  \label{eq:regseg-pdf}
   \end{align}
@@
 In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
-  the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized. In \autoref{fig:model},
+  the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized. In \autoref{fig:regseg-model},
   this minimization is illustrated for the segmentation of the \gls*{fa} and the \gls*{adc} maps
@@
   the complete bivariate data.}
-}\label{fig:model}
+}\label{fig:regseg-model}
 \todo[inline]{RV\#1(C.5)}
@@
   p_1(\vec{u}) &= \mathcal{N}(  \nabla \vec{u} \mid 0, \mathbf{B}^{-1}),
-  \label{eq:thikonov}
+  \label{eq:regseg-thikonov}
   \end{align}
@@
   \const + \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, +   \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
-  \label{eq:energy}
+  \label{eq:regseg-energy}
   \end{align}
@@
 \subsection{Numerical Implementation}
-\label{sec:numerical_implementation}
+\label{sec:regseg-numerical_implementation}
 
-\paragraph*{Deformation model}\label{sec:deformation_model}
+\paragraph*{Deformation model}\label{sec:regseg-deformation_model}
 Since the vertices of the surfaces $\{\vec{v}_i: \vec{v}_i \subset \gammaset \}_{i=1 \ldots N_V}$
@@
   $\{\vec{u}_k\}_{k=1 \ldots K}$.
-Densification is achieved using a set of associated basis functions $\psi_k$ \eqref{eq:nodes_tfm}.
+Densification is achieved using a set of associated basis functions $\psi_k$ \eqref{eq:regseg-nodes_tfm}.
 In our implementation, $\psi_k$ is a tensor-product B-spline kernel of degree three.
@@
   \vec{v}_i' = \vec{v}_i + \vec{u}_i = \vec{v}_i + \sum_k \psi_k(\vec{r}) \: \vec{u}_k.
-  \label{eq:nodes_tfm}
+  \label{eq:regseg-nodes_tfm}
   \end{equation}
@@
 \paragraph*{Optimization}
-\label{sec:gradient_descent}
-To find the minimum of the energy functional \eqref{eq:energy},
+\label{sec:regseg-gradient_descent}
+To find the minimum of the energy functional \eqref{eq:regseg-energy},
   we propose a gradient-descent approach with respect to the underlying
@@
   \frac{\partial \vec{u}(\vec{r},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}_k},
-  \label{eq:general_gradient_descent}
+  \label{eq:regseg-general_gradient_descent}
   \end{equation}
@@
 Let us assume that the anisotropy is aligned with the imaging axes to simplify
-  \eqref{eq:energy} as expression \eqref{eq:app_energy} in \ref{app:reg_term},
-  and thus to compute its derivative \eqref{eq:general_gradient_descent}:
+  \eqref{eq:regseg-energy} as expression \eqref{eq:regseg-app_energy} in \ref{app:reg_term},
+  and thus to compute its derivative \eqref{eq:regseg-general_gradient_descent}:
 
@@
   \Big\},
-  \label{eq:gradient_descent}
+  \label{eq:regseg-gradient_descent}
   \end{align}
@@
   derivatives.
-The computation of the explicit shape gradients at each $\vec{v}_i'$ is illustrated in \autoref{fig:method}.
-Then, \eqref{eq:gradient_descent} can be reformulated as (see \suppl{equations SM6-SM12}):
+The computation of the explicit shape gradients at each $\vec{v}_i'$ is illustrated in \autoref{fig:regseg-method}.
+Then, \eqref{eq:regseg-gradient_descent} can be reformulated as (see \suppl{equations SM6-SM12}):
 
@@
   \vec{g}_k  + \boldsymbol{\alpha} \cdot \vec{u}_k - \boldsymbol{\beta} \cdot (\Delta \vec{u}_k).
-  \label{eq:final_gradient}
+  \label{eq:regseg-final_gradient}
   \end{align}
@@
                     {\mathcal{F}\{(\delta^{-1} + \boldsymbol{\alpha})\, I - \boldsymbol{\beta}\Delta\}} \right\},
-  \label{eq:update_equation}
+  \label{eq:regseg-update_equation}
   \end{align}
@@
 \paragraph*{Implementation details and convergence}
-\label{sec:conv_report}
+\label{sec:regseg-conv_report}
 The \regseg{} tool includes a multiresolution strategy on the free-form deformation field.
@@
 
-\subsection{Evaluation protocol}\label{sec:evaluation_protocol}
+\subsection{Evaluation protocol}\label{sec:regseg-evaluation_protocol}
 In order to assess the performance of \regseg{}, we defined the following general
@@
 
-\subsection{Simulated phantoms}\label{sec:digital_phantoms}
+\subsection{Simulated phantoms}\label{sec:regseg-digital_phantoms}
 The workflow required to simulate the digital phantoms and to assess the performance of
-  \regseg{} with them is presented in \autoref{fig:evphantoms}.
+  \regseg{} with them is presented in \autoref{fig:regseg-evphantoms}.
 A set of four binary objects (i.e., ``box'', ``ball'', ``L'',
@@
 In total, 1200 experiments (four phantom types $\times$ 150 random warpings $\times$ two resolutions) were
-  performed according to the workflow illustrated in \autoref{fig:evphantoms}.
+  performed according to the workflow illustrated in \autoref{fig:regseg-evphantoms}.
 
@@
   1) The reference surfaces $\gammaset_R$ are triangularized meshes extracted from the four binary shapes (i.e., ``box'', ``ball'', ``L'', ``gyrus'').
-  2) A ground-truth displacement field was generated as described in \autoref{sec:digital_phantoms}, and applied to warp
+  2) A ground-truth displacement field was generated as described in \autoref{sec:regseg-digital_phantoms}, and applied to warp
       $\gammaset_R$, thereby obtaining $\gammaset_\text{true}$.
@@
      visually using automatically generated visual reports and quantitativelly with the Hausdorff distance between the
-     corresponding surfaces.}\label{fig:evphantoms}
+     corresponding surfaces.}\label{fig:regseg-evphantoms}
 \end{figure*}
@@
   and an outer surface wrapping a \gls*{gm}-like layer.
-The outside is filled with uniform background (see \autoref{fig:evphantoms}).
+The outside is filled with uniform background (see \autoref{fig:regseg-evphantoms}).
 All the experimental settings used for the phantoms are made available in
@@
 
-\subsection{Real datasets} \label{sec:human_connectome}
-The general experimental framework for the real datasets is presented in \autoref{fig:evworkflows},
+\subsection{Real datasets} \label{sec:regseg-human_connectome}
+The general experimental framework for the real datasets is presented in \autoref{fig:regseg-evworkflows},
 which extends our previous evaluation \citep{esteban_simulationbased_2014} of distortions
@@
   2) For use as the ground truth, we generated a plausible synthetic distortion $U_\text{true}$
-    from the field map with \eqref{eq:fieldmap}.
+    from the field map with \eqref{eq:regseg-fieldmap}.
   3) The \gls*{dmri} data were warped using $U_\text{true}$ to reproduce the effects of real
@@
     should be aligned with a surface, showing how distortion breaks the mapping with the structural space in which
-    the contours are defined.}}\label{fig:evworkflows}
+    the contours are defined.}}\label{fig:regseg-evworkflows}
   \todo[inline]{RV\#2(C.23)}%
@@
   reference surfaces.
-The segmentation model corresponding to this partition is shown in \autoref{fig:model}
+The segmentation model corresponding to this partition is shown in \autoref{fig:regseg-model}
   and discussed in greater detail in \suppl{section S4}.
@@
   u_\text{PE} = \frac{\gamma \, T_{acq}\, s_\text{PE}}{2\pi}\Delta B_0(\vec{r})\text{ [mm]},
-  \label{eq:fieldmap}
+  \label{eq:regseg-fieldmap}
   \end{equation}
@@
 The field map was unwrapped\footnote{fieldmaps are phase maps, which are intrinsically clipped in the interval
-  of $[-\pi, \pi)$ [rads] or [rads/s].} and smoothed before applying \eqref{eq:fieldmap}.
+  of $[-\pi, \pi)$ [rads] or [rads/s].} and smoothed before applying \eqref{eq:regseg-fieldmap}.
 Next, the original \gls*{dmri} was warped using the resulting displacement field and fed into
@@
   $\hat{U} = \epsilon \cdot U_\text{true} = \vec{r} + \epsilon \, u_\text{PE}$
-  (with $\epsilon \in [-1.1, 1.1]$ and $u_\text{PE}$ from \eqref{eq:fieldmap}),
-  and we evaluated the data term of the cost function \eqref{eq:energy}.
-The minimum of the cost function (\autoref{sec:methods_map}) was consistently located at
+  (with $\epsilon \in [-1.1, 1.1]$ and $u_\text{PE}$ from \eqref{eq:regseg-fieldmap}),
+  and we evaluated the data term of the cost function \eqref{eq:regseg-energy}.
+The minimum of the cost function (\autoref{sec:regseg-methods_map}) was consistently located at
   $\epsilon=0.0$ (the ground-truth) for all of the cases (\suppl{Figure S2}).
@@
 \newcomment[RV\#2(C.21)]{%
-A dual workflow to the general evaluation framework used for \regseg{} (\autoref{fig:evworkflows}),
+A dual workflow to the general evaluation framework used for \regseg{} (\autoref{fig:regseg-evworkflows}),
   was employed to integrate the alternate \gls*{t2b} registration scheme.
@@
 
-\paragraph*{Error measurement}\label{sec:experiments_evaluation}
+\paragraph*{Error measurement}\label{sec:regseg-experiments_evaluation}
 Distortion only occurs along the \gls*{pe} axis of the image, so we computed the
@@
   \vec{v}_i - \hat{\vec{v}}_i \|,
-  \label{eq:swindex}
+  \label{eq:regseg-swindex}
   \end{equation}

### 2015-10-01 Publications/2015-NeuroImage/05-results.tex
fix labels with prefix

@@
 \section{Results}
-\label{sec:results}
+\label{sec:regseg-results}
 
 \subsection{Proof of concept using digital phantoms}
-\label{sec:results_phantom}
+\label{sec:regseg-results_phantom}
 
 The results demonstrated that the accuracy was consistent and high, and below the image resolution.
-\autoref{fig:phantom} (block C) shows the violin plots for each model type corresponding
+\autoref{fig:regseg-phantom} (block C) shows the violin plots for each model type corresponding
   to the two sets of resolutions for the generated phantoms.
@@
     consistently below the size of the voxels.
-    }\label{fig:phantom}
+    }\label{fig:regseg-phantom}
 \end{figure*}
@@
     (1.25 mm), thereby supporting the improved results obtained by \regseg{} with the proposed settings.
-  }\label{fig:results_real}
+  }\label{fig:regseg-results_real}
 \end{figure*}
 
-\subsection{Evaluation using real datasets and cross-comparison}\label{sec:results_hcp}
+\subsection{Evaluation using real datasets and cross-comparison}\label{sec:regseg-results_hcp}
 Finally, we compared the performance of \regseg{} with that of the standard \gls*{t2b}
@@
 Summary reports for visual assessment of the 16 cases are included in \suppl{section S5}.
-In \autoref{fig:results_real}, box A, the visual report is shown for one subject.
-We computed the \gls*{swindex} \eqref{eq:swindex} of every surface after registration
+In \autoref{fig:regseg-results_real}, box A, the visual report is shown for one subject.
+We computed the \gls*{swindex} \eqref{eq:regseg-swindex} of every surface after registration
   using both the \regseg{} and \gls*{t2b} methods.
@@
   (a non-parametric alternative to ANOVA) on the warping indices for the three surfaces of
-  interest selected in \autoref{sec:experiments_evaluation}
+  interest selected in \autoref{sec:regseg-experiments_evaluation}
   ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$).
@@
   \gls*{t2b} were significantly different, and the violin plots in box B of
-  \autoref{fig:results_real} demonstrate that the errors were always larger with \gls*{t2b}.
+  \autoref{fig:regseg-results_real} demonstrate that the errors were always larger with \gls*{t2b}.
 We also show the 95\% CIs of the \gls*{swindex} for these surfaces.

### 2015-10-01 Publications/2015-NeuroImage/06-discussion.tex
fix labels with prefix

@@
 \section{Discussion}
-\label{sec:discussion}
+\label{sec:regseg-discussion}
 
 \newcomment[RV\#1(C.3)]{%
-\paragraph*{Contributions}\label{sec:related_work} %
+\paragraph*{Contributions}\label{sec:regseg-related_work} %
 Previously, joint segmentation and registration have been applied successfully to other problems
@@
 The distributions of errors along surfaces vary importantly depending on the shape of the
-  phantom (see \autoref{fig:phantom}B).
+  phantom (see \autoref{fig:regseg-phantom}B).
 The misregistration error of the ``gyrus'' phantom showed a much lower spread than that
@@
 We evaluated \regseg{} in a real environment using the experimental framework presented
-  in \autoref{fig:evworkflows}.
+  in \autoref{fig:regseg-evworkflows}.
 We processed 16 subjects from the \gls*{hcp} database using both \regseg{}
@@
 Visual inspections of all the results (\suppl{section S5}) and the violin plots in
-  \autoref{fig:results_real} confirmed that \regseg{} performed better than the \gls*{t2b} method
+  \autoref{fig:regseg-results_real} confirmed that \regseg{} performed better than the \gls*{t2b} method
   in our settings.

### 2015-10-01 Publications/2015-NeuroImage/07-conclusion.tex
fix labels with prefix

@@
 \section*{Conclusion}
-\label{sec:conclusion}
+\label{sec:regseg-conclusion}
 
@@
 \section*{Availability and reproducibility statement}
-\label{sec:availability}
+\label{sec:regseg-availability}
 We considered the reproducibility of our results as a design requirement.
@@
   Segmentation Toolkit, \url{http://www.itk.org}).
-The evaluation instruments (\autoref{fig:evworkflows}) were implemented using
+The evaluation instruments (\autoref{fig:regseg-evworkflows}) were implemented using
   \emph{nipype} \citep{gorgolewski_nipype_2011} to assess their reproducibility.

### 2015-10-01 Publications/2015-NeuroImage/ELS285_article_edited.tex
fix labels with prefix

@@
 \linenumbers
-\makeatletter{}\section{Introduction}\label{sec:introduction}
+\makeatletter{}\section{Introduction}\label{sec:regseg-introduction}
 The accurate delineation of \gls*{wm} in \gls*{dmri} and the fusion of prior
@@
 The nonlinear distortion is aligned with one of the imaging axes (see 
-  \autoref{sec:human_connectome}), so \regseg{} includes an anisotropic regularizer for
+  \autoref{sec:regseg-human_connectome}), so \regseg{} includes an anisotropic regularizer for
   the displacements field proposed by \cite{nagel_investigation_1986}.
@@
 \makeatletter{}\section{Methods}
-\label{sec:methods}
+\label{sec:regseg-methods}
 Let $\gammaset_R = \{\Gamma_m: m \in \mathbb{N}, m \leq N_c\}$ be the set of $N_c$ surfaces
@@
 
-\subsection{Cost-function derivation}\label{sec:methods_map}
+\subsection{Cost-function derivation}\label{sec:regseg-methods_map}
 In a Bayesian framework for registration \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014},
-  the mappings $U$ in \eqref{eq:transform} are
+  the mappings $U$ in \eqref{eq:regseg-transform} are
   evaluated based on their posterior probability given the observed data
@@
 \subsection{Numerical Implementation}
-\label{sec:numerical_implementation}
+\label{sec:regseg-numerical_implementation}
 
@@
     the total area of surface $\Gamma_0$.
-			}\label{fig:method}
+			}\label{fig:regseg-method}
 \end{figure}
 
-\paragraph*{Deformation model}\label{sec:deformation_model}
+\paragraph*{Deformation model}\label{sec:regseg-deformation_model}
 Since the vertices of the surfaces $\{\vec{v}_i: \vec{v}_i \subset \gammaset \}_{i=1 \ldots N_v}$
@@
   $\{\vec{u}_k\}_{k=1 \ldots K}$.
-Densification is achieved using a set of associated basis functions $\psi_k$ \eqref{eq:nodes_tfm}.
+Densification is achieved using a set of associated basis functions $\psi_k$ \eqref{eq:regseg-nodes_tfm}.
 In our implementation, $\psi_k$ is selected as a tensor-product B-spline kernel
@@
   \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\delta^{-1} \, \vec{u}_k^t - \vec{g}_k\} }                  {\mathcal{F}\{(\delta^{-1} + \boldsymbol{\alpha})\, I - \boldsymbol{\beta}\Delta\}} \right\},
-  \label{eq:update_equation}
+  \label{eq:regseg-update_equation}
   \end{align}
@@
 \paragraph*{Implementation details, settings, and convergence}
-\label{sec:conv_report}
+\label{sec:regseg-conv_report}
 The \regseg{} tool includes a multiresolution strategy on the free-form deformation field.
@@
   6) The agreement between the surfaces fitted by \regseg{} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$
-     using automatically generated visual reports was assessed by computing the Hausdorff distance between the corresponding surfaces.}\label{fig:evphantoms}
+     using automatically generated visual reports was assessed by computing the Hausdorff distance between the corresponding surfaces.}\label{fig:regseg-evphantoms}
 \end{figure*}
 
-\subsection{Real datasets} \label{sec:human_connectome}
-The general experimental framework for the real datasets is presented in \autoref{fig:evworkflows},
+\subsection{Real datasets} \label{sec:regseg-human_connectome}
+The general experimental framework for the real datasets is presented in \autoref{fig:regseg-evworkflows},
 which extends our previous evaluation \citep{esteban_simulationbased_2014} of distortions
@@
   2) For use as the ground truth, we generated a plausible synthetic distortion $U_{true}$
-    from the field map with \eqref{eq:fieldmap}.
+    from the field map with \eqref{eq:regseg-fieldmap}.
   3) The \gls*{dmri} data were warped using $U_{true}$ to reproduce the effects of real
@@
     the ground-truth deformation.
-  5) The results were evaluated visually and quantitatively.}\label{fig:evworkflows}
+  5) The results were evaluated visually and quantitatively.}\label{fig:regseg-evworkflows}
 \end{figure*}
@@
   the \gls*{fa} and \gls*{adc}, respectively.
-}\label{fig:model}
+}\label{fig:regseg-model}
 \end{figure}
@@
 The field map was unwrapped\footnote{fieldmaps are phase maps, which have been clipped in the interval
-  of $[-\pi, \pi)$ [rads] or [rads/s].} and smoothed before applying \eqref{eq:fieldmap}.
+  of $[-\pi, \pi)$ [rads] or [rads/s].} and smoothed before applying \eqref{eq:regseg-fieldmap}.
 Next, the original \gls*{dmri} was warped using the resulting displacement field and fed into
@@
   $\hat{U} = \epsilon \cdot U_{true} = \vec{r} + \epsilon \, u_\text{PE}$
-  (with $\epsilon \in [-1.1, 1.1]$ and $u_\text{PE}$ from \eqref{eq:fieldmap}),
-  and we evaluated the data term of the cost function \eqref{eq:energy}.
+  (with $\epsilon \in [-1.1, 1.1]$ and $u_\text{PE}$ from \eqref{eq:regseg-fieldmap}),
+  and we evaluated the data term of the cost function \eqref{eq:regseg-energy}.
 The minimum metric was consistently at the ground-truth location ($\epsilon=0.0$)
@@
 \paragraph*{Cross-comparison}
-A similar workflow to the general evaluation framework used for \autoref{fig:evworkflows}
+A similar workflow to the general evaluation framework used for \autoref{fig:regseg-evworkflows}
   was employed to integrate the alternate \gls*{t2b} registration scheme.
@@
 \makeatletter{}\section{Results}
-\label{sec:results}
+\label{sec:regseg-results}
 
 \subsection{Proof of concept using digital phantoms}
-\label{sec:results_phantom}
+\label{sec:regseg-results_phantom}
 
@@
     consistently below the size of the voxels.
-    }\label{fig:phantom}
+    }\label{fig:regseg-phantom}
 \end{figure*}
 In total, 1200 experiments (four phantom types $\times$ 150 random warpings $\times$ two resolutions) were
-  performed according to the workflow illustrated in \autoref{fig:evphantoms}.
+  performed according to the workflow illustrated in \autoref{fig:regseg-evphantoms}.
 For each experiment, the misregistration error was measured using the Hausdorff distance
-  (see \autoref{sec:experiments_evaluation}) between the theoretical $\gammaset_\text{true}$ and
+  (see \autoref{sec:regseg-experiments_evaluation}) between the theoretical $\gammaset_\text{true}$ and
   the estimate obtained by \regseg{} ($\hat{\gammaset}_{test}$).
 The results demonstrated that the accuracy was consistent and high, and below the image resolution.
-\autoref{fig:phantom} (block C) shows the violin plots for each model type corresponding
+\autoref{fig:regseg-phantom} (block C) shows the violin plots for each model type corresponding
   to the two sets of resolutions for the generated phantoms.
@@
 
-\subsection{Evaluation using real datasets and cross-comparison}\label{sec:results_hcp}
+\subsection{Evaluation using real datasets and cross-comparison}\label{sec:regseg-results_hcp}
 Finally, we compared the performance of \regseg{} with that of the standard \gls*{t2b}
@@
 Visual assessments of the 16 cases are described in \suppl{section S5}.
-In \autoref{fig:results_real}, box A, the visual report is shown for one subject.
-We computed the \gls*{swindex} \eqref{eq:swindex} of every surface after registration
+In \autoref{fig:regseg-results_real}, box A, the visual report is shown for one subject.
+We computed the \gls*{swindex} \eqref{eq:regseg-swindex} of every surface after registration
   using both the \regseg{} and \gls*{t2b} methods.
@@
   (a non-parametric alternative to ANOVA) on the warping indices for the three surfaces of 
-  interest selected in \autoref{sec:experiments_evaluation}
+  interest selected in \autoref{sec:regseg-experiments_evaluation}
   ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$).
@@
     (1.25 mm), thereby supporting the improved results obtained by \regseg{} with the proposed settings.
-	}\label{fig:results_real}
+	}\label{fig:regseg-results_real}
 \end{figure*} 
 \makeatletter{}\section{Discussion}
-\label{sec:discussion}
+\label{sec:regseg-discussion}
 
@@
 \makeatletter{}\section*{Conclusion}
-\label{sec:conclusion}
+\label{sec:regseg-conclusion}
 
@@
 \subsection{Simplifying the regularization term}\label{app:reg_term}
-The exponentials of the Thikonov regularization prior \eqref{eq:thikonov} have the general form
+The exponentials of the Thikonov regularization prior \eqref{eq:regseg-thikonov} have the general form
   $\vec{v}^T \mathbf{M} \vec{v}$.
@@
 In general, the anisotropy is aligned with the imaging axes, so 
-  $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:energy} can be simplified to diagonal matrices, such that
+  $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:regseg-energy} can be simplified to diagonal matrices, such that
   $\mathbf{A}= \, \boldsymbol{\alpha}\,\vec{I}_n$ and
   $\mathbf{B}= \, \boldsymbol{\beta}\,\vec{I}_n$.
-By substituting into equation \eqref{eq:energy}, we obtain:
+By substituting into equation \eqref{eq:regseg-energy}, we obtain:
 
@@
   E(\vec{u}) &= \const + \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} +   \int_{\Omega} \frac12 \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2} + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}.
-  \label{eq:app_energy}
+  \label{eq:regseg-app_energy}
   \end{align}

### 2015-10-01 Publications/2015-NeuroImage/expanded.tex
fix labels with prefix

--- 2015-09-07 Publications/2015-NeuroImage/expanded.tex
@@
 \linenumbers
-\makeatletter{}\section{Introduction}\label{sec:introduction}
+\makeatletter{}\section{Introduction}\label{sec:regseg-introduction}
 \Gls*{dmri} enables the mapping of microstructure \citep{basser_microstructural_1996}
@@
 Since the \gls*{epi} distortion is aligned with the \gls*{pe} axis (see
-  \autoref{sec:human_connectome}), \regseg{} includes an anisotropic regularizer for
+  \autoref{sec:regseg-human_connectome}), \regseg{} includes an anisotropic regularizer for
   the displacements field proposed by \cite{nagel_investigation_1986}.
@@
 \makeatletter{}\section{Methods}
-\label{sec:methods}
-\subsection{Related work}\label{sec:related_work}
+\label{sec:regseg-methods}
+\subsection{Related work}\label{sec:regseg-related_work}
 Previously, joint segmentation and registration have been applied successfully to other problems
@@
 
-\subsection{Registration framework and segmentation model derivation}\label{sec:methods_map}
+\subsection{Registration framework and segmentation model derivation}\label{sec:regseg-methods_map}
 Let $\gammaset_R = \{\Gamma_m: m \in \mathbb{N}, m \leq N_S\}$ be the set of $N_S$ surfaces
@@
   voxels is implicitly supported by the piecewise smooth partition of the space \omegaset{}.
-In \autoref{fig:model} it is shown that marginal distributions of data can be approximated by
+In \autoref{fig:regseg-model} it is shown that marginal distributions of data can be approximated by
   multivariate normal distributions. Therefore,
@@
   the complete bivariate data.
-}\label{fig:model}
+}\label{fig:regseg-model}
 \end{figure}
@@
     the total area of surface $\Gamma_0$.
-      }\label{fig:method}
+      }\label{fig:regseg-method}
 \end{figure}
 
-\paragraph*{Deformation model}\label{sec:deformation_model}
+\paragraph*{Deformation model}\label{sec:regseg-deformation_model}
 Since the vertices of the surfaces $\{\vec{v}_i: \vec{v}_i \subset \gammaset \}_{i=1 \ldots N_V}$
@@
      visually using automatically generated visual reports and quantitativelly with the Hausdorff distance between the
-     corresponding surfaces.}\label{fig:evphantoms}
+     corresponding surfaces.}\label{fig:regseg-evphantoms}
 \end{figure*}
 
-\subsection{Real datasets} \label{sec:human_connectome}
-The general experimental framework for the real datasets is presented in \autoref{fig:evworkflows},
+\subsection{Real datasets} \label{sec:regseg-human_connectome}
+The general experimental framework for the real datasets is presented in \autoref{fig:regseg-evworkflows},
 which extends our previous evaluation \citep{esteban_simulationbased_2014} of distortions
@@
     (1.25 mm), thereby supporting the improved results obtained by \regseg{} with the proposed settings.
-  }\label{fig:results_real}
+  }\label{fig:regseg-results_real}
 \end{figure*}  
 \makeatletter{}\section{Discussion}
-\label{sec:discussion}
+\label{sec:regseg-discussion}
 

### 2015-10-01 Publications/2015-NeuroImage/supplemental.tex
update title in supplementals
fix labels with prefix

@@
 % @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-04-21 13:22:12
+% @Last Modified time: 2015-04-29 17:53:12
 
@@
 \begin{document}
-\title{Supplemental Materials: \\ \emph{Active-contours driven registration method for structure-informed segmentation of diffusion MR images}}
+\title{Supplemental Materials: \\ \emph{Active contours-driven registration method for the structure-informed segmentation of diffusion MR images}}
 \author{Esteban O., Zosso D., Daducci A., Bach-Cuadra M., \\ Ledesma-Carbayo MJ., Thiran JP, and Santos A.}
@@
 \maketitle
-\section{Extensions to the mathematical formulation of the methods}\label{sec:maths_extension}
+\section{Extensions to the mathematical formulation of the methods}\label{sec:regseg-maths_extension}
 
-\subsection{Computing the gradients of shape-priors}\label{sec:shape_priors}
+\subsection{Computing the gradients of shape-priors}\label{sec:regseg-shape_priors}
 The computation of gradients at the locations of the active contours in the
@@
   N_{\Gamma_{l,m}}\right\rangle d\vec{r}.
-  \label{eq:shape_gradients}
+  \label{eq:regseg-shape_gradients}
   \end{equation}
 
-The equation \eqref{eq:shape_gradients} is discretized as follows.
+The equation \eqref{eq:regseg-shape_gradients} is discretized as follows.
 First, the surface between limiting regions $\{l, m\}$, $\Gamma_{l,m}$ is explicitly represented by
@@
   \hat{\vec{n}}_{i}\right\rangle.
-  \label{eq:shape_gradient_orig}
+  \label{eq:regseg-shape_gradient_orig}
   \end{align}
@@
   - \underset{i}{\sum} w_{p,i} F(\vec{v}_i) \frac{\partial \vec{v}_i}{\partial t} \cdot \hat{\vec{n}}_i.
-  \label{eq:shape_gradient_disc1}
+  \label{eq:regseg-shape_gradient_disc1}
   \end{equation}
@@
   \underbracket{w_{p,i} \, \Big[ F_{out}(\vec{v}_i) - F_{in}(\vec{v}_i) \Big] \hat{\vec{n}}_i}_{\bar{s}_i \text{ in Figure 1}}.
-  \label{eq:shape_gradient_disc2}
+  \label{eq:regseg-shape_gradient_disc2}
   \end{equation}
 
-\subsection{Gradient-descent optimization}\label{sec:gradient_descent}
+\subsection{Gradient-descent optimization}\label{sec:regseg-gradient_descent}
 The energy functional to be optimized in \emph{regseg} is presented in Eq. 7.
@@
   + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}}_{\text{Regularization term } (E_{reg})}.
-  \label{eq:energy}
+  \label{eq:regseg-energy}
   \end{equation}
@@
 To search for the minimum of $E(\vec{u})$ w.r.t. the coefficients $\vec{u}_k$, we use a gradient descent strategy.
-In Eq. (9) we introduced the derivative of \eqref{eq:energy}:
+In Eq. (9) we introduced the derivative of \eqref{eq:regseg-energy}:
   \begin{equation}
@@
   \Big\}.
-  \label{eq:gradient_descent}
+  \label{eq:regseg-gradient_descent}
   \end{equation}
 
-We split \eqref{eq:gradient_descent} in the derivatives of its data and regularization terms.
+We split \eqref{eq:regseg-gradient_descent} in the derivatives of its data and regularization terms.
 Let $\frac{\partial E_{data}}{\partial \vec{u}_k} = \vec{g}_k$ for simplicity,
@@
   \frac{ \partial }{\partial \vec{u}_k} \Big\{ \underset{l}{\sum} \underset{\vec{x} \in \Omega_l}{\sum} \, \mdist{f'}{l} \Big\},
-  \label{eq:data_derivative}
+  \label{eq:regseg-data_derivative}
 \end{equation}
-  where we can apply the shape-gradients \eqref{eq:shape_gradient_disc2} introduced
-  in \autoref{sec:shape_priors}, and ultimately avoid implementing level sets:
+  where we can apply the shape-gradients \eqref{eq:regseg-shape_gradient_disc2} introduced
+  in \autoref{sec:regseg-shape_priors}, and ultimately avoid implementing level sets:
 
@@
   \frac{\partial}{\partial \vec{u}_k} \left\{ \vec{v}_i + \sum_k \psi_k(\vec{v}_i) \vec{u}_k \right\} = \psi_k(\vec{v}_i)\, \hat{\vec{e}}.
-  \label{eq:gradient_wshape}
+  \label{eq:regseg-gradient_wshape}
   \end{align}%
@@
   \vec{g}_k = - \underset{i}{\sum} \bar{s}_i \cdot \psi_k(\vec{v}_i) \, \hat{\vec{e}} = - \underset{i}{\sum} \vec{g}_{i,k}.
-  \label{eq:shape_gradient_final}
+  \label{eq:regseg-shape_gradient_final}
 \end{equation}
 
-It is also necessary to obtain and discretize the derivatives of the regularization term of \eqref{eq:gradient_descent}:
+It is also necessary to obtain and discretize the derivatives of the regularization term of \eqref{eq:regseg-gradient_descent}:
   \begin{equation}
@@
   \Big\} = \boldsymbol{\alpha} \cdot \vec{u}_k - \boldsymbol{\beta} \cdot \Delta \vec{u}_k.
-  \label{eq:gradient_regterm}
+  \label{eq:regseg-gradient_regterm}
   \end{equation}
 
-Inserting \eqref{eq:shape_gradient_final} and \eqref{eq:gradient_regterm} into \eqref{eq:gradient_descent} we get the
+Inserting \eqref{eq:regseg-shape_gradient_final} and \eqref{eq:regseg-gradient_regterm} into \eqref{eq:regseg-gradient_descent} we get the
   final evolution equation:
@@
   \frac{\partial \vec{u}}{\partial t} &\propto - \Big( \vec{g}_k - \boldsymbol{\alpha} \cdot \vec{u}_k + \boldsymbol{\beta} \cdot \Delta \vec{u}_k \Big).
-  \label{eq:gradient_final}
+  \label{eq:regseg-gradient_final}
   \end{align}
 
-\subsection{Obtaining the update equation} To solve the differential equation in \eqref{eq:gradient_final},
+\subsection{Obtaining the update equation} To solve the differential equation in \eqref{eq:regseg-gradient_final},
   we use a semi-implicit Euler scheme, referring to the discrete step size as $\delta$ and where the
@@
   (1 + \delta \, \boldsymbol{\alpha} - \delta \, \boldsymbol{\beta} \Delta)\, \vec{u}_k^{t+1} &= \vec{u}_k^{t} - \delta\, \vec{g}_k^{t}
-  \label{eq:discrete_derivative}
+  \label{eq:regseg-discrete_derivative}
 \end{align}
@@
                   {\mathcal{F}\{(\delta^{-1} + \boldsymbol{\alpha})\, I-\boldsymbol{\beta}\Delta\}} \right\},
-  \label{eq:update_equation}
+  \label{eq:regseg-update_equation}
   \end{align}
@@
 
-%Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
+%Then, introducing \eqref{eq:regseg-intp_kernel} into \eqref{eq:regseg-nodes_tfm} and replacing
 %  $\psi$ by the actual kernel function, the transformation writes:
@@
 %      \beta_3( (\vec{v}_i - \vec{r}_k) \cdot \hat{\mathbf{e}}_d ) \right],
-%  \label{eq:transformation}
+%  \label{eq:regseg-transformation}
 %  \end{equation}
@@
 The tool generates a log-file in JSON format to easily inter-operate with secondary tools (such
-  as the convergence report generation, \autoref{sec:convergence_evidence}).
+  as the convergence report generation, \autoref{sec:regseg-convergence_evidence}).
 The evaluation workflow is independent, and implemented using \emph{nipype} \citep{gorgolewski_nipype_2011}.
@@
   $\vec{v}_i$ by the corresponding weights $\psi_{ik} = \psi_k(\mathbf{v}_i)$ of interpolating functions
-  \eqref{eq:shape_gradient_final}.
+  \eqref{eq:regseg-shape_gradient_final}.
 In order to optimize multiplications and summations, all the $\psi_{ik}$ are collected in a
@@
   \big(\vec{g}_k \big) = \boldsymbol{\Psi} \cdot \big( w_i \, \bar{s}_i \big)^T
-  \label{eq:sparse_matrix}
+  \label{eq:regseg-sparse_matrix}
 \end{equation}
@@
     (with $\epsilon \in [-1.1, 1.1]$ and $u_\text{PE}$ from Equation 13 in the
-    paper), and evaluated the data-term of the cost function \eqref{eq:energy}.
+    paper), and evaluated the data-term of the cost function \eqref{eq:regseg-energy}.
   The metric consistently displayed its minimum at the ground-truth location ($\epsilon=0.0$)
@@
   {\includegraphics[width=\linewidth]{Suppl-figure04.pdf}}
-  \label{fig:energymap}
+  \label{fig:regseg-energymap}
 \end{figure}
 
-\subsection{Interface and Settings}\label{sec:interface_settings}
+\subsection{Interface and Settings}\label{sec:regseg-interface_settings}
 
@@
 
-\subsection{Convergence evidencing}\label{sec:convergence_evidence}
+\subsection{Convergence evidencing}\label{sec:regseg-convergence_evidence}
 
@@
 Online checking is also possible as the algorithm writes to the standard output as well.
-A sample report is found in \autoref{fig:convreport}.
+A sample report is found in \autoref{fig:regseg-convreport}.
 
@@
     features to help researchers tune up the algorithm in their application.}%
-    \label{fig:convreport}
+    \label{fig:regseg-convreport}
 \end{figure}
@@
   In this figure, the composite workflow to extract the six surfaces that define the
-    segmentation model (\autoref{fig:jointplot}) is represented.
-  }\label{fig:wf_surfaces}
+    segmentation model (\autoref{fig:regseg-jointplot}) is represented.
+  }\label{fig:regseg-wf_surfaces}
 \end{figure}
@@
       computing the two feature maps of interest (\gls*{fa} and \gls*{adc}).
-    This workflow is performed in both undistorted and warped data.}\label{fig:wf_dti}}
+    This workflow is performed in both undistorted and warped data.}\label{fig:regseg-wf_dti}}
 \end{figure}
@@
     In this figure, the composite workflow to extract the six surfaces that define the
-      segmentation model (\autoref{fig:jointplot}) is represented.}\label{fig:wf_t2b}}
+      segmentation model (\autoref{fig:regseg-jointplot}) is represented.}\label{fig:regseg-wf_t2b}}
 \end{figure}
@@
   \includegraphics[width=\linewidth]{hcp_wf.pdf}
-  \caption{\textbf{The \emph{nipype} design of the workflow for real data}.}\label{fig:wf_hcp}
+  \caption{\textbf{The \emph{nipype} design of the workflow for real data}.}\label{fig:regseg-wf_hcp}
 \end{figure}
@@
 %   \includestandalone[width=\linewidth]{figure06}
-%   \caption{}label{fig:results_real}
+%   \caption{}label{fig:regseg-results_real}
 % \end{figure*}
@@
     \end{scope}
-  \end{tikzpicture}\label{fig:jointplot}
+  \end{tikzpicture}\label{fig:regseg-jointplot}
 \end{figure}
@@
 This report has been drafted and prepared by OE.
-Mathematical formulations in \autoref{sec:maths_extension} were originally devised and drafted
+Mathematical formulations in \autoref{sec:regseg-maths_extension} were originally devised and drafted
   by DZ and rewritten for this document by OE.

### 2015-10-01 Publications/2015-NeuroImage/04-methods.tex
fix typo

@@
   6) The agreement between the surfaces fitted by \regseg{} ($\hat{\gammaset}_\text{test}$) and $\gammaset_\text{true}$ were assessed
-     visually using automatically generated visual reports and quantitativelly with the Hausdorff distance between the
+     visually using automatically generated visual reports and quantitatively with the Hausdorff distance between the
      corresponding surfaces.}\label{fig:regseg-evphantoms}

### 2015-10-02 Publications/2015-NeuroImage/04-methods.tex
added comments from reviewers

@@
 $C$ is the number of channels comprised in the image $M$.
+In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
+  the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized.
 \newcomment[RV\#1(C.2)]{%
-In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
-  the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized. In \autoref{fig:regseg-model},
+In \autoref{fig:regseg-model},
   this minimization is illustrated for the segmentation of the \gls*{fa} and the \gls*{adc} maps
@@
 
-\paragraph*{Implementation details and convergence}
+\paragraph*{Implementation details and convergence}%
 \label{sec:regseg-conv_report}
@@
 
-\paragraph*{Segmentation model and settings}
+\newcomment[RV\#1(C.6)]{\paragraph*{Segmentation model and settings}
 The segmentation model of the phantoms is implicitly defined in each shape.
@@
 All the experimental settings used for the phantoms are made available in
-  a unique configuration file\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/regseg_default.json}}.
+  a unique configuration file\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/regseg_default.json}}.}
 
@@
   multi-shell \gls*{dmri} images.
+\newcomment[RV\#1(C.7)]{%
 Since we obtained the \gls*{dmri} data from the minimally preprocessed package, these
   images are corrected for \gls*{epi} distortions and spatially normalized in
-  \gls*{t1} space.
+  \gls*{t1} space.}
 
@@
 \paragraph*{Cross-comparison and settings} %
-\newcomment[RV\#2(C.21)]{%
 A dual workflow to the general evaluation framework used for \regseg{} (\autoref{fig:regseg-evworkflows}),
@@
   perform registration.
+\newcomment[RV\#2(C.21)]{%
 The deformation field is correspondingly restricted to the \gls*{pe} direction.

### 2015-10-02 Publications/2015-NeuroImage/06-discussion.tex
add missing review label

@@
   structures in \gls*{dmri} space.}
+\newcomment[New]{%
 With the definition of appropriate models, \regseg{} may be applied in other fields like
@@
   propagated to the initial timepoints, regardless the changes of the contrast and spatial
-  development between timepoints.
+  development between timepoints.}
 

### 2015-10-23 Publications/2015-NeuroImage/03-introduction.tex
fixed typos, added comment labels in manuscript
fix labels with prefix
update intro

@@
 % -*- root: 00-main.tex -*-
-\section{Introduction}\label{sec:introduction}
+\section{Introduction}\label{sec:regseg-intro}
 \Gls*{dmri} enables the mapping of microstructure \citep{basser_microstructural_1996}
@@
   tractography as edges \citep{hagmann_mapping_2008}.
-\newcomment[RV\#2(C.17)]{Since the proposal of the earliest methodologies to solve the three problems,
-  there is a need for precisely identifying the anatomy in the \gls*{dmri} space.
-In fact, current trends on reconstruction \citep{jeurissen_multitissue_2014} and
+\newcomment[RV\#2(C.17)]{%
+The methodologies to solve reconstruction, tractography and network building
+  require the delineation of the anatomy in the \gls*{dmri} space.
+Moreover, current trends on reconstruction \citep{jeurissen_multitissue_2014} and
   tractography \citep{smith_anatomicallyconstrained_2012} are increasingly using
@@
 
-\newcomment[RV\#2(C.15)]{The structural information required by pioneer tractography studies was a simple
-  \gls*{wm} mask to provide with a termination criteria for the streamlines.
-The first attempts to segment the \gls*{wm} in \gls*{dti} involved thresholding the
-  \gls*{fa}\footnote{the \gls*{fa} is a scalar parameter of diffusion derived from
-  the \gls*{dti}.} map.
+\newcomment[RV\#2(C.15)]{%
+Possibly, the earliest structural information incorporated to aid \gls*{dmri} processing
+  is the \gls*{wm} mask used as a termination criteria for tractography.
+The standardized procedure to obtain this mask was thresholding the \gls*{fa} map.
 However, the mask and subsequent analyses are highly dependent on the threshold set
   \citep{taoka_fractional_2009}.
+\todo{RV\#2(C.16)}
 To overcome the unreliability of \gls*{fa} thresholding, and to broaden
@@
   scalar maps such as \gls*{fa}, \gls*{adc} and others \citep{zhukov_level_2003,
-  rousson_level_2004,jonasson_segmentation_2005,liu_brain_2007,hadjiprocopis_unbiased_2005,
+  rousson_level_2004,jonasson_segmentation_2005,hadjiprocopis_unbiased_2005,liu_brain_2007,
   lu_segmentation_2008,han_experimental_2009}.}
-\todo{RV\#2(C.14)}However, the precise segmentation of \gls*{dmri} is difficult for several limitations.
-First, \gls{dmri} images have a resolution that is \newcomment[RV\#2(C.12)]{much lower} than that of the imaged
-  microstructural features.
+\todo{RV\#2(C.14)}%
+However, the precise segmentation of \gls*{dmri} is difficult for several reasons.
+First, \gls{dmri} images have a resolution that is \newcomment[RV\#2(C.12)]{much lower}
+  than that of the imaged microstructural features.
 Therefore, voxels located in structural discontinuities are affected by partial
@@
 An alternative route to segmentation in native \gls*{dmri} space is the mapping of the
-  structural information extracted from anatomical MR images, such as \gls*{t1}, through
-  an image registration process.
+  structural information extracted from anatomical MR images, such as \gls*{t1}, using
+  image registration techniques.
 Generally, intra-subject registration of MR images of the brain involves only a linear
   mapping to compensate for head motion between scans.
-However, \gls*{epi} images speed up the acquisition process at the cost of introducing geometrical
-  distortion, as well as signal degradation and destruction \citep{jezzard_correction_1995}.
-The so-called \emph{\gls*{epi} distortions} have a major impact on the anatomy extracted
-  from \gls*{dmri}, particularly in certain fiber bundles \citep{irfanoglu_effects_2012}.
-To overcome this problem, two broad families of solutions are available.
-First, it is possible to use an existing ``\gls*{epi} distortion correction''
-  method to compensate for the nonlinear part of the misalignment with the structural space,
-  combined with a rigid registration method.
-Second, to apply a nonlinear registration method with specific constraints derived from the
-  theoretical properties of the \gls*{epi} distortion.
-Both solutions require extra acquisitions to be included within the \gls*{dmri} scanning
-  protocol, such as fieldmaps \citep{jezzard_correction_1995}, \glspl*{dwi} with a different
-  \gls*{pe} scheme \citep{cordes_geometric_2000,chiou_simple_2000}, or \gls*{t2} images
-  \citep{kybic_unwarping_2000}.
-The nonlinear registration approach has been successfully solved with \gls*{t2b} methods,
-  which align the \gls*{t2} and \lowb{} images by exploiting their similarity
-  \citep{kybic_unwarping_2000,studholme_accurate_2000}.
-However, further rigid registration of the \gls*{t1} and \gls*{t2} images is still required to
-  map the anatomical information onto \gls*{dmri}-space.
-\Gls*{epi} corrections have been an active field of research, with numerous refinements and
-  combinations of the original methods \citep{holland_efficient_2010,andersson_comprehensive_2012,
-  irfanoglu_drbuddi_2015}.}\todo{RV\#2(C.18)}
+However, \gls*{epi} introduces a geometrical distortion \citep{jezzard_correction_1995}
+  that impedes the linear mapping from the structural space and warps the anatomy 
+  identified on native \gls*{dmri} space \citep{irfanoglu_effects_2012}.
+Numerous methods have been proposed to overcome this problem by incorporating information
+  from extra MR acquisitions such as fieldmaps \citep{jezzard_correction_1995},
+  \glspl*{dwi} with a different \gls*{pe} scheme \citep{cordes_geometric_2000,chiou_simple_2000},
+  or \gls*{t2} images \citep{kybic_unwarping_2000,studholme_accurate_2000}.
+These methods estimate the deformation field associated to \emph{\gls*{epi} distortions} and
+  resample the \glspl*{dwi} onto a corrected \gls*{dmri} space.
+The retrospective \gls*{epi} correction is an active field of research yielding frequent refinements 
+  and combinations of the original methods, such as
+  \citep{holland_efficient_2010,andersson_comprehensive_2012,
+  irfanoglu_drbuddi_2015}. \todo{RV\#2(C.18)}
+A standardized method to solve the remaining linear mapping between the corrected-\gls*{dmri}
+  and the structural spaces is \emph{bbregister} \citep{greve_accurate_2009}.%
+}
 
-Here we show a method that registers \newcomment[RV\#2(C.10)]{anatomically-correct surfaces to multivariate images},
-  thereby implicitly segmenting the \gls*{dmri}-space into those structures enclosed by the contours
-  of the subject's brain.
-% It is important that surfaces are defined in the coordinate system of the image used as
-%   source of structural information (typically, but not restricted to, \gls*{t1} images), otherwise segmentation performs
-%   correctly but the mapping found in registration does not represent any valuable information.
-The hypothesis underlying is that the registration and segmentation problems in \gls*{dmri}
-  can be solved simultaneously with three key benefits:
-  1) effectively establishing the necessary mapping to project structural information from anatomical images,
-  2) increasing the geometrical accuracy of the overall process,
-  and 3) avoiding data resampling of \gls*{epi} correction methods.
-Thus, we propose a tool called \regseg{} implementing 3D active contours (surfaces) without edges
-  \citep{chan_active_2001}, which evolve over a free-form deformation field of B-spline
-  basis functions.
-The surfaces are nested and they enclose regions with homogeneous diffusion features
-  (the \gls*{fa} and \gls*{adc} maps for the case of \gls*{dmri} data).
-Optimization is performed using a descent strategy with explicit shape gradients
+\newcomment[RV\#2(C.10)]{%
+Here we present a segmentation and surface-to-volume registration method 
+  named \regseg{}, and show its usefulness in incorporating the anatomical
+  information into native \gls*{dmri} space to aid subsequent reconstruction,
+  tractography and network building using a cortical parcellation.
+The hypothesis underlying is that the registration and segmentation problems
+  in \gls*{dmri} can be solved simultaneously.
+To implement \regseg{} we first establish an active-contours without edges
+  \citep{chan_active_2001} segmentation framework.
+Several \emph{anatomically-correct} surfaces of the same subject initialize
+  the 3D active-contours, which evolve searching for homogeneous regions in
+  the multivariate target-image.
+In the presented application to segment \gls*{dmri}, we propose a bivariate 
+  target-volume comprehending the \gls*{fa} and \gls*{adc} maps.
+The active-contours are without edges since the \gls*{fa} and \gls*{adc} maps
+  do not present steep image gradients (edges) but anatomy can be identified
+  by looking for piece-wise smooth homogeneous regions.
+The evolution of the surfaces is supported by a B-spline basis, optimized
+  iteratively using a descent approach driven by shape-gradients
   \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
-Finally, we evaluate \regseg{} using an extension of our instrumentation framework
-  \citep{esteban_simulationbased_2014}, which simulates known and realistic distortions
-  based on \gls*{dmri} data.
-The evaluation includes a comparison of \regseg{} to an in-house implementation of
-  the \gls*{t2b} correction.
+Therefore, \regseg{} establishes a registration framework that actually
+  overcomes the nonlinear warping induced by \gls*{epi} distortions.%
+}
+\Regseg{} integrates the benefits of segmentation and registration methods together and
+  exploits the multivariate nature of \gls*{dmri} data to contribute in the proposed
+  application on three key aspects:
+  1) it avoids the resampling the \glspl*{dwi} introduced by the correction of \gls*{epi}
+    distortions, since processing can be performed on the native \gls*{dmri} space;
+  2) the surfaces are typically extracted from the \gls*{t1} image of the same subject,
+    therefore \regseg{} does not require additional MR acquisitions to the minimal
+    \gls*{dmri} protocol; and
+  3) \regseg{} increases the geometrical accuracy of the overall process.
+In this paper, we first verify the functionality of the method and the \regseg{}
+  implementation using a set of digital phantoms, demonstrating the sub-pixel accuracy
+  of registration.
+We then evaluate \regseg{} on real \gls*{dmri} datasets, using a derivation of our
+  instrumentation framework \citep{esteban_simulationbased_2014} which simulates 
+  known and realistic \gls*{epi} distortions.
+We also compare \regseg{} and a nonlinear registration method to map the \lowb{} to 
+  the corresponding \gls*{t2} image of the same subject.
+This method is a derivation of the \gls*{t2b} correction method, reproducing the settings
+  of an existing implementation included within a widely used diffusion processing software
+  \citep{leemans_exploredti_2009}.
+With this comparison we demonstrate how \regseg{} achieves higher accuracy with the
+  simultaneous registration and segmentation process.

### 2015-10-23 Publications/2015-NeuroImage/04-methods.tex
update intro

@@
 For convenience and because it has been shown to be an appropriate approximation
-  \citep{leemput_automated_1999a,cuadra_comparison_2005}, we introduce two assumptions for each
+  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
   region $\Omega_l$:

### 2015-10-23 Publications/2015-NeuroImage/05-results.tex
added comments from reviewers
update intro

@@
 
-\subsection{Proof of concept using digital phantoms}
+\subsection{Verification and validation using digital phantoms}
 \label{sec:regseg-results_phantom}
 
-The results demonstrated that the accuracy was consistent and high, and below the image resolution.
-\autoref{fig:regseg-phantom} (block C) shows the violin plots for each model type corresponding
+The results summarized in \autoref{fig:regseg-phantom} demonstrated that the accuracy was
+  high and below the image resolution.
+Panel B on \autoref{fig:regseg-phantom} shows the violin plots for each model type corresponding
   to the two sets of resolutions for the generated phantoms.
@@
     Res.   & ``gyrus'' & ``ball''  & ``box''   & ``L''     & Aggreg.    \\\hline
-    1.0mm  & 0.18-0.38 & 0.31-0.45 & 0.34-0.42 & 0.34-0.40 & 0.34-0.38  \\
-    2.0mm  & 0.59-0.60 & 0.65-0.76 & 0.68-0.71 & 0.67-0.77 & 0.64-0.66  \\
+    1.0mm  & 0.18--0.38 & 0.31--0.45 & 0.34--0.42 & 0.34--0.40 & 0.34--0.38  \\
+    2.0mm  & 0.59--0.60 & 0.65--0.76 & 0.68--0.71 & 0.67--0.77 & 0.64--0.66  \\
     \hline

### 2015-10-24 Publications/2015-NeuroImage/03-introduction.tex
Working on discussion, intro almost finished

@@
   the multivariate target-image.
-In the presented application to segment \gls*{dmri}, we propose a bivariate 
+In the presented application to segment \gls*{dmri} data, we propose a bivariate 
   target-volume comprehending the \gls*{fa} and \gls*{adc} maps.
-The active-contours are without edges since the \gls*{fa} and \gls*{adc} maps
-  do not present steep image gradients (edges) but anatomy can be identified
-  by looking for piece-wise smooth homogeneous regions.
 The evolution of the surfaces is supported by a B-spline basis, optimized
@@
   implementation using a set of digital phantoms, demonstrating the sub-pixel accuracy
-  of registration.
+  in registration.
 We then evaluate \regseg{} on real \gls*{dmri} datasets, using a derivation of our

### 2015-10-24 Publications/2015-NeuroImage/06-discussion.tex
Working on discussion, intro almost finished

@@
 \label{sec:regseg-discussion}
-
-\newcomment[RV\#1(C.3)]{%
-\paragraph*{Contributions}\label{sec:regseg-related_work} %
-Previously, joint segmentation and registration have been applied successfully to other problems
-such as longitudinal object tracking \citep{paragios_level_2003} and atlas-based
-  segmentation \citep{gorthi_active_2011}.
-The most common approach to solve the problem simultaneously is optimizing a deformation model
-  (registration) that supports the evolution the active contours (segmentation), like
-  \citep{paragios_level_2003,yezzi_variational_2003}.
-\cite{unal_coupled_2005}, and later \cite{wang_joint_2006}, replaced the linear registration
-  transform of \cite{yezzi_variational_2003} with a free-form deformation field.
-\cite{droske_mumfordshah_2009} reviewed the existing techniques and proposed two different
-  approaches for applying the Mumford-Shah functional \citep{mumford_optimal_1989} during simultaneous
-  registration and segmentation by propagating the deformation field from
-  the contours onto the whole image definition.
-Finally, \cite{gorthi_active_2011} extended the existing methodologies using a multiphase
-  level set function to register several active contours during the application
-  of atlas-based segmentation.
-Unlike the aforementioned registration-and-segmentation methods, \regseg{} uses shape-gradients
-  \citep{herbulot_segmentation_2006} computed on a hierarchical set of explicit surfaces
-  (triangularized meshes) that substitute the multiphase level sets.
-Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
-  registration method in 2D using level sets and a nonlinear elasticity smoother on the
-  displacement vector field, which preserves the topology even with very large deformations.
-\Regseg{} includes an anisotropic regularizer for the displacements field proposed by
-  \cite{nagel_investigation_1986}.
-This regularization strategy conceptually falls in the midway between the Gaussian smoothing
-  generally included in most of the existing methodologies, and the complexity of
-  the elasticity smoother of \cite{guyader_combined_2011}.
+We present \regseg{}, a simultaneous segmentation and registration method that
+  maps a set of nested surfaces into a multivariate target-image.
+The nonlinear registration process evolves driven by the fitness of the
+  picewise-smooth classification of voxels in the target volume imposed
+  by the current mapping of the surfaces.
+We propose \regseg{} to incorporate anatomical information extracted from \gls*{t1}
+  images into the corresponding \gls*{dmri} of the same subject.
+Previously, joint segmentation and registration have been applied successfully to
+  other problems such as longitudinal object tracking \citep{paragios_level_2003}
+  and atlas-based segmentation \citep{gorthi_active_2011}.
+The most common approach to solve the problem simultaneously is optimizing a deformation
+  model (registration) that supports the evolution the active contours (segmentation),
+  like \cite{paragios_level_2003,yezzi_variational_2003}.
+\Regseg{} can be seen as particular case of atlas-based segmentation, replacing the atlas
+  by the structural image of the subject.
+However, the majority of atlas-based segmentation methods, such as \citep{gorthi_active_2011},
+  use a multiphase level set function to register several active contours.
+Alternatively, \regseg{} implements the active contours with a hierarchical set of explicit
+  surfaces (triangular meshes) that substitute the multiphase level sets, and registration
+  is driven by shape-gradients \citep{herbulot_segmentation_2006}.
 
@@
   correction of choice.
-It implements a linear mapping, and uses 3D active contours with edges\footnote{A three-dimensional active contour
-  is a surface and it is ``with edges'' when the contours evolve in search of abrupt intensity steps
-  (the edges) within the target image.} to search for intensity boundaries in the \lowb{} image.
+It implements a linear mapping, and uses 3D active contours \emph{with edges} to
+  search for intensity boundaries in the \lowb{} image.
 The active contours are initialized using surfaces extracted from the \gls*{t1} using
@@
   the affine transformation model.
-\Regseg{} can be seen as a derivative of \emph{bbregister} for the use of active surfaces.
-However, the deformation model is nonlinear and the surfaces in \regseg{} look for the
-  homogeneity within the regions enclosed, instead of image edges.}
+The deformation model of \regseg{} is nonlinear and the surfaces are active
+  contours \emph{without edges} \citep{chan_active_2001}.
+The contours are without edges because the \gls*{fa} and \gls*{adc} maps do
+  not present steep image gradients (edges) but the anatomy can be identified
+  by looking for piece-wise smooth homogeneous regions.
 
-\newcomment[RV\#1(C.5)]{%
-Finally, we also contribute proposing a piecewise-smooth segmentation model defined by
-  a selection of nested surfaces to partition the multispectral space
-  comprehending the \gls*{fa} and the \gls*{adc} maps and ultimately identify these
-  structures in \gls*{dmri} space.}
-\newcomment[New]{%
-With the definition of appropriate models, \regseg{} may be applied in other fields like
-  neonatal brain image segmentation in longitudinal \gls*{mri} studies like \citep{shi_neonatal_2010}
-  in which the surfaces obtained in a mature timepoint of the brain can be retrospectively
-  propagated to the initial timepoints, regardless the changes of the contrast and spatial
-  development between timepoints.}
+Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
+  registration method in 2D using level sets and a nonlinear elasticity smoother on the
+  displacement vector field, which preserves the topology even with very large deformations.
+\Regseg{} includes an anisotropic regularizer for the displacements field proposed by
+  \cite{nagel_investigation_1986}.
+This regularization strategy conceptually falls in the midway between the Gaussian smoothing
+  generally included in most of the existing methodologies, and the complexity of
+  the elasticity smoother of \cite{guyader_combined_2011}.
+Other minor features that differ from current methods in joint segmentation and registration are
+  the support of multivariate target-images and the efficient computation of the shape-gradients
+  implemented with sparse matrices.
 
-\paragraph*{Accuracy tests}
-The hypothesis tested in our study was that reliable image registration can be performed
-  by searching for the homogeneous regions defined by a set of nested surfaces in a
-  multispectral image
-The surfaces correspond to precise contours extracted from another image of the same subject
-  (i.e. a \gls*{t1} image, or a different time step).
-We demonstrated that active contours without edges can be used successfully to drive a
-  deformation supported by B-spline basis functions with digital phantoms.
+We verified that precise segmentation and registration of a set of surfaces into multivariate
+  data is possible on digital phantoms.
 We randomly deformed four different phantom models to mimic three homogeneous regions
-  (\gls*{wm}, \gls*{gm}, and \acrlong*{csf}) and we used them to simulate \gls*{t1} and \gls*{t2}
-  images at two resolution levels.
+  (\gls*{wm}, \gls*{gm}, and \acrlong*{csf}) and we used them to simulate \gls*{t1} 
+  and \gls*{t2} images at two resolution levels.
 After registration with \regseg{}, we measured the Hausdorff distance between the
@@
 
-\paragraph*{Application to real data}
+\todo{real data}
+This application is necessary in the extraction of structural connectivity from
+  \gls*{dmri}, and has been typically solved in a two-step approach.
+First, the \glspl*{dwi} are corrected for \emph{\gls*{epi} distortions} by estimating
+  the nonlinear-deformation field from extra MR acquisitions
+  \citep{jezzard_correction_1995,chiou_simple_2000,cordes_geometric_2000,
+  kybic_unwarping_2000}.
+Second, mapping the structural information from the corresponding \gls*{t1} image
+  using a linear registration tool like \emph{bbregister} \citep{greve_accurate_2009}.
+numerous attempts to delineate the structures of interest in diffusion
+  space have been proposed.
+Current activity on improving correction methods \citep{irfanoglu_drbuddi_2015} and
+  segmenting the \gls*{dmri} in its native space \citep{jeurissen_tissuetype_2015}
+  proofs the open interest of the application.
+\todo{real data old}
 We designed \regseg{} as a method for segmenting \gls*{dmri} data by exploiting the
@@
 
+\newcomment[RV\#1(C.3)]{%
+\paragraph*{Contributions}\label{sec:regseg-related_work} %
+
+}
+
+\newcomment[RV\#1(C.5)]{%
+Finally, we also contribute proposing a piecewise-smooth segmentation model defined by
+  a selection of nested surfaces to partition the multispectral space
+  comprehending the \gls*{fa} and the \gls*{adc} maps and ultimately identify these
+  structures in \gls*{dmri} space.}
+\newcomment[New]{%
+With the definition of appropriate models, \regseg{} may be applied in other fields like
+  neonatal brain image segmentation in longitudinal \gls*{mri} studies like \citep{shi_neonatal_2010}
+  in which the surfaces obtained in a mature timepoint of the brain can be retrospectively
+  propagated to the initial timepoints, regardless the changes of the contrast and spatial
+  development between timepoints.}
+
 % \paragraph*{Prospects}

### 2015-10-26 Publications/2015-NeuroImage/00-main.tex
fixes for version 1.2
updates before sending @zosso and @daducci
v1.3.0rc

@@
-% \documentclass[5p,authoryear,fleqn]{elsarticle}
+%\documentclass[5p,authoryear,fleqn]{elsarticle}
 \documentclass[3p,authoryear,fleqn]{elsarticle}
@@
 \usepackage[utf8]{inputenc}
+
+\usepackage{xifthen}
+\newboolean{review}
+\setboolean{review}{true}
 
@@
 \usepackage{multirow}
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage{xifthen}
+
+% *** PDF, URL AND HYPERLINK PACKAGES ***
+%
+\usepackage{url}
+\usepackage[colorlinks=true]{hyperref}
+% \usepackage[hyperref]{xcolor}
+\usepackage{doi}
+
 \usepackage{xspace}
@@
 \DeclareMathOperator{\const}{Const.}
-\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
+\providecommand{\e}[1]{\ensuremath{\cdot\text{10}^\text{#1}}}
 \providecommand{\mdist}[2]{ \mathcal{D}_{#2}^2(\mathbf{#1}) }
@@
 
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-\usepackage{url}
-\usepackage[colorlinks=true,linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
-\usepackage{doi}
-
 % correct bad hyphenation here
 \hyphenation{op-tical net-works semi-conduc-tor an-isot-ropy regis-tra-tion iso-tropic Free-Surfer %
-align-ed pre-sent work-flow data-set data-sets}
+align-ed pre-sent work-flow data-set data-sets micro-struc-tu-re}
 
@@
 
+\ifthenelse{\boolean{review}}
+{
+  \usepackage[colorinlistoftodos,textsize=tiny, textwidth=1.2cm]{todonotes}
+  \setlength{\marginparwidth}{1.5cm}
+  \reversemarginpar
+  % \setlength{\marginparwidth}{2.0cm}
+  \newcommand{\revcomment}[2][]{\ifthenelse{\equal{#1}{}}{}{\todo{#1}}%
+  \begingroup \color{blue!70} #2\endgroup}
+}
+{
+  \newcommand{\todo}[2][]{}
+  \newcommand{\revcomment}[2][]{#2}
+}
+
 % http://tex.stackexchange.com/questions/56105/missing-references-heading
 \def\bibsection{\section*{References}}
+
 \begin{document}
+\hypersetup{linkcolor=black!60, citecolor=black!60, urlcolor=black!60}
 \begin{frontmatter}

### 2015-10-26 Publications/2015-NeuroImage/02-abstract.tex
v1.3.0rc

--- 2015-09-30 Publications/2015-NeuroImage/02-abstract.tex
@@
 % -*- root: 00-main.tex -*-
-\newcomment[RV\#1(C.1)]{%
+\revcomment[RV\#1(C.1)\par RV\#2(C.10)]{%
 Current methods using diffusion MRI to map the microstructure and connectivity of the
@@
   randomly generated deformations, where subvoxel accuracy was achieved.
-\newcomment[RV\#2(C.11)]{%
+\revcomment[RV\#2(C.11)]{%
 We proposed a set of surfaces that define a segmentation model of the fractional anisotropy and

### 2015-10-26 Publications/2015-NeuroImage/03-introduction.tex
v1.3.0rc

@@
 Each \gls*{dwi} is sensitized with a gradient to probe proton diffusion in a certain
-  \newcomment[RV\#2(C.13)]{orientation}.
+  \revcomment[RV\#2(C.13)]{orientation}.
 Subsequent processing involves describing the local microstructure with one available
@@
   to current models such as \citep{daducci_accelerated_2015}.
-The microstructural map is then used to draw the preferential \newcomment[RV\#2(C.13)]{orientations} of diffusion
+The microstructural map is then used to draw the preferential \revcomment[RV\#2(C.13)]{orientations} of diffusion
   across the brain using tractography \citep{mori_threedimensional_1999}.
@@
   tractography as edges \citep{hagmann_mapping_2008}.
-\newcomment[RV\#2(C.17)]{%
+\revcomment[RV\#2(C.17)]{%
 The methodologies to solve reconstruction, tractography and network building
@@
 
-\newcomment[RV\#2(C.15)]{%
+\revcomment[RV\#2(C.15)]{%
 Possibly, the earliest structural information incorporated to aid \gls*{dmri} processing
@@
 However, the precise segmentation of \gls*{dmri} is difficult for several reasons.
-First, \gls{dmri} images have a resolution that is \newcomment[RV\#2(C.12)]{much lower}
+First, \gls{dmri} images have a resolution that is \revcomment[RV\#2(C.12)]{much lower}
   than that of the imaged microstructural features.
@@
 
-\newcomment[RV\#2(C.9)]{%
+\revcomment[RV\#2(C.9)]{%
 An alternative route to segmentation in native \gls*{dmri} space is the mapping of the
@@
 
-\newcomment[RV\#2(C.10)]{%
+\revcomment[RV\#2(C.10)]{%
 Here we present a segmentation and surface-to-volume registration method 

### 2015-10-26 Publications/2015-NeuroImage/04-methods.tex
v1.3.0rc

@@
 The best estimate $\hat{U}$ then satisfies the maximum a posteriori criterion
-  \citep{bishop_pattern_2006} and it aligns $\gammaset_R$ into $M$.
+  and it aligns $\gammaset_R$ into $M$.
 First, we assume independence between pixels, and thus we break down the
@@
   1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
-  distributions with the parameters $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
+  distributions \citep{esteban_mbis_2014} with parameters
+  $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
   for each region $\Omega_l$.
-\newcomment[RV\#1(C.4)]{%
+\revcomment[RV\#1(C.4)]{%
 Even though the features being segmented are not generally i.i.d., the spatial interdependency of
@@
   the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized.
-\newcomment[RV\#1(C.2)]{%
+\revcomment[RV\#1(C.2)]{%
 In \autoref{fig:regseg-model},
@@
 
-\begin{figure}
+\begin{figure*}
 \includestandalone[width=\linewidth]{figure02-model}
@@
 \caption{%
-\newcomment{%
+\revcomment{%
 Evolution of the segmentation model defined by the homogeneous regions $\Omega_l$
@@
 \todo[inline]{RV\#1(C.5)}
-\end{figure}
+\end{figure*}
 
@@
   following sections.
-\newcomment[RV\#2(C.20)]{%
+\revcomment[RV\#2(C.20)]{%
 A first proof of concept to demonstrate \regseg{} in phantoms with simple
@@
 
-\newcomment[RV\#1(C.6)]{\paragraph*{Segmentation model and settings}
+\revcomment[RV\#1(C.6)]{\paragraph*{Segmentation model and settings}
 The segmentation model of the phantoms is implicitly defined in each shape.
@@
   5) The results were evaluated visually and quantitatively.
-  \newcomment{%
+  \revcomment{%
   The arrows point to edges in the target images (yellow arrows for \gls*{fa}, blue for \gls*{adc}) that
@@
   multi-shell \gls*{dmri} images.
-\newcomment[RV\#1(C.7)]{%
+\revcomment[RV\#1(C.7)]{%
 Since we obtained the \gls*{dmri} data from the minimally preprocessed package, these
@@
 \paragraph*{\Regseg{} settings} %
-\newcomment[RV\#1(C.5)]{%
+\revcomment[RV\#1(C.5)]{%
 \Regseg{} accepts an affine mapping from surface-space to the \gls*{dmri} data as initialization.
@@
 Therefore, the initial estimation of distortion is zero in this experiment.}
-\newcomment[RV\#1(C.20)]{%
+\revcomment[RV\#1(C.20)]{%
 Since the distortion $U_\text{true}$ is aligned along the \gls*{pe} direction ($y$-axis in our settings),
   \regseg{} was configured to allow nonzero displacements only on that corresponding direction.}
-\newcomment[RV\#1(C.6)]{%
+\revcomment[RV\#1(C.6)]{%
 For the experiments on real data, \regseg{} established a multi-resolution pyramid of
@@
   perform registration.
-\newcomment[RV\#2(C.21)]{%
+\revcomment[RV\#2(C.21)]{%
 The deformation field is correspondingly restricted to the \gls*{pe} direction.

### 2015-10-26 Publications/2015-NeuroImage/06-discussion.tex
v1.3.0rc

@@
   the affine transformation model.
-The deformation model of \regseg{} is nonlinear and the surfaces are active
+Conversely, the deformation model of \regseg{} is nonlinear and the surfaces are active
   contours \emph{without edges} \citep{chan_active_2001}.
@@
   projected contours obtained using the ground-truth warping and our estimates.
-We concluded that the errors were significantly lower than the voxel resolution.
+We concluded that the errors were significantly lower than the voxel size.
 We also assessed the 95\% \gls*{ci}, which yielded an aggregate interval of
@@
 Therefore, we also concluded that the error was bounded above by half of the
-  voxel spacing.
-\newcomment[RV\#1(C.8)]{%}
+  voxel size.
+\revcomment[RV\#1(C.8)]{%}
 The distributions of errors along surfaces vary importantly depending on the shape of the
@@
 \todo{real data}
-This application is necessary in the extraction of structural connectivity from
-  \gls*{dmri}, and has been typically solved in a two-step approach.
+Once the method was validated, we applied it in the task of identifying or incorporating
+  structural information in 16 real \gls*{dmri} datasets.
+The segmentation of some anatomical structures in the diffusion dataset is necessary in
+  current methods involved in the connectome extraction, such as the structure-informed 
+  reconstruction of \gls*{dmri} data \citep{jeurissen_multitissue_2014,daducci_accelerated_2015},
+  the anatomically constrained tractography \citep{smith_anatomicallyconstrained_2012},
+  and the imposition of the cortical parcellation mapped from the \gls*{t1} image 
+  \citep{hagmann_mapping_2008}.
+The problem was firstly addressed using image segmentation approaches in the native diffusion
+  space, without definitive and compelling results.
+With the introduction of retrospective correction methods for the \emph{\gls*{epi} distortions}
+  and image registration approaches, the task has been typically solved in a two-step approach.
 First, the \glspl*{dwi} are corrected for \emph{\gls*{epi} distortions} by estimating
@@
   using a linear registration tool like \emph{bbregister} \citep{greve_accurate_2009}.
-numerous attempts to delineate the structures of interest in diffusion
-  space have been proposed.
-Current activity on improving correction methods \citep{irfanoglu_drbuddi_2015} and
-  segmenting the \gls*{dmri} in its native space \citep{jeurissen_tissuetype_2015}
-  proofs the open interest of the application.
-\todo{real data old}
-We designed \regseg{} as a method for segmenting \gls*{dmri} data by exploiting the
-  anatomical information extracted from a structural image (typically a \gls*{t1} image)
-  of the subject.
-Current pipelines using whole-brain tractography usually solve this problem with a two-step approach.
-First, the images are corrected for nonlinear distortions using auxiliary acquisitions
-  such as fieldmaps \citep{jezzard_correction_1995}, \glspl*{dwi} with reversed \gls*{pe}
-  blips \citep{chiou_simple_2000}, or \gls*{t2} images \citep{kybic_unwarping_2000}.
-Second, the segmentation is projected from a reference \gls*{t1} image using linear
-  registration \citep{greve_accurate_2009}.
+The current activity on improving correction methods \citep{irfanoglu_drbuddi_2015} and
+  the comeback of segmentation of \gls*{dmri} in its native space 
+  \citep{jeurissen_tissuetype_2015} proof the open interest of this application.
+
 \Regseg{} addresses this joint problem in a single step and it does not require any additional
@@
 This situation is found commonly in historical datasets.
-Moreover, since the structural information is projected into the native space of \gls*{dmri}
+Moreover, since the structural information is projected into the native space of \gls*{dmri},
   reconstruction and tractography can be performed without resampling data to an undistorted
   space.
-
 We evaluated \regseg{} in a real environment using the experimental framework presented
@@
 Visual inspections of all the results (\suppl{section S5}) and the violin plots in
-  \autoref{fig:regseg-results_real} confirmed that \regseg{} performed better than the \gls*{t2b} method
-  in our settings.
+  \autoref{fig:regseg-results_real} confirmed that \regseg{} achieved higher accuracy
+  than the \gls*{t2b} method in our settings.
 We carefully configured the \gls*{t2b} method using the same algorithm and the
-  same settings employed in a widely-used tool.
-However, cross-comparison experiments are prone to so-called \emph{instrumentation bias}
+  same settings employed in a widely-used tool for \gls*{dmri} processing.
+However, cross-comparison experiments are prone to the so-called \emph{instrumentation bias}
   \citep{tustison_instrumentation_2013}.
-Therefore, these results do not prove that \regseg{} \emph{is better than} \gls*{t2b}.
-Our results suggest that \regseg{} is a reliable option in this application field.
-In addition, the \gls*{t2b} may introduce an additional (and small) error during the necessary
-  registration of \gls*{t2} in the \gls*{t1} space.
-
-\newcomment[RV\#1(C.3)]{%
-\paragraph*{Contributions}\label{sec:regseg-related_work} %
-
-}
-
-\newcomment[RV\#1(C.5)]{%
-Finally, we also contribute proposing a piecewise-smooth segmentation model defined by
+Therefore, these results did not prove that \regseg{} \emph{is better than} \gls*{t2b},
+  but indicated that \regseg{} is a reliable option in this application field.
+\revcomment[RV\#1(C.5)]{%
+Finally, we also contributed proposing a piecewise-smooth segmentation model defined by
   a selection of nested surfaces to partition the multispectral space
   comprehending the \gls*{fa} and the \gls*{adc} maps and ultimately identify these
-  structures in \gls*{dmri} space.}
-\newcomment[New]{%
-With the definition of appropriate models, \regseg{} may be applied in other fields like
-  neonatal brain image segmentation in longitudinal \gls*{mri} studies like \citep{shi_neonatal_2010}
-  in which the surfaces obtained in a mature timepoint of the brain can be retrospectively
+  structures in \gls*{dmri} space.
+We also demonstrated the smoothness of the objective function on five of the real datasets
+  (\suppl{Figure S2}), taking advantage of the directional restriction of possible distortions.
+However, convergence of \regseg{} can be severely affected when surfaces are dense enough.
+Using the digital phantoms, we decimated the surfaces with large factors and
+  effectively displaced the minimum of the shape-gradients from the minimum of the
+  objective function.
+Thus, the surfaces must be densely sampled to enable the use of the presented method.
+}%
+
+Beyond the presented application on \gls*{dmri} data, \regseg{} can be indicated in situations
+  where there are precise surfaces delineating the structure, a target multivariate
+  image in which the surfaces must be fitted, and the mapping between the surfaces and
+  the volume encodes a relevant physiological information, such as the normal/abnormal
+  development or the macroscopic dynamics of organs and tissues.
+For instance, \regseg{} may be applied in fields like neonatal brain image segmentation
+  in longitudinal MRI studies of the early developmental patterns as \citep{shi_neonatal_2010}.
+In these studies, the surfaces obtained in a mature timepoint of the brain are retrospectively
   propagated to the initial timepoints, regardless the changes of the contrast and spatial
-  development between timepoints.}
+  development between them.
+More generally, \regseg{} may also be applied on the personalized study of longitudinal alteration
+  of the brain using multispectral images, for instance in the case of traumatic brain
+  injury \citep{irimia_structural_2014} or in the development of brain tumors
+  \citep{weizman_semiautomatic_2014}.
 

### 2015-10-26 Publications/2015-NeuroImage/02-abstract.tex
new abstract!

--- 2015-09-30 Publications/2015-NeuroImage/02-abstract.tex
@@
 % -*- root: 00-main.tex -*-
-\newcomment[RV\#1(C.1)]{%
-Current methods using diffusion MRI to map the microstructure and connectivity of the
-  human brain increasingly require precise delineations of anatomical structures.
-The problem is typically solved by projecting the structural information extracted
-  in T1-weighted images, by means of nonlinear registration since diffusion
-  MRI data present geometrical distortions.
-Here we propose \regseg{}, which segments homogeneous regions within multivariate
-  images by projecting a set of nested surfaces extracted from structural images.}
-The surfaces evolve through a free-form deformation field supported by the B-spline basis
-  to optimally map the contours onto the data in the target space.
-We tested the functionality of \regseg{} using four digital phantoms warped with known and
-  randomly generated deformations, where subvoxel accuracy was achieved.
-\newcomment[RV\#2(C.11)]{%
-We proposed a set of surfaces that define a segmentation model of the fractional anisotropy and
-  the apparent diffusion coefficient maps derived from diffusion MRI datasets.}
-These surfaces extracted from the same-subject T1-weighted image successfully segmented 16 real
-  diffusion MRI datasets from the Human Connectome Project that were warped by realistic and nonlinear
-  distortions.
-We computed the misregistration error of the surfaces estimated by \regseg{} with respect
-  to their theoretical location using the ground truth, thereby obtaining a 95\% CI of 0.56--0.66 mm
-  distance between corresponding mesh vertices, which was below the 1.25 mm isotropic resolution of the images.
-We also compared the performance of the proposed method with a widely used registration tool, which showed
-  that \regseg{} outperformed this method in our settings.
+\revcomment[RV\#1(C.1)\par RV\#2(C.10)]{%
+Current methods for processing \gls*{dmri} to map the connectivity of the human brain
+  require precise delineations of anatomical structures.
+Image segmentation on the native \gls*{dmri} with sub-pixel accuracy has been unsuccessful,
+  and the projection of the structural information from \gls*{t1} images is challenging
+  since \gls*{dmri} data present geometrical distortions.
+To solve this problem, we propose \regseg{}, a surfaces-to-volume nonlinear registration method
+  that segments homogeneous regions within multivariate images by projecting a set of nested
+  reference-surfaces.}
+We first verify the subpixel accuracy of \regseg{} on digital phantoms,
+\revcomment[RV\#2(C.11)]{%
+and then we evaluate the tool in the aforementioned application.
+We propose a set of surfaces extracted from the \gls*{t1} image to successfully segment
+  the bivariate volume comprehending the \gls*{fa} and the \gls*{adc} maps derived from the
+  corresponding \gls*{dmri} dataset of the same subject.}
+We present an evaluation framework, in which the target \gls*{dmri} are deformed with
+  magnetic susceptibility -derived and realistic distortions.
+Using the framework, we analyze the misregistration error of the surfaces estimated by \regseg{}
+  on 16 real datasets of the Human Connectome Project, thereby obtaining an error distribution
+  with 95\% CI of 0.56--0.66 mm, below the \gls*{dmri} resolution (1.25 mm, isotropic).
+Finally, we cross-compare \regseg{} against a widely-used volume-to-volume registration method,
+  and we obtain a significantly lower misregistration error with the proposed tool.

### 2015-10-26 Publications/2015-NeuroImage/07-conclusion.tex
new abstract!

@@
 We considered the reproducibility of our results as a design requirement.
-Therefore, we used real data obtained from a publicly available repository
-  (the Human Connectome Project \citep{essen_human_2012}) and all of the software
-  utilized in this study is also publicly available.
+Therefore, we used real data obtained from the Human Connectome Project \citep{essen_human_2012}
+  and all of the software utilized in this study is also publicly available.
 \Regseg{} was implemented on top of ITK-4.6 (Insight Registration and

### 2015-10-27 Publications/2015-NeuroImage/03-introduction.tex
fix errors and typos

@@
 \section{Introduction}\label{sec:regseg-intro}
-\Gls*{dmri} enables the mapping of microstructure \citep{basser_microstructural_1996}
+\Acrlong*{dmri} enables the mapping of microstructure \citep{basser_microstructural_1996}
   and connectivity \citep{craddock_imaging_2013} of the human brain \emph{in-vivo}.
@@
 Finally, a graph representing the corresponding structural network is built using
-  the regions of a cortical parcellation as nodes and the streamlines of
+  the regions of a cortical parcellation as nodes and the fiber paths found by
   tractography as edges \citep{hagmann_mapping_2008}.
@@
   their direct use in segmentation.
-Third, the low contrast between \gls*{gm} and \gls*{wm} in the \lowb{} volumes also makes
-  them unsuitable for brain tissue segmentation.
+Third, the low contrast between \gls*{gm} and \gls*{wm} in the \lowb{} volume also makes
+  it unsuitable for brain tissue segmentation.
 
@@
   the corresponding \gls*{t2} image of the same subject.
-This method is a derivation of the \gls*{t2b} correction method, reproducing the settings
-  of an existing implementation included within a widely used diffusion processing software
-  \citep{leemans_exploredti_2009}.
-With this comparison we demonstrate how \regseg{} achieves higher accuracy with the
+This method is a derivation of the \gls*{t2b} correction methods \citep{kybic_unwarping_2000},
+  reproducing the settings of an existing implementation included within a widely used diffusion
+  processing software (\emph{ExploreDTI}, \cite{leemans_exploredti_2009}).
+With this comparison, we demonstrate how \regseg{} achieves higher accuracy with the
   simultaneous registration and segmentation process.

### 2015-10-27 Publications/2015-NeuroImage/06-discussion.tex
fix errors and typos

@@
   images into the corresponding \gls*{dmri} of the same subject.
-Previously, joint segmentation and registration have been applied successfully to
+Previously, joint segmentation and registration has been applied successfully to
   other problems such as longitudinal object tracking \citep{paragios_level_2003}
@@
   like \cite{paragios_level_2003,yezzi_variational_2003}.
-\Regseg{} can be seen as particular case of atlas-based segmentation, replacing the atlas
-  by the structural image of the subject.
+\Regseg{} can be seen as a particular case of atlas-based segmentation, replacing the atlas
+  by the structural image of the subject (\emph{structure-informed segmentation}).
 However, the majority of atlas-based segmentation methods, such as \citep{gorthi_active_2011},
@@
 The \lowb{} image only includes a detectable frontier for the pial surface, and thus
-  \emph{bbregister} is limited to aligning the cortical layer in this
-  application.
+  \emph{bbregister} is limited to align the pial contour in this application.
 To overcome the problem of nonlinear distortions, \emph{bbregister} excludes the
@@
   the affine transformation model.
-Conversely, the deformation model of \regseg{} is nonlinear and the surfaces are active
-  contours \emph{without edges} \citep{chan_active_2001}.
+Conversely, the deformation model of \regseg{} is nonlinear and the active contours are
+  \emph{without-edges} \citep{chan_active_2001}.
 The contours are without edges because the \gls*{fa} and \gls*{adc} maps do
@@
 \revcomment[RV\#1(C.8)]{%}
-The distributions of errors along surfaces vary importantly depending on the shape of the
+The distributions of errors along surfaces varied importantly depending on the shape of the
   phantom (see \autoref{fig:regseg-phantom}B).
@@
 We argue that the symmetry of those other shapes posed difficulties in driving the contours
-  towards the appropriate region and producing some ``sliding'' effect between the
-  faces of the surfaces and their ground-truth position.
+  towards the appropriate region and produced a \emph{sliding} between the faces of the
+  surfaces and their ground-truth position.
 This effect should not be present in real datasets, thanks to the very convoluted cortical
@@
 The problem was firstly addressed using image segmentation approaches in the native diffusion
-  space, without definitive and compelling results.
+  space, without definite and compelling results.
 With the introduction of retrospective correction methods for the \emph{\gls*{epi} distortions}
@@
   \citep{jeurissen_tissuetype_2015} proof the open interest of this application.
-
 \Regseg{} addresses this joint problem in a single step and it does not require any additional
-  acquisition other than the minimal protocol using only \gls*{t1} and \gls*{dmri} images.
+  acquisition other than the minimal protocol comprehending only \gls*{t1} and \gls*{dmri} images.
 This situation is found commonly in historical datasets.
@@
   space.
+
 We evaluated \regseg{} in a real environment using the experimental framework presented
@@
   (\suppl{Figure S2}), taking advantage of the directional restriction of possible distortions.
-However, convergence of \regseg{} can be severely affected when surfaces are dense enough.
+However, convergence of \regseg{} can be severely affected when surfaces are not dense enough.
 Using the digital phantoms, we decimated the surfaces with large factors and
@@
 For instance, \regseg{} may be applied in fields like neonatal brain image segmentation
-  in longitudinal MRI studies of the early developmental patterns as \citep{shi_neonatal_2010}.
+  in longitudinal MRI studies of the early developmental patterns \citep{shi_neonatal_2010}.
 In these studies, the surfaces obtained in a mature timepoint of the brain are retrospectively
@@
   development between them.
-More generally, \regseg{} may also be applied on the personalized study of longitudinal alteration
+More generally, \regseg{} may also be applied to the personalized study of longitudinal alteration
   of the brain using multispectral images, for instance in the case of traumatic brain

### 2015-10-29 Publications/2015-NeuroImage/02-abstract.tex
one more iteration for resubmission

@@
   require precise delineations of anatomical structures.
-Image segmentation on the native \gls*{dmri} with sub-pixel accuracy has been unsuccessful,
-  and the projection of the structural information from \gls*{t1} images is challenging
-  since \gls*{dmri} data present geometrical distortions.
-To solve this problem, we propose \regseg{}, a surfaces-to-volume nonlinear registration method
-  that segments homogeneous regions within multivariate images by projecting a set of nested
-  reference-surfaces.}
-We first verify the subpixel accuracy of \regseg{} on digital phantoms,
+The requirement has been approached either segmenting the data in native \gls*{dmri} space
+  or projecting the structural information from \gls*{t1} images.
+The characteristic features of diffusion data in terms of \acrlong*{snr}, resolution, etc.
+  and the geometrical distortions caused by the inhomogeneity of magnetic susceptibility
+  across tissues render the problem challenging.
+To solve the problem in a unified approach, we propose \regseg{}, a surface-to-volume nonlinear
+  registration method that segments homogeneous regions within multivariate images by mapping
+  a set of nested reference-surfaces.}
 \revcomment[RV\#2(C.11)]{%
-and then we evaluate the tool in the aforementioned application.
-We propose a set of surfaces extracted from the \gls*{t1} image to successfully segment
-  the bivariate volume comprehending the \gls*{fa} and the \gls*{adc} maps derived from the
-  corresponding \gls*{dmri} dataset of the same subject.}
-We present an evaluation framework, in which the target \gls*{dmri} are deformed with
-  magnetic susceptibility -derived and realistic distortions.
-Using the framework, we analyze the misregistration error of the surfaces estimated by \regseg{}
-  on 16 real datasets of the Human Connectome Project, thereby obtaining an error distribution
-  with 95\% CI of 0.56--0.66 mm, below the \gls*{dmri} resolution (1.25 mm, isotropic).
-Finally, we cross-compare \regseg{} against a widely-used volume-to-volume registration method,
-  and we obtain a significantly lower misregistration error with the proposed tool.
+The surfaces can be extracted from the \gls*{t1} image of the subject, and the target image
+  is the bivariate volume comprehending the \gls*{fa} and the \gls*{adc} maps derived from the
+  \gls*{dmri} dataset.}
+We first verify the subpixel accuracy of \regseg{} on digital phantoms, to then propose an
+  evaluation framework that uses \gls*{dmri} datasets from the \gls*{hcp}.
+In this framework, the undistorted \gls*{dmri} undergo a known distortion derived from real
+  fieldmaps.
+We analyze the misregistration error of the surfaces estimated by \regseg{} on 16 datasets.
+The distribution of errors shows a 95\% CI of 0.56--0.66 mm, below the \gls*{dmri} resolution
+  (1.25 mm, isotropic).
+Finally, we cross-compare the proposed tool against a nonlinear \lowb{}-to-T2w registration
+  method, thereby obtaining a significantly lower misregistration error with \regseg{}.

### 2015-10-29 Publications/2015-NeuroImage/03-introduction.tex
one more iteration for resubmission

@@
 \revcomment[RV\#2(C.9)]{%
-An alternative route to segmentation in native \gls*{dmri} space is the mapping of the
+An alternative route to segmentation in \gls*{dmri} space is the mapping of the
   structural information extracted from anatomical MR images, such as \gls*{t1}, using
@@
 However, \gls*{epi} introduces a geometrical distortion \citep{jezzard_correction_1995}
-  that impedes the linear mapping from the structural space and warps the anatomy
-  identified on native \gls*{dmri} space \citep{irfanoglu_effects_2012}.
+  that impedes the linear mapping from the structural space.
 Numerous methods have been proposed to overcome this problem by incorporating information
@@
 Here we present a segmentation and surface-to-volume registration method
-  named \regseg{}, and show its usefulness in incorporating the anatomical
-  information into native \gls*{dmri} space to aid subsequent reconstruction,
-  tractography and network building using a cortical parcellation.
-The hypothesis underlying is that the registration and segmentation problems
+  named \regseg{}, and show its usefulness in mapping anatomical information from structural
+  into native \gls*{dmri} space to aid subsequent processing steps
+  (reconstruction, tractography and network building using a cortical parcellation).
+The underlying hypothesis is that the registration and segmentation problems
   in \gls*{dmri} can be solved simultaneously.
@@
   \citep{chan_active_2001} segmentation framework.
-Several \emph{anatomically-correct} surfaces of the same subject initialize
-  the 3D active-contours, which evolve searching for homogeneous regions in
-  the multivariate target-image.
-In the presented application to segment \gls*{dmri} data, we propose a bivariate
-  target-volume comprehending the \gls*{fa} and \gls*{adc} maps.
+Several reference surfaces of the same subject initialize the 3D active-contours,
+  which evolve searching for homogeneous regions in the multivariate target-image.
+In the presented application to segment \gls*{dmri} data, we propose to use \regseg{}
+  using a set of nested surfaces extracted from a strutrucal image (e.g. \gls*{t1}) and
+  a bivariate target-volume comprehending the \gls*{fa} and \gls*{adc} maps as inputs.
 The evolution of the surfaces is supported by a B-spline basis, optimized

### 2015-10-29 Publications/2015-NeuroImage/04-methods.tex
one more iteration for resubmission

--- 2015-10-26 Publications/2015-NeuroImage/04-methods.tex
@@
 
-\paragraph*{\Regseg{} settings} %
+\paragraph*{Settings} %
 \revcomment[RV\#1(C.5)]{%
@@
 
-\paragraph*{Cross-comparison and settings} %
+\paragraph*{Cross-comparison} %
 A dual workflow to the general evaluation framework used for \regseg{} (\autoref{fig:regseg-evworkflows}),

### 2015-10-29 Publications/2015-NeuroImage/06-discussion.tex
one more iteration for resubmission

@@
   by the current mapping of the surfaces.
-We propose \regseg{} to incorporate anatomical information extracted from \gls*{t1}
+We propose \regseg{} to map anatomical information extracted from \gls*{t1}
   images into the corresponding \gls*{dmri} of the same subject.
@@
   and atlas-based segmentation \citep{gorthi_active_2011}.
-The most common approach to solve the problem simultaneously is optimizing a deformation
-  model (registration) that supports the evolution the active contours (segmentation),
+The most common approach involves optimizing a deformation model (registration)
+  that supports the evolution of the active contours (segmentation),
   like \cite{paragios_level_2003,yezzi_variational_2003}.
-\Regseg{} can be seen as a particular case of atlas-based segmentation, replacing the atlas
-  by the structural image of the subject (\emph{structure-informed segmentation}).
-However, the majority of atlas-based segmentation methods, such as \citep{gorthi_active_2011},
-  use a multiphase level set function to register several active contours.
+\Regseg{} can be seen as a particular case of atlas-based segmentation-registration methods,
+  replacing the atlas by the structural image of the subject (\emph{structure-informed segmentation}).
+\cite{gorthi_active_2011} review the application of joint segmentation and registration in atlas-based
+  segmentation, and propose a multiphase level set function to register several active contours.
 Alternatively, \regseg{} implements the active contours with a hierarchical set of explicit
-  surfaces (triangular meshes) that substitute the multiphase level sets, and registration
+  surfaces (triangular meshes) instead of the multiphase level sets, and registration
   is driven by shape-gradients \citep{herbulot_segmentation_2006}.
@@
   correction of choice.
-It implements a linear mapping, and uses 3D active contours \emph{with edges} to
+It implements a linear mapping and uses 3D active contours \emph{with edges} to
   search for intensity boundaries in the \lowb{} image.
 The active contours are initialized using surfaces extracted from the \gls*{t1} using
-  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}: the \emph{pial} surface (exterior of the
-  cortical \gls*{gm}) and the \emph{white} surface (the \gls*{wm}/\gls*{gm} interface).
-The \lowb{} image only includes a detectable frontier for the pial surface, and thus
-  \emph{bbregister} is limited to align the pial contour in this application.
-To overcome the problem of nonlinear distortions, \emph{bbregister} excludes the
-  regions that are typically warped from the boundary search.
+  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}.
+To overcome the problem of nonlinear distortions, \emph{bbregister} excludes from the
+  boundary search those regions that are typically warped.
 Indeed, the distortion must be addressed separately because it is not supported by
@@
 Conversely, the deformation model of \regseg{} is nonlinear and the active contours are
-  \emph{without-edges} \citep{chan_active_2001}.
-The contours are without edges because the \gls*{fa} and \gls*{adc} maps do
-  not present steep image gradients (edges) but the anatomy can be identified
+  \emph{without-edges} \citep{chan_active_2001} because the \gls*{fa} and \gls*{adc} maps
+  do not present steep image gradients (edges) but the anatomy can be identified
   by looking for piece-wise smooth homogeneous regions.
@@
   and \gls*{t2} images at two resolution levels.
-After registration with \regseg{}, we measured the Hausdorff distance between the
-  projected contours obtained using the ground-truth warping and our estimates.
+We measured the Hausdorff distance between the contours projected using the
+  ground-truth warping and the estimations found with \regseg{}.
 We concluded that the errors were significantly lower than the voxel size.
@@
   0.34--0.38 [mm] for the high resolution phantoms (1.0 mm isotropic).
-Therefore, we also concluded that the error was bounded above by half of the
-  voxel size.
+Therefore, the error was bounded above by half of the voxel size.
 \revcomment[RV\#1(C.8)]{%}
@@
 We argue that the symmetry of those other shapes posed difficulties in driving the contours
-  towards the appropriate region and produced a \emph{sliding} between the faces of the
+  towards the appropriate region due to \emph{sliding} displacements between the
   surfaces and their ground-truth position.
-This effect should not be present in real datasets, thanks to the very convoluted cortical
-  layer, and the directional restriction of distortion.}
-
-\todo{real data}
-Once the method was validated, we applied it in the task of identifying or incorporating
-  structural information in 16 real \gls*{dmri} datasets.
-The segmentation of some anatomical structures in the diffusion dataset is necessary in
-  current methods involved in the connectome extraction, such as the structure-informed
-  reconstruction of \gls*{dmri} data \citep{jeurissen_multitissue_2014,daducci_accelerated_2015},
-  the anatomically constrained tractography \citep{smith_anatomicallyconstrained_2012},
-  and the imposition of the cortical parcellation mapped from the \gls*{t1} image
-  \citep{hagmann_mapping_2008}.
-The problem was firstly addressed using image segmentation approaches in the native diffusion
-  space, without definite and compelling results.
-With the introduction of retrospective correction methods for the \emph{\gls*{epi} distortions}
-  and image registration approaches, the task has been typically solved in a two-step approach.
-First, the \glspl*{dwi} are corrected for \emph{\gls*{epi} distortions} by estimating
-  the nonlinear-deformation field from extra MR acquisitions
-  \citep{jezzard_correction_1995,chiou_simple_2000,cordes_geometric_2000,
-  kybic_unwarping_2000}.
-Second, mapping the structural information from the corresponding \gls*{t1} image
-  using a linear registration tool like \emph{bbregister} \citep{greve_accurate_2009}.
-The current activity on improving correction methods \citep{irfanoglu_drbuddi_2015} and
-  the comeback of segmentation of \gls*{dmri} in its native space
-  \citep{jeurissen_tissuetype_2015} proof the open interest of this application.
-\Regseg{} addresses this joint problem in a single step and it does not require any additional
-  acquisition other than the minimal protocol comprehending only \gls*{t1} and \gls*{dmri} images.
-This situation is found commonly in historical datasets.
-Moreover, since the structural information is projected into the native space of \gls*{dmri},
-  reconstruction and tractography can be performed without resampling data to an undistorted
-  space.
+The effect is not detectable by the active contours framework, but it is controllable
+  increasing the regularization constraints.
+When \regseg{} is applied on real datasets, this surface sliding is negligible for the
+  convoluted nature of cortical surfaces and the directional restriction of the
+  distortion.}
 
@@
 
+The proposed application of the method in the task of identifying structural information
+  in \gls*{dmri} images is an active field of research \citep{jeurissen_tissuetype_2015}.
+Current processing steps involved in the connectome extraction require a precise segmentation
+  of anatomical structures in the diffusion space.
+Some examples of these processing tasks are the structure-informed reconstruction of \gls*{dmri}
+  data \citep{jeurissen_multitissue_2014,daducci_accelerated_2015}, the anatomically constrained
+  tractography \citep{smith_anatomicallyconstrained_2012}, and the imposition of the cortical
+  parcellation mapped from the \gls*{t1} image \citep{hagmann_mapping_2008}.
+The problem was firstly addressed using image segmentation approaches in the native diffusion
+  space, without definite and compelling results.
+With the introduction of retrospective correction methods for the \emph{\gls*{epi} distortions}
+  and image registration approaches, the task has been typically solved in a two-step approach.
+First, the \glspl*{dwi} are corrected for \emph{\gls*{epi} distortions} by estimating
+  the nonlinear-deformation field from extra MR acquisitions
+  \citep{jezzard_correction_1995,chiou_simple_2000,cordes_geometric_2000,kybic_unwarping_2000}.
+Second, mapping the structural information from the corresponding \gls*{t1} image
+  using a linear registration tool like \emph{bbregister} \citep{greve_accurate_2009}.
+The current activity on improving correction methods \citep{irfanoglu_drbuddi_2015} and
+  the comeback of segmentation of \gls*{dmri} in its native space
+  \citep{jeurissen_tissuetype_2015} proof the open interest of this application.
+\Regseg{} addresses this joint problem in a single step and it does not require any additional
+  acquisition other than the minimal protocol comprehending only \gls*{t1} and \gls*{dmri} images.
+This situation is found commonly in historical datasets.
+Moreover, since the structural information is projected into the native space of \gls*{dmri},
+  the whole connectome extraction process can be performed without resampling data to an
+  undistorted space.
+
 Beyond the presented application on \gls*{dmri} data, \regseg{} can be indicated in situations
@@
   \citep{weizman_semiautomatic_2014}.
-
-% \paragraph*{Prospects}
-% First extensions of this work will study more appropriate features to build the energy functional
-%   on, enabling to perform \regseg{} directly on the raw \gls*{dmri} data.
-% A second outlook covers incorporating knowledge about the distortion by initializing our method
-%   with the theory-based displacement field that can be estimated with fieldmaps.

### 2015-10-29 Publications/2015-NeuroImage/responseLetter2.tex
final edits

--- 2015-09-30 Publications/2015-NeuroImage/responseLetter2.tex
@@
 We appreciate the comments made by both reviewers, and considered them fully to improve our paper.
-Please, find in the following our point-by-point answer to the concerns raised by the reviewers,
+Please, find in the following paragraphs our point-by-point answer to the concerns raised by the reviewers,
   along with the description of the corresponding changes in the manuscript.
-We have also changed the color of text for the modifications in the manuscript, along with side notes indicating which reviewer
-  and comment suggested the edition.
+Text modifications in the manuscript are highlighted with a new font color, and attached notes indicate which reviewer comment
+  the corresponding change is addressing.
 
@@
 \reply{%
-We decided to call the method ``simultaneous registration and segmentation'' because antecedent methods
-  using active contours that evolve driven by the optimization of a deformation field use that nomenclature.
-These methods are now summarized and discussed in lines 234--252 of the current version.
+The description of the method given by both reviewers (see \autoref{com:10}) lead us to think that the paper
+  does not describe correctly the actual inputs of the method.
+In general, we have almost-completely reformulated the abstract, introduction and discussion to define this
+  point clearly.
+Particularly, the following sections have been introduced give an answer to this particular comment:
+\begin{itemize}
+\item Lines 45--53 (p. 2).
+\item Lines 245--246 (p. 13), here we introduce the previously missing discussion to call the
+  method ``simultaneous registration and segmentation''.
+\item Lines 251--253 (p. 13), we comment here on the similarity of \emph{regseg} to the atlas-based segmentation
+  using joint registration and segmentation methods.
+\end{itemize}
 }
@@
 \reply{
-We have defined clearly that the choice of surfaces to segment the target data is crucial, and define the
-  segmentation model that drives the registration process (lines 85--88 are new).
-We have also improved former Figure 4, and brought it earlier in the paper (now it is Figure 2).
+We have stated more clearly that the choice of surfaces to segment the target data is crucial (lines 46--47, p.2).
+We have better defined the segmentation model that drives the registration process (lines 94--95, p. 3).
+We have also improved former Figure 4, and brought it earlier in the paper (now it is Figure 2, page 5).
 Since the former visualization of the conditional (region-wise) joint distribution of FA and ADC was a bit clumsy,
   we have extracted the conditional distributions to the right side of each panel.
+We expect these modifications make clear the difference between the surfaces involved in driving the registration
+  process and the surfaces in which we evaluate the error of the method (lines 218--219, p. 10) are a subset of the
+  former.
 }
@@
 We have moved the related methods and our methodological contributions from the Introduction to a more appropriate
-  place in the Discussion (ll. 234--252), additionally satisfying \autoref{com:9} suggesting to include in the introduction
+  place in the Discussion (lines 248--275), additionally satisfying \autoref{com:9} suggesting to include in the introduction
   only the references related to the focus of the paper.
+The particular changes can be listed as follows:
+\begin{itemize}
+  \item Lines 253--257 (p. 13), we describe the methodological differences w.r.t. \cite{gorthi_active_2011},
+    one of the latest atlas-based segmentation methods using simultaneous registration and segmentation.
+  \item Lines 258--266 (p. 13), we compare the methodological features of our method and \emph{bbregister},
+    an antecedent and standardized tool used on the same application.
+  \item Lines 268--273 (p. 13), we introduce the methodological differences with a recent alternative method in
+    terms of regularization.
+  \item Lines 273--275 (p. 13) mention some other minor methodological contributions.
+\end{itemize}
 }
@@
 The reviewer is right, assuming that pixels are i.i.d. is wrong but a widely accepted simplification.
-We have supported the assumptions inserting the new lines 80--83.}
+We have supported the assumptions inserting the new lines 88--90 (p. 3).}
 
@@
 \reply{
-We have improved former Figure 4 (now, it is Figure 2) to clarify that the segmentation model is inherently defined by
+We have improved former Figure 4 (now, it is Figure 2, page 5) to clarify that the segmentation model is inherently defined by
   the choice of surfaces and their nesting.
-This is an actual contribution of the work, as stated in ll. 264--266.
-
-The model behaves robustly when a) the target data are simple as in our phantom settings; and b) when the distortion
-  has only one nonzero component like in our real-data settings.
-The analysis of the goal function for the phantoms (without directional restrictions) was not necessary since
-  convergence was straigthforward.
-In the case of real data, since the distortion under investigation is aligned with one axis, the Figure S1 shows
-  that the goal function is well defined.
-
-However, the reviewer is right when raising convergence issues.
-We noticed that, when decimating the meshes in the phantom experiments, the zero of shape-gradients moved away from
-  the minimum of the goal function.
-Since decimation factors applied in this experiment were very large, and it did not affect our results with
-  the meshes directly obtained with \emph{Freesurfer} (in both phantoms and real data), we did include this discussion
-  within the paper.
-
-Finally, we commented on the initialization issue with the insertion of ll. 188--190.
+This is an actual contribution of the work, as stated in lines 301--304 (p. 14).
+A comment on the convergence issues follows in lines 304--309 (p. 14).
+Finally, we described the initialization of the surfaces with the insertion of lines 196--198 (p. 10).
 }
@@
 Therefore, parameter settings are publicly available in GitHub.
-Particularly, under folder \url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data}
-  there are several configuration files.
-Some parameters are also discussed in the Supplemental Materials, Section S2.
-
 However, we understand that a better description of the parameter settings should be included within
-  the manuscript and modified it accordingly, in ll. 192--196.
+  the manuscript and modified it accordingly, in lines 200--204 (p. 10).
 }
@@
 \reply{
-As commented in line 160, we are using some ``minimally preprocessed'' images from the database of the
-  Human Connectome Project (HCP).
-Particularly, the dMRI datasets from these preprocessed images are given corrected for distortions
-  and registered in structural space.
-Therefore, there could be a residual misregistration between the T1w and the dMRI data, but it should be
-  very minimal.
-We have modified former Figure 3 (now Fig. 4) to reflect this aspect.
+We have clarified this point with the following edits:
+\begin{itemize}
+  \item Lines 171--173 (p. 8): we indicate that we use the ``minimally preprocessed'' images of the
+    HCP database, and thus, these images are acquired with specialized protocols and are
+    already corrected for the distortion.
+  \item Figure 4 (former fig. 3), page 9, has been modified to clearly indicate this feature.
+\end{itemize}
 }
@@
 \reply{
-We have added a discussion on this in the corresponding section, lines 286-291. Thanks for pointing out.
+We have added a discussion on this in the corresponding section, lines 284-290. Thanks for pointing out.
 }
@@
 We thank the reviewer for pointing out that the overall goal is not clearly explained.
-We have replaced the term ``active-contours'' by ``surfaces'' whenever possible.
-We have re-organized the introduction fully, as suggested (see reply to \autoref{com:1} for a brief explanation).
-We have included references of the main EPI distortion corrections (see lines 31-43).
-
-The \emph{regseg} tool is designed to allow any displacement, as demonstrated with the phantoms.
-However, as the reviewer points out, the deformation field is completelly restricted to the phase-encode direction
-  for the real datasets.
-We have added ll. 125--127 to indicate that the ground-truth in the experiments done with real data is directionally
-  restricted. Then, in lines 190-192 to clarify that \emph{regseg} is configured to fulfill this feature.
+We have rewritten the abstract, introduction and discussion sections almost fully.
+Particular comments included within this general comment are addressed later: \autoref{com:18}
+  as regards EPI distortions, and \autoref{com:20} for the phase-encoding direction point.
+In order to better focus the application field, we have introduced lines 11--14 (p. 2) and lines
+  310--317 (p. 14).
+We have also added potential applications in the new lines 329--338 (pp. 14-15) to better illustrate
+  the method.
 }
@@
 \end{reviewpoint}
-\reply{The manuscript has been re-phrased accordingly in the title, abstract and introduction.}
+\reply{%
+In our adaptation of ``active contours without edges'' \citep{chan_active_2001}, the active contours are
+  ``extracted from structural images''.
+The particular sentence has been removed from the abstract, and we expect that the comprehensive reformulation
+  of the descriptive sections of the paper give a successful answer to this issue.
+Regarding the actual inputs of \emph{regseg}, we have introduce modifications in the abstract, lines 45--50 (p. 2), and
+  lines 244--245 (p. 13).
+}
 
@@
 \end{reviewpoint}
-\reply{We agree with the reviewer, we have updated the text abstract, and proposed a completely new graphical abstract.}
+\reply{We agree with the reviewer. We have proposed new graphical and text abstracts.}
 
@@
 \end{reviewpoint}
-\reply{We agreed, and updated the manuscript accordingly.}
+\reply{We agree and have updated the manuscript accordingly.}
 
@@
 \reply{
-The reviewer is right, we have changed the paragraph about dMRI segmentation (lines 15-27) to
-  be more precise.
+The reviewer is right, we have changed the paragraph about dMRI segmentation (lines 15--27, page 2)
+  for improved precision.
 We agree with the reviewer that there actually exist many methods to segment fiber tracts.
-However, our method is intended to support fiber tracking methods requiring detailed anatomical
-  information, such as \citep{smith_anatomicallyconstrained_2012}.
-We have supported the applications more clearly on lines 11-14.
-Clustering tractograms could be another potential application of \emph{regseg},
-  but we agree with the reviewer that current methods perform well without
-  additional structural information.
+We have supported the applications of the method more clearly on lines 11--14 (p. 2).
+Now, it is clearly described that fiber tracks are not available for the task at hand.
+
 }
@@
 This comment is very related to \autoref{com:14}.
-We have modified the abstract and introduction accordingly.
+We have modified the abstract and introduction accordingly, especially lines 15--27 (p. 2).
 }
@@
   the references in line 24 are a bit orphaned.
-All methods reviewed in lines 19-28 cluster/segment/classify \emph{voxels} in dMRI space based on
+All methods reviewed in lines 15--27 cluster/segment/classify \emph{voxels} in dMRI space based on
   one or another feature derived from dMRI data.
-We have clarified the manuscript according to \autoref{com:14} and \autoref{com:15}, and specifically
-  rewritten lines 19-28 to explicitly state what those methods are clustering.
+We have clarified the manuscript according to \autoref{com:14} and \autoref{com:15} and this comment.
 }
@@
   and changes include abstract and introduction.
-We have stated more clearly the importance of the set of surfaces being registered to produce a reliable
-  segmentation model of the two diffusion features selected: FA and ADC
+We better support the application of our method (lines 11--14, p. 2), and we have stated more clearly the importance of the set
+  of surfaces being registered to produce a reliable segmentation model of the two diffusion features selected: FA and ADC
+  (please, see \autoref{com:2}).
 }
@@
 \reply{
-We have added the suggested reference (among other of interest) in lines 204-206.
-In lines 59-61 we have stated the differences of \emph{regseg} w.r.t. EPI distortion corrections.
+Effectively, the simulataneous registration and segmentation enables \emph{regseg}
+  to overcome the nonlinear warping introduced by EPI distortion.
+We have introduced the following changes:
+\begin{itemize}
+  \item Lines 28--36 (p. 2): Included description and references
+    \citep{jezzard_correction_1995,chiou_simple_2000,kybic_unwarping_2000,studholme_accurate_2000,cordes_geometric_2000} about EPI distortion.
+  \item Lines 36--38 (p. 2): Added the suggested reference \citep{irfanoglu_drbuddi_2015} with other
+    two relevant citations \citep{holland_efficient_2010,andersson_comprehensive_2012}.
+  \item Lines 317--328 (p. 14): we discuss our method as regards EPI correction.
+\end{itemize}
 }
@@
 \reply{
-The displacements field supporting the deformation model is not restricted to the phase encoding direction by default.
-Therefore, equations in Methods section are presented under such conditions.
-The proof of concept presented in the first experiment with phantoms are conducted without preferential directions
-  for the deformation.
-However, the point raised about the distortions derived from susceptibility inhomogeneity in dMRI data is also true.
-Therefore, experiments with real data mimicking such distortions are performed with the deformation aligned with the phase encoding
-  axis.
-New lines 203-205 now state this point clearly.
+The general functionality of \emph{regseg} does not restrict distortion if it is not requested using the
+  feature of anisotropic regularization.
+For the experiments on real data, the ground-truth deformations and the deformation field estimated by \emph{regseg}
+  are restricted to be nonzero only in the phase-encoding direction.
+To make this more clear, we have introduced the following edits:
+\begin{itemize}
+  \item Lines 132--134 (p. 7).
+  \item Lines 198--199 (p. 10).
+\end{itemize}
 }
@@
 We thank the reviewer for positively weighing our experiments and results.
-As stated in line 214, we use the settings from the \emph{ExploreDTI} software package \citep{leemans_exploredti_2009}, that
+As stated in line 211 (p. 10), we use the settings from the \emph{ExploreDTI} software package \citep{leemans_exploredti_2009}, that
   internally uses \emph{elastix} \citep{klein_elastix_2010} in registration.
-A footnote in line 215 directly links the settings for this method.
+A footnote in line 213 (p. 10) directly links the settings for this method.
 }
@@
 \end{reviewpoint}
-\reply{We hope this point is answered in \autoref{com:10}, and tryied to avoid this concept
-  coming from the 2D processing field.
-Now, we refer to the contours as surfaces whenever possible.
-Regarding the ``without edges'' feature, we have clarified how the surfaces enclose homogeneous
-  regions (and therefore, they are not used to search for image edges).
+\reply{
+Additionally to the edits in \autoref{com:10}, we have more clearly defined the concept with the discussion
+  introduced in lines 265--266 (p. 13).
 }

### 2015-10-29 Publications/2015-NeuroImage/02-abstract.tex
final edits

@@
 % -*- root: 00-main.tex -*-
-\revcomment[RV\#1(C.1)\par RV\#2(C.10)]{%
 Current methods for processing \gls*{dmri} to map the connectivity of the human brain
   require precise delineations of anatomical structures.
+\revcomment[C1 (R.1)]{%
 The requirement has been approached either segmenting the data in native \gls*{dmri} space
@@
   and the geometrical distortions caused by the inhomogeneity of magnetic susceptibility
-  across tissues render the problem challenging.
+  across tissues render the problem challenging.}
+\revcomment[C10 (R.2)]{%
 To solve the problem in a unified approach, we propose \regseg{}, a surface-to-volume nonlinear
@@
   a set of nested reference-surfaces.}
-\revcomment[RV\#2(C.11)]{%
+\revcomment[C11 (R.2)]{%
 The surfaces can be extracted from the \gls*{t1} image of the subject, and the target image

### 2015-10-29 Publications/2015-NeuroImage/03-introduction.tex
final edits

@@
 Each \gls*{dwi} is sensitized with a gradient to probe proton diffusion in a certain
-  \revcomment[RV\#2(C.13)]{orientation}.
+  \revcomment[C13 (R.2)]{orientation}.
 Subsequent processing involves describing the local microstructure with one available
@@
   to current models such as \citep{daducci_accelerated_2015}.
-The microstructural map is then used to draw the preferential \revcomment[RV\#2(C.13)]{orientations} of diffusion
+The microstructural map is then used to draw the preferential \revcomment[C13 (R.2)]{orientations} of diffusion
   across the brain using tractography \citep{mori_threedimensional_1999}.
@@
   tractography as edges \citep{hagmann_mapping_2008}.
-\revcomment[RV\#2(C.17)]{%
+\revcomment[C9 (R.2) \\ C17 (R.2)]{%
 The methodologies to solve reconstruction, tractography and network building
@@
 
-\revcomment[RV\#2(C.15)]{%
+\revcomment[C15 (R.2)]{%
 Possibly, the earliest structural information incorporated to aid \gls*{dmri} processing
@@
   \citep{taoka_fractional_2009}.
-\todo{RV\#2(C.16)}
+\todo{C16 (R.2)}
 To overcome the unreliability of \gls*{fa} thresholding, and to broaden
@@
   lu_segmentation_2008,han_experimental_2009}.}
-\todo{RV\#2(C.14)}%
+\todo{C14 (R.2)}%
 However, the precise segmentation of \gls*{dmri} is difficult for several reasons.
-First, \gls{dmri} images have a resolution that is \revcomment[RV\#2(C.12)]{much lower}
+First, \gls{dmri} images have a resolution that is \revcomment[C12 (R.2)]{much lower}
   than that of the imaged microstructural features.
@@
 
-\revcomment[RV\#2(C.9)]{%
+\revcomment[C18 (R.2)]{%
 An alternative route to segmentation in \gls*{dmri} space is the mapping of the
@@
   \citep{holland_efficient_2010,andersson_comprehensive_2012,
-  irfanoglu_drbuddi_2015}. \todo{RV\#2(C.18)}
+  irfanoglu_drbuddi_2015}.
 A standardized method to solve the remaining linear mapping between the corrected-\gls*{dmri}
@@
 
-\revcomment[RV\#2(C.10)]{%
+\revcomment[rephrased]{%
 Here we present a segmentation and surface-to-volume registration method
@@
 The underlying hypothesis is that the registration and segmentation problems
-  in \gls*{dmri} can be solved simultaneously.
+  in \gls*{dmri} can be solved simultaneously.}
+\revcomment[C1 (R.1) \\ C10 (R.2)]{%
 To implement \regseg{} we first establish an active-contours without edges
-  \citep{chan_active_2001} segmentation framework.
-Several reference surfaces of the same subject initialize the 3D active-contours,
-  which evolve searching for homogeneous regions in the multivariate target-image.
+  \citep{chan_active_2001} segmentation framework.}
+\revcomment[C2 (R.1)]{%
+A specific set of reference surfaces extracted from the same subject initialize
+  the 3D active-contours, which evolve searching for homogeneous regions in the multivariate
+  target-image.
 In the presented application to segment \gls*{dmri} data, we propose to use \regseg{}

### 2015-10-29 Publications/2015-NeuroImage/04-methods.tex
final edits

@@
   for each region $\Omega_l$.
-\revcomment[RV\#1(C.4)]{%
+\revcomment[C4 (R.1)]{%
 Even though the features being segmented are not generally i.i.d., the spatial interdependency of
@@
   the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized.
-\revcomment[RV\#1(C.2)]{%
+\revcomment[C2 (R.1)]{%
 In \autoref{fig:regseg-model},
@@
 }\label{fig:regseg-model}
-\todo[inline]{RV\#1(C.5)}
+\todo[inline]{C5 (R.1)}
 \end{figure*}
@@
   following sections.
-\revcomment[RV\#2(C.20)]{%
+\revcomment[C20 (R.2)]{%
 A first proof of concept to demonstrate \regseg{} in phantoms with simple
@@
 
-\revcomment[RV\#1(C.6)]{\paragraph*{Segmentation model and settings}
+\revcomment[C6 (R.2)]{\paragraph*{Segmentation model and settings}
 The segmentation model of the phantoms is implicitly defined in each shape.
@@
     the contours are defined.}}\label{fig:regseg-evworkflows}
-  \todo[inline]{RV\#2(C.23)}%
+  \todo[inline]{C7 (R.1) \& C23 (R.2)}%
 \end{figure*}
@@
   multi-shell \gls*{dmri} images.
-\revcomment[RV\#1(C.7)]{%
+\revcomment[C7 (R.1)]{%
 Since we obtained the \gls*{dmri} data from the minimally preprocessed package, these
@@
 \paragraph*{Settings} %
-\revcomment[RV\#1(C.5)]{%
+\revcomment[C5 (R.1)]{%
 \Regseg{} accepts an affine mapping from surface-space to the \gls*{dmri} data as initialization.
@@
 Therefore, the initial estimation of distortion is zero in this experiment.}
-\revcomment[RV\#1(C.20)]{%
+\revcomment[C20 (R.2)]{%
 Since the distortion $U_\text{true}$ is aligned along the \gls*{pe} direction ($y$-axis in our settings),
   \regseg{} was configured to allow nonzero displacements only on that corresponding direction.}
-\revcomment[RV\#1(C.6)]{%
+\revcomment[C6 (R.1)]{%
 For the experiments on real data, \regseg{} established a multi-resolution pyramid of
@@
   perform registration.
-\revcomment[RV\#2(C.21)]{%
+\revcomment[C21 (R.2)]{%
 The deformation field is correspondingly restricted to the \gls*{pe} direction.

### 2015-10-29 Publications/2015-NeuroImage/06-discussion.tex
final edits

@@
 \label{sec:regseg-discussion}
+\revcomment[C10 (R.2)]{%
 We present \regseg{}, a simultaneous segmentation and registration method that
-  maps a set of nested surfaces into a multivariate target-image.
+  maps a set of nested surfaces into a multivariate target-image.}
+\revcomment[C1 (R.1)]{%
 The nonlinear registration process evolves driven by the fitness of the
   picewise-smooth classification of voxels in the target volume imposed
-  by the current mapping of the surfaces.
+  by the current mapping of the surfaces.}
+\revcomment[moved]{%
 We propose \regseg{} to map anatomical information extracted from \gls*{t1}
@@
   that supports the evolution of the active contours (segmentation),
-  like \cite{paragios_level_2003,yezzi_variational_2003}.
+  like \cite{paragios_level_2003,yezzi_variational_2003}.}
+\revcomment[C1 (R.1)]{%
 \Regseg{} can be seen as a particular case of atlas-based segmentation-registration methods,
-  replacing the atlas by the structural image of the subject (\emph{structure-informed segmentation}).
+  replacing the atlas by the structural image of the subject (\emph{structure-informed segmentation}).}
+\revcomment[C3 (R.1)]{%
 \cite{gorthi_active_2011} review the application of joint segmentation and registration in atlas-based
@@
   surfaces (triangular meshes) instead of the multiphase level sets, and registration
-  is driven by shape-gradients \citep{herbulot_segmentation_2006}.
+  is driven by shape-gradients \citep{herbulot_segmentation_2006}.}
 
+\revcomment[C3 (R.1)]{%
 An important antecedent of \regseg{} is \emph{bbregister} \citep{greve_accurate_2009}.
@@
 Indeed, the distortion must be addressed separately because it is not supported by
-  the affine transformation model.
+  the affine transformation model.}
+\revcomment[C22 (R.2)]{%
 Conversely, the deformation model of \regseg{} is nonlinear and the active contours are
-  \emph{without-edges} \citep{chan_active_2001} because the \gls*{fa} and \gls*{adc} maps
+  \emph{without edges} \citep{chan_active_2001} because the \gls*{fa} and \gls*{adc} maps
   do not present steep image gradients (edges) but the anatomy can be identified
-  by looking for piece-wise smooth homogeneous regions.
+  by looking for piece-wise smooth homogeneous regions.}
 
+\revcomment[C3 (R.1)]{%
 Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
@@
   generally included in most of the existing methodologies, and the complexity of
-  the elasticity smoother of \cite{guyader_combined_2011}.
+  the elasticity smoother of \cite{guyader_combined_2011}.}
+\revcomment[C3 (R.1)]{%
 Other minor features that differ from current methods in joint segmentation and registration are
   the support of multivariate target-images and the efficient computation of the shape-gradients
-  implemented with sparse matrices.
+  implemented with sparse matrices.}
 
@@
 Therefore, the error was bounded above by half of the voxel size.
-\revcomment[RV\#1(C.8)]{%}
+\revcomment[C8 (R.1)]{%}
 The distributions of errors along surfaces varied importantly depending on the shape of the
@@
   but indicated that \regseg{} is a reliable option in this application field.
-\revcomment[RV\#1(C.5)]{%
+\revcomment[C5 (R.1)]{%
 Finally, we also contributed proposing a piecewise-smooth segmentation model defined by
@@
 
+\revcomment[C9 (R.2)]{%
 The proposed application of the method in the task of identifying structural information
@@
 The problem was firstly addressed using image segmentation approaches in the native diffusion
-  space, without definite and compelling results.
+  space, without definite and compelling results.}
+\revcomment[C18 (R.2)]{%
 With the introduction of retrospective correction methods for the \emph{\gls*{epi} distortions}
@@
   the whole connectome extraction process can be performed without resampling data to an
-  undistorted space.
+  undistorted space.}
 
+\revcomment[C9 (R.2)]{%
 Beyond the presented application on \gls*{dmri} data, \regseg{} can be indicated in situations
@@
   injury \citep{irimia_structural_2014} or in the development of brain tumors
-  \citep{weizman_semiautomatic_2014}.
+  \citep{weizman_semiautomatic_2014}.}

### 2015-10-30 Publications/2015-NeuroImage/06-discussion.tex
add discussion about gorthi et al. 2011

@@
 \Regseg{} can be seen as a particular case of atlas-based segmentation-registration methods,
-  replacing the atlas by the structural image of the subject (\emph{structure-informed segmentation}).}
+  replacing the atlas by the structural image of the subject (\emph{structure-informed segmentation}).
+In atlas-based segmentation, the origin of the problem is the anatomical variability hindering the
+  atlas matching process.
+Conversely, in structure-informed segmentation, the sources of variability are the geometrical distortions
+  after imaging and the anatomical evolution in longitudinal studies.}
 \revcomment[C3 (R.1)]{%
-\cite{gorthi_active_2011} review the application of joint segmentation and registration in atlas-based
-  segmentation, and propose a multiphase level set function to register several active contours.
+\cite{gorthi_active_2011} review the joint segmentation and registration methods, and propose a multiphase
+  level set function derived from a labeled atlas for active contour-based registration in the atlas-based
+  segmentation problem.
 Alternatively, \regseg{} implements the active contours with a hierarchical set of explicit
   surfaces (triangular meshes) instead of the multiphase level sets, and registration
-  is driven by shape-gradients \citep{herbulot_segmentation_2006}.}
+  is driven by shape-gradients \citep{herbulot_segmentation_2006}.
+This solution is better adapted to the structure-informed segmentation, and it is possibly
+   simpler to implement than the multiphase level set functions.}
 

### 2015-11-02 Publications/2015-NeuroImage/04-methods.tex
fix todonote
update supplementals citation

@@
     energies with respect to the two limiting regions of $M(\vec{v}_i)$, the features of the image
-    $M$ in the location of vertex $\vec{v}_i$ (see \suppl{equation SM5}).
+    $M$ in the location of vertex $\vec{v}_i$ \citepalias[see][equation SM5]{esteban_useful_2015}.
   In this figure, $\bar{s}_1$ is written in the lower
@@
 The computation of the explicit shape gradients at each $\vec{v}_i'$ is illustrated in \autoref{fig:regseg-method}.
-Then, \eqref{eq:regseg-gradient_descent} can be reformulated as (see \suppl{equations SM6-SM12}):
+Then, \eqref{eq:regseg-gradient_descent} can be reformulated as \citepalias[see][equations SM6-SM12]{esteban_useful_2015}:
 
@@
 
-Finally, to descend this gradient, we establish a semi-implicit Euler scheme (see \suppl{section S1.3}),
+Finally, to descend this gradient, we establish a semi-implicit Euler scheme
+  \citepalias[see][section S1.3]{esteban_useful_2015},
   with a step size parameter $\delta$, which we solve in the spectral domain as follows:
@@
   (such as $\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, the B-spline grid resolutions,
-  and target image smoothing) are discussed in \suppl{section S1}.
+  and target image smoothing) are discussed in \citepalias[][section S1]{esteban_useful_2015}.
 
@@
 
-\revcomment[C6 (R.2)]{\paragraph*{Segmentation model and settings}
+\paragraph*{Segmentation model and settings}
+\revcomment[C6 (R.2)]{%
 The segmentation model of the phantoms is implicitly defined in each shape.
@@
 The segmentation model corresponding to this partition is shown in \autoref{fig:regseg-model}
-  and discussed in greater detail in \suppl{section S4}.
+  and discussed in greater detail in \citepalias[section S4]{esteban_useful_2015}.
 
@@
 The minimum of the cost function (\autoref{sec:regseg-methods_map}) was consistently located at
-  $\epsilon=0.0$ (the ground-truth) for all of the cases (\suppl{Figure S2}).
+  $\epsilon=0.0$ (the ground-truth) for all of the cases \citepalias[figure S2]{esteban_useful_2015}.
 
@@
 These settings were obtained manually, driven by the feedback obtained from the post-registration
-  convergence reports (like that found in \suppl{section S1.3}).
+  convergence reports \citepalias[like that found in][section S1.3]{esteban_useful_2015}.
 We released \regseg{} along with the tool to generate such convergence reports.

### 2015-11-02 Publications/2015-NeuroImage/06-discussion.tex
update supplementals citation

@@
   (\autoref{tab:results_real}).
-Visual inspections of all the results (\suppl{section S5}) and the violin plots in
+Visual inspections of all the results \citepalias[section S5]{esteban_useful_2015} and the violin plots in
   \autoref{fig:regseg-results_real} confirmed that \regseg{} achieved higher accuracy
@@
 We also demonstrated the smoothness of the objective function on five of the real datasets
-  (\suppl{Figure S2}), taking advantage of the directional restriction of possible distortions.
+  \citepalias[figure S2]{esteban_useful_2015}, taking advantage of the directional
+  restriction of possible distortions.
 However, convergence of \regseg{} can be severely affected when surfaces are not dense enough.

### 2015-11-03 Publications/2015-NeuroImage/02-abstract.tex
recover changes in abstract

@@
 \revcomment[C1 (R.1)]{%
-The requirement has been approached either segmenting the data in native \gls*{dmri} space
-  or projecting the structural information from \gls*{t1} images.
-The characteristic features of diffusion data in terms of \acrlong*{snr}, resolution, etc.
-  and the geometrical distortions caused by the inhomogeneity of magnetic susceptibility
-  across tissues render the problem challenging.}
+This requirement has been approached either segmenting the data in native \gls*{dmri} space
+  or mapping the structural information from \gls*{t1} images.
+The characteristic features of diffusion data in terms of \acrlong*{snr}, resolution, as well
+  as the geometrical distortions caused by the inhomogeneity of magnetic susceptibility
+  across tissues hinder both solutions.}
 \revcomment[C10 (R.2)]{%
-To solve the problem in a unified approach, we propose \regseg{}, a surface-to-volume nonlinear
+Unifying the two approaches, we propose \regseg{}, a surface-to-volume nonlinear
   registration method that segments homogeneous regions within multivariate images by mapping
@@
 \revcomment[C11 (R.2)]{%
-The surfaces can be extracted from the \gls*{t1} image of the subject, and the target image
-  is the bivariate volume comprehending the \gls*{fa} and the \gls*{adc} maps derived from the
+Accurate surfaces are extracted from a \gls*{t1} image of the subject, using as target image
+  the bivariate volume comprehending the \gls*{fa} and the \gls*{adc} maps derived from the
   \gls*{dmri} dataset.}
-We first verify the subpixel accuracy of \regseg{} on digital phantoms, to then propose an
-  evaluation framework that uses \gls*{dmri} datasets from the \gls*{hcp}.
-In this framework, the undistorted \gls*{dmri} undergo a known distortion derived from real
-  fieldmaps.
-We analyze the misregistration error of the surfaces estimated by \regseg{} on 16 datasets.
-The distribution of errors shows a 95\% CI of 0.56--0.66 mm, below the \gls*{dmri} resolution
-  (1.25 mm, isotropic).
+\revcomment[rephrased]{%
+We first verify the accuracy of \regseg{} on a general context using digital phantoms.
+Then we establish an evaluation framework using undistorted \gls*{dmri} data from the \gls*{hcp}
+  and known deformations derived from real inhomogeneity fieldmaps.
+We analyze the performance of \regseg{} computing the misregistration error of the surfaces estimated
+  after being mapped with \regseg{} onto 16 datasets from the \gls*{hcp}.
+The distribution of errors shows a 95\% CI of 0.56--0.66 mm, that is below the \gls*{dmri}
+  resolution (1.25 mm, isotropic).
 Finally, we cross-compare the proposed tool against a nonlinear \lowb{}-to-T2w registration
-  method, thereby obtaining a significantly lower misregistration error with \regseg{}.
+  method, thereby obtaining a significantly lower misregistration error with \regseg{}.}

### 2015-11-03 Publications/2015-NeuroImage/03-introduction.tex
modifications for MJLC comments

@@
   named \regseg{}, and show its usefulness in mapping anatomical information from structural
-  into native \gls*{dmri} space to aid subsequent processing steps
+  space into native \gls*{dmri} space to aid subsequent processing steps
   (reconstruction, tractography and network building using a cortical parcellation).
@@
 Therefore, \regseg{} establishes a registration framework that actually
-  overcomes the nonlinear warping induced by \gls*{epi} distortions.%
+  deals with the nonlinear warping induced by \gls*{epi} distortions.%
 }

### 2015-11-03 Publications/2015-NeuroImage/04-methods.tex
modifications for MJLC comments

@@
 \revcomment[C20 (R.2)]{%
-A first proof of concept to demonstrate \regseg{} in phantoms with simple
-  geometries is introduced, using $U_\text{true}$ without directional restrictions.
-Then, a more realistic environment with $U_\text{true}$ restricted to the
-  \gls*{pe} axis of real \gls*{dmri} datasets is presented.}
+A first proof of concept is introduced to demonstrate \regseg{} in digital phantoms
+  with simple geometries, using $U_\text{true}$ without directional restrictions.
+Then, \regseg{} is evaluated in a framework using undistorted \gls*{dmri} datasets,
+  and $U_\text{true}$ is derived from the corresponding inhomogeneity fieldmap of
+  the subject.
+Therefore, the deformation field is nonzero only in the \gls*{pe} axis,
+  and resembles a real \gls*{epi} distortion.}
 
@@
   \revcomment{%
-  The arrows point to edges in the target images (yellow arrows for \gls*{fa}, blue for \gls*{adc}) that
-    should be aligned with a surface, showing how distortion breaks the mapping with the structural space in which
-    the contours are defined.}}\label{fig:regseg-evworkflows}
+  The arrows point to edges in the target images (light-yellow arrows for \gls*{fa}, blue for \gls*{adc}) that
+    should be aligned with a surface, showing how distortion limits the direct mapping from the structural space
+    in which the contours are defined.}}\label{fig:regseg-evworkflows}
   \todo[inline]{C7 (R.1) \& C23 (R.2)}%
@@
   normalization, etc.).
-We refer the reader to \citep{essen_human_2012} for exact details of the acquisition
-  parameters and \citep{glasser_minimal_2013} for the preprocessing issues.
+We refer the reader to \citet{essen_human_2012} for exact details of the acquisition
+  parameters and \citet{glasser_minimal_2013} for the preprocessing issues.
 These datasets comprise a large set of images, including \gls*{t1}, \gls*{t2}, and
@@
 For the experiments on real data, \regseg{} established a multi-resolution pyramid of
-  B-Spline basis, with control points distributed on grids of the following spacings:
-  40$\times$100$\times$40 [mm] for the first (coarser level), \isores[mm]{30} for the second,
-  and 20$\times$30$\times$10 [mm] in the third.
+  B-Spline functions, with control points distributed on grids of the following spacings:
+  40$\times$100$\times$40 [mm] for the first (coarser) level, \isores[mm]{30} for the second level,
+  and 20$\times$30$\times$10 [mm] in the third level.
 Only the first and second levels included Gaussian smoothing of the target image ($\sigma$ = [2.0, 0.5] mm,
@@
 The deformation field is correspondingly restricted to the \gls*{pe} direction.
-Settings used with \emph{elastix} are conveniently available in its format\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/t2b_elastix_y.txt}}.}
+The settings file for \emph{elastix} is also
+  available\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/t2b_elastix_y.txt}}.}
 

### 2015-11-03 Publications/2015-NeuroImage/05-results.tex
update supplementals citation
modifications for MJLC comments

@@
   \includestandalone[width=\linewidth]{figure06}
-  \caption{A. Example of a visual assessment report, which was generated automatically by the evaluation tool.
+  \caption{A. Example of a visual assessment report, which was automatically generated by the evaluation tool.
     Each view shows one component of the input image (in this case, the \gls*{fa} map), the ground-truth locations
@@
   method.
-Summary reports for visual assessment of the 16 cases are included in \suppl{section S5}.
+Summary reports for visual assessment of the 16 cases are included in \citepalias[section S5]{esteban_useful_2015}.
 In \autoref{fig:regseg-results_real}, box A, the visual report is shown for one subject.

### 2015-11-03 Publications/2015-NeuroImage/06-discussion.tex
modifications for MJLC comments

@@
   like \cite{paragios_level_2003,yezzi_variational_2003}.}
-\revcomment[C1 (R.1)]{%
+% Conversely, in structure-informed segmentation, the sources of variability are the geometrical distortions
+%   after imaging and the anatomical evolution in longitudinal studies.}
+\revcomment[C3 (R.1)]{%
 \Regseg{} can be seen as a particular case of atlas-based segmentation-registration methods,
   replacing the atlas by the structural image of the subject (\emph{structure-informed segmentation}).
-In atlas-based segmentation, the origin of the problem is the anatomical variability hindering the
-  atlas matching process.
-Conversely, in structure-informed segmentation, the sources of variability are the geometrical distortions
-  after imaging and the anatomical evolution in longitudinal studies.}
-\revcomment[C3 (R.1)]{%
-\cite{gorthi_active_2011} review the joint segmentation and registration methods, and propose a multiphase
-  level set function derived from a labeled atlas for active contour-based registration in the atlas-based
-  segmentation problem.
+The main difference of atlas-based segmentation and the application at hand is the resolution of the
+  target image.
+Atlas-based segmentation is typically applied on high-resolution images.
+A comprehensive review of joint segmentation and registration methods applied in atlas-based
+  segmentation is found in \citep{gorthi_active_2011}.
+They also propose a multiphase level-set function derived from a labeled atlas to implement the
+  active contours that drive the atlas registration.
 Alternatively, \regseg{} implements the active contours with a hierarchical set of explicit
@@
   is driven by shape-gradients \citep{herbulot_segmentation_2006}.
-This solution is better adapted to the structure-informed segmentation, and it is possibly
-   simpler to implement than the multiphase level set functions.}
+The use of explicit surfaces enables segmenting \gls*{dmri} images below pixel resolution.}
 
@@
 Conversely, the deformation model of \regseg{} is nonlinear and the active contours are
-  \emph{without edges} \citep{chan_active_2001} because the \gls*{fa} and \gls*{adc} maps
+  \emph{without edges} \citep{chan_active_2001} since the \gls*{fa} and \gls*{adc} maps
   do not present steep image gradients (edges) but the anatomy can be identified
@@
 However, convergence of \regseg{} can be severely affected when surfaces are not dense enough.
-Using the digital phantoms, we decimated the surfaces with large factors and
-  effectively displaced the minimum of the shape-gradients from the minimum of the
-  objective function.
-Thus, the surfaces must be densely sampled to enable the use of the presented method.
+Using the digital phantoms, we decimated the surfaces by large factors.
+Severely decimated surfaces introduced a bias that displaced the zero of the gradients from
+  the minimum of the objective function.
+Thus, the surfaces must be densely sampled to avoid the shape-gradients drive the algorithm
+  in a wrong direction impeding convergence.
 }%
@@
   image in which the surfaces must be fitted, and the mapping between the surfaces and
-  the volume encodes a relevant physiological information, such as the normal/abnormal
+  the volume encodes relevant physiological information, such as the normal/abnormal
   development or the macroscopic dynamics of organs and tissues.
@@
   of the brain using multispectral images, for instance in the case of traumatic brain
-  injury \citep{irimia_structural_2014} or in the development of brain tumors
+  injury \citep{irimia_structural_2014} or in the monitorization of brain tumors
   \citep{weizman_semiautomatic_2014}.}

### 2015-11-03 Publications/2015-NeuroImage/06-discussion.tex
fix sentence

@@
   restriction of possible distortions.
-However, convergence of \regseg{} can be severely affected when surfaces are not dense enough.
-Using the digital phantoms, we decimated the surfaces by large factors.
-Severely decimated surfaces introduced a bias that displaced the zero of the gradients from
-  the minimum of the objective function.
-Thus, the surfaces must be densely sampled to avoid the shape-gradients drive the algorithm
-  in a wrong direction impeding convergence.
+However, \regseg{} requires densely sampled surfaces to ensure the convergence.
+Using the digital phantoms, we severely decimated the surfaces by a large factor.
+These surfaces introduced a bias which displaced the zero of the gradients from the
+  minimum of the objective function impeding the convergence.
 }%

### 2015-11-04 Publications/2015-NeuroImage/03-introduction.tex
fix typos

@@
 In the presented application to segment \gls*{dmri} data, we propose to use \regseg{}
-  using a set of nested surfaces extracted from a strutrucal image (e.g. \gls*{t1}) and
+  using a set of nested surfaces extracted from a structural image (e.g. \gls*{t1}) and
   a bivariate target-volume comprehending the \gls*{fa} and \gls*{adc} maps as inputs.
@@
   reproducing the settings of an existing implementation included within a widely used diffusion
-  processing software (\emph{ExploreDTI}, \cite{leemans_exploredti_2009}).
+  processing software \citep[\emph{ExploreDTI},][]{leemans_exploredti_2009}.
 With this comparison, we demonstrate how \regseg{} achieves higher accuracy with the

### 2015-11-04 Publications/2015-NeuroImage/04-methods.tex
fix typos

@@
   where $P(M \mid U,\omegaset)$ is the data likelihood.
-Because $\omegaset$ are mapped by $U$, we simplify
+Since $\omegaset$ are mapped by $U$, we simplify
   $P(U, \omegaset) = P(U) \implies P(M \mid U,\omegaset) = P(M \mid U)$.
@@
   (such as $\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, the B-spline grid resolutions,
-  and target image smoothing) are discussed in \citepalias[][section S1]{esteban_useful_2015}.
+  and target image smoothing) are discussed in the \citetalias[][section S1]{esteban_useful_2015}.
 
@@
 
-The adaptation of this protocol to the simulated phantoms and real data is explained in the
-  following sections.
 \revcomment[C20 (R.2)]{%
@@
   and resembles a real \gls*{epi} distortion.}
+The adaptation of the evaluation protocol to the simulated phantoms and the real data is
+  explained in the following sections.
 
@@
 \revcomment[C6 (R.2)]{%
-The segmentation model of the phantoms is implicitly defined in each shape.
-All phantoms comprehend an inner surface enclosing a uniform \gls*{wm}-like region,
+The segmentation model of the phantoms is implicitly defined:
+  all phantoms comprehend an inner surface enclosing a uniform \gls*{wm}-like region,
   and an outer surface wrapping a \gls*{gm}-like layer.
@@
 \subsection{Real datasets} \label{sec:regseg-human_connectome}
-The general experimental framework for the real datasets is presented in \autoref{fig:regseg-evworkflows},
+The experimental framework for the real datasets is presented in \autoref{fig:regseg-evworkflows},
 which extends our previous evaluation \citep{esteban_simulationbased_2014} of distortions
@@
   2) For use as the ground truth, we generated a plausible synthetic distortion $U_\text{true}$
-    from the field map with \eqref{eq:regseg-fieldmap}.
+    from the fieldmap with \eqref{eq:regseg-fieldmap}.
   3) The \gls*{dmri} data were warped using $U_\text{true}$ to reproduce the effects of real
@@
 The segmentation model corresponding to this partition is shown in \autoref{fig:regseg-model}
-  and discussed in greater detail in \citepalias[section S4]{esteban_useful_2015}.
+  and discussed in greater detail in the \citetalias[section S4]{esteban_useful_2015}.
 
@@
 Certain \gls*{mr} sequences are designed to estimate $\Delta B_0$, thereby obtaining
-  the so-called field map.
-We derived the deformation $U_\text{true}$ from the field map image released with
+  the so-called fieldmap.
+We derived the deformation $U_\text{true}$ from the fieldmap image released with
   the corresponding packages of each dataset in the \gls*{hcp}.
-The field map was unwrapped\footnote{fieldmaps are phase maps, which are intrinsically clipped in the interval
+The fieldmap was unwrapped\footnote{fieldmaps are phase maps, which are intrinsically clipped in the interval
   of $[-\pi, \pi)$ [rads] or [rads/s].} and smoothed before applying \eqref{eq:regseg-fieldmap}.
@@
 \paragraph*{Cross-comparison} %
-A dual workflow to the general evaluation framework used for \regseg{} (\autoref{fig:regseg-evworkflows}),
+A dual workflow to the general evaluation used for \regseg{} (\autoref{fig:regseg-evworkflows}),
   was employed to integrate the alternate \gls*{t2b} registration scheme.

### 2015-11-04 Publications/2015-NeuroImage/06-discussion.tex
fix typos

@@
   and an in-house replication of the \acrfull*{t2b} method.
-\Regseg{} obtained very high accuracy, with an aggregate 95\% \gls*{ci} of 0.56--0.66 [mm], which was
+\Regseg{} obtained a high accuracy, with an aggregate 95\% \gls*{ci} of 0.56--0.66 [mm], which was
   below the pixel size of 1.25 mm.
@@
   a selection of nested surfaces to partition the multispectral space
-  comprehending the \gls*{fa} and the \gls*{adc} maps and ultimately identify these
+  comprehending the \gls*{fa} and the \gls*{adc} maps and ultimately identify anatomical
   structures in \gls*{dmri} space.

### 2015-11-04 Publications/2015-NeuroImage/04-methods.tex
add final MJLC comments

@@
   for a real dataset.
+The left plot in each panel represents a kernel density estimate of the location and spread of
+  each region $\Omega_l$ in the bivariate feature space of $M$, which comprises
+  the \gls*{fa} and \gls*{adc} maps.
+At the top and the right margins, the marginal distributions of $\Omega_l$ on the \gls*{fa} and
+  \gls*{adc} are respectively displayed.
+On the right side, the sampled distribution of each region is shown in small scatter plot, in contrast to
+  the complete bivariate data.
 A) shows the initialization of the model, when the anatomically-correct surfaces are not
@@
   \gls*{adc} maps with the original (undistorted) surfaces.
-Panel B shows the equivalent information, using the surfaces with perfect matching with the distorted space
-  of \gls*{dmri}.
+Panel B shows the equivalent information, using the surfaces that are perfectly mapped into the
+  distorted \gls*{dmri} space.
 Since the surfaces are now aligned in $M$, the spreads of the conditional distributions for each region are
@@
   are minimized with respect to panel A, and the resulting segmentation is close
-  to the theoretical situation presented in panel B.
-The left (large) plot in each panel represents a kernel density estimate of the location and spread of
-  each region $\Omega_l$ in the bivariate feature space of $M$, which comprises
-  the \gls*{fa} and \gls*{adc} maps.
-At the top and the right margins, the marginal distributions of $\Omega_l$ are also plotted for
-  the \gls*{fa} and \gls*{adc}, respectively.
-On the right, the sampled distribution of each region is shown in small scatter plot, in contrast to
-  the complete bivariate data.}
+  to the theoretical situation presented in panel B.}
 }\label{fig:regseg-model}

### 2015-11-04 Publications/2015-NeuroImage/06-discussion.tex
add final MJLC comments

@@
   target image.
-Atlas-based segmentation is typically applied on high-resolution images.
+Atlas-based segmentation is typically applied on structural and high-resolution images.
 A comprehensive review of joint segmentation and registration methods applied in atlas-based
   segmentation is found in \citep{gorthi_active_2011}.
-They also propose a multiphase level-set function derived from a labeled atlas to implement the
-  active contours that drive the atlas registration.
+They also propose a multiphase level-set function initialized from a labeled atlas to implement
+  the active contours that drive the atlas registration.
 Alternatively, \regseg{} implements the active contours with a hierarchical set of explicit
@@
   is driven by shape-gradients \citep{herbulot_segmentation_2006}.
-The use of explicit surfaces enables segmenting \gls*{dmri} images below pixel resolution.}
+As an advantage, the use of explicit surfaces enables segmenting \gls*{dmri} images
+  with accuracy below pixel size.}
 
@@
 Moreover, since the structural information is projected into the native space of \gls*{dmri},
-  the whole connectome extraction process can be performed without resampling data to an
-  undistorted space.}
+  the whole connectome extraction process can be performed on the unaltered \gls*{dmri}
+  signal (i.e. without resampling data to an undistorted space).}
 

### 2015-11-04 Publications/2015-NeuroImage/responseLetter2.tex
version proposed to coauthors

@@
 
-\usepackage[hidelinks]{hyperref}
+\usepackage[colorlinks=true]{hyperref}
 \usepackage{nameref}
@@
 \begin{document}
+\hypersetup{linkcolor=black!60, citecolor=black!60, urlcolor=black!60}
 
@@
 Text modifications in the manuscript are highlighted with a new font color, and attached notes indicate which reviewer comment
-  the corresponding change is addressing.
+  the corresponding change is addressing to.
 
@@
 \reply{%
-The description of the method given by both reviewers (see \autoref{com:10}) lead us to think that the paper
-  does not describe correctly the actual inputs of the method.
+The description of the method given by both reviewers (see \autoref{com:10}, Reviewer \#2) lead us to think that the paper
+  did not properly describe the actual inputs and outputs the proposed methodology.
 In general, we have almost-completely reformulated the abstract, introduction and discussion to define this
   point clearly.
-Particularly, the following sections have been introduced give an answer to this particular comment:
+Particularly, the following sections have been introduced to give an answer to this particular comment:
 \begin{itemize}
 \item Lines 45--53 (p. 2).
-\item Lines 245--246 (p. 13), here we introduce the previously missing discussion to call the
+\item Lines 244--246 (p. 13), here we introduce the previously missing discussion to call the
   method ``simultaneous registration and segmentation''.
-\item Lines 251--253 (p. 13), we comment here on the similarity of \emph{regseg} to the atlas-based segmentation
+\item Lines 250--252 (p. 13), we comment here on the similarity of \emph{regseg} to the atlas-based segmentation
   using joint registration and segmentation methods.
@@
 \reply{
-We have stated more clearly that the choice of surfaces to segment the target data is crucial (lines 46--47, p.2).
-We have better defined the segmentation model that drives the registration process (lines 94--95, p. 3).
-We have also improved former Figure 4, and brought it earlier in the paper (now it is Figure 2, page 5).
+To clarify this point, we have stated more clearly that the choice of surfaces to segment the target data is
+  crucial (lines 46--47, p.2).
+We have better defined the segmentation model that drives the registration process (lines 94--95, p. 5).
+We have also improved the former Figure 4 (now it is Figure 2, page 5), and brought it earlier in the paper to
+  illustrate earlier the segmentation model.
 Since the former visualization of the conditional (region-wise) joint distribution of FA and ADC was a bit clumsy,
@@
 We expect these modifications make clear the difference between the surfaces involved in driving the registration
-  process and the surfaces in which we evaluate the error of the method (lines 218--219, p. 10) are a subset of the
+  process and the surfaces in which we evaluate the error of the method (lines 217--218, p. 10) are a subset of the
   former.
@@
 We have moved the related methods and our methodological contributions from the Introduction to a more appropriate
-  place in the Discussion (lines 248--275), additionally satisfying \autoref{com:9} suggesting to include in the introduction
-  only the references related to the focus of the paper.
+  place in the Discussion (lines 248--275), additionally satisfying \autoref{com:9} (Reviewer \#2) suggesting to include
+  in the introduction only the references related to the focus of the paper.
 The particular changes can be listed as follows:
 \begin{itemize}
-  \item Lines 253--257 (p. 13), we describe the methodological differences w.r.t. \cite{gorthi_active_2011},
+  \item Lines 252--260 (p. 13), we describe the methodological differences w.r.t. \cite{gorthi_active_2011},
     one of the latest atlas-based segmentation methods using simultaneous registration and segmentation.
-  \item Lines 258--266 (p. 13), we compare the methodological features of our method and \emph{bbregister},
+  \item Lines 261--270 (p. 13), we compare the methodological features of our method and \emph{bbregister},
     an antecedent and standardized tool used on the same application.
-  \item Lines 268--273 (p. 13), we introduce the methodological differences with a recent alternative method in
+  \item Lines 271--273 (p. 13), we introduce the methodological differences with a recent alternative method in
     terms of regularization.
-  \item Lines 273--275 (p. 13) mention some other minor methodological contributions.
+  \item Lines 273--278 (p. 13) mention some other minor methodological contributions.
 \end{itemize}
@@
   the choice of surfaces and their nesting.
-This is an actual contribution of the work, as stated in lines 301--304 (p. 14).
-A comment on the convergence issues follows in lines 304--309 (p. 14).
+This is an actual contribution of the work, as stated in lines 304--307 (p. 14).
+A comment on the convergence issues follows in lines 307--311 (p. 14).
 Finally, we described the initialization of the surfaces with the insertion of lines 196--198 (p. 10).
@@
 \reply{
-We have added a discussion on this in the corresponding section, lines 284-290. Thanks for pointing out.
+We have added a discussion on this in the corresponding section, lines 287-293 (p. 14).
+Thanks for pointing this systematic error out.
 }
@@
 In order to better focus the application field, we have introduced lines 11--14 (p. 2) and lines
-  310--317 (p. 14).
-We have also added potential applications in the new lines 329--338 (pp. 14-15) to better illustrate
+  312--317 (p. 14).
+We have also added potential applications in the new lines 331--340 (p. 15) to better illustrate
   the method.
@@
   of the descriptive sections of the paper give a successful answer to this issue.
-Regarding the actual inputs of \emph{regseg}, we have introduce modifications in the abstract, lines 45--50 (p. 2), and
-  lines 244--245 (p. 13).
+Regarding the actual inputs of \emph{regseg}, we have introduced modifications in the abstract, in lines 45--50 (p. 2), and
+  in lines 243--243 (p. 13).
 }
@@
 All methods reviewed in lines 15--27 cluster/segment/classify \emph{voxels} in dMRI space based on
-  one or another feature derived from dMRI data.
+  one or another feature derived from dMRI data (lines 18--20, p. 2).
 We have clarified the manuscript according to \autoref{com:14} and \autoref{com:15} and this comment.
@@
 \reply{
-Effectively, the simulataneous registration and segmentation enables \emph{regseg}
+Effectively, the simultaneous registration and segmentation enables \emph{regseg}
   to overcome the nonlinear warping introduced by EPI distortion.
@@
     two relevant citations \citep{holland_efficient_2010,andersson_comprehensive_2012}.
-  \item Lines 317--328 (p. 14): we discuss our method as regards EPI correction.
+  \item Lines 319--330 (p. 14): we discuss our method as regards EPI correction.
 \end{itemize}
@@
 \begin{itemize}
-  \item Lines 132--134 (p. 7).
+  \item Lines 131--134 (p. 7).
   \item Lines 198--199 (p. 10).
@@
 Additionally to the edits in \autoref{com:10}, we have more clearly defined the concept with the discussion
-  introduced in lines 265--266 (p. 13).
+  introduced in lines 267--268 (p. 13).
 }

### 2015-11-04 Publications/2015-NeuroImage/03-introduction.tex
version proposed to coauthors

@@
   lu_segmentation_2008,han_experimental_2009}.}
-\todo{C14 (R.2)}%
-However, the precise segmentation of \gls*{dmri} is difficult for several reasons.
+\revcomment[C14 (R.2)]{%
+However, the precise segmentation of \gls*{dmri} is difficult for several reasons.}
 First, \gls{dmri} images have a resolution that is \revcomment[C12 (R.2)]{much lower}

### 2015-11-04 Publications/2015-NeuroImage/04-methods.tex
version proposed to coauthors

@@
   that correspond to $\vec{v}_i$.
-In practice, we only report the \gls*{swindex} for three surfaces of crucial interest in whole-brain
-  tractography.
-These three surfaces delineated the following regions in the model: $\Omega_{VdGM}$, $\Omega_{WM}$, and $\Omega_{GM}$.
+\revcomment[C2 (R.1)]{%
+In practice, we only report the \gls*{swindex} for three surfaces (\{$\Gamma_{VdGM}, \Gamma_{WM}, \Gamma_{pial}$\})
+  of crucial interest in whole-brain tractography.}

### 2015-11-04 Publications/2015-NeuroImage/06-discussion.tex
version proposed to coauthors

@@
 %   after imaging and the anatomical evolution in longitudinal studies.}
+\revcomment[C1 (R.1)]{%
+\Regseg{} can be seen as a particular case of atlas-based segmentation-registration methods,
+  replacing the atlas by the structural image of the subject (\emph{structure-informed segmentation}).}
 \revcomment[C3 (R.1)]{%
-\Regseg{} can be seen as a particular case of atlas-based segmentation-registration methods,
-  replacing the atlas by the structural image of the subject (\emph{structure-informed segmentation}).
 The main difference of atlas-based segmentation and the application at hand is the resolution of the

### 2015-11-05 Publications/2015-NeuroImage/07-conclusion.tex
fix zenodo reference

@@
 All of the research elements (data, source code, figures, manuscript sources, etc.) involved in this study
-  are publicly available under a unique package \citep{esteban_acweregistration_2015}.
+  are publicly available under a unique package \citep{esteban_regseg_2015}.

### 2015-11-05 Publications/2015-NeuroImage/ELS285_article_edited.tex
fix zenodo reference

@@
 All of the research elements (data, source code, figures, manuscript sources, etc.) involved in this study
-  are publicly available under a unique package \citep{esteban_acweregistration_2015}.
+  are publicly available under a unique package \citep{esteban_regseg_2015}.
 

### 2015-11-05 Publications/2015-NeuroImage/supplemental.tex
fix zenodo reference

@@
 All the intermediate results and figures in this paper have been encapsulated into
-  the workflows and are available in \citep{esteban_acweregistration_2015}.
+  the workflows and are available in \citep{esteban_regseg_2015}.
 

### 2015-11-05 Scripts/pyacwereg/viz.py
visualization improvements

@@ -399,7 +399,7 @@ def jointplot_data(im1data, im2data, in_seg, labels=None, out_file=None,
 
 def jointplot_real(imageX, imageY, segmentation, mask,
                    labels=None, xlabel='X', ylabel='Y',
-                   huelabel='tissue',
+                   huelabel='tissue', size=8, dpi=100, subsample=1.0,
                    xlims=None, ylims=None, out_file=None):
     import os.path as op
     import nibabel as nb
@@ -436,13 +436,15 @@ def jointplot_real(imageX, imageY, segmentation, mask,
         df_pieces.append(pd.DataFrame(d))
 
     df = pd.concat(df_pieces)
-    # df = df.sample(frac=.01, replace=True)
+
+    if subsample < 1.0:
+        df = df.sample(frac=subsample, replace=True)
 
     df = df[df.tissue != 'do-not-show']
-    sn.set_context("talk", font_scale=1.8)
+    sn.set_context("talk", font_scale=1. + 0.1*size)
 
     g = sn.ConditionalJointGrid(xlabel, ylabel, df, hue=huelabel, xlim=xlims, ylim=ylims,
-                     size=10, inline_labels=True)
+                     size=size, inline_labels=True)
     # g.plot_joint(plt.scatter)
     g.plot_joint(sn.kdeplot, linewidths=3.0)
     g.plot_marginals(sn.distplot)
@@ -452,7 +454,7 @@ def jointplot_real(imageX, imageY, segmentation, mask,
         plt.setp(ax.get_xticklabels(), visible=False)
 
     if out_file is not None:
-        plt.savefig(out_file, dpi=200, bbox_inches='tight')
+        plt.savefig(out_file, dpi=dpi, bbox_inches='tight')
     return g
 
 

### 2015-11-05 Publications/2015-NeuroImage/responseLetter.tex

@@
 \end{reviewpoint}
-\reply{}
+\reply{
+We thank the reviewer for pointing out that the overall goal is not clearly explained.
+This paper focuses on segmenting very accuratelly dMRI data, to support the subsequent processing tasks
+  such as image reconstruction (e.g. spherical deconvolution) and tractography.
+Since we are aligning very accurate surfaces representing the anatomical structure of the subject, we
+  implicitly find the mapping between structural and dMRI spaces.
+This mapping is a precious result since all the anatomical information is usually obtained in the structural
+  space (generally from T1w scans).
+
+As the reviewer points out, the misregistration between structural and dMRI spaces is generally two-fold:
+  a linear component because the rigid head-motion inside the scanner between the acquisition of the
+  T1w and the dMRI images; and a nonlinear warping aligned with the phase-encoding direction of dMRI
+  that is intrinsic to the EPI schemes used for acquisition.
+The first component of the misalignment w.r.t. the structural space is usually solved with afine registration, as
+  described in lines 29-38.
+The second component is the nonlinear warping, solved by the so-called ``EPI distortion correction'' methods.
+These methods find the deformation field underlying the distortion, and resample the dMRI dataset on the structural space
+  unwarping the data.
+There are two principal diverging points between \emph{regseg} and standard correction methods:
+  1) we do not resample the dMRI data in the structural space (meaning, we do not actually correct);
+  and 2) EPI-distortion correction methods generally require additional MR acquisitions like fieldmaps, reversed-encoding
+  DWIs, or T2w images.
+We discuss this second advantage in lines 241-243.
+Since our method can be seen as an alternative to EPI distortion correction, we did not aim at comparing with
+  these in this work, and left it for future publications.
+
+We understand that the manuscript did not clearly state the overall goal of the method, and we updated it accordingly: 
+ {\color{red} say what here}.
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+The reviewer is right, we have changed the sentence as follows:
+``These limitations prevent {\color{blue} partitioning the native dMRI space in homogeneous regions 
+  of diffusion features}''.
+We agree with the reviewer that there actually exist many methods to segment fiber tracts.
+However, our method is intended to support fiber tracking methods requiring detailed anatomical
+  information, such as \citep{smith_anatomicallyconstrained_2012}.
+Therefore, we defined our method without the support of a tractogram to evolve the contours.
+This could be a future line to be investigated.
+
+In order to clarify this point, we have modified the manuscript as follows: {\color{red} say what here}.
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+This comment is very related to \autoref{com:14}. We have made the following amendments to the manuscript:
+{\color{red} say what here}.
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+Again, the reviewer finds that the main goal of the paper is not clearly stated.
+We understand that modifications derived from the previous comments satisfy the answer to the concern,
+  and changes include abstract and introduction.
+We identified and removed some unrelated references {\color{red} say which ones here}, and added
+  some references regarding correction of EPI distortions {\color{red} say which ones here}.
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+In \autoref{com:9} we addressed the differences between EPI distortion correction and our method.
+Even though we do not aim at correcting for distortions, we understand that \emph{regseg} can replace
+  this task.
+Therefore, we cannot deny the closeness of EPI distortion correction to our field and added the
+  suggested reference.
+This way, the main three correction methods most widely used are cited in our paper (fielmap based correction \citep{jezzard_correction_1995},
+  blip-up/blip-down correction \citep{irfanoglu_drbuddi_2015}, and T2w-b0 registration \citep{kybic_unwarping_2000}).
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+We thank the reviewer for positively weighing our experiments and results.
+As stated in line 189, we use the settings from the \emph{ExploreDTI} software package \citep{leemans_exploredti_2009}, that
+  internally uses \emph{elastix} \citep{klein_elastix_2010} in registration.
+Settings for \emph{elastix} are available in \url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data},
+  and they are reproduced from the p-code version of \emph{ExploreDTI} handed over by Dr. A. Leemans on May 20th, 2014.
+However, it is possible that the information about the B0-T2 method is very scarce and we introduced
+  further explanations accordingly: {\color{red} say what here}.
+}
 

### 2015-11-05 Publications/2015-NeuroImage/00-main.tex

@@
 \usepackage[utf8]{inputenc}
-
+\defcitealias{esteban_useful_2015}{Supplemental Materials}
 \usepackage{xifthen}
@@
 \providecommand{\nvec}[1]{\hat{\mathbf{#1}}}
-\providecommand{\suppl}[1]{\href{http://figshare.com/s/459c26b4ee8211e493b306ec4bbcf141}{Supplemental Materials}, #1}
+
+% \providecommand{\suppl}[1]{\citepalias{esteban_useful_2015}  \href{http://figshare.com/s/459c26b4ee8211e493b306ec4bbcf141}{Supplemental Materials}, #1}
 \providecommand{\isores}[2][]{#2\ensuremath{\times}#2\ensuremath{\times}#2%
@@
 \title{Surface-driven registration method for the structure-informed segmentation of diffusion MR images\tnoteref{t1}}
-\tnotetext[t1]{Supplemental materials to this work will be publicly available upon publication with doi:\href{http://figshare.com/s/459c26b4ee8211e493b306ec4bbcf141}{10.6084/m9.figshare.1397502}.
-During the review process, the document is available with the private link.}
+\tnotetext[t1]{Supplemental materials to this work have been submitted to Data in Brief \citep{esteban_useful_2015}.
+During the review process, the document is available in doi:\href{http://figshare.com/s/459c26b4ee8211e493b306ec4bbcf141}{10.6084/m9.figshare.1397502}.}
 

### 2015-11-05 Publications/2015-NeuroImage/08-appendix-01.tex

--- 2014-10-21 Publications/2013-TMI/08-appendix-01.tex
@@
 
-{\color{red} {Using the region descriptors derived in \autoref{sec:methods_map}, we propose
+{\color{red} {Using the region descriptors derived in \autoref{sec:regseg-methods_map}, we propose
 an \gls{adf}-like, \gls{acwe}-based, piece-wise constant, image segmentation
@@
 \cite{chan_active_2001} with the energy functional obtained in
-\eqref{eq:final_map_energy}. This inverse problem is ill-posed
+\eqref{eq:regseg-final_map_energy}. This inverse problem is ill-posed
 \cite{bertero_illposed_1988,hadamard_sur_1902}.
@@
 + \int \vec{u}^T \, A \, \vec{u} \, d\vec{x} + \int \tr\{(\nabla \vec{u}^T)^T B (\nabla \vec{u}^T)\} d\vec{x}
-\label{eq:complete_energy}
+\label{eq:regseg-complete_energy}
 \end{multline}
@@
 It is necessary to discretize $t$ in order to obtain a numerical
-implementation of the equation \eqref{eq:gradient_final}:
+implementation of the equation \eqref{eq:regseg-gradient_final}:
 \begin{align}

### 2015-11-05 Publications/2015-NeuroImage/00-main.tex
fix nonreview version

@@
 %
+\usepackage{xcolor}
 \usepackage{url}
 \usepackage[colorlinks=true]{hyperref}
-% \usepackage[hyperref]{xcolor}
 \usepackage{doi}
@@
 \hyphenation{op-tical net-works semi-conduc-tor an-isot-ropy regis-tra-tion iso-tropic Free-Surfer %
-align-ed pre-sent work-flow data-set data-sets micro-struc-tu-re}
+align-ed pre-sent work-flow data-set data-sets micro-struc-tu-re Ma-drid non-linear de-for-ma-tion
+an-iso-tro-pic Stud-holme seg-ment-a-tion re-gis-tra-tion map-ped achiev-ed}
 
@@
 
-\linenumbers
+\ifthenelse{\boolean{review}}{\linenumbers}{}
 \input{03-introduction}

### 2015-11-05 Publications/2015-NeuroImage/04-methods.tex
fix nonreview version

@@
   P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
-  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) =   \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
+  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} )
+  \ifthenelse{\boolean{review}}{=}{= \notag\\ &=}
+  \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
   \mdist{f'}{l} \right)}},
@@
   \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) = \notag\\ &=
-  \const + \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, +   \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
+  \const + \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \,
+  \ifthenelse{\boolean{review}}{+}{+ \notag\\ &+}
+  \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
   \label{eq:regseg-energy}
@@
   \frac{ \partial }{\partial \vec{u}_k} \Big\{
-  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, +   \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
+  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \,
+  \ifthenelse{\boolean{review}}{+}{+ \notag\\ &+}
+  \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
   + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
@@
   \regseg{} with them is presented in \autoref{fig:regseg-evphantoms}.
-A set of four binary objects (i.e., ``box'', ``ball'', ``L'',
+A set of four binary objects (i.e. ``box'', ``ball'', ``L'',
   and ``gyrus'') was generated by combining the binarization of

### 2015-11-09 Publications/2015-NeuroImage/03-introduction.tex
DZ's comments

@@
   \revcomment[C13 (R.2)]{orientation}.
-Subsequent processing involves describing the local microstructure with one available
-  model, which ranges from the early \gls*{dti} proposed by \cite{basser_microstructural_1996}
+Subsequent processing involves describing the local microstructure with one of the available
+  models, which range from the early \gls*{dti} proposed by \cite{basser_microstructural_1996}
   to current models such as \citep{daducci_accelerated_2015}.
@@
 \revcomment[rephrased]{%
-Here we present a segmentation and surface-to-volume registration method
+Here, we present a segmentation and surface-to-volume registration method
   named \regseg{}, and show its usefulness in mapping anatomical information from structural
@@
 \revcomment[C1 (R.1) \\ C10 (R.2)]{%
-To implement \regseg{} we first establish an active-contours without edges
+To implement \regseg{} we first establish an active contours without edges
   \citep{chan_active_2001} segmentation framework.}
@@
 A specific set of reference surfaces extracted from the same subject initialize
-  the 3D active-contours, which evolve searching for homogeneous regions in the multivariate
+  the 3D active contours, which evolve searching for homogeneous regions in the multivariate
   target-image.

### 2015-11-09 Publications/2015-NeuroImage/06-discussion.tex
fix nonreview version
DZ's comments

@@
 \revcomment[C5 (R.1)]{%
-Finally, we also contributed proposing a piecewise-smooth segmentation model defined by
+Finally, we also proposed a piecewise-smooth segmentation model defined by
   a selection of nested surfaces to partition the multispectral space
@@
 First, the \glspl*{dwi} are corrected for \emph{\gls*{epi} distortions} by estimating
-  the nonlinear-deformation field from extra MR acquisitions
+  the nonlinear deformation field from extra MR acquisitions
   \citep{jezzard_correction_1995,chiou_simple_2000,cordes_geometric_2000,kybic_unwarping_2000}.

### 2015-11-10 Publications/2015-NeuroImage/responseLetter2.tex
suggestions from AS

@@
 
-\usepackage[numbers,sort&compress]{natbib}
+\usepackage[sort&compress]{natbib}
 \usepackage{xcolor}
@@
 To clarify this point, we have stated more clearly that the choice of surfaces to segment the target data is
-  crucial (lines 46--47, p.2).
+  crucial (lines 46--50, p.2).
 We have better defined the segmentation model that drives the registration process (lines 94--95, p. 5).
@@
   we have extracted the conditional distributions to the right side of each panel.
-We expect these modifications make clear the difference between the surfaces involved in driving the registration
+We expect that these modifications make clear the difference between the surfaces involved in driving the registration
   process and the surfaces in which we evaluate the error of the method (lines 217--218, p. 10) are a subset of the
@@
 We have moved the related methods and our methodological contributions from the Introduction to a more appropriate
-  place in the Discussion (lines 248--275), additionally satisfying \autoref{com:9} (Reviewer \#2) suggesting to include
+  place in the Discussion (lines 261--278), additionally satisfying \autoref{com:9} (Reviewer \#2) suggesting to include
   in the introduction only the references related to the focus of the paper.
@@
   of the descriptive sections of the paper give a successful answer to this issue.
-Regarding the actual inputs of \emph{regseg}, we have introduced modifications in the abstract, in lines 45--50 (p. 2), and
-  in lines 243--243 (p. 13).
+Regarding the actual inputs of \emph{regseg}, we have introduced modifications in the abstract, and in lines
+  45--50 (p. 2) and lines 243--244 (p. 13).
 }
@@
 The reviewer is right in the sense that we did not clearly state what we are segmenting, and therefore
-  the references in line 24 are a bit orphaned.
+  the references in the former line 24 were a bit orphaned.
 All methods reviewed in lines 15--27 cluster/segment/classify \emph{voxels} in dMRI space based on
@@
   internally uses \emph{elastix} \citep{klein_elastix_2010} in registration.
-A footnote in line 213 (p. 10) directly links the settings for this method.
+A footnote in line 212 (p. 10) directly links the settings for this method.
 }
@@
 Additionally to the edits in \autoref{com:10}, we have more clearly defined the concept with the discussion
-  introduced in lines 267--268 (p. 13).
+  introduced in lines 267--270 (p. 13).
 }

### 2015-11-10 Publications/2015-NeuroImage/06-discussion.tex
suggestions from AS

@@
 In these studies, the surfaces obtained in a mature timepoint of the brain are retrospectively
-  propagated to the initial timepoints, regardless the changes of the contrast and spatial
+  propagated to the initial timepoints, regardless of the changes in the contrast and spatial
   development between them.

### 2015-11-11 Publications/2015-NeuroImage/06-discussion.tex
final fixes for resubmission

@@
 The nonlinear registration process evolves driven by the fitness of the
-  picewise-smooth classification of voxels in the target volume imposed
+  piecewise-smooth classification of voxels in the target volume imposed
   by the current mapping of the surfaces.}

### 2016-01-16 Publications/2015-NeuroImage/02-abstract.tex
add ending sentence to abstract (MJLC)
remove resubmission comments

@@
   require precise delineations of anatomical structures.
-\revcomment[C1 (R.1)]{%
 This requirement has been approached either segmenting the data in native \gls*{dmri} space
@@
   as the geometrical distortions caused by the inhomogeneity of magnetic susceptibility
-  across tissues hinder both solutions.}
-\revcomment[C10 (R.2)]{%
+  across tissues hinder both solutions.
 Unifying the two approaches, we propose \regseg{}, a surface-to-volume nonlinear
   registration method that segments homogeneous regions within multivariate images by mapping
-  a set of nested reference-surfaces.}
-\revcomment[C11 (R.2)]{%
+  a set of nested reference-surfaces.
 Accurate surfaces are extracted from a \gls*{t1} image of the subject, using as target image
   the bivariate volume comprehending the \gls*{fa} and the \gls*{adc} maps derived from the
-  \gls*{dmri} dataset.}
-\revcomment[rephrased]{%
+  \gls*{dmri} dataset.
 We first verify the accuracy of \regseg{} on a general context using digital phantoms.
@@
 Finally, we cross-compare the proposed tool against a nonlinear \lowb{}-to-T2w registration
-  method, thereby obtaining a significantly lower misregistration error with \regseg{}.}
+  method, thereby obtaining a significantly lower misregistration error with \regseg{}.
+Therefore, we demonstrate that \regseg{} allows the accurate mapping of structural information
+  in \gls*{dmri} space, enabling the application of new structure-informed techniques in
+  the connectome extraction.

### 2016-01-16 Publications/2015-NeuroImage/03-introduction.tex
remove resubmission comments

@@
 Each \gls*{dwi} is sensitized with a gradient to probe proton diffusion in a certain
-  \revcomment[C13 (R.2)]{orientation}.
+  orientation.
 Subsequent processing involves describing the local microstructure with one of the available
@@
   to current models such as \citep{daducci_accelerated_2015}.
-The microstructural map is then used to draw the preferential \revcomment[C13 (R.2)]{orientations} of diffusion
+The microstructural map is then used to draw the preferential orientations of diffusion
   across the brain using tractography \citep{mori_threedimensional_1999}.
@@
   tractography as edges \citep{hagmann_mapping_2008}.
-\revcomment[C9 (R.2) \\ C17 (R.2)]{%
 The methodologies to solve reconstruction, tractography and network building
@@
   tractography \citep{smith_anatomicallyconstrained_2012} are increasingly using
-  structural information to improve the microstructural mapping and fiber-tracking.}
+  structural information to improve the microstructural mapping and fiber-tracking.
 
-\revcomment[C15 (R.2)]{%
 Possibly, the earliest structural information incorporated to aid \gls*{dmri} processing
@@
   \citep{taoka_fractional_2009}.
-\todo{C16 (R.2)}
 To overcome the unreliability of \gls*{fa} thresholding, and to broaden
@@
   rousson_level_2004,jonasson_segmentation_2005,hadjiprocopis_unbiased_2005,liu_brain_2007,
-  lu_segmentation_2008,han_experimental_2009}.}
-\revcomment[C14 (R.2)]{%
-However, the precise segmentation of \gls*{dmri} is difficult for several reasons.}
-First, \gls{dmri} images have a resolution that is \revcomment[C12 (R.2)]{much lower}
+  lu_segmentation_2008,han_experimental_2009}.
+However, the precise segmentation of \gls*{dmri} is difficult for several reasons.
+First, \gls{dmri} images have a resolution that is much lower
   than that of the imaged microstructural features.
@@
 
-\revcomment[C18 (R.2)]{%
 An alternative route to segmentation in \gls*{dmri} space is the mapping of the
@@
 A standardized method to solve the remaining linear mapping between the corrected-\gls*{dmri}
-  and the structural spaces is \emph{bbregister} \citep{greve_accurate_2009}.%
-}
+  and the structural spaces is \emph{bbregister} \citep{greve_accurate_2009}.
 
-\revcomment[rephrased]{%
 Here, we present a segmentation and surface-to-volume registration method
@@
 The underlying hypothesis is that the registration and segmentation problems
-  in \gls*{dmri} can be solved simultaneously.}
-\revcomment[C1 (R.1) \\ C10 (R.2)]{%
-To implement \regseg{} we first establish an active contours without edges
-  \citep{chan_active_2001} segmentation framework.}
-\revcomment[C2 (R.1)]{%
+  in \gls*{dmri} can be solved simultaneously.
+To implement \regseg{} we first establish an active-contours without edges
+  \citep{chan_active_2001} segmentation framework.
 A specific set of reference surfaces extracted from the same subject initialize
-  the 3D active contours, which evolve searching for homogeneous regions in the multivariate
+  the 3D active-contours, which evolve searching for homogeneous regions in the multivariate
   target-image.
@@
 Therefore, \regseg{} establishes a registration framework that actually
-  deals with the nonlinear warping induced by \gls*{epi} distortions.%
-}
+  deals with the nonlinear warping induced by \gls*{epi} distortions.
 \Regseg{} integrates the benefits of segmentation and registration methods together and

### 2016-01-16 Publications/2015-NeuroImage/04-methods.tex
preparing resubmission
remove resubmission comments

--- 2015-11-04 Publications/2015-NeuroImage/04-methods.tex
@@
   for each region $\Omega_l$.
-\revcomment[C4 (R.1)]{%
 Even though the features being segmented are not generally i.i.d., the spatial interdependency of
@@
 In \autoref{fig:regseg-model} it is shown that marginal distributions of data can be approximated by
-  multivariate normal distributions.} Therefore,
+  multivariate normal distributions. Therefore,
 
@@
   the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized.
-\revcomment[C2 (R.1)]{%
 In \autoref{fig:regseg-model},
   this minimization is illustrated for the segmentation of the \gls*{fa} and the \gls*{adc} maps
-  of one subject.}
+  of one subject.
 
@@
 \caption{%
-\revcomment{%
 Evolution of the segmentation model defined by the homogeneous regions $\Omega_l$
@@
   are minimized with respect to panel A, and the resulting segmentation is close
-  to the theoretical situation presented in panel B.}
-}\label{fig:regseg-model}
-\todo[inline]{C5 (R.1)}
+  to the theoretical situation presented in panel B.}\label{fig:regseg-model}
 \end{figure*}
@@
 
-\revcomment[C20 (R.2)]{%
 A first proof of concept is introduced to demonstrate \regseg{} in digital phantoms
@@
 Therefore, the deformation field is nonzero only in the \gls*{pe} axis,
-  and resembles a real \gls*{epi} distortion.}
+  and resembles a real \gls*{epi} distortion.
 The adaptation of the evaluation protocol to the simulated phantoms and the real data is
@@
 \paragraph*{Segmentation model and settings}
-\revcomment[C6 (R.2)]{%
 The segmentation model of the phantoms is implicitly defined:
@@
 All the experimental settings used for the phantoms are made available in
-  a unique configuration file\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/regseg_default.json}}.}
+  a unique configuration file\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/regseg_default.json}}.
 
@@
   5) The results were evaluated visually and quantitatively.
-  \revcomment{%
   The arrows point to edges in the target images (light-yellow arrows for \gls*{fa}, blue for \gls*{adc}) that
     should be aligned with a surface, showing how distortion limits the direct mapping from the structural space
-    in which the contours are defined.}}\label{fig:regseg-evworkflows}
-  \todo[inline]{C7 (R.1) \& C23 (R.2)}%
+    in which the contours are defined.}\label{fig:regseg-evworkflows}
 \end{figure*}
@@
   multi-shell \gls*{dmri} images.
-\revcomment[C7 (R.1)]{%
 Since we obtained the \gls*{dmri} data from the minimally preprocessed package, these
   images are corrected for \gls*{epi} distortions and spatially normalized in
-  \gls*{t1} space.}
+  \gls*{t1} space.
 
@@
 \paragraph*{Settings} %
-\revcomment[C5 (R.1)]{%
 \Regseg{} accepts an affine mapping from surface-space to the \gls*{dmri} data as initialization.
 However, the images provided by the \gls*{hcp} are already spatially normalized.
-Therefore, the initial estimation of distortion is zero in this experiment.}
-\revcomment[C20 (R.2)]{%
+Therefore, the initial estimation of distortion is zero in this experiment.
 Since the distortion $U_\text{true}$ is aligned along the \gls*{pe} direction ($y$-axis in our settings),
-  \regseg{} was configured to allow nonzero displacements only on that corresponding direction.}
-\revcomment[C6 (R.1)]{%
+  \regseg{} was configured to allow nonzero displacements only on that corresponding direction.
 For the experiments on real data, \regseg{} established a multi-resolution pyramid of
@@
 The actual choices of the parameter settings are publicly distributed with the source code for the
-  experiments\footnote{\url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data/regseg_hcp.json}}.}
+  experiments\footnote{\url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data/regseg_hcp.json}}.
 These settings were obtained manually, driven by the feedback obtained from the post-registration
@@
   perform registration.
-\revcomment[C21 (R.2)]{%
 The deformation field is correspondingly restricted to the \gls*{pe} direction.
 The settings file for \emph{elastix} is also
-  available\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/t2b_elastix_y.txt}}.}
+  available\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/t2b_elastix_y.txt}}.
 
@@
   that correspond to $\vec{v}_i$.
-\revcomment[C2 (R.1)]{%
 In practice, we only report the \gls*{swindex} for three surfaces (\{$\Gamma_{VdGM}, \Gamma_{WM}, \Gamma_{pial}$\})
-  of crucial interest in whole-brain tractography.}
+  of crucial interest in whole-brain tractography.

### 2016-01-16 Publications/2015-NeuroImage/06-discussion.tex
remove resubmission comments

--- 2015-11-09 Publications/2015-NeuroImage/06-discussion.tex
@@
 \label{sec:regseg-discussion}
-\revcomment[C10 (R.2)]{%
 We present \regseg{}, a simultaneous segmentation and registration method that
-  maps a set of nested surfaces into a multivariate target-image.}
-\revcomment[C1 (R.1)]{%
+  maps a set of nested surfaces into a multivariate target-image.
 The nonlinear registration process evolves driven by the fitness of the
   picewise-smooth classification of voxels in the target volume imposed
-  by the current mapping of the surfaces.}
-\revcomment[moved]{%
+  by the current mapping of the surfaces.
 We propose \regseg{} to map anatomical information extracted from \gls*{t1}
@@
   that supports the evolution of the active contours (segmentation),
-  like \cite{paragios_level_2003,yezzi_variational_2003}.}
+  like \cite{paragios_level_2003,yezzi_variational_2003}.
 % Conversely, in structure-informed segmentation, the sources of variability are the geometrical distortions
 %   after imaging and the anatomical evolution in longitudinal studies.}
-\revcomment[C1 (R.1)]{%
 \Regseg{} can be seen as a particular case of atlas-based segmentation-registration methods,
-  replacing the atlas by the structural image of the subject (\emph{structure-informed segmentation}).}
-\revcomment[C3 (R.1)]{%
+  replacing the atlas by the structural image of the subject (\emph{structure-informed segmentation}).
 The main difference of atlas-based segmentation and the application at hand is the resolution of the
@@
 As an advantage, the use of explicit surfaces enables segmenting \gls*{dmri} images
-  with accuracy below pixel size.}
+  with accuracy below pixel size.
 
-\revcomment[C3 (R.1)]{%
 An important antecedent of \regseg{} is \emph{bbregister} \citep{greve_accurate_2009}.
@@
 Indeed, the distortion must be addressed separately because it is not supported by
-  the affine transformation model.}
-\revcomment[C22 (R.2)]{%
+  the affine transformation model.
 Conversely, the deformation model of \regseg{} is nonlinear and the active contours are
@@
   do not present steep image gradients (edges) but the anatomy can be identified
-  by looking for piece-wise smooth homogeneous regions.}
+  by looking for piece-wise smooth homogeneous regions.
 
-\revcomment[C3 (R.1)]{%
 Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
@@
   generally included in most of the existing methodologies, and the complexity of
-  the elasticity smoother of \cite{guyader_combined_2011}.}
-\revcomment[C3 (R.1)]{%
+  the elasticity smoother of \cite{guyader_combined_2011}.
 Other minor features that differ from current methods in joint segmentation and registration are
   the support of multivariate target-images and the efficient computation of the shape-gradients
-  implemented with sparse matrices.}
+  implemented with sparse matrices.
 
@@
 Therefore, the error was bounded above by half of the voxel size.
-\revcomment[C8 (R.1)]{%}
 The distributions of errors along surfaces varied importantly depending on the shape of the
@@
   convoluted nature of cortical surfaces and the directional restriction of the
-  distortion.}
+  distortion.
 
@@
   but indicated that \regseg{} is a reliable option in this application field.
-\revcomment[C5 (R.1)]{%
-Finally, we also proposed a piecewise-smooth segmentation model defined by
+Finally, we also contributed proposing a piecewise-smooth segmentation model defined by
   a selection of nested surfaces to partition the multispectral space
@@
   minimum of the objective function impeding the convergence.
-}%
 
-\revcomment[C9 (R.2)]{%
 The proposed application of the method in the task of identifying structural information
@@
 The problem was firstly addressed using image segmentation approaches in the native diffusion
-  space, without definite and compelling results.}
-\revcomment[C18 (R.2)]{%
+  space, without definite and compelling results.
 With the introduction of retrospective correction methods for the \emph{\gls*{epi} distortions}
@@
 First, the \glspl*{dwi} are corrected for \emph{\gls*{epi} distortions} by estimating
-  the nonlinear deformation field from extra MR acquisitions
+  the nonlinear-deformation field from extra MR acquisitions
   \citep{jezzard_correction_1995,chiou_simple_2000,cordes_geometric_2000,kybic_unwarping_2000}.
@@
   the whole connectome extraction process can be performed on the unaltered \gls*{dmri}
-  signal (i.e. without resampling data to an undistorted space).}
+  signal (i.e. without resampling data to an undistorted space).
 
-\revcomment[C9 (R.2)]{%
 Beyond the presented application on \gls*{dmri} data, \regseg{} can be indicated in situations
@@
   injury \citep{irimia_structural_2014} or in the monitorization of brain tumors
-  \citep{weizman_semiautomatic_2014}.}
+  \citep{weizman_semiautomatic_2014}.

### 2016-01-16 Publications/2015-NeuroImage/responseLetter3.tex
modify new response letter

\documentclass[9pt]{memoir}

\usepackage[numbers,sort&compress]{natbib}
\usepackage{xcolor}
\usepackage[T1]{fontenc}
\usepackage{charter}

\usepackage{xcite}
\externalcitedocument{00-main}

\usepackage[colorlinks=true]{hyperref}
\usepackage{nameref}

\usepackage{amsmath}
\usepackage{amssymb}

\newcounter{reviewpoint}
\makeatletter
\newenvironment{reviewpoint}%
{\refstepcounter{reviewpoint}\par\medskip\vspace{3ex}\hrule\vspace{1.5ex}\par\noindent%
   {\fontseries{b}\selectfont Comment \arabic{reviewpoint}:}
   \begingroup%
   \color{black!60}
   \fontshape{it}\selectfont %

}
{\endgroup\label{com:\thereviewpoint}\par\medskip}
\def\reviewpointautorefname{Comment}
\makeatother

\newcommand{\reply}{\par\fontshape{n}\selectfont \noindent \textbf{Reply}:\ }

\setlrmarginsandblock{3cm}{2.5cm}{*}
\setulmarginsandblock{2.5cm}{2.5cm}{*}
\checkandfixthelayout

\begin{document}
\hypersetup{linkcolor=black!60, citecolor=black!60, urlcolor=black!60}

\section*{Response letter to the manuscript no. NIMG-15-1151}
\medskip
Original title: \emph{Active contours-driven registration method for the structure-informed segmentation of
  diffusion MR images}.

\noindent New title: \emph{{\color{blue!70} Surface}-driven registration method for the structure-informed segmentation of
  diffusion MR images}.

\bigskip
\noindent We thank the editors and the reviewers for revising our manuscript no. NIMG-15-1151 submitted to NeuroImage.
We appreciate the comments made by both reviewers, and considered them fully to improve our paper.
Please, find in the following paragraphs our point-by-point answer to the concerns raised by the reviewers,
  along with the description of the corresponding changes in the manuscript.
Text modifications in the manuscript are highlighted with a new font color, and attached notes indicate which reviewer comment
  the corresponding change is addressing to.

\bigskip
\bigskip
\subsection*{Reviewer \#1:}
\begin{reviewpoint}
This paper presents a method for simultaneous registration and segmentation of diffusion images in the presence of geometric distortions. The method uses an anisotropically regularised active contour method to co­align tissue segmentations generated from T1 images to distorted multi­channel diffusion data. It has the advantage that it has no requirements for extra scans to be acquired for correction of the distortion. The method demonstrates improved performance over a previous b0­T2 registration approach for the alignment of tissue segmentation maps. Improvements are demonstrated on simulated data, and real data with simulated distortions. The advantage of these results is that the ground truth is known.
I am satisfied that the active contour alignment approach used to align the surfaces is sound, and that the results show theoretic improvements over the T2B method. I am convinced that the method approach is effective for the alignment task. In general the methods are well explained with the exception of a few minor points (below).
My main concerns lie with whether the results in the paper prove what the authors set out in their abstract. Specifically, a stated goal is for the use of anatomically constrained tractography. However, there are no results that show this method would improve connectome extraction. In particular, the authors state that ``this method avoids any requirement for collecting additional scans nor resampling for diffusion unwarping''. I see this has advantages, but what about the consequences of not unwarping on the diffusion tracts themselves see [1][2][3]. In my opinion the authors should address this point possibly with additional results to show how they intend this to be used within a tracking framework.
I also have some concern that the method is not applied to real distortions nor compared to state­of­the­art correction methods such as EDDY [4]. This was used for the HCP corrections therefore it might provide a good basis comparison basis on real rather than modeled distortions for brain data. If this could not be done for some reason I feel it should be discussed explicitly. It should be possible to compare to bbregister for the unwarped data. Why is this not done?
[1] Jones, Derek K., and Mara Cercignani. ``Twenty‐five pitfalls in the analysis of diffusion MRI data.'' NMR in Biomedicine 23.7 (2010): 803­820.
[2] Lee, J., et al. ``Correction of Bo EPI distortions in diffusion tensor imaging and white matter tractography.'' Proc. Int. Soc. Magn. Reson. Med. Vol. 11. 2004.
[3] Andersson, J. L., et al. ``Effects of susceptibility distortions on tractography.''Proceedings of International Society of Magnetic Resonance in Medicine. Vol. 11. 2004.
[4] A comprehensive Gaussian Process framework for correcting distortions and movements in diffusion images Jesper L. R. Andersson et al ISMRM
Minor points:
Pg 5 regularisation: `` requires that the distortion and its gradient have zero mean and variance governed by the matrices A and B'' suggest adding a comma here for easier interpretation i.e. ``requires that the distortion and its gradient have zero mean, and variance governed by the matrices A and B''
Pg 7, line 4 `` anisotropy is aligned''. What is meant by anisotropy in this context, the anisotropy of the deformation gradient.
Figure 2. This looks nice, but its not easy to interpret. I can't see a clear improvement through A­C. I think there needs to be direct comparisons between the conditions represented by A­C on one plot. Why not have 6 plots for each tissue class comparing distributions for each condition.
\end{reviewpoint}
\reply{%

}
% ------------------------------------------------------------------- Reviewer 3
\bigskip
\bigskip
\subsection*{Reviewer \#3:}
\begin{reviewpoint}
This paper presents a new method for correction of EPI distortion in diffusion MRI data by performing simultaneous registration and segmentation of a set of nested surfaces, using a multi­channel FA/MD image with a registration target of an undistorted T1 image. The primary advantage over other methods is that cortical surfaces and white­gray matter boundaries are apparently aligned well by this method. Results show that it outperforms one implementation of the more standard baseline to T2 registration. Overall this is a quite interesting work, and it will be a good addition to the literature and also to the body of open­source code for diffusion MRI.
I have some comments for clarification of the paper.
Overall, I am slightly confused by some aspects of the equations, and I think that some information from the supplement (which I have not had time to read) should be moved into the paper to clarify missing points. The paper should be able to be read independently of the supplement, which does not seem to be the case currently. Specifically, the relationship of the b­spline registration framework to the evolution of the active contours is a bit confusing, with a main equation given inside Figure 1, and apparently all other detail in the supplement. This could be as simple as an additional paragraph or two.
Please state that the goal of the paper is to improve connectome­based diffusion MRI research. The processing description on pages 1­2 assumes that the connectome­based approach is the only approach to analysis of diffusion MRI. But this is clearly not the case (surgical planning, TBSS, etc).
``to current models such as (Dadicci et al)'' is confusing. Please give the name(s) of the current model(s).
￼￼On page 3, please define N (section 2.1)
In Figure 1, it is a bit confusing that active contours appear, because the abstract did not mention active contours directly, and the evolution following the inward normals at each vertex is not explicit in the equations of the paper.
It seems the evolution at $s_1$ is in the direction opposite the inward normal, so the caption of Figure 1 is a bit confusing, since it directly states the inward normals are followed, rather than just the normal to the curve.
I recommend adding a little bit of text to unify figure 1 and the surface evolution framework with the concepts presented in the registration equations section. I think this information may be in the supplement.
On page 5, equation (4), I think f should be f prime on the first line.
Please describe A and B in a sentence after equation 5. It is not clear how these were chosen (or at least refer to the appendix).
On page 5, equation (7), the ­log is removed from the second line. I would expect a sum of log probabilities, with the log preserved, rather than removing the log altogether. It seems this is an oversight. Or do the authors mean to use a proportional symbol instead of an equals symbol? Please clarify.
In equation (7), the equation that is split onto two lines (why?) has two plus symbols in a row, one on each line. In equation (7), the second integral is over omega, while the first is over $omega_l$. Why?
In equation (7) please clarify with brackets what the sum over l is applied to.
The two pluses are present again in equation (10).
Please explain in a sentence what alpha and beta are in equation (10) (expected variances along each axis of the coordinate system, or something).
$g_k$ is completely undefined in equation (11). I see there is information in a supplement but at a minimum each variable should be defined in a sentence in the main paper.
pg 7. Please define PE acronym or write out phase encoding.
eq (14). Please define what is $a_i$.
Figure 5. What are the little numbers (26, 30, 33) in the lower left corner of the phantom images? Figure 5. Please state clearly in the caption that this is digital phantom data, to help the reader.
Figure 5. Over what is the average error of surfaces computed? Average over the surface? Then the variability in the plot is across experiments? Or the other way around with the average at each surface point across experiments?? It is not clear.
Figure 5 A. Why are there so many phantom images shown? They all look the same to me. Please explain and/or remove the duplicates.
Figure 6. Are the plots in B from one or many subjects? Please state in the caption.
In Table 2 caption how many subjects were input to this experiment? This is given as 16 in the abstract, but that info is not very near the actual results and would increase their impact.
pg 13. ``We also show the 95\% CIs of the sWI for these surfaces.'' sounds like it is referring to Figure 6. However it is apparently in Table 2, which should be referenced in this sentence. Or the sentences about Table 2 should be reorganized to sound separate from Figure 6.
in Appendix (and also elsewhere): ``the anisotropy is aligned with the imaging axes'' is quite confusing. Usually anisotropy refers to FA or similar. I believe this is the anisotropy of the covariance matrices, really that their eigenvectors are assumed aligned with the voxel coordinate system. Please clarify the usage of ``anisotropy'' in the registration framework.
Close
What values are used for alpha and beta in the appendix?
minor grammatical issues:
abstract: ``approached either segmenting'' ­> ``approached by either segmenting'' pg 2: ``threshold set'' sounds like a mathematical set ­> ``threshold that is chosen'' pg 5. ``displacements field'' ­> ``displacement field''
pg. 15 ``monitorization'' is not a word ­> monitoring

\end{reviewpoint}
\reply{%
}

\bibliographystyle{mystyle}
\bibliography{Remote}

\end{document}

### 2016-01-16 Publications/2015-NeuroImage/responseLetter.tex

--- 2015-09-03 Publications/2015-NeuroImage/responseLetter.tex
@@
 
-\usepackage[hidelinks]{hyperref}
+\usepackage[colorlinks=true]{hyperref}
 \usepackage{nameref}
+
+\usepackage{amsmath}
+\usepackage{amssymb}
 
@@
 
-\noindent New title: \emph{\color{red} Segmentation of homogeneous regions of diffusion MR features by mapping
-  same-subject surfaces extracted in structural space}.
+\noindent New title: \emph{{\color{blue!70} Surface}-driven registration method for the structure-informed segmentation of
+  diffusion MR images}.
 
@@
 We appreciate the comments made by both reviewers, and considered them fully to improve our paper.
-Please, find in the following our point-by-point answer to the concerns raised by the reviewers,
+Please, find in the following paragraphs our point-by-point answer to the concerns raised by the reviewers,
   along with the description of the corresponding changes in the manuscript.
+Text modifications in the manuscript are highlighted with a new font color, and attached notes indicate which reviewer comment
+  the corresponding change is addressing to.
 
@@
 \end{reviewpoint}
-\reply{
-The first sentence of the comment is a rather accurate interpretation of our work.
-However, it is important to clarify that our method maps anatomically-correct surfaces of the same subject into dMRI space,
-  thereby segmenting this target space.
-The method works regardless how the surfaces were extracted, or if they were extracted from T2w images or manually delineated.
-In this work, the surfaces are extracted from T1w for two reasons: 1) it is the widely-accepted MR sequence for such extraction,
-  and 2) the structural information used in reconstruction, tractography and network definition of connectivity extraction pipelines
-  is usually defined in T1w-space as well.
-We understand that the manuscript was not clear to this respect, and therefore included the following changes:
-\begin{enumerate}
-\item Changed the title
-\item Changed abstract: {\color{red} say what here}
-\item Changed introduction: {\color{red} say what here}
-\end{enumerate}
-
-Regarding the second sentence, we agree on that our method can be seen as a registration of an atlas to the
-  image being segmented.
-However, we need to point out some particular features of the method.
-First is that surfaces in our method play the role of \emph{strong} priors, since the method does not allow
-  for changes in their topology (affecting the segmentation side of the approach) and they need to be
-  extremely accurate so that the mapping from structural space is correct (registration).
-This might not be the case on atlas-based segmentation, which generally needs to deal with problems by the
-  inappropriate topology of surfaces in the atlas, and the result of registration is relatively tolerant
-  to the imperfect location of surfaces from the template.
-Our method can be seen as a segmentation framework using active contours without edges in which the evolution
-  of the contours is governed by an elastic registration method;
-or as a nonlinear registration method which cost function is built on top of a segmentation model of active contours
-  without edges.
-In our opinion, the presented approach is mainly a segmentation method.
-However, since the surfaces are supposed to be very accurate and defined in a different coordinate system,
-  the method implicitly registers that space to the target image.
-Some comments of Reviewer \#2 are very similar to this one (particularly \autoref{com:11}).
-
-In order to clarify this issue, we have modified the manuscript as follows: {\color{red} say what here}.
+\reply{%
+The description of the method given by both reviewers (see \autoref{com:10}, Reviewer \#2) lead us to think that the paper
+  did not properly describe the actual inputs and outputs the proposed methodology.
+In general, we have almost-completely reformulated the abstract, introduction and discussion to define this
+  point clearly.
+Particularly, the following sections have been introduced to give an answer to this particular comment:
+\begin{itemize}
+\item Lines 45--53 (p. 2).
+\item Lines 244--246 (p. 13), here we introduce the previously missing discussion to call the
+  method ``simultaneous registration and segmentation''.
+\item Lines 250--252 (p. 13), we comment here on the similarity of \emph{regseg} to the atlas-based segmentation
+  using joint registration and segmentation methods.
+\end{itemize}
 }
@@
 \reply{
-We have noticed that the way we explained the surfaces necessary to segment real dMRI datasets
-  was not appropriate.
-On one hand, in this paper we propose a segmentation model to cluster the multivariate space of features
-  provided by the FA and ADC map.
-As correctly interpreted by reviewer \#2 mentioning it explicitly, and potentially reviewer \#1 noticed,
-  the model is given by a set of nested surfaces.
-However, it was probably unclear in the manuscript that those surfaces and their nesting is the cornerstone
-  of the method, while the set of surfaces selected in validation are only those that will have an important
-  impact in downstream processing, such as reconstruction or tractography.
-
-In order to separate the concepts of validation and the definition of the model, we have modified the
-  manuscript as follows: {\color{red} say what here}.
+To clarify this point, we have stated more clearly that the choice of surfaces to segment the target data is
+  crucial (lines 46--50, p. 2).
+We have better defined the segmentation model that drives the registration process (lines 94--95, p. 5).
+We have also improved the former Figure 4 (now it is Figure 2, p. 5), and brought it earlier in the paper to
+  illustrate earlier the segmentation model.
+Since the former visualization of the conditional (region-wise) joint distribution of FA and ADC was a bit clumsy,
+  we have extracted the conditional distributions to the right side of each panel.
+We expect that these modifications make clear the difference between the surfaces involved in driving the registration
+  process and the surfaces in which we evaluate the error of the method (lines 217--218, p. 10) are a subset of the
+  former.
 }
@@
 \reply{
-We agree with the reviewer on that the novelty of the methodological elements of our approach is very limited.
-However, to our knowledge there exists no method with a paired combination of the following elements:
+We agree with the reviewer on that the novelty of the methodological elements of our approach is limited.
+However, to our knowledge there exists no alternative method combining the same elements to segment
+  multivariate data.
+We have moved the related methods and our methodological contributions from the Introduction to a more appropriate
+  place in the Discussion (lines 261--278), additionally satisfying \autoref{com:9} (Reviewer \#2) suggesting to include
+  in the introduction only the references related to the focus of the paper.
+The particular changes can be listed as follows:
 \begin{itemize}
-\item Region based (i.e. without edges) active contours model
-\item The use of surfaces as prior knowledge to initialize the segmentation model,
-  with property of allowing precise registration if the surface is accurately representative
-  of structures being aligned.
-  Moreover, the multi-phase definition of levelsets is dual to our surface-nesting model.
-\item The anisotropic regularization model of choice, since most of the levelset frameworks
-  use Gaussian smoothing of the deformation field.
-\item Even though the shape-gradients theory was proposed back in the early 2000's, atlas-based
-  registration, and registration-segmentation methods have been generally implemented on levelsets.
+  \item Lines 252--260 (p. 13), we describe the methodological differences w.r.t. \cite{gorthi_active_2011},
+    one of the latest atlas-based segmentation methods using simultaneous registration and segmentation.
+  \item Lines 261--270 (p. 13), we compare the methodological features of our method and \emph{bbregister},
+    an antecedent and standardized tool used on the same application.
+  \item Lines 271--273 (p. 13), we introduce the methodological differences with a recent alternative method in
+    terms of regularization.
+  \item Lines 273--278 (p. 13) mention some other minor methodological contributions.
 \end{itemize}
-To our knowledge, the application of a framework as such is new and an important contribution is done
-  with the proposed segmentation model of the FA and ADC maps.
-We would have appreciated if the reviewers would have provided specific references being close to our approach
-  beyond those given in lines 43-57 of the original paper, so that we could point to the differences of our
-  method.
-In this regard, we have investigated some more potential work related to ours, modifying the manuscript as
-  follows: {\color{red} say what here}.
 }
@@
 \reply{
-  The reviewer is right, assuming that pixels are i.i.d. is wrong but a widely accepted simplification
-that performs correctly with the following details:
-\begin{enumerate}
-\item Pixels are not generally independent, and this issue is generally addressed imposing prior
-  information over neighboring, such as Markov Random Fields.
-In our framework, local dependency of pixels is implicitly addressed by the nested surfaces that impose
-  the piecewise smoothness of the model.
-\item Pixels are not identically distributed, and this issue posed important problems when defining
-  our segmentation model.
-However, we empirically found that the surfaces described in {\color{red} say line here} and their
-  nesting made a good representation of the distribution of each region.
-Actually, this reason may be at basis of the poor performance of most of the methods originally referenced
-  in lines 19-28.
-Our segmentation model ensures that pixels belonging to a certain region are ``almost'' identically distributed.
-The bivariate approach using FA and ADC facilitates this, as demonstrated in Figure 3.
-\end{enumerate}
-
-Nonetheless, we have specifically introduced this discussion in our manuscript {\color{red} say where}
-  and improved Figure 3 to better describe this point.
-}
+The reviewer is right, assuming that pixels are i.i.d. is wrong but a widely accepted simplification.
+We have supported the assumptions inserting the new lines 88--90 (p. 3).}
 
@@
 \reply{
-
+We have improved former Figure 4 (now, it is Figure 2, p. 5) to clarify that the segmentation model is inherently defined by
+  the choice of surfaces and their nesting.
+This is an actual contribution of the work, as stated in lines 304--307 (p. 14).
+A comment on the convergence issues follows in lines 307--311 (p. 14).
+Finally, we described the initialization of the surfaces with the insertion of lines 196--198 (p. 10).
 }
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+We have made a strong effort in providing reproducible experiments using openly available data.
+Therefore, parameter settings are publicly available in GitHub.
+However, we understand that a better description of the parameter settings should be included within
+  the manuscript and modified it accordingly, in lines 200--204 (p. 10).
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+We have clarified this point with the following edits:
+\begin{itemize}
+  \item Lines 171--173 (p. 8): we indicate that we use the ``minimally preprocessed'' images of the
+    HCP database, and thus, these images are acquired with specialized protocols and are
+    already corrected for the distortion.
+  \item Figure 4 (former fig. 3), p. 9, has been modified to clearly indicate this feature.
+\end{itemize}
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+We have added a discussion on this in the corresponding section, lines 287-293 (p. 14).
+Thanks for pointing this systematic error out.
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+We thank the reviewer for pointing out that the overall goal is not clearly explained.
+We have rewritten the abstract, introduction and discussion sections almost fully.
+Particular comments included within this general comment are addressed later: \autoref{com:18}
+  as regards EPI distortions, and \autoref{com:20} for the phase-encoding direction point.
+In order to better focus the application field, we have introduced lines 11--14 (p. 2) and lines
+  312--317 (p. 14).
+We have also added potential applications in the new lines 331--340 (p. 15) to better illustrate
+  the method.
+}
 
@@
 \end{reviewpoint}
-\reply{Active contours without edges \citep{chan_active_2001} is a framework for subpixel segmentation of images in
-  which the contours evolve enclosing homogeneous regions inside.
-Region-based active contours is an alternative terminology.
-As previously described in \autoref{com:1} and \autoref{com:2}, we have made an effort to clearly state that the outputs
-  to our method are a set of nested surfaces and a target multivariate image, that for the case of dMRI data comprehends
-  FA and ADC.
-
-However, and for the sake of clarity, we have modified the manuscript as follows:  {\color{red} say what here}.}
+\reply{%
+In our adaptation of ``active contours without edges'' \citep{chan_active_2001}, the active contours are
+  ``extracted from structural images''.
+The particular sentence has been removed from the abstract, and we expect that the comprehensive reformulation
+  of the descriptive sections of the paper give a successful answer to this issue.
+Regarding the actual inputs of \emph{regseg}, we have introduced modifications in the abstract, and in lines
+  45--50 (p. 2) and lines 243--244 (p. 13).
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+The reviewer is right, we have changed the paragraph about dMRI segmentation (lines 15--27, p. 2)
+  for improved precision.
+We agree with the reviewer that there actually exist many methods to segment fiber tracts.
+We have supported the applications of the method more clearly on lines 11--14 (p. 2).
+Now, it is clearly described that fiber tracks are not available for the task at hand.
+
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+This comment is very related to \autoref{com:14}.
+We have modified the abstract and introduction accordingly, especially lines 15--27 (p. 2).
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+This comment is a consequence of the two previous comments.
+The reviewer is right in the sense that we did not clearly state what we are segmenting, and therefore
+  the references in the former line 24 were a bit orphaned.
+All methods reviewed in lines 15--27 cluster/segment/classify \emph{voxels} in dMRI space based on
+  one or another feature derived from dMRI data (lines 18--20, p. 2).
+We have clarified the manuscript according to \autoref{com:14} and \autoref{com:15} and this comment.
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+Again, the reviewer finds that the main goal of the paper is not clearly stated.
+We understand that modifications derived from the previous comments satisfy the answer to the concern,
+  and changes include abstract and introduction.
+We better support the application of our method (lines 11--14, p. 2), and we have stated more clearly the importance of the set
+  of surfaces being registered to produce a reliable segmentation model of the two diffusion features selected: FA and ADC
+  (please, see \autoref{com:2}).
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+Effectively, the simultaneous registration and segmentation enables \emph{regseg}
+  to overcome the nonlinear warping introduced by EPI distortion.
+We have introduced the following changes:
+\begin{itemize}
+  \item Lines 28--36 (p. 2): Included description and references
+    \citep{jezzard_correction_1995,chiou_simple_2000,kybic_unwarping_2000,studholme_accurate_2000,cordes_geometric_2000} about EPI distortion.
+  \item Lines 36--38 (p. 2): Added the suggested reference \citep{irfanoglu_drbuddi_2015} with other
+    two relevant citations \citep{holland_efficient_2010,andersson_comprehensive_2012}.
+  \item Lines 319--330 (p. 14): we discuss our method as regards EPI correction.
+\end{itemize}
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+The general functionality of \emph{regseg} does not restrict distortion if it is not requested using the
+  feature of anisotropic regularization.
+For the experiments on real data, the ground-truth deformations and the deformation field estimated by \emph{regseg}
+  are restricted to be nonzero only in the phase-encoding direction.
+To make this more clear, we have introduced the following edits:
+\begin{itemize}
+  \item Lines 131--134 (p. 7).
+  \item Lines 198--199 (p. 10).
+\end{itemize}
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+We thank the reviewer for positively weighing our experiments and results.
+As stated in line 211 (p. 10), we use the settings from the \emph{ExploreDTI} software package \citep{leemans_exploredti_2009}, that
+  internally uses \emph{elastix} \citep{klein_elastix_2010} in registration.
+A footnote in line 212 (p. 10) directly links the settings for this method.
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+Additionally to the edits in \autoref{com:10}, we have more clearly defined the concept with the discussion
+  introduced in lines 267--270 (p. 13).
+}
 
@@
 \end{reviewpoint}
-\reply{}
+\reply{
+We have modified the caption of former Figure 3 (now Fig. 4) to describe the meaning of the arrows correctly.
+}
 

### 2016-01-16 Publications/2015-NeuroImage/05-results.tex

@@
 All of the distributions of errors in the tests (four phantom types $\times$ two resolutions) were
-  non-normal with $p<$ 0.001.
-Consequently, we used the non-parametric Wilcoxon signed-rank test with the Bonferroni
+  nonnormal with $p<$ 0.001.
+Consequently, we used the nonparametric Wilcoxon signed-rank test with the Bonferroni
   correction for multiple comparisons ($N$=150, for each phantom type).
@@
   method.
-Summary reports for visual assessment of the 16 cases are included in \citepalias[section S5]{esteban_useful_2015}.
+Summary reports for visual assessment of the 16 cases are included in the
+  \citetalias[section S5]{esteban_useful_2015}.
 In \autoref{fig:regseg-results_real}, box A, the visual report is shown for one subject.
@@
 Finally, to compare the results, we performed Kruskal-Wallis H-tests
-  (a non-parametric alternative to ANOVA) on the warping indices for the three surfaces of
+  (a nonparametric alternative to ANOVA) on the warping indices for the three surfaces of
   interest selected in \autoref{sec:regseg-experiments_evaluation}

### 2016-01-16 Publications/2015-NeuroImage/06-discussion.tex

@@
 The nonlinear registration process evolves driven by the fitness of the
-  picewise-smooth classification of voxels in the target volume imposed
+  piecewise-smooth classification of voxels in the target volume imposed
   by the current mapping of the surfaces.
@@
   but indicated that \regseg{} is a reliable option in this application field.
-Finally, we also contributed proposing a piecewise-smooth segmentation model defined by
+Finally, we also proposed a piecewise-smooth segmentation model defined by
   a selection of nested surfaces to partition the multispectral space

### 2016-01-16 Publications/2015-NeuroImage/cover_letter.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text

\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font
\usepackage{graphicx}
\usepackage[all]{background}

% Margins
\topmargin=-0.5in % Moves the top of the document 1 inch above the default
\textheight=8.5in % Total height of the text on the page before text goes on to the next page, this can be increased in a longer letter
\oddsidemargin=-10pt % Position of the left margin, can be negative or positive if you want more or less room
\textwidth=6.5in % Total width of the text, increase this if the left margin was decreased and vice-versa

%\usepackage{showframe}
\usepackage{tikz}

\newcommand{\MyGraphicLogo}{% For imported graphic logo
\begin{tikzpicture}[remember picture,overlay,yshift=-2cm, xshift=4cm]
  \node at (0,0) {\includegraphics[height=1.80in]{/home/oesteban/Pictures/logocei.png}};
 \end{tikzpicture}
}

\SetBgContents{\MyGraphicLogo}% Select included image
\SetBgPosition{current page.north west}% Select location
\SetBgOpacity{1.0}% Select opacity
\SetBgAngle{0.0}% Select roation of logo
\SetBgScale{1.0}% Select scale factor of logo

\begin{document}

%----------------------------------------------------------------------------------------
%	ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{\textbf{Stephen M. Smith}, \\
Senior Editor, \\
NeuroImage
}

%----------------------------------------------------------------------------------------
%	YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

%\includegraphics[height=1.83in]{../Figures/logocei}

\begin{flushright}
\textbf{Oscar Esteban}, \\ % Your name
Researcher at Biomedical Image Technologies (BIT), \\
Universidad Polit\'ecnica de Madrid \\
Av. Complutense, 30 - ETSI Telecomunicaci\'on C203 \\
E-28040 Madrid (Spain) \\
Tlf.: +34 915 495 700 ext.4234,\\
E-mail: oesteban@die.upm.es % Your address and phone number
\end{flushright}
\vfill

\signature{Oscar Esteban} % Your name for the signature at the bottom

%----------------------------------------------------------------------------------------
%	LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\opening{Dear Editor,}

Please, find enclosed our manuscript, \emph{Surface-driven registration method for the structure-informed segmentation
  of diffusion MR images}, by O. Esteban et al., which we would like to submit for publication in \emph{NeuroImage}.
\textbf{This work is a resubmission of the previously rejected paper, NIMG-15-1151}, as suggested in its
  corresponding decision letter.

For convenience, a point-by-point response for the reviewers of the submission NIMG-15-1151 is attached, and the
  paper is sent in two versions, one annotated for the reviewers and another without annotations.

We confirm that this manuscript has not been published elsewhere and is not under consideration by another journal.
The pre-print of the work has been made available in bioRxiv (doi 10.1101/018945), after checking there
  is no collision with the policies of NeuroImage.

We look forward to hearing from you at your earliest convenience.

\closing{Sincerely yours,}

%\encl{Curriculum vitae, employment form} % List your enclosed documents here, comment this out to get rid of the "encl:"

%----------------------------------------------------------------------------------------
\vfill
\end{letter}

\end{document}

### 2016-01-17 Publications/2015-NeuroImage/responseLetter2.tex
starting revision

--- 2015-11-04 Publications/2015-NeuroImage/responseLetter2.tex
@@
 
-\section*{Response letter to the manuscript no. NIMG-15-1151}
-\medskip
-Original title: \emph{Active contours-driven registration method for the structure-informed segmentation of
-  diffusion MR images}.
-
-\noindent New title: \emph{{\color{blue!70} Surface}-driven registration method for the structure-informed segmentation of
-  diffusion MR images}.
+\section*{Response letter to the manuscript no. NIMG-15-2877}
 
 \bigskip
-\noindent We thank the editors and the reviewers for revising our manuscript no. NIMG-15-1151 submitted to NeuroImage.
+\noindent We thank the editors and the reviewers for revising our manuscript no. NIMG-15-2877 submitted to NeuroImage.
 We appreciate the comments made by both reviewers, and considered them fully to improve our paper.
@@
 \begin{reviewpoint}
-The paper describes an approach for the delineation of selected anatomical structures ([...]) in multi-shell diffusion MR images
-  by registering a segmented T1 weighted reference image with fractional anisotropy and apparent diffusion coefficient maps derived
-  from the diffusion weighted images.
+This paper presents a method for simultaneous registration and segmentation of diffusion images in the presence of geometric distortions.
+The method uses an anisotropically regularised active contour method to coalign tissue segmentations generated from T1 images to distorted multichannel diffusion data.
+It has the advantage that it has no requirements for extra scans to be acquired for correction of the distortion.
+The method demonstrates improved performance over a previous b0T2 registration approach for the alignment of tissue segmentation maps.
+Improvements are demonstrated on simulated data, and real data with simulated distortions.
+The advantage of these results is that the ground truth is known.
 
-The approach corresponds fully to the usual registration of an atlas to the image to be segmented, leaving me rather clueless why do
-  they call the method ``simultaneous registration and segmentation''.
+I am satisfied that the active contour alignment approach used to align the surfaces is sound, and that the results show theoretic improvements over the T2B method.
+I am convinced that the method approach is effective for the alignment task.
+In general the methods are well explained with the exception of a few minor points (below).
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+My main concerns lie with whether the results in the paper prove what the authors set out in their abstract.
+Specifically, a stated goal is for the use of anatomically constrained tractography.
+However, there are no results that show this method would improve connectome extraction.
+In particular, the authors state that ``this method avoids any requirement for collecting additional scans nor resampling for diffusion unwarping''.
+I see this has advantages, but what about the consequences of not unwarping on the diffusion tracts themselves see [1][2][3].
+In my opinion the authors should address this point possibly with additional results to show how they intend this to be used within a tracking framework.
+
+[1] Jones, Derek K., and Mara Cercignani. ``Twentyfive pitfalls in the analysis of diffusion MRI data.'' NMR in Biomedicine 23.7 (2010): 803820.
+
+[2] Lee, J., et al. ``Correction of Bo EPI distortions in diffusion tensor imaging and white matter tractography.'' Proc. Int. Soc. Magn. Reson. Med. Vol. 11. 2004.
+
+[3] Andersson, J. L., et al. ``Effects of susceptibility distortions on tractography.''Proceedings of International 
+Society of Magnetic Resonance in Medicine. Vol. 11. 2004.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+I also have some concern that the method is not applied to real distortions nor compared to stateoftheart correction methods such as EDDY [4].
+This was used for the HCP corrections therefore it might provide a good basis comparison basis on real rather than modeled distortions for brain data.
+If this could not be done for some reason I feel it should be discussed explicitly.
+It should be possible to compare to bbregister for the unwarped data.
+Why is this not done?
+
+[4] A comprehensive Gaussian Process framework for correcting distortions and movements in diffusion images Jesper L. R. Andersson et al ISMRM
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+Pg 5 regularisation: `` requires that the distortion and its gradient have zero mean and variance governed by the matrices A and B'' suggest adding a comma here for easier interpretation i.e.
+``requires that the distortion and its gradient have zero mean, and variance governed by the matrices A and B''
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+Pg 7, line 4 `` anisotropy is aligned''. What is meant by anisotropy in this context, the anisotropy of the deformation gradient.
+\end{reviewpoint}
+\reply{%
+The reviewer is totally right. We have modified the text accordingly.
+This comment is also made by the reviewer \#3 in the \autoref{com:31}.
+We refer the reader to this comment for the listing of actual changes in the paper.
+}
+
+\begin{reviewpoint}
+Figure 2.
+This looks nice, but its not easy to interpret.
+I can't see a clear improvement through AC.
+I think there needs to be direct comparisons between the conditions represented by AC on one plot.
+Why not have 6 plots for each tissue class comparing distributions for each condition.
+\end{reviewpoint}
+\reply{%
+}
+% ------------------------------------------------------------------- Reviewer 3
+\bigskip
+\bigskip
+\subsection*{Reviewer \#3:}
+\begin{reviewpoint}
+This paper presents a new method for correction of EPI distortion in diffusion MRI data by performing simultaneous registration and segmentation of a set of nested surfaces, using a multichannel FA/MD image with a registration target of an undistorted T1 image.
+The primary advantage over other methods is that cortical surfaces and whitegray matter boundaries are apparently aligned well by this method.
+Results show that it outperforms one implementation of the more standard baseline to T2 registration.
+Overall this is a quite interesting work, and it will be a good addition to the literature and also to the body of opensource code for diffusion MRI.
+
+I have some comments for clarification of the paper.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+Overall, I am slightly confused by some aspects of the equations, and I think that some information from the supplement (which I have not had time to read) should be moved into the paper to clarify missing points.
+The paper should be able to be read independently of the supplement, which does not seem to be the case currently.
+Specifically, the relationship of the bspline registration framework to the evolution of the active contours is a bit confusing, with a main equation given inside Figure 1, and apparently all other detail in the supplement.
+This could be as simple as an additional paragraph or two.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+Please state that the goal of the paper is to improve connectomebased diffusion MRI research.
+The processing description on pages 12 assumes that the connectomebased approach is the only approach to analysis of diffusion MRI.
+But this is clearly not the case (surgical planning, TBSS, etc).
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+``to current models such as (Dadicci et al)'' is confusing. Please give the name(s) of the current model(s).
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+On page 3, please define N (section 2.1)
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+In Figure 1, it is a bit confusing that active contours appear, because the abstract did not mention active contours directly, and the evolution following the inward normals at each vertex is not explicit in the equations of the paper.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+It seems the evolution at $s_1$ is in the direction opposite the inward normal, so the caption of Figure 1 is a bit confusing, since it directly states the inward normals are followed, rather than just the normal to the curve.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+I recommend adding a little bit of text to unify figure 1 and the surface evolution framework with the concepts presented in the registration equations section.
+I think this information may be in the supplement.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+On page 5, equation (4), I think f should be f prime on the first line.
+\end{reviewpoint}
+\reply{%
+The reviewer is right. We have fixed the manuscript accordingly.
+}
+
+\begin{reviewpoint}
+Please describe A and B in a sentence after equation 5. It is not clear how these were chosen (or at least refer to the appendix).
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+On page 5, equation (7), the log is removed from the second line.
+I would expect a sum of log probabilities, with the log preserved, rather than removing the log altogether.
+It seems this is an oversight.
+Or do the authors mean to use a proportional symbol instead of an equals symbol?
+Please clarify.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+In equation (7), the equation that is split onto two lines (why?) has two plus symbols in a row, one on each line.
+In equation (7), the second integral is over omega, while the first is over $omega_l$.
+Why?
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+In equation (7) please clarify with brackets what the sum over l is applied to.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+The two pluses are present again in equation (10).
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+Please explain in a sentence what alpha and beta are in equation (10) (expected variances along each axis of the coordinate system, or something).
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+$g_k$ is completely undefined in equation (11). I see there is information in a supplement but at a minimum each variable should be defined in a sentence in the main paper.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+pg 7. Please define PE acronym or write out phase encoding.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+eq (14). Please define what is $a_i$.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+Figure 5. What are the little numbers (26, 30, 33) in the lower left corner of the phantom images? Figure 5. Please state clearly in the caption that this is digital phantom data, to help the reader.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+Figure 5.
+Over what is the average error of surfaces computed?
+Average over the surface?
+Then the variability in the plot is across experiments?
+Or the other way around with the average at each surface point across experiments??
+It is not clear.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+Figure 5 A. Why are there so many phantom images shown? They all look the same to me. Please explain and/or remove the duplicates.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+Figure 6. Are the plots in B from one or many subjects? Please state in the caption.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+In Table 2 caption how many subjects were input to this experiment? This is given as 16 in the abstract, but that info is not very near the actual results and would increase their impact.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+pg 13.
+``We also show the 95\% CIs of the sWI for these surfaces.'' sounds like it is referring to Figure 6.
+However it is apparently in Table 2, which should be referenced in this sentence.
+Or the sentences about Table 2 should be reorganized to sound separate from Figure 6.
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+in Appendix (and also elsewhere): ``the anisotropy is aligned with the imaging axes'' is quite confusing.
+Usually anisotropy refers to FA or similar.
+I believe this is the anisotropy of the covariance matrices, really that their eigenvectors are assumed aligned with the voxel coordinate system.
+Please clarify the usage of ``anisotropy'' in the registration framework.
+\end{reviewpoint}
+\reply{%
+The reviewer is totally right. We have modified the text accordingly.
+This comment is also made by the reviewer \#1 in the \autoref{com:5}.
+}
+
+\begin{reviewpoint}
+What values are used for alpha and beta in the appendix?
+\end{reviewpoint}
+\reply{%
+}
+
+\begin{reviewpoint}
+minor grammatical issues:
+
+abstract: ``approached either segmenting'' > ``approached by either segmenting''
+
+pg 2: ``threshold set'' sounds like a mathematical set > ``threshold that is chosen''
+
+pg 5. ``displacements field'' > ``displacement field''
+
+pg. 15 ``monitorization'' is not a word > monitoring
 
@@
 \reply{%
-The description of the method given by both reviewers (see \autoref{com:10}, Reviewer \#2) lead us to think that the paper
-  did not properly describe the actual inputs and outputs the proposed methodology.
-In general, we have almost-completely reformulated the abstract, introduction and discussion to define this
-  point clearly.
-Particularly, the following sections have been introduced to give an answer to this particular comment:
-\begin{itemize}
-\item Lines 45--53 (p. 2).
-\item Lines 244--246 (p. 13), here we introduce the previously missing discussion to call the
-  method ``simultaneous registration and segmentation''.
-\item Lines 250--252 (p. 13), we comment here on the similarity of \emph{regseg} to the atlas-based segmentation
-  using joint registration and segmentation methods.
-\end{itemize}
-}
-
-\begin{reviewpoint}
-The paper describes an approach [...] (quite some are listed as possible examples but results are only reported for the white matter,
-  deep gray matter and the pial surface) in multi-shell [...].
-\end{reviewpoint}
-\reply{
-To clarify this point, we have stated more clearly that the choice of surfaces to segment the target data is
-  crucial (lines 46--47, p.2).
-We have better defined the segmentation model that drives the registration process (lines 94--95, p. 5).
-We have also improved the former Figure 4 (now it is Figure 2, page 5), and brought it earlier in the paper to
-  illustrate earlier the segmentation model.
-Since the former visualization of the conditional (region-wise) joint distribution of FA and ADC was a bit clumsy,
-  we have extracted the conditional distributions to the right side of each panel.
-We expect these modifications make clear the difference between the surfaces involved in driving the registration
-  process and the surfaces in which we evaluate the error of the method (lines 217--218, p. 10) are a subset of the
-  former.
-}
-
-\begin{reviewpoint}
-The proposed algorithm is rather standard, relying on a region-based elastic contour deformation approach.
-The cost function is derived in the usual Bayesian framework, relying on a multivariate normal distribution prior for the image
-  intensities and a Tikhonov regularizer enforcing a predefined normal distribution over the distortions along the imaging axes.
-In this respect the novelty of the paper can be questioned, while the application context and the very promising results may still
-  justify a publication in NeuroImage.
-\end{reviewpoint}
-\reply{
-We agree with the reviewer on that the novelty of the methodological elements of our approach is limited.
-However, to our knowledge there exists no alternative method combining the same elements to segment
-  multivariate data.
-We have moved the related methods and our methodological contributions from the Introduction to a more appropriate
-  place in the Discussion (lines 248--275), additionally satisfying \autoref{com:9} (Reviewer \#2) suggesting to include
-  in the introduction only the references related to the focus of the paper.
-The particular changes can be listed as follows:
-\begin{itemize}
-  \item Lines 252--260 (p. 13), we describe the methodological differences w.r.t. \cite{gorthi_active_2011},
-    one of the latest atlas-based segmentation methods using simultaneous registration and segmentation.
-  \item Lines 261--270 (p. 13), we compare the methodological features of our method and \emph{bbregister},
-    an antecedent and standardized tool used on the same application.
-  \item Lines 271--273 (p. 13), we introduce the methodological differences with a recent alternative method in
-    terms of regularization.
-  \item Lines 273--278 (p. 13) mention some other minor methodological contributions.
-\end{itemize}
-}
-
-\begin{reviewpoint}
-However, in such a case the paper should be thoroughly revised to address the following major issues:
-1. The assumption of independence between pixels is quite usual but obviously wrong.
-The authors may want to comment on the consequences.
-\end{reviewpoint}
-\reply{
-The reviewer is right, assuming that pixels are i.i.d. is wrong but a widely accepted simplification.
-We have supported the assumptions inserting the new lines 88--90 (p. 3).}
-
-\begin{reviewpoint}
-2. It is unclear, under which conditions enforcing the optimality condition guarantees that the region identified corresponds
-  to the targeted one (it probably strongly depends on its intensity properties and those of neighbouring structures).
-A corresponding analysis of the goal function and the presence of (undesired) local minima is missing.
-Figure S1 is not particularly helpful in this respect.
-Some information about the initialization of the distortion map should also be given.
-\end{reviewpoint}
-\reply{
-We have improved former Figure 4 (now, it is Figure 2, page 5) to clarify that the segmentation model is inherently defined by
-  the choice of surfaces and their nesting.
-This is an actual contribution of the work, as stated in lines 304--307 (p. 14).
-A comment on the convergence issues follows in lines 307--311 (p. 14).
-Finally, we described the initialization of the surfaces with the insertion of lines 196--198 (p. 10).
-}
-
-\begin{reviewpoint}
-3. I miss a reasonable description of the parameter settings necessary for generating the results presented.
-It is not even clear if all experiments have been performed by the same parameter settings.
-A reasonable sensitivity analysis is also missing.
-\end{reviewpoint}
-\reply{
-We have made a strong effort in providing reproducible experiments using openly available data.
-Therefore, parameter settings are publicly available in GitHub.
-However, we understand that a better description of the parameter settings should be included within
-  the manuscript and modified it accordingly, in lines 200--204 (p. 10).
-}
-
-\begin{reviewpoint}
-4. I could not understand how the ground truth can be generated for the real datasets.
-As far as I can see, in these cases there is a (probably unknown) distortion between the T1 weighted and the
-  diffusion weighted images.
-How an emulated distortion can get around this fact?
-\end{reviewpoint}
-\reply{
-We have clarified this point with the following edits:
-\begin{itemize}
-  \item Lines 171--173 (p. 8): we indicate that we use the ``minimally preprocessed'' images of the
-    HCP database, and thus, these images are acquired with specialized protocols and are
-    already corrected for the distortion.
-  \item Figure 4 (former fig. 3), page 9, has been modified to clearly indicate this feature.
-\end{itemize}
-}
-
-\begin{reviewpoint}
-5. The vertex-wise error distributions are not only skewed, but quite clearly structured im many cases
-  (especially for the phantom images).
-What is the underlying reason?
-\end{reviewpoint}
-\reply{
-We have added a discussion on this in the corresponding section, lines 287-293 (p. 14).
-Thanks for pointing this systematic error out.
-}
-
-\bigskip
-\bigskip
-\subsection*{Reviewer \#2:}
-\begin{reviewpoint}
-This paper seems to present a method that can perform wM-GM segmentation and simultaneously estimate a deformation field to
-  correct for EPI distortion.
-Apparently, the idea is to use nested WM-GM surfaces, as well as other surfaces, to provide better cortical registration than
-  a competing method, which is not clearly described.
-The overall goal and motivation of the paper are not clearly explained, and almost all related references for EPI distortion
-  correction are missing.
-It is unclear if the deformation field solution is restricted to the phase encode direction.
-It is possible that this is a very useful method, as the results seem to indicate, but the paper is unfortunately too unclear to
-  make a full judgement about the method.
-\end{reviewpoint}
-\reply{
-We thank the reviewer for pointing out that the overall goal is not clearly explained.
-We have rewritten the abstract, introduction and discussion sections almost fully.
-Particular comments included within this general comment are addressed later: \autoref{com:18}
-  as regards EPI distortions, and \autoref{com:20} for the phase-encoding direction point.
-In order to better focus the application field, we have introduced lines 11--14 (p. 2) and lines
-  312--317 (p. 14).
-We have also added potential applications in the new lines 331--340 (p. 15) to better illustrate
-  the method.
-}
-
-\begin{reviewpoint}
-abstract: Not immediately clear what this means: ``active contours without edges extracted from structural images''
-  (Where do the edges come from? What modalities are actually input to regseg?)
-\end{reviewpoint}
-\reply{%
-In our adaptation of ``active contours without edges'' \citep{chan_active_2001}, the active contours are
-  ``extracted from structural images''.
-The particular sentence has been removed from the abstract, and we expect that the comprehensive reformulation
-  of the descriptive sections of the paper give a successful answer to this issue.
-Regarding the actual inputs of \emph{regseg}, we have introduced modifications in the abstract, in lines 45--50 (p. 2), and
-  in lines 243--243 (p. 13).
-}
-
-\begin{reviewpoint}
-I note that the text in the abstract seems to have no correspondence with the graphical abstract.
-The text abstract mentions nothing about diffusion MRI, FA, ADC, or phase maps, so it is unclear how these may be used in the method.
-These images figure prominently in the graphical abstract, which may be a better overview than the text abstract.
-\end{reviewpoint}
-\reply{We agree with the reviewer. We have proposed new graphical and text abstracts.}
-
-\begin{reviewpoint}
-Intro line 6-7. I am sure you mean to say that the resolution of dMRI is much LOWER than the images microstructure (not higher).
-  larger voxels == lower resolution
-\end{reviewpoint}
-\reply{Absolutely, we have fixed the manuscript accordingly.}
-
-\begin{reviewpoint}
-Intro 9. Orientations is the usual term for diffusion weighting, rather than ``angles''.
-\end{reviewpoint}
-\reply{We agree and have updated the manuscript accordingly.}
-
-\begin{reviewpoint}
-``These limitations prevent segmentation in the native dMRI space''  I disagree.
-In fact, many methods have been used to segment fiber tracts, then to quantitatively measure them,
-  in the dMRI space.
-It is true the correspondence with T1 is imperfect, but segmentation of fiber tracts is possible.
-\end{reviewpoint}
-\reply{
-The reviewer is right, we have changed the paragraph about dMRI segmentation (lines 15--27, page 2)
-  for improved precision.
-We agree with the reviewer that there actually exist many methods to segment fiber tracts.
-We have supported the applications of the method more clearly on lines 11--14 (p. 2).
-Now, it is clearly described that fiber tracks are not available for the task at hand.
-
-}
-
-\begin{reviewpoint}
-Please define what you mean here by segmentation.
-It seems there is an assumption that it means specifically segmentation via registration, where the
-  structures are defined using T1 images, but that is not stated anywhere.
-
-Now on line 19 it seems the goal of the segmentation may simply be the segmentation of the WM.
-So far, I have no idea what the overall goal of the approach is.
-\end{reviewpoint}
-\reply{
-This comment is very related to \autoref{com:14}.
-We have modified the abstract and introduction accordingly, especially lines 15--27 (p. 2).
-}
-
-\begin{reviewpoint}
-Line 24. clustering of what?
-\end{reviewpoint}
-\reply{
-This comment is a consequence of the two previous comments.
-The reviewer is right in the sense that we did not clearly state what we are segmenting, and therefore
-  the references in line 24 are a bit orphaned.
-All methods reviewed in lines 15--27 cluster/segment/classify \emph{voxels} in dMRI space based on
-  one or another feature derived from dMRI data (lines 18--20, p. 2).
-We have clarified the manuscript according to \autoref{com:14} and \autoref{com:15} and this comment.
-}
-
-\begin{reviewpoint}
-Now in the paragraph beginning on line 29, it seems the segmentation of WM/GM interface is of interest.
-Still, I have no idea if this is the goal of the approach, or why.
-
-Line 39. I note that there are many approaches, in addition to nonlinear T2->b0, that have been proposed
-  for correction of EPI distortions in dMRI.
-If that is the goal of this approach, the entire intro should be rewritten focusing on this related work.
-However, I do not know yet what the goal is.
-
-The entire Introduction should be rewritten to improve organization and clearly state the paper's goal much
-  earlier (in abstract and introduction).
-Truly related work should be described in more detail, and all of the unrelated references should be removed.
-
-Line 58. It seems the paper goal may be stated here. It is still completely unclear what type of segmentation should be achieved.
-\end{reviewpoint}
-\reply{
-Again, the reviewer finds that the main goal of the paper is not clearly stated.
-We understand that modifications derived from the previous comments satisfy the answer to the concern,
-  and changes include abstract and introduction.
-We better support the application of our method (lines 11--14, p. 2), and we have stated more clearly the importance of the set
-  of surfaces being registered to produce a reliable segmentation model of the two diffusion features selected: FA and ADC
-  (please, see \autoref{com:2}).
-}
-
-\begin{reviewpoint}
-It seems by line 69 that EPI distortion correction and segmentation by registration are to be performed jointly
-  in some fashion.
-There are many, many references for distortion correction that are missing from this introduction and that must
-  be added.
-See for example multiple papers by Carlo Pierpaoli's group, including this most recent:
-
-Irfanoglu, M. Okan, et al. ``DR-BUDDI (Diffeomorphic Registration for Blip-Up blip-Down Diffusion Imaging) method for correcting echo planar imaging distortions.''
-  NeuroImage 106 (2015): 284-299.
-\end{reviewpoint}
-\reply{
-Effectively, the simultaneous registration and segmentation enables \emph{regseg}
-  to overcome the nonlinear warping introduced by EPI distortion.
-We have introduced the following changes:
-\begin{itemize}
-  \item Lines 28--36 (p. 2): Included description and references
-    \citep{jezzard_correction_1995,chiou_simple_2000,kybic_unwarping_2000,studholme_accurate_2000,cordes_geometric_2000} about EPI distortion.
-  \item Lines 36--38 (p. 2): Added the suggested reference \citep{irfanoglu_drbuddi_2015} with other
-    two relevant citations \citep{holland_efficient_2010,andersson_comprehensive_2012}.
-  \item Lines 319--330 (p. 14): we discuss our method as regards EPI correction.
-\end{itemize}
-}
-
-\begin{reviewpoint}
-The majority of the EPI distortion correction methods use T2 and B0 due to similar contrast. Why was T1 chosen in this framework?
-
-Suddenly in Section 2 it becomes clear that there are surfaces being registered.
-Perhaps this should have been obvious from the title, which mentioned active contours, but before line 72 it was not clearly
-  stated that the method was posed as a surface registration problem.
-\end{reviewpoint}
-\reply{
-We have updated the title to represent this aspect.
-We have also included all previous changes to fulfill this request.
-}
-
-\begin{reviewpoint}
-Was the warp restricted to the phase encode direction?
-If so, the r in eq (1) should be restricted to this direction.
-I do not see this mentioned anywhere near the equations in the Methods section.
-If the solution found for the deformation field was not restricted to the phase encode direction,
-  then this is not a reasonable way to perform EPI distortion correction. Please clarify.
-
-[...]
-
-I still see nothing in the results indicating that the estimated distortions were restricted to the phase encode direction.
-If not, the added degrees of freedom might result in improved surface segmentation, but the distortion field would not correspond at
-  all to the physical problem under study.
-Furthermore, the dMRI gradients would have to be re-oriented independently at each voxel, and this is not currently possible in any
-  software I know of.
-\end{reviewpoint}
-\reply{
-The general functionality of \emph{regseg} does not restrict distortion if it is not requested using the
-  feature of anisotropic regularization.
-For the experiments on real data, the ground-truth deformations and the deformation field estimated by \emph{regseg}
-  are restricted to be nonzero only in the phase-encoding direction.
-To make this more clear, we have introduced the following edits:
-\begin{itemize}
-  \item Lines 131--134 (p. 7).
-  \item Lines 198--199 (p. 10).
-\end{itemize}
-}
-
-\begin{reviewpoint}
-The experiments seem reasonable and clearly presented in general. But far too little detail is given on the B0-T2 comparison method.
-\end{reviewpoint}
-\reply{
-We thank the reviewer for positively weighing our experiments and results.
-As stated in line 211 (p. 10), we use the settings from the \emph{ExploreDTI} software package \citep{leemans_exploredti_2009}, that
-  internally uses \emph{elastix} \citep{klein_elastix_2010} in registration.
-A footnote in line 213 (p. 10) directly links the settings for this method.
-}
-
-\begin{reviewpoint}
-Please define the phrase ``active contours without edges''.
-\end{reviewpoint}
-\reply{
-Additionally to the edits in \autoref{com:10}, we have more clearly defined the concept with the discussion
-  introduced in lines 267--268 (p. 13).
-}
-
-\begin{reviewpoint}
-Figure 3: What are the arrows in images 5? DEformations? Indications of important regions?
-\end{reviewpoint}
-\reply{
-We have modified the caption of former Figure 3 (now Fig. 4) to describe the meaning of the arrows correctly.
+All the issues have been corrected following the suggestions of the reviewer.
 }

### 2016-01-17 Publications/2015-NeuroImage/02-abstract.tex
starting revision

@@
   require precise delineations of anatomical structures.
-This requirement has been approached either segmenting the data in native \gls*{dmri} space
-  or mapping the structural information from \gls*{t1} images.
+This requirement has been approached \revcomment[R\#3-C.33]{by} either segmenting the data in
+  native \gls*{dmri} space or mapping the structural information from \gls*{t1} images.
 The characteristic features of diffusion data in terms of \acrlong*{snr}, resolution, as well

### 2016-01-17 Publications/2015-NeuroImage/04-methods.tex
starting revision

@@
 
+\revcomment[R\#3-C.15]{%
   \begin{align}
   P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
-  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} )
+  \mathcal{N} (\vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} )
   \ifthenelse{\boolean{review}}{=}{= \notag\\ &=}
@@
   \label{eq:regseg-pdf}
-  \end{align}
-  using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
+  \end{align}%
+  }%
+  using $\mdist{f'}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}'$ with respect
   to the descriptors of region $l$ as
-  $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
+  $\mdist{f'}{l} = (\vec{f}' - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f}' - \boldsymbol{\mu}_l)$.
 $C$ is the number of channels comprised in the image $M$.
@@
 \paragraph*{Regularization}
-The smoothness of the resulting displacements field is induced by a Thikonov regularization
-  prior:
+The smoothness of the resulting \revcomment[R\#3-C.33]{displacement field} is induced
+  by a Thikonov regularization prior:
 

### 2016-01-17 Publications/2015-NeuroImage/06-discussion.tex
starting revision

@@
   of the brain using multispectral images, for instance in the case of traumatic brain
-  injury \citep{irimia_structural_2014} or in the monitorization of brain tumors
+  injury \citep{irimia_structural_2014} or in \revcomment[R\#3-C.33]{monitoring} brain tumors
   \citep{weizman_semiautomatic_2014}.

### 2016-01-17 Publications/2015-NeuroImage/responseLetter2.tex
add reply

@@
 \reply{%
+The equation (7) was splited in two lines because in just one line it exceeded the paper margins, and we
+  considered the equal sign the most convenient point to break it.
+Unfortunatelly, we could not interpret what the reviewer means with ``has two plus symbols in a row,
+  one on each line''.
+Regarding the last part of the question, the second integral is defined directly along the whole image
+  domain $\Omega$ because it does not depend on the corresponding partition $\{\Omega_l\}$.
+This integral corresponds to the regularization term.
+The first integral is computed region-wise, since the energy functional has different
+  parameters $\Theta_l$ for each of the regions $\Omega_l$.
 }

### 2016-01-17 Publications/2015-NeuroImage/responseLetter2.tex
revising (rv1)

@@
 \reply{%
+We apologize since the paper could be misleading in its former version, driving
+  the reader to identify its goal restricted to support anatomically constrained
+  tractography.
+We have modified certain points in the abstract, introduction and discussion to
+  clarify that this method is intended to support any diffusion processing task that
+  requires a precise delineation of the anatomy in the diffusion space.
+
+In the field of connectivity analysis, this need has posed important problems in the
+  standard methodologies to build the connectivity network\footnote{generally imposing
+  a parcellation mapped from T1-space}.
+With the emergence of novel reconstruction and tractography methods (see lines ),
+  these tasks may also be improved with the application of more accurate registration
+  methods like \emph{regseg}.
+
+Regarding the given references, [1] is structured in a list of common pitfalls in the
+  analysis of dMRI.
+Particularly, pitfalls \#1, \#2, and \#14 apply to our work.
+Pitfall \#1 points out the problem of modulation in the unwrapped signal.
+Since we promote analysis in the original space, this problem does not apply in practice.
+Pitfall \#2 particularly points out the problem of misplaced tracks (and cites
+  the other two references [2, 3]).
+As we have included in the discussion (see line XXX), the tracks may be unwarped using
+  the deformation found with \emph{regseg} since it is revertible.
+Checking the correctness of tracks computed on unwarped dMRI data vs. tracks computed
+  on the original dMRI data and then unwarped with \emph{regseg} or whatever other
+  technique is out of the scope of the paper, but a very interesting lead for
+  future research.
+Finally, pitfall \#14 has side relationships with our method, since \emph{regseg} is
+  definitely intended to increase the accuracy of the ROIs imposed on the dMRI data.
+In this case, Jones et al. blame it on the misregistration of the \emph{b0} and the
+  diffusion weighted volumes, but the problem applies equally when the ROIs are
+  not correctly aligned with the \emph{b0} itself.
 }
@@
 \begin{reviewpoint}
-I also have some concern that the method is not applied to real distortions nor compared to stateoftheart correction methods such as EDDY [4].
+I also have some concern that the method is not applied to real distortions nor compared to state of the art correction methods such as EDDY [4].
 This was used for the HCP corrections therefore it might provide a good basis comparison basis on real rather than modeled distortions for brain data.
@@
 \reply{%
+As mentioned in \autoref{com:1}, the method is actually applied on real distortions.
+
+Regarding the proposed (and also some other possible) comparisons:
+\begin{itemize}
+\item \emph{Eddy} [4] is a software tool in FSL that simultaneously corrects for
+    head motion artifacts and Eddy currents -derived distortions.
+  Regarding susceptibility-derived distortions, it accepts the output of the tool 
+    \emph{topup} to reduce the number of resamplings to just one, but it does not
+    actually perform any correction for this artifact.
+  The missing comparison to \emph{topup} follows.
+\item \emph{topup} is effectively the tool used to correct diffusion images for the
+  susceptibility artifact in the HCP pipeline.
+  This tool uses one image with a certain phase-encoding scheme and a second image
+    (typically only one \emph{b0} volume) with different phase-encoding scheme (typically
+    reversed encoding blips in the same encoding axis).
+  As it is shown in Fig. 4, the \emph{b0} warped with reversed encoding blips is not
+    computed, and thus this comparison is not possible.
+  Adding this experiment would include a significant overhead that may displace the focus
+    of the paper from ``an accurate nonlinear registration method'' to ``yet another 
+    correction method for susceptibility-derived distortions''.
+  For these two reasons, we disregarded the implementation of such experiment.
+\item \emph{fieldmap-based correction strategies}, since we use the theoretical basis of these
+  methods to generate our realistic distortions, comparison to these methods is not specially
+  interesting.
+\item \emph{bbregister} is an affine registration method intended to be used with diffusion
+  data.
+  However, the ``minimally preprocessed'' datasets from the HCP that we use here are already
+  spatially normalized.
+  \emph{bbregister} is designed to find the normalization, and thus it is not applicable
+  here.
+\end{itemize}
+For all these reasons, we understood that a fair and interesting comparison would include
+  another registration method.
+The closest nonlinear registration method that is widely used in this framework happens
+  to be the \emph{b0}-to-T2w registration, as included in the experiments of the paper.
 }

### 2016-01-17 Publications/2015-NeuroImage/06-discussion.tex
revising (rv1)

@@
 This situation is found commonly in historical datasets.
-Moreover, since the structural information is projected into the native space of \gls*{dmri},
-  the whole connectome extraction process can be performed on the unaltered \gls*{dmri}
-  signal (i.e. without resampling data to an undistorted space).
+\revcomment[R\#1-C.2]{%
+We envision \regseg{} to be integrated in diffusion processing pipelines, after a preliminary
+  \gls*{dti} computation and before anatomically-informed reconstruction and tractography
+  methods.
+Since the structural information is projected into the native space of \gls*{dmri},
+  these two processes and the matrix building task can be performed on the unaltered
+  \gls*{dmri} signal (i.e. without resampling data to an undistorted space).
+For analyses other than connectivity, like \gls*{tbss}, the deformation estimated by \regseg{}
+  can be used to map the tracts into reference space.}
 

### 2016-01-20 Publications/2015-NeuroImage/responseLetter2.tex
some new answers

@@
 \reply{%
+We thank the reviewer for the time dedicated in his/her review and these positive comments.
+We are pleased that this summary of the paper matches the idea we want to transmit.
+
+We have modified the paper to clarify the nature of the distortions we use to generate the
+  ground truth (see \autoref{com:3}).
 }
@@
 \reply{%
-As mentioned in \autoref{com:1}, the method is actually applied on real distortions.
+We have modified the paper to clearly reflect that the method is actually applied
+  on realistic distortions (comes from \autoref{com:1}).
+For the phantoms, as the reviewer mentions, the distortion maps are synthetic and
+  generated with random coefficients.
+For the real datasets, however, we use realistic distortions derived from the real fieldmap
+  corresponding to each dataset.
+The fieldmaps are processed to remove the noise present in the air cavities and the background,
+  and are applied low-pass filtering to remove abrupt gradients in the resulting deformation
+  field.
 

### 2016-01-20 Publications/2015-NeuroImage/02-abstract.tex
some new answers

@@
   \gls*{dmri} dataset.
-We first verify the accuracy of \regseg{} on a general context using digital phantoms.
+We first verify the accuracy of \regseg{} on a general context using digital phantoms 
+  \revcomment[R\#1-C1]{distorted with synthetic and random deformations}.
 Then we establish an evaluation framework using undistorted \gls*{dmri} data from the \gls*{hcp}
-  and known deformations derived from real inhomogeneity fieldmaps.
+  and \revcomment[R\#1-C1]{realistic} deformations derived from 
+  \revcomment[R\#1-C1]{the inhomogeneity fieldmap corresponding to each subject}.
 We analyze the performance of \regseg{} computing the misregistration error of the surfaces estimated
@@
   method, thereby obtaining a significantly lower misregistration error with \regseg{}.
-Therefore, we demonstrate that \regseg{} allows the accurate mapping of structural information
-  in \gls*{dmri} space, enabling the application of new structure-informed techniques in
-  the connectome extraction.
+The accurate mapping of structural information in \gls*{dmri} space 
+  \revcomment[R\#1-C2]{is fundamental to increase the reliability of network building
+  in connectivity analyses, and to improve the performance the emerging structure-informed
+  techniques for \gls*{dmri} data processing}.

### 2016-01-24 Publications/2015-NeuroImage/responseLetter2.tex
revised paper

@@
 
-\usepackage[numbers,sort&compress]{natbib}
+\usepackage[sort&compress]{natbib}
 \usepackage{xcolor}
 \usepackage[T1]{fontenc}
+\usepackage[utf8]{inputenc}
 \usepackage{charter}
@@
 \usepackage[colorlinks=true]{hyperref}
+\usepackage{doi}
 \usepackage{nameref}
@@
   along with the description of the corresponding changes in the manuscript.
-Text modifications in the manuscript are highlighted with a new font color, and attached notes indicate which reviewer comment
-  the corresponding change is addressing to.
+Text modifications in the manuscript are highlighted with a new font color, and attached notes indicate 
+  which reviewer and comment the corresponding change is addressing to.
 
@@
 It has the advantage that it has no requirements for extra scans to be acquired for correction of the distortion.
-The method demonstrates improved performance over a previous b0T2 registration approach for the alignment of tissue segmentation maps.
+The method demonstrates improved performance over a previous b0--T2 registration approach for the alignment of tissue segmentation maps.
 Improvements are demonstrated on simulated data, and real data with simulated distortions.
@@
 
-We have modified the paper to clarify the nature of the distortions we use to generate the
-  ground truth (see \autoref{com:3}).
+We would like to point out that the distortions applied to the real datasets are derived from
+  real fieldmaps, and therefore they are not ``modeled'' in a strict sense.
+We have modified the paper to clarify the nature of these distortions (see \autoref{com:3}) in
+  both experiments (simulated and real datasets).
 }
@@
   requires a precise delineation of the anatomy in the diffusion space.
+We have also looked for the phrase cited by the reviewer\footnote{``this method avoids any
+ requirement for collecting additional scans nor resampling for diffusion unwarping''}
+ to reformulate it, however we did not found such exact statement in the paper.
 
@@
 
-Regarding the given references, [1] is structured in a list of common pitfalls in the
+The given references are relevant to the domain of diffusion image processing,
+  specially that provided as [1], and propose the following discussion that
+  is very interesting but goes beyond the objectives of this paper.
+\cite{jones_twentyfive_2010} present a list of common pitfalls in the
   analysis of dMRI.
-Particularly, pitfalls \#1, \#2, and \#14 apply to our work.
-Pitfall \#1 points out the problem of modulation in the unwrapped signal.
-Since we promote analysis in the original space, this problem does not apply in practice.
-Pitfall \#2 particularly points out the problem of misplaced tracks (and cites
-  the other two references [2, 3]).
-As we have included in the discussion (see line XXX), the tracks may be unwarped using
-  the deformation found with \emph{regseg} since it is revertible.
-Checking the correctness of tracks computed on unwarped dMRI data vs. tracks computed
+Particularly, pitfalls \#1, \#2, and \#14 apply to our work:
+\begin{itemize}
+\item Pitfall \#1 points to the problem of modulation in the unwrapped signal and the
+  appearance of unrealistic structures due to interpolation.
+Since we promote to perform the analysis in the original space, this problem does not 
+  apply in practice.
+\item Closely related, pitfall \#2 raises the problem of misplaced tracks (and cites
+  the other two references given [2, 3]).
+We have included in the discussion (see line XXX) some text indicating that the tracks
+  can be unwarped using the deformation found with \emph{regseg} since it is revertible.
+However, checking the correctness of tracks computed on unwarped dMRI data vs. tracks computed
   on the original dMRI data and then unwarped with \emph{regseg} or whatever other
@@
   future research.
-Finally, pitfall \#14 has side relationships with our method, since \emph{regseg} is
+As a matter of fact, we have recently published a whole-brain digital phantom of dMRI 
+  data \citep{esteban_diffantom_2016} to perform evaluations like the one suggested 
+  here \citep[for instance, please see][Chapter 5]{esteban_image_2015}.
+\item Finally, pitfall \#14 has side relationships with our method, since \emph{regseg} is
   definitely intended to increase the accuracy of the ROIs imposed on the dMRI data.
@@
   not correctly aligned with the \emph{b0} itself.
+\end{itemize}
 }
@@
 The fieldmaps are processed to remove the noise present in the air cavities and the background,
-  and are applied low-pass filtering to remove abrupt gradients in the resulting deformation
+  and undergo a low-pass filtering to remove abrupt gradients in the resulting deformation
   field.
+But they essentially reproduce the original (and real) distortion.
 
-Regarding the proposed (and also some other possible) comparisons:
+\bigskip
+
+About the proposed (and also some other possible) comparisons, we analyzed them before
+  the original submission, concluding the following:
 \begin{itemize}
@@
 \item \emph{topup} is effectively the tool used to correct diffusion images for the
-  susceptibility artifact in the HCP pipeline.
+  susceptibility artifact in the HCP pipeline \citep{glasser_minimal_2013}.
   This tool uses one image with a certain phase-encoding scheme and a second image
@@
 \item \emph{fieldmap-based correction strategies}, since we use the theoretical basis of these
-  methods to generate our realistic distortions, comparison to these methods is not specially
-  interesting.
+  methods to generate our realistic distortions from real inhomogeneity fieldmaps, 
+  the comparison to these methods is not interesting/fair.
 \item \emph{bbregister} is an affine registration method intended to be used with diffusion
-  data.
+  data in the spatial normalization to the anatomical space of the T1w image.
   However, the ``minimally preprocessed'' datasets from the HCP that we use here are already
   spatially normalized.
-  \emph{bbregister} is designed to find the normalization, and thus it is not applicable
-  here.
+  Therefore \emph{bbregister} is not applicable here.
 \end{itemize}
@@
 \reply{%
+We appreciate the suggestion and introduced the comma accordingly.
 }
@@
 The reviewer is totally right. We have modified the text accordingly.
-This comment is also made by the reviewer \#3 in the \autoref{com:31}.
+This comment is similar to one made by the reviewer \#3 in the \autoref{com:31}.
 We refer the reader to this comment for the listing of actual changes in the paper.
@@
 \reply{%
+We have modified the figure as suggested.
 }
@@
 \reply{%
+We thank the reviewer for the time dedicated in his/her review and these positive comments.
+Again, we are pleased to confirm that the summary of the reviewer agrees with our objectives in this
+  work.
 }
@@
 \reply{%
+We have introduced one Appendix explaining how the shape-gradients are applied and how the equation given
+  in the Figure 1 relates to the registration framework.
+As the reviewer suggests, this text was part of the Supplemental Materials.
 }
@@
 \reply{%
+The reviewer is right in the sense that our method was devised for its application on the extraction of
+  the human connectome from dMRI data.
+However, and with certain adaptations, the registration method is also useful in other approaches such those
+  suggested.
+Therefore, we have relaxed those statements restricting the use of this method to the connectome
+  extraction.
 }
@@
 \reply{%
+We have introduced the name of the model referred to.
 }
@@
 \reply{%
+We have introduced the symbol $\mathcal{N}$ before eq. (4) in which it appears for first time.
 }
@@
 \reply{%
+The direction of the normals is explicitly described in the new Appendix 1.
 }
@@
 \reply{%
+The evolution at $s_1$ can be in either direction, depending on the direction of the gradient.
+Again, with the insertion of Appendix 1 we understand that this issue is clarified.
 }
@@
 \reply{%
+We have included two equations (A.6, A.7) from the Supplemental Materials in the new Appendix 1.
+Additionally, some text has been added to grant the necessary consistency and connect the formulations in
+  the manuscript to the Figure 1.
 }
@@
 \reply{%
+Added a sentence describing A and B.
 }
@@
 \reply{%
+We have introduced an intermediate expression to clarify the derivation of the energy functional.
 }
@@
 \reply{%
-The equation (7) was splited in two lines because in just one line it exceeded the paper margins, and we
-  considered the equal sign the most convenient point to break it.
-Unfortunatelly, we could not interpret what the reviewer means with ``has two plus symbols in a row,
-  one on each line''.
+The eq. (7) was split thinking of the final two-column format of the journal.
+We have removed the line break and also the duplicated plus symbols.
+
 Regarding the last part of the question, the second integral is defined directly along the whole image
@@
 \reply{%
+Added large braces.
 }
@@
 \reply{%
+Added brackets as in \autoref{com:19} and fixed two pluses as in \autoref{com:18}.
 }
@@
 \reply{%
+Added explanation.
 }
@@
 \begin{reviewpoint}
-$g_k$ is completely undefined in equation (11). I see there is information in a supplement but at a minimum each variable should be defined in a sentence in the main paper.
+$\mathbf{g}_k$ is completely undefined in equation (11). I see there is information in a supplement but at a minimum each variable should be defined in a sentence in the main paper.
 \end{reviewpoint}
 \reply{%
+Equations (A.6) and (A.7) introduced in the new Appendix 1 define $\mathbf{g}_k$.
 }
@@
 \reply{%
+Although it was defined in pg. 2, we have introduced the long form again to make the text more readable.
 }
@@
 \reply{%
+A bit of text has been added defining $a_i$.
 }
@@
 \reply{%
+We have modified the caption accordingly.
 }
@@
 \reply{%
+We have modified the caption to clarify this point (also related to \autoref{com:25}).
 }
@@
 \reply{%
+We have updated the caption accordingly.
 }
@@
 \reply{%
+Reference added to that sentence.
 }
@@
 \reply{%
+In order to ensure the reproducibility of the presented experiments, the actual values
+  used in the experiments are provided in the open source repository in the 
+  human-readable JSON file format.
 }

### 2016-01-24 Publications/2015-NeuroImage/03-introduction.tex
starting revision
revised paper

@@
   models, which range from the early \gls*{dti} proposed by \cite{basser_microstructural_1996}
-  to current models such as \citep{daducci_accelerated_2015}.
+  to current models such as \revcomment[R\#3-C.10]{\gls*{amico}}.
 The microstructural map is then used to draw the preferential orientations of diffusion
@@
 The standardized procedure to obtain this mask was thresholding the \gls*{fa} map.
-However, the mask and subsequent analyses are highly dependent on the threshold set
-  \citep{taoka_fractional_2009}.
+However, the mask and subsequent analyses are highly dependent on the 
+  \revcomment[R\#3-C.33]{threshold that is chosen} \citep{taoka_fractional_2009}.
 To overcome the unreliability of \gls*{fa} thresholding, and to broaden

### 2016-01-24 Publications/2015-NeuroImage/04-methods.tex
revised paper

@@
   is presented in \autoref{fig:regseg-method}.
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figure01}
-  \caption{The active contours are defined as the interfacing surfaces of the competing
-    \glspl{roi} $\Omega_l$, which
-  are represented in green and dark blue in this close-up.
-  They evolve iteratively following their inward normals $\hat{n}_i$ at each vertex
-    $\vec{v}_i$ of the mesh.
-  The gradient speeds $\bar{s}_i$ drive registration, which are computed as the disparity of the data
-    energies with respect to the two limiting regions of $M(\vec{v}_i)$, the features of the image
-    $M$ in the location of vertex $\vec{v}_i$ \citepalias[see][equation SM5]{esteban_useful_2015}.
-  In this figure, $\bar{s}_1$ is written in the lower
-    box, with $\Omega_{wm}$ in the inner limiting region, $\Omega_{gm}$ the outer region, and
-    $w_{0,1}$ is the relative area associated with vertex $\vec{v}_1$ with respect to
-    the total area of surface $\Gamma_0$.
-      }\label{fig:regseg-method}
-\end{figure}
 
@@
   1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
-  distributions \citep{esteban_mbis_2014} with parameters
+  distributions 
+\revcomment[R\#3-C.11]{%
+  $\mathcal{N} (\vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} )$}, with parameters
   $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
-  for each region $\Omega_l$.
-Even though the features being segmented are not generally i.i.d., the spatial interdependency of
-  voxels is implicitly supported by the piecewise smooth partition of the space \omegaset{}.
-In \autoref{fig:regseg-model} it is shown that marginal distributions of data can be approximated by
-  multivariate normal distributions. Therefore,
+  for each region $\Omega_l$ \citep{esteban_mbis_2014}:
 
@@
 $C$ is the number of channels comprised in the image $M$.
+\todo{ordering} Even though the features being segmented are not generally i.i.d., the 
+  spatial interdependency of voxels is implicitly supported by the piecewise smooth
+  partition of the space \omegaset{}.
 In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
   the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized.
-In \autoref{fig:regseg-model},
-  this minimization is illustrated for the segmentation of the \gls*{fa} and the \gls*{adc} maps
+The \autoref{fig:regseg-model} shows \revcomment[authors]{how the joint distribution of 
+  the input images} is approximated with a mixture of multivariate normal distributions,
+  and this minimization is illustrated for the segmentation of the \gls*{fa} and the \gls*{adc} maps
   of one subject.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figure01}
+  \caption{The {\color{blue} interfacing surfaces $\Gamma_m$ between the competing \glspl{roi} $\Omega_l$,
+  play the role of active contours which drive the registration process.}
+  They evolve iteratively following {\color{blue} the normal $\hat{\mathbf{n}}_i$ of the surface} at each vertex
+    $\vec{v}_i$ of the mesh.
+  The gradient speeds $\bar{s}_i$ drive registration, which are computed as the disparity of the data
+    energies with respect to the two limiting regions of $M(\vec{v}_i)$, the features of the image
+    $M$ in the location of vertex $\vec{v}_i$.
+  {\color{blue} The computation of shape-gradients is developed in \ref{app:shape_gradients}.}
+  In this figure, the $\bar{s}_1$ derivated from \autoref{eq:regseg-shape_gradient_disc2} is 
+    written in the lower box, with $\Omega_{wm}$ in the inner limiting region, $\Omega_{gm}$
+    the outer region, and $w_{0,1}$ is the relative area associated with vertex $\vec{v}_1$
+    with respect to the total area of surface $\Gamma_0$.
+  }\label{fig:regseg-method}
+\end{figure}
 
@@
   \end{align}
- which requires that the distortion and its gradient have zero
-  mean and variance governed by the matrices $\mathbf{A}$ and $\mathbf{B}$.
+ which requires that the distortion and its gradient have zero mean\revcomment[R\#1-C4]{,}
+   and variance governed by the matrices $\mathbf{A}$ and $\mathbf{B}$.
+\revcomment[R\#3-C.16]{%
+Therefore, $\mathbf{A}$ and $\mathbf{B}$ are tensors that modulate the regularization, and
+  produce deformations with preferential directions}.
 Finally, the maximum a posteriori problem is adapted to a variational problem where we search for
@@
 
+\revcomment[R\#3\\C.17--C.19]{%
   \begin{align}
@@
   \underset{\vec{r} \in \Omega_l}{\prod}
-  \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) = \notag\\ &=
-  \const + \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \,
-  \ifthenelse{\boolean{review}}{+}{+ \notag\\ &+}
-  \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
+  \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) =
+\notag\\ &= - \underset{l}{\sum} \int_{\Omega_l} \Big\{ \log \left[ \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right) \right] \, + \log \left[ p_0( \vec{u})\,p_1( \vec{u}) \right] \Big\} \, d\vec{r}
+   = \notag\\ &=
+  \const + \underset{l}{\sum} \Big\{ \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \Big\} \,
+  + \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
   \label{eq:regseg-energy}
-  \end{align}
+  \end{align}%
+}
 This expression is the dual of the Mumford-Shah functional that corresponds
@@
   $\hat{U}$ of the transformation at the current time point.
-Let us assume that the anisotropy is aligned with the imaging axes to simplify
-  \eqref{eq:regseg-energy} as expression \eqref{eq:regseg-app_energy} in \ref{app:reg_term},
-  and thus to compute its derivative \eqref{eq:regseg-general_gradient_descent}:
+Let us assume that \revcomment[R\#1-C.5\\R\#3-C.31]{the preferential directions of
+  the displacement are aligned} with the imaging axes to simplify \eqref{eq:regseg-energy} as expression 
+  \eqref{eq:regseg-app_energy} in \ref{app:reg_term}, and thus to compute its
+  derivative \eqref{eq:regseg-general_gradient_descent}:
 
+\revcomment[R\#3-C.20]{%
   \begin{align}
@@
   \frac{ \partial }{\partial \vec{u}_k} \Big\{
-  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \,
-  \ifthenelse{\boolean{review}}{+}{+ \notag\\ &+}
-  \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
+  \underset{l}{\sum} \big[ \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \big] \,
+  + \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
   + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
@@
   \label{eq:regseg-gradient_descent}
-  \end{align}
-  where $\vec{u}^{\circ2} = \vec{u}^T \cdot \vec{u}$.
+  \end{align}}
+  where $\vec{u}^{\circ2} = \vec{u}^T \cdot \vec{u}$, \revcomment[R\#3-C.21]{%
+  and $\{\boldsymbol{\alpha}, \boldsymbol{\beta}\}$ are the expected variances along
+  the imaging axes of the displacements field and its gradient, respectively.}.
 Then, the data and regularization terms are split and discretized to compute their
   derivatives.
-The computation of the explicit shape gradients at each $\vec{v}_i'$ is illustrated in \autoref{fig:regseg-method}.
-Then, \eqref{eq:regseg-gradient_descent} can be reformulated as \citepalias[see][equations SM6-SM12]{esteban_useful_2015}:
+\revcomment[R\#3-C.14]{%
+The derivative of the data term is computed using explicit shape gradients (see \ref{app:shape_gradients}),
+  that finally lead to obtain vertex-wise speeds of the gradient $\bar{s}_i$ as}
+  illustrated in \autoref{fig:regseg-method}.
+\revcomment[R\#3-C.14]{%
+Using the expressions \eqref{eq:regseg-gradient_wshape} and 
+  \eqref{eq:regseg-shape_gradient_final} given in \ref{app:shape_gradients}
+  and introducing the analytical derivative of the regularization term, then
+  the \autoref{eq:regseg-gradient_descent} is reformulated as:}
 
@@
   the subject.
-Therefore, the deformation field is nonzero only in the \gls*{pe} axis,
+Therefore, the deformation field is nonzero only in the \revcomment[R\#3-C.23]{\acrfull*{pe}} axis,
   and resembles a real \gls*{epi} distortion.
@@
   \end{equation}
-  where $\vec{v}_i \subset \gammaset_\text{true}$ are the locations of the total $N_V$ vertices
+  where $\vec{v}_i \subset \gammaset_\text{true}$ are the locations of the total $N_V$ vertices,
+  \revcomment[R\#3-C.24]{$a_i$ is the area corresponding to each vertex $\vec{v}_i$},
   and $\hat{\vec{v}}_i \subset \hat{\gammaset}_\text{test}$ are the recovered locations

### 2016-01-24 Publications/2015-NeuroImage/05-results.tex
revised paper

@@
   \includestandalone[width=\linewidth]{figure05}
-  \caption{A. Visual assessment of the results obtained with the low resolution sets:
+  \caption{A. Visual assessment of the results obtained on the {\color{blue} digital phantoms.
+    The panel shows three coronal slices corresponding to the small numbers annotated on the lower
+      left corner of each cut, for} the low resolution sets:
     ``gyrus'' (top left), ``L'' (top right), ``ball'' (bottom left),
@@
   Specific regions where \regseg{} outperformed \gls*{t2b} are enlarged.
-  B. Violin plots of the error distributions for each surface, which show the voxel size of the \gls*{dmri} images
-    (1.25 mm), thereby supporting the improved results obtained by \regseg{} with the proposed settings.
+  B. Violin plots of the error distributions for each surface {\color{blue} across the 16 subjects}, which show the 
+    voxel size of the \gls*{dmri} images (1.25 mm), thereby supporting the improved results obtained
+    by \regseg{} with the proposed settings.
   }\label{fig:regseg-results_real}
@@
   \autoref{fig:regseg-results_real} demonstrate that the errors were always larger with \gls*{t2b}.
-We also show the 95\% CIs of the \gls*{swindex} for these surfaces.
+We also show the 95\% CIs of the \gls*{swindex} for these surfaces 
+  \revcomment[R\#3-C.30]{(\autoref{tab:results_real})}.
 The aggregate CI for \regseg{} was 0.56--0.66 [mm], whereas the \gls*{t2b} method
@@
     \end{tabular}
-    \caption{Statistical analysis of results obtained using real data, which show that \regseg{} performed better than
-    the alternative \acrfull{t2b} method.
+    \caption{Statistical analysis of results obtained using {\color{blue} 16 real datasets from 
+    the \gls*{hcp}}, which show that \regseg{} performed better than the alternative \acrfull{t2b} method.
     The distribution of the errors computed for the surfaces of interest ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$)

### 2016-01-24 Publications/2015-NeuroImage/06-discussion.tex
revised paper

@@
   in \gls*{dmri} images is an active field of research \citep{jeurissen_tissuetype_2015}.
-Current processing steps involved in the connectome extraction require a precise segmentation
-  of anatomical structures in the diffusion space.
+\revcomment[R\#3-C.9]{%
+Current processing of \gls{dmri} involved in the connectome extraction and other applications
+  (such as \gls*{tbss} or surgical planning)} require a precise segmentation
+  of the anatomical structures in the diffusion space.
 Some examples of these processing tasks are the structure-informed reconstruction of \gls*{dmri}
@@
   these two processes and the matrix building task can be performed on the unaltered
-  \gls*{dmri} signal (i.e. without resampling data to an undistorted space).
+  \gls*{dmri} signal (i.e. without resampling data to an undistorted space).}
+\revcomment[R\#3-C.9]{%
 For analyses other than connectivity, like \gls*{tbss}, the deformation estimated by \regseg{}

### 2016-01-24 Publications/2015-NeuroImage/08-appendix-00-shapepriors.tex
revised paper

@@
 \section*{Appendix}
+
+\subsection{Application of the shape-gradients}\label{app:shape_gradients}
+\revcomment[R\#3-C.8]{%
+The computation of gradients at the locations of the active contours in the
+  instant $t$ is based on the work of \cite{herbulot_segmentation_2006}.
+Let $F(\vec{r})$ be an ``arbitrary'' function over the image domain
+  $\Omega = \Omega_l \cup \Omega_m$ split in two regions $l$ and
+  $m$, and $\Gamma_{l,m}$ a closed boundary between them.
+We now derive the domain integral w.r.t. $t$:
+
+  \begin{equation}
+  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) d\vec{r} =
+  \int_\Omega \frac{\partial}{\partial t}F(\vec{r}) d\vec{r}
+  - \int_{\Gamma_{l,m}} F(\vec{r}) \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
+  N_{\Gamma_{l,m}}\right\rangle d\vec{r},
+  \end{equation}
+%
+  where $\left\langle\frac{\partial\Gamma_{l,m}}{\partial t}, N_{\Gamma_{l,m}}\right\rangle$ is
+  the projection of the boundary movement on the unit inward normal $N_{\Gamma_{l,m}}$.
+Assuming that the region descriptors $\{\boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l\}$ vary slowly enough, we can consider
+  that $\frac{\partial}{\partial t} F(\vec{r}) = 0$ and thus:
+
+  \begin{equation}
+  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) d\vec{r} =
+  - \int_{\Gamma_{l,m}} F(\vec{r}) \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
+  N_{\Gamma_{l,m}}\right\rangle d\vec{r}.
+  \label{eq:regseg-shape_gradients}
+  \end{equation}
+
+The equation \eqref{eq:regseg-shape_gradients} is discretized as follows.
+First, the surface between limiting regions $\{l, m\}$, $\Gamma_{l,m}$ is explicitly represented by
+  a discrete set of vertices $\vec{v}_i$, with $i \in \{0, \ldots, N_p -1 \}$.
+Consequently, the inwards normal of the surface $N_{\Gamma_{l,m}}$ is represented by the discrete
+  set of normals $\hat{\vec{n}}_i$ at each vertex of the mesh.
+The resulting summation is, therefore, discrete and the integral operator is replaced by the sum:
+  \begin{align}
+  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) d\vec{r} &=
+  \underbracket{\cancel{\int_\Omega \frac{\partial}{\partial t}F(\vec{r}) d\vec{r} }}_{\text{Functional's evolution}}
+  - \underbracket{\int_{\Gamma_{l,m}} F(\vec{r}) \left\langle \frac{\partial \Gamma_{l,m}}{\partial t},
+  N_{\Gamma_{l,m}}\right\rangle d\vec{r}}_{\text{Shape's evolution}} \notag \\
+  & = - \underset{p}{\sum} \frac{1}{A_p} \underset{i}{\sum} \, a_i \, F(\vec{v}_i) \left\langle \underbracket{\frac{\partial \vec{v}_i}{\partial t}}_{\text{speed of }\vec{v}_i},
+  \hat{\vec{n}}_{i}\right\rangle.
+  \label{eq:regseg-shape_gradient_orig}
+  \end{align}
+where $a_i$ is the area corresponding to vertex $\vec{v}_i$, and $A_p = \sum_i a_i$ is the total area of surface $p$.
+In the following, we will refer as $w_{p,i} = a_i / A_p $ to the area contribution of $\vec{v}_i$ to the
+  total area of the surface it belongs to.
+For simplicity, the sum over $p$ can be also removed, as the vertices belong to only one of the total $P$ contours.
+
+Then, the speed of $\vec{v}_i$ is discretized using the artificial time-step parameter $\delta$, as the displacement
+  $\frac{\partial \vec{v}_i}{\partial t} = \vec{v}_i(\delta = t+1) - \vec{v}_i(\delta = t)$:
+  \begin{equation}
+  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) \, d\vec{r} =
+  - \underset{i}{\sum} w_{p,i} F(\vec{v}_i) \frac{\partial \vec{v}_i}{\partial t} \cdot \hat{\vec{n}}_i.
+  \label{eq:regseg-shape_gradient_disc1}
+  \end{equation}
+
+Since the energy functional is defined over competing regions, the displacement of $\vec{v}_i$ will cause
+  an energy exchange between the limiting regions, and therefore $F(\vec{r})$ must be split in
+  two terms, $F_{in}(\vec{r})$ corresponding to the interior region and $F_{out}(\vec{r})$ to the exterior:
+  \begin{equation}
+  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) \, d\vec{r} =
+  - \underset{i}{\sum} \, \frac{\partial \vec{v}_i}{\partial t} \cdot
+  \underbracket{w_{p,i} \, \Big[ F_{out}(\vec{v}_i) - F_{in}(\vec{v}_i) \Big] \hat{\vec{n}}_i}_{\bar{s}_i \text{ in Figure 1}}.
+  \label{eq:regseg-shape_gradient_disc2}
+  \end{equation}}
+
+\revcomment[R\#3-C.22]{%
+Then, we identify the shape gradient contribution $\vec{g}_k$ on the coefficients $\vec{u}_k$ of the B-Spline grid:
+  \begin{equation}
+  \label{eq:regseg-gradient_wshape}
+  \begin{split}
+  \vec{g}_k &= - \underset{i}{\sum} \left\langle \frac{\partial \vec{v}_i'}{\partial \vec{u}_k}, \bar{s}_i'\right\rangle \\
+  \text{with }
+  \bar{s}_i' &= w_i \left[ \mdist{f_i'}{out} - \mdist{f_i'}{in} \right] \, \hat{\vec{n}}_i, \\
+  \text{and }
+  \frac{\partial \vec{v}_i'}{\partial \vec{u}_k} &=
+  \frac{\partial}{\partial \vec{u}_k} \left\{ \vec{v}_i + \sum_k \psi_k(\vec{v}_i) \vec{u}_k \right\} = \psi_k(\vec{v}_i)\, \hat{\vec{e}},
+  \end{split}
+  \end{equation}%
+  where $\hat{\vec{e}}$ is the coordinates system's unit vector.
+Therefore, the shape gradients projected to the grid of B-Spline control points is:
+\begin{equation}
+  \vec{g}_k = - \underset{i}{\sum} \bar{s}_i \cdot \psi_k(\vec{v}_i) \, \hat{\vec{e}}.
+  \label{eq:regseg-shape_gradient_final}
+\end{equation}}
 
@@
 
-In general, the anisotropy is aligned with the imaging axes, so 
-  $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:regseg-energy} can be simplified to diagonal matrices, such that
+In general, the anisotropy \revcomment[R\#3-C.31]{of the distortion field} is aligned with the 
+  \revcomment[R\#3-C.31]{voxel coordinate system}, so
+  $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:regseg-energy} can be simplified to diagonal matrices
+  \revcomment[R\#3-C.31]{to regularize the regitration process}, such that
   $\mathbf{A}= \, \boldsymbol{\alpha}\,\vec{I}_n$ and
@@
   \begin{align}
-  E(\vec{u}) &= \const + \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} +   \int_{\Omega} \frac12 \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2} + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}.
+  E(\vec{u}) &= \const + \, \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \,
+  \ifthenelse{\boolean{review}}{+}{+ \notag\\ &+}
+  \int_{\Omega} \frac12 \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2} + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}.
   \label{eq:regseg-app_energy}

### 2016-01-24 Publications/2015-NeuroImage/supplemental.tex
revised paper

@@
 % @Date:   2015-02-20 15:40:09
-% @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-04-29 17:53:12
+% @Last Modified by:   oesteban
+% @Last Modified time: 2015-11-11 13:14:36
 
@@
 \begin{document}
-\title{Supplemental Materials: \\ \emph{Active contours-driven registration method for the structure-informed segmentation of diffusion MR images}}
+\title{Supplemental Materials: \\ \emph{Surface-driven registration method for the structure-informed segmentation of diffusion MR images}}
 \author{Esteban O., Zosso D., Daducci A., Bach-Cuadra M., \\ Ledesma-Carbayo MJ., Thiran JP, and Santos A.}
@@
 Let $F(\vec{r})$ be an ``arbitrary'' function over the image domain
-  $\Omega = \Omega_l \cup \Omega_m$ splitted in two regions $l$ and
+  $\Omega = \Omega_l \cup \Omega_m$ split in two regions $l$ and
   $m$, and $\Gamma_{l,m}$ a closed boundary between them.

### 2016-01-27 Publications/2015-NeuroImage/04-methods.tex
revised the paper

@@
 \begin{figure*}
-\includestandalone[width=\linewidth]{figure02-model}
-% \includegraphics[width=\linewidth]{figure04A}
+% \includestandalone[width=\linewidth]{figure02-model}
+\includegraphics[width=\linewidth]{figure02}
 % \includegraphics[width=\linewidth]{figure04B}
@@
 \caption{%
-Evolution of the segmentation model defined by the homogeneous regions $\Omega_l$
-  for a real dataset.
-The left plot in each panel represents a kernel density estimate of the location and spread of
-  each region $\Omega_l$ in the bivariate feature space of $M$, which comprises
-  the \gls*{fa} and \gls*{adc} maps.
-At the top and the right margins, the marginal distributions of $\Omega_l$ on the \gls*{fa} and
-  \gls*{adc} are respectively displayed.
-On the right side, the sampled distribution of each region is shown in small scatter plot, in contrast to
-  the complete bivariate data.
-A) shows the initialization of the model, when the anatomically-correct surfaces are not
-  mapped into the \gls*{dmri} space.
-The panel shows the resulting conditional distributions of segmenting the (distorted) \gls*{fa} and
-  \gls*{adc} maps with the original (undistorted) surfaces.
-Panel B shows the equivalent information, using the surfaces that are perfectly mapped into the
-  distorted \gls*{dmri} space.
-Since the surfaces are now aligned in $M$, the spreads of the conditional distributions for each region are
-  reduced with respect to those in A.
-Panel C, after \regseg{} registration, shows how the covariance matrices $\boldsymbol{\Sigma}_l$
-  are minimized with respect to panel A, and the resulting segmentation is close
-  to the theoretical situation presented in panel B.}\label{fig:regseg-model}
+Evolution of the segmentation model defined by the homogeneous regions $\Omega_l$,
+  for one real dataset.
+{\color{blue}%
+Panel (A, left) shows the joint distribution of the \gls*{fa} and \gls*{adc} conditioned to
+  the segmentation \omegaset{} defined by the surfaces $\gammaset_R$ extracted from
+  the \gls*{t1} image.
+The plot was generated for reference using undistorted diffusion data, and
+  therefore, $\gammaset_R$ is aligned with the \gls*{fa} and the \gls*{adc}.
+The problem arises when the diffusion data present deformation, and the
+  contours $\gammaset_R$ do not fit within the data (A, center).
+After registration with \regseg{}, the contours are mapped onto the
+  diffusion data (A, right), and the joint density plot is closer to the reference
+  situation.
+In panel (B), the three plots in (A) are decomposed tissue-wise.
+Using filled contours, the bivariate distribution of each tissue is highlighted in its designated color,
+  and represented over the remaining tissues (in gray colors).
+To help assessment, dashed contours in black-to-white colors represent
+  the corresponding distribution in the reference plot.
+The registration process optimizes the segmentation model of \regseg{},
+  and thus, the distribution of each region after registration is located
+  closer to that corresponding in the reference situation, the shape of the distribution
+  is more similar to the reference, and their spread is also reduced.
+The effects of optimization are more noticeable on the \gls*{gm} ($\Omega_\text{GM}$) and
+  the \gls*{wm} ($\Omega_\text{WM}$).
+Particularly, the \gls*{wm} typically shows a bimodal distribution when the contours
+  \gammaset{} do not fit the data.}
+% The left plot in each panel represents a kernel density estimate of the location and spread of
+%   each region $\Omega_l$ in the bivariate feature space of $M$, which comprises
+%   the \gls*{fa} and \gls*{adc} maps.
+% At the top and the right margins, the marginal distributions of $\Omega_l$ on the \gls*{fa} and
+%   \gls*{adc} are respectively displayed.
+% On the right side, the sampled distribution of each region is shown in small scatter plot, in contrast to
+%   the complete bivariate data.
+% A) shows the initialization of the model, when the anatomically-correct surfaces are not
+%   mapped into the \gls*{dmri} space.
+% The panel shows the resulting conditional distributions of segmenting the (distorted) \gls*{fa} and
+%   \gls*{adc} maps with the original (undistorted) surfaces.
+% Panel B shows the equivalent information, using the surfaces that are perfectly mapped into the
+%   distorted \gls*{dmri} space.
+% Since the surfaces are now aligned in $M$, the spreads of the conditional distributions for each region are
+%   reduced with respect to those in A.
+% Panel C, after \regseg{} registration, shows how the covariance matrices $\boldsymbol{\Sigma}_l$
+%   are minimized with respect to panel A, and the resulting segmentation is close
+%   to the theoretical situation presented in panel B.}
+}\label{fig:regseg-model}
 \end{figure*}
@@
 \begin{figure*}
-\includegraphics[width=\linewidth]{figure02}
+\includegraphics[width=\linewidth]{figure03}
 \caption{Evaluation of \regseg{} using phantom data according to the following instrumental workflow.
@@
 \begin{figure*}
-\includegraphics[width=\linewidth]{figure03}
+\includegraphics[width=\linewidth]{figure04}
 \caption{Experimental workflow employed to process real data from the \acrfull*{hcp}.

### 2016-01-27 Scripts/pyacwereg/viz.py

@@ -398,7 +398,7 @@ def jointplot_data(im1data, im2data, in_seg, labels=None, out_file=None,
     sn.set_context("talk", font_scale=1.8)
     g = sn.ConditionalJointGrid(f1name, f2name, df, hue='tissue',
                                 xlim=f1lims, ylim=f2lims, size=20,
-                                inline_labels=True)
+                                inline_labels=True, splitgrid=splitgrid)
     g.plot_joint(sn.kdeplot, linewidths=3.0)
     g.plot_marginals(sn.distplot)
 
@@ -417,7 +417,7 @@ def jointplot_data(im1data, im2data, in_seg, labels=None, out_file=None,
 def jointplot_compare(
         imageX, imageY, segmentation, mask, imrefX, imrefY, refsegm, refmask,
         labels=None, xlabel='X', ylabel='Y', tlabel=None, huelabel='tissue', size=8, dpi=100,
-        subsample=.2, xlims=None, ylims=None, out_file=None):
+        subsample=.2, xlims=None, ylims=None, out_prefix=None):
     import os.path as op
     import nibabel as nb
     import numpy as np
@@ -437,6 +437,9 @@ def jointplot_compare(
     if ylims is None:
         ylims = (md.min(), md.max())
 
+    fa_m = np.median(fa[np.where((fa > xlims[0]) & (fa < xlims[1]))])
+    md_m = np.median(md[np.where((md > ylims[0]) & (md < ylims[1]))])
+
     md[md < ylims[0]] = np.nan
     md[md > ylims[1]] = np.nan
 
@@ -445,6 +448,9 @@ def jointplot_compare(
     md_ref = nb.load(imrefY).get_data().astype(np.float32).reshape(-1)
     seg_ref = nb.load(refsegm).get_data().astype(np.uint8).reshape(-1)
     msk_ref = nb.load(refmask).get_data().astype(np.uint8).reshape(-1)
+
+    fa_ref += fa_m - np.median(fa_ref[np.where((fa_ref > xlims[0]) & (fa_ref < xlims[1]))])
+    md_ref += md_m - np.median(md_ref[np.where((md_ref > ylims[0]) & (md_ref < ylims[1]))])
 
     md_ref[md_ref < ylims[0]] = np.nan
     md_ref[md_ref > ylims[1]] = np.nan
@@ -481,23 +487,26 @@ def jointplot_compare(
     dfref = dfref[dfref.tissue != 'do-not-show']
     sn.set_context("talk", font_scale=1. + 0.1*size)
 
[…]

### 2016-02-02 Publications/2015-NeuroImage/responseLetter2.tex
added fixes of all coauthors, and spell check

@@
   real fieldmaps, and therefore they are not ``modeled'' in a strict sense.
-We have modified the paper to clarify the nature of these distortions (see \autoref{com:3}) in
-  both experiments (simulated and real datasets).
+We have modified the abstract, to clarify the nature of these distortions in both experiments (simulated and real datasets).
+Additionally, some other editions in the paper with respect to this aspect are
+  described in the response to \autoref{com:3}.
 }
@@
   tractography.
-We have modified certain points in the abstract, introduction and discussion to
+We have modified the last sentence in the abstract and lines 339--344 in the discussion to
   clarify that this method is intended to support any diffusion processing task that
   requires a precise delineation of the anatomy in the diffusion space.
-We have also looked for the phrase cited by the reviewer\footnote{``this method avoids any
- requirement for collecting additional scans nor resampling for diffusion unwarping''}
- to reformulate it, however we did not found such exact statement in the paper.
+Although we have not found the exact phrase cited by the reviewer\footnote{``this method avoids any
+ requirement for collecting additional scans nor resampling for diffusion unwarping''},
+ we have modified our close statements in lines 55--60 to clarify both features
+ of the method.
 
-In the field of connectivity analysis, this need has posed important problems in the
+In the field of connectivity analysis, the issue of distortion has posed important problems in the
   standard methodologies to build the connectivity network\footnote{generally imposing
   a parcellation mapped from T1-space}.
-With the emergence of novel reconstruction and tractography methods (see lines ),
+With the emergence of novel reconstruction and tractography methods (see lines 5--15),
   these tasks may also be improved with the application of more accurate registration
@@
   the other two references given [2, 3]).
-We have included in the discussion (see line XXX) some text indicating that the tracks
+We have included in the discussion (see lines 343-344) some text indicating that the tracks
   can be unwarped using the deformation found with \emph{regseg} since it is revertible.
@@
   on the original dMRI data and then unwarped with \emph{regseg} or whatever other
-  technique is out of the scope of the paper, but a very interesting lead for
+  technique is out of the scope of the paper, but certainly a very interesting lead for
   future research.
@@
   field.
-But they essentially reproduce the original (and real) distortion.
+But they essentially reproduce the original (and real) distortion, as described in lines 192--200
+  of the manuscript.
+As introduced in \autoref{com:1}, we have modified the abstract to better show this aspect.
 
@@
   Adding this experiment would include a significant overhead that may displace the focus
-    of the paper from ``an accurate nonlinear registration method'' to ``yet another 
-    correction method for susceptibility-derived distortions''.
+    of the paper from ``an accurate nonlinear registration method to map the anatomical
+    structures into dMRI space'' to ``yet another correction method for susceptibility-derived
+    distortions''.
   For these two reasons, we disregarded the implementation of such experiment.
@@
   However, the ``minimally preprocessed'' datasets from the HCP that we use here are already
-  spatially normalized.
+  spatially normalized in this sense.
   Therefore \emph{bbregister} is not applicable here.
@@
 \reply{%
-The reviewer is totally right. We have modified the text accordingly.
-This comment is similar to one made by the reviewer \#3 in the \autoref{com:31}.
-We refer the reader to this comment for the listing of actual changes in the paper.
+The reviewer is totally right. We have modified the text accordingly (lines 117--118,
+  and lines 412--414 in the Appendix). 
+This comment is close and consistent with \autoref{com:31}.
 }
@@
 \reply{%
-We have modified the figure as suggested.
+We have modified the figure and its caption as suggested.
 }
@@
 \reply{%
-We have introduced one Appendix explaining how the shape-gradients are applied and how the equation given
+We have introduced one Appendix (starting by line 396) explaining how the shape-gradients are applied and how the equation given
   in the Figure 1 relates to the registration framework.
-As the reviewer suggests, this text was part of the Supplemental Materials.
+As the reviewer indicated, this text was part of the Supplemental Materials.
 }
@@
   the human connectome from dMRI data.
-However, and with certain adaptations, the registration method is also useful in other approaches such those
+However, and with certain adaptations, the registration method is also useful in other approaches such as those
   suggested.
 Therefore, we have relaxed those statements restricting the use of this method to the connectome
-  extraction.
+  extraction (in the abstract, and lines 323--324) and included brief indications on how this could be applied in
+  other dMRI applications (lines 343--344).
 }
@@
 \reply{%
-We have introduced the name of the model referred to.
+We have introduced the name of the model referred to (line 8).
 }
@@
 \reply{%
-We have introduced the symbol $\mathcal{N}$ before eq. (4) in which it appears for first time.
+We have introduced the symbol $\mathcal{N}$ before eq. (4) in which it appears for first time (line 90).
 }
@@
 \reply{%
-The direction of the normals is explicitly described in the new Appendix 1.
+The active contours are explicitly introduced in line 46, and an implicit reference to this methodological foundation
+is found in the abstract: ``\emph{we propose regseg, a surface-to-volume nonlinear registration method that segments homogeneous regions within multivariate images by mapping a set of nested reference-surfaces}''.
+The reviewer is right regarding the missing details about the direction of normals.
+In the revised manuscript, this is described in the new Appendix 1 (equations A.1--4).
 }
@@
 We have included two equations (A.6, A.7) from the Supplemental Materials in the new Appendix 1.
-Additionally, some text has been added to grant the necessary consistency and connect the formulations in
+Additionally, some text has been added (lines 122--125) to ensure the necessary consistency and connect the formulations in
   the manuscript to the Figure 1.
@@
 \reply{%
-Added a sentence describing A and B.
+Added a sentence describing A and B (lines 103--104).
 }
@@
 \reply{%
-We have introduced an intermediate expression to clarify the derivation of the energy functional.
+We have introduced an intermediate expression to clarify the derivation of the energy functional (Equation 7).
 }
@@
 The first integral is computed region-wise, since the energy functional has different
-  parameters $\Theta_l$ for each of the regions $\Omega_l$.
+  parameters $\{\boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l\}$ for each of the regions $\Omega_l$.
 }
@@
 \reply{%
-Added explanation.
+Added explanation in lines 120--121.
 }
@@
 \reply{%
-Equations (A.6) and (A.7) introduced in the new Appendix 1 define $\mathbf{g}_k$.
+Equations (A.6) and (A.7) introduced in the new Appendix 1 define $\mathbf{g}_k$ (starts by line 406).
+Additionally, a reference to the Appendix and corresponding equations (lines 124--125) and a brief
+  explanation of $\mathbf{g}_k$ (line 126) have been added.
 }
@@
 \reply{%
-Although it was defined in pg. 2, we have introduced the long form again to make the text more readable.
+Although it was defined in pg. 2, we have introduced the long form again (line 143) to make the text more readable.
 }
@@
 \reply{%
-A bit of text has been added defining $a_i$.
+A bit of text has been added defining $a_i$ (line 226).
 }
@@
 \reply{%
+The reviewer is right, this is not clear in the manuscript.
+We have modified the caption of figure 5 to state the the violin plots represent the average error on each surface,
+  across experiments.
 }
@@
 \reply{%
-Reference added to that sentence.
+The reference to Table 2 has been added to that sentence (line 249).
 }
@@
 \reply{%
-The reviewer is totally right. We have modified the text accordingly.
+The reviewer is totally right. We have modified the text accordingly (lines 117--118,
+  and lines 412--414 in the Appendix).
 This comment is also made by the reviewer \#1 in the \autoref{com:5}.
@@
 \reply{%
-In order to ensure the reproducibility of the presented experiments, the actual values
+We agree with the reviewer on the interest of the actual values for $\boldsymbol{\alpha}$
+  and $\boldsymbol{\beta}$ to replicate our experiments or derive new works.
+In order to ensure the reproducibility of the presented experiments, the values
   used in the experiments are provided in the open source repository in the 
   human-readable JSON file format.
+Please find the references to the parameters files in the footnotes of pages 9 and 10.
+We also have included typical ranges for these values in lines 132--133.
 }

### 2016-02-02 Publications/2015-NeuroImage/03-introduction.tex
added fixes of all coauthors, and spell check

@@
   application on three key aspects:
-  1) it avoids the resampling the \glspl*{dwi} introduced by the correction of \gls*{epi}
-    distortions, since processing can be performed on the native \gls*{dmri} space;
-  2) the surfaces are typically extracted from the \gls*{t1} image of the same subject,
+  1) the surfaces are typically extracted from the \gls*{t1} image of the same subject,
     therefore \regseg{} does not require additional MR acquisitions to the minimal
-    \gls*{dmri} protocol; and
+    \gls*{dmri} protocol \revcomment[R\#1-C.2]{in order to estimate the deformation field};
+  2) \revcomment[R\#1-C.2]{depending on the application, the resampling of the \glspl*{dwi}
+    introduced by the correction method can be avoided, either by performing the posterior
+    processing on the native \gls*{dmri} space or applying the unwarping on the tractography
+    itself}; and
   3) \regseg{} increases the geometrical accuracy of the overall process.

### 2016-02-02 Publications/2015-NeuroImage/04-methods.tex
added fixes of all coauthors, and spell check

@@
   {\color{blue} The computation of shape-gradients is developed in \ref{app:shape_gradients}.}
-  In this figure, the $\bar{s}_1$ derivated from \autoref{eq:regseg-shape_gradient_disc2} is 
+  In this figure, the $\bar{s}_1$ derived from \autoref{eq:regseg-shape_gradient_disc2} is 
     written in the lower box, with $\Omega_{wm}$ in the inner limiting region, $\Omega_{gm}$
@@
 % \includestandalone[width=\linewidth]{figure02-model}
-\includegraphics[width=\linewidth]{figure02}
+\begin{minipage}{\textwidth}
+\includegraphics[width=\linewidth]{figure02}\\%
+\vskip-2.0em
+\includegraphics[width=\linewidth]{legend}%x
+\end{minipage}
 % \includegraphics[width=\linewidth]{figure04B}
@@
   and $\{\boldsymbol{\alpha}, \boldsymbol{\beta}\}$ are the expected variances along
-  the imaging axes of the displacements field and its gradient, respectively.}.
+  the imaging axes of the displacements field and its gradient, respectively.}
 Then, the data and regularization terms are split and discretized to compute their
@@
   \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
-  \vec{g}_k  + \boldsymbol{\alpha} \cdot \vec{u}_k - \boldsymbol{\beta} \cdot (\Delta \vec{u}_k).
+  \vec{g}_k  + \boldsymbol{\alpha} \cdot \vec{u}_k - \boldsymbol{\beta} \cdot (\Delta \vec{u}_k),
   \label{eq:regseg-final_gradient}
   \end{align}
-
+\revcomment[R\#3-C22]{%
+where $\vec{g}_k$ are the shape-gradient contributions on the coefficients $\vec{u}_k$ of the B-spline grid.}
 Finally, to descend this gradient, we establish a semi-implicit Euler scheme
@@
   functions for each level of the multiresolution strategy.
+\revcomment[R\#3-C32]{As a rule of thumb, for a $\delta =$ 1.0 both $\boldsymbol{\alpha}$
+  and $\boldsymbol{\beta}$ will typically be in the range [0.0, 1.0].}
 The implementation details as well as other features (such as the sparse matrix approach
   to fast interpolation) and the main parameters
-  (such as $\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, the B-spline grid resolutions,
+  ($\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, the B-spline grid resolutions,
   and target image smoothing) are discussed in the \citetalias[][section S1]{esteban_useful_2015}.
@@
 \caption{Evaluation of \regseg{} using phantom data according to the following instrumental workflow.
-  1) The reference surfaces $\gammaset_R$ are triangularized meshes extracted from the four binary shapes (i.e., ``box'', ``ball'', ``L'', ``gyrus'').
+  1) The reference surfaces $\gammaset_R$ are triangle meshes extracted from the four binary shapes (i.e., ``box'', ``ball'', ``L'', ``gyrus'').
   2) A ground-truth displacement field was generated as described in \autoref{sec:regseg-digital_phantoms}, and applied to warp
@@
 For the experiments on real data, \regseg{} established a multi-resolution pyramid of
-  B-Spline functions, with control points distributed on grids of the following spacings:
+  B-spline functions, with control points distributed on grids of the following spacings:
   40$\times$100$\times$40 [mm] for the first (coarser) level, \isores[mm]{30} for the second level,

### 2016-02-02 Publications/2015-NeuroImage/05-results.tex
added fixes of all coauthors, and spell check

@@
   \caption{A. Visual assessment of the results obtained on the {\color{blue} digital phantoms.
-    The panel shows three coronal slices corresponding to the small numbers annotated on the lower
-      left corner of each cut, for} the low resolution sets:
+    The panel shows three coronal slices (indices indicated on the lower left corner) of each phantom
+    volume at low resolution}:
     ``gyrus'' (top left), ``L'' (top right), ``ball'' (bottom left),
@@
     clustering methods, but they were successfully segmented with \regseg{}.
-  B. Quantitative evaluation of registration errors in terms of the average Hausdorff distances between
-    surfaces at low (left) and high (right) resolutions, which demonstrate that the errors were
-    consistently below the size of the voxels.
+  B. Quantitative evaluation: {\color{blue} the violin plot shows the variability across experiments
+    of the average Hausdorff distance measured in each vertex of the corresponding surface, for the}
+    low (left) and high (right) resolutions. 
+    {\color{blue} Error averages were} consistently below the size of the voxel.
     }\label{fig:regseg-phantom}

### 2016-02-02 Publications/2015-NeuroImage/06-discussion.tex
added fixes of all coauthors, and spell check

@@
 For analyses other than connectivity, like \gls*{tbss}, the deformation estimated by \regseg{}
-  can be used to map the tracts into reference space.}
+  can be used to map the tracts into structural space.}
 

### 2016-02-02 Publications/2015-NeuroImage/08-appendix-00-shapepriors.tex
added fixes of all coauthors, and spell check

@@
 \revcomment[R\#3-C.22]{%
-Then, we identify the shape gradient contribution $\vec{g}_k$ on the coefficients $\vec{u}_k$ of the B-Spline grid:
+Then, we identify the shape gradient contribution $\vec{g}_k$ on the coefficients $\vec{u}_k$ of the B-spline grid:
   \begin{equation}
@@
   where $\hat{\vec{e}}$ is the coordinates system's unit vector.
-Therefore, the shape gradients projected to the grid of B-Spline control points is:
+Therefore, the shape gradients projected to the grid of B-spline control points is:
 \begin{equation}
@@
   $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:regseg-energy} can be simplified to diagonal matrices
-  \revcomment[R\#3-C.31]{to regularize the regitration process}, such that
+  \revcomment[R\#3-C.31]{to regularize the registration process}, such that
   $\mathbf{A}= \, \boldsymbol{\alpha}\,\vec{I}_n$ and

### 2016-02-02 Publications/2015-NeuroImage/08-appendix-01.tex
added fixes of all coauthors, and spell check

@@
 In the probabilistic approach to image segmentation, the smoothness constraints
-can be introduced into the \gls{map} criterium as a \gls{mrf}. Accordingly to the
+can be introduced into the \gls{map} criterion as a \gls{mrf}. Accordingly to the
 Hammersley-Clifford theorem (\textbf{FIXME}: citations needed), an \gls{mrf} can
@@
 \end{equation}
-The normalizer probability $P(I)$ has been ommited for simplicity.
+The normalizer probability $P(I)$ has been omitted for simplicity.
 

### 2016-02-02 Publications/2015-NeuroImage/02-abstract.tex
fixed error in abstract

@@
   \revcomment[R\#1-C2]{is fundamental to increase the reliability of network building
-  in connectivity analyses, and to improve the performance the emerging structure-informed
+  in connectivity analyses, and to improve the performance of the emerging structure-informed
   techniques for \gls*{dmri} data processing}.

### 2016-02-03 Publications/2015-NeuroImage/responseLetter2.tex
revised paper

@@
 We thank the reviewer for the time dedicated in his/her review and these positive comments.
-We are pleased that this summary of the paper matches the idea we want to transmit.
+We are pleased to confirm that this summary of the paper matches the idea we want to transmit.
 
@@
   tractography.
-We have modified the last sentence in the abstract and lines 339--344 in the discussion to
+We have modified the last sentence in the abstract and lines 57--58 (introduction) to
   clarify that this method is intended to support any diffusion processing task that
   requires a precise delineation of the anatomy in the diffusion space.
-Although we have not found the exact phrase cited by the reviewer\footnote{``this method avoids any
+In particular, although we have not found the exact phrase cited by the
+ reviewer\footnote{``this method avoids any
  requirement for collecting additional scans nor resampling for diffusion unwarping''},
- we have modified our close statements in lines 55--60 to clarify both features
+ we have revised our close statements in lines 55--60 to clarify both features
  of the method.
@@
   not correctly aligned with the \emph{b0} itself.
+
+Regarding how the method integrates within the diffusion processing pipeline,
+  we have included the new lines 339--344.
 \end{itemize}
@@
 \reply{%
-The reviewer is totally right. We have modified the text accordingly (lines 117--118,
-  and lines 412--414 in the Appendix). 
+The reviewer is totally right. We have modified the text accordingly (lines 116--117,
+  and lines 400--402 in the Appendix 1). 
 This comment is close and consistent with \autoref{com:31}.
@@
 \reply{%
-We have introduced one Appendix (starting by line 396) explaining how the shape-gradients are applied and how the equation given
+We have introduced one Appendix (\emph{Appendix 2. Application of the shape-gradients}, starting by line 402) to
+  show how the shape-gradients are applied and how the equation given
   in the Figure 1 relates to the registration framework.
@@
 \reply{%
-We have introduced the symbol $\mathcal{N}$ before eq. (4) in which it appears for first time (line 90).
+We have introduced the symbol $\mathcal{N}$ before eq. (4) in which it appears for first time (line 89).
 }
@@
 The reviewer is right regarding the missing details about the direction of normals.
-In the revised manuscript, this is described in the new Appendix 1 (equations A.1--4).
+In the revised manuscript, this is described in the new Appendix 1 (equations A.1--4), and the caption of
+  Figure 1 has been also modified.
 }
@@
 The evolution at $s_1$ can be in either direction, depending on the direction of the gradient.
-Again, with the insertion of Appendix 1 we understand that this issue is clarified.
+Again, with modification of the caption of Fig. 1 and the insertion of Appendix 1 we understand that this issue is clarified.
 }
@@
 \reply{%
-Added a sentence describing A and B (lines 103--104).
+Added a sentence describing A and B (lines 102--103).
 }
@@
 \reply{%
-Added explanation in lines 120--121.
+Added explanation in lines 119--120.
 }
@@
 \reply{%
-Equations (A.6) and (A.7) introduced in the new Appendix 1 define $\mathbf{g}_k$ (starts by line 406).
-Additionally, a reference to the Appendix and corresponding equations (lines 124--125) and a brief
-  explanation of $\mathbf{g}_k$ (line 126) have been added.
+Equations (A.6) and (A.7) introduced in the new Appendix 2 define $\mathbf{g}_k$ (starts by line 423).
+Additionally, a reference to the Appendix and corresponding equations (lines 123--124) and a brief
+  explanation of $\mathbf{g}_k$ (line 125) have been added.
 }
@@
 \reply{%
-The reviewer is totally right. We have modified the text accordingly (lines 117--118,
-  and lines 412--414 in the Appendix).
+The reviewer is totally right. We have modified the text accordingly (lines 116--117,
+  and lines 400--402 in the Appendix 1).
 This comment is also made by the reviewer \#1 in the \autoref{com:5}.
@@
   and $\boldsymbol{\beta}$ to replicate our experiments or derive new works.
+We have included the typical range for these parameters in lines 131--133.
 In order to ensure the reproducibility of the presented experiments, the values
@@
 Please find the references to the parameters files in the footnotes of pages 9 and 10.
-We also have included typical ranges for these values in lines 132--133.
 }
@@
 All the issues have been corrected following the suggestions of the reviewer.
+In particular, som additional appearances of ``displacements field'' have been corrected.
 }

### 2016-02-03 Publications/2015-NeuroImage/03-introduction.tex
revised paper

@@
 Here, we present a segmentation and surface-to-volume registration method
-  named \regseg{}, and show its usefulness in mapping anatomical information from structural
+  called \regseg{}, and show its usefulness in mapping anatomical information from structural
   space into native \gls*{dmri} space to aid subsequent processing steps
@@
   target-image.
-In the presented application to segment \gls*{dmri} data, we propose to use \regseg{}
-  using a set of nested surfaces extracted from a structural image (e.g. \gls*{t1}) and
-  a bivariate target-volume comprehending the \gls*{fa} and \gls*{adc} maps as inputs.
+\revcomment[authors]{We apply \regseg{} to segment \gls*{dmri} data by mapping
+  a} set of nested surfaces extracted from a structural image (e.g. \gls*{t1}) to
+  a bivariate target-volume comprehending the \gls*{fa} and \gls*{adc} maps.
 The evolution of the surfaces is supported by a B-spline basis, optimized
@@
   in registration.
-We then evaluate \regseg{} on real \gls*{dmri} datasets, using a derivation of our
+Then, we evaluate \regseg{} on real \gls*{dmri} datasets, using a derivation of our
   instrumentation framework \citep{esteban_simulationbased_2014} which simulates
@@
   the corresponding \gls*{t2} image of the same subject.
-This method is a derivation of the \gls*{t2b} correction methods \citep{kybic_unwarping_2000},
-  reproducing the settings of an existing implementation included within a widely used diffusion
+\revcomment[authors]{This approach is the first step of the} \gls*{t2b} correction
+  methods \citep{kybic_unwarping_2000}.
+\revcomment[authors]{We reproduce the settings and implementation of a} widely used diffusion
   processing software \citep[\emph{ExploreDTI},][]{leemans_exploredti_2009}.

### 2016-02-03 Publications/2015-NeuroImage/04-methods.tex
revised paper

@@
   \begin{equation}
-  P(M \mid U) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
+  P(M \mid U) = \underset{l}{\prod} \underset{\vec{r}'\in \Omega_l}{\prod}
     P\left( \vec{f}' \mid U \right),
@@
   \begin{align}
-  P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
+  P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r}' \in \Omega_l}{\prod}
   \mathcal{N} (\vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} )
   \ifthenelse{\boolean{review}}{=}{= \notag\\ &=}
-  \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
+  \underset{l}{\prod} \underset{\vec{r}' \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
   \mdist{f'}{l} \right)}},
@@
   play the role of active contours which drive the registration process.}
-  They evolve iteratively following {\color{blue} the normal $\hat{\mathbf{n}}_i$ of the surface} at each vertex
+  They evolve iteratively {\color{blue} along the normal $\hat{\mathbf{n}}_i$ of the surface} at each vertex
     $\vec{v}_i$ of the mesh.
@@
   E(\vec{u}) &= -\log \underset{l}{\prod}
-  \underset{\vec{r} \in \Omega_l}{\prod}
+  \underset{\vec{r}' \in \Omega_l}{\prod}
   \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) =
-\notag\\ &= - \underset{l}{\sum} \int_{\Omega_l} \Big\{ \log \left[ \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right) \right] \, + \log \left[ p_0( \vec{u})\,p_1( \vec{u}) \right] \Big\} \, d\vec{r}
+\notag\\ &= - \underset{l}{\sum} \int_{\Omega_l} \Big\{ \log \left[ \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right) \right] \, + \log \left[ p_0( \vec{u})\,p_1( \vec{u}) \right] \Big\} \, d\vec{r}'
    = \notag\\ &=
   \const + \underset{l}{\sum} \Big\{ \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \Big\} \,
-  + \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
+  + \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}'.
   \label{eq:regseg-energy}
@@
   \frac{ \partial }{\partial \vec{u}_k} \Big\{
-  \underset{l}{\sum} \big[ \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \big] \,
+  \underset{l}{\sum} \big[ \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r}' \big] \,
   + \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
-  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
+  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}'
   \Big\},
@@
   and $\{\boldsymbol{\alpha}, \boldsymbol{\beta}\}$ are the expected variances along
-  the imaging axes of the displacements field and its gradient, respectively.}
+  the imaging axes of the displacement field and its gradient, respectively.}
 Then, the data and regularization terms are split and discretized to compute their
@@
 The derivative of the data term is computed using explicit shape gradients (see \ref{app:shape_gradients}),
-  that finally lead to obtain vertex-wise speeds of the gradient $\bar{s}_i$ as}
+  which finally lead to obtain vertex-wise speeds of the gradient $\bar{s}_i$ as}
   illustrated in \autoref{fig:regseg-method}.
@@
   functions for each level of the multiresolution strategy.
-\revcomment[R\#3-C32]{As a rule of thumb, for a $\delta =$ 1.0 both $\boldsymbol{\alpha}$
-  and $\boldsymbol{\beta}$ will typically be in the range [0.0, 1.0].}
-The implementation details as well as other features (such as the sparse matrix approach
-  to fast interpolation) and the main parameters
-  ($\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, the B-spline grid resolutions,
-  and target image smoothing) are discussed in the \citetalias[][section S1]{esteban_useful_2015}.
+\revcomment[R\#3-C32]{As a rule of thumb, for a $\delta =$ 1.0, both $\boldsymbol{\alpha}$
+  and $\boldsymbol{\beta}$ will typically be in the range [0.0, 1.0].
+The parameters used ($\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, the B-spline grid resolutions,
+  and target image smoothing), the implementation details, and other features such as the sparse matrix
+  approach to fast interpolation} are discussed in the \citetalias[][section S1]{esteban_useful_2015}.
 
@@
 Therefore, the deformation field is nonzero only in the \revcomment[R\#3-C.23]{\acrfull*{pe}} axis,
-  and resembles a real \gls*{epi} distortion.
+  and \revcomment[R\#1-C3]{reproduces a} real \gls*{epi} distortion.
 The adaptation of the evaluation protocol to the simulated phantoms and the real data is
@@
 After generating the ground-truth deformation, the original surfaces
-  were warped by interpolating the displacements field at each vertex.
+  were warped by interpolating the displacement field at each vertex.
 The warped surfaces $\gammaset_\text{true}$ were binarized to generate tissue fractions
   at low (\isores[mm]{2.0}) and high (\isores[mm]{1.0}) resolutions.
-Using a \gls*{mr} simulator \citep{caruyer_phantomas_2014}, we synthesized
+Using an \acrshort*{mri} simulator \citep{caruyer_phantomas_2014}, we synthesized
   \gls*{t1} (TE/TR= 10/1500 ms) and \gls*{t2} images (TE/TR= 90/5000 ms), which
@@
      i.e., \isores[mm]{2.0} and \isores[mm]{1.0}, thereby producing sets of tissue fractions maps.
-  4) The tissue fractions were fed into a \acrfull*{mr} simulator, which generated \acrfull*{t1} and \acrfull*{t2} -like images at the
+  4) The tissue fractions were fed into an \acrshort*{mri} simulator, which generated \acrfull*{t1} and \acrfull*{t2} -like images at the
      two possible resolutions.
@@
 \paragraph*{Ground-truth generation}
-Realistic deformation was achieved by generating displacements fields that satisfy the theoretical
+Realistic deformation was achieved by generating displacement fields that satisfy the theoretical
   properties of distortion.
@@
   $s_\text{PE}$ is the pixel size along \gls*{pe}.
-Certain \gls*{mr} sequences are designed to estimate $\Delta B_0$, thereby obtaining
+Certain \acrshort*{mri} sequences are designed to estimate $\Delta B_0$, thereby obtaining
   the so-called fieldmap.

### 2016-02-03 Publications/2015-NeuroImage/06-discussion.tex
revised paper

@@
   displacement vector field, which preserves the topology even with very large deformations.
-\Regseg{} includes an anisotropic regularizer for the displacements field proposed by
+\Regseg{} includes an anisotropic regularizer for the displacement field proposed by
   \cite{nagel_investigation_1986}.
@@
 \revcomment[R\#1-C.2]{%
-We envision \regseg{} to be integrated in diffusion processing pipelines, after a preliminary
+We envision \regseg{} to be integrated in diffusion processing pipelines, after a preliminar
   \gls*{dti} computation and before anatomically-informed reconstruction and tractography

### 2016-02-03 Publications/2015-NeuroImage/08-appendix-00-shapepriors.tex
revised paper

@@
 \section*{Appendix}
+
+\subsection{Simplifying the regularization term}\label{app:reg_term}
+The exponentials of the Thikonov regularization prior \eqref{eq:regseg-thikonov} have the general form
+  $\vec{v}^T \mathbf{M} \vec{v}$.
+If $\mathbf{M}$ is a $n \times n$ diagonal matrix such that $\mathbf{M} = \vec{m} \, \mathbf{I}_n$,
+  then:
+
+\begin{equation*}
+\vec{v}^T \mathbf{M} \vec{v} = \vec{m} \cdot (\vec{v}^T \mathbf{I}_n \vec{v}) = \vec{m} \cdot \vec{v}^{\circ2},
+\end{equation*}
+  where we have introduced the Hadamard power notation\footnote{The Hadamard power of a matrix or a vector
+  is the power of its elements $\mathbf{M}^{\circ p} = ({m_{ij}}^{p})$}.
+
+In general, the anisotropy \revcomment[R\#3-C.31]{of the distortion field} is aligned with the 
+  \revcomment[R\#3-C.31]{voxel coordinate system}, so
+  $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:regseg-energy} can be simplified to diagonal matrices
+  \revcomment[R\#3-C.31]{to regularize the registration process}, such that
+  $\mathbf{A}= \, \boldsymbol{\alpha}\,\vec{I}_n$ and
+  $\mathbf{B}= \, \boldsymbol{\beta}\,\vec{I}_n$.
+By substituting into equation \eqref{eq:regseg-energy}, we obtain:
+
+  \begin{align}
+  E(\vec{u}) &= \const + \, \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \,
+  \ifthenelse{\boolean{review}}{+}{+ \notag\\ &+}
+  \int_{\Omega} \frac12 \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2} + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}.
+  \label{eq:regseg-app_energy}
+  \end{align}
 
@@
 The equation \eqref{eq:regseg-shape_gradients} is discretized as follows.
-First, the surface between limiting regions $\{l, m\}$, $\Gamma_{l,m}$ is explicitly represented by
+First, the surface between limiting regions $l$ and $m$ ($\Gamma_{l,m}$) is explicitly represented by
   a discrete set of vertices $\vec{v}_i$, with $i \in \{0, \ldots, N_p -1 \}$.
@@
 The resulting summation is, therefore, discrete and the integral operator is replaced by the sum:
+
   \begin{align}
@@
   $\frac{\partial \vec{v}_i}{\partial t} = \vec{v}_i(\delta = t+1) - \vec{v}_i(\delta = t)$:
+
   \begin{equation}
@@
   two terms, $F_{in}(\vec{r})$ corresponding to the interior region and $F_{out}(\vec{r})$ to the exterior:
+
   \begin{equation}
@@
 Then, we identify the shape gradient contribution $\vec{g}_k$ on the coefficients $\vec{u}_k$ of the B-spline grid:
+
   \begin{equation}
@@
   where $\hat{\vec{e}}$ is the coordinates system's unit vector.
-Therefore, the shape gradients projected to the grid of B-spline control points is:
+Therefore, the shape gradients projected to the grid of B-spline control points read:
+
 \begin{equation}
@@
 \end{equation}}
-
-\subsection{Simplifying the regularization term}\label{app:reg_term}
-The exponentials of the Thikonov regularization prior \eqref{eq:regseg-thikonov} have the general form
-  $\vec{v}^T \mathbf{M} \vec{v}$.
-If $\mathbf{M}$ is a $n \times n$ diagonal matrix such that $\mathbf{M} = \vec{m} \, \mathbf{I}_n$,
-  then:
-
-\begin{equation*}
-\vec{v}^T \mathbf{M} \vec{v} = \vec{m} \cdot (\vec{v}^T \mathbf{I}_n \vec{v}) = \vec{m} \cdot \vec{v}^{\circ2},
-\end{equation*}
-  where we have introduced the Hadamard power notation\footnote{The Hadamard power of a matrix or a vector
-  is the power of its elements $\mathbf{M}^{\circ p} = ({m_{ij}}^{p})$}.
-
-In general, the anisotropy \revcomment[R\#3-C.31]{of the distortion field} is aligned with the 
-  \revcomment[R\#3-C.31]{voxel coordinate system}, so
-  $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:regseg-energy} can be simplified to diagonal matrices
-  \revcomment[R\#3-C.31]{to regularize the registration process}, such that
-  $\mathbf{A}= \, \boldsymbol{\alpha}\,\vec{I}_n$ and
-  $\mathbf{B}= \, \boldsymbol{\beta}\,\vec{I}_n$.
-By substituting into equation \eqref{eq:regseg-energy}, we obtain:
-
-  \begin{align}
-  E(\vec{u}) &= \const + \, \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \,
-  \ifthenelse{\boolean{review}}{+}{+ \notag\\ &+}
-  \int_{\Omega} \frac12 \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2} + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}.
-  \label{eq:regseg-app_energy}
-  \end{align}

### 2016-02-03 Publications/2015-NeuroImage/cover_letter2.tex
revised paper

--- 2016-01-16 Publications/2015-NeuroImage/cover_letter.tex
@@
 \documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text
-
-\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
+\usepackage{charter}
+%\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
 %\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font
@@
 \begin{tikzpicture}[remember picture,overlay,yshift=-2cm, xshift=4cm]
-  \node at (0,0) {\includegraphics[height=1.80in]{/home/oesteban/Pictures/logocei.png}};
+  \node at (0,0) {\includegraphics[height=1.80in]{/Users/oesteban/Pictures/logocei.png}};
  \end{tikzpicture}
@@
 E-28040 Madrid (Spain) \\
-Tlf.: +34 915 495 700 ext.4234,\\
-E-mail: oesteban@die.upm.es % Your address and phone number
+Tlf.: +1 (707) 492 61 40,\\
+E-mail: \texttt{phd@oscaresteban.es} % Your address and phone number
 \end{flushright}
@@
   of diffusion MR images}, by O. Esteban et al., which we would like to submit for publication in \emph{NeuroImage}.
-\textbf{This work is a resubmission of the previously rejected paper, NIMG-15-1151}, as suggested in its
-  corresponding decision letter.
-
-For convenience, a point-by-point response for the reviewers of the submission NIMG-15-1151 is attached, and the
-  paper is sent in two versions, one annotated for the reviewers and another without annotations.
+This manuscript \textbf{is a revision of the submission NIMG-15-2877}.
+For convenience, a point-by-point response for the reviewers is attached.
 

### 2016-02-03 Publications/2015-NeuroImage/visual_report.sh

@@ -4,7 +4,8 @@ rm -f suppl-realdata.tex
 
 for i in supplemental/HCP*.pdf; do
     fname=$(basename $i)
-    fname=${fname::-4}
+    fname=${fname:0:9}
+    echo $fname
     echo "\\begin{figure}[!ht]" >> suppl-realdata.tex
     echo "  \\centering" >> suppl-realdata.tex
     echo "  \\textbf{$fname}" >> suppl-realdata.tex

### 2016-03-02 Publications/2015-NeuroImage/04-methods.tex
version submitted to NI

@@
   \includegraphics[width=\linewidth]{figure01}
-  \caption{The {\color{blue} interfacing surfaces $\Gamma_m$ between the competing \glspl{roi} $\Omega_l$,
+  \caption{The {\color{revcolor} interfacing surfaces $\Gamma_m$ between the competing \glspl{roi} $\Omega_l$,
   play the role of active contours which drive the registration process.}
-  They evolve iteratively {\color{blue} along the normal $\hat{\mathbf{n}}_i$ of the surface} at each vertex
+  They evolve iteratively {\color{revcolor} along the normal $\hat{\mathbf{n}}_i$ of the surface} at each vertex
     $\vec{v}_i$ of the mesh.
@@
     $M$ in the location of vertex $\vec{v}_i$.
-  {\color{blue} The computation of shape-gradients is developed in \ref{app:shape_gradients}.}
+  {\color{revcolor} The computation of shape-gradients is developed in \ref{app:shape_gradients}.}
   In this figure, the $\bar{s}_1$ derived from \autoref{eq:regseg-shape_gradient_disc2} is 
@@
 \vskip-2.0em
-\includegraphics[width=\linewidth]{legend}%x
+\includegraphics[width=\linewidth]{legend}%
 \end{minipage}
@@
   for one real dataset.
-{\color{blue}%
+{\color{revcolor}%
 Panel (A, left) shows the joint distribution of the \gls*{fa} and \gls*{adc} conditioned to

### 2016-03-02 Publications/2015-NeuroImage/05-results.tex
version submitted to NI

@@
   \includestandalone[width=\linewidth]{figure05}
-  \caption{A. Visual assessment of the results obtained on the {\color{blue} digital phantoms.
+  \caption{A. Visual assessment of the results obtained on the {\color{revcolor} digital phantoms.
     The panel shows three coronal slices (indices indicated on the lower left corner) of each phantom
@@
     clustering methods, but they were successfully segmented with \regseg{}.
-  B. Quantitative evaluation: {\color{blue} the violin plot shows the variability across experiments
+  B. Quantitative evaluation: {\color{revcolor} the violin plot shows the variability across experiments
     of the average Hausdorff distance measured in each vertex of the corresponding surface, for the}
     low (left) and high (right) resolutions. 
-    {\color{blue} Error averages were} consistently below the size of the voxel.
+    {\color{revcolor} Error averages were} consistently below the size of the voxel.
     }\label{fig:regseg-phantom}
@@
   Specific regions where \regseg{} outperformed \gls*{t2b} are enlarged.
-  B. Violin plots of the error distributions for each surface {\color{blue} across the 16 subjects}, which show the 
+  B. Violin plots of the error distributions for each surface {\color{revcolor} across the 16 subjects}, which show the 
     voxel size of the \gls*{dmri} images (1.25 mm), thereby supporting the improved results obtained
@@
     \end{tabular}
-    \caption{Statistical analysis of results obtained using {\color{blue} 16 real datasets from 
+    \caption{Statistical analysis of results obtained using {\color{revcolor} 16 real datasets from 
     the \gls*{hcp}}, which show that \regseg{} performed better than the alternative \acrfull{t2b} method.

### 2016-03-02 Publications/2015-NeuroImage/02-abstract.tex
remove comments from text

--- 2016-01-17 Publications/2015-NeuroImage/02-abstract.tex
@@
   require precise delineations of anatomical structures.
-This requirement has been approached \revcomment[R\#3-C.33]{by} either segmenting the data in
+This requirement has been approached by either segmenting the data in
   native \gls*{dmri} space or mapping the structural information from \gls*{t1} images.
@@
   \gls*{dmri} dataset.
-We first verify the accuracy of \regseg{} on a general context using digital phantoms.
+We first verify the accuracy of \regseg{} on a general context using digital phantoms 
+  distorted with synthetic and random deformations.
 Then we establish an evaluation framework using undistorted \gls*{dmri} data from the \gls*{hcp}
-  and known deformations derived from real inhomogeneity fieldmaps.
+  and realistic deformations derived from 
+  the inhomogeneity fieldmap corresponding to each subject.
 We analyze the performance of \regseg{} computing the misregistration error of the surfaces estimated
@@
   method, thereby obtaining a significantly lower misregistration error with \regseg{}.
-Therefore, we demonstrate that \regseg{} allows the accurate mapping of structural information
-  in \gls*{dmri} space, enabling the application of new structure-informed techniques in
-  the connectome extraction.
+The accurate mapping of structural information in \gls*{dmri} space 
+  is fundamental to increase the reliability of network building
+  in connectivity analyses, and to improve the performance of the emerging structure-informed
+  techniques for \gls*{dmri} data processing.

### 2016-03-02 Publications/2015-NeuroImage/03-introduction.tex
remove comments from text

@@
   models, which range from the early \gls*{dti} proposed by \cite{basser_microstructural_1996}
-  to current models such as \revcomment[R\#3-C.10]{\gls*{amico}}.
+  to current models such as \gls*{amico}.
 The microstructural map is then used to draw the preferential orientations of diffusion
@@
 However, the mask and subsequent analyses are highly dependent on the 
-  \revcomment[R\#3-C.33]{threshold that is chosen} \citep{taoka_fractional_2009}.
+  threshold that is chosen \citep{taoka_fractional_2009}.
 To overcome the unreliability of \gls*{fa} thresholding, and to broaden
@@
   target-image.
-\revcomment[authors]{We apply \regseg{} to segment \gls*{dmri} data by mapping
-  a} set of nested surfaces extracted from a structural image (e.g. \gls*{t1}) to
+We apply \regseg{} to segment \gls*{dmri} data by mapping
+  a set of nested surfaces extracted from a structural image (e.g. \gls*{t1}) to
   a bivariate target-volume comprehending the \gls*{fa} and \gls*{adc} maps.
@@
     therefore \regseg{} does not require additional MR acquisitions to the minimal
-    \gls*{dmri} protocol \revcomment[R\#1-C.2]{in order to estimate the deformation field};
-  2) \revcomment[R\#1-C.2]{depending on the application, the resampling of the \glspl*{dwi}
+    \gls*{dmri} protocol in order to estimate the deformation field;
+  2) depending on the application, the resampling of the \glspl*{dwi}
     introduced by the correction method can be avoided, either by performing the posterior
     processing on the native \gls*{dmri} space or applying the unwarping on the tractography
-    itself}; and
+    itself; and
   3) \regseg{} increases the geometrical accuracy of the overall process.
@@
   the corresponding \gls*{t2} image of the same subject.
-\revcomment[authors]{This approach is the first step of the} \gls*{t2b} correction
+This approach is the first step of the \gls*{t2b} correction
   methods \citep{kybic_unwarping_2000}.
-\revcomment[authors]{We reproduce the settings and implementation of a} widely used diffusion
+We reproduce the settings and implementation of a widely used diffusion
   processing software \citep[\emph{ExploreDTI},][]{leemans_exploredti_2009}.

### 2016-03-02 Publications/2015-NeuroImage/04-methods.tex
remove comments from text

@@
   distributions 
-\revcomment[R\#3-C.11]{%
-  $\mathcal{N} (\vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} )$}, with parameters
+  $\mathcal{N} (\vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} )$, with parameters
   $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
@@
 
-\revcomment[R\#3-C.15]{%
   \begin{align}
@@
   \end{align}%
-  }%
   using $\mdist{f'}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}'$ with respect
@@
 $C$ is the number of channels comprised in the image $M$.
-\todo{ordering} Even though the features being segmented are not generally i.i.d., the 
+Even though the features being segmented are not generally i.i.d., the 
   spatial interdependency of voxels is implicitly supported by the piecewise smooth
@@
   the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized.
-The \autoref{fig:regseg-model} shows \revcomment[authors]{how the joint distribution of 
-  the input images} is approximated with a mixture of multivariate normal distributions,
+The \autoref{fig:regseg-model} shows how the joint distribution of 
+  the input images is approximated with a mixture of multivariate normal distributions,
   and this minimization is illustrated for the segmentation of the \gls*{fa} and the \gls*{adc} maps
@@
   \includegraphics[width=\linewidth]{figure01}
-  \caption{The {\color{revcolor} interfacing surfaces $\Gamma_m$ between the competing \glspl{roi} $\Omega_l$,
-  play the role of active contours which drive the registration process.}
-  They evolve iteratively {\color{revcolor} along the normal $\hat{\mathbf{n}}_i$ of the surface} at each vertex
+  \caption{The interfacing surfaces $\Gamma_m$ between the competing \glspl{roi} $\Omega_l$,
+  play the role of active contours which drive the registration process.
+  They evolve iteratively along the normal $\hat{\mathbf{n}}_i$ of the surface at each vertex
     $\vec{v}_i$ of the mesh.
@@
     $M$ in the location of vertex $\vec{v}_i$.
-  {\color{revcolor} The computation of shape-gradients is developed in \ref{app:shape_gradients}.}
+  The computation of shape-gradients is developed in \ref{app:shape_gradients}.
   In this figure, the $\bar{s}_1$ derived from \autoref{eq:regseg-shape_gradient_disc2} is 
@@
   for one real dataset.
-{\color{revcolor}%
 Panel (A, left) shows the joint distribution of the \gls*{fa} and \gls*{adc} conditioned to
@@
 Particularly, the \gls*{wm} typically shows a bimodal distribution when the contours
-  \gammaset{} do not fit the data.}
+  \gammaset{} do not fit the data.
 % The left plot in each panel represents a kernel density estimate of the location and spread of
@@
 \paragraph*{Regularization}
-The smoothness of the resulting \revcomment[R\#3-C.33]{displacement field} is induced
+The smoothness of the resulting displacement field is induced
   by a Thikonov regularization prior:
@@
   \end{align}
- which requires that the distortion and its gradient have zero mean\revcomment[R\#1-C4]{,}
+ which requires that the distortion and its gradient have zero mean,
    and variance governed by the matrices $\mathbf{A}$ and $\mathbf{B}$.
-\revcomment[R\#3-C.16]{%
 Therefore, $\mathbf{A}$ and $\mathbf{B}$ are tensors that modulate the regularization, and
-  produce deformations with preferential directions}.
+  produce deformations with preferential directions.
 Finally, the maximum a posteriori problem is adapted to a variational problem where we search for
@@
 
-\revcomment[R\#3\\C.17--C.19]{%
   \begin{align}
@@
   \end{align}%
-}
 This expression is the dual of the Mumford-Shah functional that corresponds
@@
   $\hat{U}$ of the transformation at the current time point.
-Let us assume that \revcomment[R\#1-C.5\\R\#3-C.31]{the preferential directions of
-  the displacement are aligned} with the imaging axes to simplify \eqref{eq:regseg-energy} as expression 
+Let us assume that the preferential directions of
+  the displacement are aligned with the imaging axes to simplify \eqref{eq:regseg-energy} as expression 
   \eqref{eq:regseg-app_energy} in \ref{app:reg_term}, and thus to compute its
@@
 
-\revcomment[R\#3-C.20]{%
   \begin{align}
@@
   \label{eq:regseg-gradient_descent}
-  \end{align}}
-  where $\vec{u}^{\circ2} = \vec{u}^T \cdot \vec{u}$, \revcomment[R\#3-C.21]{%
+  \end{align}
+  where $\vec{u}^{\circ2} = \vec{u}^T \cdot \vec{u}$,
   and $\{\boldsymbol{\alpha}, \boldsymbol{\beta}\}$ are the expected variances along
-  the imaging axes of the displacement field and its gradient, respectively.}
+  the imaging axes of the displacement field and its gradient, respectively.
 Then, the data and regularization terms are split and discretized to compute their
   derivatives.
-\revcomment[R\#3-C.14]{%
 The derivative of the data term is computed using explicit shape gradients (see \ref{app:shape_gradients}),
-  which finally lead to obtain vertex-wise speeds of the gradient $\bar{s}_i$ as}
+  which finally lead to obtain vertex-wise speeds of the gradient $\bar{s}_i$ as
   illustrated in \autoref{fig:regseg-method}.
-\revcomment[R\#3-C.14]{%
 Using the expressions \eqref{eq:regseg-gradient_wshape} and 
@@
   and introducing the analytical derivative of the regularization term, then
-  the \autoref{eq:regseg-gradient_descent} is reformulated as:}
+  the \autoref{eq:regseg-gradient_descent} is reformulated as:
 
@@
   \end{align}
-\revcomment[R\#3-C22]{%
-where $\vec{g}_k$ are the shape-gradient contributions on the coefficients $\vec{u}_k$ of the B-spline grid.}
+where $\vec{g}_k$ are the shape-gradient contributions on the coefficients $\vec{u}_k$ of the B-spline grid.
 Finally, to descend this gradient, we establish a semi-implicit Euler scheme
@@
   functions for each level of the multiresolution strategy.
-\revcomment[R\#3-C32]{As a rule of thumb, for a $\delta =$ 1.0, both $\boldsymbol{\alpha}$
+As a rule of thumb, for a $\delta =$ 1.0, both $\boldsymbol{\alpha}$
   and $\boldsymbol{\beta}$ will typically be in the range [0.0, 1.0].
@@
   and target image smoothing), the implementation details, and other features such as the sparse matrix
-  approach to fast interpolation} are discussed in the \citetalias[][section S1]{esteban_useful_2015}.
+  approach to fast interpolation are discussed in the \citetalias[][section S1]{esteban_useful_2015}.
 
@@
   the subject.
-Therefore, the deformation field is nonzero only in the \revcomment[R\#3-C.23]{\acrfull*{pe}} axis,
-  and \revcomment[R\#1-C3]{reproduces a} real \gls*{epi} distortion.
+Therefore, the deformation field is nonzero only in the \acrfull*{pe} axis,
+  and reproduces a real \gls*{epi} distortion.
 The adaptation of the evaluation protocol to the simulated phantoms and the real data is
@@
   where $\vec{v}_i \subset \gammaset_\text{true}$ are the locations of the total $N_V$ vertices,
-  \revcomment[R\#3-C.24]{$a_i$ is the area corresponding to each vertex $\vec{v}_i$},
+  $a_i$ is the area corresponding to each vertex $\vec{v}_i$,
   and $\hat{\vec{v}}_i \subset \hat{\gammaset}_\text{test}$ are the recovered locations

### 2016-03-02 Publications/2015-NeuroImage/05-results.tex
remove comments from text

--- 2016-01-24 Publications/2015-NeuroImage/05-results.tex
@@
   \includestandalone[width=\linewidth]{figure05}
-  \caption{A. Visual assessment of the results obtained on the {\color{blue} digital phantoms.
-    The panel shows three coronal slices corresponding to the small numbers annotated on the lower
-      left corner of each cut, for} the low resolution sets:
+  \caption{A. Visual assessment of the results obtained on the digital phantoms.
+    The panel shows three coronal slices (indices indicated on the lower left corner) of each phantom
+    volume at low resolution:
     ``gyrus'' (top left), ``L'' (top right), ``ball'' (bottom left),
@@
     clustering methods, but they were successfully segmented with \regseg{}.
-  B. Quantitative evaluation of registration errors in terms of the average Hausdorff distances between
-    surfaces at low (left) and high (right) resolutions, which demonstrate that the errors were
-    consistently below the size of the voxels.
+  B. Quantitative evaluation: the violin plot shows the variability across experiments
+    of the average Hausdorff distance measured in each vertex of the corresponding surface, for the
+    low (left) and high (right) resolutions. 
+    Error averages were consistently below the size of the voxel.
     }\label{fig:regseg-phantom}
@@
   Specific regions where \regseg{} outperformed \gls*{t2b} are enlarged.
-  B. Violin plots of the error distributions for each surface {\color{blue} across the 16 subjects}, which show the 
+  B. Violin plots of the error distributions for each surface across the 16 subjects, which show the 
     voxel size of the \gls*{dmri} images (1.25 mm), thereby supporting the improved results obtained
@@
 We also show the 95\% CIs of the \gls*{swindex} for these surfaces 
-  \revcomment[R\#3-C.30]{(\autoref{tab:results_real})}.
+  (\autoref{tab:results_real}).
 The aggregate CI for \regseg{} was 0.56--0.66 [mm], whereas the \gls*{t2b} method
@@
     \end{tabular}
-    \caption{Statistical analysis of results obtained using {\color{blue} 16 real datasets from 
-    the \gls*{hcp}}, which show that \regseg{} performed better than the alternative \acrfull{t2b} method.
+    \caption{Statistical analysis of results obtained using 16 real datasets from 
+    the \gls*{hcp}, which show that \regseg{} performed better than the alternative \acrfull{t2b} method.
     The distribution of the errors computed for the surfaces of interest ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$)

### 2016-03-02 Publications/2015-NeuroImage/06-discussion.tex
remove comments from text

@@
   in \gls*{dmri} images is an active field of research \citep{jeurissen_tissuetype_2015}.
-\revcomment[R\#3-C.9]{%
 Current processing of \gls{dmri} involved in the connectome extraction and other applications
-  (such as \gls*{tbss} or surgical planning)} require a precise segmentation
+  (such as \gls*{tbss} or surgical planning) require a precise segmentation
   of the anatomical structures in the diffusion space.
@@
 This situation is found commonly in historical datasets.
-\revcomment[R\#1-C.2]{%
 We envision \regseg{} to be integrated in diffusion processing pipelines, after a preliminar
@@
   these two processes and the matrix building task can be performed on the unaltered
-  \gls*{dmri} signal (i.e. without resampling data to an undistorted space).}
-\revcomment[R\#3-C.9]{%
+  \gls*{dmri} signal (i.e. without resampling data to an undistorted space).
 For analyses other than connectivity, like \gls*{tbss}, the deformation estimated by \regseg{}
-  can be used to map the tracts into structural space.}
+  can be used to map the tracts into structural space.
 
@@
   of the brain using multispectral images, for instance in the case of traumatic brain
-  injury \citep{irimia_structural_2014} or in \revcomment[R\#3-C.33]{monitoring} brain tumors
+  injury \citep{irimia_structural_2014} or in monitoring brain tumors
   \citep{weizman_semiautomatic_2014}.

### 2016-03-02 Publications/2015-NeuroImage/08-appendix-00-shapepriors.tex
version submitted to NI
remove comments from text

@@
 % -*- root: 00-main.tex -*-
+
 \renewcommand{\theequation}{A.\arabic{equation}}
@@
 
-In general, the anisotropy \revcomment[R\#3-C.31]{of the distortion field} is aligned with the 
-  \revcomment[R\#3-C.31]{voxel coordinate system}, so
+In general, the anisotropy of the distortion field is aligned with the 
+  voxel coordinate system, so
   $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:regseg-energy} can be simplified to diagonal matrices
-  \revcomment[R\#3-C.31]{to regularize the registration process}, such that
+  to regularize the registration process, such that
   $\mathbf{A}= \, \boldsymbol{\alpha}\,\vec{I}_n$ and
@@
 \subsection{Application of the shape-gradients}\label{app:shape_gradients}
-\revcomment[R\#3-C.8]{%
 The computation of gradients at the locations of the active contours in the
@@
   \label{eq:regseg-shape_gradient_disc2}
-  \end{equation}}
+  \end{equation}
 
-\revcomment[R\#3-C.22]{%
 Then, we identify the shape gradient contribution $\vec{g}_k$ on the coefficients $\vec{u}_k$ of the B-spline grid:
@@
   \label{eq:regseg-shape_gradient_final}
-\end{equation}}
+\end{equation}

### 2016-03-02 Publications/2015-NeuroImage/responseLetter3.tex
getting submission ready

@@
 \makeatletter
+\newenvironment{reviewintro}%
+{\begingroup%
+ \color{black!60}
+ \fontshape{it}\selectfont %
+}
+{\endgroup\par\medskip}
+
 \newenvironment{reviewpoint}%
@@
 
-\section*{Response letter to the manuscript no. NIMG-15-1151}
-\medskip
-Original title: \emph{Active contours-driven registration method for the structure-informed segmentation of
-  diffusion MR images}.
-
-\noindent New title: \emph{{\color{blue!70} Surface}-driven registration method for the structure-informed segmentation of
-  diffusion MR images}.
+\section*{Response letter to the manuscript no. NIMG-15-2877R1}
 
 \bigskip
-\noindent We thank the editors and the reviewers for revising our manuscript no. NIMG-15-1151 submitted to NeuroImage.
+\noindent We thank the editors and the reviewers for revising our manuscript no. NIMG-15-2877R1 submitted to NeuroImage.
 We appreciate the comments made by both reviewers, and considered them fully to improve our paper.
@@
 \subsection*{Reviewer \#1:}
+\begin{reviewintro}
+In general, I think the authors have done a good job of restructuring the paper to qualify the goals and implementation of the method. I have a few points.
+\end{reviewintro}
 \begin{reviewpoint}
-This paper presents a method for simultaneous registration and segmentation of diffusion images in the presence of geometric distortions.
-The method uses an anisotropically regularised active contour method to co­align tissue segmentations generated from T1 images to distorted multi­channel diffusion data.
-It has the advantage that it has no requirements for extra scans to be acquired for correction of the distortion.
-The method demonstrates improved performance over a previous b0­T2 registration approach for the alignment of tissue segmentation maps.
-Improvements are demonstrated on simulated data, and real data with simulated distortions.
-The advantage of these results is that the ground truth is known.
-I am satisfied that the active contour alignment approach used to align the surfaces is sound, and that the results show theoretic improvements over the T2B method.
-I am convinced that the method approach is effective for the alignment task.
-In general the methods are well explained with the exception of a few minor points (below).
-My main concerns lie with whether the results in the paper prove what the authors set out in their abstract.
-Specifically, a stated goal is for the use of anatomically constrained tractography.
-However, there are no results that show this method would improve connectome extraction.
-In particular, the authors state that ``this method avoids any requirement for collecting additional scans nor resampling for diffusion unwarping''.
-I see this has advantages, but what about the consequences of not unwarping on the diffusion tracts themselves see [1][2][3].
-In my opinion the authors should address this point possibly with additional results to show how they intend this to be used within a tracking framework.
-I also have some concern that the method is not applied to real distortions nor compared to state­of­the­art correction methods such as EDDY [4].
-This was used for the HCP corrections therefore it might provide a good basis comparison basis on real rather than modeled distortions for brain data.
-If this could not be done for some reason I feel it should be discussed explicitly.
-It should be possible to compare to bbregister for the unwarped data.
-Why is this not done?
-[1] Jones, Derek K., and Mara Cercignani. ``Twenty‐five pitfalls in the analysis of diffusion MRI data.'' NMR in Biomedicine 23.7 (2010): 803­820.
-[2] Lee, J., et al. ``Correction of Bo EPI distortions in diffusion tensor imaging and white matter tractography.'' Proc. Int. Soc. Magn. Reson. Med. Vol. 11. 2004.
-[3] Andersson, J. L., et al. ``Effects of susceptibility distortions on tractography.''Proceedings of International Society of Magnetic Resonance in Medicine. Vol. 11. 2004.
-[4] A comprehensive Gaussian Process framework for correcting distortions and movements in diffusion images Jesper L. R. Andersson et al ISMRM
-Minor points:
-Pg 5 regularisation: `` requires that the distortion and its gradient have zero mean and variance governed by the matrices A and B'' suggest adding a comma here for easier interpretation i.e.
-``requires that the distortion and its gradient have zero mean, and variance governed by the matrices A and B''
-Pg 7, line 4 `` anisotropy is aligned''. What is meant by anisotropy in this context, the anisotropy of the deformation gradient.
-Figure 2.
-This looks nice, but its not easy to interpret.
-I can't see a clear improvement through A­C.
-I think there needs to be direct comparisons between the conditions represented by A­C on one plot.
-Why not have 6 plots for each tissue class comparing distributions for each condition.
+The authors suggest that diffusion tractography is typically performed by mapping segmentations to the data.
+This is becoming less true.
+The HCP for one performs whole brain connectivity analysis.
+Nevertheless, these frameworks still require registration to structural space for projection to surface mesh representations and/or coaligning data.
+This may be worth mentioning or clarifying.
 \end{reviewpoint}
 \reply{%
+We have introduced this discussion in lines 351--355.
+}
+\begin{reviewpoint}
+I still find figure 2 really difficult to read anything from.
+It's not easy to read the labels in the coloured text, nor see the different histograms in figure 2 A.
+This is partly because many of the colours are light very similar i.e.
+for WM and bst.
+It does occur to me that not all distributions change that greatly with registration.
+Maybe stripping back figure A to only include WM, GM and possibly VdGM might help?
+\end{reviewpoint}
+\reply{%
+We have made editions in this figure to make it more clear.
+}
+\begin{reviewpoint}
+I find the clarification of how tratography might be unwarped in the response very helpful, but some clarification of the manuscript might be helpful.
+What do the authors mean by ``the resampling of the DWIs introduced by the correction method can be avoided, either by performing the posterior processing on the native dMRI space''.
+How does the native dMRI space differ from the DWIs here, or what is meant by posterior processing.
+\end{reviewpoint}
+\reply{%
+We have reformulated that particular sentence (lines 57--59) and added a corresponding discussion in lines 348--356.
+These editions are highly related to Comments \ref{com:1}, \ref{com:5} and \ref{com:6}.
+}
+\begin{reviewpoint}
+In the appendix I think the B-spline variable (Eqs A.7 A.8)  should be defined in the text again
+\end{reviewpoint}
+\reply{%
+We have defined introduced the definition of $\mathbf{g}_k$ in lines 122--123, right before the
+  reference to the mentioned equations.
+}
 
-}
 % ------------------------------------------------------------------- Reviewer 3
@@
 \subsection*{Reviewer \#3:}
+
+\begin{reviewintro}
+This manuscript is much improved, in my opinion.
+However, there are some high-level clarifications that are present in the author response letter, but that are not clear to the reader in the manuscript.
+I also noticed a few small things to fix.
+My suggestions follow.
+\end{reviewintro}
+
 \begin{reviewpoint}
-This paper presents a new method for correction of EPI distortion in diffusion MRI data by performing simultaneous registration and segmentation of a set of nested surfaces, using a multi­channel FA/MD image with a registration target of an undistorted T1 image.
-The primary advantage over other methods is that cortical surfaces and white­gray matter boundaries are apparently aligned well by this method.
-Results show that it outperforms one implementation of the more standard baseline to T2 registration.
-Overall this is a quite interesting work, and it will be a good addition to the literature and also to the body of open­source code for diffusion MRI.
-I have some comments for clarification of the paper.
-Overall, I am slightly confused by some aspects of the equations, and I think that some information from the supplement (which I have not had time to read) should be moved into the paper to clarify missing points.
-The paper should be able to be read independently of the supplement, which does not seem to be the case currently.
-Specifically, the relationship of the b­spline registration framework to the evolution of the active contours is a bit confusing, with a main equation given inside Figure 1, and apparently all other detail in the supplement.
-This could be as simple as an additional paragraph or two.
-Please state that the goal of the paper is to improve connectome­based diffusion MRI research.
-The processing description on pages 1­2 assumes that the connectome­based approach is the only approach to analysis of diffusion MRI.
-But this is clearly not the case (surgical planning, TBSS, etc).
-``to current models such as (Dadicci et al)'' is confusing. Please give the name(s) of the current model(s).
-￼￼On page 3, please define N (section 2.1)
-In Figure 1, it is a bit confusing that active contours appear, because the abstract did not mention active contours directly, and the evolution following the inward normals at each vertex is not explicit in the equations of the paper.
-It seems the evolution at $s_1$ is in the direction opposite the inward normal, so the caption of Figure 1 is a bit confusing, since it directly states the inward normals are followed, rather than just the normal to the curve.
-I recommend adding a little bit of text to unify figure 1 and the surface evolution framework with the concepts presented in the registration equations section.
-I think this information may be in the supplement.
-On page 5, equation (4), I think f should be f prime on the first line.
-Please describe A and B in a sentence after equation 5. It is not clear how these were chosen (or at least refer to the appendix).
-On page 5, equation (7), the ­log is removed from the second line.
-I would expect a sum of log probabilities, with the log preserved, rather than removing the log altogether.
-It seems this is an oversight.
-Or do the authors mean to use a proportional symbol instead of an equals symbol?
-Please clarify.
-In equation (7), the equation that is split onto two lines (why?) has two plus symbols in a row, one on each line.
-In equation (7), the second integral is over omega, while the first is over $omega_l$.
-Why?
-In equation (7) please clarify with brackets what the sum over l is applied to.
-The two pluses are present again in equation (10).
-Please explain in a sentence what alpha and beta are in equation (10) (expected variances along each axis of the coordinate system, or something).
-$g_k$ is completely undefined in equation (11). I see there is information in a supplement but at a minimum each variable should be defined in a sentence in the main paper.
-pg 7. Please define PE acronym or write out phase encoding.
-eq (14). Please define what is $a_i$.
-Figure 5. What are the little numbers (26, 30, 33) in the lower left corner of the phantom images? Figure 5. Please state clearly in the caption that this is digital phantom data, to help the reader.
-Figure 5.
-Over what is the average error of surfaces computed?
-Average over the surface?
-Then the variability in the plot is across experiments?
-Or the other way around with the average at each surface point across experiments??
-It is not clear.
-Figure 5 A. Why are there so many phantom images shown? They all look the same to me. Please explain and/or remove the duplicates.
-Figure 6. Are the plots in B from one or many subjects? Please state in the caption.
-In Table 2 caption how many subjects were input to this experiment? This is given as 16 in the abstract, but that info is not very near the actual results and would increase their impact.
-pg 13.
-``We also show the 95\% CIs of the sWI for these surfaces.'' sounds like it is referring to Figure 6.
-However it is apparently in Table 2, which should be referenced in this sentence.
-Or the sentences about Table 2 should be reorganized to sound separate from Figure 6.
-in Appendix (and also elsewhere): ``the anisotropy is aligned with the imaging axes'' is quite confusing.
-Usually anisotropy refers to FA or similar.
-I believe this is the anisotropy of the covariance matrices, really that their eigenvectors are assumed aligned with the voxel coordinate system.
-Please clarify the usage of ``anisotropy'' in the registration framework.
-Close
-What values are used for alpha and beta in the appendix?
-minor grammatical issues:
-abstract: ``approached either segmenting'' ­> ``approached by either segmenting'' pg 2: ``threshold set'' sounds like a mathematical set ­> ``threshold that is chosen'' pg 5.
-``displacements field'' ­> ``displacement field''
-pg. 15 ``monitorization'' is not a word ­> monitoring
-
+Please clarify that the authors do not consider this to be an EPI distortion correction method, but rather a surface alignment method or similar.
+This was not clear to me from reading the paper the first time.
+I assumed EPI distortion correction was a primary goal, but it seems rather that the primary goal is to accurately segment the distorted diffusion images.
+In fact, the distortions are not corrected in the EPI data.
 \end{reviewpoint}
 \reply{%
+We have explicitly stated this in a new sentence (lines 348--349).
+}
+\begin{reviewpoint}
+Please discuss the fact that when this method is applied, the tractography is done using data that have not been corrected for EPI distortions.
+\end{reviewpoint}
+\reply{%
+Right after the previous edition, in lines 349--350 we introduce the proposal of this methodological variation.
+The discussion and proposal of a future research line is then done in lines 355--356, after some further details added
+  to address the \autoref{com:1}.
+}
+\begin{reviewpoint}
+For the reader, the overall comparison between the two methods is confusing in the introduction, where it is proposed to register surfaces to diffusion image space, yet the comparison is b0 to T2.
+A more appropriate way to present this comparison is t2 to b0, so that the result is also in the diffusion image space, if this is what was done.
+This is in fact a common processing step, so I assume this is how the experiments were performed.
+Please clarify.
+\end{reviewpoint}
+\reply{%
+}
+\begin{reviewpoint}
+In Figure 6 A, is this output coordinate system the diffusion coordinate system as expected from the description of the method? So then the output in T2B is T2 to b0?
+\end{reviewpoint}
+\reply{%
+}
+\begin{reviewpoint}
+The ``undistorted'' HCP data is not fully defined in the ``Real datasets'' section.
+Because this undistortion method serves as the ground truth for the presented experiments, it should be explained briefly how the undistortion was done for the HCP data.
+One sentence is sufficient, and will allow the reader better understand the results of this paper.
+\end{reviewpoint}
+\reply{%
+We have introduced this description in lines 183--187, along with including the detail of the actual HCP release of the data used in l. 177.
+}
+\begin{reviewpoint}
+Minor points:
+
+The graphical abstract says subpixel accuracy. I think you mean subvoxel (there are no pixels really in 3D MRI images).
+
+``preliminar DTI'' -> preliminary
+\end{reviewpoint}
+\reply{%
+We have updated the graphical abstract with the term ``subvoxel''.
+We have replaced the six appearances of the term ``pixel'' by ``voxel'', including the ``subpixel'' present in line 61.
+We also have fixed the misspelling of ``preliminar'' of line 339.
 }

### 2016-03-02 Publications/2015-NeuroImage/03-introduction.tex
getting submission ready

@@
     \gls*{dmri} protocol in order to estimate the deformation field;
-  2) depending on the application, the resampling of the \glspl*{dwi}
-    introduced by the correction method can be avoided, either by performing the posterior
-    processing on the native \gls*{dmri} space or applying the unwarping on the tractography
-    itself; and
+  2) \revcomment[R\#1-C2]{%
+    alternatively to the typical design of the processing flow, the information from the
+    reference \gls*{t1} can precisely be mapped onto the distorted \gls*{dmri} space,
+    avoiding the interpolation of the \glspl*{dwi} required by unwarping the diffusion data; and} 
+%    depending on the application, the resampling of the \glspl*{dwi}
+%    introduced by the correction method can be avoided, either by performing the posterior
+%    processing on the native \gls*{dmri} space or applying the unwarping on the tractography
+%    itself; and
   3) \regseg{} increases the geometrical accuracy of the overall process.
 In this paper, we first verify the functionality of the method and the \regseg{}
-  implementation using a set of digital phantoms, demonstrating the sub-pixel accuracy
-  in registration.
+  implementation using a set of digital phantoms, demonstrating the
+  \revcomment[R\#3-C10]{subvoxel} accuracy in registration.
 Then, we evaluate \regseg{} on real \gls*{dmri} datasets, using a derivation of our

### 2016-03-02 Publications/2015-NeuroImage/04-methods.tex
getting submission ready

@@
   and it aligns $\gammaset_R$ into $M$.
-First, we assume independence between pixels, and thus we break down the
-  global data likelihood into a product of pixel-wise conditional probabilities:
+First, we assume independence between \revcomment[R\#3-C10]{voxel}s, and thus we break down the
+  global data likelihood into a product of \revcomment[R\#3-C10]{voxel}-wise conditional probabilities:
 
@@
   illustrated in \autoref{fig:regseg-method}.
-Using the expressions \eqref{eq:regseg-gradient_wshape} and 
-  \eqref{eq:regseg-shape_gradient_final} given in \ref{app:shape_gradients}
-  and introducing the analytical derivative of the regularization term, then
+\revcomment[R\#1-C4]{%
+The shape gradient contributions $\vec{g}_k$ on the field coefficients $\vec{u}_k$
+  can then be computed }%
+using the expressions \eqref{eq:regseg-gradient_wshape} and 
+  \eqref{eq:regseg-shape_gradient_final} given in \ref{app:shape_gradients}.
+Finally, introducing the analytical derivative of the regularization term,
   the \autoref{eq:regseg-gradient_descent} is reformulated as:
@@
 To evaluate \regseg{} using real \gls*{dmri} data obtained from human brains,
-  we collected 16 subjects from the \gls*{hcp} database.
+  we collected 16 subjects from the \revcomment[R\#3-C9]{Q3 Release of the}
+  \gls*{hcp} database.
 The original acquisitions are released within ``unprocessed'' packages, whereas
@@
   \gls*{t1} space.
+\revcomment[R\#3-C9]{In the Q3 Release of the \gls*{hcp}, the \gls*{dmri} session includes
+  six runs, two runs for each of three different gradient tables, and each table is acquired
+  once with right-to-left and left-to-right encoding polarities, respectively.
+Then, the diffusion datasets with opposed polarities are corrected for susceptibility
+  distortions using the \emph{TOPUP} tool \citep{andersson_how_2003}
+  before producing the released ``minimally preprocessed'' data.}.
 
@@
 where $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time, and
-  $s_\text{PE}$ is the pixel size along \gls*{pe}.
+  $s_\text{PE}$ is the \revcomment[R\#3-C10]{voxel} size along \gls*{pe}.
 Certain \acrshort*{mri} sequences are designed to estimate $\Delta B_0$, thereby obtaining
@@
   available\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/t2b_elastix_y.txt}}.
+\revcomment[R\#3-C7\\R\#3-C7]{%
+In this registration scheme, the \gls*{t2} image is the \emph{reference} and
+  the \lowb{} plays the role of \emph{moving} image.
+Therefore, the transform is defined in the coordinate system of the \gls*{t2} image,
+  and for each point in this space provides the location of the corresponding feature
+  in the \lowb{} image.
+Since the surfaces are defined in the \gls*{t2} --\emph{reference}-- space, their coordinates
+  can be mapped to the \lowb{} space using this transform, obtaining the 
+  \emph{distorted surfaces} corresponding to the \lowb{}-to-\gls*{t2} registration.}
 

### 2016-03-02 Publications/2015-NeuroImage/06-discussion.tex
getting submission ready

@@
 As an advantage, the use of explicit surfaces enables segmenting \gls*{dmri} images
-  with accuracy below pixel size.
+  with accuracy below \revcomment[R\#3-C10]{voxel} size.
 
@@
 \Regseg{} obtained a high accuracy, with an aggregate 95\% \gls*{ci} of 0.56--0.66 [mm], which was
-  below the pixel size of 1.25 mm.
+  below the \revcomment[R\#3-C10]{voxel} size of 1.25 mm.
 The misregistration error that remained after \regseg{} was significantly lower ($p <$ 0.01) than the
@@
 This situation is found commonly in historical datasets.
-We envision \regseg{} to be integrated in diffusion processing pipelines, after a preliminar
+We envision \regseg{} to be integrated in diffusion processing pipelines, after a 
+  \revcomment[R\#3-C10]{preliminary}
   \gls*{dti} computation and before anatomically-informed reconstruction and tractography
@@
   can be used to map the tracts into structural space.
+\revcomment[R\#3-C5]{%
+Even though we apply \regseg{} to the problem of susceptibility distortion,
+  \emph{it is not a distortion correction method}, but rather a surface alignment method.
+In fact, the distortions are not corrected in the \gls*{epi} data.}
+\revcomment[R\#3-C6]{%
+Therefore, we propose here to perform the reconstruction and tractography processes in the
+  original (distorted) diffusion data.}
+\revcomment[R\#1-C1]{%
+Depending on the application (like \gls*{tbss}, or connectivity analyses that do not build the
+  network matrix upon anatomical information mapped from a reference image), the obtained tracks
+  may be geometrically unwarped using the estimation of the deformation found by \regseg{}.
+Conversely, when the information necessary in the diffusion analysis is projected from the reference
+  space, \regseg{} allows to avoid resampling and/or unwarping.}
+\revcomment[R\#3-C6]{%
+This methodological variation should be further investigated, to ensure which processing design
+  yields the most accurate tractography results.}
 

### 2016-03-02 Publications/2015-NeuroImage/08-appendix-01.tex
version submitted to NI
getting submission ready

@@
-% if have a single appendix:
-%\appendix[Proof of the Zonklar Equations]
-% or
-%\appendix  % for no appendix heading
-% do not use \section anymore after \appendix, only \section*
-% is possibly needed
-
-% use appendices with more than one appendix
-% then use \section to start each appendix
-% you must declare a \section before using any
-% \subsection or using \label (\appendices by itself
-% starts a section numbered zero.)
-%
-
+% -*- root: 00-main.tex -*-
+\onecolumn
 \appendix
@@
 \section{Proof of duality between the probabilistic and the presented frameworks}
-Here the demonstration that we are preparing.
-
-% you can choose not to have a title for an appendix
-% if you want by leaving the argument blank
-%\section{}
-%Appendix two text goes here.
 
@@
 In the probabilistic approach, the underlying concept to segmenting an image is the
-Bayes' rule. For a discrete image of $N$ pixels indexed by $i$, this rule reads as
+Bayes' rule. For a discrete image of $N$ \revcomment[R\#3-C10]{voxel}s indexed by $i$, this rule reads as
 follows:
@@
 where $\mathbf{s} = I(\mathbf{x})$, $\mathbf{s} \in \mathbb{R}^C$ (a vector of $C$ scalar features).
-Additionally, the discrete grid of $N$ pixels has been converted to a continuous space by assuming
+Additionally, the discrete grid of $N$ \revcomment[R\#3-C10]{voxel}s has been converted to a continuous space by assuming
 $d\mathbf{x}$ as infinitesimal bin size.

### 2016-03-10 Publications/2015-NeuroImage/responseLetter3.tex
updated paper

@@
 \reply{%
-We have made editions in this figure to make it more clear.
+We have made the following editions in this figure to make it more clear:
+\begin{itemize}
+\item Changed color palette to the ``Set1'' palette of the ColorBrewer tool\footnote{\url{http://colorbrewer2.org/}}, which has
+  darker colors.
+\item The darker colors allowed for thiner lines to represent the contours of the joint density plot in panel A, improving
+  readability.
+\item Better positioned labels in the panel A of this figure, using the Fruchterman-Reingold force-directed algorithm.
+\item Matched the fonts of axis labels with the font used in the NeuroImage journal.
+\end{itemize}
 }
@@
 
-\bibliographystyle{mystyle}
-\bibliography{Remote}
+% \bibliographystyle{mystyle}
+% \bibliography{Remote}
 

### 2016-03-10 Publications/2015-NeuroImage/figure02.tex
updated paper

% @Author: Oscar Esteban
% @Date:   2015-04-09 13:06:32
% @Last Modified by:   oesteban
% @Last Modified time: 2016-03-09 16:30:39

\documentclass[tikz]{standalone}
\usepackage{graphicx}
\graphicspath{{figures/}}

\usepackage[T1]{fontenc}
\usepackage{charter}

\begin{document}
\noindent
\begin{minipage}[t]{14in}
\noindent
\begin{tikzpicture}[block/.style={}]
  \node[inner sep=0pt](fig02a) at (0,14)
    {%
    \noindent
    \begin{minipage}[t]{.2\textwidth}
    \centering
	\vskip 20pt
    \includegraphics[width=\textwidth]{figure02A-1}
    \vskip 10pt
    \parbox{.9\textwidth}{%
    \textbf{Reference}: segmentation by the surfaces
    extracted in T1w and original (undistorted)
    diffusion data.}
    \end{minipage}
    \hfill
    \begin{minipage}[t]{.2\textwidth}
	\vskip 20pt
    \centering
    \includegraphics[width=\textwidth]{figure02A-2}
    \vskip 10pt
    \parbox{.9\textwidth}{%
    \textbf{Before registration}, the
	contours from the T1w are
	not aligned with the
	distorted diffusion data.}
    \end{minipage}
    \hfill
    \begin{minipage}[t]{.2\textwidth}
	\vskip 20pt
    \centering
    \includegraphics[width=\textwidth]{figure02A-3}
    \vskip 10pt
    \parbox{.9\textwidth}{%
    \textbf{After registration}, the
	contours are now aligned
	with the distorted diffusion
	data.}
    \end{minipage}};
  \node[text=black!75, inner sep=0] at (-4.7,18.7) {\Large \textbf{A. OPTIMIZATION OF THE SEGMENTATION MODEL}};
  % \node[inner sep=10pt, draw=black!40](fig05c) at (8.,0)
  %   {\hspace*{20pt}};
  % \node[circle, text=black!75, inner sep=0] at (-10.65,4.05) {\Huge \textbf{B}};
  %
  %
  %
  %
  \node[inner sep=0pt](fig02b) at (0,0)
    {%
    	\begin{minipage}[t]{.095\textwidth}
    	\centering
    	\includegraphics[width=\textwidth]{figure02B-100}
    	\includegraphics[width=\textwidth]{figure02B-101}
    	\includegraphics[width=\textwidth]{figure02B-102}
    	\end{minipage}
    	\begin{minipage}[t]{.095\textwidth}
    	\centering
    	\includegraphics[width=\textwidth]{figure02B-200}
    	\includegraphics[width=\textwidth]{figure02B-201}
    	\includegraphics[width=\textwidth]{figure02B-202}
    	\end{minipage}
    	\begin{minipage}[t]{.095\textwidth}
    	\centering
    	\includegraphics[width=\textwidth]{figure02B-300}
    	\includegraphics[width=\textwidth]{figure02B-301}
    	\includegraphics[width=\textwidth]{figure02B-302}
    	\end{minipage}
	    \hskip 15pt
    	\begin{minipage}[t]{.095\textwidth}
    	\centering
    	\includegraphics[width=\textwidth]{figure02B-103}
    	\includegraphics[width=\textwidth]{figure02B-104}
    	\includegraphics[width=\textwidth]{figure02B-105}
    	\end{minipage}
    	\begin{minipage}[t]{.095\textwidth}
    	\centering
    	\includegraphics[width=\textwidth]{figure02B-203}
    	\includegraphics[width=\textwidth]{figure02B-204}
    	\includegraphics[width=\textwidth]{figure02B-205}
    	\end{minipage}
    	\hfill
    	\begin{minipage}[t]{.095\textwidth}
    	\centering
		\includegraphics[width=\textwidth]{figure02B-303}
    	\includegraphics[width=\textwidth]{figure02B-304}
    	\includegraphics[width=\textwidth]{figure02B-305}
    	\end{minipage}%
	};
  \node[inner sep=3pt, draw=black!20](fig02c) at (0,-6.4)
    {\includegraphics[width=.57\textwidth]{legend}};
  \node[text=black!75, inner sep=0] at (-4.7,7.0) {\Large \textbf{B. DETAIL OF TISSUE-WISE JOINT-DISTRIBUTIONS}};
  \node[text=black!75, inner sep=0] at (-9.0,6.3) {\textbf{Reference}};
  \node[text=black!75, inner sep=0] at (-5.5,6.3) {\textbf{Pre-registration}};
  \node[text=black!75, inner sep=0] at (-2.0,6.3) {\textbf{Post-registration}};
  \node[text=black!75, inner sep=0] at (1.8,6.3) {\textbf{Reference}};
  \node[text=black!75, inner sep=0] at (5.4,6.3) {\textbf{Pre-registration}};
  \node[text=black!75, inner sep=0] at (8.8,6.3) {\textbf{Post-registration}};
\end{tikzpicture}
\end{minipage}
\end{document}

### 2016-03-10 Publications/2015-NeuroImage/figure05.tex
updated paper

@@
 % @Date:   2015-04-09 13:06:32
-% @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-04-09 15:52:17
+% @Last Modified by:   oesteban
+% @Last Modified time: 2016-03-09 13:49:09
 
@@
 \begin{tikzpicture}
-  \node[inner sep=10pt, draw=black!40](fig03b) at (0,8.8)
+  \node[inner sep=10pt, draw=black!40](fig05b) at (0,8.8)
     {\hspace*{20pt}\includegraphics[width=21cm]{figure05A}};
-  \node[inner sep=10pt, draw=black!40](fig03c) at (0,0)
+  \node[inner sep=10pt, draw=black!40](fig05c) at (0,0)
     {\hspace*{20pt}\includegraphics[width=21cm]{figure05B}};

### 2016-03-10 Publications/2015-NeuroImage/figure06.tex
updated paper

@@
 % @Date:   2015-04-09 14:39:30
-% @Last Modified by:   Oscar Esteban
-% @Last Modified time: 2015-04-09 15:54:48
+% @Last Modified by:   oesteban
+% @Last Modified time: 2016-03-09 13:48:53
 
@@
   \begin{tikzpicture}
-    \node[inner sep=15pt, draw=black!40](fig05a) at (0,12.0)
+    \node[inner sep=15pt, draw=black!40](fig06a) at (0,12.0)
       {\hspace*{20pt}\includegraphics[width=21cm]{figure06A}};
-    \node[inner sep=15pt, draw=black!40](fig05b) at (0,0)
+    \node[inner sep=15pt, draw=black!40](fig06b) at (0,0)
       {\hspace*{20pt}\includegraphics[width=21cm]{figure06B}};

### 2016-03-10 Publications/2015-NeuroImage/responseLetter3.tex
latest editions to the paper

--- 2016-03-02 Publications/2015-NeuroImage/responseLetter3.tex
@@
 \reply{%
+The reviewer is right, we may have oversimplified the details about the 
+  \emph{b0}-to-T2w registration.
+We have added the text in lines 227--232 to describe how the registration works.
+The direction of the transform is \emph{b0}-to-T2w (as implemented in \emph{ExploreDTI})
+  and therefore, the result is a mathematical transformation which maps the coordinates
+  of the T2w image to coordinates in the \emph{b0}.
+This way, it is possible to obtain the corresponding intensity of the \emph{b0} and
+  resample it in the reference space of the T2w.
+Since the surfaces are defined by their vertices in reference space, the transform
+  obtained in the \emph{b0}-to-T2w direction of mapping can be used to warp the
+  locations of the vertices into \emph{b0} space.
+Therefore, we have the surfaces on reference (undistorted) space, and we warp them
+  into diffusion space using three transforms: 1) the ground truth given by the realistic
+  fieldmaps, 2) \emph{regseg} and 3) the \emph{b0}-to-T2w registration method.
+The error computations are done in diffusion space, 2) and 3) against 1).
+This detail of which coordinate system supports the error measurement is included
+  in the new version in lines 238--239.
 }
@@
 \reply{%
+As described in \autoref{com:7}, yes, the coordinate system in which the surfaces are
+  compared is the diffusion space.
+Generally, when we speak of \emph{b0}-to-T2w registration, we identify the direction of
+  the mapping with the displacement of the information of the \emph{b0} image into the
+  T2w-image grid.
+However, the transformation underlying works in the opposite way: for each point in
+  the reference space computes the corresponding point in the moving image.
+The objects defined explicitly, like the surfaces, in the reference space can be directly
+  mapped to moving space using this transform.
+This gives the impression that the transform goes in the opposite direction, and we
+  understand that this is potentially what was not completely clear in the former
+  version of the manuscript.
 }
@@
 
-\bibliographystyle{mystyle}
-\bibliography{Remote}
+%\bibliographystyle{mystyle}
+%\bibliography{Remote}
 

### 2016-03-10 Publications/2015-NeuroImage/04-methods.tex
updated paper
latest editions to the paper

@@
   available\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/t2b_elastix_y.txt}}.
-\revcomment[R\#3-C7\\R\#3-C7]{%
+\revcomment[R\#3-C7\\R\#3-C8]{%
 In this registration scheme, the \gls*{t2} image is the \emph{reference} and
@@
   of crucial interest in whole-brain tractography.
+\revcomment[R\#3-C7\\R\#3-C8]{%
+The \gls*{swindex} is always computed on the \gls*{dmri} space.}

### 2016-03-10 Publications/2015-NeuroImage/04-methods.tex
latest editions prior submission

--- 2016-03-02 Publications/2015-NeuroImage/04-methods.tex
@@
 \begin{minipage}{\textwidth}
-\includegraphics[width=\linewidth]{figure02}\\%
-\vskip-2.0em
-\includegraphics[width=\linewidth]{legend}%
+\includestandalone[width=\linewidth]{figure02}%
 \end{minipage}
@@
   according to the following six regions:
-  1) thalamus ($\Omega_{Tha}$);
-  2) ventricular system and deep \gls*{gm} structures ($\Omega_{VdGM}$);
-  3) cerebral \gls*{wm} ($\Omega_{WM}$);
-  4) brain stem and cerebellar \gls*{wm} ($\Omega_{bst}$);
-  5) cerebellar \gls*{gm} ($\Omega_{cbGM}$); and
-  6) cortical \gls*{gm} ($\Omega_{GM}$).
+  1) thalamus ($\Omega_\text{Tha}$);
+  2) ventricular system and deep \gls*{gm} structures ($\Omega_\text{VdGM}$);
+  3) cerebral \gls*{wm} ($\Omega_\text{WM}$);
+  4) brain stem and cerebellar \gls*{wm} ($\Omega_\text{bst}$);
+  5) cerebellar \gls*{gm} ($\Omega_\text{cbGM}$); and
+  6) cortical \gls*{gm} ($\Omega_\text{GM}$).
 Using tools in \emph{FreeSurfer} and appropriate selections of labels in the

### 2016-03-10 Publications/2015-NeuroImage/04-methods.tex

@@
 \begin{figure*}
-% \includestandalone[width=\linewidth]{figure02-model}
-\begin{minipage}{\textwidth}
-\includestandalone[width=\linewidth]{figure02}%
-\end{minipage}
-% \includegraphics[width=\linewidth]{figure04B}
-% \includegraphics[width=\linewidth]{figure04C}
+\includestandalone[width=\textwidth]{figure02}
 \caption{%

### 2016-03-10 Publications/2015-NeuroImage/figure02.tex

@@
 % @Last Modified by:   oesteban
-% @Last Modified time: 2016-03-09 16:30:39
+% @Last Modified time: 2016-03-10 09:23:55
 
@@
     	\centering
-    	\includegraphics[width=\textwidth]{figure02B-100}
     	\includegraphics[width=\textwidth]{figure02B-101}
     	\includegraphics[width=\textwidth]{figure02B-102}
+    	\includegraphics[width=\textwidth]{figure02B-103}
     	\end{minipage}
@@
     	\centering
-    	\includegraphics[width=\textwidth]{figure02B-200}
     	\includegraphics[width=\textwidth]{figure02B-201}
     	\includegraphics[width=\textwidth]{figure02B-202}
+    	\includegraphics[width=\textwidth]{figure02B-203}
     	\end{minipage}
@@
     	\centering
-    	\includegraphics[width=\textwidth]{figure02B-300}
     	\includegraphics[width=\textwidth]{figure02B-301}
     	\includegraphics[width=\textwidth]{figure02B-302}
+    	\includegraphics[width=\textwidth]{figure02B-303}
     	\end{minipage}
@@
     	\centering
-    	\includegraphics[width=\textwidth]{figure02B-103}
     	\includegraphics[width=\textwidth]{figure02B-104}
     	\includegraphics[width=\textwidth]{figure02B-105}
+    	\includegraphics[width=\textwidth]{figure02B-100}
     	\end{minipage}
@@
     	\centering
-    	\includegraphics[width=\textwidth]{figure02B-203}
     	\includegraphics[width=\textwidth]{figure02B-204}
     	\includegraphics[width=\textwidth]{figure02B-205}
+    	\includegraphics[width=\textwidth]{figure02B-200}
     	\end{minipage}
-    	\hfill
     	\begin{minipage}[t]{.095\textwidth}
     	\centering
-		\includegraphics[width=\textwidth]{figure02B-303}
     	\includegraphics[width=\textwidth]{figure02B-304}
     	\includegraphics[width=\textwidth]{figure02B-305}
+		\includegraphics[width=\textwidth]{figure02B-300}
     	\end{minipage}%
@@
   \node[text=black!75, inner sep=0] at (8.8,6.3) {\textbf{Post-registration}};
+
+  \node[anchor=west,text width=1.4cm, text centered, fill=white, draw=black!70,row sep=0.1, inner sep=1, rounded corners=5, scale=0.8] (newmode0) at (4.8, 4.0) {New mode};
+  \node[anchor=west,text width=1.4cm, text centered, fill=white, draw=black!70,row sep=0.1, inner sep=1, rounded corners=5, scale=0.8] (newmode1) at (-5.6, -4.0) {New mode};
+  \node[anchor=west,text width=1.4cm, text centered, fill=white, draw=black!70,row sep=0.1, inner sep=1, rounded corners=5, scale=0.8] (spread0) at (4.0, -5.1) {Larger spread};
+  \node[anchor=west,text width=1.4cm, text centered, fill=white, draw=black!70,row sep=0.1, inner sep=1, rounded corners=5, scale=0.8] (spread1) at (-5.6, 0.0) {Larger spread};
+  \node[anchor=west,text width=1.4cm, text centered, fill=white, draw=black!70,row sep=0.1, inner sep=1, rounded corners=5, scale=0.8] (spread2) at (-5.6, 4.0) {Larger spread};
+
+    \draw[-stealth, line width=2pt, black]
+    (newmode0.south) -- ++(-0.85, -0.5);
+    \draw[-stealth, line width=2pt, black]
+    (newmode1.west) -- ++(-1.1, 0.6);
+    \draw[-stealth, line width=2pt, black]
+    (spread0.north east) -- ++(.2, .2);
+    \draw[-stealth, line width=2pt, black]
+    (spread1.south) -- ++(-0.7, -0.8);
+    \draw[-stealth, line width=2pt, black]
+    (spread2.south) -- ++(0, -1.25);
+    \draw[-stealth, line width=2pt, black]
+    (spread2.west) -- ++(-1.25, -0.1);
 \end{tikzpicture}

### 2016-03-11 Publications/2015-NeuroImage/responseLetter3.tex
new figure, MBC comments

@@
 \reply{%
-We have introduced this discussion in lines 351--355.
+The reviewer is right: despite new whole-brain connectivity strategies, registration from/to structural space is still required.
+We have added this clarification in lines 359--363, in the
+  discussion section.
 }
@@
 \item Matched the fonts of axis labels with the font used in the NeuroImage journal.
+\item Rearranged the tissues in panel B to match the hierarchical ordering of tissues.
+\item Added annotations to capture the reader's attention and point it to those significant areas of the plot, highlighting the
+effects of misregistration in each tissue distribution.
 \end{itemize}
@@
 \reply{%
-We have reformulated that particular sentence (lines 57--59) and added a corresponding discussion in lines 348--356.
-These editions are highly related to Comments \ref{com:1}, \ref{com:5} and \ref{com:6}.
+We have reformulated that particular sentence (lines 57--59), in a way that
+  it avoids mentioning posterior processing.
+How this methodological variation would affect posterior processing (analyzing tracks, interpolating FA on them, building the connectivity matrix, etc.) is discussed in the paragraph containing lines 358--359.
+Particularly, these lines 358--359 were inserted to address this comment, which is highly related to Comments \ref{com:1}, \ref{com:5} and \ref{com:6}.
 }
+\newpage
 \begin{reviewpoint}
@@
 \reply{%
-We have explicitly stated this in a new sentence (lines 348--349).
+We have explicitly stated this in a new sentence (lines 355--357).
 }
@@
 \reply{%
-Right after the previous edition, in lines 349--350 we introduce the proposal of this methodological variation.
-The discussion and proposal of a future research line is then done in lines 355--356, after some further details added
+Right after the previous edition, in lines 356--357 we introduce the proposal of this methodological variation.
+The discussion and proposal of a future research line is then done in lines 363--364, after some further details added
   to address the \autoref{com:1}.
@@
 }
+\newpage
 \begin{reviewpoint}
@@
 \reply{%
-We have introduced this description in lines 183--187, along with including the detail of the actual HCP release of the data used in l. 177.
+We have introduced this description in lines 184--188, along with including the detail of the actual HCP release of the data used in l. 178.
 }
@@
 We have replaced the six appearances of the term ``pixel'' by ``voxel'', including the ``subpixel'' present in line 61.
-We also have fixed the misspelling of ``preliminar'' of line 339.
+We also have fixed the misspelling of ``preliminar'' of line 350.
 }

### 2016-03-11 Publications/2015-NeuroImage/06-discussion.tex
new figure, MBC comments

@@
   original (distorted) diffusion data.}
+\revcomment[R\#1-C3]{%
+\Regseg{} allows to avoid resampling and/or unwarping because the information necessary in the diffusion analysis is projected
+  from the reference space.}
 \revcomment[R\#1-C1]{%
-Depending on the application (like \gls*{tbss}, or connectivity analyses that do not build the
-  network matrix upon anatomical information mapped from a reference image), the obtained tracks
-  may be geometrically unwarped using the estimation of the deformation found by \regseg{}.
-Conversely, when the information necessary in the diffusion analysis is projected from the reference
-  space, \regseg{} allows to avoid resampling and/or unwarping.}
+Certain applications (like \gls*{tbss}) and methodologies (like building the connectivity matrix
+  by clustering the tracks) may not perform on the native (distorted) diffusion space correctly,
+  but they still need a mapping to the reference space.
+Using \regseg{}, the tracks obtained in native space can be unwarped using the resulting estimation
+  of the deformation field.}
 \revcomment[R\#3-C6]{%

### 2016-03-11 Publications/2015-NeuroImage/04-methods.tex
minor fixes

@@
   distortions using the \emph{TOPUP} tool \citep{andersson_how_2003}
-  before producing the released ``minimally preprocessed'' data.}.
+  before producing the released ``minimally preprocessed'' data}.
 
@@
   that correspond to $\vec{v}_i$.
-In practice, we only report the \gls*{swindex} for three surfaces (\{$\Gamma_{VdGM}, \Gamma_{WM}, \Gamma_{pial}$\})
+In practice, we only report the \gls*{swindex} for three surfaces (\{$\Gamma_\text{VdGM}, \Gamma_\text{WM}, \Gamma_\text{pial}$\})
   of crucial interest in whole-brain tractography.

### 2016-03-11 Publications/2015-NeuroImage/05-results.tex
latest editions to the paper
minor fixes

@@
   interest selected in \autoref{sec:regseg-experiments_evaluation}
-  ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$).
+  ($\Gamma_\text{VdGM}$, $\Gamma_\text{WM}$, $\Gamma_\text{pial}$).
 All of the statistical tests showed that the error distributions obtained with \regseg{} and
@@
     \begin{tabular}{cccccc}
-    & & $\Gamma_{VdGM}$  & $\Gamma_{WM}$ & $\Gamma_{pial}$ & Aggreg. \\
+    & & $\Gamma_\text{VdGM}$  & $\Gamma_\text{WM}$ & $\Gamma_\text{pial}$ & Aggreg. \\
     \hline
@@
     the \gls*{hcp}, which show that \regseg{} performed better than the alternative \acrfull{t2b} method.
-    The distribution of the errors computed for the surfaces of interest ($\Gamma_{VdGM}$, $\Gamma_{WM}$, $\Gamma_{pial}$)
+    The distribution of the errors computed for the surfaces of interest ($\Gamma_\text{VdGM}$, $\Gamma_\text{WM}$, $\Gamma_\text{pial}$)
       and the aggregate of all surfaces (Aggreg. column) had lower 95\% CIs with \regseg{}.

### 2016-03-24 Publications/2015-NeuroImage/responseLetter3.tex
final changes to be submitted

@@
 The reviewer is right: despite new whole-brain connectivity strategies, registration from/to structural space is still required.
-We have added this clarification in lines 359--363, in the
+We have added this clarification in lines 361--364, in the
   discussion section.
@@
 \item Better positioned labels in the panel A of this figure, using the Fruchterman-Reingold force-directed algorithm.
+  This functionality has been included in the code release of the paper.
 \item Matched the fonts of axis labels with the font used in the NeuroImage journal.
@@
 \end{itemize}
+Additionally, we have included in the Supplemental Materials (new figures S8 and S9) zoomed versions of all the joint plots for
+  a better interpretation.
 }
@@
   it avoids mentioning posterior processing.
-How this methodological variation would affect posterior processing (analyzing tracks, interpolating FA on them, building the connectivity matrix, etc.) is discussed in the paragraph containing lines 358--359.
-Particularly, these lines 358--359 were inserted to address this comment, which is highly related to Comments \ref{com:1}, \ref{com:5} and \ref{com:6}.
+How this methodological variation would affect posterior processing (analyzing tracks, interpolating FA on them, building the connectivity matrix, etc.) is discussed in the paragraph containing lines 359--360.
+Particularly, these lines 359--360 were inserted to address this comment, which is highly related to Comments \ref{com:1} (Reviewer \#1), \ref{com:5} (R\#3) and \ref{com:6} (R\#3).
 }
@@
 \reply{%
-We have defined introduced the definition of $\mathbf{g}_k$ in lines 122--123, right before the
-  reference to the mentioned equations.
+We have introduced the definition of $\mathbf{g}_k$ in lines 122--125, introducing the former Eq. (A.8) defining the
+gradients in the text as new Equation (11).
+Also, a clarification has been added in the appendix itself (lines 443--445).
 }
@@
 \reply{%
-We have explicitly stated this in a new sentence (lines 355--357).
+To strongly support the focus of the paper defined in lines 42--43, we have explicitly stated this in a new sentence (lines 356--358).
+Additionally, the edits in lines 57--59 introduced in response to \autoref{com:2} also contribute to clarify the primary goal of
+  image registration.
 }
@@
 \reply{%
-Right after the previous edition, in lines 356--357 we introduce the proposal of this methodological variation.
-The discussion and proposal of a future research line is then done in lines 363--364, after some further details added
-  to address the \autoref{com:1}.
+Right after the previous edition, in lines 358--359 we introduce the proposal of this methodological variation.
+The discussion and proposal of a future research line is then done in lines 364--365.
+Some other details are added to cover Comments \ref{com:1} and \ref{com:2} (R\#1) as well.
 }
@@
 \reply{%
-As described in \autoref{com:7}, yes, the coordinate system in which the surfaces are
+As described in \autoref{com:7} (R\#3), yes, the coordinate system in which the surfaces are
   compared is the diffusion space.
@@
   T2w-image grid.
-However, the transformation underlying works in the opposite way: for each point in
-  the reference space computes the corresponding point in the moving image.
+However, the underlying transformation works in the opposite way: for each point in
+  the reference space (T2w) computes the corresponding point in the moving image (\emph{b0}).
 The objects defined explicitly, like the surfaces, in the reference space can be directly
@@
   version of the manuscript.
+New text (lines 227--232) in response to \autoref{com:7} was also included to clarify this point.
 }
@@
 \reply{%
-We have introduced this description in lines 184--188, along with including the detail of the actual HCP release of the data used in l. 178.
+We have introduced this description in lines 184--188, along with including the detail of the actual HCP release of the data used in line 178.
 }
@@
 We have replaced the six appearances of the term ``pixel'' by ``voxel'', including the ``subpixel'' present in line 61.
-We also have fixed the misspelling of ``preliminar'' of line 350.
+We also have fixed the misspelling of ``preliminar'' of line 351.
 }

### 2016-03-24 Publications/2015-NeuroImage/03-introduction.tex
final changes to be submitted

@@
     alternatively to the typical design of the processing flow, the information from the
-    reference \gls*{t1} can precisely be mapped onto the distorted \gls*{dmri} space,
+    reference \gls*{t1} can be precisely mapped onto the distorted \gls*{dmri} space,
     avoiding the interpolation of the \glspl*{dwi} required by unwarping the diffusion data; and} 
@@
   the corresponding \gls*{t2} image of the same subject.
-This approach is the first step of the \gls*{t2b} correction
-  methods \citep{kybic_unwarping_2000}.
+This approach is the first step of the above-mentioned \gls*{t2b} correction
+  methods.
 We reproduce the settings and implementation of a widely used diffusion

### 2016-03-24 Publications/2015-NeuroImage/04-methods.tex
final changes to be submitted

@@
 
+\begin{figure}
+  \includegraphics[width=\linewidth]{figure01}
+  \caption{The interfacing surfaces $\Gamma_m$ between the competing \glspl{roi} $\Omega_l$,
+  play the role of active contours which drive the registration process.
+  They evolve iteratively along the normal $\hat{\mathbf{n}}_i$ of the surface at each vertex
+    $\vec{v}_i$ of the mesh.
+  The gradient speeds $\bar{s}_i$ drive registration, which are computed as the disparity of the data
+    energies with respect to the two limiting regions of $M(\vec{v}_i)$, the features of the image
+    $M$ in the location of vertex $\vec{v}_i$.
+  The computation of shape-gradients is developed in \ref{app:shape_gradients}.
+  In this figure, the $\bar{s}_1$ derived from \autoref{eq:regseg-shape_gradient_disc2} is 
+    written in the lower box, with $\Omega_{wm}$ being the inner limiting region, $\Omega_{gm}$
+    the outer region, and $w_{0,1}$ the relative area associated with vertex $\vec{v}_1$
+    with respect to the total area of surface $\Gamma_0$.
+  }\label{fig:regseg-method}
+\end{figure}
+
 \paragraph*{Cost-function derivation}
@@
   where $P(M \mid U,\omegaset)$ is the data likelihood.
-Since $\omegaset$ are mapped by $U$, we simplify
+Since $\omegaset$ is mapped by $U$, we simplify
   $P(U, \omegaset) = P(U) \implies P(M \mid U,\omegaset) = P(M \mid U)$.
@@
 
-\begin{figure}
-  \includegraphics[width=\linewidth]{figure01}
-  \caption{The interfacing surfaces $\Gamma_m$ between the competing \glspl{roi} $\Omega_l$,
-  play the role of active contours which drive the registration process.
-  They evolve iteratively along the normal $\hat{\mathbf{n}}_i$ of the surface at each vertex
-    $\vec{v}_i$ of the mesh.
-  The gradient speeds $\bar{s}_i$ drive registration, which are computed as the disparity of the data
-    energies with respect to the two limiting regions of $M(\vec{v}_i)$, the features of the image
-    $M$ in the location of vertex $\vec{v}_i$.
-  The computation of shape-gradients is developed in \ref{app:shape_gradients}.
-  In this figure, the $\bar{s}_1$ derived from \autoref{eq:regseg-shape_gradient_disc2} is 
-    written in the lower box, with $\Omega_{wm}$ in the inner limiting region, $\Omega_{gm}$
-    the outer region, and $w_{0,1}$ is the relative area associated with vertex $\vec{v}_1$
-    with respect to the total area of surface $\Gamma_0$.
-  }\label{fig:regseg-method}
-\end{figure}
-
 \begin{figure*}
@@
   \gammaset{} do not fit the data.
-% The left plot in each panel represents a kernel density estimate of the location and spread of
-%   each region $\Omega_l$ in the bivariate feature space of $M$, which comprises
-%   the \gls*{fa} and \gls*{adc} maps.
-% At the top and the right margins, the marginal distributions of $\Omega_l$ on the \gls*{fa} and
-%   \gls*{adc} are respectively displayed.
-% On the right side, the sampled distribution of each region is shown in small scatter plot, in contrast to
-%   the complete bivariate data.
-% A) shows the initialization of the model, when the anatomically-correct surfaces are not
-%   mapped into the \gls*{dmri} space.
-% The panel shows the resulting conditional distributions of segmenting the (distorted) \gls*{fa} and
-%   \gls*{adc} maps with the original (undistorted) surfaces.
-% Panel B shows the equivalent information, using the surfaces that are perfectly mapped into the
-%   distorted \gls*{dmri} space.
-% Since the surfaces are now aligned in $M$, the spreads of the conditional distributions for each region are
-%   reduced with respect to those in A.
-% Panel C, after \regseg{} registration, shows how the covariance matrices $\boldsymbol{\Sigma}_l$
-%   are minimized with respect to panel A, and the resulting segmentation is close
-%   to the theoretical situation presented in panel B.}
+{\color{revcolor} The plots in (A) and (B) are provided at full-size in the
+  \citetalias[figures S8, S9]{esteban_useful_2016}.}
 }\label{fig:regseg-model}
@@
   can then be computed }%
-using the expressions \eqref{eq:regseg-gradient_wshape} and 
-  \eqref{eq:regseg-shape_gradient_final} given in \ref{app:shape_gradients}.
-Finally, introducing the analytical derivative of the regularization term,
+using the expression \eqref{eq:regseg-gradient_wshape}
+\revcomment[R\#1-C4]{of \ref{app:shape_gradients}, obtaining:
+
+\begin{equation}
+  \vec{g}_k = - \underset{i}{\sum} \bar{s}_i \cdot \psi_k(\vec{v}_i) \, \hat{\vec{e}}.
+  \label{eq:regseg-shape_gradient_final}
+\end{equation}
+
+Then,} introducing the analytical derivative of the regularization term,
   the \autoref{eq:regseg-gradient_descent} is reformulated as:
@@
   \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
-  \vec{g}_k  + \boldsymbol{\alpha} \cdot \vec{u}_k - \boldsymbol{\beta} \cdot (\Delta \vec{u}_k),
+  \vec{g}_k  + \boldsymbol{\alpha} \cdot \vec{u}_k - \boldsymbol{\beta} \cdot (\Delta \vec{u}_k).
   \label{eq:regseg-final_gradient}
   \end{align}
-where $\vec{g}_k$ are the shape-gradient contributions on the coefficients $\vec{u}_k$ of the B-spline grid.
 Finally, to descend this gradient, we establish a semi-implicit Euler scheme
-  \citepalias[see][section S1.3]{esteban_useful_2015},
+  \citepalias[see][section S1.3]{esteban_useful_2016},
   with a step size parameter $\delta$, which we solve in the spectral domain as follows:
@@
   and target image smoothing), the implementation details, and other features such as the sparse matrix
-  approach to fast interpolation are discussed in the \citetalias[][section S1]{esteban_useful_2015}.
+  approach to fast interpolation are discussed in the \citetalias[][section S1]{esteban_useful_2016}.
 
@@
   5) The results were evaluated visually and quantitatively.
-  The arrows point to edges in the target images (light-yellow arrows for \gls*{fa}, blue for \gls*{adc}) that
-    should be aligned with a surface, showing how distortion limits the direct mapping from the structural space
-    in which the contours are defined.}\label{fig:regseg-evworkflows}
+  The arrows in the step (5) point to edges in the target images (light-yellow arrows for \gls*{fa}, blue for 
+    \gls*{adc}) that should be aligned with a surface, showing how distortion limits the direct mapping from
+    the structural space in which the contours are defined.}\label{fig:regseg-evworkflows}
 \end{figure*}
@@
 The segmentation model corresponding to this partition is shown in \autoref{fig:regseg-model}
-  and discussed in greater detail in the \citetalias[section S4]{esteban_useful_2015}.
+  and discussed in greater detail in the \citetalias[section S4]{esteban_useful_2016}.
 
@@
 The minimum of the cost function (\autoref{sec:regseg-methods_map}) was consistently located at
-  $\epsilon=0.0$ (the ground-truth) for all of the cases \citepalias[figure S2]{esteban_useful_2015}.
+  $\epsilon=0.0$ (the ground-truth) for all of the cases \citepalias[figure S2]{esteban_useful_2016}.
 
@@
 These settings were obtained manually, driven by the feedback obtained from the post-registration
-  convergence reports \citepalias[like that found in][section S1.3]{esteban_useful_2015}.
+  convergence reports \citepalias[like that found in][section S1.3]{esteban_useful_2016}.
 We released \regseg{} along with the tool to generate such convergence reports.
@@
 Therefore, the transform is defined in the coordinate system of the \gls*{t2} image,
-  and for each point in this space provides the location of the corresponding feature
+  and for each point in this space it provides the location of the corresponding feature
   in the \lowb{} image.

### 2016-03-24 Publications/2015-NeuroImage/06-discussion.tex
final changes to be submitted

@@
   displacement vector field, which preserves the topology even with very large deformations.
-\Regseg{} includes an anisotropic regularizer for the displacement field proposed by
+\Regseg{} includes an anisotropic regularizer for the displacement field described by
   \cite{nagel_investigation_1986}.
@@
 The misregistration error that remained after \regseg{} was significantly lower ($p <$ 0.01) than the
-  error corresponding to the \gls*{t2b} correction according to Kruskal-Wallis H-tests
+  error corresponding to the \gls*{t2b} method according to Kruskal-Wallis H-tests
   (\autoref{tab:results_real}).
-Visual inspections of all the results \citepalias[section S5]{esteban_useful_2015} and the violin plots in
+Visual inspections of all the results \citepalias[section S5]{esteban_useful_2016} and the violin plots in
   \autoref{fig:regseg-results_real} confirmed that \regseg{} achieved higher accuracy
@@
 We also demonstrated the smoothness of the objective function on five of the real datasets
-  \citepalias[figure S2]{esteban_useful_2015}, taking advantage of the directional
+  \citepalias[figure S2]{esteban_useful_2016}, taking advantage of the directional
   restriction of possible distortions.
@@
   acquisition other than the minimal protocol comprehending only \gls*{t1} and \gls*{dmri} images.
-This situation is found commonly in historical datasets.
+This situation is commonly found in historical datasets.
+
 We envision \regseg{} to be integrated in diffusion processing pipelines, after a 
@@
 \revcomment[R\#3-C6]{%
-Therefore, we propose here to perform the reconstruction and tractography processes in the
+Therefore, we suggest here to perform the reconstruction and tractography processes in the
   original (distorted) diffusion data.}
 \revcomment[R\#1-C3]{%
-\Regseg{} allows to avoid resampling and/or unwarping because the information necessary in the diffusion analysis is projected
-  from the reference space.}
+\Regseg{} allows to avoid resampling and/or unwarping of the diffusion signal because the 
+  structural information necessary in the diffusion analysis is mapped from the
+  reference space.}
 \revcomment[R\#1-C1]{%
 Certain applications (like \gls*{tbss}) and methodologies (like building the connectivity matrix
-  by clustering the tracks) may not perform on the native (distorted) diffusion space correctly,
-  but they still need a mapping to the reference space.
+  by clustering the tracks) may not be performed correctly on the native (distorted) diffusion space
+  because they still need a mapping to the undistorted space.
 Using \regseg{}, the tracks obtained in native space can be unwarped using the resulting estimation
@@
 \revcomment[R\#3-C6]{%
-This methodological variation should be further investigated, to ensure which processing design
+This methodological variation will be further investigated, to ensure which processing design
   yields the most accurate tractography results.}
@@
   in longitudinal MRI studies of the early developmental patterns \citep{shi_neonatal_2010}.
-In these studies, the surfaces obtained in a mature timepoint of the brain are retrospectively
-  propagated to the initial timepoints, regardless of the changes in the contrast and spatial
+In these studies, the surfaces obtained in a mature time point of the brain are retrospectively
+  propagated to the initial time points, regardless of the changes in the contrast and spatial
   development between them.

### 2016-03-24 Publications/2015-NeuroImage/08-appendix-00-shapepriors.tex
final changes to be submitted

@@
   \end{align}
-where $a_i$ is the area corresponding to vertex $\vec{v}_i$, and $A_p = \sum_i a_i$ is the total area of surface $p$.
+where $a_i$ is the area corresponding to vertex $\vec{v}_i$, and $A_p = \sum_i a_i$ is the total area of the surface $p$.
 In the following, we will refer as $w_{p,i} = a_i / A_p $ to the area contribution of $\vec{v}_i$ to the
@@
 
-Then, we identify the shape gradient contribution $\vec{g}_k$ on the coefficients $\vec{u}_k$ of the B-spline grid:
+Then, we identify the shape gradient contribution $\vec{g}_k$ on the coefficients $\vec{u}_k$ of the B-spline grid 
+\revcomment[R\#1-C4]{to obtain the definition of $\vec{g}_k$ given in \eqref{eq:regseg-final_gradient}}:
 
@@
   \end{equation}%
-  where $\hat{\vec{e}}$ is the coordinates system's unit vector.
-Therefore, the shape gradients projected to the grid of B-spline control points read:
-
-\begin{equation}
-  \vec{g}_k = - \underset{i}{\sum} \bar{s}_i \cdot \psi_k(\vec{v}_i) \, \hat{\vec{e}}.
-  \label{eq:regseg-shape_gradient_final}
-\end{equation}
+\revcomment[R\#1-C4]{where $\psi_k$ and $\vec{u}_k$ define our B-spline deformation model (\autoref{eq:regseg-nodes_tfm})} and
+  $\hat{\vec{e}}$ is the coordinates system's unit vector.

### 2016-03-24 Publications/2015-NeuroImage/supplemental.tex
added fixes of all coauthors, and spell check
final changes to be submitted

@@
 % @Last Modified by:   oesteban
-% @Last Modified time: 2015-11-11 13:14:36
+% @Last Modified time: 2016-03-23 17:11:28
 
@@
 \author{Esteban O., Zosso D., Daducci A., Bach-Cuadra M., \\ Ledesma-Carbayo MJ., Thiran JP, and Santos A.}
-\date{April 2015}
+\date{March 2016}
 
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=.3\linewidth]{figure02A-1}
+  \hfill
+  \includegraphics[width=.3\linewidth]{figure02A-2}
+  \hfill
+  \includegraphics[width=.3\linewidth]{figure02A-3}
+  \caption{Detail of the evolution of the model (all tissue types).}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.3\linewidth]{figure02B-101}
+  \hfill
+  \includegraphics[width=.3\linewidth]{figure02B-201}
+  \hfill
+  \includegraphics[width=.3\linewidth]{figure02B-301}
+%
+  \includegraphics[width=.3\linewidth]{figure02B-102}
+  \hfill
+  \includegraphics[width=.3\linewidth]{figure02B-202}
+  \hfill
+  \includegraphics[width=.3\linewidth]{figure02B-302}
+%  \caption{Detail of the evolution of the model (tissue-wise distribution).}
+\end{figure}
+\begin{figure}
+  \centering
+  \includegraphics[width=.3\linewidth]{figure02B-103}
+  \hfill
+  \includegraphics[width=.3\linewidth]{figure02B-203}
+  \hfill
+  \includegraphics[width=.3\linewidth]{figure02B-303}
+%
+  \includegraphics[width=.3\linewidth]{figure02B-104}
+  \hfill
+  \includegraphics[width=.3\linewidth]{figure02B-204}
+  \hfill
+  \includegraphics[width=.3\linewidth]{figure02B-304}
+%
+  \includegraphics[width=.3\linewidth]{figure02B-105}
+  \hfill
+  \includegraphics[width=.3\linewidth]{figure02B-205}
+  \hfill
+  \includegraphics[width=.3\linewidth]{figure02B-305}
+%
+  \includegraphics[width=.3\linewidth]{figure02B-100}
+  \hfill
+  \includegraphics[width=.3\linewidth]{figure02B-200}
+  \hfill
+  \includegraphics[width=.3\linewidth]{figure02B-300}
+  \caption{Detail of the evolution of the model (tissue-wise distribution).}
+\end{figure}
+
 \clearpage

### 2016-05-03 Publications/2015-NeuroImage/YNIMG13167.tex
typesetting version

\documentclass[3p,authoryear,fleqn]{elsarticle}
\journal{NeuroImage (Accepted paper YNIMG13167)}
\usepackage[utf8]{inputenc}
\defcitealias{esteban_useful_2016}{Supplemental Materials}
\usepackage{xifthen}
\newboolean{review}
\setboolean{review}{false}

\usepackage{tikz}
\usepackage[mode=buildnew]{standalone}
\usepackage{gincltex}
\usepackage{glossaries}
\usepackage{booktabs} \usepackage{multirow}

\usepackage{xcolor}
\usepackage{url}
\usepackage[colorlinks=true]{hyperref}
\usepackage{doi}

\usepackage{xspace}
\usepackage{graphicx}
\graphicspath{{YNIMG13167_figures/}}

\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cancel}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\argmax}{argmax}
\DeclareMathOperator{\argmin}{argmin}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\const}{Const.}
\providecommand{\e}[1]{\ensuremath{\cdot\text{10}^\text{#1}}}
\providecommand{\mdist}[2]{ \mathcal{D}_{#2}^2(\mathbf{#1}) }
\providecommand{\omegaset}{\ensuremath{\boldsymbol{\Omega}}}
\providecommand{\gammaset}{\ensuremath{\boldsymbol{\Gamma}}}
\providecommand{\regseg}{\emph{regseg}}
\providecommand{\Regseg}{\emph{Regseg}}
\providecommand{\lowb}{\textit{b0}}

\let\oldhat\hat
\renewcommand{\vec}[1]{\mathbf{#1}}

\providecommand{\nvec}[1]{\hat{\mathbf{#1}}}

\providecommand{\isores}[2][]{#2\ensuremath{\times}#2\ensuremath{\times}#2\ifthenelse{\equal{#1}{}}{}{ [#1]}\xspace}

\hyphenation{op-tical net-works semi-conduc-tor an-isot-ropy regis-tra-tion iso-tropic Free-Surfer align-ed pre-sent work-flow data-set data-sets micro-struc-tu-re Ma-drid non-linear de-for-ma-tion
an-iso-tro-pic Stud-holme seg-ment-a-tion re-gis-tra-tion map-ped achiev-ed}

\newacronym{mr}{MR}{magnetic resonance}
\newacronym{mri}{MRI}{magnetic resonance imaging}
\newacronym{dwi}{DWI}{diffusion weighted image}
\newacronym{dmri}{dMRI}{diffusion MRI}
\newacronym{dw}{DW}{diffusion weighted}
\newacronym{dti}{DTI}{diffusion tensor imaging}
\newacronym{t1}{T1w}{T1-weighted}
\newacronym{t2}{T2w}{T2-weighted}
\newacronym{csf}{CSF}{cerebrospinal fluid}
\newacronym{wm}{WM}{white matter}
\newacronym{gm}{GM}{gray matter}
\newacronym{epi}{EPI}{echo-planar imaging}
\newacronym{gre}{GRE}{gradient echo sequence}
\newacronym{fa}{FA}{fractional anisotropy}
\newacronym{md}{MD}{mean diffusivity}
\newacronym{adc}{ADC}{apparent diffusion coefficient}
\newacronym{acwe}{ACWE}{active contours without edges}
\newacronym{adf}{ADF}{active deformation field}
\newacronym{map}{MAP}{maximum a posteriori}
\newacronym{snr}{SNR}{signal-to-noise ratio}
\newacronym{pve}{PVE}{partial volume effect}
\newacronym{roi}{ROI}{region of interest}
\newacronym{mrf}{MRF}{Markov Random Field}
\newacronym{pde}{PDE}{partial differential equation}
\newacronym{swindex}{sWI}{surface warping index}
\newacronym{wi}{WI}{Warping index}
\newacronym{nof}{NoF}{number of fibers}
\newacronym{t2b}{T2B}{T2w-registration based}
\newacronym{fft}{FFT}{fast Fourier transform}
\newacronym{hcp}{HCP}{Human Connectome Project}
\newacronym{pe}{PE}{phase-encoding}
\newacronym{ci}{CI}{confidence interval}
\newacronym[first={tract-based spatial statistics --TBSS--}]{tbss}{TBSS}{tract-based spatial statistics}

\newacronym[first={AMICO \cite[accelerated microstructure imaging via convex optimization,][]{daducci_accelerated_2015}}]{amico}{AMICO}{accelerated microstructure imaging via convex optimization}

\makeglossaries

\usepackage[T1]{fontenc}
\usepackage{charter}

\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\usepackage{ifpdf}
\usepackage[switch]{lineno}

\ifthenelse{\boolean{review}}
{
  \usepackage[colorinlistoftodos,textsize=tiny, textwidth=1.2cm]{todonotes}
  \colorlet{revcolor}{blue!70}
  \setlength{\marginparwidth}{1.5cm}
  \reversemarginpar
    \newcommand{\revcomment}[2][]{\ifthenelse{\equal{#1}{}}{}{\todo{#1}}  \begingroup \color{revcolor} #2\endgroup}
}
{
  \definecolor{revcolor}{gray}{0.0}
  \newcommand{\todo}[2][]{}
  \newcommand{\revcomment}[2][]{#2}
}

\def\bibsection{\section*{References}}

\begin{document}
\hypersetup{linkcolor=black!60, citecolor=black!60, urlcolor=black!60}
\begin{frontmatter}

\title{Surface-driven registration method for the structure-informed segmentation of diffusion MR images\tnoteref{t1}}
\tnotetext[t1]{Supplemental materials to this work have been submitted to Data in Brief \citep{esteban_useful_2016}.
During the review process, the document is available in doi:\href{http://figshare.com/s/459c26b4ee8211e493b306ec4bbcf141}{10.6084/m9.figshare.1397502}.}

\author[bit,ciber]{Oscar~Esteban\corref{corrauthor}}
\cortext[corrauthor]{Corresponding author}
\ead{phd@oscaresteban.es}
\author[ucla]{Dominique~Zosso}
\author[lts5]{Alessandro~Daducci}
\author[chuv,lts5]{Meritxell~Bach-Cuadra}
\author[bit,ciber]{Mar\'ia-J.~Ledesma-Carbayo}
\author[lts5,chuv]{Jean-Philippe~Thiran}
\author[bit,ciber]{Andres~Santos}

\address[bit]{Biomedical Image Technologies (BIT), ETSI Telecomunicaci\'on, Universidad Polit\'ecnica de Madrid, Madrid, Spain}
\address[ciber]{Centro de Investigaci\'on Biom\'edica en Red en Bioingenier\'ia, Biomateriales y Nanomedicina (CIBER-BBN), Spain}
\address[ucla]{Department of Mathematics, University of California,
Los Angeles (UCLA), Los Angeles, CA, US}
\address[lts5]{Signal Processing Laboratory (LTS5), \'Ecole Polytechnique
F\'ed\'erale de Lausanne (EPFL), Lausanne, Switzerland}
\address[chuv]{Dept. of Radiology, CIBM, University
Hospital Center (CHUV) and University of Lausanne (UNIL), Lausanne, Switzerland}

\begin{abstract}
Current methods for processing \gls*{dmri} to map the connectivity of the human brain
  require precise delineations of anatomical structures.
This requirement has been approached by either segmenting the data in
  native \gls*{dmri} space or mapping the structural information from \gls*{t1} images.
The characteristic features of diffusion data in terms of \acrlong*{snr}, resolution, as well
  as the geometrical distortions caused by the inhomogeneity of magnetic susceptibility
  across tissues hinder both solutions.
Unifying the two approaches, we propose \regseg{}, a surface-to-volume nonlinear
  registration method that segments homogeneous regions within multivariate images by mapping
  a set of nested reference-surfaces.
Accurate surfaces are extracted from a \gls*{t1} image of the subject, using as target image
  the bivariate volume comprehending the \gls*{fa} and the \gls*{adc} maps derived from the
  \gls*{dmri} dataset.
We first verify the accuracy of \regseg{} on a general context using digital phantoms 
  distorted with synthetic and random deformations.
Then we establish an evaluation framework using undistorted \gls*{dmri} data from the \gls*{hcp}
  and realistic deformations derived from 
  the inhomogeneity fieldmap corresponding to each subject.
We analyze the performance of \regseg{} computing the misregistration error of the surfaces estimated
  after being mapped with \regseg{} onto 16 datasets from the \gls*{hcp}.
The distribution of errors shows a 95\% CI of 0.56--0.66 mm, that is below the \gls*{dmri}
  resolution (1.25 mm, isotropic).
Finally, we cross-compare the proposed tool against a nonlinear \lowb{}-to-T2w registration
  method, thereby obtaining a significantly lower misregistration error with \regseg{}.
The accurate mapping of structural information in \gls*{dmri} space 
  is fundamental to increase the reliability of network building
  in connectivity analyses, and to improve the performance of the emerging structure-informed
  techniques for \gls*{dmri} data processing. \end{abstract}

\begin{keyword}
active~surfaces \sep cortical~parcellation \sep diffusion MRI \sep nonlinear~registration \sep segmentation \sep susceptibility~distortion.
\end{keyword}

\end{frontmatter}

\ifthenelse{\boolean{review}}{\linenumbers}{}
\section{Introduction}\label{sec:regseg-intro}
\Acrlong*{dmri} enables the mapping of microstructure \citep{basser_microstructural_1996}
  and connectivity \citep{craddock_imaging_2013} of the human brain \emph{in-vivo}.
It is generally acquired using \gls*{epi} schemes, since they are very fast at
  scanning a large sequence of images called \glspl*{dwi}.
Each \gls*{dwi} is sensitized with a gradient to probe proton diffusion in a certain
  orientation.
Subsequent processing involves describing the local microstructure with one of the available
  models, which range from the early \gls*{dti} proposed by \cite{basser_microstructural_1996}
  to current models such as \gls*{amico}.
The microstructural map is then used to draw the preferential orientations of diffusion
  across the brain using tractography \citep{mori_threedimensional_1999}.
Finally, a graph representing the corresponding structural network is built using
  the regions of a cortical parcellation as nodes and the fiber paths found by
  tractography as edges \citep{hagmann_mapping_2008}.
The methodologies to solve reconstruction, tractography and network building
  require the delineation of the anatomy in the \gls*{dmri} space.
Moreover, current trends on reconstruction \citep{jeurissen_multitissue_2014} and
  tractography \citep{smith_anatomicallyconstrained_2012} are increasingly using
  structural information to improve the microstructural mapping and fiber-tracking.

Possibly, the earliest structural information incorporated to aid \gls*{dmri} processing
  is the \gls*{wm} mask used as a termination criteria for tractography.
The standardized procedure to obtain this mask was thresholding the \gls*{fa} map.
However, the mask and subsequent analyses are highly dependent on the 
  threshold that is chosen \citep{taoka_fractional_2009}.
To overcome the unreliability of \gls*{fa} thresholding, and to broaden
  \gls*{wm} segmentation to brain tissue segmentation, a large number of
  methods have been proposed using \glspl*{dwi}, the \lowb{}, and \gls*{dti}-derived
  scalar maps such as \gls*{fa}, \gls*{adc} and others \citep{zhukov_level_2003,
  rousson_level_2004,jonasson_segmentation_2005,hadjiprocopis_unbiased_2005,liu_brain_2007,
  lu_segmentation_2008,han_experimental_2009}.
However, the precise segmentation of \gls*{dmri} is difficult for several reasons.
First, \gls{dmri} images have a resolution that is much lower
  than that of the imaged microstructural features.
Therefore, voxels located in structural discontinuities are affected by partial
  voluming of the signal sources.
Second, the extremely low \gls*{snr} and the high dimensionality of the \glspl*{dwi} prevent
  their direct use in segmentation.
Third, the low contrast between \gls*{gm} and \gls*{wm} in the \lowb{} volume also makes
  it unsuitable for brain tissue segmentation.

An alternative route to segmentation in \gls*{dmri} space is the mapping of the
  structural information extracted from anatomical MR images, such as \gls*{t1}, using
  image registration techniques.
Generally, intra-subject registration of MR images of the brain involves only a linear
  mapping to compensate for head motion between scans.
However, \gls*{epi} introduces a geometrical distortion \citep{jezzard_correction_1995}
  that impedes the linear mapping from the structural space.
Numerous methods have been proposed to overcome this problem by incorporating information
  from extra MR acquisitions such as fieldmaps \citep{jezzard_correction_1995},
  \glspl*{dwi} with a different \gls*{pe} scheme \citep{cordes_geometric_2000,chiou_simple_2000},
  or \gls*{t2} images \citep{kybic_unwarping_2000,studholme_accurate_2000}.
These methods estimate the deformation field associated to \emph{\gls*{epi} distortions} and
  resample the \glspl*{dwi} onto a corrected \gls*{dmri} space.
The retrospective \gls*{epi} correction is an active field of research yielding frequent refinements
  and combinations of the original methods, such as
  \citep{holland_efficient_2010,andersson_comprehensive_2012,
  irfanoglu_drbuddi_2015}.
A standardized method to solve the remaining linear mapping between the corrected-\gls*{dmri}
  and the structural spaces is \emph{bbregister} \citep{greve_accurate_2009}.

Here, we present a segmentation and surface-to-volume registration method
  called \regseg{}, and show its usefulness in mapping anatomical information from structural
  space into native \gls*{dmri} space to aid subsequent processing steps
  (reconstruction, tractography and network building using a cortical parcellation).
The underlying hypothesis is that the registration and segmentation problems
  in \gls*{dmri} can be solved simultaneously.
To implement \regseg{} we first establish an active-contours without edges
  \citep{chan_active_2001} segmentation framework.
A specific set of reference surfaces extracted from the same subject initialize
  the 3D active contours, which evolve searching for homogeneous regions in the multivariate
  target-image.
We apply \regseg{} to segment \gls*{dmri} data by mapping
  a set of nested surfaces extracted from a structural image (e.g. \gls*{t1}) to
  a bivariate target-volume comprehending the \gls*{fa} and \gls*{adc} maps.
The evolution of the surfaces is supported by a B-spline basis, optimized
  iteratively using a descent approach driven by shape-gradients
  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
Therefore, \regseg{} establishes a registration framework that actually
  deals with the nonlinear warping induced by \gls*{epi} distortions.
\Regseg{} integrates the benefits of segmentation and registration methods together and
  exploits the multivariate nature of \gls*{dmri} data to contribute in the proposed
  application on three key aspects:
  1) the surfaces are typically extracted from the \gls*{t1} image of the same subject,
    therefore \regseg{} does not require additional MR acquisitions to the minimal
    \gls*{dmri} protocol in order to estimate the deformation field;
  2) \revcomment[R\#1-C2]{    alternatively to the typical design of the processing flow, the information from the
    reference \gls*{t1} can be precisely mapped onto the distorted \gls*{dmri} space,
    avoiding the interpolation of the \glspl*{dwi} required by unwarping the diffusion data; and} 
  3) \regseg{} increases the geometrical accuracy of the overall process.
In this paper, we first verify the functionality of the method and the \regseg{}
  implementation using a set of digital phantoms, demonstrating the
  \revcomment[R\#3-C10]{subvoxel} accuracy in registration.
Then, we evaluate \regseg{} on real \gls*{dmri} datasets, using a derivation of our
  instrumentation framework \citep{esteban_simulationbased_2014} which simulates
  known and realistic \gls*{epi} distortions.
We also compare \regseg{} and a nonlinear registration method to map the \lowb{} to
  the corresponding \gls*{t2} image of the same subject.
This approach is the first step of the above-mentioned \gls*{t2b} correction
  methods.
We reproduce the settings and implementation of a widely used diffusion
  processing software \citep[\emph{ExploreDTI},][]{leemans_exploredti_2009}.
With this comparison, we demonstrate how \regseg{} achieves higher accuracy with the
  simultaneous registration and segmentation process.
 \section{Methods}\label{sec:regseg-methods}

\subsection{Registration framework and segmentation model}\label{sec:regseg-methods_map}
Let $\gammaset_R = \{\Gamma_m: m \in \mathbb{N}, m \leq N_S\}$ be the set of $N_S$ surfaces
  extracted from the undistorted \gls*{t1} image (the reference space $R$).
We reformulate the segmentation of the distorted \gls*{dmri} images (the moving space $M$)
  as a registration problem where we search for an underlying deformation field $U$ such that
  the structures in $R$ defined by $\gammaset_R$ align optimally with their corresponding
  structures in $M$:

  \begin{align}
  U\colon R \subset \mathbb{R}^n &\to M \subset \mathbb{R}^n \notag\\
  \vec{r} &\mapsto \vec{r}' =\vec{r}+\vec{u}(\vec{r}),
  \label{eq:regseg-transform}
  \end{align}
  where $\vec{r}$ denotes a position in $R$, $\vec{r}'$ is
  its corresponding location in $M$, and $n$ denotes the dimensionality of images.
Finally, $\vec{u} = \vec{u}(\vec{r})$ is the displacement of every point with respect
  to the reference domain.
The general overview of how the surfaces interact with the registration framework
  is presented in \autoref{fig:regseg-method}.

\begin{figure}
  \includegraphics[width=\linewidth]{figure01}
  \caption{The interfacing surfaces $\Gamma_m$ between the competing \glspl{roi} $\Omega_l$,
  play the role of active contours which drive the registration process.
  They evolve iteratively along the normal $\hat{\mathbf{n}}_i$ of the surface at each vertex
    $\vec{v}_i$ of the mesh.
  The gradient speeds $\bar{s}_i$ drive registration, which are computed as the disparity of the data
    energies with respect to the two limiting regions of $M(\vec{v}_i)$, the features of the image
    $M$ in the location of vertex $\vec{v}_i$.
  The computation of shape-gradients is developed in \ref{app:shape_gradients}.
  In this figure, the $\bar{s}_1$ derived from \autoref{eq:regseg-shape_gradient_disc2} is 
    written in the lower box, with $\Omega_{wm}$ being the inner limiting region, $\Omega_{gm}$
    the outer region, and $w_{0,1}$ the relative area associated with vertex $\vec{v}_1$
    with respect to the total area of surface $\Gamma_0$.
  }\label{fig:regseg-method}
\end{figure}

\paragraph*{Cost-function derivation}
In a Bayesian framework for registration \citep{wyatt_map_2003,pohl_bayesian_2006,gass_simultaneous_2014},
  the mappings $U$ in \eqref{eq:regseg-transform} are
  evaluated based on their posterior probability given the observed data
  $M$.
Let $\omegaset = \{\Omega_l: l \in \mathbb{N}, l \leq N_L\}$ be the set of $N_L$ competing regions in
  which $M$ is partitioned by the projection of $\gammaset_R$.
Using Bayes' rule, the posterior likelihood is computed as:

  \begin{equation}
  P(U \mid M, \omegaset ) = \frac{P(M \mid U,\omegaset )\, P(U)}{P(M)},
  \label{eq:regseg-bayes_rule}
  \end{equation}
  where $P(M \mid U,\omegaset)$ is the data likelihood.
Since $\omegaset$ is mapped by $U$, we simplify
  $P(U, \omegaset) = P(U) \implies P(M \mid U,\omegaset) = P(M \mid U)$.
The best estimate $\hat{U}$ then satisfies the maximum a posteriori criterion
  and it aligns $\gammaset_R$ into $M$.
First, we assume independence between \revcomment[R\#3-C10]{voxel}s, and thus we break down the
  global data likelihood into a product of \revcomment[R\#3-C10]{voxel}-wise conditional probabilities:

  \begin{equation}
  P(M \mid U) = \underset{l}{\prod} \underset{\vec{r}'\in \Omega_l}{\prod}
    P\left( \vec{f}' \mid U \right),
  \label{eq:regseg-bayes_aposteriori}
  \end{equation}
  where $\vec{f}' = M(\vec{r}')$ is the feature vector at the displaced
  position $\vec{r}'$ \eqref{eq:regseg-transform} in the moving image.
For convenience and because it has been shown to be an appropriate approximation
  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
  region $\Omega_l$:
  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
  distributions 
  $\mathcal{N} (\vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} )$, with parameters
  $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
  for each region $\Omega_l$ \citep{esteban_mbis_2014}:

  \begin{align}
  P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r}' \in \Omega_l}{\prod}
  \mathcal{N} (\vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} )
  \ifthenelse{\boolean{review}}{=}{= \notag\\ &=}
  \underset{l}{\prod} \underset{\vec{r}' \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
  \mdist{f'}{l} \right)}},
  \label{eq:regseg-pdf}
  \end{align}  using $\mdist{f'}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}'$ with respect
  to the descriptors of region $l$ as
  $\mdist{f'}{l} = (\vec{f}' - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f}' - \boldsymbol{\mu}_l)$.
$C$ is the number of channels comprised in the image $M$.
Even though the features being segmented are not generally i.i.d., the 
  spatial interdependency of voxels is implicitly supported by the piecewise smooth
  partition of the space \omegaset{}.
In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
  the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized.
The \autoref{fig:regseg-model} shows how the joint distribution of 
  the input images is approximated with a mixture of multivariate normal distributions,
  and this minimization is illustrated for the segmentation of the \gls*{fa} and the \gls*{adc} maps
  of one subject.

\begin{figure*}
\includegraphics[width=\textwidth]{figure02}
\caption{Evolution of the segmentation model defined by the homogeneous regions $\Omega_l$,
  for one real dataset.
Panel (A, left) shows the joint distribution of the \gls*{fa} and \gls*{adc} conditioned to
  the segmentation \omegaset{} defined by the surfaces $\gammaset_R$ extracted from
  the \gls*{t1} image.
The plot was generated for reference using undistorted diffusion data, and
  therefore, $\gammaset_R$ is aligned with the \gls*{fa} and the \gls*{adc}.
The problem arises when the diffusion data present deformation, and the
  contours $\gammaset_R$ do not fit within the data (A, center).
After registration with \regseg{}, the contours are mapped onto the
  diffusion data (A, right), and the joint density plot is closer to the reference
  situation.
In panel (B), the three plots in (A) are decomposed tissue-wise.
Using filled contours, the bivariate distribution of each tissue is highlighted in its designated color,
  and represented over the remaining tissues (in gray colors).
To help assessment, dashed contours in black-to-white colors represent
  the corresponding distribution in the reference plot.
The registration process optimizes the segmentation model of \regseg{},
  and thus, the distribution of each region after registration is located
  closer to that corresponding in the reference situation, the shape of the distribution
  is more similar to the reference, and their spread is also reduced.
The effects of optimization are more noticeable on the \gls*{gm} ($\Omega_\text{GM}$) and
  the \gls*{wm} ($\Omega_\text{WM}$).
Particularly, the \gls*{wm} typically shows a bimodal distribution when the contours
  \gammaset{} do not fit the data.
{\color{revcolor} The plots in (A) and (B) are provided at full-size in the
  \citetalias[figures S8, S9]{esteban_useful_2016}.}
}\label{fig:regseg-model}
\end{figure*}

\paragraph*{Regularization}
The smoothness of the resulting displacement field is induced
  by a Thikonov regularization prior:

  \begin{align}
  P(U) = \underset{\vec{r}}{\prod}\, p(\vec{u}) &=
  \underset{\vec{r}}{\prod}\, p_0(\vec{u}) \, p_1(\vec{u}), \text{ with} \\
  p_0(\vec{u}) &= \mathcal{N}( \vec{u} \mid 0, \mathbf{A}^{-1}), \notag\\
  p_1(\vec{u}) &= \mathcal{N}(  \nabla \vec{u} \mid 0, \mathbf{B}^{-1}),
  \label{eq:regseg-thikonov}
  \end{align}
 which requires that the distortion and its gradient have zero mean,
   and variance governed by the matrices $\mathbf{A}$ and $\mathbf{B}$.
Therefore, $\mathbf{A}$ and $\mathbf{B}$ are tensors that modulate the regularization, and
  produce deformations with preferential directions.
Finally, the maximum a posteriori problem is adapted to a variational problem where we search for
  the minimum of an energy functional by applying $E(\vec{u}) = -\log \{P( M \mid U) \, P(U)\}$:

  \begin{align}
  E(\vec{u}) &= -\log \underset{l}{\prod}
  \underset{\vec{r}' \in \Omega_l}{\prod}
  \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) =
\notag\\ &= - \underset{l}{\sum} \int_{\Omega_l} \Big\{ \log \left[ \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right) \right] \, + \log \left[ p_0( \vec{u})\,p_1( \vec{u}) \right] \Big\} \, d\vec{r}'
   = \notag\\ &=
  \const + \underset{l}{\sum} \Big\{ \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \Big\} \,
  + \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}'.
  \label{eq:regseg-energy}
  \end{align}This expression is the dual of the Mumford-Shah functional that corresponds
  to the framework of \acrlong*{acwe} \citep{chan_active_2001}
  with the anisotropic regularization term of \cite{nagel_investigation_1986}.

\subsection{Numerical Implementation}
\label{sec:regseg-numerical_implementation}

\paragraph*{Deformation model}\label{sec:regseg-deformation_model}
Since the vertices of the surfaces $\{\vec{v}_i: \vec{v}_i \subset \gammaset \}_{i=1 \ldots N_V}$
  are probably located off-grid, it is necessary to derive $\vec{u}_i = \vec{u}(\vec{v}_i)$ from a discrete set of parameters
  $\{\vec{u}_k\}_{k=1 \ldots K}$.
Densification is achieved using a set of associated basis functions $\psi_k$ \eqref{eq:regseg-nodes_tfm}.
In our implementation, $\psi_k$ is a tensor-product B-spline kernel of degree three.

  \begin{equation}
  \vec{v}_i' = \vec{v}_i + \vec{u}_i = \vec{v}_i + \sum_k \psi_k(\vec{r}) \: \vec{u}_k.
  \label{eq:regseg-nodes_tfm}
  \end{equation}

\paragraph*{Optimization}
\label{sec:regseg-gradient_descent}
To find the minimum of the energy functional \eqref{eq:regseg-energy},
  we propose a gradient-descent approach with respect to the underlying
  deformation field using the following \gls*{pde}:

  \begin{equation}
  \frac{\partial \vec{u}(\vec{r},t)}{\partial t} \propto - \frac{\partial E(\vec{u})}{\partial \vec{u}_k},
  \label{eq:regseg-general_gradient_descent}
  \end{equation}
  where $t$ is an artificial time parameter of the contour
  evolution and $\vec{u}_k$ are the parameters that support the estimate
  $\hat{U}$ of the transformation at the current time point.
Let us assume that the preferential directions of
  the displacement are aligned with the imaging axes to simplify \eqref{eq:regseg-energy} as expression 
  \eqref{eq:regseg-app_energy} in \ref{app:reg_term}, and thus to compute its
  derivative \eqref{eq:regseg-general_gradient_descent}:

  \begin{align}
  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \Big\{
  \underset{l}{\sum} \big[ \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r}' \big] \,
  + \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}'
  \Big\},
  \label{eq:regseg-gradient_descent}
  \end{align}
  where $\vec{u}^{\circ2} = \vec{u}^T \cdot \vec{u}$,
  and $\{\boldsymbol{\alpha}, \boldsymbol{\beta}\}$ are the expected variances along
  the imaging axes of the displacement field and its gradient, respectively.
Then, the data and regularization terms are split and discretized to compute their
  derivatives.
The derivative of the data term is computed using explicit shape gradients (see \ref{app:shape_gradients}),
  which finally lead to obtain vertex-wise speeds of the gradient $\bar{s}_i$ as
  illustrated in \autoref{fig:regseg-method}.
\revcomment[R\#1-C4]{The shape gradient contributions $\vec{g}_k$ on the field coefficients $\vec{u}_k$
  can then be computed }using the expression \eqref{eq:regseg-gradient_wshape}
\revcomment[R\#1-C4]{of \ref{app:shape_gradients}, obtaining:

\begin{equation}
  \vec{g}_k = - \underset{i}{\sum} \bar{s}_i \cdot \psi_k(\vec{v}_i) \, \hat{\vec{e}}.
  \label{eq:regseg-shape_gradient_final}
\end{equation}

Then,} introducing the analytical derivative of the regularization term,
  the \autoref{eq:regseg-gradient_descent} is reformulated as:

  \begin{align}
  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
  \vec{g}_k  + \boldsymbol{\alpha} \cdot \vec{u}_k - \boldsymbol{\beta} \cdot (\Delta \vec{u}_k).
  \label{eq:regseg-final_gradient}
  \end{align}
Finally, to descend this gradient, we establish a semi-implicit Euler scheme
  \citepalias[see][section S1.3]{esteban_useful_2016},
  with a step size parameter $\delta$, which we solve in the spectral domain as follows:

  \begin{align}
  \vec{u}_k^{t+1} = \mathcal{F}^{-1}\left\{ \frac{\mathcal{F}\{\delta^{-1} \, \vec{u}_k^t - \vec{g}_k\} }
                    {\mathcal{F}\{(\delta^{-1} + \boldsymbol{\alpha})\, I - \boldsymbol{\beta}\Delta\}} \right\},
  \label{eq:regseg-update_equation}
  \end{align}
  where $I$ denotes the identity operator.

\paragraph*{Implementation details and convergence}\label{sec:regseg-conv_report}
The \regseg{} tool includes a multiresolution strategy on the free-form deformation field.
Registration pyramids are created by setting the spacing between the control points of the B-spline basis
  functions for each level of the multiresolution strategy.
As a rule of thumb, for a $\delta =$ 1.0, both $\boldsymbol{\alpha}$
  and $\boldsymbol{\beta}$ will typically be in the range [0.0, 1.0].
The parameters used ($\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, the B-spline grid resolutions,
  and target image smoothing), the implementation details, and other features such as the sparse matrix
  approach to fast interpolation are discussed in the \citetalias[][section S1]{esteban_useful_2016}.

\subsection{Evaluation protocol}\label{sec:regseg-evaluation_protocol}
In order to assess the performance of \regseg{}, we defined the following general
  evaluation protocol:
1) Extract the set of undistorted surfaces $\gammaset_R$;
2) Compute a ground-truth field of displacements $U_\text{true}$, which is applied to
  generate warped images ($M$) for segmentation;
3) Execute \regseg{} with $\gammaset_R$ and use the warped data as inputs; and
4) Perform a visual assessment and compute the error metrics.

A first proof of concept is introduced to demonstrate \regseg{} in digital phantoms
  with simple geometries, using $U_\text{true}$ without directional restrictions.
Then, \regseg{} is evaluated in a framework using undistorted \gls*{dmri} datasets,
  and $U_\text{true}$ is derived from the corresponding inhomogeneity fieldmap of
  the subject.
Therefore, the deformation field is nonzero only in the \acrfull*{pe} axis,
  and reproduces a real \gls*{epi} distortion.
The adaptation of the evaluation protocol to the simulated phantoms and the real data is
  explained in the following sections.

\subsection{Simulated phantoms}\label{sec:regseg-digital_phantoms}
The workflow required to simulate the digital phantoms and to assess the performance of
  \regseg{} with them is presented in \autoref{fig:regseg-evphantoms}.
A set of four binary objects (i.e. ``box'', ``ball'', ``L'',
  and ``gyrus'') was generated by combining the binarization of
  analytical shapes and mathematical morphology.
The reference surfaces $\gammaset_R$ were extracted from the binary shapes
  using \emph{FreeSurfer} tools \citep{fischl_freesurfer_2012}.
The ground-truth distortion was generated using a chain of two displacement
  fields supported by grids of B-spline basis functions.
The coefficients of the basis functions were generated randomly for
  both levels in their three dimensions.
The three components of the displacements $\vec{u} = (u_d)$
  were bounded above by 40\% of the separation between the control points
  at each level to obtain diffeomorphic transforms
  after concatenation \citep{rueckert_diffeomorphic_2006}.
The first deformation field was applied to generate large warpings
  with control points separated by 50.50 mm in the three dimensions
  ($u_d\leq$ 20.20 mm).
With the second warping, we aimed to obtain a field with smoothness
  close to that found in a typical distortion field of \gls*{dmri} data
  \citep{irfanoglu_susceptibility_2011}.
Therefore, the control points were separated by 25.25 mm ($u_d\leq$ 10.10 mm).
After generating the ground-truth deformation, the original surfaces
  were warped by interpolating the displacement field at each vertex.
The warped surfaces $\gammaset_\text{true}$ were binarized to generate tissue fractions
  at low (\isores[mm]{2.0}) and high (\isores[mm]{1.0}) resolutions.
Using an \acrshort*{mri} simulator \citep{caruyer_phantomas_2014}, we synthesized
  \gls*{t1} (TE/TR= 10/1500 ms) and \gls*{t2} images (TE/TR= 90/5000 ms), which
  corresponded to each phantom type, with each at two resolutions
  (1.0 mm and 2.0 mm isotropic).
The field of view at both resolutions was \isores[mm]{100}.
Next, \regseg{} was applied to map $\gammaset_R$ onto the warped phantoms to
  obtain the registered surfaces ($\hat{\gammaset}_\text{test}$).
To quantify the misregistration error, we computed the Hausdorff distance between
 $\hat{\gammaset}_\text{test}$ and $\gammaset_\text{true}$ using \citep{commandeur_vtk_2011}.
In total, 1200 experiments (four phantom types $\times$ 150 random warpings $\times$ two resolutions) were
  performed according to the workflow illustrated in \autoref{fig:regseg-evphantoms}.

\begin{figure*}
\includegraphics[width=\linewidth]{figure03}
\caption{Evaluation of \regseg{} using phantom data according to the following instrumental workflow.
  1) The reference surfaces $\gammaset_R$ are triangle meshes extracted from the four binary shapes (i.e., ``box'', ``ball'', ``L'', ``gyrus'').
  2) A ground-truth displacement field was generated as described in \autoref{sec:regseg-digital_phantoms}, and applied to warp
      $\gammaset_R$, thereby obtaining $\gammaset_\text{true}$.
  3) After being warped, $\gammaset_\text{true}$ were projected onto the corresponding discrete 3D volume and downsampled to create partial volume effects at two resolutions,
     i.e., \isores[mm]{2.0} and \isores[mm]{1.0}, thereby producing sets of tissue fractions maps.
  4) The tissue fractions were fed into an \acrshort*{mri} simulator, which generated \acrfull*{t1} and \acrfull*{t2} -like images at the
     two possible resolutions.
  5) The \regseg{} tool was applied using the warped test images as multispectral moving images and $\gammaset_R$ as shape priors.
  6) The agreement between the surfaces fitted by \regseg{} ($\hat{\gammaset}_\text{test}$) and $\gammaset_\text{true}$ were assessed
     visually using automatically generated visual reports and quantitatively with the Hausdorff distance between the
     corresponding surfaces.}\label{fig:regseg-evphantoms}
\end{figure*}

\paragraph*{Segmentation model and settings}
The segmentation model of the phantoms is implicitly defined:
  all phantoms comprehend an inner surface enclosing a uniform \gls*{wm}-like region,
  and an outer surface wrapping a \gls*{gm}-like layer.
The outside is filled with uniform background (see \autoref{fig:regseg-evphantoms}).
All the experimental settings used for the phantoms are made available in
  a unique configuration file\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/regseg_default.json}}.

\subsection{Real datasets} \label{sec:regseg-human_connectome}
The experimental framework for the real datasets is presented in \autoref{fig:regseg-evworkflows},
which extends our previous evaluation \citep{esteban_simulationbased_2014} of distortions
  using \gls*{dmri} phantoms.

\begin{figure*}
\includegraphics[width=\linewidth]{figure04}
\caption{Experimental workflow employed to process real data from the \acrfull*{hcp}.
  1) $\gammaset_R$ were extracted from the anatomical reference (\gls*{t1} image).
  2) For use as the ground truth, we generated a plausible synthetic distortion $U_\text{true}$
    from the fieldmap with \eqref{eq:regseg-fieldmap}.
  3) The \gls*{dmri} data were warped using $U_\text{true}$ to reproduce the effects of real
    susceptibility-derived distortions.
  Target diffusion scalars (\gls*{fa} and \gls*{adc}) were computed with the distorted data and
    stacked to feed the multivariate input required by \regseg{}.
  4) The method was run to obtain $U_\text{test} = \hat{U}_\text{true}$, i.e., the estimate of
    the ground-truth deformation.
  5) The results were evaluated visually and quantitatively.
  The arrows in the step (5) point to edges in the target images (light-yellow arrows for \gls*{fa}, blue for 
    \gls*{adc}) that should be aligned with a surface, showing how distortion limits the direct mapping from
    the structural space in which the contours are defined.}\label{fig:regseg-evworkflows}
\end{figure*}

\paragraph*{Data}
To evaluate \regseg{} using real \gls*{dmri} data obtained from human brains,
  we collected 16 subjects from the \revcomment[R\#3-C9]{Q3 Release of the}
  \gls*{hcp} database.
The original acquisitions are released within ``unprocessed'' packages, whereas
  the ``minimally preprocessed'' packages contain the corresponding images after
  some processing (correction for several artifacts, brain-extraction, spatial
  normalization, etc.).
We refer the reader to \citet{essen_human_2012} for exact details of the acquisition
  parameters and \citet{glasser_minimal_2013} for the preprocessing issues.
These datasets comprise a large set of images, including \gls*{t1}, \gls*{t2}, and
  multi-shell \gls*{dmri} images.
Since we obtained the \gls*{dmri} data from the minimally preprocessed package, these
  images are corrected for \gls*{epi} distortions and spatially normalized in
  \gls*{t1} space.
\revcomment[R\#3-C9]{In the Q3 Release of the \gls*{hcp}, the \gls*{dmri} session includes
  six runs, two runs for each of three different gradient tables, and each table is acquired
  once with right-to-left and left-to-right encoding polarities, respectively.
Then, the diffusion datasets with opposed polarities are corrected for susceptibility
  distortions using the \emph{TOPUP} tool \citep{andersson_how_2003}
  before producing the released ``minimally preprocessed'' data}.

\paragraph*{Segmentation model}
Based on our experience   and previous studies \citep{ennis_orthogonal_2006},
  we defined the moving image as a stack of the \gls*{fa} and \gls*{adc} maps derived
  from \gls*{dmri} data.
After evaluating several alternative models, we empirically defined a partition \omegaset{}
  according to the following six regions:
  1) thalamus ($\Omega_\text{Tha}$);
  2) ventricular system and deep \gls*{gm} structures ($\Omega_\text{VdGM}$);
  3) cerebral \gls*{wm} ($\Omega_\text{WM}$);
  4) brain stem and cerebellar \gls*{wm} ($\Omega_\text{bst}$);
  5) cerebellar \gls*{gm} ($\Omega_\text{cbGM}$); and
  6) cortical \gls*{gm} ($\Omega_\text{GM}$).
Using tools in \emph{FreeSurfer} and appropriate selections of labels in the
  \emph{aparc} segmentation released with the \gls*{hcp} data, we extracted the $\gammaset_R$ set for the
  reference surfaces.
The segmentation model corresponding to this partition is shown in \autoref{fig:regseg-model}
  and discussed in greater detail in the \citetalias[section S4]{esteban_useful_2016}.

\paragraph*{Ground-truth generation}
Realistic deformation was achieved by generating displacement fields that satisfy the theoretical
  properties of distortion.
The displacements along the \gls*{pe} axis of the \gls*{dmri} image are related to the local
  deviation of the field $\Delta B_0(\vec{r})$ from its nominal value $B_0$  \citep{jezzard_correction_1995},
  as follows:

  \begin{equation}
  u_\text{PE} = \frac{\gamma \, T_{acq}\, s_\text{PE}}{2\pi}\Delta B_0(\vec{r})\text{ [mm]},
  \label{eq:regseg-fieldmap}
  \end{equation}
where $\gamma$ is the gyromagnetic ratio, $T_{acq}$ is the readout time, and
  $s_\text{PE}$ is the \revcomment[R\#3-C10]{voxel} size along \gls*{pe}.
Certain \acrshort*{mri} sequences are designed to estimate $\Delta B_0$, thereby obtaining
  the so-called fieldmap.
We derived the deformation $U_\text{true}$ from the fieldmap image released with
  the corresponding packages of each dataset in the \gls*{hcp}.
The fieldmap was unwrapped\footnote{fieldmaps are phase maps, which are intrinsically clipped in the interval
  of $[-\pi, \pi)$ [rads] or [rads/s].} and smoothed before applying \eqref{eq:regseg-fieldmap}.
Next, the original \gls*{dmri} was warped using the resulting displacement field and fed into
  a pipeline to process the corresponding \gls*{dti}, thereby computing the derived scalars of
  interest (\gls*{fa} and \gls*{adc}) using \emph{MRtrix} \citep{tournier_mrtrix_2012}.

\paragraph*{Metric assessment}
Initially, we investigated the appropriateness of the segmentation model.
For five test datasets, we uniformly sampled the space of distortions
  $\hat{U} = \epsilon \cdot U_\text{true} = \vec{r} + \epsilon \, u_\text{PE}$
  (with $\epsilon \in [-1.1, 1.1]$ and $u_\text{PE}$ from \eqref{eq:regseg-fieldmap}),
  and we evaluated the data term of the cost function \eqref{eq:regseg-energy}.
The minimum of the cost function (\autoref{sec:regseg-methods_map}) was consistently located at
  $\epsilon=0.0$ (the ground-truth) for all of the cases \citepalias[figure S2]{esteban_useful_2016}.

\paragraph*{Settings} \Regseg{} accepts an affine mapping from surface-space to the \gls*{dmri} data as initialization.
However, the images provided by the \gls*{hcp} are already spatially normalized.
Therefore, the initial estimation of distortion is zero in this experiment.
Since the distortion $U_\text{true}$ is aligned along the \gls*{pe} direction ($y$-axis in our settings),
  \regseg{} was configured to allow nonzero displacements only on that corresponding direction.
For the experiments on real data, \regseg{} established a multi-resolution pyramid of
  B-spline functions, with control points distributed on grids of the following spacings:
  40$\times$100$\times$40 [mm] for the first (coarser) level, \isores[mm]{30} for the second level,
  and 20$\times$30$\times$10 [mm] in the third level.
Only the first and second levels included Gaussian smoothing of the target image ($\sigma$ = [2.0, 0.5] mm,
  respectively).
The actual choices of the parameter settings are publicly distributed with the source code for the
  experiments\footnote{\url{https://github.com/oesteban/RegSeg/tree/master/Scripts/pyacwereg/data/regseg_hcp.json}}.
These settings were obtained manually, driven by the feedback obtained from the post-registration
  convergence reports \citepalias[like that found in][section S1.3]{esteban_useful_2016}.
We released \regseg{} along with the tool to generate such convergence reports.

\paragraph*{Cross-comparison} A dual workflow to the general evaluation used for \regseg{} (\autoref{fig:regseg-evworkflows}),
  was employed to integrate the alternate \gls*{t2b} registration scheme.
We reproduced the solution and settings provided with \emph{ExploreDTI}
  \citep{leemans_exploredti_2009}, which is a widely used toolkit for tractography analysis of
  \gls*{dti}.
\emph{ExploreDTI} internally employs \emph{elastix} \citep{klein_elastix_2010} to
  perform registration.
The deformation field is correspondingly restricted to the \gls*{pe} direction.
The settings file for \emph{elastix} is also
  available\footnote{\url{https://github.com/oesteban/RegSeg/blob/master/Scripts/pyacwereg/data/t2b_elastix_y.txt}}.
\revcomment[R\#3-C7\\R\#3-C8]{In this registration scheme, the \gls*{t2} image is the \emph{reference} and
  the \lowb{} plays the role of \emph{moving} image.
Therefore, the transform is defined in the coordinate system of the \gls*{t2} image,
  and for each point in this space it provides the location of the corresponding feature
  in the \lowb{} image.
Since the surfaces are defined in the \gls*{t2} --\emph{reference}-- space, their coordinates
  can be mapped to the \lowb{} space using this transform, obtaining the 
  \emph{distorted surfaces} corresponding to the \lowb{}-to-\gls*{t2} registration.}

\paragraph*{Error measurement}\label{sec:regseg-experiments_evaluation}
Distortion only occurs along the \gls*{pe} axis of the image, so we computed the
  \gls*{swindex} as the area-weighted distance between the corresponding vertices of
  $\gammaset_\text{true}$ and their estimate obtained by the method under the test $\hat{\gammaset}_\text{test}$:

  \begin{equation}
  sWI = \frac{1}{\sum_i a_i} \sum\limits_i^{N_V} a_i\,\|
  \vec{v}_i - \hat{\vec{v}}_i \|,
  \label{eq:regseg-swindex}
  \end{equation}
  where $\vec{v}_i \subset \gammaset_\text{true}$ are the locations of the total $N_V$ vertices,
  $a_i$ is the area corresponding to each vertex $\vec{v}_i$,
  and $\hat{\vec{v}}_i \subset \hat{\gammaset}_\text{test}$ are the recovered locations
  that correspond to $\vec{v}_i$.
In practice, we only report the \gls*{swindex} for three surfaces (\{$\Gamma_\text{VdGM}, \Gamma_\text{WM}, \Gamma_\text{pial}$\})
  of crucial interest in whole-brain tractography.
\revcomment[R\#3-C7\\R\#3-C8]{The \gls*{swindex} is always computed on the \gls*{dmri} space.}   \section{Results}
\label{sec:regseg-results}

\subsection{Verification and validation using digital phantoms}
\label{sec:regseg-results_phantom}

The results summarized in \autoref{fig:regseg-phantom} demonstrated that the accuracy was
  high and below the image resolution.
Panel B on \autoref{fig:regseg-phantom} shows the violin plots for each model type corresponding
  to the two sets of resolutions for the generated phantoms.
In order to relate the average misregistration error to the resolution of the moving image,
  we proceeded as follows.
First, we confirmed that the vertex-wise error distributions were skewed by using Shapiro-Wilk's test of
  normality.
All of the distributions of errors in the tests (four phantom types $\times$ two resolutions) were
  nonnormal with $p<$ 0.001.
Consequently, we used the nonparametric Wilcoxon signed-rank test with the Bonferroni
  correction for multiple comparisons ($N$=150, for each phantom type).
The average errors were significantly lower than the voxel size with $p <$ (0.001 / 150)
  in all tests (four phantom types $\times$ two resolutions).
Statistical tests might not be sufficiently conclusive, so we also computed the confidence intervals,
  as shown in \autoref{tab:ci_phantom}.

\begin{figure*}[!h]
  \centering
  \includegraphics[width=\linewidth]{figure05}
  \caption{A. Visual assessment of the results obtained on the digital phantoms.
    The panel shows three coronal slices (indices indicated on the lower left corner) of each phantom
    volume at low resolution:
    ``gyrus'' (top left), ``L'' (top right), ``ball'' (bottom left),
    and ``box'' at (bottom right).
  The contours recovered after registration are represented in yellow.
  \Regseg{} achieved high accuracy because it determined the almost exact locations of the registered
    contours with respect to their ground truth positions (shown in green).
  The partial volume effect makes segmentation of the sulci a challenging problem with voxel-wise
    clustering methods, but they were successfully segmented with \regseg{}.
  B. Quantitative evaluation: the violin plot shows the variability across experiments
    of the average Hausdorff distance measured in each vertex of the corresponding surface, for the
    low (left) and high (right) resolutions. 
    Error averages were consistently below the size of the voxel.
    }\label{fig:regseg-phantom}
\end{figure*}
\begin{figure*}[!h]
  \centering
  \includegraphics[width=\linewidth]{figure06}
  \caption{A. Example of a visual assessment report, which was automatically generated by the evaluation tool.
    Each view shows one component of the input image (in this case, the \gls*{fa} map), the ground-truth locations
    of the surfaces (green contours), and the resulting surfaces obtained with the test method (yellow contours).
  The first two rows show axial slices for \regseg{} and the \acrfull*{t2b} method, while the last two rows
    show the corresponding sagittal views.
  The coronal view is omitted because it was the least informative due to the directional property
    of the distortions.
  Specific regions where \regseg{} outperformed \gls*{t2b} are enlarged.
  B. Violin plots of the error distributions for each surface across the 16 subjects, which show the 
    voxel size of the \gls*{dmri} images (1.25 mm), thereby supporting the improved results obtained
    by \regseg{} with the proposed settings.
  }\label{fig:regseg-results_real}
\end{figure*}

\subsection{Evaluation using real datasets and cross-comparison}\label{sec:regseg-results_hcp}
Finally, we compared the performance of \regseg{} with that of the standard \gls*{t2b}
  method.
Summary reports for visual assessment of the 16 cases are included in the
  \citetalias[section S5]{esteban_useful_2016}.
In \autoref{fig:regseg-results_real}, box A, the visual report is shown for one subject.
We computed the \gls*{swindex} \eqref{eq:regseg-swindex} of every surface after registration
  using both the \regseg{} and \gls*{t2b} methods.
Finally, to compare the results, we performed Kruskal-Wallis H-tests
  (a nonparametric alternative to ANOVA) on the warping indices for the three surfaces of
  interest selected in \autoref{sec:regseg-experiments_evaluation}
  ($\Gamma_\text{VdGM}$, $\Gamma_\text{WM}$, $\Gamma_\text{pial}$).
All of the statistical tests showed that the error distributions obtained with \regseg{} and
  \gls*{t2b} were significantly different, and the violin plots in box B of
  \autoref{fig:regseg-results_real} demonstrate that the errors were always larger with \gls*{t2b}.
We also show the 95\% CIs of the \gls*{swindex} for these surfaces 
  (\autoref{tab:results_real}).
The aggregate CI for \regseg{} was 0.56--0.66 [mm], whereas the \gls*{t2b} method
  yielded an aggregate CI of 2.05--2.39 [mm].
The results of the statistical tests and the CIs are summarized in \autoref{tab:results_real}.

\begin{table}
\centering
\footnotesize
\tabcolsep=0.1cm
\begin{minipage}{.48\textwidth}
    \vskip-4.7ex
    \begin{tabular}{lccccc}
    Res.   & ``gyrus'' & ``ball''  & ``box''   & ``L''     & Aggreg.    \\\hline
    1.0mm  & 0.18--0.38 & 0.31--0.45 & 0.34--0.42 & 0.34--0.40 & 0.34--0.38  \\
    2.0mm  & 0.59--0.60 & 0.65--0.76 & 0.68--0.71 & 0.67--0.77 & 0.64--0.66  \\
    \hline
    \end{tabular}
    \caption{The distributions of vertex-wise Hausdorff distances between the ground-truth surfaces and their
    corresponding estimates obtained with \regseg{} presented a 95\% CI below the half-voxel size for all of
    the phantom types.
    The CIs were computed by bootstrapping using 10$^\text{4}$ samples, with the median as the location statistic.}\label{tab:ci_phantom}
\end{minipage}
\hfill
\begin{minipage}{.48\textwidth}
    \begin{tabular}{cccccc}
    & & $\Gamma_\text{VdGM}$  & $\Gamma_\text{WM}$ & $\Gamma_\text{pial}$ & Aggreg. \\
    \hline
    \multirow{2}{*}{CI}
       & \regseg{} & 0.50--0.78 & 0.50--0.55 & 0.66--0.73 & 0.56--0.66 \\
       & T2B       & 1.78--2.58 & 1.94--2.36 & 2.16--2.58 & 2.05--2.39 \\
    \hline
    \multirow{2}{*}{H-tests}
       & p-value  & 4.1\e{-6} & 2.3\e{-6} & 2.3\e{-6} & 1.8\e{-16} \\
       & H-stat   & 21.20 & 22.31 & 22.31 & 67.85 \\
    \hline
    \end{tabular}
    \caption{Statistical analysis of results obtained using 16 real datasets from 
    the \gls*{hcp}, which show that \regseg{} performed better than the alternative \acrfull{t2b} method.
    The distribution of the errors computed for the surfaces of interest ($\Gamma_\text{VdGM}$, $\Gamma_\text{WM}$, $\Gamma_\text{pial}$)
      and the aggregate of all surfaces (Aggreg. column) had lower 95\% CIs with \regseg{}.
   The CIs in this table were computed by bootstrapping using the mean as the location statistic and with 10$^\text{4}$ samples.
    The Kruskal-Wallis H-tests indicated that there was a significant difference between the results obtained using \regseg{} and
      the \gls*{t2b} method.
    }\label{tab:results_real}
\end{minipage}
\end{table} \section{Discussion}
\label{sec:regseg-discussion}
We present \regseg{}, a simultaneous segmentation and registration method that
  maps a set of nested surfaces into a multivariate target-image.
The nonlinear registration process evolves driven by the fitness of the
  piecewise-smooth classification of voxels in the target volume imposed
  by the current mapping of the surfaces.
We propose \regseg{} to map anatomical information extracted from \gls*{t1}
  images into the corresponding \gls*{dmri} of the same subject.
Previously, joint segmentation and registration has been applied successfully to
  other problems such as longitudinal object tracking \citep{paragios_level_2003}
  and atlas-based segmentation \citep{gorthi_active_2011}.
The most common approach involves optimizing a deformation model (registration)
  that supports the evolution of the active contours (segmentation),
  like \cite{paragios_level_2003,yezzi_variational_2003}.
\Regseg{} can be seen as a particular case of atlas-based segmentation-registration methods,
  replacing the atlas by the structural image of the subject (\emph{structure-informed segmentation}).
The main difference of atlas-based segmentation and the application at hand is the resolution of the
  target image.
Atlas-based segmentation is typically applied on structural and high-resolution images.
A comprehensive review of joint segmentation and registration methods applied in atlas-based
  segmentation is found in \citep{gorthi_active_2011}.
They also propose a multiphase level-set function initialized from a labeled atlas to implement
  the active contours that drive the atlas registration.
Alternatively, \regseg{} implements the active contours with a hierarchical set of explicit
  surfaces (triangular meshes) instead of the multiphase level sets, and registration
  is driven by shape-gradients \citep{herbulot_segmentation_2006}.
As an advantage, the use of explicit surfaces enables segmenting \gls*{dmri} images
  with accuracy below \revcomment[R\#3-C10]{voxel} size.

An important antecedent of \regseg{} is \emph{bbregister} \citep{greve_accurate_2009}.
The tool has been widely adopted as the standard registration method to be used along with the \gls*{epi}
  correction of choice.
It implements a linear mapping and uses 3D active contours \emph{with edges} to
  search for intensity boundaries in the \lowb{} image.
The active contours are initialized using surfaces extracted from the \gls*{t1} using
  \emph{FreeSurfer} \citep{fischl_freesurfer_2012}.
To overcome the problem of nonlinear distortions, \emph{bbregister} excludes from the
  boundary search those regions that are typically warped.
Indeed, the distortion must be addressed separately because it is not supported by
  the affine transformation model.
Conversely, the deformation model of \regseg{} is nonlinear and the active contours are
  \emph{without edges} \citep{chan_active_2001} since the \gls*{fa} and \gls*{adc} maps
  do not present steep image gradients (edges) but the anatomy can be identified
  by looking for piece-wise smooth homogeneous regions.

Recently, \cite{guyader_combined_2011} proposed a simultaneous segmentation and
  registration method in 2D using level sets and a nonlinear elasticity smoother on the
  displacement vector field, which preserves the topology even with very large deformations.
\Regseg{} includes an anisotropic regularizer for the displacement field described by
  \cite{nagel_investigation_1986}.
This regularization strategy conceptually falls in the midway between the Gaussian smoothing
  generally included in most of the existing methodologies, and the complexity of
  the elasticity smoother of \cite{guyader_combined_2011}.
Other minor features that differ from current methods in joint segmentation and registration are
  the support of multivariate target-images and the efficient computation of the shape-gradients
  implemented with sparse matrices.

We verified that precise segmentation and registration of a set of surfaces into multivariate
  data is possible on digital phantoms.
We randomly deformed four different phantom models to mimic three homogeneous regions
  (\gls*{wm}, \gls*{gm}, and \acrlong*{csf}) and we used them to simulate \gls*{t1}
  and \gls*{t2} images at two resolution levels.
We measured the Hausdorff distance between the contours projected using the
  ground-truth warping and the estimations found with \regseg{}.
We concluded that the errors were significantly lower than the voxel size.
We also assessed the 95\% \gls*{ci}, which yielded an aggregate interval of
  0.64--0.66 [mm] for the low resolution phantoms (2.0 mm isotropic voxel) and
  0.34--0.38 [mm] for the high resolution phantoms (1.0 mm isotropic).
Therefore, the error was bounded above by half of the voxel size.
The distributions of errors along surfaces varied importantly depending on the shape of the
  phantom (see \autoref{fig:regseg-phantom}B).
The misregistration error of the ``gyrus'' phantom showed a much lower spread than that
  for the other shapes.
We argue that the symmetry of those other shapes posed difficulties in driving the contours
  towards the appropriate region due to \emph{sliding} displacements between the
  surfaces and their ground-truth position.
The effect is not detectable by the active contours framework, but it is controllable
  increasing the regularization constraints.
When \regseg{} is applied on real datasets, this surface sliding is negligible for the
  convoluted nature of cortical surfaces and the directional restriction of the
  distortion.

We evaluated \regseg{} in a real environment using the experimental framework presented
  in \autoref{fig:regseg-evworkflows}.
We processed 16 subjects from the \gls*{hcp} database using both \regseg{}
  and an in-house replication of the \acrfull*{t2b} method.
\Regseg{} obtained a high accuracy, with an aggregate 95\% \gls*{ci} of 0.56--0.66 [mm], which was
  below the \revcomment[R\#3-C10]{voxel} size of 1.25 mm.
The misregistration error that remained after \regseg{} was significantly lower ($p <$ 0.01) than the
  error corresponding to the \gls*{t2b} method according to Kruskal-Wallis H-tests
  (\autoref{tab:results_real}).
Visual inspections of all the results \citepalias[section S5]{esteban_useful_2016} and the violin plots in
  \autoref{fig:regseg-results_real} confirmed that \regseg{} achieved higher accuracy
  than the \gls*{t2b} method in our settings.
We carefully configured the \gls*{t2b} method using the same algorithm and the
  same settings employed in a widely-used tool for \gls*{dmri} processing.
However, cross-comparison experiments are prone to the so-called \emph{instrumentation bias}
  \citep{tustison_instrumentation_2013}.
Therefore, these results did not prove that \regseg{} \emph{is better than} \gls*{t2b},
  but indicated that \regseg{} is a reliable option in this application field.
Finally, we also proposed a piecewise-smooth segmentation model defined by
  a selection of nested surfaces to partition the multispectral space
  comprehending the \gls*{fa} and the \gls*{adc} maps and ultimately identify anatomical
  structures in \gls*{dmri} space.
We also demonstrated the smoothness of the objective function on five of the real datasets
  \citepalias[figure S2]{esteban_useful_2016}, taking advantage of the directional
  restriction of possible distortions.
However, \regseg{} requires densely sampled surfaces to ensure the convergence.
Using the digital phantoms, we severely decimated the surfaces by a large factor.
These surfaces introduced a bias which displaced the zero of the gradients from the
  minimum of the objective function impeding the convergence.

The proposed application of the method in the task of identifying structural information
  in \gls*{dmri} images is an active field of research \citep{jeurissen_tissuetype_2015}.
Current processing of \gls{dmri} involved in the connectome extraction and other applications
  (such as \gls*{tbss} or surgical planning) require a precise segmentation
  of the anatomical structures in the diffusion space.
Some examples of these processing tasks are the structure-informed reconstruction of \gls*{dmri}
  data \citep{jeurissen_multitissue_2014,daducci_accelerated_2015}, the anatomically constrained
  tractography \citep{smith_anatomicallyconstrained_2012}, and the imposition of the cortical
  parcellation mapped from the \gls*{t1} image \citep{hagmann_mapping_2008}.
The problem was firstly addressed using image segmentation approaches in the native diffusion
  space, without definite and compelling results.
With the introduction of retrospective correction methods for the \emph{\gls*{epi} distortions}
  and image registration approaches, the task has been typically solved in a two-step approach.
First, the \glspl*{dwi} are corrected for \emph{\gls*{epi} distortions} by estimating
  the nonlinear deformation field from extra MR acquisitions
  \citep{jezzard_correction_1995,chiou_simple_2000,cordes_geometric_2000,kybic_unwarping_2000}.
Second, mapping the structural information from the corresponding \gls*{t1} image
  using a linear registration tool like \emph{bbregister} \citep{greve_accurate_2009}.
The current activity on improving correction methods \citep{irfanoglu_drbuddi_2015} and
  the comeback of segmentation of \gls*{dmri} in its native space
  \citep{jeurissen_tissuetype_2015} proof the open interest of this application.
\Regseg{} addresses this joint problem in a single step and it does not require any additional
  acquisition other than the minimal protocol comprehending only \gls*{t1} and \gls*{dmri} images.
This situation is commonly found in historical datasets.

We envision \regseg{} to be integrated in diffusion processing pipelines, after a 
  \revcomment[R\#3-C10]{preliminary}
  \gls*{dti} computation and before anatomically-informed reconstruction and tractography
  methods.
Since the structural information is projected into the native space of \gls*{dmri},
  these two processes and the matrix building task can be performed on the unaltered
  \gls*{dmri} signal (i.e. without resampling data to an undistorted space).
For analyses other than connectivity, like \gls*{tbss}, the deformation estimated by \regseg{}
  can be used to map the tracts into structural space.
\revcomment[R\#3-C5]{Even though we apply \regseg{} to the problem of susceptibility distortion,
  \emph{it is not a distortion correction method}, but rather a surface alignment method.
In fact, the distortions are not corrected in the \gls*{epi} data.}
\revcomment[R\#3-C6]{Therefore, we suggest here to perform the reconstruction and tractography processes in the
  original (distorted) diffusion data.}
\revcomment[R\#1-C3]{\Regseg{} allows to avoid resampling and/or unwarping of the diffusion signal because the 
  structural information necessary in the diffusion analysis is mapped from the
  reference space.}
\revcomment[R\#1-C1]{Certain applications (like \gls*{tbss}) and methodologies (like building the connectivity matrix
  by clustering the tracks) may not be performed correctly on the native (distorted) diffusion space
  because they still need a mapping to the undistorted space.
Using \regseg{}, the tracks obtained in native space can be unwarped using the resulting estimation
  of the deformation field.}
\revcomment[R\#3-C6]{This methodological variation will be further investigated, to ensure which processing design
  yields the most accurate tractography results.}

Beyond the presented application on \gls*{dmri} data, \regseg{} can be indicated in situations
  where there are precise surfaces delineating the structure, a target multivariate
  image in which the surfaces must be fitted, and the mapping between the surfaces and
  the volume encodes relevant physiological information, such as the normal/abnormal
  development or the macroscopic dynamics of organs and tissues.
For instance, \regseg{} may be applied in fields like neonatal brain image segmentation
  in longitudinal MRI studies of the early developmental patterns \citep{shi_neonatal_2010}.
In these studies, the surfaces obtained in a mature time point of the brain are retrospectively
  propagated to the initial time points, regardless of the changes in the contrast and spatial
  development between them.
More generally, \regseg{} may also be applied to the personalized study of longitudinal alteration
  of the brain using multispectral images, for instance in the case of traumatic brain
  injury \citep{irimia_structural_2014} or in monitoring brain tumors
  \citep{weizman_semiautomatic_2014}.

\section*{Conclusion}
\label{sec:regseg-conclusion}

\Regseg{} is a variational framework for the simultaneous segmentation and
  registration of 3D \gls*{dmri} data obtained from the human brain, where within-subject
  anatomical information is used as a reference.
The registration method segments the target multivariate image into several competing regions, which are
  defined explicitly by their limiting surfaces.
The surfaces are active and they evolve on a free-form deformation field supported by the B-spline basis.
A descent optimization strategy is guided by shape gradients computed on the current partition
  of the target image.
\Regseg{} uses active contours without edges and it searches for
  homogeneous regions within the image.
We tested \regseg{} using digital phantoms by simulating \gls*{t1} and \gls*{t2} \acrshort*{mri}
  warped with smooth and random deformations.
The resulting misregistration of the contours was significantly lower than the image resolution
  of the phantoms.

We proposed \regseg{} for simultaneously segmenting and registering \gls*{dmri} data to
  their corresponding \gls*{t1} image from the same subject.
We demonstrated the accuracy of the proposed method based on visual assessments of the results
  obtained by \regseg{} and cross-comparisons with a widely used technique.
Moreover, \regseg{} does not require any images in addition to the minimal acquisition protocol,
  which only utilizes \gls*{t1} and \gls*{dmri}.
As well as the proposed application to \gls*{dmri} data, other potential uses of \regseg{} are
  atlas-based segmentation and tracking objects in time-series.

\section*{Availability and reproducibility statement}
\label{sec:regseg-availability}
We considered the reproducibility of our results as a design requirement.
Therefore, we used real data obtained from the Human Connectome Project \citep{essen_human_2012}
  and all of the software utilized in this study is also publicly available.
\Regseg{} was implemented on top of ITK-4.6 (Insight Registration and
  Segmentation Toolkit, \url{http://www.itk.org}).
The evaluation instruments (\autoref{fig:regseg-evworkflows}) were implemented using
  \emph{nipype} \citep{gorgolewski_nipype_2011} to assess their reproducibility.
All of the research elements (data, source code, figures, manuscript sources, etc.) involved in this study
  are publicly available under a unique package \citep{esteban_regseg_2015}.

\section*{Author contributions}
All the authors contributed to this study.
OE implemented the method, designed and conducted the experiments, wrote the paper,
  simulated the phantoms, and prepared the real data.
DZ devised and drafted the registration method, generated early phantom datasets, and
  collaborated in the implementation of the method.
AD, MBC, and MJLC interpreted the results.
AD, MBC, MJLC, JPT, and AS advised on all aspects of the study.

\section*{Acknowledgments}
The authors thank Y. Alem\'an and G. Wollny for their thorough reviews of the manuscript,
  V. Estellers for early discussions at the beginning of this project,
  and L. A. Vese for her support during OE's research visits to her laboratory.
We also thank A. Leemans for kindly sharing a p-code version of \emph{ExploreDTI} from
  which the settings for \emph{elastix} could be extracted.

DZ was supported by the Swiss National Science Foundation under grants PBELP2-137727, 
  P300P2-147778, and NSF-DMS 1418812.
This study was supported by the Spanish Ministry of Science and Innovation
  (projects TEC-2013-48251-C2-2-R and INNPACTO XIORT), Comunidad de Madrid (TOPUS) and
  European Regional Development Funds, the Center for Biomedical Imaging
  (CIBM) of the Geneva and Lausanne Universities and the EPFL, as well as the
  Leenaards and Louis Jeantet Foundations.

\onecolumn
\appendix
\numberwithin{equation}{section}

\renewcommand{\theequation}{A.\arabic{equation}}
\renewcommand{\thesubsection}{Appendix \arabic{subsection}}

\section*{Appendix}

\subsection{Simplifying the regularization term}\label{app:reg_term}
The exponentials of the Thikonov regularization prior \eqref{eq:regseg-thikonov} have the general form
  $\vec{v}^T \mathbf{M} \vec{v}$.
If $\mathbf{M}$ is a $n \times n$ diagonal matrix such that $\mathbf{M} = \vec{m} \, \mathbf{I}_n$,
  then:

\begin{equation*}
\vec{v}^T \mathbf{M} \vec{v} = \vec{m} \cdot (\vec{v}^T \mathbf{I}_n \vec{v}) = \vec{m} \cdot \vec{v}^{\circ2},
\end{equation*}
  where we have introduced the Hadamard power notation\footnote{The Hadamard power of a matrix or a vector
  is the power of its elements $\mathbf{M}^{\circ p} = ({m_{ij}}^{p})$}.

In general, the anisotropy of the distortion field is aligned with the 
  voxel coordinate system, so
  $\mathbf{A}$ and $\mathbf{B}$ of \eqref{eq:regseg-energy} can be simplified to diagonal matrices
  to regularize the registration process, such that
  $\mathbf{A}= \, \boldsymbol{\alpha}\,\vec{I}_n$ and
  $\mathbf{B}= \, \boldsymbol{\beta}\,\vec{I}_n$.
By substituting into equation \eqref{eq:regseg-energy}, we obtain:

  \begin{align}
  E(\vec{u}) &= \const + \, \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \,
  \ifthenelse{\boolean{review}}{+}{+ \notag\\ &+}
  \int_{\Omega} \frac12 \left[ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2} + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} \right] \,d\vec{r}.
  \label{eq:regseg-app_energy}
  \end{align}

\subsection{Application of the shape-gradients}\label{app:shape_gradients}
The computation of gradients at the locations of the active contours in the
  instant $t$ is based on the work of \cite{herbulot_segmentation_2006}.
Let $F(\vec{r})$ be an ``arbitrary'' function over the image domain
  $\Omega = \Omega_l \cup \Omega_m$ split in two regions $l$ and
  $m$, and $\Gamma_{l,m}$ a closed boundary between them.
We now derive the domain integral w.r.t. $t$:

  \begin{equation}
  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) d\vec{r} =
  \int_\Omega \frac{\partial}{\partial t}F(\vec{r}) d\vec{r}
  - \int_{\Gamma_{l,m}} F(\vec{r}) \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
  N_{\Gamma_{l,m}}\right\rangle d\vec{r},
  \end{equation}
  where $\left\langle\frac{\partial\Gamma_{l,m}}{\partial t}, N_{\Gamma_{l,m}}\right\rangle$ is
  the projection of the boundary movement on the unit inward normal $N_{\Gamma_{l,m}}$.
Assuming that the region descriptors $\{\boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l\}$ vary slowly enough, we can consider
  that $\frac{\partial}{\partial t} F(\vec{r}) = 0$ and thus:

  \begin{equation}
  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) d\vec{r} =
  - \int_{\Gamma_{l,m}} F(\vec{r}) \left\langle \frac{\partial \Gamma_{l,m} }{\partial t},
  N_{\Gamma_{l,m}}\right\rangle d\vec{r}.
  \label{eq:regseg-shape_gradients}
  \end{equation}

The equation \eqref{eq:regseg-shape_gradients} is discretized as follows.
First, the surface between limiting regions $l$ and $m$ ($\Gamma_{l,m}$) is explicitly represented by
  a discrete set of vertices $\vec{v}_i$, with $i \in \{0, \ldots, N_p -1 \}$.
Consequently, the inwards normal of the surface $N_{\Gamma_{l,m}}$ is represented by the discrete
  set of normals $\hat{\vec{n}}_i$ at each vertex of the mesh.
The resulting summation is, therefore, discrete and the integral operator is replaced by the sum:

  \begin{align}
  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) d\vec{r} &=
  \underbracket{\cancel{\int_\Omega \frac{\partial}{\partial t}F(\vec{r}) d\vec{r} }}_{\text{Functional's evolution}}
  - \underbracket{\int_{\Gamma_{l,m}} F(\vec{r}) \left\langle \frac{\partial \Gamma_{l,m}}{\partial t},
  N_{\Gamma_{l,m}}\right\rangle d\vec{r}}_{\text{Shape's evolution}} \notag \\
  & = - \underset{p}{\sum} \frac{1}{A_p} \underset{i}{\sum} \, a_i \, F(\vec{v}_i) \left\langle \underbracket{\frac{\partial \vec{v}_i}{\partial t}}_{\text{speed of }\vec{v}_i},
  \hat{\vec{n}}_{i}\right\rangle.
  \label{eq:regseg-shape_gradient_orig}
  \end{align}
where $a_i$ is the area corresponding to vertex $\vec{v}_i$, and $A_p = \sum_i a_i$ is the total area of the surface $p$.
In the following, we will refer as $w_{p,i} = a_i / A_p $ to the area contribution of $\vec{v}_i$ to the
  total area of the surface it belongs to.
For simplicity, the sum over $p$ can be also removed, as the vertices belong to only one of the total $P$ contours.

Then, the speed of $\vec{v}_i$ is discretized using the artificial time-step parameter $\delta$, as the displacement
  $\frac{\partial \vec{v}_i}{\partial t} = \vec{v}_i(\delta = t+1) - \vec{v}_i(\delta = t)$:

  \begin{equation}
  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) \, d\vec{r} =
  - \underset{i}{\sum} w_{p,i} F(\vec{v}_i) \frac{\partial \vec{v}_i}{\partial t} \cdot \hat{\vec{n}}_i.
  \label{eq:regseg-shape_gradient_disc1}
  \end{equation}

Since the energy functional is defined over competing regions, the displacement of $\vec{v}_i$ will cause
  an energy exchange between the limiting regions, and therefore $F(\vec{r})$ must be split in
  two terms, $F_{in}(\vec{r})$ corresponding to the interior region and $F_{out}(\vec{r})$ to the exterior:

  \begin{equation}
  \frac{\partial}{\partial t} \int_\Omega F(\vec{r}) \, d\vec{r} =
  - \underset{i}{\sum} \, \frac{\partial \vec{v}_i}{\partial t} \cdot
  \underbracket{w_{p,i} \, \Big[ F_{out}(\vec{v}_i) - F_{in}(\vec{v}_i) \Big] \hat{\vec{n}}_i}_{\bar{s}_i \text{ in Figure 1}}.
  \label{eq:regseg-shape_gradient_disc2}
  \end{equation}

Then, we identify the shape gradient contribution $\vec{g}_k$ on the coefficients $\vec{u}_k$ of the B-spline grid 
\revcomment[R\#1-C4]{to obtain the definition of $\vec{g}_k$ given in \eqref{eq:regseg-final_gradient}}:

  \begin{equation}
  \label{eq:regseg-gradient_wshape}
  \begin{split}
  \vec{g}_k &= - \underset{i}{\sum} \left\langle \frac{\partial \vec{v}_i'}{\partial \vec{u}_k}, \bar{s}_i'\right\rangle \\
  \text{with }
  \bar{s}_i' &= w_i \left[ \mdist{f_i'}{out} - \mdist{f_i'}{in} \right] \, \hat{\vec{n}}_i, \\
  \text{and }
  \frac{\partial \vec{v}_i'}{\partial \vec{u}_k} &=
  \frac{\partial}{\partial \vec{u}_k} \left\{ \vec{v}_i + \sum_k \psi_k(\vec{v}_i) \vec{u}_k \right\} = \psi_k(\vec{v}_i)\, \hat{\vec{e}},
  \end{split}
  \end{equation}\revcomment[R\#1-C4]{where $\psi_k$ and $\vec{u}_k$ define our B-spline deformation model (\autoref{eq:regseg-nodes_tfm})} and
  $\hat{\vec{e}}$ is the coordinates system's unit vector.

 \ifthenelse{\boolean{review}}
{}
{\twocolumn}

\bibliographystyle{mystyle}
\bibliography{Remote}

\end{document}
